### 2023-03-28 alloscore_talk/pres.qmd
getting started

---
title: "alloscore talk"
format:
  revealjs: 
    width: 1500
editor: visual
---

## Motivation

Gov to allocate scarce hospital supplies across 50 states

-   given many (probabilistic) forecasts of need, which should they use?

Raises question: how will chosen forecast be used?

One possibility: to minimize $E_F \sum_{l=1}^{51} \max(0, Y_i - x_i)$

-   $Y_{1:51} \sim F$ are future needs ($F$ unknown)
-   $x_{1:51}$ are allocated quantities
-   only $K$ units of supplies are available

## A constrained stochastic optimization problem

Let $\overline{s}_F(\mathbf{x}) = E_F \sum \max(0, Y_i - x_i)$

Lagrange multiplier equation: $\nabla \overline{s}_F(\mathbf{x}) + \lambda =0$

$$
\frac{d}{dx_i}E_F \max(0, Y_i - x_i) = \frac{d}{dx_i}\int_{x_i}^{\infty} (y - x_i)f_i(y)dy = F_i(x_i)-1 = -\lambda\\
x_i = F_i^{-1}(1-\lambda)
$$

Solve $\sum_{l=1}^{51} F_i^{-1}(1-\lambda) = K$ for $\lambda^{\star} \in (0,1)$ to get best allocation $x_i^{\star} = F_i^{-1}(1-\lambda^{\star})$

## Solution for location-scale family $F_i$'s

$$
\sum F_i^{-1}(1-\lambda) = \sum \mu_i + \sigma_i \Phi^{-1}(1-\lambda )= K \\
\implies \lambda^{\star} = 1 - \Phi\left(\frac{K-\sum \mu_j}{\sum \sigma_j}\right) \\
\implies x_i^{\star} = 
F_i^{-1}\circ\Phi\left(\frac{K-\sum \mu_j}{\sum \sigma_j}\right)
=\mu_i+\sigma_i\left(\frac{K-\sum \mu_j}{\sum \sigma_j}\right)
$$

Interpretation: underdispersion at loc $i$ -\> too little of "excess or shortage in mean" $K-\sum \mu_j$ given to or removed from $i$

## Allocation as forecast perfomance

$\mathbf{x}_F^{\star}$ is a **Bayes act** for $F$ under loss $$
s(\mathbf{x}, \mathbf{y}) = \sum \max(0,y_i - x_i)
$$ The quality of forecast $F$ then can be estimated by how much loss using it actually incurs, i.e., $$
S(\mathbf{F}, \mathbf{y}) = \sum \max(0,y_i - x_{F,i}^{\star})
$$

-   or perhaps $S(\mathbf{F}, \mathbf{y}) - \max(0, \sum y_i-K)$, performance against an oracle

## Scoring function for LS case

$$
s_K(\mathbf{Q}, \mathbf{y})=\sum \mathbb{1}\left\{\frac{K-\sum \mu_i}{\sum \sigma_i} \leq \frac{y_i-\mu_i}{\sigma_i}\right\}\left(y_i-x_{F,i}^{\star}\right)
$$

## Scoring Functions in general

Forecaster $A$ asked to estimate a **functional** $T(F_Y)$ for a RV $Y\sim F_Y$.

-   mean, median, variance, etc.
-   $A$ believes that $Y \sim F$

After $A$ gives us $x$ and nature realizes $Y=y$, we judge performance with a **scoring function** $s(x,y)$

$s$ is **(strictly) consistent** for $T$ if$A$ cannot expect to do better than offer $x=T(F)$ as a forecast of $T(F_Y)$, i.e,

$$
E_F[s(T(F)),Y] (<) \leq E_F[s(x,Y)] \text{ for all } F,x
$$

## Elicitability

A functional is **elicitable** if it has a strictly consistent scoring function.

Basic example are $\alpha$ quantiles, with SF:

$$
\begin{aligned}
s(x, y)=s_{O, U}(x, y) &=O(g(x)-g(y))_{+}+U(g(x)-g(y))_{-} . \\
&=\kappa\left((1-\alpha)(g(x)-g(y))_{+}+\alpha(g(x)-g(y))_{-}\right) \\
&=\kappa(\mathbb{1}\{x>y\}-\alpha)(g(x)-g(y)) \\
&:=s_{\kappa, \alpha}(x, y):=\kappa s_\alpha(x, y)
\end{aligned}
$$

$\alpha=U /(U+O), \kappa = O+U$

## Reformulation of allocation problem

Hosp allocation problem of this form in each location, with $O_i=0, U_i=1$

With constraint $x \leq K$ Bayes act is just $Q=\min(K,F^{-1} (U/(U+O)))=K$

But taken jointly, have non-trivial BA $Q_i = F_i^{-1}\left(\alpha_i\left(1-\lambda_K / U_i\right)\right)$ or(!) 0

-   $\lambda_K =0$ corresponds to case when $\sum Q_i(\alpha_i) \leq K$

## Domain versions

### Newsvendor problem

Ordering decision faced by a retailer of a perishable good (such as newspapers) when customer demand is uncertain.

-   Allocation problem appears as **multi-product NVP with stocking constraint**
-   Well-studied in OR literature

### Meteoroligist/Epidemiologist Cost/Loss Problem

-   have not seen much in allocation form

No literature regarding allocation solution as scoring rule

## Solving Allocation Problem

Straightforward in simple situations

-   all $F_i$ smooth with same support starting at 0
-   $K$ not too much smaller than $\mathbf{w}^T \mathbf{Q}(\alpha)$

$$
\nabla Z(\mathbf{x}) + \lambda \mathbf{w} =0 \\
\implies \kappa_i(F_i(Q_i) - \alpha) + \lambda w_i = 0 \\
\implies Q_i(\lambda) = F_i^{-1}(\alpha_i(1 - \lambda w_i / U_i))
$$

Now solve $\sum_{i=1}^N w_i Q_i(\lambda)=K$ for $\lambda^{\star}$ to get optimal allocation $\mathbf{Q}(\lambda^{\star})$.

## Complications

If some $Q_i$ need to be 0, we don't know in general which set of 0's is best

-   problem becomes intractable for large $N$
-   need a non-linear programming algorithm... `alloscore` implements one

## Scoring Rules

-   Should be regret-based, i.e., how well you do against an oracle that knows $y$'s and never wastes any resources

## Next Steps

-   Use Evan's package [`distfromq`](https://github.com/reichlab/distfromq) to interpolate quantile functions and distribtuions from hub quantile forecasts and score across $K$ range

-   extend this to no overprediction penalty i.e. $\alpha_i = 1$ scenario

## Aspirations

-   Allocation skill could be an indicator for peak/outbreak detection skill

### 2023-03-28 random_undocumented_experiments/iter_pic.R

@@ -1,0 +1,147 @@
+library(tidyverse)
+library(patchwork)
+library(devtools)
+load_all()
+
+(dat <- tibble(
+  alpha = c(.2,.6,.3,.9,.5),
+  w = c(3,2,3,5,2),
+  dists_and_params = list(
+    list(dist = "norm", mean = 3, sd = 1),
+    list(dist = "norm", mean = 2, sd = 4),
+    list(dist = "exp", rate = .5),
+    list(dist = "gamma", shape = 1, rate = .3),
+    list(dist = "unif", min = 5, max = 6)
+    #list(dist = "distfromq", ps = c(.1,.3,.6,.9,.95), qs = c(2,3,3,4,5))
+  ),
+  name = map(dists_and_params,
+             ~toString(paste0(.[["dist"]], ", ",
+                              toString(paste(names(.)[-1], .[-1], sep = "=")))))) %>%
+  unnest_wider(col = dists_and_params) %>%
+  add_pdqr_funs(types = c("p", "q")) %>% mutate(
+    q_scale = map2_dbl(Q, alpha, exec),
+    # add marginal expected benefit functions scaled to fit on single [0,1] interval,
+    Lambda_scaled = pmap(., margexb_fun),
+    # and unscaled as well, which are not yet being used
+    Lambda = pmap(.[names(.) != "q_scale"], margexb_fun),
+    Eloss = pmap(.[c("alpha", "F")], gpl_loss_exp_fun)
+  )
+)
+
+make_iter_pic(2, dat)
+
+make_iter_pic <- function(K, dat) {
+  (dat_allo <- dat %>%
+     mutate(
+       allo = allocate(
+         F = F,
+         Q = Q,
+         alpha = alpha,
+         w = w,
+         K = K,
+         eps_K = .01,
+         eps_lam = .0001
+       ),
+       allo_q_scaled = allo/q_scale
+     ) %>%
+     relocate(allo, allo_q_scaled, q_scale)
+  )
+
+  # get all iterations of lambda from allocation algorithm
+  long <- with(dat_allo,
+               allocate(
+                 F = F,
+                 Q = Q,
+                 alpha = alpha,
+                 w = w,
+                 K = K,
+                 eps_K = .01,
+                 eps_lam = .0001,
+                 Trace = TRUE
+               ))
+  lam_iters <- long$lambdas
+
+
[…]

### 2023-03-28 random_undocumented_experiments/plotting.R

@@ -1,0 +1,79 @@
+
+plot_loss <- function(
+    p = NULL,
+    alphas,
+    lossfun,
+    parms,
+    pfun,
+    qfun,
+    mag,
+    xl,
+    xr,
+    yb,
+    yexpand = c(0,.1)) {
+  if (is.null(p)) {
+    p <- ggplot() + xlim(xl,xr)
+  }
+  p <- p + map(
+    alphas,
+    function(alpha) {
+      geom_function(aes(color = as_factor(alpha)),
+                    fun = rlang::exec(lossfun, alpha, !!!parms), n = 500)
+    }
+  ) + map(
+    alphas,
+    function(alpha) {
+      geom_segment(aes(color = as_factor(alpha),
+                       x = rlang::exec(qfun, alpha, !!!parms),
+                       y = yb,
+                       xend = rlang::exec(qfun, alpha, !!!parms),
+                       yend = mag*alpha),
+                   alpha = .5)
+    }
+  ) + map(
+    alphas,
+    function(alpha) {
+      geom_segment(aes(color = as_factor(alpha),
+                       x = rlang::exec(qfun, alpha, !!!parms),
+                       y = mag*alpha,
+                       xend = xl,
+                       yend = mag*alpha),
+                   alpha = .5)
+    }
+  ) +
+    geom_function(fun = function(q) mag*rlang::exec(pfun, q, !!!parms), n=500) +
+    scale_y_continuous(expand = yexpand,
+                       labels = function(y) y/mag,
+                       breaks = mag*alphas,
+                       limits = c(yb, mag*1.1)) +
+    scale_x_continuous(expand = c(0,.1), limits = c(xl,xr)) +
+    theme_classic() +
+    theme(legend.position = "none",
+          axis.title = element_blank())
+  return(p)
+}
+
+
+ggplot() + map(datu$F, function(F) geom_function(fun = F, n = 1000)) + xlim(c(-10,15)) +
+  map(datu$Eloss, function(Eloss) geom_function(fun = Eloss, n = 1000))
+
+ggplot() + map(dat$f, function(f) geom_function(fun = f, n = 1000)) + xlim(c(-10,15))
+
+xl = 0
+xr = 10
+plt_dat <- datu
+p <- ggplot(data = data.frame(
[…]

### 2023-03-28 alloscore_manuscript/allocation.rmd
hasty first draft of outline

---
title: "Evaluating infectious disease forecasts with allocation scoring rules"
date: "`r Sys.Date()`"
output: 
    pdf_document:
        latex_engine: xelatex
urlcolor: blue
bibliography: allocation.bib 
header-includes:
   - \usepackage{amsmath}
   - \usepackage{algorithm}
   - \usepackage[noend]{algpseudocode}
   - \usepackage{amsthm}
   - \usepackage{bm}
   - \usepackage{cases}
   - \usepackage{caption}
   - \usepackage{unicode-math}
   - \usepackage{hyperref}
   - \DeclareMathOperator*{\argmin}{argmin}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE, 
    message = FALSE, 
    warning = FALSE,
    fig.width=10, 
    fig.height=6)
library(tidyverse)
library(patchwork)
```
<!-- $ ...to toggle Latexyz math scope off -->
\newcommand{\del}[2]{\frac{\partial {#1} }{\partial {#2}} }
\newcommand{\dby}[2]{\frac{d {#1} }{d {#2}} }
\newcommand{\sbar}{\overline{s}}
\newtheorem{proposition}{Proposition}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\section{Introduction}

Infectious disease forecasts have been used as an input to public health decision-making processes such as ...

In this manuscript we argue that for many decision-making problems, it is possible and valuable to develop scoring rules for probabilistic forecasts that are responsive to that decision-making context.

\subsection{Literature review}

Previous work in infectious disease forecasting where applications of forecasting to decision-making problems have been developed

Theoretic work on decision-theoretic set up for forecast evaluation

Previous applications of decision-theoretic evaluation to fields like economics

Previous decision-theoretic work in infectious disease forecast evaluation.  I think there are at least a couple of papers out there at some level of formality.  Maybe Kat Shea has something?

operations research work on constrained allocation

\section{Methods}

\subsection{A review of probabilistic forecasting and the decision-theoretic set up for scoring rules}

General notation. scoring functions, scoring rules.

A high-level/intuitive discussion of the three-step recipe for obtaining scoring rules from loss/utility in a decision-making context.
\begin{enumerate}
\item Specify down a scoring function that measures the loss associated with taking action x when outcome y materializes
\item Given a probabilistic forecast F, determine the Bayes act
\item The scoring rule calculates the score as the loss incurred when the Bayes act was used
\end{enumerate}

\subsection{A review of quantile loss, CRPS, and the weighted interval score}

Set up of decision-making problem with costs for over-/under-prediction. Quantile/pinball loss.

Statement that this loss elicits quantiles as the Bayes act, derivations deferred to a supplement.

Interpretation of CRPS as an integral of quantile loss over quantile levels, WIS as a discrete approximation of CRPS.

\subsection{Allocation scoring rules}

Set of decision-making problem with costs for over-/under-prediction across multiple locations and a constraint on the total allocation. Allocation loss.

Description of the Bayes act at some level (derivations go in a supplement):
\begin{itemize}
\item divide available resources across locations so that according to $F$, the expected benefit of 1 additional unit is the same everywhere.
\item Statement of the result if $g_i(x) = x$.
\end{itemize}

A figure with one or two sub-panels to illustrate.  What figures?

Description of what happens if there is uncertainty about the costs $O$, $U$, or the constraint $K$.  Integration across those results in something like a forecaster-specific weighted CRPS.

\section{Applications}

\subsection{Application with a quantile elicitation problem}

Consider one of the following problems?
\begin{itemize}
\item "should we expand hospital bed capacity"
\item "should impose a stay at home order".
\end{itemize}
for individual locations with specific costs for taking the action and possible costs for "under-acting". Both are I think hard to operationalize in the sense that it's not immediate what specifically those costs should be.

\subsection{Application with an allocation problem}

Hospital admissions in the U.S.

Case study heading into the Omicron wave. Some more detailed discussion of implications of bad forecasts for specific decision-making purposes -- take a "deep dive" into one or two example states like FL.

Look at results over a broader range of time.

\section{Discussion}

We often conceive of infectious disease forecasts as being useful for decision-making purposes, but it is rare for forecast evaluation to be tied directly to the value of the forecasts for informing those decisions. This work seeks to address that gap.

We have demonstrated that evaluation methods that are tied to decision-making context can yield model rankings that are substantively different from generic measures of forecast skill like WIS.

In practice, there are many users of forecasts with many different decision-making problems. Not all can be easily quantified. Those that can be easily quantified may differ enough that no single score is appropriate for all users. We suggest reporting multiple scores. This may be tricky to operationalize in the setting of a general forecast hub. It matters how you elicit and represent probabilistic forecasts (quantiles? samples? cdfs?).

The allocation score we developed here does not directly account for important considerations such as fairness/equity of allocations.

The allocation score we developed also does not attempt to capture the broader context of decision-making. For example, in practice it may be possible to increase the resource constraint $K$ by shifting funding from other disease mitigation measures.

There is much more to do in this general area.

\section{References}

### 2023-03-28 alloscore_manuscript/allocation.rmd

@@
 
+Note: it may make sense to talk about less theory in this article?
+
 \subsection{A review of probabilistic forecasting and the decision-theoretic set up for scoring rules}
@@
 
-\section{Applications}
+\section{Application}
 
-\subsection{Application with a quantile elicitation problem}
-
-Consider one of the following problems?
-\begin{itemize}
-\item "should we expand hospital bed capacity"
-\item "should impose a stay at home order".
-\end{itemize}
-for individual locations with specific costs for taking the action and possible costs for "under-acting".
-Both are I think hard to operationalize in the sense that it's not immediate what specifically those costs should be.
-
-\subsection{Application with an allocation problem}
-
-Hospital admissions in the U.S.
+We illustrate with an application to hospital admissions in the U.S., considering the problem of allocation of a limited supply of medical resources to the states.
 

### 2023-03-29 random_undocumented_experiments/example_application.R
switch on verbose for alloscore

@@ -1,9 +1,27 @@
 library(tidyverse)
 library(covidHubUtils)
+if (!require("distfromq")) {
+  devtools::install_github("reichlab/distfromq")
+}
 library(distfromq)
+if (!require("alloscore")) {
+  devtools::install_github("aaronger/alloscore")
+}
 library(alloscore)
 
-hub_repo_path <- "~/research/epi/covid/covid19-forecast-hub/"
+# Load the gh package to check for recent and probably crucial commits to alloscore
+if (!require("gh")) {
+  install.packages("gh")
+}
+if (
+  gh::gh("/repos/aaronger/alloscore/commits", .limit = 1)[[1]]$commit$committer$date > packageDate("alloscore")
+) {
+  message("Alloscore may be out of date...")
+}
+
+
+#hub_repo_path <- "~/research/epi/covid/covid19-forecast-hub/"
+hub_repo_path <- "~/covid/covid19-forecast-hub/"
 
 inc_hosp_targets <- paste(0:30, "day ahead inc hosp")
 forecasts_hosp <- load_forecasts(
@@ -92,6 +110,7 @@ names(Kdf) <- paste0("K=",Ks)
         dg = 1,
         eps_K = .01,
         eps_lam = 1e-5,
+        verbose = TRUE,
         against_oracle = TRUE
         ))
   ))

### 2023-03-30 alloscore_manuscript/allocation.rmd
start at draft of methods section

@@
 
-Note: it may make sense to talk about less theory in this article?
+We first give a high-level review of a general procedure for developing proper scoring rules that are tailored to a specific decision-making task, and then use that procedure to develop a score that is suitable for evaluation of forecasts in the context of decisions about allocation of limited resources across multiple locations.
 
-\subsection{A review of probabilistic forecasting and the decision-theoretic set up for scoring rules}
+\subsection{The decision-theoretic setup for forecast evaluation}
 
-General notation. scoring functions, scoring rules.
+In this section, we give an overview of the decision-theoretic setup for developing proper scoring rules that measure the value of a forecast as an input to decision making.
+We keep the discussion here at a somewhat informal level; we refer the reader to [some subset of Brehmer and Gneiting; Grünwald and Dawid; Dawid; Granger and Pesaran 2000; Granger and Machina 2006; Ehm et al.
+2016] for more technically precise discussion.
 
-A high-level/intuitive discussion of the three-step recipe for obtaining scoring rules from loss/utility in a decision-making context.
+In the framework of decision theory, a decision corresponds to the selection of an action $x$ from some set of possible actions $\mathcal{X}$.
+For example, $x$ may correspond to the level of investment in a measure designed to mitigate severe disease outcomes, with $\mathcal{X}$ being the set of all possible levels of investment that we might select.
+The quality of a decision to take a particular action $x$ is measured in relation to an outcome $y$ that is unknown at the time the decision is made.
+For example, $y$ may correspond to the number of individuals who eventually become sick and would benefit from the mitigation measure, and informally, an action $x$ is successful to the extent that it meets the realized demand.
+In the face of uncertainty, a decision-maker may use a forecast $F$ of the random variable $Y$ to help inform the selection of the action to take.
+We measure the value of a forecast as an input to this decision-making process by the quality of the decisions that it leads to.
+
+We can formalize the preceding discussion with the following three-step procedure for developing scoring rules for probabilistic forecasts:
 \begin{enumerate}
-\item Specify down a scoring function that measures the loss associated with taking action x when outcome y materializes
-\item Given a probabilistic forecast F, determine the Bayes act
-\item The scoring rule calculates the score as the loss incurred when the Bayes act was used
+\item Specify a \textit{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$ eventually occurs.
+We use the letter $s$ for this function to align with the literature on forecast evaluation, in which context $s$ may be used as a \textit{scoring function}.
+\item Given a probabilistic forecast $F$, determine the \textit{Bayes act} $x^F$ that minimizes the expected loss under the distribution $F$.
+\item The \textit{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: $S(F, y) = s(x^F, y)$.
 \end{enumerate}
+This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss function.
+Subject to certain technical conditions, scoring rules obtained from this procedure are proper (cite cite).
+
+<!-- , and we illustrate the ideas with two running examples: (1) some problem that results in quantile elicitation, (2) some allocation problem.
+
+A forecaster is asked to recommend, that is, \emph{predict}, the appropriate supply $x$ of some good or investment in some precautionary measure intended to satisfy a random future demand or need $Y$.
+As a running example, we will consider the provision of a limited supply of a limited health care resource such as oxygen or ventilators.
+Suppose there is an incremental loss $O \geq 0$ incurred when over-prediction leads to unused supply and an incremental loss $U > 0$ incurred when under-prediction leads to unmet demand or need.
+-->
 
@@
 
-Set up of decision-making problem with costs for over-/under-prediction. Quantile/pinball loss.
+I haven't written anything here.  Fill it in, or delete this section?
+
+<!-- Set up of decision-making problem with costs for over-/under-prediction. Quantile/pinball loss.
 
@@
 
-Interpretation of CRPS as an integral of quantile loss over quantile levels, WIS as a discrete approximation of CRPS.
+Interpretation of CRPS as an integral of quantile loss over quantile levels, WIS as a discrete approximation of CRPS. -->
 
-\subsection{Allocation scoring rules}
+\subsection{The allocation score}
 
-Set of decision-making problem with costs for over-/under-prediction across multiple locations and a constraint on the total allocation. Allocation loss.
+We now develop a scoring rule for probabilistic forecasts that measures the value of a forecast as an input to decision making about how to allocate limited resources to meet demand across multiple locations.
+As a concrete example, we take the resource to be a good such as ventilators or oxygen supply.
+An administrator is tasked with determining where to send these resources so as to meet demand among hospital patients in different facilities or states.
 
-Description of the Bayes act at some level (derivations go in a supplement):
-\begin{itemize}
-\item divide available resources across locations so that according to $F$, the expected benefit of 1 additional unit is the same everywhere.
-\item Statement of the result if $g_i(x) = x$.
-\end{itemize}
+We first specialize our notation to this decision-making setting.
+We define an action $\mathbf{x} = (x_1, \ldots, x_n)$ as a vector specifying the amount that is allocated to each of the $n$ locations.
+We require that each $x_i$ is non-negative, and that the total allocation across all locations does not exceed a constraint $K$ on the total available resources: $\sum_{i=1}^n x_i \leq K$.
+The set $\mathcal{X}$ collects all possible allocations that satisfy these constraints.
+The eventually realized resource demand in each location is denoted by $\mathbf{y} = (y_1, \ldots, y_n)$.
+At the time that a decision-maker sets the resource allocation, the demand $\mathbf{y}$ is not yet known.
+We therefore define the random vector $Y = (Y_1, \ldots, Y_n)$ where $Y_i$ represents the as-yet-unknown level of resource demand in location $i$.
+The forecast $F = (F_1, \ldots, F_n)$ collects forecasts of demand in each location.
+Here we identify $F_i$ with its cumulative distribution function (CDF), and $F_i^{-1}$ denotes the quantile function.
+We assume that the forecasts do not allow for the possibility of negative demand, i.e.
+the support of each $F_i$ is a subset of $\mathbb{R}^+$.
+With this notation in place, we can proceed to develop a proper scoring rule following the outline in the previous section.
 
-A figure with one or two sub-panels to illustrate.  What figures?
+\paragraph{Step 1: specify a loss function.} The loss associated with a particular allocation is calculated by comparing the amount allocated to each location to the realized resource demand in that location.
+Specifically, suppose that there is a marginal cost $L$ that accrues for each unit of demand that is not met.
+We can calculate the allocation loss across all locations as
+\begin{equation}
+s(\mathbf{x}, \mathbf{y}) = \sum_{i=1}^n L (x_i - y_i)_-, \label{eqn:loss_fn}
+\end{equation}
+where $(x_i - y_i)_- := \max(-(x_i - y_i),0)$ is $0$ if the amount $x_i$ allocated to unit $i$ is greater than or equal to the realized demand $y_i$ in that location; otherwise, it is $y_i - x_i$, the amount of unmet need in that location.
+We note that a number of generalizations to this loss specification have been formulated in the literature, including an allowance for costs for over-allocation to a particular unit (e.g.
+if there are storage costs for unused resources), differing losses different units (e.g.
+if a unit of unmet demand imposes more severe costs in one location than another), and the introduction of a convex function that controls the rate at which costs accrue depending on the scale of need.
+We consider these and other generalizations in other work, but for the present exposition we restrict our attention to the relatively simple loss formulation of Equation \eqref{eqn:loss_fn}.
 
-Description of what happens if there is uncertainty about the costs $O$, $U$, or the constraint $K$.  Integration across those results in something like a forecaster-specific weighted CRPS.
+\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act is the allocation that minimizes the expected loss:
+\begin{align}
+    \mathbf{x}^F &= \underset{\mathbf{x}\in \mathbb{R}^N, 0 \leq \mathbf{x}}{\mathrm{arg min}}\,\, \sbar_F(\mathbf{x}) \text{ subject to } 
+    \sum_{i=1}^N x_i \leq K \text{, where} \label{AP} \\
+    \sbar_F(\mathbf{x}) &= \mathbb{E}_{F} s(\mathbf{x}, \mathbf{Y}) = \sum_{i=1}^{N} \mathbb{E}_{F_i}[s(x_i, Y_i)] \label{eqn:sbar}
+\end{align}
+It can be shown that with the loss function given in Equation \eqref{eqn:loss_fn}, the Bayes act has $x^F_i = F_i^{-1}(1 - \lambda^{\star}/L)$, where $\lambda^{\star}$ is chosen so as to satisfy the equation
+\begin{align}
+\sum_{i=1}^{N}F_i^{-1}(1 - \lambda^{\star}/L) = K.
+\end{align}
+This partial solution to the allocation problem seems to have first appeared in [@hadleywhitin1963]; see the supplemental materials for a derivation.
+
+One interpretation of this result is that the Bayes act sets the allocation in each location $i$ to a quantile of the forecast distribution $F_i$ for that location.
+The quantile is at a probability level $(1 - \lambda^{\star}/L)$ that is the same for all locations, and is chosen such that the constraint is satisfied.
+An alternative interpretation comes from noting that for each location $i$, $\frac{\partial}{\partial x_i} \sbar_F(\mathbf{x}) \vert_{\mathbf{x}=\mathbf{x}^F} = \lambda^{\star}$ (see the supplement for a proof).
+<!-- 
+$\frac{\partial}{\partial x_i} \sbar_F(\mathbf{x}) = L[F_i(x_i) - 1]$ (see the supplement for a derivation), and evaluating at the Bayes allocation we arrive at
+\begin{align}
+\frac{\partial}{\partial x_i} \sbar_F(\mathbf{x}) \vert_{\mathbf{x}=\mathbf{x}^F} &= L[F_i(x^F_i) - 1] = L[F_i(F_i^{-1}(1 - \lambda^{\star}/L)) - 1] = \lambda^{\star}.
+\end{align} -->
+In words, at the allocation given by the Bayes act, the rate of change of the expected score as a function of the amount allocated to location $i$ is given by $\lambda^{\star}$.
+This derivative is the same for all locations, so the optimal allocation divides the available resources across all locations in such a way that according to $F$, the expected benefit of 1 additional unit of resources is the same in all locations.
+
+\paragraph{Step 3: Define the scoring rule.} We can now define a proper scoring rule for the probabilistic forecast $F$ as
+\begin{equation}
+S(F, y) = s(\mathbf{x}^F, y) = \sum_{i=1}^n L (F_i^{-1}(1 - \lambda^{\star}/L) - y_i)_-
+\end{equation}
+This score measures the total unmet need across all locations that results from using the Bayes allocation associated with the forecast $F$ when the actual level of need in each location is observed to be $y_i$.
+
+\textbf{ELR: This section would likely benefit from a figure or two to illustrate the ideas...?}
+
+Description of what happens if there is uncertainty about the the constraint $K$.
+Integration across $K$ results in something like a forecaster-specific weighted CRPS.
+It might help to introduce notation like $\lambda^{\star}(K)$ or $S_K(F, y)$ throughout all of the above discussion?
+Then if we have a prior $p(K)$ on $K$, we get to something like
+$$S(F, y) = \int S_K(F,y) p(K) \, d K$$
+as our final score.
 
@@
 
+Forecaster's dilemma: a successful forecast may lead to decisions that change the distribution of the outcome $Y$. Our framework cannot be used in those settings.
+
 There is much more to do in this general area.

### 2023-03-31 notes/allocation_notes.rmd
some reorg...

---
title: " Allocation scoring rules as a generalization of quantile scoring rules"
date: "`r Sys.Date()`"
output: 
    pdf_document:
        latex_engine: xelatex
urlcolor: blue
bibliography: allocation.bib 
header-includes:
   - \usepackage{amsmath}
   - \usepackage{algorithm}
   - \usepackage[noend]{algpseudocode}
   - \usepackage{amsthm}
   - \usepackage{bm}
   - \usepackage{cases}
   - \usepackage{caption}
   - \usepackage{unicode-math}
   - \usepackage{hyperref}
   - \DeclareMathOperator*{\argmin}{argmin}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE, 
    message = FALSE, 
    warning = FALSE,
    fig.width=10, 
    fig.height=6)
library(tidyverse)
library(patchwork)
```
<!-- $ ...to toggle Latexyz math scope off -->
\newcommand{\del}[2]{\frac{\partial {#1} }{\partial {#2}} }
\newcommand{\dby}[2]{\frac{d {#1} }{d {#2}} }
\newcommand{\sbar}{\overline{s}}
\newtheorem{proposition}{Proposition}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\section{Quantile Scoring Rules}

A forecaster is asked to recommend, that is, \emph{predict}, the appropriate supply $x$ of some good or investment in some precautionary measure intended to satisfy a random future demand or need $Y$. As a running example, we will consider the provision of a limited supply of a limited health care resource such as oxygen or ventilators. Suppose there is an incremental loss $O \geq 0$ incurred when over-prediction leads to unused supply and an incremental loss $U > 0$ incurred when under-prediction leads to unmet demand or need. Let $g$ be a non-decreasing function that expresses a utility associated with the good or misfortune to which $Y$ refers. This function can be thought of as specifying the rate at which costs accrue depending on the scale of need, through a mechanism that will be made precise below.

For notation, we will let
\begin{itemize}
\item $\mathcal{F_0}$ be the class of probability measures on the Borel-Lebesgue sets of $\mathbb{R}$
\item a probability measure $P \in \mathcal{F_0}$ induced by a random variable $X$ be identified by its cumulative distribution function $F_X$
\item $F$ denote the forecast distribution for the level of future need $Y$
\item $Y \sim G$
\item $Q_F := F^{-1}: [0,1] \rightrightarrows \mathbb{R}$ be the (set-valued) quantile function which maps a probability level $\alpha \in [0, 1]$ to the corresponding quantile set $\left[q_{\alpha, F}^{-}, q_{\alpha, F}^{+}\right]$ of $F$ where $q_{\alpha, F}^{-} := \sup \{x \mid F(x) < \alpha\}$ and 
$q_{\alpha, F}^{+} := \sup \{x \mid F(x) \leq \alpha\}$
\item $Q_\alpha: \mathcal{F_0} \rightrightarrows \mathbb{R}$ be the functional which maps a distribution $F$ to its $\alpha$-level quantile set
\item $q_{\alpha, F}$ be a (usually unique) element of $Q_F(\alpha) = Q_{\alpha}(F)$. 
\end{itemize}

We measure the forecaster's performance when the realized need is $y$ with the scoring function
\begin{align}
s(x,y)=s_{O,U}(x,y) &= O(g(x) - g(y))\sb{+} + U(g(x)-g(y))\sb{-}. \label{eqn:scalar_score} \\
&= \kappa((1-\alpha)(g(x) - g(y))\sb{+} + \alpha(g(x)-g(y))\sb{-}) \label{eqn:scalar_score_alpha} \\
&= \kappa(\mathbb{1}\{x>y\} - \alpha)(g(x)-g(y)) \label{eqn:scalar_score_ident}\\
&:= s_{\kappa, \alpha}(x,y) := \kappa s_{\alpha}(x,y),
\end{align}
where $\alpha = U/(U+O)$, $\kappa = U+O$,
$u_+ := \max(u,0)$, $u_- := \max(-u,0) = (-u)_+$, and $\mathbb{1}\{A\}$ is the indicator function of the event $A$.
This scoring function $s$ is negatively oriented, i.e., a smaller score indicates better performance corresponding to a smaller cost due to under- or over-prediction.

By making the forecaster aware of $s$ prior to prediction, we elicit an $\alpha$-level quantile $q_{\alpha,F}$ of the forecast distribution $F$ for $Y$. That is, if a forecaster believes that $Y \sim F$, then according to this belief they minimize their expected score
\begin{align} 
\sbar_F(x) = E_F[s(x,Y)]  &= O E_F[(g(x) - g(Y))\sb{+}] + U E_F[(g(x) - g(Y))\sb{-}] \label{eqn:scalar_obj}\\
&= O\int_{-\infty}^x(g(x) - g(y))f(y)dy + U\int_x^{\infty}(g(y) - g(x))f(y)dy \label{eqn:scalar_obj_int}
\end{align}
by forecasting a quantile $q_{\alpha, F} \in Q_F\left(\alpha\right)$ which solves the first order equation
\begin{align}
\nonumber 0 =\frac{d}{dx}\sbar_F(x) 
&= O E_F\left[\frac{d}{dx}(g(x) - g(Y))\sb{+}\right] + U E_F\left[\frac{d}{dx}(g(x) - g(Y))\sb{-}\right] \\
&= O E_F[g^{\prime}(x)\mathbb{1}\{Y < x\}] + U E_F[-g^{\prime}(x)\mathbb{1}\{Y \geq x\}] \\
&= O g^{\prime}(x) \mathbb{P}_F(Y<x) - U g^{\prime}(x) \mathbb{P}_F(Y\geq x) \\
&= g^{\prime}(x)\left( O F(x) - U (1-F(x))\right) \\
&= \kappa g^{\prime}(x)\left(F(x) - \alpha\right) \label{scalar_Z_deriv}.
\end{align}
(See \hyperref[scalar_Z_deriv_detail]{appendix} for an alternative derivation in terms of a density.)
The critical point $q_{\alpha, F}$ is actually a minimum since \eqref{scalar_Z_deriv} is non-positive on $\{x<q_{\alpha, F}\}$ and non-negative on $\{x>q_{\alpha, F}\}$.
 When $g^{\prime}(q)>0$, this minimum is unique since
\begin{align}
   \frac{d^2}{dx^2}\sbar_F(x)\big|_{x=q_{\alpha, F}} &= 
   \kappa\left(g^{\prime}(x)f(x) + g^{\prime \prime}(x)\left(F(x) - \alpha\right)\right)\big|_{x=q_{\alpha, F}} \label{scalar_Z_deriv2}\\
   &=\kappa(g^{\prime}(q_{\alpha, F})f(q_{\alpha, F}) + g^{\prime \prime}(q_{\alpha, F})\cdot 0) >0.
\end{align}

If $g^{\prime}(q_{\alpha, F})=0$ or does not exist, or $Q_F(\alpha)$ is not a singleton set, we'll need to adopt additional criteria for whether $q_{\alpha, F}$ is optimal. One possibility is to take $q_{\alpha, F}=\min \left\{x \mid F_X(x) \geq \alpha\right\}$, and if $g^{\prime}(q_{\alpha, F})=0$ we choose the prediction $x$ to be the infimum of the region around $q_{\alpha, F}$ where $g$ is flat. One example of such a $g$ is the \emph{power curve} of a wind turbine which becomes constant above a certain wind speed. Here $x$ would be a forecast of future wind speed $Y$ at time $t$ and $g(x)$ would be the power a turbine operator commits to supplying the electric grid at $t$. Taking $g$ to be constant below a certain point could also serve to express the futility or lack of meaning of forecasts below that point, such as when a forecast user is only able to sell a commodity for which demand can become negative, turning into supply.

Note that $\frac{d}{dx}\sbar_F(x) =\frac{d}{dx}E_F[s(x,Y)] = h(x)E_F[V(x,Y)]$, where $h(x) = \kappa g^{\prime}(x)$ and 
\begin{align}
V(x,y) &\coloneq (1-\alpha)\mathbb{1}\{y<x\} - \alpha \mathbb{1}\{y \geq x\} \mathrel{\phantom{=}}\\
&= (1-\alpha)\mathbb{1}\{y<x\} - \alpha \left(1 - \mathbb{1}\{y<x\}\right) \\
&= \mathbb{1}\{y<x\} - \alpha. 
\end{align}
By virtue of $E_F[V(Q_F(\alpha),Y)]=0$, $V$ is said to be an \emph{identification function} for the $\alpha$ quantile. The fact that, generally speaking, any elicitable functional (such as a quantile) has an identification function is known as \emph{Osband's principle}.

Forecasting the minimizer of $\sbar_F(x) = E_F[s_{\kappa,\alpha}(x,Y)]$ in this situation is known as the \emph{Bayes act} $a_F$ under $s_\alpha$ for the forecaster. Assuming a forecaster is informed, rational, and risk-neutral enough to take $a_F$, we can evaluate $F$, implicitly, as a distributional forecast, by the \emph{scoring rule} $S_{\alpha}$ induced by $s_{\alpha}$:
\begin{align}
S_{\alpha}(F,y) := s_{\alpha}(Q_F(\alpha), y).
\end{align}
If $y$ is generated by $Y \sim G$ then $S_{\alpha}(F,y)$ is an estimate of 
\begin{align}
S_{\alpha}(F,G) &:= E_G[S_{\alpha}(F,Y)] = E_G[s_{\alpha}(Q_{F}(\alpha),Y)].
\end{align}
This defines a functional on $\mathcal{F_0}$, the class of measures on the Borel-Lebesgue sets of $\mathbb{R}$ by
\begin{align}
\mathcal{\overline{S}}_{G,\alpha}(F) := \sbar_G(Q_{\alpha}(F))
\end{align}
where $Q_{\alpha}$ is the $\alpha$ quantile functional.

Using $u_- = u_+ - u$, the objective function \eqref{eqn:scalar_obj} can also be helpfully rewritten as
\begin{align}
\nonumber \sbar_F(x) &= OE_F[(g(x) - g(Y))\sb{+}] + U E_F[(g(x) - g(Y))\sb{+}] - U E_F[(g(x) - g(Y))] \\ 
\nonumber  &= (O+U)E_F[(g(x) - g(Y))\sb{+}] - U E_F[(g(x) - g(Y))] \\
&= (O+U) \int_{-\infty}^{x} (g(x) - g(y))f(y)dy - Ug(x)+ U \overline{g}\sb{F} \label{obj_integral}\\
&= (O+U)\left(g(x)F(x) - \overline{g}_F(x) \right) - Ug(x) + U\overline{g}\sb{F}) \label{obj_sep}\\
&= \kappa\left[g(x)(F(x) - \alpha) -\overline{g}_F(x) + \alpha \overline{g}\sb{F}\right] \label{obj_pexp} 
\end{align}
where $\overline{g}_F(x) = E_F[g(Y)\mathbb{1}\{Y \leq x\}]$, and $\overline{g}\sb{F} = E_F[g(Y)]$.

The function $\overline{g}_F(x)$ sometimes goes by the name of ``partial expectation" (Schlaifer Raiffa, p. 109) and can often be made more explicit for certain $F$ and $g$.

##  Alternative Notation Possibilities

Let $Y \sim F,F_i$ be various forecaster beliefs and $Y \sim G$ be reality.
\begin{align}
\overline{s}_{F,\kappa,\alpha}(x) &:= s_{\kappa,\alpha}(x,F) := E_F[s_{\kappa,\alpha}(x,Y)] \\
\overline{s}_{\kappa,\alpha}(x) &:= s_{\kappa,\alpha}(x,G) \\
s_{\symbfit{\kappa},\symbfit{\alpha}}(\mathbf{x}, \mathbf{F}) &:= 
\overline{s}_{\mathbf{F},\symbfit{\kappa},\symbfit{\alpha}}(\mathbf{x}) = 
\sum_{i}^{\mathrm{dim}\mathbf{x}} s_{\kappa_i,\alpha_i}(x_i, F_i) \\
S(F,y) &:= s(q_{F,\alpha}, y) \\
S(F_1,F_2) &:= E_{F_2}[S(F_1,Y)] = s(q_{F_1,\alpha}, F_2) \\
\overline{S}(F) &:= E_G[s(q_{F, \alpha}, Y)]
\end{align}

Define the $s$-entropy of an arbitrary $F$ 
\begin{align}
e_s(F) &:= E_F[S(F, Y)] = E_F[s(q_{F,\alpha}, Y)] = s(q_{F,\alpha}, F) \\
\end{align}
and of the RV $Y$ as
\begin{align}
e_s(Y) &:= e_s(G).
\end{align}
Define the $s$-divergence between two forecasts as
\begin{align}
d_s(F_1,F_2) &:= S(F_1,F_2) - S(F_2,F_2) = S(F_1,F_2) - e_s(F_2)
\end{align}
and the $s$-divergence from $F$ to $Y$ as
\begin{align}
d_s(F,Y) &:= S(F,G) - e(Y).
\end{align}

Thus, $\overline{S}(F) = d_s(F,Y) + e_s(Y)$.

## Examples for various forecaster beliefs

### 1) $Y \sim U[a,b]$

Here $F(x) = \mathbb{1}\{a<x<b\}(x-a)/(b-a) + \mathbb{1}\{b \leq x\}$ in \eqref{obj_pexp} gives 
<!--  
&= 
\begin{cases}
\frac{U}{b-a}\int_a^b (g(y) - g(x))dy & x\leq a \\[8pt]
\frac{O}{b-a}\int_a^x (g(x) - g(y))dy + \frac{U}{b-a}\int_x^b (g(y) - g(x))dy & a < x < b \\[8pt]
\frac{O}{b-a}\int_a^b (g(x) - g(y))dy & b \leq x 
\end{cases} \\ 
-->
\begin{align}
\sbar_F(x) &= 
\begin{cases}
\frac{\kappa}{b-a}\left( - \alpha(b-a)g(x) + \alpha\int_a^b g(y)dy\right) & x\leq a \\[8pt]
\frac{\kappa}{b-a}\left((x-a- \alpha(b-a))g(x) - \int_a^x g(y)dy + \alpha\int_a^b g(y)dy\right) & a < x < b \\[8pt]
\frac{\kappa}{b-a}\left((1-\alpha)(b-a)g(x) -(1-\alpha)\int_a^b g(y)dy \right) & b \leq x
\end{cases} 
\end{align}

When $g(x)=x$ this becomes
\begin{align}
\sbar_F(x) &= 
\begin{cases}
\frac{\kappa}{b-a}\left( - \alpha(b-a)x + \frac{\alpha}{2}(b^2-a^2)\right) & x\leq a \\[8pt]
\frac{\kappa}{b-a}\left((x-a- \alpha(b-a))x - \frac{1}{2}(x^2 - a^2) + \frac{\alpha}{2}(b^2-a^2)\right) & a < x < b \\[8pt]
\frac{\kappa}{b-a}\left((1-\alpha)(b-a)x -\frac{(1-\alpha)}{2}(b^2-a^2)\right) & b \leq x
\end{cases} \\
&= 
\begin{cases}
-\kappa\alpha\left[x - \frac{a+b}{2}\right] & x\leq a \\[8pt]
\frac{\kappa}{2}\left[\frac{1}{b-a}(x - (a+\alpha(b-a)))^2 +\alpha(1-\alpha)(b-a)\right] & a < x < b \\[8pt]
\kappa(1-\alpha)\left[ x - \frac{a+b}{2}\right] & b \leq x
\end{cases}
\end{align}

Thus $\sbar_F(x)$ has the form of pinball loss centered on the mean $\mu_Y = \frac{a+b}{2}$ and interpolated on $(a,b)$ 
 by a quadratic centered at the quantile $a+\alpha(b-a)$ with slopes matching the linear tails at $a$ and $b$.

```{r echo = FALSE, cache=TRUE}
unif_loss <- function(alpha, min, max) {
  f <- function(x){
    (x <= min)*(-alpha)*(x - (max+min)/2) +
    ((min < x) & (x < max))*.5*((1/(max-min))*(x - (min+alpha*(max-min)))^2 + alpha*(1-alpha)*(max-min)) +
    (x >= max)*(1-alpha)*(x - (max+min)/2)
  }
  return(f)
}
plot_loss <- function(
    alphas,
    lossfun, 
    parms, 
    pfun, 
    qfun,
    mag, 
    xl, 
    xr,
    yb,
    yexpand = c(0,.1)) {
  p <- ggplot() + xlim(xl,xr)+ map(
    alphas,
    function(alpha) {
      geom_function(aes(color = as_factor(alpha)), 
                    fun = rlang::exec(lossfun, alpha, !!!parms), n = 500)
    }
  ) + map(
    alphas,
    function(alpha) {   
      geom_segment(aes(color = as_factor(alpha), 
                       x = rlang::exec(qfun, alpha, !!!parms), 
                       y = yb,
                       xend = rlang::exec(qfun, alpha, !!!parms),
                       yend = mag*alpha),
                   alpha = .5)
    }
  ) + map(
    alphas,
    function(alpha) {
      geom_segment(aes(color = as_factor(alpha),
                       x = rlang::exec(qfun, alpha, !!!parms),
                       y = mag*alpha,
                       xend = xl, 
                       yend = mag*alpha),
                   alpha = .5)
    }
  ) + 
    geom_function(fun = function(q) mag*rlang::exec(pfun, q, !!!parms), n=500) +
    scale_y_continuous(expand = yexpand,
                       labels = function(y) y/mag, 
                       breaks = mag*alphas, 
                       limits = c(yb, mag*1.1)) +
    scale_x_continuous(expand = c(0,.1), limits = c(xl,xr)) +
    theme_classic() +
    theme(legend.position = "none",
          axis.title = element_blank())
  return(p)
}

alphas = c(.1,.4,.5,.9)
p_unif <- plot_loss(
  alphas = alphas,
  lossfun = unif_loss, 
  parms = list(min = 0, max = 1), 
  pfun = punif, 
  qfun = qunif,
  mag = 1, 
  xl = -1.5, 
  xr = 2.5,
  yb = 0
  ) + ggtitle("Unif(0,1)")
```

### 2) $g(x)=x$ and $F_Y=F_{\mu, \sigma}$ lies in a location-scale family

That is, $Y = \mu + \sigma Z$ and $Z \sim F_Z = F_0$. Then $f_Y(y) = \dby{}{z} F_Z\left(\frac{y-\mu}{\sigma}\right)\dby{z}{y} = f_Z(z)/\sigma$ and so
\begin{align}
\overline{g}_F(x) = \int_a^x y f_Y(y)dy &= \int_{\frac{a-\mu}{\sigma}}^{\frac{x-\mu}{\sigma}}(\mu + \sigma z)f_Z(z)dz \\
& = \mu F_Z\left(\frac{x-\mu}{\sigma}\right) + \sigma \int_{\frac{a-\mu}{\sigma}}^{\frac{x-\mu}{\sigma}}zf_Z(z)dz\\
& =\mu F_Y(x) + \sigma \mu_{F_0}\left(\frac{x-\mu}{\sigma}\right).
\end{align}
where $\mu_{F_0}(u) = \int_{\frac{a-\mu}{\sigma}}^{u} zf_0(z)dz$. This gives
<!-- With the change of variables $y=Q_F(\alpha) = F^{-1}(\alpha), dy =  d(F^{-1}(\alpha)) = \frac{d\alpha}{f(y)}$, we have
\begin{align}
\overline{g}_F(x) = \int_a^x y f_Y(y)dy = \int_{0}^{F(x)}Q_F(\alpha)d\alpha 
& = \mu F(x) + \sigma \int_0^{F(x)} Q_0(\alpha)d\alpha \\
& =\mu F(x) + \sigma \mu_{F_0}\left(\frac{x-\mu}{\sigma}\right).
\end{align} -->
\begin{align}
\nonumber \sbar_F(x) &= \kappa\left[x(F(x) - \alpha) - \mu F(x) - 
\nonumber \sigma \mu_{F_0}\left(\frac{x-\mu}{\sigma}\right) + \alpha \mu \right]  \\
&= \kappa\left[\left(F(x) - \alpha\right)(x-\mu)- \sigma \mu_{F_0}\left(\frac{x-\mu}{\sigma}\right) \right]
\label{lc-form}
\end{align}

### 3) $g(x)=x$ and $Y \sim N(\mu,\sigma)$

When $F=\Phi_{\mu,\sigma}$, $\mu_{F_0}\left(\frac{x-\mu}{\sigma}\right)=-\sigma \varphi_{\mu, \sigma}(x)$, giving $q(\alpha)$ as the minimum of 
$$
Z_F(x) = \kappa\left[(\Phi_{\mu,\sigma}(x) - \alpha)(x-\mu) + \sigma^2 \varphi_{\mu, \sigma}(x)\right].
$$

```{r echo = FALSE, cache=TRUE}
norm_loss <- function(alpha, mean, sd) {
  f <- function(x){
    alpha^(-1)*((pnorm(x, mean, sd)-alpha)*(x-mean) + sd^2*dnorm(x,mean,sd))
  }
  return(f)
}
parms = list(mean = 0, sd = 2)
p_norm <- plot_loss(
    alphas = alphas,
    lossfun = norm_loss, 
    parms = parms, 
    pfun = pnorm, 
    qfun = qnorm,
    mag = 10, 
    xl = -6, 
    xr = 6,
    yb = 0
) + ggtitle("Normal(0,2)")
```

### 4) $g(x)=x$ and $Y \sim \mathrm{Exp}(1/\sigma)$

Taking $\mu \equiv 0, Y = \sigma Z$ and $F_Z(x) = F_0(x) = \mathbb{1}\{0 \leq x\}(1 - e^{-x})$ we have
 $\mu_{F_0}(x) = \mathbb{1}\{0 \leq x\}(1 - e^{-x}(1+x))$. Then \eqref{lc-form} is

\begin{align}
\nonumber \sbar_F(x) &= 
\mathbb{1}\{0 \leq x\}\kappa\left[F(x)x - \sigma \mu_{F_0}\left(\frac{x}{\sigma}\right) \right] - \kappa\alpha x\\
&= \mathbb{1}\{0 \leq x\}\kappa
\left[
\left(1 - e^{-x/\sigma}\right)x - \sigma \left(1- e^{-x/\sigma}\left(1+\frac{x}{\sigma}\right) \right)
\right] - \kappa\alpha x\\
&= \mathbb{1}\{0 \leq x\}\kappa\left[ \sigma e^{-x/\sigma} + x - \sigma \right] - \kappa\alpha x
\end{align}

```{r echo = FALSE, cache=TRUE}
exp_loss <- function(alpha, rate) {
  f <- function(x){
    ((exp(-x*rate)*rate^(-1) + x - rate^(-1)))*(x>0) - alpha*x
  }
  return(f)
}
parms = list(rate = 1)
p_exp <- plot_loss(
    alphas = alphas,
    lossfun = exp_loss, 
    parms = parms, 
    pfun = pexp, 
    qfun = qexp,
    mag = 3, 
    xl = -2, 
    xr = 5,
    yb = -1
) + ggtitle("Exp(1)")
```

```{r, fig.cap="Plots of CDFs and expected quantile scores at levels .1, .4, .5, and .9"}
p_unif + p_norm + p_exp + plot_layout(ncol = 2)
```

## Another form for $\sbar_F(x)$

There also may be situations where it is convenient to integrate the $(O+U)$-term in \eqref{obj_sep} by parts to get 
\begin{align}
\sbar_F(x) = \kappa\left[\int_a^x g^{\prime}(y)F(y)dy - \alpha(g(x) - \overline{g}\sb{F})\right]. \label{obj_intF}
\end{align}

## Newsvendor parameters

In the inventory management literature, minimization of an expectation of the form of $\sbar_F(x)$ for a given $F$ arises in the \emph{newsvendor problem}.  This involves the ordering decision faced by a retailer of a perishable good (such as newpapers) when customer demand is uncertain. Because the focus here is on revenues and expenditures, terms other than just the over- and underprediction penalties in \eqref{eqn:scalar_score} often appear, giving a retailer's scoring function as
\begin{align}
s_n(x,y) & = a_x x +a_y y +a_{+}(x-y)_+ + a_{-}(x-y)_-.
\end{align}
We might have, for example, the score given by net value transfers
\begin{align}
s_n(x,y) = cx-py - u(x-y)_+ + (p+r)(x-y)_-
\end{align}
where $x$ newspapers are ordered at price $c$, $y - (x-y)_-$ are sold at retail price $p$, $(x-y)_+$ are sold at salvage price $u$, 
and $r(x-y)_-$ of customer "goodwill" is lost. 
This can be rewritten as
\begin{align}
s_n(x,y) &= (a_{+} + a_x)(x-y)_+ + (a_{-} - a_x)(x-y)_- +(a_x + a_y)y \\
&= (c-u)(x-y)_+ + (p+r-c)(x-y)_- -(p-c)y
\end{align}
giving the retailer's objective function under a demand distribution $F$ as
\begin{align}
\sbar_{n,F}(x) = (a_{+} + a_x)E_F[(x-Y)_+] + (a_{-} - a_x)E_F[(x-Y)_-] +(a_x + a_y)E[Y].
\end{align}
This differs by a constant from our $\sbar_F(x)$ in \eqref{eqn:scalar_obj} after taking $g(x)=x$, $O=a_{+} + a_x=c-u$ 
and $U=a_{-} - a_=p+r-c$, and so has the same minimizer   
\begin{align}
Q = F^{-1}\left(\frac{U}{O+U}\right) =  F^{-1}\left(\frac{a_{-} - a_x}{a_{-} + a_{+}}\right) 
= F^{-1}\left(\frac{p+r-c}{p+r-u}\right).
\end{align}
Note, for reference, that in the literature $\sbar_{n,F}(x)$ is often encountered in the form of \eqref{obj_integral}:
\begin{align}
&(O+U) \int_{0}^{x} (x - y)f(y)dy - Ux+ (U+a_x+a_y)E[Y] \\
=\,&(p+r-u) \int_{0}^{x} (x - y)f(y)dy - (p+r-c)x+ rE[Y]
\end{align}

## Meteorologist parameters

Similarly, a weather forecaster might be judged by the cost $Cx$ of recommended protection $x$ against a level $y$ of adverse weather (e.g., rainfall) added to any loss $L(x-y)_-$ resulting from underprediction, leading to the scoring function 
\begin{align}
s_m(x,y) = Cx + L(x-y)_-,
\end{align}
which rearranges to 
\begin{align}
s_m(x,y) = C(x-y)_+ + (L-C)(x-y)_- + Cy.
\end{align}
The minimizer of $\sbar_{m,F}(x) = E_F[s_m(x,Y)]$ given the forecaster's belief $Y \sim F$ is now
\begin{align}
Q = F^{-1}\left(\frac{L-C}{C+L -C}\right) =F^{-1}\left(1-\frac{C}{L}\right).
\end{align}
In particular, faced with the classical binary decision problem of whether to recommend an additional unit of protection given a current level of protection $x$, the forecaster's optimal decision rule under $s_m$ is to recommend adding protection if $x < F^{-1}\left(1-\frac{C}{L}\right)$, that is, if 
\begin{align}
1-F(x) = \mathbb{P}_F\{y>x\} > \frac{C}{L},
\end{align}
the \emph{cost-loss ratio} of the problem.

# Multipoint Quantile Scoring Rules

Suppose we ask a forecaster to provide a set of point forecasts $\{x_1,x_2\}$ to be rewarded according to the aggregate scoring function 
\begin{align}
s(x_1,x_2,y) &= s_{O_1,U_1}(x_1,y) + s_{O_2,U_2}(x_2,y).
\end{align}

# Allocation scoring rules
We now develop \emph{allocation scoring functions} that elicit forecasts $x_i$ for outcomes $y_i, i = 1,\ldots,N$, each with it's own incremental cost $O_i$ and loss $U_i$ as for a quantile scoring rule, but with the additional constraint 
\begin{align} 
 \sum_{i=1}^N w_i x_i = \mathbf{w}^T \mathbf{x} \leq K
\end{align} 
on the total provision available with $w_i>0$. We assume $K>0$, that only non-negative forecasts $x_i$, i.e., recommended allocations, are accepted, and that 
\begin{align}
 g_i^{\prime}(x_i)>0 \text{ for } x_i \geq 0,
\end{align} 
though it may be interesting to introduce regions where $g_i(x_i)$ is constant, such as for the wind speed power curve mentioned above.

According to \eqref{eqn:scalar_obj_int} and \eqref{obj_pexp} the objective function is now
\begin{align}
\sbar_F(\mathbf{x})&= \sum_{i=1}^{N} \sbar_{F_i}(x_i) = \sum_{i=1}^{N} \left\{O_i\int_{-\infty}^{x_i}(g_i(x_i) - g_i(y))f_i(y)dy + 
U_i\int_{x_i}^{\infty}(g_i(y) - g_i(x_i))f_i(y)dy \right\} \\
&= \sum_{i=1}^{N} \kappa_i
\left[
g_i(x_i)(F_i(x_i) - \alpha_i) -\overline{g_i}_{F_i}(x_i) + \alpha_i \overline{g_i}\sb{F_i}
\right] 
\end{align}

and the elicited forecasts solve the allocation problem (AP)
\begin{align}
    \underset{\mathbf{x}\in \mathbb{R}^N, 0 \leq \mathbf{x}}{\mathrm{minimize}}\,\, \sbar_F(\mathbf{x}) \text{ subject to } 
    \mathbf{w}^T\mathbf{x} \leq K. \label{AP}
\end{align}

As a separable problem, this is amenable to being solved via its dual (see [@ruszczynski2011nonlinear], section 4.4). With the Lagrangian
\begin{align}
L(\mathbf{x}, \lambda; K) = \sum_{i=1}^{N} \sbar_{F_i}(x_i) + \lambda \left(\sum_{i=1}^{N}w_i x_i - K\right) \label{Lagrangian}
\end{align}
we have the dual problem of maximizing over $\lambda \geq 0$ the objective function
\begin{align}
L_D(\lambda; K) &= \min_{0 \leq \mathbf{x}} L(\mathbf{x}, \lambda) = 
-\lambda K + \sum_{i=1}^{N} \underset{0 \leq x_i}{\min} \left\{\lambda w_i x_i + \sbar_{F,i}(x_i)\right\} \label{dual_eq}
\end{align}
which gives lower bounds on $\sbar_F(\mathbf{x}^{\star})$ at a solution $\mathbf{x}^{\star}$ to the primal AP. In particular, when $O_i>0$ for all $i$, $L_D(0)$ gives the total score $\sbar_F(\mathbf{q}_{\symbfit{\alpha}, F})$ when the constraint does not prevent the forecaster from giving the quantiles in each component. When the constraint is active, the forecaster will generally not be able to attain a total score this low and the greatest lower bound on their score will correspond the maximizer $\lambda^{\star} = \lambda^{\star}(K)$ of \eqref{dual_eq}.

The advantage of this dual formulation under separability is that when considering a candidate $\lambda$ for $\lambda^{\star}$,
 we can calculate each summand 
$L_{D,i}(\lambda):=\underset{0 \leq x}{\min} \left\{\lambda w_i x + \sbar_{F,i}(x)\right\}$ independently. 
To this end, let $Z_{\lambda,i}$ be the set of $x \in \mathbb{R}$ where the weighted subgradient of $-\sbar_{F_i}(x)$ contains 
$\lambda$ (c.f. \eqref{scalar_Z_deriv}):
\begin{align}
Z_{\lambda,i} := \left\{x \mid 
\lambda \in -w_i^{-1}\partial\sbar_{F_i}(x) = w_i^{-1}\kappa_i g_i'(x)(\alpha_i - [F_i(x-), F_i(x)])\right\}. \label{dual_comp_deriv}
\end{align}
Then for $\lambda>0$ we have
\begin{align}
L_{D,i}(\lambda) = \min \left\{\lambda w_i x + \sbar_{F_i}(x) \mid x \in \{0\} \cup \{Z_{\lambda,i} \cap (0,\infty)\}  \right\}
\end{align}
Note that 
\begin{itemize}
\item $Z_{\lambda,i}$ will contain any (necessarily closed) interval where $\sbar_{F_i}$ is linear with slope $-\lambda w_i$,
\item unless $\lambda=0=O_i$, we have $\sup Z_{\lambda,i} < \infty$ since in this case $\lambda w_i x + \sbar_{F,i}(x)$ is 
eventually monotonically increasing in $x$,
\item $Z_{\lambda,i}\cap [0,\infty) = \varnothing$ for $\lambda > \max_{x\geq 0}\{w_i^{-1}\kappa_i g_i'(x)(\alpha_i - F_i(x))\}$.
\end{itemize}

In the case that we will usually restrict to where $g_i'$ is non-vanishing and non-increasing, $Z_{\lambda,i}$ will consist at most of a single interval 
(usually of width 0), and in the typical "pinball loss" case of $g_i'(x) \equiv 1$,
\begin{align}
Z_{\lambda,i}= 
\begin{cases}
Q_i(\alpha_i - \kappa_i^{-1}w_i\lambda)  & \text{ if } \alpha_i - \kappa_i^{-1}w_i\lambda > 0 \\
\varnothing & \text{ otherwise}.
\end{cases}
\end{align}

Let $Z_{\lambda} = \{\mathbf{x} \mid x_i \in Z_{\lambda,i}\}$ and $\Delta_{\sbar_{F},K,\pm}(\lambda)= \mathbf{w}^{T}\mathbf{x}_{\pm}(\lambda)-K$ where
\begin{align}
x_{-,i}(\lambda) = \min Z_{\lambda,i}\cap [0,\infty) \text{  and  } x_{+,i}(\lambda) = \max Z_{\lambda,i}\cap [0,\infty)
\end{align}
and write $\Delta_{\sbar_{F},K,\pm}(\lambda)=\Delta_{\sbar_{F},K}(\lambda)$ when $\# Z_{\lambda} = 1$.
If $\Delta_{\sbar_{F},K,\pm}(\lambda) > 0$, the constraint is violated everywhere in $Z_{\lambda}\cap [0,\infty)^N$, and from \eqref{Lagrangian} we see that 
$0 > \sup(\partial L_D(\lambda))$ so that $\lambda^{\star} > \lambda$, whereas
if $\Delta_{\sbar_{F},K,\pm}(\lambda) < 0$ there is under-utilization everywhere in $Z_{\lambda}\cap [0,\infty)^N$, 
$0 < \inf(\partial L_D(\lambda))$, and $\lambda^{\star} < \lambda$. But if 
$\Delta_{\sbar_{F},K,-}(\lambda) \leq 0 \leq \Delta_{\sbar_{F},K,+}(\lambda)$, we have $\lambda^{\star} = \lambda$ and a solution $\mathbf{x}^{\star}$ to the
AP is obtained by solving $\mathbf{w}^{T}\mathbf{x}=K$ on $Z_{\lambda}\cap [0,\infty)^N$. For example, take $N=2$, $F_1 = \mathrm{Unif}[a,b]$, 
$F_2 = N(a,a/2)$ with $a>0$, $\alpha_1 = .5, \alpha_2 = .6$, $\kappa_i=g_i'=w_i=1$, and let $q=Q_{N(a,a/2)}(.1)$. Then the allocation 
$(a,q)\in Z_{.5}$ violates
a constraint of $K=a$, but the allocation with $(0,q)\in Z_{.5}$ leaves slack. Thus the solution $(a-q,q)$ lies in $Z_{.5}$.

<!-- At $\lambda=0$, we have
\begin{align}
L_{D,i}(0) = 
\begin{cases}
 \sbar_{F_i}(\max \{0, q_{\alpha_i,F_i}^{+}\}) & \text{ if }O_i >0 \\
 \underset{x\to \infty}{\lim}U_i\int_{x}^{\infty}(g_i(y) - g_i(x))f_i(y)dy = 0 & \text{ if } O_i = 0
\end{cases}
\end{align}
(assuming a reasonable bound on the growth rate of $g_i$).
 -->

As such, we can solve, at least approximately, the AP as follows.  We first determine, from representations of the $F_i$, the finite set 
\begin{align}
I_{\sbar_{F}} := \left\{\lambda \mid \# Z_{\lambda} > 1\right\}
\end{align}
and check whether for any $\lambda \in I_{\sbar_{F}}$ we can solve $\mathbf{w}^T \mathbf{x} = K$ with $\mathbf{x}\in Z_{\lambda}$. If possible,
we have found the maximizer $\lambda^{\star}$ as well as a solution $\mathbf{x}$ to the AP.

If $I_{\sbar_{F}}$ does not yield an immediate solution, we proceed to a binary search for $\lambda^{\star}$
working inward from the interval $[\lambda_L, \lambda_U]$ where
\begin{align}
\lambda_L &= \max \left\{ \lambda \in \bar{I}_{\sbar_{F}} \mid \Delta_{\sbar_{F},K,-}(\lambda_{\tau})>0 \right\} \\
\lambda_U &= \min \left\{ \lambda \in \bar{I}_{\sbar_{F}} \mid \Delta_{\sbar_{F},K,+}(\lambda_{\tau})<0 \right\} \\
\bar{I}_{\sbar_{F}} &= \{I_{\sbar_{F}}, 0, \max_{i,x}\{w_i^{-1}\kappa_i g_i'(x)(\alpha_i - F_i(x))\} \}.
\end{align}
At each step $\tau$ of the search, we take 
$\lambda_{\tau}= 1/2(\lambda_{U,\tau}+\lambda_{L,\tau})$ and
either discard the upper half of the current search interval if $\Delta_{\sbar_{F},K}(\lambda_{\tau})<0$, setting $\lambda_{U,\tau+1}= \lambda_{\tau}$,
 or discard the bottom half if not, setting $\lambda_{L,\tau+1}= \lambda_{\tau}$.
We continue until $\Delta_{\sbar_{F},K}(\lambda_{\tau})$ is sufficiently small.

If we also impose convexity on $\sbar_F(\mathbf{x})$ by requiring that all $g_i$ have non-positive second derivatives, then since the constraint 
$\mathbf{w}^T \mathbf{x} \leq K$ is affine and the domain $\{\mathbf{x} \geq 0 \}$ is convex, we have strong duality so that the optimal value 
$L_D(\lambda^{\star})$ of the dual problem is also the attained minimum of the primal AP. An optimal allocation is then given by taking
\begin{align}
q_{i,K} \in \argmin \left\{ \lambda^{\star} w_i x + \sbar_{F_i}(x) \mid x \in Z_{\lambda^{\star},i}\right\}.
\end{align}
Note that in the "easy" case of $g_i (x)=x$ for all $i$ and $K$ not too small, this becomes
\begin{align}
q_{i,K} = Q_i(\alpha_i - \kappa_i^{-1}w_i\lambda^{\star})
\end{align}
where $\lambda^{\star}$ is found as a root of
\begin{align}
\sum_{i=1}^{N}w_i Q_i(\alpha_i - \kappa_i^{-1}w_i\lambda^{\star}) = K.
\end{align}
This partial solution to the AP seems to have first appeared in [@hadleywhitin1963].

See the \hyperref[KKT_derivation]{appendix} for a derivation and discussion of this solution method via Karush-Kuhn-Tucker equations rather than duality theory.

### Marginal benefit interpretation of the dual variable

Equation \eqref{dual_comp_deriv} can be viewed as defining functions 
\begin{align}
\lambda_i(x_i): = -\frac{1}{w_i}\del{\sbar_F}{x_i}(x_i)= 
w_i^{-1}\kappa_i g_i^{\prime}(x_i)(\alpha_i - F_i(x_i)) \label{eqn:marg}
\end{align}
which can each be interpeted as a \emph{marginal expected unit benefit} of $x_i$. That is,
if $F_i(x_i) < \alpha_i$ we receive the positive "benefit" of reducing $\sbar_F$ at a rate of $\lambda_i(x_i)$ 
per additional unit of capacity allocated to the $i$'th target at "price" $w_i$. Note that with $\mathcal{S}_i$ the support of $F_i$ we have
\begin{align}
 \lambda_i(x_i) =
\begin{cases}
  \alpha_i w_i^{-1}\kappa_i g_i^{\prime}(x_i) & x_i \leq \inf(\mathcal{S}_i) \\
 -(1-\alpha_i)w_i^{-1}\kappa_i g_i^{\prime}(x_i) & x_i \geq \sup(\mathcal{S}_i),
 \end{cases}
\end{align}
as in the [uniform][1) $Y \sim U[a,b]$] and [exponential][4) $g(x)=x$ and $Y \sim \mathrm{Exp}(1/\sigma)$] examples.

We can demonstrate the KKT condition \eqref{eqn:lam_Qnz} directly in terms of the $\lambda_i$:

\begin{proposition}
\label{LM_prop}{}
For a solution $\mathbf{Q} = (Q_1,\ldots,Q_N)$ of the AP we have
\begin{align}
 \lambda_i(Q_i) = \lambda_j(Q_j) \text{ whenever } Q_i, Q_j > 0. \label{lambda_eq}
\end{align}
\end{proposition}
\begin{proof}
Assume instead we had $\lambda_i(Q_i) > \lambda_j(Q_j)$ and $Q_i, Q_j > 0$. 
Moving away from $\mathbf{Q}$ in the direction $\mathbf{d}_{ij} = w_i^{-1}\mathbf{e}_i - w_j^{-1}\mathbf{e}_j$ we remain 
in $\mathbb{R}^N_+$ while respecting the constraint ($\langle \nabla  \mathbf{w}^T \mathbf{x}, \mathbf{d}_{ij}\rangle = 0$) but reducing $\sbar_F$:
\begin{align}
    \langle \nabla \sbar_F(\mathbf{Q}), \mathbf{d}_{ij} \rangle = 
    w_i^{-1} \frac{\partial \sbar_F}{\partial x_i}(Q_i) -
    w_j^{-1}\frac{\partial \sbar_F}{\partial x_j}(Q_j)  =  \lambda_j(Q_j) - \lambda_i(Q_i)  < 0. \label{descent}
\end{align}
Therefore $\mathbf{Q}$ cannot be a constrained minimum of $\sbar_F$. 
\end{proof}
But in line with KKT complementary slack conditions, the argument does not apply when $Q_j=0$ and $\mathbf{d}_{ij}$ 
takes us immediately out of $\mathbb{R}^N_+$. This shows again how the set of indices $\{i \big| Q_i = 0\}$ arises as another variable of the problem.

<!-- 
We can, however, confine our search to two classes of solutions.
\begin{proposition}
\label{two_classes_prop}
For a solution $\mathbf{Q} = (Q_1,\ldots,Q_N)$ of the AP we have $F_i(Q_i) \leq \alpha_i$ for all $i$.  
Also, either 
\begin{itemize}
    \item[(I)]  for all $i$, $F_i(Q_i) = \alpha_i$, that is, 
        $\mathbf{Q} = \mathbf{Q}(\symbfit{\alpha}) = (Q_1(\alpha_1),\ldots,Q_N(\alpha_N))$ is a vector of quantiles of the $F_i$
    \item[] or
    \item[(II)] $\mathbf{w}^T \mathbf{Q}=K$ and for all $i$, $F_i(Q_i) < \alpha_i$ whenever $Q_i>0$.
\end{itemize}
\end{proposition}
\begin{proof}
    First note that $F_i(Q_i) > \alpha_i$ is not possible for any $i$ since this would make decrease along the coordinate $x_i$  
    a feasible direction of descent. In other words, just as for quantile scoring rules, forecasters do not benefit in increasing a forecast beyond their quantile for any target.

    If $F_i(Q_i) = \alpha_i$ for all $i$ and $\sum w_i Q_i \leq K$, then $\mathbf{Q}$ is a feasible point 
    where $\nabla \sbar_F=0$ (cf. \eqref{scalar_Z_deriv}) and is therefore a minimum.

    Now assume $F_i(Q_i) < \alpha_i$ for some $i$. 
    If $F_j(Q_j)=\alpha_j$ and $Q_j > 0$ for some $j$ then, as in \eqref{descent}, we would have a feasible direction 
    of descent $\mathbf{d}_{ij}$. This forces either $Q_j = 0$ or $F_j(Q_j)<\alpha_j$ for all $j$. Finally, if in this case 
    we had $\mathbf{w}^T \mathbf{Q}<K$, then increase along any coordinate $x_j$ with $F_j(Q_j)<\alpha_j$ would be a feasible direction of descent. If there is no such coordinate, we are in case I.
\end{proof} -->

\subsubsection{Allocation scoring rule definition}

Collect the AP parameters into $\symbfit{\beta}:= \left\{\mathbf{w}, K, \symbfit{\alpha}, \symbfit{\kappa}\right\}$. The resulting $Q_i^{\symbfit{\beta}}= Q_i(\symbfit{\beta})$ form the Bayes act for the the forecast $\mathbf{F}$ and the realized losses and costs defines the scoring rule evaluated on $\mathbf{F}$:
\begin{align}
S_{\symbfit{\beta}}(\mathbf{F},\mathbf{y}) = 
s_{\symbfit{\beta}}(\mathbf{Q}^{\symbfit{\beta}}, \mathbf{y})&=\sum O_i(g_i(Q_i^{\symbfit{\beta}}) - g_i(y_i))_+ + U_i(g_i(Q_i^{\symbfit{\beta}})-g_i(y_i))_-
\\
&= \sum((U_i+O_i)\mathbb{1}\{Q_i^{\symbfit{\beta}}>y_i\} - U_i)(g_i(Q_i^{\symbfit{\beta}})- g_i(y_i)) \\
&= \sum \kappa_i (\mathbb{1}\{Q_i^{\symbfit{\beta}}>y_i\}-\alpha_i)(g_i(Q_i^{\symbfit{\beta}})- g_i(y_i))
\end{align}

Remarks:

- Quantile rules allow you to vary cost ratios while allocation rules add capacity as a parameter.

- While quantile rules are undefined for zero overprediction costs, allocation rules (with $K<\infty$) are defined even when some $\alpha_i = 1$. This covers one of the problems originally motivating this project: find the Bayes act for allocation $\mathbf{x}$ of hospital supplies to locations $l_i, i=1,\ldots,N$ given distributions $F_i$ of need $Y_i$, a total available stock $K$ of supplies, and a loss function
\begin{align}
l(\mathbf{x}, \mathbf{y}) = U\sum_{i=1}^{N} (x_i - y_i)_-.
\end{align}

- $s_{\symbfit{\beta}}$ is consistent by definition for the functional $\mathbf{Q}_{\symbfit{\beta}}$ and $S_{\symbfit{\beta}}$ is proper since
\begin{align}
S_{\symbfit{\beta}}(\mathbf{F}, \mathbf{F}) = s_{\symbfit{\beta}}(\mathbf{q}_{\symbfit{\beta}, \mathbf{F}},\mathbf{F}) \leq 
 s_{\symbfit{\beta}}(\mathbf{q}_{\symbfit{\beta}, \mathbf{F}'},\mathbf{F}) = S_{\symbfit{\beta}}(\mathbf{F}', \mathbf{F})
\end{align}
but is not strictly proper

\subsubsection{An Oracle Adjustment}

Differences of $s_{\symbfit{\beta}}$ and $S_{\symbfit{\beta}}$ between forecasts may become difficult to compare as $K$ becomes small. To remedy this we can adjust scores by that of an "oracle" who sees into the future and allocates according to the point mass forecast $\delta_{\mathbf{y}}$. That is, the oracle solves the deterministic version of the AP with $\sbar_F(\mathbf{x})$ replaced by $s(\mathbf{x},\mathbf{y})$. Since the oracle knows there is no benefit to allocating more than $y_i$ to the $i$'th component, we can pose this problem as
\begin{align}
    \underset{\mathbf{x}\in \mathbb{R}^N, 0 \leq \mathbf{x}\leq \mathbf{y}}{\mathrm{minimize}}\,\, 
    s(\mathbf{x},\mathbf{y})\vert_{\mathbf{x}\leq \mathbf{y}} = \sum_{i=1}^{N} -U_i (g_i(x_i)-g_i(y_i))
    \text{ subject to } 
    \sum_{i=1}^{N} w_i x_i \leq K. \label{DAP}
\end{align}
which we can again solve via its dual. The Lagrangian is now
\begin{align}
L(\mathbf{x}, \lambda) = \sum_{i=1}^{N} -U_i (g_i(x_i)-g_i(y_i)) + \lambda \left(\sum_{i=1}^{N}w_i x_i - K\right)
\end{align}
giving the dual problem of maximizing
\begin{align}
L_D(\lambda) &= \min_{0 \leq \mathbf{x}\leq \mathbf{y}} L(\mathbf{x}, \lambda) \\
&= -\lambda K + \sum_{i=1}^{N} \underset{0 \leq x_i \leq y_i}{\min} \left(\lambda w_i x_i - U_i g_i(x_i) + U_ig_i(y_i)\right)
\end{align}\
subject to $\lambda \geq 0$. We have
\begin{align}
L_{D,i}(\lambda) &:= \underset{0 \leq x_i \leq y_i}{\min} \left(\lambda w_i x_i - U_i g_i(x_i) + U_ig_i(y_i)\right) \\
&= \min \left\{ \lambda w_i x - U_i g_i(x) + U_ig_i(y_i) \mid x \in Z_{o,\lambda,i} \right\}
\end{align}
where
\begin{align}
 Z_{o,\lambda,i} = \{0, y_i\} \cup \{Z_{o,\lambda,i}' \cap (0,y_i)\}
\end{align}
and $Z_{o,\lambda, i}'$ is the possibly empty set of critical points solving
\begin{align}
\lambda = w_i^{-1}U_i g_i'(x) = w_i^{-1}\kappa_i \alpha_i g_i'(x).
\end{align}
Note that $L_D(0) = L(\mathbf{y}, 0) = 0$ is the oracle's score when the outcome $\mathbf{y}$ does not activate the constraint.

We can again find a best lower bound on the oracle's score by approximating a maximizer $\lambda^{\star}$ with a binary search starting at $\lambda_L=0$ and 
$\lambda_U > \max_{i,x} \{w_i^{-1}\kappa_i \alpha_i g_i'(x)\}$. And with our standard assumption that $g_i''(x) \leq 0$ for all $i$, under which the primal objective is convex and $\max_{i,x} \{w_i^{-1}\kappa_i \alpha_i g_i'(x)\} = \max_{i} \{w_i^{-1}\kappa_i \alpha_i g_i'(0)\}$, strong duality gives the oracle's score as $L_D(\lambda^{\star})$, corresponding to an allocation
\begin{align}
q_{o,i,K} \in \argmin \left\{ \lambda^{\star} w_i x - U_i g_i(x_i) \mid x \in Z_{o,\lambda^{\star},i}\right\}.
\end{align}

In the pinball loss case of $g_i(x)=x$ for all $i$, oracle allocation reduces to a linear program with piecewise linear dual
\begin{align}
L_D(\lambda) &= -\lambda K + \sum_{i=1}^{N} \underset{0 \leq x_i \leq y_i}{\min} \left(\lambda w_i x_i - U_i x_i + U_i y_i\right) \\
&=  -\lambda K + \sum_{i=1}^{N} \min \left\{ U_i y_i, \lambda w_i y_i\right\} \\
&= -\lambda K + \sum_{i=1}^{N} \left(U_i\mathbb{1}\{\lambda \geq U_i / w_i \} + \lambda w_i \mathbb{1}\{\lambda < U_i / w_i \}\right)y_i.
\end{align}\
The maximizer $\lambda^{\star}$ of $L_D(\lambda)$ for $\lambda \geq 0$ can now be found by evaluating at the $N+1$ points $\{\lambda_0 = 0, \lambda_i = U_i/w_i\}$. Again, $\lambda^{\star}=0$ corresponds to an inactive constraint under which the oracle can allocate $q_{o,i,K}=y_i$. 
When $\lambda^\star = U_{i^\star}/w_{i^\star}$, we get
\begin{align}
q_{o,i,K} &= \argmin_{0 \leq x \leq y_i} \left\{\frac{U_{i^\star}}{w_{i^\star}} w_i x - U_i x + U_i y_i\right\} \\
&= \argmin_{0 \leq x \leq y_i} \left\{\left(\frac{U_{i^\star}}{w_{i^\star}} w_i  - U_i\right) x\right\} \\
&= \begin{cases}
0 & \text{ if } \frac{U_{i^\star}}{w_{i^\star}} > \frac{U_{i}}{w_{i}} \\
y_i & \text{ if } \frac{U_{i^\star}}{w_{i^\star}} < \frac{U_{i}}{w_{i}}.
\end{cases}
\end{align}
If there is a single $i^{\star}$ for which $\lambda^\star = U_{i}/w_{i}$, the oracle's allocation in this coordinate is determined by the (now active) constraint as
\begin{align}
q_{o,i^\star,K} = w_{i^\star}^{-1}\left(K - \sum_{i \mid \frac{U_{i^\star}}{w_{i^\star}} < \frac{U_{i}}{w_{i}}} w_i y_i \right).
\end{align}
If $I_K = \{i \mid \lambda^\star = U_{i}/w_{i}\}$ includes $n_K > 1$ indices, then the solutions $q_{o,i,K} \in [0,y_i]$ for $i \in I_K$
are not unique because the oracle's score will be constant on
\begin{align}
\left\{ \mathbf{x} \mid \sum_{i \in I_K}w_i x_i = K -  \sum_{\frac{U_{i^\star}}{w_{i^\star}} < \frac{U_{i}}{w_{i}}}w_i y_i,\,\, x_i = q_{o,i,K} \text{ for } i \notin I_K \right\}.
\end{align}
(Note that restricted to the coordinates in $I_K$, the score level sets are parallel to the affine constraint hyperplane.)
If a specific solution is desired, additional criteria could be invoked.  A fairness criterion, for example, might require 
the remaining resources $K - \sum_{\frac{U_{i^\star}}{w_{i^\star}} < \frac{U_{i}}{w_{i}}}w_i y_i$ to be divided equally among the $n_K$ recipients.

## Special Case I: $F_i$ from same location-scale family and $w_i = 1$, $O_i = O, U_i = 1$.

Also assume the constraint is active, i.e., 
\begin{align}
K < \sum F_i^{-1}(\alpha)= \sum \mu_i + \sigma_i\Phi^{-1}(\alpha)
\end{align}
where $\Phi$ is the standard CDF for the family and $\alpha = (1+O)^{-1}$. Then setting
\begin{align}
\sum F_i^{-1}(\alpha - \lambda) = \sum \mu_i + \sigma_i\Phi^{-1}(\alpha - \lambda) = K
\end{align}
we get
\begin{align}
\alpha - \lambda = 
\Phi\left(\frac{K - \sum \mu_j}{\sum \sigma_j} \right):= F(K)
\end{align}

and the Bayes act when the constraint is set to $K$ is 
\begin{align} 
 \nonumber Q_i(K) = F_i^{-1}(F(K)) &= \mu_i +\sigma_i\Phi^{-1}(F(K)) \\ 
 \nonumber &= \mu_i + \sigma_i\left(\frac{K - \sum \mu_j}{\sum \sigma_j} \right) \\
 &= \mu_i + \tilde{\sigma}_i(K - \sum \mu_j)
\end{align}

where $\tilde{\sigma_i}$ is the proportion of $F$'s scale (e.g. SD) "due" to $F_i$.

Remarks:

- $Q_i(K)$ here does not depend on $O$ unlike when the constraint is inactive or the $O_i$ differ.

- An "excess or shortage in mean" $K - \sum \mu_j$ is divided among the components in proportion to their scale factors as adjustmments up or down from their locations $\mu_i$.  This suggests an interpretation of under-dispersion of a forecast in one component as leading to that component not receiving as much additional recources as it should when there is an excess or other components not recieving as much of scarce resources as they should when there is a shortage.

The score for the forecast $\{F_i\}$ is

\begin{align}
s_K(\mathbf{Q}, \mathbf{y})&=
 \sum(1-(1+O)\mathbb{1}
\left\{\frac{K - \sum \mu_i}{\sum \sigma_i}> 
\frac{y_i - \mu_i}{\sigma_i}\right\})
(g_i(y_i) -g_i(Q_i(K))) 
\end{align}

In the motivating case that $O=0$, this becomes
\begin{align}
s_K(\mathbf{Q}, \mathbf{y})=
 \sum \mathbb{1}
\left\{\frac{K - \sum \mu_i}{\sum \sigma_i}\leq 
\frac{y_i - \mu_i}{\sigma_i}\right\}
(g_i(y_i) -g_i(Q_i(K))) 
\end{align}

That is, a $Q_i$ is only penalized when the standardized observed excess demand in component $i$ exceeds the standardized excess in resources or the standardized observed shortfall in demand is not as large as the standardized shortage of resources.

## Special Case II: $F_i$ uniform on $[a_i,b_i]$

\noindent\fbox{
\parbox{\textwidth}{
\textbf{Problem:}
This does not seem to work. Take $F_1=F_2=F_{[0,1]}, \alpha_1=\alpha_2=.5, w_1=1, w_2=2$. Then
$$
    \lambda = \frac{3/2 - K}{1/2 + 4/2} = \frac{3-2K}{5} > \frac{1}{2}
$$
for $K<1/4$, so that $Q_2(\alpha_2(1-w_2 \lambda))$ is undefined. The minimizer here is $Q_1=K, Q_2=0$, which does satisfy the KKT equations \eqref{eqn:lam_Qnz} and \eqref{eqn:mu_Qz}. $Q_1=0, Q_2=K/2$ does not.
}
}

Following [@abdel2004exact], we can directly evaluate \eqref{lambda_eq} to solve for $\lambda$ in the case of the uniform location-scale family ($\mu_i = a_i, \sigma = b_i - a_i$):
\begin{align}
K &=\sum w_i F_{[a_i,b_i]}^{-1}\left(\alpha_i(1 - w_i \lambda)\right) \\
&=\sum w_i Q_{[a_i,b_i]}\left(\alpha_i(1 - w_i \lambda)\right) \\
&= \sum w_i\left[ a_i + (b_i-a_i)(\alpha_i -  w_i \alpha_i\lambda)\right] \\
&= \sum w_i Q_{[a_i,b_i]}(\alpha_i) - (b_i-a_i)w_i^2 \alpha_i \lambda \label{lambda_eval}
\end{align}

Writing $D = \sum w_i Q_{[a_i,b_i]}(\alpha_i) -K$ for the ``deficit" of the unconstrained solution, we have

\begin{align}
\lambda &= \frac{D}{\sum (b_j-a_j)w_j^2 \alpha_j} \\
\end{align}

and from ??,
\begin{align}
Q_i &= F_{[a_i,b_i]}^{-1}\left(\alpha_i\left(1- \frac{w_iD}{\sum_j (b_j-a_j)w_j^2 \alpha_j}\right)\right) \\
&= Q_{[a_i,b_i]}(\alpha_i) - \frac{(b_i-a_i)w_i \alpha_i D}{\sum (b_j-a_j)w_j^2 \alpha_j}.
\end{align}

That is, each constrained $Q_i$ is obtained by reducing the $\alpha_i$ quantile by an amount of the deficit proportional to the product of spread, resource weighting, and cost/loss ratio (what should this be called?) of the corresponding prediction target.

## General solution method

\noindent\fbox{
\parbox{\textwidth}{
Is there anything we can salvage here?
}
}

Again following [@abdel2004exact], we can use this solution for uniform $F_i$ as the basis of an iterative procedure for approximating the $Q_i$ under a constraint for general predictive distributions. Assuming that the unconstrained solution 
$\{Q_i(\alpha_i)\}$ is not feasible and that we can evaluate $F_i$, $F_i^{-1}$ and $f_i$ at least approximately at arbitrary $x_i$, we 
first linearize the equations \eqref{lambda_eq} at 
$\{(\alpha_i,Q_i(\alpha_i))\}$ by replacing the $F_i^{-1}$ with the tangential approximations
\begin{align}
\left(F_{i}^{(1)}\right)^{-1}(p) := Q_i(\alpha_i) + \frac{p - \alpha_i}{f_i(Q_i(\alpha_i))},
\end{align}
that is, the quantile functions of the uniform cdf's
\begin{align}
F_{i}^{(1)} := \min(0,\max(1,\alpha_i + f_i(Q_i(\alpha_i))(x_i - Q_i(\alpha_i)))).
\end{align}

As before, these equations 
\begin{align}
   K &= \left(F_{i}^{(1)}\right)^{-1}(\alpha_i(1-w_i \lambda)) \\
     &= \sum w_i Q_i(\alpha_i) - \frac{w_i^2 \alpha_i \lambda}{f_i(Q_i(\alpha_i))}  
\end{align}
are linear in $\lambda$ and have the solution
\begin{align}
    \lambda^{(1)} = \frac{D}{\sum \frac{w_j^2 \alpha_j}{f_j(Q_j(\alpha_j))}}.
\end{align}
This provides the first iterate
\begin{align}
    Q_i^{(1)} = F_i^{-1}(\alpha_i(1-w_i \lambda^{(1)}))
\end{align}
and a new deficit (or negative surplus?) 
\begin{align}
    D^{(1)} = \sum w_i Q_i^{(1)} - K.
\end{align}
Now we can repeat the process with new tengential approximations
\begin{align}
\left(F_{i}^{(2)}\right)^{-1}(p) := Q_i^{(1)} + \frac{p - \alpha_i^{(1)}}{f_i\left(Q_i^{(1)}\right)}
\end{align}
where $\alpha_i^{(1)}:=\alpha_i(1-w_i \lambda^{(1)})$ to get the next $\lambda$ iterate
\begin{align}
    \lambda^{(2)} = \frac{D^{(1)}}{\sum \frac{w_j^2 \alpha_j^{(1)}}{f_j\left(Q_j^{(1)}\right)}},
\end{align}
continuing with $\alpha_i^{(\tau)}:=\alpha_i^{(\tau -1)}(1-w_i \lambda^{(\tau)})$ until the relative error at the $\tau$'th iteration
\begin{align}
    \varepsilon^{(\tau)} = D^{(\tau)}/K
\end{align}
is sufficiently small.

<!-- ## Requiring non-negative allocations

Nothing in this procedure imposes the constraint that the $Q_i^{(\tau)}$ be non-negative which they will often be despite the support of the $F_i$ being non-negative.  We can, however, check before beginning the procedure whether this is the case using the lower bound on the budget required by non-negativity of all forecasts:
\begin{align}
    \sum w_i F_i^{-1}(\alpha_i(1-w_i \lambda^{-}))
\end{align}
where
\begin{align}
    \lambda^{-} := \min(\lambda_i(0)),
\end{align}
the minumum of the marginal expected benifits \eqref{marg} at $x_i=0$...  Following [@abdel2005analysis]
 -->
## References

<div id="refs"></div>`

\newcounter{appendix}
\refstepcounter{appendix}
\setcounter{equation}{0}

\renewcommand{\theequation}{\Alph{appendix}.\arabic{equation}}
\renewcommand{\theHequation}{\Alph{appendix}.\arabic{equation}}

\section{Appendix A. Derivation Details}

\subsubsection{Differentiation of $\sbar_F(x)$ in terms of a density $f$: \label{scalar_Z_deriv_detail}}
\begin{align}
\frac{d}{dx}\sbar_F(x) &= \frac{d}{dx} \left[ O\int_{-\infty}^x(g(x) - g(y))f(y)dy + 
U\int_x^{\infty}(g(y) - g(x))f(y)dy \right]  \\
&= O(g(x) - g(x))f(x) + O\int_{-\infty}^x\frac{d}{dx} \left[(g(x) - g(y))f(y)\right]dy \\
&- U(g(x) - g(x))f(x) + U\int_x^{\infty}\frac{d}{dx} \left[(g(y) - g(x))f(y) \right]dy \\
&= O\int_{-\infty}^x g^{\prime}(x) f(y)dy - U\int_x^{\infty} g^{\prime}(x)f(y) dy \\
&= g^{\prime}(x)\left(OF(x) - U(1-F(x)) \right) \\
&= g^{\prime}(x)(O+U)\left(F(x) - \frac{U}{O+U}\right) \\
&= \kappa g^{\prime}(x)\left(F(x) - \alpha\right) 
\end{align}

\subsubsection{Comments on the dual problem formulation: \label{dual_setup}}

In the main text, we noted that after defining
\begin{align}
L(\mathbf{x}, \lambda) = \sum_{i=1}^{N} \sbar_{F_i}(x_i) + \lambda \left(\sum_{i=1}^{N}w_i x_i - K\right),
\end{align}
we have the dual problem of maximizing over $\lambda \geq 0$ the objective function
\begin{align}
L_D(\lambda) &= \min_{0 \leq \mathbf{x}} L(\mathbf{x}, \lambda) = 
-\lambda K + \sum_{i=1}^{N} \underset{0 \leq x_i}{\min} \left\{\lambda w_i x_i + \sbar_{F,i}(x_i)\right\}.
\end{align}
The solution to this dual problem gives lower bounds on a solution to the primal AP; we verify this in more detail here.

First, note that for any $\mathbf{x}$ that is a feasible value for the AP, the constraint $\mathbf{w}^T\mathbf{x} \leq K$ holds, so that $\mathbf{w}^T\mathbf{x} - K \leq 0$ and for any $\lambda \geq 0$, $\lambda \left(\sum_{i=1}^{N}w_i x_i - K\right) \leq 0$. Therefore, for any feasible $\mathbf{x}$ and $\lambda \geq 0$,

\begin{align}
L(\mathbf{x}, \lambda) &= \sum_{i=1}^{N} \sbar_{F_i}(x_i) + \lambda \left(\sum_{i=1}^{N}w_i x_i - K\right) \leq \sbar_{\mathbf{F}}(\mathbf{x}).
\end{align}

Thus, for a fixed $\lambda \geq 0$,

\begin{align}
\min_{0 \leq \mathbf{x}, \mathbf{w}^T\mathbf{x} \leq K} L(\mathbf{x}, \lambda) \leq \min_{0 \leq \mathbf{x}, \mathbf{w}^T\mathbf{x} \leq K} \sbar_{\mathbf{F}}(\mathbf{x})
\end{align}

Now since $\{\mathbf{x}: 0 \leq \mathbf{x}, \mathbf{w}^T\mathbf{x} \leq K\} \subseteq \{\mathbf{x}: 0 \leq \mathbf{x}\}$ and using the fact that if $B \subseteq A$ then $\min_{x \in A} f(x) \leq \min_{x \in B} f(x)$, we have

\begin{align}
L_D(\lambda) &= \min_{0 \leq \mathbf{x}} L(\mathbf{x}, \lambda) \\
&\leq \min_{0 \leq \mathbf{x}, \mathbf{w}^T\mathbf{x} \leq K} L(\mathbf{x}, \lambda) \\
&\leq \min_{0 \leq \mathbf{x}, \mathbf{w}^T\mathbf{x} \leq K} \sbar_{\mathbf{F}}(\mathbf{x}), \label{eqn:appendix_weak_duality}
\end{align}
where the right hand side of Equation \eqref{eqn:appendix_weak_duality} is the optimal value of the AP.
Since this holds for all $\lambda$ and in particular for the maximizer $\lambda^\star$, we have $L_D(\lambda^\star) \leq \sbar_{\mathbf{F}}(\mathbf{x}^\star)$, where $\mathbf{x}^\star$ denotes the AP solution. This inequality goes by the name of weak duality.

\subsubsection{Solution of the AP via KKT equations: \label{KKT_derivation}}

To describe forecasts solving the AP \eqref{AP} without the direct use of duality theory, we begin by stating the necessary Karush-Kuhn–Tucker (KKT) conditions for a point $\mathbf{Q} \in C_K$ to be a solution of the AP.
Since the feasible set $C_K$ is a polyhedron on which $\sbar_F$ is smooth, these are: there exists
a vector of multipliers $\symbfit{\mu} = (\lambda, \mu_1, \ldots,\mu_N) \in \mathbb{R}^{1+N}$ such that 
\begin{align}
\lambda, \mu_1, \ldots,\mu_N &\geq 0 \label{mult_pos} \\
-\nabla \sbar_F (\mathbf{Q}) &= D^T \symbfit{\mu} \label{lagrange_eq}\\
\lambda (\mathbf{w}^T \mathbf{Q} - K) &= 0 \label{comp_slack_lam}\\
\mu_i Q_i &= 0,\quad i = 1,\ldots,N \label{comp_slack_mu}
\end{align}
(see for example, [@Nocedal2006], Theorem 12.1). These equations express the fact that at a minimizer, either the gradient
 $\nabla \sbar_F$ must vanish (so that $\symbfit{\mu}=\mathbf{0}$), or the direction of greatest descent $-\nabla \sbar_F$ must 
 lie in the normal (outwardly pointing) cone of $C_K$, which is the non-negative span of some subset (generically a single one) of the columns of $D^T$. This subset corresponds to which constraints (i.e. "sides" of $C_K$) are active, allowing the corresponding multipliers in the "complementary slackness" equations \eqref{comp_slack_lam} and \eqref{comp_slack_mu} to be positive.

From \eqref{scalar_Z_deriv}, using $\mathbf{e}_i$ to denote the $i$th coordinate vector, the "Lagrange multiplier" equation \eqref{lagrange_eq} becomes
\begin{align}
-\sum_{i=1}^{N}\kappa_i g^{\prime}_i(Q_i) (F_i(Q_i) - \alpha_i)\mathbf{e}_i = 
[\mathbf{w}\big| -\mathrm{Id}_N]\symbfit{\mu} =
\sum_{i=1}^{N} (\lambda w_i - \mu_i)\mathbf{e}_i \label{eqn:lagrange_coords}
\end{align}
which decomposes to 
\begin{align}
\lambda = w_i^{-1}(\kappa_i g_i^{\prime}(Q_i)(\alpha_i - F_i(Q_i)) + \mu_i), \quad i = 1,\ldots,N. 
\label{eqn:lagrange_decomp}
\end{align}
Using \eqref{comp_slack_mu}, we can write this as
\begin{numcases} {\lambda = }
w_i^{-1} \kappa_i g_i^{\prime}(Q_i)(\alpha_i - F_i(Q_i)) & for  $Q_i > 0$ \label{eqn:lam_Qnz}\\
w_i^{-1} (\kappa_i g_i^{\prime}(0)(\alpha_i-F_i(0)) + \mu_i) & for $Q_i = 0$. \label{eqn:mu_Qz}
\end{numcases}

This leaves us with two alternatives. 
\begin{itemize}
\item[(I)]$\lambda = 0$, i.e., the constraint $K$ on the total resources allocated is not active. Then \eqref{eqn:lam_Qnz} forces
\begin{align}
F_i(Q_i) = \alpha_i \text{  for } Q_i > 0,
\end{align}
and \eqref{eqn:mu_Qz} forces
\begin{align}
F_i(0) = \alpha_i + \mu_i/\kappa_i g_i^{\prime}(0) \geq \alpha_i \text{  for } Q_i = 0.
\end{align}
That is, $\mathbf{Q}$ is a vector with entries $Q_i = \max(Q_i(\alpha_i),0)$.
\item[(II)] $\lambda > 0$, so that the constraint $K$ is active. In this case \eqref{eqn:lam_Qnz} forces
\begin{align}
F_i(Q_i) < \alpha_i \text{  for } Q_i > 0,
\end{align}
and for any $i$ with $Q_i=0$, \eqref{eqn:mu_Qz} along with $\mu_i \geq 0$ gives a lower bound on $\lambda$,
\begin{align}
\lambda \geq w_i^{-1}\kappa_i g_i^{\prime}(0)(\alpha_i-F_i(0)).
\end{align}
And from \eqref{comp_slack_lam}, the constraint must be active, i.e., 
\begin{align}
\mathbf{w}^{T}\mathbf{Q} = K. \label{active_constraint}
\end{align}
\end{itemize}

Now since $\sbar_F$ is a continuous function on the compact set $C_K$, some solution $\mathbf{Q}\in C_K$ of the AP exists.
Uniqueness, however, requires some additional conditions on the $g_i$ over the simplex
$\{\mathbf{x} \in C_K \mid \mathbf{w}^T\mathbf{x}=K\}$. One such set of conditions -- that for convenience we'll require over all of $C_K$
-- is that the $g_i$ have non-positive second derivatives, 
\begin{align}
g_i^{\prime \prime}(x_i) \leq 0 \text{  whenever  } F_i(x_i) < \alpha_i \label{g_pos_cond}
\end{align}
so that from \eqref{scalar_Z_deriv2}, the second partial derivatives $\sbar_{F,x_i x_i}$ are all non-negative on $C_K$. Because $\sbar_{F,x_i x_j}=0, i \neq j$, the Hessian $\nabla^2 \sbar_F$ is then positive semi-definite on $C_K$ so that $\sbar_F$ in convex on $C_K$ and any local minimizer for the AP is actually global. (Note that we might have to consider minimizing sets if any of the densities $f_i$ vanish at a minimizer, but we'll set aside this possibility for now.) And along with \eqref{eqn:lagrange_decomp}, requiring \eqref{g_pos_cond} gives 
the upper bound
\begin{align}
\lambda \leq \min_i w_i^{-1}(\kappa_i g_i^{\prime}(0)(\alpha_i-F_i(0)) + \mu_i). \label{eqn:lam_upper_bound}
\end{align}

Finding this unique $\mathbf{Q}$ can sometimes be straightforward. Suppose, for example, there are no point masses,
$F_i:[0,\varepsilon_i] \rightarrow [0,\delta_i]$ are bijective for some $\varepsilon_i, \delta_i>0$ and all $i$, and $g_i^{\prime}$ are all identically 1, and that the constraint is active with $\mathbf{w}^T \mathbf{Q}_{\mathbf{F}}(\symbfit{\alpha}) = K_1 > K$.
Then we can write the equations \eqref{eqn:lam_Qnz} as
\begin{align}
Q_i(\lambda) = F_i^{-1}(\alpha_i- \lambda w_i/ \kappa_i ) = F_i^{-1}(\alpha_i(1 - \lambda w_i/ U_i ))
\end{align}
which are all defined for $\lambda$ in the interval $I=[0,\lambda_0 = \min (U_i/w_i )]$. By our assumptions, the left hand side of 
the constraint equation \eqref{active_constraint}
\begin{align}
\sum_{i=1}^{N} w_i Q_i(\lambda) = K.
\end{align}
is a continuous function 
mapping $I$ to $[K_1, \sum_{i=1}^{N} w_i Q_i(\lambda_0)]$.  If $K$ is sufficently close to $K_1$, 
we will have a root $0< \lambda^{\star} < \lambda_0$ which can be easily found
using a formal or numerical search in $I$.  And since our assumptions imply that $Q_i(\lambda^{\star})>0$, the KKT conditions \eqref{eqn:lam_Qnz} and \eqref{eqn:mu_Qz} are met.

But if the constraint is too tight or the lower bounds of the supports of the $F_i$ differ, we will potentially need to repeat the root search for each possible set $\{i \mid Q_i = 0\}$, facing intractability for large $N$. Our convexity assumption 
\eqref{g_pos_cond}, however, besides ensuring existence and uniqueness, makes available a 
binary search method of [@zhang2009binary] which is only of polynomial complexity in $N$. In particular \eqref{g_pos_cond} 
guarantees that the functions \eqref{eqn:marg} are decreasing in $x_i$ (cf. \eqref{scalar_Z_deriv2}) from
\begin{align}
 \lambda_i(0) = w_i^{-1}\kappa_i g_i^{\prime}(0)(\alpha_i -F_i(0)) \quad \text{to} \quad \lambda_i(Q_i(\alpha_i)) = 0.
\end{align}
They can therefore can be inverted to give decreasing functions $x_i(\lambda_i)$ with 
\begin{align}
x_i(0) = Q_i(\alpha_i)\quad \text{and} \quad x_i( w_i^{-1}\kappa_i g_i^{\prime}(0)(\alpha_i -F_i(0))) = 0.
\end{align}

\newcommand\comwidth{.4}
\begin{algorithm}
\begin{algorithmic}
\caption*{ZXH binary search algorithm}
\State{\textbf{Input:} $F_i, Q_i, \kappa_i, \alpha_i, w_i, g_i, i=1,\ldots,N$,} 
\State{$K, \varepsilon_K, \varepsilon_\lambda, \mathrm{root}(f,I)$}
\Comment{\parbox[t]{\comwidth\linewidth}{$\mathrm{root}$ a function returning a root of function $f$ on interval $I$}}
\For{$i=1$ to $N$}
    \State $x_i \gets \max(0,\min(Q_i(\alpha_i), w_i^{-1}2K))$ 
\Comment{\parbox[t]{\comwidth\linewidth}{ensure constraint is violated if any $\alpha_i=1$}}
\EndFor 
\If{$\mathbf{w}^T \mathbf{x} \leq K$} 
\Return $\mathbf{Q}=\mathbf{x}$ 
\Comment{\parbox[t]{\comwidth\linewidth}{return quantiles if they satisfy constraint}}
\EndIf 
\State $\lambda_1 \gets \lambda_L \gets 0$,
$\lambda_U \gets \max_i{\lambda_i(0)}$, $\tau \gets 2$ 
\If{$\lambda_U \leq 0$} \State \Return $\mathbf{Q}=\mathbf{0}$
\Comment{\parbox[t]{\comwidth\linewidth}{handle case where all expected component scores increasing at 0}}
\EndIf 
\While{$\lvert(\mathbf{w}^T \mathbf{x} - K)/ K \lvert > \varepsilon_K$ or $\lambda_U - \lambda_L > \varepsilon_\lambda$}
    \State $\lambda_\tau \gets (\lambda_U + \lambda_L)/2$
    \For{$i=1$ to $N$}
        \If{$\lambda_\tau < \lambda_i(0)$}
        \Comment{\parbox[t]{\comwidth\linewidth}{that is, if, from \eqref{eqn:mu_Qz} and \eqref{g_pos_cond}, $\lambda_\tau$ cannot be a multiplier for $x_i=0$}}
            \If{$\lambda_\tau < \lambda_{\tau-1}$}
            \State{$I_{i,\tau} \gets [x_i, Q_i(\alpha_i)]$} 
            \Else \State{$I_{i,\tau} \gets [0, x_i]$}
            \EndIf
            \State{$x_i \gets \mathrm{root}(\lambda_i(x)-\lambda_\tau, I_{i,\tau})$}
            \Comment{\parbox[t]{\comwidth\linewidth}{well-defined since $\lambda_i$, defined by \eqref{eqn:marg}, is decreasing}}
        \Else \State{$x_i \gets 0$} 
        \Comment{\parbox[t]{\comwidth\linewidth}{prevent decrease past $0$ if $\lambda_\tau > \lambda_{\tau-1}$ 
        and maintain at $0$ if $\lambda_i(0) < \lambda_\tau < \lambda_{\tau-1}$}}
        \EndIf
    \EndFor
        \If{$\mathbf{w}^T \mathbf{x} < K$}
        \State{$\lambda_U \gets \lambda_\tau$}
        \Comment{\parbox[t]{\comwidth\linewidth}{unused capacity so $\lambda_{\tau+1}<\lambda_\tau$ and any $x_i$ could be increased but none will decrease}}
        \Else \State{$\lambda_L \gets \lambda_\tau$}
        \Comment{\parbox[t]{\comwidth\linewidth}{capacity exceeded so $\lambda_{\tau+1}>\lambda_\tau$ and no $x_i$ will increase so any set to $0$ will remain so}}
        \EndIf
    \State{$\tau \gets \tau+1$}
\EndWhile
\State \Return $\mathbf{Q}=\mathbf{x}$ 
\end{algorithmic}
\end{algorithm}

### 2023-03-31 alloscore_manuscript/allocation.rmd
some reorg...

@@
 
+\textbf{1st attempt by APG to rephrase last paragraph}
+
+A key observation here is that the Bayes act results from using a \textbf{shared} probability level, $1 - \lambda^{\star}/L$, to determine allocations as the corresponding quantiles of the location-specific forecast distributions $F_i$ which sum to the constraint $K$.
+
 \paragraph{Step 3: Define the scoring rule.} We can now define a proper scoring rule for the probabilistic forecast $F$ as

### 2023-03-31 alloscore_manuscript/allocation.rmd
attempt to sum up central mathematical ideas

--- 2023-03-30 alloscore_manuscript/allocation.rmd
@@
 
+\textbf{1st attempt by APG to rephrase last paragraph}
+
+The central mathematical ideas in this construction are that 
+\begin{itemize}
+\item optimization under a \emph{single} constraint requires the rates of change of (a probabilistic forecast's) expected benefit with respect to our decision variables, $x_i$, to be the same for all locations
+\item these rates of change can be identified with probabilities
+\item and therefore the Bayes act results from using a \emph{shared} probability level, $1 - \lambda^{\star}/L$, to determine allocations as the corresponding quantiles of the location-specific forecast distributions $F_i$ which satisfy the constraint.
+\end{itemize}
+(See the supplement for detailed discussion and derivations of these points.)
+
 \paragraph{Step 3: Define the scoring rule.} We can now define a proper scoring rule for the probabilistic forecast $F$ as

### 2023-04-14 notes/crps_connection_exp.Rmd
working through connections between allocation score and CRPS with exp forecasts

---
title: "Connection between allocation scoring rules an CRPS: example with exponential forecasts"
date: "`r Sys.Date()`"
output: 
    pdf_document:
        latex_engine: xelatex
urlcolor: blue
bibliography: allocation.bib 
header-includes:
   - \usepackage{amsmath}
   - \usepackage{algorithm}
   - \usepackage[noend]{algpseudocode}
   - \usepackage{amsthm}
   - \usepackage{bm}
   - \usepackage{cases}
   - \usepackage{caption}
   - \usepackage{unicode-math}
   - \usepackage{hyperref}
   - \DeclareMathOperator*{\argmin}{argmin}
---

\newcommand{\sbar}{\overline{s}}

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
```

Consider a simplified setting where $O_i = 0$ and $U_i = U$ are shared for all $i$, and $g_i(x_i) = x_i$. This is the setting we have laid out in the initial alloscore manuscript. We consider three settings:

 1. The decision maker is fixed with a fixed, known, constraint on the total allocation, $K$.
 2. The decision maker has some uncertainty about the total constraint $K$.
 3. We would like to understand the relationship between the allocation score and CRPS by exhibiting the distribution on $K$ that would lead to an "equally-weighted" CRPS.

Throughout, we will suppose there are $n = 2$ locations, and a forecaster produces the forecasts $Y_1 \sim Exp(1/\sigma_1)$ with $\sigma_1 = 1$ and $Y_2 \sim Exp(1 / \sigma_2)$ with $\sigma_2 = 5$. The quantile functions corresponding to these forecasts are given by $F_i^{-1}(\tau) = -\sigma_i \log(1 - \tau)$, where $\tau$ is a probability level which we take to be in $(0, 1)$.

If I've correctly translated from the notes document, the expected loss function is
\begin{equation*}
\bar{s}_F(\mathbf{x}) = \sum_{i=1}^2 \mathbb{1}\{0 \leq x_i\}\left[ \sigma e^{-x_i/\sigma} - \sigma \right]
\end{equation*}

I'm going with this below because I think it's close, but I don't actually think it's correct because it leads to negative loss values.

\section{Fixed $K$}

At a fixed constraint $K$, in our simplified setting the solution of the allocation problem is given by the quantiles $(F_1^{-1}(\tau), F_2^{-1}(\tau))$ at a probability level $\tau$ (connecting to notation elsewhere, $\tau = 1 - \lambda^\star$) such that
\begin{align*}
K &= F_1^{-1}(\tau) + F_2^{-1}(\tau) \\
  &= -\log(1 - \tau) (\sigma_1 + \sigma_2) \\
  &= h(\tau)
\end{align*}
We can rearrange to obtain
\begin{equation}
\tau = h^{-1}(K) = 1 - \exp[-K/(\sigma_1 + \sigma_2)]. \label{eqn:tau_fn_K}
\end{equation}

We can visualize this in terms of the expected loss function (shown in shades of blue and yellow) as follows:

```{r, cache = TRUE}
sigma_1 <- 1
sigma_2 <- 5

K <- 5

tau <- 1 - exp(-K/(sigma_1 + sigma_2))

x_1 <- qexp(tau, rate = 1 / sigma_1)
x_2 <- qexp(tau, rate = 1 / sigma_2)

n_grid <- 1001
grid_l <- 0.01
grid_u <- 6.0
y_grid <- tidyr::expand_grid(
    y_1 = seq(from = grid_l, to = grid_u, length.out = n_grid),
    y_2 = seq(from = grid_l, to = grid_u, length.out = n_grid)
)

joint_dist <- y_grid %>%
    dplyr::mutate(
        s_bar = sigma_1 * exp(-y_1 / sigma_1) - sigma_1 +
                sigma_2 * exp(-y_2 / sigma_2) - sigma_2
    )

ggplot(data = joint_dist) +
    geom_raster(aes(x = y_1, y = y_2, fill = s_bar)) +
    geom_contour(mapping = aes(x = y_1, y = y_2, z = s_bar)) +
    geom_abline(intercept = K, slope = -1) +
    geom_point(x = x_1, y = x_2, color = "orange") +
    scale_fill_viridis_c() +
    xlim(0, 6) +
    ylim(0, 6) +
    theme_bw()
```

\section{Obtaining the CRPS weighting associated with a distribution on $K$}

Suppose that the decision maker has some uncertainty about the level of the constraint $K$, expressed by the distribution $F_K$.  For concreteness, here we take this distribution to be $Gamma(500, 0.01)$ using a shape and scale parameterization. Here's a picture of this distribution, which is concentrated near $K = 5$:

```{r}
curve(dgamma(x, shape = 500, scale = 0.01), from = 0, to = 10, n = 1001)
```

As discussed above, given fixed forecasts $F_1$ and $F_2$, for each value of the constraint $K$, a quantile probability level $\tau = h^{-1}(K) = 1 - \exp[-K/(\sigma_1 + \sigma_2)]$ is determined. We can therefore use a change of variables to obtain a density for $\tau$ from the density for $K$ as follows:

\begin{align*}
f_T(\tau) &= f_{K}(h(\tau)) \left\vert \frac{d}{d\tau} h(\tau) \right\vert \\
&= f_{K}(-\log(1 - \tau) (\sigma_1 + \sigma_2)) \frac{(\sigma_1 + \sigma_2)}{1 - \tau} \\
\end{align*}

Here's a plot of this induced density on $\tau$:

```{r}
tau <- seq(from = 0.001, to = 0.999, length.out = 10000)

d_tau <- function(tau) {
    dgamma(-log(1 - tau) * (sigma_1 + sigma_2), shape = 500, scale = 0.01) *
                ((sigma_1 + sigma_2) / (1 - tau))
}

plot(tau, d_tau(tau), "l")

# note that this is a density
sum(d_tau(tau) * diff(tau)[1])
```

We can think of the allocation score determined by $F_K$ as corresponding to a weighted CRPS with the weighting expressed by the above density on quantile levels. However, note that this interpretation is specific to this forecast. A different forecast would translate to a different weighting on quantile levels.

\section{Reproducing equally weighted CRPS}

Going in the other direction, given forecasts $F_1$ and $F_2$, we can determine the distribution on values of the constraint $K$ that corresponds to an equal weighting of all quantile levels, as is done in CRPS.  Again, this implied distribution on $K$ depends on the forecasts and will be different for different forecasters.

In this setting, we start with $\tau \sim Unif(0, 1)$, and note that given a value of $\tau$ we can calculate $K = h(\tau) = -\log(1 - \tau) (\sigma_1 + \sigma_2)$. This induces a distribution on $K$ with density
\begin{align*}
f_K(K) &= f_{T}(h^{-1}(K)) \left\vert \frac{d}{dK} h^{-1}(K) \right\vert \\
&= 1 \cdot (\sigma_1 + \sigma_2)^{-1} \exp[-K/(\sigma_1 + \sigma_2)]
\end{align*}
This is the density of an $Exp(\sigma_1 + \sigma_2)$ random variable.

Thus, with our forecasts, CRPS corresponds to the weighting on resource constraints illustrated in the following figure:

```{r}
curve(dexp(x, rate = 1 / (sigma_1 + sigma_2)), 0, 20)
```

This corresponds to placing large mass on small values of the constraint $K$, which may not correspond well to actual knowledge about the resource constraints.

### 2023-04-17 notes/allocation_notes.rmd
rewriting/correcting expected score derivations

@@
 urlcolor: blue
-bibliography: allocation.bib 
+bibliography: allocation_notes.bib 
 header-includes:
@@
    - \DeclareMathOperator*{\argmin}{argmin}
-
 ---
@@
 By making the forecaster aware of $s$ prior to prediction, we elicit an $\alpha$-level quantile $q_{\alpha,F}$ of the forecast distribution $F$ for $Y$.
-That is, if a forecaster believes that $Y \sim F$, then according to this belief they minimize their expected score
-\begin{align} 
-\sbar_F(x) = E_F[s(x,Y)]  &= O E_F[(g(x) - g(Y))\sb{+}] + U E_F[(g(x) - g(Y))\sb{-}] \label{eqn:scalar_obj}\\
-&= O\int_{-\infty}^x(g(x) - g(y))f(y)dy + U\int_x^{\infty}(g(y) - g(x))f(y)dy \label{eqn:scalar_obj_int}
-\end{align}
-by forecasting a quantile $q_{\alpha, F} \in Q_F\left(\alpha\right)$ which solves the first order equation
+To see this, notice that if a forecaster believes that $Y \sim F$, they will calculate their expected score to be
+\begin{align} 
+\sbar_F(x) = E_F[s(x,Y)]  &= O E_F[(g(x) - g(Y))\sb{+}] + U E_F[(g(x) - g(Y))\sb{-}]. \label{eqn:scalar_obj}
+\end{align}
+In order to minimize this expectation, they must forecast a solution of the first order equation
 \begin{align}
@@
 &= g^{\prime}(x)\left( O F(x) - U (1-F(x))\right) \\
-&= \kappa g^{\prime}(x)\left(F(x) - \alpha\right) \label{scalar_Z_deriv}.
-\end{align}
+&= \kappa g^{\prime}(x)\left(F(x) - \alpha\right), \label{scalar_Z_deriv}
+\end{align}
+that is, a quantile $q_{\alpha, F} \in Q_F\left(\alpha\right)$.
 (See \hyperref[scalar_Z_deriv_detail]{appendix} for an alternative derivation in terms of a density.)
@@
 \begin{align}
-V(x,y) &\coloneq (1-\alpha)\mathbb{1}\{y<x\} - \alpha \mathbb{1}\{y \geq x\} \mathrel{\phantom{=}}\\
-&= (1-\alpha)\mathbb{1}\{y<x\} - \alpha \left(1 - \mathbb{1}\{y<x\}\right) \\
-&= \mathbb{1}\{y<x\} - \alpha. 
+V(x,y) &\coloneq (1-\alpha)\mathbb{1}\{y < x\} - \alpha \mathbb{1}\{y \geq x\} \mathrel{\phantom{=}}\\
+&= (1-\alpha)\mathbb{1}\{y < x\} - \alpha \left(1 - \mathbb{1}\{y < x\}\right) \\
+&= \mathbb{1}\{y < x\} - \alpha. 
 \end{align}
@@
 
-Using $u_- = u_+ - u$, the objective function \eqref{eqn:scalar_obj} can also be helpfully rewritten as
-\begin{align}
-\nonumber \sbar_F(x) &= OE_F[(g(x) - g(Y))\sb{+}] + U E_F[(g(x) - g(Y))\sb{+}] - U E_F[(g(x) - g(Y))] \\ 
-\nonumber  &= (O+U)E_F[(g(x) - g(Y))\sb{+}] - U E_F[(g(x) - g(Y))] \\
-&= (O+U) \int_{-\infty}^{x} (g(x) - g(y))f(y)dy - Ug(x)+ U \overline{g}\sb{F} \label{obj_integral}\\
-&= (O+U)\left(g(x)F(x) - \overline{g}_F(x) \right) - Ug(x) + U\overline{g}\sb{F}) \label{obj_sep}\\
-&= \kappa\left[g(x)(F(x) - \alpha) -\overline{g}_F(x) + \alpha \overline{g}\sb{F}\right] \label{obj_pexp} 
-\end{align}
-where $\overline{g}_F(x) = E_F[g(Y)\mathbb{1}\{Y \leq x\}]$, and $\overline{g}\sb{F} = E_F[g(Y)]$.
-
-The function $\overline{g}_F(x)$ sometimes goes by the name of ``partial expectation" (Schlaifer Raiffa, p. 109) and can often be made more explicit for certain $F$ and $g$.
-
-##  Alternative Notation Possibilities
-
-Let $Y \sim F,F_i$ be various forecaster beliefs and $Y \sim G$ be reality.
-\begin{align}
-\overline{s}_{F,\kappa,\alpha}(x) &:= s_{\kappa,\alpha}(x,F) := E_F[s_{\kappa,\alpha}(x,Y)] \\
-\overline{s}_{\kappa,\alpha}(x) &:= s_{\kappa,\alpha}(x,G) \\
-s_{\symbfit{\kappa},\symbfit{\alpha}}(\mathbf{x}, \mathbf{F}) &:= 
-\overline{s}_{\mathbf{F},\symbfit{\kappa},\symbfit{\alpha}}(\mathbf{x}) = 
-\sum_{i}^{\mathrm{dim}\mathbf{x}} s_{\kappa_i,\alpha_i}(x_i, F_i) \\
-S(F,y) &:= s(q_{F,\alpha}, y) \\
-S(F_1,F_2) &:= E_{F_2}[S(F_1,Y)] = s(q_{F_1,\alpha}, F_2) \\
-\overline{S}(F) &:= E_G[s(q_{F, \alpha}, Y)]
-\end{align}
-
-Define the $s$-entropy of an arbitrary $F$ 
-\begin{align}
-e_s(F) &:= E_F[S(F, Y)] = E_F[s(q_{F,\alpha}, Y)] = s(q_{F,\alpha}, F) \\
-\end{align}
-and of the RV $Y$ as
-\begin{align}
-e_s(Y) &:= e_s(G).
-\end{align}
-Define the $s$-divergence between two forecasts as
-\begin{align}
-d_s(F_1,F_2) &:= S(F_1,F_2) - S(F_2,F_2) = S(F_1,F_2) - e_s(F_2)
-\end{align}
-and the $s$-divergence from $F$ to $Y$ as
-\begin{align}
-d_s(F,Y) &:= S(F,G) - e(Y).
-\end{align}
-
-Thus, $\overline{S}(F) = d_s(F,Y) + e_s(Y)$.
-
-## Examples for various forecaster beliefs
-
-### 1) $Y \sim U[a,b]$
-
-Here $F(x) = \mathbb{1}\{a<x<b\}(x-a)/(b-a) + \mathbb{1}\{b \leq x\}$ in \eqref{obj_pexp} gives 
-<!--  
-&= 
-\begin{cases}
-\frac{U}{b-a}\int_a^b (g(y) - g(x))dy & x\leq a \\[8pt]
-\frac{O}{b-a}\int_a^x (g(x) - g(y))dy + \frac{U}{b-a}\int_x^b (g(y) - g(x))dy & a < x < b \\[8pt]
-\frac{O}{b-a}\int_a^b (g(x) - g(y))dy & b \leq x 
-\end{cases} \\ 
--->
+## Examples of forecaster expected scores
+
+Corresponding to the two expressions of an expectation $E_F[h(y)\mathbb{1}\{a < y<b\}]$ as either side of the Riemann-Stieltjes integration-by-parts 
+formula $\int_{a}^{b}h(y)dF(y) = h(y)F(y) \big|_{a}^{b} - \int_{a}^{b}F(y)dh(y)$, the expected score \eqref{eqn:scalar_obj}
+can be written as either
+\begin{align}
+\sbar_F(x) &= O \int_{\mathbb{R}} \int_{\mathbb{R}} \mathbb{1}\{Y \leq u < x\}dg(u) dF(y) + U \int_{\mathbb{R}} \int_{\mathbb{R}} \mathbb{1}\{x \leq u < Y\}dg(u) dF(y) \\
+&= O \int_{\mathbb{R}} \mathbb{1}\{u < x\} F(u) dg(u) + U \int_{\mathbb{R}} \mathbb{1}\{u \geq x\} (1-F(u)) dg(u) \\
+&= O \int_{-\infty}^x F(u) dg(u) + U \int_{x}^{\infty}(1-F(u)) dg(u), \label{eqn:scalar_obj_cdf_form} 
+\end{align}
+or
+\begin{align}
+\nonumber \sbar_F(x) &= O\int_{-\infty}^x(g(x) - g(y))dF(y) + U\int_x^{\infty}(g(y) - g(x))dF(y)\\ 
+&= O\left(g(x)F(x) - \int_{-\infty}^x g(y)dF(y) \right) + U \left( \int_x^{\infty} g(y)dF(y) - g(x)(1-F(x))\right) \\
+&= g(x)\left((O+U)F(x) - U\right) -  (O+U)\int_{-\infty}^x g(y)dF(y) + U \int_{\mathbb{R}} g(y)dF(y) \\
+&= \kappa\left[g(x)(F(x) - \alpha) -\overline{g}_{F}(x) + \alpha \overline{g}\sb{F}\right] \label{eqn:obj_pexp} 
+\end{align}
+where 
+\begin{align}
+\overline{g}_F(x) &= E_{F}[g(Y)\mathbb{1}\{Y \leq x\}] \label{eqn:gbarx} \\
+\overline{g}\sb{F} &= E_{F}[g(Y)]. \label{eqn:gbar}
+\end{align}
+
+The function $\overline{g}_F(x)$ sometimes goes by the name of ``partial expectation" (Schlaifer Raiffa, p. 109) and can often be made more explicit
+for certain $F$ and $g$.
+
+Taking, for example, $g(x)=x$ and $F = \mathrm{Exp}(1/\sigma)$, we have
+\begin{align}
+\overline{g}_{F}(x) = \int_{0}^{\max(0,x)} y \frac{e^{-y/\sigma}}{\sigma}dy = \mathbb{1}\{x \geq 0\} \left[\sigma-e^{-x / \sigma}(\sigma+x)\right]
+\end{align}
+so that
+\begin{align}
+\sbar_{\mathrm{Exp}(1/\sigma)}(x) 
+&= \kappa\left[x(\mathbb{1}\{x \geq 0\} \left[1 -e^{-x / \sigma}\right]  - \alpha) - \mathbb{1}\{x \geq 0\} \left[\sigma-e^{-x / \sigma}(\sigma+x)\right] + \alpha\sigma\right] \\
+&= \kappa\left[\mathbb{1}\{x \geq 0\}\sigma e^{-x / \sigma} + (\mathbb{1}\{x \geq 0\} - \alpha)(x-\sigma)\right].
+\end{align}
+
+If, on the other hand, $F=F_{\mu, \sigma}$ lies in a location-scale family, that is, $F(x)= F_0\left(\frac{x-\mu}{\sigma}\right)$, we have
+\begin{align}
+\int_{-\infty}^{x} y dF(y) &= \int_{-\infty}^{\frac{x-\mu}{\sigma}}(\mu + \sigma z)dF_0(z) \\
+& = \mu F(x) + \sigma \int_{-\infty}^{\frac{x-\mu}{\sigma}}zdF_0(z).
+\end{align}
+Then \eqref{eqn:obj_pexp} becomes
+\begin{align}
+\sbar_F(x) &= \kappa\left[x(F(x) - \alpha) - \mu F(x) - \sigma \int_{-\infty}^{\frac{x-\mu}{\sigma}}zdF_0(z)  + \alpha E_F[Y]\right] \\
+&= \kappa\left[F(x)(x-\mu) - \alpha(x - E_F[Y]) - \sigma \int_{-\infty}^{\frac{x-\mu}{\sigma}}zdF_0(z)\right] \label{eqn:lc-form}
+\end{align}
+
+For example, if $F = U[a,b]$ is uniform we can take $F_0 = U[0,1], \mu = a, \sigma = b-a$, and \eqref{eqn:lc-form} is
 \begin{align}
@@
 \begin{cases}
-\frac{\kappa}{b-a}\left( - \alpha(b-a)g(x) + \alpha\int_a^b g(y)dy\right) & x\leq a \\[8pt]
-\frac{\kappa}{b-a}\left((x-a- \alpha(b-a))g(x) - \int_a^x g(y)dy + \alpha\int_a^b g(y)dy\right) & a < x < b \\[8pt]
-\frac{\kappa}{b-a}\left((1-\alpha)(b-a)g(x) -(1-\alpha)\int_a^b g(y)dy \right) & b \leq x
-\end{cases} 
-\end{align}
-
-When $g(x)=x$ this becomes
-\begin{align}
-\sbar_F(x) &= 
-\begin{cases}
-\frac{\kappa}{b-a}\left( - \alpha(b-a)x + \frac{\alpha}{2}(b^2-a^2)\right) & x\leq a \\[8pt]
-\frac{\kappa}{b-a}\left((x-a- \alpha(b-a))x - \frac{1}{2}(x^2 - a^2) + \frac{\alpha}{2}(b^2-a^2)\right) & a < x < b \\[8pt]
-\frac{\kappa}{b-a}\left((1-\alpha)(b-a)x -\frac{(1-\alpha)}{2}(b^2-a^2)\right) & b \leq x
+-\alpha\kappa\left[x - \frac{a+b}{2}\right] & x\leq a \\[8pt]
+\kappa\left[\frac{x-a}{b-a}(x-a) -\alpha\left(x - \frac{a+b}{2}\right) -  (b-a) \frac{1}{2}\left(\frac{x-a}{b-a}\right)^2\right] & a < x < b \\[8pt]
+\kappa\left[x - a -\alpha\left(x - \frac{a+b}{2}\right) - \frac{b-a}{2}\right] & b \leq x
 \end{cases} \\
@@
 
-Thus $\sbar_F(x)$ has the form of pinball loss centered on the mean $\mu_Y = \frac{a+b}{2}$ and interpolated on $(a,b)$ 
- by a quadratic centered at the quantile $a+\alpha(b-a)$ with slopes matching the linear tails at $a$ and $b$.
+Thus, outside $(a,b)$, $\sbar_{U[a.b]}(x)$ has the form of pinball loss centered on the mean $m_F = \frac{a+b}{2}$ and is interpolated on $(a,b)$ 
+ by a quadratic centered at the quantile $a+\alpha(b-a)$ with slopes matching the linear tails at $a$ and $b$ (see figure 1).
+
+Or if $F=\Phi_{\mu,\sigma}$ is normal with location $\mu = E_F[Y]$ and scale $\sigma = \sqrt{\mathrm{Var}(Y)}$, 
+we have 
+\begin{align}
+\int_{-\infty}^{\frac{x-\mu}{\sigma}}zdF_0(z)=-\frac{1}{\sqrt{2\pi}}\exp\left\{-\frac{(x-\mu)^2}{2 \sigma^2}\right\} = - \sigma\varphi_{\mu, \sigma}(x)
+\end{align}
+so that
+\begin{align}
+\sbar_F(x) = \kappa\left[(\Phi_{\mu,\sigma}(x) - \alpha)(x-\mu) + \sigma^2 \varphi_{\mu, \sigma}(x)\right].
+\end{align}
 
@@
   ) + ggtitle("Unif(0,1)")
-```
-
-### 2) $g(x)=x$ and $F_Y=F_{\mu, \sigma}$ lies in a location-scale family
-
-That is, $Y = \mu + \sigma Z$ and $Z \sim F_Z = F_0$. Then $f_Y(y) = \dby{}{z} F_Z\left(\frac{y-\mu}{\sigma}\right)\dby{z}{y} = f_Z(z)/\sigma$ and so
-\begin{align}
-\overline{g}_F(x) = \int_a^x y f_Y(y)dy &= \int_{\frac{a-\mu}{\sigma}}^{\frac{x-\mu}{\sigma}}(\mu + \sigma z)f_Z(z)dz \\
-& = \mu F_Z\left(\frac{x-\mu}{\sigma}\right) + \sigma \int_{\frac{a-\mu}{\sigma}}^{\frac{x-\mu}{\sigma}}zf_Z(z)dz\\
-& =\mu F_Y(x) + \sigma \mu_{F_0}\left(\frac{x-\mu}{\sigma}\right).
-\end{align}
-where $\mu_{F_0}(u) = \int_{\frac{a-\mu}{\sigma}}^{u} zf_0(z)dz$. This gives
-<!-- With the change of variables $y=Q_F(\alpha) = F^{-1}(\alpha), dy =  d(F^{-1}(\alpha)) = \frac{d\alpha}{f(y)}$, we have
-\begin{align}
-\overline{g}_F(x) = \int_a^x y f_Y(y)dy = \int_{0}^{F(x)}Q_F(\alpha)d\alpha 
-& = \mu F(x) + \sigma \int_0^{F(x)} Q_0(\alpha)d\alpha \\
-& =\mu F(x) + \sigma \mu_{F_0}\left(\frac{x-\mu}{\sigma}\right).
-\end{align} -->
-\begin{align}
-\nonumber \sbar_F(x) &= \kappa\left[x(F(x) - \alpha) - \mu F(x) - 
-\nonumber \sigma \mu_{F_0}\left(\frac{x-\mu}{\sigma}\right) + \alpha \mu \right]  \\
-&= \kappa\left[\left(F(x) - \alpha\right)(x-\mu)- \sigma \mu_{F_0}\left(\frac{x-\mu}{\sigma}\right) \right]
-\label{lc-form}
-\end{align}
-
-### 3) $g(x)=x$ and $Y \sim N(\mu,\sigma)$
-
-When $F=\Phi_{\mu,\sigma}$, $\mu_{F_0}\left(\frac{x-\mu}{\sigma}\right)=-\sigma \varphi_{\mu, \sigma}(x)$, giving $q(\alpha)$ as the minimum of 
-$$
-Z_F(x) = \kappa\left[(\Phi_{\mu,\sigma}(x) - \alpha)(x-\mu) + \sigma^2 \varphi_{\mu, \sigma}(x)\right].
-$$
-
-```{r echo = FALSE, cache=TRUE}
+
 norm_loss <- function(alpha, mean, sd) {
@@
 ) + ggtitle("Normal(0,2)")
-```
-
-### 4) $g(x)=x$ and $Y \sim \mathrm{Exp}(1/\sigma)$
-
-Taking $\mu \equiv 0, Y = \sigma Z$ and $F_Z(x) = F_0(x) = \mathbb{1}\{0 \leq x\}(1 - e^{-x})$ we have
- $\mu_{F_0}(x) = \mathbb{1}\{0 \leq x\}(1 - e^{-x}(1+x))$. Then \eqref{lc-form} is
-
-\begin{align}
-\nonumber \sbar_F(x) &= 
-\mathbb{1}\{0 \leq x\}\kappa\left[F(x)x - \sigma \mu_{F_0}\left(\frac{x}{\sigma}\right) \right] - \kappa\alpha x\\
-&= \mathbb{1}\{0 \leq x\}\kappa
-\left[
-\left(1 - e^{-x/\sigma}\right)x - \sigma \left(1- e^{-x/\sigma}\left(1+\frac{x}{\sigma}\right) \right)
-\right] - \kappa\alpha x\\
-&= \mathbb{1}\{0 \leq x\}\kappa\left[ \sigma e^{-x/\sigma} + x - \sigma \right] - \kappa\alpha x
-\end{align}
-
-```{r echo = FALSE, cache=TRUE}
+
 exp_loss <- function(alpha, rate) {
   f <- function(x){
-    ((exp(-x*rate)*rate^(-1) + x - rate^(-1)))*(x>0) - alpha*x
+    exp(-x*rate)*(rate^(-1))*(x>=0) + ((x>=0) - alpha)*(x - rate^(-1))
   }
@@
 
-## Another form for $\sbar_F(x)$
-
-There also may be situations where it is convenient to integrate the $(O+U)$-term in \eqref{obj_sep} by parts to get 
-\begin{align}
-\sbar_F(x) = \kappa\left[\int_a^x g^{\prime}(y)F(y)dy - \alpha(g(x) - \overline{g}\sb{F})\right]. \label{obj_intF}
-\end{align}
+We finally note that for non-trivial $g$'s it may be helpful to make the substitution $v=g(u)$ in \eqref{eqn:scalar_obj_cdf_form} to get
+\begin{align}
+\sbar_F(x) &=  O \int_{-\infty}^{g(x)} F(g^{-1}(u)) du + U \int_{g(x)}^{\infty}(1-F(g^{-1}(u))) du
+\end{align}
+and in \eqref{eqn:gbarx} to get
+\begin{align}
+\overline{g}_F(x) &= \int_{-\infty}^{g(x)}v d(F \circ g^{-1})(v).
+\end{align}
+
+##  Alternative Notation Possibilities
+
+Let $Y \sim F,F_i$ be various forecaster beliefs and $Y \sim G$ be reality.
+\begin{align}
+\overline{s}_{F,\kappa,\alpha}(x) &:= s_{\kappa,\alpha}(x,F) := E_F[s_{\kappa,\alpha}(x,Y)] \\
+\overline{s}_{\kappa,\alpha}(x) &:= s_{\kappa,\alpha}(x,G) \\
+s_{\symbfit{\kappa},\symbfit{\alpha}}(\mathbf{x}, \mathbf{F}) &:= 
+\overline{s}_{\mathbf{F},\symbfit{\kappa},\symbfit{\alpha}}(\mathbf{x}) = 
+\sum_{i}^{\mathrm{dim}\mathbf{x}} s_{\kappa_i,\alpha_i}(x_i, F_i) \\
+S(F,y) &:= s(q_{F,\alpha}, y) \\
+S(F_1,F_2) &:= E_{F_2}[S(F_1,Y)] = s(q_{F_1,\alpha}, F_2) \\
+\overline{S}(F) &:= E_G[s(q_{F, \alpha}, Y)]
+\end{align}
+
+Define the $s$-entropy of an arbitrary $F$ 
+\begin{align}
+e_s(F) &:= E_F[S(F, Y)] = E_F[s(q_{F,\alpha}, Y)] = s(q_{F,\alpha}, F) \\
+\end{align}
+and of the RV $Y$ as
+\begin{align}
+e_s(Y) &:= e_s(G).
+\end{align}
+Define the $s$-divergence between two forecasts as
+\begin{align}
+d_s(F_1,F_2) &:= S(F_1,F_2) - S(F_2,F_2) = S(F_1,F_2) - e_s(F_2)
+\end{align}
+and the $s$-divergence from $F$ to $Y$ as
+\begin{align}
+d_s(F,Y) &:= S(F,G) - e(Y).
+\end{align}
+
+Thus, $\overline{S}(F) = d_s(F,Y) + e_s(Y)$.
 
@@
 \end{align}
-Note, for reference, that in the literature $\sbar_{n,F}(x)$ is often encountered in the form of \eqref{obj_integral}:
-\begin{align}
-&(O+U) \int_{0}^{x} (x - y)f(y)dy - Ux+ (U+a_x+a_y)E[Y] \\
+Note, for reference, that in the literature $\sbar_{n,F}(x)$ is often encountered in the form
+\begin{align}
+&(O+U) \int_{0}^{x} (x - y)f(y)dy - Ux+ (U+a_x+a_y)E[Y] \\{}
 =\,&(p+r-u) \int_{0}^{x} (x - y)f(y)dy - (p+r-c)x+ rE[Y]
@@
 the \emph{cost-loss ratio} of the problem.
+
+There is considerable opportunity here for confusion as to the role of $x$ if one is following (as we mostly are) the notation of [@ehm2016quantiles].
 
@@
 
-According to \eqref{eqn:scalar_obj_int} and \eqref{obj_pexp} the objective function is now
+According to \eqref{eqn:scalar_obj} and \eqref{eqn:obj_pexp} the objective function is now
 \begin{align}

### 2023-04-17 notes/allocation_notes.rmd
slight edits of notes

@@
 \end{align}
-so that
+so that \eqref{eqn:obj_pexp} is
 \begin{align}
@@
 
-If, on the other hand, $F=F_{\mu, \sigma}$ lies in a location-scale family, that is, $F(x)= F_0\left(\frac{x-\mu}{\sigma}\right)$, we have
+And if $F=F_{\mu, \sigma}$ lies in a location-scale family, that is, $F(x)= F_0\left(\frac{x-\mu}{\sigma}\right)$, we have
 \begin{align}

### 2023-04-18 notes/crps_connection_exp.Rmd

@@
 
-If I've correctly translated from the notes document, the expected loss function is
+The expected loss function is
 \begin{equation*}
-\bar{s}_F(\mathbf{x}) = \sum_{i=1}^2 \mathbb{1}\{0 \leq x_i\}\left[ \sigma e^{-x_i/\sigma} - \sigma \right]
+\bar{s}_F(\mathbf{x}) = \sum_{i=1}^2 \mathbb{1}\{x_i < 0\}(\sigma - x_i) + \mathbb{1}\{0 \leq x_i\}\sigma e^{-x_i/\sigma}.
 \end{equation*}
-
-I'm going with this below because I think it's close, but I don't actually think it's correct because it leads to negative loss values.
 
@@
 
-x_1 <- qexp(tau, rate = 1 / sigma_1)
-x_2 <- qexp(tau, rate = 1 / sigma_2)
+x_1star <- qexp(tau, rate = 1 / sigma_1)
+x_2star <- qexp(tau, rate = 1 / sigma_2)
 
-n_grid <- 1001
-grid_l <- 0.01
-grid_u <- 6.0
+n_grid <- 101
+grid_l <- -2.0
+grid_u <- 9.0
 y_grid <- tidyr::expand_grid(
@@
 
+s_bar <- function(x_1, x_2) {
+    sigma_1 * exp(-x_1 / sigma_1)*(x_1>=0) + sigma_2 * exp(-x_2 / sigma_2)*(x_2>=0) +
+    (sigma_1 - x_1)*(x_1<0) + (sigma_2 - x_2)*(x_2<0) 
+}
+
 joint_dist <- y_grid %>%
     dplyr::mutate(
-        s_bar = sigma_1 * exp(-y_1 / sigma_1) - sigma_1 +
-                sigma_2 * exp(-y_2 / sigma_2) - sigma_2
+        s_bar = s_bar(y_1, y_2)
     )
@@
     geom_raster(aes(x = y_1, y = y_2, fill = s_bar)) +
-    geom_contour(mapping = aes(x = y_1, y = y_2, z = s_bar)) +
+    geom_contour(mapping = aes(x = y_1, y = y_2, z = s_bar), breaks = s_bar(x_1star, x_2star)%%1 + seq(0, 10, by = .5)) +
     geom_abline(intercept = K, slope = -1) +
-    geom_point(x = x_1, y = x_2, color = "orange") +
+    geom_point(x = x_1star, y = x_2star, color = "orange") +
     scale_fill_viridis_c() +
-    xlim(0, 6) +
-    ylim(0, 6) +
+    xlim(grid_l, grid_u) +
+    ylim(grid_l, grid_u) +
     theme_bw()

### 2023-04-20 alloscore_manuscript/allocation.rmd
add Bertsimas et al. to introduction

@@
 
-Infectious disease forecasts have been used as an input to public health decision-making processes such as ...
-
+Infectious disease forecasts have been used to inform decision-making about a wide variety of measures designed to reduce disease spread and/or mitigate the severity of disease outcomes, including the allocation of limited medical supplies such as ventilators [@bertsimas2021predictionsCOVID], implementation of social distancing measures such as stay-at-home policies, and planning site selection for vaccine trials [@bertsimas2021predictionsCOVID].
 In this manuscript we argue that for many decision-making problems, it is possible and valuable to develop scoring rules for probabilistic forecasts that are responsive to that decision-making context.

### 2023-04-24 alloscore_manuscript/allocation.rmd
some work on introduction of allocation mansucript

--- 2023-03-31 alloscore_manuscript/allocation.rmd
@@
 
+\section{Abstract}
+
+The COVID-19 pandemic has led to rapid innovation in methods for eliciting and evaluating forecasts of infectious disease burdens, with a primary goal being to help public health workers make informed decisions about how to manage these burdens.
+However, explicit descriptions or quantifications of the value forecasts add to society through the decisions they support are elusive.
+Moreover, there has only been limited discussion of how predominant forecast evaluation metrics might indicate the success of policies based in part those forecasts.
+
+Here we pursue one possible tether between multivariate forecasts and policy: the allocation of limited medical resources in response to COVID-19 hospitalizations in various regions so as to minimize expected unmet need.
+Given probabilistic forecasts of hospitalizations in each region, we formulate an allocation algorithm following techniques developed in operations research.
+We then score forecasts according to how much unmet need their associated allocations would have allowed.
+We illustrate this scheme with quantile forecasts of COVID-19 hospitalizations in the US at the state level that are recorded in the COVID-19 Forecast Hub, with the goal of determining the allocation of a hypothetical limited resource across the states.
+The forecast skill ranking given by this allocation scoring rule can vary substantially from the ranking given by the weighted interval score now used by the CDC, especially during surges in hospitalizations such as in late 2021 as the Omicron wave began.
+We see this as strong evidence that the allocation scoring rule detects forecast value that is missed by traditional accuracy measures and that the general strategy of designing scoring rules directly linked to policy performance is a promising research direction for epidemic forecast evaluation.
+
 \section{Introduction}
 
-Infectious disease forecasts have been used as an input to public health decision-making processes such as ...
+High level points to cover in introduction:
+\begin{itemize}
+\item People are using infectious disease forecasts as an input to decision making
+\item There are standard ways to evaluate forecasts that are responsive to decision making context, and use of those methods is relatively common in other fields like economics and meteorology
+\item However, there's not much work in infectious disease that does this
+\item In practice, infectious disease forecasts have typically been evaluated with "off the shelf" scoring rules such as the WIS which is an approximation to CRPS, log score, and so on.
+\item In this work, our goal is to begin to address that gap. We focus on a resource allocation setting.
+\item There is past work focusing on resource allocation in the operations research literature, but it doesn't take the step of getting to a measure of forecast skill.
+\end{itemize}
 
-In this manuscript we argue that for many decision-making problems, it is possible and valuable to develop scoring rules for probabilistic forecasts that are responsive to that decision-making context.
+Infectious disease forecasts have been used to inform decision-making about a wide variety of measures designed to reduce disease spread and/or mitigate the severity of disease outcomes.
+Such decision-making applications include the allocation of limited medical supplies such as ventilators [@bertsimas2021predictionsCOVID], implementation of social distancing measures such as stay-at-home policies, planning site selection for vaccine trials [@bertsimas2021predictionsCOVID], and strategies for public health communication campaigns.
 
-\subsection{Literature review}
+In decision-making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to.
+We give an overview of these procedures in Section \ref{sec:methods.decisiontheory}.
+These methods have been applied with some regularity in fields such as economics and meteorology.
+[review previous applications of decision-theoretic evaluation to fields like economics and high level description of scoring procedures they use]
 
-Previous work in infectious disease forecasting where applications of forecasting to decision-making problems have been developed
+However, we are aware of only a limited body of work that explicitly attempts to measure the value of infectious disease forecasts through their impact on policy, and much of this discussion has proceeded informally.
+For example, [@ioannidis2022forecastingCOVIDfailed] discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts.
+Separately, there is a thread of literature that does quantify the link between infectious disease modeling and policy making, but this work has been done outside of a forecasting context.
+As an example, [@Probert2016decisionMakingFootMouth] develop measures of the cost of actions designed to control a hypothetical outbreak of foot-and-mouth disease and use this framework to explore policy recommendations from a variety of simulation-based projection models.
 
-Theoretic work on decision-theoretic set up for forecast evaluation
+In practice, probablistic infectious disease forecasts have most often been evaluated with standard, off-the-shelf scoring rules such as the log score, continuous ranked probability score (CRPS), or weighted interval score (WIS).
+[cite some examples] While some of these scores can be interpreted through the lens of decision theory [thinking here of WIS/quantile loss], these connections are not a common focus of infectious disease forecast evaluation.
 
-Previous applications of decision-theoretic evaluation to fields like economics
-
-Previous decision-theoretic work in infectious disease forecast evaluation.  I think there are at least a couple of papers out there at some level of formality.  Maybe Kat Shea has something?
+In this work, we address this gap between the ways in which infectious disease forecasts have been used to support public health policy and the ways in which they have traditionally been evaluated.
+We work with a motivating example where forecasts are used to help set the allocation of a limited quantity of medical supplies across multiple regions.
 
@@
 
+The remainder of this article is organized as follows.
+In Section \ref{sec:methods}, we review the general framework for developing scoring rules for probabilistic forecasts using the tools of decision theory, develop a novel scoring rule that is motivated by the problem of allocating limited medical supplies, and explore the relationship between the proposed allocation score and existing scoring rules such as CRPS.
+We then illustrate the scoring rule through an application to short-term forecasts of COVID-19 hospital admissions in the US in section \ref{sec:application}.
+Section \ref{sec:discussion} summarizes our contributions and discusses opportunities for further extensions in future work.
+
 \section{Methods}
+\label{sec:methods}
 
@@
 \subsection{The decision-theoretic setup for forecast evaluation}
+\label{sec:methods.decisiontheory}
 
@@
 \section{Application}
+\label{sec:application}
 
@@
 \section{Discussion}
+\label{sec:discussion}
 

### 2023-04-25 alloscore_manuscript/allocation.rmd
two very small but potentially crucial additions

@@
 \item There is past work focusing on resource allocation in the operations research literature, but it doesn't take the step of getting to a measure of forecast skill.
+\item That is, while there is plenty of work on 
+\begin{itemize}
+\item DM under risk, where probabilities are taken as known and expected utility is maximized
+\item DM under uncertainty, where probabilities are taken as unknown and robust (i.e. maximin) decisions are sought
+\end{itemize}
+little seems to have been said on how forecasts can be evaluated by how well they convert a problem of DM under uncertainty to a problem of DM under risk.
 \end{itemize}

### 2023-04-28 alloscore_manuscript/methods_new.rmd
new methods draft

---
title: "Binary to continuous methodology draft"
date: "`r Sys.Date()`"
output: 
    pdf_document:
        latex_engine: xelatex
urlcolor: blue
bibliography: allocation.bib 
header-includes:
   - \usepackage{amsmath}
   - \usepackage{algorithm}
   - \usepackage[noend]{algpseudocode}
   - \usepackage{amsthm}
   - \usepackage{bm}
   - \usepackage{cases}
   - \usepackage{caption}
   - \usepackage{unicode-math}
   - \usepackage{hyperref}
   - \DeclareMathOperator*{\argmin}{argmin}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE, 
    message = FALSE, 
    warning = FALSE,
    fig.width=10, 
    fig.height=6)
library(tidyverse)
library(patchwork)
```
<!-- $ ...to toggle Latexyz math scope off -->
\newcommand{\del}[2]{\frac{\partial {#1} }{\partial {#2}} }
\newcommand{\dby}[2]{\frac{d {#1} }{d {#2}} }
\newcommand{\sbar}{\overline{s}}
\newtheorem{proposition}{Proposition}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}

We begin with a basic task in public health: triage at the unit level, that is, the problem of deciding to whom, where, and when a unit of scarce health-care resource is allocated. We assume that the resource unit, e.g., a bed, ventillator, dose, or clinician, is indivisible, and that a potential recipient is a single person during a single period of time (say, a day).  We also focus on the scenario of a recipient that is not yet under care, such as a person that may be infected with SARS-COV-2 in the future and consequently require hospitalization or intubation.

A decision problem is...

Unit-level decision problem: allocate to potential case in location $a$ or $b$. With a decision variable $d$ we write this as $d=a$ or $d=b$.

- elicit forecast $P(Y_a=1) = p$ where $Y_a=1$ if a case in addition to what is currently provided for at horizon $h$ occurs during the day at horizon $h$ in $a$, and 0 otherwise.  $p$ can in particular come from a distributional incidence forecast $F_a$ via $p = 1-F_a(x)$ where $x$ is the number of cases currently provided for in $h$ days.

- fix prob $\beta$ of additional case in $b$, via

	- baseline/persistence/climatology, or

	- $F_b$ taken as a given, not to be evaluated, or

	- $F_{\sum i \neq a}(x) = \int_{\mathbb{R}^n}\mathbb{1}\{\sum_{i \neq a} x_i \leq x\}dF$ if $b$ is a collection of locations and $F$ is a given joint distribution for all coations other than $a$.

We define a loss function for the decision problem that encodes the dilemma faced when balancing the risk of a case of unmet need in $a$ incurred when $d=b$ is chosen against the guaranteed "average" unmet need of $\beta$ in $b$ that is incurred when $d=a$ is chosen. This is the function of the decision $d$ and outcome $Y_a$ 
$$
l(d,Y) = \mathbb{1}\{d=a\}\beta + \mathbb{1}\{d=b\}Y_a,
$$
which picks one of the 4 entries in the table
$$
\begin{array}{rcc}
\text{New case in $a$:}& \text{Yes} & \text{No} \\
\hline
\begin{array}{cr}
\text{allocate to:} & 
\begin{array}{cc}
a \\
b
\end{array}
\end{array}
&
\begin{array}{cc}
\beta \\
1
\end{array}
&
\begin{array}{cc}
\beta \\
0
\end{array}
\end{array}
$$

Associated with this loss function is the *regret* function 
$$
r(d,Y) = \mathbb{1}\{d=a\}\beta + (\mathbb{1}\{d=b\} - \beta)Y_a,
$$
with tabular form
$$
\begin{array}{rcc}
\text{New case in $a$:}& \text{Yes} & \text{No} \\
\hline
\begin{array}{cr}
\text{allocate to:} & 
\begin{array}{cc}
a \\
b
\end{array}
\end{array}
&
\begin{array}{cc}
0 \\
1 - \beta
\end{array}
&
\begin{array}{cc}
\beta \\
0
\end{array}
\end{array}
$$
This encodes the allocation dilemma in terms of how much worse our decision is than that of an oracle that only allocates to $a$ when $Y_a=1$.

According to the forecast $p$, the decision has expected loss and regret
\begin{align}
E_{p}[l(d)] &=  \mathbb{1}\{d=a\}\beta + \mathbb{1}\{d=b\}p \\
E_{p}[r(d)] &=  \mathbb{1}\{d=a\}\beta + (\mathbb{1}\{d=b\} - \beta)p 
\end{align}
which are minimized by the decision rule
$$
d(p) = 
\begin{cases}
a & \text{if  } p > \beta \\
b & \text{otherwise}.
\end{cases}
$$
This decision --- or *Bayes* --- rule allows us to convert $l$ and $r$ into a scoring functions
\begin{align}
s_l(p,Y_a) &= l(d(p),Y_a) = \mathbb{1}\{p > \beta\}\beta +\mathbb{1}\{p \leq \beta\}Y_a  \\
s_r(p,Y_a) &= s_l(p,Y_a) - \beta Y_a
\end{align}

- Show, a la Ehm eqs 13-16 that integrating over $x$ gives a Brier score for any $\beta$

	- relate to quantile score for $F_a$

- Show how looking at different $\beta$ with $x$ fixed gives a *value* or *Murphy* curve

	- integrating over $x$ now gives CRPS...

### 2023-05-01 alloscore_manuscript/allocation.tex
switch manuscript from rmd to tex

--- 2023-04-25 alloscore_manuscript/allocation.rmd
@@
----
-title: "Evaluating infectious disease forecasts with allocation scoring rules"
-date: "`r Sys.Date()`"
-output: 
-    pdf_document:
-        latex_engine: xelatex
-urlcolor: blue
-bibliography: allocation.bib 
-header-includes:
-   - \usepackage{amsmath}
-   - \usepackage{algorithm}
-   - \usepackage[noend]{algpseudocode}
-   - \usepackage{amsthm}
-   - \usepackage{bm}
-   - \usepackage{cases}
-   - \usepackage{caption}
-   - \usepackage{unicode-math}
-   - \usepackage{hyperref}
-   - \DeclareMathOperator*{\argmin}{argmin}
+\documentclass{article}
 
----
+\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-    echo = FALSE, 
-    message = FALSE, 
-    warning = FALSE,
-    fig.width=10, 
-    fig.height=6)
-library(tidyverse)
-library(patchwork)
-```
-<!-- $ ...to toggle Latexyz math scope off -->
+\usepackage{amsmath, amsfonts, amssymb}
+\usepackage{graphicx}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+\usepackage{amsthm}
+\usepackage{bm}
+\usepackage{cases}
+\usepackage{caption}
+\usepackage{hyperref}
+
+\DeclareMathOperator*{\argmin}{argmin}
+
+\usepackage{biblatex}
+\addbibresource{allocation.bib}
+
+\title{Evaluating infectious disease forecasts with allocation scoring rules}
+\author{Authors}
+
+\begin{document}
+
 \newcommand{\del}[2]{\frac{\partial {#1} }{\partial {#2}} }
@@
 
-\section{Abstract}
+\maketitle
+
+\begin{abstract}
 
@@
 We see this as strong evidence that the allocation scoring rule detects forecast value that is missed by traditional accuracy measures and that the general strategy of designing scoring rules directly linked to policy performance is a promising research direction for epidemic forecast evaluation.
+
+\end{abstract}
 
@@
 Infectious disease forecasts have been used to inform decision-making about a wide variety of measures designed to reduce disease spread and/or mitigate the severity of disease outcomes.
-Such decision-making applications include the allocation of limited medical supplies such as ventilators [@bertsimas2021predictionsCOVID], implementation of social distancing measures such as stay-at-home policies, planning site selection for vaccine trials [@bertsimas2021predictionsCOVID], and strategies for public health communication campaigns.
+Such decision-making applications include the allocation of limited medical supplies such as ventilators \cite{bertsimas2021predictionsCOVID}, implementation of social distancing measures such as stay-at-home policies, planning site selection for vaccine trials \cite{bertsimas2021predictionsCOVID}, and strategies for public health communication campaigns.
 
@@
 However, we are aware of only a limited body of work that explicitly attempts to measure the value of infectious disease forecasts through their impact on policy, and much of this discussion has proceeded informally.
-For example, [@ioannidis2022forecastingCOVIDfailed] discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts.
+For example, \cite{ioannidis2022forecastingCOVIDfailed} discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts.
 Separately, there is a thread of literature that does quantify the link between infectious disease modeling and policy making, but this work has been done outside of a forecasting context.
-As an example, [@Probert2016decisionMakingFootMouth] develop measures of the cost of actions designed to control a hypothetical outbreak of foot-and-mouth disease and use this framework to explore policy recommendations from a variety of simulation-based projection models.
+As an example, \cite{Probert2016decisionMakingFootMouth} develop measures of the cost of actions designed to control a hypothetical outbreak of foot-and-mouth disease and use this framework to explore policy recommendations from a variety of simulation-based projection models.
 
@@
 The remainder of this article is organized as follows.
-In Section \ref{sec:methods}, we review the general framework for developing scoring rules for probabilistic forecasts using the tools of decision theory, develop a novel scoring rule that is motivated by the problem of allocating limited medical supplies, and explore the relationship between the proposed allocation score and existing scoring rules such as CRPS.
+In section \ref{sec:methods}, we review the general framework for developing scoring rules for probabilistic forecasts using the tools of decision theory, develop a novel scoring rule that is motivated by the problem of allocating limited medical supplies, and explore the relationship between the proposed allocation score and existing scoring rules such as CRPS.
 We then illustrate the scoring rule through an application to short-term forecasts of COVID-19 hospital admissions in the US in section \ref{sec:application}.
@@
 \begin{enumerate}
-\item Specify a \textit{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$ eventually occurs.
-We use the letter $s$ for this function to align with the literature on forecast evaluation, in which context $s$ may be used as a \textit{scoring function}.
-\item Given a probabilistic forecast $F$, determine the \textit{Bayes act} $x^F$ that minimizes the expected loss under the distribution $F$.
-\item The \textit{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: $S(F, y) = s(x^F, y)$.
+\item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$ eventually occurs.
+We use the letter $s$ for this function to align with the literature on forecast evaluation, in which context $s$ may be used as a \emph{scoring function}.
+\item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under the distribution $F$.
+\item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: $S(F, y) = s(x^F, y)$.
 \end{enumerate}
@@
 
-<!-- , and we illustrate the ideas with two running examples: (1) some problem that results in quantile elicitation, (2) some allocation problem.
+% , and we illustrate the ideas with two running examples: (1) some problem that results in quantile elicitation, (2) some allocation problem.
 
-A forecaster is asked to recommend, that is, \emph{predict}, the appropriate supply $x$ of some good or investment in some precautionary measure intended to satisfy a random future demand or need $Y$.
+%A forecaster is asked to recommend, that is, \emph{predict}, the appropriate supply $x$ of some good or investment in some precautionary measure intended to satisfy a random future demand or need $Y$.
 As a running example, we will consider the provision of a limited supply of a limited health care resource such as oxygen or ventilators.
 Suppose there is an incremental loss $O \geq 0$ incurred when over-prediction leads to unused supply and an incremental loss $U > 0$ incurred when under-prediction leads to unmet demand or need.
--->
 
@@
 
-<!-- Set up of decision-making problem with costs for over-/under-prediction. Quantile/pinball loss.
+%Set up of decision-making problem with costs for over-/under-prediction. Quantile/pinball loss.
 
-Statement that this loss elicits quantiles as the Bayes act, derivations deferred to a supplement.
+%Statement that this loss elicits quantiles as the Bayes act, derivations deferred to a supplement.
 
-Interpretation of CRPS as an integral of quantile loss over quantile levels, WIS as a discrete approximation of CRPS. -->
+%Interpretation of CRPS as an integral of quantile loss over quantile levels, WIS as a discrete approximation of CRPS.
 
@@
 An alternative interpretation comes from noting that for each location $i$, $\frac{\partial}{\partial x_i} \sbar_F(\mathbf{x}) \vert_{\mathbf{x}=\mathbf{x}^F} = \lambda^{\star}$ (see the supplement for a proof).
-<!-- 
-$\frac{\partial}{\partial x_i} \sbar_F(\mathbf{x}) = L[F_i(x_i) - 1]$ (see the supplement for a derivation), and evaluating at the Bayes allocation we arrive at
-\begin{align}
-\frac{\partial}{\partial x_i} \sbar_F(\mathbf{x}) \vert_{\mathbf{x}=\mathbf{x}^F} &= L[F_i(x^F_i) - 1] = L[F_i(F_i^{-1}(1 - \lambda^{\star}/L)) - 1] = \lambda^{\star}.
-\end{align} -->
+%$\frac{\partial}{\partial x_i} \sbar_F(\mathbf{x}) = L[F_i(x_i) - 1]$ (see the supplement for a derivation), and evaluating at the Bayes allocation we arrive at
+%\begin{align}
+%\frac{\partial}{\partial x_i} \sbar_F(\mathbf{x}) \vert_{\mathbf{x}=\mathbf{x}^F} &= L[F_i(x^F_i) - 1] = L[F_i(F_i^{-1}(1 - \lambda^{\star}/L)) - 1] = \lambda^{\star}.
+%\end{align}
 In words, at the allocation given by the Bayes act, the rate of change of the expected score as a function of the amount allocated to location $i$ is given by $\lambda^{\star}$.
@@
 \section{References}
+
+\printbibliography
+
+\end{document}

### 2023-05-01 alloscore_manuscript/allocation.tex
minor updates to manuscript text

@@
 \addbibresource{allocation.bib}
+
+\usepackage{soul}
+\usepackage{color}
+\def\elr#1{{\color{cyan}\textbf{ELR:[#1]}}}
 
@@
 
-We first give a high-level review of a general procedure for developing proper scoring rules that are tailored to a specific decision-making task, and then use that procedure to develop a score that is suitable for evaluation of forecasts in the context of decisions about allocation of limited resources across multiple locations.
+We give a high-level review of a general procedure for developing proper scoring rules that are tailored to a specific decision-making task in section \ref{sec:methods.decisiontheory}.
+In section \ref{sec:methods.quantileloss} we review how quantile loss can be obtained within this framework in a setting where a decision-maker is required to determine the quantity of a good to procure while balancing the cost of purchasing an additional unit of the good with loss that may result from under-procurement.
+We then discuss how the continuous ranked probability score (CRPS) can be obtained as an integral of the quantile loss across values of the cost/loss ratio.
+These developments mirror the structure of section \ref{sec:methods.allocation}.
+There, we develop a novel \emph{allocation score} that is analogous to the quantile score but is suitable for evaluation of forecasts in the context of decisions about allocation of limited resources across multiple locations when the resource constraint is known.
+We then describe an \emph{integrated allocation score} that is analogous to the CRPS and is obtained by integrating the allocation score across values of the resource constraint.
 
@@
 In the framework of decision theory, a decision corresponds to the selection of an action $x$ from some set of possible actions $\mathcal{X}$.
-For example, $x$ may correspond to the level of investment in a measure designed to mitigate severe disease outcomes, with $\mathcal{X}$ being the set of all possible levels of investment that we might select.
+For example, $x$ may correspond to the level of investment in a measure designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with $\mathcal{X}$ being the set of all possible levels of investment that we might select.
 The quality of a decision to take a particular action $x$ is measured in relation to an outcome $y$ that is unknown at the time the decision is made.
-For example, $y$ may correspond to the number of individuals who eventually become sick and would benefit from the mitigation measure, and informally, an action $x$ is successful to the extent that it meets the realized demand.
+For example, $y$ may correspond to the number of individuals who eventually become sick and would benefit from the mitigation measure, and informally, an action $x$ is successful to the extent that it meets the realized need.
 In the face of uncertainty, a decision-maker may use a forecast $F$ of the random variable $Y$ to help inform the selection of the action to take.
@@
 \item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$ eventually occurs.
-We use the letter $s$ for this function to align with the literature on forecast evaluation, in which context $s$ may be used as a \emph{scoring function}.
 \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under the distribution $F$.
@@
 \end{enumerate}
+We use the letter $s$ for the loss function to align with the literature on evaluation of forecasts of continuous outcomes, in which context we can often identify the action $x$ with a functional (i.e., a numeric summary such as a mean or a quantile) of the forecast distribution $F$.
+In this context, $s$ may be used as a \emph{scoring function}.
+\elr{Consider moving preceding sentences to a footnote or just deleting them?}
 This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss function.
@@
 \subsection{A review of quantile loss, CRPS, and the weighted interval score}
+\label{sec:methods.quantileloss}
+
+\subsubsection{Quantile loss}
 
@@
 \subsection{The allocation score}
+\label{sec:methods.allocation}
 

### 2023-05-01 alloscore_manuscript/allocation.tex
attempt to describe quantile loss, CRPS, and WIS

@@
 
-\subsubsection{Quantile loss}
+We review here how quantile loss arises from a particular decision-making problem, and how CRPS can be obtained by integrating across values of the parameters of that decision-making problem.
+These results have been thoroughly discussed in the literature [cite cite cite].
 
-I haven't written anything here.  Fill it in, or delete this section?
+Suppose that a decision-maker is tasked with determining the quantity $x$ of a protective measure to procure; for example, $x$ might represent the number of hospital beds or amount of medicine to purchase.
+Additionally, suppose that each unit of this good has cost $C$ so that the total cost of procurement is $Cx$.
+The variable $y$ denotes the eventual realized need for this resource, e.g.
+the number of patients in need of a hospital bed or the amount of medication that is needed.
+We assume that each unit of unmet need incurs a loss denoted by $L$, so that if the selected procurement level $x$ is less than the realized need $y$, a loss of $L(x-y)$ results.
+With this formalization of a decision-making task, we can proceed to develop a proper scoring rule using the procedure outlined in section \ref{sec:methods.decisiontheory}.
 
-%Set up of decision-making problem with costs for over-/under-prediction. Quantile/pinball loss.
+\paragraph{Step 1: specify a loss function.} Combining the cost of procuring goods at level $x$ with losses due to unmet need, we arrive at the overall loss function
+\begin{align}
+s_Q(x,y; C, L) = Cx + L(x-y)_-.
+\end{align}
+Here, $(x - y)_- := \max(-(x - y),0)$ is $0$ if the amount procured, $x$, is greater than or equal to the realized demand $y$; otherwise, it is $y - x$, the amount of unmet need.
 
-%Statement that this loss elicits quantiles as the Bayes act, derivations deferred to a supplement.
+\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} It can be shown that under the loss function $s_Q$, the Bayes act is a quantile of the forecast distribution at the probability level $\alpha = 1 - C/L$:
+\begin{align}
+x^F = F^{-1}(\alpha).
+\end{align}
+See the supplement for a verification of this result.
+%In particular, faced with the classical binary decision problem of whether to recommend an additional unit of protection given a current level of protection $x$, the forecaster's optimal decision rule under $s_m$ is to recommend adding protection if $x < F^{-1}\left(1-\frac{C}{L}\right)$, that is, if
+%\begin{align}
+%1-F(x) = \mathbb{P}_F\{y>x\} > \frac{C}{L},
+%\end{align}
+%the \emph{cost-loss ratio} of the problem.
 
-%Interpretation of CRPS as an integral of quantile loss over quantile levels, WIS as a discrete approximation of CRPS.
+\paragraph{Step 3: Define the scoring rule.} Following the procedures outlined above, we could score the forecast distribution $F$ with the scoring rule
+\begin{align}
+S_Q(F, y; C, L) &= s_Q(x^F, y; C, L) = Cx^F + L(x^F-y)_- \\
+&= C F^{-1}(\alpha) + L \left(F^{-1}(\alpha) - y\right)_-
+\end{align}
+
+This scoring rule evaluates the forecast distribution $F$ only through its $\alpha$ quantile.
+While this is faithful to the context of the decision-making problem, it may not be satisfying as a measure of the quality of the full forecast distribution.
+For this purpose, one option is to integrate the quantile loss across different values of the quantile level $\alpha$, or, equivalently, across different values of the decision problem's parameters $C$ and $L$.
+This gives rise to the (weighted) CRPS:
+\begin{align*}
+    S_{CRPS}(F, y) &= \int S_Q(F, y; C, L) h(C, L) \, dC dL
+\end{align*}
+The usual CRPS results from taking a distribution $h$ such that the induced distribution on the probability level $\alpha = 1 - C/L$ is $\text{Uniform}(0,1)$.
+The weighted interval score (WIS) can be used when the full forecast distribution $F$ is not available, and amounts to an approximation of CRPS where the distribution $h$ is discrete with point masses at the quantile levels corresponding to the endpoints of a finite set of prediction intervals.
+
+% The below was an unsuccessful attempt to connect to a more usual notation for CRPS just in terms of integrating across quantile levels.
+It resulted in a notational quagmire, but I'm keeping it here for now in case it seems like a better way to go...
+%
+%To make the connection to the usual notation, we can divide the loss function $s_Q$ by $L$ to obtain an expression in terms of only $\alpha$.
+%\begin{align*}
+%s_Q(x,y; \alpha) &= s_Q(x,y; C, L)/L \\
+%&= (C/L)x + (x-y)_- \\
+%&= (1 - \alpha)x + (x-y)_-
+%\end{align*}
+%Because these loss functions are equal up to a constant of proportionality, the Bayes act is the same for both.
+We can how obtain a weighted CRPS by integrating the quantile score across values of $C$ and $L$ or across values of $\alpha$:
+%\begin{align*}
+%S_{CRPS}(F, y) &= \int S_Q(F, y; C, L) h(C, L) \, dC dL %\\
+%&= \int S_Q(F, y; \alpha) h(\alpha) \, d\alpha
+%\end{align*}
+%The usual CRPS results from taking $h$ such that the distribution on the probability level $\alpha$ is $\text{Uniform}(0,1)$.
 

### 2023-05-01 alloscore_manuscript/allocation.tex
update section on allocation score to have better flow after new quantile score section

@@
 We assume that each unit of unmet need incurs a loss denoted by $L$, so that if the selected procurement level $x$ is less than the realized need $y$, a loss of $L(x-y)$ results.
-With this formalization of a decision-making task, we can proceed to develop a proper scoring rule using the procedure outlined in section \ref{sec:methods.decisiontheory}.
+At the time that a decision-maker determines the amount $x$ to procure, the demand $y$ is not yet known.
+We therefore define the random variable $Y$ that represents the as-yet-unknown level of demand.
+The forecast $F$ specifies a predictive distribution for $Y$.
+Here we identify $F$ with its cumulative distribution function (CDF), and $F^{-1}$ denotes the quantile function.
+With this formalization of the decision-making task, we can proceed to develop a proper scoring rule using the procedure outlined in section \ref{sec:methods.decisiontheory}.
 
@@
 An administrator is tasked with determining where to send these resources so as to meet demand among hospital patients in different facilities or states.
+In contrast to the decision-making problem in the previous section, the administrator is not able to control the total amount of supply; rather, their task is to determine how to allocate the fixed supply to different locations.
 
-We first specialize our notation to this decision-making setting.
-We define an action $\mathbf{x} = (x_1, \ldots, x_n)$ as a vector specifying the amount that is allocated to each of the $n$ locations.
-We require that each $x_i$ is non-negative, and that the total allocation across all locations does not exceed a constraint $K$ on the total available resources: $\sum_{i=1}^n x_i \leq K$.
-The set $\mathcal{X}$ collects all possible allocations that satisfy these constraints.
+In this decision-making setting, an action $\mathbf{x} = (x_1, \ldots, x_n)$ is a vector specifying the amount that is allocated to each of the $n$ locations.
+We require that each $x_i$ is non-negative and that the total allocation across all locations does not exceed a constraint $K$ on the total available resources: $\sum_{i=1}^n x_i \leq K$.
+\elr{Since we got rid of the C parameter, maybe we should just make this a hard constraint here, $\sum_i x_i = K$?} The set $\mathcal{X}$ collects all possible allocations that satisfy these constraints.
 The eventually realized resource demand in each location is denoted by $\mathbf{y} = (y_1, \ldots, y_n)$.
-At the time that a decision-maker sets the resource allocation, the demand $\mathbf{y}$ is not yet known.
-We therefore define the random vector $Y = (Y_1, \ldots, Y_n)$ where $Y_i$ represents the as-yet-unknown level of resource demand in location $i$.
-The forecast $F = (F_1, \ldots, F_n)$ collects forecasts of demand in each location.
-Here we identify $F_i$ with its cumulative distribution function (CDF), and $F_i^{-1}$ denotes the quantile function.
+Again, these levels of demand are not known at the time of decision making, so we define the random vector $Y = (Y_1, \ldots, Y_n)$ where $Y_i$ represents the as-yet-unknown level of resource demand in location $i$.
+Forecasts of demand in each location are collected in $F = (F_1, \ldots, F_n)$.
 We assume that the forecasts do not allow for the possibility of negative demand, i.e.
 the support of each $F_i$ is a subset of $\mathbb{R}^+$.
-With this notation in place, we can proceed to develop a proper scoring rule following the outline in the previous section.
+As in the previous section, we assume that a loss $L$ is incurred for each unit of unmet need.
 
-\paragraph{Step 1: specify a loss function.} The loss associated with a particular allocation is calculated by comparing the amount allocated to each location to the realized resource demand in that location.
-Specifically, suppose that there is a marginal cost $L$ that accrues for each unit of demand that is not met.
-We can calculate the allocation loss across all locations as
-\begin{equation}
-s(\mathbf{x}, \mathbf{y}) = \sum_{i=1}^n L (x_i - y_i)_-, \label{eqn:loss_fn}
-\end{equation}
-where $(x_i - y_i)_- := \max(-(x_i - y_i),0)$ is $0$ if the amount $x_i$ allocated to unit $i$ is greater than or equal to the realized demand $y_i$ in that location; otherwise, it is $y_i - x_i$, the amount of unmet need in that location.
 We note that a number of generalizations to this loss specification have been formulated in the literature, including an allowance for costs for over-allocation to a particular unit (e.g.
@@
 if a unit of unmet demand imposes more severe costs in one location than another), and the introduction of a convex function that controls the rate at which costs accrue depending on the scale of need.
-We consider these and other generalizations in other work, but for the present exposition we restrict our attention to the relatively simple loss formulation of Equation \eqref{eqn:loss_fn}.
+We consider these and other generalizations in other work.
+
+With this notation in place, we can develop a proper scoring rule following the outline in section \ref{sec:methods.decisiontheory}.
+
+\paragraph{Step 1: specify a loss function.} The loss associated with a particular allocation is calculated by summing contributions from unmet demand in each location:
+\begin{equation}
+s_A(\mathbf{x}, \mathbf{y}; L) = \sum_{i=1}^n L (x_i - y_i)_-. \label{eqn:loss_fn}
+\end{equation}
+As in the previous section, $(x_i - y_i)_-$ is $0$ if the amount $x_i$ allocated to unit $i$ is greater than or equal to the realized demand $y_i$ in that location; otherwise, it is $y_i - x_i$, the amount of unmet need in that location.
 
@@
 \begin{align}
-    \mathbf{x}^F &= \underset{\mathbf{x}\in \mathbb{R}^N, 0 \leq \mathbf{x}}{\mathrm{arg min}}\,\, \sbar_F(\mathbf{x}) \text{ subject to } 
+    \mathbf{x}^F &= \underset{\mathbf{x}\in \mathbb{R}^N, 0 \leq \mathbf{x}}{\mathrm{arg min}}\,\, \sbar_A^F(\mathbf{x}; L) \text{ subject to } 
     \sum_{i=1}^N x_i \leq K \text{, where} \label{AP} \\
-    \sbar_F(\mathbf{x}) &= \mathbb{E}_{F} s(\mathbf{x}, \mathbf{Y}) = \sum_{i=1}^{N} \mathbb{E}_{F_i}[s(x_i, Y_i)] \label{eqn:sbar}
+    \sbar_A^F(\mathbf{x}; L) &= \mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y}; L)] = \sum_{i=1}^{N} \mathbb{E}_{F_i}[s_A(x_i, Y_i; L)] \label{eqn:sbar}
 \end{align}
@@
 \end{align}
-This partial solution to the allocation problem seems to have first appeared in [@hadleywhitin1963]; see the supplemental materials for a derivation.
+This partial solution to the allocation problem seems to have first appeared in \cite{hadleywhitin1963}; see the supplemental materials for a derivation.
 
 One interpretation of this result is that the Bayes act sets the allocation in each location $i$ to a quantile of the forecast distribution $F_i$ for that location.
-The quantile is at a probability level $(1 - \lambda^{\star}/L)$ that is the same for all locations, and is chosen such that the constraint is satisfied.
-An alternative interpretation comes from noting that for each location $i$, $\frac{\partial}{\partial x_i} \sbar_F(\mathbf{x}) \vert_{\mathbf{x}=\mathbf{x}^F} = \lambda^{\star}$ (see the supplement for a proof).
+The quantile is at a probability level $\alpha = 1 - \lambda^{\star}/L$ that is the same for all locations, and is chosen such that the constraint is satisfied.
+An alternative interpretation comes from noting that for each location $i$, $\frac{\partial}{\partial x_i} \sbar_A^F(\mathbf{x}) \vert_{\mathbf{x}=\mathbf{x}^F} = \lambda^{\star}$ (see the supplement for a proof).
 %$\frac{\partial}{\partial x_i} \sbar_F(\mathbf{x}) = L[F_i(x_i) - 1]$ (see the supplement for a derivation), and evaluating at the Bayes allocation we arrive at

### 2023-05-01 alloscore_manuscript/allocation.tex
some stuff about link between integrated allocation score and crps

@@
 \def\elr#1{{\color{cyan}\textbf{ELR:[#1]}}}
+\def\apg#1{{\color{red}\textbf{APG:[#1]}}}
 
@@
 \begin{align*}
-    S_{CRPS}(F, y) &= \int S_Q(F, y; C, L) h(C, L) \, dC dL
+    S_{CRPS}(F, y) &= \int S_Q(F, y; C, L) p(C, L) \, dC dL
 \end{align*}
-The usual CRPS results from taking a distribution $h$ such that the induced distribution on the probability level $\alpha = 1 - C/L$ is $\text{Uniform}(0,1)$.
-The weighted interval score (WIS) can be used when the full forecast distribution $F$ is not available, and amounts to an approximation of CRPS where the distribution $h$ is discrete with point masses at the quantile levels corresponding to the endpoints of a finite set of prediction intervals.
+The usual CRPS results from taking $p$ to be the density of a distribution such that the induced distribution on the probability level $\alpha = 1 - C/L$ is $\text{Uniform}(0,1)$.
+The weighted interval score (WIS) can be used when the full forecast distribution $F$ is not available, and amounts to an approximation of CRPS where the distribution $p$ is discrete with point masses at the quantile levels corresponding to the endpoints of a finite set of prediction intervals.
+
+\elr{maybe better to use different notation for this integral that's less specific about whether $p$ is discrete or continuous?}
 
@@
 \end{align}
-It can be shown that with the loss function given in Equation \eqref{eqn:loss_fn}, the Bayes act has $x^F_i = F_i^{-1}(1 - \lambda^{\star}/L)$, where $\lambda^{\star}$ is chosen so as to satisfy the equation
+It can be shown that with the loss function given in Equation \eqref{eqn:loss_fn}, the Bayes act has $x^F_i = F_i^{-1}(1 - \lambda^{\star}/L)$, where $\lambda^{\star}$ depends on the problem parameters $K$ and $L$ as well as the forecast distributions and is chosen so as to satisfy the equation
 \begin{align}
@@
 This partial solution to the allocation problem seems to have first appeared in \cite{hadleywhitin1963}; see the supplemental materials for a derivation.
+Figure *** panel (a) \elr{TODO: first figure from notes crps\_connection\_exp.Rmd} illustrates the expected loss function $\sbar_A^F$ and the allocation given by the Bayes act in a simple example with $n = 2$ locations, $L = 1$, $K = 5$, and forecasts $Y_1 \sim \text{Exp}(1 / \sigma_1)$ and $Y_2 \sim \text{Exp}(1 / \sigma_2)$, using a scale parameterization of the exponential distribution with $\sigma_1 = 1$ and $\sigma_2 = 5$.
 
@@
 
-\textbf{1st attempt by APG to rephrase last paragraph}
-
-The central mathematical ideas in this construction are that 
+\apg{
+first attempt by APG to rephrase last paragraph: The central mathematical ideas in this construction are that 
 \begin{itemize}
@@
 (See the supplement for detailed discussion and derivations of these points.)
+}
 
@@
 \begin{equation}
-S(F, y) = s(\mathbf{x}^F, y) = \sum_{i=1}^n L (F_i^{-1}(1 - \lambda^{\star}/L) - y_i)_-
+S_A(F, y; L, K) = s_A(\mathbf{x}^F, y; L) = \sum_{i=1}^n L (F_i^{-1}(1 - \lambda^{\star}/L) - y_i)_-
 \end{equation}
@@
 
-\textbf{ELR: This section would likely benefit from a figure or two to illustrate the ideas...?}
+As with the quantile score $S_Q$, the allocation score $S_A$ measures the skill of the forecast distributions $F$ based on a single probability level $\alpha$.
+By analogy to the method for obtaining CRPS by integrating the quantile score, we develop an \emph{integrated allocation score} (IAS) \elr{open to better names} that integrates the allocation score across values of the problem parameters, weighting by a distribution $p$:
+$$S_{IAS}(F, y) = \int S_A(F,y; L, K) p(L, K) \, dL dK$$
 
-Description of what happens if there is uncertainty about the the constraint $K$.
-Integration across $K$ results in something like a forecaster-specific weighted CRPS.
-It might help to introduce notation like $\lambda^{\star}(K)$ or $S_K(F, y)$ throughout all of the above discussion?
-Then if we have a prior $p(K)$ on $K$, we get to something like
-$$S(F, y) = \int S_K(F,y) p(K) \, d K$$
-as our final score.
+We illustrate the relationship between the IAS and the CRPS in our example with two locations and forecasts given by $\text{Exp}(1/\sigma_i)$ distributions with $\sigma_1 = 1$ and $\sigma_2 = 5$.
+Note that the quantile functions corresponding to these forecasts are given by $F_i^{-1}(\alpha) = -\sigma_i \log(1 - \alpha)$.
+For simplicity, we keep $L = 1$ fixed (i.e., $p$ places probability 1 on $L = 1$), and only address varying $K$.
+As discussed above, each value of the constraint $K$ determines a quantile probability level $\alpha$ corresponding to the Bayes act such that $K = F_1^{-1}(\alpha) + F_2^{-1}(\alpha) = -\log(1 - \alpha) (\sigma_1 + \sigma_2)$; solving for $\alpha$, we obtain $\alpha = 1 - \exp[-K/(\sigma_1 + \sigma_2)]$.
+This link between the constraint level $K$ and the probability level $\alpha$ is the key to the link between the IAS and the CRPS.
+
+We use this link to explore the relationship between IAS and CRPS from two directions.
+First, suppose the decision-maker has some uncertainty about the value of $K$, which they express through $p$.
+Because $\alpha$ can be regarded as a function of $K$, this distribution on the resource constraint induces a distribution on quantile levels.
+Figure *** panel (b) \elr{TODO, figures from crps\_connection\_exp} illustrates with a $\text{Gamma}(500, 0.01)$ distribution for $K$, and the implied distribution on quantile levels is shown in panel (c).
+The IAS determined by $p$ corresponds to a weighted CRPS with this induced weighting on quantile levels.
+However, note that this weighting is specific to this pair of Exponential forecasts; a different pair of forecasts would translate to a different weighting on quantile levels.
+
+Figure *** panel (d) \elr{TODO, figures from crps\_connection\_exp} illustrates this link by going in the other direction: given a forecast $F$, we exhibit the distribution on $K$ that would lead to equally weighted CRPS.
+Now we use the fact that $K$ can be written as a function of $\alpha$ to obtain the distribution on $K$ that corresponds to a $\text{Uniform}(0,1)$ distribution on $\alpha$.
+In this example, the implied distribution is $K \sim \text{Exp}(\sigma_1 + \sigma_2)$.
+We observe that this is a right-skewed distribution that places much of its mass on constraint values near 0, which may not correspond well to actual knowledge about the resource constraints.
+Again, the distribution on resource constraints that corresponds to unweighted CRPS depends on the forecast distributions.
 

### 2023-05-01 alloscore_manuscript/allocation.tex
small comment

@@
 \end{align*}
+\apg{Quick note that I don't think this will work without some more identifications.
+The integral needs to be over probability levels and thresholds and without the some discussion like the one commented out (currently) at line 123 or (better in my mind) one like I'm trying to do with binary choices at the unit-level, I don't think we have a threshold.}
 The usual CRPS results from taking $p$ to be the density of a distribution such that the induced distribution on the probability level $\alpha = 1 - C/L$ is $\text{Uniform}(0,1)$.

### 2023-05-01 notes/allocation_notes.rmd
two very small but potentially crucial additions
small comment

@@
 
-Or if $F=\Phi_{\mu,\sigma}$ is normal with location $\mu = E_F[Y]$ and scale $\sigma = \sqrt{\mathrm{Var}(Y)}$, 
+Or if $F=\Phi_{\mu,\sigma}$ is normal with location $\mu = E_F[Y]$ and scale $\sigma = \sqrt{\mathrm{Var}_F(Y)}$, 
 we have 
@@
 \end{align}
+
 In particular, faced with the classical binary decision problem of whether to recommend an additional unit of protection given a current level of protection $x$, the forecaster's optimal decision rule under $s_m$ is to recommend adding protection if $x < F^{-1}\left(1-\frac{C}{L}\right)$, that is, if

### 2023-05-08 alloscore_manuscript/allocation.tex
update section on quantile score and crps

--- 2023-05-01 alloscore_manuscript/allocation.tex
@@
 
-\subsection{A review of quantile loss, CRPS, and the weighted interval score}
+\subsection{A review of quantile score, CRPS, and the weighted interval score}
 \label{sec:methods.quantileloss}
 
-We review here how quantile loss arises from a particular decision-making problem, and how CRPS can be obtained by integrating across values of the parameters of that decision-making problem.
+We review how the quantile score arises from a particular decision-making problem in section \ref{sec:methods.quantileloss.quantile}, and how CRPS can be obtained by integrating across values of the parameters of that decision-making problem, as well as the connection to WIS, in section \ref{sec:methods.quantileloss.crps}.
 These results have been thoroughly discussed in the literature [cite cite cite].
+
+\subsubsection{Decision-theoretic origins of the quantile score}
+\label{sec:methods.quantileloss.quantile}
 
@@
 \paragraph{Step 3: Define the scoring rule.} Following the procedures outlined above, we could score the forecast distribution $F$ with the scoring rule
-\begin{align}
-S_Q(F, y; C, L) &= s_Q(x^F, y; C, L) = Cx^F + L(x^F-y)_- \\
-&= C F^{-1}(\alpha) + L \left(F^{-1}(\alpha) - y\right)_-
-\end{align}
+\begin{align*}
+S_Q(F, y; C, L) &= s_Q(x^F, y; C, L) = Cx^F + L(x^F-y)_- \nonumber \\
+&= C F^{-1}(\alpha) + L \left(F^{-1}(\alpha) - y\right)_- \label{eqn:S_Q_CL}
+\end{align*}
 
-This scoring rule evaluates the forecast distribution $F$ only through its $\alpha$ quantile.
+We have set up the problem here in terms of the cost and loss parameters $C$ and $L$, which has the benefit of an intuitive connection to the decision-making context.
+However, to clarify the connection to the usual notation for the quantile loss, we can divide the loss function $s_Q$ by $L$ to obtain an expression in terms of only $\alpha$:
+\begin{align*}
+s_Q(x,y; \alpha) &= s_Q(x,y; C, L)/L \\
+&= (C/L)x + (x-y)_- \\
+&= (1 - \alpha)x + (x-y)_-.
+\end{align*}
+Because these loss functions are equal up to a constant of proportionality, the Bayes act is the same for both. The associated quantile scoring rule expressed in terms of $\alpha$ is
+\begin{equation}
+S_Q(F, y; \alpha) = (1 - \alpha)F^{-1}(\alpha) + \left(F^{-1}(\alpha) - y\right)_-. \label{eqn:S_Q_alpha}
+\end{equation}
+In either formulation, the key observation is that the Bayes act is the quantile of the forecast distribution $F$ at the probability level given by one minus the cost/loss ratio $C/L$.
+
+\subsubsection{CRPS as an integrated quantile score}
+\label{sec:methods.quantileloss.crps}
+
+The scoring rule $S_Q$ of Equation \eqref{eqn:S_Q_alpha} evaluates the forecast distribution $F$ only through its $\alpha$ quantile.
 While this is faithful to the context of the decision-making problem, it may not be satisfying as a measure of the quality of the full forecast distribution.
-For this purpose, one option is to integrate the quantile loss across different values of the quantile level $\alpha$, or, equivalently, across different values of the decision problem's parameters $C$ and $L$.
-This gives rise to the (weighted) CRPS:
+For this purpose, one option is to integrate the quantile scoring rule across different values of the probability level $\alpha$, weighting the probability levels according to a specified distribution $p$.
+This yields a weighted CRPS:
 \begin{align*}
-    S_{CRPS}(F, y) &= \int S_Q(F, y; C, L) p(C, L) \, dC dL
+S_{CRPS}(F, y; p) &= \int S_Q(F, y; \alpha) p(\alpha) \, d\alpha.
 \end{align*}
-The usual CRPS results from taking $p$ to be the density of a distribution such that the induced distribution on the probability level $\alpha = 1 - C/L$ is $\text{Uniform}(0,1)$.
-The weighted interval score (WIS) can be used when the full forecast distribution $F$ is not available, and amounts to an approximation of CRPS where the distribution $p$ is discrete with point masses at the quantile levels corresponding to the endpoints of a finite set of prediction intervals.
+This weighted form of CRPS has appeared in the literature before, e.g.
+see \textcite{gneiting2011weightedScoringRules}.
+The usual CRPS results from taking $p$ corresponding to a $\text{Uniform}(0,1)$ distribution, equally weighting all probability levels.
 
-\elr{maybe better to use different notation for this integral that's less specific about whether $p$ is discrete or continuous?}
+We emphasize that because $\alpha = 1 - C/L$, the distribution $p(\alpha)$ can be interpreted as expressing incomplete knowledge about the cost/loss ratio in the decision-making problem.
+The equal weighting used by the ordinary CRPS may be appropriate in the absence of any knowledge about the context in which forecasts will be used to support decision-making, but may be inappropriate if more information is known about the cost/loss ratio for a specific decision-making task.
 
-% The below was an unsuccessful attempt to connect to a more usual notation for CRPS just in terms of integrating across quantile levels.
-It resulted in a notational quagmire, but I'm keeping it here for now in case it seems like a better way to go...
-%
-%To make the connection to the usual notation, we can divide the loss function $s_Q$ by $L$ to obtain an expression in terms of only $\alpha$.
-%\begin{align*}
-%s_Q(x,y; \alpha) &= s_Q(x,y; C, L)/L \\
-%&= (C/L)x + (x-y)_- \\
-%&= (1 - \alpha)x + (x-y)_-
-%\end{align*}
-%Because these loss functions are equal up to a constant of proportionality, the Bayes act is the same for both.
-We can how obtain a weighted CRPS by integrating the quantile score across values of $C$ and $L$ or across values of $\alpha$:
-%\begin{align*}
-%S_{CRPS}(F, y) &= \int S_Q(F, y; C, L) h(C, L) \, dC dL %\\
-%&= \int S_Q(F, y; \alpha) h(\alpha) \, d\alpha
-%\end{align*}
-%The usual CRPS results from taking $h$ such that the distribution on the probability level $\alpha$ is $\text{Uniform}(0,1)$.
+The weighted interval score (WIS) is often used when the full forecast distribution $F$ is not available, as in the U.S.
+Forecast Hub and similar efforts where forecasts are represented by a collection of prediction intervals.
+WIS is a discrete approximation to CRPS, and can be obtained by using a distribution $p$ that has point masses at the probability levels corresponding to the endpoints of a finite set of prediction intervals.
+
+%\elr{Leaving the comment below in the tex, but noting that it was in response to an older version of the write up.
+I've since attempted to address it by moving to a discussion in terms of a version of the scoring function where we've divided by L to work in terms of the cost/loss ratio.}
+%\apg{Quick note that I don't think this will work without some more identifications.
+The integral needs to be over probability levels and thresholds and without the some discussion like the one commented out (currently) at line 123 or (better in my mind) one like I'm trying to do with binary choices at the unit-level, I don't think we have a threshold.}
+%The usual CRPS results from taking $p$ to be the density of a distribution such that the induced distribution on the probability level $\alpha = 1 - C/L$ is $\text{Uniform}(0,1)$.
 
@@
 \label{sec:methods.allocation}
+
+\elr{If we like the organization of the previous subsection into two subsubsections about the quantile score and the CRPS, we should replicate that here.}
 

### 2023-05-15 alloscore_manuscript/allocation.tex
adding update for intro paragaph 4. latex still not working quite right for me tho.

@@
 
-In practice, probablistic infectious disease forecasts have most often been evaluated with standard, off-the-shelf scoring rules such as the log score, continuous ranked probability score (CRPS), or weighted interval score (WIS).
-[cite some examples] While some of these scores can be interpreted through the lens of decision theory [thinking here of WIS/quantile loss], these connections are not a common focus of infectious disease forecast evaluation.
+In practice, probablistic infectious disease forecasts have most often been evaluated with standard, ``off-the-shelf'' scoring rules.
+For example, seasonal influenza forecasts in the US and dengue forecasts for Peru and Puerto Rico used log-score to evaluate forecast skill \cite{mcgowan_collaborative_2019,reich_collaborative_2019,johansson_open_2019}.
+Pandemic COVID-19 forecasts in the US and Europe were evaluated using the weighted interval score (WIS), an approximation of the continuous ranked probability score (CRPS), and prediction interval coverage\cite{cramer_evaluation_2022}.
+Similarly, CRPS was also used to assess probabilistic forecasts of dengue outbreaks in Vietnam\cite{colon-gonzalez_probabilistic_2021}.
+While some of these scores can be interpreted through the lens of decision theory, and all of the application-specific papers cited above had authors from public health agencies, none of them make explicit connections between a forecast evaluation and how a forecast was used in practice.
+In general, explicit links between how forecasts are evaluated and how they are scored are not a common focus of infectious disease forecast evaluations.
 

### 2023-05-17 _targets.R
adding a multi-date example.

@@ -2,9 +2,13 @@
 
 # Load packages required to define the pipeline:
 library(targets)
+library(tarchetypes)
+library(tibble)
+library(crew)
 
 # Set target options:
 tar_option_set(
+  controller = crew_controller_local(workers = 4),
   packages = c("tidyverse", "covidHubUtils", "distfromq", "alloscore", "gh"), # packages that your targets need to run
   format = "rds" # default storage format
 )
@@ -14,17 +18,17 @@ tar_option_set(
 # remotes::install_github("reichlab/distfromq")
 # remotes::install_github("reichlab/covidHubUtils")
 
-# tar_make_clustermq() configuration (okay to leave alone):
-options(clustermq.scheduler = "multicore")
-
 # Run the R scripts in the R/ folder:
 tar_source()
+
+## create a group of alloscore targets
+values <- tibble(forecast_dates = as.character(seq.Date(as.Date("2021-12-13"), as.Date("2021-12-27"), by = "7 days")))
 
 # List of targets:
 list(
   tar_target(
     name = forecast_data,
-    command = get_forecast_data("2021-12-27")
+    command = get_forecast_data(values$forecast_dates)
   ),
   tar_target(
     name = truth_data,
@@ -34,13 +38,15 @@ list(
     name = score_data,
     command = get_forecast_scores(forecast_data, truth_data)
   ),
-  tar_target(
-    name = alloscores,
-    command = run_alloscore(forecast_data, truth_data)
+  tar_map(
+    values = values,
+    tar_target(alloscore, run_alloscore(forecast_data, truth_data, forecast_dates))
   ),
   tar_target(
-    name = figure_K_v_alloscore,
-    command = plot_K_v_alloscore(alloscores),
+    name = figure_K_v_alloscore_2021.12.13,
+    command = plot_K_v_alloscore(alloscore_2021.12.13),
     format = "file"
   )
+
+
 )

### 2023-05-17 R/run-alloscore.R
adding a multi-date example.

@@ -1,4 +1,4 @@
-run_alloscore <- function(forecast_data, truth_data){
+run_alloscore <- function(forecast_data, truth_data, one_forecast_date){
   require(tidyverse)
   require(alloscore)
   require(distfromq)
@@ -17,6 +17,8 @@ run_alloscore <- function(forecast_data, truth_data){
 
   ## process forecast data, adding distfromq output
   forecast_data_processed <- forecast_data |>
+    ## forecast dates are different but reference dates are Mondays
+    dplyr::filter(reference_date == one_forecast_date) |>
     dplyr::select(-type) |>
     nest(ps = quantile, qs = value) |>
     relocate(ps, qs) |>

### 2023-05-17 _targets.R
adding more robust example, including a failure at 2022-01-31 date.

@@ -8,7 +8,7 @@ library(crew)
 
 # Set target options:
 tar_option_set(
-  controller = crew_controller_local(workers = 4),
+  controller = crew_controller_local(workers = 8),
   packages = c("tidyverse", "covidHubUtils", "distfromq", "alloscore", "gh"), # packages that your targets need to run
   format = "rds" # default storage format
 )
@@ -22,7 +22,7 @@ tar_option_set(
 tar_source()
 
 ## create a group of alloscore targets
-values <- tibble(forecast_dates = as.character(seq.Date(as.Date("2021-12-13"), as.Date("2021-12-27"), by = "7 days")))
+values <- tibble(forecast_dates = as.character(seq.Date(as.Date("2021-11-22"), as.Date("2022-02-28"), by = "7 days")))
 
 # List of targets:
 list(

### 2023-05-18 random_undocumented_experiments/debugging5_17.R
debugging script

@@ -1,0 +1,50 @@
+library(targets)
+tar_load(forecast_data)
+
+## create a group of alloscore targets
+values <- tibble(forecast_dates = as.character(seq.Date(as.Date("2021-11-22"), as.Date("2022-02-28"), by = "7 days")))
+one_forecast_date <- values$forecast_dates[11]
+
+forecast_data_processed <- forecast_data |>
+  ## forecast dates are different but reference dates are Mondays
+  dplyr::filter(reference_date == one_forecast_date) |>
+  dplyr::select(-type) |>
+  nest(ps = quantile, qs = value) |>
+  relocate(ps, qs) |>
+  mutate(
+    ps = map(ps, deframe),
+    qs = map(qs, deframe)
+  ) |>
+  dplyr::filter(model %in% mkeep) %>%
+  dplyr::mutate(
+    model = ifelse(model == "COVIDhub-4_week_ensemble", "COVIDhub-ensemble", model)
+  ) |>
+  add_pdqr_funs(dist = "distfromq", types = c("p", "q")) |>
+  relocate(dist, F, Q) |>
+  left_join(
+    truth_data |> select(location, target_end_date, value),
+    by = c("location", "target_end_date"))
+
+ascores <- forecast_data_processed %>% filter(model == "BPagano-RtDriven")
+with(ascores[5:50,] ,
+      allocate(
+  F = F,
+  Q = Q,
+  w = 1,
+  K = 5000,
+  kappa = 1,
+  alpha = 1,
+  dg = 1,
+  eps_K = .01,
+  eps_lam = 1e-5,
+  verbose = TRUE
+))
+
+nrow(ascores)
+ggplot() + map(2:5, function(i) {
+  geom_function(fun = ascores$F[[i]], aes(color = ascores$abbreviation[i]), linewidth =2)
+  }) +
+  map(c(1,6:51), function(i) geom_function(fun = ascores$F[[i]], alpha = .5)) + lims(x=c(0,500)) +
+  labs(color = "problem states", y = "distfromq F")
+
+

### 2023-05-18 alloscore_manuscript/allocation.tex
adding updated manuscript file, including revisions to the method used for bibliography.

--- 2023-05-08 alloscore_manuscript/allocation.tex
@@
 
-\usepackage{biblatex}
-\addbibresource{allocation.bib}
+\usepackage{setspace}
+\onehalfspacing
 
@@
 
-High level points to cover in introduction:
-\begin{itemize}
-\item People are using infectious disease forecasts as an input to decision making
-\item There are standard ways to evaluate forecasts that are responsive to decision making context, and use of those methods is relatively common in other fields like economics and meteorology
-\item However, there's not much work in infectious disease that does this
-\item In practice, infectious disease forecasts have typically been evaluated with "off the shelf" scoring rules such as the WIS which is an approximation to CRPS, log score, and so on.
-\item In this work, our goal is to begin to address that gap. We focus on a resource allocation setting.
-\item There is past work focusing on resource allocation in the operations research literature, but it doesn't take the step of getting to a measure of forecast skill.
-\item That is, while there is plenty of work on 
-\begin{itemize}
-\item DM under risk, where probabilities are taken as known and expected utility is maximized
-\item DM under uncertainty, where probabilities are taken as unknown and robust (i.e. maximin) decisions are sought
-\end{itemize}
-little seems to have been said on how forecasts can be evaluated by how well they convert a problem of DM under uncertainty to a problem of DM under risk.
-\end{itemize}
+%High level points to cover in introduction:
+%\begin{itemize}
+%\item People are using infectious disease forecasts as an input to decision making
+%\item There are standard ways to evaluate forecasts that are responsive to decision making context, and use of those methods is relatively common in other fields like economics and meteorology
+%\item However, there's not much work in infectious disease that does this
+%\item In practice, infectious disease forecasts have typically been evaluated with "off the shelf" scoring rules such as the WIS which is an approximation to CRPS, log score, and so on.
+%\item In this work, our goal is to begin to address that gap. We focus on a resource allocation setting.
+%\item There is past work focusing on resource allocation in the operations research literature, but it doesn't take the step of getting to a measure of forecast skill.
+%\item That is, while there is plenty of work on
+%\begin{itemize}
+%\item DM under risk, where probabilities are taken as known and expected utility is maximized
+%\item DM under uncertainty, where probabilities are taken as unknown and robust (i.e. maximin) decisions are sought
+%\end{itemize}
+%little seems to have been said on how forecasts can be evaluated by how well they convert a problem of DM under uncertainty to a problem of DM under risk.
+%\end{itemize}
 
@@
 However, we are aware of only a limited body of work that explicitly attempts to measure the value of infectious disease forecasts through their impact on policy, and much of this discussion has proceeded informally.
-For example, \cite{ioannidis2022forecastingCOVIDfailed} discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts.
+For example, Ioannidis et al.
+\cite{ioannidis2022forecastingCOVIDfailed} discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts.
 Separately, there is a thread of literature that does quantify the link between infectious disease modeling and policy making, but this work has been done outside of a forecasting context.
-As an example, \cite{Probert2016decisionMakingFootMouth} develop measures of the cost of actions designed to control a hypothetical outbreak of foot-and-mouth disease and use this framework to explore policy recommendations from a variety of simulation-based projection models.
+As an example, Probert et al.
+\cite{Probert2016decisionMakingFootMouth} develop measures of the cost of actions designed to control a hypothetical outbreak of foot-and-mouth disease and use this framework to explore policy recommendations from a variety of simulation-based projection models.
 
-In practice, probablistic infectious disease forecasts have most often been evaluated with standard, off-the-shelf scoring rules such as the log score, continuous ranked probability score (CRPS), or weighted interval score (WIS).
-[cite some examples] While some of these scores can be interpreted through the lens of decision theory [thinking here of WIS/quantile loss], these connections are not a common focus of infectious disease forecast evaluation.
+In practice, probabilistic infectious disease forecasts have most often been made for observations that emerge from public health surveillance systems and evaluated with standard, ``off-the-shelf'' scoring rules.
+For example, seasonal influenza forecasts in the US and dengue forecasts for Peru and Puerto Rico targeted public health surveillance measures of incidence over time and space, and used log-score and mean absolute errors to evaluate forecast skill \cite{mcgowan_collaborative_2019,reich_collaborative_2019,johansson_open_2019}.
+Pandemic COVID-19 forecasts of observed cases, hospitalizations and/or deaths in the US and Europe, as reported by municipal, state, or federal surveillance systems, were evaluated using the weighted interval score (WIS, which is an approximation of the continuous ranked probability score, or CRPS), and prediction interval coverage\cite{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.
+Similarly, CRPS was also used to assess probabilistic forecasts of dengue incidence at the district level in Vietnam\cite{colon-gonzalez_probabilistic_2021}.
+While some of these scores can be interpreted through the lens of decision theory, and all of the application-specific papers cited above had authors from public health agencies, none of them (including some for which we were co-authors) make explicit connections between forecast evaluation and how a forecast was used in practice.
 
-In this work, we address this gap between the ways in which infectious disease forecasts have been used to support public health policy and the ways in which they have traditionally been evaluated.
-We work with a motivating example where forecasts are used to help set the allocation of a limited quantity of medical supplies across multiple regions.
+The motivating application of this work is one where forecasts of standard public health surveillance metrics (1) are used to help determine the allocation of a limited quantity of medical supplies across multiple regions, and (2) are later evaluated based on their recommended allocations.
+The analysis presented is ``synthetic'' in that it does not correspond to an actual analysis that supported decision-making in real-time.
+However, the framework described in this paper corresponds to real-world decisions that must be made by public health administrators around the globe, and could be adapted in the future for such real-time situations.
+For example, forecasts for districts in Sierra Leone of bed demand to care for patients with Ebola was the subject of a real-time modeling study in late 2014 and early 2015\cite{camacho2015-ebola-bed}.
+And, in 2020, a model developed by an academic research group turned predictions of COVID-19 hospitalizations into estimates of ventilator usage and and shortages.
+This framework was used by the Hartford HealthCare system in Connecticut ``to align ventilator supply with projected demand at a time where the [COVID-19] pandemic was on the rise''\cite{bertsimas2021predictionsCOVID}.
+These examples illustrate the potential for forecasts to inform decisions about how to allocate limited supplies such as temporary hospital beds, ventilators, personal protective equipment, or other supplies that are known to be effective at reducing transmission or severity of disease.
 
-operations research work on constrained allocation
+As described above, there exists both a rich literature on outbreak forecast evaluation and an emerging body of work on turning forecasts into resource allocation decisions. 
+However, there has been far less attention paid (no papers that we could identify) to exploring whether standard forecast accuracy metrics are sufficient for evaluating forecasts that are used to make allocation decisions.
+In this work, we address the gap between the ways in which infectious disease forecasts have been used to support public health policy and the ways in which they have traditionally been evaluated. 
+Specifically, we define a new forecast scoring rule --the {\em allocation score}-- that explicitly evaluates forecasts based on how accurate their implied allocations would turn out to be.
+This stands in contrast to standard evaluation metrics that evaluate the accuracy of a forecast relative to the surveillance data being forecasted.
 
 The remainder of this article is organized as follows. 
-In section \ref{sec:methods}, we review the general framework for developing scoring rules for probabilistic forecasts using the tools of decision theory, develop a novel scoring rule that is motivated by the problem of allocating limited medical supplies, and explore the relationship between the proposed allocation score and existing scoring rules such as CRPS.
-We then illustrate the scoring rule through an application to short-term forecasts of COVID-19 hospital admissions in the US in section \ref{sec:application}.
+In Section \ref{sec:methods}, we review the general framework for developing scoring rules for probabilistic forecasts using the tools of decision theory, develop a novel scoring rule that is motivated by the problem of allocating limited medical supplies, and explore the relationship between the proposed allocation score and existing scoring rules such as CRPS.
+In Section \ref{sec:application} we then illustrate the scoring rule through an application to short-term forecasts of COVID-19 hospital admissions in the US. 
 Section \ref{sec:discussion} summarizes our contributions and discusses opportunities for further extensions in future work.
@@
 This weighted form of CRPS has appeared in the literature before, e.g.
-see \textcite{gneiting2011weightedScoringRules}.
+see \cite{gneiting2011weightedScoringRules}.
 The usual CRPS results from taking $p$ corresponding to a $\text{Uniform}(0,1)$ distribution, equally weighting all probability levels.
@@
 
-\printbibliography
+\bibliographystyle{unsrt}
+\bibliography{allocation.bib}
 

### 2023-05-19 notes/allocation_notes.rmd
a few lambda discussion edits

@@
 which gives lower bounds on $\sbar_F(\mathbf{x}^{\star})$ at a solution $\mathbf{x}^{\star}$ to the primal AP.
-In particular, when $O_i>0$ for all $i$, $L_D(0)$ gives the total score $\sbar_F(\mathbf{q}_{\symbfit{\alpha}, F})$ when the constraint does not prevent the forecaster from giving the quantiles in each component.
-When the constraint is active, the forecaster will generally not be able to attain a total score this low and the greatest lower bound on their score will correspond the maximizer $\lambda^{\star} = \lambda^{\star}(K)$ of \eqref{dual_eq}.
+In particular, when $O_i>0$ for all $i$, $L_D(0; K)$ gives the total score $\sbar_F(\mathbf{q}_{\symbfit{\alpha}, F})$ when the constraint does not prevent the forecaster from giving the quantiles in each component.
+When the constraint is active, the forecaster will generally not be able to attain a total score this low and the greatest lower bound on their score will
+ be $L_D(\lambda^{\star}; K)$ where $\lambda^{\star} = \lambda^{\star}(K)$ is the maximizer of \eqref{dual_eq}.
 

### 2023-05-19 random_undocumented_experiments/debugging5_17.R
CU Kansas problem

@@ -3,7 +3,7 @@ tar_load(forecast_data)
 
 ## create a group of alloscore targets
 values <- tibble(forecast_dates = as.character(seq.Date(as.Date("2021-11-22"), as.Date("2022-02-28"), by = "7 days")))
-one_forecast_date <- values$forecast_dates[11]
+one_forecast_date <- values$forecast_dates[14]
 
 forecast_data_processed <- forecast_data |>
   ## forecast dates are different but reference dates are Mondays
@@ -25,8 +25,8 @@ forecast_data_processed <- forecast_data |>
     truth_data |> select(location, target_end_date, value),
     by = c("location", "target_end_date"))
 
-ascores <- forecast_data_processed %>% filter(model == "BPagano-RtDriven")
-with(ascores[5:50,] ,
+ascores <- forecast_data_processed %>% filter(model == "CU-select")
+with(ascores,
       allocate(
   F = F,
   Q = Q,
@@ -39,6 +39,10 @@ with(ascores[5:50,] ,
   eps_lam = 1e-5,
   verbose = TRUE
 ))
+
+map2_dbl(ascores$Q[1:16], .5, exec)
+ascores$Q[[17]](.5)
+map2_dbl(ascores$Q[18:51], .5, exec)
 
 nrow(ascores)
 ggplot() + map(2:5, function(i) {

### 2023-05-19 R/data-ingestion.R
adding pass at full commit for 15 dates worth of alloscores.

@@ -33,3 +33,10 @@ get_forecast_scores <- function(forecast_data, truth_data){
   )
 }
 
+filter_problematic_data <- function(forecast_data){
+  require(tidyverse)
+  forecast_data |>
+    ## this removes two state-week-model that has a deterministically zero forecast.
+    filter(!(abbreviation=="KS" & reference_date=="2022-02-21" & model=="CU-select")) |>
+    filter(!(abbreviation=="WA" & reference_date=="2022-02-28" & model=="CU-select"))
+}

### 2023-05-19 R/run-alloscore.R

@@ -37,7 +37,7 @@ run_alloscore <- function(forecast_data, truth_data, one_forecast_date){
       by = c("location", "target_end_date"))
 
   ## run alloscore
-  Ks <- seq(from = 5000, to = 80000, by = 500)
+  Ks <- seq(from = 500, to = 80000, by = 500)
   Kdf <- data.frame(matrix(Ks,nrow = 1))
   names(Kdf) <- paste0("K=",Ks)
 

### 2023-05-19 _targets.R

@@ -30,6 +30,10 @@ list(
     name = forecast_data,
     command = get_forecast_data(values$forecast_dates)
   ),
+  # tar_target(
+  #   name = filtered_forecast_data,
+  #   command = filter_problematic_data(forecast_data)
+  # ),
   tar_target(
     name = truth_data,
     command = get_truth_data()

### 2023-05-19 _targets.R
updating targets.

@@ -51,6 +51,4 @@ list(
     command = plot_K_v_alloscore(alloscore_2021.12.13),
     format = "file"
   )
-
-
 )

### 2023-05-19 R/run-alloscore.R
updating targets.

@@ -68,7 +68,8 @@ run_alloscore <- function(forecast_data, truth_data, one_forecast_date){
       names_prefix = "K=",
       names_to = "K"
     ) |>
-    dplyr::mutate(K = as.numeric(K))
+    dplyr::mutate(K = as.numeric(K),
+                  forecast_date = one_forecast_date)
 
   return(ascores_long)
 }

### 2023-05-19 R/multi-date-analysis.R
adding script for multi-day analysis.

@@ -1,0 +1,55 @@
+library(targets)
+library(tidyverse)
+library(ggbump)
+
+theme_set(theme_bw())
+
+tar_load(starts_with("alloscore"))
+
+alloscore_data <- do.call(rbind, mget(ls()))
+
+alloscore_data_summaries <- alloscore_data |>
+  mutate(Kdiff = abs(ytot-K)) |>
+  group_by(forecast_date) |>
+  filter(Kdiff == min(Kdiff))|>
+  mutate(allo_rank = rank(value, ties.method = "random"))
+
+p1 <- ggplot(alloscore_data_summaries,
+       aes(x=as.Date(forecast_date), y=allo_rank, color=model)) +
+  geom_bump() +
+  xlab(NULL)
+p2 <- ggplot(alloscore_data_summaries,
+       aes(x=as.Date(forecast_date), y=value, color=model)) +
+  geom_point() + geom_line()+
+  xlab(NULL)
+
+
+tar_load("score_data")
+
+alloscore_models <- unique(alloscore_data$model)
+score_data_summaries <- score_data |>
+  mutate(reference_date = target_end_date - 14) |>
+  filter(model %in% alloscore_models) |>
+  group_by(model, reference_date) |>
+  summarize(mwis = mean(wis)) |>
+  ungroup() |>
+  group_by(reference_date) |>
+  mutate(mwis_rank = rank(mwis, ties.method = "random"))
+
+p3 <- ggplot(score_data_summaries,
+       aes(x=as.Date(reference_date), y=mwis_rank, color=model)) +
+  geom_bump() +
+  xlab(NULL)
+
+p4 <- ggplot(score_data_summaries,
+       aes(x=as.Date(reference_date), y=mwis, color=model)) +
+  geom_point() + geom_line() +
+  xlab(NULL)
+
+gridExtra::grid.arrange(p1, p2, p3, p4, nrow=2)
+
+all_data <- left_join(alloscore_data_summaries |> mutate(forecast_date = as.Date(forecast_date)),
+                      score_data_summaries |> rename(forecast_date = reference_date))
+
+ggplot(all_data, aes(x=mwis_rank, y=allo_rank, color=model)) +
+  geom_jitter()

### 2023-05-21 notes/application_K_choice.rmd
comments on application set up

---
title: "Comments on set up for application"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    latex_engine: xelatex
urlcolor: blue
header-includes:
   - \usepackage{amsmath}
   - \usepackage{algorithm}
   - \usepackage[noend]{algpseudocode}
   - \usepackage{amsthm}
   - \usepackage{bm}
   - \usepackage{cases}
   - \usepackage{caption}
   - \usepackage{unicode-math}
   - \usepackage{hyperref}
   - \DeclareMathOperator*{\argmin}{argmin}
---
<!-- $ ...to toggle Latexyz math scope off -->
\newcommand{\del}[2]{\frac{\partial {#1} }{\partial {#2}} }
\newcommand{\dby}[2]{\frac{d {#1} }{d {#2}} }
\newcommand{\sbar}{\overline{s}}
\newtheorem{proposition}{Proposition}

Notation:
\begin{itemize}
\item $s$ is the date a forecast is made
\item $t$ is the target date for a forecast
\item $y_t = (y_{1,t}, \ldots, y_{l,t}, \ldots, y_{L,t})$ is the vector of realized resource need in each of $L$ locations on the target date $t$
\item $K_t$ is the constraint on available resources that is in effect for target date $t$. For example, this may change over time if the planning agency is going to ramp up (or down) production of the resource in question over time.
\item $x_t = (x_{1,t}, \ldots, x_{l,t}, \ldots, x_{L,t})$ is the vector of resource allocation levels for target date $t$, with $\sum_l x_{l,t} \leq K_t$.
\end{itemize}

Preliminary comments:
\begin{itemize}
\item The introduction of the time indices $s$ and $t$ is new in this document.  Everything in our main manuscript is framed for a single forecast (so, one value of $s$) for a single target date $t$.  But our application now considers multiple forecast dates and target end dates, so we need this notation to think carefully about what's going on.
\item Our goal is to figure out how to apply the ideas we've already developed to multiple forecast dates in a justifiable way.  Other work (e.g. Bertsimas et al.) has more carefully considered how to set up the allocation problem in a way that incorporates planning over time, but we don't want to go there for this paper.
\end{itemize}

Concerns, high level:
\begin{enumerate}
\item Currently, a working decision was made in the application to take ``the alloscore value that was for the $K$ closest to the actual total observations," which I'll denote here as $y^{tot}_t = \sum_l y_{l,t}$. I think there are two problems with this:
    \begin{enumerate}
    \item I believe that this choice for $K_t$ likely yields an improper score. The basic reason is that the score depends on the value of the random variable/vector that is being forecasted and generally this is not allowed. A little more below.
    \item It's hard to see how to justify this choice for $K_t$ in a way that corresponds to the set up of the decision-making problem. The implication is that at each time $t$, we have exactly enough resources to meet what will be the eventually-realized need (if an oracle forecast is used). But this is not realistic.
    \end{enumerate}
\item Closely related to 1 (b): How are we thinking about the set up of the decision making problem across multiple dates? Is there a way we could summarize overall model performance across the time span under consideration with a single number? To the extent possible, I think we should try to start from a plausible multi-time-step decision making set up that can fit within our set up and go from there.
    \begin{enumerate}
    \item As described in 1.b., I don't think setting $K_t = y^{tot}_t = \sum_l y_{l,t}$ can be justified as corresponding to any natural decision making set up. In general, $K_t$ will not be equal to $y^{tot}_t$; our resource planning is not that good/lucky.
    \item If I remember right, on our call a few weeks ago Ben suggested something like ``Suppose you get one chance to allocate resources with a constraint $K$, and then those allocation levels are fixed and carry through the wave under consideration." This seems like a justifiable idea, but it doesn't line up well with the kind of analysis we think we'd like to do in our application where we consider multiple forecast dates $s$. Looking at multiple forecast dates only makes sense if we can update our allocations based on the information in those forecasts.
    \item Another idea is to say that we have $K$ units of resources and on each forecast date $s$, we get to decide how they will be allocated on target date $t = s + h$.
        \begin{itemize}
        \item This is a little dicey if the resource is something fixed like ventilators or hospital beds because some of the resources we had previously allocated might still be in use by the time $t$ rolls around (e.g. if someone is hospitalized or on a ventilator for multiple weeks)
        \item Could make more sense for something like supply of oxygen or medication, if $K$ new units of the resource are manufactured each week.
        \item Recognize that maybe we don't want to talk too much about how this connects to real life in the manuscript...
        \item But I don't have any great ideas for how we might motivate any particular fixed value of $K$ of interest ahead of time without trying to connect to context (e.g. look up some information somewhere about total oxygen supplies and average amounts used per admitted hospital patient...)
        \item The strategy I took when talking through things in london was to try not to fixate too much on any particular value of K, but to note that model rankings were fairly stable across values of $K$.
        \end{itemize}
    \item Returning to the question I raised above about how we might summarize overall model performance across the time span under consideration: it seems like maybe if we pre-specify a grid of values of $K$ that are used for all target dates $t$, for each of those values of $K$ maybe we could average scores across the target dates? I think this is basically Ben's suggestion, but without the idea that we only get to specify an allocation once.  We'd maybe end up with something like the peaked plot, but with multiple peaks, one for each target date...?  If you had enough target dates, maybe you'd get to something relatively smooth/domed.  This is still not a single number, and there is still the question of how to summarize across $K$'s to get to a single number.
    \end{enumerate}
\end{enumerate}

A few more thoughts about issues with propriety when setting $K_t = y^{tot}_t$:
\begin{enumerate}
\item Intuition:
    \begin{itemize}
    \item I think a slightly more careful (but still informal) definition of propriety than the informal one we usually use goes something like: in expectation, a proper scoring rule maximally rewards a forecast that corresponds to the conditional distribution of the random variable in question given all information available at the time $s$ that the forecast is produced.
    \item $y^{tot}_t$ was not observable at time $s$. A ``forecast" that knows $y^{tot}_t$ and conditions on that quantity in addition to other available information at time $s$ will be different (in general, probably narrower/more concentrated) than a forecast that conditions only on the information available at time $s$.
    \item Rephrasing the previous point: if I am omnipotent and I tell a forecaster, ``you will be scored using an allocation score using $K_t = 22,000$ because there will be 22,000 hospital admissions on the target date," I have leaked information about the future to that forecaster, and they will produce a forecast that differs from the forecast that would have been produced using only information available as of time $s$.
    \item In expectation, my score will likely reward the information-privileged forecaster more than the forecast corresponding to the conditional distribution of the random variable in question given all information available at time $s$. Therefore, it's not a proper score.
    \item The intuition in the above points is pretty rough.  I'd like to be able to state/prove this more precisely.
    \item I think the fact that we did this in our initial analysis and nobody immediately raised any red flags is a sign of how unclear all of this is. It feels kind of important that we say something about the fact that you can't do this in the manuscript, to warn readers/users off of this kind of analysis (assuming of course that I'm right about the resulting score being improper).
    \end{itemize}
\item More generally, for the integrated allocation score we probably need to pre-specify what the distribution over $K$ is. This makes perfect sense from the real-life perspective of a public health decision maker, it just gets tricky when you try to think about how to set up an evaluation exercise in an abstract setting without specific knowledge about $K$...
\item Rough start at an attempt to be a little more formal:
    \begin{itemize}
    \item Let $\mathcal{F}_s$ denote the sub $\sigma$-algebra generated by the collection $O_t$ of everything that's observable as of time $s$. Intuitively, this captures the set of all available information as of the time the forecast is generated.
    \item Let $\tilde{\mathcal{F}}_s$ denote the sub $\sigma$-algebra generated by $\tilde{O}_s = O_s \cup Y^{tot}_t$.  Noting that $Y^{tot}_t$ was not observable at time $s$, $\mathcal{F}_s \subset \tilde{\mathcal{F}}_s$.
    \item Maybe propriety is a statement about maximally rewarding the distribution of the random variable $Y_t$ given the sub $\sigma$ algebra $\mathcal{F}_s$, which is different from its distribution given $\tilde{\mathcal{F}}_s$? Where ``maximally rewarding" is in expectation with respect to... unclear \textemdash $\mathcal{F}_s$??
    \item I only half remember how this stuff works; will have to think a bit more to put the pieces together.
    \end{itemize}
\end{enumerate}

### 2023-05-22 notes/application_K_choice.rmd
updates/corrections to thoughts r.e. choice of K

@@
     \begin{enumerate}
-    \item I believe that this choice for $K_t$ likely yields an improper score.
-The basic reason is that the score depends on the value of the random variable/vector that is being forecasted and generally this is not allowed.
-A little more below.
+    \item It's not clear to me whether or not this choice for $K_t$ likely yields an improper score.
+The basic reason is that the score depends on the value of the random variable/vector that is being forecasted.
+This may or may not be OK?
     \item It's hard to see how to justify this choice for $K_t$ in a way that corresponds to the set up of the decision-making problem.
@@
 
-A few more thoughts about issues with propriety when setting $K_t = y^{tot}_t$:
-\begin{enumerate}
-\item Intuition:
+\section{More on propriety}
+
+Some general set up that's somewhere between Gerding et al. and Gneiting and Raftery:
 \begin{itemize}
-    \item I think a slightly more careful (but still informal) definition of propriety than the informal one we usually use goes something like: in expectation, a proper scoring rule maximally rewards a forecast that corresponds to the conditional distribution of the random variable in question given all information available at the time $s$ that the forecast is produced.
-    \item $y^{tot}_t$ was not observable at time $s$.
-A ``forecast" that knows $y^{tot}_t$ and conditions on that quantity in addition to other available information at time $s$ will be different (in general, probably narrower/more concentrated) than a forecast that conditions only on the information available at time $s$.
-    \item Rephrasing the previous point: if I am omnipotent and I tell a forecaster, ``you will be scored using an allocation score using $K_t = 22,000$ because there will be 22,000 hospital admissions on the target date," I have leaked information about the future to that forecaster, and they will produce a forecast that differs from the forecast that would have been produced using only information available as of time $s$.
-    \item In expectation, my score will likely reward the information-privileged forecaster more than the forecast corresponding to the conditional distribution of the random variable in question given all information available at time $s$.
-Therefore, it's not a proper score.
-    \item The intuition in the above points is pretty rough.  I'd like to be able to state/prove this more precisely.
-    \item I think the fact that we did this in our initial analysis and nobody immediately raised any red flags is a sign of how unclear all of this is.
-It feels kind of important that we say something about the fact that you can't do this in the manuscript, to warn readers/users off of this kind of analysis (assuming of course that I'm right about the resulting score being improper).
+\item $(\mathcal{Y}, \mathcal{A})$ is a measurable space
+\item $\mathcal{P}$ is a convex class of probability measures on $(\mathcal{Y}, \mathcal{A})$
+\item A probabilistic forecast is any probability measure $F \in \mathcal{P}$
+\item $S(F, y)$ is a scoring rule
+\item $\overline{S}(F, G) = \mathbb{E}_{Y \sim G}[S(F, Y)]$ is the expected score for $F$ under the distribution $G$
+\item The (negatively oriented) scoring rule $S$ is proper relative to $\mathcal{P}$ if $\overline{S}(G, G) \leq \overline{S}(F, G)$ whenever $F, G \in \mathcal{P}$
+\item $s(x, y)$ is a loss function [or scoring function if $x$ is a functional of $F$]
+\item We can obtain a scoring rule from a scoring function by setting $S(F, y) = s(x^F, y)$ where $x^F = \argmin_x \mathbb{E}_{Y \sim F} [s(x, Y)]$
 \end{itemize}
-\item More generally, for the integrated allocation score we probably need to pre-specify what the distribution over $K$ is.
-This makes perfect sense from the real-life perspective of a public health decision maker, it just gets tricky when you try to think about how to set up an evaluation exercise in an abstract setting without specific knowledge about $K$...
-\item Rough start at an attempt to be a little more formal:
+
+Specifics for our set up for allocation:
 \begin{itemize}
-    \item Let $\mathcal{F}_s$ denote the sub $\sigma$-algebra generated by the collection $O_t$ of everything that's observable as of time $s$.
-Intuitively, this captures the set of all available information as of the time the forecast is generated.
-    \item Let $\tilde{\mathcal{F}}_s$ denote the sub $\sigma$-algebra generated by $\tilde{O}_s = O_s \cup Y^{tot}_t$.
-Noting that $Y^{tot}_t$ was not observable at time $s$, $\mathcal{F}_s \subset \tilde{\mathcal{F}}_s$.
-    \item Maybe propriety is a statement about maximally rewarding the distribution of the random variable $Y_t$ given the sub $\sigma$ algebra $\mathcal{F}_s$, which is different from its distribution given $\tilde{\mathcal{F}}_s$?
-Where ``maximally rewarding" is in expectation with respect to...
-unclear \textemdash $\mathcal{F}_s$??
-    \item I only half remember how this stuff works; will have to think a bit more to put the pieces together.
+\item Our scoring function is $s_K(x,y) = \sum_{i=1}^n L (y_i - x_i)_+$ (subject to constraints that each $x_i$ is non-negative and $\sum_i x_i \leq K$, which I suppose we could build into $s_K$)
+\item $x^F$ has entries $x^F_i = F_i^{-1}(1 - \lambda^{\star,F}(K) / L)$ where $\lambda^{\star,F}(K)$ is chosen such that $\sum_i x^F_i = K$
+\item Our scoring rule is $S_K(F, y) = s_K(x^F, y) = \sum_{i=1}^n L (y_i - F_i^{-1}(1 - \lambda^{\star,F}(K) / L))_+$
 \end{itemize}
-\end{enumerate}
+
+Since $S_K$ is derived from a scoring rule using Bayes actions, it is a proper scoring rule. Writing this out a bit, we have to check that for all $F, G \in \mathcal{P}$:
+\begin{align*}
+&\overline{S}(G, G) \leq \overline{S}(F, G) \\
+\Leftrightarrow \quad &\mathbb{E}_{Y \sim G}\left[ S_K(G, Y) \right] \leq \mathbb{E}_{Y \sim G}\left[ S_K(F, Y) \right] \\
+\Leftrightarrow \quad &\mathbb{E}_{Y \sim G}\left[ \sum_{i=1}^n L (Y_i - G_i^{-1}(1 - \lambda^{\star,G}(K) / L))_+ \right] \leq \mathbb{E}_{Y \sim G}\left[ \sum_{i=1}^n L (Y_i - F_i^{-1}(1 - \lambda^{\star,F}(K) / L))_+ \right].
+\end{align*}
+But for any given $K$ this inequality holds by construction, because we chose the Bayes act allocation by minimizing the left hand side.
+
+What about when $K = Y^{tot} = \sum_i Y_i$? Continuing from the last line above, we have
+\begin{align*}
+\quad &\mathbb{E}_{Y \sim G}\left[ \sum_{i=1}^n L (Y_i - G_i^{-1}(1 - \lambda^{\star,G}(Y^{tot}) / L))_+ \right] \leq \mathbb{E}_{Y \sim G}\left[ \sum_{i=1}^n L (Y_i - F_i^{-1}(1 - \lambda^{\star,F}(Y^{tot}) / L))_+ \right]
+\end{align*}
+
+I can't decide:
+\begin{itemize}
+\item Is the result now immediate through some kind of iterated expectations thing?
+\item Or is this now complicated because $Y^{tot}$ (which is a function of all of the $Y_i$) appears in a nonlinear way in each of the terms of the summation, so that we can no longer separate into terms by location?
+\end{itemize}

### 2023-05-22 R/multi-date-analysis.R
updating first pass at multi-date analysis with K = y_tot

@@ -12,10 +12,14 @@ alloscore_data_summaries <- alloscore_data |>
   mutate(Kdiff = abs(ytot-K)) |>
   group_by(forecast_date) |>
   filter(Kdiff == min(Kdiff))|>
-  mutate(allo_rank = rank(value, ties.method = "random"))
+  mutate(nmodels = n(),
+         allo_rank = rank(value, ties.method = "random"),
+         allo_rank_std = 1- (allo_rank-1)/(nmodels-1)) |>
+  ungroup() |>
+  complete(model, forecast_date)
 
 p1 <- ggplot(alloscore_data_summaries,
-       aes(x=as.Date(forecast_date), y=allo_rank, color=model)) +
+       aes(x=as.Date(forecast_date), y=allo_rank_std, color=model)) +
   geom_bump() +
   xlab(NULL)
 p2 <- ggplot(alloscore_data_summaries,
@@ -34,10 +38,14 @@ score_data_summaries <- score_data |>
   summarize(mwis = mean(wis)) |>
   ungroup() |>
   group_by(reference_date) |>
-  mutate(mwis_rank = rank(mwis, ties.method = "random"))
+  mutate(nmodels = n(),
+         mwis_rank = rank(mwis, ties.method = "random"),
+         mwis_rank_std = 1- (mwis_rank-1)/(nmodels-1)) |>
+  ungroup() |>
+  complete(model, reference_date)
 
 p3 <- ggplot(score_data_summaries,
-       aes(x=as.Date(reference_date), y=mwis_rank, color=model)) +
+       aes(x=as.Date(reference_date), y=mwis_rank_std, color=model)) +
   geom_bump() +
   xlab(NULL)
 
@@ -51,5 +59,37 @@ gridExtra::grid.arrange(p1, p2, p3, p4, nrow=2)
 all_data <- left_join(alloscore_data_summaries |> mutate(forecast_date = as.Date(forecast_date)),
                       score_data_summaries |> rename(forecast_date = reference_date))
 
-ggplot(all_data, aes(x=mwis_rank, y=allo_rank, color=model)) +
-  geom_jitter()
+ggplot(all_data, aes(x=mwis_rank_std, y=allo_rank_std, color=model)) +
+  geom_point() + xlim(0,1) + ylim(0,1) +
+  scale_color_brewer(palette="Paired")
+
+## faceted by model, colored by model
[…]

### 2023-05-23 R/data-ingestion.R
trying to update some more targets code.

@@ -1,18 +1,18 @@
-get_forecast_data <- function(forecast_dates){
+get_forecast_data <- function(forecast_dates, models, locations){
   require(covidHubUtils)
   load_forecasts(
     dates = forecast_dates,
     date_window_size = 6,
+    models = models,
+    locations = locations,
     types = c("quantile"),
-    targets = paste(0:30, "day ahead inc hosp"),
+    targets = paste(14:20, "day ahead inc hosp"),
     source = "zoltar",
     verbose = FALSE,
     as_of = NULL,
     hub = c("US")) |>
     align_forecasts() |>
-    dplyr::filter(
-      relative_horizon == 14,
-      location < 57)
+    dplyr::filter(relative_horizon == 14)
 }
 
 get_truth_data <- function(){
@@ -31,12 +31,4 @@ get_forecast_scores <- function(forecast_data, truth_data){
     truth = truth_data,
     use_median_as_point = TRUE
   )
-}
-
-filter_problematic_data <- function(forecast_data){
-  require(tidyverse)
-  forecast_data |>
-    ## this removes two state-week-model that has a deterministically zero forecast.
-    filter(!(abbreviation=="KS" & reference_date=="2022-02-21" & model=="CU-select")) |>
-    filter(!(abbreviation=="WA" & reference_date=="2022-02-28" & model=="CU-select"))
 }

### 2023-05-23 R/multi-date-analysis.R

@@ -11,9 +11,10 @@ alloscore_data <- do.call(rbind, mget(ls()))
 alloscore_data_summaries <- alloscore_data |>
   mutate(Kdiff = abs(ytot-K)) |>
   group_by(forecast_date) |>
-  filter(Kdiff == min(Kdiff))|>
+  #filter(Kdiff == min(Kdiff))|>
+  filter(K == 15000)|>
   mutate(nmodels = n(),
-         allo_rank = rank(value, ties.method = "random"),
+         allo_rank = rank(value, ties.method = "min"),
          allo_rank_std = 1- (allo_rank-1)/(nmodels-1)) |>
   ungroup() |>
   complete(model, forecast_date)
@@ -39,7 +40,7 @@ score_data_summaries <- score_data |>
   ungroup() |>
   group_by(reference_date) |>
   mutate(nmodels = n(),
-         mwis_rank = rank(mwis, ties.method = "random"),
+         mwis_rank = rank(mwis, ties.method = "min"),
          mwis_rank_std = 1- (mwis_rank-1)/(nmodels-1)) |>
   ungroup() |>
   complete(model, reference_date)
@@ -55,6 +56,8 @@ p4 <- ggplot(score_data_summaries,
   xlab(NULL)
 
 gridExtra::grid.arrange(p1, p2, p3, p4, nrow=2)
+
+gridExtra::grid.arrange(p2, p4, nrow=2)
 
 all_data <- left_join(alloscore_data_summaries |> mutate(forecast_date = as.Date(forecast_date)),
                       score_data_summaries |> rename(forecast_date = reference_date))

### 2023-05-23 R/run-alloscore.R

@@ -37,7 +37,7 @@ run_alloscore <- function(forecast_data, truth_data, one_forecast_date){
       by = c("location", "target_end_date"))
 
   ## run alloscore
-  Ks <- seq(from = 500, to = 80000, by = 500)
+  Ks <- seq(from = 15000, to = 15000, by = 500)
   Kdf <- data.frame(matrix(Ks,nrow = 1))
   names(Kdf) <- paste0("K=",Ks)
 

### 2023-05-23 _targets.R

@@ -5,6 +5,8 @@ library(targets)
 library(tarchetypes)
 library(tibble)
 library(crew)
+library(covidHubUtils)
+data("hub_locations")
 
 # Set target options:
 tar_option_set(
@@ -24,16 +26,21 @@ tar_source()
 ## create a group of alloscore targets
 values <- tibble(forecast_dates = as.character(seq.Date(as.Date("2021-11-22"), as.Date("2022-02-28"), by = "7 days")))
 
+## set of required locations: all states + DC
+reqd_locs <- hub_locations |>
+  dplyr::filter(geo_type == "state", !(geo_value %in% c("us", "as", "gu", "mp", "pr", "um", "vi"))) |>
+  dplyr::pull(fips)
+
 # List of targets:
 list(
   tar_target(
-    name = forecast_data,
-    command = get_forecast_data(values$forecast_dates)
+    name = eligible_models,
+    command = determine_eligible_models(values$forecast_dates, locations = reqd_locs)
   ),
-  # tar_target(
-  #   name = filtered_forecast_data,
-  #   command = filter_problematic_data(forecast_data)
-  # ),
+  tar_target(
+    name = forecast_data,
+    command = get_forecast_data(values$forecast_dates, models = eligible_models, locations = reqd_locs)
+  ),
   tar_target(
     name = truth_data,
     command = get_truth_data()
@@ -45,10 +52,10 @@ list(
   tar_map(
     values = values,
     tar_target(alloscore, run_alloscore(forecast_data, truth_data, forecast_dates))
-  ),
-  tar_target(
-    name = figure_K_v_alloscore_2021.12.13,
-    command = plot_K_v_alloscore(alloscore_2021.12.13),
-    format = "file"
   )
+  # tar_target(
+  #   name = figure_K_v_alloscore_2021.12.13,
+  #   command = plot_K_v_alloscore(alloscore_2021.12.13),
+  #   format = "file"
+  # )
 )

### 2023-05-23 _targets.R
updates to use one K and eligible models.

@@ -48,7 +48,7 @@ list(
   tar_target(
     name = score_data,
     command = get_forecast_scores(forecast_data, truth_data)
-  ),
+   ),
   tar_map(
     values = values,
     tar_target(alloscore, run_alloscore(forecast_data, truth_data, forecast_dates))

### 2023-05-23 alloscore_manuscript/allocation.tex
some progress on reviewing applications of decision theoretic forecast eval

@@
 
+\usepackage{natbib}
+\bibliographystyle{unsrtnat}
+
 \title{Evaluating infectious disease forecasts with allocation scoring rules}
-\author{Authors}
+\author{Aaron Gerding, Nicholas G. Reich, Benjamin Rogers, Evan L. Ray}
 
@@
 Infectious disease forecasts have been used to inform decision-making about a wide variety of measures designed to reduce disease spread and/or mitigate the severity of disease outcomes.
-Such decision-making applications include the allocation of limited medical supplies such as ventilators \cite{bertsimas2021predictionsCOVID}, implementation of social distancing measures such as stay-at-home policies, planning site selection for vaccine trials \cite{bertsimas2021predictionsCOVID}, and strategies for public health communication campaigns.
+Such decision-making applications include the allocation of limited medical supplies such as ventilators \citep{bertsimas2021predictionsCOVID}, implementation of social distancing measures such as stay-at-home policies, planning site selection for vaccine trials \citep{bertsimas2021predictionsCOVID}, and strategies for public health communication campaigns.
 
@@
 We give an overview of these procedures in Section \ref{sec:methods.decisiontheory}.
-These methods have been applied with some regularity in fields such as economics and meteorology.
-[review previous applications of decision-theoretic evaluation to fields like economics and high level description of scoring procedures they use]
+There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision-making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
+We review this work only briefly here, and we refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \elr{TODO: identify relevant review or book style summary for meteorology} for discussions focused on applications to economics and meteorology, respectively.
+In finance, the value of forecasts can often be measured by the profits generated by trading decisions informed by the forecasts, perhaps adjusted for risk levels \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
+In applications to supply chain management and meteorology, the value of forecasts has typically been operationalized by considering the costs associated with decisions regarding the amount of inventory to hold or the level of protection against the impacts of extreme weather events to enact {Peter Catt (2007), Fotios Petropoulos and colleagues (2019), Nada Sanders and Gregory Graman (2009), T.N.
+Palmer (2002), Florian Pappenberger and colleagues (2015)}.
+For example, in supply chain management these decisions may incur costs related to holding inventory, labor, or providing poor service, while in meteorology we may need to balance the costs of implementing protective measures with the costs of potentially preventable weather damages.
+In this framework, a forecast has value if it leads to decisions with low total costs.
+In all of these fields, analyses have consistently found that common measures of statistical forecast accuracy do not necessarily correspond directly to measures of the value of forecasts as an input to decision-making \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
+% Consistently, but not uniformly
 
 However, we are aware of only a limited body of work that explicitly attempts to measure the value of infectious disease forecasts through their impact on policy, and much of this discussion has proceeded informally.
-For example, Ioannidis et al.
-\cite{ioannidis2022forecastingCOVIDfailed} discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts.
+For example, \cite{ioannidis2022forecastingCOVIDfailed} discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts.
 Separately, there is a thread of literature that does quantify the link between infectious disease modeling and policy making, but this work has been done outside of a forecasting context.
-As an example, Probert et al.
-\cite{Probert2016decisionMakingFootMouth} develop measures of the cost of actions designed to control a hypothetical outbreak of foot-and-mouth disease and use this framework to explore policy recommendations from a variety of simulation-based projection models.
+As an example, \cite{Probert2016decisionMakingFootMouth} develop measures of the cost of actions designed to control a hypothetical outbreak of foot-and-mouth disease and use this framework to explore policy recommendations from a variety of simulation-based projection models.
 
 In practice, probabilistic infectious disease forecasts have most often been made for observations that emerge from public health surveillance systems and evaluated with standard, ``off-the-shelf'' scoring rules.
-For example, seasonal influenza forecasts in the US and dengue forecasts for Peru and Puerto Rico targeted public health surveillance measures of incidence over time and space, and used log-score and mean absolute errors to evaluate forecast skill \cite{mcgowan_collaborative_2019,reich_collaborative_2019,johansson_open_2019}.
-Pandemic COVID-19 forecasts of observed cases, hospitalizations and/or deaths in the US and Europe, as reported by municipal, state, or federal surveillance systems, were evaluated using the weighted interval score (WIS, which is an approximation of the continuous ranked probability score, or CRPS), and prediction interval coverage\cite{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.
-Similarly, CRPS was also used to assess probabilistic forecasts of dengue incidence at the district level in Vietnam\cite{colon-gonzalez_probabilistic_2021}.
+For example, seasonal influenza forecasts in the US and dengue forecasts for Peru and Puerto Rico targeted public health surveillance measures of incidence over time and space, and used log-score and mean absolute errors to evaluate forecast skill \citep{mcgowan_collaborative_2019,reich_collaborative_2019,johansson_open_2019}.
+Pandemic COVID-19 forecasts of observed cases, hospitalizations and/or deaths in the US and Europe, as reported by municipal, state, or federal surveillance systems, were evaluated using the weighted interval score (WIS, which is an approximation of the continuous ranked probability score, or CRPS), and prediction interval coverage \citep{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.
+Similarly, CRPS was also used to assess probabilistic forecasts of dengue incidence at the district level in Vietnam \citep{colon-gonzalez_probabilistic_2021}.
 While some of these scores can be interpreted through the lens of decision theory, and all of the application-specific papers cited above had authors from public health agencies, none of them (including some for which we were co-authors) make explicit connections between forecast evaluation and how a forecast was used in practice.
@@
 
-\bibliographystyle{unsrt}
-\bibliography{allocation.bib}
+\bibliography{allocation}
 

### 2023-05-24 alloscore_manuscript/allocation.tex
Added a bit to paragraph 1, as well as some notes for other citations I will add in -BWR

@@
 
-Infectious disease forecasts have been used to inform decision-making about a wide variety of measures designed to reduce disease spread and/or mitigate the severity of disease outcomes.
-Such decision-making applications include the allocation of limited medical supplies such as ventilators \citep{bertsimas2021predictionsCOVID}, implementation of social distancing measures such as stay-at-home policies, planning site selection for vaccine trials \citep{bertsimas2021predictionsCOVID}, and strategies for public health communication campaigns.
+% People using infectious disease forecasting as an input to decision making:
+
+%Center for Forecasting and Outbreak Analytics (CDC) https://www.cdc.gov/forecast-outbreak-analytics/annual-report/2023/key-accomplishments.html#omicron-response anual report
+%  Bertsimas gives data-driven approach to provide clincal insights and support ventilator allocation for policy makers,
+
+% "Our results have been used at the clinical level by several hospitals to triage patients, guide care management, plan ICU capacity, and re-distribute ventilators.
+At the policy level, they are currently supporting safe back-to-work policies at a major institution and vaccine trial location planning at Janssen Pharmaceuticals, and have been integrated into the US Center for Disease Control’s pandemic forecast."
+
+%Grohskopf "Prevention and Control of Seasonal Influenza with Vaccines: Recommendations of the Advisory Committee on Immunization Practices — United States, 2022–23 Influenza Season" - Current CDC recommendations state that "Timing, intensity, and severity of the 2022-23 influenze season cannot be predicted"
+
+% Rainisch G, Shankar M, Wellman M, Merlin T, Meltzer MI.
+Regional spread of Ebola virus, West Africa, 2014.
+Emerg Infect Dis 2015;21:444–7 Forecast of geographic spread of Ebola which helped aid decision-makers in allocation of surveillance resources.
+
+% MI Meltzer - Estimating the Future Number of Cases in the Ebola (example of Ebola prediction informing policy to increase ERU capacities among other measures)
+
+%https://www.cdc.gov/mmwr/preview/mmwrhtml/su6303a1.htm
+
+Infectious disease forecasting models have emerged as important tools in public health decision making.
+Predictions of disease dynamics have been used to inform decision-making about a wide variety of measures designed to reduce disease spread and/or mitigate the severity of disease outcomes.
+In a simple example, estimates of expected onset of flu season have been used to inform national vaccination strategies  \cite{igboh2023timing}, and forecasts of ebola dynamics have been used to allocate surveillance resources \cite{meltzer2014estimating}.
+Other examples of past decision making applications of infectious disease forecasts include the allocation of limited medical supplies such as ventilators \cite{bertsimas2021predictionsCOVID}, implementation of social distancing measures such as stay-at-home policies, planning site selection for vaccine trials \cite{bertsimas2021predictionsCOVID}, and strategies for public health communication campaigns.
+
+For example, estimates of expected onset of flu have been used to inform national flu vaccination strategies \cite{igboh2023timing}, and allocation of ebola surveillance resources \cite{meltzer2014estimating}.
+Other past decision making applications of infectious disease forecasts include the allocation of limited medical supplies such as ventilators \cite{bertsimas2021predictionsCOVID}, implementation of social distancing measures such as stay-at-home policies, planning site selection for vaccine trials \cite{bertsimas2021predictionsCOVID}, and strategies for public health communication campaigns.
 

### 2023-05-26 R/run-alloscore.R
adding a run with multi-date analysis, all models.

@@ -3,18 +3,6 @@ run_alloscore <- function(forecast_data, truth_data, one_forecast_date){
   require(alloscore)
   require(distfromq)
 
-  # keep selected models
-  mkeep <- c("BPagano-RtDriven",
-             "COVIDhub-4_week_ensemble",
-             "COVIDhub-baseline",
-             "CU-select",
-             "IHME-CurveFit",
-             "JHUAPL-Bucky",
-             "JHUAPL-Gecko",
-             "MUNI-ARIMA",
-             "USC-SI_kJalpha",
-             "UVA-Ensemble")
-
   ## process forecast data, adding distfromq output
   forecast_data_processed <- forecast_data |>
     ## forecast dates are different but reference dates are Mondays
@@ -25,10 +13,6 @@ run_alloscore <- function(forecast_data, truth_data, one_forecast_date){
     mutate(
       ps = map(ps, deframe),
       qs = map(qs, deframe)
-    ) |>
-    dplyr::filter(model %in% mkeep) %>%
-    dplyr::mutate(
-      model = ifelse(model == "COVIDhub-4_week_ensemble", "COVIDhub-ensemble", model)
     ) |>
     add_pdqr_funs(dist = "distfromq", types = c("p", "q")) |>
     relocate(dist, F, Q) |>

### 2023-05-26 random_undocumented_experiments/multi-date-analysis.R

@@ -63,8 +63,11 @@ all_data <- left_join(alloscore_data_summaries |> mutate(forecast_date = as.Date
                       score_data_summaries |> rename(forecast_date = reference_date))
 
 ggplot(all_data, aes(x=mwis_rank_std, y=allo_rank_std, color=model)) +
-  geom_point() + xlim(0,1) + ylim(0,1) +
-  scale_color_brewer(palette="Paired")
+  geom_point() + xlim(0,1) + ylim(0,1)
+
+ggplot(all_data, aes(x=mwis, y=value, color=model)) +
+  geom_point(alpha=.7)
+
 
 ## faceted by model, colored by model
 ggplot(all_data, aes(x=mwis_rank_std, y=allo_rank_std, color=model)) +
@@ -86,10 +89,21 @@ ggplot(all_data, aes(x=mwis_rank_std, y=allo_rank_std, color=ytot)) +
   facet_wrap(.~model) +
   geom_abline(slope=1, intercept=0, linetype=2, color="gray")
 
+## sized by ytot
 ggplot(all_data, aes(x=mwis_rank_std, y=allo_rank_std, color=forecast_date, size=ytot)) +
   geom_point(alpha=.8) + xlim(0,1) + ylim(0,1) +
   facet_wrap(.~model) +
   geom_abline(slope=1, intercept=0, linetype=2, color="gray")
+
+## scatterplot of wis and alloscore
+ggplot(all_data, aes(x=mwis, y=value, color=forecast_date, size=ytot)) +
+  geom_point(alpha=.8) + scale_y_log10() + scale_x_log10() +
+  facet_wrap(.~model) +
+  geom_abline(slope=1, intercept=0, linetype=2, color="gray")
+
+## distribution of alloscores, by model
+ggplot(all_data, aes(x=value, y=model)) +
+  geom_boxplot()
 
 
 ggplot(all_data, aes(x=mwis_rank_std, y=allo_rank_std, color=model)) +

### 2023-05-26 alloscore_manuscript/allocation.tex
Uptaded intro paragraph 1

@@
 \def\apg#1{{\color{red}\textbf{APG:[#1]}}}
+\def\bwr#1{{\color{violet}\textbf{BWR:[#1]}}}
 
@@
 
-%Grohskopf "Prevention and Control of Seasonal Influenza with Vaccines: Recommendations of the Advisory Committee on Immunization Practices — United States, 2022–23 Influenza Season" - Current CDC recommendations state that "Timing, intensity, and severity of the 2022-23 influenze season cannot be predicted"
-
 % Rainisch G, Shankar M, Wellman M, Merlin T, Meltzer MI.
@@
 Predictions of disease dynamics have been used to inform decision-making about a wide variety of measures designed to reduce disease spread and/or mitigate the severity of disease outcomes.
-In a simple example, estimates of expected onset of flu season have been used to inform national vaccination strategies  \cite{igboh2023timing}, and forecasts of ebola dynamics have been used to allocate surveillance resources \cite{meltzer2014estimating}.
-Other examples of past decision making applications of infectious disease forecasts include the allocation of limited medical supplies such as ventilators \cite{bertsimas2021predictionsCOVID}, implementation of social distancing measures such as stay-at-home policies, planning site selection for vaccine trials \cite{bertsimas2021predictionsCOVID}, and strategies for public health communication campaigns.
-
-For example, estimates of expected onset of flu have been used to inform national flu vaccination strategies \cite{igboh2023timing}, and allocation of ebola surveillance resources \cite{meltzer2014estimating}.
-Other past decision making applications of infectious disease forecasts include the allocation of limited medical supplies such as ventilators \cite{bertsimas2021predictionsCOVID}, implementation of social distancing measures such as stay-at-home policies, planning site selection for vaccine trials \cite{bertsimas2021predictionsCOVID}, and strategies for public health communication campaigns.
+For example, estimates of expected onset of flu season were used to inform national vaccination strategies  \citep{igboh2023timing}, and forecasts of ebola dynamics were used to allocate surveillance resources \citep{meltzer2014estimating, rainisch2015regional}.
+In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision-making \citep{cdc2022cfa}.
+\cite{bertsimas2021predictionsCOVID} developed tools to inform decision making from infectious disease forecasts, which have been used to to inform allocation of limited medical supplies such as ventilators, ICU capacity planning, and vaccine distribution strategy.
+Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping \citep{utnews2022}.
+\bwr{Figure out these news site citations!}
 

### 2023-05-27 random_undocumented_experiments/example_application_BWR.R
Update comment colors latex, and added an example allocation file that works on my comp

@@ -1,3 +1,5 @@
+#Changed example allocation as a couple of covidhubUtil functions were not working.
+
 library(tidyverse)
 library(covidHubUtils)
 if (!require("distfromq")) {
@@ -19,9 +21,105 @@ if (
   message("Alloscore may be out of date...")
 }
 
-# hub_repo_path <- "~/Documents/Research/covid19-forecast-hub/"
+
+########################
+### Working versions of required covidHubUtils functions:
+
+align_forecasts_one_temporal_resolution <- function(
+    forecasts,
+    reference_dates,
+    reference_weekday,
+    reference_windows,
+    drop_nonpos_relative_horizons
+) {
+  if (length(unique(forecasts$temporal_resolution)) > 1) {
+    stop("standardize_forecasts_one_temporal_resolution only supports forecasts at a single temporal resolution.")
+  }
+  
+  if (is.null(reference_windows)) {
+    if (reference_weekday == "Saturday") {
+      reference_windows <- -4:2
+    } else if (reference_weekday == "Monday") {
+      reference_windows <- -6:0
+    } else {
+      stop("Reference windows undefined")
+    }
+  }
+  
+  if (!is.list(reference_windows)) {
+    reference_windows <- list(reference_windows)
+  }
+  
+  if (!is.null(reference_dates)) {
+    # ensure we have dates
+    reference_dates <- as.Date(reference_dates)
+  } else {
+    # every date from that of first forecast - diameter of first window
+    # to that of last forecast + diameter of last window
+    all_dates <- seq(
+      min(forecasts$forecast_date) - (
+        max(sort(reference_windows[[1]])) -
+          min(sort(reference_windows[[1]]))
+      ),
+      max(forecasts$forecast_date) + (
+        max(sort(reference_windows[[length(reference_windows)]])) -
+          min(sort(reference_windows[[length(reference_windows)]])) 
+      ),
+      by = 1
+    )
+    
+    # keep the dates identified above that are the specified reference_weekday
+    reference_dates <- all_dates[weekdays(all_dates) == reference_weekday]
+  }
+  
+  # create a tibble where each row contains:
+  # - a possible forecast date
+  # - a reference date to which that forecast date should be assigned
+  ref_df <- tibble(
+    reference_date = reference_dates,
+    forecast_date = purrr::map2(
+      reference_date, 
+      reference_windows, 
+      ~.x+.y
+    )
+  ) %>% unnest(cols = forecast_date)
+  
[…]

### 2023-06-05 notes/allocation_notes.rmd
Restructuring and elaboration of Quantile Scoring Rules section in notes

@@
 header-includes:
+   - \usepackage{hyperref}
    - \usepackage{amsmath}
@@
    - \usepackage{unicode-math}
-   - \usepackage{hyperref}
    - \DeclareMathOperator*{\argmin}{argmin}
@@
 \begin{align}
-s(x,y)=s_{O,U}(x,y) &= O(g(x) - g(y))\sb{+} + U(g(x)-g(y))\sb{-}. \label{eqn:scalar_score} \\
-&= \kappa((1-\alpha)(g(x) - g(y))\sb{+} + \alpha(g(x)-g(y))\sb{-}) \label{eqn:scalar_score_alpha} \\
-&= \kappa(\mathbb{1}\{x>y\} - \alpha)(g(x)-g(y)) \label{eqn:scalar_score_ident}\\
-&:= s_{\kappa, \alpha}(x,y) := \kappa s_{\alpha}(x,y),
-\end{align}
-where $\alpha = U/(U+O)$, $\kappa = U+O$,
-$u_+ := \max(u,0)$, $u_- := \max(-u,0) = (-u)_+$, and $\mathbb{1}\{A\}$ is the indicator function of the event $A$.
-This scoring function $s$ is negatively oriented, i.e., a smaller score indicates better performance corresponding to a smaller cost due to under- or over-prediction.
-
-By making the forecaster aware of $s$ prior to prediction, we elicit an $\alpha$-level quantile $q_{\alpha,F}$ of the forecast distribution $F$ for $Y$.
-To see this, notice that if a forecaster believes that $Y \sim F$, they will calculate their expected score to be
-\begin{align} 
-\sbar_F(x) = E_F[s(x,Y)]  &= O E_F[(g(x) - g(Y))\sb{+}] + U E_F[(g(x) - g(Y))\sb{-}]. \label{eqn:scalar_obj}
+s(x,y)=s_{O,U}(x,y) &= O(g(x) - g(y))\sb{+} + U(g(y)-g(x))\sb{+} \label{eqn:scalar_score}
+\end{align}
+where $u_+ := u \mathbb{1}\{u > 0\}$ and $\mathbb{1}\{A\}$ is the indicator function of the event $A$. This is a \emph{negatively oriented} scoring function.
+That is, a smaller score indicates better performance corresponding to a smaller cost due to under- or over-prediction of the realized $y$.
+
+By making the forecaster aware of $s$ prior to prediction, we elicit an $\alpha$-level quantile $q_{\alpha,F}$ of the forecast distribution $F$ for $Y$, 
+where $\alpha = U/(U+O)$. To see this, notice that if a forecaster believes that $Y \sim F$, they will calculate their expected score to be
+\begin{align} 
+\sbar_F(x) = E_F[s(x,Y)]  &= O E_F[(g(x) - g(Y))\sb{+}] + U E_F[(g(Y) - g(x))\sb{+}]. \label{eqn:scalar_obj}
 \end{align}
@@
 \nonumber 0 =\frac{d}{dx}\sbar_F(x) 
-&= O E_F\left[\frac{d}{dx}(g(x) - g(Y))\sb{+}\right] + U E_F\left[\frac{d}{dx}(g(x) - g(Y))\sb{-}\right] \\
+&= O E_F\left[\frac{d}{dx}(g(x) - g(Y))\sb{+}\right] + U E_F\left[\frac{d}{dx}(g(Y) - g(x))\sb{+}\right] \\
 &= O E_F[g^{\prime}(x)\mathbb{1}\{Y < x\}] + U E_F[-g^{\prime}(x)\mathbb{1}\{Y \geq x\}] \\
@@
 &= g^{\prime}(x)\left( O F(x) - U (1-F(x))\right) \\
-&= \kappa g^{\prime}(x)\left(F(x) - \alpha\right), \label{scalar_Z_deriv}
+&= (O+U) g^{\prime}(x)\left(F(x) - \alpha\right), \label{scalar_Z_deriv}
 \end{align}
@@
 The critical point $q_{\alpha, F}$ is actually a minimum since \eqref{scalar_Z_deriv} is non-positive on $\{x<q_{\alpha, F}\}$ and non-negative on $\{x>q_{\alpha, F}\}$.
- When $g^{\prime}(q)>0$, this minimum is unique since
-\begin{align}
-   \frac{d^2}{dx^2}\sbar_F(x)\big|_{x=q_{\alpha, F}} &= 
-   \kappa\left(g^{\prime}(x)f(x) + g^{\prime \prime}(x)\left(F(x) - \alpha\right)\right)\big|_{x=q_{\alpha, F}} \label{scalar_Z_deriv2}\\
-   &=\kappa(g^{\prime}(q_{\alpha, F})f(q_{\alpha, F}) + g^{\prime \prime}(q_{\alpha, F})\cdot 0) >0.
-\end{align}
-
-If $g^{\prime}(q_{\alpha, F})=0$ or does not exist, or $Q_F(\alpha)$ is not a singleton set, we'll need to adopt additional criteria for whether $q_{\alpha, F}$ is optimal.
-One possibility is to take $q_{\alpha, F}=\min \left\{x \mid F_X(x) \geq \alpha\right\}$, and if $g^{\prime}(q_{\alpha, F})=0$ we choose the prediction $x$ to be the infimum of the region around $q_{\alpha, F}$ where $g$ is flat.
-One example of such a $g$ is the \emph{power curve} of a wind turbine which becomes constant above a certain wind speed.
+ With the assumptions that $g^{\prime}(q)>0$ and that $Q_F\left(\alpha\right)$ is the singleton $\{q_{\alpha, F}\}$, this minimum is unique since
+\begin{align}
+\frac{d^2}{dx^2}\sbar_F(x)\big|_{x=q_{\alpha, F}} 
+&= (O+U)\left(g^{\prime}(x)f(x) + g^{\prime \prime}(x)\left(F(x) - \alpha\right)\right)\big|_{x=q_{\alpha, F}} \label{scalar_Z_deriv2}\\
+&= (O+U)(g^{\prime}(q_{\alpha, F})f(q_{\alpha, F}) + g^{\prime \prime}(q_{\alpha, F})\cdot 0) >0.
+\end{align}
+
+By reparametrizing to $\alpha = U/(U+O)$ and $\kappa = U+O$, we can express $s$ in several alternative forms:
+\begin{align}
+s_{O,U}(x,y) &= \kappa((1-\alpha)(g(x) - g(y))\sb{+} + \alpha(g(y)-g(x))\sb{+}) & \label{eqn:scalar_score_alpha} \\
+&= \kappa(\mathbb{1}\{y \leq x\} - \alpha)(g(x)-g(y)) & \label{eqn:scalar_score_ident}\\
+&= -\kappa\left[\alpha g(x) + \mathbb{1}\{y \leq x\}(g(y)-g(x)) + h_1(y)\right] & (\text{with} \quad h_1(y) = - \alpha g(y))  \label{eqn:scalar_score_Gnieting_Raftery}\\
+&= \kappa\left[(1-\alpha) g(x) + \mathbb{1}\{y > x\}(g(y)-g(x)) + h_2(y)\right] & (\text{with} \quad h_2(y) = (\alpha - 1) g(y)) \label{eqn:scalar_score_OR}\\
+&:= s_{\kappa, \alpha}(x,y) := \kappa s_{\alpha}(x,y).&
+\end{align}
+
+Since $h_1$ and $h_2$ are functions of only $y$, they do not enter the first order equation \eqref{scalar_Z_deriv} and can be regarded as nuisance functions carrying information about the application irrelevant to quantile elicitation.
+The forms \eqref{eqn:scalar_score_ident} and \eqref{eqn:scalar_score_Gnieting_Raftery} appear as equations (41) and (40) in [@gneitingStrictlyProperScoring2007a].\footnote{Or rather the negatives of these equations since Gneiting and Raftery use positively oriented scoring functions.
+Also note that their $x$ plays the role of our $y$.}
+ The form \eqref{eqn:scalar_score_OR} is often used in operations research and meteorology literature, and is discussed later in the section \nameref{subsec:metparams}.
+
+Note from \eqref{eqn:scalar_score_ident} that $\frac{d}{dx}\sbar_F(x) =\frac{d}{dx}E_F[s(x,Y)] = k(x)E_F[V(x,Y)]$, where $k(x) = \kappa g^{\prime}(x)$ and 
+\begin{align}
+V(x,y) 
+% &\coloneq (1-\alpha)\mathbb{1}\{y < x\} - \alpha \mathbb{1}\{y \geq x\} \mathrel{\phantom{=}}\\
+% &= (1-\alpha)\mathbb{1}\{y < x\} - \alpha \left(1 - \mathbb{1}\{y < x\}\right) \\
+&= \mathbb{1}\{y \leq x\} - \alpha. 
+\end{align}
+By virtue of $E_F[V(Q_F(\alpha),Y)]=0$, $V$ is said to be an \emph{identification function} for the $\alpha$ quantile.
+The fact that, generally speaking, any elicitable functional (such as a quantile) has an identification function is known as \emph{Osband's principle}.
+
+Forecasting the minimizer of $\sbar_F(x) = E_F[s_{\kappa,\alpha}(x,Y)]$ in this situation is known as the \emph{Bayes act} $a_F$ under $s_\alpha$ for the forecaster.
+Assuming a forecaster is informed, rational, and risk-neutral enough to take $a_F$, we can evaluate $F$, implicitly, as a distributional forecast, by the \emph{scoring rule} $S_{\alpha}$ induced by $s_{\alpha}$:
+\begin{align}
+S_{\alpha}(F,y) := s_{\alpha}(Q_F(\alpha), y).
+\end{align}
+If $y$ is generated by $Y \sim G$ then $S_{\alpha}(F,y)$ is an estimate of 
+\begin{align}
+S_{\alpha}(F,G) &:= E_G[S_{\alpha}(F,Y)] = E_G[s_{\alpha}(Q_{F}(\alpha),Y)].
+\end{align}
+The estimand here is the value at $G$ of the functional $\mathcal{\overline{S}}$ on $\mathcal{F_0}$ given by
+\begin{align}
+\mathcal{\overline{S}}_{G,\alpha}(F) := \sbar_G(Q_{\alpha}(F))
+\end{align}
+where $Q_{\alpha}$ is the $\alpha$ quantile functional. This is the expected value under $s$ of the forecaster's $\alpha$ quantile under the \emph{actual} data generating process $G$.
+
+\begin{remark}
+Relaxing either of the assumptions that $g^{\prime}(q_{\alpha, F})>0$ or that $Q_F(\alpha)$ is a singleton set would require adopting additional criteria for whether $q_{\alpha, F}$ is optimal.
+One possibility would be to take $q_{\alpha, F}=\min \left\{x \mid F_X(x) \geq \alpha\right\}$, and in case $g^{\prime}(q_{\alpha, F})=0$ or fails to exist, choose the prediction $x$ to be the infimum of the region around $q_{\alpha, F}$ where $g$ is flat.
+An example of such a $g$ is the \emph{power curve} of a wind turbine which becomes constant for wind speeds above the turbine's maximum operating speed.
 Here $x$ would be a forecast of future wind speed $Y$ at time $t$ and $g(x)$ would be the power a turbine operator commits to supplying the electric grid at $t$.
 Taking $g$ to be constant below a certain point could also serve to express the futility or lack of meaning of forecasts below that point, such as when a forecast user is only able to sell a commodity for which demand can become negative, turning into supply.
-
-Note that $\frac{d}{dx}\sbar_F(x) =\frac{d}{dx}E_F[s(x,Y)] = h(x)E_F[V(x,Y)]$, where $h(x) = \kappa g^{\prime}(x)$ and 
-\begin{align}
-V(x,y) &\coloneq (1-\alpha)\mathbb{1}\{y < x\} - \alpha \mathbb{1}\{y \geq x\} \mathrel{\phantom{=}}\\
-&= (1-\alpha)\mathbb{1}\{y < x\} - \alpha \left(1 - \mathbb{1}\{y < x\}\right) \\
-&= \mathbb{1}\{y < x\} - \alpha. 
-\end{align}
-By virtue of $E_F[V(Q_F(\alpha),Y)]=0$, $V$ is said to be an \emph{identification function} for the $\alpha$ quantile.
-The fact that, generally speaking, any elicitable functional (such as a quantile) has an identification function is known as \emph{Osband's principle}.
-
-Forecasting the minimizer of $\sbar_F(x) = E_F[s_{\kappa,\alpha}(x,Y)]$ in this situation is known as the \emph{Bayes act} $a_F$ under $s_\alpha$ for the forecaster.
-Assuming a forecaster is informed, rational, and risk-neutral enough to take $a_F$, we can evaluate $F$, implicitly, as a distributional forecast, by the \emph{scoring rule} $S_{\alpha}$ induced by $s_{\alpha}$:
-\begin{align}
-S_{\alpha}(F,y) := s_{\alpha}(Q_F(\alpha), y).
-\end{align}
-If $y$ is generated by $Y \sim G$ then $S_{\alpha}(F,y)$ is an estimate of 
-\begin{align}
-S_{\alpha}(F,G) &:= E_G[S_{\alpha}(F,Y)] = E_G[s_{\alpha}(Q_{F}(\alpha),Y)].
-\end{align}
-This defines a functional on $\mathcal{F_0}$, the class of measures on the Borel-Lebesgue sets of $\mathbb{R}$ by
-\begin{align}
-\mathcal{\overline{S}}_{G,\alpha}(F) := \sbar_G(Q_{\alpha}(F))
-\end{align}
-where $Q_{\alpha}$ is the $\alpha$ quantile functional.
-
-## Examples of forecaster expected scores
+\end{remark}
+
+\subsection{Examples of forecaster expected scores}
 
@@
 
-##  Alternative Notation Possibilities
+\subsection{Alternative Notation Possibilities}
 
@@
 
-## Newsvendor parameters
+\subsection{Newsvendor parameters}
 
@@
 \begin{align}
-s_n(x,y) & = a_x x +a_y y +a_{+}(x-y)_+ + a_{-}(x-y)_-.
+s_n(x,y) & = a_x x +a_y y +a_{+}(x-y)\sb{+} + a_{-}(y-x)\sb{+}.
 \end{align}
@@
 \begin{align}
-s_n(x,y) = cx-py - u(x-y)_+ + (p+r)(x-y)_-
-\end{align}
-where $x$ newspapers are ordered at price $c$, $y - (x-y)_-$ are sold at retail price $p$, $(x-y)_+$ are sold at salvage price $u$, 
-and $r(x-y)_-$ of customer "goodwill" is lost. 
+s_n(x,y) = cx-py - u(x-y)\sb{+} + (p+r)(y-x)\sb{+}
+\end{align}
+where $x$ newspapers are ordered at price $c$, $y - (y-x)\sb{+}$ are sold at retail price $p$, $(x-y)\sb{+}$ are sold at salvage price $u$, 
+and $r(y-x)\sb{+}$ of customer "goodwill" is lost. 
 This can be rewritten as
 \begin{align}
-s_n(x,y) &= (a_{+} + a_x)(x-y)_+ + (a_{-} - a_x)(x-y)_- +(a_x + a_y)y \\
-&= (c-u)(x-y)_+ + (p+r-c)(x-y)_- -(p-c)y
+s_n(x,y) &= (a_{+} + a_x)(x-y)\sb{+} + (a_{-} - a_x)(y-x)\sb{+} +(a_x + a_y)y \\
+&= (c-u)(x-y)\sb{+} + (p+r-c)(y-x)\sb{+} -(p-c)y
 \end{align}
@@
 \begin{align}
-\sbar_{n,F}(x) = (a_{+} + a_x)E_F[(x-Y)_+] + (a_{-} - a_x)E_F[(x-Y)_-] +(a_x + a_y)E[Y].
+\sbar_{n,F}(x) = (a_{+} + a_x)E_F[(x-Y)\sb{+}] + (a_{-} - a_x)E_F[(Y-x)\sb{+}] +(a_x + a_y)E[Y].
 \end{align}
@@
 
-## Meteorologist parameters
-
-Similarly, a weather forecaster might be judged by the cost $Cx$ of recommended protection $x$ against a level $y$ of adverse weather (e.g., rainfall) added to any loss $L(x-y)_-$ resulting from underprediction, leading to the scoring function
-\begin{align}
-s_m(x,y) = Cx + L(x-y)_-,
+\subsection{Meteorologist parameters}\label{subsec:metparams}
+
+Similarly, a weather forecaster might be judged by the cost $Cx$ of recommended protection $x$ against a level $y$ of adverse weather (e.g., rainfall) added to any loss $L(y-x)\sb{+}$ resulting from underprediction, leading to the scoring function
+\begin{align}
+s_m(x,y) = Cx + L(y-x)\sb{+},
 \end{align}
@@
 \begin{align}
-s_m(x,y) = C(x-y)_+ + (L-C)(x-y)_- + Cy.
+s_m(x,y) = C(x-y)\sb{+} + (L-C)(y-x)\sb{+} + Cy.
 \end{align}
@@
 
-# Multipoint Quantile Scoring Rules
+\section{Multipoint Quantile Scoring Rules}
 
@@
 
-# Allocation scoring rules
+\section{Allocation Scoring Rules}
+
 We now develop \emph{allocation scoring functions} that elicit forecasts $x_i$ for outcomes $y_i, i = 1,\ldots,N$, each with it's own incremental cost $O_i$ and loss $U_i$ as for a quantile scoring rule, but with the additional constraint
@@
 S_{\symbfit{\beta}}(\mathbf{F},\mathbf{y}) = 
-s_{\symbfit{\beta}}(\mathbf{Q}^{\symbfit{\beta}}, \mathbf{y})&=\sum O_i(g_i(Q_i^{\symbfit{\beta}}) - g_i(y_i))_+ + U_i(g_i(Q_i^{\symbfit{\beta}})-g_i(y_i))_-
+s_{\symbfit{\beta}}(\mathbf{Q}^{\symbfit{\beta}}, \mathbf{y})&=\sum O_i(g_i(Q_i^{\symbfit{\beta}}) - g_i(y_i))\sb{+} + U_i(g_i(y_i)-g_i(Q_i^{\symbfit{\beta}}))\sb{+}
 \\
@@
 \begin{align}
-l(\mathbf{x}, \mathbf{y}) = U\sum_{i=1}^{N} (x_i - y_i)_-.
+l(\mathbf{x}, \mathbf{y}) = U\sum_{i=1}^{N} (y_i - x_i)\sb{+}.
 \end{align}

### 2023-06-05 R/run-alloscore.R
first pass at updated targets pipeline

@@ -1,4 +1,4 @@
-run_alloscore <- function(forecast_data, truth_data, one_forecast_date){
+run_alloscore <- function(forecast_data, truth_data, Ks){
   require(tidyverse)
   require(alloscore)
   require(distfromq)
@@ -17,8 +17,6 @@ run_alloscore <- function(forecast_data, truth_data, one_forecast_date){
 
   ## process forecast data, adding distfromq output
   forecast_data_processed <- forecast_data |>
-    ## forecast dates are different but reference dates are Mondays
-    dplyr::filter(reference_date == one_forecast_date) |>
     dplyr::select(-type) |>
     nest(ps = quantile, qs = value) |>
     relocate(ps, qs) |>
@@ -37,41 +35,16 @@ run_alloscore <- function(forecast_data, truth_data, one_forecast_date){
       by = c("location", "target_end_date"))
 
   ## run alloscore
-  Ks <- seq(from = 500, to = 80000, by = 500)
-  Kdf <- data.frame(matrix(Ks,nrow = 1))
-  names(Kdf) <- paste0("K=",Ks)
+  ascores <- forecast_data_processed %>%
+  group_by(model, reference_date) %>%
+  group_modify(~ alloscore(
+    .x,
+    K = Ks,
+    y = .x$value,
+    target_names = "abbreviation"
+  ))
 
-  ascores <- forecast_data_processed  |>
-      bind_cols(Kdf) |>
-      group_by(model) |>
-      summarise(
-        ytot = sum(value),
-        across(starts_with("K"), ~alloscore:::alloscore(
-          y = value,
-          F = F,
-          Q = Q,
-          w = 1,
-          K = unique(.x),
-          kappa = 1,
-          alpha = 1,
-          dg = 1,
-          eps_K = .01,
-          eps_lam = 1e-5,
-          verbose = TRUE,
-          against_oracle = TRUE
-        ))
-      )
-
-  ascores_long <- ascores |>
-    tidyr::pivot_longer(
-      starts_with("K="),
-      names_prefix = "K=",
-      names_to = "K"
-    ) |>
-    dplyr::mutate(K = as.numeric(K),
-                  forecast_date = one_forecast_date)
-
-  return(ascores_long)
+  return(ascores)
 }
 
 

### 2023-06-05 _targets.R

@@ -10,6 +10,7 @@ library(crew)
 tar_option_set(
   controller = crew_controller_local(workers = 8),
   packages = c("tidyverse", "covidHubUtils", "distfromq", "alloscore", "gh"), # packages that your targets need to run
+  imports = c("alloscore"),
   format = "rds" # default storage format
 )
 
@@ -19,10 +20,16 @@ tar_option_set(
 # remotes::install_github("reichlab/covidHubUtils")
 
 # Run the R scripts in the R/ folder:
-tar_source()
+tar_source(files = c("R/data-ingestion.R", "R/plot-alloscores.R", "R/run-alloscore.R"))
 
 ## create a group of alloscore targets
-values <- tibble(forecast_dates = as.character(seq.Date(as.Date("2021-11-22"), as.Date("2022-02-28"), by = "7 days")))
+values <-
+  tibble(forecast_dates = as.character(seq.Date(
+    as.Date("2021-11-22"), as.Date("2022-02-28"), by = "7 days"
+  )))
+values <- values[c(12,13,14),]
+
+Ks <- seq(from = 2000, to = 7000, by = 500)
 
 # List of targets:
 list(
@@ -42,13 +49,13 @@ list(
     name = score_data,
     command = get_forecast_scores(forecast_data, truth_data)
   ),
-  tar_map(
-    values = values,
-    tar_target(alloscore, run_alloscore(forecast_data, truth_data, forecast_dates))
+  tar_target(
+    name = alloscore_df,
+    command = run_alloscore(forecast_data, truth_data, K = Ks)
   ),
   tar_target(
-    name = figure_K_v_alloscore_2021.12.13,
-    command = plot_K_v_alloscore(alloscore_2021.12.13),
+    name = figure_K_v_alloscore,
+    command = plot_K_v_alloscore(alloscore_df),
     format = "file"
   )
 )

### 2023-06-05 R/plot-alloscores.R
first pass at updated targets pipeline

@@ -19,23 +19,27 @@ plot_K_v_alloscore <- function(alloscores) {
                "#AA4499",
                "#000000")
 
+  palette <- palette[1:length(models)]
+
   names(palette) <- c(models[models != "COVIDhub-baseline"], "COVIDhub-baseline")
 
   solid_models <- c("COVIDhub-baseline", "MUNI-ARIMA", "CU-select")
-  linetypes <- c(2, 4, 5, 6, 2, 4, 5, rep(1, 3))
+  linetypes <- c(2, 4, 5, 6, 2, 4, 5, rep(1, 3))[1:length(models)]
   names(linetypes) <- c(models[!(models %in% solid_models)], solid_models)
 
   p <- ggplot(
     data = alloscores,
-    mapping = aes(x = K, y = value, color = model, linetype = model)
+    mapping = aes(x = K, y = score, color = model, linetype = model)
   ) +
     geom_line() +
     scale_color_manual(values = palette) +
     scale_linetype_manual(values = linetypes) +
-    geom_vline(xintercept = unique(alloscores$ytot), linetype = 2) +
+    geom_vline(aes(xintercept = ytot), linetype = 2) +
     xlab("Total Allocation Constraint (K)") +
     ylab("Allocation Score\n(Excess unmet need)") +
-    theme_bw()
+    facet_wrap(vars(reference_date)) +
+    theme_bw() +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
   pdf("figures/allo_scores_wide.pdf", width = 7, height = 3.5)
   print(p)

### 2023-06-06 notes/allocation_notes.rmd
another correction in quantile scoring rule section

@@
 (See \hyperref[scalar_Z_deriv_detail]{appendix} for an alternative derivation in terms of a density.)
-The critical point $q_{\alpha, F}$ is actually a minimum since \eqref{scalar_Z_deriv} is non-positive on $\{x<q_{\alpha, F}\}$ and non-negative on $\{x>q_{\alpha, F}\}$.
+The critical point $q_{\alpha, F}$ is actually a minimum since \eqref{scalar_Z_deriv} is non-positive on $\{x< \inf\left(Q_F\left(\alpha\right)\right)\}$ and non-negative on $\{x>\sup\left(Q_F\left(\alpha\right)\right)\}$.
  With the assumptions that $g^{\prime}(q)>0$ and that $Q_F\left(\alpha\right)$ is the singleton $\{q_{\alpha, F}\}$, this minimum is unique since

### 2023-06-08 R/run-alloscore-overk.R
Added a target to sum alloscore over a grid of Ks, error on week of 2.21 in CU-select forecast

@@ -1,0 +1,66 @@
+run_alloscore_overk <- function(forecast_data, truth_data, one_forecast_date){
+  require(tidyverse)
+  require(alloscore)
+  require(distfromq)
+
+  ## process forecast data, adding distfromq output
+  forecast_data_processed <- forecast_data |>
+    ## forecast dates are different but reference dates are Mondays
+    dplyr::filter(reference_date == one_forecast_date) |>
+    dplyr::select(-type) |>
+    nest(ps = quantile, qs = value) |>
+    relocate(ps, qs) |>
+    mutate(
+      ps = map(ps, deframe),
+      qs = map(qs, deframe)
+    ) |>
+    add_pdqr_funs(dist = "distfromq", types = c("p", "q")) |>
+    relocate(dist, F, Q) |>
+    left_join(
+      truth_data |> select(location, target_end_date, value),
+      by = c("location", "target_end_date"))
+
+  ## run alloscore
+  Ks <- seq(from = 500, to = 60000, by = 500)
+  Kdf <- data.frame(matrix(Ks,nrow = 1))
+  names(Kdf) <- paste0("K=",Ks)
+
+  ascores <- forecast_data_processed  |>
+      bind_cols(Kdf) |>
+      group_by(model) |>
+      summarise(
+        ytot = sum(value),
+        across(starts_with("K"), ~alloscore:::alloscore(
+          y = value,
+          F = F,
+          Q = Q,
+          w = 1,
+          K = unique(.x),
+          kappa = 1,
+          alpha = 1,
+          dg = 1,
+          eps_K = .01,
+          eps_lam = 1e-5,
+          verbose = TRUE,
+          against_oracle = TRUE
+        ))
+      )
+
+  ascores_long  <- ascores |>
+    tidyr::pivot_longer(
+      starts_with("K="),
+      names_prefix = "K=",
+      names_to = "K"
+    ) |>
+    dplyr::mutate(K = as.numeric(K),
+                  forecast_date = one_forecast_date)
+  
+  #Note: there are some negative scores going as low as -23.4. largest negatives are at K=2500 and 3,000
+  ascores_overk <- ascores_long  %>% 
+    dplyr::group_by(model)%>% 
+    dplyr::summarise(mean_score=mean(value))
+
+  return(ascores_overk )
+}
+
+

### 2023-06-08 _targets.R

@@ -53,6 +53,10 @@ list(
     values = values,
     tar_target(alloscore, run_alloscore(forecast_data, truth_data, forecast_dates))
   )
+  # tar_map(
+  #   values=values,
+  #   tar_target(alloscore_overk, run_alloscore_overk(forecast_data, truth_data, forecast_dates))
+  # )
   # tar_target(
   #   name = figure_K_v_alloscore_2021.12.13,
   #   command = plot_K_v_alloscore(alloscore_2021.12.13),

### 2023-06-09 notes/allocation_notes.rmd
More work on entropy divergence reliability resolution and newsvendor

@@
 \item a probability measure $P \in \mathcal{F_0}$ induced by a random variable $X$ be identified by its cumulative distribution function $F_X$
-\item $F$ denote the forecast distribution for the level of future need $Y$
-\item $Y \sim G$
+\item $F$ denote some forecast distribution for the level of future need $Y$ which is known entirely or up to some parameters by its forecaster
+\item $G$ denote the unknowable distribution of $Y$
 \item $Q_F := F^{-1}: [0,1] \rightrightarrows \mathbb{R}$ be the (set-valued) quantile function which maps a probability level $\alpha \in [0, 1]$ to the corresponding quantile set $\left[q_{\alpha, F}^{-}, q_{\alpha, F}^{+}\right]$ of $F$ where $q_{\alpha, F}^{-} := \sup \{x \mid F(x) < \alpha\}$ and
@@
 \item $Q_\alpha: \mathcal{F_0} \rightrightarrows \mathbb{R}$ be the functional which maps a distribution $F$ to its $\alpha$-level quantile set
-\item $q_{\alpha, F}$ be a (usually unique) element of $Q_F(\alpha) = Q_{\alpha}(F)$. 
+\item $q_{\alpha, F}$ be a (usually unique) element of $Q_F(\alpha) = Q_{\alpha}(F)$
+\item $q_{\alpha, Y}$ be an element of the unknowable set $Q_G(\alpha) = Q_{\alpha}(G)$
 \end{itemize}
 
-We measure the forecaster's performance when the realized need is $y$ with the scoring function
-\begin{align}
-s(x,y)=s_{O,U}(x,y) &= O(g(x) - g(y))\sb{+} + U(g(y)-g(x))\sb{+} \label{eqn:scalar_score}
+Let us measure the forecaster's performance when the realized need is $y$ with the scoring function
+\begin{align}
+s_{O,U}(x,y) &= O(g(x) - g(y))\sb{+} + U(g(y)-g(x))\sb{+} \label{eqn:scalar_score}
 \end{align}
@@
 
-By making the forecaster aware of $s$ prior to prediction, we elicit an $\alpha$-level quantile $q_{\alpha,F}$ of the forecast distribution $F$ for $Y$, 
-where $\alpha = U/(U+O)$. To see this, notice that if a forecaster believes that $Y \sim F$, they will calculate their expected score to be
-\begin{align} 
-\sbar_F(x) = E_F[s(x,Y)]  &= O E_F[(g(x) - g(Y))\sb{+}] + U E_F[(g(Y) - g(x))\sb{+}]. \label{eqn:scalar_obj}
-\end{align}
-In order to minimize this expectation, they must forecast a solution of the first order equation
-\begin{align}
-\nonumber 0 =\frac{d}{dx}\sbar_F(x) 
+The scoring function $s_{O,U}$ has the special property that by giving the forecaster prior notice that $s_{O,U}$ will be used to measure performance, 
+we \emph{elicit} an $\alpha$-level quantile $q_{\alpha,Y}$ for $Y$, 
+where $\alpha = U/(U+O)$. That is, we create a situation where a forecaster believing that $Y \sim F$ and trying only to minimize their expected score 
+\begin{align} 
+\sbar_{F,O,U}(x) := E_F[s_{O,U}(x,Y)]  &= O E_F[(g(x) - g(Y))\sb{+}] + U E_F[(g(Y) - g(x))\sb{+}] \label{eqn:scalar_obj}
+\end{align}
+will report their quantile $q_{\alpha,F}$. To see this, notice that in order to minimize this expectation, they must forecast a solution of the first order equation
+\begin{align}
+\nonumber 0 =\frac{d}{dx}\sbar_{F,O,U}(x) 
 &= O E_F\left[\frac{d}{dx}(g(x) - g(Y))\sb{+}\right] + U E_F\left[\frac{d}{dx}(g(Y) - g(x))\sb{+}\right] \\
@@
 \begin{align}
-\frac{d^2}{dx^2}\sbar_F(x)\big|_{x=q_{\alpha, F}} 
+\frac{d^2}{dx^2}\sbar_{F,O,U}(x)\big|_{x=q_{\alpha, F}} 
 &= (O+U)\left(g^{\prime}(x)f(x) + g^{\prime \prime}(x)\left(F(x) - \alpha\right)\right)\big|_{x=q_{\alpha, F}} \label{scalar_Z_deriv2}\\
@@
 
-Note from \eqref{eqn:scalar_score_ident} that $\frac{d}{dx}\sbar_F(x) =\frac{d}{dx}E_F[s(x,Y)] = k(x)E_F[V(x,Y)]$, where $k(x) = \kappa g^{\prime}(x)$ and 
-\begin{align}
-V(x,y) 
+Note from \eqref{eqn:scalar_score_ident} that $\frac{d}{dx}\sbar_{F,\alpha}(x) =\frac{d}{dx}E_F[s_{\alpha}(x,Y)] = k(x)E_F[V_{\alpha}(x,Y)]$, where $k(x) = \kappa g^{\prime}(x)$ and 
+\begin{align}
+V_{\alpha}(x,y) 
 % &\coloneq (1-\alpha)\mathbb{1}\{y < x\} - \alpha \mathbb{1}\{y \geq x\} \mathrel{\phantom{=}}\\
@@
 \end{align}
-By virtue of $E_F[V(Q_F(\alpha),Y)]=0$, $V$ is said to be an \emph{identification function} for the $\alpha$ quantile.
+By virtue of $E_F[V_{\alpha}(Q_F(\alpha),Y)]=0$, $V_{\alpha}$ is said to be an \emph{identification function} for the $\alpha$ quantile.
 The fact that, generally speaking, any elicitable functional (such as a quantile) has an identification function is known as \emph{Osband's principle}.
 
-Forecasting the minimizer of $\sbar_F(x) = E_F[s_{\kappa,\alpha}(x,Y)]$ in this situation is known as the \emph{Bayes act} $a_F$ under $s_\alpha$ for the forecaster.
-Assuming a forecaster is informed, rational, and risk-neutral enough to take $a_F$, we can evaluate $F$, implicitly, as a distributional forecast, by the \emph{scoring rule} $S_{\alpha}$ induced by $s_{\alpha}$:
+Forecasting the minimizer of $\sbar_{F, \kappa,\alpha}(x) = E_F[s_{\kappa,\alpha}(x,Y)]$---that is, the forecaster quantile $q_{\alpha, F}$---is known as the \emph{Bayes act} $a_F$ under $s_\alpha$ for the forecaster.
+Assuming a forecaster is informed, rational, and risk-neutral enough to take $a_F$, we can evaluate $F$, implicitly, as a distributional forecast, by the \emph{scoring rule} $S_{\alpha}$ induced by $s_{\alpha}$
 \begin{align}
@@
 \end{align}
+It follows from this definition that
+\begin{align}
+S_{\alpha}(F,\tilde{F}) &:= E_{\tilde{F}}[S_{\alpha}(F,Y)] \geq S_{\alpha}(F,F) \quad\text{for all}\quad F, \tilde{F} \in  \mathcal{F_0}. \label{eq:properness}
+\end{align}
+A scoring rule $S(F,y)$ is said to be \emph{proper} when it satisfies \eqref{eq:properness} and \emph{strictly proper} when the inequality is strict.
+$S_{\alpha}$ is not strictly proper since \eqref{eq:properness} is an equality for any $F$ and $\tilde{F}$ sharing an $\alpha$ quantile.
+$S_{\alpha}$ is however sometimes
+said to be strictly proper for the $\alpha$ quantile since the inequality becomes strict whenever $q_{\alpha, F} \neq q_{\alpha, \tilde{F}}$, (see e.g., [@joseEvaluatingQuantileAssessments2009]).
+Another way to refer to this property is to call the scoring function $s_{\alpha}$ \emph{strictly consistent} for the
+$\alpha$ quantile functional [@gneitingMakingEvaluatingPoint2011]. Having a strictly consistent scoring function makes the $\alpha$ quantile \emph{elicitable}.
+
 If $y$ is generated by $Y \sim G$ then $S_{\alpha}(F,y)$ is an estimate of 
 \begin{align}
-S_{\alpha}(F,G) &:= E_G[S_{\alpha}(F,Y)] = E_G[s_{\alpha}(Q_{F}(\alpha),Y)].
-\end{align}
-The estimand here is the value at $G$ of the functional $\mathcal{\overline{S}}$ on $\mathcal{F_0}$ given by
-\begin{align}
-\mathcal{\overline{S}}_{G,\alpha}(F) := \sbar_G(Q_{\alpha}(F))
-\end{align}
-where $Q_{\alpha}$ is the $\alpha$ quantile functional. This is the expected value under $s$ of the forecaster's $\alpha$ quantile under the \emph{actual} data generating process $G$.
+S_{\alpha}(F,G) &:= E_G[S_{\alpha}(F,Y)] = E_G[s_{\alpha}(Q_{F}(\alpha),Y)],
+\end{align}
+the expected value under $s_{\alpha}$ of the forecaster's $\alpha$ quantile under the \emph{actual} data generating process $G$.
+More abstractly, this estimand is the value at $F$ of the functional $\mathcal{\overline{S}}_{G,\alpha}$ on $\mathcal{F_0}$ given by
+\begin{align}
+\mathcal{\overline{S}}_{G,\alpha}(F) := \sbar_{G,\alpha}(Q_{\alpha}(F))
+\end{align}
+where $Q_{\alpha}$ is the $\alpha$ quantile functional.
 
@@
 \begin{align}
-\sbar_F(x) &= O \int_{\mathbb{R}} \int_{\mathbb{R}} \mathbb{1}\{Y \leq u < x\}dg(u) dF(y) + U \int_{\mathbb{R}} \int_{\mathbb{R}} \mathbb{1}\{x \leq u < Y\}dg(u) dF(y) \\
+\sbar_{F,O,U}(x) &= O \int_{\mathbb{R}} \int_{\mathbb{R}} \mathbb{1}\{Y \leq u < x\}dg(u) dF(y) + U \int_{\mathbb{R}} \int_{\mathbb{R}} \mathbb{1}\{x \leq u < Y\}dg(u) dF(y) \\
 &= O \int_{\mathbb{R}} \mathbb{1}\{u < x\} F(u) dg(u) + U \int_{\mathbb{R}} \mathbb{1}\{u \geq x\} (1-F(u)) dg(u) \\
@@
 \begin{align}
-\nonumber \sbar_F(x) &= O\int_{-\infty}^x(g(x) - g(y))dF(y) + U\int_x^{\infty}(g(y) - g(x))dF(y)\\ 
+\nonumber \sbar_{F,O,U}(x) &= O\int_{-\infty}^x(g(x) - g(y))dF(y) + U\int_x^{\infty}(g(y) - g(x))dF(y)\\ 
 &= O\left(g(x)F(x) - \int_{-\infty}^x g(y)dF(y) \right) + U \left( \int_x^{\infty} g(y)dF(y) - g(x)(1-F(x))\right) \\
@@
 \begin{align}
-\sbar_{\mathrm{Exp}(1/\sigma)}(x) 
+\sbar_{\mathrm{Exp}(1/\sigma), \kappa, \alpha}(x) 
 &= \kappa\left[x(\mathbb{1}\{x \geq 0\} \left[1 -e^{-x / \sigma}\right]  - \alpha) - \mathbb{1}\{x \geq 0\} \left[\sigma-e^{-x / \sigma}(\sigma+x)\right] + \alpha\sigma\right] \\
@@
 \begin{align}
-\sbar_F(x) &= \kappa\left[x(F(x) - \alpha) - \mu F(x) - \sigma \int_{-\infty}^{\frac{x-\mu}{\sigma}}zdF_0(z)  + \alpha E_F[Y]\right] \\
+\sbar_{F,\kappa,\alpha}(x) &= \kappa\left[x(F(x) - \alpha) - \mu F(x) - \sigma \int_{-\infty}^{\frac{x-\mu}{\sigma}}zdF_0(z)  + \alpha E_F[Y]\right] \\
 &= \kappa\left[F(x)(x-\mu) - \alpha(x - E_F[Y]) - \sigma \int_{-\infty}^{\frac{x-\mu}{\sigma}}zdF_0(z)\right] \label{eqn:lc-form}
@@
 \begin{align}
-\sbar_F(x) &= 
+\sbar_{F, \kappa,\alpha}(x) &= 
 \begin{cases}
@@
 
-Thus, outside $(a,b)$, $\sbar_{U[a.b]}(x)$ has the form of pinball loss centered on the mean $m_F = \frac{a+b}{2}$ and is interpolated on $(a,b)$ 
+Thus, outside $(a,b)$, $\sbar_{U[a.b],\kappa,\alpha}(x)$ has the form of pinball loss centered on the mean $m_F = \frac{a+b}{2}$ and is interpolated on $(a,b)$ 
  by a quadratic centered at the quantile $a+\alpha(b-a)$ with slopes matching the linear tails at $a$ and $b$ (see figure 1).
@@
 \begin{align}
-\sbar_F(x) = \kappa\left[(\Phi_{\mu,\sigma}(x) - \alpha)(x-\mu) + \sigma^2 \varphi_{\mu, \sigma}(x)\right].
+\sbar_{F,\kappa,\alpha}(x) = \kappa\left[(\Phi_{\mu,\sigma}(x) - \alpha)(x-\mu) + \sigma^2 \varphi_{\mu, \sigma}(x)\right].
 \end{align}
@@
 \begin{align}
-\sbar_F(x) &=  O \int_{-\infty}^{g(x)} F(g^{-1}(u)) du + U \int_{g(x)}^{\infty}(1-F(g^{-1}(u))) du
+\sbar_{F,O,U}(x) &=  O \int_{-\infty}^{g(x)} F(g^{-1}(u)) du + U \int_{g(x)}^{\infty}(1-F(g^{-1}(u))) du
 \end{align}
@@
 
-\subsection{Alternative Notation Possibilities}
-
-Let $Y \sim F,F_i$ be various forecaster beliefs and $Y \sim G$ be reality.
-\begin{align}
+\subsection{Entropy and Divergence}
+
+Let $F,F_i$ be general distributions such as forecaster beliefs or reports about $Y$, and $G$ be the true distribution of $Y$. 
+For a general scoring rule $S$, the $S$-entropy of $F$ is defined as 
+\begin{align}
+e_S(F) &:= S(F, F) = E_F[S(F, Y)].
+\end{align}
+Applying this to $S_\alpha$, we can define the $g,\alpha$-entropy of $F$ as
+\begin{align}
+e_{\alpha}(F) &:= S_{\alpha}(F, F) = E_F[s_{\alpha}(q_{F,\alpha}, Y)] .
+\end{align}
+If a forecaster believes that $Y \sim F$ then they expect to receive a score of $e_{\alpha}(F)$ by reporting their $\alpha$ quantile.
+According to \eqref{eqn:scalar_obj_cdf_form} and \eqref{eqn:obj_pexp},
+\begin{align}
+e_{\alpha}(F)
+&= (1-\alpha) \int_{-\infty}^{q_{\alpha,F}} F(u) dg(u) + \alpha \int_{{q_{\alpha,F}}}^{\infty}(1-F(u)) dg(u) \\
+&= \alpha \overline{g}\sb{F}-\overline{g}_{F}(x).
+\end{align}
+Similarly, we define the $g,\alpha$-divergence between two distributions as
+\begin{align}
+d_{\alpha}(F_1,F_2) &:= S_{\alpha}(F_1,F_2) - S_{\alpha}(F_2,F_2) = S_{\alpha}(F_1,F_2) - e_{\alpha}(F_2). \label{egn:divergence}
+\end{align}
+Again from \eqref{eqn:scalar_obj_cdf_form} we get
+\begin{align}
+d_{\alpha}(F_1,F_2)
+&= (1-\alpha) \int_{-\infty}^{q_{\alpha,F_1}} F_2(u) dg(u) + \alpha \int_{{q_{\alpha,F_1}}}^{\infty}(1-F_2(u)) dg(u) \\
+&-(1-\alpha) \int_{-\infty}^{q_{\alpha,F_2}} F_2(u) dg(u) - \alpha \int_{{q_{\alpha,F_2}}}^{\infty}(1-F_2(u)) dg(u) \\
+&= (1-\alpha) \int_{q_{\alpha,F_2}}^{q_{\alpha,F_1}} F_2(u) dg(u) + \alpha \int_{{q_{\alpha,F_1}}}^{q_{\alpha,F_2}}(1-F_2(u)) dg(u).
+\end{align}
+And from \eqref{eqn:obj_pexp},
+\begin{align}
+d_{\alpha}(F_1,F_2) &= g(q_{\alpha,F_1})(F_2(q_{\alpha,F_1}) - \alpha) -\overline{g}_{F_2}(q_{\alpha,F_1}) + \alpha \overline{g}\sb{F_2}  - \alpha \overline{g}\sb{F_2} + \overline{g}_{F_2}(q_{\alpha,F_2}) \\
+&= g(q_{\alpha,F_1})(F_2(q_{\alpha,F_1}) - \alpha) + \int_{q_{\alpha,F_1}}^{q_{\alpha,F_2}}g(y)dF_2(y)
+\end{align}
+To emphasize the reality of $G$ we write $e_{\alpha}(G)$ as $e_{\alpha}(Y)$, which with \eqref{egn:divergence}
+gives 
+\begin{align}
+S_{\alpha}(F,G) = d_{\alpha}(F,G) + e_{\alpha}(Y). \label{eqn:marg_score_decomp}
+\end{align}
+This expresses the true expected score of $F$ 
+(i.e., not that calculated by the forecaster) as the sum of the failure of $F$ to match $G$ and the inherent uncertainty of $Y$.
+
+We can also replace $G$ in \eqref{eqn:marg_score_decomp} by $G|F$, the true distribution of $Y$ \emph{given} that the forecaster 
+believes $Y \sim F$---a belief presumably based on some knowledge of leading indicators for $Y$. This gives
+\begin{align}
+S_{\alpha}(F,G|F) &= d_{\alpha}(F,G|F) + e_{\alpha}(Y|F)
+\end{align}
+Using
+\begin{align}
+d_{\alpha}(G,G|F) = S_{\alpha}(G,G|F) - e_{\alpha}(Y|F)
+\end{align}
+this can be expanded to
+\begin{align}
+S_{\alpha}(F,G|F) &= d_{\alpha}(F,G|F) - d_{\alpha}(G,G|F) + S_{\alpha}(G,G|F). \label{eqn:cond_score_decomp}
+\end{align}
+Now view $F$ explicitly as a realization of a distribution-valued random variable $\mathcal{F}$ associated with the forecaster. 
+To measure the forecaster's perfomance in general and not just on the occasion when they form 
+the belief $Y\sim F$, we can take the $\mathcal{F}$ expectation of \eqref{eqn:cond_score_decomp} to get a decomposition
+\begin{align}
+E_{\mathcal{F}} [S_{\alpha}(F,G|F)] &= E_{\mathcal{F}} [d_{\alpha}(F,G|F)] - E_{\mathcal{F}}[d_{\alpha}(G,G|F)] + e_{\alpha}(Y).
+\end{align}
+The first two terms on the RHS are known as the \emph{unreliability} and the \emph{resolution} of the forecaster.
+
+<!-- {}
+\begin{align}{}
 \overline{s}_{F,\kappa,\alpha}(x) &:= s_{\kappa,\alpha}(x,F) := E_F[s_{\kappa,\alpha}(x,Y)] \\
@@
 \end{align} 
-
-Define the $s$-entropy of an arbitrary $F$ 
-\begin{align}
-e_s(F) &:= E_F[S(F, Y)] = E_F[s(q_{F,\alpha}, Y)] = s(q_{F,\alpha}, F) \\
-\end{align}
-and of the RV $Y$ as
-\begin{align}
-e_s(Y) &:= e_s(G).
-\end{align}
-Define the $s$-divergence between two forecasts as
-\begin{align}
-d_s(F_1,F_2) &:= S(F_1,F_2) - S(F_2,F_2) = S(F_1,F_2) - e_s(F_2)
-\end{align}
-and the $s$-divergence from $F$ to $Y$ as
-\begin{align}
-d_s(F,Y) &:= S(F,G) - e(Y).
-\end{align}
-
-Thus, $\overline{S}(F) = d_s(F,Y) + e_s(Y)$.
+-->
 
@@
 \begin{align}
-s_n(x,y) & = a_x x +a_y y +a_{+}(x-y)\sb{+} + a_{-}(y-x)\sb{+}.
-\end{align}
-We might have, for example, the score given by net value transfers
-\begin{align}
-s_n(x,y) = cx-py - u(x-y)\sb{+} + (p+r)(y-x)\sb{+}
-\end{align}
-where $x$ newspapers are ordered at price $c$, $y - (y-x)\sb{+}$ are sold at retail price $p$, $(x-y)\sb{+}$ are sold at salvage price $u$, 
-and $r(y-x)\sb{+}$ of customer "goodwill" is lost. 
-This can be rewritten as
-\begin{align}
-s_n(x,y) &= (a_{+} + a_x)(x-y)\sb{+} + (a_{-} - a_x)(y-x)\sb{+} +(a_x + a_y)y \\
-&= (c-u)(x-y)\sb{+} + (p+r-c)(y-x)\sb{+} -(p-c)y
+s_{\mathbf{n}}(x,y) & = c x - p \min(x,y) - u(x-y)\sb{+} + r(y-x)\sb{+} \label{eqn:news_score}
+\end{align}
+where $\mathbf{n} = \{c,p,u,r\}$.
+This score is netting the monetary transfers that occur when
+where $x$ newspapers are ordered at price $c$, $\min(y,x)$ are sold at retail price $p$, $(x-y)\sb{+}$ are sold at salvage price $u$, 
+and $r(y-x)\sb{+}$ of monetarily valued customer "goodwill" is lost. 
+Using $x = y + (x-y)_+ - (y-x)_+$ and $\min(x,y) = y - (y-x)_+$ \eqref{eqn:news_score} can be rewritten as
+\begin{align}
+s_{\mathbf{n}}(x,y) &= (c-u)(x-y)\sb{+} + (p-c+r)(y-x)\sb{+} -(p-c)y
 \end{align}
@@
 \begin{align}
-\sbar_{n,F}(x) = (a_{+} + a_x)E_F[(x-Y)\sb{+}] + (a_{-} - a_x)E_F[(Y-x)\sb{+}] +(a_x + a_y)E[Y].
-\end{align}
-This differs by a constant from our $\sbar_F(x)$ in \eqref{eqn:scalar_obj} after taking $g(x)=x$, $O=a_{+} + a_x=c-u$ 
-and $U=a_{-} - a_=p+r-c$, and so has the same minimizer   
-\begin{align}
-Q = F^{-1}\left(\frac{U}{O+U}\right) =  F^{-1}\left(\frac{a_{-} - a_x}{a_{-} + a_{+}}\right) 
-= F^{-1}\left(\frac{p+r-c}{p+r-u}\right).
-\end{align} 
-Note, for reference, that in the literature $\sbar_{n,F}(x)$ is often encountered in the form
-\begin{align}
-&(O+U) \int_{0}^{x} (x - y)f(y)dy - Ux+ (U+a_x+a_y)E[Y] \\{}
-=\,&(p+r-u) \int_{0}^{x} (x - y)f(y)dy - (p+r-c)x+ rE[Y]
+\sbar_{F, {\mathbf{n}}}(x) = (c-u)E_F[(x-Y)\sb{+}] + (p-c+r)E_F[(Y-x)\sb{+}] -(p-c)E[Y].
+\end{align}
+This differs by a constant from our $\sbar_{F,O,U}(x)$ in \eqref{eqn:scalar_obj} with $g(x)=x$, $O=c-u$ 
+and $U=p-c+r$, and so has the same minimizer   
+\begin{align}
+Q = F^{-1}\left(\frac{U}{O+U}\right) = F^{-1}\left(\frac{p-c+r}{p-u + r}\right).
+\end{align} 
+Note, for reference, that in the literature $\sbar_{F, {\mathbf{n}}}(x)$ is often encountered in the form
+\begin{align}
+&(O+U) \int_{0}^{x} (x - y)dF(y) - Ux+ (U-u+r)E[Y] \\{}
+=\,&(p+r-u) \int_{0}^{x} (x - y)dF(y) - (p-c+r)x+ rE[Y]
 \end{align}
@@
 
-As a separable problem, this is amenable to being solved via its dual (see [@ruszczynski2011nonlinear], section 4.4). With the Lagrangian
+As a separable problem, this is amenable to being solved via its dual (see [@ruszczynskiNonlinearOptimization2011], section 4.4). With the Lagrangian
 \begin{align}

### 2023-06-12 ISF_talk/talk.qmd
add Norm1, a normal simulation experiment

---
title: "alloscore, ISF 43"
format:
  revealjs: 
    width: 1500
editor: visual
---

Comparitive vs absolute assessment/evaluation

illustrate algo with a lambda over xs plot against tau

### 2023-06-12 random_undocumented_experiments/Norm1.R

@@ -1,0 +1,69 @@
+library(alloscore)
+library(tidyverse)
+
+dgp_params <- tibble(
+  model = "dgp",
+  target_names = paste0("site", 1:2),
+  dist = "norm",
+  mean = c(100,25),
+  sd = 5)
+
+dgp <- dgp_params %>% add_pdqr_funs()
+
+y_gen <- function() {with(dgp, map2_dbl(r, 1, exec) %>% set_names(target_names))}
+
+modA <- dgp_params %>% mutate(
+  model = "A",
+  mean = mean[c(2,1)],
+  sd = 2,
+) %>% add_pdqr_funs(types = c("p","q"))
+
+modB <- dgp_params %>% mutate(
+  model = "B",
+  mean = .5 * mean,
+  sd = 20
+) %>% add_pdqr_funs(types = c("p","q"))
+
+modC <- dgp_params %>% mutate(
+  model = "C",
+  sd = c(3,20),
+) %>% add_pdqr_funs(types = c("p","q"))
+
+mods <- bind_rows(dgp, modA, modB, modC)
+
+Ks <- seq(10,500, by = 10)
+allos <- mods %>% group_nest(model) %>%
+  mutate(allocated = map(data, ~allocate(.x, K= Ks)))
+allos <- allos %>% mutate(slim = map(allocated, make_slim))
+
+plot_allos <- allos %>% select(-c(data, allocated)) %>%
+  unnest(slim) %>%
+  select(-score_fun) %>%
+  pivot_wider(values_from = x, names_from = target_names) %>%
+  mutate(excess = site1 + site2 - K)
+
+plot_allos %>% ggplot(aes(x=site1, y=site2, color = model)) +
+  geom_line() + theme(aspect.ratio = 1)
+
+ys <- map(1:1000, ~y_gen())
+
+allos <- allos %>% mutate(slim_scored = map(slim, ~alloscore(., ys)))
+
+plot_dat <- allos %>%
+  mutate(scores = map(slim_scored, ~select(.,-xdf) %>% unnest(scores))) %>%
+  select(-c(data, allocated, slim, slim_scored)) %>%
+  unnest(scores)
+
+(p <- plot_dat %>% filter(K %in% seq(20, 180, by = 20)) %>% ggplot(aes(x = score, fill = model)) +
+    geom_histogram(alpha = 0.5, position = 'identity', bins = 100) +
[…]

### 2023-06-13 notes/allocation_notes.rmd

@@
 Let $F,F_i$ be general distributions such as forecaster beliefs or reports about $Y$, and $G$ be the true distribution of $Y$. 
-For a general scoring rule $S$, the $S$-entropy of $F$ is defined as 
-\begin{align}
-e_S(F) &:= S(F, F) = E_F[S(F, Y)].
+For a general scoring rule $S$, we define the $S$-entropy of $F$ as 
+\begin{align}
+e_S(F) &:= S(F, F) = E_F[S(F, Y)]
+\end{align}
+and the $S$-entropy of the random variable $Y$ as
+\begin{align}
+e_S(Y) &:= e_S(G).
 \end{align}
@@
 \end{align}
-If a forecaster believes that $Y \sim F$ then they expect to receive a score of $e_{\alpha}(F)$ by reporting their $\alpha$ quantile.
+A forecaster believing that $Y \sim F$ will expect to receive a score of $e_{\alpha}(F)$ by reporting their $\alpha$ quantile, while an
+omniscient forecaster \emph{knowing} that $Y\sim G$ will expect a score of $e_S(Y)$ for reporting $q_{\alpha,Y}$.
 According to \eqref{eqn:scalar_obj_cdf_form} and \eqref{eqn:obj_pexp},
@@
 &= (1-\alpha) \int_{-\infty}^{q_{\alpha,F}} F(u) dg(u) + \alpha \int_{{q_{\alpha,F}}}^{\infty}(1-F(u)) dg(u) \\
-&= \alpha \overline{g}\sb{F}-\overline{g}_{F}(x).
+&= \alpha \overline{g}\sb{F}-\overline{g}_{F}(q_{\alpha,F}).
 \end{align}
@@
 \end{align}
-To emphasize the reality of $G$ we write $e_{\alpha}(G)$ as $e_{\alpha}(Y)$, which with \eqref{egn:divergence}
-gives 
-\begin{align}
-S_{\alpha}(F,G) = d_{\alpha}(F,G) + e_{\alpha}(Y). \label{eqn:marg_score_decomp}
-\end{align}
-This expresses the true expected score of $F$ 
+With \eqref{egn:divergence}, we get 
+\begin{align}
+S_{\alpha}(F,G) = d_{\alpha}(F,G) + e_{\alpha}(Y), \label{eqn:marg_score_decomp}
+\end{align}
+expressing the true expected score of $F$ 
 (i.e., not that calculated by the forecaster) as the sum of the failure of $F$ to match $G$ and the inherent uncertainty of $Y$.
@@
 \begin{align}
-s_{\mathbf{n}}(x,y) & = c x - p \min(x,y) - u(x-y)\sb{+} + r(y-x)\sb{+} \label{eqn:news_score}
+s_{\mathbf{n}}(x,y) & = c x - p \min(x,y) - u(x-y)\sb{+} + r(y-x)\sb{+} \label{eqn:news_score} \\
+&= \begin{cases}
+cx - px + r(y-x), & x \leq y \\
+cx - py - u(x-y), & x > y
+\end{cases}
 \end{align}
@@
 where $x$ newspapers are ordered at price $c$, $\min(y,x)$ are sold at retail price $p$, $(x-y)\sb{+}$ are sold at salvage price $u$, 
-and $r(y-x)\sb{+}$ of monetarily valued customer "goodwill" is lost. 
+and the "goodwill", monetarily valued at $r$, of $(y-x)\sb{+}$ customers is lost. 
 Using $x = y + (x-y)_+ - (y-x)_+$ and $\min(x,y) = y - (y-x)_+$ \eqref{eqn:news_score} can be rewritten as
@@
 \begin{align}
-\sbar_{F, {\mathbf{n}}}(x) = (c-u)E_F[(x-Y)\sb{+}] + (p-c+r)E_F[(Y-x)\sb{+}] -(p-c)E[Y].
+\sbar_{F, {\mathbf{n}}}(x) = (c-u)E_F[(x-Y)\sb{+}] + (p-c+r)E_F[(Y-x)\sb{+}] -(p-c)E_F[Y].
 \end{align}
@@
 \begin{align}
-Q = F^{-1}\left(\frac{U}{O+U}\right) = F^{-1}\left(\frac{p-c+r}{p-u + r}\right).
+q_{F,O,U} = F^{-1}\left(\frac{U}{O+U}\right) = F^{-1}\left(\frac{p-c+r}{p-u + r}\right).
 \end{align} 
@@
 \begin{align}
-&(O+U) \int_{0}^{x} (x - y)dF(y) - Ux+ (U-u+r)E[Y] \\{}
-=\,&(p+r-u) \int_{0}^{x} (x - y)dF(y) - (p-c+r)x+ rE[Y]
+&(O+U) E_F[(x-Y)\sb{+}] + U E_F[Y] E_F[(Y-x)\sb{+}-(x-Y)\sb{+}] -(p-c)E_F[Y] \\
+&(O+U) E_F[(x-Y)\sb{+}] + U (E_F[Y] - x) -(p-c)E_F[Y]\\
+=\,&(p+r-u) \int_{0}^{x} (x - y)dF(y) - (p-c+r)x+ rE_F[Y]
 \end{align}

### 2023-06-14 ISF_talk/talk.qmd
more entropy/divergence work.  getting slide started.

@@
 
-Comparitive vs absolute assessment/evaluation
+## The backstory
+
+Over last decade, multicenter effort to collect forecasts via hubs for ID outbreaks (epi/pan/endemic)
+with central purpose to "inform public health responses"
+
+With uncertainty quantification being well-established as needed DM input, 
+hubs have trended toward eliciting probabilistic forecasts
+
+Basic element of any such response is an *allocation decision* regarding
+ - medical supplies
+ - personnel 
+ - facility capacity
+ - R&D funding
+ - public tolerance for NPIs
+
+Allocations can be wrt space, time, demographic
+
+Similar to other contexts where there are large-scale institutional efforts to use expert judgment to assess and manage social risk
+ - finance
+ - meteorology
+ - military conflict
+
+Even with complex game-theoretic approachs to collective decision making,
+*allocation* almost always a requirement to strategy development
+
+## Scoring rules and social utility
+
+Hubs have tried to use SRs to rank and combine forecasts in ways that offer best suppport for PH responders
+
+This can be cast in decision-theoretic framework of utility, loss functions, and Bayes acts.
+
+The DM needs to decide what action $a$ to take regarding an uncertain outcome $Y$ such as the number of severe cases in various locations.
+Suppose the DM also has
+a loss function $l(a,y)$, e.g., the unmet need when $a$ beds are allocated for $y$ severe cases,
+and a utility function U quantifying, e.g., the ratio of the value of preventing 1000 unhospitalized severe cases versus the value of preventing 10 unhospitalized severe cases.
+That is U specifies numbers such as U(l(10,20)) / U(l(1000,2000)).
+Then the Bayes act relative to a forecast $Y \sim F$ is the action $a^F$ that will maximize the DM’s expected utility assuming that in reality $Y \sim F$.
+
+## Basic computational approach
+
+Adopt an iterative Lagrange multiplier method following long history in inventory management of solving
+*constrained multi-product newsvendor* problems
 
 illustrate algo with a lambda over xs plot against tau
+
+Comparitive vs absolute assessment/evaluation

### 2023-06-14 notes/allocation_notes.rmd
more entropy/divergence work.  getting slide started.

@@
 \begin{align}
-e_{\alpha}(F) &:= S_{\alpha}(F, F) = E_F[s_{\alpha}(q_{F,\alpha}, Y)] .
-\end{align}
-A forecaster believing that $Y \sim F$ will expect to receive a score of $e_{\alpha}(F)$ by reporting their $\alpha$ quantile, while an
-omniscient forecaster \emph{knowing} that $Y\sim G$ will expect a score of $e_S(Y)$ for reporting $q_{\alpha,Y}$.
+e_{\alpha}(F) &:= S_{\alpha}(F, F) = E_F[s_{\alpha}(q_{\alpha,F}, Y)] .
+\end{align}
+A forecaster believing that $Y \sim F$ will expect to receivq_e a score of $e_{\alpha}(F)$ by reporting their $\alpha$ quantile, while an
+omniscient forecaster \emph{knowing} that $Y\sim G$ will expect a score of $e_S(Y)$ for reporting $q_{\alpha,Y}$. Note that if 
+$F_{y_0} = \mathbb{1}\{x \geq y_0\}$ is deterministic, then 
+\begin{align}
+e_{\alpha}(F_{y_0}) = E_{F_{y_0}}[s_{\alpha}(q_{\alpha, F_{y_0}}, Y)] = s_{\alpha}(y_0, y_0) = 0. \label{eqn:entropy_determ}
+\end{align}
+Similarly, if $F_{\mathbf{y}_n} = \mathrm{Unif}\{y_1, \ldots, y_n\}$ is the empirical distribution of a sample $\mathbf{y}_n = \{y_1, \ldots, y_n\}$, then
+\begin{align}
+e_{\alpha}(F_{\mathbf{y}_n}) = E_{F_{\mathbf{y}_n}}[s_{\alpha}(q_{\alpha, F_{\mathbf{y}_n}}, Y)] 
+& = \frac{1}{n}\sum_{i=1}^{n} s_{\alpha}(y_{\lceil \alpha n \rceil}, y_i) \\
+& = \frac{1}{n} \left\{(1-\alpha)\sum_{i=\lceil \alpha n \rceil}^{n}g(y_i) - g(y_{\lceil \alpha n \rceil}) + 
+\alpha \sum_{i=1}^{\lceil \alpha n \rceil} g(y_{\lceil \alpha n \rceil}) - g(y_i) \right\}
+\end{align}
+which we also call the $g,\alpha$-entropy $e_{\alpha}(\mathbf{y}_n)$ of the sample $\mathbf{y}_n$.
+
 According to \eqref{eqn:scalar_obj_cdf_form} and \eqref{eqn:obj_pexp},
@@
 \end{align}
+
 Similarly, we define the $g,\alpha$-divergence between two distributions as
@@
 \end{align}
-Again from \eqref{eqn:scalar_obj_cdf_form} we get
+If $F_2 = F_{y_0}$ is deterministic, then from \eqref{eqn:entropy_determ} we have 
+\begin{align}
+d_{\alpha}(F_1, F_{y_0}) = S_{\alpha}(F_1, F_{y_0}) = E_{F_{y_0}}[S_{\alpha}(F_1, Y)] = S_{\alpha}(F_1, y_0) = s_{\alpha}(q_{\alpha, F_1}, y_0)
+\end{align}
+whereas if $F_1 = F_{y_0}$, we get
+\begin{align}
+d_{\alpha}(F_{y_0}, F_2) = E_{F_2}[s_{\alpha}(y_0,Y)] - e_{\alpha}(F_2).
+\end{align}
+When $F_2 = F_{\mathbf{y}_n}$ is an empirical distribution, then 
+\begin{align}
+d_{\alpha}(F_1, F_{\mathbf{y}_n}) 
+&= E_{F_{\mathbf{y}_n}}[S_{\alpha}(F_1, Y)] - e_{\alpha}(\mathbf{y}_n) \\
+&= \frac{1}{n}\sum_{i=1}^{n} s_{\alpha}(q_{\alpha,F_1}, y_i) - s_{\alpha}(y_{\lceil \alpha n \rceil}, y_i)
+\end{align}
+whereas inserting the $F_{\mathbf{y}_n}$ in the first position gives
+\begin{align}
+d_{\alpha}(F_{\mathbf{y}_n}, F_2) = E_{F_2}[s_{\alpha}(y_{\lceil \alpha n \rceil},Y)] - e_{\alpha}(F_2).
+\end{align}
+
+<!--
+&= (1-\alpha)(g(q_{\alpha, F_1}) - g(y_{\lceil \alpha n \rceil})) \\
+&+ \frac{1}{n} \left\{\sum_{i \in I_1} g(y_i) - g(q_{\alpha, F_1})
++\sum_{i \in I_2} g(q_{\alpha, F_1}) - g(y_i) 
++\sum_{i \in I_3} g(y_{\lceil \alpha n \rceil})) - g(q_{\alpha, F_1}) \right\}
+ \text{ using \eqref{eqn:scalar_score_OR}}
+where $I_1 = \{i | q_{\alpha, F_1} \leq y_i \leq y_{\lceil \alpha n \rceil}\}$, 
+$I_2 = \{i | y_{\lceil \alpha n \rceil} \leq y_i \leq q_{\alpha, F_1}\}$, and 
+$I_3 = \{i | y_{\lceil \alpha n \rceil}, q_{\alpha, F_1} \leq y_i\}$.
+-->
+
+As for entropy, we also can get more explicit forms for the divergence in terms of distributions $F_1, F_2$. 
+From \eqref{eqn:scalar_obj_cdf_form} we get
 \begin{align}
@@
 d_{\alpha}(F_1,F_2) &= g(q_{\alpha,F_1})(F_2(q_{\alpha,F_1}) - \alpha) -\overline{g}_{F_2}(q_{\alpha,F_1}) + \alpha \overline{g}\sb{F_2}  - \alpha \overline{g}\sb{F_2} + \overline{g}_{F_2}(q_{\alpha,F_2}) \\
-&= g(q_{\alpha,F_1})(F_2(q_{\alpha,F_1}) - \alpha) + \int_{q_{\alpha,F_1}}^{q_{\alpha,F_2}}g(y)dF_2(y)
-\end{align}
+&= g(q_{\alpha,F_1})(F_2(q_{\alpha,F_1}) - \alpha) + \int_{q_{\alpha,F_1}}^{q_{\alpha,F_2}}g(y)dF_2(y).
+\end{align}
+This offers a more explicit form of the divergence of a distribution from s sample $\mathbf{y}_n$:
+\begin{align}
+d_{\alpha}(F_1,F_{\mathbf{y}_n}) 
+= g(q_{\alpha,F_1})\left(\frac{1}{n}\sum_{1=1}^n \mathbb{1}\{y_i \leq q_{\alpha,F_1}\} - \alpha\right) 
++ \frac{1}{n}\sum_{1=1}^{\lceil \alpha n \rceil} \mathbb{1}\{y_i > q_{\alpha,F_1}\}g(y_i).
+\end{align}
+
 With \eqref{egn:divergence}, we get 
@@
 \sum_{i}^{\mathrm{dim}\mathbf{x}} s_{\kappa_i,\alpha_i}(x_i, F_i) \\
-S(F,y) &:= s(q_{F,\alpha}, y) \\
-S(F_1,F_2) &:= E_{F_2}[S(F_1,Y)] = s(q_{F_1,\alpha}, F_2) \\
-\overline{S}(F) &:= E_G[s(q_{F, \alpha}, Y)]
+S(F,y) &:= s(q_{\alpha, F}, y) \\
+S(F_1,F_2) &:= E_{F_2}[S(F_1,Y)] = s(q_{\alpha, F_1}, F_2) \\
+\overline{S}(F) &:= E_G[s(q_{ \alpha,F}, Y)]
 \end{align} 
@@
 \begin{align}
-q_{F,O,U} = F^{-1}\left(\frac{U}{O+U}\right) = F^{-1}\left(\frac{p-c+r}{p-u + r}\right).
+q_{O,U, F} = F^{-1}\left(\frac{U}{O+U}\right) = F^{-1}\left(\frac{p-c+r}{p-u + r}\right).
 \end{align} 

### 2023-06-14 random_undocumented_experiments/iter_plot.R
more entropy/divergence work.  getting slide started.

@@ -1,0 +1,20 @@
+
+iters_bp <- adf %>% filter(model == "BPagano-RtDriven", K == 2000) %>%
+  select(model, reference_date, xs) %>% unnest(xs) %>%
+  mutate(init_alloc = ntile(`1`, 5), .after = abbreviation)
+
+keep_states <- iters_bp %>% filter(reference_date==min(reference_date)) %>%
+  filter(min_rank(desc(`1`))<40) %>% pull(abbreviation)
+
+iters_bp %>% select(-qs) %>%
+  tidyr::pivot_longer(
+    cols = -c(1:4),
+    names_to = "iteration",
+    values_to = "allocation") %>%
+  mutate(
+    iteration = as.numeric(iteration)
+  ) %>%
+  filter(iteration < 8, abbreviation %in% keep_states) %>%
+  ggplot(aes(x = iteration, y = allocation, color = abbreviation, alpha = init_alloc)) +
+  facet_grid(cols = vars(reference_date)) +
+  geom_line() + geom_point() + theme_classic()

### 2023-06-16 ISF_talk/talk.qmd
current state of isf talk

--- 2023-06-12 ISF_talk/talk.qmd
@@
 
-Comparitive vs absolute assessment/evaluation
+## Timeline
+
+::: notes
+West Nile mosq control decisions
+:::
+
+## The Backstory
+
+Starting \~ 2010? Outbreak forecasting hubs developed to "inform public health responses"
+
+Trend toward elicitation of probabilistic forecasts
+
+ - uncertainty quantification often essential to a public health **DM** (decision maker)
+
+But in the inimitable words of chatGPT:
+
+> It's important to note that while these efforts represent significant advances in the field, forecasting the spread and impact of infectious diseases is a very complex task with many challenges, and these models should be interpreted with caution.
+
+::: notes
+Over last two decades, multicenter effort to collect forecasts via hubs for ID outbreaks (epi/pan/endemic) with central purpose to "inform public health responses"
+
+With uncertainty quantification being well-established as needed DM input, 
+:::
+
+## The fundamental need to decide allocations
+
+::: columns
+::: {.column width="40%"}
+Basic element of any such response is an *allocation decision* regarding
+
+ - medical supplies 
+ - personnel 
+ - facility capacity 
+ - R&D funding 
+ - public tolerance for NPIs
+:::
+
+::: {.column width="60%"}
+Similar to other contexts where there are large-scale institutional efforts to use expert judgment to assess and manage social risk
+
+ - finance 
+ - meteorology 
+ - military conflict
+:::
+:::
+
+::: notes
+Allocations can be wrt space, time, demographic
+
+Even with complex game-theoretic approachs to collective decision making, *allocation* almost always a requirement to strategy development
+:::
+
+## Scoring rules and social utility
+
+Hubs have tried to use SRs to rank and combine forecasts in ways that offer best suppport for PH responders
+
+This can be cast in decision-theoretic framework of utility, loss functions, and Bayes acts.
+
+The DM needs to decide take an action $a$ wrt an outcome $Y$
+
+- such as the number of severe cases in various locations.
+
+Suppose the DM also has a *loss function* $l(a,y)$ - e.g., the lost social/political/economic/psychological/moral utility caused by unmet need when $a < y$ beds are allocated and $y$ severe cases occur - this must somehow acknowledge risk attitude heterogeneity among various DM's that a forecast hub might serve
+
+Then the *Bayes act* relative to a forecast $Y \sim F$ is the action $a^F$ that will minimize the DM's expected loss assuming that in reality $Y \sim F$.
+
+------------------------------------------------------------------------
+
+A Scoring Rule attempts to transfer or "share" this loss function with the forecaster through a numerical reward system where FP receives
+
+$$
+S(F,y)
+$$
+
+units of money, goodwill, prestige, etc.
+A central goal in designing $S$ is to create a scenario where the FP beleives she cam maximizes her own expected utility by giving the DM the forecast that the FP believes in actually true and/or best suited for leading the DM toward actions that will maximize their utility.
+An SR that acheieves this is called *proper* and Bayes acts form a linchpin of their operation since the establish an explicit vacabulary of loss in place of vague efforts to rank preformance of various statisticla funcitonals adoptic form statistic inference
+
+## Loss vs Utility
+
+Change in loss but not utility from change in - variant severity (assuming target is infections and decision is number of beds)
+
+Change in utility but not in loss from change in - supply and personnel cost structures - public opinion regarding importance of fatality prevention - knowledge of demographic disparities in consequences of outcomes
+
+## Basic computational approach
+
+Adopt an iterative Lagrange multiplier method following long history in inventory management of solving *constrained multi-product newsvendor* problems
 
 illustrate algo with a lambda over xs plot against tau
+
+## Scoring against an oracle
+
+To get more interpretable allocation scores we adjust by the score of an oracle that allocates *fairly* based on what demands actually occur
+
+Note: The oracle is not the DGP, it is a time-traveller
+
+Comparitive vs absolute assessment/evaluation

### 2023-06-16 R/run-alloscore.R

@@ -1,22 +1,12 @@
-run_alloscore <- function(forecast_data, truth_data, Ks){
+run_alloscore <- function(forecast_data, truth_data, Ks, mkeep, reference_dates){
   require(tidyverse)
   require(alloscore)
   require(distfromq)
 
-  # keep selected models
-  mkeep <- c("BPagano-RtDriven",
-             "COVIDhub-4_week_ensemble",
-             "COVIDhub-baseline",
-             "CU-select",
-             "IHME-CurveFit",
-             "JHUAPL-Bucky",
-             "JHUAPL-Gecko",
-             "MUNI-ARIMA",
-             "USC-SI_kJalpha",
-             "UVA-Ensemble")
-
   ## process forecast data, adding distfromq output
   forecast_data_processed <- forecast_data |>
+    ## forecast dates are different but reference dates are Mondays
+    dplyr::filter(reference_date %in% reference_dates) |>
     dplyr::select(-type) |>
     nest(ps = quantile, qs = value) |>
     relocate(ps, qs) |>

### 2023-06-16 _targets.R

@@ -23,19 +23,31 @@ tar_option_set(
 tar_source(files = c("R/data-ingestion.R", "R/plot-alloscores.R", "R/run-alloscore.R"))
 
 ## create a group of alloscore targets
-values <-
-  tibble(forecast_dates = as.character(seq.Date(
+forecast_dates = as.character(seq.Date(
     as.Date("2021-11-22"), as.Date("2022-02-28"), by = "7 days"
-  )))
-values <- values[c(12,13,14),]
+  ))
+
+forecast_dates_quick <- forecast_dates[c(12,13,14)]
 
 Ks <- seq(from = 2000, to = 7000, by = 500)
+
+# keep selected models
+mkeep <- c("BPagano-RtDriven",
+           "COVIDhub-4_week_ensemble",
+           "COVIDhub-baseline",
+           "CU-select",
+           "IHME-CurveFit",
+           "JHUAPL-Bucky",
+           "JHUAPL-Gecko",
+           "MUNI-ARIMA",
+           "USC-SI_kJalpha",
+           "UVA-Ensemble")
 
 # List of targets:
 list(
   tar_target(
     name = forecast_data,
-    command = get_forecast_data(values$forecast_dates)
+    command = get_forecast_data(forecast_dates)
   ),
   # tar_target(
   #   name = filtered_forecast_data,
@@ -51,7 +63,8 @@ list(
   ),
   tar_target(
     name = alloscore_df,
-    command = run_alloscore(forecast_data, truth_data, K = Ks)
+    command = run_alloscore(forecast_data, truth_data, K = Ks,
+                            mkeep = mkeep, reference_dates = forecast_dates_quick)
   ),
   tar_target(
     name = figure_K_v_alloscore,

### 2023-06-16 random_undocumented_experiments/iter_plot.R

@@ -1,20 +1,25 @@
 
+library(targets)
+tar_load(alloscore_df)
+adf <- alloscore_df
+
 iters_bp <- adf %>% filter(model == "BPagano-RtDriven", K == 2000) %>%
   select(model, reference_date, xs) %>% unnest(xs) %>%
   mutate(init_alloc = ntile(`1`, 5), .after = abbreviation)
 
 keep_states <- iters_bp %>% filter(reference_date==min(reference_date)) %>%
-  filter(min_rank(desc(`1`))<40) %>% pull(abbreviation)
+  filter(min_rank(desc(`1`))<5) %>% pull(abbreviation)
 
-iters_bp %>% select(-qs) %>%
+plot_iter_dat <- iters_bp %>% select(-qs) %>%
   tidyr::pivot_longer(
     cols = -c(1:4),
     names_to = "iteration",
     values_to = "allocation") %>%
-  mutate(
-    iteration = as.numeric(iteration)
-  ) %>%
-  filter(iteration < 8, abbreviation %in% keep_states) %>%
-  ggplot(aes(x = iteration, y = allocation, color = abbreviation, alpha = init_alloc)) +
-  facet_grid(cols = vars(reference_date)) +
-  geom_line() + geom_point() + theme_classic()
+  mutate(iteration = as.numeric(iteration)) %>%
+  mutate(give_color = ifelse(abbreviation %in% keep_states, abbreviation, NA))
+
+(p_iter <- plot_iter_dat %>%
+  ggplot(aes(x = iteration, y = allocation, color = give_color, alpha = 1-.5*is.na(give_color))) +
+  facet_grid(cols = vars(reference_date)) + geom_point() +
+  geom_line() + theme_classic() +
+  scale_alpha_continuous(guide = "none"))

### 2023-06-16 random_undocumented_experiments/Norm1.R
current state of isf talk

@@ -1,40 +1,62 @@
 library(alloscore)
 library(tidyverse)
+library(scoringRules)
+library(patchwork)
+library(ggridges)
+source("random_undocumented_experiments/scoring_helpers.R")
 
 dgp_params <- tibble(
   model = "dgp",
   target_names = paste0("site", 1:2),
   dist = "norm",
-  mean = c(100,25),
-  sd = 5)
+  mean = c(200, 100),
+  sd = 50)
 
-dgp <- dgp_params %>% add_pdqr_funs()
+dgp <- dgp_params %>%
+  add_pdqr_funs() %>%
+  add_crps_funs()
 
 y_gen <- function() {with(dgp, map2_dbl(r, 1, exec) %>% set_names(target_names))}
 
 modA <- dgp_params %>% mutate(
   model = "A",
   mean = mean[c(2,1)],
-  sd = 2,
-) %>% add_pdqr_funs(types = c("p","q"))
+  sd = 10) %>%
+  add_pdqr_funs() %>%
+  add_crps_funs()
 
 modB <- dgp_params %>% mutate(
   model = "B",
   mean = .5 * mean,
-  sd = 20
-) %>% add_pdqr_funs(types = c("p","q"))
+  sd = 20) %>%
+  add_pdqr_funs() %>%
+  add_crps_funs()
 
 modC <- dgp_params %>% mutate(
   model = "C",
-  sd = c(3,20),
-) %>% add_pdqr_funs(types = c("p","q"))
+  sd = c(3,20)) %>%
+  add_pdqr_funs() %>%
+  add_crps_funs()
 
 mods <- bind_rows(dgp, modA, modB, modC)
 
-Ks <- seq(10,500, by = 10)
+# select(mods, model, target_names, f) %>%
+#   pivot_wider(values_from = f, names_from = c(model, target_names))
+
+ys <- tibble(samp = 1:200, ys = map(seq_along(samp), ~y_gen()))
+ytot = map_vec(ys$ys, sum)
+p_y_tot <- data.frame(x = ytot) %>% ggplot(aes(x=x)) +
+  geom_histogram() + labs(x = "ytot")
+p_y_dens <- ys %>% unnest_wider(ys) %>% ggplot(aes(x=site1,y=site2)) + geom_point()
+(y_plot <- p_y_tot / p_y_dens)
+
+K_lims <- quantile(ytot, probs = c(.1, .9)) %>% round(-1)
[…]

### 2023-06-16 ISF_talk/talk.qmd
more isf edits

@@
 
-Suppose the DM also has a *loss function* $l(a,y)$ - e.g., the lost social/political/economic/psychological/moral utility caused by unmet need when $a < y$ beds are allocated and $y$ severe cases occur - this must somehow acknowledge risk attitude heterogeneity among various DM's that a forecast hub might serve
+Suppose the DM also has a **loss function** $l(a,y)$
+
+ - unmet need when $a < y$ beds are allocated and $y$ severe cases occur
 
 Then the *Bayes act* relative to a forecast $Y \sim F$ is the action $a^F$ that will minimize the DM's expected loss assuming that in reality $Y \sim F$.
+
+::: notes
+social/political/economic/psychological/moral utility
+
+this must somehow acknowledge risk attitude heterogeneity among various DM's that a forecast hub might serve
+:::
 

### 2023-06-16 talks/talk2/pres2.qmd

---
title: "alloscore talk 2"
format:
  revealjs: 
    width: 1500
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Agenda

Intuitive understanding of expected unmet/overmet need

-   i.e. of "hockey-stick" or "piece-wise linear" loss $\max(0,Y - x)= (Y-x)_+$ and $\min(0,Y - x)= (Y-x)_-$

-   more broadly, quantile scores and therefore also WIS, CRPS, and who knows what else

-   a meta-theorem going back to Savage: all loss functions (for estimation or prediction) are enhancements of either the squared-error or piece-wise linear error.

My own experience: the more I tried to think about this, the more I realized I did not understand what a mean of a RV is.

## The "survival formula" for expectation

For a non-negative RV $Y$

$$
\mathbb{E}[Y]=\int_0^{\infty} \mathbb{P}(Y>y) \mathrm{d} y = \int_0^{\infty}(1- F_Y(y))dy
$$

Appears (unhelpfully) as an exercise at the beginning of many math-stats books

## Basic idea

Take sample points $y_i = ih$ and assume $Nh$ large enough that $F_Y(Nh)$ close to 1. 
$$
\mathbb{E}[Y] \approx \sum_{i=1}^{N} y_if_Y(y_i) = \sum_{i=1}^{N}ih f_Y(ih) = \\
\begin{array} 
& & & & & & \vdots \\
& & & & hf_Y(3h) + \cdots\\
& & hf_Y(2h) &+ &hf_Y(3h) + \cdots\\
 hf_Y(h) &+ &hf_Y(2h) &+ &hf_Y(3h) + \cdots
\end{array}\\
\ \\
\approx \int_0^{\infty}(1- F_Y(y))dy
$$

## Resclicing picture

```{r echo=FALSE}
f <- function(y) 5*(.3*dnorm(y-3, sd = .8) + .7*dnorm(y-7,sd=2))
fplt <- function(y) 10*f(y)
x0 <- -2
xl <- 4
xu <- 15
width <- xu-xl
N <- 20
h <- width/N

# dat <- tibble(yis = seq(x0,xu,by=h), heights = pmax(0,yis-xl), fis = f(yis)*(heights>0), fyis = heights*fis)
dat <- tibble(
  yis = seq(xl+h,xu,by=h), 
  heights = yis-xl, 
  fis = f(yis), 
  fyis = heights*fis
  )
y1 <- dat$yis[1]
ylast <- tail(dat$yis,1)

p <- ggplot(data = dat) + geom_function(fun = fplt, n = 1000) +
  geom_function(fun = function(y) pmax(0,y-xl)) +
  scale_y_continuous(breaks = dat$heights) +
  scale_x_continuous(breaks = dat$yis, limits = c(x0,xu)) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        panel.grid.major.y = element_line(color = "lightgrey")
        )
p

p <- p + geom_rect(aes(xmin=yis-.1*h, xmax=yis+.1*h, ymin=0, ymax=heights), alpha=.2)
p
p <- p + geom_rect(aes(xmin=yis+.15*h, xmax=yis+.35*h, ymin=0, ymax=fyis)) 
p
p <- p + map(1:N, function(idx){
      geom_rect(data = dat %>% slice_tail(n = N + 1 - idx), 
                      aes(xmin=.data$yis+.4*h, 
                          xmax=.data$yis+.6*h, 
                          ymin=.data$heights[1] - h, 
                          ymax=.data$heights[1] - h + h*.data$fis))
    })
p

# p <- p + map(1:N, ~geom_line(
#     data = tibble(x=seq(xl+h*., ylast, length.out = 500), y = dat$yis[.]-y1+h*f(x)),
#     aes(x=x,y=y)
#     )) 
# p
```

##

So 
$$
\mathbb{E}[(Y-x)_+]= \int_x^{\infty}(1- F_Y(y))dy
$$

### 2023-06-17 R/run-alloscore.R
fix date glitch in run_alloscore, remake

@@ -6,7 +6,7 @@ run_alloscore <- function(forecast_data, truth_data, Ks, mkeep, reference_dates)
   ## process forecast data, adding distfromq output
   forecast_data_processed <- forecast_data |>
     ## forecast dates are different but reference dates are Mondays
-    dplyr::filter(reference_date %in% reference_dates) |>
+    dplyr::filter(reference_date %in% as.Date(reference_dates)) |>
     dplyr::select(-type) |>
     nest(ps = quantile, qs = value) |>
     relocate(ps, qs) |>

### 2023-06-17 random_undocumented_experiments/iter_plot.R

@@ -3,6 +3,8 @@ library(targets)
 tar_load(alloscore_df)
 adf <- alloscore_df
 
+# make a plot of the xs iterations for one model across multiple dates
+
 iters_bp <- adf %>% filter(model == "BPagano-RtDriven", K == 2000) %>%
   select(model, reference_date, xs) %>% unnest(xs) %>%
   mutate(init_alloc = ntile(`1`, 5), .after = abbreviation)

### 2023-06-20 talks/ISF_talk/talk.qmd
fix date glitch in run_alloscore, remake
talk update

--- 2023-06-16 ISF_talk/talk.qmd
@@
     width: 1500
+revealjs-plugins:
+  - attribution
 editor: visual
@@
 
-## Timeline
+## {.nostretch}
+
+![hubs](hub_timeline.png){}
+
+::: {.attribution}
+Figure credits: Alex Vespignani and Nicole Samay
+:::
 
@@
 
-------------------------------------------------------------------------
+## Sharing Business
 
@@
 
+## Data workflow
+
+```
+# A tibble: 12 × 9
+    time model forecast         ys            K xdf              score_raw score_oracle score
+   <int> <chr> <list>           <list>    <dbl> <list>               <dbl>        <dbl> <dbl>
+ 1     1 m1    <tibble [4 × 7]> <dbl [4]>   400 <tibble [4 × 8]>      62.5          0    62.5
+ 2     1 m1    <tibble [4 × 7]> <dbl [4]>   450 <tibble [4 × 8]>      50.0          0    50.0
+ 3     2 m1    <tibble [4 × 7]> <dbl [4]>   400 <tibble [4 × 8]>     150.          27.3 122. 
+ 4     2 m1    <tibble [4 × 7]> <dbl [4]>   450 <tibble [4 × 8]>     137.           0   137. 
+ 5     3 m1    <tibble [4 × 7]> <dbl [4]>   400 <tibble [4 × 8]>      62.2          0    62.2
+ 6     3 m1    <tibble [4 × 7]> <dbl [4]>   450 <tibble [4 × 8]>      49.7          0    49.7
+ 7     1 m2    <tibble [4 × 7]> <dbl [4]>   400 <tibble [4 × 8]>      54.8          0    54.8
+ 8     1 m2    <tibble [4 × 7]> <dbl [4]>   450 <tibble [4 × 8]>      40.2          0    40.2
+ 9     2 m2    <tibble [4 × 7]> <dbl [4]>   400 <tibble [4 × 8]>     157.          27.3 130. 
+10     2 m2    <tibble [4 × 7]> <dbl [4]>   450 <tibble [4 × 8]>     149.           0   149. 
+11     3 m2    <tibble [4 × 7]> <dbl [4]>   400 <tibble [4 × 8]>      74.9          0    74.9
+12     3 m2    <tibble [4 × 7]> <dbl [4]>   450 <tibble [4 × 8]>      70.9          0    70.9
+```
+
+```
+> a_series_slim$xdf[[9]]
+# A tibble: 4 × 8
+  target_names     x score_fun     y oracle components_raw components_oracle components
+  <chr>        <dbl> <list>    <dbl>  <dbl>          <dbl>             <dbl>      <dbl>
+1 A             80.4 <fn>       55.0   51.5             0               3.52      -3.52
+2 B             92.2 <fn>      250.   234.            157.             16.0      141.  
+3 C            106.  <fn>       89.2   83.5             0               5.70      -5.70
+4 D            122.  <fn>       33.5   31.3             0               2.14      -2.14
+```
+
 ## Scoring against an oracle
@@
 Comparitive vs absolute assessment/evaluation
+
+## General Disclaimer
+
+This is **post-hoc** analysis
+
+Hub forecasters were unaware of
+
+- an allocation score (on joint forecast)
+- any allocation based loss
+- our quantile interpolation/extrapolation methods (`distfromq`)
+
+  - might be especially important for tails

### 2023-06-20 random_undocumented_experiments/Norm1.R

@@ -54,8 +54,9 @@ K_lims <- quantile(ytot, probs = c(.1, .9)) %>% round(-1)
 (Ks <- seq(K_lims[1], K_lims[2], length.out = 5))
 allos <- mods %>% group_nest(model) %>%
   mutate(allocated = map(data, ~allocate(.x, K= Ks)))
-allos <- allos %>% mutate(slim = map(allocated, make_slim))
+allos <- allos %>% mutate(slim = map(allocated, slim))
 
+# compute intensive:
 allos <- allos %>% mutate(slim_scored = map(slim, ~alloscore(., ys$ys)))
 
 plot_allos <- allos %>% select(-c(data, allocated)) %>%
@@ -76,7 +77,8 @@ mods_crps <- mods %>%
   select(model, mean_crps) %>% unnest(mean_crps)
 
 mods_crps %>% pivot_wider(values_from = mean_crps, names_from = model)
-mods_crps %>% ggplot(aes(x = mean_crps, color = model)) + geom_histogram()
+mods_crps %>% ggplot(aes(x = mean_crps, fill = model)) +
+  geom_histogram(position = "dodge")
 
 plot_dat <- allos %>%
   mutate(scores = map(slim_scored, ~select(.,-xdf) %>% unnest(scores))) %>%
@@ -102,11 +104,15 @@ p + geom_density(data = plot_dat %>% filter(K %in% c("mean CRPS", Ks_to_show), m
 
 p_ridges <- plot_dat %>% filter(K %in% c("mean CRPS", Ks_to_show)) %>% mutate(K = fct_rev(K)) %>%
   ggplot(aes(x = score, y = model, fill = model)) + geom_density_ridges(color = NA) +
-  facet_grid(rows = vars(K)) +
+  facet_grid(
+    rows = vars(K),
+    labeller = as_labeller(function (K) {
+      ifelse(K == "mean CRPS", K, paste0("K = ", K))
+    })) +
     labs(y = "Density of scores") +
     guides(fill = guide_legend(reverse = TRUE)) +
-   scale_alpha_discrete(guide = FALSE)
-    theme(axis.ticks.y = element_blank(),
+  scale_alpha_discrete(guide = "none") +
+  theme(axis.ticks.y = element_blank(),
           axis.text.y = element_blank(),
           strip.text.y = element_text(angle = 0))
 

### 2023-06-21 R/run-alloscore.R
finish alloscore run, work on plot functions for ISF

@@ -1,4 +1,4 @@
-run_alloscore <- function(forecast_data, truth_data, Ks, mkeep, reference_dates){
+run_alloscore <- function(forecast_data, truth_data, one_reference_date, one_model){
   require(tidyverse)
   require(alloscore)
   require(distfromq)
@@ -6,7 +6,7 @@ run_alloscore <- function(forecast_data, truth_data, Ks, mkeep, reference_dates)
   ## process forecast data, adding distfromq output
   forecast_data_processed <- forecast_data |>
     ## forecast dates are different but reference dates are Mondays
-    dplyr::filter(reference_date %in% as.Date(reference_dates)) |>
+    dplyr::filter(reference_date %in% as.Date(one_reference_date)) |>
     dplyr::select(-type) |>
     nest(ps = quantile, qs = value) |>
     relocate(ps, qs) |>
@@ -14,7 +14,7 @@ run_alloscore <- function(forecast_data, truth_data, Ks, mkeep, reference_dates)
       ps = map(ps, deframe),
       qs = map(qs, deframe)
     ) |>
-    dplyr::filter(model %in% mkeep) %>%
+    dplyr::filter(model == one_model) %>%
     dplyr::mutate(
       model = ifelse(model == "COVIDhub-4_week_ensemble", "COVIDhub-ensemble", model)
     ) |>
@@ -24,16 +24,29 @@ run_alloscore <- function(forecast_data, truth_data, Ks, mkeep, reference_dates)
       truth_data |> select(location, target_end_date, value),
       by = c("location", "target_end_date"))
 
+  if (nrow(forecast_data_processed) > 0) {
+  ytot <- sum(forecast_data_processed$value)
+  Ks <- seq(200, min(60000, 4*ytot), by = 200)
+
   ## run alloscore
   ascores <- forecast_data_processed %>%
-  group_by(model, reference_date) %>%
-  group_modify(~ alloscore(
-    .x,
+  alloscore(
     K = Ks,
-    y = .x$value,
+    y = .[["value"]],
     target_names = "abbreviation"
-  ))
-
+  )
+  ascores <- ascores %>% dplyr::mutate(
+    reference_date = one_reference_date,
+    model = one_model,
+    .before = 1
+  )
+  } else {
+    ascores <- tibble(
+      reference_date = one_reference_date,
+      model = one_model,
+      message = "no forecasts"
+    )
+  }
   return(ascores)
 }
 

### 2023-06-21 _targets.R

@@ -22,16 +22,6 @@ tar_option_set(
 # Run the R scripts in the R/ folder:
 tar_source(files = c("R/data-ingestion.R", "R/plot-alloscores.R", "R/run-alloscore.R"))
 
-## create a group of alloscore targets
-forecast_dates = as.character(seq.Date(
-    as.Date("2021-11-22"), as.Date("2022-02-28"), by = "7 days"
-  ))
-
-forecast_dates_quick <- forecast_dates[c(12,13,14)]
-
-Ks <- seq(from = 2000, to = 7000, by = 500)
-
-# keep selected models
 mkeep <- c("BPagano-RtDriven",
            "COVIDhub-4_week_ensemble",
            "COVIDhub-baseline",
@@ -42,6 +32,14 @@ mkeep <- c("BPagano-RtDriven",
            "MUNI-ARIMA",
            "USC-SI_kJalpha",
            "UVA-Ensemble")
+
+forecast_dates = as.character(seq.Date(as.Date("2021-10-18"), as.Date("2022-02-28"), by = "7 days"))
+
+## create a group of alloscore targets
+values <- tidyr::expand_grid(models = mkeep, forecast_dates = forecast_dates)
+
+forecast_dates_quick <- forecast_dates[c(12,13)]
+values_test <- tidyr::expand_grid(forecast_dates = forecast_dates_quick, models = mkeep[1])
 
 # List of targets:
 list(
@@ -61,14 +59,22 @@ list(
     name = score_data,
     command = get_forecast_scores(forecast_data, truth_data)
   ),
-  tar_target(
-    name = alloscore_df,
-    command = run_alloscore(forecast_data, truth_data, K = Ks,
-                            mkeep = mkeep, reference_dates = forecast_dates_quick)
-  ),
-  tar_target(
-    name = figure_K_v_alloscore,
-    command = plot_K_v_alloscore(alloscore_df),
-    format = "file"
-  )
+  # tar_target(
+  #   name = alloscore_df,
+  #   command = run_alloscore(forecast_data, truth_data, K = Ks,
+  #                           mkeep = mkeep, reference_dates = forecast_dates_quick)
+  # ),
+  tar_map(values = values,
+          tar_target(
+            alloscore,
+            run_alloscore(forecast_data,
+                          truth_data,
+                          forecast_dates,
+                          models)
+          ))
+  # tar_target(
+  #   name = figure_K_v_alloscore,
+  #   command = plot_K_v_alloscore(alloscore_df),
+  #   format = "file"
+  # )
 )

### 2023-06-21 talks/ISF_talk/plotting_code/plot_functions.R

@@ -1,0 +1,73 @@
+
+
+
+get_intervals <- function(df) {
+  df %>% dplyr::filter(quantile %in% c(0.025, 0.1, 0.25, 0.75, 0.9, 0.975)) %>%
+    dplyr::mutate(
+      lu = ifelse(quantile < 0.5, "lower", "upper"),
+      interval_level_num = ifelse(quantile < 0.5,
+                                  100 * (1 - 2 * quantile),
+                                  100 * (1 - 2 * (1 - quantile))),
+      interval_level = paste0(as.integer(interval_level_num), "%")
+    ) %>%
+    dplyr::select(target_end_date, value, lu, interval_level, interval_level_num,
+                  any_of(c("model", "location", "geo_value", "xmin", "xmax"))) %>%
+    tidyr::pivot_wider(names_from = "lu", values_from = "value")
+}
+
+
+plot_one_hosp <- function(
+    start_date = "2021-09-01",
+    stop_date = "2022-01-31",
+    f_colors = c("mediumblue", "royalblue", "dodgerblue"),
+    f_width1 = 2,
+    space = .5,
+    f_width2 = f_width1 + .5,
+    models = "COVIDhub-4_week_ensemble",
+    location = "US"
+) {
+  truth <- truth %>% filter(location == !!location,
+                            target_end_date >= start_date,
+                            target_end_date <= stop_date)
+  total_width <- (length(models) - 1)*(f_width1 + space) + f_width1
+  half_width <- total_width/2
+  fc_dat <- forecasts_hosp %>%
+    dplyr::filter(location == !!location, model %in% models) %>%
+    dplyr::mutate(
+      xmin = target_end_date - half_width + (which(model == models) - 1)*(f_width1 + space),
+      xmax = xmin + f_width1
+    )
+  p <- ggplot(data = truth) +
+    geom_line(mapping = aes(x = target_end_date, y = value)) +
+    geom_vline(
+      data = data.frame(
+        tedate = as.Date(c("2021-12-27", "2022-01-10")),
+        Date = c("Forecast Date", "Target Date")
+      ),
+      mapping = aes(xintercept = tedate, linetype = Date),
+      linewidth = 0.5) +
+    geom_segment(
+      data = fc_dat %>% filter(quantile == 0.5),
+      mapping = aes(
+        x = target_end_date - f_width2,
+        xend = target_end_date + f_width2,
+        y = value, yend = value), size = 1.5) +
[…]

### 2023-06-21 talks/ISF_talk/plotting_code/plot_hosps.R

@@ -1,0 +1,129 @@
+library(covidHubUtils)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(geofacet)
+
+#hub_repo_path <- "~/research/epi/covid/covid19-forecast-hub/"
+hub_repo_path <- "~/covid/covid19-forecast-hub/"
+
+inc_hosp_targets <- paste(0:30, "day ahead inc hosp")
+forecasts_hosp <- load_forecasts(
+  #  models = c("COVIDhub-ensemble", "CMU-TimeSeries"),
+  #   dates = "2022-11-08",
+  dates = "2021-12-27",
+  date_window_size = 6,
+  #locations = "US",
+  types = c("quantile"),
+  targets = inc_hosp_targets,
+  source = "local_hub_repo",
+  hub_repo_path = hub_repo_path,
+  verbose = FALSE,
+  as_of = NULL,
+  hub = c("US")
+) %>%
+  covidHubUtils::align_forecasts() %>%
+  dplyr::filter(relative_horizon == 14)
+
+truth <- load_truth(
+  truth_source = "HealthData",
+  target_variable = "inc hosp"
+)
+
+mkeep <- c("BPagano-RtDriven",
+           "COVIDhub-4_week_ensemble",
+           "COVIDhub-baseline",
+           "CU-select",
+           "IHME-CurveFit",
+           "JHUAPL-Bucky",
+           "JHUAPL-Gecko",
+           "MUNI-ARIMA",
+           "USC-SI_kJalpha",
+           "UVA-Ensemble")
+
+us_ens_fc <- forecasts_hosp %>%
+  dplyr::filter(location == "US", model == "COVIDhub-4_week_ensemble")
+state_ens_fc <- forecasts_hosp %>%
+  dplyr::filter(location < 57, model == "COVIDhub-4_week_ensemble")
+
+forecasts_hosp %>%
+  dplyr::filter(location < 57, model == "COVIDhub-4_week_ensemble")
+
+
+plot_one_hosp()
+plot_one_hosp(models = c("COVIDhub-4_week_ensemble", "MUNI-ARIMA"), location = "10")
+
+ggtitle("US National Level Hospitalizations - Ensemble Forecast, 14 day horizon")
+
+pdf("nat_hosps.pdf", width = 7, height = 3.5)
+print(p_nat)
+dev.off()
+
+
+plot_state_fc <- function(
+    models = "COVIDhub-4_week_ensemble",
+    start_date = "2021-12-01",
[…]

### 2023-06-21 tar_test.R

@@ -1,0 +1,7 @@
+library(targets)
+list(tar_target(name = forecast_data, command = get_forecast_data(forecast_dates)), 
+    tar_target(name = truth_data, command = get_truth_data()), 
+    tar_target(name = score_data, command = get_forecast_scores(forecast_data, 
+        truth_data)), tar_map(values = values, tar_target(alloscore, 
+        run_alloscore(forecast_data, truth_data, forecast_dates, 
+            models))))

### 2023-06-22 talks/ISF_talk/plotting_code/plot_functions.R
plot_hosp for ISF

@@ -11,47 +11,97 @@ get_intervals <- function(df) {
       interval_level = paste0(as.integer(interval_level_num), "%")
     ) %>%
     dplyr::select(target_end_date, value, lu, interval_level, interval_level_num,
-                  any_of(c("model", "location", "geo_value", "xmin", "xmax"))) %>%
+                  any_of(c("model", "location", "code", "xmin", "xmax"))) %>%
     tidyr::pivot_wider(names_from = "lu", values_from = "value")
 }
 
+pal1 = c("mediumblue", "goldenrod", "#CD2626")
 
-plot_one_hosp <- function(
+
+plot_hosp <- function(
     start_date = "2021-09-01",
     stop_date = "2022-01-31",
-    f_colors = c("mediumblue", "royalblue", "dodgerblue"),
+    f_date = "2021-12-27",
+    te_date = "2022-01-10",
+    f_colors = pal1,
+    f_alpha = .5,
     f_width1 = 2,
     space = .5,
-    f_width2 = f_width1 + .5,
     models = "COVIDhub-4_week_ensemble",
-    location = "US"
+    locations = "US",
+    key_width = .15
 ) {
-  truth <- truth %>% filter(location == !!location,
-                            target_end_date >= start_date,
-                            target_end_date <= stop_date)
+  locations <- locations %>% purrr::map(
+    function(loc) {
+      if (grepl("^[A-Z]{2}$", loc)) {
+        hub_locations %>%
+          dplyr::filter(abbreviation == loc, nchar(fips) == 2) %>% pull(fips)
+      } else {
+        sprintf("%02d", as.integer(loc))
+      }
+    })
+  # compute aggregate rect and segment widths.
   total_width <- (length(models) - 1)*(f_width1 + space) + f_width1
   half_width <- total_width/2
+
+  truth <- truth %>%
+  dplyr::filter(location %in% locations,
+    target_end_date >= start_date,
+    target_end_date <= stop_date) %>%
+  dplyr::mutate(
+    validation = ifelse(target_end_date > as.Date(f_date), "Validation", "Historical"),
+    code = toupper(geo_value),
+    xmin = target_end_date - half_width - space,
+    xmax = xmin + total_width + 2 * space)
+
+  target_dates <- data.frame(
+    Date = as.Date(c(f_date, te_date)),
+    Date_name = c("Forecast Date", "Target Date"))
+
   fc_dat <- forecasts_hosp %>%
-    dplyr::filter(location == !!location, model %in% models) %>%
[…]

### 2023-06-22 talks/ISF_talk/plotting_code/plot_hosps.R

@@ -50,71 +50,42 @@ forecasts_hosp %>%
   dplyr::filter(location < 57, model == "COVIDhub-4_week_ensemble")
 
 
-plot_one_hosp()
-plot_one_hosp(models = c("COVIDhub-4_week_ensemble", "MUNI-ARIMA"), location = "10")
+plot_hosp(models = c("CU-select"), f_width1 = 3, locations = "MA")
+plot_hosp(models = c("CU-select","COVIDhub-4_week_ensemble", "MUNI-ARIMA"),
+          space = 1, f_colors = pal1)
 
-ggtitle("US National Level Hospitalizations - Ensemble Forecast, 14 day horizon")
+nat_hosps <- plot_hosp(f_width1 = 4) +
+  ggtitle("US National Level Hospitalizations - Ensemble Forecast, 14 day horizon")
+nat_hosps
+
+p <- plot_hosp(
+  start_date = "2021-12-01",
+  stop_date = "2022-01-31",
+  locations = 1:57,
+  models = c("JHUAPL-Bucky","COVIDhub-4_week_ensemble", "MUNI-ARIMA"),
+  f_colors = c("mediumblue", "goldenrod", "#CD2626"),
+  f_alpha = .4,
+  f_width1 = 4,
+  space = 1)
+
+p + theme_bw()
 
 pdf("nat_hosps.pdf", width = 7, height = 3.5)
 print(p_nat)
 dev.off()
 
+p_state <- plot_hosp(
+  start_date = "2021-12-01",
+  stop_date = "2022-01-31",
+  locations = 1:57,
+  models = c("JHUAPL-Bucky","COVIDhub-4_week_ensemble", "MUNI-ARIMA"),
+  f_colors = c("#4A708B", "goldenrod", "#CD2626"),
+  f_alpha = .4,
+  f_width1 = 5,
+  key_width = .05,
+  space = 1) + theme_bw()
 
-plot_state_fc <- function(
-    models = "COVIDhub-4_week_ensemble",
-    start_date = "2021-12-01",
-    stop_date = "2022-01-31",
-    f_colors = c("mediumblue", "royalblue", "dodgerblue"),
-    f_width = 3,
-    f_space = 1
-    ) {
-  fc_dat <- forecasts_hosp %>%
-    dplyr::filter(location < 57, model %in% models)
-  truth_dat <- truth %>%
-    dplyr::filter(location < 57,
-                  target_end_date >= start_date,
-                  target_end_date <= stop_date) %>%
[…]

### 2023-06-24 talks/ISF_talk/talk.qmd
finish alloscore run, work on plot functions for ISF
more plots, working on iter_plot

@@
 editor: visual
+css: styles.css
 ---
 
-## {.nostretch}
+## Evolution of ID Forecasting Hubs {.nostretch}
 
-![hubs](hub_timeline.png){}
+![](hub_timeline.png){}
 
@@
 
-## The Backstory
+## The Role of IDF Hubs
 
-Starting \~ 2010? Outbreak forecasting hubs developed to "inform public health responses"
+Outbreak forecasting hubs developed to **"inform public health responses"**
 
-Trend toward elicitation of probabilistic forecasts
+- with trend toward elicitation of probabilistic forecasts
 
--   uncertainty quantification often essential to a public health **DM** (decision maker)
+- uncertainty quantification essential to a public health **DM** (decision maker)
 
@@
 
+Basic element of any such response is an **allocation decision** regarding
+
 ::: columns
+
+::: {.column width="10%"}
+
+<div style="height: 200px; width: 50px;">
+  <svg height="100%" width="100%" viewBox="0 0 10 20">
+    <line x1="5" y1="0" x2="5" y2="15" style="stroke:black;stroke-width:2" />
+    <polygon points="0,20 10,20 5,15" style="fill:black;" />
+  </svg>
+</div>
+
+:::
+
 ::: {.column width="40%"}
-Basic element of any such response is an *allocation decision* regarding
 
@@
 
-::: {.column width="60%"}
+::: {.column width="50%"}
 Similar to other contexts where there are large-scale institutional efforts to use expert judgment to assess and manage social risk
@@
 
-Even with complex game-theoretic approachs to collective decision making, *allocation* almost always a requirement to strategy development
+Even with complex game-theoretic approachs to collective decision making, allocation almost always a requirement to strategy development
 :::
@@
 
-Hubs have tried to use SRs to rank and combine forecasts in ways that offer best suppport for PH responders
+*Very* briefly, a **scoring rule** $S$ assigns a number $S(F,y)$ to the forecast $F$ of $Y$ when $Y=y$ is realized.
 
-This can be cast in decision-theoretic framework of utility, loss functions, and Bayes acts.
+$S$ if **proper** if $\displaystyle{E_{F}[S(F,Y)] \leq E_{G}S(F,Y)]}$ for all $F, G$
 
-The DM needs to decide take an action $a$ wrt an outcome $Y$
+- ensures that forecasters must give honest forecasts to maximize expected score
 
--   such as the number of severe cases in various locations.
+Hubs have tried to use SRs to rank and combine forecasts in ways that offer best suppport for PH response to ID outbreaks
 
-Suppose the DM also has a **loss function** $l(a,y)$
+Primary scoring rule used currently is the **Weighted Interval Score** which is a discrete approximation to the 
+**Continuous Ranked Probability Score**
 
--   unmet need when $a < y$ beds are allocated and $y$ severe cases occur
+- adopted for convenient scoring of quantile forecasts.
 
-Then the *Bayes act* relative to a forecast $Y \sim F$ is the action $a^F$ that will minimize the DM's expected loss assuming that in reality $Y \sim F$.
+## Bayes Acts
+
+Scoring rules can be cast in decision-theoretic framework of **Bayes acts**.
+
+::: {.important}
+
+The **Bayes act** relative to a forecast $Y \sim F$ is the action $x^F$ 
+by which a DM will minimize **expected loss** $E_Y[l]$ assuming that in reality $Y \sim F$.
+
+:::
+
+ - loss $l$ is a function of the outcome $y$ **and** the action $x$ DM takes in anticipation of $y$
+
+Given $l$, we have an *automatically proper* scoring rule $S(F,y) := l(x^F, y)$.
+
+::: {.callout-note appearance="simple" icon=false}
+
+### Our running example:
+
+ - $x$ and $y$ are vectors in $\mathbb{N}_0^{52}$
+ - $y =$ number of severe cases in US states and territories
+ - $l(x,y) =$ unmet need when $x < y$ beds are allocated and $y$ severe cases occur
+
+:::
+
+## Formally
+
+$$
+\begin{align}
+l(x,y) &= \sum_{i=1}^{52} \max(y_i - x_i, 0) \\
+x^F &\in \left\{ \text{possible } x' \,\middle| \, E_{F}[l(x',Y)] \leq E_{F}[l(x,Y)] \text{ for any possible } x \right\}
+\end{align}
+$$
+
+Novelty of our approach for ID forecasting is how we define $x$ being 
+**possible**
+
+ - as satisfying some resource constraint $\displaystyle{\sum x_i \leq K}$
+
+::: {.callout-note appearance="simple" icon=false}
+
+### Note:
+
+Our loss function and (single) constraint are very simple prototypes meant 
+for proof of concept.
+
+:::
 
@@
 this must somehow acknowledge risk attitude heterogeneity among various DM's that a forecast hub might serve
+
+gpl parameters, weights, multiple constraints
 :::
@@
 
-A Scoring Rule attempts to transfer or "share" this loss function with the forecaster through a numerical reward system where FP receives
+A Scoring Rule attempts to "share" DM loss with the forecaster through $S(F,y)$
+
+- idea goes back to Savage
+
+- may be in units of money, goodwill, prestige
+
+::: {.callout-note appearance="Important" icon=false}
+
+## A central goal in design of scoring rules for social good
+
+Create scenario where the forecaster believes maximizing her own expected utility aligns 
+with leading the DM toward actions that maximize social utility.
+
+:::
+
+Using an explicit vacabulary of social loss in ID forecasting hub activities has instrinsic advantage
+over use of "off-the-shelf" statistical functionals from statistic inference
+
+::: notes
+$S$ a penalty for us
+
+inference has own decision theory and ideas of utility, loss, Bayes acts
+that are related but distinct from the fundamentally econommic/policy focussed
+ones we use here
+:::
+
+## A Lagrangian formulation for the Bayes act
+
+Given $F = \{F_i\}_{i=1}^{N}$, obtaining $x^F$ is a constrained stochastic optimization problem
 
 $$ 
-S(F,y)
+\begin{align}
+L(x, \lambda ; K, F) &= \sum_{i=1}^N E_{F_i}\left[l(x_i, Y_i)]\right)+\lambda\left(\sum_{i=1}^N x_i-K\right) \\  
+\nabla_{x} L = 0 &\implies 
+E_F \left[\frac{d}{dx_i}(Y_i - x_i)_{+}\right] + \lambda 
+= F_i(x_i)-1 + \lambda = 0 \\
+&\implies x_i = F_i^{-1}(1-\lambda)
+\end{align}
 $$
 
-units of money, goodwill, prestige, etc.
-A central goal in designing $S$ is to create a scenario where the FP beleives she cam maximizes her own expected utility by giving the DM the forecast that the FP believes in actually true and/or best suited for leading the DM toward actions that will maximize their utility.
-An SR that acheieves this is called *proper* and Bayes acts form a linchpin of their operation since the establish an explicit vacabulary of loss in place of vague efforts to rank preformance of various statisticla funcitonals adoptic form statistic inference
-
-## Loss vs Utility
-
-Change in loss but not utility from change in - variant severity (assuming target is infections and decision is number of beds)
-
-Change in utility but not in loss from change in - supply and personnel cost structures - public opinion regarding importance of fatality prevention - knowledge of demographic disparities in consequences of outcomes
+Solve $\sum_{l=1}^{N} F_{i}^{-1}(1-\lambda) = K$ to get $x_{i}^F = F_i^{-1}(1-\lambda^{\star})$
 
@@
 
-Adopt an iterative Lagrange multiplier method following long history in inventory management of solving *constrained multi-product newsvendor* problems
+Even for simple parametric $F_i$, this rarely has an analytic solution.
+
+Given quantile forecasts $q_{i,j}, j = 1,\ldots,23$, we interpolate full $F_i$ to form $L(x, \lambda ; K, F)$
+
+Adopt an iterative Lagrange multiplier method following long history in inventory management of solving **constrained multi-product newsvendor** problems
 
 illustrate algo with a lambda over xs plot against tau
+
+## Application
+
+December 2021: Omicron wave clearly started US but forecast teams unsure of severity given uncertainty about $R_0$, cross-protection by vaccination, previous infection, etc.
+
+___
+
+![](plots/state_hosps_ens_only.jpeg)
 
@@
   - might be especially important for tails
+
+<!-- ## Loss vs Utility
+
+Change in loss but not utility from change in - variant severity (assuming target is infections and decision is number of beds)
+
+Change in utility but not in loss from change in - supply and personnel cost structures - public opinion regarding importance of fatality prevention - knowledge of demographic disparities in consequences of outcomes -->

### 2023-06-24 _targets.R

@@ -8,7 +8,7 @@ library(crew)
 
 # Set target options:
 tar_option_set(
-  controller = crew_controller_local(workers = 8),
+  controller = crew_controller_local(workers = 4),
   packages = c("tidyverse", "covidHubUtils", "distfromq", "alloscore", "gh"), # packages that your targets need to run
   imports = c("alloscore"),
   format = "rds" # default storage format
@@ -18,7 +18,6 @@ tar_option_set(
 # remotes::install_github("aaronger/alloscore")
 # remotes::install_github("reichlab/distfromq")
 # remotes::install_github("reichlab/covidHubUtils")
-
 # Run the R scripts in the R/ folder:
 tar_source(files = c("R/data-ingestion.R", "R/plot-alloscores.R", "R/run-alloscore.R"))
 
@@ -35,11 +34,41 @@ mkeep <- c("BPagano-RtDriven",
 
 forecast_dates = as.character(seq.Date(as.Date("2021-10-18"), as.Date("2022-02-28"), by = "7 days"))
 
+forecast_dates_quick <- forecast_dates[c(12,13)]
+
 ## create a group of alloscore targets
+values_test <- tidyr::expand_grid(forecast_dates = forecast_dates_quick, models = mkeep[1])
 values <- tidyr::expand_grid(models = mkeep, forecast_dates = forecast_dates)
 
-forecast_dates_quick <- forecast_dates[c(12,13)]
-values_test <- tidyr::expand_grid(forecast_dates = forecast_dates_quick, models = mkeep[1])
+# hard coding the names from tar_map
+# this seems like it should be its own target but I can't figure this out right now
+a_target_names <- values %>% dplyr::mutate(
+  target_name = paste0("alloscore_",
+                       gsub(x = models, pattern = "-", replacement = "."),
+                       "_",
+                       gsub(x = forecast_dates, pattern = "-", replacement = ".")
+  )) %>% dplyr::select(-forecast_dates)
+
+a_target_names_nested <- a_target_names %>% tidyr::nest(target_list = target_name)
+
+# function to create a single slim df for each model's alloscore time series
+# Don't use this! Will blow up your computer.
+# slim_dfs <- function(target_list) {
+#   target_list <- target_list %>% unlist(use.names = F)
+#   targets::tar_load(target_list)
+#   purrr::map(target_list, get) %>%
+#     purrr::map(purrr::possibly(~ alloscore::slim(., id_cols = c(
+#       "reference_date", "model"
[…]

### 2023-06-24 random_undocumented_experiments/iter_plot.R

@@ -1,27 +1,0 @@
-
-library(targets)
-tar_load(alloscore_df)
-adf <- alloscore_df
-
-# make a plot of the xs iterations for one model across multiple dates
-
-iters_bp <- adf %>% filter(model == "BPagano-RtDriven", K == 2000) %>%
-  select(model, reference_date, xs) %>% unnest(xs) %>%
-  mutate(init_alloc = ntile(`1`, 5), .after = abbreviation)
-
-keep_states <- iters_bp %>% filter(reference_date==min(reference_date)) %>%
-  filter(min_rank(desc(`1`))<5) %>% pull(abbreviation)
-
-plot_iter_dat <- iters_bp %>% select(-qs) %>%
-  tidyr::pivot_longer(
-    cols = -c(1:4),
-    names_to = "iteration",
-    values_to = "allocation") %>%
-  mutate(iteration = as.numeric(iteration)) %>%
-  mutate(give_color = ifelse(abbreviation %in% keep_states, abbreviation, NA))
-
-(p_iter <- plot_iter_dat %>%
-  ggplot(aes(x = iteration, y = allocation, color = give_color, alpha = 1-.5*is.na(give_color))) +
-  facet_grid(cols = vars(reference_date)) + geom_point() +
-  geom_line() + theme_classic() +
-  scale_alpha_continuous(guide = "none"))

### 2023-06-24 talks/ISF_talk/iter_plot.R

@@ -1,0 +1,112 @@
+
+library(targets)
+adf <- tar_read(alloscore_BPagano.RtDriven_2022.01.03)
+
+tar_load(truth_data)
+
+# copying from run_alloscore.R
+
+require(tidyverse)
+require(alloscore)
+require(distfromq)
+
+one_model <- mkeep[1]
+one_reference_date <- "2021-12-27"
+
+## process forecast data, adding distfromq output
+forecast_data_processed <- forecast_data |>
+  ## forecast dates are different but reference dates are Mondays
+  dplyr::filter(reference_date %in% as.Date(one_reference_date)) |>
+  dplyr::select(-type) |>
+  nest(ps = quantile, qs = value) |>
+  relocate(ps, qs) |>
+  mutate(
+    ps = map(ps, deframe),
+    qs = map(qs, deframe)
+  ) |>
+  dplyr::filter(model == one_model) %>%
+  dplyr::mutate(
+    model = ifelse(model == "COVIDhub-4_week_ensemble", "COVIDhub-ensemble", model)
+  ) |>
+  add_pdqr_funs(dist = "distfromq", types = c("p", "q")) |>
+  relocate(dist, F, Q) |>
+  left_join(
+    truth_data |> select(location, target_end_date, value),
+    by = c("location", "target_end_date"))
+
+  ytot <- sum(forecast_data_processed$value)
+  Ks <- seq(2000, 30000, by = 2000)
+
+# Lets plot the distfromq Fs
+
+dist_plot <- forecast_data_processed %>% select(abbreviation, F) %>%
+    mutate(x_values = list(seq(0, 1000, length.out = 1000)),
+           y_values = map2(F, x_values, function(F, xl) F(xl))) %>%
+    unnest(-F) %>%
+    ggplot(aes(x = x_values, y = y_values, color = abbreviation)) +
+    geom_line()
+
+dist_plot + xlim(c(0, 1000))
+
+  ## run alloscore
+  ascores <- forecast_data_processed %>%
+    alloscore(
+      starting_K = 30000,
+      K = 4000,
+      y = .[["value"]],
+      target_names = "abbreviation"
+    )
+  ascores <- ascores %>% dplyr::mutate(
+    reference_date = one_reference_date,
+    model = one_model,
+    .before = 1
+  )
+
+# make a plot of the xs iterations for one model across multiple dates
+
+iters_bp <- ascores %>% filter(model == "BPagano-RtDriven", K == 4000) %>%
[…]

### 2023-06-24 talks/ISF_talk/plot_allocs.R

@@ -1,0 +1,65 @@
+library(covidHubUtils)
+library(alloscore)
+library(tidyverse)
+library(geofacet)
+
+source("_targets.R")
+
+file_names <- paste0("talks/ISF_talk/slim_dfs/slim_", mkeep)
+# create an empty list to store the loaded objects
+# slim_dfs <- list()
+# for(i in seq_along(file_names)){
+#   slim_dfs[[mkeep[i]]] <- readRDS(file_names[i])
+# }
+
+slim_dfs$`COVIDhub-4_week_ensemble`$model <- "COVIDhub-ensemble"
+
+get_ytot <- function(df) {
+  if ("xdf" %in% names(df)) {
+    df %>% group_by(reference_date) %>% slice(1) %>%
+    transmute(map_vec(xdf, ~ summarise(., ytot = sum(y))))
+  } else {
+    df %>% group_by(reference_date) %>%
+    summarise(ytot = sum(y))
+  }
+}
+add_ytot <- function(df) {
+  ytot <- get_ytot(df)
+  df %>% left_join(ytot, by = "reference_date")
+}
+
+plot_components <- function(...) {
+  alloscore::plot_components_slim(
+    ...,
+    origin_time_col_name = "reference_date",
+    model_col_name = "model",
+    target_col_name = "abbreviation"
+  )
+}
+
+K1 <- slim_dfs$`COVIDhub-4_week_ensemble` %>%
+  filter(reference_date == forecast_dates[14])
+
+K1_long <- K1 %>% unnest(xdf)
+
+K1_long
+K1_long_plus <- K1_long %>% mutate(components = components - min(components))
+
+p1 <- plot_components(K1_long_plus, show_oracle = FALSE, order_at_K = 15800, show_raw = FALSE)
+
+p1 + lims(x=c(1000,25000)) + theme_bw()
+
+df <- bind_rows(slim_dfs$`MUNI-ARIMA`, slim_dfs$`JHUAPL-Bucky`)
+
+p <- plot_components(df %>% filter(K == 15000, reference_date == "2021-12-27") %>% unnest(xdf))
+
+
+
+
+# Create a repeating vector of these colors that matches the number of levels in your fill variable
+color_vector <- rep(my_colors, length.out = 51)
+
+p +  scale_fill_manual(values = color_vector)
+
+
+

### 2023-06-24 talks/ISF_talk/plot_scores.R

@@ -1,0 +1,34 @@
+library(covidHubUtils)
+library(alloscore)
+library(tidyverse)
+library(geofacet)
+
+source("_targets.R")
+
+score_viz <- tribble(
+  ~palette,  ~linetypes, ~model,
+  "#004488", 2,          "BPagano-RtDriven",
+  "#117733", 4,          "COVIDhub-ensemble",
+  "#44AA99", 6,          "CU-select",
+  "#88CCEE", 2,          "IHME-CurveFit",
+  "#DDCC77", 4,          "JHUAPL-Bucky",
+  "#999933", 5,          "JHUAPL-Gecko",
+  "#CC6677", 1,          "MUNI-ARIMA",
+  "#882255", 1,          "USC-SI_kJalpha",
+  "#AA4499", 1,          "UVA-Ensemble",
+  "#000000", 1,          "COVIDhub-baseline"
+)
+
+score_palette <- deframe(score_viz[c("model", "palette")])
+score_linetypes <- deframe(score_viz[c("model", "linetypes")])
+
+peak_day <- bind_rows(slim_dfs) %>% filter(reference_date == "2021-12-27")
+
+three_days <- bind_rows(slim_dfs) %>% filter(reference_date %in% forecast_dates[13:15])
+
+plot_scores_slim(peak_day)
+plot_scores_slim(three_days) + lims(x=c(500, 30000))
+
+p_peak <- plot_scores_slim(peak_day)
+
+p_peak + theme_bw()

### 2023-06-24 talks/ISF_talk/plotting_code/plot_functions.R

@@ -29,6 +29,7 @@ plot_hosp <- function(
     space = .5,
     models = "COVIDhub-4_week_ensemble",
     locations = "US",
+    geofacet = FALSE,
     key_width = .15
 ) {
   locations <- locations %>% purrr::map(
@@ -56,7 +57,7 @@ plot_hosp <- function(
 
   target_dates <- data.frame(
     Date = as.Date(c(f_date, te_date)),
-    Date_name = c("Forecast Date", "Target Date"))
+    Date_name = c("Forecast Date, 2021-12-27", "Target Date, 2022-01-10"))
 
   fc_dat <- forecasts_hosp %>%
     dplyr::filter(location %in% locations, model %in% models) %>%
@@ -84,7 +85,7 @@ plot_hosp <- function(
       mapping = aes(
         x = xmin,
         xend = xmax,
-        y = value, yend = value, color = "Observed Value"), size = 1) +
+        y = value, yend = value, color = "Observed Value"), size = .5) +
     # Dummy layer for legend
     geom_blank(data = data.frame(validation = "Observed Value"), aes(color = validation)) +
     scale_color_manual(
@@ -133,11 +134,13 @@ plot_hosp <- function(
       alpha = guide_legend(order = 1),
       fill = guide_legend(override.aes = list(alpha = 1), order = 2),
       color = guide_legend(order = 3))
-  if (length(locations) > 1) {
-    p <- p + facet_geo(~ code, grid = geofacet::us_state_grid1) +
+  if (geofacet) {
+    p <- p + facet_geo(~ code, grid = geofacet::us_state_grid2) +
     scale_x_date(breaks = as.Date(c("2021-12-01", "2022-01-01")),
                  date_labels = "%b %y") +
     theme(axis.title = element_blank())
+  } else if (length(locations) > 1) {
+    p <- p + facet_wrap(~ code)
   }
   p <- p + theme(
       legend.key.width = unit(f_width1*key_width, "cm"),

### 2023-06-24 talks/ISF_talk/plotting_code/plot_hosps.R

@@ -4,31 +4,15 @@ library(tidyr)
 library(ggplot2)
 library(geofacet)
 
+source("talks/ISF_talk/plotting_code/plot_functions.R")
+save_dir <- file.path("talks/ISF_talk/plots/")
+
 #hub_repo_path <- "~/research/epi/covid/covid19-forecast-hub/"
 hub_repo_path <- "~/covid/covid19-forecast-hub/"
 
-inc_hosp_targets <- paste(0:30, "day ahead inc hosp")
-forecasts_hosp <- load_forecasts(
-  #  models = c("COVIDhub-ensemble", "CMU-TimeSeries"),
-  #   dates = "2022-11-08",
-  dates = "2021-12-27",
-  date_window_size = 6,
-  #locations = "US",
-  types = c("quantile"),
-  targets = inc_hosp_targets,
-  source = "local_hub_repo",
-  hub_repo_path = hub_repo_path,
-  verbose = FALSE,
-  as_of = NULL,
-  hub = c("US")
-) %>%
-  covidHubUtils::align_forecasts() %>%
-  dplyr::filter(relative_horizon == 14)
-
-truth <- load_truth(
-  truth_source = "HealthData",
-  target_variable = "inc hosp"
-)
+forecasts_hosp <- targets::tar_read(forecast_data) %>%
+  filter(reference_date == "2021-12-27")
+truth <- targets::tar_read(truth_data)
 
 mkeep <- c("BPagano-RtDriven",
            "COVIDhub-4_week_ensemble",
@@ -41,15 +25,6 @@ mkeep <- c("BPagano-RtDriven",
            "USC-SI_kJalpha",
            "UVA-Ensemble")
 
-us_ens_fc <- forecasts_hosp %>%
-  dplyr::filter(location == "US", model == "COVIDhub-4_week_ensemble")
-state_ens_fc <- forecasts_hosp %>%
-  dplyr::filter(location < 57, model == "COVIDhub-4_week_ensemble")
-
-forecasts_hosp %>%
-  dplyr::filter(location < 57, model == "COVIDhub-4_week_ensemble")
-
-
 plot_hosp(models = c("CU-select"), f_width1 = 3, locations = "MA")
 plot_hosp(models = c("CU-select","COVIDhub-4_week_ensemble", "MUNI-ARIMA"),
           space = 1, f_colors = pal1)
[…]

### 2023-06-24 talks/ISF_talk/slim_script.R

@@ -1,0 +1,55 @@
+
+library(tidyverse)
+library(alloscore)
+library(targets)
+
+args <- commandArgs(trailingOnly = TRUE)
+j <- as.integer(args[1])
+print(j)
+
+
+mkeep <- c("BPagano-RtDriven",
+           "COVIDhub-4_week_ensemble",
+           "COVIDhub-baseline",
+           "CU-select",
+           "IHME-CurveFit",
+           "JHUAPL-Bucky",
+           "JHUAPL-Gecko",
+           "MUNI-ARIMA",
+           "USC-SI_kJalpha",
+           "UVA-Ensemble")
+
+forecast_dates = as.character(seq.Date(as.Date("2021-10-18"), as.Date("2022-02-28"), by = "7 days"))
+
+values <- tidyr::expand_grid(models = mkeep, forecast_dates = forecast_dates)
+
+# hard coding the names from tar_map
+# this seems like it should be its own target but I can't figure this out right now
+a_target_names <- values %>% dplyr::mutate(
+  target_name = paste0("alloscore_",
+                       gsub(x = models, pattern = "-", replacement = "."),
+                       "_",
+                       gsub(x = forecast_dates, pattern = "-", replacement = ".")
+  )) %>% dplyr::select(-forecast_dates)
+
+a_target_names_nested <- a_target_names %>% tidyr::nest(target_list = target_name)
+
+slim_dfs <- function(i) {
+  target_list <- purrr::map(a_target_names_nested$target_list[[i]][[1]], ~tar_read_raw(.))
+
+  saveRDS(target_list %>%
+    purrr::map(possibly(~ alloscore::slim(., id_cols = c("reference_date", "model")))) %>%
+    bind_rows(), file = paste0("talks/ISF_talk/slim_dfs/slim_", a_target_names_nested$models[i]))
+
+  rm(target_list)
+}
+
+# Get the names of all objects in the global environment
+#
+# for (i in 6:10) {
+#   slim_dfs(i)
+# }
+
+slim_dfs(j)
+
+

### 2023-06-24 tar_test.R

@@ -1,7 +1,52 @@
 library(targets)
-list(tar_target(name = forecast_data, command = get_forecast_data(forecast_dates)), 
-    tar_target(name = truth_data, command = get_truth_data()), 
-    tar_target(name = score_data, command = get_forecast_scores(forecast_data, 
-        truth_data)), tar_map(values = values, tar_target(alloscore, 
-        run_alloscore(forecast_data, truth_data, forecast_dates, 
-            models))))
+library(tidyverse)
+library(alloscore)
+library(distfromq)
+library(lobstr)
+source("_targets.R")
+
+tar_load(names = c("forecast_data", "truth_data"))
+tar_load(starts_with("alloscore_BPagano.RtDriven"))
+
+
+.mkeep <- sub(x = mkeep, pattern = "-", replacement = ".")
+
+tar_names <- tar_manifest(fields = name)
+
+BPagano.RtDriven <- mget(grep(paste0("alloscore_", .mkeep[1]), tar_names$name, value = TRUE)) %>% 
+  map(possibly(~slim(., id_cols = c("reference_date", "model")))) %>% 
+  bind_rows()
+
+obj_size(BPagano.RtDriven)
+
+run_alloscore(forecast_data = forecast_data, 
+              truth_data = truth_data, K = Ks,
+              mkeep = mkeep, reference_dates = forecast_dates_quick)
+
+alloscore_BPagano.RtDriven_2021.12.27$xdf[[1]]
+as1 <- slim(alloscore_BPagano.RtDriven_2021.12.27, id_cols = c("reference_date", "model"))
+
+as1$xdf[[1]]
+as2 <- as1 %>% mutate(xdf = map(xdf, ~select(.,-score_fun)))
+
+obj_size(as1)
+obj_size(as2) 
+
+
+
+
+
+obj_size(alloscore_df)
+
+tar_load(alloscore_CU.select_2022.01.03)
+
+alloscore_CU.select_2022.01.03
+
+
+obj_size(s1)
+obj_size(s1_slim)
+
+
+
+
+

### 2023-06-28 talks/ISF_talk/talk.qmd
post ISF

@@
 ---
-title: "alloscore, ISF 43"
+title: |
+  | Infectious Disease Forecast Evaluation Based On Realized Social Utility:
+  | Allocation Scores
+  | ISF 43
+author: Aaron Gerding, Nick Reich, Ben Rogers, Evan Ray
+institute: "School of Public Health and Health Sciences, UMass, Amherst"
 format:
   revealjs: 
-    width: 1500
+    width: 1050
+    height: 700
 revealjs-plugins:
@@
 
+## Introduction
+
+::: columns
+
+::: {.column width="50%"}
+
+About me:
+
+- PhD in pure math
+- transitioned to public health data science 
+- started as RA with Nick and Evan at [reichlab.io](reichlab.io) at UMass in 2021
+- maintain COVID-19 Forecast Hub ensemble
+- current research goals center on ensemble methodology
+  - focus on scoring rules, decision theory, forecast-policy alignment
+
+:::
+
+::: {.column width="50%"}
+
+![](Reichlab.png)
+
+![](C19FH.png)
+
+:::
+
+:::
+
 ## Evolution of ID Forecasting Hubs {.nostretch}
 
-![](hub_timeline.png){}
+![](hub_timeline.png){width=80%}
 
@@
 
+New directions: "Hubverse" infrastructure project [https://hubdocs.readthedocs.io/](), genomic targets
+
 ::: notes
@@
 
-## The Role of IDF Hubs
+## Basic products of ID Forecasting Hubs
+
+![](plots/p_VA_CU_IHME.png)
+
+## The Role of ID forecasting Hubs
 
@@
 
-- with trend toward elicitation of probabilistic forecasts
+- trend from point and probability (of bins) toward quantile reporting
 
@@
 
-But in the inimitable words of chatGPT:
+- substantial buy-in from policy makers
 
-> It's important to note that while these efforts represent significant advances in the field, forecasting the spread and impact of infectious diseases is a very complex task with many challenges, and these models should be interpreted with caution.
+> It's important to note that while these efforts represent significant advances in the field, forecasting the spread and impact of infectious diseases is a very complex task with many challenges, and these models should be **interpreted with caution**.
+
+Basic example of **interpretation** is an **allocation decision** regarding
+
+-   medical supplies
+-   facility capacity
+-   personnel
+-   R&D funding
+-   public tolerance for NPIs
 
 ::: notes
-Over last two decades, multicenter effort to collect forecasts via hubs for ID outbreaks (epi/pan/endemic) with central purpose to "inform public health responses"
+quantiles flexible interpretable - closer to describing full distribution which DM probably always wants
+
+multicenter effort to collect forecasts via hubs for ID outbreaks (epi/pan/endemic)
 
 With uncertainty quantification being well-established as needed DM input,
-:::
 
-## The fundamental need to decide allocations
-
-Basic element of any such response is an **allocation decision** regarding
-
-::: columns
-
-::: {.column width="10%"}
-
-<div style="height: 200px; width: 50px;">
-  <svg height="100%" width="100%" viewBox="0 0 10 20">
-    <line x1="5" y1="0" x2="5" y2="15" style="stroke:black;stroke-width:2" />
-    <polygon points="0,20 10,20 5,15" style="fill:black;" />
-  </svg>
-</div>
-
-:::
-
-::: {.column width="40%"}
-
--   medical supplies
--   personnel
--   facility capacity
--   R&D funding
--   public tolerance for NPIs
-:::
-
-::: {.column width="50%"}
-Similar to other contexts where there are large-scale institutional efforts to use expert judgment to assess and manage social risk
-
--   finance
--   meteorology
--   military conflict
-:::
-:::
-
-::: notes
 Allocations can be wrt space, time, demographic
 
-Even with complex game-theoretic approachs to collective decision making, allocation almost always a requirement to strategy development
+Next: formalize this.
 :::
@@
 
-$S$ if **proper** if $\displaystyle{E_{F}[S(F,Y)] \leq E_{G}S(F,Y)]}$ for all $F, G$
+$S$ if **proper** if $\displaystyle{E_{F}[S(F,Y)] \leq E_{F}S(G,Y)]}$ for all $F, G$
 
@@
 
-Hubs have tried to use SRs to rank and combine forecasts in ways that offer best suppport for PH response to ID outbreaks
+Hubs have tried to use proper SRs to rank and combine forecasts in ways that offer best suppport for PH response to ID outbreaks
 
 Primary scoring rule used currently is the **Weighted Interval Score** which is a discrete approximation to the 
-**Continuous Ranked Probability Score**
+**Continuous Ranked Probability Score** (a probabilistic generalization of MAE)
 
 - adopted for convenient scoring of quantile forecasts.
+
+<br>
+
+#### A central goal in design of scoring rules for social good
+
+::: box
+
+Create scenario where the forecaster believes maximizing her own expected utility aligns 
+with leading the DM toward actions that maximize social utility.
+
+:::
 
@@
 
-::: {.important}
+<br>
+
+::: box
 
 The **Bayes act** relative to a forecast $Y \sim F$ is the action $x^F$ 
-by which a DM will minimize **expected loss** $E_Y[l]$ assuming that in reality $Y \sim F$.
+by which a DM will minimize **expected loss** $E_Y[l]$ 
+assuming that in reality $Y \sim F$.
 
@@
 
-Given $l$, we have an *automatically proper* scoring rule $S(F,y) := l(x^F, y)$.
+::: box
 
-::: {.callout-note appearance="simple" icon=false}
+Key point: Given $l$, we have an *automatically proper* scoring rule 
+$S(F,y) := l(x^F, y)$
 
-### Our running example:
+:::
+
+<br>
+
+#### Our basic example:
 
@@
 
-:::
-
 ## Formally
@@
 \begin{align}
-l(x,y) &= \sum_{i=1}^{52} \max(y_i - x_i, 0) \\
+l(x,y) &= \sum_{i=1}^{52} \max(y_i - x_i, 0) = \sum (y_i - x_i)_{+}\\
 x^F &\in \left\{ \text{possible } x' \,\middle| \, E_{F}[l(x',Y)] \leq E_{F}[l(x,Y)] \text{ for any possible } x \right\}
@@
 
-Novelty of our approach for ID forecasting is how we define $x$ being 
-**possible**
+<br>
 
- - as satisfying some resource constraint $\displaystyle{\sum x_i \leq K}$
+#### Novelty of our approach for ID forecast evaluation:
 
-::: {.callout-note appearance="simple" icon=false}
+::: box
 
-### Note:
+Define set of **possible** $x$ as those actions which satisfy $\displaystyle{\sum x_i \leq K}$
+
+:::
+
+<br>
+
+#### Note:
 
@@
 for proof of concept.
-
-:::
 
@@
 
-## Sharing Business
-
-A Scoring Rule attempts to "share" DM loss with the forecaster through $S(F,y)$
-
-- idea goes back to Savage
-
-- may be in units of money, goodwill, prestige
-
-::: {.callout-note appearance="Important" icon=false}
-
-## A central goal in design of scoring rules for social good
-
-Create scenario where the forecaster believes maximizing her own expected utility aligns 
-with leading the DM toward actions that maximize social utility.
-
-:::
-
-Using an explicit vacabulary of social loss in ID forecasting hub activities has instrinsic advantage
-over use of "off-the-shelf" statistical functionals from statistic inference
-
-::: notes
-$S$ a penalty for us
-
-inference has own decision theory and ideas of utility, loss, Bayes acts
-that are related but distinct from the fundamentally econommic/policy focussed
-ones we use here
-:::
-
 ## A Lagrangian formulation for the Bayes act
@@
 
-$$ 
 \begin{align}
 L(x, \lambda ; K, F) &= \sum_{i=1}^N E_{F_i}\left[l(x_i, Y_i)]\right)+\lambda\left(\sum_{i=1}^N x_i-K\right) \\
-\nabla_{x} L = 0 &\implies 
-E_F \left[\frac{d}{dx_i}(Y_i - x_i)_{+}\right] + \lambda 
-= F_i(x_i)-1 + \lambda = 0 \\
-&\implies x_i = F_i^{-1}(1-\lambda)
+
+\nabla_{x} L(x^F, \lambda) = 0 &\implies 
+E_F \left[\frac{d}{dx_i}(Y_i - x_i^F)_{+}\right] + \lambda \\
+&= F_i(x_i^F)-1 + \lambda = 0 
+\implies x_i^F = F_i^{-1}(1-\lambda) \\
+
+\nabla_{\lambda} L(x^F, \lambda^{\star}) = 0 
+&\implies \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star}) - K  = 0
+\implies 
+\boxed{x_{i}^F = F_i^{-1}(1-\lambda^{\star})}
 \end{align}
-$$
 
-Solve $\sum_{l=1}^{N} F_{i}^{-1}(1-\lambda) = K$ to get $x_{i}^F = F_i^{-1}(1-\lambda^{\star})$
+Our allocation scoring rule is then given by the recipe:
+
+1. Use forecast $F$ to form Lagrangian $L$, for expected loss under a resource constraint
+
+2. Solve for optimizer $x^F$ using $L$
+
+3. Plug $x^F$ and outcome $y$ into loss function to get $S(F,y) = l(x^F,y)$
 
@@
 
-Even for simple parametric $F_i$, this rarely has an analytic solution.
+But getting $x^F$ not so simple.
 
-Given quantile forecasts $q_{i,j}, j = 1,\ldots,23$, we interpolate full $F_i$ to form $L(x, \lambda ; K, F)$
+Even for simple parametric $F_i$, there is rarely an analytic solution for $\lambda^{\star}$.
 
-Adopt an iterative Lagrange multiplier method following long history in inventory management of solving **constrained multi-product newsvendor** problems
+- and we need to reconstruct $F_i$'s from quantile forecasts $q_{i,j}, j = 1,\ldots,23$ to form $L(x, \lambda ; K, F)$
 
-illustrate algo with a lambda over xs plot against tau
+![](plots/dist_pqs_plot_OH.png)
+
+## Iterative methods
+
+So we resort to a binary search for $\lambda^{\star}$ following long history in inventory management of solving **constrained multi-product newsvendor** 
+problems^[See e.g., Zhang, Bin, Xiaoyan Xu, and Zhongsheng Hua. 2009. “A Binary Solution Method for the Multi-Product
+Newsboy Problem with Budget Constraint.”]
+
+![](plots/iter_combo.png)
 
@@
 
-___
+![](plots/nat_hosps.png)
 
-![](plots/state_hosps_ens_only.jpeg)
+## Results
+
+::: columns
+
+::: {.column width="30%"}
+
+After adjusting for an oracle the score of an oracle that allocates *fairly* based on what demands actually occur, scores resemble "value" plots from meteorology
+
+- also the Murphy curves of Gneiting et. al.
+
+- extreme shortages or surplusses diminish the oracle's advantage
+
+- ranking at peak seems to be consistent with that in large $K$
+region.
+
+:::
+
+::: {.column width="70%"}
+
+![](plots/peak_alloscore.png)
+
+:::
+
+:::
+
+::: aside
+Ehm, Werner, Tilmann Gneiting, Alexander Jordan, and Fabian Krüger. 2016. “Of Quantiles and Expectiles: Consistent Scoring Functions, Choquet Representations and Forecast Rankings.”
+:::
+
+::: notes
+Note: The oracle is not the DGP, it is a time-traveller
+:::
+
+## Alloscore and WIS rank models differently
+
+![](plots/p_wis_v_alloscore.png)
+
+## Explanations?
+
+![](plots/mvbucky_dist_alloc.png)
+
+## Limitations
+
+This is **post-hoc** analysis
+
+Hub forecasters were unaware of
+
+- an allocation score (on joint forecast)
+- any allocation based loss
+- our quantile interpolation/extrapolation methods (`distfromq`)
+
+  - might be especially important for tails
+
+We hope/think that allocation scoring is sensitive to implicit dependence structures in forecasts, but all work so far only refers directly to marginals
+- nothing yet with copulas, etc.
+
+<br>
+
+### Thank you!
+
+A very rough R package I wrote to implement scoring procedures:
+
+[https://github.com/aaronger/alloscore](https://github.com/aaronger/alloscore)
+
+A less rough package Evan wrote to implement cdf reconstruction
+
+[https://github.com/reichlab/distfromq](https://github.com/reichlab/distfromq)
+
+<!-- ## Loss vs Utility
+
+Change in loss but not utility from change in - variant severity (assuming target is infections and decision is number of beds)
+
+Change in utility but not in loss from change in - supply and personnel cost structures - public opinion regarding importance of fatality prevention - knowledge of demographic disparities in consequences of outcomes -->
 
@@
 
-## Scoring against an oracle
+## National picture
 
-To get more interpretable allocation scores we adjust by the score of an oracle that allocates *fairly* based on what demands actually occur
+![](plots/state_hosps_ens_only.jpeg){width=100%}
 
-Note: The oracle is not the DGP, it is a time-traveller
+## Sharing Business
 
-Comparitive vs absolute assessment/evaluation
+A Scoring Rule attempts to "share" DM loss with the forecaster through $S(F,y)$
 
-## General Disclaimer
+- idea goes back to Savage
 
-This is **post-hoc** analysis
+- may be in units of money, goodwill, prestige
 
-Hub forecasters were unaware of
+<br>
 
-- an allocation score (on joint forecast)
-- any allocation based loss
-- our quantile interpolation/extrapolation methods (`distfromq`)
+<br>
 
-  - might be especially important for tails
+An explicit vacabulary of social loss in ID forecast evaluation has instrinsic advantage
+over "off-the-shelf" metrics from statistic inference
 
-<!-- ## Loss vs Utility
+::: notes
+$S$ a penalty for us
 
-Change in loss but not utility from change in - variant severity (assuming target is infections and decision is number of beds)
+I do not have a clean definition of "social loss/good/risk/regret" but wanted to use term becasue it is in the session title
 
-Change in utility but not in loss from change in - supply and personnel cost structures - public opinion regarding importance of fatality prevention - knowledge of demographic disparities in consequences of outcomes -->
+inference has own decision theory and ideas of utility, loss, Bayes acts
+that are related but distinct from the fundamentally econommic/policy focussed
+ones we use here
+:::

### 2023-06-28 talks/ISF_talk/iter_plot.R

@@ -1,8 +1,9 @@
-
+library(scales)
 library(targets)
 adf <- tar_read(alloscore_BPagano.RtDriven_2022.01.03)
 
 tar_load(truth_data)
+tar_load(forecast_data)
 
 # copying from run_alloscore.R
 
@@ -10,7 +11,7 @@ require(tidyverse)
 require(alloscore)
 require(distfromq)
 
-one_model <- mkeep[1]
+one_model <- mkeep[4]
 one_reference_date <- "2021-12-27"
 
 ## process forecast data, adding distfromq output
@@ -39,20 +40,81 @@ forecast_data_processed <- forecast_data |>
 
 # Lets plot the distfromq Fs
 
-dist_plot <- forecast_data_processed %>% select(abbreviation, F) %>%
-    mutate(x_values = list(seq(0, 1000, length.out = 1000)),
+dist_plot_dat <- forecast_data_processed %>% select(abbreviation, F) %>%
+  mutate(x_values = list(c(seq(0, 3, length.out = 300), seq(3, 1500, length.out = 1000))),
            y_values = map2(F, x_values, function(F, xl) F(xl))) %>%
-    unnest(-F) %>%
-    ggplot(aes(x = x_values, y = y_values, color = abbreviation)) +
-    geom_line()
+  select(-F) %>% unnest(-abbreviation)
+
+dist_plot <- dist_plot_dat %>%
+  ggplot(aes(x = x_values, y = y_values, color = abbreviation)) + geom_line()
 
 dist_plot + xlim(c(0, 1000))
 
+# with quantiles for a subset of states
+
+make_dist_pqs_plot <- function(states) {
+pqs <- forecast_data_processed %>% filter(abbreviation %in% states) %>%
+  select(ps, qs, abbreviation, value) %>% unnest(cols = c(ps, qs))
+
+dist_pqs_plot <- pqs %>% ggplot() +
+  geom_rect(
+    xmin = pqs$qs[pqs$ps == .025],
+    xmax = pqs$qs[pqs$ps == .975],
+    ymin = -Inf,
+    ymax = .1,
+    fill = "grey",
+    alpha = .01
+  ) +
+  geom_rect(
+    xmin = pqs$qs[pqs$ps == .1],
+    xmax = pqs$qs[pqs$ps == .9],
+    ymin = -Inf,
+    ymax = .1,
+    fill = "grey",
+    alpha = .02
[…]

### 2023-06-28 talks/ISF_talk/plot_allocs.R

@@ -7,26 +7,14 @@ source("_targets.R")
 
 file_names <- paste0("talks/ISF_talk/slim_dfs/slim_", mkeep)
 # create an empty list to store the loaded objects
-# slim_dfs <- list()
-# for(i in seq_along(file_names)){
-#   slim_dfs[[mkeep[i]]] <- readRDS(file_names[i])
-# }
-
-slim_dfs$`COVIDhub-4_week_ensemble`$model <- "COVIDhub-ensemble"
-
-get_ytot <- function(df) {
-  if ("xdf" %in% names(df)) {
-    df %>% group_by(reference_date) %>% slice(1) %>%
-    transmute(map_vec(xdf, ~ summarise(., ytot = sum(y))))
-  } else {
-    df %>% group_by(reference_date) %>%
-    summarise(ytot = sum(y))
+if (!exists("slim_dfs")) {
+  slim_dfs <- list()
+  for (i in seq_along(file_names)) {
+    slim_dfs[[mkeep[i]]] <- readRDS(file_names[i])
   }
 }
-add_ytot <- function(df) {
-  ytot <- get_ytot(df)
-  df %>% left_join(ytot, by = "reference_date")
-}
+
+slim_dfs$`COVIDhub-4_week_ensemble`$model <- "COVIDhub-ensemble"
 
 plot_components <- function(...) {
   alloscore::plot_components_slim(
@@ -61,5 +49,20 @@ color_vector <- rep(my_colors, length.out = 51)
 
 p +  scale_fill_manual(values = color_vector)
 
+# bar plot for muni-v-bucky
 
+mvbucky <- slim_dfs[c("JHUAPL-Bucky", "MUNI-ARIMA")] %>% bind_rows() %>%
+  filter(reference_date == "2021-12-27", K == 22000) %>% unnest(xdf) %>%
+  filter(abbreviation %in% c("FL", "CA", "TX", "GA", "SC", "NC", "PA", "VA", "WA"))
+
+p_mvb_bar <- plot_components(
+  mvbucky,
+  order_at_K = 22000,
+  order_at_model = "JHUAPL-Bucky",
+  bar_positioning = "dodge")
+p_mvb_bar + scale_y_continuous(expand = expansion(mult = c(0, 0.1))) + theme_bw() +
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank()) +
[…]

### 2023-06-28 talks/ISF_talk/plot_scores.R

@@ -2,21 +2,23 @@ library(covidHubUtils)
 library(alloscore)
 library(tidyverse)
 library(geofacet)
+library(latex2exp)
 
 source("_targets.R")
+save_dir <- file.path("talks/ISF_talk/plots/")
 
 score_viz <- tribble(
   ~palette,  ~linetypes, ~model,
-  "#004488", 2,          "BPagano-RtDriven",
-  "#117733", 4,          "COVIDhub-ensemble",
-  "#44AA99", 6,          "CU-select",
-  "#88CCEE", 2,          "IHME-CurveFit",
-  "#DDCC77", 4,          "JHUAPL-Bucky",
-  "#999933", 5,          "JHUAPL-Gecko",
-  "#CC6677", 1,          "MUNI-ARIMA",
-  "#882255", 1,          "USC-SI_kJalpha",
-  "#AA4499", 1,          "UVA-Ensemble",
-  "#000000", 1,          "COVIDhub-baseline"
+  "#004488", "dashed",   "BPagano-RtDriven",
+  "#117733", "dotdash",  "COVIDhub-ensemble",
+  "#694A49", "twodash",  "CU-select",
+  "#88CCEE", "dashed",   "IHME-CurveFit",
+  "#4A708B", "solid",  "JHUAPL-Bucky",
+  "#999933", "longdash", "JHUAPL-Gecko",
+  "#CD2626", "solid",    "MUNI-ARIMA",
+  "#882255", "longdash", "USC-SI_kJalpha",
+  "#FF8C00", "twodash",    "UVA-Ensemble",
+  "#000000", "solid",    "COVIDhub-baseline"
 )
 
 score_palette <- deframe(score_viz[c("model", "palette")])
@@ -31,4 +33,88 @@ plot_scores_slim(three_days) + lims(x=c(500, 30000))
 
 p_peak <- plot_scores_slim(peak_day)
 
-p_peak + theme_bw()
+p_peak_out <- p_peak + theme_bw() +
+  scale_x_continuous(expand = c(0,0), limits = c(5000,40000)) +
+  annotate(geom = "text", x = 22000, y = -250, size = 3,
+           label = paste("Total hospitalizations on 2022-01-10 →"), hjust = 1) +
+  labs(
+    x = expression(paste("Total Allocation Constraint,  ", italic("K"))),
+    y = TeX("Allocation Score = $\\sum (y_i - x_i^F)_{+}$")) +
+  theme(axis.title.x = element_text(size = 15),
[…]

### 2023-06-28 talks/ISF_talk/plotting_code/plot_functions.R

@@ -24,13 +24,17 @@ plot_hosp <- function(
     f_date = "2021-12-27",
     te_date = "2022-01-10",
     f_colors = pal1,
+    st_colors = NULL,
     f_alpha = .5,
     f_width1 = 2,
     space = .5,
     models = "COVIDhub-4_week_ensemble",
     locations = "US",
     geofacet = FALSE,
-    key_width = .15
+    key_width = .15,
+    allocations = FALSE,
+    forecast_hosp = forecasts_hosp,
+    free_y = FALSE
 ) {
   locations <- locations %>% purrr::map(
     function(loc) {
@@ -67,6 +71,15 @@ plot_hosp <- function(
       xmin = target_end_date - half_width + (match(model, models) - 1)*(f_width1 + space),
       xmax = xmin + f_width1)
 
+  if (!is.null(st_colors)) {
+    fc_dat <- fc_dat %>% mutate(
+      code = as.factor(code),
+      code = fct_relevel(code, names(st_colors)))
+    truth <- truth %>% mutate(
+      code = as.factor(code),
+      code = fct_relevel(code, names(st_colors)))
+  }
+
   p <- ggplot(data = truth) +
     # truth data
     geom_line(mapping = aes(x = target_end_date, y = value, color = validation)) +
@@ -101,9 +114,21 @@ plot_hosp <- function(
       mapping = aes(
         x = xmin,
         xend = xmax,
-        y = value, yend = value), size = 1.5) +
+        y = value, yend = value), size = 1.5)
     # predictive interval rects
-    geom_rect(
+    if (!is.null(st_colors)) {
+      p <- p + geom_rect(
+      data = get_intervals(fc_dat),
+      mapping = aes(
+        xmin = xmin,
+        xmax = xmax,
+        ymin = lower,
+        ymax = upper,
+        fill = code),
+      alpha = f_alpha) +
+        scale_fill_manual(values = st_colors, name = "State")
+   } else {
+    p <- p + geom_rect(
       data = get_intervals(fc_dat),
       mapping = aes(
         xmin = xmin,
@@ -111,8 +136,9 @@ plot_hosp <- function(
         ymin = lower,
         ymax = upper,
         fill = model),
-      alpha = f_alpha) +
-    scale_fill_manual(values = f_colors, name = "Model") +
[…]

### 2023-06-28 talks/ISF_talk/plotting_code/plot_hosps.R

@@ -11,6 +11,22 @@ save_dir <- file.path("talks/ISF_talk/plots/")
 hub_repo_path <- "~/covid/covid19-forecast-hub/"
 
 forecasts_hosp <- targets::tar_read(forecast_data) %>%
+  bind_rows(
+    load_forecasts(
+      dates = "2021-12-27",
+      date_window_size = 6,
+      locations = "US",
+      types = c("quantile"),
+      targets = paste(0:30, "day ahead inc hosp"),
+      source = "local_hub_repo",
+      hub_repo_path = hub_repo_path,
+      verbose = FALSE,
+      as_of = NULL,
+      hub = c("US")) |>
+      align_forecasts() |>
+      dplyr::filter(
+        relative_horizon == 14)
+  ) %>%
   filter(reference_date == "2021-12-27")
 truth <- targets::tar_read(truth_data)
 
@@ -25,13 +41,37 @@ mkeep <- c("BPagano-RtDriven",
            "USC-SI_kJalpha",
            "UVA-Ensemble")
 
-plot_hosp(models = c("CU-select"), f_width1 = 3, locations = "MA")
-plot_hosp(models = c("CU-select","COVIDhub-4_week_ensemble", "MUNI-ARIMA"),
+three_colors <- c("#4A708B", "goldenrod", "#CD2626")
+
+p_VA_CU_IHME <- plot_hosp(models = c("CU-select","COVIDhub-4_week_ensemble", "MUNI-ARIMA"),
+                     f_width1 = 3, f_alpha = .4, locations = "VA")
+
+sc = 1
+ggsave(plot = p_VA_CU_IHME, filename = "p_VA_CU_IHME.png",
+       width = 11*sc, height = 6*sc, path = save_dir)
+
+
+p_VA_CU_IHME
+### plot with allocations
+plot_hosp(models = c("CU-select", "UVA-Ensemble"), f_width1 = 4, f_alpha = .4, locations = "MA",
+          allocations = TRUE, f_colors = c("blue", "red"))
+
+plot_hosp(models = c("CU-select","COVIDhub-4_week_ensemble", "MUNI-ARIMA"),locations = c("CA", "MA"),
           space = 1, f_colors = pal1)
 
-nat_hosps <- plot_hosp(f_width1 = 4) +
-  ggtitle("US National Level Hospitalizations - Ensemble Forecast, 14 day horizon")
-nat_hosps
+
+# First plot for application section
+
+{
[…]

### 2023-07-10 alloscore_manuscript/allocation.tex
Update comment colors latex, and added an example allocation file that works on my comp
fix iter_pic, start allocation score discussion for paper intro

@@
 \usepackage{soul}
-\usepackage{color}
+\usepackage{xcolor}
 \def\elr#1{{\color{cyan}\textbf{ELR:[#1]}}}
@@
 Section \ref{sec:discussion} summarizes our contributions and discusses opportunities for further extensions in future work.
+
+\subsection{Allocation Scoring}
+
+In a nutshell, the allocation score of a multivariate forecast $F$ is the realized unmet demand when $F$ is used to allocate so as to minimize expected unmet 
+need --- that is, when we distribute a limited amount $K$ of our resource so that total expected unmet need according to $F$,
+\begin{align}
+\sum_i E_{F_i}[\text{unmet need in location $i$}]
+\end{align}
+is minimized, where the $F_i$ are the marginal distributions of $F$.
+This stochastic allocation problem has an intuitively appealing solution: allocate so that the probabilities of need exceeding allocation 
+in various locations are as close to each other as possible (see methods).
+In general, this will lead to the allocations provided by $F$ being either zero or the quantiles of $F_i$ for some \emph{single} probability level $\tau$.
+
+For example, suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Unif}(0,4)$ and $F_1 = \mathrm{Unif}(4,8)$. If we have $K = 10$ units
+of our resource available, the optimal allocation according to $F$ would be 3 units in location 1 and 7 units in location 2.
+Then both locations would be allocated at the quantiles for probability level $\tau = .75$.
+If, however, we only have $K = 3$ units available, all 3 would go to location 2 since it is not
+possible to match probabilities of need exceeding allocation given our resource constraint.
 

### 2023-07-10 notes/allocation_notes.rmd
fix iter_pic, start allocation score discussion for paper intro

@@
 \end{align}
-A forecaster believing that $Y \sim F$ will expect to receivq_e a score of $e_{\alpha}(F)$ by reporting their $\alpha$ quantile, while an
+A forecaster believing that $Y \sim F$ will expect to receiv`e a score of $e_{\alpha}(F)$ by reporting their $\alpha$ quantile, while an
 omniscient forecaster \emph{knowing} that $Y\sim G$ will expect a score of $e_S(Y)$ for reporting $q_{\alpha,Y}$. Note that if 
@@
 $L_{D,i}(\lambda):=\underset{0 \leq x}{\min} \left\{\lambda w_i x + \sbar_{F,i}(x)\right\}$ independently. 
-To this end, let $Z_{\lambda,i}$ be the set of $x \in \mathbb{R}$ where the weighted subgradient of $-\sbar_{F_i}(x)$ contains 
-$\lambda$ (c.f. \eqref{scalar_Z_deriv}):
-\begin{align}
-Z_{\lambda,i} := \left\{x \mid 
-\lambda \in -w_i^{-1}\partial\sbar_{F_i}(x) = w_i^{-1}\kappa_i g_i'(x)(\alpha_i - [F_i(x-), F_i(x)])\right\}. \label{dual_comp_deriv}
-\end{align}
-Then for $\lambda>0$ we have
+To this end, let $Z_{\lambda,i}$ be the set of critical points in the $i$th coordinate corresponding to the value $\lambda$, that is,
+the set
+\begin{align}
+Z_{\lambda,i} := \left\{x \in \mathbb{R} \cup \infty \mid 
+\lambda \in -w_i^{-1}\partial\sbar_{F_i}(x) = w_i^{-1}\kappa_i g_i'(x)(\alpha_i - [F_i(x-), F_i(x)])\right\} \label{dual_comp_deriv}
+\end{align}
+of $x$ where the weighted subgradient of $-\sbar_{F_i}(x)$ contains 
+$\lambda$ (c.f. \eqref{scalar_Z_deriv}). (Allowing $x=\infty$ accommodates the case when $\lambda=0$ and $\alpha_i = 1$.)
+Since $\lambda w_i x + \sbar_{F,i}(x) \geq 0$ for $\lambda \geq 0$, we have
 \begin{align}
@@
 eventually monotonically increasing in $x$,
-\item $Z_{\lambda,i}\cap [0,\infty) = \varnothing$ for $\lambda > \max_{x\geq 0}\{w_i^{-1}\kappa_i g_i'(x)(\alpha_i - F_i(x))\}$.
+\item $Z_{\lambda,i}\cap [0,\infty) = \varnothing$ for $\lambda > \max_{x\geq 0}\{w_i^{-1}\kappa_i g_i'(x)(\alpha_i - F_i(x))\}$, 
+and in this case we have $L_{D,i}(\lambda) = \sbar_{F_i}(0)$.
 \end{itemize}
@@
 
-## Special Case I: $F_i$ from same location-scale family and $w_i = 1$, $O_i = O, U_i = 1$.
+## Special Case I: $F_i$ from same location-scale family and $w_i = 1$, $\alpha_i = \alpha$.
 
@@
 \end{align}
-where $\Phi$ is the standard CDF for the family and $\alpha = (1+O)^{-1}$. Then setting
-\begin{align}
-\sum F_i^{-1}(\alpha - \lambda) = \sum \mu_i + \sigma_i\Phi^{-1}(\alpha - \lambda) = K
+where $\Phi$ is the base cdf for the family. Then solving for $\lambda^{\star}$ in
+\begin{align}
+\sum F_i^{-1}(\alpha - \lambda^{\star}) = \sum \mu_i + \sigma_i\Phi^{-1}(\alpha - \lambda^{\star}) = K
 \end{align}
@@
 \begin{align}
-\alpha - \lambda = 
-\Phi\left(\frac{K - \sum \mu_j}{\sum \sigma_j} \right):= F(K)
-\end{align}
-
-and the Bayes act when the constraint is set to $K$ is 
-\begin{align} 
- \nonumber Q_i(K) = F_i^{-1}(F(K)) &= \mu_i +\sigma_i\Phi^{-1}(F(K)) \\ 
+\alpha - \lambda^{\star} = 
+\Phi\left(\frac{K - \sum \mu_j}{\sum \sigma_j} \right):= F(K).
+\end{align}
+When $\alpha = 1$, we might interpet $F(K)$ as the forecast probability under an independence assumption that $K$ will be sufficient to meet demand.
+
+We can now use the assumption that the $F_i$ have a common base distribution to express the Bayes act as
+\begin{align} 
+ \nonumber x_i^{F,K} = F_i^{-1}(F(K)) &= \mu_i +\sigma_i\Phi^{-1}(F(K)) \\ 
  \nonumber &= \mu_i + \sigma_i\left(\frac{K - \sum \mu_j}{\sum \sigma_j} \right) \\
@@
 
-- $Q_i(K)$ here does not depend on $O$ unlike when the constraint is inactive or the $O_i$ differ.
+- The $x_i^{F,K}$ here do not depend on $\alpha$ unlike when the constraint is inactive or the $\alpha_i$ differ. Intuitively, when the distibutions $F_i$
+all have the same intrinsic shape, a change in $\alpha$ changes the level of marginal expected benefit identically in each coordinate, so that an equality
+of these expected benefits is maintained.
 

### 2023-07-10 random_undocumented_experiments/iter_pic.R

@@ -1,125 +1,35 @@
 library(tidyverse)
 library(patchwork)
 library(devtools)
-load_all()
+library(alloscore)
+source("random_undocumented_experiments/iter_pic_funs.R")
 
-(dat <- tibble(
-  alpha = c(.2,.6,.3,.9,.5),
-  w = c(3,2,3,5,2),
-  dists_and_params = list(
-    list(dist = "norm", mean = 3, sd = 1),
-    list(dist = "norm", mean = 2, sd = 4),
-    list(dist = "exp", rate = .5),
-    list(dist = "gamma", shape = 1, rate = .3),
-    list(dist = "unif", min = 5, max = 6)
-    #list(dist = "distfromq", ps = c(.1,.3,.6,.9,.95), qs = c(2,3,3,4,5))
-  ),
-  name = map(dists_and_params,
-             ~toString(paste0(.[["dist"]], ", ",
-                              toString(paste(names(.)[-1], .[-1], sep = "=")))))) %>%
-  unnest_wider(col = dists_and_params) %>%
-  add_pdqr_funs(types = c("p", "q")) %>% mutate(
-    q_scale = map2_dbl(Q, alpha, exec),
-    # add marginal expected benefit functions scaled to fit on single [0,1] interval,
-    Lambda_scaled = pmap(., margexb_fun),
-    # and unscaled as well, which are not yet being used
-    Lambda = pmap(.[names(.) != "q_scale"], margexb_fun),
-    Eloss = pmap(.[c("alpha", "F")], gpl_loss_exp_fun)
-  )
-)
+make_iter_pic(K = 2,
+              alpha = c(.2,.6,.3,.9,.5),
+              w = c(3,2,3,5,2),
+              dists_and_params = list(
+                list(dist = "norm", mean = 3, sd = 1),
+                list(dist = "norm", mean = 2, sd = 4),
+                list(dist = "exp", rate = .5),
+                list(dist = "gamma", shape = 1, rate = .3),
+                list(dist = "unif", min = 5, max = 6)
+                #list(dist = "distfromq", ps = c(.1,.3,.6,.9,.95), qs = c(2,3,3,4,5))
+              ))
 
-make_iter_pic(2, dat)
+make_iter_pic(K = 20,
+              dists_and_params = list(
+                list(dist = "unif", min = 0, max = 7),
+                list(dist = "unif", min = 7, max = 15)
+              ))
 
[…]

### 2023-07-10 random_undocumented_experiments/iter_pic_funs.R

@@ -1,0 +1,93 @@
+
+
+# make function of lambda that gives total resource use at "lambda-quantiles"
+# which are regular quantiles when lambda = 0.
+# When `!is.null(forecasts))`, the total use is for the subset of rows given by
+# vector `forecasts`
+K_fun_factory <- function(df, forecasts = NULL) {
+  if (!is.null(forecasts)) df <- df[forecasts,]
+  Vectorize(with(df, function(lambda) {
+    alpha_lam <- alpha-w*lambda
+    positives <- alpha_lam>0
+    qs_plus <- pmax(0, map2_dbl(Q, alpha_lam*positives, exec))
+    qs <- qs_plus*positives
+    sum(w * qs)
+  }))
+}
+
+
+#' Plot the iterations of `allocate` relative to meb's and total resource use
+#'
+#' @param K
+#' @param alpha
+#' @param w
+#' @param dists_and_params list of named lists each containing a `dist` entry indicating
+#'  distribution family (e.g. "norm" or "distfromq") and entries for the required parameters
+#'  (e.g. "mean" and "sd" or "ps" and "qs)
+#'
+#' @return
+#' @export
+#'
+#' @examples
+make_iter_pic <- function(K, alpha = 1, w = 1, dists_and_params) {
+  dat <- make_param_df(alpha = alpha, w = w, dists_and_params = dists_and_params)
+  dat_allo <- dat %>% allocate(K = K)
+  ymax = max(map2_dbl(dat$Lambda,0,exec))*(1.1)
+  xmax = max(dat_allo$xs[[1]][,2])*1.1
+  n = nrow(dat_allo$xdf[[1]])
+  K_fun <- K_fun_factory(dat)
+  with(dat_allo,
+       {
+         p1 <- ggplot() + xlim(c(0,xmax)) + ylim(0, ymax) +
+           geom_hline(yintercept = 0) +
+           map(lam_seq[[1]], ~ geom_hline(yintercept = ., alpha = .3)) +
+           geom_label(data = enframe(lam_seq[[1]]), aes(x=0,y=value,label=name)) +
+           map(1:n, ~ geom_function(aes(color = dat$name[[.]]), fun = dat$Lambda[[.]])) +
+           map(1:n, ~ geom_vline(aes(xintercept = x[[1]][.], color = dat$name[[.]]), alpha=.5)) +
+           labs(color = "Distribution",
+                y = expression(lambda(x)),
+                x = expression(x[i])) +
+           theme_classic() +
[…]

### 2023-07-10 talks/ISF_talk/iter_plot.R

@@ -2,6 +2,7 @@ library(scales)
 library(targets)
 adf <- tar_read(alloscore_BPagano.RtDriven_2022.01.03)
 
+source("_targets.R")
 tar_load(truth_data)
 tar_load(forecast_data)
 
@@ -11,7 +12,7 @@ require(tidyverse)
 require(alloscore)
 require(distfromq)
 
-one_model <- mkeep[4]
+one_model <- mkeep[1]
 one_reference_date <- "2021-12-27"
 
 ## process forecast data, adding distfromq output
@@ -45,10 +46,10 @@ dist_plot_dat <- forecast_data_processed %>% select(abbreviation, F) %>%
            y_values = map2(F, x_values, function(F, xl) F(xl))) %>%
   select(-F) %>% unnest(-abbreviation)
 
-dist_plot <- dist_plot_dat %>%
+dist_plot <- dist_plot_dat %>% filter(abbreviation == "AK") %>%
   ggplot(aes(x = x_values, y = y_values, color = abbreviation)) + geom_line()
 
-dist_plot + xlim(c(0, 1000))
+dist_plot + xlim(c(0, 5))
 
 # with quantiles for a subset of states
 
@@ -108,36 +109,56 @@ sc = 1.4
 ggsave(plot = dist_pqs_plot_OH, filename = "dist_pqs_plot_OH.png",
        width = 7*sc, height = 3*sc, path = save_dir)
 
-  ## run alloscore
-one_K = 10000
 
+plot_iterations <- function(
+    one_K = 10000,
+    model_name = one_model,
+    itnum = NULL,
+    states_to_keep = 6,
+    default_colors = scales::hue_pal()(l)[sample(1:l, l)]) {
+  ## run alloscore
   ascores <- forecast_data_processed %>%
     alloscore(
-      starting_K = 60000,
       K = one_K,
       y = .[["value"]],
       target_names = "abbreviation"
     )
   ascores <- ascores %>% dplyr::mutate(
     reference_date = one_reference_date,
-    model = one_model,
+    model = model_name,
     .before = 1
   )
 
 # make a plot of the xs iterations for one model across multiple dates
 
-iters_bp <- ascores %>% filter(model == one_model, K == one_K) %>%
+ascores <- ascores %>% filter(model == model_name, K == one_K)
+
+iters_bp <- ascores %>%
   select(model, reference_date, K, xs) %>% unnest(xs) %>%
[…]

### 2023-07-10 alloscore_manuscript/allocation.tex
Continue allocation score discussion in intro.

--- 2023-05-26 alloscore_manuscript/allocation.tex
@@
 Section \ref{sec:discussion} summarizes our contributions and discusses opportunities for further extensions in future work.
+
+\subsection{Allocation Scoring}
+
+In a nutshell, the allocation score of a multivariate forecast $F$ is the unnecessary realized unmet demand when $F$ is used to allocate so as to minimize expected unmet 
+need --- that is, when we distribute a limited amount $K$ of our resource so that total expected unmet need according to $F$,
+\begin{align}
+\sum_i E_{F_i}[\text{unmet need in location $i$}]
+\end{align}
+is minimized, where the $F_i$ are the marginal distributions of $F$.
+By ``unnecesary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
+This stochastic allocation problem has an intuitively appealing solution: allocate so that the probabilities of need exceeding allocation 
+in various locations are as close to each other as possible (see methods).  This will lead to the allocations provided by $F$ being quantiles of $F_i$ for some \emph{single} probability level $\tau$.
+
+For example, suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Unif}(0,8)$ and $F_2 = \mathrm{Unif}(4,8)$. If we have $K = 10$ units
+of our resource available, the optimal allocation according to $F$ would be 4 units in location 1 and 6 units in location 2.
+This allocates to both locations at their forecast medians.
+If, on the other hand, we only have $K = 3$ units available, we will necessarily have to allocate in location 2 at a quantile for $\tau = 0$, and therefore all 3 would go to location 2 since allocating any resources to location 1 would lead to unbalanced probabilities of need exceeding allocation.
+\footnote{Matching probabilities will not always be possible if we allow point masses in the $F_i$.
+For example, if $F_2$ were the point forecast of 8 interpreted as a degenerate probabilistic forecast with mass 1 at 8, the optimal allocation for $K = 10$ would be 2 in location 1 and 8 in location 2.
+Notice though that both allocations are still technically medians.}
+
+Next suppose that we observe needs of 8 and 3 in locations 1 and 2, respectively. With $K = 10$ units of the resource, the oracle would be able to prevent all
+but 1 of the total 11 units of needs from going unmet.
+The forecast $F$'s allocation results, at the same time, in $8 - 6 = 2$ units of unmet need in location 1 despite leaving no need unmet in location 2.
+The allocation score for the forecast when $K = 10$ would therfore be
+1 (= 2 realized $-$  1 unavoidable) in units of unnecessary unmet need.
+When $K = 3$, however, the forecast does not ``over-allocate'' in either location which leads to the same unmet need of $11 - 3 = 8$ as does the oracle's allocations.
+The allocation score for $F$ when $K = 3$ is therefore 0.
+
+Note the way in which $F$ incurred a positive (i.e., non-optimal) allocation score of 1 for $F$ when $K = 10$. It was not directly directly due to individual misalignments 
+of the marginal forecasts $F_i$ with the observed needs, but rather because the allocations and observed needs were not proportional as vectors.
+This illustrates a fundamental property of the allocation score.  For a given probability level $\tau$, if the quantiles $Q_i(\tau)$ of the marginal forecasts
+are proportional to the observed needs $y_i$, then the allocation score is zero for the resource constraint level $K = \sum Q_i(\tau)$.
+This stands in marked constrast to other common scoring methods for multi-variate forecasts that aggegate univariate scores such as CRPS or WIS for the marginal forecasts where a
+misalignment in one coordinate is penalized regardless of alignments in other coordinates. \apg{Should this be illustrated here, or elsewhere?}
+
+It is therefore often straighforward to constuct forecasts $F$ and $\widetilde{F}$ for a given outcome distribution that switch rankings under the allocation and traditional scores by ensuring that the marginal forecasts of $F$ center sharply around allocations that are proportionally similar to the central tendencies of the outcome distribution but are strongly biased.
+
+While multivariate scoring rules have not seen wide application in infectious disease forecast evalution, it seems that the allocation score would have a similar relationship with them, since while bias penalties in multi-variate scores can be offset by better forecasting of dependence structure, the proportional biases that allocation scoring are insensitive to are not \emph{per se} tolerated by multivariate scoring rules such as the energy score, variogram score, or Dawid-Sebastiani score.
 

### 2023-07-11 alloscore_manuscript/allocation.tex
1st attempt at integrating allocation score intro... we probably don't want subsections though

@@
 
+\subsection{Allocation Scoring}
+
 As described above, there exists both a rich literature on outbreak forecast evaluation and an emerging body of work on turning forecasts into resource allocation decisions. 
-However, there has been far less attention paid (no papers that we could identify) to exploring whether standard forecast accuracy metrics are sufficient for evaluating forecasts that are used to make allocation decisions.
-In this work, we address the gap between the ways in which infectious disease forecasts have been used to support public health policy and the ways in which they have traditionally been evaluated. 
-Specifically, we define a new forecast scoring rule --the {\em allocation score}-- that explicitly evaluates forecasts based on how accurate their implied allocations would turn out to be.
-This stands in contrast to standard evaluation metrics that evaluate the accuracy of a forecast relative to the surveillance data being forecasted.
-
-The remainder of this article is organized as follows. 
-In Section \ref{sec:methods}, we review the general framework for developing scoring rules for probabilistic forecasts using the tools of decision theory, develop a novel scoring rule that is motivated by the problem of allocating limited medical supplies, and explore the relationship between the proposed allocation score and existing scoring rules such as CRPS.
-In Section \ref{sec:application} we then illustrate the scoring rule through an application to short-term forecasts of COVID-19 hospital admissions in the US. 
-Section \ref{sec:discussion} summarizes our contributions and discusses opportunities for further extensions in future work.
-
-\subsection{Allocation Scoring}
+However, we have not found literature exploring whether standard forecast accuracy metrics are sufficient for evaluating how well forecasts support allocation decisions.
+We hope to begin in this work to fill this gap between the ways in which infectious disease forecasts have been used to support public health policy and the ways in which they have traditionally been evaluated.
+To this end, we define a new forecast scoring rule --- the {\em allocation score} --- that explicitly evaluates forecasts based on how beneficial their implied allocations would turn out to be rather than some standard statisitical measure of their accuracy relative to the surveillance data being forecasted.
 
@@
 While multivariate scoring rules have not seen wide application in infectious disease forecast evalution, it seems that the allocation score would have a similar relationship with them, since while bias penalties in multi-variate scores can be offset by better forecasting of dependence structure, the proportional biases that allocation scoring are insensitive to are not \emph{per se} tolerated by multivariate scoring rules such as the energy score, variogram score, or Dawid-Sebastiani score.
+
+\subsection{Organization}
+
+The remainder of this article is organized as follows. 
+In Section \ref{sec:methods}, we review the general framework for developing scoring rules for probabilistic forecasts using the tools of decision theory, develop a novel scoring rule that is motivated by the problem of allocating limited medical supplies, and explore the relationship between the proposed allocation score and existing scoring rules such as CRPS.
+In Section \ref{sec:application} we then illustrate the scoring rule through an application to short-term forecasts of COVID-19 hospital admissions in the US. 
+Section \ref{sec:discussion} summarizes our contributions and discusses opportunities for further extensions in future work.
 

### 2023-07-12 alloscore_manuscript/allocation.tex
Begin allocation solution algorithm description

@@
 
+\subsection{Solving the Allocation Problem}
+
+As the solution to a constrained optimization problem, the optimal allocation $\bf{x}^F$ according to a forecast $F$ can be described formally as the first component of a sationary point $(\bf{x}^F, \lambda^{\star})$ for the Lagrangian
+\begin{align}
+\mathcal{L}(\mathbf{x},\lambda; K, L) = \sum_{i=1}^{N} \mathbb{E}_{F_i}[s_A(x_i, Y_i; L)] + \lambda\left(\sum_{i=1}^{N} x_i - K\right).
+\end{align}
+That is, optimality requires that both
+\begin{align}
+0 = \nabla_{x} \mathcal{L}(\mathbf{x}^F, \lambda) &= \sum_{i=1}^{N} E_F \left[\frac{d}{dx_i}(Y_i - x_i^F)_{+}\right] + \lambda \\
+& = \sum_{i=1}^{N} E_F \left[-\mathbf{1}(Y_i \geq x_i^F)\right] + \lambda  = \sum_{i=1}^{N} F_i(x_i^F)-1 + \lambda  \label{eqn:x_stationary}\\
+\text{ and} \\
+0 = \nabla_{\lambda} \mathcal{L}(\mathbf{x}^F, \lambda^{\star}) &= \sum_{i=1}^{N} x_i^F - K\\
+& = \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star}) - K \quad \text{ (by \eqref{eqn:x_stationary})}. \label{eqn:lambda_stationary}
+\end{align}
+These equations give formulae $x_i^F = F_i^{-1}(1-\lambda^{\star})$, but in many situations of interest for us the $F_i^{-1}$ are not single-valued, and even 
+when they are the solution $\lambda^{\star}$ of \eqref{eqn:lambda_stationary} can be obtained analytically only in special cases (but see \ref{sec:methods.ls.example}).
+We therefore approximate allocations using a binary search for $\lambda^{\star}$ beginning with the interval $\lambda \in [0,1]$ and
+checking at each step the sign of the inequality that results from substituting the midpoint $\lambda_{\tau}$ into \eqref{eqn:lambda_stationary} for endpoints of the sets $F_i^{-1}(1-\lambda_{\tau})$.
+If $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) > K$, the quantiles for $1-\lambda_{\tau}$ exceed our resource constraint so we set $\lambda_{\tau}$ as the lower endpoint of the $\tau + 1$ search interval.
+Similarly, if $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) < K$, the quantiles for
+$1-\lambda_{\tau}$ do not make ful use of our resource and we set $\lambda_{\tau}$ as the lower endpoint of the $\tau + 1$ search interval.
+This process halves the search interval width at each step and we continue until this width falls below a given tolerance level $\varepsilon_{\tau}$.
+
 \section{Application}
@@
 
-We illustrate with an application to hospital admissions in the U.S., considering the problem of allocation of a limited supply of medical resources to the states.
+We illustrate with an application to hospital admissions in the U.S., considering the problem of allocation of a limited supply of medical resources to {}the states.
 

### 2023-07-13 alloscore_manuscript/allocation.tex
More on allocation algo...

--- 2023-07-11 alloscore_manuscript/allocation.tex
@@
 
+\subsection{Solving the Allocation Problem}
+
+As the solution to a constrained optimization problem, the optimal allocation $\bf{x}^F$ according to a forecast $F$ can be described formally as the first component of a stationary point $(\bf{x}^F, \lambda^{\star})$ for the Lagrangian
+\begin{align}
+\mathcal{L}(\mathbf{x},\lambda; K, L) = \sum_{i=1}^{N} \mathbb{E}_{F_i}[s_A(x_i, Y_i; L)] + \lambda\left(\sum_{i=1}^{N} x_i - K\right).
+\end{align}
+That is, optimality requires that both
+\begin{align}
+0 = \nabla_{x} \mathcal{L}(\mathbf{x}^F, \lambda) &= \sum_{i=1}^{N} E_F \left[\frac{d}{dx_i}(Y_i - x_i^F)_{+}\right] + \lambda \\
+& = \sum_{i=1}^{N} E_F \left[-\mathbf{1}\{Y_i \geq x_i^F\}\right] + \lambda  = \sum_{i=1}^{N} F_i(x_i^F)-1 + \lambda  \label{eqn:x_stationary}\\
+\text{ and} \\
+0 = \nabla_{\lambda} \mathcal{L}(\mathbf{x}^F, \lambda^{\star}) &= \sum_{i=1}^{N} x_i^F - K\\
+& = \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star}) - K \quad \text{ (by \eqref{eqn:x_stationary})}. \label{eqn:lambda_stationary}
+\end{align}
+These equations give formulae $x_i^F = F_i^{-1}(1-\lambda^{\star})$, but in many situations of interest for us the $F_i^{-1}$ are not single-valued, and even 
+when they are, a solution $\lambda^{\star}$ of \eqref{eqn:lambda_stationary} can only be obtained analytically in special cases (but see \ref{sec:methods.ls.example}).
+We therefore typically calculate allocation scores using approximately optimal allocations found via a binary search for $\lambda^{\star}$.
+This search begins with the interval $\lambda \in [0,1]$ and
+checks at each step the sign of the inequality that results from substituting the interval midpoint $\lambda_{\tau}$ into \eqref{eqn:lambda_stationary} for endpoints of the sets $F_i^{-1}(1-\lambda_{\tau})$.
+If $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) > K$, the quantiles for $1-\lambda_{\tau}$ exceed our resource constraint so we set $\lambda_{\tau}$ as the lower endpoint of the $\tau + 1$ search interval.
+Similarly, if $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) < K$, the quantiles for
+$1-\lambda_{\tau}$ do not make full use of our resource and we set $\lambda_{\tau}$ as the upper endpoint of the $\tau + 1$ search interval.
+This process halves the search interval width at each step and we continue until this width falls below a given tolerance level $\varepsilon_{\tau}$ yielding an
+approximation $\lambda^{\star, \varepsilon}$. When quantiles $F_i^{-1}(1-\lambda^{\star, \varepsilon})$ are singly defined and the forecast densities $f_i$
+not too close to zero at these quantiles, the discrepancy $\Delta(\lambda^{\star, \varepsilon})  = \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star, \varepsilon}) - K$ is of order $\varepsilon_{\lambda}$ and we can take $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ as approximations to $x_i^F$.
+
+Often though, $1-\lambda^{\star, \varepsilon}$ will be the probability level of a ``plateau'' for one or more $F_i$ and the $\lambda$ search algorithm will fail to drive $\Delta(\lambda^{\star, \varepsilon})$ close enough to zero for the $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ to be acceptable approximations of the optimal allocations.
+In this event we follow-up with a ``post-processing'' step that identifies lower and upper boundaries $x_{i,L}$ and $x_{i,U}$ of any
+such low-density plateau containing points in $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$. If $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ is single-valued 
+with the density $f_i(F_{i}^{-1}(1-\lambda^{\star, \varepsilon}))$ not close to zero, we simply take $x_{i,L} = F_{i}^{-1}(1-\lambda^{\star, \varepsilon}) = x_{i,U}$.
+Defining $\mathbf{x}_L = (x_{i,L})$ and $\mathbf{x}_U = (x_{i,U})$ we then interpolate an approximately optimal allocation
+$\mathbf{x} = (1-t^{\star})\mathbf{x}_L + t^{\star}\mathbf{x}_U$ where $t^{\star}$ solves 
+$\sum_{i=1}^{N }(1-t^{\star})x_{i,L} + t^{\star}x_{i,U} = K$.
+
+We note though that by using this linear interpolation (rather than some other point in the plateau satisfying the constraint) as the allocation associated with $F$ in the scoring process we are making a choice that is not
+guided by any expected utility maximization principle. As such, this choice may imply strong and potentially questionable assumptions about fairness and
+equity principles for the decision maker when forecasts tend to be ``clumped,'' leaving many regions with very low forecasted probability.
+
 \section{Application}

### 2023-07-13 R/run-alloscore.R
use slim dfs for alloscoring

@@ -33,7 +33,8 @@ run_alloscore <- function(forecast_data, truth_data, one_reference_date, one_mod
   alloscore(
     K = Ks,
     y = .[["value"]],
-    target_names = "abbreviation"
+    target_names = "abbreviation",
+    slim = TRUE
   )
   ascores <- ascores %>% dplyr::mutate(
     reference_date = one_reference_date,

### 2023-07-15 R/run-alloscore.R
update run_alloscore to give slim output and add refit "slim" targets

@@ -1,4 +1,9 @@
-run_alloscore <- function(forecast_data, truth_data, one_reference_date, one_model){
+run_alloscore <- function(
+    forecast_data,
+    truth_data,
+    one_reference_date,
+    one_model,
+    one_K = NULL){
   require(tidyverse)
   require(alloscore)
   require(distfromq)
@@ -25,8 +30,12 @@ run_alloscore <- function(forecast_data, truth_data, one_reference_date, one_mod
       by = c("location", "target_end_date"))
 
   if (nrow(forecast_data_processed) > 0) {
-  ytot <- sum(forecast_data_processed$value)
-  Ks <- seq(200, min(60000, 4*ytot), by = 200)
+    if (!is.null(one_K)) {
+      Ks <- one_K
+    } else {
+      ytot <- sum(forecast_data_processed$value)
+      Ks <- seq(200, min(60000, 4 * ytot), by = 200)
+    }
 
   ## run alloscore
   ascores <- forecast_data_processed %>%

### 2023-07-15 alloscore_manuscript/allocation.tex
add positivity req in allocation algo section

@@
 
-As the solution to a constrained optimization problem, the optimal allocation $\bf{x}^F$ according to a forecast $F$ can be described formally as the first component of a stationary point $(\bf{x}^F, \lambda^{\star})$ for the Lagrangian
+As the solution to a constrained optimization problem, the optimal allocation $\bf{x}^F$ according to a forecast $F$ can be described formally as the first component of a stationary point $(\bf{x}^F, \lambda^{\star})$ in the region $\{\mathbf{x} \geq 0, \lambda \geq 0\}$ for the Lagrangian
 \begin{align}

### 2023-07-17 alloscore_manuscript/allocation.tex
minor update to intro

@@
 However, explicit descriptions or quantifications of the value forecasts add to society through the decisions they support are elusive.
-Moreover, there has only been limited discussion of how predominant forecast evaluation metrics might indicate the success of policies based in part those forecasts.
+Moreover, there has only been limited discussion of how predominant forecast evaluation metrics might indicate the success of policies based in part on those forecasts.
 

### 2023-07-17 alloscore_manuscript/allocation.tex
add some uncommitted commented text from another branch, related to review of forecast eval via decision theory

@@
 The forecast skill ranking given by this allocation scoring rule can vary substantially from the ranking given by the weighted interval score now used by the CDC, especially during surges in hospitalizations such as in late 2021 as the Omicron wave began.
-We see this as strong evidence that the allocation scoring rule detects forecast value that is missed by traditional accuracy measures and that the general strategy of designing scoring rules directly linked to policy performance is a promising research direction for epidemic forecast evaluation.
+We see this as strong evidence that the allocation scoring rule detects forecast value that is missed by traditional accuracy measures and that the general strategy of designing scoring rules that are directly linked to policy performance is a promising research direction for epidemic forecast evaluation.
 
@@
 For example, estimates of expected onset of flu season were used to inform national vaccination strategies  \citep{igboh2023timing}, and forecasts of ebola dynamics were used to allocate surveillance resources \citep{meltzer2014estimating, rainisch2015regional}.
-In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision-making \citep{cdc2022cfa}.
-\cite{bertsimas2021predictionsCOVID} developed tools to inform decision making from infectious disease forecasts, which have been used to to inform allocation of limited medical supplies such as ventilators, ICU capacity planning, and vaccine distribution strategy.
+\cite{bertsimas2021predictionsCOVID} developed tools to inform decision making from infectious disease forecasts, which have been used to inform allocation of limited medical supplies such as ventilators, ICU capacity planning, and vaccine distribution strategy.
 Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping \citep{utnews2022}.
+In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision-making \citep{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
 \bwr{Figure out these news site citations!}
@@
 
+% Other important early contributions include Nelson (1972) and White (1966), who are amongst the first to consider the problem of forecast evaluation from the perspective of decision theory and dynamic programming.
+Nelson, in particular, articulates the concept of the “value of a (probability) fore- cast” to an individual decision-maker as the expected value of his/her payoff using the probability forecasts relative to the payoff expected if unconditional probability estimates are used.
+
+% This has been somewhat rectified during the last decade as the observation that financial series are predictable has led to an attempt to understand the economic significance of this finding.
+See, for example, Breen, Glosten, and Jagannathan (1989), Leitch and Tanner (1991), McCulloch and Rossi (1990), West, Edison, and Cho (1993), Pesaran and Timmermann (1994, 1995), Satchell and Timmermann (1995), and Skouras (1998).
+
+% Theil (1960), White (1966)
+
 However, we are aware of only a limited body of work that explicitly attempts to measure the value of infectious disease forecasts through their impact on policy, and much of this discussion has proceeded informally.
@@
 
-In practice, probabilistic infectious disease forecasts have most often been made for observations that emerge from public health surveillance systems and evaluated with standard, ``off-the-shelf'' scoring rules.
+In practice, probabilistic infectious disease forecasts have most often been made for observations that emerge from public health surveillance systems and have typically been evaluated with standard, ``off-the-shelf'' scoring rules.
 For example, seasonal influenza forecasts in the US and dengue forecasts for Peru and Puerto Rico targeted public health surveillance measures of incidence over time and space, and used log-score and mean absolute errors to evaluate forecast skill \citep{mcgowan_collaborative_2019,reich_collaborative_2019,johansson_open_2019}.
@@
 Similarly, CRPS was also used to assess probabilistic forecasts of dengue incidence at the district level in Vietnam \citep{colon-gonzalez_probabilistic_2021}.
-While some of these scores can be interpreted through the lens of decision theory, and all of the application-specific papers cited above had authors from public health agencies, none of them (including some for which we were co-authors) make explicit connections between forecast evaluation and how a forecast was used in practice.
+While some of these scores can be interpreted through the lens of decision theory, and all of the application-specific papers cited above had authors from public health agencies, none of them make explicit connections between forecast evaluation and how a forecast was used in practice.
 
-The motivating application of this work is one where forecasts of standard public health surveillance metrics (1) are used to help determine the allocation of a limited quantity of medical supplies across multiple regions, and (2) are later evaluated based on their recommended allocations.
-The analysis presented is ``synthetic'' in that it does not correspond to an actual analysis that supported decision-making in real-time.
+In this work, we hope to begin to fill this gap between the ways in which infectious disease forecasts have been used to support public health policy and the ways in which they have traditionally been evaluated.
+For concreteness, we consider a motivating application in which forecasts of public health surveillance signals are used to help determine the allocation of a limited quantity of medical supplies across multiple regions.
+We define a new forecast scoring rule --- the {\em allocation score} --- that explicitly evaluates forecasts based on how beneficial their implied allocations would turn out to be.
+Unlike the traditional measures of forecast skill such as WIS, CRPS, and log score, the proposed allocation score is rooted in this decision making context and measures the suitability of a forecast as an input to that specific decision making task.
+
+We present an illustrative analysis using the allocation score to evaluate forecasts of hospital admissions in the US leading up to the Omicron wave in winter 2022.
+This analysis is ``synthetic'' in that it does not correspond to an actual analysis that supported decision-making in real-time.
 However, the framework described in this paper corresponds to real-world decisions that must be made by public health administrators around the globe, and could be adapted in the future for such real-time situations.
-For example, forecasts for districts in Sierra Leone of bed demand to care for patients with Ebola was the subject of a real-time modeling study in late 2014 and early 2015\cite{camacho2015-ebola-bed}.
+For example, forecasts for districts in Sierra Leone of bed demand to care for patients with Ebola was the subject of a real-time modeling study in late 2014 and early 2015 \citep{camacho2015-ebola-bed}.
 And, in 2020, a model developed by an academic research group turned predictions of COVID-19 hospitalizations into estimates of ventilator usage and and shortages.
-This framework was used by the Hartford HealthCare system in Connecticut ``to align ventilator supply with projected demand at a time where the [COVID-19] pandemic was on the rise''\cite{bertsimas2021predictionsCOVID}.
+This framework was used by the Hartford HealthCare system in Connecticut ``to align ventilator supply with projected demand at a time where the [COVID-19] pandemic was on the rise'' \citep{bertsimas2021predictionsCOVID}.
 These examples illustrate the potential for forecasts to inform decisions about how to allocate limited supplies such as temporary hospital beds, ventilators, personal protective equipment, or other supplies that are known to be effective at reducing transmission or severity of disease.
+However, we emphasize again that these studies did not take the step of evaluating forecasts based on the quality of the allocation decisions that they supported or could have been used to support.
+
+% As described above, there exists both a rich literature on outbreak forecast evaluation and an emerging body of work on turning forecasts into resource allocation decisions. 
+% However, we have not found literature exploring whether standard forecast accuracy metrics are sufficient for evaluating how well forecasts support allocation decisions.
+
+% \subsection{Organization}
+
+The remainder of this article is organized as follows. 
+In Section \ref{sec:methods}, we review the general framework for developing scoring rules for probabilistic forecasts using the tools of decision theory, develop a novel scoring rule that is motivated by the problem of allocating limited medical supplies, and explore the relationship between the proposed allocation score and existing scoring rules such as CRPS.
+In Section \ref{sec:application} we illustrate the scoring rule through an application to short-term forecasts of COVID-19 hospital admissions in the US. 
+Section \ref{sec:discussion} summarizes our contributions and discusses opportunities for further extensions in future work.
+
+\section{Methods}
+\label{sec:methods}
+
+We give a high-level review of a general procedure for developing proper scoring rules that are tailored to a specific decision-making task in section \ref{sec:methods.decisiontheory}.
+In section \ref{sec:methods.quantileloss} we review how quantile loss can be obtained within this framework in a setting where a decision-maker is required to determine the quantity of a good to procure while balancing the cost of purchasing an additional unit of the good with loss that may result from under-procurement.
+We then discuss how the continuous ranked probability score (CRPS) can be obtained as an integral of the quantile loss across values of the cost/loss ratio.
+These developments mirror the structure of section \ref{sec:methods.allocation}.
+There, we develop a novel \emph{allocation score} that is analogous to the quantile score but is suitable for evaluation of forecasts in the context of decisions about allocation of limited resources across multiple locations when the resource constraint is known.
+We then describe an \emph{integrated allocation score} that is analogous to the CRPS and is obtained by integrating the allocation score across values of the resource constraint.
 
 \subsection{Allocation Scoring}
-
-As described above, there exists both a rich literature on outbreak forecast evaluation and an emerging body of work on turning forecasts into resource allocation decisions. 
-However, we have not found literature exploring whether standard forecast accuracy metrics are sufficient for evaluating how well forecasts support allocation decisions.
-We hope to begin in this work to fill this gap between the ways in which infectious disease forecasts have been used to support public health policy and the ways in which they have traditionally been evaluated.
-To this end, we define a new forecast scoring rule --- the {\em allocation score} --- that explicitly evaluates forecasts based on how beneficial their implied allocations would turn out to be rather than some standard statisitical measure of their accuracy relative to the surveillance data being forecasted.
 
@@
 
-While multivariate scoring rules have not seen wide application in infectious disease forecast evalution, it seems that the allocation score would have a similar relationship with them, since while bias penalties in multi-variate scores can be offset by better forecasting of dependence structure, the proportional biases that allocation scoring are insensitive to are not \emph{per se} tolerated by multivariate scoring rules such as the energy score, variogram score, or Dawid-Sebastiani score.
-
-\subsection{Organization}
-
-The remainder of this article is organized as follows. 
-In Section \ref{sec:methods}, we review the general framework for developing scoring rules for probabilistic forecasts using the tools of decision theory, develop a novel scoring rule that is motivated by the problem of allocating limited medical supplies, and explore the relationship between the proposed allocation score and existing scoring rules such as CRPS.
-In Section \ref{sec:application} we then illustrate the scoring rule through an application to short-term forecasts of COVID-19 hospital admissions in the US. 
-Section \ref{sec:discussion} summarizes our contributions and discusses opportunities for further extensions in future work.
-
-\section{Methods}
-\label{sec:methods}
-
-We give a high-level review of a general procedure for developing proper scoring rules that are tailored to a specific decision-making task in section \ref{sec:methods.decisiontheory}.
-In section \ref{sec:methods.quantileloss} we review how quantile loss can be obtained within this framework in a setting where a decision-maker is required to determine the quantity of a good to procure while balancing the cost of purchasing an additional unit of the good with loss that may result from under-procurement.
-We then discuss how the continuous ranked probability score (CRPS) can be obtained as an integral of the quantile loss across values of the cost/loss ratio.
-These developments mirror the structure of section \ref{sec:methods.allocation}.
-There, we develop a novel \emph{allocation score} that is analogous to the quantile score but is suitable for evaluation of forecasts in the context of decisions about allocation of limited resources across multiple locations when the resource constraint is known.
-We then describe an \emph{integrated allocation score} that is analogous to the CRPS and is obtained by integrating the allocation score across values of the resource constraint.
+While multivariate scoring rules have not seen wide application in infectious disease forecast evaluation, it seems that the allocation score would have a similar relationship with them, since while bias penalties in multi-variate scores can be offset by better forecasting of dependence structure, the proportional biases that allocation scoring are insensitive to are not \emph{per se} tolerated by multivariate scoring rules such as the energy score, variogram score, or Dawid-Sebastiani score.
 

### 2023-07-17 _targets.R
rebuilding and updating all targets.

@@ -34,41 +34,9 @@ mkeep <- c("BPagano-RtDriven",
 
 forecast_dates = as.character(seq.Date(as.Date("2021-10-18"), as.Date("2022-02-28"), by = "7 days"))
 
-forecast_dates_quick <- forecast_dates[c(12,13)]
-
 ## create a group of alloscore targets
-values_test <- tidyr::expand_grid(forecast_dates = forecast_dates_quick, models = mkeep[1])
 values <- tidyr::expand_grid(models = mkeep, forecast_dates = forecast_dates)
 
-# hard coding the names from tar_map
-# this seems like it should be its own target but I can't figure this out right now
-a_target_names <- values %>% dplyr::mutate(
-  target_name = paste0("alloscore_",
-                       gsub(x = models, pattern = "-", replacement = "."),
-                       "_",
-                       gsub(x = forecast_dates, pattern = "-", replacement = ".")
-  )) %>% dplyr::select(-forecast_dates)
-
-a_target_names_nested <- a_target_names %>% tidyr::nest(target_list = target_name)
-
-# function to create a single slim df for each model's alloscore time series
-# Don't use this! Will blow up your computer.
-# slim_dfs <- function(target_list) {
-#   target_list <- target_list %>% unlist(use.names = F)
-#   targets::tar_load(target_list)
-#   purrr::map(target_list, get) %>%
-#     purrr::map(purrr::possibly(~ alloscore::slim(., id_cols = c(
-#       "reference_date", "model"
-#     )))) %>%
-#     dplyr::bind_rows()
-#   rm(list = target_list)
-# }
-
-slim_df <- function(adf) {
-  require(tidyverse)
-  require(alloscore)
-  alloscore::slim(adf, id_cols = c("reference_date", "model"))
-}
 
 # List of targets:
 list(
@@ -76,10 +44,6 @@ list(
     name = forecast_data,
     command = get_forecast_data(forecast_dates)
   ),
-  # tar_target(
-  #   name = filtered_forecast_data,
-  #   command = filter_problematic_data(forecast_data)
-  # ),
   tar_target(
     name = truth_data,
     command = get_truth_data()
@@ -96,10 +60,6 @@ list(
                           forecast_dates,
                           models)
           )
-  # tar_map(
-  #   values = a_target_names,
-  #   names = target_name,
-  #   tar_target(slim, slim_df(target_name))
   )
   # tar_target(
   #   name = figure_K_v_alloscore,

### 2023-07-18 R/run-alloscore.R
adding first pass at aggregating slim alloscore data

@@ -1,3 +1,4 @@
+## engine to run the alloscores
 run_alloscore <- function(
     forecast_data,
     truth_data,
@@ -61,4 +62,21 @@ run_alloscore <- function(
 }
 
 
+## take estimated alloscores and put them in one clean dataset
+## returned object has a row for each model, reference_date, K, state
+assemble_alloscores <- function() {
+  require(dplyr)
+
+  ## load all alloscore targets and get their names
+  tar_load(starts_with("alloscore"))
+  ascore_tars <- ls(pattern="alloscore*")
+
+  ## bind all alloscore dataframes together and filter out ones with no forecasts
+  alloscores_with_data <- do.call(bind_rows, mget(ascore_tars)) |>
+    mutate(target = ascore_tars) |>
+    dplyr::filter(is.na(message))
+
+  ## extract alloscore matrices
+  alloscore::slim(alloscores_with_data, id_cols = c("model", "reference_date"))
+}
 

### 2023-07-18 _targets.R

@@ -60,6 +60,10 @@ list(
                           forecast_dates,
                           models)
           )
+  ),
+  tar_target(
+    name = alloscore_data,
+    command = assemble_alloscores
   )
   # tar_target(
   #   name = figure_K_v_alloscore,

### 2023-07-21 alloscore_manuscript/allocation.tex
add a proposal for a one-paragraph summary of what the allocation score does in intro

@@
 Unlike the traditional measures of forecast skill such as WIS, CRPS, and log score, the proposed allocation score is rooted in this decision making context and measures the suitability of a forecast as an input to that specific decision making task.
+
+Briefly, the allocation score of a forecast is the unnecessary realized unmet demand that results from using that forecast to set resource allocations by minimizing expected unmet need --- that is, when we distribute a limited amount $K$ of our resource so that total expected unmet need according to the forecast,
+\begin{align}
+\sum_i E_{F_i}[\text{unmet need in location $i$}]
+\end{align}
+is minimized, where the $F_i$ are the marginal forecast distributions for each location.
+By ``unnecesary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
+According to the allocation score, one forecast is better than another if it would lead decision-makers to an allocation that results in less unmet need.
 

### 2023-07-26 alloscore_manuscript/allocation.tex

--- 2023-07-17 alloscore_manuscript/allocation.tex
@@
 Unlike the traditional measures of forecast skill such as WIS, CRPS, and log score, the proposed allocation score is rooted in this decision making context and measures the suitability of a forecast as an input to that specific decision making task.
+
+Briefly, the allocation score of a forecast is the unnecessary unmet demand that results from using that forecast to set resource allocations by minimizing expected unmet need.
+In a little more detail, the total expected unmet need according to the forecast is
+\begin{align}
+\sum_i E_{F_i}[\text{unmet need in location $i$}],
+\end{align}
+where the $F_i$ are the marginal forecast distributions of demand in each location; for example, these might be forecasts of the level of need for medical resources in each state or hospital system.
+If there is a limited amount $K$ of a resource that is available to distribute, a decision-maker could choose an allocation of that resource across locations that minimizes this expected unmet need.
+%By ``unnecessary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
+According to the allocation score, one forecast is better than another if it would lead decision-makers to an allocation that results in less unmet need.
+We elaborate on these ideas in Section~\ref{sec:methods}.
 

### 2023-07-27 alloscore_manuscript/allocation.tex
add some stuff related to exponential distributions example

--- 2023-07-17 alloscore_manuscript/allocation.tex
@@
 
-\subsection{Allocation Scoring}
+\subsection{Overview of Allocation Scoring}
 
-In a nutshell, the allocation score of a multivariate forecast $F$ is the unnecessary realized unmet demand when $F$ is used to allocate so as to minimize expected unmet 
+The allocation score of a multivariate forecast $F$ is the unnecessary realized unmet demand when $F$ is used to allocate so as to minimize expected unmet 
 need --- that is, when we distribute a limited amount $K$ of our resource so that total expected unmet need according to $F$,
 \begin{align}
-\sum_i E_{F_i}[\text{unmet need in location $i$}]
+\sum_{i=1}^n E_{F_i}[\text{unmet need in location $i$}]
 \end{align}
 is minimized, where the $F_i$ are the marginal distributions of $F$.
-By ``unnecesary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
+By ``unnecessary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
 This stochastic allocation problem has an intuitively appealing solution: allocate so that the probabilities of need exceeding allocation 
 in various locations are as close to each other as possible (see methods).  This will lead to the allocations provided by $F$ being quantiles of $F_i$ for some \emph{single} probability level $\tau$.
+
+\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$ and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, with $\sigma_1 = 1 and $\sigma_2 = 5$.
+If we have $K = 5$ units of our resource available, the optimal allocation according to $F$ would be 4 units in location 1 and 6 units in location 2.
+This allocates to both locations at their forecast medians.
+If, on the other hand, we only have $K = 3$ units available, we will necessarily have to allocate in location 2 at a quantile for $\tau = 0$, and therefore all 3 would go to location 2 since allocating any resources to location 1 would lead to unbalanced probabilities of need exceeding allocation.
+
+suppose we have $n = 2$ locations, with a fixed total of $K=5$ units of available resources. to the forecasts $Y_1 \sim $ with $\sigma_1 = 1$ and $Y_2 \sim Exp(1 / \sigma_2)$ with $\sigma_2 = 5$.
 
@@
 The weighted interval score (WIS) is often used when the full forecast distribution $F$ is not available, as in the U.S.
-Forecast Hub and similar efforts where forecasts are represented by a collection of prediction intervals.
+COVID-19 Forecast Hub and similar efforts where forecasts are represented by a collection of prediction intervals.
 WIS is a discrete approximation to CRPS, and can be obtained by using a distribution $p$ that has point masses at the probability levels corresponding to the endpoints of a finite set of prediction intervals.
@@
 In contrast to the decision-making problem in the previous section, the administrator is not able to control the total amount of supply; rather, their task is to determine how to allocate the fixed supply to different locations.
+Following the structure of the discussion for the quantile score and CRPS, we first discuss the set up for a single value of the resource constraint $K$, which plays a role that is analogous to the cost/loss ratio for the quantile score, and then we discuss the possibility of integrating (or averaging) across values of $K$.
 
-In this decision-making setting, an action $\mathbf{x} = (x_1, \ldots, x_n)$ is a vector specifying the amount that is allocated to each of the $n$ locations.
+\subsubsection{The allocation score for a fixed resource constraint}
+
+In the decision-making setting that we consider, an action $\mathbf{x} = (x_1, \ldots, x_n)$ is a vector specifying the amount that is allocated to each of $n$ locations.
 We require that each $x_i$ is non-negative and that the total allocation across all locations does not exceed a constraint $K$ on the total available resources: $\sum_{i=1}^n x_i \leq K$.

### 2023-07-28 alloscore_manuscript/allocation.tex

--- 2023-07-27 alloscore_manuscript/allocation.tex
@@
 
+Briefly, the allocation score of a forecast is the unnecessary unmet demand that results from using that forecast to set resource allocations by minimizing expected unmet need.
+For example, suppose that a decision-maker is provided with forecasts of the level of need for medical resources in each of several states or hospital systems.
+If there is a limited amount $K$ of the medical resource that is available to distribute, a decision-maker could choose an allocation of that resource across locations that minimizes the expected unmet need according to that forecast.
+%By ``unnecessary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
+As measured by the allocation score, one forecast is better than another if it would lead decision-makers to an allocation that results in less unmet need.
+We elaborate on these ideas in Section~\ref{sec:methods}.
+
 We present an illustrative analysis using the allocation score to evaluate forecasts of hospital admissions in the US leading up to the Omicron wave in winter 2022.
@@
 
-The allocation score of a multivariate forecast $F$ is the unnecessary realized unmet demand when $F$ is used to allocate so as to minimize expected unmet 
-need --- that is, when we distribute a limited amount $K$ of our resource so that total expected unmet need according to $F$,
+\elr{I have abandoned the attempt to get a succinct one-sentence definition of the allocation score at the beginning here, because it felt like we were trying to pack too many ideas into one sentence.
+But I'm open to revisiting.
+Maybe the definition of alloscore can be moved earlier?}
+
+Suppose that a decision-maker is tasked with determining how much of a resource such as a medical supply should be allocated to each of $n$ locations, subject to the constraint that only $K$ units of the resource are available in total.
+If the decision-maker is provided with a multivariate forecast $F$ that predicts resource demand in each of the $n$ locations, one possibility is that they may choose the resource allocation that minimizes the expected total unmet demand according to the forecast.
+We will give a more precise mathematical statement in section \ref{sec:methods.allocation}, but at a high level, the total expected unmet demand according to the forecast is
 \begin{align}
-\sum_{i=1}^n E_{F_i}[\text{unmet need in location $i$}]
+\sum_{i=1}^n E_{F_i}[\text{unmet need in location $i$}],
 \end{align}
-is minimized, where the $F_i$ are the marginal distributions of $F$.
+where $F_i$ is the marginal forecast distribution of demand in location $i$.
+This stochastic allocation problem has an intuitively appealing solution: allocate so that the probabilities of need exceeding allocation in various locations are as close to each other as possible (see methods).
+This will lead to the allocations provided by $F$ being quantiles of the marginal distributions $F_i$ for some \emph{single} probability level $\tau$ that is shared in common for all locations.
+
+After time passes, the quality of a selected allocation can be measured by comparing the actual observed demand in each location to the amount of resources that were sent there.
+Specifically, we compute the total unmet demand that resulted from the selected allocation:
+\begin{align}
+    \sum_{i=1}^n \text{unmet demand in location $i$ based on the selected allocation},
+\end{align}
+One allocation is better than another if it results in lower total unmet demand.
+
+The \textbf{allocation score} of the forecast $F$ is the unnecessary unmet demand that results from using the allocation that minimizes the expected unmet need according to that forecast.
 By ``unnecessary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
-This stochastic allocation problem has an intuitively appealing solution: allocate so that the probabilities of need exceeding allocation 
-in various locations are as close to each other as possible (see methods).  This will lead to the allocations provided by $F$ being quantiles of $F_i$ for some \emph{single} probability level $\tau$.
 
-\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$ and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, with $\sigma_1 = 1 and $\sigma_2 = 5$.
-If we have $K = 5$ units of our resource available, the optimal allocation according to $F$ would be 4 units in location 1 and 6 units in location 2.
-This allocates to both locations at their forecast medians.
-If, on the other hand, we only have $K = 3$ units available, we will necessarily have to allocate in location 2 at a quantile for $\tau = 0$, and therefore all 3 would go to location 2 since allocating any resources to location 1 would lead to unbalanced probabilities of need exceeding allocation.
+\paragraph{Example 1} Suppose we have a forecast $F$ for demand in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$ and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, with $\sigma_1 = 2$ and $\sigma_2 = 3$.
+If we have $K = 5$ units of our resource available, the optimal allocation according to $F$ would be 2 units in location 1 and 3 units in location 2.
+If, on the other hand, we have $K = 10$ units available, we will allocate 4 units of resources in location 1 and 6 units in location 2.
+Figure~\ref{fig:exp_alloc_example} illustrates the situation, and derivations of the optimal allocations are given in section *** of the supplemental materials.
 
-suppose we have $n = 2$ locations, with a fixed total of $K=5$ units of available resources. to the forecasts $Y_1 \sim $ with $\sigma_1 = 1$ and $Y_2 \sim Exp(1 / \sigma_2)$ with $\sigma_2 = 5$.
+\begin{figure}
+    \includegraphics[width=\textwidth]{../figures/exponential_pred_expected_loss.pdf}
+    \caption{An illustration of the resource allocation problem in Example 1.
+There are $n = 2$ locations, with predictive distributions $F_1 = \mathrm{Exp}(1/2)$ and $F_2 = \mathrm{Exp}(1/3)$, illustrated in the panels at bottom and right.
+In the center panel, the background shading corresponds to the expected loss according to these forecasts.
+Diagonal black lines indicate resource constraints at $K=5$ and $K=10$ units.
+For these forecasts, the optimal allocations are $(2,3)$ for $K=5$ and $(4,6)$ at $K=10$.
+These allocations are at the point on the constraint line where the expected loss is smallest, which also corresponds to the point where a level set of the expected loss surface (blue curve) is tangent to the constraint.}
+    \label{fig:exp_alloc_example}
+\end{figure}
 
-For example, suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Unif}(0,8)$ and $F_2 = \mathrm{Unif}(4,8)$. If we have $K = 10$ units
-of our resource available, the optimal allocation according to $F$ would be 4 units in location 1 and 6 units in location 2.
-This allocates to both locations at their forecast medians.
-If, on the other hand, we only have $K = 3$ units available, we will necessarily have to allocate in location 2 at a quantile for $\tau = 0$, and therefore all 3 would go to location 2 since allocating any resources to location 1 would lead to unbalanced probabilities of need exceeding allocation.
-\footnote{Matching probabilities will not always be possible if we allow point masses in the $F_i$.
-For example, if $F_2$ were the point forecast of 8 interpreted as a degenerate probabilistic forecast with mass 1 at 8, the optimal allocation for $K = 10$ would be 2 in location 1 and 8 in location 2.
-Notice though that both allocations are still technically medians.}
+Next suppose that we observe needs of 2 and 9 in locations 1 and 2, respectively. With $K = 10$ units of the resource, the oracle would be able to prevent all
+but 1 of the total 11 units of needs from going unmet, for example by allocating 2 units of resources to location 1 and the remaining 8 units of resources to location 2.
+However, the forecast $F$'s allocation results in $9 - 6 = 3$ units of unmet need in location 2 despite leaving no need unmet in location 1.
+The allocation score for the forecast when $K = 10$ would therefore be 2 (= 3 realized $-$ 1 unavoidable) in units of unnecessary unmet need.
+When $K = 5$, however, the forecast does not ``over-allocate'' in either location which leads to the same unmet need of $11 - 5 = 8$ as does the oracle's allocations.
+The allocation score for $F$ when $K = 5$ is therefore 0.
 
-Next suppose that we observe needs of 8 and 3 in locations 1 and 2, respectively. With $K = 10$ units of the resource, the oracle would be able to prevent all
-but 1 of the total 11 units of needs from going unmet.
-The forecast $F$'s allocation results, at the same time, in $8 - 6 = 2$ units of unmet need in location 1 despite leaving no need unmet in location 2.
-The allocation score for the forecast when $K = 10$ would therfore be
-1 (= 2 realized $-$  1 unavoidable) in units of unnecessary unmet need.
-When $K = 3$, however, the forecast does not ``over-allocate'' in either location which leads to the same unmet need of $11 - 3 = 8$ as does the oracle's allocations.
-The allocation score for $F$ when $K = 3$ is therefore 0.
+Note the way in which $F$ incurred a positive (i.e., non-optimal) allocation score of 2 when $K = 10$. It was not directly directly due to individual misalignments 
+of the marginal forecasts $F_i$ with the observed needs, but rather because the allocations and observed needs were not proportional as vectors.
+Restating: as far as allocation decisions are concerned, with a fixed constraint $K=10$ the fundamental problem with the forecast $F$ is not that it predicts a mean total resource demand of $5$ units; it is that the realized demand was 4.5 times as large in location 2 as in location 1, but the forecast only indicated that the resource allocation for location 2 should be 1.5 times the allocation to location 1.
 
-Note the way in which $F$ incurred a positive (i.e., non-optimal) allocation score of 1 for $F$ when $K = 10$. It was not directly directly due to individual misalignments 
-of the marginal forecasts $F_i$ with the observed needs, but rather because the allocations and observed needs were not proportional as vectors.
-This illustrates a fundamental property of the allocation score.  For a given probability level $\tau$, if the quantiles $Q_i(\tau)$ of the marginal forecasts
-are proportional to the observed needs $y_i$, then the allocation score is zero for the resource constraint level $K = \sum Q_i(\tau)$.
+This illustrates a fundamental property of the allocation score: at its core, it measures whether the forecast accurately captures the relative magnitudes of resource demand across different locations, which is precisely the information that is needed to allocate resources to those locations subject to a fixed resource constraint.
+On the other hand, the allocation score is not directly sensitive to whether the forecasts in each location correctly capture the magnitude of resource demand in each individual location.
+%For a given probability level $\tau$, if the quantiles $Q_i(\tau)$ of the marginal forecasts are proportional to the observed needs $y_i$, then the allocation score is zero for the resource constraint level $K = \sum Q_i(\tau)$.
 This stands in marked constrast to other common scoring methods for multi-variate forecasts that aggegate univariate scores such as CRPS or WIS for the marginal forecasts where a 
-misalignment in one coordinate is penalized regardless of alignments in other coordinates. \apg{Should this be illustrated here, or elsewhere?}
+misalignment in one coordinate is penalized regardless of alignments in other coordinates.
+\apg{Should this be illustrated here, or elsewhere?} \elr{Not sure what I think.
+Comparisons with WIS/etc seem critical, but some of the text I've added here feels like text for the discussion rather than up here in the ``methods"...}
+Note that we do not claim the allocation score is generically preferable to these other scores \textemdash rather, it provides a view of forecast performance that is specifically tuned to the context of decision-making about resource allocations.
 
-It is therefore often straighforward to constuct forecasts $F$ and $\widetilde{F}$ for a given outcome distribution that switch rankings under the allocation and traditional scores by ensuring that the marginal forecasts of $F$ center sharply around allocations that are proportionally similar to the central tendencies of the outcome distribution but are strongly biased.
+%It is therefore often straighforward to constuct forecasts $F$ and $\widetilde{F}$ for a given outcome distribution that switch rankings under the allocation and traditional scores by ensuring that the marginal forecasts of $F$ center sharply around allocations that are proportionally similar to the central tendencies of the outcome distribution but are strongly biased.
 
-While multivariate scoring rules have not seen wide application in infectious disease forecast evaluation, it seems that the allocation score would have a similar relationship with them, since while bias penalties in multi-variate scores can be offset by better forecasting of dependence structure, the proportional biases that allocation scoring are insensitive to are not \emph{per se} tolerated by multivariate scoring rules such as the energy score, variogram score, or Dawid-Sebastiani score.
+%While multivariate scoring rules have not seen wide application in infectious disease forecast evaluation, it seems that the allocation score would have a similar relationship with them, since while bias penalties in multi-variate scores can be offset by better forecasting of dependence structure, the proportional biases that allocation scoring are insensitive to are not \emph{per se} tolerated by multivariate scoring rules such as the energy score, variogram score, or Dawid-Sebastiani score.
 

### 2023-07-28 notes/crps_connection_exp.Rmd

@@
 
+Suppose we adjust the scale parameters in both locations by the multiplicative factor $c$, so that our updated forecast distributions are $Y_1 \sim Exp(1/(c\sigma_1))$ and $Y_2 \sim Exp(1 / (c\sigma_2))$.
+As a function of $c$, the optimal allocations are at the quantile level
+
+$$\tau(c) = h^{-1}(K) = 1 - \exp[-K/(c\sigma_1 + c\sigma_2)]$$
+
+The optimal allocations are then
+\begin{align*}
+F_i^{-1}(\tau(c)) &= -c\sigma_i \log(1 - \tau(c)) \\
+&= -c \sigma_i -K/(c\sigma_1 + c\sigma_2) \\
+&= \frac{\sigma_i}{\sigma_1 + \sigma_2} K
+\end{align*}
+
+Note that these allocations do not depend on the parameter $c$, only the relative scales in the two locations.
+Since the allocations are the same for any pairs of forecast distributions with the same relative scales, the allocation score is also the same.
+While this result is specific to exponentially-distributed forecasts, it illustrates an important feature of the allocation score: in general, the allocation score does not measure whether forecast distributions in individual locations are correct in an absolute sense; instead, it measures whether the relative magnitudes in different locations are forecasted correctly.
+It is these relative magnitudes that determine whether or not forecast-informed allocations send resources to the correct locations.
+
 \section{Obtaining the CRPS weighting associated with a distribution on $K$}

### 2023-08-03 alloscore_manuscript/allocation.tex
some writing and bib edits to paragraph 1

@@
 
-% Rainisch G, Shankar M, Wellman M, Merlin T, Meltzer MI.
-Regional spread of Ebola virus, West Africa, 2014.
-Emerg Infect Dis 2015;21:444–7 Forecast of geographic spread of Ebola which helped aid decision-makers in allocation of surveillance resources.
-
-% MI Meltzer - Estimating the Future Number of Cases in the Ebola (example of Ebola prediction informing policy to increase ERU capacities among other measures)
-
 %https://www.cdc.gov/mmwr/preview/mmwrhtml/su6303a1.htm
 
-Infectious disease forecasting models have emerged as important tools in public health decision making.
-Predictions of disease dynamics have been used to inform decision-making about a wide variety of measures designed to reduce disease spread and/or mitigate the severity of disease outcomes.
-For example, estimates of expected onset of flu season were used to inform national vaccination strategies  \citep{igboh2023timing}, and forecasts of ebola dynamics were used to allocate surveillance resources \citep{meltzer2014estimating, rainisch2015regional}.
+Infectious disease forecasting models have emerged as important tools in public health.
+Predictions of disease dynamics have been used to inform decision-making about a wide variety of measures to reduce disease spread and/or mitigate the severity of disease outcomes.
+For example, estimates of expected onset of flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola dynamics have been used to allocate surveillance resources \citep{meltzer2014estimating, rainisch2015regional}.
 \cite{bertsimas2021predictionsCOVID} developed tools to inform decision making from infectious disease forecasts, which have been used to inform allocation of limited medical supplies such as ventilators, ICU capacity planning, and vaccine distribution strategy.
-Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping \citep{utnews2022}.
-In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision-making \citep{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
-\bwr{Figure out these news site citations!}
+Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping (University of Texas, 2022) \nocite{utnews2022}.
+In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision-making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
+The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid} or weighted interval score (WIS) \citep{bracher2021evaluating}, however, recently authors have been calling for evaluating forecasts through their impact on policy \citep{marshall2023predictions}.
 

### 2023-08-07 alloscore_manuscript/application.Rnw
adding an unknit Rnw file for an application section. having trouble getting it to knit offline with targets. will troubleshoot later.

\documentclass{article}

\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

\usepackage{amsmath, amsfonts, amssymb}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{cases}
\usepackage{caption}
\usepackage{hyperref}

\usepackage{setspace}
\onehalfspacing

\usepackage{parskip}

\usepackage{soul}
\usepackage{xcolor}
\def\elr#1{{\color{cyan}\textbf{ELR:[#1]}}}
\def\apg#1{{\color{red}\textbf{APG:[#1]}}}
\def\bwr#1{{\color{violet}\textbf{BWR:[#1]}}}
\def\ngr#1{{\color{blue}\textbf{NGR:[#1]}}}

\title{Using the allocation scoring rule to evaluate short-term COVID-19 hospitalization forecasts during the 2021-2022 Omicron wave}
\author{Aaron Gerding, Nicholas G. Reich, Benjamin Rogers, Evan L. Ray}

\begin{document}

\maketitle

<<setup, tidy=FALSE, echo=FALSE, message=FALSE>>=
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
@

<<environment-setup>>=
library(targets)
library(tidyverse)
library(ggbump)

theme_set(theme_bw())
@

\section{Methods}

\ngr{TODO: add a figure showing some forecasts and ground truth data.}

\subsection{Hospitalization data}
Starting in the summer of 2020, the US Health and Human Services began reporting counts of daily new admissions to hospitals for individuals with COVID-19.(cite HHS Protect)
These data served as the source of ``ground truth'' data for the US COVID-19 Forecast Hub which, starting in December 2020, collected short-term forecasts of new hospital admissions at the daily scale.
These daily counts were available for the US as a whole, and all states and several additional jurisdictions such as Puerto Rico and Washington DC.
The data were updated daily and were available for download by the public through the HHS HealthData.gov website.
For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data.(cite package) \ngr{can we use tar\_manifest() to reproducibly display the date the data were stored?}

\subsection{Forecast data}
The US COVID-19 Forecast Hub, a consortium funded by the US CDC and led by a research group at UMass-Amherst, collected short-term forecasts of hospitalizations starting in December 2020.(cite Data descriptor paper)
Any team that with appropriately formatted forecasts could submit them to the Forecast Hub data repository on GitHub.(cite repo site)
Forecasts were time-stamped by GitHub upon submission and passed validation checks that ensured correct formatting and that the forecasts were being submitted only for dates in the future, not for data that had already been observed.

Forecast submission followed a weekly cycle and culminated in the creation of an ensemble forecast.
Forecasts could be submitted on any day during the week.
However once a week on Mondays, the Forecast Hub would collect the most recent forecasts submitted by all teams that met certain inclusion criteria and create an ensemble forecast using quantile averaging.(cite Evan's paper)
An ensemble that treated all models equally was created (COVIDhub-ensemble) as was a model that created weights of submitted models based on performance in the past 12 weeks (COVIDhub-trained\_ensemble).
One other model that combined multiple forecasts from different teams but used a different ensembling algorithm, a linear pooling method with tail extrapolation, was also included in our analyses (JHUAPL-SLPHospEns).
Several other models have ``ensemble'' in their name, but this refers to combinations of different variations of models that the specific team created, not to a multi-model ensemble combining different submitted forecasts to the Forecast Hub.

All forecasts, including the ensemble, were submitted as probabilistic predictions about the number of new hospital admissions on a particular day in the future, in a specific jurisdiction of the US (national level, state, or territory).
Probability distributions were specified, per the requirements established by the Forecast Hub, as a set of 23 quantiles for each individual prediction.
The submitted quantiles included a median (treated as a ``point'' prediction) and defined 11 central prediction intervals, from a 99\% to a 10\% prediction interval.

The analysis in this work focuses on forecasts made before and during the first wave of the Omicron SARS-CoV-2 variant in the US.
As such, we downloaded forecasts for the 15 weeks starting with Monday 2021-11-22 through Monday 2022-02-28.

We established a set of inclusion criteria to determine which forecasts and models to include in our analysis.
Models were eligible to be included in the analysis if they were considered a ``primary'' model from a team. (If a team submitted multiple versions of similar models, they were required to designate one as ``primary''.)
For a model to have a complete, eligible submission in a given week, it had to have a 14 day-ahead forecast for all 50 states plus Washington DC.
Models had to have a complete forecast for at least 4 of the 15 weeks in the analysis to be considered eligible for inclusion.

\subsection{Evaluation metrics}

This manuscript focuses on two proper forecast scores, the allocation score and the weighted interval score (WIS).(cite Gerding and Bracher)
Propriety of forecast scoring rules is desirable as it ensures that forecasters are not incentivized to modify their forecast distribution to achieve a better score.(cite Gneiting and Raftery or Raftery and Gneiting)

\subsubsection{Allocation Score}

\ngr{TODO: general introduction to the alloscore}

For the analysis, we fixed a resource constraint $K$ to be 15,000, based roughly on a reported number of ventilators available for reallocation in the US.(cite paper)
For each week, we computed a the allocation score for each 14 day-ahead forecast based on $K=15,000$.
\ngr{it appears that for the last two weeks selected (forecast dates: 2022-02-14 and 2022-02-28) the upper limit of K was 11400 and 9200 respectively, so we do not have alloscores computed for K=15000. Maybe they would have all been zero, since the limits were likely set based on the number of cases being low. What is the rule used to determine the max K computed? Don't think we need to change it, but would be nice to state in the manuscript what it is.}

We also computed a standardized rank for the allocation score for each model $m$ and week $w$.
First, we computed the number of models that forecasted that week ($n_w$) and the rank of model $m$ among the $n_w$ models ($r_{m,w}^{AS}$).
The model with the best allocation score received a rank of 1 and the worst received a rank of $n_w$.
In the case of a tie between one or more models, all models received the better rank.
We then rescaled these rankings to compute the allocation score standardized rank ($sr_{m,w}^AS$) between 0 and 1, where 0 corresponds to the worst rank and 1 to the best.

\begin{equation}
sr_{m,w}^{AS} = 1 - \frac{r^{AS}_{m,w}-1}{n_w-1}
\end{equation}

\subsubsection{Weighted Interval Score (WIS)}

\ngr{TODO: general introduction to WIS and define MWIS as mean WIS.}

We computed standardized ranks for MWIS ($sr_{m,w}^{MWIS}$) using the same procedure as for allocation scores described above.

\subsection{Data and code availability}

\section{Results}

<<score-computation>>=
tar_load("all_alloscore_data")

alloscore_data_summaries <- all_alloscore_data |>
  filter(K == 15000)|>
  group_by(model, reference_date) |>
  summarize(alloscore = sum(components), yobs = sum(y)) |>
  group_by(reference_date) |>
  mutate(nmodels = n(),
         allo_rank = rank(alloscore, ties.method = "min"),
         allo_rank_std = 1- (allo_rank-1)/(nmodels-1)) |>
  ungroup() |>
  complete(model, reference_date) |>
  mutate(target_end_date = as.Date(reference_date) + 14)

tar_load("score_data")

alloscore_models <- unique(alloscore_data_summaries$model)
score_data_summaries <- score_data |>
  mutate(reference_date = target_end_date - 14) |>
  filter(model %in% alloscore_models) |>
  group_by(model, reference_date) |>
  summarize(mwis = mean(wis)) |>
  ungroup() |>
  group_by(reference_date) |>
  mutate(nmodels = n(),
         mwis_rank = rank(mwis, ties.method = "min"),
         mwis_rank_std = 1- (mwis_rank-1)/(nmodels-1)) |>
  ungroup() |>
  complete(model, reference_date) |>
  mutate(target_end_date = as.Date(reference_date) + 14)
@

\subsection{Metrics as a function of time}
<<grob-hosps-over-time>>=
tar_load("truth_data")
dat_to_plot <- truth_data |>
  filter(location == "US",
         target_end_date %in% score_data_summaries$target_end_date)
p1 <- ggplot(dat_to_plot,
             aes(x=target_end_date, y=value)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept=15000, linetype=2) +
  ylab("hospital admissions") +
  xlab(NULL) +
  ylim(c(0, NA))
@

<<metrics-over-time, fig.height=9, fig.width=7, fig.cap="Hospital admissions and evaluation metrics over time. Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022. These are the days for which forecasts were made and evaluated. A horizontal dashed line at 15,000 shows the assumed resource constraint $K$. Panel B shows allocation scores for each model's 14-day ahead forecast, across all US states. The x-axis corresponds to the date for which the prediction was made. Allocation scores typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down). Panel C shows the MWIS metric across weeks, averaged across all states. Similarly to panel B, the x-axis corresponds to the date for which the prediction was made. MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the pean of the Omicron wave.">>=
p2 <- ggplot(alloscore_data_summaries,
             aes(x=target_end_date, y=alloscore, color=model, shape=model)) +
  scale_x_date(limits=range(score_data_summaries$target_end_date)) +
  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
  ylab("Allocation score") +
  geom_point() + geom_line()+
  theme(legend.pos="none") +
  xlab(NULL)

p4 <- ggplot(score_data_summaries,
             aes(x=target_end_date, y=mwis, color=model,shape=model)) +
  geom_point() + geom_line() +
  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
  ylab("MWIS")+
  theme(legend.pos="bottom")+
  xlab(NULL)

cowplot::plot_grid(p1, p2, p4,
                   nrow=3,
                   align="v",
                   rel_heights = c(1, 1, 1.5),
                   labels = "AUTO")
@

Allocation scores varied substantially by date and by model (Figure XX).
For predictions made for the first three Mondays in December 2021 and the last three Mondays in February 2022 all models had allocation scores under 500 (and the mean across all models was less than 100), indicating that the unnecessary unmet need was fairly low relative to the total number of hospital admissions on those days.
The allocation scores are on the whole highest when the observed number of new hospital admissions is closest to the resource threshold of 15,000, as this is the time when any mistakes in allocation are costly in terms of wasting resources in one location that could have been used in another.
Predictions made during the peak week and just after showed the highest variation in allocation scores, with some models having values under 1000 and others having values over 3500.
<<>>=
## some analysis to support above statements
alloscore_data_summaries |>
  group_by(target_end_date) |>
  summarize(meanAS = mean(alloscore, na.rm=TRUE),
            medianAS = median(alloscore, na.rm=TRUE),
            minAS = min(alloscore, na.rm=TRUE),
            maxAS = max(alloscore, na.rm=TRUE),
            range = maxAS-minAS)
@

Mean weighted interval scores (MWIS) also varied by date and model, and more clearly were dependent on the scale of the observed data.
MWIS values were low (all models under 100) for all Mondays in December 2021 and the final four Mondays evaluated.
Across all models both the average and median MWIS value for every Monday in January was above 100, with the largest errors occurring one and two weeks after the peak was observed.

<<>>=
## some analysis to support above statements
score_data_summaries |>
  group_by(target_end_date) |>
  summarize(meanWIS = mean(mwis, na.rm=TRUE),
            medianWIS = median(mwis, na.rm=TRUE),
            minWIS = min(mwis, na.rm=TRUE),
            maxWIS = max(mwis, na.rm=TRUE),
            range = maxWIS-minWIS)
@

\subsection{correlation between WIS and allocation score}

<<metrics-correlation, fig.width=7, fig.height=6, fig.cap="Association of standardized ranks for MWIS and allocation score by model and week. Each facet of the plot corresponds to one model. Within each facet, each point corresponds to a week. The x- and y-values correspond to the MWIS standardized rank and the allocation score standardized rank for that week. Points corresponding to earlier dates have darker shading. The size of the point corresponds to the observed value on the date for which the prediction was made. Models show different degrees of association between the two metrics.">>=
all_data <- left_join(alloscore_data_summaries, score_data_summaries,
                      by=c("model", "target_end_date"))

ggplot(all_data,
       aes(x=mwis_rank_std, y=allo_rank_std, color=target_end_date, size=yobs)) +
  geom_point(alpha=.8) +
  scale_x_continuous("MWIS standardized rank", limits=c(0,1)) +
  scale_y_continuous("Allocation score standardized rank", limits=c(0,1)) +
  facet_wrap(.~model) +
  geom_abline(slope=1, intercept=0, linetype=2, color="gray")

@

 - bubble figure with ranks sized by total observations

 - different models show different levels of correlation between the metrics

\subsection{Integrated allocation score across values of K}

\section{Discussion}

 - WIS is scale dependent, alloscore not as much

\end{document}

### 2023-08-07 alloscore_manuscript/application.Rnw
some additional content.

@@
        aes(x=mwis_rank_std, y=allo_rank_std, color=target_end_date, size=yobs)) +
-  geom_point(alpha=.8) +
+  geom_point(alpha=.7) +
   scale_x_continuous("MWIS standardized rank", limits=c(0,1)) +
@@
 
+Models show differing levels of correlation between their allocation scores and MWIS values (Figure XX).
+Here are some examples of the different patterns observed:
+\begin{itemize}
+\item positive association between allocation score and MWIS ranks: \texttt{Karlen-pypm} and \texttt{USC-SI_kJalpha} models
+\item consistently strong MWIS ranks, highly variable allocation score ranks, no clear association: \texttt{JHUAPL-SLPHospEns} and \texttt{CU-select}
+\item consistently strong ranks for both metrics, no clear association: \texttt{COVIDhub-ensemble}
+\end{itemize}
+Additionally, the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, had the highest rank for allocation score in six weeks, more than any other model except the \texttt{COVIDhub-ensemble} which also had six.
+
+<<>>=
+alloscore_data_summaries |>
+  group_by(model) |>
+  summarize(top_rank = sum(allo_rank_std==1))
+@
+
  - bubble figure with ranks sized by total observations
-
- - different models show different levels of correlation between the metrics
 
@@
 
+ - something here from Ben's analysis?
+
 \section{Discussion}
+
+ - It is clear that the metrics are capturing different aspects of performance
 
@@
 
+ - the fact that it is hard to beat the baseline for allocation suggests that we are not consistently adding value over just looking at the current levels.
+\ngr{although is there something more to it about how the baseline accounts for uncertainty?}
+
 \end{document}

### 2023-08-09 alloscore_manuscript/allocation.tex
updates to description of alloscore in intro and overview

@@
 
-In this work, we hope to begin to fill this gap between the ways in which infectious disease forecasts have been used to support public health policy and the ways in which they have traditionally been evaluated.
-For concreteness, we consider a motivating application in which forecasts of public health surveillance signals are used to help determine the allocation of a limited quantity of medical supplies across multiple regions.
-We define a new forecast scoring rule --- the {\em allocation score} --- that explicitly evaluates forecasts based on how beneficial their implied allocations would turn out to be.
-Unlike the traditional measures of forecast skill such as WIS, CRPS, and log score, the proposed allocation score is rooted in this decision making context and measures the suitability of a forecast as an input to that specific decision making task.
+In this work, we begin to fill this gap between the ways that infectious disease forecasts have traditionally been evaluated and the ways that they have been used to support public health policy.
+We consider a setting in which forecasts are used to help determine the allocation of a limited quantity of medical supplies across multiple regions.
+We define a new forecast scoring rule --- the {\em allocation score} --- that evaluates forecasts based on how beneficial resource allocations derived from them would turn out to be.
+%Unlike the traditional measures of forecast skill such as WIS, CRPS, and log score, the proposed allocation score measures the suitability of a forecast as an input to this specific decision making task.
 
-Briefly, the allocation score of a forecast is the unnecessary unmet demand that results from using that forecast to set resource allocations by minimizing expected unmet need.
+Briefly, the allocation score of a forecast is the avoidable unmet demand that results from using that forecast to set resource allocations by minimizing expected unmet need.
 For example, suppose that a decision-maker is provided with forecasts of the level of need for medical resources in each of several states or hospital systems.
-If there is a limited amount $K$ of the medical resource that is available to distribute, a decision-maker could choose an allocation of that resource across locations that minimizes the expected unmet need according to that forecast.
+If there is a limited amount of the medical resource that is available to distribute, a decision-maker could choose an allocation of that resource across locations that minimizes the expected unmet need according to that forecast.
+As measured by the allocation score, one forecast is better than another if it would lead decision-makers to an allocation that results in less unmet need.
 %By ``unnecessary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
-As measured by the allocation score, one forecast is better than another if it would lead decision-makers to an allocation that results in less unmet need.
+If the amount of resources that is available to distribute is less than the actual need, some amount of unmet demand is unavoidable.
+The allocation score for a forecast does not include the unmet demand that was unavoidable given the resource constraint, and so it measures only the amount of unmet demand that could have been prevented by using a different allocation of available resources than that suggested by the forecast.
 We elaborate on these ideas in Section~\ref{sec:methods}.
@@
 
-\elr{I have abandoned the attempt to get a succinct one-sentence definition of the allocation score at the beginning here, because it felt like we were trying to pack too many ideas into one sentence.
-But I'm open to revisiting.
-Maybe the definition of alloscore can be moved earlier?}
-
 Suppose that a decision-maker is tasked with determining how much of a resource such as a medical supply should be allocated to each of $n$ locations, subject to the constraint that only $K$ units of the resource are available in total.
-If the decision-maker is provided with a multivariate forecast $F$ that predicts resource demand in each of the $n$ locations, one possibility is that they may choose the resource allocation that minimizes the expected total unmet demand according to the forecast.
+If the decision-maker is provided with a multivariate forecast $F$ that predicts resource demand in each location, one option is to choose the resource allocation that minimizes the expected total unmet demand according to the forecast.
 We will give a more precise mathematical statement in section \ref{sec:methods.allocation}, but at a high level, the total expected unmet demand according to the forecast is
@@
 where $F_i$ is the marginal forecast distribution of demand in location $i$.
-This stochastic allocation problem has an intuitively appealing solution: allocate so that the probabilities of need exceeding allocation in various locations are as close to each other as possible (see methods).
+This allocation problem has an intuitively appealing solution: allocate so that the probabilities of need exceeding allocation in various locations are as close to each other as possible.
 This will lead to the allocations provided by $F$ being quantiles of the marginal distributions $F_i$ for some \emph{single} probability level $\tau$ that is shared in common for all locations.
 
-After time passes, the quality of a selected allocation can be measured by comparing the actual observed demand in each location to the amount of resources that were sent there.
+After time passes and the actual level of resource need has been observed, the quality of a selected allocation can be measured by comparing the actual demand in each location to the amount of resources that were sent there.
 Specifically, we compute the total unmet demand that resulted from the selected allocation:
 \begin{align}
-    \sum_{i=1}^n \text{unmet demand in location $i$ based on the selected allocation},
+    \sum_{i=1}^n \text{unmet demand in location $i$ based on the selected allocation}.
 \end{align}
@@
 
-The \textbf{allocation score} of the forecast $F$ is the unnecessary unmet demand that results from using the allocation that minimizes the expected unmet need according to that forecast.
-By ``unnecessary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
+The \textbf{allocation score} of the forecast $F$ is the avoidable unmet demand that results from using the allocation that minimizes the expected unmet need according to that forecast.
+By ``avoidable unmet demand'', we mean that the allocation score does not include the amount of unmet demand that was inevitable simply because the amount of available resources $K$ was less than the demand for resources.
+Thus, the allocation score measures the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
+An allocation score of 0 is optimal, and indicates that no other allocation of resources could have met demand better than the allocation suggested by $F$.
+A larger allocation score indicates that it would have been possible to improve upon the allocation suggested by $F$.
 
-\paragraph{Example 1} Suppose we have a forecast $F$ for demand in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$ and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, with $\sigma_1 = 2$ and $\sigma_2 = 3$.
-If we have $K = 5$ units of our resource available, the optimal allocation according to $F$ would be 2 units in location 1 and 3 units in location 2.
-If, on the other hand, we have $K = 10$ units available, we will allocate 4 units of resources in location 1 and 6 units in location 2.
-Figure~\ref{fig:exp_alloc_example} illustrates the situation, and derivations of the optimal allocations are given in section *** of the supplemental materials.
+\paragraph{Example 1} Suppose we have a forecast $F$ for demand in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$ and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$.
+When the marginal forecasts are exponential distributions, it can be shown that the optimal allocation divides the available resources among the locations proportionally to the scale parameters $\sigma_i$ (see section *** of the supplemental materials).
+If we have $K = 5$ units of our resource available, the optimal allocation according to $F$ would be 1 unit of resources in location 1 and 4 units of resources in location 2.
+If, on the other hand, we have $K = 10$ units available, we will allocate 2 units of resources to location 1 and 8 units to location 2.
+Figure~\ref{fig:exp_alloc_example} illustrates the situation.
 
@@
     \caption{An illustration of the resource allocation problem in Example 1.
-There are $n = 2$ locations, with predictive distributions $F_1 = \mathrm{Exp}(1/2)$ and $F_2 = \mathrm{Exp}(1/3)$, illustrated in the panels at bottom and right.
+There are $n = 2$ locations, with predictive distributions $F_1 = \mathrm{Exp}(1)$ and $F_2 = \mathrm{Exp}(1/4)$.
+The cdfs of these distributions are illustrated in the panels at bottom and right.
 In the center panel, the background shading corresponds to the expected loss according to these forecasts.
-Diagonal black lines indicate resource constraints at $K=5$ and $K=10$ units.
-For these forecasts, the optimal allocations are $(2,3)$ for $K=5$ and $(4,6)$ at $K=10$.
+Diagonal black lines indicate resource constraints at $K=5$ and $K=10$ units; any point along those lines corresponds to an allocation that meets the resource constraint.
+For these forecasts, the optimal allocations are $(1, 4)$ for $K=5$ and $(2, 8)$ at $K=10$.
 These allocations are at the point on the constraint line where the expected loss is smallest, which also corresponds to the point where a level set of the expected loss surface (blue curve) is tangent to the constraint.}
@@
 
-Next suppose that we observe needs of 2 and 9 in locations 1 and 2, respectively. With $K = 10$ units of the resource, the oracle would be able to prevent all
-but 1 of the total 11 units of needs from going unmet, for example by allocating 2 units of resources to location 1 and the remaining 8 units of resources to location 2.
-However, the forecast $F$'s allocation results in $9 - 6 = 3$ units of unmet need in location 2 despite leaving no need unmet in location 1.
-The allocation score for the forecast when $K = 10$ would therefore be 2 (= 3 realized $-$ 1 unavoidable) in units of unnecessary unmet need.
-When $K = 5$, however, the forecast does not ``over-allocate'' in either location which leads to the same unmet need of $11 - 5 = 8$ as does the oracle's allocations.
-The allocation score for $F$ when $K = 5$ is therefore 0.
+Next suppose that we observe needs of 1 and 10 in locations 1 and 2, respectively.
+Based on these observed needs, we can measure the quality of the allocation suggested by the forecast by calculating the amount of unmet need that resulted from that allocation over and above what was unavoidable given the resource constraint.
+With $K = 5$ units of the resource, the allocation based on the forecast exactly meets the observed demand in location 1, but it leaves 6 units of need unmet in location 2.
+However, working within the resource constraint, no other allocation could have done better: for example, allocating 0 units of resources to location 1 and 5 to location 2 still results in a total unmet need of 6 across both locations.
+Therefore, the forecast's allocation score is 0 with $K = 5$.
+On the other hand, when $K = 10$, the forecast $F$'s allocation results in $10 - 8 = 2$ units of unmet need in location 2 despite leaving no need unmet in location 1.
+In this case, the oracle would be able to prevent all but 1 of the total 11 units of need from going unmet, for example by allocating 1 unit of resources to location 1 and the remaining 9 units of resources to location 2.
+The allocation score for the forecast when $K = 10$ would therefore be 1 (= 2 realized $-$ 1 unavoidable) in units of avoidable unmet need.
 
-Note the way in which $F$ incurred a positive (i.e., non-optimal) allocation score of 2 when $K = 10$. It was not directly directly due to individual misalignments 
-of the marginal forecasts $F_i$ with the observed needs, but rather because the allocations and observed needs were not proportional as vectors.
-Restating: as far as allocation decisions are concerned, with a fixed constraint $K=10$ the fundamental problem with the forecast $F$ is not that it predicts a mean total resource demand of $5$ units; it is that the realized demand was 4.5 times as large in location 2 as in location 1, but the forecast only indicated that the resource allocation for location 2 should be 1.5 times the allocation to location 1.
+These scores illustrate a general result: allocation scores for a forecast will tend to be larger when the resource constraint is close to the observed demand, because this is when it matters most which locations are allocated more or less resources.
+If the resource constraint is very small, any allocation of those limited resources will result in a large amount of unmet need.
+If the resource constraint is very large, it becomes less important which locations receive relatively more or less resources because all locations will receive enough resources to meet their demand.
+
+\paragraph{Example 2} Now consider a different forecast that also has exponential distributions for resource demand in each location, but that has the scale parameters $\sigma_1 = 2$ and $\sigma_2 = 8$, twice as large as the scale parameters of the forecast in Example 1.
+Because the optimal allocation is proportional to the scale parameters, this forecast would lead to the same optimal allocations as the forecast in Example 1, and would therefore be assigned the same allocation score.
+
+Note the way in which these forecasts incurred a positive (i.e., non-optimal) allocation score of 1 when $K = 10$.
+It was not directly directly due to individual misalignments of the marginal forecasts $F_i$ with the observed needs, but rather because the allocations and observed needs were not proportional as vectors.
+Restating: as far as allocation decisions are concerned, with a fixed constraint $K=10$ the fundamental problem with the forecast $F$ in Example 1 is not that it predicts a mean total resource demand of $5$ units; it is that the realized demand was 10 times as large in location 2 as in location 1, but the forecast only indicated that the resource allocation for location 2 should be 4 times the allocation for location 1.
 
@@
 %For a given probability level $\tau$, if the quantiles $Q_i(\tau)$ of the marginal forecasts are proportional to the observed needs $y_i$, then the allocation score is zero for the resource constraint level $K = \sum Q_i(\tau)$.
-This stands in marked constrast to other common scoring methods for multi-variate forecasts that aggegate univariate scores such as CRPS or WIS for the marginal forecasts where a 
+This stands in marked constrast to other common scoring methods for multivariate forecasts that aggegate univariate scores such as CRPS or WIS for the marginal forecasts where a 
 misalignment in one coordinate is penalized regardless of alignments in other coordinates.
-\apg{Should this be illustrated here, or elsewhere?} \elr{Not sure what I think.
-Comparisons with WIS/etc seem critical, but some of the text I've added here feels like text for the discussion rather than up here in the ``methods"...}
-Note that we do not claim the allocation score is generically preferable to these other scores \textemdash rather, it provides a view of forecast performance that is specifically tuned to the context of decision-making about resource allocations.
+Note that we do not claim that the allocation score is generically preferable to these other scores \textemdash rather, it provides a view of forecast performance that is specifically tuned to the context of decision-making about resource allocations.
 

### 2023-08-09 alloscore_manuscript/allocation.tex

@@
 %For a given probability level $\tau$, if the quantiles $Q_i(\tau)$ of the marginal forecasts are proportional to the observed needs $y_i$, then the allocation score is zero for the resource constraint level $K = \sum Q_i(\tau)$.
-This stands in marked constrast to other common scoring methods for multivariate forecasts that aggegate univariate scores such as CRPS or WIS for the marginal forecasts where a 
+This stands in marked contrast to other common scoring methods for multivariate forecasts that aggegate univariate scores such as CRPS or WIS for the marginal forecasts where a 
 misalignment in one coordinate is penalized regardless of alignments in other coordinates.

### 2023-08-10 alloscore_manuscript/allocation.tex
attempt to reorganize section 2.3 to present quantile score and allocation score simultaneously

@@
 \end{enumerate}
-We use the letter $s$ for the loss function to align with the literature on evaluation of forecasts of continuous outcomes, in which context we can often identify the action $x$ with a functional (i.e., a numeric summary such as a mean or a quantile) of the forecast distribution $F$.
+%We use the letter $s$ for the loss function to align with the literature on evaluation of forecasts of continuous outcomes, in which context we can often identify the action $x$ with a functional (i.e., a numeric summary such as a mean or a quantile) of the forecast distribution $F$.
 In this context, $s$ may be used as a \emph{scoring function}.
@@
 
-\subsection{A review of quantile score, CRPS, and the weighted interval score}
+\subsection{A decision theoretic derivation of quantile score and allocation score}
 \label{sec:methods.quantileloss}
 
-We review how the quantile score arises from a particular decision-making problem in section \ref{sec:methods.quantileloss.quantile}, and how CRPS can be obtained by integrating across values of the parameters of that decision-making problem, as well as the connection to WIS, in section \ref{sec:methods.quantileloss.crps}.
-These results have been thoroughly discussed in the literature [cite cite cite].
+In this section, we demonstrate that the quantile score and the allocation score can both be derived from the procedure outlined in section \ref{sec:methods.decisiontheory} based on two different, but closely related, decision-making problems.
+The results for quantile score are well-known in the literature (cite cite cite).
 
-\subsubsection{Decision-theoretic origins of the quantile score}
-\label{sec:methods.quantileloss.quantile}
+\paragraph{Problem definitions and loss functions.}
+In both decision-making problems, the decision-maker must determine the level $x_i$ of a resource, such as a medical supply, that will be available to people who may need it at some future time in locations indexed by $i$.
+We use the variable $y_i$ to denote the eventual realized need for the resource in that location, e.g.
+the number of patients in need of a hospital bed or the amount of medication that is needed.
+Each unit of unmet need incurs a loss $L$, so that if the selected resource level $x_i$ is less than the realized need $y_i$, a loss of $L(y_i - x_i)$ results.
+At the time that a decision must be made, the demand $y_i$ is not yet known.
+We therefore define the random variable $Y_i$ that represents the as-yet-unknown level of demand in location $i$.
+The forecast $F_i$ specifies a predictive distribution for $Y_i$.
+Here we identify $F_i$ with its cumulative distribution function (CDF), and $F_i^{-1}$ denotes the quantile function.
+We use the notation $x = (x_1, \ldots, x_n)$, $y = (y_1, \ldots, y_n)$, $Y = (Y_1, \ldots, Y_n)$, and $F = (F_1, \ldots, F_n)$ to collect resource levels, resource demand, and forecasts across all locations.
 
-Suppose that a decision-maker is tasked with determining the quantity $x$ of a protective measure to procure; for example, $x$ might represent the number of hospital beds or amount of medicine to purchase.
-Additionally, suppose that each unit of this good has cost $C$ so that the total cost of procurement is $Cx$.
-The variable $y$ denotes the eventual realized need for this resource, e.g.
-the number of patients in need of a hospital bed or the amount of medication that is needed.
-We assume that each unit of unmet need incurs a loss denoted by $L$, so that if the selected procurement level $x$ is less than the realized need $y$, a loss of $L(x-y)$ results.
-At the time that a decision-maker determines the amount $x$ to procure, the demand $y$ is not yet known.
-We therefore define the random variable $Y$ that represents the as-yet-unknown level of demand.
-The forecast $F$ specifies a predictive distribution for $Y$.
-Here we identify $F$ with its cumulative distribution function (CDF), and $F^{-1}$ denotes the quantile function.
-With this formalization of the decision-making task, we can proceed to develop a proper scoring rule using the procedure outlined in section \ref{sec:methods.decisiontheory}.
+We note that the problem formulation here assumes that the resource in question does not impact the amount of demand $y_i$ that will materialize, but rather it is a resource that satisfies that demand.
+In the context of infectious disease, this means that we do not consider resources such as vaccine doses or masks, which are intended to reduce the number of people who will become sick.
+Our set up addresses resources like hospital beds, oxygen supply, or ventilators, which are intended to meet the medical needs of patients who are sick.
 
-\paragraph{Step 1: specify a loss function.} Combining the cost of procuring goods at level $x$ with losses due to unmet need, we arrive at the overall loss function
+%In both decision-making problems, the decision-maker is required to select a course of action that is designed to meet the future demand for a resource, which is unknown at the time that a decision must be made.
+The primary difference between these problems is the type of actions available to the decision-maker and the challenge (cost or resource constraint).
+TODO: clarify goal, minimize total costs.
+
+The differences between the two decision-making problems primarily lie in what actions are available to the decision-maker and what trade-offs they must consider between different alternatives.
+In decision problem 1, which will lead to the quantile score, the challenge faced by the decision-maker is that each unit of the resource has cost $C$, so that the total cost of procurement for location $i$ is $C x_i$.
+Summing across locations and combining the costs of procurement with the losses associated with procuring fewer resources than are needed, we arrive at the following loss function (subscripted by $Q$ for \emph{q}uantile):
+\begin{equation}
+s_Q(x, y; C, L) = \sum_{i=1}^n C x_i + L(y_i - x_i)_+. \label{eqn:quantile_loss_fn}
+\end{equation}
+The decision-maker's goal is to minimize this loss, but there is no hard limit on the available resources or the amount of funds that can be spent to purchase resources.
+It follows that although we have written a single loss function that sums across all locations, this loss function is separable and in practice decisions about the quantity of resources to procure can be made separately for each location.
+
+In decision problem 2, which will lead to the allocation score, a fixed quantity $K$ of the resource is available, and the problem is how best to allocate those resources to different locations.
+Thus, unlike in problem 1, here we must consider all locations simultaneously and any valid decision is required to satisfy the constraint $\sum_i x_i \leq K$.
+On the other hand, since the total amount of available resources is fixed, it is not relevant to consider the cost of procuring resources.
+When the resource constraint is satisfied, the overall loss function (subscripted by $A$ for \emph{a}llocation) is:
+\begin{equation}
+s_A(\mathbf{x}, \mathbf{y}; L) = \sum_{i=1}^n L (y_i - x_i)_+. \label{eqn:alloc_loss_fn}
+\end{equation}
+
+\paragraph{Bayes acts.}
+Now that we have formalized the decision-making tasks with loss functions that measure the quality of a selected action $x$ in relation to realized demand levels $y$, we can identify the Bayes acts corresponding to these loss functions.
+It can be shown that under each of these loss functions, the Bayes act is a quantile of the forecast distribution in each location (see supplemental section *** for the derivations):
 \begin{align}
-s_Q(x,y; C, L) = Cx + L(x-y)_-.
+    x^{F}_i = F_i^{-1}(\alpha).
 \end{align}    
-Here, $(x - y)_- := \max(-(x - y),0)$ is $0$ if the amount procured, $x$, is greater than or equal to the realized demand $y$; otherwise, it is $y - x$, the amount of unmet need.
+The probability level $\alpha$ for this quantile is the same across all locations, but it is different for the Bayes acts derived from the quantile loss and the allocation loss.
 
-\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} It can be shown that under the loss function $s_Q$, the Bayes act is a quantile of the forecast distribution at the probability level $\alpha = 1 - C/L$:
+Under the quantile loss, the Bayes act sets resource levels to be quantiles of the forecast distributions at the probability level $\alpha_Q = 1 - C/L$.
+\elr{Not confident in the following sentences' correctness, or in their value in this manuscript.
+We could just try to gloss over all of this or summarize with a less precise statement about the cost-loss ratio?
+$C/L > 1$ is also a possibility that it's awkward to address...}
+With this choice, the probability that demand exceeds the selected resource supply is equal to $C/L$.
+It does not make sense to purchase more resources than this because the cost $C$ of an additional unit of resources would be greater than $L$ times the forecasted probability of an additional unit of resource demand.
+In the edge case where the cost $C$ of purchasing resources is greater than the loss $L$ associated with under-purchasing resources, so that $\alpha_Q < 0$, the decision-maker will not purchase any resources.
+
+Under the allocation loss, the Bayes act chooses quantiles at a probability level that depends on the resource constraint $K$, the loss $L$, and the forecast distributions, and is selected so that the allocations do not exceed the resource constraint:
 \begin{align}
-x^F = F^{-1}(\alpha).
+\alpha_A &= 1 - \lambda^{\star}/L \text{, where $\lambda^\star$ is selected so that} \\
+\sum_{i=1}^{N}F_i^{-1}(1 - \lambda^{\star}/L) &= K.
 \end{align}
-See the supplement for a verification of this result.
+This partial solution to the allocation problem seems to have first appeared in \cite{hadleywhitin1963}.
+One interpretation of this result is that the Bayes act divides the available resources among the locations in such a way that the forecasted probability that resource demand will exceed the available resources is the same in all locations.
+
+%One interpretation of this result is that the Bayes act sets the allocation in each location $i$ to a quantile of the forecast distribution $F_i$ for that location.
+The quantile is at a probability level $\alpha = 1 - \lambda^{\star}/L$ that is the same for all locations, and is chosen such that the constraint is satisfied.
+An alternative interpretation comes from noting that for each location $i$, $\frac{\partial}{\partial x_i} \sbar_A^F(\mathbf{x}) \vert_{\mathbf{x}=\mathbf{x}^F} = \lambda^{\star}$ (see the supplement for a proof).
+%$\frac{\partial}{\partial x_i} \sbar_F(\mathbf{x}) = L[F_i(x_i) - 1]$ (see the supplement for a derivation), and evaluating at the Bayes allocation we arrive at
+%\begin{align}
+%\frac{\partial}{\partial x_i} \sbar_F(\mathbf{x}) \vert_{\mathbf{x}=\mathbf{x}^F} &= L[F_i(x^F_i) - 1] = L[F_i(F_i^{-1}(1 - \lambda^{\star}/L)) - 1] = \lambda^{\star}.
+%\end{align}
+%In words, at the allocation given by the Bayes act, the rate of change of the expected score as a function of the amount allocated to location $i$ is given by $\lambda^{\star}$.
+%This derivative is the same for all locations, so the optimal allocation divides the available resources across all locations in such a way that according to $F$, the expected benefit of 1 additional unit of resources is the same in all locations.
+
+\apg{
+first attempt by APG to rephrase last paragraph: The central mathematical ideas in this construction are that
+\begin{itemize}
+\item optimization under a \emph{single} constraint requires the rates of change of (a probabilistic forecast's) expected benefit with respect to our decision variables, $x_i$, to be the same for all locations
+\item these rates of change can be identified with probabilities
+\item and therefore the Bayes act results from using a \emph{shared} probability level, $1 - \lambda^{\star}/L$, to determine allocations as the corresponding quantiles of the location-specific forecast distributions $F_i$ which satisfy the constraint.
+\end{itemize}
+(See the supplement for detailed discussion and derivations of these points.)
+}
+
 %In particular, faced with the classical binary decision problem of whether to recommend an additional unit of protection given a current level of protection $x$, the forecaster's optimal decision rule under $s_m$ is to recommend adding protection if $x < F^{-1}\left(1-\frac{C}{L}\right)$, that is, if
@@
 
-\paragraph{Step 3: Define the scoring rule.} Following the procedures outlined above, we could score the forecast distribution $F$ with the scoring rule
+\paragraph{Scoring rules.} We can now define scoring rules for a forecast $F$ by evaluating the loss function at the Bayes act corresponding to that loss function.
+This leads to the following scoring rule for quantile forecasts:
 \begin{align*}
-S_Q(F, y; C, L) &= s_Q(x^F, y; C, L) = Cx^F + L(x^F-y)_- \nonumber \\
-&= C F^{-1}(\alpha) + L \left(F^{-1}(\alpha) - y\right)_- \label{eqn:S_Q_CL}
+S_Q(F, y; C, L) &= \sum_{i=1}^n s_Q(x^F, y; C, L) = \sum_{i=1}^n Cx_i^F + L(y_i - x_i^F)_+ \nonumber \\
+&= C F_i^{-1}(\alpha_Q) + L \left(y_i - F_i^{-1}(\alpha_Q)\right)_+ \label{eqn:S_Q_CL}
 \end{align*}
+This score measures the total cost of resource purchases and losses incurred due to unmet need that results from using the Bayes resource procurements associated with the forecast $F$ when the actual level of need in each location is observed to be $y_i$.
 
-We have set up the problem here in terms of the cost and loss parameters $C$ and $L$, which has the benefit of an intuitive connection to the decision-making context.
-However, to clarify the connection to the usual notation for the quantile loss, we can divide the loss function $s_Q$ by $L$ to obtain an expression in terms of only $\alpha$:
-\begin{align*}
-s_Q(x,y; \alpha) &= s_Q(x,y; C, L)/L \\
-&= (C/L)x + (x-y)_- \\
-&= (1 - \alpha)x + (x-y)_-.
-\end{align*}
-Because these loss functions are equal up to a constant of proportionality, the Bayes act is the same for both. The associated quantile scoring rule expressed in terms of $\alpha$ is
+On the other hand, the allocation scoring rule is:
+\begin{equation}
+S_A(F, y; L, K) = s_A(\mathbf{x}^F, y; L) = \sum_{i=1}^n L (y_i - F_i^{-1}(1 - \lambda^{\star}/L))_+
+\end{equation}
+This score measures the total unmet need across all locations that results from using the Bayes allocation associated with the forecast $F$ when the observed need is $y_i$.
+
+We have defined the quantile scoring rule $S_Q$ in terms of the cost and loss parameters to connect to the decision-making context. 
+To connect to the usual notation for the quantile score, we can divide through by $L$ to obtain an expression in terms of only $\alpha$:
+% \begin{align*}
+% s_Q(x,y; \alpha) &= s_Q(x,y; C, L)/L \\
+% &= (C/L)x + (x-y)_- \\
+% &= (1 - \alpha)x + (x-y)_-.
+% \end{align*}
+% Because these loss functions are equal up to a constant of proportionality, the Bayes act is the same for both. The associated quantile scoring rule expressed in terms of $\alpha$ is
 \begin{equation}
@@
 \end{equation}
-In either formulation, the key observation is that the Bayes act is the quantile of the forecast distribution $F$ at the probability level given by one minus the cost/loss ratio $C/L$.
+These two formulations are equivalent.
+%In either formulation, the key observation is that the Bayes act is the quantile of the forecast distribution $F$ at the probability level given by one minus the cost/loss ratio $C/L$.
 
-\subsubsection{CRPS as an integrated quantile score}
+\elr{It's a little tempting to just divide by the loss parameter $L$ from the start and simplify throughout?}
+
+\subsubsection{Integrated scores}
 \label{sec:methods.quantileloss.crps}
+
+\elr{Not updated yet, but in this subsection the plan would be to talk about CRPS as integrated quantile score, integrated allocation score, and connections between them.}
 
@@
 
-\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act is the allocation that minimizes the expected loss:
-\begin{align}
-    \mathbf{x}^F &= \underset{\mathbf{x}\in \mathbb{R}^N, 0 \leq \mathbf{x}}{\mathrm{arg min}}\,\, \sbar_A^F(\mathbf{x}; L) \text{ subject to }
-    \sum_{i=1}^N x_i \leq K \text{, where} \label{AP} \\
-    \sbar_A^F(\mathbf{x}; L) &= \mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y}; L)] = \sum_{i=1}^{N} \mathbb{E}_{F_i}[s_A(x_i, Y_i; L)] \label{eqn:sbar}
-\end{align}
-It can be shown that with the loss function given in Equation \eqref{eqn:loss_fn}, the Bayes act has $x^F_i = F_i^{-1}(1 - \lambda^{\star}/L)$, where $\lambda^{\star}$ depends on the problem parameters $K$ and $L$ as well as the forecast distributions and is chosen so as to satisfy the equation
-\begin{align}
-\sum_{i=1}^{N}F_i^{-1}(1 - \lambda^{\star}/L) = K.
-\end{align}
-This partial solution to the allocation problem seems to have first appeared in \cite{hadleywhitin1963}; see the supplemental materials for a derivation.
-Figure *** panel (a) \elr{TODO: first figure from notes crps\_connection\_exp.Rmd} illustrates the expected loss function $\sbar_A^F$ and the allocation given by the Bayes act in a simple example with $n = 2$ locations, $L = 1$, $K = 5$, and forecasts $Y_1 \sim \text{Exp}(1 / \sigma_1)$ and $Y_2 \sim \text{Exp}(1 / \sigma_2)$, using a scale parameterization of the exponential distribution with $\sigma_1 = 1$ and $\sigma_2 = 5$.
-
-One interpretation of this result is that the Bayes act sets the allocation in each location $i$ to a quantile of the forecast distribution $F_i$ for that location.
-The quantile is at a probability level $\alpha = 1 - \lambda^{\star}/L$ that is the same for all locations, and is chosen such that the constraint is satisfied.
-An alternative interpretation comes from noting that for each location $i$, $\frac{\partial}{\partial x_i} \sbar_A^F(\mathbf{x}) \vert_{\mathbf{x}=\mathbf{x}^F} = \lambda^{\star}$ (see the supplement for a proof).
-%$\frac{\partial}{\partial x_i} \sbar_F(\mathbf{x}) = L[F_i(x_i) - 1]$ (see the supplement for a derivation), and evaluating at the Bayes allocation we arrive at
-%\begin{align}
-%\frac{\partial}{\partial x_i} \sbar_F(\mathbf{x}) \vert_{\mathbf{x}=\mathbf{x}^F} &= L[F_i(x^F_i) - 1] = L[F_i(F_i^{-1}(1 - \lambda^{\star}/L)) - 1] = \lambda^{\star}.
-%\end{align}
-In words, at the allocation given by the Bayes act, the rate of change of the expected score as a function of the amount allocated to location $i$ is given by $\lambda^{\star}$.
-This derivative is the same for all locations, so the optimal allocation divides the available resources across all locations in such a way that according to $F$, the expected benefit of 1 additional unit of resources is the same in all locations.
-
-\apg{
-first attempt by APG to rephrase last paragraph: The central mathematical ideas in this construction are that
-\begin{itemize}
-\item optimization under a \emph{single} constraint requires the rates of change of (a probabilistic forecast's) expected benefit with respect to our decision variables, $x_i$, to be the same for all locations
-\item these rates of change can be identified with probabilities
-\item and therefore the Bayes act results from using a \emph{shared} probability level, $1 - \lambda^{\star}/L$, to determine allocations as the corresponding quantiles of the location-specific forecast distributions $F_i$ which satisfy the constraint.
-\end{itemize}
-(See the supplement for detailed discussion and derivations of these points.)
-}
-
-\paragraph{Step 3: Define the scoring rule.} We can now define a proper scoring rule for the probabilistic forecast $F$ as
-\begin{equation}
-S_A(F, y; L, K) = s_A(\mathbf{x}^F, y; L) = \sum_{i=1}^n L (F_i^{-1}(1 - \lambda^{\star}/L) - y_i)_-
-\end{equation}
-This score measures the total unmet need across all locations that results from using the Bayes allocation associated with the forecast $F$ when the actual level of need in each location is observed to be $y_i$.
+\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} 
+\paragraph{Step 3: Define the scoring rule.}
 

### 2023-08-18 alloscore_manuscript/methods_apg.tex
adding in new methods section as separate file

\documentclass{article}

\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

\usepackage{amsmath, amsfonts, amssymb}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{cases}
\usepackage{caption}
\usepackage{hyperref}

\DeclareMathOperator*{\argmin}{argmin}

\usepackage{setspace}
\onehalfspacing

\usepackage{soul}
\usepackage{xcolor}
\def\elr#1{{\color{cyan}\textbf{ELR:[#1]}}}
\def\apg#1{{\color{red}\textbf{APG:[#1]}}}
\def\bwr#1{{\color{violet}\textbf{BWR:[#1]}}}

\usepackage{natbib}
\bibliographystyle{unsrtnat}

\title{Allocation scores, WIS, and CRPS via decision theory}
\author{Aaron Gerding, Nicholas G. Reich, Benjamin Rogers, Evan L. Ray}

\begin{document}

\newcommand{\del}[2]{\frac{\partial {#1} }{\partial {#2}} }
\newcommand{\dby}[2]{\frac{d {#1} }{d {#2}} }
\newcommand{\sbar}{\overline{s}}
\newtheorem{proposition}{Proposition}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\maketitle

\begin{abstract}

A somewhat open-ended attempt to motivate and construct quantile scores, WIS, CRPS, and allocation scores within a unified decision theoretic framwork.

\end{abstract}

\section{Introduction}

In this section/report/paper/rambling manifesto, we present a methodological perspective on the use scoring rules for forecast evaluation that we see as well-adapted to our motivating research goal to understand and describe the social utility of infectious disease forecasts. Our approach follows the common practice in decision theory of giving central roles to the foundational concepts of state, action, and consequence. From these primitives, we first derive the familiar measures of quantile score (QS), continuous ranked probability score (CRPS) and their aggregates across forecast targets.  We then introduce new scoring rules, the \emph{specific} and \emph{integrated allocation scores}, as reparametrizations of QS and CRPS aggregates that result when part of our definition of consequence is shifted into our definition of action. Specifically, we will trade the regret a decision maker experiences as a consequence of seeing allocated resources go unused for the requirement that a decision maker choose an action that allocates no more than a certain amount in total. This exchange mirrors the duality in mathematical optimization theory between constraints on "decision variables" and the "Lagrange multiplier variables" which quantify the sensitivity of solutions to the constraints, that is, the regret (or relief) experienced when a change in the constraint leads to a change in the optimal cost.
Our computational methods (see ..) in fact make use of this connection.
Something about Bayes risks...

\section{Scoring forecasts via decision problems}

Broadly speaking, we score a probabilistic forecast $F$ by identifying it with a solution to a \emph{decision problem} made up of pre-specified actions, states and consequences and rating the utility of that solution.  That is, $F$ will receive a "better" or "worse" score in a particular instance when the consequence is better or worse of a rational decision maker using $F$ (and essentially only $F$) to select an action in the face of uncertainty about a future state.  It should be noted immediately that a better or worse score in a single instance may tell us almost nothing about the intrinsic value of the forecasting method that produced $F$.  But as scores accumulate, we can use their distributions across forecasting instances to make inferences as to whether one forecasting method is better or worse than another in a decision theoretic sense.

\subsection{The ingredients of a decision problem with examples}
The basic example of a future state we are considering is the number of individuals that become sick enough from an infectious disease to be in need of some potentially limited resource such as hospital beds, ventilators, medication, or medical staff.  We write $y$ to refer to a particular state such as the quantity of need in a single location or the quantities across multiple locations or times and $Y$ for a future uncertain state which is a random variable with a set of realizable values $\mathcal{Y}$. The forecast $F$ to be scored is a claim of how $Y$ is (or will be) distributed. We also will refer to states as outcomes, which helps to emphasize their future realization.  But we want to clearly distinguish an outcome from a consequence, which as we discuss below, is the result of taking an action under a certain outcome.

The basic example of an action we consider is the specific amount or set of amounts of such a resource that a public health worker decides to make available in either a single location or across multiple locations or times.  We write an action as $x$ and assume that it takes values in a set of possible actions $\mathcal{X}$ that is possibly defined by some constraint such as the total amount of resource available to the public health worker for allocation.  Note that $x$ must be selected prior to the realization of $y$.  A decision maker might use a set rule to select $x$ based on inputs, such as $F$, but the observed value of $y$ cannot be such an input.\footnote{In a dynamical decision theoretic approach, the action $x$ could be a time series depending at each step on previous values of $y$.  This could be a very fruitful framework for infectious disease forecast evaluation but we do not pursue it is this work.}

And when actions are assigned to states, there are consequences, which we abstractly identify with an element $c$ of a set $\mathcal{C}$. We have two basic examples in mind of what such a $c$ might be.   First, we consider some amount of the regret mentioned above that resources made available in a place or time are not actually needed at that place or time. We emphasize that the nature of this regret is potentially quite amorphous.  In business contexts (e.g., the "newsvendor problem" from inventory management) this regret is often easily quantifiable as monetary waste, but in a public health context, it could... (A desire to better understand over-allocation regret and its relevance to infectious disease forecast evaluation was actually a primary impetus for the present work.) 
Our second and more concrete example of a consequence is the loss incurred when a sick individual is not able to access a resource because it was not made available to them where and when they needed it.   We emphasize here that while we formally identify such loss as a consequence of a particular action being taken under a particular state, it may or may not be an avoidable loss.  That is, it could just as well result from a factor outside a decision maker's control such as a severe resource constraint as from a choice to withhold a resource, motivated, for example, by a desire to reduce waste.

\subsection{The formal procedure}

With a view towards a decision problem defined by these examples of sets $\mathcal{X}$, $\mathcal{Y}$, and $\mathcal{C}$ of actions, outcomes, and consequences, we now give a formal three-step procedure for defining and evaluating a \emph{scoring rule} for a forecast $F$ for a future value $y \in \mathcal{Y}$ of $Y$.

\begin{enumerate}
\item Specify a \emph{loss function} $s(x,y)$ that quantifies the utility or disutility of the consequence $c$ of taking action $x$ under outcome $y$.
\item Map the given probabilistic forecast $F$ for $Y$ to an action $x^F$ which ``solves" the decision problem by minimizing the expected loss $E_{F}[s(x,Y)]$ over all $x \in \mathcal{X}$ under the distribution $F$.  Following \cite{dawid2007geometry}, we call $x^F$ the \emph{Bayes act} for $F$. Note that $x^F$ depends not only on $F$ and $s$, but also on the definition of $\mathcal{X}$. For example, having more available resources usually allows for actions incurring less loss.   
\item Assign $F$ the score $s(x^F,y)$, that is, the numerical quantification via $s$ of the consequence under outcome $y$ of accepting $x^F$ as a solution to the decision problem.
\end{enumerate}{}

Thus, given definitions of $\mathcal{X}$, $\mathcal{Y}$, $\mathcal{C}$, and a loss function $s$, we produce a scoring rule $S(F,y):= s(x^F,y)$ that can be applied to any $F$ for any outcome $y \in \mathcal{Y}$.  We emphasize again a key feature of this formulation: the scoring rule $S$ is defined entirely in terms of what we take to be possible actions, states, and consequences along with a function $s$ that assigns values to the possible consequences.

Another key feature of this procedure is that the scoring rule $S$ it yields is proper by definition.  This is because a forecaster that believes that $Y$ has distribution $F$ and understands the decision problem and loss function $s$ must also believe that $x^F$ is the optimal action to take.  This, in turn, implies the belief that there is no forecast that will give a better expected score than $F$.  Therefore, there is no incentive to reveal a forecast at odds with a forecaster's beliefs about $Y$.

\begin{remark} For there to be an incentive \emph{not} to reveal a forecast other than $F$ we would need a \emph{strictly} proper scoring rule, which would require more conditions on the decision problem and/or $s$. In the literature, a functional $T(F)$ (such as the mean or median of $F$) is called \emph{elicitable} if there is a strictly proper scoring rule $S(F,y) = s(x^F,y)$ for which $T(F)=x^F$.  But the focus in work on elicitability often seems to be the functional $T$ as a starting point rather than the decision problem, which is our primary concern here.
\end{remark}

The following sections explain in detail how the QS, WIS, CRPS (for various parameters and weightings) as well as extensions to higher dimnsional setting can be seen as outputs of our procedure.

\section{Step 1: Defining decision problems and loss functions.}

We begin by imagining a public health worker trying to decide at time $t_0$ what quantity $x$ to procure of a protective resource such as beds, staff, or medicine in anticipation of $y$ new infections that will be observed (and create resource demand) at time $t_1>t_0$. The resource has a unit-cost of $C$, and we assign a loss of $L$ to the consequence of a sick individual not having access to the resource. This specifies a loss function  
\begin{align}
s_Q(x,y) = Cx + L(y-x)_+ \label{eqn:quantile_loss}
\end{align}
where $u_+ = \max(u,0)$.  We use the subscript $Q$ because it will be shown below that this loss function yields the quantile score for the the probability level $\alpha = 1 - C/L$.  Loss functions of this general type are referred to as \emph{piece-wise linear}, distinguishing them from smooth non-linear loss functions such as squared error.
The underlying decision problem here consists of action and outcome sets $\mathcal{X},\mathcal{Y}$ both equal to $\mathbb{R}_+$, the non-negative reals.  (We ignore for now any discreteness of the resource.) The consequence set $\mathcal{C}$ is abstractly a set of over-expenditure regrets and numbers of sick individuals with unmet resource need, but the loss function $s$ maps $\mathcal{C}$ into $\mathbb{R}_+$. We also from now on make the assumption that $C < L$ since otherwise -- at least according to this loss function -- there is no ``cost-effectiveness'' in trying to prevent even a single case of unmet need. This would imply a Bayes act $x^F$ that is identically 0, telling us nothing about the value of $F$.

This decision problem may however be complicated by uncertainty at time $t_0$ about the unit-cost of the resource. We take such uncertainty as requiring the worker to commit in advance to procurement levels $x_i$ for a range of potential costs $C_i < L$ where $i \in 1, \ldots, m$.  Even though only one of the $x_i$ will end up needing to be procured, there does not seem to be any intrinsically decision theoretic principle by which our loss function should ignore counter-factual but exactly specified consequences of the decision under unrealized costs. If we wanted to commit to a \emph{locality} principle for prediction where losses and scores only depend on predictive densities in a neighborhood of the relevant observation, this attempt would falter, and we would need to frame our methods around the log score and it's higher order relatives. But we do not see that as an appropriate principle in public health resourse allocation where  
probabilistic forecasts can be valuable by pushing decision makers in the "right direction" even when they assign low probability to observed outcomes. We instead adopt a ``sensitivity-to-distance'' principle from which piece-wise linear loss functions and the QS can be shown to naturally emerge. It is this principle that makes the loss under alternative costs seem relevant. \apg{These are obviously incomplete thoughts but I feel like there is something here that I've been grasping for for a lng time.}

A simple way to take account of the consequences of actions under counterfactual costs is to use the mean of all losses under certainty about the cost:
\begin{align}
s_{\mathrm{WIS}}(x,y) = \frac{1}{m}\sum_{i=1}^{m}C_ix_i + L(x_i - y)_+ \quad (C_i < L \text{ for all } i)
\end{align}
where we here use $x$ to express the decision vector $(x_1,\ldots, x_m)$. Or if there are prior beliefs about the probabilities $p_i$ of the $C_i$, we could use a weighted sum
\begin{align}
s_{w\mathrm{WIS}}(x,y) = \frac{1}{m}\sum_{i=1}^{m}p_i(C_ix_i + L(x_i - y)_+). \label{eqn:sum_CL_loss}
\end{align}
As for $Q$, the $WIS$ and $wWIS$ subscripts indicate that, for particular values of the $\{C_i\}$ and $L$, this loss function produces the WIS or a version that weights intervals (i.e. the average of QS's for $C_i$ and $L-C_i$) with non-standard weights (combining $p_i$ and $C_i/L$).

We can also consider $C$ a continuous random variable and accommodate this by letting $m \to \infty$ in \eqref{eqn:sum_CL_loss} which (with some mathematical care) converts the loss into an integral
\begin{align}
s_{\mathrm{CRPS}}(x,y) = \frac{1}{L}\int_{0}^{L} (cx(c) + L(x(c) - y)_+) dc
\end{align}
over all relevant costs $c \in [0,L]$, or a weighted version
\begin{align}
s_{w\mathrm{CRPS}}(x,y) = \int_{0}^{L} (cx(c) + L(x(c) - y)_+) f_C(c)dc{}
\end{align}
that uses a prior density $f_C$ on the possible costs. These loss functions will yield the traditional CRPS and the wieghted version of \cite{gneiting2011weightedScoringRules}.

It is important to note that these generalizations change not only the loss function but also the structure of the decision problem by using higher dimensional action sets $\mathcal{X}= \mathbb{R}_+^m$ or the set $\mathcal{X} = \{x(c):[0,\infty) \to \mathbb{R}_+\}$ of $\mathbb{R}_+$ valued functions  on $[0,\infty)$.

We can also expand our decision problem to include multiple actions that will all, in fact (not counterfactually), have to be taken.  This includes the motivating scenario for our work, the simultaneous procurement of resources across several locations (U.S. states and territories) during a pandemic. And again, summing the losses  \eqref{eqn:quantile_loss} over the locations (for a particular $C$ and $L$) creates a loss function that is reactive to the overall procurement decision. Let the locations be indexed by $l \in 1,\ldots,N$.  The action sets for the decision problem corresponding to a single certain cost $C$, a discrete cost distribution over $\{C_i\}, i = 1,\ldots,m$, and a continuous cost distribution $f_C$ with support on $[0,L]$ are now
\begin{align}
\mathcal{X}_{cert} &= \mathbb{R}_{+}^N \\
\mathcal{X}_{disc} &= (\mathbb{R}_{+}^m)^N\\
\mathcal{X}_{cont} &= \{x(c):[0,\infty) \to \mathbb{R}_{+}^N\}.
\end{align}
And for these "vector-action" problems we can define \emph{total} versions of our loss functions
\begin{align}
s_{t\mathrm{Q}}(x,y) &= \sum_{l=1}^N Cx_l + L(y_l-x_l)_+ \\
s_{t\mathrm{WIS}}(x,y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}C_i x_{l,i} + L(x_{l,i} - y_l)_+ \\
s_{tw\mathrm{WIS}}(x,y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}p_i(C_i x_{l,i} + L(x_{l,i} - y_l)_+)\\
s_{t\mathrm{CRPS}}(x,y) &= \frac{1}{L} \sum_{l=1}^N\int_{0}^{L} (cx_l(c) + L(x_l(c) - y_l)_+) dc \\
s_{tw\mathrm{CRPS}}(x,y) &= \sum_{l=1}^N \int_{0}^{L} (cx_l(c) + L(x_l(c) - y_l)_+) f_C(c)dc.
\end{align}
A key point for our presentation will be that the same $C$, set $\{C_i\}$, probabilities $\{p_i\}$, and density $f_C$ are used for all locations simultaneously.  We believe that our methods could be adapted to variations across locations (which could be important in practice), but we leave this generalization for later work.

The loss functions (and corresponding decision problems) defined so far  have all been thouroughly covered, from one perspective or another, in various strands of forecast evaluation literature.  We now introduce one that, while certainly familiar in operations research and related fields, has not, to our knowledge, been explicitly discussed in a forecast evaluation context.

Consider again the simultaneous resource procurement decision a public health worker must make in attempting to mitigate a pandemic across multiple locations. 
We now add a constraint $K$ on the total amount of the resource available for allocation across all locations. This presents the worker with a new decision problem we call the \emph{allocation problem} (AP or $\mathrm{AP}(K)$) with a smaller action set $\mathcal{X}_K = \{x \in \mathbb{R}_{+}^N | \sum x_l \leq K\} \subset \mathcal{X}_{cert}\}$ from which to choose (and analogously restricted version of the sets $\mathcal{X}_{disc}$ and $\mathcal{X}_{cont}$ for the decision problems with uncertain cost). Loss functions with the basic $Cx + L(y-x)_{+}$ structure we have used up until now continue to generate meaningful scoring rules with respect to the AP following our procedure.  But with a view toward developing a scoring rule that accommodates uncertainty about both $C$ and $K$ in a non-redundant (and more interpretable and computationally feasible) way, we augment the action set of the AP to include, in addition to the procurement levels $x_l$ for all locations, a ``reserve'' quantity $z_K$ defined as the available amount of resource that the worker decides not to allocate anywhere.
In optimization theory quantities defined in this manner are known as \emph{slack} variables, and are introduced in order to re-express contraint inequalities as equalities.  In our case, we get
\begin{align}
\mathcal{X}_{K} = \{(x,z) \in \mathbb{R}_{+}^{N+1} \mid \sum_{l=1}^{N}x_l + z = K \}.
\end{align}
In these action set variables, the loss function we used for the unconstrained $N$ location decision problem with action set $\mathcal{X}_{cert}$ now has the expression
\begin{align}
s_{tQ}\left((x,z), y\right) = C(K-z) + \sum_{l=1}^{N} L(y_l - x_l)_{+}.
\end{align}
This isolates -- at least formally -- the explicit contribution of the procurement part, $x$, of the decision to the second term which we call the \emph{allocation loss function}:
\begin{align}
s_{A,K}(x,y):= L\sum_{l=1}^{N} (y_l - x_l)_{+}.
\end{align}
It will turn out, as explained in the following sections, that for a given forecast $F$, a scoring rule derived with respect to the AP from $s_{tQ}$ under uncertainty about $C$ -- that is, from $s_{tWIS}$ or $s_{tCRPS}$ -- will have the same values for all outcomes $y$ as a scoring rule that is a weighted average or integration of scoring rules derived from allocation losses $s_{A,K}$ with respect the AP's for a range of $K$'s.

\section{Step 2: Derive the Bayes act for a forecast $F$}

I think we should not hide the formula 
\begin{align}
\dby{}{x_{il}} E_F[s_{.}(x,y)] &= E_{F}[\dby{}{x_{il}} s_{.}(x,y)] = C - LE_{F_l}[\mathbf{1}(y_l > x_{il})] = C + L(F_l(x_{il}) - 1) \\
\implies x_{il}^{F_l} &= F_l^{-1}(1 - C/L) \\
&\text{ or for CRPS} \\
x_{l}^{F_l}(c) &= F_l^{-1}(1 - c/L)
\end{align}

Fundamental point: 
\begin{itemize}
\item When $z_K > 0$, $K$ has no influence on the $\mathrm{AP}(K)$ Bayes act, i.e., $\partial_K x^{F,C,K} = 0$.  
\item When $z_K=0$ and $x^F$ are not $1-C/L$ quantiles, $\partial_C x^{F,C,K} = 0$.
\end{itemize}

Picture: For $K$ fixed, $x^{F,K}(c)$ will be a constant point $x^{F,K}$ on $\{\sum x_l = K\}$ for $c \in [0, c_1]$, 
and then wander back to 0 for $c \in (c_1,L]$. When $C$ is fixed, $x^{F}(k)$ will be a point on $\{\sum x_l = k\}$ (which does not depend on $C$)
for $k \in [0, \sum F_l^{-1}(1-C/L)]$, and then constant at $\mathbf{F}^{-1}(1-C/L)$ for $k \in (\sum F_l^{-1}(1-C/L), \infty)$.

\section{Step 3: Use the Bayes act to score the forecast $F$ against outcomes.}

Deal with alloscore by noticing that simultaneous quantile loss has same minimizer as alloscore with sum of quantiles as constraint level.

\begin{align}
s_{t\mathrm{WIS}}(x^{F,t\mathrm{WIS}},y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}s_{Q}(x_{l,i}^{F_l,\mathrm{Q}, C_i}, y_l)_+ \\ 
&= \frac{1}{m} \sum_{i=1}^{m}s_{\mathrm{AS}}(x^{F,\mathrm{AS},  K(C_i, F, L)},y) 
\end{align}

\begin{align}
s_{\mathrm{IAS}}(x^{F, \mathrm{IAS}}(k), y) &= \int_0^{\infty} s_{\mathrm{AS}}(x^{F, \mathrm{AS}(k)}, y) dk \\
&= \int_0^{\infty} \sum_{l=1}^N s_{\mathrm{Q}}(x^{F_l, \mathrm{Q}(c(k, F))}, y) dk \\
&= \sum_{l=1}^N \int_0^{\infty} s_{\mathrm{Q}}(x^{F_l, \mathrm{Q}(c(k, F))}, y) dk
\end{align}

But 
\begin{align}
k = k(c,F) &= \sum_{l=1}^{N} F_l^{-1}(1-c/L) \\
\del{k}{c}& = \sum_{l=1}^{N} \frac{-1}{L f_l(q_{F_l,\tau(c)})} := -\mathrm{MAI}(c,F)\label{eqn:dkdc}
\end{align}
(where we assume that the $F_l$ have non-negative support). The name MAI refers to this derivative being the marginal total allocation increase for the total quantile loss Bayes act $x^{F,Q(C)}$ that results from a reduction of the cost $C$ of the resource by 1. Or in terms of the probability level $\tau(c)=1-c/L$,
\begin{align}
\del{k}{\tau} = \del{k}{c}\dby{c}{\tau} = -\mathrm{MAI}(c,F) \cdot (-L\tau) = L\mathrm{MAI}(c,F) \label{eqn:dkdtau}
\end{align}

Now a constraint $K = 0$ corresponds to cost-loss parity $C=L$, i.e. $\tau = 0$, for which the Bayes act is the zero vector.  And $K \to \infty$ corresponds to the cost of over-allocation vanishing, i.e. $\tau \to 1$. So making a change of variables with \eqref{eqn:dkdc} or \eqref{eqn:dkdtau} we get
\begin{align}
s_{\mathrm{IAS}}(x^{F, \mathrm{IAS}}(k), y) 
&= \sum_{l=1}^N \int_0^{L} s_{\mathrm{Q}}(x^{F_l, \mathrm{Q}(c)}, y) \mathrm{MAI}(c,F) dc \\
&= \sum_{l=1}^N \int_0^{1} s_{\mathrm{Q}}(x^{F_l, \tau}, y) L\mathrm{MAI}(L(1-\tau),F) d\tau \\
&= tw\mathrm{CRPS}(F,y)
\end{align}
where MAI provides the (identical) weighting in each coordinate. Crucially though, this weighting \emph{depends} on $F$ so that IAS's of two forecasts $F$ and $\tilde{F}$ \emph{cannot} be interpreted as result of applying the same $w$CRPS to the two forecasts.

More about SR's:
\begin{itemize}
\item QS: meteorologist loss v newsvendor
\item WIS/CRPS: multiple scenario loss (Gruschka-Cockayne et al, Fissler and Ziegel, Jose and Winkler)
	- Least squares analogy
\item bring in Brier?
\item take stab at explaing log score following Dawid with Bayes act being inverse of $x^F(c)$ i.e., $F$ itself.
\end{itemize}

\subsection{Solving the Allocation Problem}

As the solution to a constrained optimization problem, the optimal allocation $x^F$ according to a forecast $F$ can be described formally as the first component of a stationary point $(x^F, z_K, \lambda^{\star})$ in the region $\{\mathbf{x} \geq 0, \lambda \geq 0\}$ for the Lagrangian 
\begin{align}
\mathcal{L}(\mathbf{x},\lambda; K, C, L) = L\sum_{l=1}^{N} E_{F_l}[(Y_l - x_l)_{+}] + C(K-z_K) + \lambda\left(z_K + \sum_{l=1}^{N} x_l - K\right).
\end{align}
That is, optimality at $x^F$ requires that
\begin{align}
0 = \frac{\partial}{\partial x_l} \mathcal{L}(x^F, z_K, \lambda) &= L E_F \left[\frac{d}{dx_l}(Y_l - x_l^F)_{+}\right] + \lambda \\
& = L E_F \left[-\mathbf{1}\{Y_i \geq x_i^F\}\right] + \lambda \\
& = L(F_i(x_l^F)-1) + \lambda \label{eqn:x_stationary}
\end{align}
for any $x_l^F > 0$;
\begin{align}
0 = \frac{\partial}{\partial z_K} \mathcal{L}(x^F, z_K, \lambda) &= \lambda - C
\end{align}
if $z_K>0$; and
\begin{align}
0 = \del{}{\lambda} \mathcal{L}(\mathbf{x}^F, \lambda^{\star}) &= \sum_{i=1}^{N} x_i^F - K\\
& = \sum_{\{i | x_i^F > 0\}} F_{i}^{-1}(1-\lambda^{\star}) - K \quad \text{ (by \eqref{eqn:x_stationary})}. \label{eqn:lambda_stationary}
\end{align}
These equations give formulae $x_i^F = F_i^{-1}(1-\lambda^{\star})$, but in many situations of interest for us the $F_i^{-1}$ are not single-valued, and even 
when they are, a solution $\lambda^{\star}$ of \eqref{eqn:lambda_stationary} can only be obtained analytically in special cases (but see ...). We therefore typically calculate allocation scores using approximately optimal allocations found via a binary search for $\lambda^{\star}$.  This search begins with the interval $\lambda \in [0,1]$ and 
checks at each step the sign of the inequality that results from substituting the interval midpoint $\lambda_{\tau}$ into \eqref{eqn:lambda_stationary} for endpoints of the sets $F_i^{-1}(1-\lambda_{\tau})$.  If $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) > K$, the quantiles for $1-\lambda_{\tau}$ exceed our resource constraint so we set $\lambda_{\tau}$ as the lower endpoint of the $\tau + 1$ search interval.  Similarly, if $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) < K$, the quantiles for 
$1-\lambda_{\tau}$ do not make full use of our resource and we set $\lambda_{\tau}$ as the upper endpoint of the $\tau + 1$ search interval.  This process halves the search interval width at each step and we continue until this width falls below a given tolerance level $\varepsilon_{\tau}$ yielding an 
approximation $\lambda^{\star, \varepsilon}$. When quantiles $F_i^{-1}(1-\lambda^{\star, \varepsilon})$ are singly defined and the forecast densities $f_i$
not too close to zero at these quantiles, the discrepancy $\Delta(\lambda^{\star, \varepsilon})  = \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star, \varepsilon}) - K$ is of order $\varepsilon_{\lambda}$ and we can take $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ as approximations to $x_i^F$.

Often though, $1-\lambda^{\star, \varepsilon}$ will be the probability level of a ``plateau'' for one or more $F_i$ and the $\lambda$ search algorithm will fail to drive $\Delta(\lambda^{\star, \varepsilon})$ close enough to zero for the $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ to be acceptable approximations of the optimal allocations. In this event we follow-up with a ``post-processing'' step that identifies lower and upper boundaries $x_{i,L}$ and $x_{i,U}$ of any
such low-density plateau containing points in $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$. If $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ is single-valued 
with the density $f_i(F_{i}^{-1}(1-\lambda^{\star, \varepsilon}))$ not close to zero, we simply take $x_{i,L} = F_{i}^{-1}(1-\lambda^{\star, \varepsilon}) = x_{i,U}$. Defining $\mathbf{x}_L = (x_{i,L})$ and $\mathbf{x}_U = (x_{i,U})$ we then interpolate an approximately optimal allocation 
$\mathbf{x} = (1-t^{\star})\mathbf{x}_L + t^{\star}\mathbf{x}_U$ where $t^{\star}$ solves 
$\sum_{i=1}^{N }(1-t^{\star})x_{i,L} + t^{\star}x_{i,U} = K$.

We note though that by using this linear interpolation (rather than some other point in the plateau satisfying the constraint) as the allocation associated with $F$ in the scoring process we are making a choice that is not
guided by any expected utility maximization principle. As such, this choice may imply strong and potentially questionable assumptions about fairness and
equity principles for the decision maker when forecasts tend to be ``clumped,'' leaving many regions with very low forecasted probability.

\section{Discussion}
\label{sec:discussion}

\section{References}

\bibliography{allocation}

\end{document}

### 2023-08-20 alloscore_manuscript/methods_apg.tex
add z to ias/crps

@@
 \begin{align}
-s_{t\mathrm{WIS}}(x^{F,t\mathrm{WIS}},y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}s_{Q}(x_{l,i}^{F_l,\mathrm{Q}, C_i}, y_l)_+ \\ 
-&= \frac{1}{m} \sum_{i=1}^{m}s_{\mathrm{AS}}(x^{F,\mathrm{AS},  K(C_i, F, L)},y) 
+s_{t\mathrm{WIS}}(x^{F,t\mathrm{WIS}},y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}(x_{l,i}^{F_l,\mathrm{Q}, C_i}- y_l)_+  + C_i x_{l,i}^{F_l,\mathrm{Q}, C_i}\\ 
+&= \frac{1}{m} \sum_{i=1}^{m}s_{\mathrm{AS}}(x^{F,\mathrm{AS},  K(C_i, F, L)},y) + C_i K(C_i, F, L)
 \end{align}
@@
 s_{\mathrm{IAS}}(x^{F, \mathrm{IAS}}(k), y) &= \int_0^{\infty} s_{\mathrm{AS}}(x^{F, \mathrm{AS}(k)}, y) dk \\
-&= \int_0^{\infty} \sum_{l=1}^N s_{\mathrm{Q}}(x^{F_l, \mathrm{Q}(c(k, F))}, y) dk \\
-&= \sum_{l=1}^N \int_0^{\infty} s_{\mathrm{Q}}(x^{F_l, \mathrm{Q}(c(k, F))}, y) dk
+&= \int_0^{\infty} \sum_{l=1}^N s_{\mathrm{Q}}(x^{F_l, \mathrm{Q}(c(k, F))}, y) - c(k, F)kdk \\
+&= \sum_{l=1}^N \int_0^{\infty} s_{\mathrm{Q}}(x^{F_l, \mathrm{Q}(c(k, F))}, y) dk -  \int_0^{\infty} c(k, F)kdk
 \end{align}

### 2023-08-20 alloscore_manuscript/methods_apg.tex
add z to ias/crps

@@
 
+$s_{tQ}(x^{F,C,K}) \mid C = c, K = k$ in $\mathrm{AP}(k)$ coordinates:
+\begin{align}
+s_{tQ}(x^{F,c,k})(y) &= 
+\begin{cases}
+\sum_{l=1}^{N}L(y_l - x_l^{F,k})_{+} + ck, & k < \sum F_l^{-1}(1-c/L) \\
+\sum_{l=1}^{N}L(y_l - x_l^{F,Q,c})_{+} + c(k - z_k(c)), & \text{otherwise, with } 
+\end{cases} \\
+z_k(c) &= (k -  \sum F_l^{-1}(1-c/L))_{+}
+\end{align}
+
+Let $k_1 = \sum F_l^{-1}(1-c_1/L)$ so that $x^{F,Q,c_1} = x^{F, k_1}$ and $z_{k_1}(\{c \leq c_1\}) = 0$, and take $0<c_a < c_1 < c_b < L$ and $0 < k_a < k_1 < k_b < K_2$.
+
+Then with $K = k_1$ fixed, the path of functions $s_{k_1}(c) = s_{tQ}(x^{F,c,k_1})$ parametrized by $c \in [0,L]$ takes values
+\begin{align}
+s_{k_1}(0)  &= L\sum ( y_l-x_l^{F,k_1})_{+}  \\
+s_{k_1}(c_a)  &= L\sum ( y_l-x_l^{F,k_1})_{+} + c_a k_1\\
+s_{k_1}(c_1) &= L\sum (y_l - x_l^{F,k_1})_{+} + c_1 k_1  \\
+s_{k_1}(c_b) &= L\sum (y_l - x_l^{F,Q, c_b})_{+} + c_b(k_1 - z_{k_1}(c_b)) \\
+&= L\sum (y_l - x_l^{F,Q, c_b})_{+} + c_b \sum F_l^{-1}(1-c_b/L)\\
+s_{k_1}(L)  &= L\sum {y_l}_{+}.
+\end{align}
+
+If instead $C = c_1$ is fixed, then $s_{c_1}(k) = s_{tQ}(x^{F,c_1,k})$ parameterized by 
+$k \in [0,K_2]$ take values
+\begin{align}
+s_{c_1}(0) & = L \sum {y_l}_{+} \\
+s_{c_1}(k_a) & = L \sum (y_l - x_l^{F,k_a})_{+} + c_1 k_a\\
+& = L \sum (y_l - x_l^{F,Q,c(k_a)})_{+} + c_1 \sum F_l^{-1}(1-c(k_a)/L)\\
+s_{c_1}(k_1) & = L \sum (y_l - x_l^{F,k_1})_{+} + c_1 k_1 \\
+s_{c_1}(k_b) & = L \sum (y_l - x_l^{F,k_1})_{+} + c_1 (k_b - z_{k_b}(c_1)) \\
+ & = L \sum (y_l - x_l^{F,k_1})_{+} + c_1 \sum F_l^{-1}(1-c(k_1)/L)\\
+s_{c_1}(K_2) & = L \sum (y_l - x_l^{F,k_1})_{+} + c_1 (K_2 - z_{K_2}(c_1))
+\end{align}
+
+Writing $k(c) = \sum F_l^{-1}(1-c/L)$,
+\begin{align}
+s_{k_1}(c) - s_{c_1}(k(c)) &= (c-c_1)\sum F_l^{-1}(1-\max(c,c_1)/L) \\
+s_{c_1}(k) - s_{k_1}(c(k)) &=  
+\end{align}
+
 But 
 \begin{align}
-k = k(c,F) &= \sum_{l=1}^{N} F_l^{-1}(1-c/L) \\
 \del{k}{c}& = \sum_{l=1}^{N} \frac{-1}{L f_l(q_{F_l,\tau(c)})} := -\mathrm{MAI}(c,F)\label{eqn:dkdc}

### 2023-08-23 alloscore_manuscript/methods_apg.tex
reorganizing in terms of scoring rules rather than procedure steps

@@
 
-Broadly speaking, we score a probabilistic forecast $F$ by identifying it with a solution to a \emph{decision problem} made up of pre-specified actions, states and consequences and rating the utility of that solution.
-That is, $F$ will receive a "better" or "worse" score in a particular instance when the consequence is better or worse of a rational decision maker using $F$ (and essentially only $F$) to select an action in the face of uncertainty about a future state.
+Broadly speaking, we score a probabilistic forecast $F$ by identifying it with a solution to a \emph{decision problem} and rating the utility of that solution.
+The term ``decision problem'' refers to a pre-specified set of permissable actions, possible future states, and resulting consequences of actions under various states.
+Accordingly, $F$ will receive a "better" or "worse" score in a particular instance when the consequence is better or worse of a rational decision maker using $F$ (and essentially only $F$) to select an action in the face of uncertainty about a future state.
 It should be noted immediately that a better or worse score in a single instance may tell us almost nothing about the intrinsic value of the forecasting method that produced $F$.
@@
 
-\section{Step 1: Defining decision problems and loss functions.}
+\section{The quantile score}
 
@@
 \begin{align}
-s_Q(x,y) = Cx + L(y-x)_+ \label{eqn:quantile_loss}
+s_{Q,C}(x,y) = Cx + L(y-x)_+ \label{eqn:quantile_loss}
 \end{align}
-where $u_+ = \max(u,0)$.
-We use the subscript $Q$ because it will be shown below that this loss function yields the quantile score for the the probability level $\alpha = 1 - C/L$.
-Loss functions of this general type are referred to as \emph{piece-wise linear}, distinguishing them from smooth non-linear loss functions such as squared error.
+where $u_+ = \max(u,0)$.  Loss functions of this general form are referred to as \emph{piece-wise linear}, distinguishing them from smooth non-linear loss functions such as squared error.
 The underlying decision problem here consists of action and outcome sets $\mathcal{X},\mathcal{Y}$ both equal to $\mathbb{R}_+$, the non-negative reals.
@@
 
-This decision problem may however be complicated by uncertainty at time $t_0$ about the unit-cost of the resource.
+The Bayes act for $C \in (0,L)$ is the solution of the first order equation 
+\begin{align}
+0 = \dby{}{x} E_F[s_{Q,C}(x,y)] &= E_{F}[\dby{}{x}s_{Q,C}(x,y)] \\
+&= C + LE_{F}[\dby{}{x}(y-x)_+] \\
+&= C - LE_{F}[\mathbf{1}\{y > x\}] \\
+&= C + L(F(x) - 1), \label{eqn:q_deriv}
+\end{align}
+that is, the quantile $x^F = F_l^{-1}(1 - C/L) = q_{F,\tau}$ for the probability level $\tau = 1-C/L$.
+(That the critical point $q_{F,\tau}$ is actually a minimum follows from \eqref{eqn:q_deriv} being non-positive to its left and non-negtive to its right, and assuming as we do for now that $F$ is smooth and strictly increasing at $q_{F,\tau}$, it is a unique minimum.) The quantile solves the decision problem with respect to the forecast $F$ by being the action which best balances expected cost against expected loss among all actions available to the decision maker.
+
+Finally, we evaluate the scoring rule $S_Q$ on $F$ when outcome $y$ realizes by measuring the loss that occurs when the action $q_{F,\tau}$ is taken:
+\begin{align}
+S_Q(F,y) = s_Q(q_{F,\tau}, y) = Cq_{F,\tau} + L(y- q_{F,\tau})_{+}.
+\end{align}
+
+\section{WIS and CRPS as quantile scores under cost uncertainty}
+
+The decision problem leading to the QS may however be complicated by uncertainty at time $t_0$ about the cost $C$ of the resource.
 We take such uncertainty as requiring the worker to commit in advance to procurement levels $x_i$ for a range of potential costs $C_i < L$ where $i \in 1, \ldots, m$.
@@
 
-A simple way to take account of the consequences of actions under counterfactual costs is to use the mean of all losses under certainty about the cost:
+We take into account the consequences of actions under a set $\mathcal{C} = \{C_i\}$ of possible costs by expanding the loss function to be a weighted mean of quantile losses under certainty about the cost:
 \begin{align}
-s_{\mathrm{WIS}}(x,y) = \frac{1}{m}\sum_{i=1}^{m}C_ix_i + L(x_i - y)_+ \quad (C_i < L \text{ for all } i)
+s_{Q,\mathcal{C}, p_C}(x,y) = \frac{1}{m}\sum_{i=1}^{m}p_i(C_ix_i + L(x_i - y)_+). \label{eqn:sum_CL_loss}
 \end{align}
-where we here use $x$ to express the decision vector $(x_1,\ldots, x_m)$. Or if there are prior beliefs about the probabilities $p_i$ of the $C_i$, we could use a weighted sum
+Here, $p_C = \{p_i\}, i = 1,\ldots,m$ is a prior distribution on $\mathcal{C}$ and $x = (x_1,\ldots,x_m)$ is the decision vector corresponding to all possible costs.
+$p_C$ would be provided by the decision maker (or perhaps a separate expert), and we emphasize that we are considering it here as part of the loss function $s_{Q,\mathcal{C}, p_C}$ that would be formulated in an effort to adapt forecast evaluation to the specifics of a public health decision problem.
+In practice, $p_C$ has usually been taken to be uniform, which we will show for a particular $\mathcal{C}$ (given $L$) leads to the WIS familiar in ID forecasting hubs.
+This might be interpreted as deferring to an uninformative prior.
+
+We can also consider $C$ a continuous random variable with a prior density $f_C$ and accommodate this by letting $m \to \infty$ in \eqref{eqn:sum_CL_loss} 
+which (with some mathematical care) converts the weighted mean into an integral
 \begin{align}
-s_{w\mathrm{WIS}}(x,y) = \frac{1}{m}\sum_{i=1}^{m}p_i(C_ix_i + L(x_i - y)_+). \label{eqn:sum_CL_loss}
+s_{Q,f_C}(x,y) = \int_{0}^{L} (cx(c) + L(x(c) - y)_+) f_C(c)dc{}.
 \end{align}
-As for $Q$, the $WIS$ and $wWIS$ subscripts indicate that, for particular values of the $\{C_i\}$ and $L$, this loss function produces the WIS or a version that weights intervals (i.e.
-the average of QS's for $C_i$ and $L-C_i$) with non-standard weights (combining $p_i$ and $C_i/L$).
-
-We can also consider $C$ a continuous random variable and accommodate this by letting $m \to \infty$ in \eqref{eqn:sum_CL_loss} which (with some mathematical care) converts the loss into an integral
-\begin{align}
-s_{\mathrm{CRPS}}(x,y) = \frac{1}{L}\int_{0}^{L} (cx(c) + L(x(c) - y)_+) dc
-\end{align}
-over all relevant costs $c \in [0,L]$, or a weighted version
-\begin{align}
-s_{w\mathrm{CRPS}}(x,y) = \int_{0}^{L} (cx(c) + L(x(c) - y)_+) f_C(c)dc{}
-\end{align}
-that uses a prior density $f_C$ on the possible costs. These loss functions will yield the traditional CRPS and the wieghted version of \cite{gneiting2011weightedScoringRules}.
 
@@
 
-We can also expand our decision problem to include multiple actions that will all, in fact (not counterfactually), have to be taken.
+The Bayes acts for losses $s_{Q,\mathcal{C}, p_C}$ and $s_{Q,f_C}$ therefore correspond not just to the vanishing of the derivative with respect to a single
+decision variable $x$ but to the vanishing of the partial derivatives with respect to all $m$ decision variables $\{x_i\}$ in the first case or the infinitely many variables $\{x(c)\}$ in the second case.
+But since the losses are sums of single varianble losses, passing to solutions of these higher dominsional equations is a mere formality of the viewing the set of Bayes acts, i.e.
+quantiles, for costs $c \in \mathcal{C}$ or $(0,L)$ as a vector
+$(x^F(1),\ldots,x^F(m))=(q_{F,\tau_1},\ldots,q_{F,\tau_m})$ or a function $x(c) = q_{F,\tau(c)}$.
+
+Similarly, given an outcome $y$, evaluation of the scoring rules $s_{Q,\mathcal{C}, p_C}$ and $s_{Q,f_C}$ on the forecast $F$ amounts to vectorizing the
+formula for $S_Q(F,y)$. That is,
+\begin{align}
+S_{Q,\mathcal{C}, p_C}(F,y) &= \frac{1}{m}\sum_{i=1}^{m}p_i(C_i q_{F,\tau_i} + L(q_{F,\tau_i} - y)_+) \\
+S_{Q, f_C}(F,y) &= \int_{0}^{L}(cq_{F,\tau(c)} + L(q_{F,\tau(c)} - y)_+) f_C(c)dc.
+\end{align}
+
+Again, the hierarchy we have introduced here describing WIS and CRPS as parallels to the quantile score for higher dimensional action sets is in itself a mere formality employing the linearity of differentiation with respect to the expectation operator $E_{C}$ for $C$ considered as a random variable.
+But as we move now to decision problems with multiple outcomes, it will turn out to provide a key shift in perspective that allows the scoring rules we define in terms of allocation under scarcity to be seen as a natural and non-trivial extension of the hierarchy.
+
+\section{Aggregating QS, WIS, and CRPS across dimensions}
+
+We next expand our decision problem to include multiple actions that will all, in fact (not counterfactually), have to be taken.
 This includes the motivating scenario for our work, the simultaneous procurement of resources across several locations (U.S.
@@
 \mathcal{X}_{cert} &= \mathbb{R}_{+}^N \\
-\mathcal{X}_{disc} &= (\mathbb{R}_{+}^m)^N\\
+\mathcal{X}_{disc} &= (\mathbb{R}_{+}^m)^N = \{x(i):\{1,\ldots,m\} \to \mathbb{R}_{+}^N\}\\
 \mathcal{X}_{cont} &= \{x(c):[0,\infty) \to \mathbb{R}_{+}^N\}.
 \end{align}
-And for these "vector-action" problems we can define \emph{total} versions of our loss functions
+And for these ``vector-action'' problems we can define \emph{total} versions of our loss functions
 \begin{align}
-s_{t\mathrm{Q}}(x,y) &= \sum_{l=1}^N Cx_l + L(y_l-x_l)_+ \\
-s_{t\mathrm{WIS}}(x,y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}C_i x_{l,i} + L(x_{l,i} - y_l)_+ \\
-s_{tw\mathrm{WIS}}(x,y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}p_i(C_i x_{l,i} + L(x_{l,i} - y_l)_+)\\
-s_{t\mathrm{CRPS}}(x,y) &= \frac{1}{L} \sum_{l=1}^N\int_{0}^{L} (cx_l(c) + L(x_l(c) - y_l)_+) dc \\
-s_{tw\mathrm{CRPS}}(x,y) &= \sum_{l=1}^N \int_{0}^{L} (cx_l(c) + L(x_l(c) - y_l)_+) f_C(c)dc.
+s_{tQ}(x,y) &= \sum_{l=1}^N Cx_l + L(y_l-x_l)_+ \\
+s_{tQ, \mathcal{C}, p_C}(x,y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}p_i(C_i x_{l,i} + L(x_{l,i} - y_l)_+)\\
+s_{tQ, f_C}(x,y) &= \sum_{l=1}^N \int_{0}^{L} (cx_l(c) + L(x_l(c) - y_l)_+) f_C(c)dc.
 \end{align}
@@
 
+\section{The allocation score}
+
 The loss functions (and corresponding decision problems) defined so far  have all been thouroughly covered, from one perspective or another, in various strands of forecast evaluation literature.
@@
 
-Consider again the simultaneous resource procurement decision a public health worker must make in attempting to mitigate a pandemic across multiple locations. 
+Consider again the resource procurement decision $x(c) \in \mathcal{X}_{cont}$ (or $\mathcal{X}_{disc}$ if $c = C_i, i \in 1,\ldots,m$) a public health worker must make in 
+attempting to mitigate a pandemic across multiple locations $l \in 1,\ldots,N$.
+But now we ask the worker to extend their decision to cover the contingency that the total amount of the resource $\sum x_l(C)$ that they specify for the realized cost $C$ turns out to not be available.
+This could be due to demand-side factors such as a budget cut, or supply-side factors such as manufacturing or shipping failures.
+In either case, we assume that these factors are undetermined at time $t_0$ when decisions must be made and model them with a positive random variable $K$ giving the total amount of resource that will be available
+at time $t_1$.
+This adds, for each $k \in \mathbb{R}_{+}$, a component to our overall decision problem that asks the question: If the total amount $K$ available of the resourse is insufficient allow the procurements $x_l(C)$ you specify for the realized cost $C$, what alternate procurement levels $x_l^{k,C}$ would you choose?
+The action set for this new component is
+\begin{align}
+\mathcal{X}_k = \{x \in \mathbb{R}_{+}^N \mid \sum_{l=1}^{N} x_l = k\}.
+\end{align}
+Note that because this decision is only relevant when $\sum x_l(C)>k$ we do not include actions which use less than $k$ total of the resource. (Why?)
+
+In these action set variables, the loss function we used for the unconstrained $N$ location decision problem with action set $\mathcal{X}_{cert}$ now has the expression
+\begin{align}
+s_{tQ}\left(x, y\right) = Ck + \sum_{l=1}^{N} L(y_l - x_l)_{+}.
+\end{align}
+We see then that the cost $C$ will not effect the Bayes act for this component of the decision problem motivating the definition of a new single loss function (component), the \emph{allocation loss function}
+\begin{align}
+s_{A}(x,y):= L\sum_{l=1}^{N} (y_l - x_l)_{+},
+\end{align}
+we use for all $\mathcal{X}_k$.
+
+\subsubsection*{Old allocation intro}
 We now add a constraint $K$ on the total amount of the resource available for allocation across all locations.
@@
 It will turn out, as explained in the following sections, that for a given forecast $F$, a scoring rule derived with respect to the AP from $s_{tQ}$ under uncertainty about $C$ -- that is, from $s_{tWIS}$ or $s_{tCRPS}$ -- will have the same values for all outcomes $y$ as a scoring rule that is a weighted average or integration of scoring rules derived from allocation losses $s_{A,K}$ with respect the AP's for a range of $K$'s.
-
-\section{Step 2: Derive the Bayes act for a forecast $F$}
-
-I think we should not hide the formula 
-\begin{align}
-\dby{}{x_{il}} E_F[s_{.}(x,y)] &= E_{F}[\dby{}{x_{il}} s_{.}(x,y)] = C - LE_{F_l}[\mathbf{1}(y_l > x_{il})] = C + L(F_l(x_{il}) - 1) \\
-\implies x_{il}^{F_l} &= F_l^{-1}(1 - C/L) \\
-&\text{ or for CRPS} \\
-x_{l}^{F_l}(c) &= F_l^{-1}(1 - c/L)
-\end{align}
 
@@
 
-\subsection{Solving the Allocation Problem}
+\section{Solving the Allocation Problem}
 
-As the solution to a constrained optimization problem, the optimal allocation $x^F$ according to a forecast $F$ can be described formally as the first component of a stationary point $(x^F, z_K, \lambda^{\star})$ in the region $\{\mathbf{x} \geq 0, \lambda \geq 0\}$ for the Lagrangian
+As the solution to a constrained optimization problem, the optimal allocation $x^F$ according to a forecast $F$ can be described formally as the first component of a stationary point $(x^F, z_K^F, \lambda^{\star})$ in the region $\{x, z \geq 0, \lambda \in \mathbb{R}\}$ for the Lagrangian
 \begin{align}
-\mathcal{L}(\mathbf{x},\lambda; K, C, L) = L\sum_{l=1}^{N} E_{F_l}[(Y_l - x_l)_{+}] + C(K-z_K) + \lambda\left(z_K + \sum_{l=1}^{N} x_l - K\right).
+\mathcal{L}(x, z_K, \lambda; K, C, L) = L\sum_{l=1}^{N} E_{F_l}[(Y_l - x_l)_{+}] + C(K-z_K) + \lambda\left(z_K + \sum_{l=1}^{N} x_l - K\right).
 \end{align}
@@
 0 = \frac{\partial}{\partial x_l} \mathcal{L}(x^F, z_K, \lambda) &= L E_F \left[\frac{d}{dx_l}(Y_l - x_l^F)_{+}\right] + \lambda \\
-& = L E_F \left[-\mathbf{1}\{Y_i \geq x_i^F\}\right] + \lambda \\
-& = L(F_i(x_l^F)-1) + \lambda \label{eqn:x_stationary}
+& = L E_F \left[-\mathbf{1}\{Y_l \geq x_l^F\}\right] + \lambda \\
+& = L(F_l(x_l^F)-1) + \lambda \label{eqn:x_stationary}
 \end{align}
@@
 \begin{align}
-0 = \del{}{\lambda} \mathcal{L}(\mathbf{x}^F, \lambda^{\star}) &= \sum_{i=1}^{N} x_i^F - K\\
-& = \sum_{\{i | x_i^F > 0\}} F_{i}^{-1}(1-\lambda^{\star}) - K \quad \text{ (by \eqref{eqn:x_stationary})}. \label{eqn:lambda_stationary}
+0 = \del{}{\lambda} \mathcal{L}(x^F, z_K^F, \lambda^{\star}) &= \sum_{l=1}^{N} x_l^F + z_K^F - K\\
+& = \sum_{\{l | x_l^F > 0\}} F_{l}^{-1}(1-\lambda^{\star}) + z_K^F - K \quad \text{ (by \eqref{eqn:x_stationary})}. \label{eqn:lambda_stationary}
 \end{align}

### 2023-08-24 alloscore_manuscript/methods_apg.tex
continuing to reorient toward AS and auxiliary to CRPS

@@
 In this section/report/paper/rambling manifesto, we present a methodological perspective on the use scoring rules for forecast evaluation that we see as well-adapted to our motivating research goal to understand and describe the social utility of infectious disease forecasts.
-Our approach follows the common practice in decision theory of giving central roles to the foundational concepts of state, action, and consequence.
+Our approach follows the common practice in decision theory of giving central roles to the foundational concepts of state, action, and result as the basic components of a decision problem.
 From these primitives, we first derive the familiar measures of quantile score (QS), continuous ranked probability score (CRPS) and their aggregates across forecast targets.
-We then introduce new scoring rules, the \emph{specific} and \emph{integrated allocation scores}, as reparametrizations of QS and CRPS aggregates that result when part of our definition of consequence is shifted into our definition of action.
-Specifically, we will trade the regret a decision maker experiences as a consequence of seeing allocated resources go unused for the requirement that a decision maker choose an action that allocates no more than a certain amount in total.
-This exchange mirrors the duality in mathematical optimization theory between constraints on "decision variables" and the "Lagrange multiplier variables" which quantify the sensitivity of solutions to the constraints, that is, the regret (or relief) experienced when a change in the constraint leads to a change in the optimal cost.
+We then introduce new scoring rules, the \emph{specific} and \emph{integrated allocation scores} designed to reflect aspects of the decision problem corresponding to the CRPS that emerge when multiple instances are confronted simultaneously.
+Specifically, we will consider not only whether using more or less of a resource leads to a better result in a specific problem, but also whether a limited resource is better allocated to one problem rather than another.
+This added concern brings into play the duality in mathematical optimization theory between constraints on ``decision variables" and the ``Lagrange multiplier variables" which quantify the sensitivity of solutions to constraints, that is, the degree to which loosening constraints improves results under the optimal action..
 Our computational methods (see ..) in fact make use of this connection.
@@
 Broadly speaking, we score a probabilistic forecast $F$ by identifying it with a solution to a \emph{decision problem} and rating the utility of that solution.
-The term ``decision problem'' refers to a pre-specified set of permissable actions, possible future states, and resulting consequences of actions under various states.
-Accordingly, $F$ will receive a "better" or "worse" score in a particular instance when the consequence is better or worse of a rational decision maker using $F$ (and essentially only $F$) to select an action in the face of uncertainty about a future state.
+The term ``decision problem'' refers to a pre-specified set of allowed actions, possible future states, and the possible results of allowed actions under those states.
+Accordingly, $F$ will receive a ``better" or ``worse" score in a particular instance when the result is better or worse of a rational decision maker using $F$ (and essentially only $F$) to select an action in the face of uncertainty about a future state.
 It should be noted immediately that a better or worse score in a single instance may tell us almost nothing about the intrinsic value of the forecasting method that produced $F$.
@@
 We also will refer to states as outcomes, which helps to emphasize their future realization.
-But we want to clearly distinguish an outcome from a consequence, which as we discuss below, is the result of taking an action under a certain outcome.
+But we want to clearly distinguish an outcome from a result, which as we discuss below, which is determined by an action under a certain outcome.
 
@@
 
-And when actions are assigned to states, there are consequences, which we abstractly identify with an element $c$ of a set $\mathcal{C}$.
-We have two basic examples in mind of what such a $c$ might be.
+And when actions are assigned to states, there are results, which we abstractly identify with an element $r$ of a set $\mathcal{R}$.
+We have two basic examples in mind of what such a $r$ might be.
 First, we consider some amount of the regret mentioned above that resources made available in a place or time are not actually needed at that place or time.
 We emphasize that the nature of this regret is potentially quite amorphous.
-In business contexts (e.g., the "newsvendor problem" from inventory management) this regret is often easily quantifiable as monetary waste, but in a public health context, it could...
+In business contexts (e.g., the ``newsvendor problem" from inventory management) this regret is often easily quantifiable as monetary waste, but in a public health context, it could...
 (A desire to better understand over-allocation regret and its relevance to infectious disease forecast evaluation was actually a primary impetus for the present work.)
-Our second and more concrete example of a consequence is the loss incurred when a sick individual is not able to access a resource because it was not made available to them where and when they needed it.
-We emphasize here that while we formally identify such loss as a consequence of a particular action being taken under a particular state, it may or may not be an avoidable loss.
+Our second and more concrete example of a result is the loss incurred when a sick individual is not able to access a resource because it was not made available to them where and when they needed it.
+We emphasize here that while we formally identify such loss as a result of a particular action being taken under a particular state, it may or may not be an avoidable loss.
 That is, it could just as well result from a factor outside a decision maker's control such as a severe resource constraint as from a choice to withhold a resource, motivated, for example, by a desire to reduce waste.
@@
 
-With a view towards a decision problem defined by these examples of sets $\mathcal{X}$, $\mathcal{Y}$, and $\mathcal{C}$ of actions, outcomes, and consequences, we now give a formal three-step procedure for defining and evaluating a \emph{scoring rule} for a forecast $F$ for a future value $y \in \mathcal{Y}$ of $Y$.
+With a view towards a decision problem defined by these examples of sets $\mathcal{X}$, $\mathcal{Y}$, and $\mathcal{R}$ of actions, outcomes, and results, we now give a formal three-step procedure for defining and evaluating a \emph{scoring rule} for a forecast $F$ for a future value $y \in \mathcal{Y}$ of $Y$.
 
 \begin{enumerate}
-\item Specify a \emph{loss function} $s(x,y)$ that quantifies the utility or disutility of the consequence $c$ of taking action $x$ under outcome $y$.
+\item Specify a \emph{loss function} $s(x,y)$ that quantifies the utility or disutility of the result $r$ of taking action $x$ under outcome $y$.
 \item Map the given probabilistic forecast $F$ for $Y$ to an action $x^F$ which ``solves" the decision problem by minimizing the expected loss $E_{F}[s(x,Y)]$ over all $x \in \mathcal{X}$ under the distribution $F$.
@@
 For example, having more available resources usually allows for actions incurring less loss.
-\item Assign $F$ the score $s(x^F,y)$, that is, the numerical quantification via $s$ of the consequence under outcome $y$ of accepting $x^F$ as a solution to the decision problem.
+\item Assign $F$ the score $s(x^F,y)$, that is, the numerical quantification via $s$ of the result under outcome $y$ of accepting $x^F$ as a solution to the decision problem.
 \end{enumerate}{}
@@
 Thus, given definitions of $\mathcal{X}$, $\mathcal{Y}$, $\mathcal{C}$, and a loss function $s$, we produce a scoring rule $S(F,y):= s(x^F,y)$ that can be applied to any $F$ for any outcome $y \in \mathcal{Y}$.
-We emphasize again a key feature of this formulation: the scoring rule $S$ is defined entirely in terms of what we take to be possible actions, states, and consequences along with a function $s$ that assigns values to the possible consequences.
+We emphasize again a key feature of this formulation: the scoring rule $S$ is defined entirely in terms of what we take to be possible actions, states, and results along with a function $s$ that assigns values to the possible results.
 
@@
 We begin by imagining a public health worker trying to decide at time $t_0$ what quantity $x$ to procure of a protective resource such as beds, staff, or medicine in anticipation of $y$ new infections that will be observed (and create resource demand) at time $t_1>t_0$.
-The resource has a unit-cost of $C$, and we assign a loss of $L$ to the consequence of a sick individual not having access to the resource.
+The resource has a unit-cost of $C$, and we assign a loss of $L$ to the result that a sick individual cannot access the resource.
 This specifies a loss function
@@
 The underlying decision problem here consists of action and outcome sets $\mathcal{X},\mathcal{Y}$ both equal to $\mathbb{R}_+$, the non-negative reals.
-(We ignore for now any discreteness of the resource.) The consequence set $\mathcal{C}$ is abstractly a set of over-expenditure regrets and numbers of sick individuals with unmet resource need, but the loss function $s$ maps $\mathcal{C}$ into $\mathbb{R}_+$.
+(We ignore for now any discreteness of the resource.) The results set $\mathcal{R}$ is abstractly a set of over-expenditure regrets and numbers of sick individuals with unmet resource need, but the loss function $s$ maps $\mathcal{C}$ into $\mathbb{R}_+$.
 We also from now on make the assumption that $C < L$ since otherwise -- at least according to this loss function -- there is no ``cost-effectiveness'' in trying to prevent even a single case of unmet need.
@@
 We take such uncertainty as requiring the worker to commit in advance to procurement levels $x_i$ for a range of potential costs $C_i < L$ where $i \in 1, \ldots, m$.
-Even though only one of the $x_i$ will end up needing to be procured, there does not seem to be any intrinsically decision theoretic principle by which our loss function should ignore counter-factual but exactly specified consequences of the decision under unrealized costs.
+Even though only one of the $x_i$ will end up needing to be procured, there does not seem to be any intrinsically decision theoretic principle by which our loss function should ignore counter-factual but exactly specified results of the decision under unrealized costs.
 If we wanted to commit to a \emph{locality} principle for prediction where losses and scores only depend on predictive densities in a neighborhood of the relevant observation, this attempt would falter, and we would need to frame our methods around the log score and it's higher order relatives.
@@
 
-We take into account the consequences of actions under a set $\mathcal{C} = \{C_i\}$ of possible costs by expanding the loss function to be a weighted mean of quantile losses under certainty about the cost:
+We take into account the results of actions under a set $\mathcal{C} = \{C_i\}$ of possible costs by expanding the loss function to be a weighted mean of quantile losses under certainty about the cost:
 \begin{align}
@@
 \begin{align}
-s_{Q,f_C}(x,y) = \int_{0}^{L} (cx(c) + L(x(c) - y)_+) f_C(c)dc{}.
+s_{Q,f_C}(x,y) = \frac{1}{L}\int_{0}^{L} (cx(c) + L(x(c) - y)_+) f_C(c)dc{}.
 \end{align}
@@
 S_{Q,\mathcal{C}, p_C}(F,y) &= \frac{1}{m}\sum_{i=1}^{m}p_i(C_i q_{F,\tau_i} + L(q_{F,\tau_i} - y)_+) \\
-S_{Q, f_C}(F,y) &= \int_{0}^{L}(cq_{F,\tau(c)} + L(q_{F,\tau(c)} - y)_+) f_C(c)dc.
+S_{Q, f_C}(F,y) &= \frac{1}{L}\int_{0}^{L}(cq_{F,\tau(c)} + L(q_{F,\tau(c)} - y)_+) f_C(c)dc. \label{eqn:CRPS_cform}
 \end{align}
+
+How do the WIS and CRPS arise from this construction? Defining $\tau = 1-C/L$ as a random variable gives it the density 
+$f_{\tau}(\tau) = -f_C\left(L(1-\tau)\right)/L$, 
+and \eqref{eqn:CRPS_cform} becomes 
+\begin{align}
+S_{Q, f_C}(F,y) &= \int_{0}^{1}((1-\tau)q_{F,\tau} + (y-q_{F,\tau})_{+})f_{\tau}(\tau)d\tau \\
+&= \int_{0}^{1}(\mathbf{1}\{y \leq q_{F,\tau}\}-\tau)(q_{F,\tau} - y)f_{\tau}(\tau)d\tau + \int_{0}^{1}(1-\tau)y f_{\tau}(\tau)d\tau \\
+&= wCRPS(F,y) + (1-E[\tau])y
+\end{align}
+where $w$CRPS is the weighted CRPS of \cite{gneiting2011weightedScoringRules} with quantile weight function $\nu(\tau) = \frac{1}{2}f_{\tau}(\tau)$. 
+(The $(1-E[\tau])y$ that results from our loss function $s_Q$ adding a $Cy$ term to the standard quantile loss function does not effect the dependence of the scoring rule of $F$.)
+
+Writing $s_{Q,\tau}(x,y) = s_{Q,C}(x,y)/L = (1-\tau)x + (y-x)_{+}$, we have
+\begin{align}
+s_{Q,\tau}(q_{F,\tau},y) + s_{Q,1-\tau}(q_{F,1-\tau},y) &= q_{F,\tau} + \tau(q_{F,1-\tau} - q_{F,\tau}) + (y - q_{F,\tau})_{+} + (y - q_{F,1-\tau})_{+} \\
+&= y + \tau\left(q_{F,1-\tau} - q_{F,\tau} + \frac{1}{\tau}(q_{F,\tau} - y)_{+} + \frac{1}{\tau}(y - q_{F,1-\tau})_{+}\right) \\
+&= y + \tau IS_{2\tau}(F,y)
+\end{align}
+where $IS_{\alpha}(F,y)$ is the \emph{interval score} suggested by expression (43) in \cite{gneiting2007strictly} and formalized by \cite{bracher2021evaluating}.
+
+So when $\mathcal{C} = \mathcal{C}_1\cup (L-\mathcal{C}_1)$ and $p_\mathcal{C}$ is uniform, we get
 

### 2023-08-24 alloscore_manuscript/methods_apg.tex
add WIS in sec4

@@
 &= \int_{0}^{1}(\mathbf{1}\{y \leq q_{F,\tau}\}-\tau)(q_{F,\tau} - y)f_{\tau}(\tau)d\tau + \int_{0}^{1}(1-\tau)y f_{\tau}(\tau)d\tau \\
-&= wCRPS(F,y) + (1-E[\tau])y
+&= w\mathrm{CRPS}(F,y) + (1-E[\tau])y
 \end{align}
@@
 &= y + \tau\left(q_{F,1-\tau} - q_{F,\tau} + \frac{1}{\tau}(q_{F,\tau} - y)_{+} + \frac{1}{\tau}(y - q_{F,1-\tau})_{+}\right) \\
-&= y + \tau IS_{2\tau}(F,y)
+&= y + \tau \mathrm{IS}_{2\tau}(F,y)
 \end{align}
-where $IS_{\alpha}(F,y)$ is the \emph{interval score} suggested by expression (43) in \cite{gneiting2007strictly} and formalized by \cite{bracher2021evaluating}.
-
-So when $\mathcal{C} = \mathcal{C}_1\cup (L-\mathcal{C}_1)$ and $p_\mathcal{C}$ is uniform, we get
+The scoring rule $\mathrm{IS}_{\alpha}(F,y)$ is the \emph{interval score} suggested by expression (43) in \cite{gneiting2007strictly} 
+and formalized by \cite{bracher2021evaluating} as a means for defining the weighted interval score 
+\begin{align}
+\mathrm{WIS}_{\alpha_{\{0: J\}}}(F, y)=\frac{1}{J+1 / 2}\left(w_0 |y-q_{F,.5}|+\sum_{j=1}^J w_j \operatorname{IS}_{\alpha_j}(F, y)\right).
+\end{align}
+So by taking $\mathcal{C} = \{(1 \pm \alpha_j)L/2 \mid j > 0\}\cup \{L/2\}$ and $p_\mathcal{C}$ symmetric across $L/2$, we get 
+$S_{Q,\mathcal{C}, p_C}(F,y) = \mathrm{WIS}_{\alpha_{\{0: J\}}}(F, y)$ with $w_j = p_{j}\alpha_j/2$ and $w_0 = p_{L/2}$.
 

### 2023-08-25 alloscore_manuscript/methods_apg.tex
fix

@@
 So by taking $\mathcal{C} = \{(1 \pm \alpha_j)L/2 \mid j > 0\}\cup \{L/2\}$ and $p_\mathcal{C}$ symmetric across $L/2$, we get 
-$S_{Q,\mathcal{C}, p_C}(F,y) = \mathrm{WIS}_{\alpha_{\{0: J\}}}(F, y)$ with $w_j = p_{j}\alpha_j/2$ and $w_0 = p_{L/2}$.
+$S_{Q,\mathcal{C}, p_C}(F,y) = \mathrm{WIS}_{\alpha_{\{0: J\}}}(F, y) + E_{p_{\mathcal{C}}}[y]$ with $w_j = p_{j}\alpha_j/2$ and $w_0 = p_{L/2}$.
 
@@
 attempting to mitigate a pandemic across multiple locations $l \in 1,\ldots,N$.
-But now we ask the worker to extend their decision to cover the contingency that the total amount of the resource $\sum x_l(C)$ that they specify for the realized cost $C$ turns out to not be available.
+But now we ask the worker to extend their decision to cover the contingency that there turns out to only be a total amount $k$ of the resource available for allocation, less than the amount $\sum x_l(C)$ that they would choose to distribute if balancing only the realized cost $C$ against potential unmet demand.
 This could be due to demand-side factors such as a budget cut, or supply-side factors such as manufacturing or shipping failures.
@@
 at time $t_1$.
-This adds, for each $k \in \mathbb{R}_{+}$, a component to our overall decision problem that asks the question: If the total amount $K$ available of the resourse is insufficient allow the procurements $x_l(C)$ you specify for the realized cost $C$, what alternate procurement levels $x_l^{k,C}$ would you choose?
+This adds, for each $k \in \mathbb{R}_{+}$, a component to our overall decision problem that asks the question: If the total amount $K$ available of the resource is insufficient allow the procurements $x_l(C)$ you specify for the realized cost $C$, what alternate procurement levels $x_l^{k,C}$ would you choose?
 The action set for this new component is
@@
 \end{align}
-we use for all $\mathcal{X}_k$.
+we use in deriving Bayes acts for all $\mathcal{X}_k$.
 

### 2023-08-25 alloscore_manuscript/methods_apg.tex
start adjusting lagrange method to equality constraint

@@
 
-As the solution to a constrained optimization problem, the optimal allocation $x^F$ according to a forecast $F$ can be described formally as the first component of a stationary point $(x^F, z_K^F, \lambda^{\star})$ in the region $\{x, z \geq 0, \lambda \in \mathbb{R}\}$ for the Lagrangian
+As the solution to a constrained optimization problem, the optimal allocation $x^F$ according to a forecast $F$ can be described formally as the first component of a stationary point $(x^{F,k}, \lambda^{\star})$ in the region $\{x \geq 0, \lambda \in \mathbb{R}\}$ for the Lagrangian
 \begin{align}
-\mathcal{L}(x, z_K, \lambda; K, C, L) = L\sum_{l=1}^{N} E_{F_l}[(Y_l - x_l)_{+}] + C(K-z_K) + \lambda\left(z_K + \sum_{l=1}^{N} x_l - K\right).
+\mathcal{L}(x, \lambda; F, k, L) = L\sum_{l=1}^{N} E_{F_l}[(Y_l - x_l)_{+}] + \lambda\left(\sum_{l=1}^{N} x_l - k\right).
 \end{align}

### 2023-08-29 alloscore_manuscript/allocation.tex
an attempt to remove enough content to get to something suitable for 'paper 1'

--- 2023-08-09 alloscore_manuscript/allocation.tex
@@
 
-Briefly, the allocation score of a forecast is the avoidable unmet demand that results from using that forecast to set resource allocations by minimizing expected unmet need.
+Briefly, the allocation score of a forecast is the avoidable unmet need that results from using that forecast to set resource allocations by minimizing expected unmet need.
 For example, suppose that a decision-maker is provided with forecasts of the level of need for medical resources in each of several states or hospital systems.
@@
 %By ``unnecessary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
-If the amount of resources that is available to distribute is less than the actual need, some amount of unmet demand is unavoidable.
-The allocation score for a forecast does not include the unmet demand that was unavoidable given the resource constraint, and so it measures only the amount of unmet demand that could have been prevented by using a different allocation of available resources than that suggested by the forecast.
+If the amount of resources that is available to distribute is less than the actual need, some amount of unmet need is unavoidable.
+The allocation score for a forecast does not include the unmet need that was unavoidable given the resource constraint, and so it measures only the amount of unmet need that could have been prevented by using a different allocation of available resources than that suggested by the forecast.
 We elaborate on these ideas in Section~\ref{sec:methods}.
@@
 
-We give a high-level review of a general procedure for developing proper scoring rules that are tailored to a specific decision-making task in section \ref{sec:methods.decisiontheory}.
-In section \ref{sec:methods.quantileloss} we review how quantile loss can be obtained within this framework in a setting where a decision-maker is required to determine the quantity of a good to procure while balancing the cost of purchasing an additional unit of the good with loss that may result from under-procurement.
-We then discuss how the continuous ranked probability score (CRPS) can be obtained as an integral of the quantile loss across values of the cost/loss ratio.
-These developments mirror the structure of section \ref{sec:methods.allocation}.
-There, we develop a novel \emph{allocation score} that is analogous to the quantile score but is suitable for evaluation of forecasts in the context of decisions about allocation of limited resources across multiple locations when the resource constraint is known.
-We then describe an \emph{integrated allocation score} that is analogous to the CRPS and is obtained by integrating the allocation score across values of the resource constraint.
+We begin with an informal description of the allocation score and some examples illustrating its key characteristics in section \ref{sec:methods.overview}.
+In section \ref{sec:methods.detailed} we develop the allocation score more carefully, building on decision-theoretic procedures for deriving proper scoring rules.
+We then comment on some connections between the allocation score that we propose and other common scores that can be derived from decision-theoretic foundations, such as the quantile score, WIS, and CRPS, in section \ref{sec:methods.related}.
 
 \subsection{Overview of Allocation Scoring}
+\label{sec:methods.overview}
 
-Suppose that a decision-maker is tasked with determining how much of a resource such as a medical supply should be allocated to each of $n$ locations, subject to the constraint that only $K$ units of the resource are available in total.
-If the decision-maker is provided with a multivariate forecast $F$ that predicts resource demand in each location, one option is to choose the resource allocation that minimizes the expected total unmet demand according to the forecast.
-We will give a more precise mathematical statement in section \ref{sec:methods.allocation}, but at a high level, the total expected unmet demand according to the forecast is
+Suppose that a decision-maker is tasked with determining how much of a resource such as a medical supply should be allocated to each of $N$ locations, subject to the constraint that only $K$ units of the resource are available.
+If the decision-maker is provided with a multivariate forecast $F$ that predicts resource need in each location, one option is to choose the resource allocation that minimizes the expected total unmet need according to the forecast.
+We will give a more precise mathematical statement in section \ref{sec:methods.allocation}, but informally, the total expected unmet need according to the forecast is
 \begin{align}
-\sum_{i=1}^n E_{F_i}[\text{unmet need in location $i$}],
+\sum_{i=1}^N E_{F_i}[\text{unmet need in location $i$}], \label{eqn:informal_objective}
 \end{align}
-where $F_i$ is the marginal forecast distribution of demand in location $i$.
+where $F_i$ is the marginal forecast distribution of need in location $i$ and the unmet need in a particular location is the difference between resource need in that location and the number of resources that were allocated there.
 This allocation problem has an intuitively appealing solution: allocate so that the probabilities of need exceeding allocation in various locations are as close to each other as possible.
@@
 
-After time passes and the actual level of resource need has been observed, the quality of a selected allocation can be measured by comparing the actual demand in each location to the amount of resources that were sent there.
-Specifically, we compute the total unmet demand that resulted from the selected allocation:
+After time passes and the actual level of resource need has been observed, the quality of a selected allocation can be measured by comparing the actual need in each location to the amount of resources that were sent there.
+Specifically, we compute the total unmet need that resulted from the selected allocation:
 \begin{align}
-    \sum_{i=1}^n \text{unmet demand in location $i$ based on the selected allocation}.
+    \sum_{i=1}^N \text{unmet need in location $i$}. \label{eqn:informal_loss}
 \end{align}
-One allocation is better than another if it results in lower total unmet demand.
+We emphasize that in Equation \eqref{eqn:informal_loss} the calculation of unmet need is based on the actual resource need that was realized in each location, while in Equation \eqref{eqn:informal_objective} the calculation of unmet need was based on a hypothetical potential future level of resource need.
+Once the actual levels of resource need have been observed, we can obtain a quantitative measure of the quality of alternative allocation decisions: one allocation is better than another if it results in lower total unmet need.
 
-The \textbf{allocation score} of the forecast $F$ is the avoidable unmet demand that results from using the allocation that minimizes the expected unmet need according to that forecast.
-By ``avoidable unmet demand'', we mean that the allocation score does not include the amount of unmet demand that was inevitable simply because the amount of available resources $K$ was less than the demand for resources.
-Thus, the allocation score measures the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
-An allocation score of 0 is optimal, and indicates that no other allocation of resources could have met demand better than the allocation suggested by $F$.
+The \textbf{allocation score} of the forecast $F$ is the avoidable unmet need that results from using the allocation that minimizes the expected unmet need according to that forecast.
+By ``avoidable unmet need'', we mean that the allocation score does not include the amount of unmet need that was inevitable simply because the amount of available resources $K$ was less than the need for resources.
+Rather, the allocation score measures the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
+An allocation score of 0 is optimal, and indicates that no other allocation of resources could have met need better than the allocation suggested by $F$.
 A larger allocation score indicates that it would have been possible to improve upon the allocation suggested by $F$.
 
-\paragraph{Example 1} Suppose we have a forecast $F$ for demand in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$ and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$.
+\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$ and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$.
 When the marginal forecasts are exponential distributions, it can be shown that the optimal allocation divides the available resources among the locations proportionally to the scale parameters $\sigma_i$ (see section *** of the supplemental materials).
@@
     \caption{An illustration of the resource allocation problem in Example 1.
-There are $n = 2$ locations, with predictive distributions $F_1 = \mathrm{Exp}(1)$ and $F_2 = \mathrm{Exp}(1/4)$.
+There are $N = 2$ locations, with predictive distributions $F_1 = \mathrm{Exp}(1)$ and $F_2 = \mathrm{Exp}(1/4)$.
 The cdfs of these distributions are illustrated in the panels at bottom and right.
@@
 Based on these observed needs, we can measure the quality of the allocation suggested by the forecast by calculating the amount of unmet need that resulted from that allocation over and above what was unavoidable given the resource constraint.
-With $K = 5$ units of the resource, the allocation based on the forecast exactly meets the observed demand in location 1, but it leaves 6 units of need unmet in location 2.
+With $K = 5$ units of the resource, the allocation based on the forecast exactly meets the observed need in location 1, but it leaves 6 units of need unmet in location 2.
 However, working within the resource constraint, no other allocation could have done better: for example, allocating 0 units of resources to location 1 and 5 to location 2 still results in a total unmet need of 6 across both locations.
@@
 
-These scores illustrate a general result: allocation scores for a forecast will tend to be larger when the resource constraint is close to the observed demand, because this is when it matters most which locations are allocated more or less resources.
+These scores illustrate a general result: allocation scores for a forecast will tend to be larger when the resource constraint is close to the observed need, because this is when it matters most which locations are allocated more or less resources.
 If the resource constraint is very small, any allocation of those limited resources will result in a large amount of unmet need.
-If the resource constraint is very large, it becomes less important which locations receive relatively more or less resources because all locations will receive enough resources to meet their demand.
+If the resource constraint is very large, it becomes less important which locations receive relatively more or less resources because all locations will receive enough resources to meet their need.
 
-\paragraph{Example 2} Now consider a different forecast that also has exponential distributions for resource demand in each location, but that has the scale parameters $\sigma_1 = 2$ and $\sigma_2 = 8$, twice as large as the scale parameters of the forecast in Example 1.
+\paragraph{Example 2} Now consider a different forecast that also has exponential distributions for resource need in each location, but that has the scale parameters $\sigma_1 = 2$ and $\sigma_2 = 8$, twice as large as the scale parameters of the forecast in Example 1.
 Because the optimal allocation is proportional to the scale parameters, this forecast would lead to the same optimal allocations as the forecast in Example 1, and would therefore be assigned the same allocation score.
@@
 It was not directly directly due to individual misalignments of the marginal forecasts $F_i$ with the observed needs, but rather because the allocations and observed needs were not proportional as vectors.
-Restating: as far as allocation decisions are concerned, with a fixed constraint $K=10$ the fundamental problem with the forecast $F$ in Example 1 is not that it predicts a mean total resource demand of $5$ units; it is that the realized demand was 10 times as large in location 2 as in location 1, but the forecast only indicated that the resource allocation for location 2 should be 4 times the allocation for location 1.
+Restating: as far as allocation decisions are concerned, with a fixed constraint $K=10$ the fundamental problem with the forecast $F$ in Example 1 is not that it predicts a mean total resource need of $5$ units; it is that the realized need was 10 times as large in location 2 as in location 1, but the forecast only indicated that the resource allocation for location 2 should be 4 times the allocation for location 1.
 
-This illustrates a fundamental property of the allocation score: at its core, it measures whether the forecast accurately captures the relative magnitudes of resource demand across different locations, which is precisely the information that is needed to allocate resources to those locations subject to a fixed resource constraint.
-On the other hand, the allocation score is not directly sensitive to whether the forecasts in each location correctly capture the magnitude of resource demand in each individual location.
+This illustrates a fundamental property of the allocation score: at its core, it measures whether the forecast accurately captures the relative magnitudes of resource need across different locations, which is precisely the information that is needed to allocate resources to those locations subject to a fixed resource constraint.
+On the other hand, the allocation score is not directly sensitive to whether the forecasts in each location correctly capture the magnitude of resource need in each individual location.
 %For a given probability level $\tau$, if the quantiles $Q_i(\tau)$ of the marginal forecasts are proportional to the observed needs $y_i$, then the allocation score is zero for the resource constraint level $K = \sum Q_i(\tau)$.
-This stands in marked contrast to other common scoring methods for multivariate forecasts that aggegate univariate scores such as CRPS or WIS for the marginal forecasts where a 
+This stands in marked contrast to other common scoring methods for multivariate forecasts that aggregate univariate scores such as CRPS or WIS for the marginal forecasts where a 
 misalignment in one coordinate is penalized regardless of alignments in other coordinates.
@@
 
-\subsection{The decision-theoretic setup for forecast evaluation}
-\label{sec:methods.decisiontheory}
+\subsection{A decision-theoretic development of the allocation score}
+\label{sec:methods.detailed}
 
-In this section, we give an overview of the decision-theoretic setup for developing proper scoring rules that measure the value of a forecast as an input to decision making.
+We give a high-level review of a general procedure for developing proper scoring rules that are tailored to specific decision-making tasks in section \ref{sec:methods.detailed.decisiontheory}, and then in section \ref{sec:methods.detailed.allocation} we apply that procedure to develop the allocation score based on the task of deciding on how to allocate a fixed supply of resources across multiple locations.
+
+\subsubsection{The decision-theoretic setup for forecast evaluation}
+\label{sec:methods.detailed.decisiontheory}
+
+%In this section, we give an overview of the decision-theoretic setup for developing proper scoring rules that measure the value of a forecast as an input to decision making.
 We keep the discussion here at a somewhat informal level; we refer the reader to [some subset of Brehmer and Gneiting; Grünwald and Dawid; Dawid; Granger and Pesaran 2000; Granger and Machina 2006; Ehm et al.
@@
 This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss function.
-Subject to certain technical conditions, scoring rules obtained from this procedure are proper (cite cite).
+Subject to certain technical conditions, scoring rules obtained from this procedure are proper.
+We refer the reader to [some subset of Brehmer and Gneiting; Grünwald and Dawid; Dawid; Granger and Pesaran 2000; Granger and Machina 2006; Ehm et al.
+2016] for more technically precise discussion.
 
@@
 
-\subsection{A review of quantile score, CRPS, and the weighted interval score}
-\label{sec:methods.quantileloss}
+% \subsection{A review of quantile score, CRPS, and the weighted interval score}
+% \label{sec:methods.quantileloss}
 
-We review how the quantile score arises from a particular decision-making problem in section \ref{sec:methods.quantileloss.quantile}, and how CRPS can be obtained by integrating across values of the parameters of that decision-making problem, as well as the connection to WIS, in section \ref{sec:methods.quantileloss.crps}.
+% We review how the quantile score arises from a particular decision-making problem in section \ref{sec:methods.quantileloss.quantile}, and how CRPS can be obtained by integrating across values of the parameters of that decision-making problem, as well as the connection to WIS, in section \ref{sec:methods.quantileloss.crps}.
 These results have been thoroughly discussed in the literature [cite cite cite].
 
-\subsubsection{Decision-theoretic origins of the quantile score}
-\label{sec:methods.quantileloss.quantile}
+% \subsubsection{Decision-theoretic origins of the quantile score}
+% \label{sec:methods.quantileloss.quantile}
 
-Suppose that a decision-maker is tasked with determining the quantity $x$ of a protective measure to procure; for example, $x$ might represent the number of hospital beds or amount of medicine to purchase.
+% Suppose that a decision-maker is tasked with determining the quantity $x$ of a protective measure to procure; for example, $x$ might represent the number of hospital beds or amount of medicine to purchase.
 Additionally, suppose that each unit of this good has cost $C$ so that the total cost of procurement is $Cx$.
@@
 
-\paragraph{Step 1: specify a loss function.} Combining the cost of procuring goods at level $x$ with losses due to unmet need, we arrive at the overall loss function
-\begin{align}
-s_Q(x,y; C, L) = Cx + L(x-y)_-.
-\end{align}
-Here, $(x - y)_- := \max(-(x - y),0)$ is $0$ if the amount procured, $x$, is greater than or equal to the realized demand $y$; otherwise, it is $y - x$, the amount of unmet need.
+% \paragraph{Step 1: specify a loss function.} Combining the cost of procuring goods at level $x$ with losses due to unmet need, we arrive at the overall loss function
+% \begin{align}
+% s_Q(x,y; C, L) = Cx + L(x-y)_-.
+% \end{align}
+% Here, $(x - y)_- := \max(-(x - y),0)$ is $0$ if the amount procured, $x$, is greater than or equal to the realized demand $y$; otherwise, it is $y - x$, the amount of unmet need.
 
-\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} It can be shown that under the loss function $s_Q$, the Bayes act is a quantile of the forecast distribution at the probability level $\alpha = 1 - C/L$:
-\begin{align}
-x^F = F^{-1}(\alpha).
-\end{align}
-See the supplement for a verification of this result.
-%In particular, faced with the classical binary decision problem of whether to recommend an additional unit of protection given a current level of protection $x$, the forecaster's optimal decision rule under $s_m$ is to recommend adding protection if $x < F^{-1}\left(1-\frac{C}{L}\right)$, that is, if
-%\begin{align}
-%1-F(x) = \mathbb{P}_F\{y>x\} > \frac{C}{L},
-%\end{align}
-%the \emph{cost-loss ratio} of the problem.
+% \paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} It can be shown that under the loss function $s_Q$, the Bayes act is a quantile of the forecast distribution at the probability level $\alpha = 1 - C/L$:
+% \begin{align}
+% x^F = F^{-1}(\alpha).
+% \end{align}
+% See the supplement for a verification of this result.
+% %In particular, faced with the classical binary decision problem of whether to recommend an additional unit of protection given a current level of protection $x$, the forecaster's optimal decision rule under $s_m$ is to recommend adding protection if $x < F^{-1}\left(1-\frac{C}{L}\right)$, that is, if
+% %\begin{align}
+% %1-F(x) = \mathbb{P}_F\{y>x\} > \frac{C}{L},
+% %\end{align}
+% %the \emph{cost-loss ratio} of the problem.
 
-\paragraph{Step 3: Define the scoring rule.} Following the procedures outlined above, we could score the forecast distribution $F$ with the scoring rule
-\begin{align*}
-S_Q(F, y; C, L) &= s_Q(x^F, y; C, L) = Cx^F + L(x^F-y)_- \nonumber \\
-&= C F^{-1}(\alpha) + L \left(F^{-1}(\alpha) - y\right)_- \label{eqn:S_Q_CL}
-\end{align*}
+% \paragraph{Step 3: Define the scoring rule.} Following the procedures outlined above, we could score the forecast distribution $F$ with the scoring rule
+% \begin{align*}
+% S_Q(F, y; C, L) &= s_Q(x^F, y; C, L) = Cx^F + L(x^F-y)_- \nonumber \\
+% &= C F^{-1}(\alpha) + L \left(F^{-1}(\alpha) - y\right)_- \label{eqn:S_Q_CL}
+% \end{align*}
 
-We have set up the problem here in terms of the cost and loss parameters $C$ and $L$, which has the benefit of an intuitive connection to the decision-making context.
+% We have set up the problem here in terms of the cost and loss parameters $C$ and $L$, which has the benefit of an intuitive connection to the decision-making context.
 However, to clarify the connection to the usual notation for the quantile loss, we can divide the loss function $s_Q$ by $L$ to obtain an expression in terms of only $\alpha$:
-\begin{align*}
-s_Q(x,y; \alpha) &= s_Q(x,y; C, L)/L \\
-&= (C/L)x + (x-y)_- \\
-&= (1 - \alpha)x + (x-y)_-.
-\end{align*}
-Because these loss functions are equal up to a constant of proportionality, the Bayes act is the same for both. The associated quantile scoring rule expressed in terms of $\alpha$ is
-\begin{equation}
-S_Q(F, y; \alpha) = (1 - \alpha)F^{-1}(\alpha) + \left(F^{-1}(\alpha) - y\right)_-. \label{eqn:S_Q_alpha}
-\end{equation}
-In either formulation, the key observation is that the Bayes act is the quantile of the forecast distribution $F$ at the probability level given by one minus the cost/loss ratio $C/L$.
+% \begin{align*}
+% s_Q(x,y; \alpha) &= s_Q(x,y; C, L)/L \\
+% &= (C/L)x + (x-y)_- \\
+% &= (1 - \alpha)x + (x-y)_-.
+% \end{align*}
+% Because these loss functions are equal up to a constant of proportionality, the Bayes act is the same for both. The associated quantile scoring rule expressed in terms of $\alpha$ is
+% \begin{equation}
+% S_Q(F, y; \alpha) = (1 - \alpha)F^{-1}(\alpha) + \left(F^{-1}(\alpha) - y\right)_-. \label{eqn:S_Q_alpha}
+% \end{equation}
+% In either formulation, the key observation is that the Bayes act is the quantile of the forecast distribution $F$ at the probability level given by one minus the cost/loss ratio $C/L$.
 
-\subsubsection{CRPS as an integrated quantile score}
-\label{sec:methods.quantileloss.crps}
+% \subsubsection{CRPS as an integrated quantile score}
+% \label{sec:methods.quantileloss.crps}
 
-The scoring rule $S_Q$ of Equation \eqref{eqn:S_Q_alpha} evaluates the forecast distribution $F$ only through its $\alpha$ quantile.
+% The scoring rule $S_Q$ of Equation \eqref{eqn:S_Q_alpha} evaluates the forecast distribution $F$ only through its $\alpha$ quantile.
 While this is faithful to the context of the decision-making problem, it may not be satisfying as a measure of the quality of the full forecast distribution.
@@
 This yields a weighted CRPS:
-\begin{align*}
-S_{CRPS}(F, y; p) &= \int S_Q(F, y; \alpha) p(\alpha) \, d\alpha.
-\end{align*}
-This weighted form of CRPS has appeared in the literature before, e.g.
+% \begin{align*}
+% S_{CRPS}(F, y; p) &= \int S_Q(F, y; \alpha) p(\alpha) \, d\alpha.
+% \end{align*}
+% This weighted form of CRPS has appeared in the literature before, e.g.
 see \cite{gneiting2011weightedScoringRules}.
@@
 
-We emphasize that because $\alpha = 1 - C/L$, the distribution $p(\alpha)$ can be interpreted as expressing incomplete knowledge about the cost/loss ratio in the decision-making problem.
+% We emphasize that because $\alpha = 1 - C/L$, the distribution $p(\alpha)$ can be interpreted as expressing incomplete knowledge about the cost/loss ratio in the decision-making problem.
 The equal weighting used by the ordinary CRPS may be appropriate in the absence of any knowledge about the context in which forecasts will be used to support decision-making, but may be inappropriate if more information is known about the cost/loss ratio for a specific decision-making task.
 
-The weighted interval score (WIS) is often used when the full forecast distribution $F$ is not available, as in the U.S.
+% The weighted interval score (WIS) is often used when the full forecast distribution $F$ is not available, as in the U.S.
 COVID-19 Forecast Hub and similar efforts where forecasts are represented by a collection of prediction intervals.
@@
 
-\subsection{The allocation score}
-\label{sec:methods.allocation}
+% \subsection{The allocation score}
+% \label{sec:methods.allocation}
 
-\elr{If we like the organization of the previous subsection into two subsubsections about the quantile score and the CRPS, we should replicate that here.}
+%\elr{If we like the organization of the previous subsection into two subsubsections about the quantile score and the CRPS, we should replicate that here.}
 
-We now develop a scoring rule for probabilistic forecasts that measures the value of a forecast as an input to decision making about how to allocate limited resources to meet demand across multiple locations.
-As a concrete example, we take the resource to be a good such as ventilators or oxygen supply.
+% We now give a more formal derivation of the allocation score starting from the set up of a decision making problem about how to allocate limited resources to meet demand across multiple locations.
+In section \ref{sec:methods.allocation.special}, we consider a setting where the resource constraint is known, which will result in a score that is specialized to a particular quantile of the forecasts.
+Then in sction \ref{sec.methods.allocation.integrated}, we allow for the possibility of decision making under uncertainty about the precise value of the resource constraint, which will lead to averaging the specialized allocation score across multiple quantile levels.
+
+%As a concrete example, we take the resource to be a good such as ventilators or oxygen supply.
 An administrator is tasked with determining where to send these resources so as to meet demand among hospital patients in different facilities or states.
@@
 \subsubsection{The allocation score for a fixed resource constraint}
+\label{sec:methods.detailed.specific_allocation}
 
-In the decision-making setting that we consider, an action $\mathbf{x} = (x_1, \ldots, x_n)$ is a vector specifying the amount that is allocated to each of $n$ locations.
-We require that each $x_i$ is non-negative and that the total allocation across all locations does not exceed a constraint $K$ on the total available resources: $\sum_{i=1}^n x_i \leq K$.
-\elr{Since we got rid of the C parameter, maybe we should just make this a hard constraint here, $\sum_i x_i = K$?} The set $\mathcal{X}$ collects all possible allocations that satisfy these constraints.
-The eventually realized resource demand in each location is denoted by $\mathbf{y} = (y_1, \ldots, y_n)$.
-Again, these levels of demand are not known at the time of decision making, so we define the random vector $Y = (Y_1, \ldots, Y_n)$ where $Y_i$ represents the as-yet-unknown level of resource demand in location $i$.
-Forecasts of demand in each location are collected in $F = (F_1, \ldots, F_n)$.
-We assume that the forecasts do not allow for the possibility of negative demand, i.e.
+In the decision-making setting that we consider, an action $\mathbf{x} = (x_1, \ldots, x_N)$ is a vector specifying the amount that is allocated to each of $N$ locations.
+We require that each $x_i$ is non-negative and that the total allocation across all locations equals the amount of available resources, $K$: $\sum_{i=1}^N x_i = K$.
+The set $\mathcal{X}$ consists of all possible allocations that satisfy these constraints.
+The eventually realized resource need in each location is denoted by $\mathbf{y} = (y_1, \ldots, y_N)$.
+These levels of need are not known at the time of decision making, so we define the random vector $Y = (Y_1, \ldots, Y_N)$ where $Y_i$ represents the as-yet-unknown level of resource need in location $i$.
+Forecasts of need in each location are collected in $F = (F_1, \ldots, F_N)$.
+We assume that resource need is non-negative and the forecasts reflect that, i.e.
 the support of each $F_i$ is a subset of $\mathbb{R}^+$.
-As in the previous section, we assume that a loss $L$ is incurred for each unit of unmet need.
+Finally, we assume that each unit of unmet need incurs a loss denoted by $L$, so that if the selected resource level $x_i$ in location $i$ is less than the realized need $y_i$, a loss of $L(y_i - x_i)$ results.
 
-We note that a number of generalizations to this loss specification have been formulated in the literature, including an allowance for costs for over-allocation to a particular unit (e.g.
-if there are storage costs for unused resources), differing losses different units (e.g.
-if a unit of unmet demand imposes more severe costs in one location than another), and the introduction of a convex function that controls the rate at which costs accrue depending on the scale of need.
-We consider these and other generalizations in other work.
+We note that the problem formulation here assumes that the resource in question does not impact the amount of demand $y_i$ that will materialize, but rather it is a resource that satisfies that demand.
+In the context of infectious disease, this means that we do not consider resources such as vaccine doses or masks, which are intended to reduce the number of people who will become sick.
+Our set up addresses resources like hospital beds, oxygen supply, or ventilators, which are intended to meet the medical needs of patients who are sick.
 
-With this notation in place, we can develop a proper scoring rule following the outline in section \ref{sec:methods.decisiontheory}.
+With this notation in place, we can develop a proper scoring rule following the outline in section \ref{sec:methods.detailed.decisiontheory}.
 
-\paragraph{Step 1: specify a loss function.} The loss associated with a particular allocation is calculated by summing contributions from unmet demand in each location:
+\paragraph{Step 1: specify a loss function.} The loss associated with a particular allocation is calculated by summing contributions from unmet need in each location:
 \begin{equation}
-s_A(\mathbf{x}, \mathbf{y}; L) = \sum_{i=1}^n L (x_i - y_i)_-. \label{eqn:loss_fn}
+s_A(\mathbf{x}, \mathbf{y}; L) = \sum_{i=1}^N L (y_i - x_i)_+. \label{eqn:loss_fn}
 \end{equation}
-As in the previous section, $(x_i - y_i)_-$ is $0$ if the amount $x_i$ allocated to unit $i$ is greater than or equal to the realized demand $y_i$ in that location; otherwise, it is $y_i - x_i$, the amount of unmet need in that location.
+Here, $(y_i - x_i)_+$ is the unmet need in location $i$, which is given by $y_i - x_i$ if the realized need $y_i$ in location $i$ is greater than the amount $x_i$ allocated to that location, or $0$ if the amount $x_i$ allocated to unit $i$ is greater than or equal to the realized need.
 
-\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act is the allocation that minimizes the expected loss:
+\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the forecast, $\mathbf{x}^F$, is the allocation that minimizes the expected loss:
 \begin{align}
-    \mathbf{x}^F &= \underset{\mathbf{x}\in \mathbb{R}^N, 0 \leq \mathbf{x}}{\mathrm{arg min}}\,\, \sbar_A^F(\mathbf{x}; L) \text{ subject to }
-    \sum_{i=1}^N x_i \leq K \text{, where} \label{AP} \\
-    \sbar_A^F(\mathbf{x}; L) &= \mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y}; L)] = \sum_{i=1}^{N} \mathbb{E}_{F_i}[s_A(x_i, Y_i; L)] \label{eqn:sbar}
+    \mathbf{x}^F &= \mathrm{arg min}\,\, \sbar_A^F(\mathbf{x}; L) \text{ subject to }
+    0 \leq x_i \text{ for all $i$ and} \, \sum_{i=1}^N x_i = K \text{, where} \label{AP} \\
+    \sbar_A^F(\mathbf{x}; L) &= \mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y}; L)] = \sum_{i=1}^{N} L \mathbb{E}_{F_i}[(Y_i - x_i)_+] \label{eqn:sbar}
 \end{align}
@@
 This partial solution to the allocation problem seems to have first appeared in \cite{hadleywhitin1963}; see the supplemental materials for a derivation.
-Figure *** panel (a) \elr{TODO: first figure from notes crps\_connection\_exp.Rmd} illustrates the expected loss function $\sbar_A^F$ and the allocation given by the Bayes act in a simple example with $n = 2$ locations, $L = 1$, $K = 5$, and forecasts $Y_1 \sim \text{Exp}(1 / \sigma_1)$ and $Y_2 \sim \text{Exp}(1 / \sigma_2)$, using a scale parameterization of the exponential distribution with $\sigma_1 = 1$ and $\sigma_2 = 5$.
+%Figure *** panel (a) \elr{TODO: first figure from notes crps\_connection\_exp.Rmd} illustrates the expected loss function $\sbar_A^F$ and the allocation given by the Bayes act in a simple example with $n = 2$ locations, $L = 1$, $K = 5$, and forecasts $Y_1 \sim \text{Exp}(1 / \sigma_1)$ and $Y_2 \sim \text{Exp}(1 / \sigma_2)$, using a scale parameterization of the exponential distribution with $\sigma_1 = 1$ and $\sigma_2 = 5$.
 
@@
 The quantile is at a probability level $\alpha = 1 - \lambda^{\star}/L$ that is the same for all locations, and is chosen such that the constraint is satisfied.
+According to this interpretation, we divide the available resources among the locations in such a way that the forecasted probability that need will exceed demand is the same in all locations.
 An alternative interpretation comes from noting that for each location $i$, $\frac{\partial}{\partial x_i} \sbar_A^F(\mathbf{x}) \vert_{\mathbf{x}=\mathbf{x}^F} = \lambda^{\star}$ (see the supplement for a proof).
@@
 
-\apg{
-first attempt by APG to rephrase last paragraph: The central mathematical ideas in this construction are that
-\begin{itemize}
-\item optimization under a \emph{single} constraint requires the rates of change of (a probabilistic forecast's) expected benefit with respect to our decision variables, $x_i$, to be the same for all locations
-\item these rates of change can be identified with probabilities
-\item and therefore the Bayes act results from using a \emph{shared} probability level, $1 - \lambda^{\star}/L$, to determine allocations as the corresponding quantiles of the location-specific forecast distributions $F_i$ which satisfy the constraint.
-\end{itemize}
-(See the supplement for detailed discussion and derivations of these points.)
-}
+% ELR: I made some more edits to the previous paragraph, but still may not have captured everything that Aaron think is important here.
+% \apg{
+% first attempt by APG to rephrase last paragraph: The central mathematical ideas in this construction are that
+% \begin{itemize}
+% \item optimization under a \emph{single} constraint requires the rates of change of (a probabilistic forecast's) expected benefit with respect to our decision variables, $x_i$, to be the same for all locations
+% \item these rates of change can be identified with probabilities
+% \item and therefore the Bayes act results from using a \emph{shared} probability level, $1 - \lambda^{\star}/L$, to determine allocations as the corresponding quantiles of the location-specific forecast distributions $F_i$ which satisfy the constraint.
+% \end{itemize}
+% (See the supplement for detailed discussion and derivations of these points.)
+% }
 
@@
 \begin{equation}
-S_A(F, y; L, K) = s_A(\mathbf{x}^F, y; L) = \sum_{i=1}^n L (F_i^{-1}(1 - \lambda^{\star}/L) - y_i)_-
+S_A(F, y; L, K) = s_A(\mathbf{x}^F, y; L) = \sum_{i=1}^n L (y_i - F_i^{-1}(1 - \lambda^{\star}/L))_+
 \end{equation}
@@
 
-As with the quantile score $S_Q$, the allocation score $S_A$ measures the skill of the forecast distributions $F$ based on a single probability level $\alpha$.
-By analogy to the method for obtaining CRPS by integrating the quantile score, we develop an \emph{integrated allocation score} (IAS) \elr{open to better names} that integrates the allocation score across values of the problem parameters, weighting by a distribution $p$:
+\subsection{Integrating the allocation score across resource constraint levels}
+\label{sec:methods.detailed.integrated_allocation}
+
+\elr{I did a little work here but did not really finish it off. Can we just integrate across $K$? And maybe there's a better way to set this up.  Aiming for just 1-2 paragraphs, though.}
+
+The allocation score $S_A$ that we developed in the previous section measures the skill of the forecast distributions $F$ based on a single probability level $\alpha = 1 - \lambda^{\star}/L$.
+This is appropriate if the resource constraint $K$ is a known constant.
+However, if $K$ is not precisely known at the time of decision-making or there is interest in measuring the value of forecasts across a range of decision-making scenarios, we can use an \emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a distribution $p$:
 $$S_{IAS}(F, y) = \int S_A(F,y; L, K) p(L, K) \, dL dK$$
 
-We illustrate the relationship between the IAS and the CRPS in our example with two locations and forecasts given by $\text{Exp}(1/\sigma_i)$ distributions with $\sigma_1 = 1$ and $\sigma_2 = 5$.
+%We illustrate the relationship between the IAS and the CRPS in our example with two locations and forecasts given by $\text{Exp}(1/\sigma_i)$ distributions with $\sigma_1 = 1$ and $\sigma_2 = 5$.
 Note that the quantile functions corresponding to these forecasts are given by $F_i^{-1}(\alpha) = -\sigma_i \log(1 - \alpha)$.
@@
 
-We use this link to explore the relationship between IAS and CRPS from two directions.
+%We use this link to explore the relationship between IAS and CRPS from two directions.
 First, suppose the decision-maker has some uncertainty about the value of $K$, which they express through $p$.
@@
 
-Figure *** panel (d) \elr{TODO, figures from crps\_connection\_exp} illustrates this link by going in the other direction: given a forecast $F$, we exhibit the distribution on $K$ that would lead to equally weighted CRPS.
+%Figure *** panel (d) \elr{TODO, figures from crps\_connection\_exp} illustrates this link by going in the other direction: given a forecast $F$, we exhibit the distribution on $K$ that would lead to equally weighted CRPS.
 Now we use the fact that $K$ can be written as a function of $\alpha$ to obtain the distribution on $K$ that corresponds to a $\text{Uniform}(0,1)$ distribution on $\alpha$.
@@
 
-\subsection{Solving the Allocation Problem}
+\subsection{Generalizations and Connections to Other Scores}
+\label{sec:methods.related}
 
-As the solution to a constrained optimization problem, the optimal allocation $\bf{x}^F$ according to a forecast $F$ can be described formally as the first component of a stationary point $(\bf{x}^F, \lambda^{\star})$ in the region $\{\mathbf{x} \geq 0, \lambda \geq 0\}$ for the Lagrangian
-\begin{align}
-\mathcal{L}(\mathbf{x},\lambda; K, L) = \sum_{i=1}^{N} \mathbb{E}_{F_i}[s_A(x_i, Y_i; L)] + \lambda\left(\sum_{i=1}^{N} x_i - K\right).
-\end{align}
-That is, optimality requires that both
-\begin{align}
-0 = \nabla_{x} \mathcal{L}(\mathbf{x}^F, \lambda) &= \sum_{i=1}^{N} E_F \left[\frac{d}{dx_i}(Y_i - x_i^F)_{+}\right] + \lambda \\
-& = \sum_{i=1}^{N} E_F \left[-\mathbf{1}\{Y_i \geq x_i^F\}\right] + \lambda  = \sum_{i=1}^{N} F_i(x_i^F)-1 + \lambda  \label{eqn:x_stationary}\\
-\text{ and} \\
-0 = \nabla_{\lambda} \mathcal{L}(\mathbf{x}^F, \lambda^{\star}) &= \sum_{i=1}^{N} x_i^F - K\\
-& = \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star}) - K \quad \text{ (by \eqref{eqn:x_stationary})}. \label{eqn:lambda_stationary}
-\end{align}
-These equations give formulae $x_i^F = F_i^{-1}(1-\lambda^{\star})$, but in many situations of interest for us the $F_i^{-1}$ are not single-valued, and even 
-when they are, a solution $\lambda^{\star}$ of \eqref{eqn:lambda_stationary} can only be obtained analytically in special cases (but see \ref{sec:methods.ls.example}).
+We have described how the allocation score arises from a standard procedure for developing proper scoring rules when the decision relates to how a fixed quantity of resources should be allocated to multiple locations.
+In fields such as meteorology and supply chain management, a great deal of attention has been given to a different problem where a decision must be made about the quantity of a resource to purchase for a single location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each unit of unmet need.
+This decision-making problem leads to the pinball loss that is often used in quantile regression, and the optimal quantity of resources to purchase is the predictive quantile at the probability level $1 - \frac{C}{L}$.
+
+This result bears a strong resemblance to the results for the allocation problem, where the optimal solution also sets the resource level in each location to a quantile of the respective forecast distributions.
+These decision-making problems differ according to the challenge faced by the decision maker: a fixed constraint on the available resources for the allocation problem, or a cost per unit of resources in the resource purchasing problem.
+In fact, it is possible to combine these into a more general problem where the decision-maker must decide on both the total level of resources to purchase and an allocation of those resources across multiple locations, subject to a cost per unit of resources and a constraint on the total resources that can be purchased.
+The quantile loss and the allocation score presented in this work are both be special cases of the score that emerges from that more general decision-making problem.
+We pursue this direction further in other work (cite arxiv version of paper 2).
+
+Similarly, it is possible to see both the integrated allocation score and common scores such as the  continuous ranked probability score (CRPS) and weighted interval score (WIS) as special cases of a more general score that integrates across resource cost and constraint parameters together.
+\elr{previous sentence may be controversial -- how to frame?} It is well known in the literature that WIS is equivalent to a weighted average of quantile scores across multiple quantile levels (cite Bracher et al.) and CRPS integrates quantile scores across multiple quantile levels (cite cite).
+By integrating the combined resource acquisition and allocation score that we described in the previous paragraph across the parameters for both resource costs and constraints, we obtain a score that combines features of both IAS and CRPS and contains those scores as special cases.
+Again, we defer a more detailed discussion of these results to (cite arxiv version of paper 2).
+
+%Finally, we note that a number of generalizations to the loss specification that we introduced for the allocation problem  have been formulated in the literature, including an allowance for costs for over-allocation to a particular unit (e.g.
+if there are storage costs for unused resources), differing losses different units (e.g.
+if a unit of unmet need imposes more severe costs in one location than another), and the introduction of a convex function that controls the rate at which costs accrue depending on the scale of need.
+We consider these and other generalizations in other work.
+
+% \subsection{Solving the Allocation Problem}
+
+% As the solution to a constrained optimization problem, the optimal allocation $\bf{x}^F$ according to a forecast $F$ can be described formally as the first component of a stationary point $(\bf{x}^F, \lambda^{\star})$ in the region $\{\mathbf{x} \geq 0, \lambda \geq 0\}$ for the Lagrangian
+% \begin{align}
+% \mathcal{L}(\mathbf{x},\lambda; K, L) = \sum_{i=1}^{N} \mathbb{E}_{F_i}[s_A(x_i, Y_i; L)] + \lambda\left(\sum_{i=1}^{N} x_i - K\right).
+% \end{align}
+% That is, optimality requires that both
+% \begin{align}
+% 0 = \nabla_{x} \mathcal{L}(\mathbf{x}^F, \lambda) &= \sum_{i=1}^{N} E_F \left[\frac{d}{dx_i}(Y_i - x_i^F)_{+}\right] + \lambda \\
+% & = \sum_{i=1}^{N} E_F \left[-\mathbf{1}\{Y_i \geq x_i^F\}\right] + \lambda  = \sum_{i=1}^{N} F_i(x_i^F)-1 + \lambda  \label{eqn:x_stationary}\\
+% \text{ and} \\
+% 0 = \nabla_{\lambda} \mathcal{L}(\mathbf{x}^F, \lambda^{\star}) &= \sum_{i=1}^{N} x_i^F - K\\
+% & = \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star}) - K \quad \text{ (by \eqref{eqn:x_stationary})}. \label{eqn:lambda_stationary}
+% \end{align}
+% These equations give formulae $x_i^F = F_i^{-1}(1-\lambda^{\star})$, but in many situations of interest for us the $F_i^{-1}$ are not single-valued, and even 
+% when they are, a solution $\lambda^{\star}$ of \eqref{eqn:lambda_stationary} can only be obtained analytically in special cases (but see \ref{sec:methods.ls.example}).
 We therefore typically calculate allocation scores using approximately optimal allocations found via a binary search for $\lambda^{\star}$.
 This search begins with the interval $\lambda \in [0,1]$ and
-checks at each step the sign of the inequality that results from substituting the interval midpoint $\lambda_{\tau}$ into \eqref{eqn:lambda_stationary} for endpoints of the sets $F_i^{-1}(1-\lambda_{\tau})$.
+% checks at each step the sign of the inequality that results from substituting the interval midpoint $\lambda_{\tau}$ into \eqref{eqn:lambda_stationary} for endpoints of the sets $F_i^{-1}(1-\lambda_{\tau})$.
 If $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) > K$, the quantiles for $1-\lambda_{\tau}$ exceed our resource constraint so we set $\lambda_{\tau}$ as the lower endpoint of the $\tau + 1$ search interval.
 Similarly, if $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) < K$, the quantiles for
-$1-\lambda_{\tau}$ do not make full use of our resource and we set $\lambda_{\tau}$ as the upper endpoint of the $\tau + 1$ search interval.
+% $1-\lambda_{\tau}$ do not make full use of our resource and we set $\lambda_{\tau}$ as the upper endpoint of the $\tau + 1$ search interval.
 This process halves the search interval width at each step and we continue until this width falls below a given tolerance level $\varepsilon_{\tau}$ yielding an
-approximation $\lambda^{\star, \varepsilon}$. When quantiles $F_i^{-1}(1-\lambda^{\star, \varepsilon})$ are singly defined and the forecast densities $f_i$
-not too close to zero at these quantiles, the discrepancy $\Delta(\lambda^{\star, \varepsilon})  = \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star, \varepsilon}) - K$ is of order $\varepsilon_{\lambda}$ and we can take $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ as approximations to $x_i^F$.
+% approximation $\lambda^{\star, \varepsilon}$. When quantiles $F_i^{-1}(1-\lambda^{\star, \varepsilon})$ are singly defined and the forecast densities $f_i$
+% not too close to zero at these quantiles, the discrepancy $\Delta(\lambda^{\star, \varepsilon})  = \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star, \varepsilon}) - K$ is of order $\varepsilon_{\lambda}$ and we can take $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ as approximations to $x_i^F$.
 
-Often though, $1-\lambda^{\star, \varepsilon}$ will be the probability level of a ``plateau'' for one or more $F_i$ and the $\lambda$ search algorithm will fail to drive $\Delta(\lambda^{\star, \varepsilon})$ close enough to zero for the $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ to be acceptable approximations of the optimal allocations.
+% Often though, $1-\lambda^{\star, \varepsilon}$ will be the probability level of a ``plateau'' for one or more $F_i$ and the $\lambda$ search algorithm will fail to drive $\Delta(\lambda^{\star, \varepsilon})$ close enough to zero for the $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ to be acceptable approximations of the optimal allocations.
 In this event we follow-up with a ``post-processing'' step that identifies lower and upper boundaries $x_{i,L}$ and $x_{i,U}$ of any
-such low-density plateau containing points in $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$. If $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ is single-valued 
-with the density $f_i(F_{i}^{-1}(1-\lambda^{\star, \varepsilon}))$ not close to zero, we simply take $x_{i,L} = F_{i}^{-1}(1-\lambda^{\star, \varepsilon}) = x_{i,U}$.
+% such low-density plateau containing points in $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$. If $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ is single-valued 
+% with the density $f_i(F_{i}^{-1}(1-\lambda^{\star, \varepsilon}))$ not close to zero, we simply take $x_{i,L} = F_{i}^{-1}(1-\lambda^{\star, \varepsilon}) = x_{i,U}$.
 Defining $\mathbf{x}_L = (x_{i,L})$ and $\mathbf{x}_U = (x_{i,U})$ we then interpolate an approximately optimal allocation
-$\mathbf{x} = (1-t^{\star})\mathbf{x}_L + t^{\star}\mathbf{x}_U$ where $t^{\star}$ solves 
-$\sum_{i=1}^{N }(1-t^{\star})x_{i,L} + t^{\star}x_{i,U} = K$.
+% $\mathbf{x} = (1-t^{\star})\mathbf{x}_L + t^{\star}\mathbf{x}_U$ where $t^{\star}$ solves 
+% $\sum_{i=1}^{N }(1-t^{\star})x_{i,L} + t^{\star}x_{i,U} = K$.
 
-We note though that by using this linear interpolation (rather than some other point in the plateau satisfying the constraint) as the allocation associated with $F$ in the scoring process we are making a choice that is not
-guided by any expected utility maximization principle. As such, this choice may imply strong and potentially questionable assumptions about fairness and
-equity principles for the decision maker when forecasts tend to be ``clumped,'' leaving many regions with very low forecasted probability.
+% We note though that by using this linear interpolation (rather than some other point in the plateau satisfying the constraint) as the allocation associated with $F$ in the scoring process we are making a choice that is not
+% guided by any expected utility maximization principle. As such, this choice may imply strong and potentially questionable assumptions about fairness and
+% equity principles for the decision maker when forecasts tend to be ``clumped,'' leaving many regions with very low forecasted probability.
 

### 2023-08-30 alloscore_manuscript/methods_apg.tex
Getting closer to finishing dual lambda search explanation

@@
 \item Specify a \emph{loss function} $s(x,y)$ that quantifies the utility or disutility of the result $r$ of taking action $x$ under outcome $y$.
-\item Map the given probabilistic forecast $F$ for $Y$ to an action $x^F$ which ``solves" the decision problem by minimizing the expected loss $E_{F}[s(x,Y)]$ over all $x \in \mathcal{X}$ under the distribution $F$.
+\item Map the given probabilistic forecast $F$ for $Y$ to an action $x^F$ which ``solves" the decision problem by minimizing the expected loss $E_F\left[s(x,Y)\right]$ over all $x \in \mathcal{X}$ under the distribution $F$.
 Following \cite{dawid2007geometry}, we call $x^F$ the \emph{Bayes act} for $F$.
@@
 \begin{align}
-0 = \dby{}{x} E_F[s_{Q,C}(x,y)] &= E_{F}[\dby{}{x}s_{Q,C}(x,y)] \\
-&= C + LE_{F}[\dby{}{x}(y-x)_+] \\
-&= C - LE_{F}[\mathbf{1}\{y > x\}] \\
+0 = \dby{}{x} E_F\left[s_{Q,C}(x,y)\right] &= E_F\left[\dby{}{x}s_{Q,C}(x,y)\right] \\
+&= C + LE_F\left[\dby{}{x}(y-x)_+\right] \\
+&= C - LE_F\left[\mathbf{1}\{y > x\}\right] \\
 &= C + L(F(x) - 1), \label{eqn:q_deriv}
@@
 So by taking $\mathcal{C} = \{(1 \pm \alpha_j)L/2 \mid j > 0\}\cup \{L/2\}$ and $p_\mathcal{C}$ symmetric across $L/2$, we get 
-$S_{Q,\mathcal{C}, p_C}(F,y) = \mathrm{WIS}_{\alpha_{\{0: J\}}}(F, y) + E_{p_{\mathcal{C}}}[y]$ with $w_j = p_{j}\alpha_j/2$ and $w_0 = p_{L/2}$.
+\begin{align}	
+S_{Q,\mathcal{C}, p_C}(F,y) = \mathrm{WIS}_{\alpha_{\{0: J\}}}(F, y) + y \sum_{j=0}^{J} w_j
+\end{align}	
+with $w_j = p_{j}\alpha_j/2$ and $w_0 = p_{L/2}$.
 
@@
 \begin{align}
-s_{tQ}(x,y) &= \sum_{l=1}^N Cx_l + L(y_l-x_l)_+ \\
-s_{tQ, \mathcal{C}, p_C}(x,y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}p_i(C_i x_{l,i} + L(x_{l,i} - y_l)_+)\\
-s_{tQ, f_C}(x,y) &= \sum_{l=1}^N \int_{0}^{L} (cx_l(c) + L(x_l(c) - y_l)_+) f_C(c)dc.
+s_{TQ}(x,y) &= \sum_{l=1}^N Cx_l + L(y_l-x_l)_+ \label{eqn:TQ_loss} \\
+s_{TQ, \mathcal{C}, p_C}(x,y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}p_i(C_i x_{l,i} + L(x_{l,i} - y_l)_+)\\
+s_{TQ, f_C}(x,y) &= \sum_{l=1}^N \int_{0}^{L} (cx_l(c) + L(x_l(c) - y_l)_+) f_C(c)dc.
 \end{align}
@@
 This adds, for each $k \in \mathbb{R}_{+}$, a component to our overall decision problem that asks the question: If the total amount $K$ available of the resource is insufficient allow the procurements $x_l(C)$ you specify for the realized cost $C$, what alternate procurement levels $x_l^{k,C}$ would you choose?
-The action set for this new component is
+Writing $T_x = \sum_{l=1}^{N} x_l$, the action set for this new component is
 \begin{align}
-\mathcal{X}_k = \{x \in \mathbb{R}_{+}^N \mid \sum_{l=1}^{N} x_l = k\}.
+\mathcal{X}_k = \{x \in \mathbb{R}_{+}^N \mid T_x = k\}.
 \end{align}
@@
 
-\subsubsection*{Old allocation intro}
-We now add a constraint $K$ on the total amount of the resource available for allocation across all locations.
+% \subsubsection*{Old allocation intro}
+% We now add a constraint $K$ on the total amount of the resource available for allocation across all locations.
 This presents the worker with a new decision problem we call the \emph{allocation problem} (AP or $\mathrm{AP}(K)$) with a smaller action set $\mathcal{X}_K = \{x \in \mathbb{R}_{+}^N | \sum x_l \leq K\} \subset \mathcal{X}_{cert}\}$ from which to choose (and analogously restricted version of the sets $\mathcal{X}_{disc}$ and $\mathcal{X}_{cont}$ for the decision problems with uncertain cost).
@@
 But with a view toward developing a scoring rule that accommodates uncertainty about both $C$ and $K$ in a non-redundant (and more interpretable and computationally feasible) way, we augment the action set of the AP to include, in addition to the procurement levels $x_l$ for all locations, a ``reserve'' quantity $z_K$ defined as the available amount of resource that the worker decides not to allocate anywhere.
-In optimization theory quantities defined in this manner are known as \emph{slack} variables, and are introduced in order to re-express contraint inequalities as equalities.  In our case, we get
-\begin{align}
-\mathcal{X}_{K} = \{(x,z) \in \mathbb{R}_{+}^{N+1} \mid \sum_{l=1}^{N}x_l + z = K \}.
-\end{align}
-In these action set variables, the loss function we used for the unconstrained $N$ location decision problem with action set $\mathcal{X}_{cert}$ now has the expression
-\begin{align}
-s_{tQ}\left((x,z), y\right) = C(K-z) + \sum_{l=1}^{N} L(y_l - x_l)_{+}.
-\end{align}
-This isolates -- at least formally -- the explicit contribution of the procurement part, $x$, of the decision to the second term which we call the \emph{allocation loss function}:
-\begin{align}
-s_{A,K}(x,y):= L\sum_{l=1}^{N} (y_l - x_l)_{+}.
-\end{align}
-It will turn out, as explained in the following sections, that for a given forecast $F$, a scoring rule derived with respect to the AP from $s_{tQ}$ under uncertainty about $C$ -- that is, from $s_{tWIS}$ or $s_{tCRPS}$ -- will have the same values for all outcomes $y$ as a scoring rule that is a weighted average or integration of scoring rules derived from allocation losses $s_{A,K}$ with respect the AP's for a range of $K$'s.
+% In optimization theory quantities defined in this manner are known as \emph{slack} variables, and are introduced in order to re-express contraint inequalities as equalities.  In our case, we get
+% \begin{align}
+% \mathcal{X}_{K} = \{(x,z) \in \mathbb{R}_{+}^{N+1} \mid \sum_{l=1}^{N}x_l + z = K \}.
+% \end{align}
+% In these action set variables, the loss function we used for the unconstrained $N$ location decision problem with action set $\mathcal{X}_{cert}$ now has the expression
+% \begin{align}
+% s_{tQ}\left((x,z), y\right) = C(K-z) + \sum_{l=1}^{N} L(y_l - x_l)_{+}.
+% \end{align}
+% This isolates -- at least formally -- the explicit contribution of the procurement part, $x$, of the decision to the second term which we call the \emph{allocation loss function}:
+% \begin{align}
+% s_{A,K}(x,y):= L\sum_{l=1}^{N} (y_l - x_l)_{+}.
+% \end{align}
+% It will turn out, as explained in the following sections, that for a given forecast $F$, a scoring rule derived with respect to the AP from $s_{tQ}$ under uncertainty about $C$ -- that is, from $s_{tWIS}$ or $s_{tCRPS}$ -- will have the same values for all outcomes $y$ as a scoring rule that is a weighted average or integration of scoring rules derived from allocation losses $s_{A,K}$ with respect the AP's for a range of $K$'s.
 
-Fundamental point: 
-\begin{itemize}
-\item When $z_K > 0$, $K$ has no influence on the $\mathrm{AP}(K)$ Bayes act, i.e., $\partial_K x^{F,C,K} = 0$.  
-\item When $z_K=0$ and $x^F$ are not $1-C/L$ quantiles, $\partial_C x^{F,C,K} = 0$.
-\end{itemize}
+% Fundamental point: 
+% \begin{itemize}
+% \item When $z_K > 0$, $K$ has no influence on the $\mathrm{AP}(K)$ Bayes act, i.e., $\partial_K x^{F,C,K} = 0$.  
+% \item When $z_K=0$ and $x^F$ are not $1-C/L$ quantiles, $\partial_C x^{F,C,K} = 0$.
+% \end{itemize}
 
-Picture: For $K$ fixed, $x^{F,K}(c)$ will be a constant point $x^{F,K}$ on $\{\sum x_l = K\}$ for $c \in [0, c_1]$, 
-and then wander back to 0 for $c \in (c_1,L]$. When $C$ is fixed, $x^{F}(k)$ will be a point on $\{\sum x_l = k\}$ (which does not depend on $C$)
-for $k \in [0, \sum F_l^{-1}(1-C/L)]$, and then constant at $\mathbf{F}^{-1}(1-C/L)$ for $k \in (\sum F_l^{-1}(1-C/L), \infty)$.
+% Picture: For $K$ fixed, $x^{F,K}(c)$ will be a constant point $x^{F,K}$ on $\{\sum x_l = K\}$ for $c \in [0, c_1]$, 
+% and then wander back to 0 for $c \in (c_1,L]$. When $C$ is fixed, $x^{F}(k)$ will be a point on $\{\sum x_l = k\}$ (which does not depend on $C$)
+% for $k \in [0, \sum F_l^{-1}(1-C/L)]$, and then constant at $\mathbf{F}^{-1}(1-C/L)$ for $k \in (\sum F_l^{-1}(1-C/L), \infty)$.
 
-\section{Step 3: Use the Bayes act to score the forecast $F$ against outcomes.}
+% \section{Step 3: Use the Bayes act to score the forecast $F$ against outcomes.}
 
-Deal with alloscore by noticing that simultaneous quantile loss has same minimizer as alloscore with sum of quantiles as constraint level.
+% Deal with alloscore by noticing that simultaneous quantile loss has same minimizer as alloscore with sum of quantiles as constraint level.
 
-\begin{align}
-s_{t\mathrm{WIS}}(x^{F,t\mathrm{WIS}},y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}(x_{l,i}^{F_l,\mathrm{Q}, C_i}- y_l)_+  + C_i x_{l,i}^{F_l,\mathrm{Q}, C_i}\\ 
-&= \frac{1}{m} \sum_{i=1}^{m}s_{\mathrm{AS}}(x^{F,\mathrm{AS},  K(C_i, F, L)},y) + C_i K(C_i, F, L)
-\end{align}
+% \begin{align}
+% s_{t\mathrm{WIS}}(x^{F,t\mathrm{WIS}},y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}(x_{l,i}^{F_l,\mathrm{Q}, C_i}- y_l)_+  + C_i x_{l,i}^{F_l,\mathrm{Q}, C_i}\\ 
+% &= \frac{1}{m} \sum_{i=1}^{m}s_{\mathrm{AS}}(x^{F,\mathrm{AS},  K(C_i, F, L)},y) + C_i K(C_i, F, L)
+% \end{align}
 
-\begin{align}
-s_{\mathrm{IAS}}(x^{F, \mathrm{IAS}}(k), y) &= \int_0^{\infty} s_{\mathrm{AS}}(x^{F, \mathrm{AS}(k)}, y) dk \\
-&= \int_0^{\infty} \sum_{l=1}^N s_{\mathrm{Q}}(x^{F_l, \mathrm{Q}(c(k, F))}, y) - c(k, F)kdk \\
-&= \sum_{l=1}^N \int_0^{\infty} s_{\mathrm{Q}}(x^{F_l, \mathrm{Q}(c(k, F))}, y) dk -  \int_0^{\infty} c(k, F)kdk
-\end{align}
+% \begin{align}
+% s_{\mathrm{IAS}}(x^{F, \mathrm{IAS}}(k), y) &= \int_0^{\infty} s_{\mathrm{AS}}(x^{F, \mathrm{AS}(k)}, y) dk \\
+% &= \int_0^{\infty} \sum_{l=1}^N s_{\mathrm{Q}}(x^{F_l, \mathrm{Q}(c(k, F))}, y) - c(k, F)kdk \\
+% &= \sum_{l=1}^N \int_0^{\infty} s_{\mathrm{Q}}(x^{F_l, \mathrm{Q}(c(k, F))}, y) dk -  \int_0^{\infty} c(k, F)kdk
+% \end{align}
 
-$s_{tQ}(x^{F,C,K}) \mid C = c, K = k$ in $\mathrm{AP}(k)$ coordinates:
-\begin{align}
-s_{tQ}(x^{F,c,k})(y) &= 
-\begin{cases}
-\sum_{l=1}^{N}L(y_l - x_l^{F,k})_{+} + ck, & k < \sum F_l^{-1}(1-c/L) \\
-\sum_{l=1}^{N}L(y_l - x_l^{F,Q,c})_{+} + c(k - z_k(c)), & \text{otherwise, with } 
-\end{cases} \\
-z_k(c) &= (k -  \sum F_l^{-1}(1-c/L))_{+}
-\end{align}
+% $s_{tQ}(x^{F,C,K}) \mid C = c, K = k$ in $\mathrm{AP}(k)$ coordinates:
+% \begin{align}
+% s_{tQ}(x^{F,c,k})(y) &= 
+% \begin{cases}
+% \sum_{l=1}^{N}L(y_l - x_l^{F,k})_{+} + ck, & k < \sum F_l^{-1}(1-c/L) \\
+% \sum_{l=1}^{N}L(y_l - x_l^{F,Q,c})_{+} + c(k - z_k(c)), & \text{otherwise, with } 
+% \end{cases} \\
+% z_k(c) &= (k -  \sum F_l^{-1}(1-c/L))_{+}
+% \end{align}
 
-Let $k_1 = \sum F_l^{-1}(1-c_1/L)$ so that $x^{F,Q,c_1} = x^{F, k_1}$ and $z_{k_1}(\{c \leq c_1\}) = 0$, and take $0<c_a < c_1 < c_b < L$ and $0 < k_a < k_1 < k_b < K_2$.
+% Let $k_1 = \sum F_l^{-1}(1-c_1/L)$ so that $x^{F,Q,c_1} = x^{F, k_1}$ and $z_{k_1}(\{c \leq c_1\}) = 0$, and take $0<c_a < c_1 < c_b < L$ and $0 < k_a < k_1 < k_b < K_2$.
 
-Then with $K = k_1$ fixed, the path of functions $s_{k_1}(c) = s_{tQ}(x^{F,c,k_1})$ parametrized by $c \in [0,L]$ takes values
-\begin{align}
-s_{k_1}(0)  &= L\sum ( y_l-x_l^{F,k_1})_{+}  \\
-s_{k_1}(c_a)  &= L\sum ( y_l-x_l^{F,k_1})_{+} + c_a k_1\\
-s_{k_1}(c_1) &= L\sum (y_l - x_l^{F,k_1})_{+} + c_1 k_1  \\
-s_{k_1}(c_b) &= L\sum (y_l - x_l^{F,Q, c_b})_{+} + c_b(k_1 - z_{k_1}(c_b)) \\
-&= L\sum (y_l - x_l^{F,Q, c_b})_{+} + c_b \sum F_l^{-1}(1-c_b/L)\\
-s_{k_1}(L)  &= L\sum {y_l}_{+}.
-\end{align}
+% Then with $K = k_1$ fixed, the path of functions $s_{k_1}(c) = s_{tQ}(x^{F,c,k_1})$ parametrized by $c \in [0,L]$ takes values
+% \begin{align}
+% s_{k_1}(0)  &= L\sum ( y_l-x_l^{F,k_1})_{+}  \\
+% s_{k_1}(c_a)  &= L\sum ( y_l-x_l^{F,k_1})_{+} + c_a k_1\\
+% s_{k_1}(c_1) &= L\sum (y_l - x_l^{F,k_1})_{+} + c_1 k_1  \\
+% s_{k_1}(c_b) &= L\sum (y_l - x_l^{F,Q, c_b})_{+} + c_b(k_1 - z_{k_1}(c_b)) \\
+% &= L\sum (y_l - x_l^{F,Q, c_b})_{+} + c_b \sum F_l^{-1}(1-c_b/L)\\
+% s_{k_1}(L)  &= L\sum {y_l}_{+}.
+% \end{align}
 
-If instead $C = c_1$ is fixed, then $s_{c_1}(k) = s_{tQ}(x^{F,c_1,k})$ parameterized by 
-$k \in [0,K_2]$ take values
-\begin{align}
-s_{c_1}(0) & = L \sum {y_l}_{+} \\
-s_{c_1}(k_a) & = L \sum (y_l - x_l^{F,k_a})_{+} + c_1 k_a\\
-& = L \sum (y_l - x_l^{F,Q,c(k_a)})_{+} + c_1 \sum F_l^{-1}(1-c(k_a)/L)\\
-s_{c_1}(k_1) & = L \sum (y_l - x_l^{F,k_1})_{+} + c_1 k_1 \\
-s_{c_1}(k_b) & = L \sum (y_l - x_l^{F,k_1})_{+} + c_1 (k_b - z_{k_b}(c_1)) \\
- & = L \sum (y_l - x_l^{F,k_1})_{+} + c_1 \sum F_l^{-1}(1-c(k_1)/L)\\
-s_{c_1}(K_2) & = L \sum (y_l - x_l^{F,k_1})_{+} + c_1 (K_2 - z_{K_2}(c_1))
-\end{align}
+% If instead $C = c_1$ is fixed, then $s_{c_1}(k) = s_{tQ}(x^{F,c_1,k})$ parameterized by 
+% $k \in [0,K_2]$ take values
+% \begin{align}
+% s_{c_1}(0) & = L \sum {y_l}_{+} \\
+% s_{c_1}(k_a) & = L \sum (y_l - x_l^{F,k_a})_{+} + c_1 k_a\\
+% & = L \sum (y_l - x_l^{F,Q,c(k_a)})_{+} + c_1 \sum F_l^{-1}(1-c(k_a)/L)\\
+% s_{c_1}(k_1) & = L \sum (y_l - x_l^{F,k_1})_{+} + c_1 k_1 \\
+% s_{c_1}(k_b) & = L \sum (y_l - x_l^{F,k_1})_{+} + c_1 (k_b - z_{k_b}(c_1)) \\
+%  & = L \sum (y_l - x_l^{F,k_1})_{+} + c_1 \sum F_l^{-1}(1-c(k_1)/L)\\
+% s_{c_1}(K_2) & = L \sum (y_l - x_l^{F,k_1})_{+} + c_1 (K_2 - z_{K_2}(c_1))
+% \end{align}
 
-Writing $k(c) = \sum F_l^{-1}(1-c/L)$,
-\begin{align}
-s_{k_1}(c) - s_{c_1}(k(c)) &= (c-c_1)\sum F_l^{-1}(1-\max(c,c_1)/L) \\
-s_{c_1}(k) - s_{k_1}(c(k)) &=  
-\end{align}
+% Writing $k(c) = \sum F_l^{-1}(1-c/L)$,
+% \begin{align}
+% s_{k_1}(c) - s_{c_1}(k(c)) &= (c-c_1)\sum F_l^{-1}(1-\max(c,c_1)/L) \\
+% s_{c_1}(k) - s_{k_1}(c(k)) &=  
+% \end{align}
 
-But 
-\begin{align}
-\del{k}{c}& = \sum_{l=1}^{N} \frac{-1}{L f_l(q_{F_l,\tau(c)})} := -\mathrm{MAI}(c,F)\label{eqn:dkdc}
-\end{align}
-(where we assume that the $F_l$ have non-negative support).
+% But 
+% \begin{align}
+% \del{k}{c}& = \sum_{l=1}^{N} \frac{-1}{L f_l(q_{F_l,\tau(c)})} := -\mathrm{MAI}(c,F)\label{eqn:dkdc}
+% \end{align}
+% (where we assume that the $F_l$ have non-negative support).
 The name MAI refers to this derivative being the marginal total allocation increase for the total quantile loss Bayes act $x^{F,Q(C)}$ that results from a reduction of the cost $C$ of the resource by 1.
 Or in terms of the probability level $\tau(c)=1-c/L$,
-\begin{align}
-\del{k}{\tau} = \del{k}{c}\dby{c}{\tau} = -\mathrm{MAI}(c,F) \cdot (-L\tau) = L\mathrm{MAI}(c,F) \label{eqn:dkdtau}
-\end{align}
+% \begin{align}
+% \del{k}{\tau} = \del{k}{c}\dby{c}{\tau} = -\mathrm{MAI}(c,F) \cdot (-L\tau) = L\mathrm{MAI}(c,F) \label{eqn:dkdtau}
+% \end{align}
 
-Now a constraint $K = 0$ corresponds to cost-loss parity $C=L$, i.e.
+% Now a constraint $K = 0$ corresponds to cost-loss parity $C=L$, i.e.
 $\tau = 0$, for which the Bayes act is the zero vector.
@@
 So making a change of variables with \eqref{eqn:dkdc} or \eqref{eqn:dkdtau} we get
-\begin{align}
-s_{\mathrm{IAS}}(x^{F, \mathrm{IAS}}(k), y) 
-&= \sum_{l=1}^N \int_0^{L} s_{\mathrm{Q}}(x^{F_l, \mathrm{Q}(c)}, y) \mathrm{MAI}(c,F) dc \\
-&= \sum_{l=1}^N \int_0^{1} s_{\mathrm{Q}}(x^{F_l, \tau}, y) L\mathrm{MAI}(L(1-\tau),F) d\tau \\
-&= tw\mathrm{CRPS}(F,y)
-\end{align}
-where MAI provides the (identical) weighting in each coordinate.
+% \begin{align}
+% s_{\mathrm{IAS}}(x^{F, \mathrm{IAS}}(k), y) 
+% &= \sum_{l=1}^N \int_0^{L} s_{\mathrm{Q}}(x^{F_l, \mathrm{Q}(c)}, y) \mathrm{MAI}(c,F) dc \\
+% &= \sum_{l=1}^N \int_0^{1} s_{\mathrm{Q}}(x^{F_l, \tau}, y) L\mathrm{MAI}(L(1-\tau),F) d\tau \\
+% &= tw\mathrm{CRPS}(F,y)
+% \end{align}
+% where MAI provides the (identical) weighting in each coordinate.
 Crucially though, this weighting \emph{depends} on $F$ so that IAS's of two forecasts $F$ and $\tilde{F}$ \emph{cannot} be interpreted as result of applying the same $w$CRPS to the two forecasts.
 
-More about SR's:
-\begin{itemize}
-\item QS: meteorologist loss v newsvendor
-\item WIS/CRPS: multiple scenario loss (Gruschka-Cockayne et al, Fissler and Ziegel, Jose and Winkler)
-	- Least squares analogy
-\item bring in Brier?
-\item take stab at explaing log score following Dawid with Bayes act being inverse of $x^F(c)$ i.e., $F$ itself.
-\end{itemize}
+% More about SR's:
+% \begin{itemize}
+% \item QS: meteorologist loss v newsvendor
+% \item WIS/CRPS: multiple scenario loss (Gruschka-Cockayne et al, Fissler and Ziegel, Jose and Winkler)
+% 	- Least squares analogy
+% \item bring in Brier?
+% \item take stab at explaing log score following Dawid with Bayes act being inverse of $x^F(c)$ i.e., $F$ itself.
+% \end{itemize}
 
@@
 
-As the solution to a constrained optimization problem, the optimal allocation $x^F$ according to a forecast $F$ can be described formally as the first component of a stationary point $(x^{F,k}, \lambda^{\star})$ in the region $\{x \geq 0, \lambda \in \mathbb{R}\}$ for the Lagrangian
+The optimal allocation $x^{F,k}$ according to a forecast $F$ is a solution to the constrained optimization, or \emph{allocation} problem
 \begin{align}
-\mathcal{L}(x, \lambda; F, k, L) = L\sum_{l=1}^{N} E_{F_l}[(Y_l - x_l)_{+}] + \lambda\left(\sum_{l=1}^{N} x_l - k\right).
+    (\mathrm{AP}) \quad \underset{x \in \mathbb{R}_{+}^N}{\mathrm{minimize}}\,\, E_{F}\left[s_A(Y,x)\right] \text{ subject to } 
+    T_x = k. \label{AP}
 \end{align}
-That is, optimality at $x^F$ requires that
+The AP has a convex objective function and affine constraint, so from general convex optimization theory 
+(see e.g. \cite{ruszczynski2011nonlinear}, Chapter 4), it is equivalent to the
+problem of finding a saddle point $(x^{F,k}, \lambda^{\star})$ of the associated Lagrangian 
 \begin{align}
-0 = \frac{\partial}{\partial x_l} \mathcal{L}(x^F, z_K, \lambda) &= L E_F \left[\frac{d}{dx_l}(Y_l - x_l^F)_{+}\right] + \lambda \\
-& = L E_F \left[-\mathbf{1}\{Y_l \geq x_l^F\}\right] + \lambda \\
-& = L(F_l(x_l^F)-1) + \lambda \label{eqn:x_stationary}
+\mathcal{L}(x, \lambda) = \mathcal{L}(x, \lambda; F, k, L) 
+&= E_F\left[s_A(Y,x)\right] + \lambda(T_x - k) \\
+&= L\sum_{l=1}^{N} E_{F_l}[(Y_l - x_l)_{+}] + \lambda\left(\sum_{l=1}^{N} x_l - k\right).  
 \end{align}
-for any $x_l^F > 0$;
+in the region $\{x \geq 0, \lambda \in \mathbb{R}\}$.\footnote{The Lagrangian in this context has the interpretation as the best (lowest) objective function value we can achieve if we forced to choose $x$ but can buy or get credit for changing the constraint level $k$
+at the ``price'' $\lambda$.}
+The Lagrangian reformulation in particular recasts the constraint $T_x=k$ of the original problem as a first-order condition on $\mathcal{L}$ with respect to the new \emph{Lagrange multiplier} variable $\lambda$:
 \begin{align}
-0 = \frac{\partial}{\partial z_K} \mathcal{L}(x^F, z_K, \lambda) &= \lambda - C
+0 = \del{}{\lambda} \mathcal{L}(x^{F,k}, \lambda^{\star}) &= \sum_{l=1}^{N} x_l^{F,k} - k. \label{eqn:lambda_stationary}
 \end{align}
-if $z_K>0$; and
+But for $(x^{F,k}, \lambda^{\star})$ to be a saddle point we also need
 \begin{align}
-0 = \del{}{\lambda} \mathcal{L}(x^F, z_K^F, \lambda^{\star}) &= \sum_{l=1}^{N} x_l^F + z_K^F - K\\
-& = \sum_{\{l | x_l^F > 0\}} F_{l}^{-1}(1-\lambda^{\star}) + z_K^F - K \quad \text{ (by \eqref{eqn:x_stationary})}. \label{eqn:lambda_stationary}
+0 = \frac{\partial}{\partial x_l} \mathcal{L}(x^{F.k}, \lambda) &= L(F_l(x_l^{F,k})-1) + \lambda^{\star} 
+\quad (\text{cf. \eqref{eqn:q_deriv}}) \label{eqn:x_stationary}
 \end{align}
-These equations give formulae $x_i^F = F_i^{-1}(1-\lambda^{\star})$, but in many situations of interest for us the $F_i^{-1}$ are not single-valued, and even 
-when they are, a solution $\lambda^{\star}$ of \eqref{eqn:lambda_stationary} can only be obtained analytically in special cases (but see ...).
-We therefore typically calculate allocation scores using approximately optimal allocations found via a binary search for $\lambda^{\star}$.
-This search begins with the interval $\lambda \in [0,1]$ and
-checks at each step the sign of the inequality that results from substituting the interval midpoint $\lambda_{\tau}$ into \eqref{eqn:lambda_stationary} for endpoints of the sets $F_i^{-1}(1-\lambda_{\tau})$.
-If $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) > K$, the quantiles for $1-\lambda_{\tau}$ exceed our resource constraint so we set $\lambda_{\tau}$ as the lower endpoint of the $\tau + 1$ search interval.
-Similarly, if $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) < K$, the quantiles for
-$1-\lambda_{\tau}$ do not make full use of our resource and we set $\lambda_{\tau}$ as the upper endpoint of the $\tau + 1$ search interval.
-This process halves the search interval width at each step and we continue until this width falls below a given tolerance level $\varepsilon_{\tau}$ yielding an
-approximation $\lambda^{\star, \varepsilon}$. When quantiles $F_i^{-1}(1-\lambda^{\star, \varepsilon})$ are singly defined and the forecast densities $f_i$
-not too close to zero at these quantiles, the discrepancy $\Delta(\lambda^{\star, \varepsilon})  = \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star, \varepsilon}) - K$ is of order $\varepsilon_{\lambda}$ and we can take $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ as approximations to $x_i^F$.
+for any $x_l^{F,k} > 0$, which gives a formula 
+\begin{align}
+x_l^{F,k} = F_l^{-1}(1-\lambda^{\star}/L) \label{eqn:xfk_formula}
+\end{align}
+for any non-zero element $x_l^{F,k}$ of the solution. Thus \eqref{eqn:lambda_stationary} becomes
+\begin{align}
+0 = \sum_{\{l | x_l^{F,k} > 0\}} F_{l}^{-1}(1-\lambda^{\star}/L) - k. \label{eqn:lambda_stationary2}
+\end{align}
 
-Often though, $1-\lambda^{\star, \varepsilon}$ will be the probability level of a ``plateau'' for one or more $F_i$ and the $\lambda$ search algorithm will fail to drive $\Delta(\lambda^{\star, \varepsilon})$ close enough to zero for the $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ to be acceptable approximations of the optimal allocations.
+Since convexity and compactness guarantee existence of a solution, the problem is now reduced to solving \eqref{eqn:lambda_stationary} for the set 
+$\{l \mid x_l^{F,k} > 0\}$ and value $\lambda^{\star}$. Note, however, that this is a more complicated task than solving just 
+for $\lambda^{\star}$ under the assumption that \eqref{eqn:xfk_formula} holds for all coordinates of $x^{F,k}$.\footnote{Such an assumption would exclude fairly simple examples in which $F$ has support bounded away from $0$ in one coordinate but not in another.} The task is simplified though by bringing in the full saddle point condition for $\lambda^{\star}$ which implies not only \eqref{eqn:lambda_stationary} but also that
+$\lambda^{\star}$ is in fact the unique maximizer of the \emph{dual function} 
+$\mathcal{L}_D(\lambda) := \inf_{x \in \mathbb{R}_{+}^N} \mathcal{L}(x, \lambda)$.
+This approach is especially suitable in our situation because our original objective function $E_F\left[s_A(Y,x)\right]$ is decomposable in the coodinates $x_l$, allowing the dual function to be computed separately with respect to each coordinate:
+\begin{align}
+\mathcal{L}_D(\lambda) &= \inf_{x \in \mathbb{R}_{+}^N} L\sum_{l=1}^{N} E_{F_l}[(Y_l - x_l)_{+}] + \lambda\left(\sum_{l=1}^{N} x_l - k\right) \\
+&= - \lambda k + \sum_{l=1}^{N} \inf_{x_l \in \mathbb{R}_{+}}\left\{ L E_{F_l}[(Y_l - x_l)_{+}] + \lambda x_l\right\}  \label{eqn:dual_decomp} \\
+&= - \lambda k + \sum_{l=1}^{N}  LE_{F_l}[(Y_l - x_l(\lambda))_{+}] + \lambda x_l(\lambda).
+\end{align}
+Here, $x_l(\lambda)$ is any member of the set 
+\begin{align}
+\chi_l(\lambda) = \underset{x_l \in \mathbb{R}_{+}}{\mathrm{arginf}} \left\{ L E_{F_l}[(Y_l - x_l)_{+}] + \lambda x_l\right\}
+\end{align}
+which for $F_l$ in a standard parametric family will often have simple the form $\chi_l(\lambda) = \{F_l^{-1}(1-\lambda/L)\}$ but could an interval containing $0$ or $\infty$.
+From general theory, $\mathcal{L}_D(\lambda)$ is concave, and while it may not be differentiable when the $x_l$'s attaining the infima in \eqref{eqn:dual_decomp} are zero, it does have as a subdifferential the well-defined interval
+\begin{align}
+\partial \mathcal{L}_D(\lambda) 
+&= \sum_{l=1}^N \chi_l(\lambda) - k.
+\end{align}
+This of course looks very similar to the right hand side of \eqref{eqn:lambda_stationary2}, but the dual perspective has the key advantage of not requiring any guesses of the set $\{l \mid x_l^{F,k} > 0\}$ when initiating a search algorithm for $\lambda^{\star}$.
+Instead, given a candidate $\lambda_0$ for $\lambda^{\star}$, we proceed by calculating the sets $\chi_l(\lambda_0), l = 1,\ldots,N$, and using them to find the (usually zero-width) interval $\partial \mathcal{L}_D(\lambda_0)$.
+If $\partial \mathcal{L}_D(\lambda_0) \subset (-\infty,0)$, we can eliminate $\{\lambda \geq \lambda_0\}$ from our search and similarly if $\partial \mathcal{L}_D(\lambda_0) \subset (0, \infty)$, we can eliminate $\{\lambda \leq \lambda_0\}$.
+
+Intutively, we are testing to see whether the total amount $T(\lambda_0)$ of resources that would be procured at a cost $C_0=\lambda_0$ in the unconstrained decision problem with loss function \eqref{eqn:TQ_loss} leaves a deficit or surplus relative to $k$.
+$T$ is a sum over a set of quantiles and both the set and the quantiles are non-decreasing with cost.
+(This is essentially why $\mathcal{L}_D$ is concave.) Thus by evaluating $T$ at a mid-point we cut the search interval width in half at each step, arriving quickly at a close approximation $\lambda^{\star, \varepsilon}$ of the cost corresponding to $k$.
+The strategy of analysing constrained optimization problems by describing the ``prices'' at which unconstrained decisions match constrained ones is ubiquitous in optimization theory and its applications, especially economics.
+
+If the $F_l$ increase steadily enough toward 1 after leaving a neigborhood of 0, we can expect $\partial \mathcal{L}_D(\lambda)$ to be bounded away from 0 for all $\lambda \neq \lambda^{\star}$, allowing a binary search to converge without ambiguity.
+But when some $F_l$ exhibit internal ``plateaus'' where their densities stay at or near 0, we can and often do in our applications encounter $\overline{\lambda}$ for which $\partial \mathcal{L}_D(\overline{\lambda})$---or at least our numerical approximation of it---contains 0 as an interior point.
+That is, $\overline{\lambda}$ is a price at which an uncontrained decision maker informed only by $F$ can increase from total provision $T_1<k$ to another $T_2>k$ while maintaining an optimal expected utility by making (at least approximately) one-to-one trade-offs between increased resource cost and risk reduction in the one or more locations where $F_l$ has a plateau of height near $1-\overline{\lambda}/L$.
+
 In this event we follow-up with a ``post-processing'' step that identifies lower and upper boundaries $x_{i,L}$ and $x_{i,U}$ of any

### 2023-08-30 alloscore_manuscript/allocation.tex
easy edits in response to Aaron's comments

@@
 Infectious disease forecasting models have emerged as important tools in public health.
-Predictions of disease dynamics have been used to inform decision-making about a wide variety of measures to reduce disease spread and/or mitigate the severity of disease outcomes.
+Predictions of disease dynamics have been used to inform decision making about a wide variety of measures to reduce disease spread and/or mitigate the severity of disease outcomes.
 For example, estimates of expected onset of flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola dynamics have been used to allocate surveillance resources \citep{meltzer2014estimating, rainisch2015regional}.
@@
 Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping (University of Texas, 2022) \nocite{utnews2022}.
-In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision-making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
+In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
 The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid} or weighted interval score (WIS) \citep{bracher2021evaluating}, however, recently authors have been calling for evaluating forecasts through their impact on policy \citep{marshall2023predictions}.
 
-In decision-making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to.
+In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to.
 We give an overview of these procedures in Section \ref{sec:methods.decisiontheory}.
-There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision-making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
+There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
 We review this work only briefly here, and we refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \elr{TODO: identify relevant review or book style summary for meteorology} for discussions focused on applications to economics and meteorology, respectively.
@@
 In this framework, a forecast has value if it leads to decisions with low total costs.
-In all of these fields, analyses have consistently found that common measures of statistical forecast accuracy do not necessarily correspond directly to measures of the value of forecasts as an input to decision-making \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
+In all of these fields, analyses have consistently found that common measures of statistical forecast accuracy do not necessarily correspond directly to measures of the value of forecasts as an input to decision making \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
 % Consistently, but not uniformly
@@
 % Other important early contributions include Nelson (1972) and White (1966), who are amongst the first to consider the problem of forecast evaluation from the perspective of decision theory and dynamic programming.
-Nelson, in particular, articulates the concept of the “value of a (probability) fore- cast” to an individual decision-maker as the expected value of his/her payoff using the probability forecasts relative to the payoff expected if unconditional probability estimates are used.
+Nelson, in particular, articulates the concept of the “value of a (probability) fore- cast” to an individual decision maker as the expected value of his/her payoff using the probability forecasts relative to the payoff expected if unconditional probability estimates are used.
 
@@
 Briefly, the allocation score of a forecast is the avoidable unmet need that results from using that forecast to set resource allocations by minimizing expected unmet need.
-For example, suppose that a decision-maker is provided with forecasts of the level of need for medical resources in each of several states or hospital systems.
-If there is a limited amount of the medical resource that is available to distribute, a decision-maker could choose an allocation of that resource across locations that minimizes the expected unmet need according to that forecast.
-As measured by the allocation score, one forecast is better than another if it would lead decision-makers to an allocation that results in less unmet need.
+For example, suppose that a decision maker is provided with forecasts of the level of need for medical resources in each of several states or hospital systems.
+If there is a limited amount of the medical resource that is available to distribute, a decision maker could choose an allocation of that resource across locations that minimizes the expected unmet need according to that forecast.
+As measured by the allocation score, one forecast is better than another if it would lead decision makers to an allocation that results in less unmet need.
 %By ``unnecessary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
@@
 We present an illustrative analysis using the allocation score to evaluate forecasts of hospital admissions in the US leading up to the Omicron wave in winter 2022.
-This analysis is ``synthetic'' in that it does not correspond to an actual analysis that supported decision-making in real-time.
+This analysis is ``synthetic'' in that it does not correspond to an actual analysis that supported decision making in real-time.
 However, the framework described in this paper corresponds to real-world decisions that must be made by public health administrators around the globe, and could be adapted in the future for such real-time situations.
@@
 We begin with an informal description of the allocation score and some examples illustrating its key characteristics in section \ref{sec:methods.overview}.
-In section \ref{sec:methods.detailed} we develop the allocation score more carefully, building on decision-theoretic procedures for deriving proper scoring rules.
-We then comment on some connections between the allocation score that we propose and other common scores that can be derived from decision-theoretic foundations, such as the quantile score, WIS, and CRPS, in section \ref{sec:methods.related}.
+In section \ref{sec:methods.detailed} we develop the allocation score more carefully, building on decision theoretic procedures for deriving proper scoring rules.
+We then comment on some connections between the allocation score that we propose and other common scores that can be derived from decision theoretic foundations, such as the quantile score, WIS, and CRPS, in section \ref{sec:methods.related}.
 
@@
 
-Suppose that a decision-maker is tasked with determining how much of a resource such as a medical supply should be allocated to each of $N$ locations, subject to the constraint that only $K$ units of the resource are available.
-If the decision-maker is provided with a multivariate forecast $F$ that predicts resource need in each location, one option is to choose the resource allocation that minimizes the expected total unmet need according to the forecast.
+Suppose that a decision maker is tasked with determining how to allocate $K$ available units of a resource across $N$ locations.
+If the decision maker is provided with a multivariate forecast $F$ where each marginal forecast distribution $F_i$ predicts resource need in a particular location, one option is to choose the resource allocation that minimizes the expected total unmet need according to the forecast.
 We will give a more precise mathematical statement in section \ref{sec:methods.allocation}, but informally, the total expected unmet need according to the forecast is
@@
 misalignment in one coordinate is penalized regardless of alignments in other coordinates.
-Note that we do not claim that the allocation score is generically preferable to these other scores \textemdash rather, it provides a view of forecast performance that is specifically tuned to the context of decision-making about resource allocations.
+Note that we do not claim that the allocation score is generically preferable to these other scores \textemdash rather, it provides a view of forecast performance that is specifically tuned to the context of decision making about resource allocations.
 
@@
 
-\subsection{A decision-theoretic development of the allocation score}
+\subsection{A decision theoretic development of the allocation score}
 \label{sec:methods.detailed}
 
-We give a high-level review of a general procedure for developing proper scoring rules that are tailored to specific decision-making tasks in section \ref{sec:methods.detailed.decisiontheory}, and then in section \ref{sec:methods.detailed.allocation} we apply that procedure to develop the allocation score based on the task of deciding on how to allocate a fixed supply of resources across multiple locations.
+We give a high-level review of a general procedure for developing proper scoring rules that are tailored to specific decision making tasks in section \ref{sec:methods.detailed.decisiontheory}, and then in section \ref{sec:methods.detailed.allocation} we apply that procedure to develop the allocation score based on the task of deciding on how to allocate a fixed supply of resources across multiple locations.
 
-\subsubsection{The decision-theoretic setup for forecast evaluation}
+\subsubsection{The decision theoretic setup for forecast evaluation}
 \label{sec:methods.detailed.decisiontheory}
 
-%In this section, we give an overview of the decision-theoretic setup for developing proper scoring rules that measure the value of a forecast as an input to decision making.
+%In this section, we give an overview of the decision theoretic setup for developing proper scoring rules that measure the value of a forecast as an input to decision making.
 We keep the discussion here at a somewhat informal level; we refer the reader to [some subset of Brehmer and Gneiting; Grünwald and Dawid; Dawid; Granger and Pesaran 2000; Granger and Machina 2006; Ehm et al.
@@
 For example, $y$ may correspond to the number of individuals who eventually become sick and would benefit from the mitigation measure, and informally, an action $x$ is successful to the extent that it meets the realized need.
-In the face of uncertainty, a decision-maker may use a forecast $F$ of the random variable $Y$ to help inform the selection of the action to take.
-We measure the value of a forecast as an input to this decision-making process by the quality of the decisions that it leads to.
+In the face of uncertainty, a decision maker may use a forecast $F$ of the random variable $Y$ to help inform the selection of the action to take.
+We measure the value of a forecast as an input to this decision making process by the quality of the decisions that it leads to.
 
@@
 Subject to certain technical conditions, scoring rules obtained from this procedure are proper.
-We refer the reader to [some subset of Brehmer and Gneiting; Grünwald and Dawid; Dawid; Granger and Pesaran 2000; Granger and Machina 2006; Ehm et al.
-2016] for more technically precise discussion.
+We refer the reader to discussion in (cite paper 2 on arxiv) for a more technically precise discussion.
 
@@
 
-% We review how the quantile score arises from a particular decision-making problem in section \ref{sec:methods.quantileloss.quantile}, and how CRPS can be obtained by integrating across values of the parameters of that decision-making problem, as well as the connection to WIS, in section \ref{sec:methods.quantileloss.crps}.
+% We review how the quantile score arises from a particular decision making problem in section \ref{sec:methods.quantileloss.quantile}, and how CRPS can be obtained by integrating across values of the parameters of that decision making problem, as well as the connection to WIS, in section \ref{sec:methods.quantileloss.crps}.
 These results have been thoroughly discussed in the literature [cite cite cite].
 
-% \subsubsection{Decision-theoretic origins of the quantile score}
+% \subsubsection{decision theoretic origins of the quantile score}
 % \label{sec:methods.quantileloss.quantile}
 
-% Suppose that a decision-maker is tasked with determining the quantity $x$ of a protective measure to procure; for example, $x$ might represent the number of hospital beds or amount of medicine to purchase.
+% Suppose that a decision maker is tasked with determining the quantity $x$ of a protective measure to procure; for example, $x$ might represent the number of hospital beds or amount of medicine to purchase.
 Additionally, suppose that each unit of this good has cost $C$ so that the total cost of procurement is $Cx$.
@@
 We assume that each unit of unmet need incurs a loss denoted by $L$, so that if the selected procurement level $x$ is less than the realized need $y$, a loss of $L(x-y)$ results.
-At the time that a decision-maker determines the amount $x$ to procure, the demand $y$ is not yet known.
+At the time that a decision maker determines the amount $x$ to procure, the demand $y$ is not yet known.
 We therefore define the random variable $Y$ that represents the as-yet-unknown level of demand.
@@
 Here we identify $F$ with its cumulative distribution function (CDF), and $F^{-1}$ denotes the quantile function.
-With this formalization of the decision-making task, we can proceed to develop a proper scoring rule using the procedure outlined in section \ref{sec:methods.decisiontheory}.
+With this formalization of the decision making task, we can proceed to develop a proper scoring rule using the procedure outlined in section \ref{sec:methods.decisiontheory}.
 
@@
 
-% We have set up the problem here in terms of the cost and loss parameters $C$ and $L$, which has the benefit of an intuitive connection to the decision-making context.
+% We have set up the problem here in terms of the cost and loss parameters $C$ and $L$, which has the benefit of an intuitive connection to the decision making context.
 However, to clarify the connection to the usual notation for the quantile loss, we can divide the loss function $s_Q$ by $L$ to obtain an expression in terms of only $\alpha$:
@@
 % The scoring rule $S_Q$ of Equation \eqref{eqn:S_Q_alpha} evaluates the forecast distribution $F$ only through its $\alpha$ quantile.
-While this is faithful to the context of the decision-making problem, it may not be satisfying as a measure of the quality of the full forecast distribution.
+While this is faithful to the context of the decision making problem, it may not be satisfying as a measure of the quality of the full forecast distribution.
 For this purpose, one option is to integrate the quantile scoring rule across different values of the probability level $\alpha$, weighting the probability levels according to a specified distribution $p$.
@@
 
-% We emphasize that because $\alpha = 1 - C/L$, the distribution $p(\alpha)$ can be interpreted as expressing incomplete knowledge about the cost/loss ratio in the decision-making problem.
-The equal weighting used by the ordinary CRPS may be appropriate in the absence of any knowledge about the context in which forecasts will be used to support decision-making, but may be inappropriate if more information is known about the cost/loss ratio for a specific decision-making task.
+% We emphasize that because $\alpha = 1 - C/L$, the distribution $p(\alpha)$ can be interpreted as expressing incomplete knowledge about the cost/loss ratio in the decision making problem.
+The equal weighting used by the ordinary CRPS may be appropriate in the absence of any knowledge about the context in which forecasts will be used to support decision making, but may be inappropriate if more information is known about the cost/loss ratio for a specific decision making task.
 
@@
 An administrator is tasked with determining where to send these resources so as to meet demand among hospital patients in different facilities or states.
-In contrast to the decision-making problem in the previous section, the administrator is not able to control the total amount of supply; rather, their task is to determine how to allocate the fixed supply to different locations.
+In contrast to the decision making problem in the previous section, the administrator is not able to control the total amount of supply; rather, their task is to determine how to allocate the fixed supply to different locations.
 Following the structure of the discussion for the quantile score and CRPS, we first discuss the set up for a single value of the resource constraint $K$, which plays a role that is analogous to the cost/loss ratio for the quantile score, and then we discuss the possibility of integrating (or averaging) across values of $K$.
@@
 
-In the decision-making setting that we consider, an action $\mathbf{x} = (x_1, \ldots, x_N)$ is a vector specifying the amount that is allocated to each of $N$ locations.
+In the decision making setting that we consider, an action $\mathbf{x} = (x_1, \ldots, x_N)$ is a vector specifying the amount that is allocated to each of $N$ locations.
 We require that each $x_i$ is non-negative and that the total allocation across all locations equals the amount of available resources, $K$: $\sum_{i=1}^N x_i = K$.
@@
 The quantile is at a probability level $\alpha = 1 - \lambda^{\star}/L$ that is the same for all locations, and is chosen such that the constraint is satisfied.
-According to this interpretation, we divide the available resources among the locations in such a way that the forecasted probability that need will exceed demand is the same in all locations.
+According to this interpretation, we divide the available resources among the locations in such a way that the forecasted probability that need will exceed the available resources is the same in all locations.
 An alternative interpretation comes from noting that for each location $i$, $\frac{\partial}{\partial x_i} \sbar_A^F(\mathbf{x}) \vert_{\mathbf{x}=\mathbf{x}^F} = \lambda^{\star}$ (see the supplement for a proof).
@@
 This is appropriate if the resource constraint $K$ is a known constant.
-However, if $K$ is not precisely known at the time of decision-making or there is interest in measuring the value of forecasts across a range of decision-making scenarios, we can use an \emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a distribution $p$:
+However, if $K$ is not precisely known at the time of decision making or there is interest in measuring the value of forecasts across a range of decision making scenarios, we can use an \emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a distribution $p$:
 $$S_{IAS}(F, y) = \int S_A(F,y; L, K) p(L, K) \, dL dK$$
@@
 %We use this link to explore the relationship between IAS and CRPS from two directions.
-First, suppose the decision-maker has some uncertainty about the value of $K$, which they express through $p$.
+First, suppose the decision maker has some uncertainty about the value of $K$, which they express through $p$.
 Because $\alpha$ can be regarded as a function of $K$, this distribution on the resource constraint induces a distribution on quantile levels.
@@
 In fields such as meteorology and supply chain management, a great deal of attention has been given to a different problem where a decision must be made about the quantity of a resource to purchase for a single location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each unit of unmet need.
-This decision-making problem leads to the pinball loss that is often used in quantile regression, and the optimal quantity of resources to purchase is the predictive quantile at the probability level $1 - \frac{C}{L}$.
+This decision making problem leads to the pinball loss that is often used in quantile regression, and the optimal quantity of resources to purchase is the predictive quantile at the probability level $1 - \frac{C}{L}$.
 
 This result bears a strong resemblance to the results for the allocation problem, where the optimal solution also sets the resource level in each location to a quantile of the respective forecast distributions.
-These decision-making problems differ according to the challenge faced by the decision maker: a fixed constraint on the available resources for the allocation problem, or a cost per unit of resources in the resource purchasing problem.
-In fact, it is possible to combine these into a more general problem where the decision-maker must decide on both the total level of resources to purchase and an allocation of those resources across multiple locations, subject to a cost per unit of resources and a constraint on the total resources that can be purchased.
-The quantile loss and the allocation score presented in this work are both be special cases of the score that emerges from that more general decision-making problem.
+These decision making problems differ according to the challenge faced by the decision maker: a fixed constraint on the available resources for the allocation problem, or a cost per unit of resources in the resource purchasing problem.
+In fact, it is possible to combine these into a more general problem where the decision maker must decide on both the total level of resources to purchase and an allocation of those resources across multiple locations, subject to a cost per unit of resources and a constraint on the total resources that can be purchased.
+The quantile loss and the allocation score presented in this work are both be special cases of the score that emerges from that more general decision making problem.
 We pursue this direction further in other work (cite arxiv version of paper 2).
@@
 Similarly, it is possible to see both the integrated allocation score and common scores such as the  continuous ranked probability score (CRPS) and weighted interval score (WIS) as special cases of a more general score that integrates across resource cost and constraint parameters together.
-\elr{previous sentence may be controversial -- how to frame?} It is well known in the literature that WIS is equivalent to a weighted average of quantile scores across multiple quantile levels (cite Bracher et al.) and CRPS integrates quantile scores across multiple quantile levels (cite cite).
+It is well known in the literature that WIS is equivalent to a weighted average of quantile scores across multiple quantile levels (cite Bracher et al.) and CRPS integrates quantile scores across multiple quantile levels (cite cite).
 By integrating the combined resource acquisition and allocation score that we described in the previous paragraph across the parameters for both resource costs and constraints, we obtain a score that combines features of both IAS and CRPS and contains those scores as special cases.
@@
 Case study heading into the Omicron wave.
-Some more detailed discussion of implications of bad forecasts for specific decision-making purposes -- take a "deep dive" into one or two example states like FL.
+Some more detailed discussion of implications of bad forecasts for specific decision making purposes -- take a "deep dive" into one or two example states like FL.
 
@@
 
-We often conceive of infectious disease forecasts as being useful for decision-making purposes, but it is rare for forecast evaluation to be tied directly to the value of the forecasts for informing those decisions.
+We often conceive of infectious disease forecasts as being useful for decision making purposes, but it is rare for forecast evaluation to be tied directly to the value of the forecasts for informing those decisions.
 This work seeks to address that gap.
 
-We have demonstrated that evaluation methods that are tied to decision-making context can yield model rankings that are substantively different from generic measures of forecast skill like WIS.
+We have demonstrated that evaluation methods that are tied to decision making context can yield model rankings that are substantively different from generic measures of forecast skill like WIS.
 
-In practice, there are many users of forecasts with many different decision-making problems.
+In practice, there are many users of forecasts with many different decision making problems.
 Not all can be easily quantified.
@@
 
-The allocation score we developed also does not attempt to capture the broader context of decision-making.
+The allocation score we developed also does not attempt to capture the broader context of decision making.
 For example, in practice it may be possible to increase the resource constraint $K$ by shifting funding from other disease mitigation measures.

### 2023-08-30 alloscore_manuscript/allocation.tex
minor update to description of IAS

@@
 
-\elr{I did a little work here but did not really finish it off. Can we just integrate across $K$? And maybe there's a better way to set this up.  Aiming for just 1-2 paragraphs, though.}
-
 The allocation score $S_A$ that we developed in the previous section measures the skill of the forecast distributions $F$ based on a single probability level $\alpha = 1 - \lambda^{\star}/L$.
 This is appropriate if the resource constraint $K$ is a known constant.
-However, if $K$ is not precisely known at the time of decision making or there is interest in measuring the value of forecasts across a range of decision making scenarios, we can use an \emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a distribution $p$:
-$$S_{IAS}(F, y) = \int S_A(F,y; L, K) p(L, K) \, dL dK$$
+However, if $K$ is not precisely known at the time of decision making or there is interest in measuring the value of forecasts across a range of decision making scenarios with different resource constraints, we can use an \emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a distribution $p$:
+$$S_{IAS}(F, y; L) = \int S_A(F,y; L, K) p(K) \, dK$$
 

### 2023-08-30 alloscore_manuscript/allocation.tex
lost paragraph from https://github.com/aaronger/utility-eval-papers/pull/31

@@
 
-%A forecaster is asked to recommend, that is, \emph{predict}, the appropriate supply $x$ of some good or investment in some precautionary measure intended to satisfy a random future demand or need $Y$.
-As a running example, we will consider the provision of a limited supply of a limited health care resource such as oxygen or ventilators.
-Suppose there is an incremental loss $O \geq 0$ incurred when over-prediction leads to unused supply and an incremental loss $U > 0$ incurred when under-prediction leads to unmet demand or need.
-
-% \subsection{A review of quantile score, CRPS, and the weighted interval score}
-% \label{sec:methods.quantileloss}
-
-% We review how the quantile score arises from a particular decision making problem in section \ref{sec:methods.quantileloss.quantile}, and how CRPS can be obtained by integrating across values of the parameters of that decision making problem, as well as the connection to WIS, in section \ref{sec:methods.quantileloss.crps}.
-These results have been thoroughly discussed in the literature [cite cite cite].
-
-% \subsubsection{decision theoretic origins of the quantile score}
-% \label{sec:methods.quantileloss.quantile}
-
-% Suppose that a decision maker is tasked with determining the quantity $x$ of a protective measure to procure; for example, $x$ might represent the number of hospital beds or amount of medicine to purchase.
-Additionally, suppose that each unit of this good has cost $C$ so that the total cost of procurement is $Cx$.
-The variable $y$ denotes the eventual realized need for this resource, e.g.
-the number of patients in need of a hospital bed or the amount of medication that is needed.
-We assume that each unit of unmet need incurs a loss denoted by $L$, so that if the selected procurement level $x$ is less than the realized need $y$, a loss of $L(x-y)$ results.
-At the time that a decision maker determines the amount $x$ to procure, the demand $y$ is not yet known.
-We therefore define the random variable $Y$ that represents the as-yet-unknown level of demand.
-The forecast $F$ specifies a predictive distribution for $Y$.
-Here we identify $F$ with its cumulative distribution function (CDF), and $F^{-1}$ denotes the quantile function.
-With this formalization of the decision making task, we can proceed to develop a proper scoring rule using the procedure outlined in section \ref{sec:methods.decisiontheory}.
-
-% \paragraph{Step 1: specify a loss function.} Combining the cost of procuring goods at level $x$ with losses due to unmet need, we arrive at the overall loss function
-% \begin{align}
-% s_Q(x,y; C, L) = Cx + L(x-y)_-.
-% \end{align}
-% Here, $(x - y)_- := \max(-(x - y),0)$ is $0$ if the amount procured, $x$, is greater than or equal to the realized demand $y$; otherwise, it is $y - x$, the amount of unmet need.
-
-% \paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} It can be shown that under the loss function $s_Q$, the Bayes act is a quantile of the forecast distribution at the probability level $\alpha = 1 - C/L$:
-% \begin{align}
-% x^F = F^{-1}(\alpha).
-% \end{align}
-% See the supplement for a verification of this result.
-% %In particular, faced with the classical binary decision problem of whether to recommend an additional unit of protection given a current level of protection $x$, the forecaster's optimal decision rule under $s_m$ is to recommend adding protection if $x < F^{-1}\left(1-\frac{C}{L}\right)$, that is, if
-% %\begin{align}
-% %1-F(x) = \mathbb{P}_F\{y>x\} > \frac{C}{L},
-% %\end{align}
-% %the \emph{cost-loss ratio} of the problem.
-
-% \paragraph{Step 3: Define the scoring rule.} Following the procedures outlined above, we could score the forecast distribution $F$ with the scoring rule
-% \begin{align*}
-% S_Q(F, y; C, L) &= s_Q(x^F, y; C, L) = Cx^F + L(x^F-y)_- \nonumber \\
-% &= C F^{-1}(\alpha) + L \left(F^{-1}(\alpha) - y\right)_- \label{eqn:S_Q_CL}
-% \end{align*}
-
-% We have set up the problem here in terms of the cost and loss parameters $C$ and $L$, which has the benefit of an intuitive connection to the decision making context.
-However, to clarify the connection to the usual notation for the quantile loss, we can divide the loss function $s_Q$ by $L$ to obtain an expression in terms of only $\alpha$:
-% \begin{align*}
-% s_Q(x,y; \alpha) &= s_Q(x,y; C, L)/L \\
-% &= (C/L)x + (x-y)_- \\
-% &= (1 - \alpha)x + (x-y)_-.
-% \end{align*}
-% Because these loss functions are equal up to a constant of proportionality, the Bayes act is the same for both. The associated quantile scoring rule expressed in terms of $\alpha$ is
-% \begin{equation}
-% S_Q(F, y; \alpha) = (1 - \alpha)F^{-1}(\alpha) + \left(F^{-1}(\alpha) - y\right)_-. \label{eqn:S_Q_alpha}
-% \end{equation}
-% In either formulation, the key observation is that the Bayes act is the quantile of the forecast distribution $F$ at the probability level given by one minus the cost/loss ratio $C/L$.
-
-% \subsubsection{CRPS as an integrated quantile score}
-% \label{sec:methods.quantileloss.crps}
-
-% The scoring rule $S_Q$ of Equation \eqref{eqn:S_Q_alpha} evaluates the forecast distribution $F$ only through its $\alpha$ quantile.
-While this is faithful to the context of the decision making problem, it may not be satisfying as a measure of the quality of the full forecast distribution.
-For this purpose, one option is to integrate the quantile scoring rule across different values of the probability level $\alpha$, weighting the probability levels according to a specified distribution $p$.
-This yields a weighted CRPS:
-% \begin{align*}
-% S_{CRPS}(F, y; p) &= \int S_Q(F, y; \alpha) p(\alpha) \, d\alpha.
-% \end{align*}
-% This weighted form of CRPS has appeared in the literature before, e.g.
-see \cite{gneiting2011weightedScoringRules}.
-The usual CRPS results from taking $p$ corresponding to a $\text{Uniform}(0,1)$ distribution, equally weighting all probability levels.
-
-% We emphasize that because $\alpha = 1 - C/L$, the distribution $p(\alpha)$ can be interpreted as expressing incomplete knowledge about the cost/loss ratio in the decision making problem.
-The equal weighting used by the ordinary CRPS may be appropriate in the absence of any knowledge about the context in which forecasts will be used to support decision making, but may be inappropriate if more information is known about the cost/loss ratio for a specific decision making task.
-
-% The weighted interval score (WIS) is often used when the full forecast distribution $F$ is not available, as in the U.S.
-COVID-19 Forecast Hub and similar efforts where forecasts are represented by a collection of prediction intervals.
-WIS is a discrete approximation to CRPS, and can be obtained by using a distribution $p$ that has point masses at the probability levels corresponding to the endpoints of a finite set of prediction intervals.
-
-%\elr{Leaving the comment below in the tex, but noting that it was in response to an older version of the write up.
-I've since attempted to address it by moving to a discussion in terms of a version of the scoring function where we've divided by L to work in terms of the cost/loss ratio.}
-%\apg{Quick note that I don't think this will work without some more identifications.
-The integral needs to be over probability levels and thresholds and without the some discussion like the one commented out (currently) at line 123 or (better in my mind) one like I'm trying to do with binary choices at the unit-level, I don't think we have a threshold.}
-%The usual CRPS results from taking $p$ to be the density of a distribution such that the induced distribution on the probability level $\alpha = 1 - C/L$ is $\text{Uniform}(0,1)$.
-
-% \subsection{The allocation score}
-% \label{sec:methods.allocation}
-
-%\elr{If we like the organization of the previous subsection into two subsubsections about the quantile score and the CRPS, we should replicate that here.}
-
-% We now give a more formal derivation of the allocation score starting from the set up of a decision making problem about how to allocate limited resources to meet demand across multiple locations.
-In section \ref{sec:methods.allocation.special}, we consider a setting where the resource constraint is known, which will result in a score that is specialized to a particular quantile of the forecasts.
-Then in sction \ref{sec.methods.allocation.integrated}, we allow for the possibility of decision making under uncertainty about the precise value of the resource constraint, which will lead to averaging the specialized allocation score across multiple quantile levels.
-
-%As a concrete example, we take the resource to be a good such as ventilators or oxygen supply.
-An administrator is tasked with determining where to send these resources so as to meet demand among hospital patients in different facilities or states.
-In contrast to the decision making problem in the previous section, the administrator is not able to control the total amount of supply; rather, their task is to determine how to allocate the fixed supply to different locations.
-Following the structure of the discussion for the quantile score and CRPS, we first discuss the set up for a single value of the resource constraint $K$, which plays a role that is analogous to the cost/loss ratio for the quantile score, and then we discuss the possibility of integrating (or averaging) across values of $K$.
-
 \subsubsection{The allocation score for a fixed resource constraint}
@@
 We note that the problem formulation here assumes that the resource in question does not impact the amount of demand $y_i$ that will materialize, but rather it is a resource that satisfies that demand.
-In the context of infectious disease, this means that we do not consider resources such as vaccine doses or masks, which are intended to reduce the number of people who will become sick.
-Our set up addresses resources like hospital beds, oxygen supply, or ventilators, which are intended to meet the medical needs of patients who are sick.
+In the context of infectious disease, this means that we do not consider resources that are intended to reduce the number of people who will become sick at some point in the future, such as a preventative influenza or COVID-19 vaccine.
+Our set up addresses resources like hospital beds, oxygen supply, ventilators, or rabies vaccines which are intended to meet the medical needs of patients who are already sick.
 
@@
 
-% ELR: I made some more edits to the previous paragraph, but still may not have captured everything that Aaron think is important here.
-% \apg{
-% first attempt by APG to rephrase last paragraph: The central mathematical ideas in this construction are that
-% \begin{itemize}
-% \item optimization under a \emph{single} constraint requires the rates of change of (a probabilistic forecast's) expected benefit with respect to our decision variables, $x_i$, to be the same for all locations
-% \item these rates of change can be identified with probabilities
-% \item and therefore the Bayes act results from using a \emph{shared} probability level, $1 - \lambda^{\star}/L$, to determine allocations as the corresponding quantiles of the location-specific forecast distributions $F_i$ which satisfy the constraint.
-% \end{itemize}
-% (See the supplement for detailed discussion and derivations of these points.)
-% }
-
 \paragraph{Step 3: Define the scoring rule.} We can now define a proper scoring rule for the probabilistic forecast $F$ as
@@
 $$S_{IAS}(F, y; L) = \int S_A(F,y; L, K) p(K) \, dK$$
-
-%We illustrate the relationship between the IAS and the CRPS in our example with two locations and forecasts given by $\text{Exp}(1/\sigma_i)$ distributions with $\sigma_1 = 1$ and $\sigma_2 = 5$.
-Note that the quantile functions corresponding to these forecasts are given by $F_i^{-1}(\alpha) = -\sigma_i \log(1 - \alpha)$.
-For simplicity, we keep $L = 1$ fixed (i.e., $p$ places probability 1 on $L = 1$), and only address varying $K$.
-As discussed above, each value of the constraint $K$ determines a quantile probability level $\alpha$ corresponding to the Bayes act such that $K = F_1^{-1}(\alpha) + F_2^{-1}(\alpha) = -\log(1 - \alpha) (\sigma_1 + \sigma_2)$; solving for $\alpha$, we obtain $\alpha = 1 - \exp[-K/(\sigma_1 + \sigma_2)]$.
-This link between the constraint level $K$ and the probability level $\alpha$ is the key to the link between the IAS and the CRPS.
-
-%We use this link to explore the relationship between IAS and CRPS from two directions.
-First, suppose the decision maker has some uncertainty about the value of $K$, which they express through $p$.
-Because $\alpha$ can be regarded as a function of $K$, this distribution on the resource constraint induces a distribution on quantile levels.
-Figure *** panel (b) \elr{TODO, figures from crps\_connection\_exp} illustrates with a $\text{Gamma}(500, 0.01)$ distribution for $K$, and the implied distribution on quantile levels is shown in panel (c).
-The IAS determined by $p$ corresponds to a weighted CRPS with this induced weighting on quantile levels.
-However, note that this weighting is specific to this pair of Exponential forecasts; a different pair of forecasts would translate to a different weighting on quantile levels.
-
-%Figure *** panel (d) \elr{TODO, figures from crps\_connection\_exp} illustrates this link by going in the other direction: given a forecast $F$, we exhibit the distribution on $K$ that would lead to equally weighted CRPS.
-Now we use the fact that $K$ can be written as a function of $\alpha$ to obtain the distribution on $K$ that corresponds to a $\text{Uniform}(0,1)$ distribution on $\alpha$.
-In this example, the implied distribution is $K \sim \text{Exp}(\sigma_1 + \sigma_2)$.
-We observe that this is a right-skewed distribution that places much of its mass on constraint values near 0, which may not correspond well to actual knowledge about the resource constraints.
-Again, the distribution on resource constraints that corresponds to unweighted CRPS depends on the forecast distributions.
 
@@
 We consider these and other generalizations in other work.
-
-% \subsection{Solving the Allocation Problem}
-
-% As the solution to a constrained optimization problem, the optimal allocation $\bf{x}^F$ according to a forecast $F$ can be described formally as the first component of a stationary point $(\bf{x}^F, \lambda^{\star})$ in the region $\{\mathbf{x} \geq 0, \lambda \geq 0\}$ for the Lagrangian
-% \begin{align}
-% \mathcal{L}(\mathbf{x},\lambda; K, L) = \sum_{i=1}^{N} \mathbb{E}_{F_i}[s_A(x_i, Y_i; L)] + \lambda\left(\sum_{i=1}^{N} x_i - K\right).
-% \end{align}
-% That is, optimality requires that both
-% \begin{align}
-% 0 = \nabla_{x} \mathcal{L}(\mathbf{x}^F, \lambda) &= \sum_{i=1}^{N} E_F \left[\frac{d}{dx_i}(Y_i - x_i^F)_{+}\right] + \lambda \\
-% & = \sum_{i=1}^{N} E_F \left[-\mathbf{1}\{Y_i \geq x_i^F\}\right] + \lambda  = \sum_{i=1}^{N} F_i(x_i^F)-1 + \lambda  \label{eqn:x_stationary}\\
-% \text{ and} \\
-% 0 = \nabla_{\lambda} \mathcal{L}(\mathbf{x}^F, \lambda^{\star}) &= \sum_{i=1}^{N} x_i^F - K\\
-% & = \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star}) - K \quad \text{ (by \eqref{eqn:x_stationary})}. \label{eqn:lambda_stationary}
-% \end{align}
-% These equations give formulae $x_i^F = F_i^{-1}(1-\lambda^{\star})$, but in many situations of interest for us the $F_i^{-1}$ are not single-valued, and even 
-% when they are, a solution $\lambda^{\star}$ of \eqref{eqn:lambda_stationary} can only be obtained analytically in special cases (but see \ref{sec:methods.ls.example}).
-We therefore typically calculate allocation scores using approximately optimal allocations found via a binary search for $\lambda^{\star}$.
-This search begins with the interval $\lambda \in [0,1]$ and
-% checks at each step the sign of the inequality that results from substituting the interval midpoint $\lambda_{\tau}$ into \eqref{eqn:lambda_stationary} for endpoints of the sets $F_i^{-1}(1-\lambda_{\tau})$.
-If $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) > K$, the quantiles for $1-\lambda_{\tau}$ exceed our resource constraint so we set $\lambda_{\tau}$ as the lower endpoint of the $\tau + 1$ search interval.
-Similarly, if $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) < K$, the quantiles for
-% $1-\lambda_{\tau}$ do not make full use of our resource and we set $\lambda_{\tau}$ as the upper endpoint of the $\tau + 1$ search interval.
-This process halves the search interval width at each step and we continue until this width falls below a given tolerance level $\varepsilon_{\tau}$ yielding an
-% approximation $\lambda^{\star, \varepsilon}$. When quantiles $F_i^{-1}(1-\lambda^{\star, \varepsilon})$ are singly defined and the forecast densities $f_i$
-% not too close to zero at these quantiles, the discrepancy $\Delta(\lambda^{\star, \varepsilon})  = \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star, \varepsilon}) - K$ is of order $\varepsilon_{\lambda}$ and we can take $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ as approximations to $x_i^F$.
-
-% Often though, $1-\lambda^{\star, \varepsilon}$ will be the probability level of a ``plateau'' for one or more $F_i$ and the $\lambda$ search algorithm will fail to drive $\Delta(\lambda^{\star, \varepsilon})$ close enough to zero for the $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ to be acceptable approximations of the optimal allocations.
-In this event we follow-up with a ``post-processing'' step that identifies lower and upper boundaries $x_{i,L}$ and $x_{i,U}$ of any
-% such low-density plateau containing points in $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$. If $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ is single-valued 
-% with the density $f_i(F_{i}^{-1}(1-\lambda^{\star, \varepsilon}))$ not close to zero, we simply take $x_{i,L} = F_{i}^{-1}(1-\lambda^{\star, \varepsilon}) = x_{i,U}$.
-Defining $\mathbf{x}_L = (x_{i,L})$ and $\mathbf{x}_U = (x_{i,U})$ we then interpolate an approximately optimal allocation
-% $\mathbf{x} = (1-t^{\star})\mathbf{x}_L + t^{\star}\mathbf{x}_U$ where $t^{\star}$ solves 
-% $\sum_{i=1}^{N }(1-t^{\star})x_{i,L} + t^{\star}x_{i,U} = K$.
-
-% We note though that by using this linear interpolation (rather than some other point in the plateau satisfying the constraint) as the allocation associated with $F$ in the scoring process we are making a choice that is not
-% guided by any expected utility maximization principle. As such, this choice may imply strong and potentially questionable assumptions about fairness and
-% equity principles for the decision maker when forecasts tend to be ``clumped,'' leaving many regions with very low forecasted probability.
 

### 2023-08-31 alloscore_manuscript/allocation.tex
fix some small typos

@@
 For example, forecasts for districts in Sierra Leone of bed demand to care for patients with Ebola was the subject of a real-time modeling study in late 2014 and early 2015 \citep{camacho2015-ebola-bed}.
-And, in 2020, a model developed by an academic research group turned predictions of COVID-19 hospitalizations into estimates of ventilator usage and and shortages.
+And, in 2020, a model developed by an academic research group turned predictions of COVID-19 hospitalizations into estimates of ventilator usage and shortages.
 This framework was used by the Hartford HealthCare system in Connecticut ``to align ventilator supply with projected demand at a time where the [COVID-19] pandemic was on the rise'' \citep{bertsimas2021predictionsCOVID}.
@@
 The remainder of this article is organized as follows. 
-In Section \ref{sec:methods}, we review the general framework for developing scoring rules for probabilistic forecasts using the tools of decision theory, develop a novel scoring rule that is motivated by the problem of allocating limited medical supplies, and explore the relationship between the proposed allocation score and existing scoring rules such as CRPS.
-In Section \ref{sec:application} we illustrate the scoring rule through an application to short-term forecasts of COVID-19 hospital admissions in the US. 
+We describe the allocation score in Section \ref{sec:methods}, and in Section \ref{sec:application} we illustrate the use of the score in an application to evaluate short-term forecasts of COVID-19 hospital admissions in the US.
 Section \ref{sec:discussion} summarizes our contributions and discusses opportunities for further extensions in future work.
@@
 We begin with an informal description of the allocation score and some examples illustrating its key characteristics in section \ref{sec:methods.overview}.
-In section \ref{sec:methods.detailed} we develop the allocation score more carefully, building on decision theoretic procedures for deriving proper scoring rules.
-We then comment on some connections between the allocation score that we propose and other common scores that can be derived from decision theoretic foundations, such as the quantile score, WIS, and CRPS, in section \ref{sec:methods.related}.
+In section \ref{sec:methods.detailed} we develop the allocation score more carefully, building on decision theoretic procedures for deriving proper scoring rules.We then comment on some connections between the allocation score that we propose and other common scores that can be derived from decision theoretic foundations, such as the quantile score, WIS, and CRPS, in section \ref{sec:methods.related}.
 
@@
 If the decision maker is provided with a multivariate forecast $F$ where each marginal forecast distribution $F_i$ predicts resource need in a particular location, one option is to choose the resource allocation that minimizes the expected total unmet need according to the forecast.
-We will give a more precise mathematical statement in section \ref{sec:methods.allocation}, but informally, the total expected unmet need according to the forecast is
+We will give a more precise mathematical statement in section \ref{sec:methods.detailed}, but informally, the total expected unmet need according to the forecast is
 \begin{align}
@@
 \end{align}
-where $F_i$ is the marginal forecast distribution of need in location $i$ and the unmet need in a particular location is the difference between resource need in that location and the number of resources that were allocated there.
+where the unmet need in a particular location is the difference between resource need in that location and the number of resources that were allocated there.
 This allocation problem has an intuitively appealing solution: allocate so that the probabilities of need exceeding allocation in various locations are as close to each other as possible.
@@
 Diagonal black lines indicate resource constraints at $K=5$ and $K=10$ units; any point along those lines corresponds to an allocation that meets the resource constraint.
-For these forecasts, the optimal allocations are $(1, 4)$ for $K=5$ and $(2, 8)$ at $K=10$.
+For these forecasts, the optimal allocations are $(1, 4)$ for $K=5$ and $(2, 8)$ for $K=10$.
 These allocations are at the point on the constraint line where the expected loss is smallest, which also corresponds to the point where a level set of the expected loss surface (blue curve) is tangent to the constraint.}
@@
 Note the way in which these forecasts incurred a positive (i.e., non-optimal) allocation score of 1 when $K = 10$.
-It was not directly directly due to individual misalignments of the marginal forecasts $F_i$ with the observed needs, but rather because the allocations and observed needs were not proportional as vectors.
+It was not directly due to individual misalignments of the marginal forecasts $F_i$ with the observed needs, but rather because the allocations and observed needs were not proportional as vectors.
 Restating: as far as allocation decisions are concerned, with a fixed constraint $K=10$ the fundamental problem with the forecast $F$ in Example 1 is not that it predicts a mean total resource need of $5$ units; it is that the realized need was 10 times as large in location 2 as in location 1, but the forecast only indicated that the resource allocation for location 2 should be 4 times the allocation for location 1.
@@
 
-We give a high-level review of a general procedure for developing proper scoring rules that are tailored to specific decision making tasks in section \ref{sec:methods.detailed.decisiontheory}, and then in section \ref{sec:methods.detailed.allocation} we apply that procedure to develop the allocation score based on the task of deciding on how to allocate a fixed supply of resources across multiple locations.
+We give a high-level review of a general procedure for developing proper scoring rules that are tailored to specific decision making tasks in section \ref{sec:methods.detailed.decisiontheory}, and then in section \ref{sec:methods.detailed.specific_allocation} we apply that procedure to develop the allocation score based on the task of deciding on how to allocate a fixed supply of resources across multiple locations.
+In \ref{sec:methods.detailed.integrated_allocation} we consider a small extension where the resource constraint is not known, or it is desired to consider the value of forecasts across a range of decision making scenarios.
+This gives rise to the \emph{integrated allocation score}.
 
@@
 
-\subsection{Integrating the allocation score across resource constraint levels}
+\subsubsection{Integrating the allocation score across resource constraint levels}
 \label{sec:methods.detailed.integrated_allocation}
@@
 These decision making problems differ according to the challenge faced by the decision maker: a fixed constraint on the available resources for the allocation problem, or a cost per unit of resources in the resource purchasing problem.
-In fact, it is possible to combine these into a more general problem where the decision maker must decide on both the total level of resources to purchase and an allocation of those resources across multiple locations, subject to a cost per unit of resources and a constraint on the total resources that can be purchased.
-The quantile loss and the allocation score presented in this work are both be special cases of the score that emerges from that more general decision making problem.
+In fact, it is possible to combine these into a more general problem where the decision maker must decide on both the total level of resources to purchase and an allocation of those resources across multiple locations, subject to a cost per unit of resources and a constraint on the total quantity of resources that can be purchased.
+The quantile loss and the allocation score presented in this work are both special cases of the score that emerges from that more general decision making problem.
 We pursue this direction further in other work (cite arxiv version of paper 2).

### 2023-08-31 alloscore_manuscript/allocation.tex
some simplifications near eqn for bayes act

@@
 \begin{align}
-    \mathbf{x}^F &= \mathrm{arg min}\,\, \sbar_A^F(\mathbf{x}; L) \text{ subject to }
-    0 \leq x_i \text{ for all $i$ and} \, \sum_{i=1}^N x_i = K \text{, where} \label{AP} \\
-    \sbar_A^F(\mathbf{x}; L) &= \mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y}; L)] = \sum_{i=1}^{N} L \mathbb{E}_{F_i}[(Y_i - x_i)_+] \label{eqn:sbar}
+    \mathbf{x}^F &= \mathrm{arg min}_{\mathbf{x}}\,\, \mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y}; L)] \text{ subject to }
+    0 \leq x_i \text{ for all $i$ and} \, \sum_{i=1}^N x_i = K, \label{AP}
 \end{align}
-It can be shown that with the loss function given in Equation \eqref{eqn:loss_fn}, the Bayes act has $x^F_i = F_i^{-1}(1 - \lambda^{\star}/L)$, where $\lambda^{\star}$ depends on the problem parameters $K$ and $L$ as well as the forecast distributions and is chosen so as to satisfy the equation
+where $\mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y}; L)] = \sum_{i=1}^{N} L \mathbb{E}_{F_i}[(Y_i - x_i)_+]$ sums the expected loss due to unmet need across all locations.
+It can be shown that the Bayes act has $x^F_i = F_i^{-1}(1 - \lambda^{\star}/L)$, where $\lambda^{\star}$ depends on the problem parameters $K$ and $L$ as well as the forecast distributions and is chosen so as to satisfy the equation
 \begin{align}
@@
 This partial solution to the allocation problem seems to have first appeared in \cite{hadleywhitin1963}; see the supplemental materials for a derivation.
-%Figure *** panel (a) \elr{TODO: first figure from notes crps\_connection\_exp.Rmd} illustrates the expected loss function $\sbar_A^F$ and the allocation given by the Bayes act in a simple example with $n = 2$ locations, $L = 1$, $K = 5$, and forecasts $Y_1 \sim \text{Exp}(1 / \sigma_1)$ and $Y_2 \sim \text{Exp}(1 / \sigma_2)$, using a scale parameterization of the exponential distribution with $\sigma_1 = 1$ and $\sigma_2 = 5$.
 
@@
 According to this interpretation, we divide the available resources among the locations in such a way that the forecasted probability that need will exceed the available resources is the same in all locations.
-An alternative interpretation comes from noting that for each location $i$, $\frac{\partial}{\partial x_i} \sbar_A^F(\mathbf{x}) \vert_{\mathbf{x}=\mathbf{x}^F} = \lambda^{\star}$ (see the supplement for a proof).
-%$\frac{\partial}{\partial x_i} \sbar_F(\mathbf{x}) = L[F_i(x_i) - 1]$ (see the supplement for a derivation), and evaluating at the Bayes allocation we arrive at
-%\begin{align}
-%\frac{\partial}{\partial x_i} \sbar_F(\mathbf{x}) \vert_{\mathbf{x}=\mathbf{x}^F} &= L[F_i(x^F_i) - 1] = L[F_i(F_i^{-1}(1 - \lambda^{\star}/L)) - 1] = \lambda^{\star}.
-%\end{align}
+An alternative interpretation comes from noting that for each location $i$, $\frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y}; L)] \vert_{\mathbf{x}=\mathbf{x}^F} = - \lambda^{\star}$ (see the supplement for a proof).
 In words, at the allocation given by the Bayes act, the rate of change of the expected score as a function of the amount allocated to location $i$ is given by $\lambda^{\star}$.

### 2023-08-31 alloscore_manuscript/allocation.tex
clean up notation around L and K

@@
 \begin{equation}
-s_A(\mathbf{x}, \mathbf{y}; L) = \sum_{i=1}^N L (y_i - x_i)_+. \label{eqn:loss_fn}
+s_A(\mathbf{x}, \mathbf{y}) = \sum_{i=1}^N L (y_i - x_i)_+. \label{eqn:loss_fn}
 \end{equation}
@@
 \begin{align}
-    \mathbf{x}^F &= \mathrm{arg min}_{\mathbf{x}}\,\, \mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y}; L)] \text{ subject to }
+    \mathbf{x}^F &= \mathrm{arg min}_{\mathbf{x}}\,\, \mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y})] \text{ subject to }
     0 \leq x_i \text{ for all $i$ and} \, \sum_{i=1}^N x_i = K, \label{AP}
 \end{align}
-where $\mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y}; L)] = \sum_{i=1}^{N} L \mathbb{E}_{F_i}[(Y_i - x_i)_+]$ sums the expected loss due to unmet need across all locations.
+where $\mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y})] = \sum_{i=1}^{N} L \mathbb{E}_{F_i}[(Y_i - x_i)_+]$ sums the expected loss due to unmet need across all locations.
 It can be shown that the Bayes act has $x^F_i = F_i^{-1}(1 - \lambda^{\star}/L)$, where $\lambda^{\star}$ depends on the problem parameters $K$ and $L$ as well as the forecast distributions and is chosen so as to satisfy the equation
@@
 According to this interpretation, we divide the available resources among the locations in such a way that the forecasted probability that need will exceed the available resources is the same in all locations.
-An alternative interpretation comes from noting that for each location $i$, $\frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y}; L)] \vert_{\mathbf{x}=\mathbf{x}^F} = - \lambda^{\star}$ (see the supplement for a proof).
-In words, at the allocation given by the Bayes act, the rate of change of the expected score as a function of the amount allocated to location $i$ is given by $\lambda^{\star}$.
+An alternative interpretation comes from noting that for each location $i$, $\frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y})] \vert_{\mathbf{x}=\mathbf{x}^F} = - \lambda^{\star}$ (see the supplement for a proof).
+In words, at the allocation given by the Bayes act, the rate of change of the expected score as a function of the amount allocated to location $i$ is given by $-\lambda^{\star}$.
 This derivative is the same for all locations, so the optimal allocation divides the available resources across all locations in such a way that according to $F$, the expected benefit of 1 additional unit of resources is the same in all locations.
@@
 \begin{equation}
-S_A(F, y; L, K) = s_A(\mathbf{x}^F, y; L) = \sum_{i=1}^n L (y_i - F_i^{-1}(1 - \lambda^{\star}/L))_+
+S_A(F, y; K) = s_A(\mathbf{x}^F, y) = \sum_{i=1}^n L (y_i - F_i^{-1}(1 - \lambda^{\star}/L))_+
 \end{equation}
@@
 However, if $K$ is not precisely known at the time of decision making or there is interest in measuring the value of forecasts across a range of decision making scenarios with different resource constraints, we can use an \emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a distribution $p$:
-$$S_{IAS}(F, y; L) = \int S_A(F,y; L, K) p(K) \, dK$$
+$$S_{IAS}(F, y) = \int S_A(F,y; K) p(K) \, dK$$
 

### 2023-08-31 alloscore_manuscript/methods_apg.tex
change location index var from l to i

@@
 
-A somewhat open-ended attempt to motivate and construct quantile scores, WIS, CRPS, and allocation scores within a unified decision theoretic framwork.
+We develop a unified decision theoretic framwork for quantile scores, WIS, CRPS, and allocation scores.
 
@@
 
-In this section/report/paper/rambling manifesto, we present a methodological perspective on the use scoring rules for forecast evaluation that we see as well-adapted to our motivating research goal to understand and describe the social utility of infectious disease forecasts.
+In this report, we present a methodological perspective on the use scoring rules for forecast evaluation that we see as well-adapted to our motivating research goal to understand and describe the social utility of infectious disease forecasts.
 Our approach follows the common practice in decision theory of giving central roles to the foundational concepts of state, action, and result as the basic components of a decision problem.
 From these primitives, we first derive the familiar measures of quantile score (QS), continuous ranked probability score (CRPS) and their aggregates across forecast targets.
-We then introduce new scoring rules, the \emph{specific} and \emph{integrated allocation scores} designed to reflect aspects of the decision problem corresponding to the CRPS that emerge when multiple instances are confronted simultaneously.
-Specifically, we will consider not only whether using more or less of a resource leads to a better result in a specific problem, but also whether a limited resource is better allocated to one problem rather than another.
-This added concern brings into play the duality in mathematical optimization theory between constraints on ``decision variables" and the ``Lagrange multiplier variables" which quantify the sensitivity of solutions to constraints, that is, the degree to which loosening constraints improves results under the optimal action..
+We then introduce new scoring rules, the \emph{specific} and \emph{integrated allocation scores} designed to reflect aspects of the decision problem beyond those defining the QR and CRPS that emerge when multiple instances of a decision problem are confronted simultaneously.
+Specifically, we will consider not only whether using more or less of a costly resource leads to a better result in a specific problem, but also whether a limited resource is better allocated to one problem rather than another.
+This added concern brings into play the duality in mathematical optimization theory between constraints on decision variables on the one hand and dual or ``Lagrange multiplier'' variables on the other.
+Dual variables quantify the sensitivity of the optimal value of an objective function to constraints, that is, the degree to which loosening constraints improves results under the optimal action.
 Our computational methods (see ..) in fact make use of this connection.
@@
 
-Broadly speaking, we score a probabilistic forecast $F$ by identifying it with a solution to a \emph{decision problem} and rating the utility of that solution.
+Broadly speaking, we view scoring a probabilistic forecast $F$ with a scoring rule as the process of identifying $F$ with a solution to a \emph{decision problem} and then rating the utility of that solution.
 The term ``decision problem'' refers to a pre-specified set of allowed actions, possible future states, and the possible results of allowed actions under those states.
 Accordingly, $F$ will receive a ``better" or ``worse" score in a particular instance when the result is better or worse of a rational decision maker using $F$ (and essentially only $F$) to select an action in the face of uncertainty about a future state.
-It should be noted immediately that a better or worse score in a single instance may tell us almost nothing about the intrinsic value of the forecasting method that produced $F$.
-But as scores accumulate, we can use their distributions across forecasting instances to make inferences as to whether one forecasting method is better or worse than another in a decision theoretic sense.
+As such scores accumulate, we can use their empirical distributions across forecasting instances as inputs to higher level decision problems: whether a status quo policy should be adjusted according to the forecasts produced by a particular model or method; whether one forecasting method is a better choice to guide policy rather than another; how different forecasting methods can best be coordinated or combined to guide policy.
+By connecting such policies with the specific decision problems we use to define
+scoring rules, we may be able to improve these higher level decisions we make about how forecasts are used.
 
@@
 We also will refer to states as outcomes, which helps to emphasize their future realization.
-But we want to clearly distinguish an outcome from a result, which as we discuss below, which is determined by an action under a certain outcome.
+But we want to clearly distinguish an outcome from a result, which as we discuss below, is determined by an action being taken under a certain outcome.
 
@@
 A decision maker might use a set rule to select $x$ based on inputs, such as $F$, but the observed value of $y$ cannot be such an input.\footnote{In a dynamical decision theoretic approach, the action $x$ could be a time series depending at each step on previous values of $y$.
-This could be a very fruitful framework for infectious disease forecast evaluation but we do not pursue it is this work.}
+This could be a very fruitful framework for infectious disease forecast evaluation but we do not pursue it is this work.} We will also require that $y$ be observed before $x$ can effect $y$.
+Thus, if we were considering actions of providing vaccines or masks to protect a (likely estimated) number $y_e$ of people exposed to an observed number $y$ of infectious individuals, we would require $y$ be observed before the vaccines or masks had any chance to reduce disease incidence.
 
@@
 \begin{remark} For there to be an incentive \emph{not} to reveal a forecast other than $F$ we would need a \emph{strictly} proper scoring rule, which would require more conditions on the decision problem and/or $s$.
-In the literature, a functional $T(F)$ (such as the mean or median of $F$) is called \emph{elicitable} if there is a strictly proper scoring rule $S(F,y) = s(x^F,y)$ for which $T(F)=x^F$.
+In the literature, a functional $M(F)$ (such as the mean or median of $F$) is called \emph{elicitable} if there is a strictly proper scoring rule $S(F,y) = s(x^F,y)$ for which $M(F)=x^F$.
 But the focus in work on elicitability often seems to be the functional $T$ as a starting point rather than the decision problem, which is our primary concern here.
@@
 \end{align}
-that is, the quantile $x^F = F_l^{-1}(1 - C/L) = q_{F,\tau}$ for the probability level $\tau = 1-C/L$.
+that is, the quantile $x^F = F_i^{-1}(1 - C/L) = q_{F,\tau}$ for the probability level $\tau = 1-C/L$.
 (That the critical point $q_{F,\tau}$ is actually a minimum follows from \eqref{eqn:q_deriv} being non-positive to its left and non-negtive to its right, and assuming as we do for now that $F$ is smooth and strictly increasing at $q_{F,\tau}$, it is a unique minimum.) The quantile solves the decision problem with respect to the forecast $F$ by being the action which best balances expected cost against expected loss among all actions available to the decision maker.
@@
 The decision problem leading to the QS may however be complicated by uncertainty at time $t_0$ about the cost $C$ of the resource.
-We take such uncertainty as requiring the worker to commit in advance to procurement levels $x_i$ for a range of potential costs $C_i < L$ where $i \in 1, \ldots, m$.
+We take such uncertainty as requiring the worker to commit in advance to procurement levels $x_j$ for a range of potential costs $C_j < L$ where $j \in 1, \ldots, m$.
 Even though only one of the $x_i$ will end up needing to be procured, there does not seem to be any intrinsically decision theoretic principle by which our loss function should ignore counter-factual but exactly specified results of the decision under unrealized costs.
 If we wanted to commit to a \emph{locality} principle for prediction where losses and scores only depend on predictive densities in a neighborhood of the relevant observation, this attempt would falter, and we would need to frame our methods around the log score and it's higher order relatives.
-But we do not see that as an appropriate principle in public health resourse allocation where
+But we do not see that as an appropriate principle in public health resource allocation where
 probabilistic forecasts can be valuable by pushing decision makers in the "right direction" even when they assign low probability to observed outcomes.
@@
 
-We take into account the results of actions under a set $\mathcal{C} = \{C_i\}$ of possible costs by expanding the loss function to be a weighted mean of quantile losses under certainty about the cost:
+We take into account the results of actions under a set $\mathcal{C} = \{C_j\}$ of possible costs by expanding the loss function to be a weighted mean of quantile losses under certainty about the cost:
 \begin{align}
-s_{Q,\mathcal{C}, p_C}(x,y) = \frac{1}{m}\sum_{i=1}^{m}p_i(C_ix_i + L(x_i - y)_+). \label{eqn:sum_CL_loss}
+s_{Q,\mathcal{C}, p_C}(x,y) = \frac{1}{m}\sum_{j=1}^{m}p_j(C_j x_j + L(x_i - y)_+). \label{eqn:sum_CL_loss}
 \end{align}
-Here, $p_C = \{p_i\}, i = 1,\ldots,m$ is a prior distribution on $\mathcal{C}$ and $x = (x_1,\ldots,x_m)$ is the decision vector corresponding to all possible costs.
+Here, $p_C = \{p_j\}, j = 1,\ldots,m$ is a prior distribution on $\mathcal{C}$ and $x = (x_1,\ldots,x_m)$ is the decision vector corresponding to all possible costs.
 $p_C$ would be provided by the decision maker (or perhaps a separate expert), and we emphasize that we are considering it here as part of the loss function $s_{Q,\mathcal{C}, p_C}$ that would be formulated in an effort to adapt forecast evaluation to the specifics of a public health decision problem.
@@
 \begin{align}
-S_{Q,\mathcal{C}, p_C}(F,y) &= \frac{1}{m}\sum_{i=1}^{m}p_i(C_i q_{F,\tau_i} + L(q_{F,\tau_i} - y)_+) \\
+S_{Q,\mathcal{C}, p_C}(F,y) &= \frac{1}{m}\sum_{j=1}^{m}p_j(C_j q_{F,\tau_j} + L(q_{F,\tau_j} - y)_+) \\
 S_{Q, f_C}(F,y) &= \frac{1}{L}\int_{0}^{L}(cq_{F,\tau(c)} + L(q_{F,\tau(c)} - y)_+) f_C(c)dc. \label{eqn:CRPS_cform}
@@
 \begin{align}
-\mathrm{WIS}_{\alpha_{\{0: J\}}}(F, y)=\frac{1}{J+1 / 2}\left(w_0 |y-q_{F,.5}|+\sum_{j=1}^J w_j \operatorname{IS}_{\alpha_j}(F, y)\right).
+\mathrm{WIS}_{\alpha_{\{1: J\}}}(F, y)=\frac{1}{J+1 / 2}\left(w_0 |y-q_{F,.5}|+\sum_{j=1}^J w_j \operatorname{IS}_{\alpha_j}(F, y)\right).
 \end{align}
@@
 \end{align}	
-with $w_j = p_{j}\alpha_j/2$ and $w_0 = p_{L/2}$.
+with $w_j = p_{j}\alpha_j/2$ and $w_0 = p_{L/2}$. Setting $p_j=1/m$ and $\alpha_j=0.9, \ldots, 0.2,0.1,0.05,0.02, j = 1,\ldots,11$ gives the WIS currently used in 
+the COVID-19 Forecast Hub.
 
@@
 And again, summing the losses  \eqref{eqn:quantile_loss} over the locations (for a particular $C$ and $L$) creates a loss function that is reactive to the overall procurement decision.
-Let the locations be indexed by $l \in 1,\ldots,N$.
-The action sets for the decision problem corresponding to a single certain cost $C$, a discrete cost distribution over $\{C_i\}, i = 1,\ldots,m$, and a continuous cost distribution $f_C$ with support on $[0,L]$ are now
+Let the locations be indexed by $i \in 1,\ldots,N$.
+The action sets for the decision problem corresponding to a single certain cost $C$, a discrete cost distribution over $\{C_j\}, j = 1,\ldots,m$, and a continuous cost distribution $f_C$ with support on $[0,L]$ are now
 \begin{align}
 \mathcal{X}_{cert} &= \mathbb{R}_{+}^N \\
-\mathcal{X}_{disc} &= (\mathbb{R}_{+}^m)^N = \{x(i):\{1,\ldots,m\} \to \mathbb{R}_{+}^N\}\\
+\mathcal{X}_{disc} &= (\mathbb{R}_{+}^m)^N = \{x(j):\{1,\ldots,m\} \to \mathbb{R}_{+}^N\}\\
 \mathcal{X}_{cont} &= \{x(c):[0,\infty) \to \mathbb{R}_{+}^N\}.
@@
 \begin{align}
-s_{TQ}(x,y) &= \sum_{l=1}^N Cx_l + L(y_l-x_l)_+ \label{eqn:TQ_loss} \\
-s_{TQ, \mathcal{C}, p_C}(x,y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}p_i(C_i x_{l,i} + L(x_{l,i} - y_l)_+)\\
-s_{TQ, f_C}(x,y) &= \sum_{l=1}^N \int_{0}^{L} (cx_l(c) + L(x_l(c) - y_l)_+) f_C(c)dc.
+s_{TQ}(x,y) &= \sum_{i=1}^N Cx_i + L(y_i-x_i)_+ \label{eqn:TQ_loss} \\
+s_{TQ, \mathcal{C}, p_C}(x,y) &= \frac{1}{m}\sum_{i=1}^N \sum_{j=1}^{m}p_j(C_i x_{i,j} + L(x_{i,j} - y_i)_+)\\
+s_{TQ, f_C}(x,y) &= \sum_{i=1}^N \int_{0}^{L} (cx_i(c) + L(x_i(c) - y_i)_+) f_C(c)dc.
 \end{align}
-A key point for our presentation will be that the same $C$, set $\{C_i\}$, probabilities $\{p_i\}$, and density $f_C$ are used for all locations simultaneously.
+A key point for our presentation will be that the same $C$, set $\{C_j\}$, probabilities $\{p_j\}$, and density $f_C$ are used for all locations simultaneously.
 We believe that our methods could be adapted to variations across locations (which could be important in practice), but we leave this generalization for later work.
@@
 
-Consider again the resource procurement decision $x(c) \in \mathcal{X}_{cont}$ (or $\mathcal{X}_{disc}$ if $c = C_i, i \in 1,\ldots,m$) a public health worker must make in 
-attempting to mitigate a pandemic across multiple locations $l \in 1,\ldots,N$.
-But now we ask the worker to extend their decision to cover the contingency that there turns out to only be a total amount $k$ of the resource available for allocation, less than the amount $\sum x_l(C)$ that they would choose to distribute if balancing only the realized cost $C$ against potential unmet demand.
+Consider again the resource procurement decision $x(c) \in \mathcal{X}_{cont}$ (or $\mathcal{X}_{disc}$ if $c = C_j, j \in 1,\ldots,m$) a public health worker must make in 
+attempting to mitigate a pandemic across multiple locations $i \in 1,\ldots,N$.
+But now we ask the worker to extend their decision to cover the contingency that there turns out to only be a total amount $k$ of the resource available for allocation, less than the amount $\sum x_i(C)$ that they would choose to distribute if balancing only the realized cost $C$ against potential unmet demand.
 This could be due to demand-side factors such as a budget cut, or supply-side factors such as manufacturing or shipping failures.
@@
 at time $t_1$.
-This adds, for each $k \in \mathbb{R}_{+}$, a component to our overall decision problem that asks the question: If the total amount $K$ available of the resource is insufficient allow the procurements $x_l(C)$ you specify for the realized cost $C$, what alternate procurement levels $x_l^{k,C}$ would you choose?
-Writing $T_x = \sum_{l=1}^{N} x_l$, the action set for this new component is
+This adds, for each $k \in \mathbb{R}_{+}$, a component to our overall decision problem that asks the question: If the total amount $K$ available of the resource is insufficient allow the procurements $x_i(C)$ you specify for the realized cost $C$, what alternate procurement levels $x_i^{k,C}$ would you choose?
+Writing $T_x = \sum_{i=1}^{N} x_i$, the action set for this new component is
 \begin{align}
@@
 \end{align}
-Note that because this decision is only relevant when $\sum x_l(C)>k$ we do not include actions which use less than $k$ total of the resource. (Why?)
+Note that because this decision is only relevant when $\sum x_i(C)>k$ we do not include actions which use less than $k$ total of the resource. (Why?)
 
@@
 \begin{align}
-s_{tQ}\left(x, y\right) = Ck + \sum_{l=1}^{N} L(y_l - x_l)_{+}.
+s_{tQ}\left(x, y\right) = Ck + \sum_{i=1}^{N} L(y_i - x_i)_{+}.
 \end{align}
@@
 \begin{align}
-s_{A}(x,y):= L\sum_{l=1}^{N} (y_l - x_l)_{+},
+s_{A}(x,y):= L\sum_{i=1}^{N} (y_i - x_i)_{+},
 \end{align}
@@
 % We now add a constraint $K$ on the total amount of the resource available for allocation across all locations.
-This presents the worker with a new decision problem we call the \emph{allocation problem} (AP or $\mathrm{AP}(K)$) with a smaller action set $\mathcal{X}_K = \{x \in \mathbb{R}_{+}^N | \sum x_l \leq K\} \subset \mathcal{X}_{cert}\}$ from which to choose (and analogously restricted version of the sets $\mathcal{X}_{disc}$ and $\mathcal{X}_{cont}$ for the decision problems with uncertain cost).
+This presents the worker with a new decision problem we call the \emph{allocation problem} (AP or $\mathrm{AP}(K)$) with a smaller action set $\mathcal{X}_K = \{x \in \mathbb{R}_{+}^N | \sum x_i \leq K\} \subset \mathcal{X}_{cert}\}$ from which to choose (and analogously restricted version of the sets $\mathcal{X}_{disc}$ and $\mathcal{X}_{cont}$ for the decision problems with uncertain cost).
 Loss functions with the basic $Cx + L(y-x)_{+}$ structure we have used up until now continue to generate meaningful scoring rules with respect to the AP following our procedure.
-But with a view toward developing a scoring rule that accommodates uncertainty about both $C$ and $K$ in a non-redundant (and more interpretable and computationally feasible) way, we augment the action set of the AP to include, in addition to the procurement levels $x_l$ for all locations, a ``reserve'' quantity $z_K$ defined as the available amount of resource that the worker decides not to allocate anywhere.
+But with a view toward developing a scoring rule that accommodates uncertainty about both $C$ and $K$ in a non-redundant (and more interpretable and computationally feasible) way, we augment the action set of the AP to include, in addition to the procurement levels $x_i$ for all locations, a ``reserve'' quantity $z_K$ defined as the available amount of resource that the worker decides not to allocate anywhere.
 % In optimization theory quantities defined in this manner are known as \emph{slack} variables, and are introduced in order to re-express contraint inequalities as equalities.  In our case, we get
 % \begin{align}
-% \mathcal{X}_{K} = \{(x,z) \in \mathbb{R}_{+}^{N+1} \mid \sum_{l=1}^{N}x_l + z = K \}.
+% \mathcal{X}_{K} = \{(x,z) \in \mathbb{R}_{+}^{N+1} \mid \sum_{i=1}^{N}x_i + z = K \}.
 % \end{align}
@@
 % \begin{align}
-% s_{tQ}\left((x,z), y\right) = C(K-z) + \sum_{l=1}^{N} L(y_l - x_l)_{+}.
+% s_{tQ}\left((x,z), y\right) = C(K-z) + \sum_{i=1}^{N} L(y_i - x_i)_{+}.
 % \end{align}
@@
 % \begin{align}
-% s_{A,K}(x,y):= L\sum_{l=1}^{N} (y_l - x_l)_{+}.
+% s_{A,K}(x,y):= L\sum_{i=1}^{N} (y_i - x_i)_{+}.
 % \end{align}
@@
 
-% Picture: For $K$ fixed, $x^{F,K}(c)$ will be a constant point $x^{F,K}$ on $\{\sum x_l = K\}$ for $c \in [0, c_1]$, 
-% and then wander back to 0 for $c \in (c_1,L]$. When $C$ is fixed, $x^{F}(k)$ will be a point on $\{\sum x_l = k\}$ (which does not depend on $C$)
-% for $k \in [0, \sum F_l^{-1}(1-C/L)]$, and then constant at $\mathbf{F}^{-1}(1-C/L)$ for $k \in (\sum F_l^{-1}(1-C/L), \infty)$.
+% Picture: For $K$ fixed, $x^{F,K}(c)$ will be a constant point $x^{F,K}$ on $\{\sum x_i = K\}$ for $c \in [0, c_1]$, 
+% and then wander back to 0 for $c \in (c_1,L]$. When $C$ is fixed, $x^{F}(k)$ will be a point on $\{\sum x_i = k\}$ (which does not depend on $C$)
+% for $k \in [0, \sum F_i^{-1}(1-C/L)]$, and then constant at $\mathbf{F}^{-1}(1-C/L)$ for $k \in (\sum F_i^{-1}(1-C/L), \infty)$.
 
@@
 % \begin{align}
-% s_{t\mathrm{WIS}}(x^{F,t\mathrm{WIS}},y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}(x_{l,i}^{F_l,\mathrm{Q}, C_i}- y_l)_+  + C_i x_{l,i}^{F_l,\mathrm{Q}, C_i}\\ 
+% s_{t\mathrm{WIS}}(x^{F,t\mathrm{WIS}},y) &= \frac{1}{m}\sum_{i=1}^N \sum_{i=1}^{m}(x_{l,i}^{F_i,\mathrm{Q}, C_i}- y_i)_+  + C_i x_{l,i}^{F_i,\mathrm{Q}, C_i}\\ 
 % &= \frac{1}{m} \sum_{i=1}^{m}s_{\mathrm{AS}}(x^{F,\mathrm{AS},  K(C_i, F, L)},y) + C_i K(C_i, F, L)
@@
 % s_{\mathrm{IAS}}(x^{F, \mathrm{IAS}}(k), y) &= \int_0^{\infty} s_{\mathrm{AS}}(x^{F, \mathrm{AS}(k)}, y) dk \\
-% &= \int_0^{\infty} \sum_{l=1}^N s_{\mathrm{Q}}(x^{F_l, \mathrm{Q}(c(k, F))}, y) - c(k, F)kdk \\
-% &= \sum_{l=1}^N \int_0^{\infty} s_{\mathrm{Q}}(x^{F_l, \mathrm{Q}(c(k, F))}, y) dk -  \int_0^{\infty} c(k, F)kdk
+% &= \int_0^{\infty} \sum_{i=1}^N s_{\mathrm{Q}}(x^{F_i, \mathrm{Q}(c(k, F))}, y) - c(k, F)kdk \\
+% &= \sum_{i=1}^N \int_0^{\infty} s_{\mathrm{Q}}(x^{F_i, \mathrm{Q}(c(k, F))}, y) dk -  \int_0^{\infty} c(k, F)kdk
 % \end{align}
@@
 % \begin{cases}
-% \sum_{l=1}^{N}L(y_l - x_l^{F,k})_{+} + ck, & k < \sum F_l^{-1}(1-c/L) \\
-% \sum_{l=1}^{N}L(y_l - x_l^{F,Q,c})_{+} + c(k - z_k(c)), & \text{otherwise, with } 
+% \sum_{i=1}^{N}L(y_i - x_i^{F,k})_{+} + ck, & k < \sum F_i^{-1}(1-c/L) \\
+% \sum_{i=1}^{N}L(y_i - x_i^{F,Q,c})_{+} + c(k - z_k(c)), & \text{otherwise, with } 
 % \end{cases} \\
-% z_k(c) &= (k -  \sum F_l^{-1}(1-c/L))_{+}
+% z_k(c) &= (k -  \sum F_i^{-1}(1-c/L))_{+}
 % \end{align}
 
-% Let $k_1 = \sum F_l^{-1}(1-c_1/L)$ so that $x^{F,Q,c_1} = x^{F, k_1}$ and $z_{k_1}(\{c \leq c_1\}) = 0$, and take $0<c_a < c_1 < c_b < L$ and $0 < k_a < k_1 < k_b < K_2$.
+% Let $k_1 = \sum F_i^{-1}(1-c_1/L)$ so that $x^{F,Q,c_1} = x^{F, k_1}$ and $z_{k_1}(\{c \leq c_1\}) = 0$, and take $0<c_a < c_1 < c_b < L$ and $0 < k_a < k_1 < k_b < K_2$.
 
@@
 % \begin{align}
-% s_{k_1}(0)  &= L\sum ( y_l-x_l^{F,k_1})_{+}  \\
-% s_{k_1}(c_a)  &= L\sum ( y_l-x_l^{F,k_1})_{+} + c_a k_1\\
-% s_{k_1}(c_1) &= L\sum (y_l - x_l^{F,k_1})_{+} + c_1 k_1  \\
-% s_{k_1}(c_b) &= L\sum (y_l - x_l^{F,Q, c_b})_{+} + c_b(k_1 - z_{k_1}(c_b)) \\
-% &= L\sum (y_l - x_l^{F,Q, c_b})_{+} + c_b \sum F_l^{-1}(1-c_b/L)\\
-% s_{k_1}(L)  &= L\sum {y_l}_{+}.
+% s_{k_1}(0)  &= L\sum ( y_i-x_i^{F,k_1})_{+}  \\
+% s_{k_1}(c_a)  &= L\sum ( y_i-x_i^{F,k_1})_{+} + c_a k_1\\
+% s_{k_1}(c_1) &= L\sum (y_i - x_i^{F,k_1})_{+} + c_1 k_1  \\
+% s_{k_1}(c_b) &= L\sum (y_i - x_i^{F,Q, c_b})_{+} + c_b(k_1 - z_{k_1}(c_b)) \\
+% &= L\sum (y_i - x_i^{F,Q, c_b})_{+} + c_b \sum F_i^{-1}(1-c_b/L)\\
+% s_{k_1}(L)  &= L\sum {y_i}_{+}.
 % \end{align}
@@
 % \begin{align}
-% s_{c_1}(0) & = L \sum {y_l}_{+} \\
-% s_{c_1}(k_a) & = L \sum (y_l - x_l^{F,k_a})_{+} + c_1 k_a\\
-% & = L \sum (y_l - x_l^{F,Q,c(k_a)})_{+} + c_1 \sum F_l^{-1}(1-c(k_a)/L)\\
-% s_{c_1}(k_1) & = L \sum (y_l - x_l^{F,k_1})_{+} + c_1 k_1 \\
-% s_{c_1}(k_b) & = L \sum (y_l - x_l^{F,k_1})_{+} + c_1 (k_b - z_{k_b}(c_1)) \\
-%  & = L \sum (y_l - x_l^{F,k_1})_{+} + c_1 \sum F_l^{-1}(1-c(k_1)/L)\\
-% s_{c_1}(K_2) & = L \sum (y_l - x_l^{F,k_1})_{+} + c_1 (K_2 - z_{K_2}(c_1))
+% s_{c_1}(0) & = L \sum {y_i}_{+} \\
+% s_{c_1}(k_a) & = L \sum (y_i - x_i^{F,k_a})_{+} + c_1 k_a\\
+% & = L \sum (y_i - x_i^{F,Q,c(k_a)})_{+} + c_1 \sum F_i^{-1}(1-c(k_a)/L)\\
+% s_{c_1}(k_1) & = L \sum (y_i - x_i^{F,k_1})_{+} + c_1 k_1 \\
+% s_{c_1}(k_b) & = L \sum (y_i - x_i^{F,k_1})_{+} + c_1 (k_b - z_{k_b}(c_1)) \\
+%  & = L \sum (y_i - x_i^{F,k_1})_{+} + c_1 \sum F_i^{-1}(1-c(k_1)/L)\\
+% s_{c_1}(K_2) & = L \sum (y_i - x_i^{F,k_1})_{+} + c_1 (K_2 - z_{K_2}(c_1))
 % \end{align}
 
-% Writing $k(c) = \sum F_l^{-1}(1-c/L)$,
+% Writing $k(c) = \sum F_i^{-1}(1-c/L)$,
 % \begin{align}
-% s_{k_1}(c) - s_{c_1}(k(c)) &= (c-c_1)\sum F_l^{-1}(1-\max(c,c_1)/L) \\
+% s_{k_1}(c) - s_{c_1}(k(c)) &= (c-c_1)\sum F_i^{-1}(1-\max(c,c_1)/L) \\
 % s_{c_1}(k) - s_{k_1}(c(k)) &=  
@@
 % \begin{align}
-% \del{k}{c}& = \sum_{l=1}^{N} \frac{-1}{L f_l(q_{F_l,\tau(c)})} := -\mathrm{MAI}(c,F)\label{eqn:dkdc}
+% \del{k}{c}& = \sum_{i=1}^{N} \frac{-1}{L f_i(q_{F_i,\tau(c)})} := -\mathrm{MAI}(c,F)\label{eqn:dkdc}
 % \end{align}
-% (where we assume that the $F_l$ have non-negative support).
+% (where we assume that the $F_i$ have non-negative support).
 The name MAI refers to this derivative being the marginal total allocation increase for the total quantile loss Bayes act $x^{F,Q(C)}$ that results from a reduction of the cost $C$ of the resource by 1.
@@
 % s_{\mathrm{IAS}}(x^{F, \mathrm{IAS}}(k), y) 
-% &= \sum_{l=1}^N \int_0^{L} s_{\mathrm{Q}}(x^{F_l, \mathrm{Q}(c)}, y) \mathrm{MAI}(c,F) dc \\
-% &= \sum_{l=1}^N \int_0^{1} s_{\mathrm{Q}}(x^{F_l, \tau}, y) L\mathrm{MAI}(L(1-\tau),F) d\tau \\
+% &= \sum_{i=1}^N \int_0^{L} s_{\mathrm{Q}}(x^{F_i, \mathrm{Q}(c)}, y) \mathrm{MAI}(c,F) dc \\
+% &= \sum_{i=1}^N \int_0^{1} s_{\mathrm{Q}}(x^{F_i, \tau}, y) L\mathrm{MAI}(L(1-\tau),F) d\tau \\
 % &= tw\mathrm{CRPS}(F,y)
@@
 &= E_F\left[s_A(Y,x)\right] + \lambda(T_x - k) \\
-&= L\sum_{l=1}^{N} E_{F_l}[(Y_l - x_l)_{+}] + \lambda\left(\sum_{l=1}^{N} x_l - k\right).  
+&= L\sum_{i=1}^{N} E_{F_i}[(Y_i - x_i)_{+}] + \lambda\left(\sum_{i=1}^{N} x_i - k\right).  
 \end{align}
-in the region $\{x \geq 0, \lambda \in \mathbb{R}\}$.\footnote{The Lagrangian in this context has the interpretation as the best (lowest) objective function value we can achieve if we forced to choose $x$ but can buy or get credit for changing the constraint level $k$
+in the region $\{x \geq 0, \lambda \in \mathbb{R}\}$.\footnote{The Lagrangian in this context has the interpretation as the best (lowest) objective function value we can achieve if we forced to choose $x$ but can buy or get credit for changes the constraint level $k$
 at the ``price'' $\lambda$.}
@@
 \begin{align}
-0 = \del{}{\lambda} \mathcal{L}(x^{F,k}, \lambda^{\star}) &= \sum_{l=1}^{N} x_l^{F,k} - k. \label{eqn:lambda_stationary}
+0 = \del{}{\lambda} \mathcal{L}(x^{F,k}, \lambda^{\star}) &= \sum_{i=1}^{N} x_i^{F,k} - k. \label{eqn:lambda_stationary}
 \end{align}
@@
 \begin{align}
-0 = \frac{\partial}{\partial x_l} \mathcal{L}(x^{F.k}, \lambda) &= L(F_l(x_l^{F,k})-1) + \lambda^{\star} 
+0 = \frac{\partial}{\partial x_i} \mathcal{L}(x^{F.k}, \lambda) &= L(F_i(x_i^{F,k})-1) + \lambda^{\star} 
 \quad (\text{cf. \eqref{eqn:q_deriv}}) \label{eqn:x_stationary}
 \end{align}
-for any $x_l^{F,k} > 0$, which gives a formula 
+for any $x_i^{F,k} > 0$, which gives a formula 
 \begin{align}
-x_l^{F,k} = F_l^{-1}(1-\lambda^{\star}/L) \label{eqn:xfk_formula}
+x_i^{F,k} = F_i^{-1}(1-\lambda^{\star}/L) \label{eqn:xfk_formula}
 \end{align}
-for any non-zero element $x_l^{F,k}$ of the solution. Thus \eqref{eqn:lambda_stationary} becomes
+for any non-zero element $x_i^{F,k}$ of the solution. Thus \eqref{eqn:lambda_stationary} becomes
 \begin{align}
-0 = \sum_{\{l | x_l^{F,k} > 0\}} F_{l}^{-1}(1-\lambda^{\star}/L) - k. \label{eqn:lambda_stationary2}
+0 = \sum_{\{l | x_i^{F,k} > 0\}} F_{l}^{-1}(1-\lambda^{\star}/L) - k. \label{eqn:lambda_stationary2}
 \end{align}
@@
 Since convexity and compactness guarantee existence of a solution, the problem is now reduced to solving \eqref{eqn:lambda_stationary} for the set 
-$\{l \mid x_l^{F,k} > 0\}$ and value $\lambda^{\star}$. Note, however, that this is a more complicated task than solving just 
+$\{l \mid x_i^{F,k} > 0\}$ and value $\lambda^{\star}$. Note, however, that this is a more complicated task than solving just 
 for $\lambda^{\star}$ under the assumption that \eqref{eqn:xfk_formula} holds for all coordinates of $x^{F,k}$.\footnote{Such an assumption would exclude fairly simple examples in which $F$ has support bounded away from $0$ in one coordinate but not in another.} The task is simplified though by bringing in the full saddle point condition for $\lambda^{\star}$ which implies not only \eqref{eqn:lambda_stationary} but also that
@@
 $\mathcal{L}_D(\lambda) := \inf_{x \in \mathbb{R}_{+}^N} \mathcal{L}(x, \lambda)$.
-This approach is especially suitable in our situation because our original objective function $E_F\left[s_A(Y,x)\right]$ is decomposable in the coodinates $x_l$, allowing the dual function to be computed separately with respect to each coordinate:
+This approach is especially suitable in our situation because our original objective function $E_F\left[s_A(Y,x)\right]$ is decomposable in the coodinates $x_i$, allowing the dual function to be computed separately with respect to each coordinate:
 \begin{align}
-\mathcal{L}_D(\lambda) &= \inf_{x \in \mathbb{R}_{+}^N} L\sum_{l=1}^{N} E_{F_l}[(Y_l - x_l)_{+}] + \lambda\left(\sum_{l=1}^{N} x_l - k\right) \\
-&= - \lambda k + \sum_{l=1}^{N} \inf_{x_l \in \mathbb{R}_{+}}\left\{ L E_{F_l}[(Y_l - x_l)_{+}] + \lambda x_l\right\}  \label{eqn:dual_decomp} \\
-&= - \lambda k + \sum_{l=1}^{N}  LE_{F_l}[(Y_l - x_l(\lambda))_{+}] + \lambda x_l(\lambda).
+\mathcal{L}_D(\lambda) &= \inf_{x \in \mathbb{R}_{+}^N} L\sum_{i=1}^{N} E_{F_i}[(Y_i - x_i)_{+}] + \lambda\left(\sum_{i=1}^{N} x_i - k\right) \\
+&= - \lambda k + \sum_{i=1}^{N} \inf_{x_i \in \mathbb{R}_{+}}\left\{ L E_{F_i}[(Y_i - x_i)_{+}] + \lambda x_i\right\}  \label{eqn:dual_decomp} \\
+&= - \lambda k + \sum_{i=1}^{N}  LE_{F_i}[(Y_i - x_i(\lambda))_{+}] + \lambda x_i(\lambda).
 \end{align}
-Here, $x_l(\lambda)$ is any member of the set 
+Here, $x_i(\lambda)$ is any member of the set 
 \begin{align}
-\chi_l(\lambda) = \underset{x_l \in \mathbb{R}_{+}}{\mathrm{arginf}} \left\{ L E_{F_l}[(Y_l - x_l)_{+}] + \lambda x_l\right\}
+\chi_i(\lambda) = \underset{x_i \in \mathbb{R}_{+}}{\mathrm{arginf}} \left\{ L E_{F_i}[(Y_i - x_i)_{+}] + \lambda x_i\right\}
 \end{align}
-which for $F_l$ in a standard parametric family will often have simple the form $\chi_l(\lambda) = \{F_l^{-1}(1-\lambda/L)\}$ but could an interval containing $0$ or $\infty$.
-From general theory, $\mathcal{L}_D(\lambda)$ is concave, and while it may not be differentiable when the $x_l$'s attaining the infima in \eqref{eqn:dual_decomp} are zero, it does have as a subdifferential the well-defined interval
+which for $F_i$ in a standard parametric family will often have simple the form $\chi_i(\lambda) = \{F_i^{-1}(1-\lambda/L)\}$ but could an interval containing $0$ or $\infty$.
+From general theory, $\mathcal{L}_D(\lambda)$ is concave, and while it may not be differentiable when the $x_i$'s attaining the infima in \eqref{eqn:dual_decomp} are zero, it does have as a subdifferential the well-defined interval
 \begin{align}
 \partial \mathcal{L}_D(\lambda) 
-&= \sum_{l=1}^N \chi_l(\lambda) - k.
+&= \sum_{i=1}^N \chi_i(\lambda) - k.
 \end{align}
-This of course looks very similar to the right hand side of \eqref{eqn:lambda_stationary2}, but the dual perspective has the key advantage of not requiring any guesses of the set $\{l \mid x_l^{F,k} > 0\}$ when initiating a search algorithm for $\lambda^{\star}$.
-Instead, given a candidate $\lambda_0$ for $\lambda^{\star}$, we proceed by calculating the sets $\chi_l(\lambda_0), l = 1,\ldots,N$, and using them to find the (usually zero-width) interval $\partial \mathcal{L}_D(\lambda_0)$.
+This of course looks very similar to the right hand side of \eqref{eqn:lambda_stationary2}, but the dual perspective has the key advantage of not requiring any guesses of the set $\{l \mid x_i^{F,k} > 0\}$ when initiating a search algorithm for $\lambda^{\star}$.
+Instead, given a candidate $\lambda_0$ for $\lambda^{\star}$, we proceed by calculating the sets $\chi_i(\lambda_0), l = 1,\ldots,N$, and using them to find the (usually zero-width) interval $\partial \mathcal{L}_D(\lambda_0)$.
 If $\partial \mathcal{L}_D(\lambda_0) \subset (-\infty,0)$, we can eliminate $\{\lambda \geq \lambda_0\}$ from our search and similarly if $\partial \mathcal{L}_D(\lambda_0) \subset (0, \infty)$, we can eliminate $\{\lambda \leq \lambda_0\}$.
@@
 
-If the $F_l$ increase steadily enough toward 1 after leaving a neigborhood of 0, we can expect $\partial \mathcal{L}_D(\lambda)$ to be bounded away from 0 for all $\lambda \neq \lambda^{\star}$, allowing a binary search to converge without ambiguity.
-But when some $F_l$ exhibit internal ``plateaus'' where their densities stay at or near 0, we can and often do in our applications encounter $\overline{\lambda}$ for which $\partial \mathcal{L}_D(\overline{\lambda})$---or at least our numerical approximation of it---contains 0 as an interior point.
-That is, $\overline{\lambda}$ is a price at which an uncontrained decision maker informed only by $F$ can increase from total provision $T_1<k$ to another $T_2>k$ while maintaining an optimal expected utility by making (at least approximately) one-to-one trade-offs between increased resource cost and risk reduction in the one or more locations where $F_l$ has a plateau of height near $1-\overline{\lambda}/L$.
+If the $F_i$ increase steadily enough toward 1 after leaving a neigborhood of 0, we can expect $\partial \mathcal{L}_D(\lambda)$ to be bounded away from 0 for all $\lambda \neq \lambda^{\star}$, allowing a binary search to converge without ambiguity.
+But when some $F_i$ exhibit internal ``plateaus'' where their densities stay at or near 0, we can and often do in our applications encounter $\overline{\lambda}$ for which $\partial \mathcal{L}_D(\overline{\lambda})$---or at least our numerical approximation of it---contains 0 as an interior point.
+That is, $\overline{\lambda}$ is a price at which an uncontrained decision maker informed only by $F$ can increase from total provision $T_1<k$ to another $T_2>k$ while maintaining an optimal expected utility by making (at least approximately) one-to-one trade-offs between increased resource cost and risk reduction in the one or more locations where $F_i$ has a plateau of height near $1-\overline{\lambda}/L$.
 

### 2023-08-31 alloscore_manuscript/allocation.tex

@@
 
-In the decision making setting that we consider, an action $\mathbf{x} = (x_1, \ldots, x_N)$ is a vector specifying the amount that is allocated to each of $N$ locations.
-We require that each $x_i$ is non-negative and that the total allocation across all locations equals the amount of available resources, $K$: $\sum_{i=1}^N x_i = K$.
+In the decision making setting that we consider, an action $x = (x_1, \ldots, x_N)$ is a vector specifying the amount that is allocated to each of $N$ locations.
+We require that $0 \leq x$, i.e., that each $x_i$ is non-negative, and that the total allocation across all locations equals the amount of available resources, $K$: $\sum_{i=1}^N x_i = K$.
 The set $\mathcal{X}$ consists of all possible allocations that satisfy these constraints.
-The eventually realized resource need in each location is denoted by $\mathbf{y} = (y_1, \ldots, y_N)$.
+The eventually realized resource need in each location is denoted by $y = (y_1, \ldots, y_N)$.
 These levels of need are not known at the time of decision making, so we define the random vector $Y = (Y_1, \ldots, Y_N)$ where $Y_i$ represents the as-yet-unknown level of resource need in location $i$.
@@
 \begin{equation}
-s_A(\mathbf{x}, \mathbf{y}) = \sum_{i=1}^N L (y_i - x_i)_+. \label{eqn:loss_fn}
+s_A(x, y) = \sum_{i=1}^N L (y_i - x_i)_+. \label{eqn:loss_fn}
 \end{equation}
@@
 
-\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the forecast, $\mathbf{x}^F$, is the allocation that minimizes the expected loss:
+\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the forecast, $x^F$, is the allocation that minimizes the expected loss:
 \begin{align}
-    \mathbf{x}^F &= \mathrm{arg min}_{\mathbf{x}}\,\, \mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y})] \text{ subject to }
-    0 \leq x_i \text{ for all $i$ and} \, \sum_{i=1}^N x_i = K, \label{AP}
+    x^F &= \underset{0 \leq x}{\mathrm{arg min}}\,\, \mathbb{E}_{F} [s_A(x, Y)] \text{ subject to }
+     \, \sum_{i=1}^N x_i = K, \label{AP}
 \end{align}
-where $\mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y})] = \sum_{i=1}^{N} L \mathbb{E}_{F_i}[(Y_i - x_i)_+]$ sums the expected loss due to unmet need across all locations.
+where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \mathbb{E}_{F_i}[(Y_i - x_i)_+]$ sums the expected loss due to unmet need across all locations.
 It can be shown that the Bayes act has $x^F_i = F_i^{-1}(1 - \lambda^{\star}/L)$, where $\lambda^{\star}$ depends on the problem parameters $K$ and $L$ as well as the forecast distributions and is chosen so as to satisfy the equation
@@
 According to this interpretation, we divide the available resources among the locations in such a way that the forecasted probability that need will exceed the available resources is the same in all locations.
-An alternative interpretation comes from noting that for each location $i$, $\frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(\mathbf{x}, \mathbf{Y})] \vert_{\mathbf{x}=\mathbf{x}^F} = - \lambda^{\star}$ (see the supplement for a proof).
+An alternative interpretation comes from noting that for each location $i$, $\frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^F} = - \lambda^{\star}$ (see the supplement for a proof).
 In words, at the allocation given by the Bayes act, the rate of change of the expected score as a function of the amount allocated to location $i$ is given by $-\lambda^{\star}$.
@@
 \begin{equation}
-S_A(F, y; K) = s_A(\mathbf{x}^F, y) = \sum_{i=1}^n L (y_i - F_i^{-1}(1 - \lambda^{\star}/L))_+
+S_A(F, y; K) = s_A(x^F, y) = \sum_{i=1}^n L (y_i - F_i^{-1}(1 - \lambda^{\star}/L))_+
 \end{equation}

### 2023-09-01 alloscore_manuscript/application.Rnw
changes to make project root dir the knitr location and fixing one typo that was making latex stop.

@@
 theme_set(theme_bw())
+
+## necessary to have project root as wd, to override document directory
+knitr::opts_knit$set(root.dir = '../')
 @
@@
 \begin{itemize}
-\item positive association between allocation score and MWIS ranks: \texttt{Karlen-pypm} and \texttt{USC-SI_kJalpha} models
+\item positive association between allocation score and MWIS ranks: \texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha} models
 \item consistently strong MWIS ranks, highly variable allocation score ranks, no clear association: \texttt{JHUAPL-SLPHospEns} and \texttt{CU-select}

### 2023-09-01 R/exponential-examples.R
adding files for the "made" exponential example

@@ -221,7 +221,7 @@ if (margin_type == "pdf") {
 }
 
 
-pdf('../figures/exponential_pred_expected_loss.pdf',
+pdf('figures/exponential_pred_expected_loss.pdf',
     width = 8,
     height = 6.2)
 if (margin_type == "pdf") {

### 2023-09-01 alloscore_manuscript/alloscore-application.Rnw
deleteing allocation.tex and application.Rnw and associated files and committing alloscore-application.Rnw and associated files in their place.

--- 2023-08-31 alloscore_manuscript/allocation.tex
@@
 
+\usepackage{parskip}
+
 \usepackage{soul}
@@
 \def\bwr#1{{\color{violet}\textbf{BWR:[#1]}}}
+\def\ngr#1{{\color{blue}\textbf{NGR:[#1]}}}
 
@@
 
+<<setup, tidy=FALSE, echo=FALSE, message=FALSE>>=
+knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
+@
+
+<<environment-setup>>=
+library(targets)
+library(tidyverse)
+library(ggbump)
+
+theme_set(theme_bw())
+
+## necessary to have project root as wd, to override document directory
+knitr::opts_knit$set(root.dir = '../')
+@
+
 \begin{abstract}
@@
 \section{Introduction}
-
-%High level points to cover in introduction:
-%\begin{itemize}
-%\item People are using infectious disease forecasts as an input to decision making
-%\item There are standard ways to evaluate forecasts that are responsive to decision making context, and use of those methods is relatively common in other fields like economics and meteorology
-%\item However, there's not much work in infectious disease that does this
-%\item In practice, infectious disease forecasts have typically been evaluated with "off the shelf" scoring rules such as the WIS which is an approximation to CRPS, log score, and so on.
-%\item In this work, our goal is to begin to address that gap. We focus on a resource allocation setting.
-%\item There is past work focusing on resource allocation in the operations research literature, but it doesn't take the step of getting to a measure of forecast skill.
-%\item That is, while there is plenty of work on
-%\begin{itemize}
-%\item DM under risk, where probabilities are taken as known and expected utility is maximized
-%\item DM under uncertainty, where probabilities are taken as unknown and robust (i.e. maximin) decisions are sought
-%\end{itemize}
-%little seems to have been said on how forecasts can be evaluated by how well they convert a problem of DM under uncertainty to a problem of DM under risk.
-%\end{itemize}
-
-% People using infectious disease forecasting as an input to decision making:
-
-%Center for Forecasting and Outbreak Analytics (CDC) https://www.cdc.gov/forecast-outbreak-analytics/annual-report/2023/key-accomplishments.html#omicron-response anual report
-%  Bertsimas gives data-driven approach to provide clincal insights and support ventilator allocation for policy makers,
-
-% "Our results have been used at the clinical level by several hospitals to triage patients, guide care management, plan ICU capacity, and re-distribute ventilators.
-At the policy level, they are currently supporting safe back-to-work policies at a major institution and vaccine trial location planning at Janssen Pharmaceuticals, and have been integrated into the US Center for Disease Control’s pandemic forecast."
-
-%https://www.cdc.gov/mmwr/preview/mmwrhtml/su6303a1.htm
 
@@
 In all of these fields, analyses have consistently found that common measures of statistical forecast accuracy do not necessarily correspond directly to measures of the value of forecasts as an input to decision making \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
-% Consistently, but not uniformly
-
-% Other important early contributions include Nelson (1972) and White (1966), who are amongst the first to consider the problem of forecast evaluation from the perspective of decision theory and dynamic programming.
-Nelson, in particular, articulates the concept of the “value of a (probability) fore- cast” to an individual decision maker as the expected value of his/her payoff using the probability forecasts relative to the payoff expected if unconditional probability estimates are used.
-
-% This has been somewhat rectified during the last decade as the observation that financial series are predictable has led to an attempt to understand the economic significance of this finding.
-See, for example, Breen, Glosten, and Jagannathan (1989), Leitch and Tanner (1991), McCulloch and Rossi (1990), West, Edison, and Cho (1993), Pesaran and Timmermann (1994, 1995), Satchell and Timmermann (1995), and Skouras (1998).
-
-% Theil (1960), White (1966)
 
@@
 
-% As described above, there exists both a rich literature on outbreak forecast evaluation and an emerging body of work on turning forecasts into resource allocation decisions. 
-% However, we have not found literature exploring whether standard forecast accuracy metrics are sufficient for evaluating how well forecasts support allocation decisions.
-
-% \subsection{Organization}
-
 The remainder of this article is organized as follows.
@@
 
-\section{Methods}
+\section{The Allocation Score}
 \label{sec:methods}
@@
 
-\section{Application}
+\section{Evaluating forecasts of COVID hospitalizations using the allocation score}
 \label{sec:application}
@@
 
+\ngr{TODO: add a figure showing some forecasts and ground truth data.}
+
+\subsection{Data}
+\subsubsection{Hospitalization data}
+Starting in the summer of 2020, the US Health and Human Services began reporting counts of daily new admissions to hospitals for individuals with COVID-19.(cite HHS Protect)
+These data served as the source of ``ground truth'' data for the US COVID-19 Forecast Hub which, starting in December 2020, collected short-term forecasts of new hospital admissions at the daily scale.
+These daily counts were available for the US as a whole, and all states and several additional jurisdictions such as Puerto Rico and Washington DC.
+The data were updated daily and were available for download by the public through the HHS HealthData.gov website.
+For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data.(cite package) \ngr{can we use tar\_manifest() to reproducibly display the date the data were stored?}
+
+\subsubsection{Forecast data}
+The US COVID-19 Forecast Hub, a consortium funded by the US CDC and led by a research group at UMass-Amherst, collected short-term forecasts of hospitalizations starting in December 2020.(cite Data descriptor paper)
+Any team that with appropriately formatted forecasts could submit them to the Forecast Hub data repository on GitHub.(cite repo site)
+Forecasts were time-stamped by GitHub upon submission and passed validation checks that ensured correct formatting and that the forecasts were being submitted only for dates in the future, not for data that had already been observed.
+
+Forecast submission followed a weekly cycle and culminated in the creation of an ensemble forecast.
+Forecasts could be submitted on any day during the week.
+However once a week on Mondays, the Forecast Hub would collect the most recent forecasts submitted by all teams that met certain inclusion criteria and create an ensemble forecast using quantile averaging.(cite Evan's paper)
+An ensemble that treated all models equally was created (COVIDhub-ensemble) as was a model that created weights of submitted models based on performance in the past 12 weeks (COVIDhub-trained\_ensemble).
+One other model that combined multiple forecasts from different teams but used a different ensembling algorithm, a linear pooling method with tail extrapolation, was also included in our analyses (JHUAPL-SLPHospEns).
+Several other models have ``ensemble'' in their name, but this refers to combinations of different variations of models that the specific team created, not to a multi-model ensemble combining different submitted forecasts to the Forecast Hub.
+
+All forecasts, including the ensemble, were submitted as probabilistic predictions about the number of new hospital admissions on a particular day in the future, in a specific jurisdiction of the US (national level, state, or territory).
+Probability distributions were specified, per the requirements established by the Forecast Hub, as a set of 23 quantiles for each individual prediction.
+The submitted quantiles included a median (treated as a ``point'' prediction) and defined 11 central prediction intervals, from a 99\% to a 10\% prediction interval.
+
+The analysis in this work focuses on forecasts made before and during the first wave of the Omicron SARS-CoV-2 variant in the US.
+As such, we downloaded forecasts for the 15 weeks starting with Monday 2021-11-22 through Monday 2022-02-28.
+
+We established a set of inclusion criteria to determine which forecasts and models to include in our analysis.
+Models were eligible to be included in the analysis if they were considered a ``primary'' model from a team.
+(If a team submitted multiple versions of similar models, they were required to designate one as ``primary''.)
+For a model to have a complete, eligible submission in a given week, it had to have a 14 day-ahead forecast for all 50 states plus Washington DC.
+Models had to have a complete forecast for at least 4 of the 15 weeks in the analysis to be considered eligible for inclusion.
+
+\subsection{Evaluation metrics}
+
+This manuscript focuses on two proper forecast scores, the allocation score and the weighted interval score (WIS).(cite Gerding and Bracher)
+Propriety of forecast scoring rules is desirable as it ensures that forecasters are not incentivized to modify their forecast distribution to achieve a better score.(cite Gneiting and Raftery or Raftery and Gneiting)
+
+\subsubsection{Allocation Score}
+
+\ngr{TODO: general introduction to the alloscore}
+
+For the analysis, we fixed a resource constraint $K$ to be 15,000, based roughly on a reported number of ventilators available for reallocation in the US.(cite paper)
+For each week, we computed a the allocation score for each 14 day-ahead forecast based on $K=15,000$.
+\ngr{it appears that for the last two weeks selected (forecast dates: 2022-02-14 and 2022-02-28) the upper limit of K was 11400 and 9200 respectively, so we do not have alloscores computed for K=15000.
+Maybe they would have all been zero, since the limits were likely set based on the number of cases being low.
+What is the rule used to determine the max K computed?
+Don't think we need to change it, but would be nice to state in the manuscript what it is.}
+
+We also computed a standardized rank for the allocation score for each model $m$ and week $w$.
+First, we computed the number of models that forecasted that week ($n_w$) and the rank of model $m$ among the $n_w$ models ($r_{m,w}^{AS}$).
+The model with the best allocation score received a rank of 1 and the worst received a rank of $n_w$.
+In the case of a tie between one or more models, all models received the better rank.
+We then rescaled these rankings to compute the allocation score standardized rank ($sr_{m,w}^AS$) between 0 and 1, where 0 corresponds to the worst rank and 1 to the best.
+
+\begin{equation}
+sr_{m,w}^{AS} = 1 - \frac{r^{AS}_{m,w}-1}{n_w-1}
+\end{equation}
+
+\subsubsection{Weighted Interval Score (WIS)}
+
+\ngr{TODO: general introduction to WIS and define MWIS as mean WIS.}
+
+We computed standardized ranks for MWIS ($sr_{m,w}^{MWIS}$) using the same procedure as for allocation scores described above.
+
+\subsection{Data and code availability}
+
+\subsection{Application results}
+
+<<score-computation>>=
+tar_load("all_alloscore_data")
+
+alloscore_data_summaries <- all_alloscore_data |>
+  filter(K == 15000)|>
+  group_by(model, reference_date) |>
+  summarize(alloscore = sum(components), yobs = sum(y)) |>
+  group_by(reference_date) |>
+  mutate(nmodels = n(),
+         allo_rank = rank(alloscore, ties.method = "min"),
+         allo_rank_std = 1- (allo_rank-1)/(nmodels-1)) |>
+  ungroup() |>
+  complete(model, reference_date) |>
+  mutate(target_end_date = as.Date(reference_date) + 14)
+
+tar_load("score_data")
+
+alloscore_models <- unique(alloscore_data_summaries$model)
+score_data_summaries <- score_data |>
+  mutate(reference_date = target_end_date - 14) |>
+  filter(model %in% alloscore_models) |>
+  group_by(model, reference_date) |>
+  summarize(mwis = mean(wis)) |>
+  ungroup() |>
+  group_by(reference_date) |>
+  mutate(nmodels = n(),
+         mwis_rank = rank(mwis, ties.method = "min"),
+         mwis_rank_std = 1- (mwis_rank-1)/(nmodels-1)) |>
+  ungroup() |>
+  complete(model, reference_date) |>
+  mutate(target_end_date = as.Date(reference_date) + 14)
+@
+
+\subsubsection{Metrics as a function of time}
+<<grob-hosps-over-time>>=
+tar_load("truth_data")
+dat_to_plot <- truth_data |>
+  filter(location == "US",
+         target_end_date %in% score_data_summaries$target_end_date)
+p1 <- ggplot(dat_to_plot,
+             aes(x=target_end_date, y=value)) +
+  geom_line() +
+  geom_point() +
+  geom_hline(yintercept=15000, linetype=2) +
+  ylab("hospital admissions") +
+  xlab(NULL) +
+  ylim(c(0, NA))
+@
+
+<<metrics-over-time, fig.height=9, fig.width=7, fig.cap="Hospital admissions and evaluation metrics over time.
+Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022.
+These are the days for which forecasts were made and evaluated.
+A horizontal dashed line at 15,000 shows the assumed resource constraint $K$.
+Panel B shows allocation scores for each model's 14-day ahead forecast, across all US states.
+The x-axis corresponds to the date for which the prediction was made.
+Allocation scores typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down).
+Panel C shows the MWIS metric across weeks, averaged across all states.
+Similarly to panel B, the x-axis corresponds to the date for which the prediction was made.
+MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the pean of the Omicron wave.">>=
+p2 <- ggplot(alloscore_data_summaries,
+             aes(x=target_end_date, y=alloscore, color=model, shape=model)) +
+  scale_x_date(limits=range(score_data_summaries$target_end_date)) +
+  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
+  ylab("Allocation score") +
+  geom_point() + geom_line()+
+  theme(legend.pos="none") +
+  xlab(NULL)
+
+p4 <- ggplot(score_data_summaries,
+             aes(x=target_end_date, y=mwis, color=model,shape=model)) +
+  geom_point() + geom_line() +
+  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
+  ylab("MWIS")+
+  theme(legend.pos="bottom")+
+  xlab(NULL)
+
+cowplot::plot_grid(p1, p2, p4,
+                   nrow=3,
+                   align="v",
+                   rel_heights = c(1, 1, 1.5),
+                   labels = "AUTO")
+@
+
+Allocation scores varied substantially by date and by model (Figure XX).
+For predictions made for the first three Mondays in December 2021 and the last three Mondays in February 2022 all models had allocation scores under 500 (and the mean across all models was less than 100), indicating that the unnecessary unmet need was fairly low relative to the total number of hospital admissions on those days.
+The allocation scores are on the whole highest when the observed number of new hospital admissions is closest to the resource threshold of 15,000, as this is the time when any mistakes in allocation are costly in terms of wasting resources in one location that could have been used in another.
+Predictions made during the peak week and just after showed the highest variation in allocation scores, with some models having values under 1000 and others having values over 3500.
+<<eval=FALSE>>=
+## some analysis to support above statements
+alloscore_data_summaries |>
+  group_by(target_end_date) |>
+  summarize(meanAS = mean(alloscore, na.rm=TRUE),
+            medianAS = median(alloscore, na.rm=TRUE),
+            minAS = min(alloscore, na.rm=TRUE),
+            maxAS = max(alloscore, na.rm=TRUE),
+            range = maxAS-minAS)
+@
+
+Mean weighted interval scores (MWIS) also varied by date and model, and more clearly were dependent on the scale of the observed data.
+MWIS values were low (all models under 100) for all Mondays in December 2021 and the final four Mondays evaluated.
+Across all models both the average and median MWIS value for every Monday in January was above 100, with the largest errors occurring one and two weeks after the peak was observed.
+
+<<eval=FALSE>>=
+## some analysis to support above statements
+score_data_summaries |>
+  group_by(target_end_date) |>
+  summarize(meanWIS = mean(mwis, na.rm=TRUE),
+            medianWIS = median(mwis, na.rm=TRUE),
+            minWIS = min(mwis, na.rm=TRUE),
+            maxWIS = max(mwis, na.rm=TRUE),
+            range = maxWIS-minWIS)
+@
+
+\subsubsection{correlation between WIS and allocation score}
+
+<<metrics-correlation, fig.width=7, fig.height=6, fig.cap="Association of standardized ranks for MWIS and allocation score by model and week.
+Each facet of the plot corresponds to one model.
+Within each facet, each point corresponds to a week.
+The x- and y-values correspond to the MWIS standardized rank and the allocation score standardized rank for that week.
+Points corresponding to earlier dates have darker shading.
+The size of the point corresponds to the observed value on the date for which the prediction was made.
+Models show different degrees of association between the two metrics.">>=
+all_data <- left_join(alloscore_data_summaries, score_data_summaries,
+                      by=c("model", "target_end_date"))
+
+ggplot(all_data,
+       aes(x=mwis_rank_std, y=allo_rank_std, color=target_end_date, size=yobs)) +
+  geom_point(alpha=.7) +
+  scale_x_continuous("MWIS standardized rank", limits=c(0,1)) +
+  scale_y_continuous("Allocation score standardized rank", limits=c(0,1)) +
+  facet_wrap(.~model) +
+  geom_abline(slope=1, intercept=0, linetype=2, color="gray")
+
+@
+
+Models show differing levels of correlation between their allocation scores and MWIS values (Figure XX).
+Here are some examples of the different patterns observed:
+\begin{itemize}
+\item positive association between allocation score and MWIS ranks: \texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha} models
+\item consistently strong MWIS ranks, highly variable allocation score ranks, no clear association: \texttt{JHUAPL-SLPHospEns} and \texttt{CU-select}
+\item consistently strong ranks for both metrics, no clear association: \texttt{COVIDhub-ensemble}
+\end{itemize}
+Additionally, the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, had the highest rank for allocation score in six weeks, more than any other model except the \texttt{COVIDhub-ensemble} which also had six.
+
+<<eval=FALSE>>=
+alloscore_data_summaries |>
+  group_by(model) |>
+  summarize(top_rank = sum(allo_rank_std==1))
+@
+
+ - bubble figure with ranks sized by total observations
+
+\subsubsection{Integrated allocation score across values of K}
+
+ - something here from Ben's analysis?
+
 \section{Discussion}
 \label{sec:discussion}
+
+Notes for the application
+
+ - It is clear that the metrics are capturing different aspects of performance
+
+ - WIS is scale dependent, alloscore not as much
+
+ - the fact that it is hard to beat the baseline for allocation suggests that we are not consistently adding value over just looking at the current levels.
+\ngr{although is there something more to it about how the baseline accounts for uncertainty?}
 

### 2023-09-01 alloscore_manuscript/alloscore-application.Rnw
some minor changes:  - adding true expectation sign  - defining L  - changing "positive part" function to max(0,...)

@@
 \begin{align}
-\sum_{i=1}^N E_{F_i}[\text{unmet need in location $i$}], \label{eqn:informal_objective}
+\sum_{i=1}^N \mathbb{E}_{F_i}[\text{unmet need in location $i$}], \label{eqn:informal_objective}
 \end{align}
@@
 \begin{equation}
-s_A(x, y) = \sum_{i=1}^N L (y_i - x_i)_+. \label{eqn:loss_fn}
+s_A(x, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i). \label{eqn:loss_fn}
 \end{equation}
-Here, $(y_i - x_i)_+$ is the unmet need in location $i$, which is given by $y_i - x_i$ if the realized need $y_i$ in location $i$ is greater than the amount $x_i$ allocated to that location, or $0$ if the amount $x_i$ allocated to unit $i$ is greater than or equal to the realized need.
+Here, $\max(0, y_i - x_i)$ is the unmet need in location $i$, which is given by $y_i - x_i$ if the realized need $y_i$ in location $i$ is greater than the amount $x_i$ allocated to that location, or $0$ if the amount $x_i$ allocated to unit $i$ is greater than or equal to the realized need.
+Also, $L$ is a constant scalar value, the same across all locations, specifying the ``cost'' of one unit of unmet need.
 
@@
 \end{align}
-where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \mathbb{E}_{F_i}[(Y_i - x_i)_+]$ sums the expected loss due to unmet need across all locations.
+where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \mathbb{E}_{F_i}[max(0, Y_i - x_i)]$ sums the expected loss due to unmet need across all locations.
 It can be shown that the Bayes act has $x^F_i = F_i^{-1}(1 - \lambda^{\star}/L)$, where $\lambda^{\star}$ depends on the problem parameters $K$ and $L$ as well as the forecast distributions and is chosen so as to satisfy the equation
@@
 \begin{equation}
-S_A(F, y; K) = s_A(x^F, y) = \sum_{i=1}^n L (y_i - F_i^{-1}(1 - \lambda^{\star}/L))_+
+S_A(F, y; K) = s_A(x^F, y) = \sum_{i=1}^n L \max(0, y_i - F_i^{-1}(1 - \lambda^{\star}/L))
 \end{equation}

### 2023-09-02 alloscore_manuscript/alloscore-application.Rnw
rewording of step 3 and removal of lambda^star

@@
 
-\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the forecast, $x^F$, is the allocation that minimizes the expected loss:
+\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the forecast, $x^{F,K}$, is the allocation that minimizes the expected loss:
 \begin{align}
-    x^F &= \underset{0 \leq x}{\mathrm{arg min}}\,\, \mathbb{E}_{F} [s_A(x, Y)] \text{ subject to }
+    x^{F,K} &= \underset{0 \leq x}{\mathrm{arg min}}\,\, \mathbb{E}_{F} [s_A(x, Y)] \text{ subject to }
      \, \sum_{i=1}^N x_i = K, \label{AP}
 \end{align}
-where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \mathbb{E}_{F_i}[max(0, Y_i - x_i)]$ sums the expected loss due to unmet need across all locations.
-It can be shown that the Bayes act has $x^F_i = F_i^{-1}(1 - \lambda^{\star}/L)$, where $\lambda^{\star}$ depends on the problem parameters $K$ and $L$ as well as the forecast distributions and is chosen so as to satisfy the equation
+where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]$ sums the expected loss due to unmet need across all locations.
+
+In the supplement we show that for a typical forecast $F$ and constraint level $K$, the components $x_i^{F,K}$ of the Bayes act are quantiles 
+$q_{i,\tau} = F_i^{-1}(\tau)$ of the marginal forecasts $F_i$ for a probability level $\tau = \tau(K,F)$ satisfying the equation
 \begin{align}
-\sum_{i=1}^{N}F_i^{-1}(1 - \lambda^{\star}/L) = K.
+\sum_{i=1}^{N}q_{i,\tau} = K.
 \end{align}
-This partial solution to the allocation problem seems to have first appeared in \cite{hadleywhitin1963}; see the supplemental materials for a derivation.
+We note that $\tau(K,F)$ depends not only on the constraint level $K$ defining the decision problem but also on the forecast $F$. This is quite
+different from the situation for quantile scores and the WIS, where the probability levels used to calculate scores are fixed.
 
-One interpretation of this result is that the Bayes act sets the allocation in each location $i$ to a quantile of the forecast distribution $F_i$ for that location.
-The quantile is at a probability level $\alpha = 1 - \lambda^{\star}/L$ that is the same for all locations, and is chosen such that the constraint is satisfied.
-According to this interpretation, we divide the available resources among the locations in such a way that the forecasted probability that need will exceed the available resources is the same in all locations.
-An alternative interpretation comes from noting that for each location $i$, $\frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^F} = - \lambda^{\star}$ (see the supplement for a proof).
-In words, at the allocation given by the Bayes act, the rate of change of the expected score as a function of the amount allocated to location $i$ is given by $-\lambda^{\star}$.
-This derivative is the same for all locations, so the optimal allocation divides the available resources across all locations in such a way that according to $F$, the expected benefit of 1 additional unit of resources is the same in all locations.
+The fact that the vector $(q_{1,\tau},\ldots,q_{N,\tau})$ of quantiles solves the allocation problem \eqref{AP} has the 
+interpretation that 
+in order to allocate optimally (according to $F$) we must divide resoruces among the locations so that the forecasted probability of need exceeding local allocation is the same everywhere.
+A related interpretation is offered by our model's expression as a partial derivative (derived in the supplement) for the reduction in expected score (i.e., increase in expected utility) in location $i$ that results from a unit increase in the resource:
+\begin{align}
+\frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^{F,K}} = L \cdot (1-\tau).
+\end{align}
+This equation tells us that when our allocations are set to the Bayes act $x^{F,K}$, shifting a small amount $\Delta x$ of the resource from one location to another would yield no change in expected utility since the gain $\Delta x \cdot L \cdot (1-\tau)$ in the receiving location would be canceled out by the loss $\Delta x \cdot L \cdot (\tau-1)$ in the location giving up the resource.
+This partial solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
+There, products and order quantities play the roles we have given to locations and resource allocation levels.
 
@@
 \begin{equation}
-S_A(F, y; K) = s_A(x^F, y) = \sum_{i=1}^n L \max(0, y_i - F_i^{-1}(1 - \lambda^{\star}/L))
+S_A(F, y; K) = s_A(x^{F,K}, y) = \sum_{i=1}^n L \max(0, y_i - F_i^{-1}(\tau(K,L,F)))
 \end{equation}
@@
 
-The allocation score $S_A$ that we developed in the previous section measures the skill of the forecast distributions $F$ based on a single probability level $\alpha = 1 - \lambda^{\star}/L$.
+The allocation score $S_A$ that we developed in the previous section measures the skill of the forecast distributions $F$ based on a single probability level $\tau$.
 This is appropriate if the resource constraint $K$ is a known constant.

### 2023-09-03 alloscore_manuscript/alloscore-application.Rnw
change allocation problem def in step 3

@@
 
-\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the forecast, $x^{F,K}$, is the allocation that minimizes the expected loss:
+\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the forecast, $x^{F,K}$, is the allocation that minimizes the expected loss, that is, the solution of the \emph{allocation problem} associated with $K$
 \begin{align}
-    x^{F,K} &= \underset{0 \leq x}{\mathrm{arg min}}\,\, \mathbb{E}_{F} [s_A(x, Y)] \text{ subject to }
+    \underset{0 \leq x}{\mathrm{minimize}}\,\, \mathbb{E}_{F} [s_A(x, Y)] \text{ subject to }
      \, \sum_{i=1}^N x_i = K, \label{AP}
@@
 In the supplement we show that for a typical forecast $F$ and constraint level $K$, the components $x_i^{F,K}$ of the Bayes act are quantiles 
-$q_{i,\tau} = F_i^{-1}(\tau)$ of the marginal forecasts $F_i$ for a probability level $\tau = \tau(K,F)$ satisfying the equation
+$q_{i,\tau} = F_i^{-1}(\tau)$ of the marginal forecasts $F_i$ for a probability level $\tau = \tau(K,F)$ that is defined by the constraint equation of the
+allocation problem
 \begin{align}

### 2023-09-07 alloscore_manuscript/methods_apg.tex
sorting out total bayes acts

@@
 \begin{align}
-s_{Q,f_C}(x,y) = \frac{1}{L}\int_{0}^{L} (cx(c) + L(x(c) - y)_+) f_C(c)dc{}.
+s_{Q,f_C}(x,y) = \frac{1}{L}\int_{0}^{L} (cx(c) + L(x(c) - y)_+) f_C(c)dc{}. \label{eqn:CRPS_loss}
 \end{align}
@@
 decision variable $x$ but to the vanishing of the partial derivatives with respect to all $m$ decision variables $\{x_i\}$ in the first case or the infinitely many variables $\{x(c)\}$ in the second case.
-But since the losses are sums of single varianble losses, passing to solutions of these higher dominsional equations is a mere formality of the viewing the set of Bayes acts, i.e.
+But since the losses are sums of single variable losses, passing to solutions of these higher diminsional equations is a mere formality of the viewing the set of Bayes acts, i.e.
 quantiles, for costs $c \in \mathcal{C}$ or $(0,L)$ as a vector
@@
 
-Similarly, given an outcome $y$, evaluation of the scoring rules $s_{Q,\mathcal{C}, p_C}$ and $s_{Q,f_C}$ on the forecast $F$ amounts to vectorizing the
-formula for $S_Q(F,y)$. That is,
+Accordingly, evaluation of the scoring rules $S_{Q,\mathcal{C}, p_C}$ and $S_{Q,f_C}$ on the forecast $F$ given an outcome $y$ amounts to summing the formula for $S_Q(F,y)$ across such a vector or function.
+That is,
 \begin{align}
@@
 
-We next expand our decision problem to include multiple actions that will all, in fact (not counterfactually), have to be taken.
-This includes the motivating scenario for our work, the simultaneous procurement of resources across several locations (U.S.
+But before introducing an allocation scoring rule, we need to expand our decision problem along one more dimension to include composite actions that address multivariate outcomes $y = (y_1,\ldots,y_N)$ being forecasted by a multivariate distribution $F$
+with marginal distributions $F_i, i = 1,\ldots,N$. 
+This includes our motivating scenario of simultaneous procurement of medical countermeasures across several locations (U.S.
 states and territories) during a pandemic.
-And again, summing the losses  \eqref{eqn:quantile_loss} over the locations (for a particular $C$ and $L$) creates a loss function that is reactive to the overall procurement decision.
-Let the locations be indexed by $i \in 1,\ldots,N$.
+And again, we simply aggregate the lower dimensional losses \eqref{eqn:quantile_loss}, \eqref{eqn:sum_CL_loss}, or \eqref{eqn:CRPS_loss} over the locations (or multiple forecast targets more generally) in order to formally obtain a single loss function that is reactive to all components of the overall procurement decision.
+
 The action sets for the decision problem corresponding to a single certain cost $C$, a discrete cost distribution over $\{C_j\}, j = 1,\ldots,m$, and a continuous cost distribution $f_C$ with support on $[0,L]$ are now
@@
 \end{align}
-And for these ``vector-action'' problems we can define \emph{total} versions of our loss functions
+For these ``vector-action'' problems we define \emph{total} versions of our loss functions
 \begin{align}
-s_{TQ}(x,y) &= \sum_{i=1}^N Cx_i + L(y_i-x_i)_+ \label{eqn:TQ_loss} \\
+s_{TQ, C}(x,y) &= \sum_{i=1}^N Cx_i + L(y_i-x_i)_+ \label{eqn:TQ_loss} \\
 s_{TQ, \mathcal{C}, p_C}(x,y) &= \frac{1}{m}\sum_{i=1}^N \sum_{j=1}^{m}p_j(C_i x_{i,j} + L(x_{i,j} - y_i)_+)\\
@@
 \end{align}
+Similarly as for the expansion to multiple cost contingencies---though now with respect to forecasts $F_i$ depending on the 
+outcome component $y_i$---first order conditions for optimality decouple under the additive separability of the loss function so that the Bayes act $x^F$ for $F$ is just a vector of quantiles $q_{F,\tau} := (q_{F_1,\tau},\ldots,q_{F_N,\tau})$ or quantile functions when $\tau = 1 - c/L$, $c \in \mathcal{C}$ or $[0,L]$.
+Note that while $x^F=q_{F,\tau}$ only depends on the marginals
+$F_i$, it is indeed the Bayes act for the joint $F$ since $E_F[s_{Q,c}(x_i, y_i)] = E_{F_i}[s_{Q,c}(x_i, y_i)]$.
+
+Combining the total loss functions with these composite Bayes acts gives us total versions of scoring rules for the composite decisiion problem
+\begin{align}
+S_{TQ,C}(F,y) &= \sum_{i=1}^{N}Cq_{F_i,\tau} + L(y_i-x_i)_+ \\
+S_{TQ,\mathcal{C}, p_C}(F,y) &= \frac{1}{m}\sum_{i=1}^{N}\sum_{j=1}^{m}p_j(C_j q_{F_i,\tau_j} + L(q_{F_i,\tau_j} - y_i)_+) \\
+S_{TQ, f_C}(F,y) &= \frac{1}{L}\sum_{i=1}^{N}\int_{0}^{L}(cq_{F_i,\tau(c)} + L(q_{F_i,\tau(c)} - y_i)_+) f_C(c)dc.
+\end{align}
+
 A key point for our presentation will be that the same $C$, set $\{C_j\}$, probabilities $\{p_j\}$, and density $f_C$ are used for all locations simultaneously.
@@
 \end{align}
-We see then that the cost $C$ will not effect the Bayes act for this component of the decision problem motivating the definition of a new single loss function (component), the \emph{allocation loss function}
+The component of the Bayes act for the contingency that $K=k$ is thus independent of the realized cost $C$, motivating the definition of a new single loss function (component), the \emph{allocation loss function}
 \begin{align}
@@
 \end{align}
-we use in deriving Bayes acts for all $\mathcal{X}_k$.
+we use in deriving these Bayes acts components for all $\mathcal{X}_k$.
 
@@
 \begin{align}
-0 = \sum_{\{l | x_i^{F,k} > 0\}} F_{l}^{-1}(1-\lambda^{\star}/L) - k. \label{eqn:lambda_stationary2}
+0 = \sum_{\{i | x_i^{F,k} > 0\}} F_{i}^{-1}(1-\lambda^{\star}/L) - k. \label{eqn:lambda_stationary2}
 \end{align}
@@
 Since convexity and compactness guarantee existence of a solution, the problem is now reduced to solving \eqref{eqn:lambda_stationary} for the set 
-$\{l \mid x_i^{F,k} > 0\}$ and value $\lambda^{\star}$. Note, however, that this is a more complicated task than solving just 
+$\{i \mid x_i^{F,k} > 0\}$ and value $\lambda^{\star}$. Note, however, that this is a more complicated task than solving just 
 for $\lambda^{\star}$ under the assumption that \eqref{eqn:xfk_formula} holds for all coordinates of $x^{F,k}$.\footnote{Such an assumption would exclude fairly simple examples in which $F$ has support bounded away from $0$ in one coordinate but not in another.} The task is simplified though by bringing in the full saddle point condition for $\lambda^{\star}$ which implies not only \eqref{eqn:lambda_stationary} but also that
@@
 \end{align}
-This of course looks very similar to the right hand side of \eqref{eqn:lambda_stationary2}, but the dual perspective has the key advantage of not requiring any guesses of the set $\{l \mid x_i^{F,k} > 0\}$ when initiating a search algorithm for $\lambda^{\star}$.
-Instead, given a candidate $\lambda_0$ for $\lambda^{\star}$, we proceed by calculating the sets $\chi_i(\lambda_0), l = 1,\ldots,N$, and using them to find the (usually zero-width) interval $\partial \mathcal{L}_D(\lambda_0)$.
+This of course looks very similar to the right hand side of \eqref{eqn:lambda_stationary2}, but the dual perspective has the key advantage of not requiring any guesses of the set $\{i \mid x_i^{F,k} > 0\}$ when initiating a search algorithm for $\lambda^{\star}$.
+Instead, given a candidate $\lambda_0$ for $\lambda^{\star}$, we proceed by calculating the sets $\chi_i(\lambda_0), i = 1,\ldots,N$, and using them to find the (usually zero-width) interval $\partial \mathcal{L}_D(\lambda_0)$.
 If $\partial \mathcal{L}_D(\lambda_0) \subset (-\infty,0)$, we can eliminate $\{\lambda \geq \lambda_0\}$ from our search and similarly if $\partial \mathcal{L}_D(\lambda_0) \subset (0, \infty)$, we can eliminate $\{\lambda \leq \lambda_0\}$.

### 2023-09-08 alloscore_manuscript/alloscore-application.Rnw
Update alloscore_manuscript/alloscore-application.Rnw
Co-authored-by: Evan Ray <elray1@users.noreply.github.com>

@@
 
-In the supplement we show that for a typical forecast $F$ and constraint level $K$, the components $x_i^{F,K}$ of the Bayes act are quantiles 
+In the supplement we show that in general, the components $x_i^{F,K}$ of the Bayes act are quantiles 
 $q_{i,\tau} = F_i^{-1}(\tau)$ of the marginal forecasts $F_i$ for a probability level $\tau = \tau(K,F)$ that is defined by the constraint equation of the

### 2023-09-08 alloscore_manuscript/alloscore-application.Rnw
Update alloscore_manuscript/alloscore-application.Rnw
Co-authored-by: Evan Ray <elray1@users.noreply.github.com>
Update alloscore_manuscript/alloscore-application.Rnw

@@
 In the supplement we show that in general, the components $x_i^{F,K}$ of the Bayes act are quantiles 
-$q_{i,\tau} = F_i^{-1}(\tau)$ of the marginal forecasts $F_i$ for a probability level $\tau = \tau(K,F)$ that is defined by the constraint equation of the
+$F_i^{-1}(\tau)$ of the marginal forecasts $F_i$ for a probability level $\tau = \tau(K,F)$ that is defined by the constraint equation of the
 allocation problem
 \begin{align}
-\sum_{i=1}^{N}q_{i,\tau} = K.
+\sum_{i=1}^N F_i^{-1}(\tau) = K.
 \end{align}

### 2023-09-08 alloscore_manuscript/alloscore-application.Rnw
Update alloscore_manuscript/alloscore-application.Rnw

@@
 In the supplement we show that in general, the components $x_i^{F,K}$ of the Bayes act are quantiles 
-$F_i^{-1}(\tau)$ of the marginal forecasts $F_i$ for a probability level $\tau = \tau(K,F)$ that is defined by the constraint equation of the
+$F_i^{-1}(\tau)$ of the marginal forecasts $F_i$ for a probability level $\tau = \tau(K,F)$ that solves the constraint equation of the
 allocation problem

### 2023-09-09 alloscore_manuscript/alloscore-application.Rnw

@@
 \end{align}
-We note that $\tau(K,F)$ depends not only on the constraint level $K$ defining the decision problem but also on the forecast $F$. This is quite
-different from the situation for quantile scores and the WIS, where the probability levels used to calculate scores are fixed.
+This partial solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
+In our context, this result has the interpretation that in order to allocate optimally (according to $F$), we must divide resources among the locations so that there is an equal forecasted probability in every location that the allocation is sufficient to meet resource need.
+We note that this probability, $\tau(K,F)$, depends not only on the constraint level $K$ defining the decision problem but also on the forecast $F$.
+This is quite different from the situation for quantile scores and the WIS, where the probability levels used to calculate scores are fixed.
 
-The fact that the vector $(q_{1,\tau},\ldots,q_{N,\tau})$ of quantiles solves the allocation problem \eqref{AP} has the 
-interpretation that 
-in order to allocate optimally (according to $F$) we must divide resoruces among the locations so that the forecasted probability of need exceeding local allocation is the same everywhere.
-A related interpretation is offered by our model's expression as a partial derivative (derived in the supplement) for the reduction in expected score (i.e., increase in expected utility) in location $i$ that results from a unit increase in the resource:
+A related interpretation comes from the observation that at the allocation given by the Bayes act, the rate of change of the expected loss with respect to the amount of resources allocated to a particular location $i$ is the same for all locations, and is given by
 \begin{align}
-\frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^{F,K}} = L \cdot (1-\tau).
+    \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^{F,K}} = L \cdot (\tau - 1).
 \end{align}
-This equation tells us that when our allocations are set to the Bayes act $x^{F,K}$, shifting a small amount $\Delta x$ of the resource from one location to another would yield no change in expected utility since the gain $\Delta x \cdot L \cdot (1-\tau)$ in the receiving location would be canceled out by the loss $\Delta x \cdot L \cdot (\tau-1)$ in the location giving up the resource.
-This partial solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
-There, products and order quantities play the roles we have given to locations and resource allocation levels.
+Therefore, when our allocations are set to the Bayes act $x^{F,K}$, shifting a small amount of the resource from one location to another would yield no change in total expected unmet need since the change in expected unmet need in the receiving location would be canceled out by the change in the location giving up the resource.
 

### 2023-09-10 alloscore_manuscript/alloscore-application.Rnw
tentative rererererewrite

--- 2023-09-08 alloscore_manuscript/alloscore-application.Rnw
@@
 
-In the supplement we show that in general, the components $x_i^{F,K}$ of the Bayes act are quantiles 
-$F_i^{-1}(\tau)$ of the marginal forecasts $F_i$ for a probability level $\tau = \tau(K,F)$ that solves the constraint equation of the
-allocation problem
+In the supplement we derive a general form of this solution by beginning with the fact that a condition for minimizing
+expected loss under the constraint is that there is no way to decrease expected loss further by shifting a small amount of the resource from one location to another.
+As a consequence, at $x^{F,K}$, all the rates of change of expected loss with respect to the amount of resources allocated to a particular location $i$ must be equal:
 \begin{align}
-\sum_{i=1}^N F_i^{-1}(\tau) = K.
+ \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^{F,K}} = \lambda
 \end{align}
-We note that $\tau(K,F)$ depends not only on the constraint level $K$ defining the decision problem but also on the forecast $F$. This is quite
-different from the situation for quantile scores and the WIS, where the probability levels used to calculate scores are fixed.
-
-The fact that the vector $(q_{1,\tau},\ldots,q_{N,\tau})$ of quantiles solves the allocation problem \eqref{AP} has the 
-interpretation that 
-in order to allocate optimally (according to $F$) we must divide resoruces among the locations so that the forecasted probability of need exceeding local allocation is the same everywhere.
-A related interpretation is offered by our model's expression as a partial derivative (derived in the supplement) for the reduction in expected score (i.e., increase in expected utility) in location $i$ that results from a unit increase in the resource:
+for some constant $\lambda$ independent of $i$. It turns out that these derivatives can also be expressed in terms of 
+the marginal forecasts $F_i$ as
 \begin{align}
-\frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^{F,K}} = L \cdot (1-\tau).
+ \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] = L \cdot (F_i(x_i) - 1)
 \end{align}
-This equation tells us that when our allocations are set to the Bayes act $x^{F,K}$, shifting a small amount $\Delta x$ of the resource from one location to another would yield no change in expected utility since the gain $\Delta x \cdot L \cdot (1-\tau)$ in the receiving location would be canceled out by the loss $\Delta x \cdot L \cdot (\tau-1)$ in the location giving up the resource.
+so that 
+\begin{align}
+F(x_i^{F,K}) = 1-\lambda/L.
+\end{align}
+This tells us that in order to allocate optimally (according to $F$), we must divide resources among the locations so that there is an equal forecasted probability in every location that the allocation is sufficient to meet resource need.
+It also allows us to write the components $x_i^{F,K}$ of the Bayes act as quantiles
+\begin{align}
+x_i^{F,K} = F_i^{-1}(1 - \lambda/L)
+\end{align}
+which can be substituted into the constraint equation of the allocation problem to get an equation for $\lambda$,
+\begin{align}
+\sum_{i=1}^N F_i^{-1}(1 - \lambda/L) = K.
+\end{align}
+A solution to this equation---which almost always must be approximated with a numerical method---can be substituted back into the quantile formulas
+to arrive at an explicit expression for the Bayes act.
 This partial solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
-There, products and order quantities play the roles we have given to locations and resource allocation levels.
+We note that $\lambda = \lambda(K,F)$ and the probability level $1-\lambda/L$ depend on both the constraint level $K$ defining the decision problem and on the forecast $F$.
+This is quite different from the situation for quantile scores and the WIS, where the probability levels used to calculate scores are fixed.
 
@@
 \begin{equation}
-S_A(F, y; K) = s_A(x^{F,K}, y) = \sum_{i=1}^n L \max(0, y_i - F_i^{-1}(\tau(K,L,F)))
+S_A(F, y; K) = s_A(x^{F,K}, y) = \sum_{i=1}^n L \max(0, y_i - F_i^{-1}(1- \lambda(K,F)/L))
 \end{equation}

### 2023-09-11 alloscore_manuscript/methods_apg.tex
AS approx idea start
Talk about KKT.

@@
 \end{align}
-The AP has a convex objective function and affine constraint, so from general convex optimization theory 
-(see e.g. \cite{ruszczynski2011nonlinear}, Chapter 4), it is equivalent to the
-problem of finding a saddle point $(x^{F,k}, \lambda^{\star})$ of the associated Lagrangian 
+The AP has a convex objective function and affine constraint, so from general convex optimization theory, a solution $x^{F,k}$ must satisfy a set of \emph{Karush-Kuhn-Tucker} (KKT) conditions.
+These include that for any $i$ such that $x_i^{F,K} > 0$,
 \begin{align}
+\frac{\partial}{\partial x_i} E_{F}\left[s_A(Y,x^{F,k})\right] &= L(F_i(x_i^{F,k})-1) = \lambda
+\quad (\text{cf. \eqref{eqn:q_deriv}}) \label{eqn:x_stationary}
+\end{align}
+for some $\lambda$ independent of $i$. These equations follow from the assumption that when $x=x^{F,k}$, the objective $E_{F}\left[s_A(Y,x)\right]$ cannot be reduced by any small
+resource shifts between locations and they provide a quantile formula for any non-zero optimal allocation level,
+\begin{align}
+x_i^{F,k} = F_i^{-1}(1-\lambda/L). \label{eqn:xfk_formula}
+\end{align}
+Another KKT condition is the constraint equation, which can now be written in terms of $\lambda$ as
+\begin{align}
+0 = \sum_{i \in I_+} F_{i}^{-1}(1-\lambda/L) - k, \label{eqn:lambda_stationary2}
+\end{align}
+where $I_+ = \{i \mid x_i^{F,k} > 0\}$.
+Since convexity and compactness guarantee existence of a solution, the problem is now reduced to solving 
+\eqref{eqn:lambda_stationary2} for the set 
+$\{i \mid x_i^{F,k} > 0\}$ and a root $\lambda^{\star}$, in terms of which we can exrpess the Bayes act as
+\begin{align}
+x_i^{F,k} = 
+\begin{cases}
+F_i^{-1}(1-\lambda^{\star}/L) &  i \in I_+ \\
+0 & \text{otherwise}.
+\end{cases}
+\end{align}
+Note, however, that this is a more complicated task than just solving \eqref{eqn:lambda_stationary2}
+for $\lambda^{\star}$ under the assumption that $I_+ = \{1,\ldots,N\}$.\footnote{Such an assumption would exclude fairly simple examples in which $F$ has support bounded away from $0$ in one coordinate but not in another.}
+
+The task can be simplified though by turning our attention to the \emph{dual} formulation of the AP.
+(See e.g.
+\cite{ruszczynski2011nonlinear}, Chapter 4, which we follow here.) This involves recasting the KKT equations \eqref{eqn:x_stationary}
+and the constraint $T_x=k$ as the first order conditions in variables $x$ and $\lambda$, respectively, for $(x^{F,k}, \lambda^{\star})$ to be a saddle point of the \emph{Lagrangian}
+\begin{align}{}
 \mathcal{L}(x, \lambda) = \mathcal{L}(x, \lambda; F, k, L) 
@@
 \end{align}
-in the region $\{x \geq 0, \lambda \in \mathbb{R}\}$.\footnote{The Lagrangian in this context has the interpretation as the best (lowest) objective function value we can achieve if we forced to choose $x$ but can buy or get credit for changes the constraint level $k$
+in the region $\{x \geq 0, \lambda \in \mathbb{R}\}$.\footnote{The Lagrangian in this context has the interpretation as the best (lowest) objective function value we can achieve if we forced to choose $x$ but can buy or get credit for changes in the constraint level $k$
 at the ``price'' $\lambda$.}
-The Lagrangian reformulation in particular recasts the constraint $T_x=k$ of the original problem as a first-order condition on $\mathcal{L}$ with respect to the new \emph{Lagrange multiplier} variable $\lambda$:
-\begin{align}
-0 = \del{}{\lambda} \mathcal{L}(x^{F,k}, \lambda^{\star}) &= \sum_{i=1}^{N} x_i^{F,k} - k. \label{eqn:lambda_stationary}
-\end{align}
-But for $(x^{F,k}, \lambda^{\star})$ to be a saddle point we also need
-\begin{align}
-0 = \frac{\partial}{\partial x_i} \mathcal{L}(x^{F.k}, \lambda) &= L(F_i(x_i^{F,k})-1) + \lambda^{\star} 
-\quad (\text{cf. \eqref{eqn:q_deriv}}) \label{eqn:x_stationary}
-\end{align}
-for any $x_i^{F,k} > 0$, which gives a formula 
-\begin{align}
-x_i^{F,k} = F_i^{-1}(1-\lambda^{\star}/L) \label{eqn:xfk_formula}
-\end{align}
-for any non-zero element $x_i^{F,k}$ of the solution. Thus \eqref{eqn:lambda_stationary} becomes
-\begin{align}
-0 = \sum_{\{i | x_i^{F,k} > 0\}} F_{i}^{-1}(1-\lambda^{\star}/L) - k. \label{eqn:lambda_stationary2}
-\end{align}
-
-Since convexity and compactness guarantee existence of a solution, the problem is now reduced to solving \eqref{eqn:lambda_stationary} for the set 
-$\{i \mid x_i^{F,k} > 0\}$ and value $\lambda^{\star}$. Note, however, that this is a more complicated task than solving just 
-for $\lambda^{\star}$ under the assumption that \eqref{eqn:xfk_formula} holds for all coordinates of $x^{F,k}$.\footnote{Such an assumption would exclude fairly simple examples in which $F$ has support bounded away from $0$ in one coordinate but not in another.} The task is simplified though by bringing in the full saddle point condition for $\lambda^{\star}$ which implies not only \eqref{eqn:lambda_stationary} but also that
-$\lambda^{\star}$ is in fact the unique maximizer of the \emph{dual function} 
-$\mathcal{L}_D(\lambda) := \inf_{x \in \mathbb{R}_{+}^N} \mathcal{L}(x, \lambda)$.
-This approach is especially suitable in our situation because our original objective function $E_F\left[s_A(Y,x)\right]$ is decomposable in the coodinates $x_i$, allowing the dual function to be computed separately with respect to each coordinate:
+The \emph{dual function} is then defined as $\mathcal{L}_D(\lambda) := \inf_{x \in \mathbb{R}_{+}^N} \mathcal{L}(x, \lambda)$. 
+From general theory, $\mathcal{L}_D(\lambda)$ is concave and $\lambda^{\star}$ is its unique maximizer.
+The reason why the dual approach is especially suitable in our situation is that 
+our original objective function $E_F\left[s_A(Y,x)\right]$ is decomposable in the coodinates $x_i$, allowing the corresponding infima in $\mathcal{L}_D$ to be computed separately:
 \begin{align}
@@
 \end{align}
-which for $F_i$ in a standard parametric family will often have simple the form $\chi_i(\lambda) = \{F_i^{-1}(1-\lambda/L)\}$ but could an interval containing $0$ or $\infty$.
-From general theory, $\mathcal{L}_D(\lambda)$ is concave, and while it may not be differentiable when the $x_i$'s attaining the infima in \eqref{eqn:dual_decomp} are zero, it does have as a subdifferential the well-defined interval
+which for $F_i$ in a standard parametric family will often simply be $\{F_i^{-1}(1-\lambda/L)\}$ but could an interval containing $0$ or $\infty$.
+While $\mathcal{L}_D(\lambda)$ may not be differentiable when the $x_i$'s attaining the infima in \eqref{eqn:dual_decomp} are zero, it does have as a well-defined subdifferential which is the sum of intervals
 \begin{align}
 \partial \mathcal{L}_D(\lambda) 
-&= \sum_{i=1}^N \chi_i(\lambda) - k.
+&= \sum_{i=1}^N \chi_i(\lambda) - k \\
+&= \left[ \sum_{i=1}^N \min(\chi_i(\lambda))  - k, \sum_{i=1}^N \max(\chi_i(\lambda))  - k\right]
 \end{align}
+
+% The Lagrangian reformulation in particular recasts the constraint $T_x=k$ of the original problem as a first-order condition on $\mathcal{L}$ with respect to the new \emph{Lagrange multiplier} variable $\lambda$:
+% \begin{align}
+% 0 = \del{}{\lambda} \mathcal{L}(x^{F,k}, \lambda^{\star}) &= \sum_{i=1}^{N} x_i^{F,k} - k. \label{eqn:lambda_stationary}
+% \end{align}
+% But for $(x^{F,k}, \lambda^{\star})$ to be a saddle point we also need
+% \begin{align}
+% 0 = \frac{\partial}{\partial x_i} \mathcal{L}(x^{F.k}, \lambda) &= L(F_i(x_i^{F,k})-1) + \lambda^{\star} 
+% \quad (\text{cf. \eqref{eqn:q_deriv}}) \label{eqn:x_stationary}
+% \end{align}
+% for any $x_i^{F,k} > 0$, which gives a formula 
+% \begin{align}
+% x_i^{F,k} = F_i^{-1}(1-\lambda^{\star}/L) \label{eqn:xfk_formula}
+% \end{align}
+% for any non-zero element $x_i^{F,k}$ of the solution. Thus \eqref{eqn:lambda_stationary} becomes
+% \begin{align}
+% 0 = \sum_{\{i | x_i^{F,k} > 0\}} F_{i}^{-1}(1-\lambda^{\star}/L) - k. \label{eqn:lambda_stationary2}
+% \end{align}
+
 This of course looks very similar to the right hand side of \eqref{eqn:lambda_stationary2}, but the dual perspective has the key advantage of not requiring any guesses of the set $\{i \mid x_i^{F,k} > 0\}$ when initiating a search algorithm for $\lambda^{\star}$.
@@
 
+\section{Approximating the allocation score}
+
+cdf of $TY_F$:
+\begin{align}
+TF(K) = \sup \left\{\tau \in [0,1] \mid \sum F_i^{-1}(\tau) \leq K \right\} 
+\end{align}
+Given a set of distinct and increasing probability levels $\boldsymbol{\alpha} = \{\alpha_1,\ldots,\alpha_J\} \subset (0,1)$, we can define a function
+\begin{align}
+j_{\boldsymbol{\alpha},-}(K) = \max \left\{j \in \{1,\ldots,J\} \mid \sum F_i^{-1}(\alpha_j) \leq K \right\} 
+\end{align}
+
+An approximation of $S_{A,K}(F,y)$ using elicited quantiles:
+\begin{align}
+S_{A, \boldsymbol{\alpha}}(F,y) := w_1 s_A(q_{F,j}, y) + w_2 s_A(q_{F,j+1}, y)
+\end{align}
+
 \section{Discussion}

### 2023-09-14 alloscore_manuscript/alloscore-application.Rnw
Update alloscore_manuscript/alloscore-application.Rnw
Co-authored-by: Evan Ray <elray1@users.noreply.github.com>

@@
 \end{align}
-for some constant $\lambda$ independent of $i$. It turns out that these derivatives can also be expressed in terms of 
+for some constant $\lambda$ that is the same for all locations $i$. It turns out that these derivatives can also be expressed in terms of 
 the marginal forecasts $F_i$ as
@@
 \end{align}
-A solution to this equation---which almost always must be approximated with a numerical method---can be substituted back into the quantile formulas
+The value of $\lambda$ that solves this equation---which almost always must be approximated with a numerical method---can be substituted back into the quantile formulas
 to arrive at an explicit expression for the Bayes act.

### 2023-09-14 alloscore_manuscript/alloscore-application.Rnw
Update alloscore_manuscript/alloscore-application.Rnw
Co-authored-by: Evan Ray <elray1@users.noreply.github.com>

@@
 This partial solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
-We note that $\lambda = \lambda(K,F)$ and the probability level $1-\lambda/L$ depend on both the constraint level $K$ defining the decision problem and on the forecast $F$.
+We note that $\lambda$ and the probability level $1-\lambda/L$ depend on both the constraint level $K$ defining the decision problem and on the forecast $F$.
 This is quite different from the situation for quantile scores and the WIS, where the probability levels used to calculate scores are fixed.

### 2023-09-15 alloscore_manuscript/alloscore-application.Rnw

@@
 expected loss under the constraint is that there is no way to decrease expected loss further by shifting a small amount of the resource from one location to another.
-As a consequence, at $x^{F,K}$, all the rates of change of expected loss with respect to the amount of resources allocated to a particular location $i$ must be equal:
+As a consequence, at $x^{F,K}$, adding resources to any particular location $i$ reduces expected loss at the same rate $\lambda$:
 \begin{align}
- \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^{F,K}} = \lambda
+\frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^{F,K}} = -\lambda.
 \end{align}
-for some constant $\lambda$ that is the same for all locations $i$. It turns out that these derivatives can also be expressed in terms of 
-the marginal forecasts $F_i$ as
+It turns out that these derivatives can also be expressed in terms of the marginal forecasts $F_i$ as
 \begin{align}

### 2023-09-28 alloscore_manuscript/methods_apg.tex
intro opportunity loss; fix WIS derivation; add copula in Norm1.R

@@
 The resource has a unit-cost of $C$, and we assign a loss of $L$ to the result that a sick individual cannot access the resource.
-This specifies a loss function
+We assume that $C < L$ since otherwise -- at least according to the quantification $L$ -- there is no ``cost-effectiveness'' in trying to prevent even a single case of unmet need.
+With $C$ and $L$, there are two natural loss function we can define.
+First, we have the \emph{absolute} loss
 \begin{align}
-s_{Q,C}(x,y) = Cx + L(y-x)_+ \label{eqn:quantile_loss}
+s_{Q,C,a}(x,y) = Cx + L(y-x)_+ \label{eqn:quantile_loss}
 \end{align}
-where $u_+ = \max(u,0)$.  Loss functions of this general form are referred to as \emph{piece-wise linear}, distinguishing them from smooth non-linear loss functions such as squared error.
+where $u_+ = \max(u,0)$.  Alternatively, we have the \emph{opportunity} loss 
+\begin{align}
+s_{Q,C,o}(x,y) &= Cx + L(y-x)_+ - Cy \\
+&= C(x-y)_+ + (L-C)(y-x)_+ \\
+&= (C - L \mathbf{1}\{y > x\})(x-y) \\
+&= L(\mathbf{1}\{y \leq x\} - 1 + C/L)(x - y)\label{eqn:quantile_loss_opp}
+\end{align}
+which considers only the loss relative to that of an oracle that is able to minimize loss \textit{ex-post} by procuring 
+exactly $y$ units of the resource.
+
+Loss functions of these general forms are referred to as \emph{piece-wise linear}, distinguishing them from smooth non-linear loss functions such as squared error.
+The opportunity loss $s_{Q,C,o}$ is sometimes called ``pinball'' loss.
+It has the convenien{}t
+{}property that $s_{Q,C,o}(z,z)=0$, but loses some of the interpretability of $s_{Q,C}$ (see \ref{} below).
 The underlying decision problem here consists of action and outcome sets $\mathcal{X},\mathcal{Y}$ both equal to $\mathbb{R}_+$, the non-negative reals.
-(We ignore for now any discreteness of the resource.) The results set $\mathcal{R}$ is abstractly a set of over-expenditure regrets and numbers of sick individuals with unmet resource need, but the loss function $s$ maps $\mathcal{C}$ into $\mathbb{R}_+$.
-We also from now on make the assumption that $C < L$ since otherwise -- at least according to this loss function -- there is no ``cost-effectiveness'' in trying to prevent even a single case of unmet need.
-This would imply a Bayes act $x^F$ that is identically 0, telling us nothing about the value of $F$.
+(We ignore for now any discreteness of the resource.) The results set $\mathcal{R}$ is abstractly a set of transfers of a monetary or similar nature and sick individuals with unmet resource need, but the loss function $s$ maps $\mathcal{R}$ into $\mathbb{R}_+$.
 
-The Bayes act for $C \in (0,L)$ is the solution of the first order equation 
-\begin{align}
+Note that the absolute and opportunity losses have the same dependence on action variable $x$, and therefore the same Bayes act. For $C \in (0,L)$ this is the solution of the first order equation 
+\begin{align}{}
 0 = \dby{}{x} E_F\left[s_{Q,C}(x,y)\right] &= E_F\left[\dby{}{x}s_{Q,C}(x,y)\right] \\
@@
 
-Finally, we evaluate the scoring rule $S_Q$ on $F$ when outcome $y$ realizes by measuring the loss that occurs when the action $q_{F,\tau}$ is taken:
+Finally, we evaluate the scoring rules $S_{Q,C,a}$ and $S_{Q,C,o}$ on $F$ when outcome $y$ realizes by measuring the loss that occurs when the action $q_{F,\tau}$ is taken:
 \begin{align}
-S_Q(F,y) = s_Q(q_{F,\tau}, y) = Cq_{F,\tau} + L(y- q_{F,\tau})_{+}.
+S_{Q,C,a}(F,y) &= s_{Q,C,a}(q_{F,\tau}, y) = Cq_{F,\tau} + L(y- q_{F,\tau})_{+} \\
+S_{Q,C,o}(F,y) &= s_{Q,C,o}(q_{F,\tau}, y) = (C - L \mathbf{1}\{y > x\})(q_{F,\tau}-y).
 \end{align}
@@
 
-We take into account the results of actions under a set $\mathcal{C} = \{C_j\}$ of possible costs by expanding the loss function to be a weighted mean of quantile losses under certainty about the cost:
+We take into account the results of actions under a set $\mathcal{C} = \{C_j\}$ of possible costs by expanding 
+(either the absolute or opportunity version of) the loss function to be a weighted mean of quantile losses under certainty about the cost:
 \begin{align}
-s_{Q,\mathcal{C}, p_C}(x,y) = \frac{1}{m}\sum_{j=1}^{m}p_j(C_j x_j + L(x_i - y)_+). \label{eqn:sum_CL_loss}
+s_{Q,\mathcal{C}, p_C}(x,y) = \sum_{j=1}^{m}p_js_{Q,C_j}(x_j,y). \label{eqn:sum_CL_loss}
 \end{align}
@@
 \begin{align}
-s_{Q,f_C}(x,y) = \frac{1}{L}\int_{0}^{L} (cx(c) + L(x(c) - y)_+) f_C(c)dc{}. \label{eqn:CRPS_loss}
+s_{Q,f_C}(x,y) = \int_{0}^{L} s_{Q,c}(x(c),y) f_C(c)dc{}. \label{eqn:CRPS_loss}
 \end{align}
@@
 \begin{align}
-S_{Q,\mathcal{C}, p_C}(F,y) &= \frac{1}{m}\sum_{j=1}^{m}p_j(C_j q_{F,\tau_j} + L(q_{F,\tau_j} - y)_+) \\
-S_{Q, f_C}(F,y) &= \frac{1}{L}\int_{0}^{L}(cq_{F,\tau(c)} + L(q_{F,\tau(c)} - y)_+) f_C(c)dc. \label{eqn:CRPS_cform}
+S_{Q,\mathcal{C}, p_C}(F,y) &= \sum_{j=1}^{m}p_js_{Q,C_j}(q_{F,\tau(C_j)},y) \\
+S_{Q, f_C}(F,y) &= \int_{0}^{L} s_{Q,c}(q_{F,\tau(c)},y) f_C(c)dc. \label{eqn:CRPS_cform}
 \end{align}
@@
 How do the WIS and CRPS arise from this construction? Defining $\tau = 1-C/L$ as a random variable gives it the density 
-$f_{\tau}(\tau) = -f_C\left(L(1-\tau)\right)/L$, 
-and \eqref{eqn:CRPS_cform} becomes 
+$f_{\tau}(\tau) = f_C\left(L(1-\tau)\right)\cdot L$, 
+and \eqref{eqn:CRPS_cform} with the opportunity loss \eqref{eqn:quantile_loss_opp} becomes 
 \begin{align} 
-S_{Q, f_C}(F,y) &= \int_{0}^{1}((1-\tau)q_{F,\tau} + (y-q_{F,\tau})_{+})f_{\tau}(\tau)d\tau \\
-&= \int_{0}^{1}(\mathbf{1}\{y \leq q_{F,\tau}\}-\tau)(q_{F,\tau} - y)f_{\tau}(\tau)d\tau + \int_{0}^{1}(1-\tau)y f_{\tau}(\tau)d\tau \\
-&= w\mathrm{CRPS}(F,y) + (1-E[\tau])y
+S_{Q, f_C}(F,y) &= \int_{0}^{L}L(\mathbf{1}\{y \leq q_{F,\tau(c)}\}-\tau(c))(q_{F,\tau(c)} - y)f_{C}(c)dc \\
+&= \int_{0}^{1}(\mathbf{1}\{y \leq q_{F,\tau}\}-\tau)(q_{F,\tau} - y)f_{\tau}(\tau)d\tau \\
+&= w\mathrm{CRPS}(F,y)
 \end{align}
 where $w$CRPS is the weighted CRPS of \cite{gneiting2011weightedScoringRules} with quantile weight function $\nu(\tau) = \frac{1}{2}f_{\tau}(\tau)$. 
-(The $(1-E[\tau])y$ that results from our loss function $s_Q$ adding a $Cy$ term to the standard quantile loss function does not effect the dependence of the scoring rule of $F$.)
+Using the absolute loss instead of opportunity loss would add a term $E[C]y$. This is the value of the absolute loss scoring rule 
+at the oracle's point mass forecast: $S_{Q, f_C, a}(\delta_y,y)$.
 
-Writing $s_{Q,\tau}(x,y) = s_{Q,C}(x,y)/L = (1-\tau)x + (y-x)_{+}$, we have
-\begin{align}
-s_{Q,\tau}(q_{F,\tau},y) + s_{Q,1-\tau}(q_{F,1-\tau},y) &= q_{F,\tau} + \tau(q_{F,1-\tau} - q_{F,\tau}) + (y - q_{F,\tau})_{+} + (y - q_{F,1-\tau})_{+} \\
-&= y + \tau\left(q_{F,1-\tau} - q_{F,\tau} + \frac{1}{\tau}(q_{F,\tau} - y)_{+} + \frac{1}{\tau}(y - q_{F,1-\tau})_{+}\right) \\
-&= y + \tau \mathrm{IS}_{2\tau}(F,y)
-\end{align}
-The scoring rule $\mathrm{IS}_{\alpha}(F,y)$ is the \emph{interval score} suggested by expression (43) in \cite{gneiting2007strictly} 
-and formalized by \cite{bracher2021evaluating} as a means for defining the weighted interval score 
+To connect with the WIS, we recall it's definition by \cite{bracher2021evaluating}:
 \begin{align}
@@
 \end{align}
-So by taking $\mathcal{C} = \{(1 \pm \alpha_j)L/2 \mid j > 0\}\cup \{L/2\}$ and $p_\mathcal{C}$ symmetric across $L/2$, we get 
+The scoring rule $\mathrm{IS}_{\alpha}(F,y)$ is the \emph{interval score} suggested by expression (43) in \cite{gneiting2007strictly} 
+and formalized by \cite{bracher2021evaluating} as
 \begin{align}
-S_{Q,\mathcal{C}, p_C}(F,y) = \mathrm{WIS}_{\alpha_{\{0: J\}}}(F, y) + y \sum_{j=0}^{J} w_j
+\operatorname{IS}_{\alpha_j}(F, y) = \frac{\alpha_j}{2}(q_{F,\alpha_j/2}-y)_+ + \frac{\alpha_j}{2}(y-q_{F,1-\alpha_j/2})_+ 
+ + (q_{F,1-\alpha_j/2} - q_{F,\alpha_j/2})
 \end{align}
-with $w_j = p_{j}\alpha_j/2$ and $w_0 = p_{L/2}$. Setting $p_j=1/m$ and $\alpha_j=0.9, \ldots, 0.2,0.1,0.05,0.02, j = 1,\ldots,11$ gives the WIS currently used in 
-the COVID-19 Forecast Hub.
+Thus the $\alpha_j$ are twice the probability level of the lower endpoint of a central
+$1-\alpha_j$ prediction interval. The COVID-19 Forecast Hub chooses $\alpha_j=0.02, 0.05, 0.1, \ldots, 0.9, j = 1,\ldots,11$.
+
+Now write $s_{Q,\tau, o}(x,y) = s_{Q,C, o}(x,y)/L = (1-\tau)(x-y)_+ + \tau(y-x)_+$, 
+and assume $\tau < 1/2$.  We have
+\begin{align}
+s_{Q,\tau,o}(q_{F,\tau},y) &+ s_{Q,1-\tau,o}(q_{F,1-\tau},y) \\
+&= (1-\tau)(q_{F,\tau}-y)_+ + \tau(y-q_{F,\tau})_+ 
++ \tau(q_{F,1-\tau}-y)_+ + (1-\tau)(y-q_{F,1-\tau})_+ \\ 
+&= (q_{F,\tau}-y)_+ + (y-q_{F,1-\tau})_+ \\
+&\quad + \tau \left[(y-q_{F,\tau})_+ - (q_{F,\tau}-y)_+ + (q_{F,1-\tau}-y)_+ - (y-q_{F,1-\tau})_+\right] \\
+&= (q_{F,\tau}-y)_+ + (y-q_{F,1-\tau})_+  + \tau(q_{F,1-\tau} - q_{F,\tau}) \\
+&= \tau \mathrm{IS}_{2\tau}(F,y). \label{eqn:pinball_IS}
+\end{align}
+
+To reproduce the WIS we set $\mathcal{C} = \{C_k\}, k = 0,\ldots 2J$ where 
+\begin{align}
+C_0 &= L/2 \\
+\alpha_k/2 &= 1-C_k/L, k=1,..J \\
+1-\alpha_{k-J}/2 &= C_k/L, k = J+1,\ldots,2J
+\end{align}
+(note that $C_k$ decrease as $k$ increases for $k>0$)
+and let $p_{\mathcal{C}}$ be a probability distribution symmetric across $L/2$. This gives us
+\begin{align}	
+S_{Q,\mathcal{C}, p_C, o}(F,y) &= p_0 s_{Q,L/2, o}(q_{F,.5},y) + \sum_{k = 1}^{2J} p_k s_{Q,C_k, o}(q_{F, 1-C_k/L},y) \\
+& =  p_0 s_{Q,L/2, o}(q_{F,.5},y) + \sum_{k = 1}^{J} L(1-C_k/L)p_k \mathrm{IS}_{2(1-C_k/L)}(F,y) \\
+& = \frac{L}{2} \left(p_0|y-q_{F,.5}| + \sum_{j = 1}^{J} \alpha_j p_j \mathrm{IS}_{\alpha_j}(F,y)\right) \\
+& \propto \mathrm{WIS}_{\alpha_{\{0: J\}}}(F, y) \label{eqn:WIS_propto}
+\end{align}	
+if we set $p_j \propto w_j/\alpha_j$ and define $\alpha_0 = 1$. The COVID-19 Forecast Hub chooses $w_j = \alpha_j/2$ 
+so that the WIS there corresponds to a uniform distribution over the $C_k$ where each cost has probability
+\begin{align}
+p = \frac{1/2}{\sum_{k=0}^{2J} 1/2} = \frac{1/2}{J+1/2}.
+\end{align}
+In this case the constant of proportionality in \eqref{eqn:WIS_propto} is $L/2$.
 
@@
 Similarly as for the expansion to multiple cost contingencies---though now with respect to forecasts $F_i$ depending on the 
-outcome component $y_i$---first order conditions for optimality decouple under the additive separability of the loss function so that the Bayes act $x^F$ for $F$ is just a vector of quantiles $q_{F,\tau} := (q_{F_1,\tau},\ldots,q_{F_N,\tau})$ or quantile functions when $\tau = 1 - c/L$, $c \in \mathcal{C}$ or $[0,L]$.
+index $i$ of the relevant outcome component $y_i$---first order conditions for optimality decouple under the additive separability of the loss function so that the Bayes act $x^F$ for $F$ is just a vector of quantiles $q_{F,\tau} := (q_{F_1,\tau},\ldots,q_{F_N,\tau})$ or quantile functions when $\tau = 1 - c/L$, $c \in \mathcal{C}$ or $[0,L]$.
 Note that while $x^F=q_{F,\tau}$ only depends on the marginals
-$F_i$, it is indeed the Bayes act for the joint $F$ since $E_F[s_{Q,c}(x_i, y_i)] = E_{F_i}[s_{Q,c}(x_i, y_i)]$.
+$F_i$, it is indeed the Bayes act for the joint $F$ since $E_F[s_{TQ,c}(x, y)] = \sum_{i=1}^{N}E_{F_i}[s_{Q,c}(x_i, y_i)]$.
 
@@
 \begin{align}
-s_{tQ}\left(x, y\right) = Ck + \sum_{i=1}^{N} L(y_i - x_i)_{+}.
+s_{TQ}\left(x, y\right) = Ck + \sum_{i=1}^{N} L(y_i - x_i)_{+}.
 \end{align}
-The component of the Bayes act for the contingency that $K=k$ is thus independent of the realized cost $C$, motivating the definition of a new single loss function (component), the \emph{allocation loss function}
+The component $x_i^{F,k}$ of the Bayes act for the contingency that $K=k$ is thus independent of the realized cost $C$, motivating the definition of a new single loss function (component), the \emph{allocation loss function}
 \begin{align}
@@
 we use in deriving these Bayes acts components for all $\mathcal{X}_k$.
+
+We show in section \ref{sec:solvingAP} that $x^{F,k}$ is defined by a single probability level $\tau$ just as was the Bayes act $x^{F,c}$ for $s_{TQ,C}$ on $\mathcal{X}_{cert}$, but that in this case, $\tau = \tau(k, F)$ depends not only on the decision problem
+parameter $k$, but also on the forecast $F$ itself.
 
@@
 \section{Solving the Allocation Problem}
+\label{sec:solvingAP}
 
@@
 \eqref{eqn:lambda_stationary2} for the set 
-$\{i \mid x_i^{F,k} > 0\}$ and a root $\lambda^{\star}$, in terms of which we can exrpess the Bayes act as
+$\{i \mid x_i^{F,k} > 0\}$ and a root $\lambda^{\star}$, in terms of which we can express the Bayes act as
 \begin{align}
@@
 Note, however, that this is a more complicated task than just solving \eqref{eqn:lambda_stationary2}
-for $\lambda^{\star}$ under the assumption that $I_+ = \{1,\ldots,N\}$.\footnote{Such an assumption would exclude fairly simple examples in which $F$ has support bounded away from $0$ in one coordinate but not in another.}
+for $\lambda^{\star}$ under the assumption that $I_+ = \{1,\ldots,N\}$.\footnote{Such an assumption would exclude fairly simple examples in which the marginal densities $f_i$ of $F$ have different non-zero limits as the $x_i \to 0$.}
 
@@
 The reason why the dual approach is especially suitable in our situation is that 
-our original objective function $E_F\left[s_A(Y,x)\right]$ is decomposable in the coodinates $x_i$, allowing the corresponding infima in $\mathcal{L}_D$ to be computed separately:
+our original objective function $E_F\left[s_A(Y,x)\right]$ is decomposable in the coodinates $x_i$, allowing the infimum
+$\mathcal{L}_D$ to be computed as a sum of separate infima in these coordinates:
 \begin{align}

### 2023-09-28 random_undocumented_experiments/Norm1.R

@@ -3,13 +3,14 @@ library(tidyverse)
 library(scoringRules)
 library(patchwork)
 library(ggridges)
+library(copula)
 source("random_undocumented_experiments/scoring_helpers.R")
 
 dgp_params <- tibble(
   model = "dgp",
   target_names = paste0("site", 1:2),
   dist = "norm",
-  mean = c(200, 100),
+  mean = c(200, 180),
   sd = 50)
 
 dgp <- dgp_params %>%
@@ -20,8 +21,9 @@ y_gen <- function() {with(dgp, map2_dbl(r, 1, exec) %>% set_names(target_names))
 
 modA <- dgp_params %>% mutate(
   model = "A",
-  mean = mean[c(2,1)],
-  sd = 10) %>%
+  mean = c(190, 190),
+  #mean = mean[c(2,1)],
+  sd = 20) %>%
   add_pdqr_funs() %>%
   add_crps_funs()
 
@@ -43,7 +45,37 @@ mods <- bind_rows(dgp, modA, modB, modC)
 # select(mods, model, target_names, f) %>%
 #   pivot_wider(values_from = f, names_from = c(model, target_names))
 
-ys <- tibble(samp = 1:200, ys = map(seq_along(samp), ~y_gen()))
+ys <- tibble(samp = 1:1000, ys = map(seq_along(samp), ~y_gen()))
+
+correlate_samples <- function(ys, rho = .5) {
+  # Extract site1 and site2 values from ys
+  Y1 <- sapply(ys, function(x) x["site1"])
+  Y2 <- sapply(ys, function(x) x["site2"])
+
+  # Transform data to have uniform marginals using empirical CDF
+  pY1 <- ecdf(Y1)
+  pY2 <- ecdf(Y2)
+  U1 <- pY1(Y1)
+  U2 <- pY2(Y2)
+
+  # Fit a copula; for simplicity, we'll use a Clayton copula
+  gcop <- normalCopula(param = rho, dim = 2)
+  fit <- fitCopula(gcop, cbind(U1, U2))
+
+  # Sample from the copula
+  copula_sample <- rCopula(length(Y1), fit@copula)
+
+  # Transform the samples back to original marginal distributions
+  Z1 <- as.numeric(quantile(Y1, copula_sample[,1]))
+  Z2 <- as.numeric(quantile(Y2, copula_sample[,2]))
+
+  # Construct the list of samples
[…]

### 2023-10-10 alloscore_manuscript/alloscore-application.Rnw
Add oracle adjustment definition.

@@
 
-\paragraph{Step 3: Define the scoring rule.} We can now define a proper scoring rule for the probabilistic forecast $F$ as
-\begin{equation}
-S_A(F, y; K) = s_A(x^{F,K}, y) = \sum_{i=1}^n L \max(0, y_i - F_i^{-1}(1- \lambda(K,F)/L))
-\end{equation}
-This score measures the total unmet need across all locations that results from using the Bayes allocation associated with the forecast $F$ when the actual level of need in each location is observed to be $y_i$.
+\paragraph{Step 3: Define the scoring rule.} We can now use the Bayes act to define a proper scoring rule for the probabilistic forecast $F$.  
+Consider first the ``raw'' score defined as
+\begin{align}
+S_A^{\text{raw}}(F, y; K) = s_A(x^{F,K}, y) = \sum_{i=1}^N L \max(0, y_i - F_i^{-1}(1- \lambda(K,F)/L))
+\end{align}
+This measures the total unmet need across all locations that results from using the Bayes allocation associated with the forecast $F$ when the actual level of need in each location is observed to be $y_i$.
+But to make this a more easily interpreted measure of
+forecast performance, we will adjust it by the the minimum loss $l_{\text{oracle}}(y;K)$ achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$ which is $l_{\text{oracle}}(y;K) = L \max(0, \sum_{i=1}^{N}y_i - K)$.
+Our scoring rule is then given by
+\begin{align}
+S_A(F, y; K) &= S_A^{\text{raw}}(F, y; K) - l_{\text{oracle}}(y;K) \\
+& = \sum_{i=1}^N L \max(0, y_i - x_i^{F,K}) - L \max(0, \sum_{i=1}^{N}y_i - K) \\
+& = 
+\begin{cases}
+\sum_{i=1}^n L \max(0, y_i - x_i^{F,K}), &  \sum_{i=1}^{N}y_i \leq K \\
+LK - L\sum_{i=1}^{N} \min(x_i^{F,K}, y_i), &  \sum_{i=1}^{N}y_i > K.
+\end{cases}
+\end{align}
+In the first case, the oracle incurs no loss so that the raw and adjusted scores coincide.
+The second case can be read as taking the $K$ resource units perfectly allocated by the oracle as a base penalty on
+the imperfect forecast $F$ and then, for each location $i$, reducing this penalty by however much of the need $y_i$ 
+is met with the Bayes act component $x_i^{F,K}$.
+The oracle adjustment aligns with a central theme in modern decision theory across disciplines including economics, computer science, and psychology which is that 
+\emph{opportunity loss} (often known as \emph{regret} or (negative) \emph{relative utility}) is often a more important quantity than absolute loss from both a descriptive and normative perspective.
 

### 2023-10-17 alloscore_manuscript/alloscore-application.Rnw
draft of section to describe connections between WIS and alloscore

@@
 In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to.
-We give an overview of these procedures in Section \ref{sec:methods.decisiontheory}.
-There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
+We give an overview of these procedures in Section \ref{sec:methods.decisiontheory}.\ngr{This section label doesn't exist.
+Should it point to methods.detailed?} There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
 We review this work only briefly here, and we refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \elr{TODO: identify relevant review or book style summary for meteorology} for discussions focused on applications to economics and meteorology, respectively.
@@
 We begin with an informal description of the allocation score and some examples illustrating its key characteristics in section \ref{sec:methods.overview}.
-In section \ref{sec:methods.detailed} we develop the allocation score more carefully, building on decision theoretic procedures for deriving proper scoring rules.We then comment on some connections between the allocation score that we propose and other common scores that can be derived from decision theoretic foundations, such as the quantile score, WIS, and CRPS, in section \ref{sec:methods.related}.
+In section \ref{sec:methods.detailed} we develop the allocation score more carefully, building on decision theoretic procedures for deriving proper scoring rules.
+We then comment on some connections between the allocation score that we propose and other common scores that can be derived from decision theoretic foundations, such as the quantile score, WIS, and CRPS, in section \ref{sec:methods.related}.
 
@@
 %For a given probability level $\tau$, if the quantiles $Q_i(\tau)$ of the marginal forecasts are proportional to the observed needs $y_i$, then the allocation score is zero for the resource constraint level $K = \sum Q_i(\tau)$.
-This stands in marked contrast to other common scoring methods for multivariate forecasts that aggregate univariate scores such as CRPS or WIS for the marginal forecasts where a
-misalignment in one coordinate is penalized regardless of alignments in other coordinates.
+This stands in marked contrast to other common scoring methods for multivariate forecasts that aggregate univariate scores such as log score, CRPS, or WIS for the marginal forecasts where a
+poor forecast made for one unit (a location, say) is penalized regardless of alignments in other units.
 Note that we do not claim that the allocation score is generically preferable to these other scores \textemdash rather, it provides a view of forecast performance that is specifically tuned to the context of decision making about resource allocations.
@@
 the support of each $F_i$ is a subset of $\mathbb{R}^+$.
-Finally, we assume that each unit of unmet need incurs a loss denoted by $L$, so that if the selected resource level $x_i$ in location $i$ is less than the realized need $y_i$, a loss of $L(y_i - x_i)$ results.
+Finally, we assume that each unit of unmet need incurs a loss denoted by $L$, so that if the selected resource level $x_i$ in location $i$ is less than the realized need $y_i$, a loss of $L \cdot (y_i - x_i)$ results.
 
@@
 \end{align}
-where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]$ sums the expected loss due to unmet need across all locations.
+where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]$ sums the expected loss due to unmet need across all locations.
 
-In the supplement we derive a general form of this solution by beginning with the fact that a condition for minimizing
-expected loss under the constraint is that there is no way to decrease expected loss further by shifting a small amount of the resource from one location to another.
-As a consequence, at $x^{F,K}$, adding resources to any particular location $i$ reduces expected loss at the same rate $\lambda$:
-\begin{align}
-\frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^{F,K}} = -\lambda.
-\end{align}
-It turns out that these derivatives can also be expressed in terms of the marginal forecasts $F_i$ as
-\begin{align}
- \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] = L \cdot (F_i(x_i) - 1)
-\end{align}
-so that 
-\begin{align}
-F(x_i^{F,K}) = 1-\lambda/L.
-\end{align}
-This tells us that in order to allocate optimally (according to $F$), we must divide resources among the locations so that there is an equal forecasted probability in every location that the allocation is sufficient to meet resource need.
-It also allows us to write the components $x_i^{F,K}$ of the Bayes act as quantiles
-\begin{align}
-x_i^{F,K} = F_i^{-1}(1 - \lambda/L)
-\end{align}
-which can be substituted into the constraint equation of the allocation problem to get an equation for $\lambda$,
-\begin{align}
-\sum_{i=1}^N F_i^{-1}(1 - \lambda/L) = K.
-\end{align}
-The value of $\lambda$ that solves this equation---which almost always must be approximated with a numerical method---can be substituted back into the quantile formulas
-to arrive at an explicit expression for the Bayes act.
-This partial solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
-We note that $\lambda$ and the probability level $1-\lambda/L$ depend on both the constraint level $K$ defining the decision problem and on the forecast $F$.
+%%%% start of large excision %%%%%
+
+% In the supplement we derive a general form of this solution by beginning with the fact that a condition for minimizing
+% expected loss under the constraint is that there is no way to decrease expected loss further by shifting a small amount of the resource from one location to another.
+% As a consequence, at $x^{F,K}$, adding resources to any particular location $i$ reduces expected loss at the same rate $\lambda$:
+% \begin{align}
+% \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^{F,K}} = -\lambda.
+% \end{align}
+% It turns out that these derivatives can also be expressed in terms of the marginal forecasts $F_i$ as
+% \begin{align}
+%  \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] = L \cdot (F_i(x_i) - 1)
+% \end{align}
+% so that
+% \begin{align}
+% F(x_i^{F,K}) = 1-\lambda/L.
+% \end{align}
+% This tells us that in order to allocate optimally (according to $F$), we must divide resources among the locations so that there is an equal forecasted probability in every location that the allocation is sufficient to meet resource need.
+% It also allows us to write the components $x_i^{F,K}$ of the Bayes act as quantiles
+% \begin{align}
+% x_i^{F,K} = F_i^{-1}(1 - \lambda/L)
+% \end{align}
+% which can be substituted into the constraint equation of the allocation problem to get an equation for $\lambda$,
+% \begin{align}
+% \sum_{i=1}^N F_i^{-1}(1 - \lambda/L) = K.
+% \end{align}
+% The value of $\lambda$ that solves this equation---which almost always must be approximated with a numerical method---can be substituted back into the quantile formulas
+% to arrive at an explicit expression for the Bayes act.
+% This partial solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
+% We note that $\lambda$ and the probability level $1-\lambda/L$ depend on both the constraint level $K$ defining the decision problem and on the forecast $F$.
 This is quite different from the situation for quantile scores and the WIS, where the probability levels used to calculate scores are fixed.
+
+%%%% end of large excision %%%%%
 
@@
 \begin{align}
-S_A^{\text{raw}}(F, y; K) = s_A(x^{F,K}, y) = \sum_{i=1}^N L \max(0, y_i - F_i^{-1}(1- \lambda(K,F)/L))
+S_A^{\text{raw}}(F, y; K) = s_A(x^{F,K}, y) = \sum_{i=1}^N L \cdot \max(0, y_i - F_i^{-1}(1- \lambda(K,F)/L))
 \end{align}
@@
 But to make this a more easily interpreted measure of
-forecast performance, we will adjust it by the the minimum loss $l_{\text{oracle}}(y;K)$ achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$ which is $l_{\text{oracle}}(y;K) = L \max(0, \sum_{i=1}^{N}y_i - K)$.
+forecast performance, we will adjust it by the the minimum loss $l_{\text{oracle}}(y;K)$ achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$ which is $l_{\text{oracle}}(y;K) = L \cdot \max(0, \sum_{i=1}^{N}y_i - K)$.
 Our scoring rule is then given by
@@
 S_A(F, y; K) &= S_A^{\text{raw}}(F, y; K) - l_{\text{oracle}}(y;K) \\
-& = \sum_{i=1}^N L \max(0, y_i - x_i^{F,K}) - L \max(0, \sum_{i=1}^{N}y_i - K) \\
+& = \sum_{i=1}^N L  \cdot \max(0, y_i - x_i^{F,K}) - L \cdot  \max(0, \sum_{i=1}^{N}y_i - K) \\
 & =
 \begin{cases}
-\sum_{i=1}^n L \max(0, y_i - x_i^{F,K}), &  \sum_{i=1}^{N}y_i \leq K \\
-LK - L\sum_{i=1}^{N} \min(x_i^{F,K}, y_i), &  \sum_{i=1}^{N}y_i > K.
+\sum_{i=1}^n L  \cdot  \max(0, y_i - x_i^{F,K}), &  \sum_{i=1}^{N}y_i \leq K \\
+L \cdot K - L \cdot \sum_{i=1}^{N} \min(x_i^{F,K}, y_i), &  \sum_{i=1}^{N}y_i > K.
 \end{cases}
@@
 
-We have described how the allocation score arises from a standard procedure for developing proper scoring rules when the decision relates to how a fixed quantity of resources should be allocated to multiple locations.
-In fields such as meteorology and supply chain management, a great deal of attention has been given to a different problem where a decision must be made about the quantity of a resource to purchase for a single location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each unit of unmet need.
-This decision making problem leads to the pinball loss that is often used in quantile regression, and the optimal quantity of resources to purchase is the predictive quantile at the probability level $1 - \frac{C}{L}$.
+We begin this section by briefly sketching how the weighted interval score (WIS), a commonly used proper scoring rule for probabilistic forecasts during the COVID-19 pandemic, can also be derived using the approach above.
+Then, we discuss similarities and differences between WIS and the allocation score, and other scores in general.
 
-This result bears a strong resemblance to the results for the allocation problem, where the optimal solution also sets the resource level in each location to a quantile of the respective forecast distributions.
+\subsubsection{The quantile loss and weighted interval score (WIS)}
+\label{sec:methods.wis}
+
+The weighted interval score (WIS), is a scoring rule first defined in 2020 to assist with scoring public forecasts that were being made in the early stages of the COVID-19 pandemic.\citep{bracher2021evaluating}
+Many COVID forecasts were stored in a quantile format, where probabilistic distributions are encoded by a fixed set of quantiles.
+%% optional detail
+This was largely a choice of convenience (e.g., all forecasts had a known, fixed data size) and was codified by the US COVID-19 Forecast Hub and other collaborative forecasting projects.\citep{cramer_united_2022}
+%%
+While pointing a reader interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS, is as its name suggests, a weighted sum of interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...).
+Larger interval scores indicate less skillful forecasts.
+Interval scores themselves consist of (a) the width of the interval, with larger intervals receiving higher scores, and (b) a penalty if the interval does not cover the eventual observation, which increases the further away the interval is from the observed value.
+We note that the most commonly used version of WIS is one that chooses a specific set of weights so that WIS approximates the continuous rank probability score (CRPS) or pinball loss, a commonly used score for probabilistic forecasts.
+Other specifications of weights for the various interval scores can be used.
+\emph{It is important to note that the version of WIS that has been commonly used for evaluating COVID forecasts was proposed because it approximates a commonly used forecast scoring rule, the CRPS, and not because it aligned with any particular public health decision-making rationale.}
+
+The decision-making setup that leads to a choice of WIS as a scoring rule can be formulated using the same recipe as shown in section \ref{sec:methods.detailed.decisiontheory}.
+In fields such as meteorology and supply chain management, a great deal of attention has been given to the problem where a decision must be made about the quantity of a resource to purchase for a single location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each unit of unmet need.\ngr{I might suggest adding a detailed Step 1/2/3 as an appendix and referencing it here.
+But I think that adding that full detail here might be too much.} This decision-making problem leads to the pinball loss that is often used in quantile regression, and thence to the CRPS (or WIS, via approximation) as a scoring rule.\ngr{can we make more explicit the link here to the section above?
+I'm not confident in my ability to describe with technical accuracy the link between quantile/pinball loss and CRPS, but I think this is a good opportunity to do so briefly.}
+
+\subsubsection{Connections between scores}
+\label{sec:methods.generalization}
+
+We have described in previous sections how both the allocation score and the weighted interval score (WIS) arise from a single standard procedure for developing proper scoring rules.
+The allocation score arises when the decision relates to how a fixed quantity of resources should be allocated to multiple locations.
+The WIS and CRPS arise when the decision relates to how much of a resource to order in the context of the cost of the resource and wanting to minimize unmet need in one or more locations.
+In both settings, the optimal solution sets the resource level in each location to a quantile of the respective forecast distributions.
+
 These decision making problems differ according to the challenge faced by the decision maker: a fixed constraint on the available resources for the allocation problem, or a cost per unit of resources in the resource purchasing problem.
 In fact, it is possible to combine these into a more general problem where the decision maker must decide on both the total level of resources to purchase and an allocation of those resources across multiple locations, subject to a cost per unit of resources and a constraint on the total quantity of resources that can be purchased.
-The quantile loss and the allocation score presented in this work are both special cases of the score that emerges from that more general decision making problem.
+The quantile loss and the allocation score presented in this work are both special cases of the score that emerges from that more general decision making problem that integrates across resource cost and constraint parameters together.
 We pursue this direction further in other work (cite arxiv version of paper 2).
-
-Similarly, it is possible to see both the integrated allocation score and common scores such as the  continuous ranked probability score (CRPS) and weighted interval score (WIS) as special cases of a more general score that integrates across resource cost and constraint parameters together.
-It is well known in the literature that WIS is equivalent to a weighted average of quantile scores across multiple quantile levels (cite Bracher et al.) and CRPS integrates quantile scores across multiple quantile levels (cite cite).
-By integrating the combined resource acquisition and allocation score that we described in the previous paragraph across the parameters for both resource costs and constraints, we obtain a score that combines features of both IAS and CRPS and contains those scores as special cases.
-Again, we defer a more detailed discussion of these results to (cite arxiv version of paper 2).
-
-%Finally, we note that a number of generalizations to the loss specification that we introduced for the allocation problem  have been formulated in the literature, including an allowance for costs for over-allocation to a particular unit (e.g.
-if there are storage costs for unused resources), differing losses different units (e.g.
-if a unit of unmet need imposes more severe costs in one location than another), and the introduction of a convex function that controls the rate at which costs accrue depending on the scale of need.
-We consider these and other generalizations in other work.
 
@@
 
+In summary, proper scoring rules for probabilistic forecasts can be traced back to a loss function motivated by a specific decision-making context.
+Often, forecast scores are used to rank models according to accuracy with that particular score, but without reference to the underlying decision-making process for which that score was designed.
+With careful thought and collaboration between modelers and public health officials, we argue that scores that are more aligned with public health decisions could be developed to inform specific problems.
+In some situations, individual or ensemble models could be developed to minimize the loss for a particular setting, although this is also an area largely unexplored in the realm of public health to date.
+
 Notes for the application

### 2023-10-18 alloscore_manuscript/alloscore-application.Rnw
updates to bibliography and citations in paper.

@@
 In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to.
-We give an overview of these procedures in Section \ref{sec:methods.decisiontheory}.\ngr{This section label doesn't exist.
-Should it point to methods.detailed?} There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
+We give an overview of these procedures in Section \ref{sec:methods.decisiontheory}.
+\ngr{This section label doesn't exist. Should it point to methods.detailed?}
+There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
 We review this work only briefly here, and we refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \elr{TODO: identify relevant review or book style summary for meteorology} for discussions focused on applications to economics and meteorology, respectively.
@@
 \subsubsection{Forecast data}
-The US COVID-19 Forecast Hub, a consortium funded by the US CDC and led by a research group at UMass-Amherst, collected short-term forecasts of hospitalizations starting in December 2020.(cite Data descriptor paper)
+The US COVID-19 Forecast Hub, a consortium funded by the US CDC and led by a research group at UMass-Amherst, collected short-term forecasts of hospitalizations starting in December 2020.\citep{cramer_united_2022}
 Any team that with appropriately formatted forecasts could submit them to the Forecast Hub data repository on GitHub.(cite repo site)
@@
 Forecasts could be submitted on any day during the week.
-However once a week on Mondays, the Forecast Hub would collect the most recent forecasts submitted by all teams that met certain inclusion criteria and create an ensemble forecast using quantile averaging.(cite Evan's paper)
+However once a week on Mondays, the Forecast Hub would collect the most recent forecasts submitted by all teams that met certain inclusion criteria and create an ensemble forecast using quantile averaging.\citep{ray_comparing_2023}
 An ensemble that treated all models equally was created (COVIDhub-ensemble) as was a model that created weights of submitted models based on performance in the past 12 weeks (COVIDhub-trained\_ensemble).

### 2023-10-18 R/plot_allocation_forecasts.R
first draft plotting function

@@ -1,0 +1,44 @@
+## plot to create a one-location allocation inset plot
+
+plot_allocation_forecasts <- function(forecast_data, truth_data, loc_abbr, tar_date, intervals = c(.50, .80, .95)) {
+  require(ggplot2)
+  require(patchwork)
+  require(tidyr)
+
+  p_medians <- forecast_data |>
+    filter(abbreviation == loc_abbr, quantile == 0.5) |>
+    ggplot(aes(x=target_end_date, y= value)) +
+    geom_point(alpha=.3) +
+    geom_point(data=filter(truth_data, location=="25", target_end_date <= as.Date("2022-03-15"), target_end_date >= as.Date("2021-12-01")),
+               aes(x=target_end_date), shape=4) +
+    xlab(NULL) +
+    ylab("New hospital admissions") +
+    ggtitle(paste("Observed and forecasted hospital admissions in", loc_abbr))
+
+
+  qtiles <- as.character(c((1-intervals)/2, 1-(1-intervals)/2))
+  fdat_focus <- forecast_data |>
+    mutate(quantile_str = as.character(quantile)) |>
+    filter(abbreviation == loc_abbr,
+           target_end_date == as.Date(tar_date),
+           quantile_str %in% qtiles) |>
+    tidyr::pivot_wider(id_cols= -quantile,
+                       names_from = quantile_str,
+                       values_from = value,
+                       names_prefix="q")
+
+  truth_value <- truth_data |>
+    filter(abbreviation == loc_abbr,
+           target_end_date == tar_date) |>
+    pull(value)
+
+  p_one_date <- ggplot(fdat_focus, aes(x=model)) +
+    geom_linerange(aes(ymin=q0.025, ymax=q0.975), linewidth=2, alpha=.3) +
+    geom_linerange(aes(ymin=q0.1, ymax=q0.9), linewidth=2, alpha=.3) +
+    geom_linerange(aes(ymin=q0.25, ymax=q0.75), linewidth=2, alpha=.3) +
+    geom_hline(yintercept = truth_value, linetype=2) +
+    scale_x_discrete(name="models", labels=NULL) +
+    ggtitle(tar_date)
+
+  p_medians + patchwork::inset_element(p_one_date, left = .6, right = .99, top = .99, bottom = .5)
+}

### 2023-10-18 R/plot-allocation-forecasts.R
adding initial plot to targets.

@@ -40,5 +40,7 @@ plot_allocation_forecasts <- function(forecast_data, truth_data, loc_abbr, tar_d
     scale_x_discrete(name="models", labels=NULL) +
     ggtitle(tar_date)
 
-  p_medians + patchwork::inset_element(p_one_date, left = .6, right = .99, top = .99, bottom = .5)
+  pdf('figures/allocation-forecasts.pdf', height=6, width=8)
+  print(p_medians + patchwork::inset_element(p_one_date, left = .6, right = .99, top = .99, bottom = .5))
+  dev.off()
 }

### 2023-10-18 _targets.R

@@ -26,7 +26,8 @@ tar_source(files = c("R/data-ingestion.R",
                      "R/plot-alloscores.R",
                      "R/run-alloscore.R",
                      "R/determine-model-eligibility.R",
-                     "R/exponential-examples.R"))
+                     "R/exponential-examples.R",
+                     "R/plot-allocation-forecasts.R"))
 
 values <- tibble(forecast_dates = as.character(seq.Date(as.Date("2021-11-22"), as.Date("2022-02-28"), by = "7 days")))
 
@@ -59,6 +60,10 @@ setup <- list(
   tar_target(
     name = exponential_example,
     command = make_exponential_example_figure()
+  ),
+  tar_target(
+    name = forecasts_plot,
+    command = plot_allocation_forecasts(forecast_data, truth_data, loc_abbr="MA", tar_date = "2022-01-24")
   )
 )
 

### 2023-10-20 alloscore_manuscript/alloscore-application.Rnw
updates to application, including adapted figure from ISF talk, citation updates, and more

@@
 
-\ngr{TODO: add a figure showing some forecasts and ground truth data.}
+\begin{figure}
+    \includegraphics[width=\textwidth]{../figures/allocation-forecasts-v2.pdf}
+    \caption{Hospitalization data and example forecasts for Virginia, US.
+New daily hospital admissions are shown by the line of data.
+The dark line indicates data up to the time the forecast was made on December 27, 2021; the grey line indicates data available in retrospect.
+The set of colored bars represent forecasts made for 14 days into the future, with a target date of January 10, 2022 (dashed vertical line).
+The horizontal dashed line indicates the level of the observed value on this day.
+Note that all three forecasts correspond to forecasts for the same day, but they are spread out for better visibility.
+}
+    \label{fig:sample-forecasts}
+\end{figure}
 
@@
 
-This manuscript focuses on two proper forecast scores, the allocation score and the weighted interval score (WIS).(cite Gerding and Bracher)
-Propriety of forecast scoring rules is desirable as it ensures that forecasters are not incentivized to modify their forecast distribution to achieve a better score.(cite Gneiting and Raftery or Raftery and Gneiting)
+This manuscript focuses on two proper forecast scores, the allocation score and the weighted interval score (WIS), both defined above.
 
@@
 
-\ngr{TODO: general introduction to the alloscore}
-
-For the analysis, we fixed a resource constraint $K$ to be 15,000, based roughly on a reported number of ventilators available for reallocation in the US.(cite paper)
+For this analysis, we fixed a resource constraint $K$ to be 15,000, based roughly on a reported number of ventilators available for reallocation in the US.\citep{ajao_assessing_2015}
 For each week, we computed a the allocation score for each 14 day-ahead forecast based on $K=15,000$.
@@
 
-\ngr{TODO: general introduction to WIS and define MWIS as mean WIS.}
-
-We computed standardized ranks for MWIS ($sr_{m,w}^{MWIS}$) using the same procedure as for allocation scores described above.
+The Weighted Interval Score (WIS), as described in earlier sections, measures the alignment of a single probabilistic forecast ($F$) with an observation ($y$).
+We computed the mean WIS across all $L$ locations for each model and each forecasted week as
+\begin{equation}
+MWIS_{m,w} = \frac{1}{L}\sum_{l=1}^L WIS(F_{l,m,w},y)
+\end{equation}
+where $F_{l,m,w}$ is the probabilistic forecast from model $m$ for location $l$ and week $w$.
+Using the same procedure as for allocation scores described above, we computed standardized ranks for MWIS ($sr_{m,w}^{MWIS}$).
 

### 2023-10-20 _targets.R

@@ -27,7 +27,7 @@ tar_source(files = c("R/data-ingestion.R",
                      "R/run-alloscore.R",
                      "R/determine-model-eligibility.R",
                      "R/exponential-examples.R",
-                     "R/plot-allocation-forecasts.R"))
+                     "R/plot_functions.R"))
 
 values <- tibble(forecast_dates = as.character(seq.Date(as.Date("2021-11-22"), as.Date("2022-02-28"), by = "7 days")))
 
@@ -61,9 +61,19 @@ setup <- list(
     name = exponential_example,
     command = make_exponential_example_figure()
   ),
+  # other version of figure with point forecasts for all dates and inset
+  # tar_target(
+  #   name = forecasts_plot,
+  #   command = plot_allocation_forecasts(forecast_data, truth_data, loc_abbr="MA", tar_date = "2022-01-24")
+  # ),
   tar_target(
-    name = forecasts_plot,
-    command = plot_allocation_forecasts(forecast_data, truth_data, loc_abbr="MA", tar_date = "2022-01-24")
+    name = application_figure_1,
+    command = plot_hosp(forecast_data,
+                        truth_data,
+                        models = c("CU-select","COVIDhub-ensemble", "MUNI-ARIMA"),
+                        f_width1 = 3,
+                        f_alpha = .4,
+                        locations = "VA")
   )
 )
 

### 2023-10-25 talks/midas2023/gerding - forecast evaluation via social utility.qmd
slides for MIDAS talk

---
title: |
  | Infectious Disease Forecast Evaluation Via Social Utility: Allocation Scores
subtitle: MIDAS 2023
author: Aaron Gerding, Nick Reich, Ben Rogers, Evan Ray
institute: "School of Public Health and Health Sciences, UMass, Amherst"
format: 
  beamer:
    navigation: horizontal
    fontsize: 9pt
    header-includes: |
      \titlegraphic{
      \begin{minipage}{0.4\linewidth}
          \includegraphics[width=\linewidth]{reichlab.png}
      \end{minipage}%
      \hspace{0.06\linewidth}
      \begin{minipage}{0.4\linewidth}
          \includegraphics[width=\linewidth]{C19FH.png}
      \end{minipage}%
      }
  # revealjs: 
  #   width: 1050
  #   height: 700
# revealjs-plugins:
#   - attribution
editor: visual
# css: styles.css
---

## The Role of ID forecasting Hubs

Outbreak forecasting hubs developed to **"inform public health responses"**

Basic examples of **responses** are **allocation decisions** regarding

-   medical supplies
-   facility capacity
-   personnel
-   R&D funding
-   public tolerance for NPIs

## Basic Hub Products

![](plots/p_VA_CU_IHME.png)

::: notes
quantiles flexible interpretable - closer to describing full distribution which DM probably always wants

multicenter effort to collect forecasts via hubs for ID outbreaks (epi/pan/endemic)

With uncertainty quantification being well-established as needed DM input,

Allocations can be wrt space, time, demographic

Next: formalize this.
:::

## Scoring rules and social welfare

A **scoring rule** $S$ assigns a number $S(F,y)$ to the forecast $F$ of $Y$ when $Y=y$.

$S$ if **proper** if $\displaystyle{E_{F}[S(F,Y)] \leq E_{F}S(G,Y)]}$ for all $F, G$

-   ensures that honesty maximizes expected score

Hub goal of increasing social welfare by ranking and combining forecasts to best support PH responses seems to require use of proper scores.

Current standard is the **Weighted Interval Score** (a discrete version of the **Continuous Ranked Probability Score**)

-   adopted for convenient scoring of quantile forecasts.

<br>

#### A central goal in design of scoring rules for social welfare

::: box
Create scenario where the forecaster believes maximizing her own expected utility aligns with leading the DM toward actions that maximize social utility.
:::

## Bayes Acts

Scoring rules can be cast in decision-theoretic framework of **Bayes acts**.

::: box
The **Bayes act** relative to a forecast $Y \sim F$ is the action $x^F$ by which a DM will minimize **expected loss** $E_F[l(x,Y)]$.
:::

-   loss $l(x,y)$ is the **consequence** of experiencing $y$ having prepared with $x$

::: box
Key point: $S(F,y) := l(x^F, y)$ is an *automatically proper* scoring rule.
:::

<br>

#### Our basic example:

-   $x$ and $y$ are vectors in $\mathbb{N}_0^{52}$
-   $y =$ number of severe cases in US states and territories
-   $l(x,y) =$ unmet need when $x$ beds allocated and $y$ severe cases occur

```{=tex}
\begin{align*}
l(x,y) &= \sum_{i=1}^{52} \max(y_i - x_i, 0)\\
x^F &= \text{minimizer of } E_{F}[l(x,Y)] \text{ over all feasible } x
\end{align*} 
```
<br>

**Our new idea:** Define **feasible** $x$ as satisfying $\displaystyle{\sum x_i \leq K}$

<br>

::: notes
social/political/economic/psychological/moral utility

this must somehow acknowledge risk attitude heterogeneity among various DM's that a forecast hub might serve

gpl parameters, weights, multiple constraints
:::

## Computation

Obtaining $x^F$ is a constrained stochastic optimization problem

-   known in inventory management as a **constrained multi-product newsvendor** problem

-   formally solvable using Lagrange multiplier method to get a quantile representation $x_i^F = F_i^{-1}(\tau(K,F)), i = 1,\ldots,N$

-   in practice, we find $x_i^F$'s via an iterative method:

![](plots/iter_combo.png)

## Application

December 2021: Omicron wave clearly started US but forecast teams unsure of severity given uncertainty about $R_0$, cross-protection by vaccination, previous infection, etc.

![](plots/nat_hosps.png)

## Oracle adjusted allocation scores near Omicron peak

![](plots/peak_alloscore.png)

## Some Observations

::: columns
::: {.column width="40%"}
-   related to the Murphy curves of Ehm, Gneiting, Jordan, Krüger, 2016

-   extreme shortage or surplus diminishes oracle's advantage

-   ranking consistent across large $K$ region.
:::

::: {.column width="60%"}
![](plots/peak_alloscore_min.png)
:::
:::

## Alloscore and WIS rank models differently

![](plots/p_wis_v_alloscore.png)

## Explanations?

![](plots/mvbucky_dist_alloc.png)

## Limitations

This is **post-hoc** analysis

Hub forecasters were unaware of

-   an allocation score (on joint forecast)

-   any allocation based loss

-   our quantile interpolation/extrapolation methods (`distfromq`)

    -   might be especially important for tails

We hope/think that allocation scoring is sensitive to implicit dependence structures in forecasts, but all work so far only refers directly to marginals - nothing yet with copulas, etc.

<br>

### Thank you!

A very rough R package I wrote to implement scoring procedures:

<https://github.com/aaronger/alloscore>

A less rough package Evan wrote to implement cdf reconstruction

<https://github.com/reichlab/distfromq>

### 2023-10-27 alloscore_manuscript/alloscore-application.tex
adding alloscore objects for now.

--- 2023-10-17 alloscore_manuscript/alloscore-application.Rnw
@@
-\documentclass{article}
+\documentclass{article}\usepackage[]{graphicx}\usepackage[]{xcolor}
+% maxwidth is the original width if it is less than linewidth
+% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 \author{Aaron Gerding, Nicholas G. Reich, Benjamin Rogers, Evan L. Ray}
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 \maketitle
-
-<<setup, tidy=FALSE, echo=FALSE, message=FALSE>>=
-knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
-@
-
-<<environment-setup>>=
-library(targets)
-library(tidyverse)
-library(ggbump)
-
-theme_set(theme_bw())
-
-## necessary to have project root as wd, to override document directory
-knitr::opts_knit$set(root.dir = '../')
-@
 
@@
 In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to.
-We give an overview of these procedures in Section \ref{sec:methods.decisiontheory}.\ngr{This section label doesn't exist.
-Should it point to methods.detailed?} There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
+We give an overview of these procedures in Section \ref{sec:methods.decisiontheory}.
+There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
 We review this work only briefly here, and we refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \elr{TODO: identify relevant review or book style summary for meteorology} for discussions focused on applications to economics and meteorology, respectively.
@@
 %For a given probability level $\tau$, if the quantiles $Q_i(\tau)$ of the marginal forecasts are proportional to the observed needs $y_i$, then the allocation score is zero for the resource constraint level $K = \sum Q_i(\tau)$.
-This stands in marked contrast to other common scoring methods for multivariate forecasts that aggregate univariate scores such as log score, CRPS, or WIS for the marginal forecasts where a
-poor forecast made for one unit (a location, say) is penalized regardless of alignments in other units.
+This stands in marked contrast to other common scoring methods for multivariate forecasts that aggregate univariate scores such as CRPS or WIS for the marginal forecasts where a
+misalignment in one coordinate is penalized regardless of alignments in other coordinates.
 Note that we do not claim that the allocation score is generically preferable to these other scores \textemdash rather, it provides a view of forecast performance that is specifically tuned to the context of decision making about resource allocations.
@@
 the support of each $F_i$ is a subset of $\mathbb{R}^+$.
-Finally, we assume that each unit of unmet need incurs a loss denoted by $L$, so that if the selected resource level $x_i$ in location $i$ is less than the realized need $y_i$, a loss of $L \cdot (y_i - x_i)$ results.
+Finally, we assume that each unit of unmet need incurs a loss denoted by $L$, so that if the selected resource level $x_i$ in location $i$ is less than the realized need $y_i$, a loss of $L(y_i - x_i)$ results.
 
@@
 \end{align}
-where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]$ sums the expected loss due to unmet need across all locations.
+where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]$ sums the expected loss due to unmet need across all locations.
 
-%%%% start of large excision %%%%%
-
-% In the supplement we derive a general form of this solution by beginning with the fact that a condition for minimizing
-% expected loss under the constraint is that there is no way to decrease expected loss further by shifting a small amount of the resource from one location to another.
-% As a consequence, at $x^{F,K}$, adding resources to any particular location $i$ reduces expected loss at the same rate $\lambda$:
-% \begin{align}
-% \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^{F,K}} = -\lambda.
-% \end{align}
-% It turns out that these derivatives can also be expressed in terms of the marginal forecasts $F_i$ as
-% \begin{align}
-%  \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] = L \cdot (F_i(x_i) - 1)
-% \end{align}
-% so that
-% \begin{align}
-% F(x_i^{F,K}) = 1-\lambda/L.
-% \end{align}
-% This tells us that in order to allocate optimally (according to $F$), we must divide resources among the locations so that there is an equal forecasted probability in every location that the allocation is sufficient to meet resource need.
-% It also allows us to write the components $x_i^{F,K}$ of the Bayes act as quantiles
-% \begin{align}
-% x_i^{F,K} = F_i^{-1}(1 - \lambda/L)
-% \end{align}
-% which can be substituted into the constraint equation of the allocation problem to get an equation for $\lambda$,
-% \begin{align}
-% \sum_{i=1}^N F_i^{-1}(1 - \lambda/L) = K.
-% \end{align}
-% The value of $\lambda$ that solves this equation---which almost always must be approximated with a numerical method---can be substituted back into the quantile formulas
-% to arrive at an explicit expression for the Bayes act.
-% This partial solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
-% We note that $\lambda$ and the probability level $1-\lambda/L$ depend on both the constraint level $K$ defining the decision problem and on the forecast $F$.
+In the supplement we derive a general form of this solution by beginning with the fact that a condition for minimizing
+expected loss under the constraint is that there is no way to decrease expected loss further by shifting a small amount of the resource from one location to another.
+As a consequence, at $x^{F,K}$, adding resources to any particular location $i$ reduces expected loss at the same rate $\lambda$:
+\begin{align}
+\frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^{F,K}} = -\lambda.
+\end{align}
+It turns out that these derivatives can also be expressed in terms of the marginal forecasts $F_i$ as
+\begin{align}
+ \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] = L \cdot (F_i(x_i) - 1)
+\end{align}
+so that 
+\begin{align}
+F(x_i^{F,K}) = 1-\lambda/L.
+\end{align}
+This tells us that in order to allocate optimally (according to $F$), we must divide resources among the locations so that there is an equal forecasted probability in every location that the allocation is sufficient to meet resource need.
+It also allows us to write the components $x_i^{F,K}$ of the Bayes act as quantiles
+\begin{align}
+x_i^{F,K} = F_i^{-1}(1 - \lambda/L)
+\end{align}
+which can be substituted into the constraint equation of the allocation problem to get an equation for $\lambda$,
+\begin{align}
+\sum_{i=1}^N F_i^{-1}(1 - \lambda/L) = K.
+\end{align}
+The value of $\lambda$ that solves this equation---which almost always must be approximated with a numerical method---can be substituted back into the quantile formulas
+to arrive at an explicit expression for the Bayes act.
+This partial solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
+We note that $\lambda$ and the probability level $1-\lambda/L$ depend on both the constraint level $K$ defining the decision problem and on the forecast $F$.
 This is quite different from the situation for quantile scores and the WIS, where the probability levels used to calculate scores are fixed.
-
-%%%% end of large excision %%%%%
 
@@
 \begin{align}
-S_A^{\text{raw}}(F, y; K) = s_A(x^{F,K}, y) = \sum_{i=1}^N L \cdot \max(0, y_i - F_i^{-1}(1- \lambda(K,F)/L))
+S_A^{\text{raw}}(F, y; K) = s_A(x^{F,K}, y) = \sum_{i=1}^N L \max(0, y_i - F_i^{-1}(1- \lambda(K,F)/L))
 \end{align}
@@
 But to make this a more easily interpreted measure of
-forecast performance, we will adjust it by the the minimum loss $l_{\text{oracle}}(y;K)$ achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$ which is $l_{\text{oracle}}(y;K) = L \cdot \max(0, \sum_{i=1}^{N}y_i - K)$.
+forecast performance, we will adjust it by the the minimum loss $l_{\text{oracle}}(y;K)$ achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$ which is $l_{\text{oracle}}(y;K) = L \max(0, \sum_{i=1}^{N}y_i - K)$.
 Our scoring rule is then given by
@@
 S_A(F, y; K) &= S_A^{\text{raw}}(F, y; K) - l_{\text{oracle}}(y;K) \\
-& = \sum_{i=1}^N L  \cdot \max(0, y_i - x_i^{F,K}) - L \cdot  \max(0, \sum_{i=1}^{N}y_i - K) \\
+& = \sum_{i=1}^N L \max(0, y_i - x_i^{F,K}) - L \max(0, \sum_{i=1}^{N}y_i - K) \\
 & = 
 \begin{cases}
-\sum_{i=1}^n L  \cdot  \max(0, y_i - x_i^{F,K}), &  \sum_{i=1}^{N}y_i \leq K \\
-L \cdot K - L \cdot \sum_{i=1}^{N} \min(x_i^{F,K}, y_i), &  \sum_{i=1}^{N}y_i > K.
+\sum_{i=1}^n L \max(0, y_i - x_i^{F,K}), &  \sum_{i=1}^{N}y_i \leq K \\
+LK - L\sum_{i=1}^{N} \min(x_i^{F,K}, y_i), &  \sum_{i=1}^{N}y_i > K.
 \end{cases}
@@
 
-We begin this section by briefly sketching how the weighted interval score (WIS), a commonly used proper scoring rule for probabilistic forecasts during the COVID-19 pandemic, can also be derived using the approach above.
-Then, we discuss similarities and differences between WIS and the allocation score, and other scores in general.
+We have described how the allocation score arises from a standard procedure for developing proper scoring rules when the decision relates to how a fixed quantity of resources should be allocated to multiple locations.
+In fields such as meteorology and supply chain management, a great deal of attention has been given to a different problem where a decision must be made about the quantity of a resource to purchase for a single location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each unit of unmet need.
+This decision making problem leads to the pinball loss that is often used in quantile regression, and the optimal quantity of resources to purchase is the predictive quantile at the probability level $1 - \frac{C}{L}$.
 
-\subsubsection{The quantile loss and weighted interval score (WIS)}
-\label{sec:methods.wis}
-
-The weighted interval score (WIS), is a scoring rule first defined in 2020 to assist with scoring public forecasts that were being made in the early stages of the COVID-19 pandemic.\citep{bracher2021evaluating}
-Many COVID forecasts were stored in a quantile format, where probabilistic distributions are encoded by a fixed set of quantiles.
-%% optional detail
-This was largely a choice of convenience (e.g., all forecasts had a known, fixed data size) and was codified by the US COVID-19 Forecast Hub and other collaborative forecasting projects.\citep{cramer_united_2022}
-%%
-While pointing a reader interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS, is as its name suggests, a weighted sum of interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...).
-Larger interval scores indicate less skillful forecasts.
-Interval scores themselves consist of (a) the width of the interval, with larger intervals receiving higher scores, and (b) a penalty if the interval does not cover the eventual observation, which increases the further away the interval is from the observed value.
-We note that the most commonly used version of WIS is one that chooses a specific set of weights so that WIS approximates the continuous rank probability score (CRPS) or pinball loss, a commonly used score for probabilistic forecasts.
-Other specifications of weights for the various interval scores can be used.
-\emph{It is important to note that the version of WIS that has been commonly used for evaluating COVID forecasts was proposed because it approximates a commonly used forecast scoring rule, the CRPS, and not because it aligned with any particular public health decision-making rationale.}
-
-The decision-making setup that leads to a choice of WIS as a scoring rule can be formulated using the same recipe as shown in section \ref{sec:methods.detailed.decisiontheory}.
-In fields such as meteorology and supply chain management, a great deal of attention has been given to the problem where a decision must be made about the quantity of a resource to purchase for a single location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each unit of unmet need.\ngr{I might suggest adding a detailed Step 1/2/3 as an appendix and referencing it here.
-But I think that adding that full detail here might be too much.} This decision-making problem leads to the pinball loss that is often used in quantile regression, and thence to the CRPS (or WIS, via approximation) as a scoring rule.\ngr{can we make more explicit the link here to the section above?
-I'm not confident in my ability to describe with technical accuracy the link between quantile/pinball loss and CRPS, but I think this is a good opportunity to do so briefly.}
-
-\subsubsection{Connections between scores}
-\label{sec:methods.generalization}
-
-We have described in previous sections how both the allocation score and the weighted interval score (WIS) arise from a single standard procedure for developing proper scoring rules.
-The allocation score arises when the decision relates to how a fixed quantity of resources should be allocated to multiple locations.
-The WIS and CRPS arise when the decision relates to how much of a resource to order in the context of the cost of the resource and wanting to minimize unmet need in one or more locations.
-In both settings, the optimal solution sets the resource level in each location to a quantile of the respective forecast distributions.
-
+This result bears a strong resemblance to the results for the allocation problem, where the optimal solution also sets the resource level in each location to a quantile of the respective forecast distributions.
 These decision making problems differ according to the challenge faced by the decision maker: a fixed constraint on the available resources for the allocation problem, or a cost per unit of resources in the resource purchasing problem.
 In fact, it is possible to combine these into a more general problem where the decision maker must decide on both the total level of resources to purchase and an allocation of those resources across multiple locations, subject to a cost per unit of resources and a constraint on the total quantity of resources that can be purchased.
-The quantile loss and the allocation score presented in this work are both special cases of the score that emerges from that more general decision making problem that integrates across resource cost and constraint parameters together.
+The quantile loss and the allocation score presented in this work are both special cases of the score that emerges from that more general decision making problem.
 We pursue this direction further in other work (cite arxiv version of paper 2).
+
+Similarly, it is possible to see both the integrated allocation score and common scores such as the  continuous ranked probability score (CRPS) and weighted interval score (WIS) as special cases of a more general score that integrates across resource cost and constraint parameters together.
+It is well known in the literature that WIS is equivalent to a weighted average of quantile scores across multiple quantile levels (cite Bracher et al.) and CRPS integrates quantile scores across multiple quantile levels (cite cite).
+By integrating the combined resource acquisition and allocation score that we described in the previous paragraph across the parameters for both resource costs and constraints, we obtain a score that combines features of both IAS and CRPS and contains those scores as special cases.
+Again, we defer a more detailed discussion of these results to (cite arxiv version of paper 2).
+
+%Finally, we note that a number of generalizations to the loss specification that we introduced for the allocation problem  have been formulated in the literature, including an allowance for costs for over-allocation to a particular unit (e.g.
+if there are storage costs for unused resources), differing losses different units (e.g.
+if a unit of unmet need imposes more severe costs in one location than another), and the introduction of a convex function that controls the rate at which costs accrue depending on the scale of need.
+We consider these and other generalizations in other work.
 
@@
 
-<<score-computation>>=
-tar_load("all_alloscore_data")
+\subsubsection{Metrics as a function of time}
 
-alloscore_data_summaries <- all_alloscore_data |>
-  filter(K == 15000)|>
-  group_by(model, reference_date) |>
-  summarize(alloscore = sum(components), yobs = sum(y)) |>
-  group_by(reference_date) |>
-  mutate(nmodels = n(),
-         allo_rank = rank(alloscore, ties.method = "min"),
-         allo_rank_std = 1- (allo_rank-1)/(nmodels-1)) |>
-  ungroup() |>
-  complete(model, reference_date) |>
-  mutate(target_end_date = as.Date(reference_date) + 14)
-
-tar_load("score_data")
-
-alloscore_models <- unique(alloscore_data_summaries$model)
-score_data_summaries <- score_data |>
-  mutate(reference_date = target_end_date - 14) |>
-  filter(model %in% alloscore_models) |>
-  group_by(model, reference_date) |>
-  summarize(mwis = mean(wis)) |>
-  ungroup() |>
-  group_by(reference_date) |>
-  mutate(nmodels = n(),
-         mwis_rank = rank(mwis, ties.method = "min"),
-         mwis_rank_std = 1- (mwis_rank-1)/(nmodels-1)) |>
-  ungroup() |>
-  complete(model, reference_date) |>
-  mutate(target_end_date = as.Date(reference_date) + 14)
-@
-
-\subsubsection{Metrics as a function of time}
-<<grob-hosps-over-time>>=
-tar_load("truth_data")
-dat_to_plot <- truth_data |>
-  filter(location == "US",
-         target_end_date %in% score_data_summaries$target_end_date)
-p1 <- ggplot(dat_to_plot,
-             aes(x=target_end_date, y=value)) +
-  geom_line() +
-  geom_point() +
-  geom_hline(yintercept=15000, linetype=2) +
-  ylab("hospital admissions") +
-  xlab(NULL) +
-  ylim(c(0, NA))
-@
-
-<<metrics-over-time, fig.height=9, fig.width=7, fig.cap="Hospital admissions and evaluation metrics over time.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\includegraphics[width=\maxwidth]{figure/metrics-over-time-1} \caption[Hospital admissions and evaluation metrics over time]{Hospital admissions and evaluation metrics over time.
 Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022.
@@
 Similarly to panel B, the x-axis corresponds to the date for which the prediction was made.
-MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the pean of the Omicron wave.">>=
-p2 <- ggplot(alloscore_data_summaries,
-             aes(x=target_end_date, y=alloscore, color=model, shape=model)) +
-  scale_x_date(limits=range(score_data_summaries$target_end_date)) +
-  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
-  ylab("Allocation score") +
-  geom_point() + geom_line()+
-  theme(legend.pos="none") +
-  xlab(NULL)
+MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the pean of the Omicron wave.}\label{fig:metrics-over-time}
+\end{figure}
 
-p4 <- ggplot(score_data_summaries,
-             aes(x=target_end_date, y=mwis, color=model,shape=model)) +
-  geom_point() + geom_line() +
-  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
-  ylab("MWIS")+
-  theme(legend.pos="bottom")+
-  xlab(NULL)
-
-cowplot::plot_grid(p1, p2, p4,
-                   nrow=3,
-                   align="v",
-                   rel_heights = c(1, 1, 1.5),
-                   labels = "AUTO")
-@
+\end{knitrout}
 
@@
 Predictions made during the peak week and just after showed the highest variation in allocation scores, with some models having values under 1000 and others having values over 3500.
-<<eval=FALSE>>=
-## some analysis to support above statements
-alloscore_data_summaries |>
-  group_by(target_end_date) |>
-  summarize(meanAS = mean(alloscore, na.rm=TRUE),
-            medianAS = median(alloscore, na.rm=TRUE),
-            minAS = min(alloscore, na.rm=TRUE),
-            maxAS = max(alloscore, na.rm=TRUE),
-            range = maxAS-minAS)
-@
 
@@
 
-<<eval=FALSE>>=
-## some analysis to support above statements
-score_data_summaries |>
-  group_by(target_end_date) |>
-  summarize(meanWIS = mean(mwis, na.rm=TRUE),
-            medianWIS = median(mwis, na.rm=TRUE),
-            minWIS = min(mwis, na.rm=TRUE),
-            maxWIS = max(mwis, na.rm=TRUE),
-            range = maxWIS-minWIS)
-@
-
 \subsubsection{correlation between WIS and allocation score}
 
-<<metrics-correlation, fig.width=7, fig.height=6, fig.cap="Association of standardized ranks for MWIS and allocation score by model and week.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\includegraphics[width=\maxwidth]{figure/metrics-correlation-1} \caption[Association of standardized ranks for MWIS and allocation score by model and week]{Association of standardized ranks for MWIS and allocation score by model and week.
 Each facet of the plot corresponds to one model.
@@
 The size of the point corresponds to the observed value on the date for which the prediction was made.
-Models show different degrees of association between the two metrics.">>=
-all_data <- left_join(alloscore_data_summaries, score_data_summaries,
-                      by=c("model", "target_end_date"))
+Models show different degrees of association between the two metrics.}\label{fig:metrics-correlation}
+\end{figure}
 
-ggplot(all_data,
-       aes(x=mwis_rank_std, y=allo_rank_std, color=target_end_date, size=yobs)) +
-  geom_point(alpha=.7) +
-  scale_x_continuous("MWIS standardized rank", limits=c(0,1)) +
-  scale_y_continuous("Allocation score standardized rank", limits=c(0,1)) +
-  facet_wrap(.~model) +
-  geom_abline(slope=1, intercept=0, linetype=2, color="gray")
-
-@
+\end{knitrout}
 
@@
 
-<<eval=FALSE>>=
-alloscore_data_summaries |>
-  group_by(model) |>
-  summarize(top_rank = sum(allo_rank_std==1))
-@
-
  - bubble figure with ranks sized by total observations
@@
 \label{sec:discussion}
-
-In summary, proper scoring rules for probabilistic forecasts can be traced back to a loss function motivated by a specific decision-making context.
-Often, forecast scores are used to rank models according to accuracy with that particular score, but without reference to the underlying decision-making process for which that score was designed.
-With careful thought and collaboration between modelers and public health officials, we argue that scores that are more aligned with public health decisions could be developed to inform specific problems.
-In some situations, individual or ensemble models could be developed to minimize the loss for a particular setting, although this is also an area largely unexplored in the realm of public health to date.
 

### 2023-10-27 random_undocumented_experiments/debugging5_17.R

@@ -69,4 +69,6 @@ ggplot() + map(2:5, function(i) {
   map(c(1,6:51), function(i) geom_function(fun = ascores$F[[i]], alpha = .5)) + lims(x=c(0,500)) +
   labs(color = "problem states", y = "distfromq F")
 
-
+for (i in 6:17) {
+as <- run_alloscore(forecast_data, truth_data, "2022-01-31", eligible_models[i])
+}

### 2023-10-30 alloscore_manuscript/alloscore-application.Rnw
adding onek with k=15000 target and integrating into manuscript

@@
 <<score-computation>>=
-tar_load("all_alloscore_data")
+tar_load("Kat15k_alloscores")
 
-alloscore_data_summaries <- all_alloscore_data |>
-  filter(K == 15000)|>
+alloscore_data_summaries <- Kat15k_alloscores |>
   group_by(model, reference_date) |>
@@
 
-Allocation scores varied substantially by date and by model (Figure XX).
+Allocation scores varied substantially by date and by model (Figure \ref{fig:metrics-over-time}).
 For predictions made for the first three Mondays in December 2021 and the last three Mondays in February 2022 all models had allocation scores under 500 (and the mean across all models was less than 100), indicating that the unnecessary unmet need was fairly low relative to the total number of hospital admissions on those days.
@@
 
-Models show differing levels of correlation between their allocation scores and MWIS values (Figure XX).
+Models show differing levels of correlation between their allocation scores and MWIS values (Figure \ref{fig:metrics-correlation}).
 Here are some examples of the different patterns observed:

### 2023-10-30 _targets.R
adding back in Ben's "over k" work from @ 7a478fd

@@ -26,8 +26,7 @@ tar_source(files = c("R/data-ingestion.R",
                      "R/plot-alloscores.R",
                      "R/run-alloscore.R",
                      "R/determine-model-eligibility.R",
-                     "R/exponential-examples.R",
-                     "R/plot_functions.R"))
+                     "R/exponential-examples.R"))
 
 values <- tibble(forecast_dates = as.character(seq.Date(as.Date("2021-11-22"), as.Date("2022-02-28"), by = "7 days")))
 
@@ -61,21 +60,20 @@ setup <- list(
     name = exponential_example,
     command = make_exponential_example_figure()
   ),
-  # other version of figure with point forecasts for all dates and inset
-  # tar_target(
-  #   name = forecasts_plot,
-  #   command = plot_allocation_forecasts(forecast_data, truth_data, loc_abbr="MA", tar_date = "2022-01-24")
-  # ),
   tar_target(
-    name = application_figure_1,
-    command = plot_hosp(forecast_data,
-                        truth_data,
-                        models = c("CU-select","COVIDhub-ensemble", "MUNI-ARIMA"),
-                        f_width1 = 3,
-                        f_alpha = .4,
-                        locations = "VA")
+    name = Kat15k_alloscores,
+    command = run_and_assemble_alloscores(forecast_data,
+                                          truth_data,
+                                          reference_dates = values$forecast_dates,
+                                          one_K = 15000)
   )
 )
+
+## from Ben's "overk" analysis
+# tar_map(
+#   values=values,
+#   tar_target(alloscore_overk, run_alloscore_overk(forecast_data, truth_data, forecast_dates))
+# )
 
 mapped <- tar_map(
   unlist = FALSE,

### 2023-11-03 talks/midas2023/gerding - forecast evaluation via social utility.qmd
separate plotting code into talks/plots/

@@
 author: Aaron Gerding, Nick Reich, Ben Rogers, Evan Ray
-institute: "School of Public Health and Health Sciences, UMass, Amherst"
+institute: |
+ | School of Public Health and Health Sciences, UMass, Amherst
+ | Department of Biostatistics and Epidemiology
 format: 
   beamer:
+    keep-tex: true
     navigation: horizontal
@@
 
-## The Role of ID forecasting Hubs
+## Outbreak Forecast Hubs and **Informed** Resource Allocation
 
-Outbreak forecasting hubs developed to **"inform public health responses"**
+Hubs developed to **"inform public health responses"**
 
-Basic examples of **responses** are **allocation decisions** regarding
+-   such as how resources are **allocated** among locations
 
--   medical supplies
--   facility capacity
--   personnel
--   R&D funding
--   public tolerance for NPIs
+-   e.g., medical supplies, facility capacity, personnel
 
-## Basic Hub Products
-
-![](plots/p_VA_CU_IHME.png)
+![](../plots/output/p_VA_JHU_MUNI.jpeg)
 
@@
 
-A **scoring rule** $S$ assigns a number $S(F,y)$ to the forecast $F$ of $Y$ when $Y=y$.
+Hubs strive to rank and combine forecasts so as to optimize social welfare via the public health decisions that forecasts inform
 
-$S$ if **proper** if $\displaystyle{E_{F}[S(F,Y)] \leq E_{F}S(G,Y)]}$ for all $F, G$
+-   uncertainty quantification, and therefore, probabilistic forecasting essential
 
--   ensures that honesty maximizes expected score
+-   basic strategy: use a **scoring rule** $S$ which assigns a loss $S(F,y)$ when a probabilistic forecast $F$ of $Y$ is chosen and $Y=y$ is observed.
 
-Hub goal of increasing social welfare by ranking and combining forecasts to best support PH responses seems to require use of proper scores.
+-   optimizing welfare requires that forecasters say what they believe; so $S$ should be **proper** meaning $\displaystyle{E_{F}[S(F,Y)] \leq E_{F}S(G,Y)]}$ for all $F, G$
 
-Current standard is the **Weighted Interval Score** (a discrete version of the **Continuous Ranked Probability Score**)
+Current standard is the **Weighted Interval Score** (discrete CRPS)
 
--   adopted for convenient scoring of quantile forecasts.
+-   adopted largely for convenient scoring of quantile forecasts.
 
-<br>
+## Tools from decision theory
 
-#### A central goal in design of scoring rules for social welfare
+### A central goal in design of scoring rules
 
-::: box
-Create scenario where the forecaster believes maximizing her own expected utility aligns with leading the DM toward actions that maximize social utility.
-:::
+Tie forecast scores directly to the benefits to society of the decisions they inform
 
-## Bayes Acts
+Key tools from decision theory for linking success/failure of forecast-informed policy actions to scoring rules:
 
-Scoring rules can be cast in decision-theoretic framework of **Bayes acts**.
+-   Let $l(x,y)$ be the **loss** of experiencing $y$ after taking policy action $x$
 
-::: box
-The **Bayes act** relative to a forecast $Y \sim F$ is the action $x^F$ by which a DM will minimize **expected loss** $E_F[l(x,Y)]$.
-:::
+-   The **Bayes risk** of a forecast $Y \sim F$ is $\min_x E_F[l(x,Y)]$
 
--   loss $l(x,y)$ is the **consequence** of experiencing $y$ having prepared with $x$
+-   A **Bayes act** for $F$ is an action $x^F$ that attains this minimum
 
-::: box
-Key point: $S(F,y) := l(x^F, y)$ is an *automatically proper* scoring rule.
-:::
+-   Losses from Bayes acts define an *automatically proper* scoring rule
 
-<br>
+$$
+S(F,y) := l(x^F, y)
+$$
 
-#### Our basic example:
+**Note:** Scoring rules are really only a means to an end.
+It is the expected score, in this case the Bayes risk, which characterize the value a forecast adds for a decision maker.
+Scoring rule sample averages estimate the expected score.
+
+## A new scoring rule via a new loss function for constrained actions
+
+### Our basic example:
 
 -   $x$ and $y$ are vectors in $\mathbb{N}_0^{52}$
+
 -   $y =$ number of severe cases in US states and territories
+
 -   $l(x,y) =$ unmet need when $x$ beds allocated and $y$ severe cases occur
@@
 
-![](plots/iter_combo.png)
+![](../plots/output/iter_combo.png)
 
@@
 
-![](plots/nat_hosps.png)
+![](../plots/output/nat_hosps.png)
 
@@
 
-![](plots/peak_alloscore.png)
+![](../plots/output/peak_alloscore.png)
 
@@
 ::: {.column width="60%"}
-![](plots/peak_alloscore_min.png)
+![](../plots/output/peak_alloscore_min.png)
 :::
@@
 
-![](plots/p_wis_v_alloscore.png)
+![](../plots/output/p_wis_v_alloscore.png)
 
@@
 
-![](plots/mvbucky_dist_alloc.png)
+![](../plots/output/mvbucky_dist_alloc.png)
 

### 2023-11-03 talks/midas2023/labnotes.qmd
separate plotting code into talks/plots/

---
title: Lab meeting notes
author: Aaron Gerding
format: 
  revealjs:
    width: 1050
    height: 700
revealjs-plugins:
  - attribution
editor: visual
css: styles.css
---

## Losses

Losses are defined on

-   actions x (wait = 0, intervene = 1) and

-   outcomes y (T \> incidence = 0, T \< incidence = 1)

Examples:

$l(x,y) = Cx + L(y-x)_+$

regret form

$C[[x>y]] + (L-C)[[y>x]]$

## Decision rules

Decision rules map current supporting info $z = z_{t_0}$ to future actions $x = x_{t_1}$ where $t_1 = t_0 + h > t_0$

Prior to $t_0$, we can view $Z=Z_{t_0}$ and $Y=Y_{t_1}$ as RV's with joint distribution $P_{ZY}$ - a decision rule $d(w)$ has an expected loss $E_d = E_{ZY}[l(d(Z),Y)]$

Alternatively we can change variables to $X=d(W)$ and define expected loss as $E_d = E_{XY}[l(X,Y)]$

Basic question for a decision maker: which $d$ has minimum $E_d$?

To answer, DM might use an estimate $\widehat{E}_d$ from time series samples from $(Z_t,Y_{t+h})$ or $(X_t,Y_{t+h})$ as stochastic processes

But! What we really care about is $E_d = E_{X_{t_0}, Y_{t_1}}[l(X_{t_0},Y_{t_1})]$ - and $\widehat{E}_d$ might not estimate this with strong stationarity or mixing assumptions

## Decision support info

Types of current info:

Forecast q of class probability - from e.g. logistic regression, or any other model a decision rule can be defined with - a threshold - or as Bayes act for a loss function

Bayes act is the optimal (minimal Eloss) action if forecast is correct

Bayes act for both $l(x,y) = Cx + L(y-x)_+$ and $C[[x>y]] + (L-C)[[y>x]]$ is $[[q > C/L]]$

$E_q[l(x,y)] = q(L-C)[[x=0]] + (1-q)C[[x=1]]$

$x^q = [[(1-q)C < q(L-C)]] = [[\frac{C/L}{1-C/L} < \frac{q}{1-q}]] = [[C/L < q]]$

## Other types of info

Forecast f of right action (deterministic)

-   a decision rule can be defined with

    -   blind faith

    -   sampling to (empirically) impute class probabilities:

        -   wait -\> $P(y=0 \mid f = \text{wait})$,

        -   intervene -\> $P(y=1 \mid f = \text{intervene})$ ,ie., calibration functions

-   also can be done for prob forecasts $rc(q) = P(y=1 \| f = q)$

Forecast of incidence quantile for prob level $1-C/L$

-   decision rule: intervene if quantile \> T

-   forecast is saying $P(y=1) = P(incidence > T) = 1 - P(incidence \leq T) > 1 - 1 + C/L = C/L$

-   assuming reliability

## Scoring rules

Choosing a decision rule for prob forecasts defines a scoring rule $S(q,y) = l(d(q), y)$

-   allows us to define the expected loss of a pforecast as $E(S(Q,Y))$

Bayes scoring rule for $l(x,y) =C[[x>y]] + (L-C)[[y>x]]$ is cost-weighted SR

$$
S_{C/L}(q,y) = y (L-C)[[1-q > 1-C/L]] + (1-y)C [[q \geq C/L]]
$$

Note: We cannot get scoring rules like

-   Brier: $S(q,y) = y(1-q)^2+(1-y) q^2$) or

-   log: $S(q,y) = -y \log (q)-(1-y) \log (1-q)$

this way, but we can write them as mixtures of cost-weighted losses

## Value scores

The 'value score' for a deterministic forecast is

$\frac{E[l(d(m_Y),Y)] - E[l(f,Y)]}{E[l(d(m_Y),Y)] - E[l(Y,Y)]} = \frac{E^l_{clm} -E^l_f}{E^l_{clm} - E^l_{oracle}} = 1-\frac{E^r_f}{E^r_{clm}}$

${clm}$ is the climatology forecast which always

-   the skill score for the loss-based scoring rule applied to q or the imputed probs

These expectations can be computed with respect to the margins ($p_f, p_Y$) and conditionals of $P_{f,Y}$

-   via calibration-refinement (ppv and npv as conditionals); resolution is how much pv depends on f

-   or likelihood-baserate (specificity/fnr, sensitivity/fpr); discrimination is how much likelihood depends on y

```{=tex}
\begin{align}
E_{cl} = 
\end{align}
```
## Deriving the weighted misclassification measure

$$
\frac{E_f}{E_{cl}} = \frac{(L-C)fnp + C fpp}{\max ((L-C)p_Y, C(1-p_Y))} = \frac{(L-C)fnr p_Y + C fpr (1-p_Y)}{\max ((L-C)p_Y, C(1-p_Y))}
$$

$$\frac{E_f}{E_{cl}} = fnr + \frac{C/L}{1-C/L}\frac{1-p_Y}{p_Y}fpr \text{ or } \frac{1-C/L}{C/L}\frac{p_Y}{1-p_Y}fnr + fpr$$

### 2023-11-03 talks/ISF_talk/plot_scores.R

@@ -3,6 +3,7 @@ library(alloscore)
 library(tidyverse)
 library(geofacet)
 library(latex2exp)
+library(here)
 
 source("_targets.R")
 save_dir <- file.path("talks/ISF_talk/plots/")
@@ -31,10 +32,11 @@ three_days <- bind_rows(slim_dfs) %>% filter(reference_date %in% forecast_dates[
 plot_scores_slim(peak_day)
 plot_scores_slim(three_days) + lims(x=c(500, 30000))
 
-p_peak <- plot_scores_slim(peak_day)
+p_peak <- plot_scores_slim(peak_day, palette = score_palette, linetypes = score_linetypes) + 
+  theme_bw() +
+  scale_x_continuous(expand = c(0,0), limits = c(5000,40000))
 
-p_peak_out <- p_peak + theme_bw() +
-  scale_x_continuous(expand = c(0,0), limits = c(5000,40000)) +
+p_peak_out <- p_peak + 
   annotate(geom = "text", x = 22000, y = -250, size = 3,
            label = paste("Total hospitalizations on 2022-01-10 →"), hjust = 1) +
   labs(
@@ -43,8 +45,14 @@ p_peak_out <- p_peak + theme_bw() +
   theme(axis.title.x = element_text(size = 15),
         axis.title.y = element_text(size = 15))
 
+p_peak_out_min <- p_peak + theme_bw() + 
+  theme(legend.position = "none")
+
 ggsave(plot = p_peak_out, filename = "peak_alloscore.png",
-       width = 8, height = 6, path = save_dir)
+       width = 8, height = 6, path = save_dir, dpi = 600)
+
+ggsave(plot = p_peak_out_min, filename = "peak_alloscore_min.png",
+       width = 6, height = 6, path = save_dir)
 
 # plot score time series
 

### 2023-11-03 talks/plots/code/plot_allocs.R

@@ -1,0 +1,68 @@
+library(covidHubUtils)
+library(alloscore)
+library(tidyverse)
+library(geofacet)
+
+source("_targets.R")
+
+file_names <- paste0("data/slim_", mkeep)
+# create an empty list to store the loaded objects
+if (!exists("slim_dfs")) {
+  slim_dfs <- list()
+  for (i in seq_along(file_names)) {
+    slim_dfs[[mkeep[i]]] <- readRDS(file_names[i])
+  }
+}
+
+slim_dfs$`COVIDhub-4_week_ensemble`$model <- "COVIDhub-ensemble"
+
+plot_components <- function(...) {
+  alloscore::plot_components_slim(
+    ...,
+    origin_time_col_name = "reference_date",
+    model_col_name = "model",
+    target_col_name = "abbreviation"
+  )
+}
+
+K1 <- slim_dfs$`COVIDhub-4_week_ensemble` %>%
+  filter(reference_date == forecast_dates[14])
+
+K1_long <- K1 %>% unnest(xdf)
+
+K1_long
+K1_long_plus <- K1_long %>% mutate(components = components - min(components))
+
+p1 <- plot_components(K1_long_plus, show_oracle = FALSE, order_at_K = 15800, show_raw = FALSE)
+
+p1 + lims(x=c(1000,25000)) + theme_bw()
+
+df <- bind_rows(slim_dfs$`MUNI-ARIMA`, slim_dfs$`JHUAPL-Bucky`)
+
+p <- plot_components(df %>% filter(K == 15000, reference_date == "2021-12-27") %>% unnest(xdf))
+
+
+
+
+# Create a repeating vector of these colors that matches the number of levels in your fill variable
+color_vector <- rep(my_colors, length.out = 51)
+
+p +  scale_fill_manual(values = color_vector)
+
+# bar plot for muni-v-bucky
+
+mvbucky <- slim_dfs[c("JHUAPL-Bucky", "MUNI-ARIMA")] %>% bind_rows() %>%
+  filter(reference_date == "2021-12-27", K == 22000) %>% unnest(xdf) %>%
+  filter(abbreviation %in% c("FL", "CA", "TX", "GA", "SC", "NC", "PA", "VA", "WA"))
+
+p_mvb_bar <- plot_components(
+  mvbucky,
+  order_at_K = 22000,
+  order_at_model = "JHUAPL-Bucky",
+  bar_positioning = "dodge")
[…]

### 2023-11-03 talks/plots/code/plot_hosps.R

@@ -1,0 +1,156 @@
+library(covidHubUtils)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(geofacet)
+
+source("plot_functions.R")
+save_dir <- file.path("talks/plots/output/")
+
+#hub_repo_path <- "~/research/epi/covid/covid19-forecast-hub/"
+hub_repo_path <- "~/covid/covid19-forecast-hub/"
+
+forecasts_hosp <- targets::tar_read(forecast_data) %>%
+  bind_rows(
+    load_forecasts(
+      dates = "2021-12-27",
+      date_window_size = 6,
+      locations = "US",
+      types = c("quantile"),
+      targets = paste(0:30, "day ahead inc hosp"),
+      source = "local_hub_repo",
+      hub_repo_path = hub_repo_path,
+      verbose = FALSE,
+      as_of = NULL,
+      hub = c("US")) |>
+      align_forecasts() |>
+      dplyr::filter(
+        relative_horizon == 14)
+  ) %>%
+  filter(reference_date == "2021-12-27")
+truth <- targets::tar_read(truth_data)
+
+mkeep <- c("BPagano-RtDriven",
+           "COVIDhub-ensemble",
+           "COVIDhub-baseline",
+           "CU-select",
+           "IHME-CurveFit",
+           "JHUAPL-Bucky",
+           "JHUAPL-Gecko",
+           "MUNI-ARIMA",
+           "USC-SI_kJalpha",
+           "UVA-Ensemble")
+
+three_colors <- c("#4A708B", "goldenrod", "#CD2626")
+
+p_VA_JHU_MUNI <- plot_hosp(models = c("JHUAPL-Bucky","COVIDhub-ensemble", "MUNI-ARIMA"),
+                     f_width1 = 3, f_alpha = .4, locations = c("VA", "GA"))
+
+p_2Ls_CU_IHME <- plot_hosp(models = c("CU-select","COVIDhub-ensemble", "MUNI-ARIMA"),
+                          f_width1 = 3, f_alpha = .4, locations = c("FL", "NY"))
+
+sc = 1
+ggsave(plot = p_VA_CU_IHME, filename = "p_VA_CU_IHME.png",
+       width = 11*sc, height = 6*sc, path = save_dir)
+ggsave(plot = p_VA_CU_IHME, filename = "p_VA_JHU_MUNI.png",
[…]

### 2023-11-03 talks/plots/code/slim_script.R

@@ -1,0 +1,55 @@
+
+library(tidyverse)
+library(alloscore)
+library(targets)
+
+args <- commandArgs(trailingOnly = TRUE)
+j <- as.integer(args[1])
+print(j)
+
+
+mkeep <- c("BPagano-RtDriven",
+           "COVIDhub-4_week_ensemble",
+           "COVIDhub-baseline",
+           "CU-select",
+           "IHME-CurveFit",
+           "JHUAPL-Bucky",
+           "JHUAPL-Gecko",
+           "MUNI-ARIMA",
+           "USC-SI_kJalpha",
+           "UVA-Ensemble")
+
+forecast_dates = as.character(seq.Date(as.Date("2021-10-18"), as.Date("2022-02-28"), by = "7 days"))
+
+values <- tidyr::expand_grid(models = mkeep, forecast_dates = forecast_dates)
+
+# hard coding the names from tar_map
+# this seems like it should be its own target but I can't figure this out right now
+a_target_names <- values %>% dplyr::mutate(
+  target_name = paste0("alloscore_",
+                       gsub(x = models, pattern = "-", replacement = "."),
+                       "_",
+                       gsub(x = forecast_dates, pattern = "-", replacement = ".")
+  )) %>% dplyr::select(-forecast_dates)
+
+a_target_names_nested <- a_target_names %>% tidyr::nest(target_list = target_name)
+
+slim_dfs <- function(i) {
+  target_list <- purrr::map(a_target_names_nested$target_list[[i]][[1]], ~tar_read_raw(.))
+
+  saveRDS(target_list %>%
+    purrr::map(possibly(~ alloscore::slim(., id_cols = c("reference_date", "model")))) %>%
+    bind_rows(), file = paste0("data/slim_", a_target_names_nested$models[i]))
+
+  rm(target_list)
+}
+
+# Get the names of all objects in the global environment
+#
+# for (i in 6:10) {
+#   slim_dfs(i)
+# }
+
+slim_dfs(j)
+
+

### 2023-11-03 alloscore_manuscript/alloscore-application.Rnw
adding updated application section and discussion.

@@
 
-We illustrate with an application to hospital admissions in the U.S., considering the problem of allocation of a limited supply of medical resources to {}the states.
-
-Case study heading into the Omicron wave.
-Some more detailed discussion of implications of bad forecasts for specific decision making purposes -- take a "deep dive" into one or two example states like FL.
-
-Look at results over a broader range of time.
+We illustrate with an application to hospital admissions in the U.S., considering the problem of allocation of a limited supply of medical resources to states.
 
@@
 \subsection{Data}
+<<get-data-dates>>=
+metadata <- tar_meta()
+@
+
 \subsubsection{Hospitalization data}
-Starting in the summer of 2020, the US Health and Human Services began reporting counts of daily new admissions to hospitals for individuals with COVID-19.(cite HHS Protect)
+Starting in the summer of 2020, the US Health and Human Services began reporting counts of daily new admissions to hospitals for individuals with COVID-19.\citep{healthdatagov_covid-19_nodate}
 These data served as the source of ``ground truth'' data for the US COVID-19 Forecast Hub which, starting in December 2020, collected short-term forecasts of new hospital admissions at the daily scale.
@@
 The data were updated daily and were available for download by the public through the HHS HealthData.gov website.
-For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data.(cite package) \ngr{can we use tar\_manifest() to reproducibly display the date the data were stored?}
+For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data.\cite{wang-covidhubutils}
+The hospitalization data were downloaded for analysis on \Sexpr{as.Date(metadata$time[which(metadata$name == "truth_data")])}.
 
@@
 As such, we downloaded forecasts for the 15 weeks starting with Monday 2021-11-22 through Monday 2022-02-28.
+Forecast data were downloaded on for analysis on \Sexpr{as.Date(metadata$time[which(metadata$name == "forecast_data")])}.
 
@@
 For each week, we computed a the allocation score for each 14 day-ahead forecast based on $K=15,000$.
-\ngr{it appears that for the last two weeks selected (forecast dates: 2022-02-14 and 2022-02-28) the upper limit of K was 11400 and 9200 respectively, so we do not have alloscores computed for K=15000.
-Maybe they would have all been zero, since the limits were likely set based on the number of cases being low.
-What is the rule used to determine the max K computed?
-Don't think we need to change it, but would be nice to state in the manuscript what it is.}
 
@@
 
-\subsubsection{Metrics as a function of time}
+\subsubsection{Forecast scores showed differences over time}
 <<grob-hosps-over-time>>=
@@
 The allocation scores are on the whole highest when the observed number of new hospital admissions is closest to the resource threshold of 15,000, as this is the time when any mistakes in allocation are costly in terms of wasting resources in one location that could have been used in another.
-Predictions made during the peak week and just after showed the highest variation in allocation scores, with some models having values under 1000 and others having values over 3500.
+Predictions made during the peak week and just after showed the highest variation in allocation scores, with some models having allocation scores under 1000 and others having values over 3500.
+
+Overall, the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, had the highest rank for allocation score in six weeks, more than any other model except the \texttt{COVIDhub-ensemble} which also had six.
+
 <<eval=FALSE>>=
@@
 
-\subsubsection{correlation between WIS and allocation score}
+\subsubsection{Model rankings for different scores did not reliably correlate}
 
-<<metrics-correlation, fig.width=7, fig.height=6, fig.cap="Association of standardized ranks for MWIS and allocation score by model and week.
+<<metrics-correlation, fig.width=10, fig.height=7.5, fig.cap="Association of standardized ranks for MWIS and allocation score by model and week.
 Each facet of the plot corresponds to one model.
@@
 
-Models show differing levels of correlation between their allocation scores and MWIS values (Figure \ref{fig:metrics-correlation}).
-Here are some examples of the different patterns observed:
+Models showed differing levels of correlation between their allocation scores and MWIS values (Figure \ref{fig:metrics-correlation}).
+Here are some examples of the different model-specific patterns observed:
 \begin{itemize}
-\item positive association between allocation score and MWIS ranks: \texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha} models
-\item consistently strong MWIS ranks, highly variable allocation score ranks, no clear association: \texttt{JHUAPL-SLPHospEns} and \texttt{CU-select}
-\item consistently strong ranks for both metrics, no clear association: \texttt{COVIDhub-ensemble}
+\item Several models showed a positive association between allocation score and MWIS ranks (e.g., \texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha}).
+\item Several models had consistently strong MWIS ranks but also had highly variable allocation score ranks with no clear association between the two (e.g., \texttt{JHUAPL-SLPHospEns} and \texttt{CU-select})
+\item One model performed consistently well for both metrics with no clear association (\texttt{COVIDhub-ensemble})
+\item One model performed consistently well for allocation score but had only middlings ranks for MWIS (\texttt{CMU-TimeSeries})
 \end{itemize}
-Additionally, the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, had the highest rank for allocation score in six weeks, more than any other model except the \texttt{COVIDhub-ensemble} which also had six.
 
@@
 
- - bubble figure with ranks sized by total observations
-
 \subsubsection{Integrated allocation score across values of K}
 
+Using the integrated allocation score summarizes allocation scores across a range of possible values of the constraint ($K$) (Section \ref{sec:methods.detailed.integrated_allocation}).
+This could be useful in situations where the actual constraint may not be known, or could only be specified as a probability distribution.
+
  - something here from Ben's analysis?
+
+\ngr{include figure here that shows relative rankings of locations?}
 
@@
 
-In summary, proper scoring rules for probabilistic forecasts can be traced back to a loss function motivated by a specific decision-making context.
-Often, forecast scores are used to rank models according to accuracy with that particular score, but without reference to the underlying decision-making process for which that score was designed.
+All proper scoring rules for probabilistic forecasts have an explicit link to a loss function, thereby placing numerical value on different kinds of errors in specifying forecasts.
+In epidemiological forecasting, well-known proper scoring rules such as the log score or variations on the continuous rank probability score (such as the weighted interval score, or WIS) have been frequently utilized to evaluate probabilistic forecasts.
+Often, these scores are used to rank models according to accuracy with that particular score, but without reference to the underlying decision-making process for which that score was designed.
 With careful thought and collaboration between modelers and public health officials, we argue that scores that are more aligned with public health decisions could be developed to inform specific problems.
-In some situations, individual or ensemble models could be developed to minimize the loss for a particular setting, although this is also an area largely unexplored in the realm of public health to date.
-
-Notes for the application
-
- - It is clear that the metrics are capturing different aspects of performance
-
- - WIS is scale dependent, alloscore not as much
-
- - the fact that it is hard to beat the baseline for allocation suggests that we are not consistently adding value over just looking at the current levels.
-\ngr{although is there something more to it about how the baseline accounts for uncertainty?}
+We have demonstrated that forecast evaluation methods that are tied to a specific decision making context can yield model rankings that are substantively different from standard measures of forecast skill.
 
@@
 This work seeks to address that gap.
+However, we do note that the decision-making context presented in this work, while motivated by examples from real-world public health resource allocation problems, has not used to inform an actual real-time decision-making process.
 
-We have demonstrated that evaluation methods that are tied to decision making context can yield model rankings that are substantively different from generic measures of forecast skill like WIS.
+Resource allocation decisions are a realistic example of the kinds of decisions that could motivate more targeted forecasting exercises.
+One real-world example is allocation of ventilators during a respiratory viral pandemic.
+This was the motivating example for the examples presented in the work above.
+However, other examples where allocation is a central concern exist as well, such as the allocation of a limited stockpile of vaccinations\citep{persad_fair_2023} or diagnostic tests\citep{du_optimal_2022}.
 
-In practice, there are many users of forecasts with many different decision making problems.
+In practice, there are many potential users of forecasts with many different decision making problems.
 Not all can be easily quantified.
-Those that can be easily quantified may differ enough that no single score is appropriate for all users.
-We suggest reporting multiple scores.
-This may be tricky to operationalize in the setting of a general forecast hub.
-It matters how you elicit and represent probabilistic forecasts (quantiles?
-samples?
-cdfs?).
+Those that can be easily quantified may differ enough that it seems likely that no single score would be appropriate for all users.
+In an ideal setting, a forecasting tool could be developed through close collaboration between modelers and public health officials.
+However, this setting may only be possible in settings with enough resources to support both an analytics and public heatlh team.
+Increasingly, collaborative modeling hubs are being used to generate ``one-size-fits-all'' forecasts for many locations at once.
+In these settings, where tailored models are not available, it still could be possible to evaluate contributed models against a set of multiple scores.
+% NGR: leaving this out for now as it feels like too much detail?
+% This may be tricky to operationalize in the setting of a general forecast hub. It matters how you elicit and represent probabilistic forecasts (quantiles? samples? cdfs?).
 
-The allocation score we developed here does not directly account for important considerations such as fairness/equity of allocations.
+In our specific application, we made several key observations that we believe should inform future work in this area.
+First, it is clear that different metrics (in our case allocation score and mean weighted interval score) captured different aspects of forecast performance.
+The MWIS was strongly dependent on the scale of the forecasted quantity (e.g., when hospitalizations were high, so was WIS).
+The allocation score was not as scale-dependent, as the highest/worst allocation scores were observed when the number of hospitalizations was closest to the allocation constraint.
+We also note that contributed models found it hard to have a better allocation score than a na\"ive baseline model that just predicted a flat line from the last previous observation with wide uncertainty.
+This suggests suggests that contributed models (other than the ensemble, which combined forecasts from all contributed models) were not consistently adding value over just extrapolating from the current levels.
 
-The allocation score we developed also does not attempt to capture the broader context of decision making.
+There are several important limitations to the current work.
+The allocation score we developed here does not directly account for important considerations such as fairness or equity of allocations.
+Nor does the proposed framework attempt to capture the broader context of decision making.
 For example, in practice it may be possible to increase the resource constraint $K$ by shifting funding from other disease mitigation measures.
+We also note that in some settings, a ``successful'' epidemiological forecast may lead to policy decisions that change the distribution of the predicted outcome $Y$.
+Our framework cannot be used to evaluate a forecast at a horizon that may have been impacted by such a decision.
 
-Forecaster's dilemma: a successful forecast may lead to decisions that change the distribution of the outcome $Y$. Our framework cannot be used in those settings.
+We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work in more closely coupling applied epidemiological forecasting with public health decision-making.
+We note that others have explored similar linkages.
+For example, \cite{bilinski_adaptive_2023} explore ways in which policymaker preferences could inform risk thresholds for predictive models.
+\ngr{Would be nice to have a few other examples here.}
+In some situations, individual or ensemble models could be developed to minimize the loss for a particular setting, although this is also an area largely unexplored in the realm of public health to date, with some initial methodological development in econometrics.\citep{loaiza-maya_focused_2021}
 
-There is much more to do in this general area.
-
-\section{References}
+In conclusion, we argue that the way modelers and policymakers view and evaluate forecasts should change depending on the specific decision-making context.
+Using standard forecast evaluation metrics can mask the utility of certain forecasts, or lead to forecasts being used to inform a particular decision that are not optimal for that particular set of values that are important for that context.
+Further collaborative work is needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or simulation exercises that could be used to inform further development of new or alternative scoring metrics.
 

### 2023-11-03 alloscore_manuscript/alloscore-application.Rnw
More general statement about who uses WIS in abstract

@@
 We illustrate this scheme with quantile forecasts of COVID-19 hospitalizations in the US at the state level that are recorded in the COVID-19 Forecast Hub, with the goal of determining the allocation of a hypothetical limited resource across the states.
-The forecast skill ranking given by this allocation scoring rule can vary substantially from the ranking given by the weighted interval score now used by the CDC, especially during surges in hospitalizations such as in late 2021 as the Omicron wave began.
+The forecast skill ranking given by this allocation scoring rule can vary substantially from the ranking given by the weighted interval score that is often used to score quantile forecasts, especially during surges in hospitalizations such as in late 2021 as the Omicron wave began.
 We see this as strong evidence that the allocation scoring rule detects forecast value that is missed by traditional accuracy measures and that the general strategy of designing scoring rules that are directly linked to policy performance is a promising research direction for epidemic forecast evaluation.

### 2023-11-03 alloscore_manuscript/alloscore-application.Rnw
updating bubble figure to remove last 4 weeks of analysis with lots of zeroes.

--- 2023-11-03 alloscore_manuscript/alloscore-application.Rnw
@@
 
-Overall, the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, had the highest rank for allocation score in six weeks, more than any other model except the \texttt{COVIDhub-ensemble} which also had six.
-
 <<eval=FALSE>>=
@@
             range = maxAS-minAS)
+@
+
+Overall, across the first 11 weeks evaluated (we excluded the last four since nearly all the models achieved an allocation score of zero), the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, had the highest rank for allocation score in four weeks, more than any other model except the \texttt{COVIDhub-ensemble} which also had four.
+
+<<eval=FALSE>>=
+## analysis for the above statements
+alloscore_data_summaries |>
+  filter(as.Date(reference_date) < as.Date("2022-02-01"), allo_rank == 1)
 @
@@
 
-ggplot(all_data,
-       aes(x=mwis_rank_std, y=allo_rank_std, color=target_end_date, size=yobs)) +
+all_data |>
+  filter(as.Date(reference_date.x) < as.Date("2022-02-01") ) |>
+  ggplot(aes(x=mwis_rank_std, y=allo_rank_std, color=target_end_date, size=yobs)) +
   geom_point(alpha=.7) +

### 2023-11-03 alloscore_manuscript/alloscore-application.Rnw
some cleanup of intro paragraph 2

--- 2023-11-03 alloscore_manuscript/alloscore-application.Rnw
@@
 We illustrate this scheme with quantile forecasts of COVID-19 hospitalizations in the US at the state level that are recorded in the COVID-19 Forecast Hub, with the goal of determining the allocation of a hypothetical limited resource across the states.
-The forecast skill ranking given by this allocation scoring rule can vary substantially from the ranking given by the weighted interval score that is often used to score quantile forecasts, especially during surges in hospitalizations such as in late 2021 as the Omicron wave began.
+The forecast skill ranking given by this allocation scoring rule can vary substantially from the ranking given by the weighted interval score now used by the CDC, especially during surges in hospitalizations such as in late 2021 as the Omicron wave began.
 We see this as strong evidence that the allocation scoring rule detects forecast value that is missed by traditional accuracy measures and that the general strategy of designing scoring rules that are directly linked to policy performance is a promising research direction for epidemic forecast evaluation.
@@
 \cite{bertsimas2021predictionsCOVID} developed tools to inform decision making from infectious disease forecasts, which have been used to inform allocation of limited medical supplies such as ventilators, ICU capacity planning, and vaccine distribution strategy.
-Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping (University of Texas, 2022) \nocite{utnews2022}.
+Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping (University of Texas, 2022)\nocite{utnews2022}.
 In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
@@
 In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to.
-We give an overview of these procedures in Section \ref{sec:methods.decisiontheory}.
-\ngr{This section label doesn't exist. Should it point to methods.detailed?}
+We give an overview of these procedures in Section \ref{sec:methods.detailed.decisiontheory}.
 There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
-We review this work only briefly here, and we refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \elr{TODO: identify relevant review or book style summary for meteorology} for discussions focused on applications to economics and meteorology, respectively.
+We review this work only briefly here, and we refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \cite{murphy1993whatisagoodforecast} for discussions focused on applications to economics and meteorology, respectively.
 In finance, the value of forecasts can often be measured by the profits generated by trading decisions informed by the forecasts, perhaps adjusted for risk levels \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
-In applications to supply chain management and meteorology, the value of forecasts has typically been operationalized by considering the costs associated with decisions regarding the amount of inventory to hold or the level of protection against the impacts of extreme weather events to enact {Peter Catt (2007), Fotios Petropoulos and colleagues (2019), Nada Sanders and Gregory Graman (2009), T.N.
-Palmer (2002), Florian Pappenberger and colleagues (2015)}.
+In applications to supply chain management and meteorology, the value of forecasts has typically been operationalized by considering the costs associated with decisions regarding the amount of inventory to hold or the level of protection against the impacts of extreme weather events to enact \cite[e.g.,][]{catt2007assessingcostofforecasterror, petropoulos2019inventoryperformanceforecasting, palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
 For example, in supply chain management these decisions may incur costs related to holding inventory, labor, or providing poor service, while in meteorology we may need to balance the costs of implementing protective measures with the costs of potentially preventable weather damages.
 In this framework, a forecast has value if it leads to decisions with low total costs.
-In all of these fields, analyses have consistently found that common measures of statistical forecast accuracy do not necessarily correspond directly to measures of the value of forecasts as an input to decision making \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
+In all of these fields, analyses have consistently found that common measures of statistical forecast accuracy do not necessarily correspond directly to measures of the value of forecasts as an input to decision making \cite[e.g.,][]{leitch1991economicForecastEval, murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}.
 

### 2023-11-06 talks/ISF_talk/talk.qmd

@@
 ::: columns
-
 ::: {.column width="50%"}
-
 About me:
 
-- PhD in pure math
-- transitioned to public health data science 
-- started as RA with Nick and Evan at [reichlab.io](reichlab.io) at UMass in 2021
-- maintain COVID-19 Forecast Hub ensemble
-- current research goals center on ensemble methodology
-  - focus on scoring rules, decision theory, forecast-policy alignment
-
+After working as a post-doc in pure math for several years I transitioned to public health data science, joining Nick Reich and Evan Ray's group (reichlab.io) at UMass in 2021 as an RA.
+I help maintain the COVID-19 Forecast Hub ensemble forecasting system and do research on ensemble methodology, scoring rules, decision theory, and forecast-policy alignment.
 :::
@@
 ::: {.column width="50%"}
-
 ![](Reichlab.png)
@@
 ![](C19FH.png)
-
 :::
-
 :::
@@
 
-![](hub_timeline.png){width=80%}
+![](hub_timeline.png){width="80%"}
 
-::: {.attribution}
+```{=html}
+<!-- ::: attribution
 Figure credits: Alex Vespignani and Nicole Samay
-:::
-
+::: -->
+```
 New directions: "Hubverse" infrastructure project [https://hubdocs.readthedocs.io/](), genomic targets
@@
 
-Primary scoring rule used currently is the **Weighted Interval Score** which is a discrete approximation to the 
-**Continuous Ranked Probability Score** (a probabilistic generalization of MAE)
+Primary scoring rule used currently is the **Weighted Interval Score** which is a discrete approximation to the **Continuous Ranked Probability Score** (a probabilistic generalization of MAE)
 
@@
 ::: box
-
-Create scenario where the forecaster believes maximizing her own expected utility aligns 
-with leading the DM toward actions that maximize social utility.
-
+Create scenario where the forecaster believes maximizing her own expected utility aligns with leading the DM toward actions that maximize social utility.
 :::
@@
 ::: box
-
-The **Bayes act** relative to a forecast $Y \sim F$ is the action $x^F$ 
-by which a DM will minimize **expected loss** $E_Y[l]$ 
-assuming that in reality $Y \sim F$.
-
+The **Bayes act** relative to a forecast $Y \sim F$ is the action $x^F$ by which a DM will minimize **expected loss** $E_Y[l]$ given that $Y \sim F$.
 :::
 
- - loss $l$ is a function of the outcome $y$ **and** the action $x$ DM takes in anticipation of $y$
+-   loss $l(x,y)$ is the **consequence** of experiencing $y$ having prepared with $x$
 
 ::: box
-
-Key point: Given $l$, we have an *automatically proper* scoring rule 
-$S(F,y) := l(x^F, y)$
-
+Key point: Given $l$, we have an *automatically proper* scoring rule $S(F,y) := l(x^F, y)$
 :::
@@
 
-$$
-\begin{align}
+```{=tex}
+\begin{align*}
 l(x,y) &= \sum_{i=1}^{52} \max(y_i - x_i, 0) = \sum (y_i - x_i)_{+}\\
 x^F &\in \left\{ \text{possible } x' \,\middle| \, E_{F}[l(x',Y)] \leq E_{F}[l(x,Y)] \text{ for any possible } x \right\}
-\end{align}
-$$
+\end{align*}
+```
 
@@
 ::: box
-
 Define set of **possible** $x$ as those actions which satisfy $\displaystyle{\sum x_i \leq K}$
-
 :::
@@
 
-Our loss function and (single) constraint are very simple prototypes meant 
-for proof of concept.
+Our loss function and (single) constraint are very simple prototypes meant for proof of concept.
 
@@
 
+```{=tex}
 \begin{align}
@@
 \end{align}
-
+```
 Our allocation scoring rule is then given by the recipe:
@@
 
-So we resort to a binary search for $\lambda^{\star}$ following long history in inventory management of solving **constrained multi-product newsvendor** 
-problems^[See e.g., Zhang, Bin, Xiaoyan Xu, and Zhongsheng Hua. 2009. “A Binary Solution Method for the Multi-Product
-Newsboy Problem with Budget Constraint.”]
+So we resort to a binary search for $\lambda^{\star}$ following long history in inventory management of solving **constrained multi-product newsvendor** problems[^1]
+
+[^1]: See e.g., Zhang, Bin, Xiaoyan Xu, and Zhongsheng Hua. 2009. "A Binary Solution Method for the Multi-Product Newsboy Problem with Budget Constraint."
 
@@
 ::: columns
-
 ::: {.column width="30%"}
-
 After adjusting for an oracle the score of an oracle that allocates *fairly* based on what demands actually occur, scores resemble "value" plots from meteorology
@@
 
-- ranking at peak seems to be consistent with that in large $K$
-region.
-
+-   ranking at peak seems to be consistent with that in large $K$ region.
 :::
@@
 ::: {.column width="70%"}
-
 ![](plots/peak_alloscore.png)
-
 :::
-
 :::
@@
 ::: aside
-Ehm, Werner, Tilmann Gneiting, Alexander Jordan, and Fabian Krüger. 2016. “Of Quantiles and Expectiles: Consistent Scoring Functions, Choquet Representations and Forecast Rankings.”
+Ehm, Werner, Tilmann Gneiting, Alexander Jordan, and Fabian Krüger. 2016. "Of Quantiles and Expectiles: Consistent Scoring Functions, Choquet Representations and Forecast Rankings."
 :::
@@
 -   an allocation score (on joint forecast)
+
 -   any allocation based loss
+
 -   our quantile interpolation/extrapolation methods (`distfromq`)
@@
 
-We hope/think that allocation scoring is sensitive to implicit dependence structures in forecasts, but all work so far only refers directly to marginals
-- nothing yet with copulas, etc.
+We hope/think that allocation scoring is sensitive to implicit dependence structures in forecasts, but all work so far only refers directly to marginals - nothing yet with copulas, etc.
 
@@
 
-[https://github.com/aaronger/alloscore](https://github.com/aaronger/alloscore)
+<https://github.com/aaronger/alloscore>
 
@@
 
-[https://github.com/reichlab/distfromq](https://github.com/reichlab/distfromq)
+<https://github.com/reichlab/distfromq>
 
+```{=html}
 <!-- ## Loss vs Utility
@@
 Change in utility but not in loss from change in - supply and personnel cost structures - public opinion regarding importance of fatality prevention - knowledge of demographic disparities in consequences of outcomes -->
-
+```
 ## Data workflow
@@
 
-![](plots/state_hosps_ens_only.jpeg){width=100%}
+![](plots/state_hosps_ens_only.jpeg){width="100%"}
 
@@
 
-An explicit vacabulary of social loss in ID forecast evaluation has instrinsic advantage
-over "off-the-shelf" metrics from statistic inference
+An explicit vacabulary of social loss in ID forecast evaluation has instrinsic advantage over "off-the-shelf" metrics from statistic inference
 
@@
 
-inference has own decision theory and ideas of utility, loss, Bayes acts
-that are related but distinct from the fundamentally econommic/policy focussed
-ones we use here
+inference has own decision theory and ideas of utility, loss, Bayes acts that are related but distinct from the fundamentally econommic/policy focussed ones we use here
 :::

### 2023-11-06 talks/midas2023/gerding---forecast-evaluation-via-social-utility.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  9pt,
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolshorizontal
% Prevent slide breaks in the middle of a paragraph
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\newif\ifbibliography
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
\usepackage{caption}
% Make caption package work with longtable
\makeatletter
\def\fnum@table{\tablename~\thetable}
\makeatother
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\titlegraphic{
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth]{reichlab.png}
\end{minipage}%
\hspace{0.06\linewidth}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth]{C19FH.png}
\end{minipage}%
}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdfauthor={Aaron Gerding, Nick Reich, Ben Rogers, Evan Ray},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Infectious Disease Forecast Evaluation Via Social Utility:
Allocation Scores}
\subtitle{MIDAS 2023}
\author{Aaron Gerding, Nick Reich, Ben Rogers, Evan Ray}
\date{}
\institute{School of Public Health and Health Sciences, UMass, Amherst\\
Department of Biostatistics and Epidemiology}

\begin{document}
\frame{\titlepage}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, interior hidden, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi

\begin{frame}{Outbreak Forecast Hubs and \textbf{Informed} Resource
Allocation}
\protect\hypertarget{outbreak-forecast-hubs-and-informed-resource-allocation}{}
Hubs developed to \textbf{``inform public health responses''}

\begin{itemize}
\item
  such as how resources are \textbf{allocated} among locations
\item
  e.g., medical supplies, facility capacity, personnel
\end{itemize}

\includegraphics{../plots/output/p_VA_JHU_MUNI.jpeg}

\note{quantiles flexible interpretable - closer to describing full
distribution which DM probably always wants

multicenter effort to collect forecasts via hubs for ID outbreaks
(epi/pan/endemic)

With uncertainty quantification being well-established as needed DM
input,

Allocations can be wrt space, time, demographic

Next: formalize this.}
\end{frame}

\begin{frame}{Scoring rules and social welfare}
\protect\hypertarget{scoring-rules-and-social-welfare}{}
Hubs strive to rank and combine forecasts so as to optimize social
welfare via the public health decisions that forecasts inform

\begin{itemize}
\item
  uncertainty quantification, and therefore, probabilistic forecasting
  essential
\item
  basic strategy: use a \textbf{scoring rule} \(S\) which assigns a loss
  \(S(F,y)\) when a probabilistic forecast \(F\) of \(Y\) is chosen and
  \(Y=y\) is observed.
\item
  optimizing welfare requires that forecasters say what they believe; so
  \(S\) should be \textbf{proper} meaning
  \(\displaystyle{E_{F}[S(F,Y)] \leq E_{F}S(G,Y)]}\) for all \(F, G\)
\end{itemize}

Current standard is the \textbf{Weighted Interval Score} (discrete CRPS)

\begin{itemize}
\tightlist
\item
  adopted largely for convenient scoring of quantile forecasts.
\end{itemize}
\end{frame}

\begin{frame}{Tools from decision theory}
\protect\hypertarget{tools-from-decision-theory}{}
\begin{block}{A central goal in design of scoring rules}
\protect\hypertarget{a-central-goal-in-design-of-scoring-rules}{}
Tie forecast scores directly to the benefits to society of the decisions
they inform

Key tools from decision theory for linking success/failure of
forecast-informed policy actions to scoring rules:

\begin{itemize}
\item
  Let \(l(x,y)\) be the \textbf{loss} of experiencing \(y\) after taking
  policy action \(x\)
\item
  The \textbf{Bayes risk} of a forecast \(Y \sim F\) is
  \(\min_x E_F[l(x,Y)]\)
\item
  A \textbf{Bayes act} for \(F\) is an action \(x^F\) that attains this
  minimum
\item
  Losses from Bayes acts define an \emph{automatically proper} scoring
  rule
\end{itemize}

\[
S(F,y) := l(x^F, y)
\]

\textbf{Note:} Scoring rules are really only a means to an end. It is
the expected score, in this case the Bayes risk, which characterize the
value a forecast adds for a decision maker. Scoring rule sample averages
estimate the expected score.
\end{block}
\end{frame}

\begin{frame}{A new scoring rule via a new loss function for constrained
actions}
\protect\hypertarget{a-new-scoring-rule-via-a-new-loss-function-for-constrained-actions}{}
\begin{block}{Our basic example:}
\protect\hypertarget{our-basic-example}{}
\begin{itemize}
\item
  \(x\) and \(y\) are vectors in \(\mathbb{N}_0^{52}\)
\item
  \(y =\) number of severe cases in US states and territories
\item
  \(l(x,y) =\) unmet need when \(x\) beds allocated and \(y\) severe
  cases occur
\end{itemize}

\begin{align*}
l(x,y) &= \sum_{i=1}^{52} \max(y_i - x_i, 0)\\
x^F &= \text{minimizer of } E_{F}[l(x,Y)] \text{ over all feasible } x
\end{align*}

\textbf{Our new idea:} Define \textbf{feasible} \(x\) as satisfying
\(\displaystyle{\sum x_i \leq K}\)

\note{social/political/economic/psychological/moral utility

this must somehow acknowledge risk attitude heterogeneity among various
DM's that a forecast hub might serve

gpl parameters, weights, multiple constraints}
\end{block}
\end{frame}

\begin{frame}{Computation}
\protect\hypertarget{computation}{}
Obtaining \(x^F\) is a constrained stochastic optimization problem

\begin{itemize}
\item
  known in inventory management as a \textbf{constrained multi-product
  newsvendor} problem
\item
  formally solvable using Lagrange multiplier method to get a quantile
  representation \(x_i^F = F_i^{-1}(\tau(K,F)), i = 1,\ldots,N\)
\item
  in practice, we find \(x_i^F\)'s via an iterative method:
\end{itemize}

\includegraphics{../plots/output/iter_combo.png}
\end{frame}

\begin{frame}{Application}
\protect\hypertarget{application}{}
December 2021: Omicron wave clearly started US but forecast teams unsure
of severity given uncertainty about \(R_0\), cross-protection by
vaccination, previous infection, etc.

\includegraphics{../plots/output/nat_hosps.png}
\end{frame}

\begin{frame}{Oracle adjusted allocation scores near Omicron peak}
\protect\hypertarget{oracle-adjusted-allocation-scores-near-omicron-peak}{}
\includegraphics{../plots/output/peak_alloscore.png}
\end{frame}

\begin{frame}{Some Observations}
\protect\hypertarget{some-observations}{}
\begin{columns}[T]
\begin{column}{0.4\textwidth}
\begin{itemize}
\item
  related to the Murphy curves of Ehm, Gneiting, Jordan, Krüger, 2016
\item
  extreme shortage or surplus diminishes oracle's advantage
\item
  ranking consistent across large \(K\) region.
\end{itemize}
\end{column}

\begin{column}{0.6\textwidth}
\includegraphics{../plots/output/peak_alloscore_min.png}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Alloscore and WIS rank models differently}
\protect\hypertarget{alloscore-and-wis-rank-models-differently}{}
\includegraphics{../plots/output/p_wis_v_alloscore.png}
\end{frame}

\begin{frame}{Explanations?}
\protect\hypertarget{explanations}{}
\includegraphics{../plots/output/mvbucky_dist_alloc.png}
\end{frame}

\begin{frame}[fragile]{Limitations}
\protect\hypertarget{limitations}{}
This is \textbf{post-hoc} analysis

Hub forecasters were unaware of

\begin{itemize}
\item
  an allocation score (on joint forecast)
\item
  any allocation based loss
\item
  our quantile interpolation/extrapolation methods (\texttt{distfromq})

  \begin{itemize}
  \tightlist
  \item
    might be especially important for tails
  \end{itemize}
\end{itemize}

We hope/think that allocation scoring is sensitive to implicit
dependence structures in forecasts, but all work so far only refers
directly to marginals - nothing yet with copulas, etc.

\begin{block}{Thank you!}
\protect\hypertarget{thank-you}{}
A very rough R package I wrote to implement scoring procedures:

\url{https://github.com/aaronger/alloscore}

A less rough package Evan wrote to implement cdf reconstruction

\url{https://github.com/reichlab/distfromq}
\end{block}
\end{frame}

\end{document}

### 2023-11-06 alloscore_manuscript/alloscore-application.Rnw

@@
 In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
-The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid} or weighted interval score (WIS) \citep{bracher2021evaluating}, however, recently authors have been calling for evaluating forecasts through their impact on policy \citep{marshall2023predictions}.
+The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid} or weighted interval score (WIS) \citep{bracher2021evaluating}.
+However, recently authors have been calling for evaluating forecasts through their impact on policy \citep{marshall2023predictions, bilinski_adaptive_2023}.
 
@@
 For example, \cite{ioannidis2022forecastingCOVIDfailed} discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts.
-Separately, there is a thread of literature that does quantify the link between infectious disease modeling and policy making, but this work has been done outside of a forecasting context.
+\cite{bilinski_adaptive_2023} explore ways in which policymaker preferences could inform risk thresholds for predictive models using a framework for measuring the costs and losses associated with taking an action that is similar to methods that have been used in meteorology and elsewhere.
+\cite{marshall2023predictions} develop a forecast scoring rule that is informally motivated by utility considerations, but the score is not derived from a decision-theoretic set up.
+Separately, there is a thread of literature that quantifies the link between infectious disease modeling and policy making outside of a forecasting context.
 As an example, \cite{Probert2016decisionMakingFootMouth} develop measures of the cost of actions designed to control a hypothetical outbreak of foot-and-mouth disease and use this framework to explore policy recommendations from a variety of simulation-based projection models.
@@
 We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work in more closely coupling applied epidemiological forecasting with public health decision-making.
-We note that others have explored similar linkages.
-For example, \cite{bilinski_adaptive_2023} explore ways in which policymaker preferences could inform risk thresholds for predictive models.
-\ngr{Would be nice to have a few other examples here.}
-In some situations, individual or ensemble models could be developed to minimize the loss for a particular setting, although this is also an area largely unexplored in the realm of public health to date, with some initial methodological development in econometrics.\citep{loaiza-maya_focused_2021}
+We note that a few papers have begun exploring similar linkages as described in the literature review \textendash but we see much room for additional work in this area.
+In some situations, individual or ensemble models could be developed to optimize scores that are attunend to a particular decision making setting.
+This area is also largely unexplored in the realm of public health to date, with some initial methodological development in econometrics \citep{loaiza-maya_focused_2021}.
 

### 2023-11-06 alloscore_manuscript/supplement.rnw
add supplement outline and gitignore tex for Rnws

--- 2023-11-03 alloscore_manuscript/alloscore-application.Rnw
@@
 
-\title{Evaluating infectious disease forecasts with allocation scoring rules}
+\title{Supplementary Material for ``Evaluating infectious disease forecasts with allocation scoring rules''}
 \author{Aaron Gerding, Nicholas G. Reich, Benjamin Rogers, Evan L. Ray}
@@
 \begin{abstract}
-
-The COVID-19 pandemic has led to rapid innovation in methods for eliciting and evaluating forecasts of infectious disease burdens, with a primary goal being to help public health workers make informed decisions about how to manage these burdens.
-However, explicit descriptions or quantifications of the value forecasts add to society through the decisions they support are elusive.
-Moreover, there has only been limited discussion of how predominant forecast evaluation metrics might indicate the success of policies based in part on those forecasts.
-
-Here we pursue one possible tether between multivariate forecasts and policy: the allocation of limited medical resources in response to COVID-19 hospitalizations in various regions so as to minimize expected unmet need.
-Given probabilistic forecasts of hospitalizations in each region, we formulate an allocation algorithm following techniques developed in operations research.
-We then score forecasts according to how much unmet need their associated allocations would have allowed.
-We illustrate this scheme with quantile forecasts of COVID-19 hospitalizations in the US at the state level that are recorded in the COVID-19 Forecast Hub, with the goal of determining the allocation of a hypothetical limited resource across the states.
-The forecast skill ranking given by this allocation scoring rule can vary substantially from the ranking given by the weighted interval score now used by the CDC, especially during surges in hospitalizations such as in late 2021 as the Omicron wave began.
-We see this as strong evidence that the allocation scoring rule detects forecast value that is missed by traditional accuracy measures and that the general strategy of designing scoring rules that are directly linked to policy performance is a promising research direction for epidemic forecast evaluation.
-
+We briefly address some technical and methodological points in the main text, referring to the forthcoming ... for 
+more thorough discussion.
 \end{abstract}
 
-\section{Introduction}
+\begin{itemize}
+	\item From 2.2.1, why are Bayes act scoring rules proper?
+	\item Explain ``All proper scoring rules for probabilistic forecasts have an explicit link to a loss function'' from discussion.
+	\item DGP as optimal for any decision problem, ref Diebold, Gunther, Tay p. 866.
+	\item For 2.2.2, how to get quantile representation of Bayes act using Lagrange multiplier, assuming smooth, never-zero densities well behaved at $x=0$.
+Work out exponetial example.
+Refer to methods paper for general case.
+	\item Derivation of quantile scoring rule with quantile as Bayes act for C/L problem, assuming never-zero densities. 
+	\item Descriptions of 
+		\begin{itemize}		
+			\item CRPS as average quantile score across $C \in [0/L]$ decision problems
+			\item IS as average of two quantile scores with a prob-width penalty
+			\item WIS as average quantile score across 23 C/L problems.
+		\end{itemize}
+	\item Sketch of scoring for decision problems involving both cost and constraint.
 
-Infectious disease forecasting models have emerged as important tools in public health.
-Predictions of disease dynamics have been used to inform decision making about a wide variety of measures to reduce disease spread and/or mitigate the severity of disease outcomes.
-For example, estimates of expected onset of flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola dynamics have been used to allocate surveillance resources \citep{meltzer2014estimating, rainisch2015regional}.
-\cite{bertsimas2021predictionsCOVID} developed tools to inform decision making from infectious disease forecasts, which have been used to inform allocation of limited medical supplies such as ventilators, ICU capacity planning, and vaccine distribution strategy.
-Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping (University of Texas, 2022)\nocite{utnews2022}.
-In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
-The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid} or weighted interval score (WIS) \citep{bracher2021evaluating}, however, recently authors have been calling for evaluating forecasts through their impact on policy \citep{marshall2023predictions}.
-
-In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to.
-We give an overview of these procedures in Section \ref{sec:methods.detailed.decisiontheory}.
-There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
-We review this work only briefly here, and we refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \cite{murphy1993whatisagoodforecast} for discussions focused on applications to economics and meteorology, respectively.
-In finance, the value of forecasts can often be measured by the profits generated by trading decisions informed by the forecasts, perhaps adjusted for risk levels \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
-In applications to supply chain management and meteorology, the value of forecasts has typically been operationalized by considering the costs associated with decisions regarding the amount of inventory to hold or the level of protection against the impacts of extreme weather events to enact \cite[e.g.,][]{catt2007assessingcostofforecasterror, petropoulos2019inventoryperformanceforecasting, palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
-For example, in supply chain management these decisions may incur costs related to holding inventory, labor, or providing poor service, while in meteorology we may need to balance the costs of implementing protective measures with the costs of potentially preventable weather damages.
-In this framework, a forecast has value if it leads to decisions with low total costs.
-In all of these fields, analyses have consistently found that common measures of statistical forecast accuracy do not necessarily correspond directly to measures of the value of forecasts as an input to decision making \cite[e.g.,][]{leitch1991economicForecastEval, murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}.
-
-However, we are aware of only a limited body of work that explicitly attempts to measure the value of infectious disease forecasts through their impact on policy, and much of this discussion has proceeded informally.
-For example, \cite{ioannidis2022forecastingCOVIDfailed} discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts.
-Separately, there is a thread of literature that does quantify the link between infectious disease modeling and policy making, but this work has been done outside of a forecasting context.
-As an example, \cite{Probert2016decisionMakingFootMouth} develop measures of the cost of actions designed to control a hypothetical outbreak of foot-and-mouth disease and use this framework to explore policy recommendations from a variety of simulation-based projection models.
-
-In practice, probabilistic infectious disease forecasts have most often been made for observations that emerge from public health surveillance systems and have typically been evaluated with standard, ``off-the-shelf'' scoring rules.
-For example, seasonal influenza forecasts in the US and dengue forecasts for Peru and Puerto Rico targeted public health surveillance measures of incidence over time and space, and used log-score and mean absolute errors to evaluate forecast skill \citep{mcgowan_collaborative_2019,reich_collaborative_2019,johansson_open_2019}.
-Pandemic COVID-19 forecasts of observed cases, hospitalizations and/or deaths in the US and Europe, as reported by municipal, state, or federal surveillance systems, were evaluated using the weighted interval score (WIS, which is an approximation of the continuous ranked probability score, or CRPS), and prediction interval coverage \citep{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.
-Similarly, CRPS was also used to assess probabilistic forecasts of dengue incidence at the district level in Vietnam \citep{colon-gonzalez_probabilistic_2021}.
-While some of these scores can be interpreted through the lens of decision theory, and all of the application-specific papers cited above had authors from public health agencies, none of them make explicit connections between forecast evaluation and how a forecast was used in practice.
-
-In this work, we begin to fill this gap between the ways that infectious disease forecasts have traditionally been evaluated and the ways that they have been used to support public health policy.
-We consider a setting in which forecasts are used to help determine the allocation of a limited quantity of medical supplies across multiple regions.
-We define a new forecast scoring rule --- the {\em allocation score} --- that evaluates forecasts based on how beneficial resource allocations derived from them would turn out to be.
-%Unlike the traditional measures of forecast skill such as WIS, CRPS, and log score, the proposed allocation score measures the suitability of a forecast as an input to this specific decision making task.
-
-Briefly, the allocation score of a forecast is the avoidable unmet need that results from using that forecast to set resource allocations by minimizing expected unmet need.
-For example, suppose that a decision maker is provided with forecasts of the level of need for medical resources in each of several states or hospital systems.
-If there is a limited amount of the medical resource that is available to distribute, a decision maker could choose an allocation of that resource across locations that minimizes the expected unmet need according to that forecast.
-As measured by the allocation score, one forecast is better than another if it would lead decision makers to an allocation that results in less unmet need.
-%By ``unnecessary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
-If the amount of resources that is available to distribute is less than the actual need, some amount of unmet need is unavoidable.
-The allocation score for a forecast does not include the unmet need that was unavoidable given the resource constraint, and so it measures only the amount of unmet need that could have been prevented by using a different allocation of available resources than that suggested by the forecast.
-We elaborate on these ideas in Section~\ref{sec:methods}.
-
-We present an illustrative analysis using the allocation score to evaluate forecasts of hospital admissions in the US leading up to the Omicron wave in winter 2022.
-This analysis is ``synthetic'' in that it does not correspond to an actual analysis that supported decision making in real-time.
-However, the framework described in this paper corresponds to real-world decisions that must be made by public health administrators around the globe, and could be adapted in the future for such real-time situations.
-For example, forecasts for districts in Sierra Leone of bed demand to care for patients with Ebola was the subject of a real-time modeling study in late 2014 and early 2015 \citep{camacho2015-ebola-bed}.
-And, in 2020, a model developed by an academic research group turned predictions of COVID-19 hospitalizations into estimates of ventilator usage and shortages.
-This framework was used by the Hartford HealthCare system in Connecticut ``to align ventilator supply with projected demand at a time where the [COVID-19] pandemic was on the rise'' \citep{bertsimas2021predictionsCOVID}.
-These examples illustrate the potential for forecasts to inform decisions about how to allocate limited supplies such as temporary hospital beds, ventilators, personal protective equipment, or other supplies that are known to be effective at reducing transmission or severity of disease.
-However, we emphasize again that these studies did not take the step of evaluating forecasts based on the quality of the allocation decisions that they supported or could have been used to support.
-
-The remainder of this article is organized as follows.
-We describe the allocation score in Section \ref{sec:methods}, and in Section \ref{sec:application} we illustrate the use of the score in an application to evaluate short-term forecasts of COVID-19 hospital admissions in the US.
-Section \ref{sec:discussion} summarizes our contributions and discusses opportunities for further extensions in future work.
-
-\section{The Allocation Score}
-\label{sec:methods}
-
-We begin with an informal description of the allocation score and some examples illustrating its key characteristics in section \ref{sec:methods.overview}.
-In section \ref{sec:methods.detailed} we develop the allocation score more carefully, building on decision theoretic procedures for deriving proper scoring rules.
-We then comment on some connections between the allocation score that we propose and other common scores that can be derived from decision theoretic foundations, such as the quantile score, WIS, and CRPS, in section \ref{sec:methods.related}.
-
-\subsection{Overview of Allocation Scoring}
-\label{sec:methods.overview}
-
-Suppose that a decision maker is tasked with determining how to allocate $K$ available units of a resource across $N$ locations.
-If the decision maker is provided with a multivariate forecast $F$ where each marginal forecast distribution $F_i$ predicts resource need in a particular location, one option is to choose the resource allocation that minimizes the expected total unmet need according to the forecast.
-We will give a more precise mathematical statement in section \ref{sec:methods.detailed}, but informally, the total expected unmet need according to the forecast is
-\begin{align}
-\sum_{i=1}^N \mathbb{E}_{F_i}[\text{unmet need in location $i$}], \label{eqn:informal_objective}
-\end{align}
-where the unmet need in a particular location is the difference between resource need in that location and the number of resources that were allocated there.
-This allocation problem has an intuitively appealing solution: allocate so that the probabilities of need exceeding allocation in various locations are as close to each other as possible.
-This will lead to the allocations provided by $F$ being quantiles of the marginal distributions $F_i$ for some \emph{single} probability level $\tau$ that is shared in common for all locations.
-
-After time passes and the actual level of resource need has been observed, the quality of a selected allocation can be measured by comparing the actual need in each location to the amount of resources that were sent there.
-Specifically, we compute the total unmet need that resulted from the selected allocation:
-\begin{align}
-    \sum_{i=1}^N \text{unmet need in location $i$}. \label{eqn:informal_loss}
-\end{align}
-We emphasize that in Equation \eqref{eqn:informal_loss} the calculation of unmet need is based on the actual resource need that was realized in each location, while in Equation \eqref{eqn:informal_objective} the calculation of unmet need was based on a hypothetical potential future level of resource need.
-Once the actual levels of resource need have been observed, we can obtain a quantitative measure of the quality of alternative allocation decisions: one allocation is better than another if it results in lower total unmet need.
-
-The \textbf{allocation score} of the forecast $F$ is the avoidable unmet need that results from using the allocation that minimizes the expected unmet need according to that forecast.
-By ``avoidable unmet need'', we mean that the allocation score does not include the amount of unmet need that was inevitable simply because the amount of available resources $K$ was less than the need for resources.
-Rather, the allocation score measures the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
-An allocation score of 0 is optimal, and indicates that no other allocation of resources could have met need better than the allocation suggested by $F$.
-A larger allocation score indicates that it would have been possible to improve upon the allocation suggested by $F$.
-
-\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$ and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$.
-When the marginal forecasts are exponential distributions, it can be shown that the optimal allocation divides the available resources among the locations proportionally to the scale parameters $\sigma_i$ (see section *** of the supplemental materials).
-If we have $K = 5$ units of our resource available, the optimal allocation according to $F$ would be 1 unit of resources in location 1 and 4 units of resources in location 2.
-If, on the other hand, we have $K = 10$ units available, we will allocate 2 units of resources to location 1 and 8 units to location 2.
-Figure~\ref{fig:exp_alloc_example} illustrates the situation.
-
-\begin{figure}
-    \includegraphics[width=\textwidth]{../figures/exponential_pred_expected_loss.pdf}
-    \caption{An illustration of the resource allocation problem in Example 1.
-There are $N = 2$ locations, with predictive distributions $F_1 = \mathrm{Exp}(1)$ and $F_2 = \mathrm{Exp}(1/4)$.
-The cdfs of these distributions are illustrated in the panels at bottom and right.
-In the center panel, the background shading corresponds to the expected loss according to these forecasts.
-Diagonal black lines indicate resource constraints at $K=5$ and $K=10$ units; any point along those lines corresponds to an allocation that meets the resource constraint.
-For these forecasts, the optimal allocations are $(1, 4)$ for $K=5$ and $(2, 8)$ for $K=10$.
-These allocations are at the point on the constraint line where the expected loss is smallest, which also corresponds to the point where a level set of the expected loss surface (blue curve) is tangent to the constraint.}
-    \label{fig:exp_alloc_example}
-\end{figure}
-
-Next suppose that we observe needs of 1 and 10 in locations 1 and 2, respectively.
-Based on these observed needs, we can measure the quality of the allocation suggested by the forecast by calculating the amount of unmet need that resulted from that allocation over and above what was unavoidable given the resource constraint.
-With $K = 5$ units of the resource, the allocation based on the forecast exactly meets the observed need in location 1, but it leaves 6 units of need unmet in location 2.
-However, working within the resource constraint, no other allocation could have done better: for example, allocating 0 units of resources to location 1 and 5 to location 2 still results in a total unmet need of 6 across both locations.
-Therefore, the forecast's allocation score is 0 with $K = 5$.
-On the other hand, when $K = 10$, the forecast $F$'s allocation results in $10 - 8 = 2$ units of unmet need in location 2 despite leaving no need unmet in location 1.
-In this case, the oracle would be able to prevent all but 1 of the total 11 units of need from going unmet, for example by allocating 1 unit of resources to location 1 and the remaining 9 units of resources to location 2.
-The allocation score for the forecast when $K = 10$ would therefore be 1 (= 2 realized $-$ 1 unavoidable) in units of avoidable unmet need.
-
-These scores illustrate a general result: allocation scores for a forecast will tend to be larger when the resource constraint is close to the observed need, because this is when it matters most which locations are allocated more or less resources.
-If the resource constraint is very small, any allocation of those limited resources will result in a large amount of unmet need.
-If the resource constraint is very large, it becomes less important which locations receive relatively more or less resources because all locations will receive enough resources to meet their need.
-
-\paragraph{Example 2} Now consider a different forecast that also has exponential distributions for resource need in each location, but that has the scale parameters $\sigma_1 = 2$ and $\sigma_2 = 8$, twice as large as the scale parameters of the forecast in Example 1.
-Because the optimal allocation is proportional to the scale parameters, this forecast would lead to the same optimal allocations as the forecast in Example 1, and would therefore be assigned the same allocation score.
-
-Note the way in which these forecasts incurred a positive (i.e., non-optimal) allocation score of 1 when $K = 10$.
-It was not directly due to individual misalignments of the marginal forecasts $F_i$ with the observed needs, but rather because the allocations and observed needs were not proportional as vectors.
-Restating: as far as allocation decisions are concerned, with a fixed constraint $K=10$ the fundamental problem with the forecast $F$ in Example 1 is not that it predicts a mean total resource need of $5$ units; it is that the realized need was 10 times as large in location 2 as in location 1, but the forecast only indicated that the resource allocation for location 2 should be 4 times the allocation for location 1.
-
-This illustrates a fundamental property of the allocation score: at its core, it measures whether the forecast accurately captures the relative magnitudes of resource need across different locations, which is precisely the information that is needed to allocate resources to those locations subject to a fixed resource constraint.
-On the other hand, the allocation score is not directly sensitive to whether the forecasts in each location correctly capture the magnitude of resource need in each individual location.
-%For a given probability level $\tau$, if the quantiles $Q_i(\tau)$ of the marginal forecasts are proportional to the observed needs $y_i$, then the allocation score is zero for the resource constraint level $K = \sum Q_i(\tau)$.
-This stands in marked contrast to other common scoring methods for multivariate forecasts that aggregate univariate scores such as log score, CRPS, or WIS for the marginal forecasts where a
-poor forecast made for one unit (a location, say) is penalized regardless of alignments in other units.
-Note that we do not claim that the allocation score is generically preferable to these other scores \textemdash rather, it provides a view of forecast performance that is specifically tuned to the context of decision making about resource allocations.
-
-%It is therefore often straighforward to constuct forecasts $F$ and $\widetilde{F}$ for a given outcome distribution that switch rankings under the allocation and traditional scores by ensuring that the marginal forecasts of $F$ center sharply around allocations that are proportionally similar to the central tendencies of the outcome distribution but are strongly biased.
-
-%While multivariate scoring rules have not seen wide application in infectious disease forecast evaluation, it seems that the allocation score would have a similar relationship with them, since while bias penalties in multi-variate scores can be offset by better forecasting of dependence structure, the proportional biases that allocation scoring are insensitive to are not \emph{per se} tolerated by multivariate scoring rules such as the energy score, variogram score, or Dawid-Sebastiani score.
-
-\subsection{A decision theoretic development of the allocation score}
-\label{sec:methods.detailed}
-
-We give a high-level review of a general procedure for developing proper scoring rules that are tailored to specific decision making tasks in section \ref{sec:methods.detailed.decisiontheory}, and then in section \ref{sec:methods.detailed.specific_allocation} we apply that procedure to develop the allocation score based on the task of deciding on how to allocate a fixed supply of resources across multiple locations.
-In \ref{sec:methods.detailed.integrated_allocation} we consider a small extension where the resource constraint is not known, or it is desired to consider the value of forecasts across a range of decision making scenarios.
-This gives rise to the \emph{integrated allocation score}.
-
-\subsubsection{The decision theoretic setup for forecast evaluation}
-\label{sec:methods.detailed.decisiontheory}
-
-%In this section, we give an overview of the decision theoretic setup for developing proper scoring rules that measure the value of a forecast as an input to decision making.
-We keep the discussion here at a somewhat informal level; we refer the reader to [some subset of Brehmer and Gneiting; Grünwald and Dawid; Dawid; Granger and Pesaran 2000; Granger and Machina 2006; Ehm et al.
-2016] for more technically precise discussion.
-
-In the framework of decision theory, a decision corresponds to the selection of an action $x$ from some set of possible actions $\mathcal{X}$.
-For example, $x$ may correspond to the level of investment in a measure designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with $\mathcal{X}$ being the set of all possible levels of investment that we might select.
-The quality of a decision to take a particular action $x$ is measured in relation to an outcome $y$ that is unknown at the time the decision is made.
-For example, $y$ may correspond to the number of individuals who eventually become sick and would benefit from the mitigation measure, and informally, an action $x$ is successful to the extent that it meets the realized need.
-In the face of uncertainty, a decision maker may use a forecast $F$ of the random variable $Y$ to help inform the selection of the action to take.
-We measure the value of a forecast as an input to this decision making process by the quality of the decisions that it leads to.
-
-We can formalize the preceding discussion with the following three-step procedure for developing scoring rules for probabilistic forecasts:
-\begin{enumerate}
-\item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$ eventually occurs.
-\item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under the distribution $F$.
-\item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: $S(F, y) = s(x^F, y)$.
-\end{enumerate}
-%We use the letter $s$ for the loss function to align with the literature on evaluation of forecasts of continuous outcomes, in which context we can often identify the action $x$ with a functional (i.e., a numeric summary such as a mean or a quantile) of the forecast distribution $F$.
-In this context, $s$ may be used as a \emph{scoring function}.
-\elr{Consider moving preceding sentences to a footnote or just deleting them?}
-This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss function.
-Subject to certain technical conditions, scoring rules obtained from this procedure are proper.
-We refer the reader to discussion in (cite paper 2 on arxiv) for a more technically precise discussion.
-
-% , and we illustrate the ideas with two running examples: (1) some problem that results in quantile elicitation, (2) some allocation problem.
-
-\subsubsection{The allocation score for a fixed resource constraint}
-\label{sec:methods.detailed.specific_allocation}
-
-In the decision making setting that we consider, an action $x = (x_1, \ldots, x_N)$ is a vector specifying the amount that is allocated to each of $N$ locations.
-We require that $0 \leq x$, i.e., that each $x_i$ is non-negative, and that the total allocation across all locations equals the amount of available resources, $K$: $\sum_{i=1}^N x_i = K$.
-The set $\mathcal{X}$ consists of all possible allocations that satisfy these constraints.
-The eventually realized resource need in each location is denoted by $y = (y_1, \ldots, y_N)$.
-These levels of need are not known at the time of decision making, so we define the random vector $Y = (Y_1, \ldots, Y_N)$ where $Y_i$ represents the as-yet-unknown level of resource need in location $i$.
-Forecasts of need in each location are collected in $F = (F_1, \ldots, F_N)$.
-We assume that resource need is non-negative and the forecasts reflect that, i.e.
-the support of each $F_i$ is a subset of $\mathbb{R}^+$.
-Finally, we assume that each unit of unmet need incurs a loss denoted by $L$, so that if the selected resource level $x_i$ in location $i$ is less than the realized need $y_i$, a loss of $L \cdot (y_i - x_i)$ results.
-
-We note that the problem formulation here assumes that the resource in question does not impact the amount of demand $y_i$ that will materialize, but rather it is a resource that satisfies that demand.
-In the context of infectious disease, this means that we do not consider resources that are intended to reduce the number of people who will become sick at some point in the future, such as a preventative influenza or COVID-19 vaccine.
-Our set up addresses resources like hospital beds, oxygen supply, ventilators, or rabies vaccines which are intended to meet the medical needs of patients who are already sick.
-
-With this notation in place, we can develop a proper scoring rule following the outline in section \ref{sec:methods.detailed.decisiontheory}.
-
-\paragraph{Step 1: specify a loss function.} The loss associated with a particular allocation is calculated by summing contributions from unmet need in each location:
-\begin{equation}
-s_A(x, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i). \label{eqn:loss_fn}
-\end{equation}
-Here, $\max(0, y_i - x_i)$ is the unmet need in location $i$, which is given by $y_i - x_i$ if the realized need $y_i$ in location $i$ is greater than the amount $x_i$ allocated to that location, or $0$ if the amount $x_i$ allocated to unit $i$ is greater than or equal to the realized need.
-Also, $L$ is a constant scalar value, the same across all locations, specifying the ``cost'' of one unit of unmet need.
-
-\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the forecast, $x^{F,K}$, is the allocation that minimizes the expected loss, that is, the solution of the \emph{allocation problem} associated with $K$
-\begin{align}
-    \underset{0 \leq x}{\mathrm{minimize}}\,\, \mathbb{E}_{F} [s_A(x, Y)] \text{ subject to }
-     \, \sum_{i=1}^N x_i = K, \label{AP}
-\end{align}
-where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]$ sums the expected loss due to unmet need across all locations.
-
-%%%% start of large excision %%%%%
-
-% In the supplement we derive a general form of this solution by beginning with the fact that a condition for minimizing
-% expected loss under the constraint is that there is no way to decrease expected loss further by shifting a small amount of the resource from one location to another.
-% As a consequence, at $x^{F,K}$, adding resources to any particular location $i$ reduces expected loss at the same rate $\lambda$:
-% \begin{align}
-% \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^{F,K}} = -\lambda.
-% \end{align}
-% It turns out that these derivatives can also be expressed in terms of the marginal forecasts $F_i$ as
-% \begin{align}
-%  \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] = L \cdot (F_i(x_i) - 1)
-% \end{align}
-% so that
-% \begin{align}
-% F(x_i^{F,K}) = 1-\lambda/L.
-% \end{align}
-% This tells us that in order to allocate optimally (according to $F$), we must divide resources among the locations so that there is an equal forecasted probability in every location that the allocation is sufficient to meet resource need.
-% It also allows us to write the components $x_i^{F,K}$ of the Bayes act as quantiles
-% \begin{align}
-% x_i^{F,K} = F_i^{-1}(1 - \lambda/L)
-% \end{align}
-% which can be substituted into the constraint equation of the allocation problem to get an equation for $\lambda$,
-% \begin{align}
-% \sum_{i=1}^N F_i^{-1}(1 - \lambda/L) = K.
-% \end{align}
-% The value of $\lambda$ that solves this equation---which almost always must be approximated with a numerical method---can be substituted back into the quantile formulas
-% to arrive at an explicit expression for the Bayes act.
-% This partial solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
-% We note that $\lambda$ and the probability level $1-\lambda/L$ depend on both the constraint level $K$ defining the decision problem and on the forecast $F$.
-This is quite different from the situation for quantile scores and the WIS, where the probability levels used to calculate scores are fixed.
-
-%%%% end of large excision %%%%%
-
-\paragraph{Step 3: Define the scoring rule.} We can now use the Bayes act to define a proper scoring rule for the probabilistic forecast $F$.
-Consider first the ``raw'' score defined as
-\begin{align}
-S_A^{\text{raw}}(F, y; K) = s_A(x^{F,K}, y) = \sum_{i=1}^N L \cdot \max(0, y_i - F_i^{-1}(1- \lambda(K,F)/L))
-\end{align}
-This measures the total unmet need across all locations that results from using the Bayes allocation associated with the forecast $F$ when the actual level of need in each location is observed to be $y_i$.
-But to make this a more easily interpreted measure of
-forecast performance, we will adjust it by the the minimum loss $l_{\text{oracle}}(y;K)$ achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$ which is $l_{\text{oracle}}(y;K) = L \cdot \max(0, \sum_{i=1}^{N}y_i - K)$.
-Our scoring rule is then given by
-\begin{align}
-S_A(F, y; K) &= S_A^{\text{raw}}(F, y; K) - l_{\text{oracle}}(y;K) \\
-& = \sum_{i=1}^N L  \cdot \max(0, y_i - x_i^{F,K}) - L \cdot  \max(0, \sum_{i=1}^{N}y_i - K) \\
-& =
-\begin{cases}
-\sum_{i=1}^n L  \cdot  \max(0, y_i - x_i^{F,K}), &  \sum_{i=1}^{N}y_i \leq K \\
-L \cdot K - L \cdot \sum_{i=1}^{N} \min(x_i^{F,K}, y_i), &  \sum_{i=1}^{N}y_i > K.
-\end{cases}
-\end{align}
-In the first case, the oracle incurs no loss so that the raw and adjusted scores coincide.
-The second case can be read as taking the $K$ resource units perfectly allocated by the oracle as a base penalty on
-the imperfect forecast $F$ and then, for each location $i$, reducing this penalty by however much of the need $y_i$
-is met with the Bayes act component $x_i^{F,K}$.
-The oracle adjustment aligns with a central theme in modern decision theory across disciplines including economics, computer science, and psychology which is that
-\emph{opportunity loss} (often known as \emph{regret} or (negative) \emph{relative utility}) is often a more important quantity than absolute loss from both a descriptive and normative perspective.
-
-\subsubsection{Integrating the allocation score across resource constraint levels}
-\label{sec:methods.detailed.integrated_allocation}
-
-The allocation score $S_A$ that we developed in the previous section measures the skill of the forecast distributions $F$ based on a single probability level $\tau$.
-This is appropriate if the resource constraint $K$ is a known constant.
-However, if $K$ is not precisely known at the time of decision making or there is interest in measuring the value of forecasts across a range of decision making scenarios with different resource constraints, we can use an \emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a distribution $p$:
-$$S_{IAS}(F, y) = \int S_A(F,y; K) p(K) \, dK$$
-
-\subsection{Generalizations and Connections to Other Scores}
-\label{sec:methods.related}
-
-We begin this section by briefly sketching how the weighted interval score (WIS), a commonly used proper scoring rule for probabilistic forecasts during the COVID-19 pandemic, can also be derived using the approach above.
-Then, we discuss similarities and differences between WIS and the allocation score, and other scores in general.
-
-\subsubsection{The quantile loss and weighted interval score (WIS)}
-\label{sec:methods.wis}
-
-The weighted interval score (WIS), is a scoring rule first defined in 2020 to assist with scoring public forecasts that were being made in the early stages of the COVID-19 pandemic.\citep{bracher2021evaluating}
-Many COVID forecasts were stored in a quantile format, where probabilistic distributions are encoded by a fixed set of quantiles.
-%% optional detail
-This was largely a choice of convenience (e.g., all forecasts had a known, fixed data size) and was codified by the US COVID-19 Forecast Hub and other collaborative forecasting projects.\citep{cramer_united_2022}
-%%
-While pointing a reader interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS, is as its name suggests, a weighted sum of interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...).
-Larger interval scores indicate less skillful forecasts.
-Interval scores themselves consist of (a) the width of the interval, with larger intervals receiving higher scores, and (b) a penalty if the interval does not cover the eventual observation, which increases the further away the interval is from the observed value.
-We note that the most commonly used version of WIS is one that chooses a specific set of weights so that WIS approximates the continuous rank probability score (CRPS) or pinball loss, a commonly used score for probabilistic forecasts.
-Other specifications of weights for the various interval scores can be used.
-\emph{It is important to note that the version of WIS that has been commonly used for evaluating COVID forecasts was proposed because it approximates a commonly used forecast scoring rule, the CRPS, and not because it aligned with any particular public health decision-making rationale.}
-
-The decision-making setup that leads to a choice of WIS as a scoring rule can be formulated using the same recipe as shown in section \ref{sec:methods.detailed.decisiontheory}.
-In fields such as meteorology and supply chain management, a great deal of attention has been given to the problem where a decision must be made about the quantity of a resource to purchase for a single location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each unit of unmet need.\ngr{I might suggest adding a detailed Step 1/2/3 as an appendix and referencing it here.
-But I think that adding that full detail here might be too much.} This decision-making problem leads to the pinball loss that is often used in quantile regression, and thence to the CRPS (or WIS, via approximation) as a scoring rule.\ngr{can we make more explicit the link here to the section above?
-I'm not confident in my ability to describe with technical accuracy the link between quantile/pinball loss and CRPS, but I think this is a good opportunity to do so briefly.}
-
-\subsubsection{Connections between scores}
-\label{sec:methods.generalization}
-
-We have described in previous sections how both the allocation score and the weighted interval score (WIS) arise from a single standard procedure for developing proper scoring rules.
-The allocation score arises when the decision relates to how a fixed quantity of resources should be allocated to multiple locations.
-The WIS and CRPS arise when the decision relates to how much of a resource to order in the context of the cost of the resource and wanting to minimize unmet need in one or more locations.
-In both settings, the optimal solution sets the resource level in each location to a quantile of the respective forecast distributions.
-
-These decision making problems differ according to the challenge faced by the decision maker: a fixed constraint on the available resources for the allocation problem, or a cost per unit of resources in the resource purchasing problem.
-In fact, it is possible to combine these into a more general problem where the decision maker must decide on both the total level of resources to purchase and an allocation of those resources across multiple locations, subject to a cost per unit of resources and a constraint on the total quantity of resources that can be purchased.
-The quantile loss and the allocation score presented in this work are both special cases of the score that emerges from that more general decision making problem that integrates across resource cost and constraint parameters together.
-We pursue this direction further in other work (cite arxiv version of paper 2).
-
-\section{Evaluating forecasts of COVID hospitalizations using the allocation score}
-\label{sec:application}
-
-We illustrate with an application to hospital admissions in the U.S., considering the problem of allocation of a limited supply of medical resources to states.
-
-\begin{figure}
-    \includegraphics[width=\textwidth]{../figures/allocation-forecasts-v2.pdf}
-    \caption{Hospitalization data and example forecasts for Virginia, US.
-New daily hospital admissions are shown by the line of data.
-The dark line indicates data up to the time the forecast was made on December 27, 2021; the grey line indicates data available in retrospect.
-The set of colored bars represent forecasts made for 14 days into the future, with a target date of January 10, 2022 (dashed vertical line).
-The horizontal dashed line indicates the level of the observed value on this day.
-Note that all three forecasts correspond to forecasts for the same day, but they are spread out for better visibility.
-}
-    \label{fig:sample-forecasts}
-\end{figure}
-
-\subsection{Data}
-<<get-data-dates>>=
-metadata <- tar_meta()
-@
-
-\subsubsection{Hospitalization data}
-Starting in the summer of 2020, the US Health and Human Services began reporting counts of daily new admissions to hospitals for individuals with COVID-19.\citep{healthdatagov_covid-19_nodate}
-These data served as the source of ``ground truth'' data for the US COVID-19 Forecast Hub which, starting in December 2020, collected short-term forecasts of new hospital admissions at the daily scale.
-These daily counts were available for the US as a whole, and all states and several additional jurisdictions such as Puerto Rico and Washington DC.
-The data were updated daily and were available for download by the public through the HHS HealthData.gov website.
-For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data.\cite{wang-covidhubutils}
-The hospitalization data were downloaded for analysis on \Sexpr{as.Date(metadata$time[which(metadata$name == "truth_data")])}.
-
-\subsubsection{Forecast data}
-The US COVID-19 Forecast Hub, a consortium funded by the US CDC and led by a research group at UMass-Amherst, collected short-term forecasts of hospitalizations starting in December 2020.\citep{cramer_united_2022}
-Any team that with appropriately formatted forecasts could submit them to the Forecast Hub data repository on GitHub.(cite repo site)
-Forecasts were time-stamped by GitHub upon submission and passed validation checks that ensured correct formatting and that the forecasts were being submitted only for dates in the future, not for data that had already been observed.
-
-Forecast submission followed a weekly cycle and culminated in the creation of an ensemble forecast.
-Forecasts could be submitted on any day during the week.
-However once a week on Mondays, the Forecast Hub would collect the most recent forecasts submitted by all teams that met certain inclusion criteria and create an ensemble forecast using quantile averaging.\citep{ray_comparing_2023}
-An ensemble that treated all models equally was created (COVIDhub-ensemble) as was a model that created weights of submitted models based on performance in the past 12 weeks (COVIDhub-trained\_ensemble).
-One other model that combined multiple forecasts from different teams but used a different ensembling algorithm, a linear pooling method with tail extrapolation, was also included in our analyses (JHUAPL-SLPHospEns).
-Several other models have ``ensemble'' in their name, but this refers to combinations of different variations of models that the specific team created, not to a multi-model ensemble combining different submitted forecasts to the Forecast Hub.
-
-All forecasts, including the ensemble, were submitted as probabilistic predictions about the number of new hospital admissions on a particular day in the future, in a specific jurisdiction of the US (national level, state, or territory).
-Probability distributions were specified, per the requirements established by the Forecast Hub, as a set of 23 quantiles for each individual prediction.
-The submitted quantiles included a median (treated as a ``point'' prediction) and defined 11 central prediction intervals, from a 99\% to a 10\% prediction interval.
-
-The analysis in this work focuses on forecasts made before and during the first wave of the Omicron SARS-CoV-2 variant in the US.
-As such, we downloaded forecasts for the 15 weeks starting with Monday 2021-11-22 through Monday 2022-02-28.
-Forecast data were downloaded on for analysis on \Sexpr{as.Date(metadata$time[which(metadata$name == "forecast_data")])}.
-
-We established a set of inclusion criteria to determine which forecasts and models to include in our analysis.
-Models were eligible to be included in the analysis if they were considered a ``primary'' model from a team.
-(If a team submitted multiple versions of similar models, they were required to designate one as ``primary''.)
-For a model to have a complete, eligible submission in a given week, it had to have a 14 day-ahead forecast for all 50 states plus Washington DC.
-Models had to have a complete forecast for at least 4 of the 15 weeks in the analysis to be considered eligible for inclusion.
-
-\subsection{Evaluation metrics}
-
-This manuscript focuses on two proper forecast scores, the allocation score and the weighted interval score (WIS), both defined above.
-
-\subsubsection{Allocation Score}
-
-For this analysis, we fixed a resource constraint $K$ to be 15,000, based roughly on a reported number of ventilators available for reallocation in the US.\citep{ajao_assessing_2015}
-For each week, we computed a the allocation score for each 14 day-ahead forecast based on $K=15,000$.
-
-We also computed a standardized rank for the allocation score for each model $m$ and week $w$.
-First, we computed the number of models that forecasted that week ($n_w$) and the rank of model $m$ among the $n_w$ models ($r_{m,w}^{AS}$).
-The model with the best allocation score received a rank of 1 and the worst received a rank of $n_w$.
-In the case of a tie between one or more models, all models received the better rank.
-We then rescaled these rankings to compute the allocation score standardized rank ($sr_{m,w}^AS$) between 0 and 1, where 0 corresponds to the worst rank and 1 to the best.
-
-\begin{equation}
-sr_{m,w}^{AS} = 1 - \frac{r^{AS}_{m,w}-1}{n_w-1}
-\end{equation}
-
-\subsubsection{Weighted Interval Score (WIS)}
-
-The Weighted Interval Score (WIS), as described in earlier sections, measures the alignment of a single probabilistic forecast ($F$) with an observation ($y$).
-We computed the mean WIS across all $L$ locations for each model and each forecasted week as
-\begin{equation}
-MWIS_{m,w} = \frac{1}{L}\sum_{l=1}^L WIS(F_{l,m,w},y)
-\end{equation}
-where $F_{l,m,w}$ is the probabilistic forecast from model $m$ for location $l$ and week $w$.
-Using the same procedure as for allocation scores described above, we computed standardized ranks for MWIS ($sr_{m,w}^{MWIS}$).
-
-\subsection{Data and code availability}
-
-\subsection{Application results}
-
-<<score-computation>>=
-tar_load("Kat15k_alloscores")
-
-alloscore_data_summaries <- Kat15k_alloscores |>
-  group_by(model, reference_date) |>
-  summarize(alloscore = sum(components), yobs = sum(y)) |>
-  group_by(reference_date) |>
-  mutate(nmodels = n(),
-         allo_rank = rank(alloscore, ties.method = "min"),
-         allo_rank_std = 1- (allo_rank-1)/(nmodels-1)) |>
-  ungroup() |>
-  complete(model, reference_date) |>
-  mutate(target_end_date = as.Date(reference_date) + 14)
-
-tar_load("score_data")
-
-alloscore_models <- unique(alloscore_data_summaries$model)
-score_data_summaries <- score_data |>
-  mutate(reference_date = target_end_date - 14) |>
-  filter(model %in% alloscore_models) |>
-  group_by(model, reference_date) |>
-  summarize(mwis = mean(wis)) |>
-  ungroup() |>
-  group_by(reference_date) |>
-  mutate(nmodels = n(),
-         mwis_rank = rank(mwis, ties.method = "min"),
-         mwis_rank_std = 1- (mwis_rank-1)/(nmodels-1)) |>
-  ungroup() |>
-  complete(model, reference_date) |>
-  mutate(target_end_date = as.Date(reference_date) + 14)
-@
-
-\subsubsection{Forecast scores showed differences over time}
-<<grob-hosps-over-time>>=
-tar_load("truth_data")
-dat_to_plot <- truth_data |>
-  filter(location == "US",
-         target_end_date %in% score_data_summaries$target_end_date)
-p1 <- ggplot(dat_to_plot,
-             aes(x=target_end_date, y=value)) +
-  geom_line() +
-  geom_point() +
-  geom_hline(yintercept=15000, linetype=2) +
-  ylab("hospital admissions") +
-  xlab(NULL) +
-  ylim(c(0, NA))
-@
-
-<<metrics-over-time, fig.height=9, fig.width=7, fig.cap="Hospital admissions and evaluation metrics over time.
-Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022.
-These are the days for which forecasts were made and evaluated.
-A horizontal dashed line at 15,000 shows the assumed resource constraint $K$.
-Panel B shows allocation scores for each model's 14-day ahead forecast, across all US states.
-The x-axis corresponds to the date for which the prediction was made.
-Allocation scores typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down).
-Panel C shows the MWIS metric across weeks, averaged across all states.
-Similarly to panel B, the x-axis corresponds to the date for which the prediction was made.
-MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the pean of the Omicron wave.">>=
-p2 <- ggplot(alloscore_data_summaries,
-             aes(x=target_end_date, y=alloscore, color=model, shape=model)) +
-  scale_x_date(limits=range(score_data_summaries$target_end_date)) +
-  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
-  ylab("Allocation score") +
-  geom_point() + geom_line()+
-  theme(legend.pos="none") +
-  xlab(NULL)
-
-p4 <- ggplot(score_data_summaries,
-             aes(x=target_end_date, y=mwis, color=model,shape=model)) +
-  geom_point() + geom_line() +
-  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
-  ylab("MWIS")+
-  theme(legend.pos="bottom")+
-  xlab(NULL)
-
-cowplot::plot_grid(p1, p2, p4,
-                   nrow=3,
-                   align="v",
-                   rel_heights = c(1, 1, 1.5),
-                   labels = "AUTO")
-@
-
-Allocation scores varied substantially by date and by model (Figure \ref{fig:metrics-over-time}).
-For predictions made for the first three Mondays in December 2021 and the last three Mondays in February 2022 all models had allocation scores under 500 (and the mean across all models was less than 100), indicating that the unnecessary unmet need was fairly low relative to the total number of hospital admissions on those days.
-The allocation scores are on the whole highest when the observed number of new hospital admissions is closest to the resource threshold of 15,000, as this is the time when any mistakes in allocation are costly in terms of wasting resources in one location that could have been used in another.
-Predictions made during the peak week and just after showed the highest variation in allocation scores, with some models having allocation scores under 1000 and others having values over 3500.
-
-Overall, the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, had the highest rank for allocation score in six weeks, more than any other model except the \texttt{COVIDhub-ensemble} which also had six.
-
-<<eval=FALSE>>=
-## some analysis to support above statements
-alloscore_data_summaries |>
-  group_by(target_end_date) |>
-  summarize(meanAS = mean(alloscore, na.rm=TRUE),
-            medianAS = median(alloscore, na.rm=TRUE),
-            minAS = min(alloscore, na.rm=TRUE),
-            maxAS = max(alloscore, na.rm=TRUE),
-            range = maxAS-minAS)
-@
-
-Mean weighted interval scores (MWIS) also varied by date and model, and more clearly were dependent on the scale of the observed data.
-MWIS values were low (all models under 100) for all Mondays in December 2021 and the final four Mondays evaluated.
-Across all models both the average and median MWIS value for every Monday in January was above 100, with the largest errors occurring one and two weeks after the peak was observed.
-
-<<eval=FALSE>>=
-## some analysis to support above statements
-score_data_summaries |>
-  group_by(target_end_date) |>
-  summarize(meanWIS = mean(mwis, na.rm=TRUE),
-            medianWIS = median(mwis, na.rm=TRUE),
-            minWIS = min(mwis, na.rm=TRUE),
-            maxWIS = max(mwis, na.rm=TRUE),
-            range = maxWIS-minWIS)
-@
-
-\subsubsection{Model rankings for different scores did not reliably correlate}
-
-<<metrics-correlation, fig.width=10, fig.height=7.5, fig.cap="Association of standardized ranks for MWIS and allocation score by model and week.
-Each facet of the plot corresponds to one model.
-Within each facet, each point corresponds to a week.
-The x- and y-values correspond to the MWIS standardized rank and the allocation score standardized rank for that week.
-Points corresponding to earlier dates have darker shading.
-The size of the point corresponds to the observed value on the date for which the prediction was made.
-Models show different degrees of association between the two metrics.">>=
-all_data <- left_join(alloscore_data_summaries, score_data_summaries,
-                      by=c("model", "target_end_date"))
-
-ggplot(all_data,
-       aes(x=mwis_rank_std, y=allo_rank_std, color=target_end_date, size=yobs)) +
-  geom_point(alpha=.7) +
-  scale_x_continuous("MWIS standardized rank", limits=c(0,1)) +
-  scale_y_continuous("Allocation score standardized rank", limits=c(0,1)) +
-  facet_wrap(.~model) +
-  geom_abline(slope=1, intercept=0, linetype=2, color="gray")
-
-@
-
-Models showed differing levels of correlation between their allocation scores and MWIS values (Figure \ref{fig:metrics-correlation}).
-Here are some examples of the different model-specific patterns observed:
-\begin{itemize}
-\item Several models showed a positive association between allocation score and MWIS ranks (e.g., \texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha}).
-\item Several models had consistently strong MWIS ranks but also had highly variable allocation score ranks with no clear association between the two (e.g., \texttt{JHUAPL-SLPHospEns} and \texttt{CU-select})
-\item One model performed consistently well for both metrics with no clear association (\texttt{COVIDhub-ensemble})
-\item One model performed consistently well for allocation score but had only middlings ranks for MWIS (\texttt{CMU-TimeSeries})
 \end{itemize}
 
-<<eval=FALSE>>=
-alloscore_data_summaries |>
-  group_by(model) |>
-  summarize(top_rank = sum(allo_rank_std==1))
-@
-
-\subsubsection{Integrated allocation score across values of K}
-
-Using the integrated allocation score summarizes allocation scores across a range of possible values of the constraint ($K$) (Section \ref{sec:methods.detailed.integrated_allocation}).
-This could be useful in situations where the actual constraint may not be known, or could only be specified as a probability distribution.
-
- - something here from Ben's analysis?
-
-\ngr{include figure here that shows relative rankings of locations?}
-
-\section{Discussion}
-\label{sec:discussion}
-
-All proper scoring rules for probabilistic forecasts have an explicit link to a loss function, thereby placing numerical value on different kinds of errors in specifying forecasts.
-In epidemiological forecasting, well-known proper scoring rules such as the log score or variations on the continuous rank probability score (such as the weighted interval score, or WIS) have been frequently utilized to evaluate probabilistic forecasts.
-Often, these scores are used to rank models according to accuracy with that particular score, but without reference to the underlying decision-making process for which that score was designed.
-With careful thought and collaboration between modelers and public health officials, we argue that scores that are more aligned with public health decisions could be developed to inform specific problems.
-We have demonstrated that forecast evaluation methods that are tied to a specific decision making context can yield model rankings that are substantively different from standard measures of forecast skill.
-
-We often conceive of infectious disease forecasts as being useful for decision making purposes, but it is rare for forecast evaluation to be tied directly to the value of the forecasts for informing those decisions.
-This work seeks to address that gap.
-However, we do note that the decision-making context presented in this work, while motivated by examples from real-world public health resource allocation problems, has not used to inform an actual real-time decision-making process.
-
-Resource allocation decisions are a realistic example of the kinds of decisions that could motivate more targeted forecasting exercises.
-One real-world example is allocation of ventilators during a respiratory viral pandemic.
-This was the motivating example for the examples presented in the work above.
-However, other examples where allocation is a central concern exist as well, such as the allocation of a limited stockpile of vaccinations\citep{persad_fair_2023} or diagnostic tests\citep{du_optimal_2022}.
-
-In practice, there are many potential users of forecasts with many different decision making problems.
-Not all can be easily quantified.
-Those that can be easily quantified may differ enough that it seems likely that no single score would be appropriate for all users.
-In an ideal setting, a forecasting tool could be developed through close collaboration between modelers and public health officials.
-However, this setting may only be possible in settings with enough resources to support both an analytics and public heatlh team.
-Increasingly, collaborative modeling hubs are being used to generate ``one-size-fits-all'' forecasts for many locations at once.
-In these settings, where tailored models are not available, it still could be possible to evaluate contributed models against a set of multiple scores.
-% NGR: leaving this out for now as it feels like too much detail?
-% This may be tricky to operationalize in the setting of a general forecast hub. It matters how you elicit and represent probabilistic forecasts (quantiles? samples? cdfs?).
-
-In our specific application, we made several key observations that we believe should inform future work in this area.
-First, it is clear that different metrics (in our case allocation score and mean weighted interval score) captured different aspects of forecast performance.
-The MWIS was strongly dependent on the scale of the forecasted quantity (e.g., when hospitalizations were high, so was WIS).
-The allocation score was not as scale-dependent, as the highest/worst allocation scores were observed when the number of hospitalizations was closest to the allocation constraint.
-We also note that contributed models found it hard to have a better allocation score than a na\"ive baseline model that just predicted a flat line from the last previous observation with wide uncertainty.
-This suggests suggests that contributed models (other than the ensemble, which combined forecasts from all contributed models) were not consistently adding value over just extrapolating from the current levels.
-
-There are several important limitations to the current work.
-The allocation score we developed here does not directly account for important considerations such as fairness or equity of allocations.
-Nor does the proposed framework attempt to capture the broader context of decision making.
-For example, in practice it may be possible to increase the resource constraint $K$ by shifting funding from other disease mitigation measures.
-We also note that in some settings, a ``successful'' epidemiological forecast may lead to policy decisions that change the distribution of the predicted outcome $Y$.
-Our framework cannot be used to evaluate a forecast at a horizon that may have been impacted by such a decision.
-
-We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work in more closely coupling applied epidemiological forecasting with public health decision-making.
-We note that others have explored similar linkages.
-For example, \cite{bilinski_adaptive_2023} explore ways in which policymaker preferences could inform risk thresholds for predictive models.
-\ngr{Would be nice to have a few other examples here.}
-In some situations, individual or ensemble models could be developed to minimize the loss for a particular setting, although this is also an area largely unexplored in the realm of public health to date, with some initial methodological development in econometrics.\citep{loaiza-maya_focused_2021}
-
-In conclusion, we argue that the way modelers and policymakers view and evaluate forecasts should change depending on the specific decision-making context.
-Using standard forecast evaluation metrics can mask the utility of certain forecasts, or lead to forecasts being used to inform a particular decision that are not optimal for that particular set of values that are important for that context.
-Further collaborative work is needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or simulation exercises that could be used to inform further development of new or alternative scoring metrics.
-
-\bibliography{allocation}
-
 \end{document}

### 2023-11-06 alloscore_manuscript/alloscore-application.tex

--- 2023-10-20 alloscore_manuscript/alloscore-application.Rnw
@@
 
-<<setup, tidy=FALSE, echo=FALSE, message=FALSE>>=
-knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
-@
-
-<<environment-setup>>=
-library(targets)
-library(tidyverse)
-library(ggbump)
-
-theme_set(theme_bw())
-
-## necessary to have project root as wd, to override document directory
-knitr::opts_knit$set(root.dir = '../')
-@
-
 \begin{abstract}
@@
 \cite{bertsimas2021predictionsCOVID} developed tools to inform decision making from infectious disease forecasts, which have been used to inform allocation of limited medical supplies such as ventilators, ICU capacity planning, and vaccine distribution strategy.
-Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping (University of Texas, 2022) \nocite{utnews2022}.
+Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping (University of Texas, 2022)\nocite{utnews2022}.
 In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
-The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid} or weighted interval score (WIS) \citep{bracher2021evaluating}, however, recently authors have been calling for evaluating forecasts through their impact on policy \citep{marshall2023predictions}.
+The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid} or weighted interval score (WIS) \citep{bracher2021evaluating}.
+However, recently authors have been calling for evaluating forecasts through their impact on policy \citep{marshall2023predictions, bilinski_adaptive_2023}.
 
 In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to.
-We give an overview of these procedures in Section \ref{sec:methods.decisiontheory}.
-\ngr{This section label doesn't exist. Should it point to methods.detailed?}
+We give an overview of these procedures in Section \ref{sec:methods.detailed.decisiontheory}.
 There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
-We review this work only briefly here, and we refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \elr{TODO: identify relevant review or book style summary for meteorology} for discussions focused on applications to economics and meteorology, respectively.
+We review this work only briefly here, and we refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \cite{murphy1993whatisagoodforecast} for discussions focused on applications to economics and meteorology, respectively.
 In finance, the value of forecasts can often be measured by the profits generated by trading decisions informed by the forecasts, perhaps adjusted for risk levels \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
-In applications to supply chain management and meteorology, the value of forecasts has typically been operationalized by considering the costs associated with decisions regarding the amount of inventory to hold or the level of protection against the impacts of extreme weather events to enact {Peter Catt (2007), Fotios Petropoulos and colleagues (2019), Nada Sanders and Gregory Graman (2009), T.N.
-Palmer (2002), Florian Pappenberger and colleagues (2015)}.
+In applications to supply chain management and meteorology, the value of forecasts has typically been operationalized by considering the costs associated with decisions regarding the amount of inventory to hold or the level of protection against the impacts of extreme weather events to enact \cite[e.g.,][]{catt2007assessingcostofforecasterror, petropoulos2019inventoryperformanceforecasting, palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
 For example, in supply chain management these decisions may incur costs related to holding inventory, labor, or providing poor service, while in meteorology we may need to balance the costs of implementing protective measures with the costs of potentially preventable weather damages.
 In this framework, a forecast has value if it leads to decisions with low total costs.
-In all of these fields, analyses have consistently found that common measures of statistical forecast accuracy do not necessarily correspond directly to measures of the value of forecasts as an input to decision making \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
+In all of these fields, analyses have consistently found that common measures of statistical forecast accuracy do not necessarily correspond directly to measures of the value of forecasts as an input to decision making \cite[e.g.,][]{leitch1991economicForecastEval, murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}.
 
@@
 \subsection{Data}
+
 \subsubsection{Hospitalization data}
-Starting in the summer of 2020, the US Health and Human Services began reporting counts of daily new admissions to hospitals for individuals with COVID-19.(cite HHS Protect)
+Starting in the summer of 2020, the US Health and Human Services began reporting counts of daily new admissions to hospitals for individuals with COVID-19.\citep{healthdatagov_covid-19_nodate}
 These data served as the source of ``ground truth'' data for the US COVID-19 Forecast Hub which, starting in December 2020, collected short-term forecasts of new hospital admissions at the daily scale.
@@
 The data were updated daily and were available for download by the public through the HHS HealthData.gov website.
-For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data.(cite package) \ngr{can we use tar\_manifest() to reproducibly display the date the data were stored?}
+For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data.\cite{wang-covidhubutils}
+The hospitalization data were downloaded for analysis on 2023-10-29.
 
@@
 As such, we downloaded forecasts for the 15 weeks starting with Monday 2021-11-22 through Monday 2022-02-28.
+Forecast data were downloaded on for analysis on 2023-10-06.
 
@@
 
-<<score-computation>>=
-tar_load("all_alloscore_data")
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-alloscore_data_summaries <- all_alloscore_data |>
-  filter(K == 15000)|>
-  group_by(model, reference_date) |>
-  summarize(alloscore = sum(components), yobs = sum(y)) |>
-  group_by(reference_date) |>
-  mutate(nmodels = n(),
-         allo_rank = rank(alloscore, ties.method = "min"),
-         allo_rank_std = 1- (allo_rank-1)/(nmodels-1)) |>
-  ungroup() |>
-  complete(model, reference_date) |>
-  mutate(target_end_date = as.Date(reference_date) + 14)
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in gzfile(file, "{}rb"{}): cannot open the connection}}
 
-tar_load("score_data")
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in group\_by(Kat15k\_alloscores, model, reference\_date): object 'Kat15k\_alloscores' not found}}
 
-alloscore_models <- unique(alloscore_data_summaries$model)
-score_data_summaries <- score_data |>
-  mutate(reference_date = target_end_date - 14) |>
-  filter(model %in% alloscore_models) |>
-  group_by(model, reference_date) |>
-  summarize(mwis = mean(wis)) |>
-  ungroup() |>
-  group_by(reference_date) |>
-  mutate(nmodels = n(),
-         mwis_rank = rank(mwis, ties.method = "min"),
-         mwis_rank_std = 1- (mwis_rank-1)/(nmodels-1)) |>
-  ungroup() |>
-  complete(model, reference_date) |>
-  mutate(target_end_date = as.Date(reference_date) + 14)
-@
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in unique(alloscore\_data\_summaries\$model): object 'alloscore\_data\_summaries' not found}}
 
-\subsubsection{Metrics as a function of time}
-<<grob-hosps-over-time>>=
-tar_load("truth_data")
-dat_to_plot <- truth_data |>
-  filter(location == "US",
-         target_end_date %in% score_data_summaries$target_end_date)
-p1 <- ggplot(dat_to_plot,
-             aes(x=target_end_date, y=value)) +
-  geom_line() +
-  geom_point() +
-  geom_hline(yintercept=15000, linetype=2) +
-  ylab("hospital admissions") +
-  xlab(NULL) +
-  ylim(c(0, NA))
-@
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in `filter()`:\\\#\# i In argument: `model \%in\% alloscore\_models`.\\\#\# Caused by error in `model \%in\% alloscore\_models`:\\\#\# !
+object 'alloscore\_models' not found}}\end{kframe}
+\end{knitrout}
 
-<<metrics-over-time, fig.height=9, fig.width=7, fig.cap="Hospital admissions and evaluation metrics over time.
-Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022.
-These are the days for which forecasts were made and evaluated.
-A horizontal dashed line at 15,000 shows the assumed resource constraint $K$.
-Panel B shows allocation scores for each model's 14-day ahead forecast, across all US states.
-The x-axis corresponds to the date for which the prediction was made.
-Allocation scores typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down).
-Panel C shows the MWIS metric across weeks, averaged across all states.
-Similarly to panel B, the x-axis corresponds to the date for which the prediction was made.
-MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the pean of the Omicron wave.">>=
-p2 <- ggplot(alloscore_data_summaries,
-             aes(x=target_end_date, y=alloscore, color=model, shape=model)) +
-  scale_x_date(limits=range(score_data_summaries$target_end_date)) +
-  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
-  ylab("Allocation score") +
-  geom_point() + geom_line()+
-  theme(legend.pos="none") +
-  xlab(NULL)
+\subsubsection{Forecast scores showed differences over time}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-p4 <- ggplot(score_data_summaries,
-             aes(x=target_end_date, y=mwis, color=model,shape=model)) +
-  geom_point() + geom_line() +
-  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
-  ylab("MWIS")+
-  theme(legend.pos="bottom")+
-  xlab(NULL)
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in `filter()`:\\\#\# i In argument: `target\_end\_date \%in\%\\\#\# \ \ score\_data\_summaries\$target\_end\_date`.\\\#\# Caused by error in `target\_end\_date \%in\% score\_data\_summaries\$target\_end\_date`:\\\#\# !
+object 'score\_data\_summaries' not found}}
 
-cowplot::plot_grid(p1, p2, p4,
-                   nrow=3,
-                   align="v",
-                   rel_heights = c(1, 1, 1.5),
-                   labels = "AUTO")
-@
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ggplot(dat\_to\_plot, aes(x = target\_end\_date, y = value)): object 'dat\_to\_plot' not found}}\end{kframe}
+\end{knitrout}
 
-Allocation scores varied substantially by date and by model (Figure XX).
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ggplot(alloscore\_data\_summaries, aes(x = target\_end\_date, y = alloscore, : object 'alloscore\_data\_summaries' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ggplot(score\_data\_summaries, aes(x = target\_end\_date, y = mwis, : object 'score\_data\_summaries' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in cowplot::plot\_grid(p1, p2, p4, nrow = 3, align = "{}v"{}, rel\_heights = c(1, : object 'p1' not found}}\end{kframe}
+\end{knitrout}
+
+Allocation scores varied substantially by date and by model (Figure \ref{fig:metrics-over-time}).
 For predictions made for the first three Mondays in December 2021 and the last three Mondays in February 2022 all models had allocation scores under 500 (and the mean across all models was less than 100), indicating that the unnecessary unmet need was fairly low relative to the total number of hospital admissions on those days.
 The allocation scores are on the whole highest when the observed number of new hospital admissions is closest to the resource threshold of 15,000, as this is the time when any mistakes in allocation are costly in terms of wasting resources in one location that could have been used in another.
-Predictions made during the peak week and just after showed the highest variation in allocation scores, with some models having values under 1000 and others having values over 3500.
-<<eval=FALSE>>=
-## some analysis to support above statements
-alloscore_data_summaries |>
-  group_by(target_end_date) |>
-  summarize(meanAS = mean(alloscore, na.rm=TRUE),
-            medianAS = median(alloscore, na.rm=TRUE),
-            minAS = min(alloscore, na.rm=TRUE),
-            maxAS = max(alloscore, na.rm=TRUE),
-            range = maxAS-minAS)
-@
+Predictions made during the peak week and just after showed the highest variation in allocation scores, with some models having allocation scores under 1000 and others having values over 3500.
+
+Overall, across the first 11 weeks evaluated (we excluded the last four since nearly all the models achieved an allocation score of zero), the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, had the highest rank for allocation score in four weeks, more than any other model except the \texttt{COVIDhub-ensemble} which also had four.
 
@@
 
-<<eval=FALSE>>=
-## some analysis to support above statements
-score_data_summaries |>
-  group_by(target_end_date) |>
-  summarize(meanWIS = mean(mwis, na.rm=TRUE),
-            medianWIS = median(mwis, na.rm=TRUE),
-            minWIS = min(mwis, na.rm=TRUE),
-            maxWIS = max(mwis, na.rm=TRUE),
-            range = maxWIS-minWIS)
-@
+\subsubsection{Model rankings for different scores did not reliably correlate}
 
-\subsubsection{correlation between WIS and allocation score}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-<<metrics-correlation, fig.width=7, fig.height=6, fig.cap="Association of standardized ranks for MWIS and allocation score by model and week.
-Each facet of the plot corresponds to one model.
-Within each facet, each point corresponds to a week.
-The x- and y-values correspond to the MWIS standardized rank and the allocation score standardized rank for that week.
-Points corresponding to earlier dates have darker shading.
-The size of the point corresponds to the observed value on the date for which the prediction was made.
-Models show different degrees of association between the two metrics.">>=
-all_data <- left_join(alloscore_data_summaries, score_data_summaries,
-                      by=c("model", "target_end_date"))
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in left\_join(alloscore\_data\_summaries, score\_data\_summaries, by = c("{}model"{}, : object 'alloscore\_data\_summaries' not found}}
 
-ggplot(all_data,
-       aes(x=mwis_rank_std, y=allo_rank_std, color=target_end_date, size=yobs)) +
-  geom_point(alpha=.7) +
-  scale_x_continuous("MWIS standardized rank", limits=c(0,1)) +
-  scale_y_continuous("Allocation score standardized rank", limits=c(0,1)) +
-  facet_wrap(.~model) +
-  geom_abline(slope=1, intercept=0, linetype=2, color="gray")
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in filter(all\_data, as.Date(reference\_date.x) < as.Date("{}2022-02-01"{})): object 'all\_data' not found}}\end{kframe}
+\end{knitrout}
 
-@
-
-Models show differing levels of correlation between their allocation scores and MWIS values (Figure XX).
-Here are some examples of the different patterns observed:
+Models showed differing levels of correlation between their allocation scores and MWIS values (Figure \ref{fig:metrics-correlation}).
+Here are some examples of the different model-specific patterns observed:
 \begin{itemize}
-\item positive association between allocation score and MWIS ranks: \texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha} models
-\item consistently strong MWIS ranks, highly variable allocation score ranks, no clear association: \texttt{JHUAPL-SLPHospEns} and \texttt{CU-select}
-\item consistently strong ranks for both metrics, no clear association: \texttt{COVIDhub-ensemble}
+\item Several models showed a positive association between allocation score and MWIS ranks (e.g., \texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha}).
+\item Several models had consistently strong MWIS ranks but also had highly variable allocation score ranks with no clear association between the two (e.g., \texttt{JHUAPL-SLPHospEns} and \texttt{CU-select})
+\item One model performed consistently well for both metrics with no clear association (\texttt{COVIDhub-ensemble})
+\item One model performed consistently well for allocation score but had only middlings ranks for MWIS (\texttt{CMU-TimeSeries})
 \end{itemize}
-Additionally, the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, had the highest rank for allocation score in six weeks, more than any other model except the \texttt{COVIDhub-ensemble} which also had six.
-
-<<eval=FALSE>>=
-alloscore_data_summaries |>
-  group_by(model) |>
-  summarize(top_rank = sum(allo_rank_std==1))
-@
-
- - bubble figure with ranks sized by total observations
 
@@
 
+Using the integrated allocation score summarizes allocation scores across a range of possible values of the constraint ($K$) (Section \ref{sec:methods.detailed.integrated_allocation}).
+This could be useful in situations where the actual constraint may not be known, or could only be specified as a probability distribution.
+
  - something here from Ben's analysis?
+
+\ngr{include figure here that shows relative rankings of locations?}
 
@@
 This work seeks to address that gap.
+However, we do note that the decision-making context presented in this work, while motivated by examples from real-world public health resource allocation problems, has not used to inform an actual real-time decision-making process.
 
-We have demonstrated that evaluation methods that are tied to decision making context can yield model rankings that are substantively different from generic measures of forecast skill like WIS.
+Resource allocation decisions are a realistic example of the kinds of decisions that could motivate more targeted forecasting exercises.
+One real-world example is allocation of ventilators during a respiratory viral pandemic.
+This was the motivating example for the examples presented in the work above.
+However, other examples where allocation is a central concern exist as well, such as the allocation of a limited stockpile of vaccinations\citep{persad_fair_2023} or diagnostic tests\citep{du_optimal_2022}.
 
-In practice, there are many users of forecasts with many different decision making problems.
+In practice, there are many potential users of forecasts with many different decision making problems.
 Not all can be easily quantified.
-Those that can be easily quantified may differ enough that no single score is appropriate for all users.
-We suggest reporting multiple scores.
-This may be tricky to operationalize in the setting of a general forecast hub.
-It matters how you elicit and represent probabilistic forecasts (quantiles?
-samples?
-cdfs?).
+Those that can be easily quantified may differ enough that it seems likely that no single score would be appropriate for all users.
+In an ideal setting, a forecasting tool could be developed through close collaboration between modelers and public health officials.
+However, this setting may only be possible in settings with enough resources to support both an analytics and public heatlh team.
+Increasingly, collaborative modeling hubs are being used to generate ``one-size-fits-all'' forecasts for many locations at once.
+In these settings, where tailored models are not available, it still could be possible to evaluate contributed models against a set of multiple scores.
+% NGR: leaving this out for now as it feels like too much detail?
+% This may be tricky to operationalize in the setting of a general forecast hub. It matters how you elicit and represent probabilistic forecasts (quantiles? samples? cdfs?).
 
-The allocation score we developed here does not directly account for important considerations such as fairness/equity of allocations.
+In our specific application, we made several key observations that we believe should inform future work in this area.
+First, it is clear that different metrics (in our case allocation score and mean weighted interval score) captured different aspects of forecast performance.
+The MWIS was strongly dependent on the scale of the forecasted quantity (e.g., when hospitalizations were high, so was WIS).
+The allocation score was not as scale-dependent, as the highest/worst allocation scores were observed when the number of hospitalizations was closest to the allocation constraint.
+We also note that contributed models found it hard to have a better allocation score than a na\"ive baseline model that just predicted a flat line from the last previous observation with wide uncertainty.
+This suggests suggests that contributed models (other than the ensemble, which combined forecasts from all contributed models) were not consistently adding value over just extrapolating from the current levels.
 
-The allocation score we developed also does not attempt to capture the broader context of decision making.
+There are several important limitations to the current work.
+The allocation score we developed here does not directly account for important considerations such as fairness or equity of allocations.
+Nor does the proposed framework attempt to capture the broader context of decision making.
 For example, in practice it may be possible to increase the resource constraint $K$ by shifting funding from other disease mitigation measures.
+We also note that in some settings, a ``successful'' epidemiological forecast may lead to policy decisions that change the distribution of the predicted outcome $Y$.
+Our framework cannot be used to evaluate a forecast at a horizon that may have been impacted by such a decision.
 
-Forecaster's dilemma: a successful forecast may lead to decisions that change the distribution of the outcome $Y$. Our framework cannot be used in those settings.
+We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work in more closely coupling applied epidemiological forecasting with public health decision-making.
+We note that a few papers have begun exploring similar linkages as described in the literature review \textendash but we see much room for additional work in this area.
+In some situations, individual or ensemble models could be developed to optimize scores that are attunend to a particular decision making setting.
+This area is also largely unexplored in the realm of public health to date, with some initial methodological development in econometrics \citep{loaiza-maya_focused_2021}.
 
-There is much more to do in this general area.
-
-\section{References}
+In conclusion, we argue that the way modelers and policymakers view and evaluate forecasts should change depending on the specific decision-making context.
+Using standard forecast evaluation metrics can mask the utility of certain forecasts, or lead to forecasts being used to inform a particular decision that are not optimal for that particular set of values that are important for that context.
+Further collaborative work is needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or simulation exercises that could be used to inform further development of new or alternative scoring metrics.
 

### 2023-11-06 alloscore_manuscript/alloscore-application.Rnw

@@
 \end{align}
-We emphasize that in Equation \eqref{eqn:informal_loss} the calculation of unmet need is based on the actual resource need that was realized in each location, while in Equation \eqref{eqn:informal_objective} the calculation of unmet need was based on a hypothetical potential future level of resource need.
+We emphasize that in Equation \eqref{eqn:informal_loss} the calculation of unmet need is based on the actual resource need that was realized in each location, while in Equation \eqref{eqn:informal_objective} the calculation of unmet need was based on the forecast distribution of future levels of resource need.
 Once the actual levels of resource need have been observed, we can obtain a quantitative measure of the quality of alternative allocation decisions: one allocation is better than another if it results in lower total unmet need.
@@
 There are $N = 2$ locations, with predictive distributions $F_1 = \mathrm{Exp}(1)$ and $F_2 = \mathrm{Exp}(1/4)$.
-The cdfs of these distributions are illustrated in the panels at bottom and right.
+The cumulative distribution functions of these distributions are illustrated in the panels at bottom and right.
 In the center panel, the background shading corresponds to the expected loss according to these forecasts.
@@
 
-Next suppose that we observe needs of 1 and 10 in locations 1 and 2, respectively.
+Next suppose that we observe resource needs of 1 and 10 in locations 1 and 2, respectively.
 Based on these observed needs, we can measure the quality of the allocation suggested by the forecast by calculating the amount of unmet need that resulted from that allocation over and above what was unavoidable given the resource constraint.
@@
 If the resource constraint is very large, it becomes less important which locations receive relatively more or less resources because all locations will receive enough resources to meet their need.
+In either of these extremes of resource availability, the avoidable unmet need that arises from the allocation suggested by a forecast (i.e., the forecast's allocation score) will tend to be small.
 
@@
 Subject to certain technical conditions, scoring rules obtained from this procedure are proper.
-We refer the reader to discussion in (cite paper 2 on arxiv) for a more technically precise discussion.
-
-% , and we illustrate the ideas with two running examples: (1) some problem that results in quantile elicitation, (2) some allocation problem.
+We refer the reader to (cite paper 2 on arxiv) for a more technically precise discussion.
 
@@
 Finally, we assume that each unit of unmet need incurs a loss denoted by $L$, so that if the selected resource level $x_i$ in location $i$ is less than the realized need $y_i$, a loss of $L \cdot (y_i - x_i)$ results.
+A variety of extensions to this setup are possible; for example, we might account for storage costs for resources that go unused, allow for a different loss per unit of unmet need in each location, or account for resource transportation costs.
+In this work, we choose to keep the loss function relatively simple to focus on the core ideas.
 
-We note that the problem formulation here assumes that the resource in question does not impact the amount of demand $y_i$ that will materialize, but rather it is a resource that satisfies that demand.
+It is helpful to clearly distinguish between the time $t_d$ when a \emph{d}ecision is made about a public health resource allocation and the time $t_r$ when \emph{r}esource needs that might be addressed by that allocation occur.
+Our setup assumes that $t_d < t_r$.
+Additionally, the structure of our loss captures a setting where the resource in question does not impact the amount of demand $y_i$ that will materialize at time $t_r$, but rather it is a resource that satisfies that demand.
 In the context of infectious disease, this means that we do not consider resources that are intended to reduce the number of people who will become sick at some point in the future, such as a preventative influenza or COVID-19 vaccine.
-Our set up addresses resources like hospital beds, oxygen supply, ventilators, or rabies vaccines which are intended to meet the medical needs of patients who are already sick.
+Instead, our set up addresses resources like hospital beds, oxygen supply, ventilators, or rabies vaccines which are intended to meet the medical needs of patients who are already sick.
+We also note that our problem formulation addresses decision-making that is related to resource needs only at the time $t_r$; we do not explicitly consider sequences of multiple decisions that are made over time or account for the impact of decisions on resource needs at any time other than $t_r$.
+We outline some opportunities to extend our work to more complex decision making settings in the discussion.
 
-With this notation in place, we can develop a proper scoring rule following the outline in section \ref{sec:methods.detailed.decisiontheory}.
+With this problem formulation in place, we can develop a proper scoring rule following the outline in section \ref{sec:methods.detailed.decisiontheory}.
 
@@
 
-\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the forecast, $x^{F,K}$, is the allocation that minimizes the expected loss, that is, the solution of the \emph{allocation problem} associated with $K$
+\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the forecast, $x^{F,K}$, is the allocation that minimizes the expected loss, that is, the solution of the \emph{allocation problem} associated with $K$:
 \begin{align}
@@
 where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]$ sums the expected loss due to unmet need across all locations.
+
+In the supplement we derive a general form of the Bayes act by beginning with the fact that a condition for minimizing expected loss subject to the resource constraint is that there is no way to decrease expected loss further by shifting a small amount of the resource from one location to another.
+From this starting point, we can show that the components of the Bayes act are quantiles $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource constraint $K$, but is shared across all locations.
+% Compare % In the supplement we derive a general form of the Bayes act by beginning with the fact that a condition for minimizing expected loss under the constraint is that there is no way to decrease expected loss further by shifting a small amount of the resource from one location to another.
+% From this, we can show that the components of the Bayes act are quantiles $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource constraint $K$, but is shared across all locations.
+% Compare with earlier text: This will lead to the allocations provided by $F$ being quantiles of the marginal distributions $F_i$ for some \emph{single} probability level $\tau$ that is shared in common for all locations.
+This probability level is the level at which the resource constraint is satisfied: $\sum_{i=1}^N F_i^{-1}(\tau^{F,K}) = K$.
+This tells us that in order to allocate optimally (according to $F$), we must divide resources among the locations so that there is an equal forecasted probability in every location that the allocation is sufficient to meet resource need.
+This solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
 
@@
 \begin{align}
-S_A^{\text{raw}}(F, y; K) = s_A(x^{F,K}, y) = \sum_{i=1}^N L \cdot \max(0, y_i - F_i^{-1}(1- \lambda(K,F)/L))
+S_A^{\text{raw}}(F, y; K) = s_A(x^{F,K}, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i^{F,K}).
 \end{align}
 This measures the total unmet need across all locations that results from using the Bayes allocation associated with the forecast $F$ when the actual level of need in each location is observed to be $y_i$.
-But to make this a more easily interpreted measure of
-forecast performance, we will adjust it by the the minimum loss $l_{\text{oracle}}(y;K)$ achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$ which is $l_{\text{oracle}}(y;K) = L \cdot \max(0, \sum_{i=1}^{N}y_i - K)$.
+
+To make this a more easily interpreted measure of forecast performance, we will adjust the raw score by subtracting the minimum loss $l_{\text{oracle}}(y;K)$ achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$.
+This minimum loss is given by $l_{\text{oracle}}(y;K) = L \cdot \max(0, \sum_{i=1}^{N}y_i - K)$: if there are enough resources to meet the total resource need ($\sum_{i=1}^{N}y_i \leq K$), the oracle can achieve a loss of 0, but if the resource constraint is less than the total needed resources ($\sum_{i=1}^{N}y_i > K$), it is not possible to do better than a loss of $L \cdot (\sum_{i=1}^{N}y_i - K)$.
+
 Our scoring rule is then given by
@@
 S_A(F, y; K) &= S_A^{\text{raw}}(F, y; K) - l_{\text{oracle}}(y;K) \\
-& = \sum_{i=1}^N L  \cdot \max(0, y_i - x_i^{F,K}) - L \cdot  \max(0, \sum_{i=1}^{N}y_i - K) \\
 & =
@@
 is met with the Bayes act component $x_i^{F,K}$.
-The oracle adjustment aligns with a central theme in modern decision theory across disciplines including economics, computer science, and psychology which is that
-\emph{opportunity loss} (often known as \emph{regret} or (negative) \emph{relative utility}) is often a more important quantity than absolute loss from both a descriptive and normative perspective.
+The oracle adjustment aligns with the idea that \emph{opportunity loss} (often known as \emph{regret} or (negative) \emph{relative utility}) is often a more important quantity than absolute loss \elr{needs citation}.
 
@@
 
-The allocation score $S_A$ that we developed in the previous section measures the skill of the forecast distributions $F$ based on a single probability level $\tau$.
+The allocation score $S_A$ that we developed in the previous section measures the skill of the forecast distributions $F$ based on a single probability level $\tau^{F,K}$.
 This is appropriate if the resource constraint $K$ is a known constant.
@@
 $$S_{IAS}(F, y) = \int S_A(F,y; K) p(K) \, dK$$
+We note that the device of considering a range of hypothetical decision makers or decision making problems with different problem parameters has been employed in the past \cite[e.g.,][]{murphy1993whatisagoodforecast}.
 

### 2023-11-07 alloscore_manuscript/alloscore-application.Rnw
updates to sec 2.3

@@
 
-We begin this section by briefly sketching how the weighted interval score (WIS), a commonly used proper scoring rule for probabilistic forecasts during the COVID-19 pandemic, can also be derived using the approach above.
+We begin this section by briefly sketching how the weighted interval score (WIS), a commonly used proper scoring rule for probabilistic forecasts during the COVID-19 pandemic, can be derived using the decision theoretic approach above.
 Then, we discuss similarities and differences between WIS and the allocation score, and other scores in general.
@@
 
-The weighted interval score (WIS), is a scoring rule first defined in 2020 to assist with scoring public forecasts that were being made in the early stages of the COVID-19 pandemic.\citep{bracher2021evaluating}
-Many COVID forecasts were stored in a quantile format, where probabilistic distributions are encoded by a fixed set of quantiles.
+The weighted interval score (WIS) was proposed in 2020 as a way to score forecasts that were being made in the early stages of the COVID-19 pandemic \citep{bracher2021evaluating}; equivalent scores had also been used in previous forecast evaluation efforts \cite[e.g.,][]{hong2016probabilisticEnergyForecasting}.
+The WIS is a proper scoring rule for forecasts that use a set of quantiles to represent a probabilistic forecast distribution.
 %% optional detail
-This was largely a choice of convenience (e.g., all forecasts had a known, fixed data size) and was codified by the US COVID-19 Forecast Hub and other collaborative forecasting projects.\citep{cramer_united_2022}
+Many early COVID-19 modeling efforts, including the US COVID-19 Forecast Hub and other collaborative forecasting projects, adopted a quantile forecast format as a matter of convenience (e.g., this format does not require the forecaster to pre-specify an upper bound for disease counts as would be required to collect forecasts in the format of bin probabilities) \citep{cramer_united_2022}.
 %%
-While pointing a reader interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS, is as its name suggests, a weighted sum of interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...).
+While pointing a reader interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS is a weighted sum of interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...).
 Larger interval scores indicate less skillful forecasts.
-Interval scores themselves consist of (a) the width of the interval, with larger intervals receiving higher scores, and (b) a penalty if the interval does not cover the eventual observation, which increases the further away the interval is from the observed value.
-We note that the most commonly used version of WIS is one that chooses a specific set of weights so that WIS approximates the continuous rank probability score (CRPS) or pinball loss, a commonly used score for probabilistic forecasts.
-Other specifications of weights for the various interval scores can be used.
-\emph{It is important to note that the version of WIS that has been commonly used for evaluating COVID forecasts was proposed because it approximates a commonly used forecast scoring rule, the CRPS, and not because it aligned with any particular public health decision-making rationale.}
+An interval score consists of (a) the width of the interval, with larger intervals receiving higher scores, and (b) a penalty if the interval does not cover the eventual observation, which increases the further away the interval is from the observed value.
+Equivalently, the WIS can also be characterized as a weighted sum of quantile scores for each individual predictive quantile.
+The quantile score for a particular quantile level assigns an asymmetric penalty to predictions that are too high or too low, with the relative sizes of the penalties set so that in expectation the score is minimized by the given quantile of the distribution.
+The most commonly used version of WIS is one that uses an equal weighting of all quantile levels, in which case WIS approximates the continuous rank probability score (CRPS), a commonly used score for probabilistic forecasts.
+\emph{It is important to note that this weighting was proposed because the resulting score approximates the CRPS, and not because it aligned with any particular public health decision-making rationale.}
 
-The decision-making setup that leads to a choice of WIS as a scoring rule can be formulated using the same recipe as shown in section \ref{sec:methods.detailed.decisiontheory}.
-In fields such as meteorology and supply chain management, a great deal of attention has been given to the problem where a decision must be made about the quantity of a resource to purchase for a single location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each unit of unmet need.\ngr{I might suggest adding a detailed Step 1/2/3 as an appendix and referencing it here.
-But I think that adding that full detail here might be too much.} This decision-making problem leads to the pinball loss that is often used in quantile regression, and thence to the CRPS (or WIS, via approximation) as a scoring rule.\ngr{can we make more explicit the link here to the section above?
+That said, the quantile score and WIS can be derived using the same decision theoretic procedure that we outlined in section \ref{sec:methods.detailed}.
+In fields such as meteorology and supply chain management, a great deal of attention has been given to the problem where a decision must be made about the quantity of a resource to purchase for a single location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each unit of unmet need.
+This leads to the quantile score for the probability level $\tau = 1 - C/L$.
+From this point, the WIS or CRPS can be obtained by averaging across a range of decision making settings with different cost and loss parameters, using a similar motivation that we used to obtain the IAS from the AS in \ref{sec:methods.detailed.integrated_allocation}.
+%\ngr{I might suggest adding a detailed Step 1/2/3 as an appendix and referencing it here. But I think that adding that full detail here might be too much.}
+%\ngr{can we make more explicit the link here to the section above?
 I'm not confident in my ability to describe with technical accuracy the link between quantile/pinball loss and CRPS, but I think this is a good opportunity to do so briefly.}
@@
 The quantile loss and the allocation score presented in this work are both special cases of the score that emerges from that more general decision making problem that integrates across resource cost and constraint parameters together.
-We pursue this direction further in other work (cite arxiv version of paper 2).
+We pursue this direction further in other work that is in progress.
 

### 2023-11-07 alloscore_manuscript/alloscore-application.Rnw
minor update

@@
 This leads to the quantile score for the probability level $\tau = 1 - C/L$.
-From this point, the WIS or CRPS can be obtained by averaging across a range of decision making settings with different cost and loss parameters, using a similar motivation that we used to obtain the IAS from the AS in \ref{sec:methods.detailed.integrated_allocation}.
+From this point, the WIS or CRPS can be obtained by averaging across a range of decision making settings with different cost and loss parameters, using a similar motivation that we used to obtain the IAS from the AS in section \ref{sec:methods.detailed.integrated_allocation}.
 %\ngr{I might suggest adding a detailed Step 1/2/3 as an appendix and referencing it here. But I think that adding that full detail here might be too much.}

### 2023-11-07 alloscore_manuscript/alloscore-application.Rnw
Update alloscore-application.Rnw
CRPS/WIS comment

@@
 The quantile score for a particular quantile level assigns an asymmetric penalty to predictions that are too high or too low, with the relative sizes of the penalties set so that in expectation the score is minimized by the given quantile of the distribution.
-The most commonly used version of WIS is one that uses an equal weighting of all quantile levels, in which case WIS approximates the continuous rank probability score (CRPS), a commonly used score for probabilistic forecasts.
+The most commonly used version of WIS is one that uses an equal weighting of all quantile levels, in which case WIS approximates the continuous ranked probability score (CRPS), a commonly used score for probabilistic forecasts.
+\apg{Maybe a rewording and/or a supplement reference here to mention that any WIS approximates the a corresponding weighted CRPS, and that WIS actually is the wCRPS for a discrete measure/weighting.}
 \emph{It is important to note that this weighting was proposed because the resulting score approximates the CRPS, and not because it aligned with any particular public health decision-making rationale.}

### 2023-11-14 R/run-alloscore.R
a pass at fixing targets. removing tar_load() from inside a target function. also removing all objects that should be ignored in _targets/objects

@@ -83,17 +83,17 @@ run_alloscore <- function(
 
 ## take estimated alloscores and put them in one clean dataset
 ## returned object has a row for each model, reference_date, K, state
-assemble_alloscores <- function(...) {
+assemble_alloscores <- function(alloscores_with_data) {
   require(targets)
   require(dplyr)
 
   ## load all alloscore targets and get their names
   ## NB: if any target starts with "alloscore" that isn't one of the fits, this will cause an error
-  tar_load(starts_with("alloscore"))
-  ascore_tars <- ls(pattern="alloscore*")
+  #tar_load(starts_with("alloscore"))
+  #ascore_tars <- ls(pattern="alloscore*")
 
   ## bind all alloscore dataframes together and filter out ones with no forecasts
-  alloscores_with_data <- do.call(bind_rows, mget(ascore_tars)) |>
+  alloscores_with_data <- alloscores_with_data |>
     dplyr::filter(is.na(.data$message))
 
   ## extract alloscore matrices

### 2023-11-14 alloscore_manuscript/supplement.rnw
starting propriety section

@@
 
-\usepackage{amsmath, amsfonts, amssymb}
+\usepackage{amsmath, amsfonts, amssymb, mathtools}
 \usepackage{graphicx}
@@
 \usepackage{caption}
+\usepackage{pgfplots}
+\usepackage{xr-hyper}
 \usepackage{hyperref}
+\externaldocument{alloscore-application}
 
 \DeclareMathOperator*{\argmin}{argmin}
+\DeclareMathOperator{\short}{sh}
+\DeclareMathOperator{\Ex}{\mathbb{E}}
 
 \usepackage{setspace}
-\onehalfspacing
 
@@
 \newcommand{\sbar}{\overline{s}}
+
 \newtheorem{proposition}{Proposition}
@@
 	\item Explain ``All proper scoring rules for probabilistic forecasts have an explicit link to a loss function'' from discussion.
-	\item DGP as optimal for any decision problem, ref Diebold, Gunther, Tay p. 866.
+	\item DGP as optimal for any decision problem, ref Diebold, Gunther, Tay p.
+866; and if forecasts are ideal, then forecasts with better information always yield better decisions, ref Holzmann and Eulert, Corr 2.
 	\item For 2.2.2, how to get quantile representation of Bayes act using Lagrange multiplier, assuming smooth, never-zero densities well behaved at $x=0$.
@@
 
+\section{Shortages}
+\label{sec:shortage}
+
+For convenience, we write $u_{+} = \max\{0,u\}$, and refer to $(y - x)_{+}$ as a \emph{shortage} in accordance with our
+typical use of $y$ for a demand or need and $x$ for an available supply. To regard shortage as a function depending on only one 
+variable $x$ or $y$, with the other being a parameter describing the dependence we can write 
+$(y - x)_{+} = \short^{y}(x) = \short_{x}(y)$.  Note that 
+$\short^{y}(x)$ and $\short_{x}(y)$ are both convex functions and ``mirror'' each other:
+
+\pgfplotsset{compat=1.16} % Set the compatibility of pgfplots
+
+\begin{tikzpicture}
+\begin{axis}[
+    axis lines=middle, % Center the axis
+    xmin=0, xmax=10, % Set the range for the x-axis
+    ymin=0, ymax=10, % Set the range for the y-axis
+    xlabel={$x$ or $y$},
+    ylabel={shortage},
+    xtick={4, 6}, % Set the position of the ticks on the x-axis
+    ytick=\empty, % No ticks on the y-axis
+    xticklabels={$x_0$, $y_0$}, % Label for the tick on the x-axis
+    clip=false,
+    every axis x label/.style={at={(current axis.right of origin)},anchor=north west},
+    every axis y label/.style={at={(current axis.above origin)},anchor=south east}
+]
+
+% Add the plot for max(0, y_0 - x)
+\addplot+[sharp plot, no marks, thick, blue] coordinates {(0,6) (6,0) (10,0)}
+node[pos=.1,anchor=south west] {$\short^{y_0}(x) = (y_0 - x)_{+}$}; % Label for the first function
+
+% Add the plot for max(0, y - x_0)
+\addplot+[sharp plot, no marks, thick, red, ->] coordinates {(0,0) (4,0) (10,6)}
+node[pos=.7,anchor=north west] {$\short_{x_0}(y) = (y - x_0)_{+}$};
+
+\end{axis}
+\end{tikzpicture}
+
+Let $Y$ be a random variable with distribution $F$. The random shortage $(Y-x)_{+}$ can be thought of as either a 
+real-valued random variable $\short_x(Y)$ for every $x$, or a function-valued random variable $\short^Y$ whose value
+for any realization $Y=y$ is a convex function $\short^y(x)$ of $x$.  We see then that the 
+\emph{expected shortage}\footnote{A more natural sounding term for $(y-x)_{+}$ might have been \emph{shortfall}. Unfortunately 
+\emph{expected shortfall} has long been used in finance to refer to quantities more 
+closely related to the \emph{conditonal} expectation $\Ex_F [Y-x \mid Y - x \geq 0] = \Ex_F [(Y-x)_{+}]/\mathbb{P}_F\{Y \geq x\}$.}
+$\Ex_F [(Y-x)_{+}] = \Ex_F[\short^Y](x)$ (assuming it exists) is also convex (and therefore continuous) in $x$ by integrating the convexity inequality for
+$\short^y(x)$ with respect to the probability measure $dF(y)$:
+\begin{align}
+	\Ex_F[\short^Y](\lambda x_1+(1-\lambda) x_2) &= \int \short^y(\lambda x_1+(1-\lambda) x_2)dF(y) \nonumber \\
+	&\leq \int \lambda\short^y(x_1) +(1-\lambda) \short^y(x_2)dF(y) \nonumber \\
+	&=\lambda\Ex_F[\short^Y](x_1) + (1-\lambda)\Ex_F[\short^Y](x_2). \label{eqn:convex-esf}
+\end{align}
+Convexity is also shown by directly exhibiting the the left and right derivatives of $\Ex_F[\short^Y](x)$:
+\begin{align}
+D_{-}\Ex_F [(Y-x)_{+}] &= \lim_{h \searrow 0}\frac{1}{h}\Ex_F [(Y - x)_{+} - (Y - x - h)_{+}] \\
+&= \lim_{h \searrow 0}\frac{1}{h}\int_{[x\mathrlap{-h, x]}} (x - h - y) dF(y) - 
+\lim_{h \searrow 0}\frac{1}{h}\int_{(x\mathrlap{,\infty)}} h dF(y)  \\
+&= \lim_{h \searrow 0}\frac{1}{h}\int_{[x\mathrlap{-h, x]}} - h dF(y) - 1 + F(x) \label{eqn:dsfllim} \\
+&= -(F(x) - F(x-)) - 1 + F(x) \quad \left(\text{where } F(x-) := \lim_{t \nearrow x} F(t)\right)\\
+&= F(x-) - 1  \label{eqn:dsfl} \\
+D_{+}\Ex_F [(Y-x)_{+}] &= \lim_{h \searrow 0}\frac{1}{h}\Ex_F [(Y - x + h)_{+} - (Y - x)_{+}] \\
+&= \lim_{h \searrow 0}\frac{1}{h}\int_{[x\mathrlap{, x+h]}} (x - y) dF(y) - 
+\lim_{h \searrow 0}\frac{1}{h}\int_{(x\mathrlap{+h,\infty)}} h dF(y)  \\
+&= \lim_{h \searrow 0}\frac{1}{h}\int_{[x\mathrlap{, x+h]}} 0 dF(y) - 1 + F(x) \label{eqn:dsfrlim} \\
+&= F(x) - 1 \label{eqn:dsfr}
+\end{align}
+where in \eqref{eqn:dsfllim} and \eqref{eqn:dsfrlim} we are able to replace the integrands with their values at $x$ because 
+they are bounded over the shrinking regions of integration $[x-h,x]$ and $[x,x+h]$. Convexity follows since 
+$D_{-}\Ex_F[\short^Y](x) \leq D_{+}\Ex_F[\short^Y](x)$ by the definition of $F(x)$ and $F(x-)$. This shows that if $F$ does not 
+have a point mass at $x$, we have 
+\begin{align}
+	\frac{d}{dx} \Ex_F [(Y-x)_{+}] = F(x)-1,
+\end{align}
+coinciding with the ``Leibniz rule'' calculation
+\begin{align}
+	\frac{d}{dx} \Ex_F [(Y-x)_{+}] &= \frac{d}{dx} \int_{x}^{\infty} (y-x) f_Y(y)dy \\
+	&= \int_{x}^{\infty} \frac{d}{dx}(y-x) f_Y(y)dy - (x-x) f_Y(x) = -\int_{x}^{\infty} f_Y(y)dy = F(x)-1.
+\end{align}
+which assumes $Y$ has an adequately well-behaved density $f_Y$.
+
+\section{Quantiles and Expected Shortage}
+\label{sec:quantiles_shortage}
+
+We recall how quantiles arise as solutions to a probabilistic decision problem. Let $Y$ be a random variable
+representing the future level of an undesirable outcome such as severe COVID incidence. Let $x$ be a decision variable
+representing the possible levels of some costly counter-measure, such as procurement of monoclonal antibody treatments,
+that can be taken at a cost $C$ per unit in preparation for $Y$.  A decision maker must decide on a level $x$ of investment in the
+counter-measure, and wishes to avoid excesses in either the expediture $Cx$ or the shortage $(y-x)_{+}$ when $Y=y$ is
+realized. To formalize the trade-off between these potential excesses we quantify the loss associated with a unit of
+shortage by a constant $L>C$ (which assumes that the counter-measure has some economic value) and combine the total shortage loss with expenditure into a \emph{loss function}\footnote
+{This does involve a confusing use of the word \emph{loss} to refer to two different quantities, but this seems to be
+an ingrained and unavoidable habit in the literature.}
+\[
+l(x,y) = Cx + L(y-x)_{+}.
+\] 
+The decision problem is then to select a random future loss $l(x,Y)$ in a way that aligns with the preference that $l(x,y)$ be
+as low as possible given any realization $Y=y$.
+
+To give the decision problem more structure we assume the decision maker either knows the distribution $F$ of $Y$, or
+wishes to proceed as if a forecast $F$ of $Y$ were true. This gives us what is known in decision theory as a decision
+problem \emph{under risk} (regarding the future value of $Y$) as opposed to one \emph{under uncertainty} where both $Y$
+as well as $F$ are unknown when the decision is to be made. A principle commonly invoked in this situation\footnote
+{Note that this priciple might be inappropriate when the decision maker is \emph{risk averse} in some way such as
+having a preference for random losses with lower variance.} is that the decision maker should or will seek to minimize the expected
+loss
+\begin{align}
+  \mathbb{E}_{F}[l(x,Y)] = Cx + L\mathbb{E}_{F}[(Y-x)_{+}]. \label{eqn:expQloss}
+\end{align}
+
+The expected loss is an affine transformation of the convex expected shortage (c.f. \eqref{eqn:convex-esf}). Therefore
+$\mathbb{E}_{F}[l(x,Y)]$ is also is convex and has right and left derivatives $D_{\pm}\mathbb{E}_{F}[l(x,Y)]$ at every $x$. 
+Because these derivatives exist everywhere, a necessary condition for $x^{\star}$ to
+minimize $\mathbb{E}_{F}[l(x,Y)]$ is that $D_{+}\mathbb{E}_{F}[l(x^{\star},Y)] \geq 0$ and 
+$D_{-}\mathbb{E}_{F}[l(x^{\star},Y)] \leq 0$, and because of convexity, this condition is also sufficient. From \eqref{eqn:dsfl} 
+and \eqref{eqn:dsfr} this means that
+\begin{align}
+D_{+}\mathbb{E}_{F}[l(x^{\star},Y)] = C + L(F(x^{\star}) - 1) \geq 0 \geq 
+D_{-}\mathbb{E}_{F}[l(x^{\star},Y)] = C + L(F(x^{\star}-) - 1) \label{eqn:expQd-ineq}
+\end{align}
+which rearranges with $\alpha = 1-C/L$ to 
+\begin{align}
+F(x^{\star}) \geq \alpha \geq F(x^{\star}-). \label{eqn:F-alpha-ineq}
+\end{align}
+Note that because $F(x)$ and $F(x-)$ are right and left continuous, repectively, the set $\{x \mid F(x) \geq \alpha\}$ is closed
+on the left and the set $\{x \mid \alpha \geq F(x-)\}$ is closed on the right.
+Therefore, \eqref{eqn:F-alpha-ineq} implies that
+\begin{align}
+\min\{x \mid F(x) \geq \alpha\}  \leq x^{\star} \leq \max\{x \mid \alpha \geq F(x-)\}. \label{eqn:set-alpha-ineq}
+\end{align}
+We call $q^{-}_{\alpha, F} := \min\{x \mid F(x) \geq \alpha\}$ and $q^{+}_{\alpha, F} := \max\{x \mid F(x-) \leq \alpha\}$
+the left and right quantiles of $F$ (for probability level $\alpha$) and any element $q_{\alpha, F} \in [q^{-}_{\alpha, F}, q^{+}_{\alpha, F}]$
+a quantile of $F$.  Thus $x^{\star}$ minimizes the expected loss \eqref{eqn:expQloss} and gives an optimal sultion to the decision problem
+if and only if it is a quantile $q_{\alpha, F}$.
+
+Quantiles equivalently arise when the decision problem is defined in terms of the random \emph{oportunity} loss 
+\begin{align}
+l_o(x,Y) := l(x,Y) - l(Y,Y) = Cx + L(Y-x)_{+} - CY \label{eqn:opp-loss}
+\end{align}
+which expresses how much more loss is realized by the decision $x$ than an oracle would have incurred, knowing to invest exactly
+the future value of $Y$. The optimal decision for $\Ex_F[l_o(x,Y)]$ is the same as for $\Ex_F[l(x,Y)]$ since the term $-C\Ex_F[Y]$ is 
+constant in $x$, leading again to the inequalities \eqref{eqn:expQd-ineq}.
+
+Opportunity loss \eqref{eqn:opp-loss} rearranges to 
+\begin{align}
+l_o(x,Y) &= C(x - Y)_{+} + (L-C)(Y-x)_{+} \\
+&= L(1-\alpha)(x - Y)_{+} + L\alpha(Y-x)_{+},
+\end{align}
+a form in which it is often called \emph{pinball} loss, despite its graph being an unlikely pinball trajectory for $\alpha \neq 1/2$.
+
+\section{Allocation Bayes acts as vectors of marginal quantiles.}
+
+Here we explain why the Bayes act $x_i^{F,K}$ for the allocation
+problem \eqref{AP} can be represented as a vector of quantiles for the marginal forecast distributions $F_i$ at a
+single probability level $\tau^{F,K}$, that is, $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$.
+
+For an arbitrary allocation vector $x \in \mathbb{R}^N_{+}$ the expected loss
+\begin{align}
+\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[(Y_i - x_i)_{+}] \label{eqn:AS_formula}
+\end{align} 
+is the sum of expected shortages (scaled by $L$) under the allocations $x_i$ in each location. We therefore
+have the following necessary condition for $x^{\star} \in \mathbb{R}^N_{+}$ to be an optimal allocation for $\mathbb
+{E}_{F} [s_A(x, Y)]$ under the constraint $\sum_{i=1}^{N} x_i = K$: if $\delta > 0$ of the $x_i^{\star}$ units of resource
+allocated to location $i$ are reallocated to location $j$, expected shortage will increase in location
+$i$ by at least as much as it decreases in location $j$. That is,
+\begin{align}
+\mathbb{E}_{F_i}[(Y_i - x^{\star}_i - \delta)_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}] \geq 
+\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - x^{\star}_j + \delta)_{+}]. \label{eqn:ASoptimal1}
+\end{align}
+Since the expected shortages in $i$ and $j$ have right and left derivatives at any $x_i$ and $x_j$ (see Section \ref{sec:shortage}), we can
+divide \eqref{eqn:ASoptimal1} by $\delta$ and take limits for $\delta \searrow 0$ to get
+\begin{align}
+-D_{-} \mathbb{E}_{F} [(Y_i - x^{\star}_i)_{+}] \geq -D_{+}\mathbb{E}_{F} [(Y_j - x^{\star}_j)_{+}]. \label{eqn:RLineq}
+\end{align}
+Note that the minus signs appear because our optimality condition addresses how a \emph{decrease} in resources will \emph{increase}
+the expected shortage in $i$ and vice versa in $j$.
+Scaling by $L$ to match the right and left partial derivatives of $\mathbb{E}_{F} [s_A(x, Y)]$ and using formulae \eqref{eqn:dsfl} and \eqref{eqn:dsfr}, \eqref{eqn:RLineq} becomes
+\begin{align}
+L(1-F_i(x^{\star}_i-)) \geq L(1-F_j(x^{\star}_j)). \label{eqn:RLineq2}
+\end{align}
+Inequalities \eqref{eqn:RLineq} and \eqref{eqn:RLineq2} remain true with $i$ and $j$ reversed. They hold with $i=j$ as well by the
+definition of $F_i(x^{\star}_i-)$. Therefore, a single number $\lambda$ (a \emph{Lagrange multiplier}) exists such that
+\begin{align}
+L(1-F_i(x^{\star}_i-)) \geq \lambda \geq L(1-F_i(x^{\star}_i)), \quad \text{for all } i \in 1,\ldots,N
+\end{align}
+that is,
+\begin{align}
+F_i(x^{\star}_i) \geq 1 - \lambda/L \geq F_i(x^{\star}_i-),
+\end{align}
+which says (c.f. discussion after \eqref{eqn:F-alpha-ineq} and \eqref{eqn:set-alpha-ineq}) that $x^{\star}_i$ is a quantile $q_{\tau,F_i}$ for
+$\tau = 1 - \lambda/L$.
+
+\section{Properties and Properness}
+
+\begin{quote}
+For a prediction to be useful, it must \textbf{proper}ly describe a \textbf{proper}ty.
+\end{quote}
+
+Expanding on the decision theoretic perspective sketched in Section \ref{sec:quantiles_shortage}, we can view a loss function 
+as a general tool for formalizing a decision problem that assigns numerical value to the \emph{result} of taking an \emph{action} $x$
+in preparation for an \emph{outcome} $y$.  A \emph{scoring rule} $S$ (or just \emph{score}) is a loss function where the action 
+is a probabilistic forecast $F$ of the outcome $y$ (or the stating by a forecaster of the forecast), written as $S(F,y)$.
+
+A scoring rule $S$ is \emph{proper} when it    the expected score $\Ex [S(F,Y)]$ is lowest 
+ invites the mistaken impression that whether a score 
+is proper might depend on the true distribution of what is being forecasted.
+
 \end{document}

### 2023-11-16 alloscore_manuscript/alloscore-application.tex

@@
 
-The weighted interval score (WIS), is a scoring rule first defined in 2020 to assist with scoring public forecasts that were being made in the early stages of the COVID-19 pandemic.\citep{bracher2021evaluating}
-Many COVID forecasts were stored in a quantile format, where probabilistic distributions are encoded by a fixed set of quantiles.
+The weighted interval score (WIS) was proposed in 2020 as a way to score forecasts that were being made in the early stages of the COVID-19 pandemic \citep{bracher2021evaluating}; equivalent scores had also been used in previous forecast evaluation efforts \cite[e.g.,][]{hong2016probabilisticEnergyForecasting}.
+The WIS is a proper scoring rule for forecasts that use a set of quantiles to represent a probabilistic forecast distribution.
 %% optional detail
-This was largely a choice of convenience (e.g., all forecasts had a known, fixed data size) and was codified by the US COVID-19 Forecast Hub and other collaborative forecasting projects.\citep{cramer_united_2022}
+Many early COVID-19 modeling efforts, including the US COVID-19 Forecast Hub and other collaborative forecasting projects, adopted a quantile forecast format as a matter of convenience (e.g., this format does not require the forecaster to pre-specify an upper bound for disease counts as would be required to collect forecasts in the format of bin probabilities) \citep{cramer_united_2022}.
 %%
-While pointing a reader interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS, is as its name suggests, a weighted sum of interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...).
+While pointing a reader interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS is a weighted sum of interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...).
 Larger interval scores indicate less skillful forecasts.
-Interval scores themselves consist of (a) the width of the interval, with larger intervals receiving higher scores, and (b) a penalty if the interval does not cover the eventual observation, which increases the further away the interval is from the observed value.
-We note that the most commonly used version of WIS is one that chooses a specific set of weights so that WIS approximates the continuous rank probability score (CRPS) or pinball loss, a commonly used score for probabilistic forecasts.
-Other specifications of weights for the various interval scores can be used.
-\emph{It is important to note that the version of WIS that has been commonly used for evaluating COVID forecasts was proposed because it approximates a commonly used forecast scoring rule, the CRPS, and not because it aligned with any particular public health decision-making rationale.}
+An interval score consists of (a) the width of the interval, with larger intervals receiving higher scores, and (b) a penalty if the interval does not cover the eventual observation, which increases the further away the interval is from the observed value.
+Equivalently, the WIS can also be characterized as a weighted sum of quantile scores for each individual predictive quantile.
+The quantile score for a particular quantile level assigns an asymmetric penalty to predictions that are too high or too low, with the relative sizes of the penalties set so that in expectation the score is minimized by the given quantile of the distribution.
+The most commonly used version of WIS is one that uses an equal weighting of all quantile levels, in which case WIS approximates the continuous ranked probability score (CRPS), a commonly used score for probabilistic forecasts.
+\apg{Maybe a rewording and/or a supplement reference here to mention that any WIS approximates the a corresponding weighted CRPS, and that WIS actually is the wCRPS for a discrete measure/weighting.}
+\emph{It is important to note that this weighting was proposed because the resulting score approximates the CRPS, and not because it aligned with any particular public health decision-making rationale.}
 
-The decision-making setup that leads to a choice of WIS as a scoring rule can be formulated using the same recipe as shown in section \ref{sec:methods.detailed.decisiontheory}.
-In fields such as meteorology and supply chain management, a great deal of attention has been given to the problem where a decision must be made about the quantity of a resource to purchase for a single location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each unit of unmet need.\ngr{I might suggest adding a detailed Step 1/2/3 as an appendix and referencing it here.
-But I think that adding that full detail here might be too much.} This decision-making problem leads to the pinball loss that is often used in quantile regression, and thence to the CRPS (or WIS, via approximation) as a scoring rule.\ngr{can we make more explicit the link here to the section above?
+That said, the quantile score and WIS can be derived using the same decision theoretic procedure that we outlined in section \ref{sec:methods.detailed}.
+In fields such as meteorology and supply chain management, a great deal of attention has been given to the problem where a decision must be made about the quantity of a resource to purchase for a single location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each unit of unmet need.
+This leads to the quantile score for the probability level $\tau = 1 - C/L$.
+From this point, the WIS or CRPS can be obtained by averaging across a range of decision making settings with different cost and loss parameters, using a similar motivation that we used to obtain the IAS from the AS in section \ref{sec:methods.detailed.integrated_allocation}.
+\apg{Refs to supp for quantile score derivation and detail about averaging.}
+%\ngr{I might suggest adding a detailed Step 1/2/3 as an appendix and referencing it here. But I think that adding that full detail here might be too much.}
+%\ngr{can we make more explicit the link here to the section above?
 I'm not confident in my ability to describe with technical accuracy the link between quantile/pinball loss and CRPS, but I think this is a good opportunity to do so briefly.}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in gzfile(file, "{}rb"{}): cannot open the connection}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in group\_by(Kat15k\_alloscores, model, reference\_date): object 'Kat15k\_alloscores' not found}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in unique(alloscore\_data\_summaries\$model): object 'alloscore\_data\_summaries' not found}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in `filter()`:\\\#\# i In argument: `model \%in\% alloscore\_models`.\\\#\# Caused by error in `model \%in\% alloscore\_models`:\\\#\# !
-object 'alloscore\_models' not found}}\end{kframe}
-\end{knitrout}
-
 \subsubsection{Forecast scores showed differences over time}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in `filter()`:\\\#\# i In argument: `target\_end\_date \%in\%\\\#\# \ \ score\_data\_summaries\$target\_end\_date`.\\\#\# Caused by error in `target\_end\_date \%in\% score\_data\_summaries\$target\_end\_date`:\\\#\# !
-object 'score\_data\_summaries' not found}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ggplot(dat\_to\_plot, aes(x = target\_end\_date, y = value)): object 'dat\_to\_plot' not found}}\end{kframe}
-\end{knitrout}
 
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\includegraphics[width=\maxwidth]{figure/metrics-over-time-1} \caption[Hospital admissions and evaluation metrics over time]{Hospital admissions and evaluation metrics over time.
+Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022.
+These are the days for which forecasts were made and evaluated.
+A horizontal dashed line at 15,000 shows the assumed resource constraint $K$.
+Panel B shows allocation scores for each model's 14-day ahead forecast, across all US states.
+The x-axis corresponds to the date for which the prediction was made.
+Allocation scores typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down).
+Panel C shows the MWIS metric across weeks, averaged across all states.
+Similarly to panel B, the x-axis corresponds to the date for which the prediction was made.
+MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the pean of the Omicron wave.}\label{fig:metrics-over-time}
+\end{figure}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ggplot(alloscore\_data\_summaries, aes(x = target\_end\_date, y = alloscore, : object 'alloscore\_data\_summaries' not found}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ggplot(score\_data\_summaries, aes(x = target\_end\_date, y = mwis, : object 'score\_data\_summaries' not found}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in cowplot::plot\_grid(p1, p2, p4, nrow = 3, align = "{}v"{}, rel\_heights = c(1, : object 'p1' not found}}\end{kframe}
 \end{knitrout}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\includegraphics[width=\maxwidth]{figure/metrics-correlation-1} \caption[Association of standardized ranks for MWIS and allocation score by model and week]{Association of standardized ranks for MWIS and allocation score by model and week.
+Each facet of the plot corresponds to one model.
+Within each facet, each point corresponds to a week.
+The x- and y-values correspond to the MWIS standardized rank and the allocation score standardized rank for that week.
+Points corresponding to earlier dates have darker shading.
+The size of the point corresponds to the observed value on the date for which the prediction was made.
+Models show different degrees of association between the two metrics.}\label{fig:metrics-correlation}
+\end{figure}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in left\_join(alloscore\_data\_summaries, score\_data\_summaries, by = c("{}model"{}, : object 'alloscore\_data\_summaries' not found}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in filter(all\_data, as.Date(reference\_date.x) < as.Date("{}2022-02-01"{})): object 'all\_data' not found}}\end{kframe}
 \end{knitrout}

### 2023-11-16 alloscore_manuscript/supplement.rnw

@@
 as a general tool for formalizing a decision problem that assigns numerical value to the \emph{result} of taking an \emph{action} $x$
-in preparation for an \emph{outcome} $y$.  A \emph{scoring rule} $S$ (or just \emph{score}) is a loss function where the action 
-is a probabilistic forecast $F$ of the outcome $y$ (or the stating by a forecaster of the forecast), written as $S(F,y)$.
+in preparation for an \emph{outcome} $y$.  A \emph{scoring rule} $S$ is a loss function where the action 
+is a probabilistic forecast $F$ of the outcome $y$ (or the statement of $F$ by a forecaster). Just as in 
+Section \ref{sec:quantiles_shortage}, given an action $F$, $S$ transforms a random outcome variable $Y$ into a random loss $S(F,Y)$. 
+We refer to the realized loss $S(F,y)$ as the \emph{score} of $F$ at $y$, and the process of evaluating $S(F,y_i)$ for
+a data set $\mathcal{Y}=\{y_i\}$ as \emph{scoring} $F$ against $\mathcal{Y}$.
 
-A scoring rule $S$ is \emph{proper} when it    the expected score $\Ex [S(F,Y)]$ is lowest 
- invites the mistaken impression that whether a score 
-is proper might depend on the true distribution of what is being forecasted.
+Decision theoretically, probabilistic forecasts are a unique kind of action in that they can be used to generate their own 
+(simulated) outcome data, against which they can be scored using $S$. $S$ therefore commits a probabilistic forecast $F$
+to the ``self-assessment'' $\Ex [S(F, Y^F)]$, where $Y^F \sim F$ is the random variable defined by sampling from $F$, as well
+to an assessment $\Ex [S(G, Y^F)]$ of any alternative forecast $G$.
+A natural consistency criterion for $S$ is that it does not commit $F$ to assessing any other 
+forecast $G$ as being better than $F$ itself, that is, that
+\begin{align}
+\Ex [S(F, Y^F)] \leq \Ex [S(G, Y^F)]
+\end{align} 
+for any $F,G$. Otherwise, the optimal decision for some forecaster would be to state a forecast $G$ other than the forecast $F$
+which they believe describes the outcome $Y$.
+A scoring rule meeting this criterion is called \emph{proper}.  The inequality can also be written 
+as $\Ex_F [S(F, Y)] \leq \Ex_F [S(G, Y)]$ where the subscript specifies the distribution of $Y$.  
+$S$ is \emph{strictly proper} when this inequality is sharp, in which case the \emph{only} optimal decision for a forecaster is to state
+the forecast they believe to be true.
+
+Another definition of $S$ being proper that is often used is that $\Ex[S(F,Y)]$ is lowest when $F$ is the true distribution of $Y$.  
+Under a flexible reading, this definition is equivalent to ours, but we find it problematic because it invites the mistaken impression that 
+the properness of a score might depend on the true distribution of what is being forecasted. Whether a score is proper is unrelated 
+to any particular forecast being scored or source of the data being used to score it.
+
+The condition of being proper is quite strong, and na\"{i}ve means of reducing a forecast distribution to a single
+number based on an observed outcome $y$ will generally define an improper scoring rule.  For example, the \emph{probability score} 
+$S_{\mathrm{Prob},c}(F,y):= -(F(y+c) - F(y-c))$ and the \emph{linear score} $S_{\mathrm{Lin}}
+(F,y):= \lim_{c\to 0} \frac{1}{2c}S_{\mathrm{Prob},c}(F,y) = -f(y)$ (where we assume $F$ has a density $f$) and are
+both improper because they commit a general $F$ to assessing an alternative forecast $G$ as being better than $F$
+itself whenever $G$ is sufficiently more concentrated than $F$ in the neighborhood $[m_f-c,m_f+c]$ of a mode $m_f$ of
+$f$. The same is true when the outcome is discrete (and ordered for $S_{\mathrm{Prob}}$) with index $y$: for a forecast
+pmf $p(y)$, moving the mass of $p$ onto a neigborhood of a mode $m_p$ to get a new forecast $p_m$ will improve the 
+expected score $\Ex[S_{\mathrm{Prob/Lin}}(p_m,Y^p)]$ according to $p$ itself. This creates the classic pathology
+of probabilistic forecasters having no incentive to express uncertainty. The meteorologist getting paid according to
+$p_{\text{rain}}\text{rain} + (1-p_{\text{rain}})(1-\text{rain})$ (the negative of the linear score) will say rain is 
+inevitable or impossible when the chances of rain appear to be 51\% or 49\%.
+
+Sometimes a scoring rule can be made proper with an added correction that penalizes
+
+But given an auxiliary decision problem $\mathcal{D}_{\mathrm{Aux}} =\{\mathcal{X}, \mathcal{Y}, l(x,y)\}$, such as the
+allocation problem from the main text or the cost-loss problem from Section \ref{sec:quantiles_shortage}, we can
+produce a scoring rule $S_l$ that is automatically proper using the Bayes act formalism introduced in Section \ref
+{sec:methods.detailed.specific_allocation}. This works by defining $S_l(F, y)$ as the value $l(x^F,y)$ already given by  
+$\mathcal{D}_{\mathrm{Aux}}$ for the Bayes act $x^F$; that is, the action which \emph{by design} will be assessed via
+the forecast $F$ to have an expected loss $\Ex[l(x^F,Y^F)]$ as low as possible.
+
+by mapping a forecast $F$ to an the action $x^F$
 

### 2023-11-16 alloscore_manuscript/supplement.rnw
clean out midas talk folder

@@
 \emph{expected shortfall} has long been used in finance to refer to quantities more 
-closely related to the \emph{conditonal} expectation $\Ex_F [Y-x \mid Y - x \geq 0] = \Ex_F [(Y-x)_{+}]/\mathbb{P}_F\{Y \geq x\}$.}
+closely related to the \emph{conditional} expectation $\Ex_F [Y-x \mid Y - x \geq 0] = \Ex_F [(Y-x)_{+}]/\mathbb{P}_F\{Y \geq x\}$.}
 $\Ex_F [(Y-x)_{+}] = \Ex_F[\short^Y](x)$ (assuming it exists) is also convex (and therefore continuous) in $x$ by integrating the convexity inequality for

### 2023-11-19 alloscore_manuscript/supplement.rnw
more on properness...

@@
 l_o(x,Y) &= C(x - Y)_{+} + (L-C)(Y-x)_{+} \\
-&= L(1-\alpha)(x - Y)_{+} + L\alpha(Y-x)_{+},
+&= L(1-\alpha)(x - Y)_{+} + L\alpha(Y-x)_{+} \\
+&= L(\alpha - \mathbf{1}\{Y < x\})(Y-x),
 \end{align}
@@
 We refer to the realized loss $S(F,y)$ as the \emph{score} of $F$ at $y$, and the process of evaluating $S(F,y_i)$ for
-a data set $\mathcal{Y}=\{y_i\}$ as \emph{scoring} $F$ against $\mathcal{Y}$.
+a data set $\mathcal{Y}=\{y_i\}$ as \emph{scoring} $F$ against $\mathcal{Y}$. Perhaps the most fundamental example of a scoring rule
+from a statistical perspective is the logarithmic score $S_{\log}(F,y) = -\log f(y)$ (where $f(y)$ is the density or the mass of $F$ at $y$
+if it exists and otherwise the mass), that is, the negative log-likelihood of $F$ interpreted as a parameter for the singleton 
+data set $\{y\}$. This frames a maximum likelihood estimate as the solution to a decision problem following the expected loss
+minimization principle introduced in Section \ref{sec:quantiles_shortage}.
 
-Decision theoretically, probabilistic forecasts are a unique kind of action in that they can be used to generate their own 
-(simulated) outcome data, against which they can be scored using $S$. $S$ therefore commits a probabilistic forecast $F$
-to the ``self-assessment'' $\Ex [S(F, Y^F)]$, where $Y^F \sim F$ is the random variable defined by sampling from $F$, as well
-to an assessment $\Ex [S(G, Y^F)]$ of any alternative forecast $G$.
+Decision theoretically, probabilistic forecasts are a unique kind of action in that they can be used to generate their
+own(simulated) outcome data, against which they can be scored using $S$. $S$ therefore commits a probabilistic forecast
+$F$ to the ``self-assessment'' $\Ex [S(F, Y^F)]$, where $Y^F \sim F$ is the random variable defined by sampling from
+$F$, as well to an assessment $\Ex [S(G, Y^F)]$ of any alternative forecast $G$. For $S_{\log}$ this self-assessment is
+the Shannon entropy $H(F) = -\int \log(f(x))f(x)dx$ of $F$, and adding to $H(F)$ the Kullback-Leibler (KL) divergence $D_
+{KL}(F,G) = -\int \log(g(x)/f(x))f(x)dx$ gives $F$'s assessment $\Ex [S_{\log}(G, Y^F)]$ of $G$. (That is, $\Ex [S_
+{\log}(G, Y^F)] = H(F) + D_{KL}(F,G)$.) The KL divergence is the degree to which $F$ perceives $G$ as divergent from
+being able to minimize expected loss in this particular forecaster decision problem (i.e. maximize expected log-likelihood).
+
 A natural consistency criterion for $S$ is that it does not commit $F$ to assessing any other 
@@
 \begin{align}
-\Ex [S(F, Y^F)] \leq \Ex [S(G, Y^F)]
+\Ex [S(F, Y^F)] \leq \Ex [S(G, Y^F)] \label{eqn:prop_ineq}
 \end{align} 
@@
 $S$ is \emph{strictly proper} when this inequality is sharp, in which case the \emph{only} optimal decision for a forecaster is to state
-the forecast they believe to be true.
+the forecast they believe to be true. The logarithmic scoring rule $S_{\log}$, for example, is strictly proper due to the 
+positivity of KL divergence.
 
@@
 
-But given an auxiliary decision problem $\mathcal{D}_{\mathrm{Aux}} =\{\mathcal{X}, \mathcal{Y}, l(x,y)\}$, such as the
+But given an auxiliary decision problem $\mathcal{D}_{\mathrm{Aux}} = \{\mathcal{X}, \mathcal{Y}, l(x,y)\}$, such as the
 allocation problem from the main text or the cost-loss problem from Section \ref{sec:quantiles_shortage}, we can
@@
 
+The study of \emph{elicitability} also concerns this construction (see, e.g., \cite{gneiting2011making}), but from a
+different vantage point than ours. The question there is whether for a given map (called a \emph{functional} in
+statistics and a \emph{property} in computer science) $T(F)=x \in \mathcal{X}$ of distributions into the auxiliary
+action space, there exists a loss function $l$ such that $T(F) = x^F$ is the Bayes act for $l$ and the associated
+scoring rule $S_l(F,x)$ is strictly proper. Such an $l$ is said to \emph{elicit} $T$.  Our focus is rather on how
+forecast evaluation proceeds via the Bayes act construction for a given loss function of subject matter interest.
+
+\bibliography{allocation}
+
 \end{document}

### 2023-11-20 alloscore_manuscript/alloscore-application.Rnw
add proportionality explanation in supp and link to this in example 1

@@
 \usepackage{caption}
+\usepackage{xr-hyper}
 \usepackage{hyperref}
+\externaldocument{supplement}
 
@@
 \paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$ and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$.
-When the marginal forecasts are exponential distributions, it can be shown that the optimal allocation divides the available resources among the locations proportionally to the scale parameters $\sigma_i$ (see section *** of the supplemental materials).
+When the marginal forecasts are exponential distributions, it can be shown that the optimal allocation divides the available resources among the locations proportionally to the scale parameters $\sigma_i$ (see section
+\ref{sec:bayes-quantiles} of the supplemental materials).
 If we have $K = 5$ units of our resource available, the optimal allocation according to $F$ would be 1 unit of resources in location 1 and 4 units of resources in location 2.
@@
 From this point, the WIS or CRPS can be obtained by averaging across a range of decision making settings with different cost and loss parameters, using a similar motivation that we used to obtain the IAS from the AS in section \ref{sec:methods.detailed.integrated_allocation}.
+\apg{Refs to supp for quantile score derivation and detail about averaging.}
 %\ngr{I might suggest adding a detailed Step 1/2/3 as an appendix and referencing it here. But I think that adding that full detail here might be too much.}

### 2023-11-20 alloscore_manuscript/supplement.rnw
add proportionality explanation in supp and link to this in example 1

@@
 \section{Allocation Bayes acts as vectors of marginal quantiles.}
+\label{sec:bayes-quantiles}
 
-Here we explain why the Bayes act $x_i^{F,K}$ for the allocation
-problem \eqref{AP} can be represented as a vector of quantiles for the marginal forecast distributions $F_i$ at a
-single probability level $\tau^{F,K}$, that is, $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$.
+Here we show that the Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$, corresponding to the
+allocation problem \eqref{eqn:loss_fn} (in Section \ref{sec:methods.detailed.specific_allocation} in the main text) can
+be represented as a vector of quantiles for the marginal forecast distributions $F_i$ at a single probability level
+$\tau^{F,K}$, that is, $x_i^{F,K} = q_{F_i,\tau^{F,K}}$. An immediate consequence used in the examples in Section \ref
+{sec:methods.overview} in the main text is that if $F_i = \mathrm{Exp}(1/\sigma_i)$ for all $i$, then the Bayes act is
+proportional to $(\sigma_1,\ldots,\sigma_N)$, since $q_{\mathrm{Exp}(1/\sigma),\tau} = -\sigma \log(1-\tau)$.
 

### 2023-11-21 alloscore_manuscript/supplement.rnw
add Bayes risk mention and DeGroot and Berger to bib

@@
 $\mathcal{D}_{\mathrm{Aux}}$ for the Bayes act $x^F$; that is, the action which \emph{by design} will be assessed via
-the forecast $F$ to have an expected loss $\Ex[l(x^F,Y^F)]$ as low as possible.
+the forecast $F$ to have the lowest possible expected loss $\Ex[l(x^F,Y^F)]$. In statistical decision theory, 
+$\Ex[l(x^F,Y^F)]= \Ex_F[l(x^F,Y)]$ is 
+sometimes\footnote{A clear exposition of basic decision theory using this terminology is Chapter 8 of \cite{degroot2005optimal}.
+Unfortunately other influential sources such as \cite{berger2013statistical} use ``Bayes risk'' to refer to the expected loss
+for a general $x$.} 
+refered to as the \emph{Bayes risk} of $F$ relative to $\mathcal{D}_{\mathrm{Aux}}$.
 

### 2023-11-21 alloscore_manuscript/supplement.rnw
Make flashier to-do-list with updates from issues and fix RL derivative signs in sec 1

@@
 \usepackage{caption}
+\usepackage{soul}
 \usepackage{pgfplots}
@@
 \externaldocument{alloscore-application}
+
+\usepackage{enumitem}
+\newlist{todolist}{itemize}{2}
+\setlist[todolist]{label=$\square$}
+\usepackage{pifont}
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
+\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
+\hspace{-2.5pt}}
+\newcommand{\wontfix}{\rlap{$\square$}{\large\hspace{1pt}\xmark}}
 
@@
 
-\begin{itemize}
-	\item From 2.2.1, why are Bayes act scoring rules proper?
+\begin{todolist}
+	\item[\done] From 2.2.1, why are Bayes act scoring rules proper?
 	\item Explain ``All proper scoring rules for probabilistic forecasts have an explicit link to a loss function'' from discussion.
@@
 866; and if forecasts are ideal, then forecasts with better information always yield better decisions, ref Holzmann and Eulert, Corr 2.
-	\item For 2.2.2, how to get quantile representation of Bayes act using Lagrange multiplier, assuming smooth, never-zero densities well behaved at $x=0$.
-Work out exponetial example.
+	\item[\done] For 2.2.2, how to get quantile representation of Bayes act using Lagrange multiplier, \st{assuming smooth, never-zero densities well behaved at $x=0$.} Work out exponential example.
 Refer to methods paper for general case.
-	\item Derivation of quantile scoring rule with quantile as Bayes act for C/L problem, assuming never-zero densities. 
+	\item[\done] Derivation of quantile scoring rule with quantile as Bayes act for C/L problem, \st{assuming never-zero densities}. 
+	\item algorithmic details
+	\begin{todolist}
+		\item use of \texttt{distfromq} to get from quantiles to distribution functions
+		\item \texttt{alloscore}
+		\item implications for propriety.
+do quantiles elicited by \texttt{distfromq} $\hookrightarrow$ \texttt{alloscore} process align with ``real'' quantiles?
+the alloscore is proper if distribution functions F are handed to us; is it still proper given our algorithm situation?
+	\end{todolist}
+
 	\item Descriptions of 
-		\begin{itemize}		
+		\begin{todolist}		
 			\item CRPS as average quantile score across $C \in [0/L]$ decision problems
@@
 			\item WIS as average quantile score across 23 C/L problems.
-		\end{itemize}
+		\end{todolist}
 	\item Sketch of scoring for decision problems involving both cost and constraint.
-
-\end{itemize}
+\end{todolist}
 
@@
 \begin{align}
-D_{-}\Ex_F [(Y-x)_{+}] &= \lim_{h \searrow 0}\frac{1}{h}\Ex_F [(Y - x)_{+} - (Y - x - h)_{+}] \\
+D_{-}\Ex_F [(Y-x)_{+}] &= \lim_{h \searrow 0}\frac{1}{h}\Ex_F [(Y - x)_{+} - (Y - (x - h))_{+}] \\
 &= \lim_{h \searrow 0}\frac{1}{h}\int_{[x\mathrlap{-h, x]}} (x - h - y) dF(y) - 
@@
 &= F(x-) - 1  \label{eqn:dsfl} \\
-D_{+}\Ex_F [(Y-x)_{+}] &= \lim_{h \searrow 0}\frac{1}{h}\Ex_F [(Y - x + h)_{+} - (Y - x)_{+}] \\
+D_{+}\Ex_F [(Y-x)_{+}] &= \lim_{h \searrow 0}\frac{1}{h}\Ex_F [(Y - (x + h))_{+} - (Y - x)_{+}] \\
 &= \lim_{h \searrow 0}\frac{1}{h}\int_{[x\mathrlap{, x+h]}} (x - y) dF(y) - 

### 2023-11-26 alloscore_manuscript/supplement.rnw
more about quantile functions

@@
 We recall how quantiles arise as solutions to a probabilistic decision problem. Let $Y$ be a random variable
-representing the future level of an undesirable outcome such as severe COVID incidence. Let $x$ be a decision variable
+representing the future level of an undesirable outcome such as severe COVID incidence. Let $x \in \mathbb{R}$ be a decision variable
 representing the possible levels of some costly counter-measure, such as procurement of monoclonal antibody treatments,
-that can be taken at a cost $C$ per unit in preparation for $Y$.  A decision maker must decide on a level $x$ of investment in the
+that can be taken at a cost $C>0$ per unit in preparation for $Y$.  A decision maker must decide on a level $x$ of investment in the
 counter-measure, and wishes to avoid excesses in either the expediture $Cx$ or the shortage $(y-x)_{+}$ when $Y=y$ is
@@
 \end{align}
-Note that because $F(x)$ and $F(x-)$ are right and left continuous, repectively, the set $\{x \mid F(x) \geq \alpha\}$ is closed
+Note that because $F(x)$ and $F(x-)$ are right and left continuous, repectively, and $0<\alpha <1$, the set 
+$\{x \mid F(x) \geq \alpha\}$ is closed
 on the left and the set $\{x \mid \alpha \geq F(x-)\}$ is closed on the right.
@@
 the left and right quantiles of $F$ (for probability level $\alpha$) and any element $q_{\alpha, F} \in [q^{-}_{\alpha, F}, q^{+}_{\alpha, F}]$
-a quantile of $F$.  Thus $x^{\star}$ minimizes the expected loss \eqref{eqn:expQloss} and gives an optimal sultion to the decision problem
-if and only if it is a quantile $q_{\alpha, F}$.
+a quantile of $F$.  The \emph{quantile function} for $F$, which we write as either $Q_F(\alpha)$ or $F^{-1}(\alpha)$, is the 
+\emph{set-valued} function that maps $\alpha \in (0,1)$ to to the set $[q^{-}_{\alpha, F}, q^{+}_{\alpha, F}]$. 
+Thus $x^{\star}$ minimizes the expected loss \eqref{eqn:expQloss} and gives an optimal solution to the decision problem
+if and only if $x^{\star} \in Q_F(\alpha)$.
 
-Quantiles equivalently arise when the decision problem is defined in terms of the random \emph{oportunity} loss 
+\subsection{Quantile functions}
+\label{sec:quantile-functions}
+
+For future reference, we record several key properties of quantile functions.
+
+The probability levels $\{\alpha_i\}$ for which $\#Q_F(\alpha_i)>1$ form a discrete subset of $(0,1)$ and correspond to
+the non-zero width intervals $[q_{\alpha_i,F}^{-},q_{\alpha_i,F}^{+}]$ where $F$ is constant with values 
+$\{\alpha_i\}$.  Conversely, if $F$ is strictly increasing on a (Borel) set $A\subset \mathbb{R}$, then the restriction
+$Q_{F}\big|_{F(A)}=F^{-1}\big|_{F(A)}$ of $Q_F$ to $F(A)$ is in fact a real-valued function which is left-continuous
+and provides an inverse to $F$ on $F(A)$. If the the support $\mathrm{supp}(F) = \{x \mid 0<F(x)<1\}$ of $F$ is such an
+$A$, then extending $F^{-1}$ by $F^{-1}(0)=\inf(\mathrm{supp}(F))$ and $F^{-1}(1)=\sup(\mathrm{supp}(F))$ provides an
+inverse to $F$ on $F(\mathbb{R})\subset [0,1]$. If $F$ has a point mass at $x\in A$, so that
+$\{\alpha \mid F(x-)< \alpha < F(x)\}$ is a non-empty set disjoint from $F(A)$, then $Q_F$ takes the constant value $x$ 
+on the closure $\{\alpha \mid F(x-)\leq \alpha \leq F(x)\}$. Conversely, if $F$ has no discrete component on $A$, 
+then $Q_{F}\big|_{F(A)}$ is strictly increasing.
+
+Moreover, it can be said that $Q_F$ is increasing as a set-valued function on $\mathbb{R}$ in the generalized sense that 
+$(q_{\alpha, F} - q_{\beta, F})(\alpha-\beta) \geq 0$ whenever
+$q_{\alpha, F} \in Q_F(\alpha)$ and $q_{\beta, F} \in Q_F(\beta)$, that is, the set 
+$\mathrm{graph}(Q_F) = \{(\alpha, q) \mid q \in Q_F(\alpha)\}$ has no downward sloping secants. Conversely, given such an 
+increasing set-valued function $Q$ on $(0,1)$, we can construct a right-continuous function $F_Q$ from $\mathbb{R}$ to $[0,1]$ 
+which will be the cdf of the random variable $Y_Q := \min(Q(U))$ where $U \sim \mathrm{Unif}[0,1]$.
+
+\subsection{Opportunity relative to an oracle}
+Quantiles equivalently arise when the decision problem is defined in terms of the random \emph{opportunity} loss 
 \begin{align}
@@
 $\tau = 1 - \lambda/L$.
+
+The constraint now determines $\tau$ (and hence the Bayes act) through $\sum_{i=1}^{N} q_{\tau,F_i} = K.$
+This equation can be restated as the requirement that 
+\begin{align}
+K \in TQ_F(\tau) \label{eqn:tau-inclusion}
+\end{align}
+where the set-valued function
+\begin{align}
+TQ_F(\tau) := \sum_{i=1}^{N} Q_{F_i}(\tau) = \left[\sum_{i=1}^{N} q^{-}_{\tau,F_i}, \sum_{i=1}^{N} q^{+}_{\tau,F_i}\right]
+\end{align}
+is defined using interval addition $[a,b] + [c,d] = [a+c, b+d]$.
+
+$TQ_F$ satisfies the conditions mentioned in section \ref{sec:quantile-functions} for being a quantile function, and so there
+is a random variable $TY_F$ with cdf $F_{T} := F_{TQ_F}$.  From this perspective, the problem of finding $\tau$ becomes the calculation 
+of $F_{T}(K) = \mathbb{P}(TY_F \leq K)$, making clear the existence of a solution $\tau^{\star}$ to \eqref{eqn:tau-inclusion}.
+
+Two awkward features of the quantile representation of the Bayes act can arise.
+First, point masses in the $F_i$ create point masses for $F_T$ which may cause $\tau^{\star}$ to not be the unique solution to \eqref{eqn:tau-inclusion}.
+Secondly, if more than
+one $Q_{F_i}(\tau^{\star})$ is a positive-width interval, then the Bayes act will not be unique in these coordinates, and generically
+not all points in the $Q_{F_i}(\tau^{\star})$ will be the coordinate of a Bayes act.
+
+It is important to note that $\lambda$ depends on the forecast $F$ and the constraint level $K$. Thus while $\lambda = L(1-\tau)$ can be
+interpreted as a kind of ``cost'' imposed by the constraint in the allocation problem which is analogous to $C = L(1-\alpha)$ in the  
+the cost-lost problem of section \ref{sec:quantiles_shortage}, it does not serve to define the allocation loss function in the way that $C$ defines \eqref{eqn:expQloss}.
+$\lambda$ is rather a parameter that must be found, given the pair $F$ and $K$.
 
@@
 sometimes\footnote{A clear exposition of basic decision theory using this terminology is Chapter 8 of \cite{degroot2005optimal}.
-Unfortunately other influential sources such as \cite{berger2013statistical} use ``Bayes risk'' to refer to the expected loss
-for a general $x$.} 
+Unfortunately other influential sources use ``Bayes risk'' to refer to a wider variety of quantities. In 
+\cite{berger2013statistical} for example, the Bayes risk can be both the optimal expected loss 
+for $F$ (see p. 17) or the expected loss for a general $x$ with respect to $F$ (see p. 6).} 
 refered to as the \emph{Bayes risk} of $F$ relative to $\mathcal{D}_{\mathrm{Aux}}$.

### 2023-11-27 alloscore_manuscript/alloscore-application.Rnw
minor grammatical changes to text, plus a clean knit with data stores working.

@@
 This work seeks to address that gap.
-However, we do note that the decision-making context presented in this work, while motivated by examples from real-world public health resource allocation problems, has not used to inform an actual real-time decision-making process.
+However, we do note that the decision-making context presented in this work, while motivated by examples from real-world public health resource allocation problems, has not been used to inform an actual real-time decision-making process.
 
@@
 First, it is clear that different metrics (in our case allocation score and mean weighted interval score) captured different aspects of forecast performance.
-The MWIS was strongly dependent on the scale of the forecasted quantity (e.g., when hospitalizations were high, so was WIS).
+The mean WIS (MWIS) was strongly dependent on the scale of the forecasted quantity (e.g., when hospitalizations were high, so were MWIS scores).
 The allocation score was not as scale-dependent, as the highest/worst allocation scores were observed when the number of hospitalizations was closest to the allocation constraint.
@@
 Using standard forecast evaluation metrics can mask the utility of certain forecasts, or lead to forecasts being used to inform a particular decision that are not optimal for that particular set of values that are important for that context.
-Further collaborative work is needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or simulation exercises that could be used to inform further development of new or alternative scoring metrics.
+New collaborative work between public health officials and modeling teams is needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or simulation exercises that could be used to inform further development of new or alternative scoring metrics.
 

### 2023-11-29 alloscore_manuscript/alloscore-application.Rnw
pass at updating application.

@@
   mutate(target_end_date = as.Date(reference_date) + 14)
+
+all_scores <- left_join(alloscore_data_summaries, score_data_summaries,
+                      by=c("model", "target_end_date"))
+@
+
+\subsubsection{Anatomy of forecast scores for one week}
+
+To illustrate the mechanics of allocation scoring, we start by focusing on how forecasts generated on or before December 20, 2021, with predictions for January 3, 2022, were scored by different metrics.
+This week was around the Omicon wave peak nationally, with individual states typically observing a peak at or after January 3, 2022.
+
+For many locations, forecasts were \ngr{fill in details here}.
+On this date, of the 10 models evaluated, the CU-select model had the most accurate forecasts according to the allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table \ref{tab:one-week-results}).
+The JHUAPL-Bucky model had the second best MWIS but the third worst allocation score.
+
+<<metrics-rankings, results = "asis">>=
+## for highlighting cells, if we want that https://stackoverflow.com/questions/37448348/color-code-cells-of-xtable
+all_scores |>
+  filter(target_end_date == "2022-01-03", !is.na(mwis)) |>
+  select(model, alloscore, allo_rank, mwis, mwis_rank) |>
+  arrange(allo_rank) |>
+  xtable::xtable(label = "tab:one-week-results",
+                 caption = "caption TBD") |>
+  print(include.rownames = FALSE)
+@
+
+Forecasts and allocations for the nine states with the highest numbers of hospitalizations on January 3, 2022 reveal mechanics about how allocations are made (Figure \ref{fig:thermometer-plot}).
+\ngr{Aaron - can you add this file to the Sweave doc? Maybe just make a target for it so the figure is generated outside the latex?}
+As described in the methods above, allocations for a given location from a specific model are generated by finding the single quantile of the probabilistic forecast across all locations that reaches the resource allocation limit (in this case, assumed to be 15,000).
+A direct comparison between the forecasts from the JHUAPL-Bucky and USC-SI\_kJalpha models show that while the JHUAPL-Bucky forecast distributions were closer to the truth across all states, the allocations suggested by those forecasts often were more inefficient than those from the USC-SI\_kJalpha model (Figure \ref{fig:thermometer-plot}).
+
+<<multi-loc-ranks>>=
+## date points to target_end_date of 2022-01-10 to highlight
+ref_date_for_plot <- "2021-12-20"
+models_to_plot <- c("JHUAPL-Bucky", "CU-select")
+
+# tar_load(forecast_data)
+# forecast_dat_to_merge <- forecast_data |>
+#   dplyr::filter(reference_date == ref_date_for_plot,
+#          model %in% c("JHUAPL-Bucky", "MUNI-ARIMA"),
+#          quantile == "0.5") |>
+#   select(model, reference_date, state = abbreviation, quantile, median = value) |>
+#   mutate(reference_date = as.character(reference_date))
+
+multiloc_dat <- Kat15k_alloscores |>
+  filter(reference_date == ref_date_for_plot,
+         ## the above date points to target_end_dateof 2022-01-10 where things were really wrong
+         model %in% models_to_plot) |>
+  rename(allocation = x, observation = y, state = abbreviation) |>
+  select(model, reference_date, state, allocation, observation) |>
+#  left_join(forecast_dat_to_merge) |>
+#  pivot_longer(cols = c("allocation", "observation", "median"))
+  pivot_longer(cols = c("allocation", "observation"))
+
+## plot all locations observations and allocations
+multiloc_plot <- ggplot(multiloc_dat, aes(x=name, y=value, group=state)) +
+  geom_line() +
+  geom_line(data=dplyr::filter(multiloc_dat, state=="FL"), color="orange") +
+  ylab("Daily hospital admissions") +
+  scale_x_discrete(name=NULL, limits=rev, expand = c(.1,.1)) +
+  facet_grid(.~model)
+
+oracle_rows <- Kat15k_alloscores |>
+  filter(reference_date == ref_date_for_plot,
+         model %in% c("JHUAPL-Bucky")) |>
+  rename(observation = y, state = abbreviation, allocation = oracle) |>
+  select(reference_date, state, allocation, observation) |>
+  pivot_longer(cols = c("allocation", "observation")) |>
+  mutate(model = "Oracle")
+
+multiloc_dat |>
+  bind_rows(oracle_rows) |>
+  mutate(state = reorder(state, -ifelse(name =="observation", value, 0)),
+         name = factor(name, levels = c("observation", "allocation"))) |>
+  arrange(name) |>
+  ggplot(aes(x=state, y=value, fill=name)) +
+  scale_fill_manual(values = c("#CD2626", "goldenrod"))+
+  geom_col(position="identity", alpha=.7) +
+  facet_grid(model~.)
 @
@@
 Models show different degrees of association between the two metrics.">>=
-all_data <- left_join(alloscore_data_summaries, score_data_summaries,
-                      by=c("model", "target_end_date"))
-
-all_data |>
+all_scores |>
   filter(as.Date(reference_date.x) < as.Date("2022-02-01") ) |>
@@
 
+Models that were able to provide more accurate relative rankings of locations according to the hospitalization counts at the future horizon in general showed better performance according to the allocation score (Figure \ref{fig:multi-loc-ranks}.
+
 \subsubsection{Integrated allocation score across values of K}
@@
  - something here from Ben's analysis?
-
-\ngr{include figure here that shows relative rankings of locations?}
 

### 2023-11-30 R/plot_functions.R
making mvbucky thermo a target

@@ -1,4 +1,4 @@
-## noting that these functions have been adapted from talks/ISF_talk/plotting_code/plot_functions.R
+
 
 
 get_intervals <- function(df) {
@@ -19,8 +19,6 @@ pal1 = c("mediumblue", "goldenrod", "#CD2626")
 
 
 plot_hosp <- function(
-    forecast_data,
-    truth_data,
     start_date = "2021-09-01",
     stop_date = "2022-01-31",
     f_date = "2021-12-27",
@@ -30,13 +28,18 @@ plot_hosp <- function(
     f_alpha = .5,
     f_width1 = 2,
     space = .5,
-    models = "COVIDhub-4_week_ensemble",
+    models = "COVIDhub-ensemble",
     locations = "US",
     geofacet = FALSE,
     key_width = .15,
     allocations = FALSE,
+    one_K = NULL,
+    forecasts_hosp = forecasts_hosp,
+    alloscores,
+    truth,
     free_y = FALSE
 ) {
+  loc_abbrevs <- locations
   locations <- locations %>% purrr::map(
     function(loc) {
       if (grepl("^[A-Z]{2}$", loc)) {
@@ -50,13 +53,13 @@ plot_hosp <- function(
   total_width <- (length(models) - 1)*(f_width1 + space) + f_width1
   half_width <- total_width/2
 
-  truth <- truth_data %>%
+  truth <- truth %>%
   dplyr::filter(location %in% locations,
     target_end_date >= start_date,
     target_end_date <= stop_date) %>%
   dplyr::mutate(
     validation = ifelse(target_end_date > as.Date(f_date), "Validation", "Historical"),
-    code = toupper(geo_value),
+    code = as.factor(toupper(geo_value)),
     xmin = target_end_date - half_width - space,
     xmax = xmin + total_width + 2 * space)
 
@@ -64,20 +67,23 @@ plot_hosp <- function(
     Date = as.Date(c(f_date, te_date)),
     Date_name = c("Forecast Date, 2021-12-27", "Target Date, 2022-01-10"))
 
-  fc_dat <- forecast_data %>%
-    dplyr::filter(location %in% locations, model %in% models, target_end_date == te_date) %>%
+  fc_dat <- forecasts_hosp %>%
+    dplyr::filter(location %in% locations, model %in% models) %>%
     dplyr::mutate(
-      code = toupper(geo_value),
+      code = as.factor(toupper(geo_value)),
[…]

### 2023-11-30 _targets.R

@@ -26,7 +26,8 @@ tar_source(files = c("R/data-ingestion.R",
                      "R/plot-alloscores.R",
                      "R/run-alloscore.R",
                      "R/determine-model-eligibility.R",
-                     "R/exponential-examples.R"))
+                     "R/exponential-examples.R",
+                     "R/plot_functions.R"))
 
 values <- tibble(forecast_dates = as.character(seq.Date(as.Date("2021-11-22"), as.Date("2022-02-28"), by = "7 days")))
 
@@ -95,4 +96,9 @@ combined <- tar_combine(
 #   format = "file"
 # )
 
-list(setup, mapped, combined)
+plots <- tar_target(
+  name = mvbucky_dist_alloc,
+  command = plot_thermo(all_alloscore_data, forecast_data, truth_data)
+)
+
+list(setup, mapped, combined, plots)

### 2023-11-30 R/plot_functions.R
plot_hosp date axis/legend fix

@@ -64,8 +64,8 @@ plot_hosp <- function(
     xmax = xmin + total_width + 2 * space)
 
   target_dates <- data.frame(
-    Date = as.Date(c(f_date, te_date)),
-    Date_name = c("Forecast Date, 2021-12-27", "Target Date, 2022-01-10"))
+    Date = as.Date(c(f_date, te_date))) %>% 
+    dplyr::mutate(Date_name = paste0(c("Forecast Date, ", "Target Date, "), Date))
 
   fc_dat <- forecasts_hosp %>%
     dplyr::filter(location %in% locations, model %in% models) %>%
@@ -157,7 +157,8 @@ plot_hosp <- function(
     scale_alpha_manual(
       values = c("95%" = .3, "80%" = .6, "50%" = .9),
       name = "Predictive Interval") +
-    scale_x_date(date_labels = "%b %y") +
+    scale_x_date(date_labels = "%b %m '%y", 
+                 breaks = as.Date(lubridate::floor_date(as.Date(te_date), unit = "month"))) +
     coord_cartesian(xlim = as.Date(c(start_date, stop_date)),
                     ylim = c(0, 1.3*max(truth$value))) +
     xlab("Date") +
@@ -223,7 +224,7 @@ plot_hosp <- function(
   p <- p + theme(
       legend.key.width = unit(f_width1*key_width, "cm"),
       panel.spacing = unit(0.1, "lines"),
-      axis.text.x = element_text(hjust = -0.1))
+      axis.text.x = element_text(hjust = -0.05))
   p
 }
 

### 2023-11-30 R/plot_functions.R
updates to thermometer plot functions.

@@ -39,22 +39,23 @@ plot_hosp <- function(
     truth,
     free_y = FALSE
 ) {
-  loc_abbrevs <- locations
-  locations <- locations %>% purrr::map(
-    function(loc) {
-      if (grepl("^[A-Z]{2}$", loc)) {
-        hub_locations %>%
-          dplyr::filter(abbreviation == loc, nchar(fips) == 2) %>% pull(fips)
-      } else {
-        sprintf("%02d", as.integer(loc))
-      }
-    })
+  ## removing below as it adds a dependency on hub_locations and I don't think is necessary if we use the abbreviations
+  # loc_abbrevs <- locations
+  # locations <- locations %>% purrr::map(
+  #   function(loc) {
+  #     if (grepl("^[A-Z]{2}$", loc)) {
+  #       hub_locations %>%
+  #         dplyr::filter(abbreviation == loc, nchar(fips) == 2) %>% pull(fips)
+  #     } else {
+  #       sprintf("%02d", as.integer(loc))
+  #     }
+  #   })
   # compute aggregate rect and segment widths.
   total_width <- (length(models) - 1)*(f_width1 + space) + f_width1
   half_width <- total_width/2
 
   truth <- truth %>%
-  dplyr::filter(location %in% locations,
+  dplyr::filter(abbreviation %in% locations,
     target_end_date >= start_date,
     target_end_date <= stop_date) %>%
   dplyr::mutate(
@@ -65,21 +66,21 @@ plot_hosp <- function(
 
   target_dates <- data.frame(
     Date = as.Date(c(f_date, te_date)),
-    Date_name = c("Forecast Date, 2021-12-27", "Target Date, 2022-01-10"))
+    Date_name = c(paste("Forecast Date,", f_date), paste("Target Date,", te_date)))
 
   fc_dat <- forecasts_hosp %>%
-    dplyr::filter(location %in% locations, model %in% models) %>%
+    dplyr::filter(abbreviation %in% locations, model %in% models) %>%
     dplyr::mutate(
       code = as.factor(toupper(geo_value)),
       model = forcats::fct_relevel(model, models),
       xmin = target_end_date - half_width + (match(model, models) - 1)*(f_width1 + space),
       xmax = xmin + f_width1)
-  
-  fc_dat <- fc_dat %>% mutate(
-    code = fct_relevel(code, loc_abbrevs))
-  truth <- truth %>% mutate(
[…]

### 2023-12-01 alloscore_manuscript/alloscore-application.Rnw
updates to application reorg

@@
 tar_load("Kat15k_alloscores")
+tar_load("score_data")
+tar_load("truth_data")
+tar_load("forecast_data")
 
@@
   mutate(target_end_date = as.Date(reference_date) + 14)
-
-tar_load("score_data")
 
@@
 all_scores |>
-  filter(target_end_date == "2022-01-03", !is.na(mwis)) |>
+  filter(target_end_date == "2022-01-10", !is.na(mwis)) |>
   select(model, alloscore, allo_rank, mwis, mwis_rank) |>
@@
 Forecasts and allocations for the nine states with the highest numbers of hospitalizations on January 3, 2022 reveal mechanics about how allocations are made (Figure \ref{fig:thermometer-plot}).
-\ngr{Aaron - can you add this file to the Sweave doc? Maybe just make a target for it so the figure is generated outside the latex?}
 As described in the methods above, allocations for a given location from a specific model are generated by finding the single quantile of the probabilistic forecast across all locations that reaches the resource allocation limit (in this case, assumed to be 15,000).
 A direct comparison between the forecasts from the JHUAPL-Bucky and USC-SI\_kJalpha models show that while the JHUAPL-Bucky forecast distributions were closer to the truth across all states, the allocations suggested by those forecasts often were more inefficient than those from the USC-SI\_kJalpha model (Figure \ref{fig:thermometer-plot}).
+\ngr{add in some specific interpretations from the figure.}
+
+<<thermometer-plot, fig.height=3.5, fig.width=7.5, fig.cap="Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the nine highest hospitalization counts.
+For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts.
+The side-by-side shaded regions show the median (solid horizontal line) and 50%, 80% and 95% prediction intervals for the two selected models.
+The forecasts were made for new hospitalizations on January 3, 2022 (vertical dashed line, with number of hospitalizations indicated by red horizontal line segment at the intersection of the dashed line and the grey line of data).
+The vertical bars with red and yellow shading show the allocations: the red bar goes from zero to the level of the observed number of hospitalizations while the yellow bar shows how many resources were suggested by the model to be allocated to that location.
+The allocations (yellow bar) may be more or less than the observed number, therefore exposed red bar indicates unmet need, while the yellow bar extending above the observed number of hospitalizations means that excess resources were suggested for that given location.">>=
+source("R/plot_functions.R")
+tar_load(alloscore_2021.12.20)
+plot_hosp(models = c("JHUAPL-Bucky", "CU-select"),
+          start_date = "2021-12-01",
+          stop_date = "2022-01-25",
+          f_date = "2021-12-20",
+          te_date = "2022-01-03",
+          ## highest locations in terms of hospitalizations
+          locations = c("FL", "NY", "TX", "CA"),
+          f_width1 = 4, f_alpha = .4,
+          allocations = TRUE,
+          forecasts_hosp = forecast_data %>%  filter(reference_date == "2021-12-20"),
+          alloscores = alloscore::slim(alloscore_2021.12.20, id_cols = c("model", "reference_date")),
+          truth = truth_data,
+          f_colors = c("#4A708B", "#CD2626"),
+          one_K = 15000) + labs(x = NULL)
+@
 
@@
 <<grob-hosps-over-time>>=
-tar_load("truth_data")
 dat_to_plot <- truth_data |>

### 2023-12-01 R/plot_functions.R

@@ -158,7 +158,7 @@ plot_hosp <- function(
     scale_alpha_manual(
       values = c("95%" = .3, "80%" = .6, "50%" = .9),
       name = "Predictive Interval") +
-    scale_x_date(date_labels = "%b %d '%y") +
+    scale_x_date(date_labels = "%b %d") +
     coord_cartesian(xlim = as.Date(c(start_date, stop_date)),
                     ylim = c(0, 1.3*max(truth$value))) +
     xlab("Date") +
@@ -209,9 +209,10 @@ plot_hosp <- function(
       scale_y_continuous(expand = c(0,0))
   }
   if (geofacet) {
+    require(geofacet)
     p <- p + facet_geo(~ code, grid = geofacet::us_state_grid2) +
     scale_x_date(breaks = as.Date(c("2021-12-01", "2022-01-01")),
-                 date_labels = "%b %d '%y") +
+                 date_labels = "%b %d") +
     theme(axis.title = element_blank())
   } else if (length(locations) > 1) {
     if (free_y) {
@@ -223,7 +224,8 @@ plot_hosp <- function(
   p <- p + theme(
       legend.key.width = unit(f_width1*key_width, "cm"),
       panel.spacing = unit(0.1, "lines"),
-      axis.text.x = element_text(hjust = -0.05))
+      #axis.text.x = element_text(hjust = -0.05)
+      )
   p
 }
 

### 2023-12-01 _targets.R
removing a few figure files and updating _targets.

@@ -96,9 +96,5 @@ combined <- tar_combine(
 #   format = "file"
 # )
 
-plots <- tar_target(
-  name = mvbucky_dist_alloc,
-  command = plot_thermo(all_alloscore_data, forecast_data, truth_data)
-)
 
-list(setup, mapped, combined, plots)
+list(setup, mapped, combined)

### 2023-12-01 alloscore_manuscript/supplement.rnw
write down TQ equation to prepare for alloscore algo discussion

@@
 is a random variable $TY_F$ with cdf $F_{T} := F_{TQ_F}$.  From this perspective, the problem of finding $\tau$ becomes the calculation 
-of $F_{T}(K) = \mathbb{P}(TY_F \leq K)$, making clear the existence of a solution $\tau^{\star}$ to \eqref{eqn:tau-inclusion}.
+of $F_{T}(K) = \mathbb{P}(TY_F \leq K)$, making clear the existence of a solution $\tau^{\star}$ to \eqref{eqn:tau-inclusion}. This also
+yields the interesting formal expression of the Bayes act
+\begin{align}
+x^{F,K} = (F_1,\ldots,F_N)^{-1}(\mathbf{1}_N F_{T}(K)),
+\end{align}
+where $\mathbf{1}_N: \mathbb{R} \to \mathbb{R}^N$ is the linear map that takes $a$ to the $N$-vector $(a,\ldots,a)^T$ and 
+$(F_1,\ldots,F_N)^{-1}$ is a map from $\mathbb{R}^N$ to $\mathbb{R}^N$. We can also arrange this as
+\begin{align}
+TQ_F\left(\frac{1}{N} \mathbf{1}^T (F_1,\ldots,F_N)(x^{F,K})\right) = K,
+\end{align}
+which leads conceptually to the iterative numerical method of solving for $x^{F,K}$ discussed next in section \ref{sec:numeric}.
 
@@
 $\lambda$ is rather a parameter that must be found, given the pair $F$ and $K$.
+
+\section{Numerical computation of allocation scores}
+\label{sec:numeric}
 

### 2023-12-01 alloscore_manuscript/alloscore-application.Rnw
manuscript and figure updates

@@
 
-To illustrate the mechanics of allocation scoring, we start by focusing on how forecasts generated on or before December 20, 2021, with predictions for January 3, 2022, were scored by different metrics.
+To illustrate the mechanics of allocation scoring, we start by focusing on how forecasts generated on or before December 20, 2021, with predictions for January 03, 2022, were scored by different metrics.
 This week was around the Omicon wave peak nationally, with individual states typically observing a peak at or after January 3, 2022.
 
-For many locations, forecasts were \ngr{fill in details here}.
-On this date, of the 10 models evaluated, the CU-select model had the most accurate forecasts according to the allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table \ref{tab:one-week-results}).
+For many locations, forecasts predicted lower values than were eventually observed, as this was during the period of steep increase of viral transmission across many states.
+Of the 10 models evaluated, the CU-select model had the most accurate forecasts according to the allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table \ref{tab:one-week-results}).
 The JHUAPL-Bucky model had the second best MWIS but the third worst allocation score.
@@
 all_scores |>
-  filter(target_end_date == "2022-01-10", !is.na(mwis)) |>
+  filter(target_end_date == "2022-01-03", !is.na(mwis)) |>
   select(model, alloscore, allo_rank, mwis, mwis_rank) |>
@@
 
-Forecasts and allocations for the nine states with the highest numbers of hospitalizations on January 3, 2022 reveal mechanics about how allocations are made (Figure \ref{fig:thermometer-plot}).
+Forecasts and allocations for a selection of states with high numbers of hospitalizations on January 3, 2022 reveal mechanics about how allocations are made (Figure \ref{fig:thermometer-plot}).
 As described in the methods above, allocations for a given location from a specific model are generated by finding the single quantile of the probabilistic forecast across all locations that reaches the resource allocation limit (in this case, assumed to be 15,000).
-A direct comparison between the forecasts from the JHUAPL-Bucky and USC-SI\_kJalpha models show that while the JHUAPL-Bucky forecast distributions were closer to the truth across all states, the allocations suggested by those forecasts often were more inefficient than those from the USC-SI\_kJalpha model (Figure \ref{fig:thermometer-plot}).
-\ngr{add in some specific interpretations from the figure.}
+A direct comparison between the forecasts from the JHUAPL-Bucky and CU-select models show that while the JHUAPL-Bucky forecast distributions were closer to the eventual observations in many states, the allocations suggested by those forecasts often were more inefficient than those from the CU-select model (Figure \ref{fig:thermometer-plot}).
+For JHUAPL-Bucky, the allocations were assigned around the 25th quantile of the predictive distribution and for CU-select it was around the 90th percentile. \ngr{could we make these more specific?}
+This reflects that in general the JHUAPL-Bucky forecasts assigned more probabilities to higher values and thus a low quantile value across all locations reached the allocation limit of 15,000.
+In California, New York and Texas, the eventually observed number of hospital admissions was closer to the predicted median from JHUAPL-Bucky and was contained in JHUAPL-Bucky's 80\% PI.
+However for each of these three states, the resource allocation was lower and therefore resulted in more hypothetical unmet need than CU-select allocations.
+In Florida, the forecast from JHUAPL-Bucky missed by a wide margin (as did the forecast from CU-select) and the allocation was nearly XX units lower than what was allocated according to the CU-select model.
+\ngr{add specific number here}
 
@@
 
-<<multi-loc-ranks>>=
+In the example of this one week, JHUAPL-Bucky had a worse allocation score than CU-select because its forecasts led to allocations that sent excess resources to several states, such as Ohio, Pennsylvania, and Michigan, that would have been more effectively allocated to states that did receive enough resources to meet their needs, such as Florida and California (Figure \ref{multi-loc-ranks}A).
+These allocation errors resulted because that model’s forecasts did not consistently capture the relative resource needs across different states.
+(Figure \ref{fig:thermometer-plot} \& \ref{multi-loc-ranks}A).
+The CU-select model made some similar errors — most prominently, over-allocating resources to Ohio — but overall, it did a better job of forecasting the relative resource demands across different locations.
+
+On the other hand, CU-select had worse performance as measured by WIS.
+Its forecasts were biased downwards, and it consistently incurred a large penalty for underprediction (Figure \ref{fig:multi-loc-rank}B).
+Predictions from the JHUAPL-Bucky model were wider, and included the observed level of daily hospital admissions more often.
+Therefore, WIS did not as often assign that model severe penalties for forecasts that underpredicted or overpredicted the actual hospitalization count.
+
+Together, these results corroborate the understanding that the allocation score and weighted interval score measure different aspects of forecast performance.
+The allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.
+
+<<multi-loc-ranks, fig.cap="Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select).
+Panel A shows the observed resource need, in this case the observed number of hospitalizations, for each state, along with the hypothetical number of resources allocated to the given location based on the forecasts from each model.
+The number of available resources was fixed at 15,000 and forecasts from each model were used to determine an optimal allocation strategy before the resource need was known.
+For most locations the resource need exceeded the resources allocated, indicated by the 'observed resource need' bar being larger than the 'allocation' bar.
+Panel B shows the breakdown of the weighted interval score (WIS) into components of underprediction, overprediction and dispersion.
+Larger values of WIS indicate more error, and the full WIS score for each location can be decomposed into the three components shown here.">>=
 ## date points to target_end_date of 2022-01-10 to highlight
@@
 
-# tar_load(forecast_data)
-# forecast_dat_to_merge <- forecast_data |>
-#   dplyr::filter(reference_date == ref_date_for_plot,
-#          model %in% c("JHUAPL-Bucky", "MUNI-ARIMA"),
-#          quantile == "0.5") |>
-#   select(model, reference_date, state = abbreviation, quantile, median = value) |>
-#   mutate(reference_date = as.character(reference_date))
-
 multiloc_dat <- Kat15k_alloscores |>
   filter(reference_date == ref_date_for_plot,
-         ## the above date points to target_end_dateof 2022-01-10 where things were really wrong
+         ## the above date points to target_end_date of 2022-01-03 where things were really wrong
          model %in% models_to_plot) |>
-  rename(allocation = x, observation = y, state = abbreviation) |>
-  select(model, reference_date, state, allocation, observation) |>
-#  left_join(forecast_dat_to_merge) |>
-#  pivot_longer(cols = c("allocation", "observation", "median"))
-  pivot_longer(cols = c("allocation", "observation"))
+  rename(allocation = x, `observed resource need` = y, state = abbreviation) |>
+  select(model, reference_date, state, allocation, `observed resource need`) |>
+  pivot_longer(cols = c("allocation", "observed resource need"))
 
 ## plot all locations observations and allocations
-multiloc_plot <- ggplot(multiloc_dat, aes(x=name, y=value, group=state)) +
-  geom_line() +
-  geom_line(data=dplyr::filter(multiloc_dat, state=="FL"), color="orange") +
-  ylab("Daily hospital admissions") +
-  scale_x_discrete(name=NULL, limits=rev, expand = c(.1,.1)) +
-  facet_grid(.~model)
 
-oracle_rows <- Kat15k_alloscores |>
-  filter(reference_date == ref_date_for_plot,
-         model %in% c("JHUAPL-Bucky")) |>
-  rename(observation = y, state = abbreviation, allocation = oracle) |>
-  select(reference_date, state, allocation, observation) |>
-  pivot_longer(cols = c("allocation", "observation")) |>
-  mutate(model = "Oracle")
+## line plot below here
+# multiloc_plot <- ggplot(multiloc_dat, aes(x=name, y=value, group=state)) +
+#   geom_line() +
+#   geom_line(data=dplyr::filter(multiloc_dat, state=="FL"), color="orange") +
+#   ylab("Daily hospital admissions") +
+#   scale_x_discrete(name=NULL, limits=rev, expand = c(.1,.1)) +
+#   facet_grid(.~model)
 
-multiloc_dat |>
-  bind_rows(oracle_rows) |>
-  mutate(state = reorder(state, -ifelse(name =="observation", value, 0)),
-         name = factor(name, levels = c("observation", "allocation"))) |>
+ordered_levels <- truth_data |>
+  filter(target_end_date == as.Date(ref_date_for_plot)+14, location != "US") |>
+  select(location, abbreviation, value) |>
+  arrange(-value)
+
+alloc_plot <- multiloc_dat |>
+  mutate(state = factor(state, levels = ordered_levels$abbreviation),
+         name = factor(name, levels = c("observed resource need", "allocation"))) |>
   arrange(name) |>
   ggplot(aes(x=state, y=value, fill=name)) +
-  scale_fill_manual(values = c("#CD2626", "goldenrod"))+
+  scale_fill_manual(name=NULL, values = c("#CD2626", "goldenrod")) +
+  theme(axis.text.x=element_text(angle=90,vjust=.5), legend.position = c(.9,.95), legend.justification = c(1,1)) +
+  ylab("Resource need and allocation") +
+  xlab(NULL) +
   geom_col(position="identity", alpha=.7) +
   facet_grid(model~.)
+
+wis <- covidHubUtils::score_forecasts(
+  forecast_data |> filter(reference_date == "2021-12-20", str_detect(model,"Buck|CU-select")),
+  truth_data,
+  use_median_as_point = TRUE,
+  metrics = "wis_components",
+  return_format="long") |>
+  filter(location != "US", score_name != "true_value") |>
+  left_join(select(ordered_levels, location, abbreviation)) |>
+  mutate(abbreviation = factor(abbreviation, levels=ordered_levels$abbreviation),
+         `WIS component` = factor(score_name, levels = c("overprediction", "dispersion", "underprediction")))
+
+## palette from
+## colorspace::divergingx_hcl(palette = "Blue-Yellow 3", n=3)
+
+wis_plot <- ggplot(data = wis, aes(x=abbreviation)) +
+  geom_col(aes(y=score_value, fill=`WIS component`)) +
+  scale_fill_manual(values = c("#9FA2FF", "gray77", "#BAAE00")) +
+  facet_grid(model~.) +
+  theme(axis.text.x=element_text(angle=90,vjust=.5), legend.position = c(.9,.95), legend.justification = c(1,1)) +
+  xlab(NULL) +
+  ylab("WIS")
+
+cowplot::plot_grid(alloc_plot, wis_plot,
+                   nrow=2,
+                   align="v",
+                   rel_heights = c(1, 1),
+                   labels = "AUTO")
+
 @

### 2023-12-01 alloscore_manuscript/alloscore-application.Rnw
adding some content for multi-k comparisons #42

@@
 tar_load(alloscore_2021.12.20)
+slim_alloscore_2021.12.20 <- alloscore::slim(alloscore_2021.12.20, id_cols = c("model", "reference_date"))
 plot_hosp(models = c("JHUAPL-Bucky", "CU-select"),
@@
           forecasts_hosp = forecast_data %>%  filter(reference_date == "2021-12-20"),
-          alloscores = alloscore::slim(alloscore_2021.12.20, id_cols = c("model", "reference_date")),
+          alloscores = slim_alloscore_2021.12.20,
           truth = truth_data,
@@
 For predictions made for the first three Mondays in December 2021 and the last three Mondays in February 2022 all models had allocation scores under 500 (and the mean across all models was less than 100), indicating that the unnecessary unmet need was fairly low relative to the total number of hospital admissions on those days.
-The allocation scores are on the whole highest when the observed number of new hospital admissions is closest to the resource threshold of 15,000, as this is the time when any mistakes in allocation are costly in terms of wasting resources in one location that could have been used in another.
+The allocation scores were on the whole highest when the observed number of new hospital admissions was closest to the resource threshold of 15,000, as those are the times when any mistakes in allocation are costly in terms of wasting resources in one location that could have been used in another.
 Predictions made during the peak week and just after showed the highest variation in allocation scores, with some models having allocation scores under 1000 and others having values over 3500.
@@
 
-\subsubsection{Model rankings for different scores did not reliably correlate}
+\subsubsection{Metrics were not consistently correlated over time}
 
@@
 
-Models that were able to provide more accurate relative rankings of locations according to the hospitalization counts at the future horizon in general showed better performance according to the allocation score (Figure \ref{fig:multi-loc-ranks}.
-
 \subsubsection{Integrated allocation score across values of K}
@@
  - something here from Ben's analysis?
+
+<<ias-comparison, results='asis'>>=
+ias <- slim_alloscore_2021.12.20 |>
+  group_by(model, K) |>
+  summarize(`Allocation score` = sum(components)) |>
+  group_by(model) |>
+  summarize(IAS = mean(`Allocation score`)) |>
+  arrange(IAS)
+
+as <- slim_alloscore_2021.12.20 |>
+  filter(K == 15000) |>
+  group_by(model) |>
+  summarize(AS = sum(components)) |>
+  arrange(AS)
+
+left_join(ias, as) |>
+  xtable::xtable() |>
+  print(include.rownames = FALSE)
+@
+
+<<multi-k, fig.height=4>>=
+slim_alloscore_2021.12.20 |>
+  group_by(model, K) |>
+  summarize(`Allocation score` = sum(components)) |>
+  mutate(model = factor(model, levels = ias$model)) |>
+  ggplot(aes(x=K, y=`Allocation score`, color=model, linetype=model)) +
+  geom_line()
+@
 

### 2023-12-03 alloscore_manuscript/supplement.rnw
some minor updates to supplement

@@
 
+\section{Introduction}
+\label{sec:intro}
+
+We briefly address some technical and methodological points in the main text.
+We begin in section \ref{sec:shortage} by formalizing the concept of a \emph{shortage} of resources and giving some key results about expected resource shortages under a distribution characterizing uncertainty about (future) levels of resource need.
+Resource shortages play a central role in the decision-making problems that give rise to the quantile loss and the allocation score, which we discuss in sections \ref{sec:quantiles_shortage} and \ref{sec:bayes-quantiles} respectively.
+Section \ref{sec:numeric} gives details on the numerical methods that we use to calculate allocation scores, including some special considerations for settings where forecasts are represented by a finite collection of predictive quantiles, such as the application to forecasts of hospitalizations due to COVID-19 in section 3 of the article.
+
 \section{Shortages}
@@
 
-For convenience, we write $u_{+} = \max\{0,u\}$, and refer to $(y - x)_{+}$ as a \emph{shortage} in accordance with our
-typical use of $y$ for a demand or need and $x$ for an available supply. To regard shortage as a function depending on only one 
+We use $y$ to denote the demand or need for resources and $x$ to denote the level of available resources.
+The resource shortage is the amount by which resource demand exceeds supply. 
+For convenience, we write $u_{+} = \max\{0,u\}$, i.e., ``the positive part'' of $u$.
+With this notation, the shortage is written as $(y - x)_{+}$. To regard shortage as a function of only one 
 variable $x$ or $y$, with the other being a parameter describing the dependence we can write 
@@
 
+\begin{figure}
 \begin{tikzpicture}
@@
 \end{tikzpicture}
+\caption{Shortage functions.
+$\short^{y_0}(x)$, shown in blue, gives the resource shortage as a function of the level of available resources, $x$, for a fixed value of resource demand $y_0$.
+$\short_{x_0}(y)$, shown in red, gives the resource shortage as a function of resource demand, $y$, for a fixed value of resource supply $x_0$.}
+\end{figure}
 
-Let $Y$ be a random variable with distribution $F$. The random shortage $(Y-x)_{+}$ can be thought of as either a 
+Let $Y$ be a random variable with distribution $F$ representing the unknown level of resource demand. The random shortage $(Y-x)_{+}$ can be thought of as either a 
 real-valued random variable $\short_x(Y)$ for every $x$, or a function-valued random variable $\short^Y$ whose value
-for any realization $Y=y$ is a convex function $\short^y(x)$ of $x$.  We see then that the 
+for any realization $Y=y$ is a convex function $\short^y(x)$ of $x$.  In the sections below, we will work with the 
 \emph{expected shortage}\footnote{A more natural sounding term for $(y-x)_{+}$ might have been \emph{shortfall}. Unfortunately 
@@
 closely related to the \emph{conditional} expectation $\Ex_F [Y-x \mid Y - x \geq 0] = \Ex_F [(Y-x)_{+}]/\mathbb{P}_F\{Y \geq x\}$.}
-$\Ex_F [(Y-x)_{+}] = \Ex_F[\short^Y](x)$ (assuming it exists) is also convex (and therefore continuous) in $x$ by integrating the convexity inequality for
+$\Ex_F [(Y-x)_{+}] = \Ex_F[\short^Y](x)$.
+Assuming that this expected value exists, which is the case as long as the distribution $F$ is well behaved, we can see that $\Ex_F[\short^Y](x)$ is also convex (and therefore continuous) in $x$ by integrating the convexity inequality for
 $\short^y(x)$ with respect to the probability measure $dF(y)$:
@@
 they are bounded over the shrinking regions of integration $[x-h,x]$ and $[x,x+h]$. Convexity follows since 
-$D_{-}\Ex_F[\short^Y](x) \leq D_{+}\Ex_F[\short^Y](x)$ by the definition of $F(x)$ and $F(x-)$. This shows that if $F$ does not 
+$D_{-}\Ex_F[\short^Y](x) \leq D_{+}\Ex_F[\short^Y](x)$ by the definition of $F(x)$ and $F(x-)$. This also shows that if $F$ does not 
 have a point mass at $x$, we have 

### 2023-12-03 alloscore_manuscript/supplement.rnw
couple more minor edits to supplement

@@
 representing the future level of an undesirable outcome such as severe COVID incidence. Let $x \in \mathbb{R}$ be a decision variable
-representing the possible levels of some costly counter-measure, such as procurement of monoclonal antibody treatments,
+representing the level of some costly counter-measure, such as procurement of monoclonal antibody treatments,
 that can be taken at a cost $C>0$ per unit in preparation for $Y$.  A decision maker must decide on a level $x$ of investment in the
@@
 realized. To formalize the trade-off between these potential excesses we quantify the loss associated with a unit of
-shortage by a constant $L>C$ (which assumes that the counter-measure has some economic value) and combine the total shortage loss with expenditure into a \emph{loss function}\footnote
+shortage by a constant $L>C$ (which assumes that it is possible to quantify costs $C$ per unit of resource and losses $L$ per unit of shortfall on the same scale).
+We then combine the total shortage loss with expenditure into a \emph{loss function}\footnote
 {This does involve a confusing use of the word \emph{loss} to refer to two different quantities, but this seems to be
@@
 \] 
-The decision problem is then to select a random future loss $l(x,Y)$ in a way that aligns with the preference that $l(x,y)$ be
-as low as possible given any realization $Y=y$.
+The decision problem is then to select an action $x$ in a way that aligns with the preference that the random future loss $l(x,Y)$ be as low as possible.
 

### 2023-12-03 alloscore_manuscript/supplement.rnw
basic idea of iteration step

--- 2023-12-01 alloscore_manuscript/supplement.rnw
@@
 \usepackage{amsmath, amsfonts, amssymb, mathtools}
+\usepackage{accents}
 \usepackage{graphicx}
@@
 	\item Sketch of scoring for decision problems involving both cost and constraint.
+	\item  Derivation of case 2 in formula for Oracle adjustment
 \end{todolist}
@@
 of $F_{T}(K) = \mathbb{P}(TY_F \leq K)$, making clear the existence of a solution $\tau^{\star}$ to \eqref{eqn:tau-inclusion}. This also
-yields the interesting formal expression of the Bayes act
+yields the interesting formal equation for the Bayes act
 \begin{align}
-x^{F,K} = (F_1,\ldots,F_N)^{-1}(\mathbf{1}_N F_{T}(K)),
+\undertilde{\bm{F}}(x^{F,K}) = \mathbf{1}_N F_{T}(K),
 \end{align}
-where $\mathbf{1}_N: \mathbb{R} \to \mathbb{R}^N$ is the linear map that takes $a$ to the $N$-vector $(a,\ldots,a)^T$ and 
-$(F_1,\ldots,F_N)^{-1}$ is a map from $\mathbb{R}^N$ to $\mathbb{R}^N$. We can also arrange this as
+where $\mathbf{1}_N: \mathbb{R} \to \mathbb{R}^N$ is the linear map that takes $a$ to the $N$-vector $(a,\ldots,a)^T$ and the vector of marginal 
+cdfs $\undertilde{\bm{F}} := (F_1,\ldots,F_N)$ is a map from $\mathbb{R}^N$ to $\mathbb{R}^N$. 
+With this notation we can write \eqref{eqn:tau-inclusion} as
 \begin{align}
-TQ_F\left(\frac{1}{N} \mathbf{1}^T (F_1,\ldots,F_N)(x^{F,K})\right) = K,
+K \in TQ_F\left(\frac{1}{N} \mathbf{1}^T \undertilde{\bm{F}}(x^{F,K})\right),
 \end{align}
-which leads conceptually to the iterative numerical method of solving for $x^{F,K}$ discussed next in section \ref{sec:numeric}.
+which leads conceptually to the iterative numerical method of solving for $\tau$ and $x^{F,K}$ discussed next in section \ref{sec:numeric}.
 
@@
 \label{sec:numeric}
+
+Suppose we have established that $\tau^{\star} = F_T(K)$ lies in the interval $I_1 = [\tau_L, \tau_U]$ with $\tau_L < \tau_U$,
+that is, $K \in [q^{-}_{F_T,\tau_L}, q^{+}_{F_T,\tau_U}]$.
+From section \ref{sec:quantile-functions}, we know that the set 
+$TQ_F(\tau_L) \cup TQ_F(\tau_U) \subset [q^{-}_{F_T,\tau_L}, q^{+}_{F_T,\tau_U}]$
+is arranged in exactly one of the following ways: 
+\begin{itemize}
+\item[($\bullet \bullet$)] $TQ_F(\tau_L) \cap TQ_F(\tau_U) = \varnothing$ (and $q^{+}_{F_T,\tau_L} < q^{-}_{F_T,\tau_U}$)
+\item[($\bullet$)] $TQ_F(\tau_L) = \{K\} = TQ_F(\tau_U)$ (a point mass at $K$)
+\item[($\bullet\!-$)] $TQ_F(\tau_L) = \{q^{-}_{F_T,\tau_U}\} \subsetneq TQ_F(\tau_U)$ (a point mass at $q^{-}_{F_T,\tau_U}$)
+\item[($-\!\bullet$)] $TQ_F(\tau_L) \supsetneq \{q^{+}_{F_T,\tau_L}\} = TQ_F(\tau_U)$ (a point mass at $q^{+}_{F_T,\tau_U}$)
+\end{itemize}
+
+In the case ($\bullet$), we can immediately take 
+$\tau^{\star}=\tau_U$ as the probability level defining the allocation Bayes act.
+In the cases
+($\bullet\!-$), and ($-\!\bullet$), which imply the presence of a point mass in one or more of the 
+component forecasts adjacent to a region of zero density in all components, we can take $\tau^{\star}=\tau_U$ or $\tau_L$, 
+respectively, as the representing probability level and find Bayes acts within the set 
+$\undertilde{\bm{F}}^{-1}(\tau^{\star})$ (by solving the linear equation 
+$\sum_{x \in \undertilde{\bm{F}}^{-1}(\tau^{\star})}x_i = K$).
+
+Otherwise, by evaluating $TQ_F$ at $\tau_M = \frac{1}{2}\left(\tau_L + \tau_U\right)$ we can replace $I_1$ with the interval
+\begin{align}
+I_2 = 
+\begin{cases}
+[\tau_L, \tau_M] & \text{ if } K \leq q^{+}_{F_T,\tau_M} \\
+[\tau_M, \tau_U] & \text{ if } K > q^{-}_{F_T,\tau_M}.
+\end{cases}
+\end{align}
 

### 2023-12-03 talks/plots/code/setup.R

@@ -24,26 +24,26 @@ tar_load(truth_data)
 tar_load(forecast_data)
 tar_load(all_alloscore_data)
 
-slim_dfs <- filter(all_alloscore_data, reference_date == "2021-12-27")
+forecasts_hosp <- slim_dfs <- filter(all_alloscore_data, reference_date == "2021-12-27")
 
-forecasts_hosp <- forecast_data %>%
-  bind_rows(
-    load_forecasts(
-      dates = "2021-12-27",
-      date_window_size = 6,
-      locations = "US",
-      types = c("quantile"),
-      targets = paste(0:30, "day ahead inc hosp"),
-      source = "local_hub_repo",
-      hub_repo_path = hub_repo_path,
-      verbose = FALSE,
-      as_of = NULL,
-      hub = c("US")) |>
-      align_forecasts() |>
-      dplyr::filter(
-        relative_horizon == 14)
-  ) %>%
-  filter(reference_date == "2021-12-27")
+# forecasts_hosp <- forecast_data %>%
+#   bind_rows(
+#     load_forecasts(
+#       dates = "2021-12-27",
+#       date_window_size = 6,
+#       locations = "US",
+#       types = c("quantile"),
+#       targets = paste(0:30, "day ahead inc hosp"),
+#       source = "local_hub_repo",
+#       hub_repo_path = hub_repo_path,
+#       verbose = FALSE,
+#       as_of = NULL,
+#       hub = c("US")) |>
+#       align_forecasts() |>
+#       dplyr::filter(
+#         relative_horizon == 14)
+#   ) %>%
+#   filter(reference_date == "2021-12-27")
 truth <- truth_data
 
 mkeep <- c("BPagano-RtDriven",

### 2023-12-03 alloscore_manuscript/supplement.rnw

@@
 respectively, as the representing probability level and find Bayes acts within the set 
-$\undertilde{\bm{F}}^{-1}(\tau^{\star})$ (by solving the linear equation 
+$\undertilde{\bm{F}}^{-1}(\tau^{\star})$ (by solving the linear program 
 $\sum_{x \in \undertilde{\bm{F}}^{-1}(\tau^{\star})}x_i = K$).
@@
 \begin{cases}
-[\tau_L, \tau_M] & \text{ if } K \leq q^{+}_{F_T,\tau_M} \\
-[\tau_M, \tau_U] & \text{ if } K > q^{-}_{F_T,\tau_M}.
+[\tau_L, \tau_M] & \text{ if } K < q^{-}_{F_T,\tau_M} \\
+[\tau_M, \tau_U] & \text{ if } K \geq q^{-}_{F_T,\tau_M}.
 \end{cases}

### 2023-12-03 alloscore_manuscript/alloscore-application.Rnw
adding multi-k weighted analysis.

@@
 <<ias-comparison, results='asis'>>=
+## for visual context
+## curve(dnorm(x, mean=15000, sd=3000), from=0, to=30000)
+ks <- seq(5000, 25000, by=200)
+kprobs <- data.frame(K = ks,
+                     weight = dnorm(ks, mean=15000, sd=3000)/sum(dnorm(ks, mean=15000, sd=3000)))
+
 ias <- slim_alloscore_2021.12.20 |>
@@
   summarize(`Allocation score` = sum(components)) |>
+  left_join(kprobs) |>
   group_by(model) |>
-  summarize(IAS = mean(`Allocation score`)) |>
+  summarize(IAS_wt = sum(`Allocation score`*weight, na.rm=TRUE),
+            IAS = mean(`Allocation score`)) |>
   arrange(IAS)
@@
   arrange(AS)
+
+p1 <- ggplot(left_join(ias, as), aes(x=AS, y=IAS)) + geom_point(color="blue")
+p2 <- ggplot(left_join(ias, as), aes(x=AS, y=IAS_wt)) + geom_point(color="red")
+
+p3 <- slim_alloscore_2021.12.20 |>
+  group_by(model, K) |>
+  summarize(`Allocation score` = sum(components)) |>
+  left_join(kprobs) |>
+  ggplot(aes(x=K)) +
+  geom_point(aes(y=weight), color = "red") +
+  geom_point(y=1/length(unique(slim_alloscore_2021.12.20$K)), color="blue") +
+  geom_vline(xintercept = 15000, color="black")
+
+cowplot::plot_grid(p2, p3, p1,
+                   nrow=2,
+                   align="v",
+                   labels = "AUTO")
 

### 2023-12-04 alloscore_manuscript/alloscore-application.Rnw
First attempt at getting detailed AS info inside Rnw

@@
 A direct comparison between the forecasts from the JHUAPL-Bucky and CU-select models show that while the JHUAPL-Bucky forecast distributions were closer to the eventual observations in many states, the allocations suggested by those forecasts often were more inefficient than those from the CU-select model (Figure \ref{fig:thermometer-plot}).
-For JHUAPL-Bucky, the allocations were assigned around the 25th quantile of the predictive distribution and for CU-select it was around the 90th percentile. \ngr{could we make these more specific?}
+
+<<detailedASinfo>>=
+source("R/run-alloscore.R")
+as_Buck <- run_alloscore(
+  forecast_data = forecast_data, 
+  truth_data = truth_data,
+  one_reference_date = "2021-12-20",
+  one_model = "JHUAPL-Bucky",
+  one_K = 15000,
+  slim = FALSE)
+
+as_CU <- run_alloscore(
+  forecast_data = forecast_data, 
+  truth_data = truth_data,
+  one_reference_date = "2021-12-20",
+  one_model = "CU-select",
+  one_K = 15000,
+  slim = FALSE)
+
+Buck_quantile <- round(1 - as_Buck$lam,2)*100
+CU_quantile <- round(1 - as_CU$lam,2)*100
+
+FL_allocs <- Kat15k_alloscores |> 
+   dplyr::filter(
+     model %in% c("JHUAPL-Bucky", "CU-select"), 
+     reference_date == "2021-12-20", 
+     abbreviation == "FL") |>
+  dplyr::select(model, x) |>
+  tibble::deframe()
+FL_diff <- round(FL_allocs["CU-select"] - FL_allocs["JHUAPL-Bucky"])
+@
+
+For JHUAPL-Bucky, the allocations were assigned at the \Sexpr{Buck_quantile}th quantile of the predictive distribution and for CU-select it was at the \Sexpr{CU_quantile}th percentile. 
 This reflects that in general the JHUAPL-Bucky forecasts assigned more probabilities to higher values and thus a low quantile value across all locations reached the allocation limit of 15,000.
@@
 However for each of these three states, the resource allocation was lower and therefore resulted in more hypothetical unmet need than CU-select allocations.
-In Florida, the forecast from JHUAPL-Bucky missed by a wide margin (as did the forecast from CU-select) and the allocation was nearly XX units lower than what was allocated according to the CU-select model.
-\ngr{add specific number here}
+In Florida, the forecast from JHUAPL-Bucky missed by a wide margin (as did the forecast from CU-select) and the allocation was \Sexpr{FL_diff} units lower than what was allocated according to the CU-select model.
 

### 2023-12-04 R/run-alloscore.R

@@ -23,7 +23,8 @@ run_alloscore <- function(
     truth_data,
     one_reference_date,
     one_model,
-    one_K = NULL){
+    one_K = NULL,
+    slim = TRUE){
   require(tidyverse)
   require(alloscore)
   require(distfromq)
@@ -63,7 +64,7 @@ run_alloscore <- function(
     K = Ks,
     y = .[["value"]],
     target_names = "abbreviation",
-    slim = TRUE
+    slim = slim
   )
   ascores <- ascores %>% dplyr::mutate(
     reference_date = one_reference_date,

### 2023-12-04 alloscore_manuscript/supplement.rnw
a little more search iteration yoga

@@
 
+Iterating this process we obtain a sequence $\{I_k\}, k=1,2,\ldots$ of intervals of widths 
+$\left|I_k \right| = 2^{1-k}\left|I_1 \right|$ which either terminates at one of the scenarios 
+($\bullet$), ($\bullet\!-$), or ($-\!\bullet$), or provides infinite sequences $\{\tau_{L,k}\}$ and $\{\tau_{U,k}\}$ converging
+to $\tau^{\star}$ from below and above.
+
+In the generic case of an infinite $\{I_k\}$, we need to define practical stopping criteria for each of three possible behaviours of 
+$F_T$ at $K$:
+\begin{samepage}
+\begin{itemize}
+\item[($\,\mathrlap{\,\bullet}{\diagup}\,$)] $\tau^{\star} = F_T(K-)$ and $TQ_F(\tau^{\star}) = \{K\}$ (that is, if $F_T$ is continuous and strictly increasing at $K$)
+\item[(\raisebox{-.5ex}{$-$}\!\!\raisebox{.5ex}{$\bullet\!-$})] $F_T$ has a point mass at $K$
+\item[($-\!\!\!\bullet\!\!\!-$)] $K \in (q^{-}_{F_T,\tau^{\star}}, q^{+}_{F_T,\tau^{\star}})$ (that is, all component forecasts $F_i$ have zero density at points $x_i \in (q^{-}_{F_i,\tau^{\star}}, q^{+}_{F_i,\tau^{\star}})$ and $\sum x_i = K$).
+\end{itemize}
+\end{samepage}
+
 \section{Properties and Properness}

### 2023-12-05 alloscore_manuscript/supplement.rnw
some supplement issue edits

@@
 \begin{align}
-\mathbb{E}_{F_i}[(Y_i - x^{\star}_i - \delta)_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}] \geq 
-\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - x^{\star}_j + \delta)_{+}]. \label{eqn:ASoptimal1}
+\mathbb{E}_{F_i}[(Y_i - (x^{\star}_i - \delta))_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}] \geq 
+\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - (x^{\star}_j + \delta))_{+}]. \label{eqn:ASoptimal1}
 \end{align}
@@
 Inequalities \eqref{eqn:RLineq} and \eqref{eqn:RLineq2} remain true with $i$ and $j$ reversed. They hold with $i=j$ as well by the
-definition of $F_i(x^{\star}_i-)$. Therefore, a single number $\lambda$ (a \emph{Lagrange multiplier}) exists such that
+definition of $F_i(x^{\star}_i-)$. Therefore, a number $\lambda$ (a \emph{Lagrange multiplier}) exists, which is independent of $i$,
+such that
 \begin{align}
-L(1-F_i(x^{\star}_i-)) \geq \lambda \geq L(1-F_i(x^{\star}_i)), \quad \text{for all } i \in 1,\ldots,N
+L(1-F_i(x^{\star}_i-)) \geq \lambda \geq L(1-F_i(x^{\star}_i)), \quad \text{for all } i \in 1,\ldots,N.
 \end{align}
-that is,
+That is,
 \begin{align}
@@
 \end{align}
-is defined using interval addition $[a,b] + [c,d] = [a+c, b+d]$.
+is defined using interval addition $[a,b] + [c,d] = [a+c, b+d]$. (Note that the letter $T$ is being used to connote a ``totalling'' 
+operation, as it often is in survey sampling literature.)
 

### 2023-12-05 alloscore_manuscript/supplement.rnw
revert a hunk from beginning of quantiles section and slightly rework

@@
 We recall how quantiles arise as solutions to a probabilistic decision problem. Let $Y$ be a random variable
-representing the future level of an undesirable outcome such as severe COVID incidence. Let $x \in \mathbb{R}$ be a decision variable
-representing the level of some costly counter-measure, such as procurement of monoclonal antibody treatments,
-that can be taken at a cost $C>0$ per unit in preparation for $Y$.  A decision maker must decide on a level $x$ of investment in the
+representing the future level of an undesirable outcome such as severe COVID incidence. Let $x \in \mathbb{R}_+$ be a decision variable
+representing levels of some costly counter-measure, such as procurement of monoclonal antibody treatments,
+that can be taken at a cost $C>0$ per unit in preparation for $Y$.\footnote{
+Quantiles could also be derived for a problem in which $x$ and $Y$ take negative values, corresponding, for instance, to a decision
+maker that operates in a resource market and a that $Y$ takes negative values when ``recoveries'' outnumber incidence.  But we do not consider
+such scenarios in this work. 
+}  A decision maker must decide on a level $x$ of investment in the
 counter-measure, and wishes to avoid excesses in either the expediture $Cx$ or the shortage $(y-x)_{+}$ when $Y=y$ is
 realized. To formalize the trade-off between these potential excesses we quantify the loss associated with a unit of
-shortage by a constant $L>C$ (which assumes that it is possible to quantify costs $C$ per unit of resource and losses $L$ per unit of shortfall on the same scale).
-We then combine the total shortage loss with expenditure into a \emph{loss function}\footnote
+shortage by a constant $L>C$ (which assumes that the counter-measure has some practical value) and combine the total shortage loss with expenditure into a \emph{loss function}\footnote
 {This does involve a confusing use of the word \emph{loss} to refer to two different quantities, but this seems to be
@@
 \] 
-The decision problem is then to select an action $x$ in a way that aligns with the preference that the random future loss $l(x,Y)$ be as low as possible.
+The decision problem is then to select a random future loss $l(x,Y)$ in a way that aligns with the preference that $l(x,y)$ be
+as low as possible given any realization $Y=y$.
 

### 2023-12-05 alloscore_manuscript/alloscore-application.Rnw
updates to manuscript with start at multi-k analysis.

@@
 
-In the example of this one week, JHUAPL-Bucky had a worse allocation score than CU-select because its forecasts led to allocations that sent excess resources to several states, such as Ohio, Pennsylvania, and Michigan, that would have been more effectively allocated to states that did receive enough resources to meet their needs, such as Florida and California (Figure \ref{multi-loc-ranks}A).
+In the example of this one week, JHUAPL-Bucky had a worse allocation score than CU-select because its forecasts led to allocations that sent excess resources to several states, such as Ohio, Pennsylvania, and Michigan, that would have been more effectively allocated to states that did receive enough resources to meet their needs, such as Florida and California (Figure \ref{fig:multi-loc-ranks}A).
 These allocation errors resulted because that model’s forecasts did not consistently capture the relative resource needs across different states.
-(Figure \ref{fig:thermometer-plot} \& \ref{multi-loc-ranks}A).
+(Figure \ref{fig:thermometer-plot} \& \ref{fig:multi-loc-ranks}A).
 The CU-select model made some similar errors — most prominently, over-allocating resources to Ohio — but overall, it did a better job of forecasting the relative resource demands across different locations.
@@
 On the other hand, CU-select had worse performance as measured by WIS.
-Its forecasts were biased downwards, and it consistently incurred a large penalty for underprediction (Figure \ref{fig:multi-loc-rank}B).
+Its forecasts were biased downwards, and it consistently incurred a large penalty for underprediction (Figure \ref{fig:multi-loc-ranks}B).
 Predictions from the JHUAPL-Bucky model were wider, and included the observed level of daily hospital admissions more often.
@@
 
-Using the integrated allocation score summarizes allocation scores across a range of possible values of the constraint ($K$) (Section \ref{sec:methods.detailed.integrated_allocation}).
+The integrated allocation score summarizes allocation scores across a range of possible values of the constraint ($K$), possibly taking weights into account for different values of $K$ that might be more or less likely (Section \ref{sec:methods.detailed.integrated_allocation}).
 This could be useful in situations where the actual constraint may not be known, or could only be specified as a probability distribution.
 
- - something here from Ben's analysis?
+This also underscores the points above that allocation scores are highest where the observed value is hear the allocation constraint, but here we are holding the date/models fixed and varying K.
 
-<<ias-comparison, results='asis'>>=
+Allocation scores are fairly stable across a range of values for K, and therefore IAS is not very sensitive to the choice of distribution on K, and does not differ much from the regular old allocation score as long as that allocation score was computed at a reasonable choice of K.
+
+Interpretation of Figure \ref{fig:multi-k} needed.
+
+<<ias-comparison>>=
 ## for visual context
@@
 
-p1 <- ggplot(left_join(ias, as), aes(x=AS, y=IAS)) + geom_point(color="blue")
-p2 <- ggplot(left_join(ias, as), aes(x=AS, y=IAS_wt)) + geom_point(color="red")
+ias_to_plot <- left_join(ias, as) |>
+  pivot_longer(cols = c(IAS, IAS_wt), names_to = "IAS_type", values_to = "IAS") |>
+  mutate(IAS_type = ifelse(IAS_type=="IAS", "uniform", "centered at 15,000"))
 
-p3 <- slim_alloscore_2021.12.20 |>
+p1 <- ggplot(ias_to_plot, aes(x=AS, y=IAS, color=IAS_type, shape=IAS_type)) +
+  geom_point() +
+  geom_smooth(method="lm", se=FALSE, linetype=2) +
+  scale_color_manual(name = "Integrated Allocation Score", values = c("#386cb0", "#fdc086")) +
+  scale_shape_manual(name = "Integrated Allocation Score", values = c(1, 2)) +
+  geom_abline(slope=1, intercept=0, color="gray") +
+  theme(legend.position = c(0.05,.9), legend.justification = c(0,1),
+        legend.background=element_rect(fill = alpha("white", 0.2)))
+
+p2 <- slim_alloscore_2021.12.20 |>
   group_by(model, K) |>
@@
   ggplot(aes(x=K)) +
-  geom_point(aes(y=weight), color = "red") +
-  geom_point(y=1/length(unique(slim_alloscore_2021.12.20$K)), color="blue") +
-  geom_vline(xintercept = 15000, color="black")
-
-cowplot::plot_grid(p2, p3, p1,
-                   nrow=2,
-                   align="v",
-                   labels = "AUTO")
-
-left_join(ias, as) |>
-  xtable::xtable() |>
-  print(include.rownames = FALSE)
+  geom_point(aes(y=weight), color = "#386cb0") +
+  geom_point(y=1/length(unique(slim_alloscore_2021.12.20$K)), color="#fdc086", shape = 2) +
+  geom_vline(xintercept = 15000, color="gray")
 @
 
-<<multi-k, fig.height=4>>=
-slim_alloscore_2021.12.20 |>
+<<multi-k-scores, results='asis'>>=
+left_join(ias, as) |>
+  rename(`IAS centered at 15k` = IAS_wt,
+         `IAS uniform` = IAS) |>
+  xtable::xtable(caption = "Allocation scores shown in Figure \\ref{fig:multi-k}C, sorted by Integrated Allocation Scores (IAS) centered at 15,000.", digits = 0) |>
+  print(include.rownames = FALSE, )
+@
+
+<<multi-k, fig.height=10, fig.width = 8, fig.cap = "Allocation scores across different resource constraints (K) for the 10 models that made forecasts on December 20, 2021.
+Panel A shows, for each model, the allocation score assigned for values of K between 200 and 60,000 at increments of 200.
+The allocation scores show a sharp peak just under 20,000 which was the eventually observed number of hospitalizations.
+It is hardest to get allocations exactly right, with no excess unmet need, when the resource constraint is exactly equal to the number of resources needed.
+Panel B shows two possible weighting functions for the Integrated Allocation Score (IAS).
+The first (dark blue circles) computes weights proportional to a normal distribution centered at 15,000 (solid vertical gray line) with a standard deviation of 3,000, and truncated to be between 5,000 and 25,000.
+The second (light orange triangles) uses a uniform weight for all possible values of K.
+Note that the Allocation Score (AS), used in earlier sections of the application, only uses the allocation scores at the `pre-defined' value of K=15,000.
+the  Panel C shows how either method of computing the IAS (y-axis, again in blue circles and orange squares) is correlated with AS (x-axis, computed at K=15,000).
+Every point represents the AS and IAS for one model.
+Note that the IAS centered at 15,000 are much more closely correlated  with the AS values, which makes sense as this version of the IAS assigns higher weights to values of the IAS that are closer to the single K at which the AS scores were computed.">>=
+p3 <- slim_alloscore_2021.12.20 |>
   group_by(model, K) |>
   summarize(`Allocation score` = sum(components)) |>
-  mutate(model = factor(model, levels = ias$model)) |>
+  mutate(model = factor(model, levels = rev(ias$model))) |>
   ggplot(aes(x=K, y=`Allocation score`, color=model, linetype=model)) +
-  geom_line()
+  geom_line() +
+  theme(legend.position = c(0.01,.99), legend.justification = c(0,1),
+        legend.background=element_rect(fill = alpha("white", 0.2)),
+        legend.title = element_blank(), legend.text = element_text(size=7))
+
+cowplot::plot_grid(p3, p2, p1,
+                   nrow=3,
+                   align = "h",
+                   labels = "AUTO")
 @

### 2023-12-06 alloscore_manuscript/supplement.rnw
more painfully obtained iteration details

@@
 \usepackage{soul}
+\usepackage{tikz}
 \usepackage{pgfplots}
@@
 
-Otherwise, by evaluating $TQ_F$ at $\tau_M = \frac{1}{2}\left(\tau_L + \tau_U\right)$ we can replace $I_1$ with the interval
+Otherwise, by evaluating $TQ_F$ at $\tau_M = \frac{1}{2}\left(\tau_L + \tau_U\right)$ we can replace $I_1$ with one of its 
+halves
 \begin{align}
@@
 [\tau_L, \tau_M] & \text{ if } K < q^{-}_{F_T,\tau_M} \\
-[\tau_M, \tau_U] & \text{ if } K \geq q^{-}_{F_T,\tau_M}.
+[\tau_M, \tau_U] & \text{ if } K \geq q^{-}_{F_T,\tau_M}
+\end{cases}
+\end{align}
+which also contains $\tau^{\star} = F_T(K)$. This follows from the definition $q^{-}_{F_T,\tau_M} := \min\{x \mid F_T(x) \geq \tau_M\}$, 
+according to which
+\begin{align}
+\begin{cases}
+K < q^{-}_{F_T,\tau_M} \text{ implies } \tau^{\star} = F_T(K) < \tau_M  \\
+K \geq q^{-}_{F_T,\tau_M} \text{ implies } \tau^{\star} = F_T(K) \geq \tau_M.
 \end{cases}
@@
 ($\bullet$), ($\bullet\!-$), or ($-\!\bullet$), or provides infinite sequences $\{\tau_{L,k}\}$ and $\{\tau_{U,k}\}$ converging
-to $\tau^{\star}$ from below and above.
+to $\tau^{\star}$ from below and above. Note that this algorithm is somewhat more delicate than the usual ``bisection'' algorithm which
+would involve bisecting intervals in the domain of $F_T$ to find a root.
 
-In the generic case of an infinite $\{I_k\}$, we need to define practical stopping criteria for each of three possible behaviours of 
-$F_T$ at $K$:
-\begin{samepage}
+In the generic case of an infinite $\{I_k\}$, we need to define practical stopping criteria for the possible limit behaviours of 
+$F_T(K \pm \varepsilon)$ as $\varepsilon \searrow 0$:
 \begin{itemize}
-\item[($\,\mathrlap{\,\bullet}{\diagup}\,$)] $\tau^{\star} = F_T(K-)$ and $TQ_F(\tau^{\star}) = \{K\}$ (that is, if $F_T$ is continuous and strictly increasing at $K$)
-\item[(\raisebox{-.5ex}{$-$}\!\!\raisebox{.5ex}{$\bullet\!-$})] $F_T$ has a point mass at $K$
-\item[($-\!\!\!\bullet\!\!\!-$)] $K \in (q^{-}_{F_T,\tau^{\star}}, q^{+}_{F_T,\tau^{\star}})$ (that is, all component forecasts $F_i$ have zero density at points $x_i \in (q^{-}_{F_i,\tau^{\star}}, q^{+}_{F_i,\tau^{\star}})$ and $\sum x_i = K$).
+\item[($\mathrlap{\,\bullet}{\,/}\,$)] $F_T(K - \varepsilon) < F_T(K-) = F_T(K) < F_T(K + \varepsilon)$
+(so that $\tau^{\star} = F_T(K-)$ and $TQ_F(\tau^{\star}) = \{K\}$)
+\item[($/\!\!\raisebox{.9ex}{$\bullet$\!-}$)] $F_T(K - \varepsilon) < F_T(K-) = F_T(K) = F_T(K + \varepsilon)$
+\item[($\raisebox{-.9ex}{-\!$\bullet$}\!\!/$)] $F_T(K - \varepsilon) = F_T(K) < F_T(K + \varepsilon)$
+\item[($\raisebox{-.9ex}{\phantom{-}\!$\bullet$}\!\!/$)] $F_T(K - \varepsilon) \leq F_T(K-) < F_T(K) < F_T(K + \varepsilon)$
+\item[($\,\bullet\!\text{-}$)] $F_T(K - \varepsilon) \leq F_T(K-) < F_T(K) = F_T(K + \varepsilon)$
+\item[($\text{-}\!\!\bullet\!\!\text{-}$)]  $F_T(K - \varepsilon) = F_T(K) = F_T(K + \varepsilon)$
+so that $K \in (q^{-}_{F_T,\tau^{\star}}, q^{+}_{F_T,\tau^{\star}})$ and all component forecasts $F_i$ have 
+zero density at points $x_i \in (q^{-}_{F_i,\tau^{\star}}, q^{+}_{F_i,\tau^{\star}})$ and $\sum x_i = K$.
 \end{itemize}
-\end{samepage}
+
+We also want to be able to decribe what further ``post-processing'' steps are required to find a sufficiently good approximation
+to a Bayes act within the set
+$\undertilde{\bm{F}}^{-1}(I_{\overline{k}})$ where $I_{\overline{k}}$ is the terminal search interval for $\tau^{\star}$.
 

### 2023-12-06 alloscore_manuscript/supplement.rnw
more painfully obtained iteration details

@@
 
-We also want to be able to decribe what further ``post-processing'' steps are required to find a sufficiently good approximation
-to a Bayes act within the set
-$\undertilde{\bm{F}}^{-1}(I_{\overline{k}})$ where $I_{\overline{k}}$ is the terminal search interval for $\tau^{\star}$.
+We also want to be able to decribes what further ``post-processing'' steps are required to find an approximate Bayes act in the set
+$\undertilde{\bm{F}}^{-1}(I_{\overline{k}})$ where $\overline{k}$ is the final iteration step index.
 

### 2023-12-06 alloscore_manuscript/supplement.rnw
another clarification regarding bisection

--- 2023-12-06 alloscore_manuscript/supplement.rnw
@@
 \end{align}
+Note that $K \leq q^{-}_{F_T,\tau_M}$ would not imply $F_T(K) \leq \tau_M$ due to the possibility of a point mass at $K$.
 
@@
 ($\bullet$), ($\bullet\!-$), or ($-\!\bullet$), or provides infinite sequences $\{\tau_{L,k}\}$ and $\{\tau_{U,k}\}$ converging
-to $\tau^{\star}$ from below and above. Note that this algorithm is somewhat more delicate than the usual ``bisection'' algorithm which
-would involve bisecting intervals in the domain of $F_T$ to find a root.
+to $\tau^{\star}$ from below and above. Note that this algorithm is somewhat more delicate than the corrsponding ``bisection'' algorithm 
+for finding a root of $F_T(x) - \tau$ for a given $\tau$.
 

### 2023-12-07 alloscore_manuscript/supplement.rnw
give some refs for quantiles as decisions

@@
 
-We recall how quantiles arise as solutions to a probabilistic decision problem. Let $Y$ be a random variable
+We recall how quantiles arise as solutions to a probabilistic decision problem, drawing on perspectives developed in the fields of
+both forecasting (see e.g., \cite{gneiting2011quantiles} and \cite{jose2009evaluating}) and stochastic optimization 
+(se e.g., \cite{royset2022optimization}, sections 1.C and 3.C).
+
+Let $Y$ be a random variable
 representing the future level of an undesirable outcome such as severe COVID incidence. Let $x \in \mathbb{R}_+$ be a decision variable
@@
 Quantiles could also be derived for a problem in which $x$ and $Y$ take negative values, corresponding, for instance, to a decision
-maker that operates in a resource market and a that $Y$ takes negative values when ``recoveries'' outnumber incidence.  But we do not consider
-such scenarios in this work.	 
+maker that both buys and sells in a resource market and a $Y$ that takes negative values when ``recoveries'' outnumber incidence.  
+But we do not consider such scenarios in this work.	 
 }  A decision maker must decide on a level $x$ of investment in the

### 2023-12-07 alloscore_manuscript/supplement.rnw
slowly tying the logic of secs 4 and 5 together

@@
 
-The constraint now determines $\tau$ (and hence the Bayes act) through $\sum_{i=1}^{N} q_{\tau,F_i} = K.$
-This equation can be restated as the requirement that 
+The constraint now determines $\tau$ (and hence the Bayes act) through 
+\begin{align}
+\sum_{i=1}^{N} q_{\tau,F_i} = K. \label{eqn:quantiles-sum-to-K}
+\end{align}
+This equation implies that 
 \begin{align}
@@
 is defined using interval addition $[a,b] + [c,d] = [a+c, b+d]$. (Note that the letter $T$ is being used to connote a ``totalling'' 
-operation, as it often is in survey sampling literature.)
+operation, as it often is in survey sampling literature.) Conversely, if $\tau^{\star}$ satisfies \eqref{eqn:tau-inclusion}, then
+we can find a solution $\bm{q}_{\tau^{\star},F} := (q_{\tau^{\star},F_1},\ldots,q_{\tau^{\star},F_N})$ to 
+\eqref{eqn:quantiles-sum-to-K} within the set 
+$\undertilde{\bm{F}}^{-1}(\tau^{\star}) \subset \mathbb{R}_{+}^N$ where $\undertilde{\bm{F}}^{-1}(\tau)$  is the 
+Cartesian product $(F_1^{-1}(\tau),\ldots,F_N^{-1}(\tau))$.
 
@@
 \label{sec:numeric}
+
+Exact computation of the allocation score $S_A(F, y; K)$ of a forecast $F$ requires an explicit choice of and expression for a 
+solution $\bm{q}_{\tau^{\star},F} \in \undertilde{\bm{F}}^{-1}(\tau^{\star})$ to \eqref{eqn:quantiles-sum-to-K} where
+$\tau^{\star}$ solves \eqref{eqn:tau-inclusion}. For practical scenarios in which the forecast $F$ has a form preventing 
+analytical solutions to \eqref{eqn:quantiles-sum-to-K}, we must develop a method to approximate solutions to \eqref{eqn:tau-inclusion}
+which lead to adequate approximations of the linear program
+\begin{align}
+\sum_{x \in \undertilde{\bm{F}}^{-1}(\tau^{\star})}x_i = K.
+\end{align}
 
@@
 ($\bullet$), ($\bullet\!-$), or ($-\!\bullet$), or provides infinite sequences $\{\tau_{L,k}\}$ and $\{\tau_{U,k}\}$ converging
-to $\tau^{\star}$ from below and above. Note that this algorithm is somewhat more delicate than the corrsponding ``bisection'' algorithm 
-for finding a root of $F_T(x) - \tau$ for a given $\tau$.
+to $\tau^{\star}$ from below and above. Such a sequence provides the basis of a ``bisection'' algorithm for finding
+the ``root'' $\tau^{\star}$ of the set condition $0 \in TQ_F(\tau) - K$.
 

### 2023-12-08 alloscore_manuscript/supplement.rnw
bayes risk error idea

@@
 Exact computation of the allocation score $S_A(F, y; K)$ of a forecast $F$ requires an explicit choice of and expression for a 
-solution $\bm{q}_{\tau^{\star},F} \in \undertilde{\bm{F}}^{-1}(\tau^{\star})$ to \eqref{eqn:quantiles-sum-to-K} where
+solution $x^{F,K} = \bm{q}_{\tau^{\star},F} \in \undertilde{\bm{F}}^{-1}(\tau^{\star})$ to \eqref{eqn:quantiles-sum-to-K} where
 $\tau^{\star}$ solves \eqref{eqn:tau-inclusion}. For practical scenarios in which the forecast $F$ has a form preventing 
-analytical solutions to \eqref{eqn:quantiles-sum-to-K}, we must develop a method to approximate solutions to \eqref{eqn:tau-inclusion}
-which lead to adequate approximations of the linear program
+analytical solutions to \eqref{eqn:quantiles-sum-to-K}, we seek a method of generating an approximation $\tilde{\tau}$ to 
+a solution $\tau^{\star}$ which offers control (at least in principle) of the Bayes risk error
 \begin{align}
-\sum_{x \in \undertilde{\bm{F}}^{-1}(\tau^{\star})}x_i = K.
+\left|\mathbb{E}_F[s_A(\bm{q}_{\tilde{\tau},F}, Y)] - \mathbb{E}_F[s_A(x^{F,K}, Y)]\right|
 \end{align}
+where $\bm{q}_{\tilde{\tau},F}$ solves the linear program
+$\sum_{x \in \undertilde{\bm{F}}^{-1}(\tilde{\tau})}x_i = K$.
 

### 2023-12-08 alloscore_manuscript/supplement.rnw
one more time.

@@
 analytical solutions to \eqref{eqn:quantiles-sum-to-K}, we seek a method of generating an approximation $\tilde{\tau}$ to 
-a solution $\tau^{\star}$ which offers control (at least in principle) of the Bayes risk error
+a solution $\tau^{\star}$ which offers control (at least in principle) of the Bayes risk error or ``optimality gap''
 \begin{align}
-\left|\mathbb{E}_F[s_A(\bm{q}_{\tilde{\tau},F}, Y)] - \mathbb{E}_F[s_A(x^{F,K}, Y)]\right|
+\left|\mathbb{E}_F\left[s_A(\bm{q}_{\tilde{\tau},F}, Y)\right] - \mathbb{E}_F[s_A(x^{F,K}, Y)]\right| = 
+L\left| \sum_{i=1}^{N}\mathbb{E}_{F_i}\left[(Y_i-q_{\tilde{\tau},F_i})_+ - (Y_i-x_i^{F,K})_+\right]\right|
 \end{align}

### 2023-12-09 alloscore_manuscript/supplement.rnw
updates to supplement related to distfromq

--- 2023-12-08 alloscore_manuscript/supplement.rnw
@@
 
-\section{Properties and Properness}
+\section{Computing allocations from finite quantile forecast representations}
 
-\begin{quote}
-For a prediction to be useful, it must \textbf{proper}ly describe a \textbf{proper}ty.
-\end{quote}
+In section 3 of the manuscript, we used the allocation score to evaluate forecasts of COVID-19 hospitalizations that have been submitted to the COVID-19 Forecast Hub.
+These forecasts are submitted to the Hub using a set of 23 quantiles of the forecast distribution at the 23 probability levels in the set $\mathcal{T} = \{0.01, 0.025, 0.05, 0.1, 0.15, \ldots, 0.9, 0.95, 0.975, 0.99\}$, which specify a predictive median and the endpoints of central $(1 - \alpha) \times 100\%$ prediction intervals at levels $\alpha = 0.02, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9$.
+For a given week and target date, we use $q_{i,k}$ to denote the submitted quantiles for location $i$ and probability level $\tau_k \in \mathcal{T}$, $k = 1, \ldots, 23$.
+
+In the event that there is some $k \in \{1, \ldots, 23\}$ for which $\sum_i q_{i,k} = K$, i.e., the provided predictive quantiles at level $\tau_k$ sum across locations to the resource constraint $K$, the solution to the allocation problem is given by those quantiles.
+However, generally this will not be the case; the optimal allocation will typically be at some probability level $\tau^\star \notin \mathcal{T}$.
+
+To address this situation and support the numerical allocation algorithm outlined in section \ref{sec:numeric}, we need a mechanism to approximate the full cumulative distribution functions $F_i$, $i = 1, \ldots, N$ based on the provided quantiles.
+We have developed functionality for this purpose in the \verb`distfromq` package for R (cite).
+This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:
+\begin{enumerate}
+  \item Identify a discrete component of the distribution consisting of zero or more point masses, and create an adjusted set of predictive quantiles for the continuous part of the distribution by subtracting the point mass probabilities and rescaling.
+  \item For the continuous part of the distribution, different approaches are used on the interior and exterior of the provided quantiles:
+  \begin{enumerate}
+    \item On the interior, a monotonic cubic spline interpolates the adjusted quantiles representing the continuous part of the distribution.
+    \item A location-scale parametric family is used to extrapolate beyond the provided quantiles.
+The location and scale parameters are estimated separately for the lower and upper tails so as to obtain a tail distribution that matches the two most extreme quantiles in each tail.
+  \end{enumerate}
+\end{enumerate}
+The resulting distributional estimate exactly matches all of the predictive quantiles provided by the forecaster.
+We use the cumulative distribution function resulting from this procedure as an input to the allocation score algorithm.
+
+We refer the reader to the \verb`distfromq` documentation for further detail.
+
+% \begin{enumerate}
+%   \item First, we handle settings where the provided quantiles imply the existence of one or more point masses.
+Specifically, if there are distinct probability levels $\tau_k < \tau_{k^\prime}$ that have the same quantiles, $q_{l,k} = q_{l, k^\prime}$, the distribution contains a point mass at $q_{l,k}$.
+We split into three cases depending on the number of distinct provided quantiles:
+%   \begin{enumerate}
+%     \item If all provided quantiles are equal to each other, we infer that the distribution consists of a single point mass at that quantile value.
+%     \item If there are two distinct quantiles across all of the probability levels $\tau_k \in \mathcal{T}$, our estimated distribution consists of a mixture of two point masses.
+The probability assigned to the point mass at $q$ is proportional to $\tau_q^\max - \tau_q^\min$, where
+%     $$\tau_q^\max = \begin{cases}
+%       \max_{\{\}} \tau_k
+%     \end{cases}$$ is the largest probability level $\tau_k$ for which $q_{l,k} = q$, and $\tau_q^\min$ is defined similarly.
+%     \item Otherwise, our final estimate will consist of a mixture of a discrete distribution with point masses at the duplicated quantiles and a continuous distribution elsewhere.
+In this step, we  In this case, the discrete distribution has point mass probabilities calculated as $d_q = \tau_q^\max - \tau_q^\min$, where $\tau_q^\max$ and $\tau_q^\min$ are defined as in the previous point.
+%   \end{enumerate}
+% \end{enumerate}
+
+\section{Propriety of the allocation score}
+\label{sec:proper}
+
+A desirable feature of a scoring rule is that it is \emph{proper}, which roughly means that it rewards a forecaster for stating their true beliefs about the events being forecasted.
+We give a more formal definition of propriety in section \ref{sec:proper_def} before addressing the propriety of the allocation score in three stages:
+\begin{itemize}
+  \item Section \ref{sec:alloscore_proper} shows that the allocation score is proper when the forecaster's actual forecast distribution $F$ is available (as is any score derived using the decision theoretic procedures outlined in section 2 of the manuscript)
+  \item Section \ref{sec:distfromq_alloscore_proper_prospective} shows that the allocation score is still proper if the forecast is represented as a member of a parametric family of distributions, so long as this representation and scoring procedure was specified prospectively by the hub.
+This includes the use of methods such as \verb`distfromq` for constructing forecast distributions as a special case.
+  \item Section \ref{sec:distfromq_alloscore_improper_post_hoc} shows that \emph{post hoc} application of this scoring procedure to quantile forecasts is improper, and makes some suggestions for organized forecasting efforts who are considering using the allocation score.
+\end{itemize}
+
+\subsection{Proper scores}
+\label{sec:proper_def}
 
@@
 as a general tool for formalizing a decision problem that assigns numerical value to the \emph{result} of taking an \emph{action} $x$
-in preparation for an \emph{outcome} $y$.  A \emph{scoring rule} $S$ is a loss function where the action 
+in preparation for an \emph{outcome} $y$. Here, we consider a setting where the individual making the decision is the forecaster, and the action is to report a particular forecast.
+
+A \emph{scoring rule} $S$ is a loss function where the action 
 is a probabilistic forecast $F$ of the outcome $y$ (or the statement of $F$ by a forecaster). Just as in 
 Section \ref{sec:quantiles_shortage}, given an action $F$, $S$ transforms a random outcome variable $Y$ into a random loss $S(F,Y)$. 
-We refer to the realized loss $S(F,y)$ as the \emph{score} of $F$ at $y$, and the process of evaluating $S(F,y_i)$ for
-a data set $\mathcal{Y}=\{y_i\}$ as \emph{scoring} $F$ against $\mathcal{Y}$. Perhaps the most fundamental example of a scoring rule
-from a statistical perspective is the logarithmic score $S_{\log}(F,y) = -\log f(y)$ (where $f(y)$ is the density or the mass of $F$ at $y$
-if it exists and otherwise the mass), that is, the negative log-likelihood of $F$ interpreted as a parameter for the singleton 
-data set $\{y\}$. This frames a maximum likelihood estimate as the solution to a decision problem following the expected loss
-minimization principle introduced in Section \ref{sec:quantiles_shortage}.
+We refer to the realized loss $S(F,y)$ as the \emph{score} of $F$ at $y$.
 
 Decision theoretically, probabilistic forecasts are a unique kind of action in that they can be used to generate their
-own(simulated) outcome data, against which they can be scored using $S$. $S$ therefore commits a probabilistic forecast
+own (simulated) outcome data, against which they can be scored using $S$. $S$ therefore commits a probabilistic forecast
 $F$ to the ``self-assessment'' $\Ex [S(F, Y^F)]$, where $Y^F \sim F$ is the random variable defined by sampling from
-$F$, as well to an assessment $\Ex [S(G, Y^F)]$ of any alternative forecast $G$. For $S_{\log}$ this self-assessment is
-the Shannon entropy $H(F) = -\int \log(f(x))f(x)dx$ of $F$, and adding to $H(F)$ the Kullback-Leibler (KL) divergence $D_
-{KL}(F,G) = -\int \log(g(x)/f(x))f(x)dx$ gives $F$'s assessment $\Ex [S_{\log}(G, Y^F)]$ of $G$. (That is, $\Ex [S_
-{\log}(G, Y^F)] = H(F) + D_{KL}(F,G)$.) The KL divergence is the degree to which $F$ perceives $G$ as divergent from
-being able to minimize expected loss in this particular forecaster decision problem (i.e. maximize expected log-likelihood).
+$F$, as well to an assessment $\Ex [S(G, Y^F)]$ of any alternative forecast $G$.
 
-A natural consistency criterion for $S$ is that it does not commit $F$ to assessing any other 
-forecast $G$ as being better than $F$ itself, that is, that
+A natural consistency criterion for $S$ is that it does not commit $F$ to assessing any other forecast $G$ as being better than $F$ itself, that is, that
 \begin{align}
@@
 $S$ is \emph{strictly proper} when this inequality is sharp, in which case the \emph{only} optimal decision for a forecaster is to state
-the forecast they believe to be true. The logarithmic scoring rule $S_{\log}$, for example, is strictly proper due to the 
-positivity of KL divergence.
+the forecast they believe to be true.
 
-Another definition of $S$ being proper that is often used is that $\Ex[S(F,Y)]$ is lowest when $F$ is the true distribution of $Y$.  
-Under a flexible reading, this definition is equivalent to ours, but we find it problematic because it invites the mistaken impression that 
-the properness of a score might depend on the true distribution of what is being forecasted. Whether a score is proper is unrelated 
-to any particular forecast being scored or source of the data being used to score it.
+\subsection{The allocation score is proper}
+\label{sec:alloscore_proper}
 
-The condition of being proper is quite strong, and na\"{i}ve means of reducing a forecast distribution to a single
-number based on an observed outcome $y$ will generally define an improper scoring rule.  For example, the \emph{probability score} 
-$S_{\mathrm{Prob},c}(F,y):= -(F(y+c) - F(y-c))$ and the \emph{linear score} $S_{\mathrm{Lin}}
-(F,y):= \lim_{c\to 0} \frac{1}{2c}S_{\mathrm{Prob},c}(F,y) = -f(y)$ (where we assume $F$ has a density $f$) and are
-both improper because they commit a general $F$ to assessing an alternative forecast $G$ as being better than $F$
-itself whenever $G$ is sufficiently more concentrated than $F$ in the neighborhood $[m_f-c,m_f+c]$ of a mode $m_f$ of
-$f$. The same is true when the outcome is discrete (and ordered for $S_{\mathrm{Prob}}$) with index $y$: for a forecast
-pmf $p(y)$, moving the mass of $p$ onto a neigborhood of a mode $m_p$ to get a new forecast $p_m$ will improve the 
-expected score $\Ex[S_{\mathrm{Prob/Lin}}(p_m,Y^p)]$ according to $p$ itself. This creates the classic pathology
-of probabilistic forecasters having no incentive to express uncertainty. The meteorologist getting paid according to
-$p_{\text{rain}}\text{rain} + (1-p_{\text{rain}})(1-\text{rain})$ (the negative of the linear score) will say rain is 
-inevitable or impossible when the chances of rain appear to be 51\% or 49\%.
+We recall the three-step decision theoretic procedure for deriving proper scoring rules that we outlined in section \ref{sec:methods.detailed.decisiontheory} of the main text:
+\begin{enumerate}
+  \item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$ eventually occurs.
+  \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under the distribution $F$.
+  \item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: $S(F, y) = s(x^F, y)$.
+\end{enumerate}
 
-Sometimes a scoring rule can be made proper with an added correction that penalizes
+By construction, this procedure yields a proper score since
+\begin{align}
+\Ex_F [S(F, Y)] &= \Ex_F [ s(x^F, Y) ] \nonumber \\
+ &\leq \Ex_F [ s(x^G, Y) ] \label{eqn:dt_proper_key} \\
+ &= \Ex_F [ S(G, Y)]. \nonumber
+\end{align}
+Here, step~\eqref{eqn:dt_proper_key} holds because the Bayes act $x^F$ was selected so that the expected loss $\Ex_F [ s(x^F, Y) ]$ for that action is less than or equal to the expected loss $\Ex_F [ s(x, Y) ]$ for any other action $x$ (and for the action $x^G$ in particular).
 
-But given an auxiliary decision problem $\mathcal{D}_{\mathrm{Aux}} = \{\mathcal{X}, \mathcal{Y}, l(x,y)\}$, such as the
-allocation problem from the main text or the cost-loss problem from Section \ref{sec:quantiles_shortage}, we can
-produce a scoring rule $S_l$ that is automatically proper using the Bayes act formalism introduced in Section \ref
-{sec:methods.detailed.specific_allocation}. This works by defining $S_l(F, y)$ as the value $l(x^F,y)$ already given by  
-$\mathcal{D}_{\mathrm{Aux}}$ for the Bayes act $x^F$; that is, the action which \emph{by design} will be assessed via
-the forecast $F$ to have the lowest possible expected loss $\Ex[l(x^F,Y^F)]$. In statistical decision theory, 
-$\Ex[l(x^F,Y^F)]= \Ex_F[l(x^F,Y)]$ is 
-sometimes\footnote{A clear exposition of basic decision theory using this terminology is Chapter 8 of \cite{degroot2005optimal}.
-Unfortunately other influential sources use ``Bayes risk'' to refer to a wider variety of quantities. In 
-\cite{berger2013statistical} for example, the Bayes risk can be both the optimal expected loss 
-for $F$ (see p. 17) or the expected loss for a general $x$ with respect to $F$ (see p. 6).} 
-refered to as the \emph{Bayes risk} of $F$ relative to $\mathcal{D}_{\mathrm{Aux}}$.
+Because the alloscore was developed using this procedure, it is a proper score.
 
-by mapping a forecast $F$ to an the action $x^F$
+\subsection{Prospectively-specified use of the allocation score to evaluate forecasts approximately represented as members of a parametric family is proper}
+\label{sec:distfromq_alloscore_proper_prospective}
 
-The study of \emph{elicitability} also concerns this construction (see, e.g., \cite{gneiting2011making}), but from a
-different vantage point than ours. The question there is whether for a given map (called a \emph{functional} in
-statistics and a \emph{property} in computer science) $T(F)=x \in \mathcal{X}$ of distributions into the auxiliary
-action space, there exists a loss function $l$ such that $T(F) = x^F$ is the Bayes act for $l$ and the associated
-scoring rule $S_l(F,x)$ is strictly proper. Such an $l$ is said to \emph{elicit} $T$.  Our focus is rather on how
-forecast evaluation proceeds via the Bayes act construction for a given loss function of subject matter interest.
+In practice, open forecasting exercises are generally not able to collect an exact representation of the forecast distribution $F$ other than in simple settings such as for a categorical variable with a relatively small number of categories.
+In settings where the outcome being forecasted is a continuous quantity (such as rates of influenza-like illness among outpatient doctor visits) or a count (such as influenza hospitalizations), forecasting exercises have therefore resorted to collecting summaries of a forecast distribution such as bin probabilities or predictive quantiles.
+This raises the question of whether the allocation score is still proper if the forecast distribution $F$ is not itself directly recorded.
+The answer we provide in this section is that it is, so long as the fact that the allocation score will be used for forecast evaluation and the method that will be used to obtain a distributional estimate of $F$ from the provided forecast representation are communicated to participating forecasters prospectively.
+
+We consider a setting where a forecasting exercise (such as a forecast hub) pre-specifies that forecasts will be represented using a parametric family of forecast distributions $F_\theta(y)$, and forecasts will be specified by providing the value of the parameters $\theta$.
+We use $\mathcal{P}$ to denote the collection of all distributions $F_\theta$ in the given parametric family.
+For instance, it has recently been proposed that mixture distributions could be used to represent forecast distributions \cite{wadsworth2023mixture}.
+Additionally, we note that the functionality in \verb`distfromq` can be viewed as specifying a parametric family of distributions where the parameters $\theta$ are given the interpretation of being the quantiles of $F_\theta$ at pre-specified probability levels, but more generally control the shape of $F_\theta$ over the full range of its support.
+
+We find it helpful now to formally distinguish between two decision making problems.
+The first is the public health decision maker's allocation problem where the task is to select an allocation $x$, with the allocation loss $s_A(x, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i)$ as described in section \ref{sec:methods.detailed.specific_allocation} of the main text.
+The second is the forecaster's reporting problem where the task is to select parameter values $\theta$ to report.
+The forecaster's loss is given by $s_R(\theta, y) = s_A(x^{F_\theta}, y)$, where $x^{F_\theta}$ is the Bayes act for the allocation problem under the distribution $F_\theta$.
+In words, the loss associated with reporting $\theta$ is equal to the loss associated with taking the Bayes allocation corresponding to the distribution $F_\theta$.
+
+Following our usual construction, the Bayes act for the forecast reporting problem is the parameter set that minimizes the forecaster's expected loss:
+\begin{align*}
+\theta^F &= \text{argmin}_\theta \Ex_F [s_R(\theta, Y)] \\
+&= \text{argmin}_\theta \Ex_F [s_A(x^{F_\theta}, Y)]
+\end{align*}
+This leads the forecaster to choose values of the parameters $\theta$ such that the allocation $x^{F_\theta}$ is as close as possible to the allocation $x^F$ under the forecaster's true forecast distribution $F$.
+
+We then arrive at the scoring rule
+$$S_R(F, y) = s_R(\theta^F, y) = s_A(x^{F_{\theta^F}}, y).$$
+\elr{Notation got pretty bad here... borderline about whether it's worth searching for a better notation for "the bayes allocation for the distribution with the bayes parameters"}
+It follows from the discussion in section \ref{sec:alloscore_proper} that this is a proper scoring rule for $F$.
+Additionally, we note that the score can be calculated from the reported parameter values $\theta^F$.
+
+We emphasize that the forecaster's true predictive distribution $F$ does not need to be a member of the specified parametric family $\mathcal{P}$ for this construction to yield a proper score.
+It is, however, necessary to specify the parametric family to use and the foundational scoring rule $s_A$ (including any relevant problem parameters such as the resource constraint $K$) in advance, so that forecasters can identify the Bayes act parameter set $theta^F$ to report.
+
+As an alternative, a forecast hub could ask forecasters to provide the Bayes allocations associated with their forecasts for one or more fixed resource constraints $K$, allowing for direct calculation of scores and narrowing the gap between model outputs and public health end users.
+
+\subsection{Post hoc evaluation of quantile forecasts using distfromq and the allocation score is improper}
+\label{sec:distfromq_alloscore_improper_post_hoc}
+
+In this section, we illustrate that a post hoc evaluation of quantile forecasts using \verb`distfromq` and the allocation score is not proper.
+In brief, this is because the Bayes parameters $\theta^F$ for the forecaster to report may not align with quantiles of the distribution $F$ at the specified probability levels.
+We illustrate this with examples (borrow from elicitation.R...)
 

### 2023-12-09 alloscore_manuscript/supplement.rnw
prune checklist

@@
 	\item[\done] From 2.2.1, why are Bayes act scoring rules proper?
-	\item Explain ``All proper scoring rules for probabilistic forecasts have an explicit link to a loss function'' from discussion.
+	%\item Explain ``All proper scoring rules for probabilistic forecasts have an explicit link to a loss function'' from discussion.
 	\item DGP as optimal for any decision problem, ref Diebold, Gunther, Tay p.
@@
 		\end{todolist}
-	\item Sketch of scoring for decision problems involving both cost and constraint.
+	%item Sketch of scoring for decision problems involving both cost and constraint.
 	\item  Derivation of case 2 in formula for Oracle adjustment

### 2023-12-09 alloscore_manuscript/supplement.rnw
some edits to last elr pr and a comment

@@
 
-A desirable feature of a scoring rule is that it is \emph{proper}, which roughly means that it rewards a forecaster for stating their true beliefs about the events being forecasted.
-We give a more formal definition of propriety in section \ref{sec:proper_def} before addressing the propriety of the allocation score in three stages:
+In order for a scoring rule to be guaranteed to yield logically consistent forecast evaluations, it must be \emph{proper}, which means,
+in essence, that a forecaster cannot perceive any forecast at odds with their beliefs about the events being forecasted as having an optimal expected score.
+We formalize this in terms of forecaster ``self-assessment'' in section \ref{sec:proper_def} and then address the propriety of the allocation score in three stages:
 \begin{itemize}
@@
   \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under the distribution $F$.
-  \item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: $S(F, y) = s(x^F, y)$.
+  \item Define a \emph{scoring rule} which assigns the loss realized by the Bayes act $x^F$ against outcome $y$ as the score of $F$ at $y$:
+  \begin{align}
+    S(F, y) = s(x^F, y). \label{eqn:bayes_sr}
+  \end{align}
 \end{enumerate}
 
-By construction, this procedure yields a proper score since
+Such scoring rules, which we call \emph{Bayes scoring rules}, are proper by construction since
 \begin{align}
 \Ex_F [S(F, Y)] &= \Ex_F [ s(x^F, Y) ] \nonumber \\
+ &= \min_{x} \Ex_F [ s(x, Y) ] \quad \text{ (by definition of $x^F$)} \\
  &\leq \Ex_F [ s(x^G, Y) ] \label{eqn:dt_proper_key} \\
@@
 \end{align}
-Here, step~\eqref{eqn:dt_proper_key} holds because the Bayes act $x^F$ was selected so that the expected loss $\Ex_F [ s(x^F, Y) ]$ for that action is less than or equal to the expected loss $\Ex_F [ s(x, Y) ]$ for any other action $x$ (and for the action $x^G$ in particular).
 
-Because the alloscore was developed using this procedure, it is a proper score.
+The allocation scoring rule is Bayes and therefore proper.
+
+We note that in the probabilistic forecasting literature (see e.g., \cite{gneiting2011making}, Theorem 3) what we have 
+termed Bayes scoring rules typically appear via \eqref{eqn:bayes_sr} where $x^F$ is some given functional of $F$ which 
+can be shown to be \emph{elicitable}, that is, to be the Bayes act for some loss function $s$.
+Such a loss function is said to be a \emph{consistent loss (or scoring) function} for the functional $F \mapsto x^F$, and many important
+recent results in the literature (e.g., \cite{fisslerziegel2016consistency}) address whether there \emph{exists} any loss 
+function for $x^F$ which is consistent. Our orientation
+is different from this insofar as we \emph{begin} by specifying a decision problem and a loss function of subject matter relevance
+and use the Bayes act only as a bridge to a proper scoring rule.  Consistency is never in doubt.
 
@@
 This leads the forecaster to choose values of the parameters $\theta$ such that the allocation $x^{F_\theta}$ is as close as possible to the allocation $x^F$ under the forecaster's true forecast distribution $F$.
+\apg{I think we need to say something about what we mean by ``close''.  The score divergence is what we are confronting here and I appreciate that 
+others may not want to broach the topic.  But I think we have to say something about ``score expectation defined distance''. This also relates directly to the notation crunch below.}
 

### 2023-12-10 alloscore_manuscript/supplement.rnw
a few more edits in sec:distfromq_alloscore_proper_prospective

@@
 
-We consider a setting where a forecasting exercise (such as a forecast hub) pre-specifies that forecasts will be represented using a parametric family of forecast distributions $F_\theta(y)$, and forecasts will be specified by providing the value of the parameters $\theta$.
+We consider a setting where a forecasting exercise (such as a forecast hub) pre-specifies that forecasts will be represented using a parametric family of forecast distributions $F_\theta(y)$, and the task of the forecaster is to select a particular parameter value $\theta$.
 We use $\mathcal{P}$ to denote the collection of all distributions $F_\theta$ in the given parametric family.
 For instance, it has recently been proposed that mixture distributions could be used to represent forecast distributions \cite{wadsworth2023mixture}.
-Additionally, we note that the functionality in \verb`distfromq` can be viewed as specifying a parametric family of distributions where the parameters $\theta$ are given the interpretation of being the quantiles of $F_\theta$ at pre-specified probability levels, but more generally control the shape of $F_\theta$ over the full range of its support.
+Additionally, we note that the functionality in \verb`distfromq` can be viewed as specifying a parametric family $\mathcal{P}_{\mathrm{dfq}}$ where the parameters $\theta$ of $F_\theta$
+are its quantiles at pre-specified probability levels, and where the shape of any $F_\theta \in \mathcal{P}_{\mathrm{dfq}}$ over the full range of its support is entirely controlled by these quantiles.
 

### 2023-12-10 alloscore_manuscript/alloscore-application.Rnw
some language for the IAS section

@@
 This could be useful in situations where the actual constraint may not be known, or could only be specified as a probability distribution.
+Unlike in the above analyses where we conducted analyses assuming that $K$ was known to be 15,000, the analyses in this subsection are conducted assuming different distributions of $K$.
 
-This also underscores the points above that allocation scores are highest where the observed value is hear the allocation constraint, but here we are holding the date/models fixed and varying K.
+<<calc-tot-y>>=
+tot_y <- slim_alloscore_2021.12.20 |>
+  filter(model == "COVIDhub-ensemble", K==200) |>
+  pull(y) |>
+  sum()
+@
 
-Allocation scores are fairly stable across a range of values for K, and therefore IAS is not very sensitive to the choice of distribution on K, and does not differ much from the regular old allocation score as long as that allocation score was computed at a reasonable choice of K.
+Allocation scores (AS) were computed for a range of $K$ from 200 to 60,000 for forecasts made on 2020-12-20 predicting levels of hospitalizations on 2021-01-03 (Figure \ref{fig:multi-k}A).
+These calculations highlight that allocation scores were highest at the $K$ nearest to the observed value of \Sexpr{format(tot_y, big.mark = ",")}.
+Rankings of allocation scores from all models were fairly stable across a range of values for K, with some crossings, especially at $K$ values further away from the observed value.
+
+Integrated allocation scores (IAS) were computed for two distributions of $K$, one uniform across the entire range and the other a symmetric distribution around 15,000 (Figure \ref{fig:multi-k}B).
+Both versions of the IAS were correlated with the original AS conducted at $K=15,000$, with the higher correlation coming from the distribution that was centered at $K=15,000$ (Figure \ref{fig:multi-k}C).
 

### 2023-12-12 alloscore_manuscript/supplement.rnw
add quantile fig, change error measure in sec:numeric

@@
 
-For future reference, we record several key properties of quantile functions.
+For future reference, we record several key properties of quantile functions. An illuminating source for these
+facts is \cite{rockafellar2014random}, from which figure \ref{fig:FQ} is adapted.
 
-The probability levels $\{\alpha_i\}$ for which $\#Q_F(\alpha_i)>1$ form a discrete subset of $(0,1)$ and correspond to
+The probability levels $\{\alpha_i\}$ for which $\#Q_F(\alpha_i)>1$ ($\{\alpha_3\}$ in the example of figure \ref{fig:FQ}) form a 
+discrete subset of $(0,1)$ and correspond to
 the non-zero width intervals $[q_{\alpha_i,F}^{-},q_{\alpha_i,F}^{+}]$ where $F$ is constant with values 
@@
 $A$, then extending $F^{-1}$ by $F^{-1}(0)=\inf(\mathrm{supp}(F))$ and $F^{-1}(1)=\sup(\mathrm{supp}(F))$ provides an
-inverse to $F$ on $F(\mathbb{R})\subset [0,1]$. If $F$ has a point mass at $x\in A$, so that
-$\{\alpha \mid F(x-)< \alpha < F(x)\}$ is a non-empty set disjoint from $F(A)$, then $Q_F$ takes the constant value $x$ 
+inverse to $F$ on $F(\mathbb{R})\subset [0,1]$. If $F$ has a point mass at $x\in A$ (e.g., $x_1 \in [0,\infty)$ in figure \ref{fig:FQ}),
+so that $\{\alpha \mid F(x-)< \alpha < F(x)\}$ (e.g., $(\alpha_1, \alpha_2)$ in figure \ref{fig:FQ}) is a non-empty set disjoint from $F(A)$,
+then $Q_F$ takes the constant value $x$ 
 on the closure $\{\alpha \mid F(x-)\leq \alpha \leq F(x)\}$. Conversely, if $F$ has no discrete component on $A$, 
-then $Q_{F}\big|_{F(A)}$ is strictly increasing.
+then $Q_{F}\big|_{F(A)}$ is strictly increasing (e.g., on $[\alpha_2, \infty)$ in figure \ref{fig:FQ}).
 
@@
 which will be the cdf of the random variable $Y_Q := \min(Q(U))$ where $U \sim \mathrm{Unif}[0,1]$.
+\begin{figure}
+\vspace{-4cm}
+\includegraphics[width=.9\textwidth]{../figures/FQ.pdf}
+\vspace{-15mm}
+\caption{Example of a cdf $F$ (with graph including solid line and filled points) superimposed on its 
+quantile function $Q_F$ (with graph including dashed and solid lines as well as both filled and empty points).}
+\label{fig:FQ}
+\end{figure}
 
@@
 analytical solutions to \eqref{eqn:quantiles-sum-to-K}, we seek a method of generating an approximation $\tilde{\tau}$ to 
-a solution $\tau^{\star}$ which offers control (at least in principle) of the Bayes risk error or ``optimality gap''
+a solution $\tau^{\star}$ which offers control (at least in principle) over a robust measure of scoring rule error such as
+the sum of absolute coordinate errors
 \begin{align}
-\left|\mathbb{E}_F\left[s_A(\bm{q}_{\tilde{\tau},F}, Y)\right] - \mathbb{E}_F[s_A(x^{F,K}, Y)]\right| = 
-L\left| \sum_{i=1}^{N}\mathbb{E}_{F_i}\left[(Y_i-q_{\tilde{\tau},F_i})_+ - (Y_i-x_i^{F,K})_+\right]\right|
+\mathrm{SAE}(Y,\tilde{\tau}) = \sum_{i=1}^{N} \left| (Y_i-q_{\tilde{\tau},F_i})_+ - (Y_i-x_i^{F,K})_+ \right|
 \end{align}
+% of the Bayes risk error or ``optimality gap''
+% \begin{align}
+% \left|\mathbb{E}_F\left[s_A(\bm{q}_{\tilde{\tau},F}, Y)\right] - \mathbb{E}_F[s_A(x^{F,K}, Y)]\right| = 
+% L\left| \sum_{i=1}^{N}\mathbb{E}_{F_i}\left[(Y_i-q_{\tilde{\tau},F_i})_+ - (Y_i-x_i^{F,K})_+\right]\right|
+% \end{align}
 where $\bm{q}_{\tilde{\tau},F}$ solves the linear program

### 2023-12-12 alloscore_manuscript/alloscore-application.Rnw
adding complete multi-k section draft to manuscript.

@@
 
-The integrated allocation score summarizes allocation scores across a range of possible values of the constraint ($K$), possibly taking weights into account for different values of $K$ that might be more or less likely (Section \ref{sec:methods.detailed.integrated_allocation}).
+The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the constraint ($K$), possibly taking weights into account for different values of $K$ that might be more or less likely (Section \ref{sec:methods.detailed.integrated_allocation}).
 This could be useful in situations where the actual constraint may not be known, or could only be specified as a probability distribution.
@@
 
-Allocation scores (AS) were computed for a range of $K$ from 200 to 60,000 for forecasts made on 2020-12-20 predicting levels of hospitalizations on 2021-01-03 (Figure \ref{fig:multi-k}A).
-These calculations highlight that allocation scores were highest at the $K$ nearest to the observed value of \Sexpr{format(tot_y, big.mark = ",")}.
-Rankings of allocation scores from all models were fairly stable across a range of values for K, with some crossings, especially at $K$ values further away from the observed value.
+AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on 2020-12-20 predicting levels of hospitalizations on 2021-01-03 (Figure \ref{fig:multi-k}A).
+These calculations highlight that AS was highest at the $K$ nearest to the observed value of \Sexpr{format(tot_y, big.mark = ",")}.
+Rankings of AS from all models were fairly stable across a range of values for K, with some crossings, especially at $K$ values further away from the observed value.
 
-Integrated allocation scores (IAS) were computed for two distributions of $K$, one uniform across the entire range and the other a symmetric distribution around 15,000 (Figure \ref{fig:multi-k}B).
+IAS was computed for two distributions of $K$, one uniform across the entire range and the other a symmetric distribution around 15,000 (Figure \ref{fig:multi-k}B).
 Both versions of the IAS were correlated with the original AS conducted at $K=15,000$, with the higher correlation coming from the distribution that was centered at $K=15,000$ (Figure \ref{fig:multi-k}C).
-
-Interpretation of Figure \ref{fig:multi-k} needed.
+Model rankings based on the AC and the centered IAS were roughly similar, with the top and bottom two models being the same for both scores (Table \ref{tab:multi-k-scores}).
 
@@
          `IAS uniform` = IAS) |>
-  xtable::xtable(caption = "Allocation scores shown in Figure \\ref{fig:multi-k}C, sorted by Integrated Allocation Scores (IAS) centered at 15,000.", digits = 0) |>
+  arrange(-`IAS centered at 15k`) |>
+  xtable::xtable(caption = "Allocation scores shown in Figure \\ref{fig:multi-k}C, sorted by Integrated Allocation Scores (IAS) centered at 15,000.",
+                 digits = 0,
+                 label = "tab:multi-k-scores") |>
   print(include.rownames = FALSE, )

### 2023-12-12 alloscore_manuscript/alloscore-application.Rnw
trimming long bibliography authorships. closes #76.

@@
 \usepackage{natbib}
-\bibliographystyle{unsrtnat}
+\bibliographystyle{unsrtnat-abbr}
 
@@
          `IAS uniform` = IAS) |>
-  arrange(-`IAS centered at 15k`) |>
+  arrange(`IAS centered at 15k`) |>
   xtable::xtable(caption = "Allocation scores shown in Figure \\ref{fig:multi-k}C, sorted by Integrated Allocation Scores (IAS) centered at 15,000.",

### 2023-12-12 alloscore_manuscript/alloscore-application.Rnw
tamed captions for 5 and 7. closes #75

@@
 Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022.
-These are the days for which forecasts were made and evaluated.
+These are the dates for which forecasts were made and evaluated.
 A horizontal dashed line at 15,000 shows the assumed resource constraint $K$.
-Panel B shows allocation scores for each model's 14-day ahead forecast, across all US states.
+Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states.
 The x-axis corresponds to the date for which the prediction was made.
-Allocation scores typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down).
+AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down).
 Panel C shows the MWIS metric across weeks, averaged across all states.
-Similarly to panel B, the x-axis corresponds to the date for which the prediction was made.
-MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the pean of the Omicron wave.">>=
+MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.">>=
 p2 <- ggplot(alloscore_data_summaries,
@@
 
-<<multi-k, fig.height=10, fig.width = 8, fig.cap = "Allocation scores across different resource constraints (K) for the 10 models that made forecasts on December 20, 2021.
-Panel A shows, for each model, the allocation score assigned for values of K between 200 and 60,000 at increments of 200.
-The allocation scores show a sharp peak just under 20,000 which was the eventually observed number of hospitalizations.
-It is hardest to get allocations exactly right, with no excess unmet need, when the resource constraint is exactly equal to the number of resources needed.
+<<multi-k, fig.height=10, fig.width = 8, fig.cap = "Allocation scores (AS) across different resource constraints (K) for 10 models that made forecasts on 2021-12-20.
+Panel A shows, for each model, the AS for values of K between 200 and 60,000 at increments of 200.
+The AS show a sharp peak just under 20,000, near the eventually observed number of hospitalizations.
 Panel B shows two possible weighting functions for the Integrated Allocation Score (IAS).
@@
 The second (light orange triangles) uses a uniform weight for all possible values of K.
-Note that the Allocation Score (AS), used in earlier sections of the application, only uses the allocation scores at the `pre-defined' value of K=15,000.
-the  Panel C shows how either method of computing the IAS (y-axis, again in blue circles and orange squares) is correlated with AS (x-axis, computed at K=15,000).
+Note that the AS used in earlier sections of the application uses the single fixed value of K=15,000.
+Panel C shows how either method of computing the IAS (y-axis) is correlated with AS at K=15,000 (x-axis).
 Every point represents the AS and IAS for one model.
-Note that the IAS centered at 15,000 are much more closely correlated  with the AS values, which makes sense as this version of the IAS assigns higher weights to values of the IAS that are closer to the single K at which the AS scores were computed.">>=
+The IAS centered at 15,000 are more closely correlated with the AS values.">>=
 p3 <- slim_alloscore_2021.12.20 |>

### 2023-12-12 alloscore_manuscript/supplement.rnw
update fig 2 to exhibit more FQ behaviors

@@
 \usepackage{pgfplots}
+\usepackage{float}
 \usepackage{xr-hyper}
@@
 $(y - x)_{+} = \short^{y}(x) = \short_{x}(y)$.  Note that 
-$\short^{y}(x)$ and $\short_{x}(y)$ are both convex functions and ``mirror'' each other:
+$\short^{y}(x)$ and $\short_{x}(y)$ are both convex functions and ``mirror'' each other (see figure \ref{fig:shortages}).
 
 \pgfplotsset{compat=1.16} % Set the compatibility of pgfplots
-
-\begin{figure}
+\begin{figure}[H]
 \begin{tikzpicture}
@@
 $\short_{x_0}(y)$, shown in red, gives the resource shortage as a function of resource demand, $y$, for a fixed value of resource supply $x_0$.}
+\label{fig:shortages}
 \end{figure}
@@
 
-The probability levels $\{\alpha_i\}$ for which $\#Q_F(\alpha_i)>1$ ($\{\alpha_3\}$ in the example of figure \ref{fig:FQ}) form a 
+The probability levels $\{\alpha_i\}$ for which $\#Q_F(\alpha_i)>1$ ($\{\alpha_1, \alpha_5\}$ in the example of figure \ref{fig:FQ}) form a 
 discrete subset of $(0,1)$ and correspond to
@@
 $A$, then extending $F^{-1}$ by $F^{-1}(0)=\inf(\mathrm{supp}(F))$ and $F^{-1}(1)=\sup(\mathrm{supp}(F))$ provides an
-inverse to $F$ on $F(\mathbb{R})\subset [0,1]$. If $F$ has a point mass at $x\in A$ (e.g., $x_1 \in [0,\infty)$ in figure \ref{fig:FQ}),
-so that $\{\alpha \mid F(x-)< \alpha < F(x)\}$ (e.g., $(\alpha_1, \alpha_2)$ in figure \ref{fig:FQ}) is a non-empty set disjoint from $F(A)$,
+inverse to $F$ on $F(\mathbb{R})\subset [0,1]$. If $F$ has a point mass at $x\in A$ (e.g., $x_3 \in [0,\infty)$ in figure \ref{fig:FQ}),
+so that $\{\alpha \mid F(x-)< \alpha < F(x)\}$ (e.g., $(\alpha_3, \alpha_5)$ in figure \ref{fig:FQ}) is a 
+non-empty set disjoint from $F(A)$,
 then $Q_F$ takes the constant value $x$ 
 on the closure $\{\alpha \mid F(x-)\leq \alpha \leq F(x)\}$. Conversely, if $F$ has no discrete component on $A$, 
-then $Q_{F}\big|_{F(A)}$ is strictly increasing (e.g., on $[\alpha_2, \infty)$ in figure \ref{fig:FQ}).
+then $Q_{F}\big|_{F(A)}$ is strictly increasing (e.g., on $[\alpha_4, \infty)$ in figure \ref{fig:FQ}).
 
@@
 \begin{figure}
-\vspace{-4cm}
-\includegraphics[width=.9\textwidth]{../figures/FQ.pdf}
+\vspace{-1cm}
+\includegraphics[width=\textwidth]{../figures/FQ.pdf}
 \vspace{-15mm}
 \caption{Example of a cdf $F$ (with graph including solid line and filled points) superimposed on its 
-quantile function $Q_F$ (with graph including dashed and solid lines as well as both filled and empty points).}
+quantile function $Q_F$ (with graph including dashed and solid lines as well as both filled and empty points).
+The notation $F(A)$ should be read as $F(x)$ for any $x\in A$.}
 \label{fig:FQ}
@@
 $TQ_F(\tau_L) \cup TQ_F(\tau_U) \subset [q^{-}_{F_T,\tau_L}, q^{+}_{F_T,\tau_U}]$
-is arranged in exactly one of the following ways: 
+is arranged in exactly one of the following ways (where e.g. refers to figure \ref{fig:FQ}): 
 \begin{itemize}
-\item[($\bullet \bullet$)] $TQ_F(\tau_L) \cap TQ_F(\tau_U) = \varnothing$ (and $q^{+}_{F_T,\tau_L} < q^{-}_{F_T,\tau_U}$)
-\item[($\bullet$)] $TQ_F(\tau_L) = \{K\} = TQ_F(\tau_U)$ (a point mass at $K$)
-\item[($\bullet\!-$)] $TQ_F(\tau_L) = \{q^{-}_{F_T,\tau_U}\} \subsetneq TQ_F(\tau_U)$ (a point mass at $q^{-}_{F_T,\tau_U}$)
-\item[($-\!\bullet$)] $TQ_F(\tau_L) \supsetneq \{q^{+}_{F_T,\tau_L}\} = TQ_F(\tau_U)$ (a point mass at $q^{+}_{F_T,\tau_U}$)
+\item[($\bullet \bullet$)] $TQ_F(\tau_L) \cap TQ_F(\tau_U) = \varnothing$ 
+(and $q^{+}_{F_T,\tau_L} < q^{-}_{F_T,\tau_U}$, e.g., $\tau_L=\alpha_1, \tau_U=\alpha_5$)
+\item[($\bullet$)] $TQ_F(\tau_L) = \{K\} = TQ_F(\tau_U)$ 
+(a point mass at $K$, e.g., $\alpha_3 \leq \tau_L \leq \tau_U \leq \alpha_5$)
+\item[($\bullet\!-$)] $TQ_F(\tau_L) = \{q^{-}_{F_T,\tau_U}\} \subsetneq TQ_F(\tau_U)$ 
+(a point mass at $q^{-}_{F_T,\tau_U}$, e.g., $\tau_L = \alpha_3, \tau_U = \alpha_5$)
+\item[($-\!\bullet$)] $TQ_F(\tau_L) \supsetneq \{q^{+}_{F_T,\tau_L}\} = TQ_F(\tau_U)$ 
+(a point mass at $q^{+}_{F_T,\tau_U}$, e.g., e.g., $\tau_L = \alpha_5, \tau_U = \alpha_4$).
 \end{itemize}

### 2023-12-12 alloscore_manuscript/alloscore-application.Rnw
adding additional Meyers citations, closes #71

@@
 For example, estimates of expected onset of flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola dynamics have been used to allocate surveillance resources \citep{meltzer2014estimating, rainisch2015regional}.
-\cite{bertsimas2021predictionsCOVID} developed tools to inform decision making from infectious disease forecasts, which have been used to inform allocation of limited medical supplies such as ventilators, ICU capacity planning, and vaccine distribution strategy.
+\cite{bertsimas2021predictionsCOVID} developed tools to guide decision making from infectious disease forecasts, which have been used to inform allocation of limited medical supplies such as ventilators, ICU capacity planning, and vaccine distribution strategy.
 Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping (University of Texas, 2022)\nocite{utnews2022}.
@@
 Resource allocation decisions are a realistic example of the kinds of decisions that could motivate more targeted forecasting exercises.
-One real-world example is allocation of ventilators during a respiratory viral pandemic.
+One real-world example is allocation of ventilators during a respiratory viral pandemic\citep{huang_stockpiling_2017}.
 This was the motivating example for the examples presented in the work above.
-However, other examples where allocation is a central concern exist as well, such as the allocation of a limited stockpile of vaccinations\citep{persad_fair_2023} or diagnostic tests\citep{du_optimal_2022}.
+However, other examples where allocation is a central concern exist as well, such as the allocation of a limited stockpile of vaccinations\citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests\citep{du_optimal_2022,pasco_covid-19_2023}.
 

### 2023-12-12 alloscore_manuscript/alloscore-application.Rnw
paragraph in discussion talking about whether the forecast adds useful information to an existing decision-making process. closes #44

@@
 
+An additional limitation of the current work and opportunity for further investigation is to more carefully evaluate whether a forecast adds value to existing decision-making processes.
+In the context of allocation-based decisions, standard procedures might involve simply extrapolating need based on a current observed data (similar to the `baseline' model presented above), with or without adjustments based on other political or real-world considerations.
+For example, in many settings public health stakeholders will synthesize multiple sources of information, e.g.
+real-time data from a wide variety of quantitative and qualitative sources coupled with expert judgment based on a situational understanding of past and current experience.
+The allocation score presented in this work assesses the optimality of the allocation based on an omniscient view of the future.
+However, what we care about is whether a given forecast adds useful information to an existing decision-making process.
+While the scoring procedures as presented do not directly address this question, they could be modified (say, by comparison to a `baseline` model) to quantify the benefit -if any- of using a forecast to inform a specific decision.
+
 We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work in more closely coupling applied epidemiological forecasting with public health decision-making.

### 2023-12-12 alloscore_manuscript/supplement.rnw
attempting to put stuff about dfq to rest

--- 2023-12-12 alloscore_manuscript/supplement.rnw
@@
 In order for a scoring rule to be guaranteed to yield logically consistent forecast evaluations, it must be \emph{proper}, which means,
-in essence, that a forecaster cannot perceive any forecast at odds with their beliefs about the events being forecasted as having an optimal expected score.
-We formalize this in terms of forecaster ``self-assessment'' in section \ref{sec:proper_def} and then address the propriety of the allocation score in three stages:
+in essence, that a forecaster cannot perceive any forecast that is at odds with their beliefs about the events being forecasted as having an optimal expected score.
+We formalize this in terms of forecaster ``self-assessment'' in section \ref{sec:proper_def} and then address the propriety of the allocation score in two stages:
 \begin{itemize}
   \item Section \ref{sec:alloscore_proper} shows that the allocation score is proper when the forecaster's actual forecast distribution $F$ is available (as is any score derived using the decision theoretic procedures outlined in section 2 of the manuscript)
-  \item Section \ref{sec:distfromq_alloscore_proper_prospective} shows that the allocation score is still proper if the forecast is represented as a member of a parametric family of distributions, so long as this representation and scoring procedure was specified prospectively by the hub.
+  \item Section \ref{sec:distfromq_alloscore_proper_prospective} addresses the propriety of an analysis that combines an approximate parametric representation of the forecast distribution and evaluation with the allocation score.
+It shows that this procedure yields a proper score so long as this representation and scoring procedure was specified prospectively by the hub.
 This includes the use of methods such as \verb`distfromq` for constructing forecast distributions as a special case.
-  \item Section \ref{sec:distfromq_alloscore_improper_post_hoc} shows that \emph{post hoc} application of this scoring procedure to quantile forecasts is improper, and makes some suggestions for organized forecasting efforts who are considering using the allocation score.
+However,  \emph{post hoc} application of this scoring procedure to quantile forecasts is improper.
 \end{itemize}
@@
 In settings where the outcome being forecasted is a continuous quantity (such as rates of influenza-like illness among outpatient doctor visits) or a count (such as influenza hospitalizations), forecasting exercises have therefore resorted to collecting summaries of a forecast distribution such as bin probabilities or predictive quantiles.
-This raises the question of whether the allocation score is still proper if the forecast distribution $F$ is not itself directly recorded.
+This raises the question of whether use of the allocation score still consitutes a proper evaluation procedure if the forecast distribution $F$ is not itself directly recorded.
 The answer we provide in this section is that it is, so long as the fact that the allocation score will be used for forecast evaluation and the method that will be used to obtain a distributional estimate of $F$ from the provided forecast representation are communicated to participating forecasters prospectively.
 
-We consider a setting where a forecasting exercise (such as a forecast hub) pre-specifies that forecasts will be represented using a parametric family of forecast distributions $F_\theta(y)$, and the task of the forecaster is to select a particular parameter value $\theta$.
-We use $\mathcal{P}$ to denote the collection of all distributions $F_\theta$ in the given parametric family.
+We consider a setting where a forecasting exercise (such as a forecast hub) pre-specifies that forecasts will be represented using a parametric family of forecast distributions $G_\theta(y)$, and the task of the forecaster is to select a particular parameter value $\theta$.
+We use $\mathcal{P}$ to denote the collection of all distributions $G_\theta$ in the given parametric family.
 For instance, it has recently been proposed that mixture distributions could be used to represent forecast distributions \cite{wadsworth2023mixture}.
-Additionally, we note that the functionality in \verb`distfromq` can be viewed as specifying a parametric family $\mathcal{P}_{\mathrm{dfq}}$ where the parameters $\theta$ of $F_\theta$
-are its quantiles at pre-specified probability levels, and where the shape of any $F_\theta \in \mathcal{P}_{\mathrm{dfq}}$ over the full range of its support is entirely controlled by these quantiles.
+Additionally, we note that the functionality in \verb`distfromq` can be viewed as specifying a parametric family $\mathcal{P}_{\mathrm{dfq}}$ where the parameters $\theta$ of $G_\theta$
+are its quantiles at pre-specified probability levels, and where the shape of any $G_\theta \in \mathcal{P}_{\mathrm{dfq}}$ over the full range of its support is entirely controlled by these quantiles.
 
@@
 The second is the forecaster's reporting problem where the task is to select parameter values $\theta$ to report.
-The forecaster's loss is given by $s_R(\theta, y) = s_A(x^{F_\theta}, y)$, where $x^{F_\theta}$ is the Bayes act for the allocation problem under the distribution $F_\theta$.
-In words, the loss associated with reporting $\theta$ is equal to the loss associated with taking the Bayes allocation corresponding to the distribution $F_\theta$.
+The forecaster's loss is given by $s_R(\theta, y) = s_A(x^{G_\theta}, y)$, where $x^{G_\theta}$ is the Bayes act for the allocation problem under the distribution $G_\theta$.
+In words, the loss associated with reporting $\theta$ is equal to the loss associated with taking the Bayes allocation corresponding to the distribution $G_\theta$.
 
-Following our usual construction, the Bayes act for the forecast reporting problem is the parameter set that minimizes the forecaster's expected loss:
+Following our usual construction, the Bayes act for the forecast reporting problem is the parameter set that minimizes the forecaster's expected loss.
+Breaking with our earlier notation for improved legibility, we use $\theta^\star(F)$ to denote this Bayes act:
 \begin{align*}
-\theta^F &= \text{argmin}_\theta \Ex_F [s_R(\theta, Y)] \\
-&= \text{argmin}_\theta \Ex_F [s_A(x^{F_\theta}, Y)]
+\theta^\star(F) &= \text{argmin}_\theta \Ex_F [s_R(\theta, Y)] \\
+&= \text{argmin}_\theta \Ex_F [s_A(x^{G_\theta}, Y)]
 \end{align*}
-This leads the forecaster to choose values of the parameters $\theta$ such that the allocation $x^{F_\theta}$ is as close as possible to the allocation $x^F$ under the forecaster's true forecast distribution $F$.
-\apg{I think we need to say something about what we mean by ``close''.  The score divergence is what we are confronting here and I appreciate that 
-others may not want to broach the topic.  But I think we have to say something about ``score expectation defined distance''. This also relates directly to the notation crunch below.}
 
 We then arrive at the scoring rule
-$$S_R(F, y) = s_R(\theta^F, y) = s_A(x^{F_{\theta^F}}, y).$$
-\elr{Notation got pretty bad here... borderline about whether it's worth searching for a better notation for "the bayes allocation for the distribution with the bayes parameters"}
+$$S_R(F, y) = s_R(\theta^\star(F), y) = s_A(x^{G_{\theta^\star(F)}}, y).$$
 It follows from the discussion in section \ref{sec:alloscore_proper} that this is a proper scoring rule for $F$.
-Additionally, we note that the score can be calculated from the reported parameter values $\theta^F$.
+Additionally, we note that the score can be calculated from the reported parameter values $\theta^\star(F)$.
 
 We emphasize that the forecaster's true predictive distribution $F$ does not need to be a member of the specified parametric family $\mathcal{P}$ for this construction to yield a proper score.
-It is, however, necessary to specify the parametric family to use and the foundational scoring rule $s_A$ (including any relevant problem parameters such as the resource constraint $K$) in advance, so that forecasters can identify the Bayes act parameter set $theta^F$ to report.
+It is, however, necessary to specify the parametric family to use and the foundational scoring rule $s_A$ (including any relevant problem parameters such as the resource constraint $K$) in advance, so that forecasters can identify the Bayes act parameter set $\theta^\star(F)$ to report.
 
-As an alternative, a forecast hub could ask forecasters to provide the Bayes allocations associated with their forecasts for one or more fixed resource constraints $K$, allowing for direct calculation of scores and narrowing the gap between model outputs and public health end users.
+For a given resource constraint $K$, if it is possible to find a member of the specified parametric family $\mathcal{P}$ with the same allocation as $F$ for all possible forecast distributions $F$ (i.e., for any $F$ there is a $G_\theta \in \mathcal{P}$ such that $x^F = x^{G_\theta}$), the scoring rule $S_R$ will yield the same score as the allocation score $S_A$.
+For the particular choice of the parametric family $\mathcal{P}_{dfq}$ (i.e., using the \verb`distfromq` package), this is the case.
+For instance, the forecaster could pick one required quantile level (such as 0.5, for which the corresponding predictions are predictive medians), and set the submitted quantiles of their forecast distribution at that level to be the desired allocations.
+However, this representation of the forecast may be quite different from the actual forecast distribution $F$.
 
-\subsection{Post hoc evaluation of quantile forecasts using distfromq and the allocation score is improper}
-\label{sec:distfromq_alloscore_improper_post_hoc}
+A key observation is that a \emph{post hoc} analysis combining \verb`distfromq` with the allocation score does not in general yield a proper evaluation method.
+This is because the forecast distribution $F$ and the distribution $G_\theta \in \mathcal{P}_{dfq}$ with the same quantiles as $F$ may determine quite different resource allocations, particularly if the tail extrapolations performed by \verb`distfromq` do not match the tail behavior of $F$.
 
-In this section, we illustrate that a post hoc evaluation of quantile forecasts using \verb`distfromq` and the allocation score is not proper.
-In brief, this is because the Bayes parameters $\theta^F$ for the forecaster to report may not align with quantiles of the distribution $F$ at the specified probability levels.
-We illustrate this with examples (borrow from elicitation.R...)
+As another alternative for practical forecasting exercises, a forecast hub could ask forecasters to provide the Bayes allocations associated with their forecasts for one or more fixed resource constraints $K$.
+This would allow for direct calculation of scores and narrowing the gap between model outputs and public health end users.
+For this to be feasible, implementations of the allocation algorithm would have to be provided to participating forecasters.
 

### 2023-12-13 alloscore_manuscript/supplement.rnw
more updates to supplement r.e. use of distfromq

@@
 
-For a given resource constraint $K$, if it is possible to find a member of the specified parametric family $\mathcal{P}$ with the same allocation as $F$ for all possible forecast distributions $F$ (i.e., for any $F$ there is a $G_\theta \in \mathcal{P}$ such that $x^F = x^{G_\theta}$), the scoring rule $S_R$ will yield the same score as the allocation score $S_A$.
-For the particular choice of the parametric family $\mathcal{P}_{dfq}$ (i.e., using the \verb`distfromq` package), this is the case.
+If the parametric family used to represent forecast distributions is flexible enough, the reporting scoring rule $S_R$ and the allocation score are equivalent in the sense that they will yield the same score for any distribution $F$.
+Suppose that for a given resource constraint $K$, for any forecast distribution $F$ it is possible to find a member $G_{\theta^\star}$ of the specified parametric family $\mathcal{P}$ with the same allocation as $F$ (i.e., $x^F = x^{G_{\theta^\star}}$).
+Then $\theta^\star$ is a Bayes act for the reporting problem since for any other parameter value $\theta$,
+\begin{align*}
+\Ex_F[s_R(\theta^\star, Y)] &= \Ex_F[ s_A(x^{G_{\theta^\star}}, Y) ] \\
+&= \Ex_F[s_A(x^F, Y)]  \quad \text{ (since $x^F = x^{G_{\theta^\star}}$)} \\
+&\leq \Ex_F[ s_A(x^{G_\theta}, Y) ]  \quad \text{ (by definition of $x^F$)} \\
+&= \Ex_F[ s_R(\theta, Y)].
+\end{align*}
+It therefore follows that
+\begin{align*}
+S_R(F, y) &= s_R(\theta^\star, y) \\
+&= s_A(x^{G_{\theta^\star}}, y) \\
+&= s_A(x^F, y) \\
+&= S_A(F, y).
+\end{align*}
+
+For the particular choice of the parametric family $\mathcal{P}_{\text{dfq}}$ (i.e., using the \verb`distfromq` package), this flexibility requirement is satisfied.
 For instance, the forecaster could pick one required quantile level (such as 0.5, for which the corresponding predictions are predictive medians), and set the submitted quantiles of their forecast distribution at that level to be the desired allocations.
@@
 A key observation is that a \emph{post hoc} analysis combining \verb`distfromq` with the allocation score does not in general yield a proper evaluation method.
-This is because the forecast distribution $F$ and the distribution $G_\theta \in \mathcal{P}_{dfq}$ with the same quantiles as $F$ may determine quite different resource allocations, particularly if the tail extrapolations performed by \verb`distfromq` do not match the tail behavior of $F$.
+This is because the forecast distribution $F$ and the distribution $G_\theta \in \mathcal{P}_{\text{dfq}}$ with the same quantiles as $F$ may determine quite different resource allocations, particularly if the tail extrapolations performed by \verb`distfromq` do not match the tail behavior of $F$.
 
-As another alternative for practical forecasting exercises, a forecast hub could ask forecasters to provide the Bayes allocations associated with their forecasts for one or more fixed resource constraints $K$.
-This would allow for direct calculation of scores and narrowing the gap between model outputs and public health end users.
-For this to be feasible, implementations of the allocation algorithm would have to be provided to participating forecasters.
+As another alternative for practical forecasting exercises, a forecast hub could ask forecasters to directly provide the Bayes allocations associated with their forecasts for one or more specified resource constraints $K$.
+At the cost of increasing the number of quantities solicited by the forecast hub, this would have several advantages: it would prevent any artificial distortion of the forecast distributions, allow for direct calculation of scores, and narrow the gap between model outputs and public health end users.
+For this to be feasible, implementations of the allocation algorithm would have to be provided to participating forecasters in the computational languages being used for modeling.
 

### 2023-12-13 alloscore_manuscript/supplement.rnw
use updated fig 2 to shorten bisection explanation

--- 2023-12-12 alloscore_manuscript/supplement.rnw
@@
 In the case ($\bullet$), we can immediately take 
-$\tau^{\star}=\tau_U$ as the probability level defining the allocation Bayes act.
+$\tau^{\star}=\tau_U$ as the probability level representing the allocation Bayes act.
 In the cases
@@
 component forecasts adjacent to a region of zero density in all components, we can take $\tau^{\star}=\tau_U$ or $\tau_L$, 
-respectively, as the representing probability level and find Bayes acts within the set 
-$\undertilde{\bm{F}}^{-1}(\tau^{\star})$ (by solving the linear program 
-$\sum_{x \in \undertilde{\bm{F}}^{-1}(\tau^{\star})}x_i = K$).
+respectively, as the representing probability level.  
+Having found our $\tau^{\star}$ we then arrive at a Bayes act $x^{F,K}$ by by solving the
+linear program $\sum_{x \in \undertilde{\bm{F}}^{-1}(\tau^{\star})}x_i = K$.
 
-Otherwise, by evaluating $TQ_F$ at $\tau_M = \frac{1}{2}\left(\tau_L + \tau_U\right)$ we can replace $I_1$ with one of its 
+In the remaining and typical case of ($\bullet \bullet$), further search is generally necessary. We can proceed 
+by evaluating $TQ_F$ at $\tau_M = \frac{1}{2}\left(\tau_L + \tau_U\right)$ and replacing $I_1$ with one of its 
 halves
@@
 In the generic case of an infinite $\{I_k\}$, we need to define practical stopping criteria for the possible limit behaviours of 
-$F_T(K \pm \varepsilon)$ as $\varepsilon \searrow 0$:
-\begin{itemize}
-\item[($\mathrlap{\,\bullet}{\,/}\,$)] $F_T(K - \varepsilon) < F_T(K-) = F_T(K) < F_T(K + \varepsilon)$
-(so that $\tau^{\star} = F_T(K-)$ and $TQ_F(\tau^{\star}) = \{K\}$)
-\item[($/\!\!\raisebox{.9ex}{$\bullet$\!-}$)] $F_T(K - \varepsilon) < F_T(K-) = F_T(K) = F_T(K + \varepsilon)$
-\item[($\raisebox{-.9ex}{-\!$\bullet$}\!\!/$)] $F_T(K - \varepsilon) = F_T(K) < F_T(K + \varepsilon)$
-\item[($\raisebox{-.9ex}{\phantom{-}\!$\bullet$}\!\!/$)] $F_T(K - \varepsilon) \leq F_T(K-) < F_T(K) < F_T(K + \varepsilon)$
-\item[($\,\bullet\!\text{-}$)] $F_T(K - \varepsilon) \leq F_T(K-) < F_T(K) = F_T(K + \varepsilon)$
-\item[($\text{-}\!\!\bullet\!\!\text{-}$)]  $F_T(K - \varepsilon) = F_T(K) = F_T(K + \varepsilon)$
-so that $K \in (q^{-}_{F_T,\tau^{\star}}, q^{+}_{F_T,\tau^{\star}})$ and all component forecasts $F_i$ have 
-zero density at points $x_i \in (q^{-}_{F_i,\tau^{\star}}, q^{+}_{F_i,\tau^{\star}})$ and $\sum x_i = K$.
-\end{itemize}
+$F_T(K \pm \varepsilon)$ as $\varepsilon \searrow 0$ which are exemplified in figure \ref{fig:FQ} at $0, x_1, x_2, x_3, (x_3 +x_4)/2$ and 
+$x_5$.
+% \begin{itemize}
+% \item[($\mathrlap{\,\bullet}{\,/}\,$)] $F_T(K - \varepsilon) < F_T(K-) = F_T(K) < F_T(K + \varepsilon)$
+% (so that $\tau^{\star} = F_T(K-)$ and $TQ_F(\tau^{\star}) = \{K\}$)
+% \item[($/\!\!\raisebox{.9ex}{$\bullet$\!-}$)] $F_T(K - \varepsilon) < F_T(K-) = F_T(K) = F_T(K + \varepsilon)$
+% \item[($\raisebox{-.9ex}{-\!$\bullet$}\!\!/$)] $F_T(K - \varepsilon) = F_T(K) < F_T(K + \varepsilon)$
+% \item[($\raisebox{-.9ex}{\phantom{-}\!$\bullet$}\!\!/$)] $F_T(K - \varepsilon) \leq F_T(K-) < F_T(K) < F_T(K + \varepsilon)$
+% \item[($\,\bullet\!\text{-}$)] $F_T(K - \varepsilon) \leq F_T(K-) < F_T(K) = F_T(K + \varepsilon)$
+% \item[($\text{-}\!\!\bullet\!\!\text{-}$)]  $F_T(K - \varepsilon) = F_T(K) = F_T(K + \varepsilon)$
+% so that $K \in (q^{-}_{F_T,\tau^{\star}}, q^{+}_{F_T,\tau^{\star}})$ and all component forecasts $F_i$ have 
+% zero density at points $x_i \in (q^{-}_{F_i,\tau^{\star}}, q^{+}_{F_i,\tau^{\star}})$ and $\sum x_i = K$.
+% \end{itemize}
 
-We also want to be able to decribe what further ``post-processing'' steps are required to find a sufficiently good approximation
-to a Bayes act within the set
-$\undertilde{\bm{F}}^{-1}(I_{\overline{k}})$ where $I_{\overline{k}}$ is the terminal search interval for $\tau^{\star}$.
+And after having obtained sufficiently a sufficient narrow terminal $I_K = [\tau_{L,K}, \tau_{U,K}]$ we need a protocol for 
+how to select a solution to the linear program $\sum_{x \in \undertilde{\bm{F}}^{-1}(I_K)}x_i = K$ as an approximate Bayes act.
 

### 2023-12-13 alloscore_manuscript/alloscore-application.Rnw
removing sentence. closes #58

@@
 For many locations, forecasts predicted lower values than were eventually observed, as this was during the period of steep increase of viral transmission across many states.
-Of the 10 models evaluated, the CU-select model had the most accurate forecasts according to the allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table \ref{tab:one-week-results}).
+Of the 10 models evaluated, the CU-select model had the most accurate forecasts according to the allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table \ref{tab:multi-k-scores}).
 The JHUAPL-Bucky model had the second best MWIS but the third worst allocation score.
 
-<<metrics-rankings, results = "asis">>=
+<<ias-comparison>>=
+tar_load(alloscore_2021.12.20)
+slim_alloscore_2021.12.20 <- alloscore::slim(alloscore_2021.12.20, id_cols = c("model", "reference_date"))
+
+## for visual context
+## curve(dnorm(x, mean=15000, sd=3000), from=0, to=30000)
+ks <- seq(5000, 25000, by=200)
+kprobs <- data.frame(K = ks,
+                     weight = dnorm(ks, mean=15000, sd=3000)/sum(dnorm(ks, mean=15000, sd=3000)))
+
+ias <- slim_alloscore_2021.12.20 |>
+  group_by(model, K) |>
+  summarize(`Allocation score` = sum(components)) |>
+  left_join(kprobs) |>
+  group_by(model) |>
+  summarize(IAS_wt = sum(`Allocation score`*weight, na.rm=TRUE),
+            IAS = mean(`Allocation score`)) |>
+  arrange(IAS)
+
+as <- slim_alloscore_2021.12.20 |>
+  filter(K == 15000) |>
+  group_by(model) |>
+  summarize(AS = sum(components)) |>
+  arrange(AS)
+@
+
+<<multi-k-scores, results='asis'>>=
+wis_vals <- all_scores |>
+  filter(target_end_date == "2022-01-03", !is.na(mwis)) |>
+  select(model,
+# can uncomment the two lines below to check they match
+#         AS = alloscore,
+#         `AS rank` = allo_rank,
+         MWIS = mwis)
+
+left_join(ias, as) |>
+  left_join(wis_vals, by="model") |>
+  rename(`IAS centered at 15k` = IAS_wt,
+         `IAS uniform` = IAS) |>
+  select(model, AS, MWIS, `IAS centered at 15k`, `IAS uniform`) |>
+  arrange(AS) |>
+  xtable::xtable(caption = "Comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS).
+All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03.
+Results are sorted by AS.
+For all metrics, lower scores indicate better accuracy.",
+                 digits = 0,
+                 label = "tab:multi-k-scores") |>
+  print(include.rownames = FALSE, )
+@
+
+<<metrics-rankings, results = "asis", eval=FALSE>>=
+## NOTE: this content now shown in multi-k-scores chunk
 ## for highlighting cells, if we want that https://stackoverflow.com/questions/37448348/color-code-cells-of-xtable
@@
   filter(target_end_date == "2022-01-03", !is.na(mwis)) |>
-  select(model, alloscore, allo_rank, mwis, mwis_rank) |>
-  arrange(allo_rank) |>
+  select(model,
+         AS = alloscore,
+         `AS rank` = allo_rank,
+         MWIS = mwis,
+         `MWIS rank` = mwis_rank) |>
+  arrange(`AS rank`) |>
   xtable::xtable(label = "tab:one-week-results",
-                 caption = "caption TBD") |>
+                 caption = "Allocation scores (AS) and mean weighted interval scores (MWIS) for each model' as forecasted's forecasts for ") |>
   print(include.rownames = FALSE)
@@
 source("R/plot_functions.R")
-tar_load(alloscore_2021.12.20)
-slim_alloscore_2021.12.20 <- alloscore::slim(alloscore_2021.12.20, id_cols = c("model", "reference_date"))
 plot_hosp(models = c("JHUAPL-Bucky", "CU-select"),
@@
 
-<<ias-comparison>>=
-## for visual context
-## curve(dnorm(x, mean=15000, sd=3000), from=0, to=30000)
-ks <- seq(5000, 25000, by=200)
-kprobs <- data.frame(K = ks,
-                     weight = dnorm(ks, mean=15000, sd=3000)/sum(dnorm(ks, mean=15000, sd=3000)))
-
-ias <- slim_alloscore_2021.12.20 |>
-  group_by(model, K) |>
-  summarize(`Allocation score` = sum(components)) |>
-  left_join(kprobs) |>
-  group_by(model) |>
-  summarize(IAS_wt = sum(`Allocation score`*weight, na.rm=TRUE),
-            IAS = mean(`Allocation score`)) |>
-  arrange(IAS)
-
-as <- slim_alloscore_2021.12.20 |>
-  filter(K == 15000) |>
-  group_by(model) |>
-  summarize(AS = sum(components)) |>
-  arrange(AS)
-
-ias_to_plot <- left_join(ias, as) |>
-  pivot_longer(cols = c(IAS, IAS_wt), names_to = "IAS_type", values_to = "IAS") |>
-  mutate(IAS_type = ifelse(IAS_type=="IAS", "uniform", "centered at 15,000"))
-
-p1 <- ggplot(ias_to_plot, aes(x=AS, y=IAS, color=IAS_type, shape=IAS_type)) +
-  geom_point() +
-  geom_smooth(method="lm", se=FALSE, linetype=2) +
-  scale_color_manual(name = "Integrated Allocation Score", values = c("#386cb0", "#fdc086")) +
-  scale_shape_manual(name = "Integrated Allocation Score", values = c(1, 2)) +
-  geom_abline(slope=1, intercept=0, color="gray") +
-  theme(legend.position = c(0.05,.9), legend.justification = c(0,1),
-        legend.background=element_rect(fill = alpha("white", 0.2)))
-
-p2 <- slim_alloscore_2021.12.20 |>
-  group_by(model, K) |>
-  summarize(`Allocation score` = sum(components)) |>
-  left_join(kprobs) |>
-  ggplot(aes(x=K)) +
-  geom_point(aes(y=weight), color = "#386cb0") +
-  geom_point(y=1/length(unique(slim_alloscore_2021.12.20$K)), color="#fdc086", shape = 2) +
-  geom_vline(xintercept = 15000, color="gray")
-@
-
-<<multi-k-scores, results='asis'>>=
-left_join(ias, as) |>
-  rename(`IAS centered at 15k` = IAS_wt,
-         `IAS uniform` = IAS) |>
-  arrange(`IAS centered at 15k`) |>
-  xtable::xtable(caption = "Allocation scores shown in Figure \\ref{fig:multi-k}C, sorted by Integrated Allocation Scores (IAS) centered at 15,000.",
-                 digits = 0,
-                 label = "tab:multi-k-scores") |>
-  print(include.rownames = FALSE, )
-@
-
 <<multi-k, fig.height=10, fig.width = 8, fig.cap = "Allocation scores (AS) across different resource constraints (K) for 10 models that made forecasts on 2021-12-20.
@@
 The IAS centered at 15,000 are more closely correlated with the AS values.">>=
+
+ias_to_plot <- left_join(ias, as) |>
+  pivot_longer(cols = c(IAS, IAS_wt), names_to = "IAS_type", values_to = "IAS") |>
+  mutate(IAS_type = ifelse(IAS_type=="IAS", "uniform", "centered at 15,000"))
+
+p1 <- ggplot(ias_to_plot, aes(x=AS, y=IAS, color=IAS_type, shape=IAS_type)) +
+  geom_point() +
+  geom_smooth(method="lm", se=FALSE, linetype=2) +
+  scale_color_manual(name = "Integrated Allocation Score", values = c("#386cb0", "#fdc086")) +
+  scale_shape_manual(name = "Integrated Allocation Score", values = c(1, 2)) +
+  geom_abline(slope=1, intercept=0, color="gray") +
+  theme(legend.position = c(0.05,.9), legend.justification = c(0,1),
+        legend.background=element_rect(fill = alpha("white", 0.2)))
+
+p2 <- slim_alloscore_2021.12.20 |>
+  group_by(model, K) |>
+  summarize(`Allocation score` = sum(components)) |>
+  left_join(kprobs) |>
+  ggplot(aes(x=K)) +
+  geom_point(aes(y=weight), color = "#386cb0") +
+  geom_point(y=1/length(unique(slim_alloscore_2021.12.20$K)), color="#fdc086", shape = 2) +
+  geom_vline(xintercept = 15000, color="gray")
+
 p3 <- slim_alloscore_2021.12.20 |>
@@
 
-All proper scoring rules for probabilistic forecasts have an explicit link to a loss function, thereby placing numerical value on different kinds of errors in specifying forecasts.
 In epidemiological forecasting, well-known proper scoring rules such as the log score or variations on the continuous rank probability score (such as the weighted interval score, or WIS) have been frequently utilized to evaluate probabilistic forecasts.

### 2023-12-13 alloscore_manuscript/alloscore-application.Rnw

@@
 \subsubsection{Hospitalization data}
-Starting in the summer of 2020, the US Health and Human Services began reporting counts of daily new admissions to hospitals for individuals with COVID-19.\citep{healthdatagov_covid-19_nodate}
-These data served as the source of ``ground truth'' data for the US COVID-19 Forecast Hub which, starting in December 2020, collected short-term forecasts of new hospital admissions at the daily scale.
+Starting in the summer of 2020, the US Health and Human Services (HHS) began reporting counts of daily new admissions to hospitals for individuals with COVID-19 \citep{healthdatagov_covid-19_nodate}.
 These daily counts were available for the US as a whole, and all states and several additional jurisdictions such as Puerto Rico and Washington DC.
 The data were updated daily and were available for download by the public through the HHS HealthData.gov website.
-For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data.\cite{wang-covidhubutils}
-The hospitalization data were downloaded for analysis on \Sexpr{as.Date(metadata$time[which(metadata$name == "truth_data")])}.
+For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data \cite{wang-covidhubutils}.
+The hospitalization data were downloaded for analysis on \Sexpr{format(as.Date(metadata$time[which(metadata$name == "truth_data")]), "%B %d, %Y")}.
 
 \subsubsection{Forecast data}
-The US COVID-19 Forecast Hub, a consortium funded by the US CDC and led by a research group at UMass-Amherst, collected short-term forecasts of hospitalizations starting in December 2020.\citep{cramer_united_2022}
-Any team that with appropriately formatted forecasts could submit them to the Forecast Hub data repository on GitHub.(cite repo site)
+The US COVID-19 Forecast Hub, a consortium funded by the US CDC and led by a research group at the University of Massachusetts-Amherst, collected short-term forecasts of new hospital admissions at the daily scale starting in December 2020, using the HHS data as a source of ``ground truth'' \citep{cramer_united_2022}.
+Any team that with appropriately formatted forecasts could submit them to the Forecast Hub data repository on GitHub (cite repo site).
 Forecasts were time-stamped by GitHub upon submission and passed validation checks that ensured correct formatting and that the forecasts were being submitted only for dates in the future, not for data that had already been observed.
@@
 Forecasts could be submitted on any day during the week.
-However once a week on Mondays, the Forecast Hub would collect the most recent forecasts submitted by all teams that met certain inclusion criteria and create an ensemble forecast using quantile averaging.\citep{ray_comparing_2023}
+However once a week on Mondays, the Forecast Hub would collect the most recent forecasts submitted by all teams that met certain inclusion criteria and create an ensemble forecast using quantile averaging \citep{ray_comparing_2023}.
 An ensemble that treated all models equally was created (COVIDhub-ensemble) as was a model that created weights of submitted models based on performance in the past 12 weeks (COVIDhub-trained\_ensemble).
@@
 All forecasts, including the ensemble, were submitted as probabilistic predictions about the number of new hospital admissions on a particular day in the future, in a specific jurisdiction of the US (national level, state, or territory).
-Probability distributions were specified, per the requirements established by the Forecast Hub, as a set of 23 quantiles for each individual prediction.
-The submitted quantiles included a median (treated as a ``point'' prediction) and defined 11 central prediction intervals, from a 99\% to a 10\% prediction interval.
+Probability distributions were represented using a set of 23 quantiles for each individual prediction.
+The submitted quantiles included a median and the lower and upper limits of 11 central prediction intervals, from a 99\% to a 10\% prediction interval.
 
 The analysis in this work focuses on forecasts made before and during the first wave of the Omicron SARS-CoV-2 variant in the US.
-As such, we downloaded forecasts for the 15 weeks starting with Monday 2021-11-22 through Monday 2022-02-28.
-Forecast data were downloaded on for analysis on \Sexpr{as.Date(metadata$time[which(metadata$name == "forecast_data")])}.
-
-We established a set of inclusion criteria to determine which forecasts and models to include in our analysis.
+As such, we analyzed forecasts for the 15 weeks starting with Monday November 22, 2021 through Monday February 28, 2022.
+%Forecast data were downloaded on for analysis on \Sexpr{format(as.Date(metadata$time[which(metadata$name == "forecast_data")]), "%B %d, %Y")}.
+
+We established a set of criteria to determine which forecasts and models to include in our analysis.
 Models were eligible to be included in the analysis if they were considered a ``primary'' model from a team.
@@
 
-For this analysis, we fixed a resource constraint $K$ to be 15,000, based roughly on a reported number of ventilators available for reallocation in the US.\citep{ajao_assessing_2015}
-For each week, we computed a the allocation score for each 14 day-ahead forecast based on $K=15,000$.
+For this analysis, we fixed the resource constraint $K$ to be 15,000, based roughly on a reported number of ventilators available for reallocation in the US \citep{ajao_assessing_2015}.
+For each week, we computed the allocation score for the 14 day-ahead forecast.
 
@@
 \subsection{Data and code availability}
+
+All forecast data used in this evaluation are available through the COVID-19 Forecast Hub (cite repository). Code for our analyses is available on GitHub (cite repository).
 
@@
 
-For JHUAPL-Bucky, the allocations were assigned at the \Sexpr{Buck_quantile}th quantile of the predictive distribution and for CU-select it was at the \Sexpr{CU_quantile}th percentile.
-This reflects that in general the JHUAPL-Bucky forecasts assigned more probabilities to higher values and thus a low quantile value across all locations reached the allocation limit of 15,000.
-In California, New York and Texas, the eventually observed number of hospital admissions was closer to the predicted median from JHUAPL-Bucky and was contained in JHUAPL-Bucky's 80\% PI.
-However for each of these three states, the resource allocation was lower and therefore resulted in more hypothetical unmet need than CU-select allocations.
-In Florida, the forecast from JHUAPL-Bucky missed by a wide margin (as did the forecast from CU-select) and the allocation was \Sexpr{FL_diff} units lower than what was allocated according to the CU-select model.
+For JHUAPL-Bucky, the allocations were assigned at the \Sexpr{Buck_quantile}th percentile of the predictive distribution and for CU-select they were at the \Sexpr{CU_quantile}th percentile.
+This reflects that in general the JHUAPL-Bucky forecasts assigned more probability to higher values and thus a low quantile value across all locations reached the allocation limit of 15,000.
+In California, New York and Texas, the eventually observed number of hospital admissions was closer to the predicted median from JHUAPL-Bucky than the median from CU-select, and was contained in JHUAPL-Bucky's 80\% PI but not in the 80\% PI from CU-select.
+However for each of these three states, the resource allocation was lower for JHUAPL-Bucky than for CU-select and therefore resulted in more hypothetical unmet need.
+In Florida, which was the state that saw the largest number of reported hospitalizations on this target date, the forecast from JHUAPL-Bucky missed by a wide margin (as did the forecast from CU-select) and the allocation was \Sexpr{FL_diff} units lower than the allocation derived from the CU-select model.
 
@@
 The number of available resources was fixed at 15,000 and forecasts from each model were used to determine an optimal allocation strategy before the resource need was known.
-For most locations the resource need exceeded the resources allocated, indicated by the 'observed resource need' bar being larger than the 'allocation' bar.
+For most locations the resource need exceeded the resources allocated, indicated by the `observed resource need' bar being larger than the `allocation' bar.
 Panel B shows the breakdown of the weighted interval score (WIS) into components of underprediction, overprediction and dispersion.
@@
 Allocation scores varied substantially by date and by model (Figure \ref{fig:metrics-over-time}).
-For predictions made for the first three Mondays in December 2021 and the last three Mondays in February 2022 all models had allocation scores under 500 (and the mean across all models was less than 100), indicating that the unnecessary unmet need was fairly low relative to the total number of hospital admissions on those days.
+For predictions made for the first three Mondays in December 2021 and the last three Mondays in February 2022 all models had allocation scores under 500 (and the mean across all models was less than 100), indicating that the unnecessary unmet need was fairly low on those days.
 The allocation scores were on the whole highest when the observed number of new hospital admissions was closest to the resource threshold of 15,000, as those are the times when any mistakes in allocation are costly in terms of wasting resources in one location that could have been used in another.
@@
 The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the constraint ($K$), possibly taking weights into account for different values of $K$ that might be more or less likely (Section \ref{sec:methods.detailed.integrated_allocation}).
-This could be useful in situations where the actual constraint may not be known, or could only be specified as a probability distribution.
-Unlike in the above analyses where we conducted analyses assuming that $K$ was known to be 15,000, the analyses in this subsection are conducted assuming different distributions of $K$.
+This could be useful in situations where the actual constraint may not be known precisely, or where we wish to consider results across a range of decision making settings with different constraints.
+Unlike in the above analyses where we conducted analyses assuming that $K$ was known to be 15,000, the analyses in this subsection are conducted assuming different distributions on $K$.
 
@@
 
-AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on 2020-12-20 predicting levels of hospitalizations on 2021-01-03 (Figure \ref{fig:multi-k}A).
+AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 (Figure \ref{fig:multi-k}A).
 These calculations highlight that AS was highest at the $K$ nearest to the observed value of \Sexpr{format(tot_y, big.mark = ",")}.
@@
 
-IAS was computed for two distributions of $K$, one uniform across the entire range and the other a symmetric distribution around 15,000 (Figure \ref{fig:multi-k}B).
+IAS was computed for two distributions on $K$, one uniform across the entire range and the other a symmetric distribution around 15,000 (Figure \ref{fig:multi-k}B).
 Both versions of the IAS were correlated with the original AS conducted at $K=15,000$, with the higher correlation coming from the distribution that was centered at $K=15,000$ (Figure \ref{fig:multi-k}C).
@@
 Those that can be easily quantified may differ enough that it seems likely that no single score would be appropriate for all users.
-In an ideal setting, a forecasting tool could be developed through close collaboration between modelers and public health officials.
-However, this setting may only be possible in settings with enough resources to support both an analytics and public heatlh team.
+Ideally, a forecasting tool could be developed through close collaboration between modelers and public health officials.
+However, this may only be possible in settings with both an analytics and a public health team.
 Increasingly, collaborative modeling hubs are being used to generate ``one-size-fits-all'' forecasts for many locations at once.
-In these settings, where tailored models are not available, it still could be possible to evaluate contributed models against a set of multiple scores.
+In these settings, where tailored models are not available, it still could be possible to evaluate contributed models using a set of multiple scores to support public health end users in understanding the value of forecasts as an input to their particular decision making contexts.
 % NGR: leaving this out for now as it feels like too much detail?
@@
 The allocation score was not as scale-dependent, as the highest/worst allocation scores were observed when the number of hospitalizations was closest to the allocation constraint.
-We also note that contributed models found it hard to have a better allocation score than a na\"ive baseline model that just predicted a flat line from the last previous observation with wide uncertainty.
+We also note that contributed models found it hard to achieve a better allocation score than a na\"ive baseline model that just predicted a flat line from the last previous observation with wide uncertainty.
 This suggests suggests that contributed models (other than the ensemble, which combined forecasts from all contributed models) were not consistently adding value over just extrapolating from the current levels.
@@
 We also note that in some settings, a ``successful'' epidemiological forecast may lead to policy decisions that change the distribution of the predicted outcome $Y$.
-Our framework cannot be used to evaluate a forecast at a horizon that may have been impacted by such a decision.
-
-An additional limitation of the current work and opportunity for further investigation is to more carefully evaluate whether a forecast adds value to existing decision-making processes.
-In the context of allocation-based decisions, standard procedures might involve simply extrapolating need based on a current observed data (similar to the `baseline' model presented above), with or without adjustments based on other political or real-world considerations.
+Our framework cannot be used to evaluate a forecast at a horizon where the outcome may have been impacted by such a decision.
+
+An additional limitation of the current work and opportunity for further investigation is to more carefully evaluate whether a forecast adds value to existing decision making processes.
+In the context of allocation-based decisions, standard procedures might involve extrapolating need based on a current observed data (similar to the `baseline' model presented above), with or without adjustments based on other political or real-world considerations.
 For example, in many settings public health stakeholders will synthesize multiple sources of information, e.g.
 real-time data from a wide variety of quantitative and qualitative sources coupled with expert judgment based on a situational understanding of past and current experience.
-The allocation score presented in this work assesses the optimality of the allocation based on an omniscient view of the future.
+The allocation score presented in this work assesses the optimality of the allocation with reference an omniscient view of the future captured by the oracle.
 However, what we care about is whether a given forecast adds useful information to an existing decision-making process.
@@
 
-We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work in more closely coupling applied epidemiological forecasting with public health decision-making.
+We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling applied epidemiological forecasting with public health decision making.
 We note that a few papers have begun exploring similar linkages as described in the literature review \textendash but we see much room for additional work in this area.
-In some situations, individual or ensemble models could be developed to optimize scores that are attunend to a particular decision making setting.
+In some situations, individual or ensemble models could be developed to optimize scores that are attuned to a particular decision making setting.
 This area is also largely unexplored in the realm of public health to date, with some initial methodological development in econometrics \citep{loaiza-maya_focused_2021}.
@@
 In conclusion, we argue that the way modelers and policymakers view and evaluate forecasts should change depending on the specific decision-making context.
-Using standard forecast evaluation metrics can mask the utility of certain forecasts, or lead to forecasts being used to inform a particular decision that are not optimal for that particular set of values that are important for that context.
+Using standard forecast evaluation metrics can mask the utility of certain forecasts, or lead to forecasts being used to inform a particular decision that are not well matched to that decision making context.
 New collaborative work between public health officials and modeling teams is needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or simulation exercises that could be used to inform further development of new or alternative scoring metrics.

### 2023-12-13 alloscore_manuscript/alloscore-application.Rnw
a few cosmetic changes

@@
 \usepackage{hyperref}
+\usepackage{xurl}
 \externaldocument{supplement}
@@
 \DeclareMathOperator*{\argmin}{argmin}
+\DeclareMathOperator{\short}{sh}
+\DeclareMathOperator{\Ex}{\mathbb{E}}
 
@@
 However, what we care about is whether a given forecast adds useful information to an existing decision-making process.
-While the scoring procedures as presented do not directly address this question, they could be modified (say, by comparison to a `baseline` model) to quantify the benefit -if any- of using a forecast to inform a specific decision.
+While the scoring procedures as presented do not directly address this question, they could be modified (say, by comparison to a `baseline' model) to quantify the benefit---if any---of using a forecast to inform a specific decision.
 
 We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling applied epidemiological forecasting with public health decision making.
-We note that a few papers have begun exploring similar linkages as described in the literature review \textendash but we see much room for additional work in this area.
+We note that a few papers have begun exploring similar linkages as described in the literature review{\textendash}but we see much room for additional work in this area.
 In some situations, individual or ensemble models could be developed to optimize scores that are attuned to a particular decision making setting.

### 2023-12-13 alloscore_manuscript/supplement.rnw
a few cosmetic changes

@@
 \caption{Shortage functions.
-$\short^{y_0}(x)$, shown in blue, gives the resource shortage as a function of the level of available resources, $x$, for a fixed value of resource demand $y_0$.
-$\short_{x_0}(y)$, shown in red, gives the resource shortage as a function of resource demand, $y$, for a fixed value of resource supply $x_0$.}
+$\mathrm{sh}^{y_0}(x)$, shown in blue, gives the resource shortage as a function of the level of available resources, $x$, for a fixed value of resource demand $y_0$.
+$\mathrm{sh}_{x_0}(y)$, shown in red, gives the resource shortage as a function of resource demand, $y$, for a fixed value of resource supply $x_0$.}
 \label{fig:shortages}
@@
 In section 3 of the manuscript, we used the allocation score to evaluate forecasts of COVID-19 hospitalizations that have been submitted to the COVID-19 Forecast Hub.
-These forecasts are submitted to the Hub using a set of 23 quantiles of the forecast distribution at the 23 probability levels in the set $\mathcal{T} = \{0.01, 0.025, 0.05, 0.1, 0.15, \ldots, 0.9, 0.95, 0.975, 0.99\}$, which specify a predictive median and the endpoints of central $(1 - \alpha) \times 100\%$ prediction intervals at levels $\alpha = 0.02, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9$.
+These forecasts are submitted to the Hub using a set of 23 quantiles of the forecast distribution at the 23 probability levels in the set $\mathcal{T} = \{0.01, 0.025, 0.05, 0.1, 0.15, \ldots, \allowbreak 0.9, 0.95, 0.975, 0.99\}$, which specify a predictive median and the endpoints of central $(1 - \alpha) \times 100\%$ prediction intervals at levels $\alpha = 0.02, 0.05, 0.1, 0.2, \allowbreak 0.3, 0.4, 0.5, 0.6, \allowbreak 0.7, 0.8, 0.9$.
 For a given week and target date, we use $q_{i,k}$ to denote the submitted quantiles for location $i$ and probability level $\tau_k \in \mathcal{T}$, $k = 1, \ldots, 23$.

### 2023-12-14 alloscore_manuscript/alloscore-application.Rnw
fixes re issues #68, #80, #81, cite for regret, and a few sec:numerics edits

@@
 To make this a more easily interpreted measure of forecast performance, we will adjust the raw score by subtracting the minimum loss $l_{\text{oracle}}(y;K)$ achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$.
-This minimum loss is given by $l_{\text{oracle}}(y;K) = L \cdot \max(0, \sum_{i=1}^{N}y_i - K)$: if there are enough resources to meet the total resource need ($\sum_{i=1}^{N}y_i \leq K$), the oracle can achieve a loss of 0, but if the resource constraint is less than the total needed resources ($\sum_{i=1}^{N}y_i > K$), it is not possible to do better than a loss of $L \cdot (\sum_{i=1}^{N}y_i - K)$.
-
-Our scoring rule is then given by
+When the oracle has sufficient resources to meet the total need, i.e., when $\sum_{i=1}^{N}y_i \leq K$, the oracle's loss
+$L \cdot \max(0, \sum_{i=1}^{N}y_i - K) = 0$ and there is no adjustment.  Thus the allocation score coincides with
+the raw score,
 \begin{align}
-S_A(F, y; K) &= S_A^{\text{raw}}(F, y; K) - l_{\text{oracle}}(y;K) \\
-& =
-\begin{cases}
-\sum_{i=1}^n L  \cdot  \max(0, y_i - x_i^{F,K}), &  \sum_{i=1}^{N}y_i \leq K \\
-L \cdot K - L \cdot \sum_{i=1}^{N} \min(x_i^{F,K}, y_i), &  \sum_{i=1}^{N}y_i > K.
-\end{cases}
+S_A(F, y; K) = S_A^{\text{raw}}(F, y; K)=\sum_{i=1}^N L \cdot \max(0, y_i - x_i^{F,K}) \quad \text{if } \sum_{i=1}^{N}y_i \leq K.
 \end{align}
-In the first case, the oracle incurs no loss so that the raw and adjusted scores coincide.
-The second case can be read as taking the $K$ resource units perfectly allocated by the oracle as a base penalty on
+On the other hand, when the oracle cannot cover all need and incurs a loss of $L \cdot \sum_{i=1}^{N}y_i - L\cdot K > 0$, we adjust
+the raw score by this loss to get
+\begin{align}
+S_A(F, y; K)  &= S_A^{\text{raw}}(F, y; K) - L \cdot \sum_{i=1}^{N}y_i + L\cdot K  \nonumber\\
+&= L \cdot \sum_{i=1}^{N} \left\{\max(0, y_i - x_i^{F,K}) - y_i\right\}  + L\cdot K \nonumber \\
+&= L \cdot K - L \cdot \sum_{i=1}^{N} \min(x_i^{F,K}, y_i) \quad \text{if } \sum_{i=1}^{N}y_i > K.
+\end{align}
+This can be read as taking the $K$ resource units perfectly allocated by the oracle as a base penalty on
 the imperfect forecast $F$ and then, for each location $i$, reducing this penalty by however much of the need $y_i$
 is met with the Bayes act component $x_i^{F,K}$.
-The oracle adjustment aligns with the idea that \emph{opportunity loss} (often known as \emph{regret} or (negative) \emph{relative utility}) is often a more important quantity than absolute loss \elr{needs citation}.
-
-\subsubsection{Integrating the allocation score across resource constraint levels}
+The oracle adjustment aligns with a common theme in economic decision theory that \emph{opportunity loss} (often known as \emph{regret} or (negative) \emph{relative utility}) is often a more important quantity than absolute loss (see e.g., \cite{DIECIDUE201788}).
+
+\subsubsection{Integrating the allocation score across resource constraint levels}{}
 \label{sec:methods.detailed.integrated_allocation}
@@
 The WIS and CRPS arise when the decision relates to how much of a resource to order in the context of the cost of the resource and wanting to minimize unmet need in one or more locations.
-In both settings, the optimal solution sets the resource level in each location to a quantile of the respective forecast distributions.
+In both settings, the score can be defined in terms of the shortage or excess of resource procurement levels against realized need where
+these levels are expressable as quantiles of the respective forecast distributions by virtue of minimizing expected shortages or excesses.
 
@@
 In fact, it is possible to combine these into a more general problem where the decision maker must decide on both the total level of resources to purchase and an allocation of those resources across multiple locations, subject to a cost per unit of resources and a constraint on the total quantity of resources that can be purchased.
-The quantile loss and the allocation score presented in this work are both special cases of the score that emerges from that more general decision making problem that integrates across resource cost and constraint parameters together.
+When
+constraints are likely to be large, and therefore of less importance, a composite score for this general scenario would converge to the WIS or CRPS.
+And if the constraints are likely to be tight, and therefore to be a determining factor, this composite score would converge to the IAS.
 We pursue this direction further in other work that is in progress.

### 2023-12-14 alloscore_manuscript/supplement.rnw
fixes re issues #68, #80, #81, cite for regret, and a few sec:numerics edits

@@
 	%item Sketch of scoring for decision problems involving both cost and constraint.
-	\item  Derivation of case 2 in formula for Oracle adjustment
+	\item[\xmark]  Derivation of case 2 in formula for Oracle adjustment (done in manuscript?)
 \end{todolist}
@@
 \] 
-The decision problem is then to select a random future loss $l(x,Y)$ in a way that aligns with the preference that $l(x,y)$ be
-as low as possible given any realization $Y=y$.
+The decision problem is then to select an action $x$ that aligns with the preference that $l(x,y)$ be
+as low as possible given any realization $Y=y$.  Equivalently, the problem is to choose a random future loss $l(x,Y)$ which
+is \emph{admissible}, meaning that there is no alternative action $\tilde{x}$ producing a random future loss $l(\tilde{x},Y)$
+which is never greater than $l(x,Y)$ and strictly lower for some realization $Y=y_0$.
 
+Admissibility with respect to the loss $l$ is not, however, by itself a very strong decision criterion.
 To give the decision problem more structure we assume the decision maker either knows the distribution $F$ of $Y$, or
@@
 
-\section{Numerical computation of allocation scores}
+\section{Numerical computation of allocation Bayes acts}
 \label{sec:numeric}
 
-Exact computation of the allocation score $S_A(F, y; K)$ of a forecast $F$ requires an explicit choice of and expression for a 
-solution $x^{F,K} = \bm{q}_{\tau^{\star},F} \in \undertilde{\bm{F}}^{-1}(\tau^{\star})$ to \eqref{eqn:quantiles-sum-to-K} where
-$\tau^{\star}$ solves \eqref{eqn:tau-inclusion}. For practical scenarios in which the forecast $F$ has a form preventing 
-analytical solutions to \eqref{eqn:quantiles-sum-to-K}, we seek a method of generating an approximation $\tilde{\tau}$ to 
-a solution $\tau^{\star}$ which offers control (at least in principle) over a robust measure of scoring rule error such as
-the sum of absolute coordinate errors
+In practice, \eqref{eqn:quantiles-sum-to-K} is rarely amenable to closed form solution.
+Using the Bayes act in explicit computations therefore requires a method of generating an approximation $\tilde{\tau}$ to 
+a solution $\tau^{\star}$ of \eqref{eqn:tau-inclusion}.  One would also strive to have some control over a robust measure of 
+of the approximation error of the random shortages which we aim to compute using the Bayes act. One possible measure is the sum
+of absolute coordinate errors
 \begin{align}
@@
 where $\bm{q}_{\tilde{\tau},F}$ solves the linear program
-$\sum_{x \in \undertilde{\bm{F}}^{-1}(\tilde{\tau})}x_i = K$.
+$\sum_{x \in \undertilde{\bm{F}}^{-1}(\tilde{\tau})}x_i = K$ approximating the true Bayes act defining linear program
+$\sum_{x \in \undertilde{\bm{F}}^{-1}(\tau^{\star})}x_i = K$. In this section we outline such a method we have implemented 
+in the \verb`R` package \verb`alloscore`.
 

### 2023-12-19 alloscore_manuscript/alloscore-application-shorter.Rnw
misc minor figure updates

--- 2023-12-13 alloscore_manuscript/alloscore-application.Rnw
@@
 
-The COVID-19 pandemic has led to rapid innovation in methods for eliciting and evaluating forecasts of infectious disease burdens, with a primary goal being to help public health workers make informed decisions about how to manage these burdens.
-However, explicit descriptions or quantifications of the value forecasts add to society through the decisions they support are elusive.
-Moreover, there has only been limited discussion of how predominant forecast evaluation metrics might indicate the success of policies based in part on those forecasts.
-
-Here we pursue one possible tether between multivariate forecasts and policy: the allocation of limited medical resources in response to COVID-19 hospitalizations in various regions so as to minimize expected unmet need.
-Given probabilistic forecasts of hospitalizations in each region, we formulate an allocation algorithm following techniques developed in operations research.
-We then score forecasts according to how much unmet need their associated allocations would have allowed.
-We illustrate this scheme with quantile forecasts of COVID-19 hospitalizations in the US at the state level that are recorded in the COVID-19 Forecast Hub, with the goal of determining the allocation of a hypothetical limited resource across the states.
-The forecast skill ranking given by this allocation scoring rule can vary substantially from the ranking given by the weighted interval score that is often used to score quantile forecasts, especially during surges in hospitalizations such as in late 2021 as the Omicron wave began.
-We see this as strong evidence that the allocation scoring rule detects forecast value that is missed by traditional accuracy measures and that the general strategy of designing scoring rules that are directly linked to policy performance is a promising research direction for epidemic forecast evaluation.
+Recent years have seen increasing efforts to forecast infectious disease burdens, with a primary goal being to help public health workers make informed policy decisions.
+However, there has only been limited discussion of how predominant forecast evaluation metrics might indicate the success of policies based in part on those forecasts.
+We explore one possible tether between forecasts and policy: the allocation of limited medical resources so as to minimize unmet need.
+We use probabilistic forecasts of disease burden in each of several regions to determine optimal resource allocations, and then we score forecasts according to how much unmet need their associated allocations would have allowed.
+We illustrate with forecasts of COVID-19 hospitalizations in the US, and we find that the forecast skill ranking given by this allocation scoring rule can vary substantially from the ranking given by the weighted interval score.
+We see this as evidence that the allocation scoring rule detects forecast value that is missed by traditional accuracy measures and that the general strategy of designing scoring rules that are directly linked to policy performance is a promising direction for epidemic forecast evaluation.
 
@@
 Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping (University of Texas, 2022)\nocite{utnews2022}.
-In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
-The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid} or weighted interval score (WIS) \citep{bracher2021evaluating}.
+%In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
+%The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid} or weighted interval score (WIS) \citep{bracher2021evaluating}.
 However, recently authors have been calling for evaluating forecasts through their impact on policy \citep{marshall2023predictions, bilinski_adaptive_2023}.
@@
 We consider a setting in which forecasts are used to help determine the allocation of a limited quantity of medical supplies across multiple regions.
-We define a new forecast scoring rule --- the {\em allocation score} --- that evaluates forecasts based on how beneficial resource allocations derived from them would turn out to be.
+In section~\ref{sec:methods} of the paper, we define a new forecast scoring rule --- the {\em allocation score} --- that evaluates forecasts based on how beneficial resource allocations derived from them would turn out to be.
 %Unlike the traditional measures of forecast skill such as WIS, CRPS, and log score, the proposed allocation score measures the suitability of a forecast as an input to this specific decision making task.
 
-Briefly, the allocation score of a forecast is the avoidable unmet need that results from using that forecast to set resource allocations by minimizing expected unmet need.
-For example, suppose that a decision maker is provided with forecasts of the level of need for medical resources in each of several states or hospital systems.
-If there is a limited amount of the medical resource that is available to distribute, a decision maker could choose an allocation of that resource across locations that minimizes the expected unmet need according to that forecast.
-As measured by the allocation score, one forecast is better than another if it would lead decision makers to an allocation that results in less unmet need.
+%Briefly, the allocation score of a forecast is the avoidable unmet need that results from using that forecast to set resource allocations by minimizing expected unmet need.
+%For example, suppose that a decision maker is provided with forecasts of the level of need for medical resources in each of several states or hospital systems.
+%If there is a limited amount of the medical resource that is available to distribute, a decision maker could choose an allocation of that resource across locations that minimizes the expected unmet need according to that forecast.
+%As measured by the allocation score, one forecast is better than another if it would lead decision makers to an allocation that results in less unmet need.
 %By ``unnecessary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
-If the amount of resources that is available to distribute is less than the actual need, some amount of unmet need is unavoidable.
-The allocation score for a forecast does not include the unmet need that was unavoidable given the resource constraint, and so it measures only the amount of unmet need that could have been prevented by using a different allocation of available resources than that suggested by the forecast.
-We elaborate on these ideas in Section~\ref{sec:methods}.
+%If the amount of resources that is available to distribute is less than the actual need, some amount of unmet need is unavoidable.
+%The allocation score for a forecast does not include the unmet need that was unavoidable given the resource constraint, and so it measures only the amount of unmet need that could have been prevented by using a different allocation of available resources than that suggested by the forecast.
+%We elaborate on these ideas in Section~\ref{sec:methods}.
 
-We present an illustrative analysis using the allocation score to evaluate forecasts of hospital admissions in the US leading up to the Omicron wave in winter 2022.
-This analysis is ``synthetic'' in that it does not correspond to an actual analysis that supported decision making in real-time.
+In section \ref{sec:application}, we present an illustrative analysis using the allocation score to evaluate forecasts of hospital admissions in the US leading up to the Omicron wave in winter 2022.
+This analysis is ``synthetic'' in that it does not correspond to an actual analysis that supported decision making in real time.
 However, the framework described in this paper corresponds to real-world decisions that must be made by public health administrators around the globe, and could be adapted in the future for such real-time situations.
@@
 This framework was used by the Hartford HealthCare system in Connecticut ``to align ventilator supply with projected demand at a time where the [COVID-19] pandemic was on the rise'' \citep{bertsimas2021predictionsCOVID}.
-These examples illustrate the potential for forecasts to inform decisions about how to allocate limited supplies such as temporary hospital beds, ventilators, personal protective equipment, or other supplies that are known to be effective at reducing transmission or severity of disease.
+%These examples illustrate the potential for forecasts to inform decisions about how to allocate limited supplies such as temporary hospital beds, ventilators, personal protective equipment, or other supplies that are known to be effective at reducing transmission or severity of disease.
 However, we emphasize again that these studies did not take the step of evaluating forecasts based on the quality of the allocation decisions that they supported or could have been used to support.
 
-The remainder of this article is organized as follows.
-We describe the allocation score in Section \ref{sec:methods}, and in Section \ref{sec:application} we illustrate the use of the score in an application to evaluate short-term forecasts of COVID-19 hospital admissions in the US.
-Section \ref{sec:discussion} summarizes our contributions and discusses opportunities for further extensions in future work.
+%The remainder of this article is organized as follows.
+%We describe the allocation score in Section \ref{sec:methods}, and in Section \ref{sec:application} we illustrate the use of the score in an application to evaluate short-term forecasts of COVID-19 hospital admissions in the US.
+%Section \ref{sec:discussion} summarizes our contributions and discusses opportunities for further extensions in future work.
 
@@
 In section \ref{sec:methods.detailed} we develop the allocation score more carefully, building on decision theoretic procedures for deriving proper scoring rules.
-We then comment on some connections between the allocation score that we propose and other common scores that can be derived from decision theoretic foundations, such as the quantile score, WIS, and CRPS, in section \ref{sec:methods.related}.
+In section \ref{sec:methods.related}, we note that another group of common scores including the quantile score, WIS, and CRPS, can also be derived from decision theoretic foundations \textemdash starting from a different decision making context.
 
@@
 \end{align}
-We emphasize that in Equation \eqref{eqn:informal_loss} the calculation of unmet need is based on the actual resource need that was realized in each location, while in Equation \eqref{eqn:informal_objective} the calculation of unmet need was based on the forecast distribution of future levels of resource need.
-Once the actual levels of resource need have been observed, we can obtain a quantitative measure of the quality of alternative allocation decisions: one allocation is better than another if it results in lower total unmet need.
+%We emphasize that in Equation \eqref{eqn:informal_loss} the calculation of unmet need is based on the actual resource need that was realized in each location, while in Equation \eqref{eqn:informal_objective} the calculation of unmet need was based on the forecast distribution of future levels of resource need.
+One allocation is better than another if it results in lower total unmet need, and one forecast is better than another if the allocation derived from it results in lower total unmet need.
 
@@
 \paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$ and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$.
+The means of these distributions are given by the scale parameters $\sigma_i$.
 When the marginal forecasts are exponential distributions, it can be shown that the optimal allocation divides the available resources among the locations proportionally to the scale parameters $\sigma_i$ (see section
 \ref{sec:bayes-quantiles} of the supplemental materials).
-If we have $K = 5$ units of our resource available, the optimal allocation according to $F$ would be 1 unit of resources in location 1 and 4 units of resources in location 2.
-If, on the other hand, we have $K = 10$ units available, we will allocate 2 units of resources to location 1 and 8 units to location 2.
+If $K = 5$ units of the resource are available, the optimal allocation according to $F$ would be 1 unit of resources in location 1 and 4 units of resources in location 2.
+If, on the other hand, $K = 10$ units are available, we will allocate 2 units of resources to location 1 and 8 units to location 2.
 Figure~\ref{fig:exp_alloc_example} illustrates the situation.
@@
 Therefore, the forecast's allocation score is 0 with $K = 5$.
-On the other hand, when $K = 10$, the forecast $F$'s allocation results in $10 - 8 = 2$ units of unmet need in location 2 despite leaving no need unmet in location 1.
+On the other hand, when $K = 10$, the forecast's allocation results in $10 - 8 = 2$ units of unmet need in location 2 despite leaving no need unmet in location 1.
 In this case, the oracle would be able to prevent all but 1 of the total 11 units of need from going unmet, for example by allocating 1 unit of resources to location 1 and the remaining 9 units of resources to location 2.
@@
 \paragraph{Example 2} Now consider a different forecast that also has exponential distributions for resource need in each location, but that has the scale parameters $\sigma_1 = 2$ and $\sigma_2 = 8$, twice as large as the scale parameters of the forecast in Example 1.
-Because the optimal allocation is proportional to the scale parameters, this forecast would lead to the same optimal allocations as the forecast in Example 1, and would therefore be assigned the same allocation score.
+Because the optimal allocation is proportional to the scale parameters, this forecast would lead to the same allocations as the forecast in Example 1, and would therefore be assigned the same allocation score.
 
-Note the way in which these forecasts incurred a positive (i.e., non-optimal) allocation score of 1 when $K = 10$.
-It was not directly due to individual misalignments of the marginal forecasts $F_i$ with the observed needs, but rather because the allocations and observed needs were not proportional as vectors.
-Restating: as far as allocation decisions are concerned, with a fixed constraint $K=10$ the fundamental problem with the forecast $F$ in Example 1 is not that it predicts a mean total resource need of $5$ units; it is that the realized need was 10 times as large in location 2 as in location 1, but the forecast only indicated that the resource allocation for location 2 should be 4 times the allocation for location 1.
+Examination of these results leads to two observations.
+First, the reason that these forecasts had a positive (i.e., non-optimal) allocation score at $K=10$ is that they did not get the relative magnitude of resource need across the two locations right: the realized need was 10 times as large in location 2 as in location 1, but the forecasts only indicated that the resource allocation for location 2 should be 4 times the allocation for location 1.
+At its core, the allocation score measures whether the forecast accurately captures the relative magnitudes of resource need across different locations, which is precisely the information that is needed to allocate resources to those locations subject to a fixed resource constraint.
 
-This illustrates a fundamental property of the allocation score: at its core, it measures whether the forecast accurately captures the relative magnitudes of resource need across different locations, which is precisely the information that is needed to allocate resources to those locations subject to a fixed resource constraint.
-On the other hand, the allocation score is not directly sensitive to whether the forecasts in each location correctly capture the magnitude of resource need in each individual location.
-%For a given probability level $\tau$, if the quantiles $Q_i(\tau)$ of the marginal forecasts are proportional to the observed needs $y_i$, then the allocation score is zero for the resource constraint level $K = \sum Q_i(\tau)$.
-This stands in marked contrast to other common scoring methods for multivariate forecasts that aggregate univariate scores such as log score, CRPS, or WIS for the marginal forecasts where a
-poor forecast made for one unit (a location, say) is penalized regardless of alignments in other units.
-Note that we do not claim that the allocation score is generically preferable to these other scores \textemdash rather, it provides a view of forecast performance that is specifically tuned to the context of decision making about resource allocations.
-
-%It is therefore often straighforward to constuct forecasts $F$ and $\widetilde{F}$ for a given outcome distribution that switch rankings under the allocation and traditional scores by ensuring that the marginal forecasts of $F$ center sharply around allocations that are proportionally similar to the central tendencies of the outcome distribution but are strongly biased.
-
-%While multivariate scoring rules have not seen wide application in infectious disease forecast evaluation, it seems that the allocation score would have a similar relationship with them, since while bias penalties in multi-variate scores can be offset by better forecasting of dependence structure, the proportional biases that allocation scoring are insensitive to are not \emph{per se} tolerated by multivariate scoring rules such as the energy score, variogram score, or Dawid-Sebastiani score.
+A second observation is that the forecasts in examples 1 and 2 predicted different mean levels of resource needs, but had the same allocation score.
+The allocation score does not directly measure whether the forecasts correctly capture the absolute magnitude of resource need in each individual location.
+This stands in contrast to other common scoring methods that aggregate scores such as log score, CRPS, or WIS for each location, where a poor forecast for one location is penalized regardless of alignments in other units.
+%Note that we do not claim that the allocation score is generically preferable to these other scores \textemdash rather, it provides a view of forecast performance that is specifically tuned to the context of decision making about resource allocations.
 
@@
 \end{enumerate}
-%We use the letter $s$ for the loss function to align with the literature on evaluation of forecasts of continuous outcomes, in which context we can often identify the action $x$ with a functional (i.e., a numeric summary such as a mean or a quantile) of the forecast distribution $F$.
-In this context, $s$ may be used as a \emph{scoring function}.
-\elr{Consider moving preceding sentences to a footnote or just deleting them?}
 This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss function.
 Subject to certain technical conditions, scoring rules obtained from this procedure are proper.
-We refer the reader to (cite paper 2 on arxiv) for a more technically precise discussion.
+We refer the reader to the supplement for a more technically precise discussion.
 
@@
 Our setup assumes that $t_d < t_r$.
-Additionally, the structure of our loss captures a setting where the resource in question does not impact the amount of demand $y_i$ that will materialize at time $t_r$, but rather it is a resource that satisfies that demand.
+Additionally, the structure of our loss captures a setting where the resource in question does not impact the amount of demand that will materialize at time $t_r$, but rather it is a resource that satisfies that demand.
 In the context of infectious disease, this means that we do not consider resources that are intended to reduce the number of people who will become sick at some point in the future, such as a preventative influenza or COVID-19 vaccine.
-Instead, our set up addresses resources like hospital beds, oxygen supply, ventilators, or rabies vaccines which are intended to meet the medical needs of patients who are already sick.
-We also note that our problem formulation addresses decision-making that is related to resource needs only at the time $t_r$; we do not explicitly consider sequences of multiple decisions that are made over time or account for the impact of decisions on resource needs at any time other than $t_r$.
+Instead, our setup addresses resources like hospital beds, oxygen supply, or ventilators which are intended to meet the medical needs of patients who are already sick.
+Additionally, our setup addresses decision-making that is related to resource needs only at the time $t_r$; we do not explicitly consider sequences of multiple decisions that are made over time or account for the impact of decisions on resource needs at any time other than $t_r$.
 We outline some opportunities to extend our work to more complex decision making settings in the discussion.
@@
 
-In the supplement we derive a general form of the Bayes act by beginning with the fact that a condition for minimizing expected loss subject to the resource constraint is that there is no way to decrease expected loss further by shifting a small amount of the resource from one location to another.
-From this starting point, we can show that the components of the Bayes act are quantiles $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource constraint $K$, but is shared across all locations.
-% Compare % In the supplement we derive a general form of the Bayes act by beginning with the fact that a condition for minimizing expected loss under the constraint is that there is no way to decrease expected loss further by shifting a small amount of the resource from one location to another.
-% From this, we can show that the components of the Bayes act are quantiles $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource constraint $K$, but is shared across all locations.
-% Compare with earlier text: This will lead to the allocations provided by $F$ being quantiles of the marginal distributions $F_i$ for some \emph{single} probability level $\tau$ that is shared in common for all locations.
+In the supplement we show that the components of the Bayes act are quantiles $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource constraint $K$, but is shared across all locations.
 This probability level is the level at which the resource constraint is satisfied: $\sum_{i=1}^N F_i^{-1}(\tau^{F,K}) = K$.
@@
 This solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
-
-%%%% start of large excision %%%%%
-
-% In the supplement we derive a general form of this solution by beginning with the fact that a condition for minimizing
-% expected loss under the constraint is that there is no way to decrease expected loss further by shifting a small amount of the resource from one location to another.
-% As a consequence, at $x^{F,K}$, adding resources to any particular location $i$ reduces expected loss at the same rate $\lambda$:
-% \begin{align}
-% \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^{F,K}} = -\lambda.
-% \end{align}
-% It turns out that these derivatives can also be expressed in terms of the marginal forecasts $F_i$ as
-% \begin{align}
-%  \frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] = L \cdot (F_i(x_i) - 1)
-% \end{align}
-% so that
-% \begin{align}
-% F(x_i^{F,K}) = 1-\lambda/L.
-% \end{align}
-% This tells us that in order to allocate optimally (according to $F$), we must divide resources among the locations so that there is an equal forecasted probability in every location that the allocation is sufficient to meet resource need.
-% It also allows us to write the components $x_i^{F,K}$ of the Bayes act as quantiles
-% \begin{align}
-% x_i^{F,K} = F_i^{-1}(1 - \lambda/L)
-% \end{align}
-% which can be substituted into the constraint equation of the allocation problem to get an equation for $\lambda$,
-% \begin{align}
-% \sum_{i=1}^N F_i^{-1}(1 - \lambda/L) = K.
-% \end{align}
-% The value of $\lambda$ that solves this equation---which almost always must be approximated with a numerical method---can be substituted back into the quantile formulas
-% to arrive at an explicit expression for the Bayes act.
-% This partial solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
-% We note that $\lambda$ and the probability level $1-\lambda/L$ depend on both the constraint level $K$ defining the decision problem and on the forecast $F$.
-This is quite different from the situation for quantile scores and the WIS, where the probability levels used to calculate scores are fixed.
-
-%%%% end of large excision %%%%%
 
@@
 To make this a more easily interpreted measure of forecast performance, we will adjust the raw score by subtracting the minimum loss $l_{\text{oracle}}(y;K)$ achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$.
-This minimum loss is given by $l_{\text{oracle}}(y;K) = L \cdot \max(0, \sum_{i=1}^{N}y_i - K)$: if there are enough resources to meet the total resource need ($\sum_{i=1}^{N}y_i \leq K$), the oracle can achieve a loss of 0, but if the resource constraint is less than the total needed resources ($\sum_{i=1}^{N}y_i > K$), it is not possible to do better than a loss of $L \cdot (\sum_{i=1}^{N}y_i - K)$.
-
-Our scoring rule is then given by
+When the oracle has sufficient resources to meet the total need, i.e., when $\sum_{i=1}^{N}y_i \leq K$, the oracle's loss is zero and there is no adjustment.  Thus the allocation score coincides with
+the raw score in that case.
+On the other hand, when the oracle cannot cover all need and incurs a loss of $L \cdot \left(\sum_{i=1}^{N}y_i - K \right) > 0$, we adjust
+the raw score by this loss.
+The oracle-adjusted score is
 \begin{align}
 S_A(F, y; K)  &= S_A^{\text{raw}}(F, y; K) - l_{\text{oracle}}(y;K) \\
-& =
-\begin{cases}
-\sum_{i=1}^n L  \cdot  \max(0, y_i - x_i^{F,K}), &  \sum_{i=1}^{N}y_i \leq K \\
-L \cdot K - L \cdot \sum_{i=1}^{N} \min(x_i^{F,K}, y_i), &  \sum_{i=1}^{N}y_i > K.
+&= \begin{cases}
+S_A^{\text{raw}}(F, y; K), &  \sum_{i=1}^{N}y_i \leq K \\
+S_A^{\text{raw}}(F, y; K) - L \cdot \left( \sum_{i=1}^{N}y_i - K\right), &  \sum_{i=1}^{N}y_i > K.
 \end{cases}
 \end{align}
-In the first case, the oracle incurs no loss so that the raw and adjusted scores coincide.
-The second case can be read as taking the $K$ resource units perfectly allocated by the oracle as a base penalty on
-the imperfect forecast $F$ and then, for each location $i$, reducing this penalty by however much of the need $y_i$
-is met with the Bayes act component $x_i^{F,K}$.
-The oracle adjustment aligns with the idea that \emph{opportunity loss} (often known as \emph{regret} or (negative) \emph{relative utility}) is often a more important quantity than absolute loss \elr{needs citation}.
+The oracle adjustment aligns with a common theme in economic decision theory that \emph{opportunity loss} (often known as \emph{regret} or (negative) \emph{relative utility}) is often a more important quantity than absolute loss (see e.g., \cite{DIECIDUE201788}).
 
-\subsubsection{Integrating the allocation score across resource constraint levels}
+\subsubsection{Integrating the allocation score across resource constraint levels}{}
 \label{sec:methods.detailed.integrated_allocation}
@@
 
-\subsection{Generalizations and Connections to Other Scores}
+\subsection{Connections to Other Scores}
 \label{sec:methods.related}
-
-We begin this section by briefly sketching how the weighted interval score (WIS), a commonly used proper scoring rule for probabilistic forecasts during the COVID-19 pandemic, can be derived using the decision theoretic approach above.
-Then, we discuss similarities and differences between WIS and the allocation score, and other scores in general.
-
-\subsubsection{The quantile loss and weighted interval score (WIS)}
-\label{sec:methods.wis}
 
@@
 The WIS is a proper scoring rule for forecasts that use a set of quantiles to represent a probabilistic forecast distribution.
-%% optional detail
-Many early COVID-19 modeling efforts, including the US COVID-19 Forecast Hub and other collaborative forecasting projects, adopted a quantile forecast format as a matter of convenience (e.g., this format does not require the forecaster to pre-specify an upper bound for disease counts as would be required to collect forecasts in the format of bin probabilities) \citep{cramer_united_2022}.
-%%
 While pointing a reader interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS is a weighted sum of interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...).
@@
 The most commonly used version of WIS is one that uses an equal weighting of all quantile levels, in which case WIS approximates the continuous ranked probability score (CRPS), a commonly used score for probabilistic forecasts.
-\apg{Maybe a rewording and/or a supplement reference here to mention that any WIS approximates the a corresponding weighted CRPS, and that WIS actually is the wCRPS for a discrete measure/weighting.}
 \emph{It is important to note that this weighting was proposed because the resulting score approximates the CRPS, and not because it aligned with any particular public health decision-making rationale.}
@@
 This leads to the quantile score for the probability level $\tau = 1 - C/L$.
-From this point, the WIS or CRPS can be obtained by averaging across a range of decision making settings with different cost and loss parameters, using a similar motivation that we used to obtain the IAS from the AS in section \ref{sec:methods.detailed.integrated_allocation}.
-\apg{Refs to supp for quantile score derivation and detail about averaging.}
-%\ngr{I might suggest adding a detailed Step 1/2/3 as an appendix and referencing it here. But I think that adding that full detail here might be too much.}
-%\ngr{can we make more explicit the link here to the section above?
-I'm not confident in my ability to describe with technical accuracy the link between quantile/pinball loss and CRPS, but I think this is a good opportunity to do so briefly.}
-
-\subsubsection{Connections between scores}
-\label{sec:methods.generalization}
-
-We have described in previous sections how both the allocation score and the weighted interval score (WIS) arise from a single standard procedure for developing proper scoring rules.
-The allocation score arises when the decision relates to how a fixed quantity of resources should be allocated to multiple locations.
-The WIS and CRPS arise when the decision relates to how much of a resource to order in the context of the cost of the resource and wanting to minimize unmet need in one or more locations.
-In both settings, the optimal solution sets the resource level in each location to a quantile of the respective forecast distributions.
-
-These decision making problems differ according to the challenge faced by the decision maker: a fixed constraint on the available resources for the allocation problem, or a cost per unit of resources in the resource purchasing problem.
-In fact, it is possible to combine these into a more general problem where the decision maker must decide on both the total level of resources to purchase and an allocation of those resources across multiple locations, subject to a cost per unit of resources and a constraint on the total quantity of resources that can be purchased.
-The quantile loss and the allocation score presented in this work are both special cases of the score that emerges from that more general decision making problem that integrates across resource cost and constraint parameters together.
-We pursue this direction further in other work that is in progress.
+From this point, the WIS or CRPS can be obtained by averaging across a range of decision making settings with different cost and loss parameters, using a similar motivation that we used to obtain the IAS from the AS in section \ref{sec:methods.detailed.integrated_allocation} \citep{gneiting2011weightedScoringRules}.
 
@@
 
-We illustrate with an application to hospital admissions in the U.S., considering the problem of allocation of a limited supply of medical resources to states.
-
-\begin{figure}
-    \includegraphics[width=\textwidth]{../figures/allocation-forecasts-v2.pdf}
-    \caption{Hospitalization data and example forecasts for Virginia, US.
-New daily hospital admissions are shown by the line of data.
-The dark line indicates data up to the time the forecast was made on December 27, 2021; the grey line indicates data available in retrospect.
-The set of colored bars represent forecasts made for 14 days into the future, with a target date of January 10, 2022 (dashed vertical line).
-The horizontal dashed line indicates the level of the observed value on this day.
-Note that all three forecasts correspond to forecasts for the same day, but they are spread out for better visibility.
-}
-    \label{fig:sample-forecasts}
-\end{figure}
+We illustrate with an application to hospital admissions in the U.S., considering a hypothetical problem of allocating a limited supply of medical resources to states.
 
@@
 
-\subsubsection{Hospitalization data}
-Starting in the summer of 2020, the US Health and Human Services (HHS) began reporting counts of daily new admissions to hospitals for individuals with COVID-19 \citep{healthdatagov_covid-19_nodate}.
-These daily counts were available for the US as a whole, and all states and several additional jurisdictions such as Puerto Rico and Washington DC.
-The data were updated daily and were available for download by the public through the HHS HealthData.gov website.
-For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data \cite{wang-covidhubutils}.
-The hospitalization data were downloaded for analysis on \Sexpr{format(as.Date(metadata$time[which(metadata$name == "truth_data")]), "%B %d, %Y")}.
+%\subsubsection{Hospitalization data}
+%Starting in the summer of 2020, the US Health and Human Services (HHS) began reporting counts of daily new admissions to hospitals for individuals with COVID-19 \citep{healthdatagov_covid-19_nodate}.
+%These daily counts were available for the US as a whole, and all states and several additional jurisdictions such as Puerto Rico and Washington DC.
+%The data were updated daily and were available for download by the public through the HHS HealthData.gov website.
+%For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data \cite{wang-covidhubutils}.
 
-\subsubsection{Forecast data}
-The US COVID-19 Forecast Hub, a consortium funded by the US CDC and led by a research group at the University of Massachusetts-Amherst, collected short-term forecasts of new hospital admissions at the daily scale starting in December 2020, using the HHS data as a source of ``ground truth'' \citep{cramer_united_2022}.
-Any team that with appropriately formatted forecasts could submit them to the Forecast Hub data repository on GitHub (cite repo site).
-Forecasts were time-stamped by GitHub upon submission and passed validation checks that ensured correct formatting and that the forecasts were being submitted only for dates in the future, not for data that had already been observed.
-
-Forecast submission followed a weekly cycle and culminated in the creation of an ensemble forecast.
-Forecasts could be submitted on any day during the week.
-However once a week on Mondays, the Forecast Hub would collect the most recent forecasts submitted by all teams that met certain inclusion criteria and create an ensemble forecast using quantile averaging \citep{ray_comparing_2023}.
-An ensemble that treated all models equally was created (COVIDhub-ensemble) as was a model that created weights of submitted models based on performance in the past 12 weeks (COVIDhub-trained\_ensemble).
-One other model that combined multiple forecasts from different teams but used a different ensembling algorithm, a linear pooling method with tail extrapolation, was also included in our analyses (JHUAPL-SLPHospEns).
-Several other models have ``ensemble'' in their name, but this refers to combinations of different variations of models that the specific team created, not to a multi-model ensemble combining different submitted forecasts to the Forecast Hub.
-
-All forecasts, including the ensemble, were submitted as probabilistic predictions about the number of new hospital admissions on a particular day in the future, in a specific jurisdiction of the US (national level, state, or territory).
-Probability distributions were represented using a set of 23 quantiles for each individual prediction.
-The submitted quantiles included a median and the lower and upper limits of 11 central prediction intervals, from a 99\% to a 10\% prediction interval.
+%\subsubsection{Forecast data}
+The US COVID-19 Forecast Hub collected short-term forecasts of daily new hospital admissions for individuals with COVID-19 starting in December 2020 \citep{cramer_united_2022}.
+The target data for these forecasts were hospital admissions as reported by the US Department of Health and Human Services through the HealthData.gov website.
+Forecasts were probabilistic predictions of the number of new hospital admissions on a particular day in the future, in a specific jurisdiction of the US (national level, state, or territory).
+Probability distributions were represented using a set of 23 quantiles for each individual prediction, including a median and the lower and upper limits of 11 central prediction intervals, from a 99\% to a 10\% prediction interval.
 
@@
 
-We established a set of criteria to determine which forecasts and models to include in our analysis.
-Models were eligible to be included in the analysis if they were considered a ``primary'' model from a team.
-(If a team submitted multiple versions of similar models, they were required to designate one as ``primary''.)
+Submission to the Forecast Hub followed a weekly cycle, and each Monday the Hub collected the most recent forecasts submitted by all teams that met certain inclusion criteria and created ensemble forecasts using quantile averaging \citep{ray_comparing_2023}.
+Our analysis includes these ensembles (COVIDhub-ensemble and COVIDhub-trained\_ensemble) as well as one other ensemble of hub models created by another team (JHUAPL-SLPHospEns) and several other individual models.
+Models were eligible to be included in the analysis if they were designated as a ``primary'' model from a team. 
 For a model to have a complete, eligible submission in a given week, it had to have a 14 day-ahead forecast for all 50 states plus Washington DC.
-Models had to have a complete forecast for at least 4 of the 15 weeks in the analysis to be considered eligible for inclusion.
+Models had to have a complete forecast for at least 4 of the 15 weeks in the analysis to be eligible for inclusion.
+
+The hospitalization data used for scoring forecasts were downloaded on \Sexpr{format(as.Date(metadata$time[which(metadata$name == "truth_data")]), "%B %d, %Y")}.
 
@@
 
-This manuscript focuses on two proper forecast scores, the allocation score and the weighted interval score (WIS), both defined above.
+We measured forecast skill using two forecast scores, the allocation score and the weighted interval score (WIS), both defined above.
+We computed these scores for the 14 day ahead forecasts made each week.
+For this analysis, we fixed the resource constraint used for the allocation score to be $K=15,000$, based roughly on a reported number of ventilators available for reallocation in the US \citep{ajao_assessing_2015}.
+We computed the mean WIS (MWIS) across all of the forecasted state-level locations.
+%For each week, we computed the allocation score for the 14 day-ahead forecast.
 
-\subsubsection{Allocation Score}
+For both scores, we also computed models' standardized rank among all models that submitted forecasts each week.
+The standardized rank is between 0 and 1, where 0 corresponds to the worst rank and 1 to the best.
+In the case of a tie between one or more models, all models received the better rank.
 
-For this analysis, we fixed the resource constraint $K$ to be 15,000, based roughly on a reported number of ventilators available for reallocation in the US \citep{ajao_assessing_2015}.
-For each week, we computed the allocation score for the 14 day-ahead forecast.
+%We also computed a standardized rank for the allocation score for each model $m$ and week $w$.
+%First, we computed the number of models that forecasted that week ($n_w$) and the rank of model $m$ among the $n_w$ models ($r_{m,w}^{AS}$).
+%The model with the best allocation score received a rank of 1 and the worst received a rank of $n_w$.
+%In the case of a tie between one or more models, all models received the better rank.
+%We then rescaled these rankings to compute the allocation score standardized rank ($sr_{m,w}^AS$) between 0 and 1, where 0 corresponds to the worst rank and 1 to the best.
 
-We also computed a standardized rank for the allocation score for each model $m$ and week $w$.
-First, we computed the number of models that forecasted that week ($n_w$) and the rank of model $m$ among the $n_w$ models ($r_{m,w}^{AS}$).
-The model with the best allocation score received a rank of 1 and the worst received a rank of $n_w$.
-In the case of a tie between one or more models, all models received the better rank.
-We then rescaled these rankings to compute the allocation score standardized rank ($sr_{m,w}^AS$) between 0 and 1, where 0 corresponds to the worst rank and 1 to the best.
+%\begin{equation}
+%sr_{m,w}^{AS} = 1 - \frac{r^{AS}_{m,w}-1}{n_w-1}
+%\end{equation}
 
-\begin{equation}
-sr_{m,w}^{AS} = 1 - \frac{r^{AS}_{m,w}-1}{n_w-1}
-\end{equation}
+%\subsubsection{Weighted Interval Score (WIS)}
 
-\subsubsection{Weighted Interval Score (WIS)}
-
-The Weighted Interval Score (WIS), as described in earlier sections, measures the alignment of a single probabilistic forecast ($F$) with an observation ($y$).
-We computed the mean WIS across all $L$ locations for each model and each forecasted week as
-\begin{equation}
-MWIS_{m,w} = \frac{1}{L}\sum_{l=1}^L WIS(F_{l,m,w},y)
-\end{equation}
-where $F_{l,m,w}$ is the probabilistic forecast from model $m$ for location $l$ and week $w$.
-Using the same procedure as for allocation scores described above, we computed standardized ranks for MWIS ($sr_{m,w}^{MWIS}$).
+%The Weighted Interval Score (WIS), as described in earlier sections, measures the alignment of a single probabilistic forecast ($F$) with an observation ($y$).
+%We computed the mean WIS across all $L$ locations for each model and each forecasted week as
+%\begin{equation}
+%MWIS_{m,w} = \frac{1}{L}\sum_{l=1}^L WIS(F_{l,m,w},y)
+%\end{equation}
+%where $F_{l,m,w}$ is the probabilistic forecast from model $m$ for location $l$ and week $w$.
+%Using the same procedure as for allocation scores described above, we computed standardized ranks for MWIS ($sr_{m,w}^{MWIS}$).
 
@@
 
-All forecast data used in this evaluation are available through the COVID-19 Forecast Hub (cite repository). Code for our analyses is available on GitHub (cite repository).
+All forecast data used in this evaluation are available through the COVID-19 Forecast Hub \citep{cramer_reichlabcovid19-forecast-hub_2021}.
+An R package implementing the allocation score is available at \url{https://github.com/aaronger/alloscore} and all code for the analyses presented in this manuscript is available at \url{https://github.com/aaronger/utility-eval-papers}.
 
@@
 
-For many locations, forecasts predicted lower values than were eventually observed, as this was during the period of steep increase of viral transmission across many states.
+%For many locations, forecasts predicted lower values than were eventually observed, as this was during the period of steep increase of viral transmission across many states.
 Of the 10 models evaluated, the CU-select model had the most accurate forecasts according to the allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table \ref{tab:multi-k-scores}).
@@
   left_join(wis_vals, by="model") |>
-  rename(`IAS centered at 15k` = IAS_wt,
+  rename(Model = model,
+         `IAS centered at 15k` = IAS_wt,
          `IAS uniform` = IAS) |>
-  select(model, AS, MWIS, `IAS centered at 15k`, `IAS uniform`) |>
+  select(Model, AS, MWIS, `IAS centered at 15k`, `IAS uniform`) |>
   arrange(AS) |>
@@
 
-Forecasts and allocations for a selection of states with high numbers of hospitalizations on January 3, 2022 reveal mechanics about how allocations are made (Figure \ref{fig:thermometer-plot}).
-As described in the methods above, allocations for a given location from a specific model are generated by finding the single quantile of the probabilistic forecast across all locations that reaches the resource allocation limit (in this case, assumed to be 15,000).
-A direct comparison between the forecasts from the JHUAPL-Bucky and CU-select models show that while the JHUAPL-Bucky forecast distributions were closer to the eventual observations in many states, the allocations suggested by those forecasts often were more inefficient than those from the CU-select model (Figure \ref{fig:thermometer-plot}).
+%Forecasts and allocations for a selection of states with high numbers of hospitalizations on January 3, 2022 reveal mechanics about how allocations are made (Figure \ref{fig:thermometer-plot}).
+%As described in the methods above, allocations for a given location from a specific model are generated by finding the single quantile of the probabilistic forecast across all locations that reaches the resource allocation limit (in this case, assumed to be 15,000).
+%A direct comparison between the forecasts from the JHUAPL-Bucky and CU-select models show that while the JHUAPL-Bucky forecast distributions were closer to the eventual observations in many states, the allocations suggested by those forecasts often were more inefficient than those from the CU-select model (Figure \ref{fig:thermometer-plot}).
 
@@
 
-For JHUAPL-Bucky, the allocations were assigned at the \Sexpr{Buck_quantile}th percentile of the predictive distribution and for CU-select they were at the \Sexpr{CU_quantile}th percentile.
-This reflects that in general the JHUAPL-Bucky forecasts assigned more probability to higher values and thus a low quantile value across all locations reached the allocation limit of 15,000.
-In California, New York and Texas, the eventually observed number of hospital admissions was closer to the predicted median from JHUAPL-Bucky than the median from CU-select, and was contained in JHUAPL-Bucky's 80\% PI but not in the 80\% PI from CU-select.
+%For JHUAPL-Bucky, the allocations were assigned at the \Sexpr{Buck_quantile}th percentile of the predictive distribution and for CU-select they were at the \Sexpr{CU_quantile}th percentile.
+%This reflects that in general the JHUAPL-Bucky forecasts assigned more probability to higher values and thus a low quantile value across all locations reached the allocation limit of 15,000.
+%In California, New York and Texas, the eventually observed number of hospital admissions was closer to the predicted median from JHUAPL-Bucky than the median from CU-select, and was contained in JHUAPL-Bucky's 80\% PI but not in the 80\% PI from CU-select.
 However for each of these three states, the resource allocation was lower for JHUAPL-Bucky than for CU-select and therefore resulted in more hypothetical unmet need.
-In Florida, which was the state that saw the largest number of reported hospitalizations on this target date, the forecast from JHUAPL-Bucky missed by a wide margin (as did the forecast from CU-select) and the allocation was \Sexpr{FL_diff} units lower than the allocation derived from the CU-select model.
+%In Florida, which was the state that saw the largest number of reported hospitalizations on this target date, the forecast from JHUAPL-Bucky missed by a wide margin (as did the forecast from CU-select) and the allocation was \Sexpr{FL_diff} units lower than the allocation derived from the CU-select model.
 
@@
 
+A comparison of the forecasts from the JHUAPL-Bucky and CU-select models shows that while the JHUAPL-Bucky forecast distributions were closer to the eventual observations in many states, the allocations suggested by those forecasts often were more inefficient than those from the CU-select model (Figure \ref{fig:thermometer-plot}).
 In the example of this one week, JHUAPL-Bucky had a worse allocation score than CU-select because its forecasts led to allocations that sent excess resources to several states, such as Ohio, Pennsylvania, and Michigan, that would have been more effectively allocated to states that did receive enough resources to meet their needs, such as Florida and California (Figure \ref{fig:multi-loc-ranks}A).
@@
 
-Together, these results corroborate the understanding that the allocation score and weighted interval score measure different aspects of forecast performance.
+%Together, these results corroborate the understanding that the allocation score and weighted interval score measure different aspects of forecast performance.
 The allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.
 
-<<multi-loc-ranks, fig.cap="Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select).
+<<multi-loc-ranks, fig.width=8, fig.cap="Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select).
 Panel A shows the observed resource need, in this case the observed number of hospitalizations, for each state, along with the hypothetical number of resources allocated to the given location based on the forecasts from each model.
@@
   geom_hline(yintercept=15000, linetype=2) +
-  ylab("hospital admissions") +
+  ylab("Hospital admissions") +
   xlab(NULL) +
@@
 
-<<metrics-over-time, fig.height=9, fig.width=7, fig.cap="Hospital admissions and evaluation metrics over time.
+<<metrics-over-time, fig.height=8, fig.width=8, fig.cap="Hospital admissions and evaluation metrics over time.
 Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022.
@@
   scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
+  xlab("Forecast target date") +
   ylab("MWIS")+
   theme(legend.pos="bottom")+
-  xlab(NULL)
+  guides(colour = guide_legend(title=NULL, ncol = 4),
+         shape = guide_legend(title=NULL, ncol = 4))
 
@@
                    align="v",
-                   rel_heights = c(1, 1, 1.5),
+                   rel_heights = c(1, 1, 1.85),
                    labels = "AUTO")
@@
 Allocation scores varied substantially by date and by model (Figure \ref{fig:metrics-over-time}).
-For predictions made for the first three Mondays in December 2021 and the last three Mondays in February 2022 all models had allocation scores under 500 (and the mean across all models was less than 100), indicating that the unnecessary unmet need was fairly low on those days.
+For predictions made for the first three Mondays in December 2021 and the last three Mondays in February 2022 all models had allocation scores under 500 (and the mean across all models was less than 100), indicating that unnecessary unmet need was fairly low on those days.
 The allocation scores were on the whole highest when the observed number of new hospital admissions was closest to the resource threshold of 15,000, as those are the times when any mistakes in allocation are costly in terms of wasting resources in one location that could have been used in another.
-Predictions made during the peak week and just after showed the highest variation in allocation scores, with some models having allocation scores under 1000 and others having values over 3500.
+%Predictions made during the peak week and just after showed the highest variation in allocation scores, with some models having allocation scores under 1000 and others having values over 3500.
 
@@
 
-Mean weighted interval scores (MWIS) also varied by date and model, and more clearly were dependent on the scale of the observed data.
-MWIS values were low (all models under 100) for all Mondays in December 2021 and the final four Mondays evaluated.
-Across all models both the average and median MWIS value for every Monday in January was above 100, with the largest errors occurring one and two weeks after the peak was observed.
+%Mean weighted interval scores (MWIS) also varied by date and model, and more clearly were dependent on the scale of the observed data.
+%MWIS values were low (all models under 100) for all Mondays in December 2021 and the final four Mondays evaluated.
+%Across all models both the average and median MWIS value for every Monday in January was above 100, with the largest errors occurring one and two weeks after the peak was observed.
 
@@
 
-<<metrics-correlation, fig.width=10, fig.height=7.5, fig.cap="Association of standardized ranks for MWIS and allocation score by model and week.
+<<metrics-correlation, fig.width=8, fig.height=9, fig.cap="Association of standardized ranks for MWIS and allocation score by model and week.
 Each facet of the plot corresponds to one model.
@@
   filter(as.Date(reference_date.x) < as.Date("2022-02-01") ) |>
+  dplyr::mutate(target_end_date = as.Date(target_end_date)) |>
   ggplot(aes(x=mwis_rank_std, y=allo_rank_std, color=target_end_date, size=yobs)) +
@@
     scale_y_continuous("Allocation score standardized rank", limits=c(0,1)) +
-  facet_wrap(.~model) +
-  geom_abline(slope=1, intercept=0, linetype=2, color="gray")
+#    scale_color_gradient("Target date") +
+    scale_color_date(
+      name="Target date",
+      breaks = seq.Date(from=as.Date("2021-12-13"), by = 14, length.out=5)) +
+    scale_size_continuous("Observed\nhospitalizations") +
+    facet_wrap(.~model, ncol=3) +
+    geom_abline(slope=1, intercept=0, linetype=2, color="gray") #+
+#    theme(legend.pos="bottom")
 
@@
 
-The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the constraint ($K$), possibly taking weights into account for different values of $K$ that might be more or less likely (Section \ref{sec:methods.detailed.integrated_allocation}).
-This could be useful in situations where the actual constraint may not be known precisely, or where we wish to consider results across a range of decision making settings with different constraints.
-Unlike in the above analyses where we conducted analyses assuming that $K$ was known to be 15,000, the analyses in this subsection are conducted assuming different distributions on $K$.
+%The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the constraint ($K$), possibly taking weights into account for different values of $K$ that might be more or less likely (Section \ref{sec:methods.detailed.integrated_allocation}).
+%This could be useful in situations where the actual constraint may not be known precisely, or where we wish to consider results across a range of decision making settings with different constraints.
+%Unlike in the above analyses where we conducted analyses assuming that $K$ was known to be 15,000, the analyses in this subsection are conducted assuming different distributions on $K$.
 
@@
 AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 (Figure \ref{fig:multi-k}A).
-These calculations highlight that AS was highest at the $K$ nearest to the observed value of \Sexpr{format(tot_y, big.mark = ",")}.
+These calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of \Sexpr{format(tot_y, big.mark = ",")} that day.
 Rankings of AS from all models were fairly stable across a range of values for K, with some crossings, especially at $K$ values further away from the observed value.
 
+The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the constraint ($K$), possibly taking weights into account for different values of $K$ that might be more or less likely (Section \ref{sec:methods.detailed.integrated_allocation}).
 IAS was computed for two distributions on $K$, one uniform across the entire range and the other a symmetric distribution around 15,000 (Figure \ref{fig:multi-k}B).
@@
   geom_smooth(method="lm", se=FALSE, linetype=2) +
-  scale_color_manual(name = "Integrated Allocation Score", values = c("#386cb0", "#fdc086")) +
-  scale_shape_manual(name = "Integrated Allocation Score", values = c(1, 2)) +
+  scale_color_manual(name = "Integrated allocation score weighting", values = c("#386cb0", "#fdc086")) +
+  scale_shape_manual(name = "Integrated allocation score weighting", values = c(1, 2)) +
   geom_abline(slope=1, intercept=0, color="gray") +
+  xlab("Allocation score") +
+  ylab("Integrated allocation score") +
   theme(legend.position = c(0.05,.9), legend.justification = c(0,1),
@@
   geom_point(y=1/length(unique(slim_alloscore_2021.12.20$K)), color="#fdc086", shape = 2) +
-  geom_vline(xintercept = 15000, color="gray")
+  geom_vline(xintercept = 15000, color="gray") +
+  ylab("Weight")
 
@@
 
-In epidemiological forecasting, well-known proper scoring rules such as the log score or variations on the continuous rank probability score (such as the weighted interval score, or WIS) have been frequently utilized to evaluate probabilistic forecasts.
-Often, these scores are used to rank models according to accuracy with that particular score, but without reference to the underlying decision-making process for which that score was designed.
+In epidemiological forecasting, well-known proper scoring rules such as the log score or variations on the continuous ranked probability score (such as the weighted interval score, or WIS) have been frequently utilized to evaluate probabilistic forecasts.
+Often, these scores are used to rank models by accuracy according to that particular score, but without reference to the underlying decision-making process for which that score was designed.
 With careful thought and collaboration between modelers and public health officials, we argue that scores that are more aligned with public health decisions could be developed to inform specific problems.
@@
 
-We often conceive of infectious disease forecasts as being useful for decision making purposes, but it is rare for forecast evaluation to be tied directly to the value of the forecasts for informing those decisions.
+%We often conceive of infectious disease forecasts as being useful for decision making purposes, but it is rare for forecast evaluation to be tied directly to the value of the forecasts for informing those decisions.
 This work seeks to address that gap.
-However, we do note that the decision-making context presented in this work, while motivated by examples from real-world public health resource allocation problems, has not been used to inform an actual real-time decision-making process.
+%However, we do note that the decision-making context presented in this work, while motivated by examples from real-world public health resource allocation problems, has not been used to inform an actual real-time decision-making process.
 
-Resource allocation decisions are a realistic example of the kinds of decisions that could motivate more targeted forecasting exercises.
-One real-world example is allocation of ventilators during a respiratory viral pandemic\citep{huang_stockpiling_2017}.
-This was the motivating example for the examples presented in the work above.
-However, other examples where allocation is a central concern exist as well, such as the allocation of a limited stockpile of vaccinations\citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests\citep{du_optimal_2022,pasco_covid-19_2023}.
+While our example used a hypothetical scenario to illustrate a potential application of the allocation score, resource allocation decisions are a realistic example of the kinds of decisions that could motivate more targeted forecasting exercises.
+One real-world example, which was the motivating example presented in the work above, is allocation of ventilators during a respiratory viral pandemic\citep{huang_stockpiling_2017}.
+Other examples include the allocation of a limited stockpile of vaccinations\citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests\citep{du_optimal_2022,pasco_covid-19_2023}.
 
@@
 Not all can be easily quantified.
-Those that can be easily quantified may differ enough that it seems likely that no single score would be appropriate for all users.
-Ideally, a forecasting tool could be developed through close collaboration between modelers and public health officials.
-However, this may only be possible in settings with both an analytics and a public health team.
+Those that can be easily quantified may differ enough that no single score would be appropriate for all users.
+Ideally, targeted forecasting tools could be developed through close collaboration between modelers and public health officials.
+However, this may only be possible in settings with sufficient staffing on both an analytics and a public health team.
 Increasingly, collaborative modeling hubs are being used to generate ``one-size-fits-all'' forecasts for many locations at once.
@@
 First, it is clear that different metrics (in our case allocation score and mean weighted interval score) captured different aspects of forecast performance.
-The mean WIS (MWIS) was strongly dependent on the scale of the forecasted quantity (e.g., when hospitalizations were high, so were MWIS scores).
-The allocation score was not as scale-dependent, as the highest/worst allocation scores were observed when the number of hospitalizations was closest to the allocation constraint.
+The allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.
+%A forecast with a good allocation score can have a poor weighted interval score and vice versa.
+%The mean WIS (MWIS) was strongly dependent on the scale of the forecasted quantity (e.g., when hospitalizations were high, so were MWIS scores).
+%The allocation score was not as scale-dependent, as the highest/worst allocation scores were observed when the number of hospitalizations was closest to the allocation constraint.
 We also note that contributed models found it hard to achieve a better allocation score than a na\"ive baseline model that just predicted a flat line from the last previous observation with wide uncertainty.
-This suggests suggests that contributed models (other than the ensemble, which combined forecasts from all contributed models) were not consistently adding value over just extrapolating from the current levels.
+This suggests suggests that contributed models (other than the ensemble, which combined forecasts from all contributed models) were not consistently adding value to allocation decisions over just extrapolating from the current levels.
 
@@
 
-An additional limitation of the current work and opportunity for further investigation is to more carefully evaluate whether a forecast adds value to existing decision making processes.
-In the context of allocation-based decisions, standard procedures might involve extrapolating need based on a current observed data (similar to the `baseline' model presented above), with or without adjustments based on other political or real-world considerations.
-For example, in many settings public health stakeholders will synthesize multiple sources of information, e.g.
-real-time data from a wide variety of quantitative and qualitative sources coupled with expert judgment based on a situational understanding of past and current experience.
-The allocation score presented in this work assesses the optimality of the allocation with reference an omniscient view of the future captured by the oracle.
-However, what we care about is whether a given forecast adds useful information to an existing decision-making process.
-While the scoring procedures as presented do not directly address this question, they could be modified (say, by comparison to a `baseline' model) to quantify the benefit---if any---of using a forecast to inform a specific decision.
+An opportunity for further investigation is to more carefully evaluate whether a forecast adds value to existing decision making processes.
+In the context of decisions about allocations, standard procedures might involve extrapolating need based on a current observed data (similar to the `baseline' model presented above), with or without adjustments based on other political or real-world considerations.
+For example, in many settings public health stakeholders will make decisions after synthesizing information from a variety of quantitative and qualitative sources coupled with expert judgment.
+The allocation score presented in this work does not directly measure whether a given forecast adds useful information to such an existing decision-making process.
+While the scoring procedures as presented do not directly address this question, they could be modified (say, by comparison to a `baseline' model or expert-elicited allocations in the absence of forecast data) to quantify the benefit of using a forecast to inform a specific decision.
 
-We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling applied epidemiological forecasting with public health decision making.
-We note that a few papers have begun exploring similar linkages as described in the literature review{\textendash}but we see much room for additional work in this area.
-In some situations, individual or ensemble models could be developed to optimize scores that are attuned to a particular decision making setting.
-This area is also largely unexplored in the realm of public health to date, with some initial methodological development in econometrics \citep{loaiza-maya_focused_2021}.
+%We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling applied epidemiological forecasting with public health decision making.
+%We note that a few papers have begun exploring similar linkages as described in the literature review{\textendash}but we see much room for additional work in this area.
+%In some situations, individual or ensemble models could be developed to optimize scores that are attuned to a particular decision making setting.
+%This area is also largely unexplored in the realm of public health to date, with some initial methodological development in econometrics \citep{loaiza-maya_focused_2021}.
 
@@
 New collaborative work between public health officials and modeling teams is needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or simulation exercises that could be used to inform further development of new or alternative scoring metrics.
+We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling applied epidemiological forecasting with public health decision making.
 

### 2023-12-19 alloscore_manuscript/alloscore-application-shorter.Rnw
update figure 2, show top 10 locations

@@
 
-<<thermometer-plot, fig.height=3.5, fig.width=7.5, fig.cap="Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the nine highest hospitalization counts.
+<<thermometer-plot, fig.height=7, fig.width=7.5, fig.cap="Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the ten highest hospitalization counts.
 For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts.
@@
 source("R/plot_functions.R")
+
+locs_to_plot <- truth_data |>
+  dplyr::filter(target_end_date == "2022-01-03", location != "US") |>
+  dplyr::arrange(desc(value)) |>
+  head(10) |>
+  dplyr::pull(abbreviation)
+
 plot_hosp(models = c("JHUAPL-Bucky", "CU-select"),
@@
           ## highest locations in terms of hospitalizations
-          locations = c("FL", "NY", "TX", "CA"),
+          locations = locs_to_plot,
           f_width1 = 4, f_alpha = .4,
+          facet_ncol=2,
           allocations = TRUE,
-          forecasts_hosp = forecast_data %>%  filter(reference_date == "2021-12-20"),
-          alloscores = slim_alloscore_2021.12.20,
-          truth = truth_data,
+          forecasts_hosp = forecast_data %>% filter(reference_date == "2021-12-20") %>%
+            dplyr::mutate(
+              abbreviation = factor(abbreviation, levels=locs_to_plot, ordered=TRUE),
+              geo_value = abbreviation),
+          alloscores = slim_alloscore_2021.12.20 %>%
+            dplyr::mutate(
+              abbreviation = factor(abbreviation, levels=locs_to_plot, ordered=TRUE)),
+          truth = truth_data %>%
+            dplyr::mutate(
+              abbreviation = factor(abbreviation, levels=locs_to_plot, ordered=TRUE),
+              geo_value = abbreviation),
           f_colors = c("#4A708B", "#CD2626"),

### 2023-12-19 R/plot_functions.R

@@ -31,6 +31,7 @@ plot_hosp <- function(
     models = "COVIDhub-ensemble",
     locations = "US",
     geofacet = FALSE,
+    facet_ncol=3,
     key_width = .15,
     allocations = FALSE,
     one_K = NULL,
@@ -60,7 +61,8 @@ plot_hosp <- function(
     target_end_date <= stop_date) %>%
   dplyr::mutate(
     validation = ifelse(target_end_date > as.Date(f_date), "Validation", "Historical"),
-    code = as.factor(toupper(geo_value)),
+    code = geo_value,
+    # code = as.factor(toupper(geo_value)),
     xmin = target_end_date - half_width - space,
     xmax = xmin + total_width + 2 * space)
 
@@ -71,7 +73,8 @@ plot_hosp <- function(
   fc_dat <- forecasts_hosp %>%
     dplyr::filter(abbreviation %in% locations, model %in% models) %>%
     dplyr::mutate(
-      code = as.factor(toupper(geo_value)),
+      code = geo_value,
+      # code = as.factor(toupper(geo_value)),
       model = forcats::fct_relevel(model, models),
       xmin = target_end_date - half_width + (match(model, models) - 1)*(f_width1 + space),
       xmax = xmin + f_width1)
@@ -81,12 +84,12 @@ plot_hosp <- function(
   # truth <- truth %>% mutate(
   #   code = fct_relevel(code, loc_abbrevs))
 
-  if (!is.null(st_colors)) {
-    fc_dat <- fc_dat %>% mutate(
-      code = fct_relevel(code, names(st_colors)))
-    truth <- truth %>% mutate(
-      code = fct_relevel(code, names(st_colors)))
-  }
+  # if (!is.null(st_colors)) {
+  #   fc_dat <- fc_dat %>% mutate(
+  #     code = fct_relevel(code, names(st_colors)))
+  #   truth <- truth %>% mutate(
+  #     code = fct_relevel(code, names(st_colors)))
+  # }
 
   p <- ggplot(data = truth) +
     # truth data
@@ -216,9 +219,9 @@ plot_hosp <- function(
     theme(axis.title = element_blank())
   } else if (length(locations) > 1) {
     if (free_y) {
-      p <- p + facet_wrap(~ code, scales = "free_y")
+      p <- p + facet_wrap(~ code, scales = "free_y", ncol=facet_ncol)
     } else {
-      p <- p + facet_wrap(~ code)
[…]

### 2023-12-19 alloscore_manuscript/alloscore-application-shorter.Rnw

--- 2023-12-19 alloscore_manuscript/alloscore-application-shorter.Rnw
@@
 \paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$ and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$.
-The means of these distributions are given by the scale parameters $\sigma_i$.
 When the marginal forecasts are exponential distributions, it can be shown that the optimal allocation divides the available resources among the locations proportionally to the scale parameters $\sigma_i$ (see section
@@
   left_join(wis_vals, by="model") |>
-  rename(Model = model,
-         `IAS centered at 15k` = IAS_wt,
+  rename(`IAS centered at 15k` = IAS_wt,
          `IAS uniform` = IAS) |>
-  select(Model, AS, MWIS, `IAS centered at 15k`, `IAS uniform`) |>
+  select(model, AS, MWIS, `IAS centered at 15k`, `IAS uniform`) |>
   arrange(AS) |>
@@
 
-<<multi-loc-ranks, fig.width=8, fig.cap="Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select).
+<<multi-loc-ranks, fig.cap="Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select).
 Panel A shows the observed resource need, in this case the observed number of hospitalizations, for each state, along with the hypothetical number of resources allocated to the given location based on the forecasts from each model.
@@
   geom_hline(yintercept=15000, linetype=2) +
-  ylab("Hospital admissions") +
+  ylab("hospital admissions") +
   xlab(NULL) +
@@
 
-<<metrics-over-time, fig.height=8, fig.width=8, fig.cap="Hospital admissions and evaluation metrics over time.
+<<metrics-over-time, fig.height=9, fig.width=7, fig.cap="Hospital admissions and evaluation metrics over time.
 Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022.
@@
   scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
-  xlab("Forecast target date") +
   ylab("MWIS")+
   theme(legend.pos="bottom")+
-  guides(colour = guide_legend(title=NULL, ncol = 4),
-         shape = guide_legend(title=NULL, ncol = 4))
+  xlab(NULL)
 
@@
                    align="v",
-                   rel_heights = c(1, 1, 1.85),
+                   rel_heights = c(1, 1, 1.5),
                    labels = "AUTO")
@@
 
-<<metrics-correlation, fig.width=8, fig.height=9, fig.cap="Association of standardized ranks for MWIS and allocation score by model and week.
+<<metrics-correlation, fig.width=10, fig.height=7.5, fig.cap="Association of standardized ranks for MWIS and allocation score by model and week.
 Each facet of the plot corresponds to one model.
@@
   filter(as.Date(reference_date.x) < as.Date("2022-02-01") ) |>
-  dplyr::mutate(target_end_date = as.Date(target_end_date)) |>
   ggplot(aes(x=mwis_rank_std, y=allo_rank_std, color=target_end_date, size=yobs)) +
@@
   scale_y_continuous("Allocation score standardized rank", limits=c(0,1)) +
-#    scale_color_gradient("Target date") +
-    scale_color_date(
-      name="Target date",
-      breaks = seq.Date(from=as.Date("2021-12-13"), by = 14, length.out=5)) +
-    scale_size_continuous("Observed\nhospitalizations") +
-    facet_wrap(.~model, ncol=3) +
-    geom_abline(slope=1, intercept=0, linetype=2, color="gray") #+
-#    theme(legend.pos="bottom")
+  facet_wrap(.~model) +
+  geom_abline(slope=1, intercept=0, linetype=2, color="gray")
 
@@
   geom_smooth(method="lm", se=FALSE, linetype=2) +
-  scale_color_manual(name = "Integrated allocation score weighting", values = c("#386cb0", "#fdc086")) +
-  scale_shape_manual(name = "Integrated allocation score weighting", values = c(1, 2)) +
+  scale_color_manual(name = "Integrated Allocation Score", values = c("#386cb0", "#fdc086")) +
+  scale_shape_manual(name = "Integrated Allocation Score", values = c(1, 2)) +
   geom_abline(slope=1, intercept=0, color="gray") +
-  xlab("Allocation score") +
-  ylab("Integrated allocation score") +
   theme(legend.position = c(0.05,.9), legend.justification = c(0,1),
@@
   geom_point(y=1/length(unique(slim_alloscore_2021.12.20$K)), color="#fdc086", shape = 2) +
-  geom_vline(xintercept = 15000, color="gray") +
-  ylab("Weight")
+  geom_vline(xintercept = 15000, color="gray")
 

### 2023-12-20 alloscore_manuscript/alloscore-application-shorter.Rnw
add main text paragraph outlining use of distfromq

--- 2023-12-19 alloscore_manuscript/alloscore-application-shorter.Rnw
@@
 
-We measured forecast skill using two forecast scores, the allocation score and the weighted interval score (WIS), both defined above.
+We measured forecast skill using two forecast scores, the allocation score (AS) and the weighted interval score (WIS), both defined above.
 We computed these scores for the 14 day ahead forecasts made each week.
-For this analysis, we fixed the resource constraint used for the allocation score to be $K=15,000$, based roughly on a reported number of ventilators available for reallocation in the US \citep{ajao_assessing_2015}.
+For this analysis, we fixed the resource constraint used for the AS to be $K=15,000$, based roughly on a reported number of ventilators available for reallocation in the US \citep{ajao_assessing_2015}.
 We computed the mean WIS (MWIS) across all of the forecasted state-level locations.
@@
 In the case of a tie between one or more models, all models received the better rank.
+
+As described above, predictions were submitted to the Forecast Hub in the form of a set of 23 quantiles of the predictive distribution.
+The WIS can be directly calculated from these quantiles.
+However, calculation of the AS requires that the full cumulative distribution functions of the forecast distribution are available for each location.
+For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution functions based on the provided quantiles.
+On the interior of the provided quantiles we used a monotonic cubic spline to interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as to match the two lowest and two highest quantiles (see the supplement for further details).
+In the supplement, we show that if this evaluation procedure had been specified prospectively, the resulting score would be proper \textemdash but a post hoc application of this procedure is improper.
+We use the procedure here to illustrate the properties of the score, and note that a forecaster or collaborative forecasting exercise interested in using the allocation score for evaluation could circumvent issues with propriety through thoughtful elicitation of representations of forecast distributions or by collecting the allocations at specified resource levels as part of forecast submissions.
 

### 2023-12-20 alloscore_manuscript/alloscore-application-shorter.tex
adding bib update

--- 2023-12-19 alloscore_manuscript/alloscore-application-shorter.Rnw
@@
 \subsection{Data}
-<<get-data-dates>>=
-metadata <- tar_meta()
-@
 
@@
 As such, we analyzed forecasts for the 15 weeks starting with Monday November 22, 2021 through Monday February 28, 2022.
-%Forecast data were downloaded on for analysis on \Sexpr{format(as.Date(metadata$time[which(metadata$name == "forecast_data")]), "%B %d, %Y")}.
+%Forecast data were downloaded on for analysis on format(as.Date(metadata$time[which(metadata$name == "forecast_data")]), "%B %d, %Y").
 
@@
 
-The hospitalization data used for scoring forecasts were downloaded on \Sexpr{format(as.Date(metadata$time[which(metadata$name == "truth_data")]), "%B %d, %Y")}.
+The hospitalization data used for scoring forecasts were downloaded on December 03, 2023.
 
@@
 
-<<score-computation>>=
-tar_load("Kat15k_alloscores")
-tar_load("score_data")
-tar_load("truth_data")
-tar_load("forecast_data")
-
-alloscore_data_summaries <- Kat15k_alloscores |>
-  group_by(model, reference_date) |>
-  summarize(alloscore = sum(components), yobs = sum(y)) |>
-  group_by(reference_date) |>
-  mutate(nmodels = n(),
-         allo_rank = rank(alloscore, ties.method = "min"),
-         allo_rank_std = 1- (allo_rank-1)/(nmodels-1)) |>
-  ungroup() |>
-  complete(model, reference_date) |>
-  mutate(target_end_date = as.Date(reference_date) + 14)
-
-alloscore_models <- unique(alloscore_data_summaries$model)
-score_data_summaries <- score_data |>
-  mutate(reference_date = target_end_date - 14) |>
-  filter(model %in% alloscore_models) |>
-  group_by(model, reference_date) |>
-  summarize(mwis = mean(wis)) |>
-  ungroup() |>
-  group_by(reference_date) |>
-  mutate(nmodels = n(),
-         mwis_rank = rank(mwis, ties.method = "min"),
-         mwis_rank_std = 1- (mwis_rank-1)/(nmodels-1)) |>
-  ungroup() |>
-  complete(model, reference_date) |>
-  mutate(target_end_date = as.Date(reference_date) + 14)
-
-all_scores <- left_join(alloscore_data_summaries, score_data_summaries,
-                      by=c("model", "target_end_date"))
-@
-
 \subsubsection{Anatomy of forecast scores for one week}
@@
 
-<<ias-comparison>>=
-tar_load(alloscore_2021.12.20)
-slim_alloscore_2021.12.20 <- alloscore::slim(alloscore_2021.12.20, id_cols = c("model", "reference_date"))
-
-## for visual context
-## curve(dnorm(x, mean=15000, sd=3000), from=0, to=30000)
-ks <- seq(5000, 25000, by=200)
-kprobs <- data.frame(K = ks,
-                     weight = dnorm(ks, mean=15000, sd=3000)/sum(dnorm(ks, mean=15000, sd=3000)))
-
-ias <- slim_alloscore_2021.12.20 |>
-  group_by(model, K) |>
-  summarize(`Allocation score` = sum(components)) |>
-  left_join(kprobs) |>
-  group_by(model) |>
-  summarize(IAS_wt = sum(`Allocation score`*weight, na.rm=TRUE),
-            IAS = mean(`Allocation score`)) |>
-  arrange(IAS)
-
-as <- slim_alloscore_2021.12.20 |>
-  filter(K == 15000) |>
-  group_by(model) |>
-  summarize(AS = sum(components)) |>
-  arrange(AS)
-@
-
-<<multi-k-scores, results='asis'>>=
-wis_vals <- all_scores |>
-  filter(target_end_date == "2022-01-03", !is.na(mwis)) |>
-  select(model,
-# can uncomment the two lines below to check they match
-#         AS = alloscore,
-#         `AS rank` = allo_rank,
-         MWIS = mwis)
-
-left_join(ias, as) |>
-  left_join(wis_vals, by="model") |>
-  rename(`IAS centered at 15k` = IAS_wt,
-         `IAS uniform` = IAS) |>
-  select(model, AS, MWIS, `IAS centered at 15k`, `IAS uniform`) |>
-  arrange(AS) |>
-  xtable::xtable(caption = "Comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS).
+% latex table generated in R 4.3.1 by xtable 1.8-4 package
+% Tue Dec 19 17:26:53 2023
+\begin{table}[ht]
+\centering
+\begin{tabular}{lrrrr}
+  \hline
+model & AS & MWIS & IAS centered at 15k & IAS uniform \\ 
+  \hline
+CU-select & 669 & 133 & 774 & 326 \\ 
+  COVIDhub-ensemble & 873 & 159 & 1067 & 438 \\ 
+  USC-SI\_kJalpha & 995 & 91 & 1216 & 1097 \\ 
+  JHUAPL-Gecko & 1034 & 164 & 1141 & 418 \\ 
+  MUNI-ARIMA & 1084 & 169 & 1248 & 440 \\ 
+  COVIDhub-trained\_ensemble & 1089 & 169 & 1271 & 823 \\ 
+  COVIDhub-baseline & 1175 & 170 & 1317 & 535 \\ 
+  JHUAPL-Bucky & 1358 & 102 & 1566 & 1214 \\ 
+  JHUAPL-SLPHospEns & 1540 & 129 & 1604 & 1102 \\ 
+  UVA-Ensemble & 2469 & 213 & 2635 & 2494 \\ 
+   \hline
+\end{tabular}
+\caption{Comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS).
 All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03.
 Results are sorted by AS.
-For all metrics, lower scores indicate better accuracy.",
-                 digits = 0,
-                 label = "tab:multi-k-scores") |>
-  print(include.rownames = FALSE, )
-@
-
-<<metrics-rankings, results = "asis", eval=FALSE>>=
-## NOTE: this content now shown in multi-k-scores chunk
-## for highlighting cells, if we want that https://stackoverflow.com/questions/37448348/color-code-cells-of-xtable
-all_scores |>
-  filter(target_end_date == "2022-01-03", !is.na(mwis)) |>
-  select(model,
-         AS = alloscore,
-         `AS rank` = allo_rank,
-         MWIS = mwis,
-         `MWIS rank` = mwis_rank) |>
-  arrange(`AS rank`) |>
-  xtable::xtable(label = "tab:one-week-results",
-                 caption = "Allocation scores (AS) and mean weighted interval scores (MWIS) for each model' as forecasted's forecasts for ") |>
-  print(include.rownames = FALSE)
-@
+For all metrics, lower scores indicate better accuracy.}
+\label{tab:multi-k-scores}
+\end{table}
 
@@
 
-<<detailedASinfo>>=
-source("R/run-alloscore.R")
-as_Buck <- run_alloscore(
-  forecast_data = forecast_data,
-  truth_data = truth_data,
-  one_reference_date = "2021-12-20",
-  one_model = "JHUAPL-Bucky",
-  one_K = 15000,
-  slim = FALSE)
-
-as_CU <- run_alloscore(
-  forecast_data = forecast_data,
-  truth_data = truth_data,
-  one_reference_date = "2021-12-20",
-  one_model = "CU-select",
-  one_K = 15000,
-  slim = FALSE)
-
-Buck_quantile <- round(1 - as_Buck$lam,2)*100
-CU_quantile <- round(1 - as_CU$lam,2)*100
-
-FL_allocs <- Kat15k_alloscores |>
-   dplyr::filter(
-     model %in% c("JHUAPL-Bucky", "CU-select"),
-     reference_date == "2021-12-20",
-     abbreviation == "FL") |>
-  dplyr::select(model, x) |>
-  tibble::deframe()
-FL_diff <- round(FL_allocs["CU-select"] - FL_allocs["JHUAPL-Bucky"])
-@
-
-%For JHUAPL-Bucky, the allocations were assigned at the \Sexpr{Buck_quantile}th percentile of the predictive distribution and for CU-select they were at the \Sexpr{CU_quantile}th percentile.
+%For JHUAPL-Bucky, the allocations were assigned at the Buck_quantileth percentile of the predictive distribution and for CU-select they were at the CU_quantileth percentile.
 %This reflects that in general the JHUAPL-Bucky forecasts assigned more probability to higher values and thus a low quantile value across all locations reached the allocation limit of 15,000.
@@
 However for each of these three states, the resource allocation was lower for JHUAPL-Bucky than for CU-select and therefore resulted in more hypothetical unmet need.
-%In Florida, which was the state that saw the largest number of reported hospitalizations on this target date, the forecast from JHUAPL-Bucky missed by a wide margin (as did the forecast from CU-select) and the allocation was \Sexpr{FL_diff} units lower than the allocation derived from the CU-select model.
+%In Florida, which was the state that saw the largest number of reported hospitalizations on this target date, the forecast from JHUAPL-Bucky missed by a wide margin (as did the forecast from CU-select) and the allocation was FL_diff units lower than the allocation derived from the CU-select model.
 
-<<thermometer-plot, fig.height=3.5, fig.width=7.5, fig.cap="Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the nine highest hospitalization counts.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\includegraphics[width=\maxwidth]{figure/thermometer-plot-1} \caption[Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the nine highest hospitalization counts]{Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the nine highest hospitalization counts.
 For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts.
-The side-by-side shaded regions show the median (solid horizontal line) and 50%, 80% and 95% prediction intervals for the two selected models.
+The side-by-side shaded regions show the median (solid horizontal line) and 50\%, 80\% and 95\% prediction intervals for the two selected models.
 The forecasts were made for new hospitalizations on January 3, 2022 (vertical dashed line, with number of hospitalizations indicated by red horizontal line segment at the intersection of the dashed line and the grey line of data).
 The vertical bars with red and yellow shading show the allocations: the red bar goes from zero to the level of the observed number of hospitalizations while the yellow bar shows how many resources were suggested by the model to be allocated to that location.
-The allocations (yellow bar) may be more or less than the observed number, therefore exposed red bar indicates unmet need, while the yellow bar extending above the observed number of hospitalizations means that excess resources were suggested for that given location.">>=
-source("R/plot_functions.R")
-plot_hosp(models = c("JHUAPL-Bucky", "CU-select"),
-          start_date = "2021-12-01",
-          stop_date = "2022-01-25",
-          f_date = "2021-12-20",
-          te_date = "2022-01-03",
-          ## highest locations in terms of hospitalizations
-          locations = c("FL", "NY", "TX", "CA"),
-          f_width1 = 4, f_alpha = .4,
-          allocations = TRUE,
-          forecasts_hosp = forecast_data %>%  filter(reference_date == "2021-12-20"),
-          alloscores = slim_alloscore_2021.12.20,
-          truth = truth_data,
-          f_colors = c("#4A708B", "#CD2626"),
-          one_K = 15000) + labs(x = NULL)
-@
+The allocations (yellow bar) may be more or less than the observed number, therefore exposed red bar indicates unmet need, while the yellow bar extending above the observed number of hospitalizations means that excess resources were suggested for that given location.}\label{fig:thermometer-plot}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<multi-loc-ranks, fig.cap="Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select).
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\includegraphics[width=\maxwidth]{figure/multi-loc-ranks-1} \caption[Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select)]{Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select).
 Panel A shows the observed resource need, in this case the observed number of hospitalizations, for each state, along with the hypothetical number of resources allocated to the given location based on the forecasts from each model.
@@
 Panel B shows the breakdown of the weighted interval score (WIS) into components of underprediction, overprediction and dispersion.
-Larger values of WIS indicate more error, and the full WIS score for each location can be decomposed into the three components shown here.">>=
-## date points to target_end_date of 2022-01-10 to highlight
-ref_date_for_plot <- "2021-12-20"
-models_to_plot <- c("JHUAPL-Bucky", "CU-select")
+Larger values of WIS indicate more error, and the full WIS score for each location can be decomposed into the three components shown here.}\label{fig:multi-loc-ranks}
+\end{figure}
 
-multiloc_dat <- Kat15k_alloscores |>
-  filter(reference_date == ref_date_for_plot,
-         ## the above date points to target_end_date of 2022-01-03 where things were really wrong
-         model %in% models_to_plot) |>
-  rename(allocation = x, `observed resource need` = y, state = abbreviation) |>
-  select(model, reference_date, state, allocation, `observed resource need`) |>
-  pivot_longer(cols = c("allocation", "observed resource need"))
-
-## plot all locations observations and allocations
-
-## line plot below here
-# multiloc_plot <- ggplot(multiloc_dat, aes(x=name, y=value, group=state)) +
-#   geom_line() +
-#   geom_line(data=dplyr::filter(multiloc_dat, state=="FL"), color="orange") +
-#   ylab("Daily hospital admissions") +
-#   scale_x_discrete(name=NULL, limits=rev, expand = c(.1,.1)) +
-#   facet_grid(.~model)
-
-ordered_levels <- truth_data |>
-  filter(target_end_date == as.Date(ref_date_for_plot)+14, location != "US") |>
-  select(location, abbreviation, value) |>
-  arrange(-value)
-
-alloc_plot <- multiloc_dat |>
-  mutate(state = factor(state, levels = ordered_levels$abbreviation),
-         name = factor(name, levels = c("observed resource need", "allocation"))) |>
-  arrange(name) |>
-  ggplot(aes(x=state, y=value, fill=name)) +
-  scale_fill_manual(name=NULL, values = c("#CD2626", "goldenrod")) +
-  theme(axis.text.x=element_text(angle=90,vjust=.5), legend.position = c(.9,.95), legend.justification = c(1,1)) +
-  ylab("Resource need and allocation") +
-  xlab(NULL) +
-  geom_col(position="identity", alpha=.7) +
-  facet_grid(model~.)
-
-wis <- covidHubUtils::score_forecasts(
-  forecast_data |> filter(reference_date == "2021-12-20", str_detect(model,"Buck|CU-select")),
-  truth_data,
-  use_median_as_point = TRUE,
-  metrics = "wis_components",
-  return_format="long") |>
-  filter(location != "US", score_name != "true_value") |>
-  left_join(select(ordered_levels, location, abbreviation)) |>
-  mutate(abbreviation = factor(abbreviation, levels=ordered_levels$abbreviation),
-         `WIS component` = factor(score_name, levels = c("overprediction", "dispersion", "underprediction")))
-
-## palette from
-## colorspace::divergingx_hcl(palette = "Blue-Yellow 3", n=3)
-
-wis_plot <- ggplot(data = wis, aes(x=abbreviation)) +
-  geom_col(aes(y=score_value, fill=`WIS component`)) +
-  scale_fill_manual(values = c("#9FA2FF", "gray77", "#BAAE00")) +
-  facet_grid(model~.) +
-  theme(axis.text.x=element_text(angle=90,vjust=.5), legend.position = c(.9,.95), legend.justification = c(1,1)) +
-  xlab(NULL) +
-  ylab("WIS")
-
-cowplot::plot_grid(alloc_plot, wis_plot,
-                   nrow=2,
-                   align="v",
-                   rel_heights = c(1, 1),
-                   labels = "AUTO")
-
-@
+\end{knitrout}
 
 \subsubsection{Forecast scores showed differences over time}
-<<grob-hosps-over-time>>=
-dat_to_plot <- truth_data |>
-  filter(location == "US",
-         target_end_date %in% score_data_summaries$target_end_date)
-p1 <- ggplot(dat_to_plot,
-             aes(x=target_end_date, y=value)) +
-  geom_line() +
-  geom_point() +
-  geom_hline(yintercept=15000, linetype=2) +
-  ylab("hospital admissions") +
-  xlab(NULL) +
-  ylim(c(0, NA))
-@
 
-<<metrics-over-time, fig.height=9, fig.width=7, fig.cap="Hospital admissions and evaluation metrics over time.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\includegraphics[width=\maxwidth]{figure/metrics-over-time-1} \caption[Hospital admissions and evaluation metrics over time]{Hospital admissions and evaluation metrics over time.
 Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022.
@@
 Panel C shows the MWIS metric across weeks, averaged across all states.
-MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.">>=
-p2 <- ggplot(alloscore_data_summaries,
-             aes(x=target_end_date, y=alloscore, color=model, shape=model)) +
-  scale_x_date(limits=range(score_data_summaries$target_end_date)) +
-  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
-  ylab("Allocation score") +
-  geom_point() + geom_line()+
-  theme(legend.pos="none") +
-  xlab(NULL)
+MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.}\label{fig:metrics-over-time}
+\end{figure}
 
-p4 <- ggplot(score_data_summaries,
-             aes(x=target_end_date, y=mwis, color=model,shape=model)) +
-  geom_point() + geom_line() +
-  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
-  ylab("MWIS")+
-  theme(legend.pos="bottom")+
-  xlab(NULL)
-
-cowplot::plot_grid(p1, p2, p4,
-                   nrow=3,
-                   align="v",
-                   rel_heights = c(1, 1, 1.5),
-                   labels = "AUTO")
-@
+\end{knitrout}
 
@@
 
-<<eval=FALSE>>=
-## some analysis to support above statements
-alloscore_data_summaries |>
-  group_by(target_end_date) |>
-  summarize(meanAS = mean(alloscore, na.rm=TRUE),
-            medianAS = median(alloscore, na.rm=TRUE),
-            minAS = min(alloscore, na.rm=TRUE),
-            maxAS = max(alloscore, na.rm=TRUE),
-            range = maxAS-minAS)
-@
-
 Overall, across the first 11 weeks evaluated (we excluded the last four since nearly all the models achieved an allocation score of zero), the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, had the highest rank for allocation score in four weeks, more than any other model except the \texttt{COVIDhub-ensemble} which also had four.
-
-<<eval=FALSE>>=
-## analysis for the above statements
-alloscore_data_summaries |>
-  filter(as.Date(reference_date) < as.Date("2022-02-01"), allo_rank == 1)
-@
 
@@
 
-<<eval=FALSE>>=
-## some analysis to support above statements
-score_data_summaries |>
-  group_by(target_end_date) |>
-  summarize(meanWIS = mean(mwis, na.rm=TRUE),
-            medianWIS = median(mwis, na.rm=TRUE),
-            minWIS = min(mwis, na.rm=TRUE),
-            maxWIS = max(mwis, na.rm=TRUE),
-            range = maxWIS-minWIS)
-@
-
 \subsubsection{Metrics were not consistently correlated over time}
 
-<<metrics-correlation, fig.width=10, fig.height=7.5, fig.cap="Association of standardized ranks for MWIS and allocation score by model and week.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\includegraphics[width=\maxwidth]{figure/metrics-correlation-1} \caption[Association of standardized ranks for MWIS and allocation score by model and week]{Association of standardized ranks for MWIS and allocation score by model and week.
 Each facet of the plot corresponds to one model.
@@
 The size of the point corresponds to the observed value on the date for which the prediction was made.
-Models show different degrees of association between the two metrics.">>=
-all_scores |>
-  filter(as.Date(reference_date.x) < as.Date("2022-02-01") ) |>
-  ggplot(aes(x=mwis_rank_std, y=allo_rank_std, color=target_end_date, size=yobs)) +
-  geom_point(alpha=.7) +
-  scale_x_continuous("MWIS standardized rank", limits=c(0,1)) +
-  scale_y_continuous("Allocation score standardized rank", limits=c(0,1)) +
-  facet_wrap(.~model) +
-  geom_abline(slope=1, intercept=0, linetype=2, color="gray")
+Models show different degrees of association between the two metrics.}\label{fig:metrics-correlation}
+\end{figure}
 
-@
+\end{knitrout}
 
@@
 
-<<eval=FALSE>>=
-alloscore_data_summaries |>
-  group_by(model) |>
-  summarize(top_rank = sum(allo_rank_std==1))
-@
-
 \subsubsection{Integrated allocation score across values of K}
@@
 
-<<calc-tot-y>>=
-tot_y <- slim_alloscore_2021.12.20 |>
-  filter(model == "COVIDhub-ensemble", K==200) |>
-  pull(y) |>
-  sum()
-@
-
 AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 (Figure \ref{fig:multi-k}A).
-These calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of \Sexpr{format(tot_y, big.mark = ",")} that day.
+These calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of 19,581 that day.
 Rankings of AS from all models were fairly stable across a range of values for K, with some crossings, especially at $K$ values further away from the observed value.
@@
 
-<<multi-k, fig.height=10, fig.width = 8, fig.cap = "Allocation scores (AS) across different resource constraints (K) for 10 models that made forecasts on 2021-12-20.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\includegraphics[width=\maxwidth]{figure/multi-k-1} \caption[Allocation scores (AS) across different resource constraints (K) for 10 models that made forecasts on 2021-12-20]{Allocation scores (AS) across different resource constraints (K) for 10 models that made forecasts on 2021-12-20.
 Panel A shows, for each model, the AS for values of K between 200 and 60,000 at increments of 200.
@@
 Every point represents the AS and IAS for one model.
-The IAS centered at 15,000 are more closely correlated with the AS values.">>=
+The IAS centered at 15,000 are more closely correlated with the AS values.}\label{fig:multi-k}
+\end{figure}
 
-ias_to_plot <- left_join(ias, as) |>
-  pivot_longer(cols = c(IAS, IAS_wt), names_to = "IAS_type", values_to = "IAS") |>
-  mutate(IAS_type = ifelse(IAS_type=="IAS", "uniform", "centered at 15,000"))
-
-p1 <- ggplot(ias_to_plot, aes(x=AS, y=IAS, color=IAS_type, shape=IAS_type)) +
-  geom_point() +
-  geom_smooth(method="lm", se=FALSE, linetype=2) +
-  scale_color_manual(name = "Integrated Allocation Score", values = c("#386cb0", "#fdc086")) +
-  scale_shape_manual(name = "Integrated Allocation Score", values = c(1, 2)) +
-  geom_abline(slope=1, intercept=0, color="gray") +
-  theme(legend.position = c(0.05,.9), legend.justification = c(0,1),
-        legend.background=element_rect(fill = alpha("white", 0.2)))
-
-p2 <- slim_alloscore_2021.12.20 |>
-  group_by(model, K) |>
-  summarize(`Allocation score` = sum(components)) |>
-  left_join(kprobs) |>
-  ggplot(aes(x=K)) +
-  geom_point(aes(y=weight), color = "#386cb0") +
-  geom_point(y=1/length(unique(slim_alloscore_2021.12.20$K)), color="#fdc086", shape = 2) +
-  geom_vline(xintercept = 15000, color="gray")
-
-p3 <- slim_alloscore_2021.12.20 |>
-  group_by(model, K) |>
-  summarize(`Allocation score` = sum(components)) |>
-  mutate(model = factor(model, levels = rev(ias$model))) |>
-  ggplot(aes(x=K, y=`Allocation score`, color=model, linetype=model)) +
-  geom_line() +
-  theme(legend.position = c(0.01,.99), legend.justification = c(0,1),
-        legend.background=element_rect(fill = alpha("white", 0.2)),
-        legend.title = element_blank(), legend.text = element_text(size=7))
-
-cowplot::plot_grid(p3, p2, p1,
-                   nrow=3,
-                   align = "h",
-                   labels = "AUTO")
-@
+\end{knitrout}
 

### 2023-12-20 alloscore_manuscript/alloscore-application-shorter.Rnw
Assorted main text (shorter) edits

@@
 \usepackage{xurl}
-\externaldocument{supplement}
+\externaldocument{supplement_short}
 
@@
 We begin with an informal description of the allocation score and some examples illustrating its key characteristics in section \ref{sec:methods.overview}.
-In section \ref{sec:methods.detailed} we develop the allocation score more carefully, building on decision theoretic procedures for deriving proper scoring rules.
+In section \ref{sec:methods.detailed} we develop the allocation score more carefully, using a basic decision theoretic procedure for deriving proper scoring rules.
+(See section \ref{sec:proper} of the supplement for a definition of a proper scoring rule
+and a more technical discussion of the procedure).
 In section \ref{sec:methods.related}, we note that another group of common scores including the quantile score, WIS, and CRPS, can also be derived from decision theoretic foundations \textemdash starting from a different decision making context.
@@
 If the decision maker is provided with a multivariate forecast $F$ where each marginal forecast distribution $F_i$ predicts resource need in a particular location, one option is to choose the resource allocation that minimizes the expected total unmet need according to the forecast.
-We will give a more precise mathematical statement in section \ref{sec:methods.detailed}, but informally, the total expected unmet need according to the forecast is
+We will give a more precise mathematical statement in section \ref{sec:methods.detailed}, but informally, the total expected unmet need according to the forecast is{}
 \begin{align}
@@
 This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss function.
-Subject to certain technical conditions, scoring rules obtained from this procedure are proper.
-We refer the reader to the supplement for a more technically precise discussion.
+We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} and in section \ref{sec:proper} of the
+supplement we show that they are proper by construction.
 
@@
 
-To make this a more easily interpreted measure of forecast performance, we will adjust the raw score by subtracting the minimum loss $l_{\text{oracle}}(y;K)$ achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$.
-When the oracle has sufficient resources to meet the total need, i.e., when $\sum_{i=1}^{N}y_i \leq K$, the oracle's loss is zero and there is no adjustment.  Thus the allocation score coincides with
-the raw score in that case.
+To make this a more easily interpreted measure of forecast performance, we will adjust the raw score by subtracting the minimum loss
+achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$.
+When the oracle has sufficient resources to meet the total need, i.e., when $\sum_{i=1}^{N}y_i \leq K$, the oracle's loss is zero 
+and allocation score coincides with the raw score.
 On the other hand, when the oracle cannot cover all need and incurs a loss of $L \cdot \left(\sum_{i=1}^{N}y_i - K \right) > 0$, we adjust
 the raw score by this loss.
-The oracle-adjusted score is
+The oracle-adjusted score can therefore be written as
 \begin{align}
-S_A(F, y; K)  &= S_A^{\text{raw}}(F, y; K) - l_{\text{oracle}}(y;K) \\
-&= \begin{cases}
-S_A^{\text{raw}}(F, y; K), &  \sum_{i=1}^{N}y_i \leq K \\
-S_A^{\text{raw}}(F, y; K) - L \cdot \left( \sum_{i=1}^{N}y_i - K\right), &  \sum_{i=1}^{N}y_i > K.
-\end{cases}
+S_A(F, y; K)  &= S_A^{\text{raw}}(F, y; K) - L \cdot \max\left(0,\sum_{i=1}^{N}y_i - K\right) \\
+&= L\left\{\sum_{i=1}^N \max(0, y_i - x_i^{F,K}) -  \max\left(0,\sum_{i=1}^{N}y_i - K\right)\right\}
 \end{align}
@@
 In epidemiological forecasting, well-known proper scoring rules such as the log score or variations on the continuous ranked probability score (such as the weighted interval score, or WIS) have been frequently utilized to evaluate probabilistic forecasts.
-Often, these scores are used to rank models by accuracy according to that particular score, but without reference to the underlying decision-making process for which that score was designed.
+Often models are ranked with respect to accuracy according to a particular score, but without reference to any underlying decision-making process for which that score was designed or from which it might be derivable as a Bayes scoring rule.
 With careful thought and collaboration between modelers and public health officials, we argue that scores that are more aligned with public health decisions could be developed to inform specific problems.
-We have demonstrated that forecast evaluation methods that are tied to a specific decision making context can yield model rankings that are substantively different from standard measures of forecast skill.
+We have demonstrated that forecast evaluation methods that are tied to a specific decision making context can yield model rankings that differ substantively from those based on standard measures of forecast skill.
 
@@
 In practice, there are many potential users of forecasts with many different decision making problems.
-Not all can be easily quantified.
-Those that can be easily quantified may differ enough that no single score would be appropriate for all users.
+Not all can be easily modeled as an effort to minimize the expectation of a transparent loss function.
+And even those that are easily framed as expected loss minimization may differ enough that no single score would be 
+appropriate for all users.
 Ideally, targeted forecasting tools could be developed through close collaboration between modelers and public health officials.
@@
 We also note that in some settings, a ``successful'' epidemiological forecast may lead to policy decisions that change the distribution of the predicted outcome $Y$.
-Our framework cannot be used to evaluate a forecast at a horizon where the outcome may have been impacted by such a decision.
+Our framework would need considerable enhancements before being applicable to forecast evaluation beyond horizons
+for which causal feedback can be neglected.
 
@@
 In conclusion, we argue that the way modelers and policymakers view and evaluate forecasts should change depending on the specific decision-making context.
-Using standard forecast evaluation metrics can mask the utility of certain forecasts, or lead to forecasts being used to inform a particular decision that are not well matched to that decision making context.
+Defaulting to standard forecast evaluation metrics can mask the utility (or disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those for which they were intended.
 New collaborative work between public health officials and modeling teams is needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or simulation exercises that could be used to inform further development of new or alternative scoring metrics.

### 2023-12-20 alloscore_manuscript/alloscore-application-shorter.tex
Assorted main text (shorter) edits

@@
 \paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$ and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$.
+The means of these distributions are given by the scale parameters $\sigma_i$.
 When the marginal forecasts are exponential distributions, it can be shown that the optimal allocation divides the available resources among the locations proportionally to the scale parameters $\sigma_i$ (see section
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Tue Dec 19 17:26:53 2023
+% Wed Dec 20 04:09:40 2023
 \begin{table}[ht]
@@
   \hline
-model & AS & MWIS & IAS centered at 15k & IAS uniform \\ 
+Model & AS & MWIS & IAS centered at 15k & IAS uniform \\ 
   \hline
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/thermometer-plot-1} \caption[Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the nine highest hospitalization counts]{Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the nine highest hospitalization counts.
+\includegraphics[width=\maxwidth]{figure/thermometer-plot-1} \caption[Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the ten highest hospitalization counts]{Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the ten highest hospitalization counts.
 For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts.

### 2023-12-20 alloscore_manuscript/supplement_short.tex
Assorted main text (shorter) edits

\documentclass{article}

\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

\usepackage{amsmath, amsfonts, amssymb, mathtools}
\usepackage{accents}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{cases}
\usepackage{caption}
\usepackage{soul}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{float}
\usepackage{xr-hyper}
\usepackage{hyperref}
\externaldocument{alloscore-application-shorter}

\usepackage{enumitem}
\newlist{todolist}{itemize}{2}
\setlist[todolist]{label=$\square$}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
\hspace{-2.5pt}}
\newcommand{\wontfix}{\rlap{$\square$}{\large\hspace{1pt}\xmark}}

\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator{\short}{sh}
\DeclareMathOperator{\Ex}{\mathbb{E}}

\usepackage{setspace}

\usepackage{parskip}

\usepackage{soul}
\usepackage{xcolor}
\def\elr#1{{\color{cyan}\textbf{ELR:[#1]}}}
\def\apg#1{{\color{red}\textbf{APG:[#1]}}}
\def\bwr#1{{\color{violet}\textbf{BWR:[#1]}}}
\def\ngr#1{{\color{blue}\textbf{NGR:[#1]}}}

\usepackage{natbib}
\bibliographystyle{unsrtnat}

\title{Supplementary Material for ``Evaluating infectious disease forecasts with allocation scoring rules''}
\author{Aaron Gerding, Nicholas G. Reich, Benjamin Rogers, Evan L. Ray}

\begin{document}

\newcommand{\del}[2]{\frac{\partial {#1} }{\partial {#2}} }
\newcommand{\dby}[2]{\frac{d {#1} }{d {#2}} }
\newcommand{\sbar}{\overline{s}}

\newtheorem{proposition}{Proposition}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\maketitle

\begin{todolist}
\item 2 actual references to supplement (re exponential quantiles and solving the allocation problem)
\item the statement that "we computed the allocation score for the 14 day-ahead forecast" which needs some answer to "how?"
\item use of "proper" a number of times
\end{todolist}

\begin{abstract}
We briefly address some technical and methodological points in the main text, referring to the forthcoming ... for 
more thorough discussion.
\end{abstract}

\section{Introduction}
\label{sec:intro}

We briefly address some technical and methodological points in the main text. We begin in section \ref{sec:shortage} by formalizing the concept of a \emph{shortage} of resources and giving some key results about expected resource shortages under a distribution characterizing uncertainty about (future) levels of resource need. Resource shortages play a central role in the decision-making problems that give rise to the quantile loss and the allocation score, which we discuss in sections \ref{sec:quantiles_shortage} and \ref{sec:bayes-quantiles} respectively. Section \ref{sec:numeric} gives details on the numerical methods that we use to calculate allocation scores, including some special considerations for settings where forecasts are represented by a finite collection of predictive quantiles, such as the application to forecasts of hospitalizations due to COVID-19 in section 3 of the article.

\section{Proper scoring rules}
\label{sec:proper}

In decision theory, a loss function $l$ is used to formalize a decision problem by assigning numerical value $l(x,y)$ to the 
\emph{result} of taking an \emph{action} $x$ in preparation for an \emph{outcome} $y$. A \emph{scoring rule} $S$ is a 
loss function for a decision problem where the action is a probabilistic forecast $F$ of the outcome $y$ (or the statement of $F$ by a forecaster). 
% As does any loss fun $S$ transforms a random outcome variable $Y$ into a random loss $S(F,Y)$. 
We refer to the realized loss $S(F,y)$ as the \emph{score} of $F$ at $y$.

Decision theoretically, probabilistic forecasts are a unique kind of action in that they can be used to generate their
own (simulated) outcome data, against which they can be scored using $S$. A probabilistic forecast
$F$ is thus committed to the ``self-assessment'' $\Ex_F [S(F, Y)] := \Ex [S(F, Y^F)]$, where $Y^F \sim F$ is the random variable defined 
by sampling from $F$, as well to an assessment $\Ex_F [S(G, Y)]$ of any alternative forecast $G$.

A natural consistency criterion for $S$ is that it does not commit $F$ to assessing any other forecast $G$ 
as being better than $F$ itself, that is, that
\begin{align}
\Ex_F [S(F, Y)] \leq \Ex_F [S(G, Y)] \label{eqn:prop_ineq}
\end{align} 
for any $F,G$. A scoring rule meeting this criterion is called \emph{proper}. If $S$ were improper, then from the perspective of 
a forecaster focussed (solely) on expected loss minimization (which we will call an ELM forecaster), the decision to state 
a forecast $G$ other than the forecast $F$ which they believe describes $Y$ could be superior to the decision to state $F$. 
$S$ is \emph{strictly proper} when 
\eqref{eqn:prop_ineq} is sharp, in which case the 
\emph{only} optimal decision for an ELM forecaster is to state the forecast they believe to be true.

\subsection{The allocation score is proper}
\label{sec:alloscore_proper}

Our primary decision theoretical procedure, outlined in section \ref{sec:methods.detailed.decisiontheory} of the main text, 
uses a decision problem with loss function $s(x,y)$ to define a scoring rule 
\begin{align}
S(F,y) := s(x^F,y) \label{eqn:bayes_sr}
\end{align}
where $x^F := \argmin_{x} \Ex_F[s(x,Y)]$ is the Bayes act for $F$ with respect to $s$. 
Such scoring rules, which we call \emph{Bayes scoring rules}, 
are proper by construction since
\begin{align}
\Ex_F [S(F, Y)] &= \Ex_F [ s(x^F, Y) ] \nonumber \\
 &= \mathrm{min}_{x} \Ex_F [ s(x, Y) ] \quad \text{ (by definition of $x^F$)} \\
 &\leq \Ex_F [ s(x^G, Y) ] \label{eqn:dt_proper_key} \\
 &= \Ex_F [ S(G, Y)]. \nonumber
\end{align}

The allocation scoring rule is Bayes and therefore proper.

We note that in the probabilistic forecasting literature (see e.g., \cite{gneiting2011making}, Theorem 3) what we have 
termed Bayes scoring rules typically appear via \eqref{eqn:bayes_sr} where $x^F$ is some given functional of $F$ which 
can be shown to be \emph{elicitable}, that is, to be the Bayes act for some loss function $s$.
Such a loss function is said to be a \emph{consistent loss (or scoring) function} for the functional $F \mapsto x^F$, and many important
recent results in the literature (e.g., \cite{fisslerziegel2016consistency}) address whether there \emph{exists} any loss 
function for $x^F$ which is consistent. Our orientation
is different from this insofar as we \emph{begin} by specifying a decision problem and a loss function of subject matter relevance
and use the Bayes act only as a bridge to a proper scoring rule.  Consistency is never in doubt.

\section{Expected shortages}
\label{sec:ex-shortage}

A key feature of loss functions used in the decision theoretic definition of quantiles and related scoring rules such as the 
CRPS and the WIS (see ...), as well as the allocation loss function presented in this work, is the presence of a \emph{shortage}: 
the amount $\max\{0,y-x\}$ by which a random resource demand 
$y$ exceeds a supply decision variable $x$, which, for convenience, we write as $(y-x)_{+}$. In particular, a quantile at 
probability level $\alpha$ of a distribution $F$ on $\mathbb{R}^1$ (which we assume to have a well-defined density $f(x)$) 
is a Bayes act for the loss function
\[
l(x,y) = Cx + L(y-x)_{+}
\] 
where $\alpha = 1-C/L$ and $C$ and $L$ can be interpreted as the cost per unit of a resource (such as medicine) and the loss
incurred when a unit of demand (such as illness) cannot be met due to the shortage $(y-x)_{+}$.  This follows because a 
Bayes act, as a minimizer of $\Ex_F[l(x,Y)]$, must also be a vanishing point of the derivative
\begin{align}
\dby{}{x} \Ex_F\left[l(x,Y)\right] &= \Ex_F\left[\dby{}{x}l(x,Y)\right] \nonumber\\
&= C + L\Ex_F\left[\dby{}{x}(Y-x)_+\right] \nonumber\\
&= C - L\Ex_F\left[\mathbf{1}\{Y > x\}\right] \nonumber\\
&= C + L(F(x) - 1), \label{eqn:q_deriv}
\end{align}
so that $1-C/L = F(x)$.
The formula $\dby{}{x}\Ex_F\left[(Y-x)_+\right] = F(x) - 1$ for the derivative of the shortage 
which we use below in deriving the Bayes act for the allocation
loss, can be seen as following, as indicated above in \eqref{eqn:q_deriv}, from a probability theoretic definition of the expectation operator on piecewise defined functions, or as application of the ``Leibniz Rule''
\begin{align}
	\frac{d}{dx} \Ex_F [(Y-x)_{+}] &= \frac{d}{dx} \int_{x}^{\infty} (y-x) f_Y(y)dy \nonumber\\
	&= \int_{x}^{\infty} \frac{d}{dx}(y-x) f_Y(y)dy - (x-x) f_Y(x) = -\int_{x}^{\infty} f_Y(y)dy = F(x)-1. \label{eqn:shortage_deriv}
\end{align}
(Note that more care is required when $F$ does not have a density.)

\section{Allocation Bayes acts as vectors of marginal quantiles.}
\label{sec:bayes-quantiles}

Here we show that the Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$, corresponding to the
allocation problem (AP) (\eqref{eqn:loss_fn} in section \ref{sec:methods.detailed.specific_allocation})
\begin{align}
    \underset{0 \leq x}{\mathrm{minimize}}\,\, \mathbb{E}_{F} [s_A(x, Y)]= \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[(Y_i - x_i)_{+}] 
     \text{ subject to }
     \, \sum_{i=1}^N x_i = K, \label{AP}
\end{align}
can be represented as a vector of quantiles for the marginal forecast distributions $F_i$ at a single probability level
$\tau^{F,K}$, that is, $x_i^{F,K} = q_{F_i,\tau^{F,K}}$. An immediate consequence used in the examples in Section \ref
{sec:methods.overview} in the main text is that if $F_i = \mathrm{Exp}(1/\sigma_i)$ for all $i$, then the Bayes act is
proportional to $(\sigma_1,\ldots,\sigma_N)$, since $q_{\mathrm{Exp}(1/\sigma),\tau} = -\sigma \log(1-\tau)$.

In order for $x^{\star} \in \mathbb{R}^N_{+}$ to solve the AP it must be true that reallocatating $\delta > 0$ of the 
$x_i^{\star}$ units of resource allocated to location $i$ to location $j$ will increase the expected shortage in location
$i$ by at least as much as it decreases the expected shortage in location $j$. Letting $\delta \searrow 0$, this implies from
\eqref{eqn:shortage_deriv} that

\begin{align}
1-F_i(x^{\star}_i) &= -\frac{d}{dx_i}\mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}] \nonumber \\
&= \lim_{\delta \searrow 0} \frac{1}{\delta}
\left\{\mathbb{E}_{F_i}[(Y_i - (x^{\star}_i - \delta))_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}]\right\} 
\text{ (increase in $i$) } \nonumber \\
&\geq 
\lim_{\delta \searrow 0} \frac{1}{\delta} 
\left\{\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - (x^{\star}_j + \delta))_{+}]\right\} 
\text{ (decrease in $j$) } \nonumber \\
 &= -\dby{}{x_j}\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] = 1-F_j(x^{\star}_j) \label{eqn:ASoptimal1}
\end{align}

Note that negative derivatives appear because our optimality condition addresses how a \emph{decrease} in resources will 
\emph{increase} the expected shortage in $i$ and vice versa in $j$. Since \eqref{eqn:ASoptimal1} holds with $i$ and $j$ 
reversed, a number $\lambda$ (a \emph{Lagrange multiplier}) exists such that
$L(1-F_k(x^{\star}_k)) = \lambda$ for all $k \in 1,\ldots,N$.
(We scale by $L$ to facilitate possible future interpretations of $\lambda$ in terms of the partial derivatives 
of $\mathbb{E}_{F} [s_A(x, Y)]$.)
That is, $x^{\star}_k$ is a quantile $q_{\tau,F_k}$ for
$\tau = 1 - \lambda/L$. The value of $\tau$ is then determined by the constraint equation 
\begin{align}
\sum_{i=1}^N q_{\tau,F_i} = K.
\end{align}
It is important to note that $\tau$ depends on $F$ and $K$ and is \emph{not} a fixed parameter
of the allocation scoring rule.
\newpage

\section{Propriety of parametric approximation} % (fold)
\label{sec:propriety_of_parametric_approximation}

For convenience we write $S(F,G) := \Ex_G[S(F,Y)]$. Let
\begin{align*}
q_{23}(F) &:= (q_{.01,F},\ldots,q_{.99,F}) \\
G^q(F) &:= G_{q_{23}(F)} \in \mathcal{P}_{\mathrm{dfq}} \\
G^{\star}(F) &:= \argmin_{G \in \mathcal{P}_{\mathrm{dfq}}} S_A(G,F)
\end{align*}
We have the ``right'' approximate allocation score
\begin{align*}
S_{R}(F,y):= S_A(G^{\star}(F),y)
\end{align*}
which is proper because 
\begin{align*}
S_R(F,F) &= S_A(G^{\star}(F), F) \\ 
&\leq S_A(G^{\star}(F), H) \quad \text{ (by definition of $G^{\star}$)} \\
&=S_R(F,H).
\end{align*}
This seems to almost be what is called ``properization'' in \cite{brehmer2020properization}.
We also have a ``wrong'' approximation 
\begin{align*}
\tilde{S}_R(F,y) := S_A(G^q(F),y)
\end{align*}
which is improper because
\begin{align*}
\tilde{S}_R(F,G^q(F)) &= S_A(G^q(F),G^q(F)) \\
 &<  S_A(G^q(F),F) \quad \text{ (because $S_A$ is proper)} \\
 &= \tilde{S}_R(F,F).
\end{align*}

% section propriety_of_parametric_approximation (end)

\bibliography{allocation}

\end{document}

### 2023-12-20 alloscore_manuscript/alloscore-application-shorter.Rnw
very minor wording updates

@@
 We begin with an informal description of the allocation score and some examples illustrating its key characteristics in section \ref{sec:methods.overview}.
-In section \ref{sec:methods.detailed} we develop the allocation score more carefully, using a basic decision theoretic procedure for deriving proper scoring rules.
+In section \ref{sec:methods.detailed} we develop the allocation score more carefully, using a decision theoretic procedure for deriving proper scoring rules.
 (See section \ref{sec:proper} of the supplement for a definition of a proper scoring rule
@@
 We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} and in section \ref{sec:proper} of the
-supplement we show that they are proper by construction.
+supplement we show that Bayes scoring rules are proper by construction.
 
@@
 The WIS can be directly calculated from these quantiles.
-However, calculation of the AS requires that the full cumulative distribution functions of the forecast distribution are available for each location.
+However, calculation of the AS requires that the full cumulative distribution functions of the forecast distributions for each location are available.
 For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution functions based on the provided quantiles.
@@
 To illustrate the mechanics of allocation scoring, we start by focusing on how forecasts generated on or before December 20, 2021, with predictions for January 03, 2022, were scored by different metrics.
-This week was around the Omicon wave peak nationally, with individual states typically observing a peak at or after January 3, 2022.
+This week was around the peak of the Omicon wave nationally, with individual states typically observing a peak at or after January 3, 2022.
 

### 2023-12-20 alloscore_manuscript/alloscore-application-shorter.tex
very minor wording updates

@@
 
-% latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Wed Dec 20 04:09:40 2023
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Wed Dec 20 11:50:37 2023
 \begin{table}[ht]

### 2023-12-20 alloscore_manuscript/alloscore-application-shorter.Rnw
remove discussion of performance relative to baseline that i didn't find convincing - open to revisiting

@@
 
-Overall, across the first 11 weeks evaluated (we excluded the last four since nearly all the models achieved an allocation score of zero), the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, had the highest rank for allocation score in four weeks, more than any other model except the \texttt{COVIDhub-ensemble} which also had four.
+%Overall, across the first 11 weeks evaluated (we excluded the last four since nearly all the models achieved an allocation score of zero), the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, had the highest rank for allocation score in four weeks, more than any other model except the \texttt{COVIDhub-ensemble} which also had four.
 
@@
 We have demonstrated that forecast evaluation methods that are tied to a specific decision making context can yield model rankings that differ substantively from those based on standard measures of forecast skill.
+In particular, the allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.
 
@@
 
-In our specific application, we made several key observations that we believe should inform future work in this area.
-First, it is clear that different metrics (in our case allocation score and mean weighted interval score) captured different aspects of forecast performance.
-The allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.
+%In our specific application, we made several key observations that we believe should inform future work in this area.
+%First, it is clear that different metrics (in our case allocation score and mean weighted interval score) captured different aspects of forecast performance.
+%The allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.
 %A forecast with a good allocation score can have a poor weighted interval score and vice versa.
@@
 %The allocation score was not as scale-dependent, as the highest/worst allocation scores were observed when the number of hospitalizations was closest to the allocation constraint.
-We also note that contributed models found it hard to achieve a better allocation score than a na\"ive baseline model that just predicted a flat line from the last previous observation with wide uncertainty.
-This suggests suggests that contributed models (other than the ensemble, which combined forecasts from all contributed models) were not consistently adding value to allocation decisions over just extrapolating from the current levels.
+%We also note that contributed models found it hard to achieve a better allocation score than a na\"ive baseline model that just predicted a flat line from the last previous observation with wide uncertainty.
+%This suggests suggests that contributed models (other than the ensemble, which combined forecasts from all contributed models) were not consistently adding value to allocation decisions over just extrapolating from the current levels.
 

### 2023-12-20 alloscore_manuscript/alloscore-application-shorter.tex
remove discussion of performance relative to baseline that i didn't find convincing - open to revisiting

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Dec 20 11:50:37 2023
+% Wed Dec 20 11:56:48 2023
 \begin{table}[ht]

### 2023-12-20 alloscore_manuscript/supplement_short.tex
add all the stuff back in...

@@
 \maketitle
+
+\tableofcontents
 
@@
 
+\section{Numerical computation of allocation Bayes acts}
+\label{sec:numeric}
+
+In practice, \eqref{eqn:quantiles-sum-to-K} is rarely amenable to closed form solution.
+Using the Bayes act in explicit computations therefore requires a method of generating an approximation $\tilde{\tau}$ to 
+a solution $\tau^{\star}$ of \eqref{eqn:tau-inclusion}.  One would also strive to have some control over a robust measure of 
+of the approximation error of the random shortages which we aim to compute using the Bayes act. One possible measure is the sum
+of absolute coordinate errors
+\begin{align}
+\mathrm{SAE}(Y,\tilde{\tau}) = \sum_{i=1}^{N} \left| (Y_i-q_{\tilde{\tau},F_i})_+ - (Y_i-x_i^{F,K})_+ \right|
+\end{align}
+% of the Bayes risk error or ``optimality gap''
+% \begin{align}
+% \left|\mathbb{E}_F\left[s_A(\bm{q}_{\tilde{\tau},F}, Y)\right] - \mathbb{E}_F[s_A(x^{F,K}, Y)]\right| = 
+% L\left| \sum_{i=1}^{N}\mathbb{E}_{F_i}\left[(Y_i-q_{\tilde{\tau},F_i})_+ - (Y_i-x_i^{F,K})_+\right]\right|
+% \end{align}
+where $\bm{q}_{\tilde{\tau},F}$ solves the linear program
+$\sum_{x \in \undertilde{\bm{F}}^{-1}(\tilde{\tau})}x_i = K$ approximating the true Bayes act defining linear program
+$\sum_{x \in \undertilde{\bm{F}}^{-1}(\tau^{\star})}x_i = K$. In this section we outline such a method we have implemented 
+in the \verb`R` package \verb`alloscore`.
+
+Suppose we have established that $\tau^{\star} = F_T(K)$ lies in the interval $I_1 = [\tau_L, \tau_U]$ with $\tau_L < \tau_U$,
+that is, $K \in [q^{-}_{F_T,\tau_L}, q^{+}_{F_T,\tau_U}]$.
+From section \ref{sec:quantile-functions}, we know that the set 
+$TQ_F(\tau_L) \cup TQ_F(\tau_U) \subset [q^{-}_{F_T,\tau_L}, q^{+}_{F_T,\tau_U}]$
+is arranged in exactly one of the following ways (where e.g. refers to figure \ref{fig:FQ}): 
+\begin{itemize}
+\item[($\bullet \bullet$)] $TQ_F(\tau_L) \cap TQ_F(\tau_U) = \varnothing$ 
+(and $q^{+}_{F_T,\tau_L} < q^{-}_{F_T,\tau_U}$, e.g., $\tau_L=\alpha_1, \tau_U=\alpha_5$)
+\item[($\bullet$)] $TQ_F(\tau_L) = \{K\} = TQ_F(\tau_U)$ 
+(a point mass at $K$, e.g., $\alpha_3 \leq \tau_L \leq \tau_U \leq \alpha_5$)
+\item[($\bullet\!-$)] $TQ_F(\tau_L) = \{q^{-}_{F_T,\tau_U}\} \subsetneq TQ_F(\tau_U)$ 
+(a point mass at $q^{-}_{F_T,\tau_U}$, e.g., $\tau_L = \alpha_3, \tau_U = \alpha_5$)
+\item[($-\!\bullet$)] $TQ_F(\tau_L) \supsetneq \{q^{+}_{F_T,\tau_L}\} = TQ_F(\tau_U)$ 
+(a point mass at $q^{+}_{F_T,\tau_U}$, e.g., e.g., $\tau_L = \alpha_5, \tau_U = \alpha_4$).
+\end{itemize}
+
+In the case ($\bullet$), we can immediately take 
+$\tau^{\star}=\tau_U$ as the probability level representing the allocation Bayes act.
+In the cases
+($\bullet\!-$), and ($-\!\bullet$), which imply the presence of a point mass in one or more of the 
+component forecasts adjacent to a region of zero density in all components, we can take $\tau^{\star}=\tau_U$ or $\tau_L$, 
+respectively, as the representing probability level.  
+Having found our $\tau^{\star}$ we then arrive at a Bayes act $x^{F,K}$ by by solving the
+linear program $\sum_{x \in \undertilde{\bm{F}}^{-1}(\tau^{\star})}x_i = K$.
+
+In the remaining and typical case of ($\bullet \bullet$), further search is generally necessary. We can proceed 
+by evaluating $TQ_F$ at $\tau_M = \frac{1}{2}\left(\tau_L + \tau_U\right)$ and replacing $I_1$ with one of its 
+halves
+\begin{align}
+I_2 = 
+\begin{cases}
+[\tau_L, \tau_M] & \text{ if } K < q^{-}_{F_T,\tau_M} \\
+[\tau_M, \tau_U] & \text{ if } K \geq q^{-}_{F_T,\tau_M}
+\end{cases}
+\end{align}
+which also contains $\tau^{\star} = F_T(K)$. This follows from the definition $q^{-}_{F_T,\tau_M} := \min\{x \mid F_T(x) \geq \tau_M\}$, 
+according to which
+\begin{align}
+\begin{cases}
+K < q^{-}_{F_T,\tau_M} \text{ implies } \tau^{\star} = F_T(K) < \tau_M  \\
+K \geq q^{-}_{F_T,\tau_M} \text{ implies } \tau^{\star} = F_T(K) \geq \tau_M.
+\end{cases}
+\end{align}
+Note that $K \leq q^{-}_{F_T,\tau_M}$ would not imply $F_T(K) \leq \tau_M$ due to the possibility of a point mass at $K$.
+
+Iterating this process we obtain a sequence $\{I_k\}, k=1,2,\ldots$ of intervals of widths 
+$\left|I_k \right| = 2^{1-k}\left|I_1 \right|$ which either terminates at one of the scenarios 
+($\bullet$), ($\bullet\!-$), or ($-\!\bullet$), or provides infinite sequences $\{\tau_{L,k}\}$ and $\{\tau_{U,k}\}$ converging
+to $\tau^{\star}$ from below and above. Such a sequence provides the basis of a ``bisection'' algorithm for finding
+the ``root'' $\tau^{\star}$ of the set condition $0 \in TQ_F(\tau) - K$.
+
+In the generic case of an infinite $\{I_k\}$, we need to define practical stopping criteria for the possible limit behaviours of 
+$F_T(K \pm \varepsilon)$ as $\varepsilon \searrow 0$ which are exemplified in figure \ref{fig:FQ} at $0, x_1, x_2, x_3, (x_3 +x_4)/2$ and 
+$x_5$.
+% \begin{itemize}
+% \item[($\mathrlap{\,\bullet}{\,/}\,$)] $F_T(K - \varepsilon) < F_T(K-) = F_T(K) < F_T(K + \varepsilon)$
+% (so that $\tau^{\star} = F_T(K-)$ and $TQ_F(\tau^{\star}) = \{K\}$)
+% \item[($/\!\!\raisebox{.9ex}{$\bullet$\!-}$)] $F_T(K - \varepsilon) < F_T(K-) = F_T(K) = F_T(K + \varepsilon)$
+% \item[($\raisebox{-.9ex}{-\!$\bullet$}\!\!/$)] $F_T(K - \varepsilon) = F_T(K) < F_T(K + \varepsilon)$
+% \item[($\raisebox{-.9ex}{\phantom{-}\!$\bullet$}\!\!/$)] $F_T(K - \varepsilon) \leq F_T(K-) < F_T(K) < F_T(K + \varepsilon)$
+% \item[($\,\bullet\!\text{-}$)] $F_T(K - \varepsilon) \leq F_T(K-) < F_T(K) = F_T(K + \varepsilon)$
+% \item[($\text{-}\!\!\bullet\!\!\text{-}$)]  $F_T(K - \varepsilon) = F_T(K) = F_T(K + \varepsilon)$
+% so that $K \in (q^{-}_{F_T,\tau^{\star}}, q^{+}_{F_T,\tau^{\star}})$ and all component forecasts $F_i$ have 
+% zero density at points $x_i \in (q^{-}_{F_i,\tau^{\star}}, q^{+}_{F_i,\tau^{\star}})$ and $\sum x_i = K$.
+% \end{itemize}
+
+And after having obtained sufficiently a sufficient narrow terminal $I_K = [\tau_{L,K}, \tau_{U,K}]$ we need a protocol for 
+how to select a solution to the linear program $\sum_{x \in \undertilde{\bm{F}}^{-1}(I_K)}x_i = K$ as an approximate Bayes act.
+
+\section{Computing allocations from finite quantile forecast representations}
+
+In section 3 of the manuscript, we used the allocation score to evaluate forecasts of COVID-19 hospitalizations that have been submitted to the COVID-19 Forecast Hub.
+These forecasts are submitted to the Hub using a set of 23 quantiles of the forecast distribution at the 23 probability levels in the set $\mathcal{T} = \{0.01, 0.025, 0.05, 0.1, 0.15, \ldots, \allowbreak 0.9, 0.95, 0.975, 0.99\}$, which specify a predictive median and the endpoints of central $(1 - \alpha) \times 100\%$ prediction intervals at levels $\alpha = 0.02, 0.05, 0.1, 0.2, \allowbreak 0.3, 0.4, 0.5, 0.6, \allowbreak 0.7, 0.8, 0.9$.
+For a given week and target date, we use $q_{i,k}$ to denote the submitted quantiles for location $i$ and probability level $\tau_k \in \mathcal{T}$, $k = 1, \ldots, 23$.
+
+In the event that there is some $k \in \{1, \ldots, 23\}$ for which $\sum_i q_{i,k} = K$, i.e., the provided predictive quantiles at level $\tau_k$ sum across locations to the resource constraint $K$, the solution to the allocation problem is given by those quantiles.
+However, generally this will not be the case; the optimal allocation will typically be at some probability level $\tau^\star \notin \mathcal{T}$.
+
+To address this situation and support the numerical allocation algorithm outlined in section \ref{sec:numeric}, we need a mechanism to approximate the full cumulative distribution functions $F_i$, $i = 1, \ldots, N$ based on the provided quantiles.
+We have developed functionality for this purpose in the \verb`distfromq` package for R (cite).
+This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:
+\begin{enumerate}
+  \item Identify a discrete component of the distribution consisting of zero or more point masses, and create an adjusted set of predictive quantiles for the continuous part of the distribution by subtracting the point mass probabilities and rescaling.
+  \item For the continuous part of the distribution, different approaches are used on the interior and exterior of the provided quantiles:
+  \begin{enumerate}
+    \item On the interior, a monotonic cubic spline interpolates the adjusted quantiles representing the continuous part of the distribution.
+    \item A location-scale parametric family is used to extrapolate beyond the provided quantiles.
+The location and scale parameters are estimated separately for the lower and upper tails so as to obtain a tail distribution that matches the two most extreme quantiles in each tail.
+  \end{enumerate}
+\end{enumerate}
+The resulting distributional estimate exactly matches all of the predictive quantiles provided by the forecaster.
+We use the cumulative distribution function resulting from this procedure as an input to the allocation score algorithm.
+
+We refer the reader to the \verb`distfromq` documentation for further detail.
+
 \section{Propriety of parametric approximation} % (fold)
 \label{sec:propriety_of_parametric_approximation}
+
+In practice, open forecasting exercises are generally not able to collect an exact representation of the forecast distribution $F$ other than in simple settings such as for a categorical variable with a relatively small number of categories.
+In settings where the outcome being forecasted is a continuous quantity (such as rates of influenza-like illness among outpatient doctor visits) or a count (such as influenza hospitalizations), forecasting exercises have therefore resorted to collecting summaries of a forecast distribution such as bin probabilities or predictive quantiles.
+This raises the question of whether use of the allocation score still consitutes a proper evaluation procedure if the forecast distribution $F$ is not itself directly recorded.
+The answer we provide in this section is that it is, so long as the fact that the allocation score will be used for forecast evaluation and the method that will be used to obtain a distributional estimate of $F$ from the provided forecast representation are communicated to participating forecasters prospectively.
+
+We consider a setting where a forecasting exercise (such as a forecast hub) pre-specifies that forecasts will be represented using a parametric family of forecast distributions $G_\theta(y)$, and the task of the forecaster is to select a particular parameter value $\theta$.
+We use $\mathcal{P}$ to denote the collection of all distributions $G_\theta$ in the given parametric family.
+For instance, it has recently been proposed that mixture distributions could be used to represent forecast distributions \cite{wadsworth2023mixture}.
+Additionally, we note that the functionality in \verb`distfromq` can be viewed as specifying a parametric family $\mathcal{P}_{\mathrm{dfq}}$ where the parameters $\theta$ of $G_\theta$
+are its quantiles at pre-specified probability levels, and where the shape of any $G_\theta \in \mathcal{P}_{\mathrm{dfq}}$ over the full range of its support is entirely controlled by these quantiles.
 
@@
 
+As another alternative for practical forecasting exercises, a forecast hub could ask forecasters to directly provide the Bayes allocations associated with their forecasts for one or more specified resource constraints $K$.
+At the cost of increasing the number of quantities solicited by the forecast hub, this would have several advantages: it would prevent any artificial distortion of the forecast distributions, allow for direct calculation of scores, and narrow the gap between model outputs and public health end users.
+For this to be feasible, implementations of the allocation algorithm would have to be provided to participating forecasters in the computational languages being used for modeling.
+
 % section propriety_of_parametric_approximation (end)

### 2023-12-21 alloscore_manuscript/supplement_short.tex
some progress on short supp

--- 2023-12-20 alloscore_manuscript/supplement_short.tex
@@
 \begin{align}
-\sum_{i=1}^N q_{\tau,F_i} = K.
+\sum_{i=1}^N q_{\tau,F_i} = K. \label{eqn:quantiles-sum-to-K}
 \end{align}
@@
 of the allocation scoring rule.
-\newpage
+
+\section{Numerical computation of allocation Bayes acts}
+\label{sec:numeric}
+
+To compute an allocation score $S_A(F,y;K) := s_A(x^{F,K},y)$, we require numerical values for a 
+Bayes act solving the AP \eqref{AP}.  Assuming we have reliable means of calculating quantiles $q_{\alpha,F_i}$ 
+of the marginal forecasts $F_i$,  
+such values are given by $q_{\tau^{\star},F_i}$ where $\tau^{\star}$ solves the equation \eqref{eqn:quantiles-sum-to-K}.
+Usually, however, this equation is not analytically tractable and we must resort to a numerical method
+for finding an approximation $\tilde{\tau}$ of $\tau^{\star}$.  This task is made simpler by $\sum_{i=1}^N q_{\tau,F_i}$ being
+an increasing function of $\tau$, and, in principle, a bisection method that evaluates the sign of $\sum_{i=1}^N q_{\tau,F_i} - K$
+at each midpoint of the search interval, narrowing it accordingly, will suffice.  We have implemented such a procedure along with 
+the resulting score computations in the 
+R package \verb`alloscore` which provided all allocation score values used in the analysis of section \ref{sec:application} 
+in the main text.
+
+Awkward subtleties can arise though when the forecast
+densities $f_i$ vanish or are very small, in which case quantiles are non-unique or highly variable at a probability level, 
+leading to ambiguity in how to evaluate $\sum_{i=1}^N q_{\tau,F_i}$. And if point masses are present in any of the $F_i$, 
+\eqref{eqn:quantiles-sum-to-K} will fail to have a unique solution for some discrete set of constraint levels $K$.
+Potentially complex conventions must be adopted for both detecting such levels and enforcing consistency in score calculations
+near them. In both situations, it appears to us that conditions guaranteeing that numerical instabilities and artifacts do not 
+threaten the validity of inferences regarding forecaster performance may be highly dependent on $N$ as well as forecast ranges
+and regularity.  We have established and coded conditions that through extensive experimentation seem appropriate for...
+but we leave a more rigorous approximation error analysis for later work.
+
+\section{Computing allocations from finite quantile forecast representations}
+
+In section 3 of the manuscript, we used the allocation score to evaluate forecasts of COVID-19 hospitalizations that have been submitted to the COVID-19 Forecast Hub.
+These forecasts are submitted to the Hub using a set of 23 quantiles of the forecast distribution at the 23 probability levels in the set $\mathcal{T} = \{0.01, 0.025, 0.05, 0.1, 0.15, \ldots, \allowbreak 0.9, 0.95, 0.975, 0.99\}$, which specify a predictive median and the endpoints of central $(1 - \alpha) \times 100\%$ prediction intervals at levels $\alpha = 0.02, 0.05, 0.1, 0.2, \allowbreak 0.3, 0.4, 0.5, 0.6, \allowbreak 0.7, 0.8, 0.9$.
+For a given week and target date, we use $q_{i,k}$ to denote the submitted quantiles for location $i$ and probability level $\tau_k \in \mathcal{T}$, $k = 1, \ldots, 23$.
+
+In the event that there is some $k \in \{1, \ldots, 23\}$ for which $\sum_i q_{i,k} = K$, i.e., the provided predictive quantiles at level $\tau_k$ sum across locations to the resource constraint $K$, the solution to the allocation problem is given by those quantiles.
+However, generally this will not be the case; the optimal allocation will typically be at some probability level $\tau^\star \notin \mathcal{T}$.
+
+To address this situation and support the numerical allocation algorithm outlined in section \ref{sec:numeric}, we need a mechanism to approximate the full cumulative distribution functions $F_i$, $i = 1, \ldots, N$ based on the provided quantiles.
+We have developed functionality for this purpose in the \verb`distfromq` package for R (cite).
+This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:
+\begin{enumerate}
+  \item Identify a discrete component of the distribution consisting of zero or more point masses, and create an adjusted set of predictive quantiles for the continuous part of the distribution by subtracting the point mass probabilities and rescaling.
+  \item For the continuous part of the distribution, different approaches are used on the interior and exterior of the provided quantiles:
+  \begin{enumerate}
+    \item On the interior, a monotonic cubic spline interpolates the adjusted quantiles representing the continuous part of the distribution.
+    \item A location-scale parametric family is used to extrapolate beyond the provided quantiles.
+The location and scale parameters are estimated separately for the lower and upper tails so as to obtain a tail distribution that matches the two most extreme quantiles in each tail.
+  \end{enumerate}
+\end{enumerate}
+The resulting distributional estimate exactly matches all of the predictive quantiles provided by the forecaster.
+We use the cumulative distribution function resulting from this procedure as an input to the allocation score algorithm.
+
+We refer the reader to the \verb`distfromq` documentation for further detail.
+
+% \begin{enumerate}
+%   \item First, we handle settings where the provided quantiles imply the existence of one or more point masses.
+Specifically, if there are distinct probability levels $\tau_k < \tau_{k^\prime}$ that have the same quantiles, $q_{l,k} = q_{l, k^\prime}$, the distribution contains a point mass at $q_{l,k}$.
+We split into three cases depending on the number of distinct provided quantiles:
+%   \begin{enumerate}
+%     \item If all provided quantiles are equal to each other, we infer that the distribution consists of a single point mass at that quantile value.
+%     \item If there are two distinct quantiles across all of the probability levels $\tau_k \in \mathcal{T}$, our estimated distribution consists of a mixture of two point masses.
+The probability assigned to the point mass at $q$ is proportional to $\tau_q^\max - \tau_q^\min$, where
+%     $$\tau_q^\max = \begin{cases}
+%       \max_{\{\}} \tau_k
+%     \end{cases}$$ is the largest probability level $\tau_k$ for which $q_{l,k} = q$, and $\tau_q^\min$ is defined similarly.
+%     \item Otherwise, our final estimate will consist of a mixture of a discrete distribution with point masses at the duplicated quantiles and a continuous distribution elsewhere.
+In this step, we  In this case, the discrete distribution has point mass probabilities calculated as $d_q = \tau_q^\max - \tau_q^\min$, where $\tau_q^\max$ and $\tau_q^\min$ are defined as in the previous point.
+%   \end{enumerate}
+% \end{enumerate}
 
@@
 \label{sec:propriety_of_parametric_approximation}
+
+\textbf{Bayes scoring rules are invariant under reparametrization of the action space.}
+
+Let $S(F,y)$ be a Bayes scoring rule derived from an action set $\mathcal{X}$ and loss $s(x,y)$.
+Let $\mathcal{W}$ be another set and $M:\mathcal{W} \to \mathcal{X}$ a map. $M$ defines a loss $s_M(\theta, y):= s(M(\theta), y)$
+on $\mathcal{W}$. This in turn defines the Bayes scoring rule
+\begin{align}
+S_M(F,y) := s_M(\theta^F,y) = s(M(\theta^F),y).
+\end{align}
+If $M$ is surjective onto the set of Bayes acts $\{x \in \mathcal{X} \mid x = x^F \text{ for some } F \in \mathcal{P}\}$, then the scores coincide since
+\begin{align}
+S_M(F,y) = s_M(\theta^F,y) & = \min_{\theta}\{s_M(\theta,F)\} \\
+&= \min_{\theta}\{s(M(\theta),F)\}  \\
+&= \min_{x}\{s(x,F)\} = s(x^F,F) = S(F,y).
+\end{align}
+(Note that this does not imply that $M(\theta^F) = x^F$ since $x^F$ might not be a unique Bayes act.)
+
+The set $\mathcal{P}_{\mathrm{dfq}}$ with map $M(G_{\theta}):= x^{G_{\theta}}$ is surjective onto the action set of allocations
+under any constraint $K$ since they can be specified as quantiles for distributions in $\mathcal{P}_{\mathrm{dfq}}$.  Therefore
+the \verb`distfromq` ``reparametrized'' allocation scoring rule coincides with the AS itself.
+
+Unfortunately,
 

### 2023-12-21 alloscore_manuscript/supplement_short.tex
some updates to first few sections of supplement

@@
 
-\begin{abstract}
-We briefly address some technical and methodological points in the main text, referring to the forthcoming ... for 
-more thorough discussion.
-\end{abstract}
-
 \section{Introduction}
 \label{sec:intro}
+
+\elr{This section needs to be updated.}
 
@@
 for any $F,G$. A scoring rule meeting this criterion is called \emph{proper}. If $S$ were improper, then from the perspective of 
-a forecaster focussed (solely) on expected loss minimization (which we will call an ELM forecaster), the decision to state 
+a forecaster focussed (solely) on expected loss minimization, the decision to state 
 a forecast $G$ other than the forecast $F$ which they believe describes $Y$ could be superior to the decision to state $F$. 
 $S$ is \emph{strictly proper} when 
-\eqref{eqn:prop_ineq} is sharp, in which case the 
-\emph{only} optimal decision for an ELM forecaster is to state the forecast they believe to be true.
+\eqref{eqn:prop_ineq} is a strict inequality, in which case the 
+\emph{only} optimal decision for a forecaster seeking to minimize their expected loss is to state the forecast they believe to be true.
 
@@
 recent results in the literature (e.g., \cite{fisslerziegel2016consistency}) address whether there \emph{exists} any loss 
-function for $x^F$ which is consistent. Our orientation
+function that is consistent for $x^F$. Our orientation
 is different from this insofar as we \emph{begin} by specifying a decision problem and a loss function of subject matter relevance
@@
 CRPS and the WIS (see ...), as well as the allocation loss function presented in this work, is the presence of a \emph{shortage}: 
-the amount $\max\{0,y-x\}$ by which a random resource demand 
+the amount $\max\{0,y-x\}$ by which a resource demand 
 $y$ exceeds a supply decision variable $x$, which, for convenience, we write as $(y-x)_{+}$. In particular, a quantile at 
@@
 so that $1-C/L = F(x)$.
-The formula $\dby{}{x}\Ex_F\left[(Y-x)_+\right] = F(x) - 1$ for the derivative of the shortage 
-which we use below in deriving the Bayes act for the allocation
-loss, can be seen as following, as indicated above in \eqref{eqn:q_deriv}, from a probability theoretic definition of the expectation operator on piecewise defined functions, or as application of the ``Leibniz Rule''
+The formula $\dby{}{x}\Ex_F\left[(Y-x)_+\right] = F(x) - 1$ for the derivative of the shortage,
+used above in \eqref{eqn:q_deriv}, can be obtained from an application of the ``Leibniz Rule'':
 \begin{align}
@@
 \end{align}
-(Note that more care is required when $F$ does not have a density.)
+Note that more care is required when $F$ does not have a density.
+We will also use this result below in deriving the Bayes act for the allocation loss.
 
@@
 
-Here we show that the Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$, corresponding to the
-allocation problem (AP) (\eqref{eqn:loss_fn} in section \ref{sec:methods.detailed.specific_allocation})
+Here we study the form of the Bayes act for the allocation problem (AP) (\eqref{eqn:loss_fn} in section \ref{sec:methods.detailed.specific_allocation}) of the text:
 \begin{align}
@@
 \end{align}
-can be represented as a vector of quantiles for the marginal forecast distributions $F_i$ at a single probability level
+We show that the Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$ and resource constraint level $K$
+is a vector of quantiles of the marginal forecast distributions $F_i$ at a single probability level
 $\tau^{F,K}$, that is, $x_i^{F,K} = q_{F_i,\tau^{F,K}}$. An immediate consequence used in the examples in Section \ref
@@
 
-In order for $x^{\star} \in \mathbb{R}^N_{+}$ to solve the AP it must be true that reallocatating $\delta > 0$ of the 
-$x_i^{\star}$ units of resource allocated to location $i$ to location $j$ will increase the expected shortage in location
-$i$ by at least as much as it decreases the expected shortage in location $j$. Letting $\delta \searrow 0$, this implies from
+In order for $x^{\star} \in \mathbb{R}^N_{+}$ to solve the AP it must be true that reallocating $\delta > 0$ units of the resource from location $i$ to location $j$ will lead to a net increase in expected shortage --- in other words, if the reallocation increases the expected shortage in location $i$ is at least as much as it decreases the expected shortage in location $j$:
+\begin{align*}
+&\mathbb{E}_{F_i}[(Y_i - (x^{\star}_i - \delta))_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}]
+\text{ (increase in $i$) } \nonumber \\
+&\qquad \geq 
+\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - (x^{\star}_j + \delta))_{+}]
+\text{ (decrease in $j$) }.
+\end{align*}
+
+Dividing by $\delta$ and letting $\delta \searrow 0$, this implies from
 \eqref{eqn:shortage_deriv} that
-
 \begin{align}
@@
 Note that negative derivatives appear because our optimality condition addresses how a \emph{decrease} in resources will 
-\emph{increase} the expected shortage in $i$ and vice versa in $j$. Since \eqref{eqn:ASoptimal1} holds with $i$ and $j$ 
+\emph{increase} the expected shortage in $i$ and vice versa in $j$. Since \eqref{eqn:ASoptimal1} also holds with $i$ and $j$ 
 reversed, a number $\lambda$ (a \emph{Lagrange multiplier}) exists such that
@@
 To compute an allocation score $S_A(F,y;K) := s_A(x^{F,K},y)$, we require numerical values for a 
-Bayes act solving the AP \eqref{AP}.  Assuming we have reliable means of calculating quantiles $q_{\alpha,F_i}$ 
+Bayes act solving the AP \eqref{AP} --- that is, we must find the specific resource allocations for each location that are determined by the forecast $F$ under the resource constraint $K$.
+Assuming we have reliable means of calculating quantiles $q_{\alpha,F_i}$ 
 of the marginal forecasts $F_i$,  
-such values are given by $q_{\tau^{\star},F_i}$ where $\tau^{\star}$ solves the equation \eqref{eqn:quantiles-sum-to-K}.
-Usually, however, this equation is not analytically tractable and we must resort to a numerical method
-for finding an approximation $\tilde{\tau}$ of $\tau^{\star}$.  This task is made simpler by $\sum_{i=1}^N q_{\tau,F_i}$ being
-an increasing function of $\tau$, and, in principle, a bisection method that evaluates the sign of $\sum_{i=1}^N q_{\tau,F_i} - K$
-at each midpoint of the search interval, narrowing it accordingly, will suffice.  We have implemented such a procedure along with 
-the resulting score computations in the 
+these allocations are given by $q_{\tau^{\star},F_i}$ where $\tau^{\star}$ solves the equation \eqref{eqn:quantiles-sum-to-K}.
+However, this equation is not analytically tractable and we must resort to a numerical method
+for finding an approximation $\tilde{\tau}$ of $\tau^{\star}$.
+
+We have implemented an iterative bisection method that makes use of the fact that $\sum_{i=1}^N q_{\tau,F_i}$
+is an increasing function of $\tau$.
+The algorithm begins with an initial search interval $[\tau_L, \tau_U]$ that is known to contain the solution $\tau^{\star}$.
+This can be confirmed by verifying that the resource allocations corresponding to the lower endpoint of the search interval do not exceed the resource constraint ($\sum_{i=1}^N q_{\tau_L,F_i} \leq K$) and the resource allocations corresponding to the upper endpoint match or exceed the available resources ($\sum_{i=1}^N q_{\tau_U,F_i} \geq K$).
+At each step of the algorithm, we evaluate the total allocation $\sum_{i=1}^N q_{\tau_{M},F_i}$ at the midpoint of the search interval, $\tau_M = \frac{1}{2}(\tau_L + \tau_U)$.
+We then narrow the search to the sub-interval that contains the solution:
+\begin{itemize}
+  \item If $\sum_{i=1}^N q_{\tau_{M},F_i} \geq K$, we continue the search on the interval $[\tau_L, \tau_M]$.
+  \item If $\sum_{i=1}^N q_{\tau_M, F_i} \leq K$, we continue the search on the interval $[\tau_M, \tau_U]$.
+\end{itemize}
+This search continues until a stopping condition, defined in terms of the width of the search interval, is met.
+We have implemented this procedure along with the resulting score computations in the 
 R package \verb`alloscore` which provided all allocation score values used in the analysis of section \ref{sec:application} 
@@
 
-Awkward subtleties can arise though when the forecast
-densities $f_i$ vanish or are very small, in which case quantiles are non-unique or highly variable at a probability level, 
-leading to ambiguity in how to evaluate $\sum_{i=1}^N q_{\tau,F_i}$. And if point masses are present in any of the $F_i$, 
-\eqref{eqn:quantiles-sum-to-K} will fail to have a unique solution for some discrete set of constraint levels $K$.
-Potentially complex conventions must be adopted for both detecting such levels and enforcing consistency in score calculations
-near them. In both situations, it appears to us that conditions guaranteeing that numerical instabilities and artifacts do not 
-threaten the validity of inferences regarding forecaster performance may be highly dependent on $N$ as well as forecast ranges
-and regularity.  We have established and coded conditions that through extensive experimentation seem appropriate for...
-but we leave a more rigorous approximation error analysis for later work.
+% Subtleties can arise when the forecast
+% densities $f_i$ vanish or are very small, in which case quantiles are non-unique or highly variable near a probability level, 
+% leading to ambiguity or numerical instabilities in the evaluation of $\sum_{i=1}^N q_{\tau,F_i}$. Additionally, if point masses are present in any of the $F_i$, 
+% \eqref{eqn:quantiles-sum-to-K} will not have a unique solution for some discrete set of constraint levels $K$.
+% We have adopted conventions for detecting such levels and enforcing consistency in score calculations
+% near them.
+Through extensive experimentation, we have determined that these conditions seem to address these challenges with the forecasts we are working with, but we leave a more rigorous approximation error analysis for later work.
 

### 2023-12-21 alloscore_manuscript/supplement_short.tex
minor wording fix

@@
 
-In order for $x^{\star} \in \mathbb{R}^N_{+}$ to solve the AP it must be true that reallocating $\delta > 0$ units of the resource from location $i$ to location $j$ will lead to a net increase in expected shortage --- in other words, if the reallocation increases the expected shortage in location $i$ is at least as much as it decreases the expected shortage in location $j$:
+In order for $x^{\star} \in \mathbb{R}^N_{+}$ to solve the AP it must be true that reallocating $\delta > 0$ units of the resource from location $i$ to location $j$ will lead to a net increase in expected shortage --- in other words, the reallocation increases the expected shortage in location $i$ is at least as much as it decreases the expected shortage in location $j$:
 \begin{align*}

### 2023-12-21 alloscore_manuscript/supplement_short.tex
reword dfq Bayes sr

--- 2023-12-21 alloscore_manuscript/supplement_short.tex
@@
 
-The set $\mathcal{P}_{\mathrm{dfq}}$ with map $M(G_{\theta}):= x^{G_{\theta}}$ is surjective onto the action set of allocations
-under any constraint $K$ since they can be specified as quantiles for distributions in $\mathcal{P}_{\mathrm{dfq}}$.  Therefore
-the \verb`distfromq` ``reparametrized'' allocation scoring rule coincides with the AS itself.
+Identifying $G_{\theta}\in \mathcal{P}_{\mathrm{dfq}}$ with its quantile parameters $\theta$, the map $M(\theta):= x^{\theta}$  
+is surjective onto the action set of allocations under any constraint $K$ since they can be specified as quantiles for 
+distributions in $\mathcal{P}_{\mathrm{dfq}}$.  Therefore
+the \verb`distfromq` ``reparametrized'' allocation scoring rule 
+$S_{A, \mathrm{dfq}}(F,y):= s_{\mathrm{dfq}}(\theta^F,y) = s_A(x^{\theta^F},y)$ coincides with the AS $S_A(F,y)$ itself.
 

### 2023-12-21 alloscore_manuscript/supplement_short.tex
minor suggested changes to supplement.

@@
 
-Decision theoretically, probabilistic forecasts are a unique kind of action in that they can be used to generate their
+Probabilistic forecasts can be seen as a unique kind of action in that they can be used to generate their
 own (simulated) outcome data, against which they can be scored using $S$. A probabilistic forecast
@@
 
-A natural consistency criterion for $S$ is that it does not commit $F$ to assessing any other forecast $G$ 
+A natural consistency criterion for $S$ is that, for observations assumed to be drawn from $F$, it will not assess any other forecast $G$
 as being better than $F$ itself, that is, that
@@
 
-Here we study the form of the Bayes act for the allocation problem (AP) (\eqref{eqn:loss_fn} in section \ref{sec:methods.detailed.specific_allocation}) of the text:
+Here we study the form of the Bayes act for the allocation problem (AP) (equation \eqref{eqn:loss_fn} in section \ref{sec:methods.detailed.specific_allocation}) of the text:
 \begin{align}
@@
 \end{align}
+where the marginal forecasts $F_i$ for $i=1,\dots,N$ represent forecasts for $N$ distinct locations.
 We show that the Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$ and resource constraint level $K$
@@
 is an increasing function of $\tau$.
-The algorithm begins with an initial search interval $[\tau_L, \tau_U]$ that is known to contain the solution $\tau^{\star}$.
-This can be confirmed by verifying that the resource allocations corresponding to the lower endpoint of the search interval do not exceed the resource constraint ($\sum_{i=1}^N q_{\tau_L,F_i} \leq K$) and the resource allocations corresponding to the upper endpoint match or exceed the available resources ($\sum_{i=1}^N q_{\tau_U,F_i} \geq K$).
-At each step of the algorithm, we evaluate the total allocation $\sum_{i=1}^N q_{\tau_{M},F_i}$ at the midpoint of the search interval, $\tau_M = \frac{1}{2}(\tau_L + \tau_U)$.
-We then narrow the search to the sub-interval that contains the solution:
-\begin{itemize}
-  \item If $\sum_{i=1}^N q_{\tau_{M},F_i} \geq K$, we continue the search on the interval $[\tau_L, \tau_M]$.
-  \item If $\sum_{i=1}^N q_{\tau_M, F_i} \leq K$, we continue the search on the interval $[\tau_M, \tau_U]$.
-\end{itemize}
-This search continues until a stopping condition, defined in terms of the width of the search interval, is met.
+The algorithm begins with an initial search interval $[\tau_{L,1}, \tau_{U,1}]$ (such as $[0,\max_{i}F_i(K)]$) that clearly contains
+the solution $\tau^{\star}$.
+% This can be confirmed by verifying that the resource allocations corresponding to the lower endpoint of the search interval do not exceed the resource constraint ($\sum_{i=1}^N q_{\tau_L,F_i} \leq K$) and the resource allocations corresponding to the upper endpoint match or exceed the available resources ($\sum_{i=1}^N q_{\tau_U,F_i} \geq K$).
+At each step $j$ of the algorithm, we evaluate the total allocation $\sum_{i=1}^N q_{\tau_{M,j},F_i}$ at the midpoint of the search interval,
+$\tau_{M,j} = \frac{1}{2}(\tau_{L,j} + \tau_{U,j})$ and continue the search on the narrowed sub-interval
+\begin{align}
+[\tau_{L,j+1},\tau_{U,j+1}] =
+\begin{cases}
+[\tau_{L,j}, \tau_{M,j}] & \text{if } \sum_{i=1}^N q_{\tau_{M,j},F_i} \geq K \\
+[\tau_{M,j}, \tau_{U,j}] & \text{if } \sum_{i=1}^N q_{\tau_{M,j},F_i} < K.
+\end{cases} \nonumber
+\end{align}
+% We then narrow the search to the sub-interval that contains the solution:
+% \begin{itemize}
+%   \item If $\sum_{i=1}^N q_{\tau_{M},F_i} \geq K$, we continue the search on the interval $[\tau_L, \tau_M]$.
+%   \item If $\sum_{i=1}^N q_{\tau_M, F_i} \leq K$, we continue the search on the interval $[\tau_M, \tau_U]$.
+% \end{itemize}
+This search continues until $\tau_{U,j+1} < (1+\varepsilon)\tau_{L,j+1}$ for a suitably small $\varepsilon>0$.
 We have implemented this procedure along with the resulting score computations in the
@@
 
-% Subtleties can arise when the forecast
-% densities $f_i$ vanish or are very small, in which case quantiles are non-unique or highly variable near a probability level, 
-% leading to ambiguity or numerical instabilities in the evaluation of $\sum_{i=1}^N q_{\tau,F_i}$. Additionally, if point masses are present in any of the $F_i$, 
-% \eqref{eqn:quantiles-sum-to-K} will not have a unique solution for some discrete set of constraint levels $K$.
-% We have adopted conventions for detecting such levels and enforcing consistency in score calculations
-% near them.
+Subtleties can arise when the forecast
+densities $f_i$ vanish or are very small, in which case quantiles are non-unique or highly variable near a probability level,
+leading to ambiguity or numerical instabilities in the evaluation of $\sum_{i=1}^N q_{\tau,F_i}$. Additionally, if point masses are present in any of the $F_i$,
+\eqref{eqn:quantiles-sum-to-K} will not have a unique solution for some discrete set of constraint levels $K$.
+We have adopted conventions for detecting such levels and enforcing consistency in score calculations
+near them.
 Through extensive experimentation, we have determined that these conditions seem to address these challenges with the forecasts we are working with, but we leave a more rigorous approximation error analysis for later work.
@@
 
-In section 3 of the manuscript, we used the allocation score to evaluate forecasts of COVID-19 hospitalizations that have been submitted to the COVID-19 Forecast Hub.
+In section \ref{sec:application} of the manuscript, we used the allocation score to evaluate forecasts of COVID-19 hospitalizations that have been submitted to the US COVID-19 Forecast Hub.
 These forecasts are submitted to the Hub using a set of 23 quantiles of the forecast distribution at the 23 probability levels in the set $\mathcal{T} = \{0.01, 0.025, 0.05, 0.1, 0.15, \ldots, \allowbreak 0.9, 0.95, 0.975, 0.99\}$, which specify a predictive median and the endpoints of central $(1 - \alpha) \times 100\%$ prediction intervals at levels $\alpha = 0.02, 0.05, 0.1, 0.2, \allowbreak 0.3, 0.4, 0.5, 0.6, \allowbreak 0.7, 0.8, 0.9$.
@@
 To address this situation and support the numerical allocation algorithm outlined in section \ref{sec:numeric}, we need a mechanism to approximate the full cumulative distribution functions $F_i$, $i = 1, \ldots, N$ based on the provided quantiles.
-We have developed functionality for this purpose in the \verb`distfromq` package for R (cite).
-This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:
+We have developed functionality for this purpose in the \verb`distfromq` package for R.\citep{ray-distfromq} This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:
 \begin{enumerate}
@@
 
-We refer the reader to the \verb`distfromq` documentation for further detail.
+We refer the reader to the \verb`distfromq` documentation for further detail.\citep{ray-distfromq}
 

### 2023-12-21 alloscore_manuscript/supplement_short.tex
thoughts on sec 7

@@
 
-Probabilistic forecasts can be seen as a unique kind of action in that they can be used to generate their
+Decision theoretically, probabilistic forecasts are a unique kind of action in that they can be used to generate their
 own (simulated) outcome data, against which they can be scored using $S$. A probabilistic forecast
@@
 
-A natural consistency criterion for $S$ is that, for observations assumed to be drawn from $F$, it will not assess any other forecast $G$
+A natural consistency criterion for $S$ is that it does not commit $F$ to assessing any other forecast $G$ 
 as being better than $F$ itself, that is, that
@@
 
-Here we study the form of the Bayes act for the allocation problem (AP) (equation \eqref{eqn:loss_fn} in section \ref{sec:methods.detailed.specific_allocation}) of the text:
+Here we study the form of the Bayes act for the allocation problem (AP) (\eqref{eqn:loss_fn} in section \ref{sec:methods.detailed.specific_allocation}) of the text:
 \begin{align}
@@
 \end{align}
-where the marginal forecasts $F_i$ for $i=1,\dots,N$ represent forecasts for $N$ distinct locations.
 We show that the Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$ and resource constraint level $K$
@@
 
-In section \ref{sec:application} of the manuscript, we used the allocation score to evaluate forecasts of COVID-19 hospitalizations that have been submitted to the US COVID-19 Forecast Hub.
+In section 3 of the manuscript, we used the allocation score to evaluate forecasts of COVID-19 hospitalizations that have been submitted to the COVID-19 Forecast Hub.
 These forecasts are submitted to the Hub using a set of 23 quantiles of the forecast distribution at the 23 probability levels in the set $\mathcal{T} = \{0.01, 0.025, 0.05, 0.1, 0.15, \ldots, \allowbreak 0.9, 0.95, 0.975, 0.99\}$, which specify a predictive median and the endpoints of central $(1 - \alpha) \times 100\%$ prediction intervals at levels $\alpha = 0.02, 0.05, 0.1, 0.2, \allowbreak 0.3, 0.4, 0.5, 0.6, \allowbreak 0.7, 0.8, 0.9$.
@@
 To address this situation and support the numerical allocation algorithm outlined in section \ref{sec:numeric}, we need a mechanism to approximate the full cumulative distribution functions $F_i$, $i = 1, \ldots, N$ based on the provided quantiles.
-We have developed functionality for this purpose in the \verb`distfromq` package for R.\citep{ray-distfromq} This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:
+We have developed functionality for this purpose in the \verb`distfromq` package for R (cite).
+This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:
 \begin{enumerate}
@@
 
-We refer the reader to the \verb`distfromq` documentation for further detail.\citep{ray-distfromq}
+We refer the reader to the \verb`distfromq` documentation for further detail.
 
@@
 \label{sec:propriety_of_parametric_approximation}
+
+In section \ref{sec:alloscore_proper} we showed that Bayes scoring rules (and the allocation score as a special case) are proper.
+However, this result was stated for a setting where the full forecast distribution $F$ is available.
+In practice, open forecasting exercises are generally not able to collect an exact representation of the forecast distribution $F$ other than in simple settings such as for a categorical variable with a relatively small number of categories.
+In settings where the outcome being forecasted is a continuous quantity (such as the proportion of outpatient doctor visits where the patient had influenza-like illness) or a count (such as influenza hospitalizations), forecasting exercises have therefore resorted to collecting summaries of a forecast distribution such as predictive quantiles.
+
+Here we consider a setting where forecast distributions are summarized by some low-dimensional summary $\theta$. We make two observations in this setting:
+\begin{enumerate}
+  \item If the hub (or forecast evaluator) specifies ahead of time how $\theta$ will be converted into a score in the presence of data y, and ...
+level set stuff ..., a scoring rule can be formulated that is equivalent to S.
+  \begin{itemize}
+    \item We have a low dimensional summary (e.g. quantiles).  Maybe that summary directly gives us the summary we want. For example, to calculate WIS, we are done!
+  \end{itemize}
+  \item On the other hand, if the loss that will be used to score a submitted $\theta$ is not specified in advance, there is no way for the forecaster to select a value of $\theta$ that is derived from their forecast distribution F in any meaningful way.
+\end{enumerate}
+
+TEXT WE MAY NOT LIKE ANYMORE:
+Here, we consider a setting where forecasts are represented using a parametric family of forecast distributions $G_\theta(y)$, and the task of the forecaster is to select a particular parameter value $\theta$.
+We use $\mathcal{P} = \{G_\theta(y), \theta \in \Theta \}$ to denote the collection of all distributions in the given parametric family.
+For instance, it has recently been proposed that mixture distributions could be used to represent forecast distributions \cite{wadsworth2023mixture}.
+Additionally, we note that the functionality in \verb`distfromq` can be viewed as specifying a parametric family $\mathcal{P}_{\mathrm{dfq}}$ where the parameters $\theta$ of $G_\theta$
+are its quantiles at pre-specified probability levels, and where the shape of any $G_\theta \in \mathcal{P}_{\mathrm{dfq}}$ over the full range of its support is entirely controlled by these quantiles.
+
+Our general result is that Bayes scoring rules are invariant under reparametrization of the action space.
+
+Main points:
+ * If we have access to the actual forecast distributions F, Bayes scoring rules (including allocation score) are proper.
+    - We already said this in section 2, remind the reader of that.
+    - We don't need to say this: the propriety of such a score doesn't depend on whether or not the forecaster knows that we'll be using that score
+ * However, if we don't have access to F, the forecaster can only share some intermediate [representation]/[low-dimensional functional] of F
+    - The hub (or forecast evaluator) must specify ahead of time how that representation will be converted into a score in the presence of data y, i.e., the mapping s^\star(theta, y) = s(M(theta), y).
+With this knowledge, the ELM forecaster can select the representation theta^F that minimizes E_F[s^\star(theta, Y)] = E_F[ s(M(theta), Y) ] (1).
+        - In the range of 0 to 1 sentences: Suppose that the set of functionals that may be submitted is Theta, and M(Theta) \subset, not equal....
+            - some forecasts F, F' with different bayes acts x^F, x^{F'} but theta^F = theta^{F'} ...
+there is no way to tell them apart from functionals, (a score that was stricly proper may be converted to a score that is proper, for example.
+)
+            - there is a level set for a theta value that does not live inside a level set for a single A value, and therefore the scores are not equivalent --- the score calcualted based on this functional is "coarser" at that value.
+    - On the other hand, if this information is not communicated to the forecaster ahead of time, they may select some non-optimal theta, theta^\prime which does not minimize (1)
+
+OLD BRAINSTORMING:
+ * However, if we don't have access to F, the forecaster can only share some intermediate [representation]/[low-dimensional functional] of F
+    - Consider the quantile score: We can't directly evaluate S_q(F, y)
+       - Option 1: but we can evaluate it exactly via the functionals that are provided, if the forecaster knew that they should provide quantiles for their functionals, i.e., if they knew that we were going to use quantile score/WIS for evaluation.
+       - Option 2: if I have a sufficient statistic theta_suff^F that allows me to calculate the quantiles of F as M_suff(theta_suff^F) "regardless of whether or not forecaster knows mapping M_suff", I can calculate S_q(F, y)
+          - But there's a step in the hub setup/elicitation process where we tell the forecaster to submit theta_suff^F
+       - Option 3: 
+    - For allocation score:
+       - One option: the forecaster submits their allocations x^F.  This is the only quantity you need to plug into s_A(x, y) in order to calculate S_A(F, y).
+       - Another option: the forecaster submits some other quantity theta^F that is (sufficient) for downstream calculations of x^F, e.g.
+via x^F = allocate(distfromq(theta^F)).
+As long as the evaluator and the forecaster agree on the mapping from theta^F to x^F, and any point in X can be reached by some theta in Theta, the resulting score is proper.
+       - Quantiles are not "sufficient" for allocations.
+       - But a vector of 23 numbers used to parameterize dfq are sufficient for allocations
 

### 2023-12-22 alloscore_manuscript/supplement_short.tex
updates to supp

--- 2023-12-21 alloscore_manuscript/supplement_short.tex
@@
 \section{Computing allocations from finite quantile forecast representations}
+\label{sec:distfromq}
 
@@
 
-\textbf{Bayes scoring rules are invariant under reparametrization of the action space.}
+In practice, open forecasting exercises are generally not able to collect a perfect description of the forecast distribution $F$ other than in simple settings such as for a categorical variable with a relatively small number of categories.
+In settings where the outcome being forecasted is a continuous quantity (such as the proportion of outpatient doctor visits where the patient has influenza-like illness) or a count (such as influenza hospitalizations), forecasting exercises have therefore resorted to collecting summaries of a forecast distribution such as bin probabilities or predictive quantiles.
+In this section, we address two practical concerns raised by this.
+First, we discuss conditions under which it is possible to calculate the allocation score when only summaries of a forecast distribution are recorded in a submission to a forecast hub.
+Second, we show that a post hoc attempt to compute the allocation score based on submitted predictive quantiles may in fact compute an alternative score that is not proper.
 
-Let $S(F,y)$ be a Bayes scoring rule derived from an action set $\mathcal{X}$ and loss $s(x,y)$.
-Let $\mathcal{W}$ be another set and $M:\mathcal{W} \to \mathcal{X}$ a map. $M$ defines a loss $s_M(\theta, y):= s(M(\theta), y)$
-on $\mathcal{W}$. This in turn defines the Bayes scoring rule
+\subsection{Propriety -- prospective}
+
+We consider a setting where a forecasting exercise (such as a forecast hub) pre-specifies that forecasts will be represented using a parametric family of forecast distributions $G_\theta(y)$, and the task of the forecaster is to select a particular parameter value $\theta$.
+We use $\mathcal{P}$ to denote the collection of all distributions $G_\theta$ in the given parametric family.
+For instance, it has recently been proposed that mixture distributions could be used to represent forecast distributions \cite{wadsworth2023mixture}.
+Additionally, we note that the functionality in \verb`distfromq` can be viewed as specifying a parametric family $\mathcal{P}_{\mathrm{dfq}}$ where the parameters $\theta$ of $G_\theta$
+are its quantiles at pre-specified probability levels, and where the shape of any $G_\theta \in \mathcal{P}_{\mathrm{dfq}}$ over the full range of its support is entirely controlled by these quantiles.
+
+We find it helpful now to formally distinguish between two decision making problems.
+The first is the public health decision maker's allocation problem where the task is to select an allocation $x$, with the allocation loss $s_A(x, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i)$ as described in section \ref{sec:methods.detailed.specific_allocation} of the main text.
+The second is the forecaster's reporting problem where the task is to select parameter values $\theta$ to report.
+The forecaster's loss is given by $s_R(\theta, y) = s_A(x^{G_\theta}, y)$, where $x^{G_\theta}$ is the Bayes act for the allocation problem under the distribution $G_\theta$.
+In words, the loss associated with reporting $\theta$ is equal to the loss associated with taking the Bayes allocation corresponding to the distribution $G_\theta$.
+
+Following our usual construction, the Bayes act for the forecast reporting problem is the parameter set that minimizes the forecaster's expected loss.
+Breaking with our earlier notation for improved legibility, we use $\theta^\star(F)$ to denote this Bayes act:
+\begin{align*}
+\theta^\star(F) &= \text{argmin}_\theta \Ex_F [s_R(\theta, Y)] \\
+&= \text{argmin}_\theta \Ex_F [s_A(x^{G_\theta}, Y)]
+\end{align*}
+
+We then arrive at the scoring rule
+$$S_R(F, y) = s_R(\theta^\star(F), y) = s_A(x^{G_{\theta^\star(F)}}, y).$$
+It follows from the discussion in section \ref{sec:alloscore_proper} that this is a proper scoring rule for $F$.
+Although the full forecast distribution $F$ is not available in the forecast submission, the score $S_R(F, y)$ can be calculated from the reported parameter values as long as the forecaster submits the optimal parameters $\theta^\star(F)$.
+
+We emphasize that the forecaster's true predictive distribution $F$ does not need to be a member of the specified parametric family $\mathcal{P}$ for this construction to yield a proper score.
+It is, however, necessary to specify the parametric family to use and the foundational scoring rule $s_A$ (including any relevant problem parameters such as the resource constraint $K$) in advance, so that forecasters can identify the Bayes act parameter set $\theta^\star(F)$ to report.
+
+If the parametric family used to represent forecast distributions is flexible enough, the reporting scoring rule $S_R$ and the allocation score are equivalent in the sense that they will yield the same score for any distribution $F$.
+Suppose that for a given resource constraint $K$, for any forecast distribution $F$ it is possible to find a member $G_{\theta^\star}$ of the specified parametric family $\mathcal{P}$ with the same allocation as $F$ (i.e., $x^F = x^{G_{\theta^\star}}$).
+Then $\theta^\star$ is a Bayes act for the reporting problem since for any other parameter value $\theta$,
+\begin{align*}
+\Ex_F[s_R(\theta^\star, Y)] &= \Ex_F[ s_A(x^{G_{\theta^\star}}, Y) ] \\
+&= \Ex_F[s_A(x^F, Y)]  \quad \text{ (since $x^F = x^{G_{\theta^\star}}$)} \\
+&\leq \Ex_F[ s_A(x^{G_\theta}, Y) ]  \quad \text{ (by definition of $x^F$)} \\
+&= \Ex_F[ s_R(\theta, Y)].
+\end{align*}
+It therefore follows that
+\begin{align*}
+S_R(F, y) &= s_R(\theta^\star, y) \\
+&= s_A(x^{G_{\theta^\star}}, y) \\
+&= s_A(x^F, y) \\
+&= S_A(F, y).
+\end{align*}
+
+For the particular choice of the parametric family $\mathcal{P}_{\text{dfq}}$ (i.e., using the \verb`distfromq` package), this flexibility requirement is satisfied.
+For instance, the forecaster could pick one required quantile level (such as 0.5, for which the corresponding predictions are predictive medians), and set the submitted quantiles of their forecast distribution in each location at that level to be the desired allocations, which sum to $K$ across all locations.
+However, this representation of the forecast may be quite different from the actual forecast distribution $F$.
+For example, for the actual forecast distribution $F$ the allocations may occur at some quantile level other than 0.5.
+
+As another alternative for practical forecasting exercises, a forecast hub could ask forecasters to directly provide the Bayes allocations associated with their forecasts for one or more specified resource constraints $K$.
+At the cost of increasing the number of quantities solicited by the forecast hub, this would have several advantages: it would prevent any artificial distortion of the forecast distributions, allow for direct calculation of scores, and narrow the gap between model outputs and public health end users.
+For this to be feasible, implementations of the allocation algorithm would have to be provided to participating forecasters in the computational languages being used for modeling.
+
+\subsection{Propriety -- post hoc}
+
+A \emph{post hoc} evaluation of quantile forecasts that combines the parametric family specified by \verb`distfromq` with the allocation score does not yield the allocation score of the forecast distribution $F$.
+Instead, it computes an alternative score that is improper.
+This is because the forecast distribution $F$ and the distribution $G^q \in \mathcal{P}_{\text{dfq}}$ with the same quantiles as $F$ may determine different resource allocations.
+In our investigations, these discrepancies appear to be relatively minor on the interior of the provided quantiles, but could be severe if the tail extrapolations performed by \verb`distfromq` do not match the tail behavior of $F$ and the allocations are in the tails of the predictive distribution.
+
+We define
+$$G^{\star}(F) := \argmin_{G \in \mathcal{P}_{\mathrm{dfq}}} E_{F}[S_A(G, Y)],$$
+the distribution corresponding to the optimal submission for the forecaster with predictive distribution $F$.
+Because the distribution $G^q(F)$ typically will not yield allocations that are optimal under the distribution $F$, in general $G^q(F)$ and $G^\star(F)$ are different distributions.
+
+When an analyst attempts a post hoc computation of the allocation score using $G^q(F)$ (implicitly assuming that $G^q(F) = G^\star(F)$), they in fact compute the alternative score
+$$\tilde{S}(F, y) = S_A(G^q(F), y) = s_A(x^{G^q(F)}, y).$$
+This score is improper because $E_{F}[S_A(G, Y)]$ is minimized by $G^\star(F)$, not $G^q(F)$.
+That is,
 \begin{align}
-S_M(F,y) := s_M(\theta^F,y) = s(M(\theta^F),y).
+E_{F}[\tilde{S}(G^\star(F), Y)] &< E_{F}[ S_A(G^q(F), Y) ] \label{eqn:tilde_s_improper} \\
+  &= E_{F}[\tilde{S}(F, Y)] \nonumber
 \end{align}
-If $M$ is surjective onto the set of Bayes acts $\{x \in \mathcal{X} \mid x = x^F \text{ for some } F \in \mathcal{P}\}$, then the scores coincide since
-\begin{align}
-S_M(F,y) = s_M(\theta^F,y) & = \min_{\theta}\{s_M(\theta,F)\} \\
-&= \min_{\theta}\{s(M(\theta),F)\}  \\
-&= \min_{x}\{s(x,F)\} = s(x^F,F) = S(F,y).
-\end{align}
-(Note that this does not imply that $M(\theta^F) = x^F$ since $x^F$ might not be a unique Bayes act.)
+The inequality in step \eqref{eqn:tilde_s_improper} is strict when $G^q(F)$ is not a minimizer of the expected allocation loss under the distribution $F$.
+At an intuitive level, this occurs when the allocations determined by $G^\star(F)$ and $G^q(F)$ are different.
 
-Identifying $G_{\theta}\in \mathcal{P}_{\mathrm{dfq}}$ with its quantile parameters $\theta$, the map $M(\theta):= x^{\theta}$
-is surjective onto the action set of allocations under any constraint $K$ since they can be specified as quantiles for
-distributions in $\mathcal{P}_{\mathrm{dfq}}$.  Therefore
-the \verb`distfromq` ``reparametrized'' allocation scoring rule
-$S_{A, \mathrm{dfq}}(F,y):= s_{\mathrm{dfq}}(\theta^F,y) = s_A(x^{\theta^F},y)$ coincides with the AS $S_A(F,y)$ itself.
+% Def. $S(F, y)$ is proper if $E_{Y \sim F}[S(F, Y)] \leq E_{Y \sim F}[S(G, Y)]$ for all $F$, $G$.
 
-Unfortunately,
+% One way to demonstrate that S is improper is to find a pair F, G such that $E_{Y \sim F}[S(F, Y)] > E_{Y \sim F}[S(G, Y)]$.
 
-For convenience we write $S(F,G) := \Ex_G[S(F,Y)]$. Let
-\begin{align*}
-q_{23}(F) &:= (q_{.01,F},\ldots,q_{.99,F}) \\
-G^q(F) &:= G_{q_{23}(F)} \in \mathcal{P}_{\mathrm{dfq}} \\
-G^{\star}(F) &:= \argmin_{G \in \mathcal{P}_{\mathrm{dfq}}} S_A(G,F)
-\end{align*}
-We have the ``right'' approximate allocation score
-\begin{align*}
-S_{R}(F,y):= S_A(G^{\star}(F),y)
-\end{align*}
-which is proper because
-\begin{align*}
-S_R(F,F) &= S_A(G^{\star}(F), F) \\
-&\leq S_A(G^{\star}(F), H) \quad \text{ (by definition of $G^{\star}$)} \\
-&=S_R(F,H).
-\end{align*}
-This seems to almost be what is called ``properization'' in \cite{brehmer2020properization}.
-We also have a ``wrong'' approximation
-\begin{align*}
-\tilde{S}_R(F,y) := S_A(G^q(F),y)
-\end{align*}
-which is improper because
-\begin{align*}
-\tilde{S}_R(F,G^q(F)) &= S_A(G^q(F),G^q(F)) \\
- &<  S_A(G^q(F),F) \quad \text{ (because $S_A$ is proper)} \\
- &= \tilde{S}_R(F,F).
-\end{align*}
+% Take $\tilde{S}(F, y)$ to be the "incorrect" scoring rule that uses quantiles rather than a minimizer of the expected DFQ+Alloscore loss:
+% $$\tilde{S}(F, y) = S_A(G^q(F), y) = s_A(x^{G^q(F)}, y)$$
+
+% What we want is a distribution F and an alternative distribution G such that $E_{Y \sim F}[S(F, Y)] > E_{Y \sim F}[S(G, Y)]$.
+
+%  - Take F to be something with non-normal tails, in which case $x^{G^q(F)}$ will be a non-optimal allocation, and
+%  - G to be $G^\star(F)$ = the minimizer of $E_{Y \sim F}[S_A(F, Y)]$
+
+%  Then we will have
+% \begin{align*}
+% E_{Y \sim F}[\tilde{S}(G, Y)] &= E_{Y \sim F}[\tilde{S}(G^\star(F), Y)] \\
+% &= E_{Y \sim F}[ S_A(G^q(G^\star(F)), Y) ] \\
+% &= E_{Y \sim F}[ S_A(G^\star(F), Y) ] \\
+% &= E_{Y \sim F}[ s_A(x^{G^\star(F)}, Y) ] \\
+% &= E_{Y \sim F}[ s_A(x^{F}, Y) ] \\
+% &< E_{Y \sim F}[ s_A(x^{G^q(F)}, Y) ] \\
+% &= E_{Y \sim F}[\tilde{S}(F, Y)]
+% \end{align*}
+
+% \begin{align*}
+%   E_{Y \sim F}[\tilde{S}(G, Y)] &= E_{Y \sim F}[\tilde{S}(G^\star(F), Y)] \\
+%   &= E_{Y \sim F}[ S_A(G^q(G^\star(F)), Y) ] \\
+%   &= E_{Y \sim F}[ S_A(G^\star(F), Y) ] \\
+%   &< E_{Y \sim F}[ S_A(G^q(F), Y) ] \\
+%   &= E_{Y \sim F}[\tilde{S}(F, Y)]
+%   \end{align*}
+
+% $G^q(G^\star(F)) = G^\star(F)$
+
+% $E_{Y \sim F}[\tilde{S}(F, Y)]$
+
+% We can show that this scoring rule is improper by taking, for example, $F = (Exp(1/1), Exp(1/4))$, or something else like this with non-normal tails.
+Find an example where the allocations derived from $G^q(F)$ are different from the allocations from $F$.
 

### 2023-12-22 alloscore_manuscript/supplement_short.tex
attempt to clean up some language around discussion of impropriety

@@
 The location and scale parameters are estimated separately for the lower and upper tails so as to obtain a tail distribution that matches the two most extreme quantiles in each tail.
+In this work, we use normal distributions for the tails.
   \end{enumerate}
@@
 This score is improper because $E_{F}[S_A(G, Y)]$ is minimized by $G^\star(F)$, not $G^q(F)$.
-That is,
+In general, we have
 \begin{align}
-E_{F}[\tilde{S}(G^\star(F), Y)] &< E_{F}[ S_A(G^q(F), Y) ] \label{eqn:tilde_s_improper} \\
+E_{F}[\tilde{S}(G^\star(F), Y)] &\leq E_{F}[ S_A(G^q(F), Y) ] \label{eqn:tilde_s_improper} \\
   &= E_{F}[\tilde{S}(F, Y)] \nonumber
 \end{align}
-The inequality in step \eqref{eqn:tilde_s_improper} is strict when $G^q(F)$ is not a minimizer of the expected allocation loss under the distribution $F$.
-At an intuitive level, this occurs when the allocations determined by $G^\star(F)$ and $G^q(F)$ are different.
+However, the inequality in step \eqref{eqn:tilde_s_improper} will typically be strict.
+For example, if $F$ has heavy upper tails (such as for a lognormal distribution), but normal distributions are used for tail extrapolations in \verb`distfromq`, then the resource allocations based on the distribution $G^q(F)$ may be quite different from the optimal allocations under the distribution $F$, leading to a strict inequality.
+This demonstrates that $\tilde{S}$ is improper.
 

### 2023-12-22 alloscore_manuscript/supplement_short.tex
update to intro of supplement

@@
 
-\begin{todolist}
-\item 2 actual references to supplement (re exponential quantiles and solving the allocation problem)
-\item the statement that "we computed the allocation score for the 14 day-ahead forecast" which needs some answer to "how?"
-\item use of "proper" a number of times
-\end{todolist}
-
 \section{Introduction}
@@
 
-\elr{This section needs to be updated.}
-
-We briefly address some technical and methodological points in the main text.
-We begin in section \ref{sec:shortage} by formalizing the concept of a \emph{shortage} of resources and giving some key results about expected resource shortages under a distribution characterizing uncertainty about (future) levels of resource need.
-Resource shortages play a central role in the decision-making problems that give rise to the quantile loss and the allocation score, which we discuss in sections \ref{sec:quantiles_shortage} and \ref{sec:bayes-quantiles} respectively.
-Section \ref{sec:numeric} gives details on the numerical methods that we use to calculate allocation scores, including some special considerations for settings where forecasts are represented by a finite collection of predictive quantiles, such as the application to forecasts of hospitalizations due to COVID-19 in section 3 of the article.
+We address some technical and methodological points from the main text.
+We begin in section \ref{sec:proper} by defining proper scoring rules and showing that the allocation score is proper.
+In section \ref{sec:ex-shortage}, we formalize the concept of a \emph{shortage} of resources, give some key results about expected resource shortages under a distribution characterizing uncertainty about (future) levels of resource need, and illustrate how quantiles arise as the Bayes act in a decision making problem about the quantity of a resource to purchase.
+Section \ref{sec:bayes-quantiles} gives a justification for the result that the Bayes act for the allocation problem is given by a vector of quantiles of the forecast distributions in each location at a shared probability level, which was stated in section \ref{sec:methods.detailed.specific_allocation} of the main text.
+We describe the algorithm that we use to compute allocations given a forecast distribution in each location in section \ref{sec:numeric}.
+In section \ref{sec:distfromq}, we describe the methods for approximating a full distribution from a set of quantiles, implemented in the R package \verb`distfromq`, that we used to support computation of allocation scores from quantile forecasts that were submitted to the COVID-19 Forecast Hub for the application in the main text.
+Finally, in section \ref{sec:propriety_of_parametric_approximation} we examine implications for propriety of an analysis that uses summaries of forecasts (such as predictive quantiles) rather than the full forecast distributions for computation of allocation scores.
 

### 2023-12-22 alloscore_manuscript/supplement_short.tex
update subsection names

@@
 
-\subsection{Propriety -- prospective}
+\subsection{Propriety when scoring methods are announced prospectively}
 
@@
 
-\subsection{Propriety -- post hoc}
+\subsection{Impropriety of post hoc allocation scoring with quantile forecasts}
 
@@
 \end{align}
-However, the inequality in step \eqref{eqn:tilde_s_improper} will typically be strict.
+However, the inequality in \eqref{eqn:tilde_s_improper} will typically be strict.
 For example, if $F$ has heavy upper tails (such as for a lognormal distribution), but normal distributions are used for tail extrapolations in \verb`distfromq`, then the resource allocations based on the distribution $G^q(F)$ may be quite different from the optimal allocations under the distribution $F$, leading to a strict inequality.

### 2023-12-22 alloscore_manuscript/supplement_short.tex
few more tweaks

@@
 
-A key feature of loss functions used in the decision theoretic definition of quantiles and related scoring rules such as the
-CRPS and the WIS (see ...), as well as the allocation loss function presented in this work, is the presence of a \emph{shortage}:
+A key feature of loss functions for decision problems used to define quantiles and related scoring rules such as the
+CRPS and the WIS (see e.g., \cite{gneiting2011quantiles}, \cite{jose2009evaluating}, 
+and \cite{royset2022optimization}, sections 1.C and 3.C), as well as the allocation loss function presented in this work, 
+is the presence of a \emph{shortage}:
 the amount $\max\{0,y-x\}$ by which a resource demand
@@
 The first is the public health decision maker's allocation problem where the task is to select an allocation $x$, with the allocation loss $s_A(x, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i)$ as described in section \ref{sec:methods.detailed.specific_allocation} of the main text.
-The second is the forecaster's reporting problem where the task is to select parameter values $\theta$ to report.
-The forecaster's loss is given by $s_R(\theta, y) = s_A(x^{G_\theta}, y)$, where $x^{G_\theta}$ is the Bayes act for the allocation problem under the distribution $G_\theta$.
+The second is the forecaster's reporting problem where the task is to select parameter values $\theta$ to report. The forecaster's loss is given by 
+\begin{align}
+s_R(\theta, y) = s_A(x^{G_\theta}, y), \label{eqn:forecaster_theta_loss}
+\end{align}
+where $x^{G_\theta}$ is the Bayes act for the allocation problem under the distribution $G_\theta$.
 In words, the loss associated with reporting $\theta$ is equal to the loss associated with taking the Bayes allocation corresponding to the distribution $G_\theta$.
@@
 We define
-$$G^{\star}(F) := \argmin_{G \in \mathcal{P}_{\mathrm{dfq}}} E_{F}[S_A(G, Y)],$$
-the distribution corresponding to the optimal submission for the forecaster with predictive distribution $F$.
-Because the distribution $G^q(F)$ typically will not yield allocations that are optimal under the distribution $F$, in general $G^q(F)$ and $G^\star(F)$ are different distributions.
+$$G^{\star}(F) := \argmin_{G \in \mathcal{P}_{\mathrm{dfq}}} E_{F}[S_A(G, Y)].$$
+Since $S_R$ is defined as the Bayes scoring rule for the forecaster's loss \eqref{eqn:forecaster_theta_loss}, $G^{\star}(F)$ coincides
+with $G_{\theta^{\star}(F)}$, the distribution in $\mathcal{P}_{\mathrm{dfq}}$ given by the 
+optimal submission parameters $\theta^{\star}(F)$ for the forecaster with predictive distribution $F$.
+In general, $G^q(F)$ and $G^\star(F)$ will be different distributions:  matching $F$ at specific quantiles
+does not require $G^q(F)$ to match $F$ at the quantiles for $\tau^{F,K}$ (c.f. \eqref{eqn:quantiles-sum-to-K}), 
+which would be necessary for it to share $x^F$ as an optimal allocation.
 

### 2023-12-22 alloscore_manuscript/supplement_short.tex
two very minor adjustments to sections 1 and 7, plus small adjustment to bibliography.

@@
 We describe the algorithm that we use to compute allocations given a forecast distribution in each location in section \ref{sec:numeric}.
-In section \ref{sec:distfromq}, we describe the methods for approximating a full distribution from a set of quantiles, implemented in the R package \verb`distfromq`, that we used to support computation of allocation scores from quantile forecasts that were submitted to the COVID-19 Forecast Hub for the application in the main text.
+In section \ref{sec:distfromq}, we describe the methods for approximating a full distribution from a set of quantiles, implemented in the R package \verb`distfromq`, that we used to support computation of allocation scores from quantile forecasts that were submitted to the US COVID-19 Forecast Hub for the application in the main text.
 Finally, in section \ref{sec:propriety_of_parametric_approximation} we examine implications for propriety of an analysis that uses summaries of forecasts (such as predictive quantiles) rather than the full forecast distributions for computation of allocation scores.
@@
 We have implemented this procedure along with the resulting score computations in the
-R package \verb`alloscore` which provided all allocation score values used in the analysis of section \ref{sec:application}
+R package \verb`alloscore` \citep{gerding-alloscore} which provided all allocation score values used in the analysis of section \ref{sec:application}
 in the main text.
@@
 densities $f_i$ vanish or are very small, in which case quantiles are non-unique or highly variable near a probability level,
-leading to ambiguity or numerical instabilities in the evaluation of $\sum_{i=1}^N q_{\tau,F_i}$. Additionally, if point masses are present in any of the $F_i$,
+leading to ambiguity or numerical instabilities{} in the evaluation of $\sum_{i=1}^N q_{\tau,F_i}$. Additionally, if point masses are present in any of the $F_i$,
 \eqref{eqn:quantiles-sum-to-K} will not have a unique solution for some discrete set of constraint levels $K$.
@@
 We use $\mathcal{P}$ to denote the collection of all distributions $G_\theta$ in the given parametric family.
-For instance, it has recently been proposed that mixture distributions could be used to represent forecast distributions \cite{wadsworth2023mixture}.
+For instance, it has recently been proposed that mixture distributions could be used to represent forecast distributions \citep{wadsworth2023mixture}.
 Additionally, we note that the functionality in \verb`distfromq` can be viewed as specifying a parametric family $\mathcal{P}_{\mathrm{dfq}}$ where the parameters $\theta$ of $G_\theta$

### 2023-12-22 alloscore_manuscript/alloscore-application-shorter.Rnw
some progress on short supp
add Acknowledgements

@@
 S_A(F, y; K)  &= S_A^{\text{raw}}(F, y; K) - L \cdot \max\left(0,\sum_{i=1}^{N}y_i - K\right) \\
-&= L\left\{\sum_{i=1}^N \max(0, y_i - x_i^{F,K}) -  \max\left(0,\sum_{i=1}^{N}y_i - K\right)\right\}
+&= L\left\{\sum_{i=1}^N \max(0, y_i - x_i^{F,K}) -  \max\left(0,\sum_{i=1}^{N}y_i - K\right)\right\}.
 \end{align}
@@
 
+\section*{Acknowledgements}
+This work has been supported by the National Institutes of General Medical Sciences (R35GM119582) and the U.S.
+CDC(1U01IP001122).
+The content is solely the responsibility of the authors and does not necessarily represent the official views of NIGMS, the National Institutes of Health, or CDC.
+
 \bibliography{allocation}

### 2023-12-22 alloscore_manuscript/alloscore-application-shorter.tex
some progress on short supp
bib fixes
add Acknowledgements

@@
 
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Dec 20 11:56:48 2023
+% latex table generated in R 4.3.1 by xtable 1.8-4 package
+% Fri Dec 22 10:45:01 2023
 \begin{table}[ht]

### 2024-01-25 alloscore_manuscript/alloscore-application-shorter.Rnw
only meaningful change here is a formatting change to make the \Sexpr render correctly for me, by moving some computation to a chunk rather than inside the expression itself.

@@
 
-The hospitalization data used for scoring forecasts were downloaded on \Sexpr{format(as.Date(metadata$time[which(metadata$name == "truth_data")]), "%B %d, %Y")}.
+<<>>=
+download_date <- format(as.Date(metadata$time[which(metadata$name == "truth_data")]), "%B %d, %Y")
+@
+
+The hospitalization data used for scoring forecasts were downloaded on \Sexpr{download_date}.
 
@@
 
-For both scores, we also computed models' standardized rank among all models that submitted forecasts each week.
+For both scores, we also computed models standardized rank among all models that submitted forecasts each week.
 The standardized rank is between 0 and 1, where 0 corresponds to the worst rank and 1 to the best.

### 2024-01-26 arxiv_submission/alloscore-application-shorter.tex
catching up work from right before arxiv submission

--- 2023-12-22 alloscore_manuscript/alloscore-application-shorter.tex
@@
 \documentclass{article}\usepackage[]{graphicx}\usepackage[]{xcolor}
+\pdfoutput=1
 % maxwidth is the original width if it is less than linewidth
@@
 \usepackage{amsmath, amsfonts, amssymb}
+\usepackage{authblk}
 \usepackage{graphicx}
@@
 
-\title{Evaluating infectious disease forecasts with allocation scoring rules}
-\author{Aaron Gerding, Nicholas G. Reich, Benjamin Rogers, Evan L. Ray}
+\author{Aaron Gerding\thanks{Corresponding Author: \texttt{agerding@umass.edu}}}
+\author{Nicholas G. Reich}
+\author{Benjamin Rogers}
+\author{Evan L. Ray}
+\affil{Department of Biostatistics and Epidemiology, School of Public Health and Health
+Sciences, University of Massachusetts at Amherst}
+
+\title{Evaluating infectious disease forecasts \\ with allocation scoring rules}
+%\author{Aaron Gerding, Nicholas G. Reich, Benjamin Rogers, Evan L. Ray}
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
@@
 \begin{figure}
-    \includegraphics[width=\textwidth]{../figures/exponential_pred_expected_loss.pdf}
+    \includegraphics[width=\textwidth]{figure/exponential_pred_expected_loss}
     \caption{An illustration of the resource allocation problem in Example 1.
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Dec 22 10:45:01 2023
+% Fri Dec 22 11:30:21 2023
 \begin{table}[ht]

### 2024-01-26 R/percap.R
try to make percap in pipeline

@@ -1,0 +1,23 @@
+
+score_per_capita_allocation <- function(
+		all_allscore_data, 
+		Kat15k_alloscores) {
+  require(tidyverse)
+
+  pops22 <- readr::read_csv("https://raw.githubusercontent.com/reichlab/flusion/main/data-raw/us-census/NST-EST2022-ALLDATA.csv") %>% 
+  	select(full_location_name = NAME, POPESTIMATE2021) %>% 
+  	inner_join(hub_locations, by = join_by(full_location_name))
+
+	percap <- bind_rows(all_alloscore_data, Kat15k_alloscores) %>% 
+	  filter(model == "COVIDhub-baseline") %>% 
+	  mutate(model = "percap") %>% 
+	  left_join(pops22[c('POPESTIMATE2021', 'abbreviation')], by = "abbreviation") %>% 
+	  slice(1, .by = c(reference_date, K ,abbreviation)) %>% 
+	  mutate(popprop = POPESTIMATE2021 / sum(POPESTIMATE2021), 
+	         .before = x, .by = c(K, reference_date))  %>% 
+	  mutate(x = K * popprop, components_raw = pmax(y-x,0), 
+	         components = components_raw - components_oracle) %>% 
+	  select(-popprop, -POPESTIMATE2021)
+
+	 return(percap)
+}

### 2024-01-26 _targets.R

@@ -27,6 +27,7 @@ tar_source(files = c("R/data-ingestion.R",
                      "R/run-alloscore.R",
                      "R/determine-model-eligibility.R",
                      "R/exponential-examples.R",
+                     "R/percap.R",
                      "R/plot_functions.R"))
 
 values <- tibble(forecast_dates = as.character(seq.Date(as.Date("2021-11-22"), as.Date("2022-02-28"), by = "7 days")))
@@ -96,5 +97,9 @@ combined <- tar_combine(
 #   format = "file"
 # )
 
+make_percap <- tar_target(
+  name = percap,
+  command = score_per_capita_allocation(all_alloscore_data, Kat15k_alloscores)
+)
 
-list(setup, mapped, combined)
+list(setup, mapped, combined, make_percap)

### 2024-01-26 alloscore_manuscript/alloscore-application-shorter.Rnw
adding some minor updates to application-shorter and bib file.

@@
 
-\title{Evaluating infectious disease forecasts with allocation scoring rules}
+\title{Evaluating infectious disease forecasts \\ with allocation scoring rules}
 \author{Aaron Gerding, Nicholas G. Reich, Benjamin Rogers, Evan L. Ray}
@@
 Predictions of disease dynamics have been used to inform decision making about a wide variety of measures to reduce disease spread and/or mitigate the severity of disease outcomes.
-For example, estimates of expected onset of flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola dynamics have been used to allocate surveillance resources \citep{meltzer2014estimating, rainisch2015regional}.
+For example, estimates of expected onset of flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola and diptheria dynamics have been used to allocate resources \citep{meltzer2014estimating, rainisch2015regional,camacho2015-ebola-bed,finger_real-time_2019}.
 \cite{bertsimas2021predictionsCOVID} developed tools to guide decision making from infectious disease forecasts, which have been used to inform allocation of limited medical supplies such as ventilators, ICU capacity planning, and vaccine distribution strategy.
@@
 This analysis is ``synthetic'' in that it does not correspond to an actual analysis that supported decision making in real time.
-However, the framework described in this paper corresponds to real-world decisions that must be made by public health administrators around the globe, and could be adapted in the future for such real-time situations.
-For example, forecasts for districts in Sierra Leone of bed demand to care for patients with Ebola was the subject of a real-time modeling study in late 2014 and early 2015 \citep{camacho2015-ebola-bed}.
-And, in 2020, a model developed by an academic research group turned predictions of COVID-19 hospitalizations into estimates of ventilator usage and shortages.
+However, the framework described in this paper corresponds to real-world decisions that must be made by public health administrators around the globe, and could be adapted in the future for such real-time situations such as informing bed demand, ventilator usage, or doses needed for specific medications.
+%For example, forecasts for districts in Sierra Leone of bed demand to care for patients with Ebola was the subject of a real-time modeling study in late 2014 and early 2015 \citep{camacho2015-ebola-bed}.
+%And, in 2020, a model developed by an academic research group turned predictions of COVID-19 hospitalizations into estimates of ventilator usage and shortages.
 This framework was used by the Hartford HealthCare system in Connecticut ``to align ventilator supply with projected demand at a time where the [COVID-19] pandemic was on the rise'' \citep{bertsimas2021predictionsCOVID}.
@@
 These scores illustrate a general result: allocation scores for a forecast will tend to be larger when the resource constraint is close to the observed need, because this is when it matters most which locations are allocated more or less resources.
-If the resource constraint is very small, any allocation of those limited resources will result in a large amount of unmet need.
-If the resource constraint is very large, it becomes less important which locations receive relatively more or less resources because all locations will receive enough resources to meet their need.
+If the available resources are small compared to the observed need, any allocation of those limited resources will result in a large amount of unmet need.
+If the available resources are comparatively larger, it becomes less important which locations receive relatively more or fewer resources because all locations will receive enough resources to meet their need.
 In either of these extremes of resource availability, the avoidable unmet need that arises from the allocation suggested by a forecast (i.e., the forecast's allocation score) will tend to be small.
@@
 Larger interval scores indicate less skillful forecasts.
-An interval score consists of (a) the width of the interval, with larger intervals receiving higher scores, and (b) a penalty if the interval does not cover the eventual observation, which increases the further away the interval is from the observed value.
+An interval score consists of (a) the width of the interval, with larger intervals receiving higher scores (higher scores indicate less accuracy), and (b) a penalty if the interval does not cover the eventual observation, which increases the further away the interval is from the observed value.
 Equivalently, the WIS can also be characterized as a weighted sum of quantile scores for each individual predictive quantile.
@@
 \section*{Acknowledgements}
+
+We wish to thank the following individuals who contributed valuable comments and feedback on early versions of this work: Matthew Biggerstaff, Rebecca Borchering, Rosa Ergas, Sebastian Funk, Melissa Kerr, and Jeffrey Shaman.
+
 This work has been supported by the National Institutes of General Medical Sciences (R35GM119582) and the U.S.

### 2024-01-29 alloscore_manuscript/alloscore-application-shorter.Rnw
removing Rosa from acknowledgments, per her request and state guidelines.

@@
 
-We wish to thank the following individuals who contributed valuable comments and feedback on early versions of this work: Matthew Biggerstaff, Rebecca Borchering, Rosa Ergas, Sebastian Funk, Melissa Kerr, and Jeffrey Shaman.
+We wish to thank the following individuals who contributed valuable comments and feedback on early versions of this work: Matthew Biggerstaff, Rebecca Borchering, Sebastian Funk, Melissa Kerr, and Jeffrey Shaman.
 

### 2024-01-29 alloscore_manuscript/alloscore-application-shorter.Rnw
adding comments addressed through section 2.

@@
 
-Infectious disease forecasting models have emerged as important tools in public health.
+Infectious disease forecasting models have emerged as important tools in public health outbreak response.
 Predictions of disease dynamics have been used to inform decision making about a wide variety of measures to reduce disease spread and/or mitigate the severity of disease outcomes.
 For example, estimates of expected onset of flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola and diptheria dynamics have been used to allocate resources \citep{meltzer2014estimating, rainisch2015regional,camacho2015-ebola-bed,finger_real-time_2019}.
-\cite{bertsimas2021predictionsCOVID} developed tools to guide decision making from infectious disease forecasts, which have been used to inform allocation of limited medical supplies such as ventilators, ICU capacity planning, and vaccine distribution strategy.
+\cite{bertsimas2021predictionsCOVID} developed tools to guide decision making from infectious disease forecasts, which have been used to inform allocation of limited resources such as ventilators, ICU capacity, and vaccine supply.
 Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping (University of Texas, 2022)\nocite{utnews2022}.
 %In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
-%The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid} or weighted interval score (WIS) \citep{bracher2021evaluating}.
-However, recently authors have been calling for evaluating forecasts through their impact on policy \citep{marshall2023predictions, bilinski_adaptive_2023}.
 
-In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to.
-We give an overview of these procedures in Section \ref{sec:methods.detailed.decisiontheory}.
-There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
-We review this work only briefly here, and we refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \cite{murphy1993whatisagoodforecast} for discussions focused on applications to economics and meteorology, respectively.
-In finance, the value of forecasts can often be measured by the profits generated by trading decisions informed by the forecasts, perhaps adjusted for risk levels \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
+In the wake of the COVID-19 pandemic, researchers have been calling for evaluating forecasts through their impact on policy \citep{marshall2023predictions, bilinski_adaptive_2023}.
+The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid}, log-score \citep{mcgowan_collaborative_2019}, or weighted interval score (WIS) \citep{bracher2021evaluating}.
+However, as we show in this paper, these metrics only evaluate one specific dimension of accuracy, and often do not have a clear correspondence to features that would make a forecast valuable in the context of a specific decision.
+While these ideas have received little attention in the context of infectious disease forecasting, the results are well known from other fields, such as economics and finance, supply chain management, and meteorology.
+
+In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to (see details in Section \ref{sec:methods.detailed.decisiontheory}).
+%There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
+We refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \cite{murphy1993whatisagoodforecast} for discussions focused on applications to economics and meteorology, respectively.
+%In finance, the value of forecasts can often be measured by the profits generated by trading decisions informed by the forecasts, perhaps adjusted for risk levels \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
 In applications to supply chain management and meteorology, the value of forecasts has typically been operationalized by considering the costs associated with decisions regarding the amount of inventory to hold or the level of protection against the impacts of extreme weather events to enact \cite[e.g.,][]{catt2007assessingcostofforecasterror, petropoulos2019inventoryperformanceforecasting, palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
-For example, in supply chain management these decisions may incur costs related to holding inventory, labor, or providing poor service, while in meteorology we may need to balance the costs of implementing protective measures with the costs of potentially preventable weather damages.
 In this framework, a forecast has value if it leads to decisions with low total costs.
-In all of these fields, analyses have consistently found that common measures of statistical forecast accuracy do not necessarily correspond directly to measures of the value of forecasts as an input to decision making \cite[e.g.,][]{leitch1991economicForecastEval, murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}.
+As one example,
+% in supply chain management these decisions may incur costs related to holding inventory, labor, or providing poor service, while
+in meteorology, policy makers need to balance the costs of implementing protective measures with the costs of potentially preventable weather damages \citep{palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
+Across many fields, analyses have consistently found that common measures of forecast accuracy do not necessarily correspond directly to measures of the value of forecasts as an input to decision making \cite[e.g.,][]{leitch1991economicForecastEval, murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}.
 
-However, we are aware of only a limited body of work that explicitly attempts to measure the value of infectious disease forecasts through their impact on policy, and much of this discussion has proceeded informally.
+We are aware of only a limited body of work that explicitly attempts to measure the value of infectious disease forecasts through their impact on policy.
 For example, \cite{ioannidis2022forecastingCOVIDfailed} discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts.
@@
 
-In practice, probabilistic infectious disease forecasts have most often been made for observations that emerge from public health surveillance systems and have typically been evaluated with standard, ``off-the-shelf'' scoring rules.
+In practice, probabilistic infectious disease forecasts have most often been made for observations that emerge from public health surveillance systems and have typically been evaluated with common scoring rules that have implementations available in existing software packages.
 For example, seasonal influenza forecasts in the US and dengue forecasts for Peru and Puerto Rico targeted public health surveillance measures of incidence over time and space, and used log-score and mean absolute errors to evaluate forecast skill \citep{mcgowan_collaborative_2019,reich_collaborative_2019,johansson_open_2019}.
@@
 Similarly, CRPS was also used to assess probabilistic forecasts of dengue incidence at the district level in Vietnam \citep{colon-gonzalez_probabilistic_2021}.
-While some of these scores can be interpreted through the lens of decision theory, and all of the application-specific papers cited above had authors from public health agencies, none of them make explicit connections between forecast evaluation and how a forecast was used in practice.
+While some of these scores can be interpreted through the lens of decision theory, and all of the application-specific papers cited above have benefited from direct collaboration with public health agencies, none of them make explicit connections between forecast evaluation and how a forecast was used in practice.
 
@@
 We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} and in section \ref{sec:proper} of the
-supplement we show that Bayes scoring rules are proper by construction.
+supplement we demonstrate that Bayes scoring rules are proper by construction (a result also shown by \cite{dawid2007geometry,gneiting2007strictly}).
 

### 2024-01-29 alloscore_manuscript/alloscore-application-shorter.tex
adding comments addressed through section 2.

@@
 
-Infectious disease forecasting models have emerged as important tools in public health.
+Infectious disease forecasting models have emerged as important tools in public health outbreak response.
 Predictions of disease dynamics have been used to inform decision making about a wide variety of measures to reduce disease spread and/or mitigate the severity of disease outcomes.
-For example, estimates of expected onset of flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola dynamics have been used to allocate surveillance resources \citep{meltzer2014estimating, rainisch2015regional}.
-\cite{bertsimas2021predictionsCOVID} developed tools to guide decision making from infectious disease forecasts, which have been used to inform allocation of limited medical supplies such as ventilators, ICU capacity planning, and vaccine distribution strategy.
+For example, estimates of expected onset of flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola and diptheria dynamics have been used to allocate resources \citep{meltzer2014estimating, rainisch2015regional,camacho2015-ebola-bed,finger_real-time_2019}.
+\cite{bertsimas2021predictionsCOVID} developed tools to guide decision making from infectious disease forecasts, which have been used to inform allocation of limited resources such as ventilators, ICU capacity, and vaccine supply.
 Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping (University of Texas, 2022)\nocite{utnews2022}.
 %In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
-%The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid} or weighted interval score (WIS) \citep{bracher2021evaluating}.
-However, recently authors have been calling for evaluating forecasts through their impact on policy \citep{marshall2023predictions, bilinski_adaptive_2023}.
 
-In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to.
-We give an overview of these procedures in Section \ref{sec:methods.detailed.decisiontheory}.
-There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
-We review this work only briefly here, and we refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \cite{murphy1993whatisagoodforecast} for discussions focused on applications to economics and meteorology, respectively.
-In finance, the value of forecasts can often be measured by the profits generated by trading decisions informed by the forecasts, perhaps adjusted for risk levels \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
+In the wake of the COVID-19 pandemic, researchers have been calling for evaluating forecasts through their impact on policy \citep{marshall2023predictions, bilinski_adaptive_2023}.
+The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid}, log-score \citep{mcgowan_collaborative_2019}, or weighted interval score (WIS) \citep{bracher2021evaluating}.
+However, as we show in this paper, these metrics only evaluate one specific dimension of accuracy, and often do not have a clear correspondence to features that would make a forecast valuable in the context of a specific decision.
+While these ideas have received little attention in the context of infectious disease forecasting, the results are well known from other fields, such as economics and finance, supply chain management, and meteorology.
+
+In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to (see details in Section \ref{sec:methods.detailed.decisiontheory}).
+%There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
+We refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \cite{murphy1993whatisagoodforecast} for discussions focused on applications to economics and meteorology, respectively.
+%In finance, the value of forecasts can often be measured by the profits generated by trading decisions informed by the forecasts, perhaps adjusted for risk levels \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
 In applications to supply chain management and meteorology, the value of forecasts has typically been operationalized by considering the costs associated with decisions regarding the amount of inventory to hold or the level of protection against the impacts of extreme weather events to enact \cite[e.g.,][]{catt2007assessingcostofforecasterror, petropoulos2019inventoryperformanceforecasting, palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
-For example, in supply chain management these decisions may incur costs related to holding inventory, labor, or providing poor service, while in meteorology we may need to balance the costs of implementing protective measures with the costs of potentially preventable weather damages.
 In this framework, a forecast has value if it leads to decisions with low total costs.
-In all of these fields, analyses have consistently found that common measures of statistical forecast accuracy do not necessarily correspond directly to measures of the value of forecasts as an input to decision making \cite[e.g.,][]{leitch1991economicForecastEval, murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}.
+As one example,
+% in supply chain management these decisions may incur costs related to holding inventory, labor, or providing poor service, while
+in meteorology, policy makers need to balance the costs of implementing protective measures with the costs of potentially preventable weather damages \citep{palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
+Across many fields, analyses have consistently found that common measures of forecast accuracy do not necessarily correspond directly to measures of the value of forecasts as an input to decision making \cite[e.g.,][]{leitch1991economicForecastEval, murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}.
 
-However, we are aware of only a limited body of work that explicitly attempts to measure the value of infectious disease forecasts through their impact on policy, and much of this discussion has proceeded informally.
+We are aware of only a limited body of work that explicitly attempts to measure the value of infectious disease forecasts through their impact on policy.
 For example, \cite{ioannidis2022forecastingCOVIDfailed} discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts.
@@
 
-% latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Dec 22 10:45:01 2023
+% latex table generated in R 4.3.2 by xtable 1.8-4 package
+% Mon Jan 29 10:57:25 2024
 \begin{table}[ht]
@@
 \section*{Acknowledgements}
+
+We wish to thank the following individuals who contributed valuable comments and feedback on early versions of this work: Matthew Biggerstaff, Rebecca Borchering, Sebastian Funk, Melissa Kerr, and Jeffrey Shaman.
+
 This work has been supported by the National Institutes of General Medical Sciences (R35GM119582) and the U.S.

### 2024-01-29 alloscore_manuscript/alloscore-application-shorter.Rnw
updates and citations for data availability statement.

@@
 All forecast data used in this evaluation are available through the COVID-19 Forecast Hub \citep{cramer_reichlabcovid19-forecast-hub_2021}.
-An R package implementing the allocation score is available at \url{https://github.com/aaronger/alloscore} and all code for the analyses presented in this manuscript is available at \url{https://github.com/aaronger/utility-eval-papers}.
+An R package implementing the allocation score is available at \url{https://github.com/aaronger/alloscore}.
+All code and data for the analyses presented in this manuscript are available at \url{https://github.com/aaronger/utility-eval-papers}.
+The analyses were generated using a reproducible workflow using \Sexpr{version$version.string} and the {\tt targets} package \citep{Rcore-2023, landau_2021_targets}.
 

### 2024-01-29 alloscore_manuscript/alloscore-application-shorter.tex
updates and citations for data availability statement.

@@
 All forecast data used in this evaluation are available through the COVID-19 Forecast Hub \citep{cramer_reichlabcovid19-forecast-hub_2021}.
-An R package implementing the allocation score is available at \url{https://github.com/aaronger/alloscore} and all code for the analyses presented in this manuscript is available at \url{https://github.com/aaronger/utility-eval-papers}.
+An R package implementing the allocation score is available at \url{https://github.com/aaronger/alloscore}.
+All code and data for the analyses presented in this manuscript are available at \url{https://github.com/aaronger/utility-eval-papers}.
+The analyses were generated using a reproducible workflow using R R version 4.3.2 (2023-10-31) and the {\tt targets} package \citep{Rcore-2023, landau_2021_targets}.
 
@@
 % latex table generated in R 4.3.2 by xtable 1.8-4 package
-% Mon Jan 29 10:57:25 2024
+% Mon Jan 29 11:09:55 2024
 \begin{table}[ht]

### 2024-01-29 alloscore_manuscript/alloscore-application-shorter.Rnw
compromise way to make percap

--- 2024-01-29 alloscore_manuscript/alloscore-application-shorter.Rnw
@@
 Predictions of disease dynamics have been used to inform decision making about a wide variety of measures to reduce disease spread and/or mitigate the severity of disease outcomes.
-For example, estimates of expected onset of flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola and diptheria dynamics have been used to allocate resources \citep{meltzer2014estimating, rainisch2015regional,camacho2015-ebola-bed,finger_real-time_2019}.
+For example, estimates of expected onset of flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola dynamics have been used to allocate surveillance resources \citep{meltzer2014estimating, rainisch2015regional}.
 \cite{bertsimas2021predictionsCOVID} developed tools to guide decision making from infectious disease forecasts, which have been used to inform allocation of limited medical supplies such as ventilators, ICU capacity planning, and vaccine distribution strategy.
@@
 This analysis is ``synthetic'' in that it does not correspond to an actual analysis that supported decision making in real time.
-However, the framework described in this paper corresponds to real-world decisions that must be made by public health administrators around the globe, and could be adapted in the future for such real-time situations such as informing bed demand, ventilator usage, or doses needed for specific medications.
-%For example, forecasts for districts in Sierra Leone of bed demand to care for patients with Ebola was the subject of a real-time modeling study in late 2014 and early 2015 \citep{camacho2015-ebola-bed}.
-%And, in 2020, a model developed by an academic research group turned predictions of COVID-19 hospitalizations into estimates of ventilator usage and shortages.
+However, the framework described in this paper corresponds to real-world decisions that must be made by public health administrators around the globe, and could be adapted in the future for such real-time situations.
+For example, forecasts for districts in Sierra Leone of bed demand to care for patients with Ebola was the subject of a real-time modeling study in late 2014 and early 2015 \citep{camacho2015-ebola-bed}.
+And, in 2020, a model developed by an academic research group turned predictions of COVID-19 hospitalizations into estimates of ventilator usage and shortages.
 This framework was used by the Hartford HealthCare system in Connecticut ``to align ventilator supply with projected demand at a time where the [COVID-19] pandemic was on the rise'' \citep{bertsimas2021predictionsCOVID}.
@@
 These scores illustrate a general result: allocation scores for a forecast will tend to be larger when the resource constraint is close to the observed need, because this is when it matters most which locations are allocated more or less resources.
-If the available resources are small compared to the observed need, any allocation of those limited resources will result in a large amount of unmet need.
-If the available resources are comparatively larger, it becomes less important which locations receive relatively more or fewer resources because all locations will receive enough resources to meet their need.
+If the resource constraint is very small, any allocation of those limited resources will result in a large amount of unmet need.
+If the resource constraint is very large, it becomes less important which locations receive relatively more or less resources because all locations will receive enough resources to meet their need.
 In either of these extremes of resource availability, the avoidable unmet need that arises from the allocation suggested by a forecast (i.e., the forecast's allocation score) will tend to be small.
@@
 Larger interval scores indicate less skillful forecasts.
-An interval score consists of (a) the width of the interval, with larger intervals receiving higher scores (higher scores indicate less accuracy), and (b) a penalty if the interval does not cover the eventual observation, which increases the further away the interval is from the observed value.
+An interval score consists of (a) the width of the interval, with larger intervals receiving higher scores, and (b) a penalty if the interval does not cover the eventual observation, which increases the further away the interval is from the observed value.
 Equivalently, the WIS can also be characterized as a weighted sum of quantile scores for each individual predictive quantile.
@@
 tar_load("forecast_data")
+tar_load("percap")
+
+Kat15k_alloscores <- bind_rows(Kat15k_alloscores, percap |> filter(K == 15000))
 
@@
 ref_date_for_plot <- "2021-12-20"
-models_to_plot <- c("JHUAPL-Bucky", "CU-select")
+models_to_plot <- c("JHUAPL-Bucky", "CU-select", "percap")
 
@@
 \section*{Acknowledgements}
-
-We wish to thank the following individuals who contributed valuable comments and feedback on early versions of this work: Matthew Biggerstaff, Rebecca Borchering, Sebastian Funk, Melissa Kerr, and Jeffrey Shaman.
-
 This work has been supported by the National Institutes of General Medical Sciences (R35GM119582) and the U.S.

### 2024-01-29 alloscore_manuscript/alloscore-application-shorter.tex
compromise way to make percap

--- 2023-12-22 alloscore_manuscript/alloscore-application-shorter.tex
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Dec 22 10:45:01 2023
+% Mon Jan 29 15:31:56 2024
 \begin{table}[ht]
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/metrics-over-time-1} \caption[Hospital admissions and evaluation metrics over time]{Hospital admissions and evaluation metrics over time.
-Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022.
-These are the dates for which forecasts were made and evaluated.
-A horizontal dashed line at 15,000 shows the assumed resource constraint $K$.
-Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states.
-The x-axis corresponds to the date for which the prediction was made.
-AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down).
-Panel C shows the MWIS metric across weeks, averaged across all states.
-MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.}\label{fig:metrics-over-time}
-\end{figure}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in `palette()`:\\\#\# ! Insufficient values in manual scale. 18 needed but only 17 provided.}}\end{kframe}
 \end{knitrout}

### 2024-01-29 R/percap.R

@@ -1,23 +1,20 @@
 
-score_per_capita_allocation <- function(
-		all_allscore_data, 
-		Kat15k_alloscores) {
+score_per_capita_allocation <- function(dat, pops, Kmax = 60000) {
   require(tidyverse)
 
-  pops22 <- readr::read_csv("https://raw.githubusercontent.com/reichlab/flusion/main/data-raw/us-census/NST-EST2022-ALLDATA.csv") %>% 
-  	select(full_location_name = NAME, POPESTIMATE2021) %>% 
-  	inner_join(hub_locations, by = join_by(full_location_name))
-
-	percap <- bind_rows(all_alloscore_data, Kat15k_alloscores) %>% 
+	percap <- dat %>% 
 	  filter(model == "COVIDhub-baseline") %>% 
-	  mutate(model = "percap") %>% 
-	  left_join(pops22[c('POPESTIMATE2021', 'abbreviation')], by = "abbreviation") %>% 
+	  select(-K) %>% 
+	  tidyr::expand_grid(K = make_K_grid(Kmax)) %>% 
+	  relocate(K, .after = reference_date) %>%
 	  slice(1, .by = c(reference_date, K ,abbreviation)) %>% 
+	  mutate(model = "percap") %>% 
+	  left_join(pops[c('POPESTIMATE2021', 'abbreviation')], by = "abbreviation") %>% 
 	  mutate(popprop = POPESTIMATE2021 / sum(POPESTIMATE2021), 
 	         .before = x, .by = c(K, reference_date))  %>% 
 	  mutate(x = K * popprop, components_raw = pmax(y-x,0), 
 	         components = components_raw - components_oracle) %>% 
 	  select(-popprop, -POPESTIMATE2021)
 
-	 return(percap)
+	return(percap)
 }

### 2024-01-29 _targets.R

@@ -68,6 +68,12 @@ setup <- list(
                                           truth_data,
                                           reference_dates = values$forecast_dates,
                                           one_K = 15000)
+  ),
+  tar_target(
+    name = pops22,
+    command = readr::read_csv("https://raw.githubusercontent.com/reichlab/flusion/main/data-raw/us-census/NST-EST2022-ALLDATA.csv") |> 
+      dplyr::select(full_location_name = NAME, POPESTIMATE2021) |>
+      dplyr::inner_join(hub_locations, by = join_by(full_location_name))
   )
 )
 
@@ -99,7 +105,7 @@ combined <- tar_combine(
 
 make_percap <- tar_target(
   name = percap,
-  command = score_per_capita_allocation(all_alloscore_data, Kat15k_alloscores)
+  command = score_per_capita_allocation(dat = Kat15k_alloscores, pops = pops22)
 )
 
 list(setup, mapped, combined, make_percap)

### 2024-01-29 alloscore_manuscript/alloscore-application-shorter.Rnw
updating code for Figure 3A

--- 2024-01-29 alloscore_manuscript/alloscore-application-shorter.Rnw
@@
 
+ordered_levels <- truth_data |>
+  filter(target_end_date == as.Date(ref_date_for_plot)+14, location != "US") |>
+  select(location, abbreviation, value) |>
+  arrange(-value)
+
 multiloc_dat <- Kat15k_alloscores |>
@@
   rename(allocation = x, `observed resource need` = y, state = abbreviation) |>
-  select(model, reference_date, state, allocation, `observed resource need`) |>
-  pivot_longer(cols = c("allocation", "observed resource need"))
+  mutate(`unmet resource need` = ifelse(allocation < `observed resource need`,
+                                        `observed resource need` - allocation,
+                                        0),
+         `need met by allocated resources` = ifelse(allocation < `observed resource need`,
+                                                    allocation,
+                                                    `observed resource need`),
+         `allocation exceeding need` = ifelse(allocation < `observed resource need`,
+                                        0,
+                                        allocation - `observed resource need`)
+         ) |>
+  # select(model, reference_date, state, allocation, `observed resource need`) |>
+  select(model, reference_date, state,
+         `unmet resource need`, `need met by allocated resources`, `allocation exceeding need`) |>
+  pivot_longer(cols = c("unmet resource need", "need met by allocated resources", "allocation exceeding need")) |>
+  mutate(state = factor(state, levels = ordered_levels$abbreviation),
+         name = factor(name, levels = c("unmet resource need",
+                                        "allocation exceeding need",
+                                        "need met by allocated resources"))) |>
+  arrange(name)
 
@@
 
-ordered_levels <- truth_data |>
-  filter(target_end_date == as.Date(ref_date_for_plot)+14, location != "US") |>
-  select(location, abbreviation, value) |>
-  arrange(-value)
-
-alloc_plot <- multiloc_dat |>
-  mutate(state = factor(state, levels = ordered_levels$abbreviation),
-         name = factor(name, levels = c("observed resource need", "allocation"))) |>
-  arrange(name) |>
-  ggplot(aes(x=state, y=value, fill=name)) +
-  scale_fill_manual(name=NULL, values = c("#CD2626", "goldenrod")) +
+alloc_plot <- ggplot(multiloc_dat, aes(x=state, y=value, fill=name)) +
+  geom_col(position="stack", alpha=1) +
+  #scale_fill_manual(name=NULL, values = c("#ca0020", "#0571b0", "purple")) +
+  #scale_fill_manual(name=NULL, values = c(  "#e41a1c", "#377eb8", "lightgray")) +
+  scale_fill_manual(name=NULL, values = c("#fb6a4a", "#9ecae1", "#807dba")) +
   theme(axis.text.x=element_text(angle=90,vjust=.5), legend.position = c(.9,.95), legend.justification = c(1,1)) +
@@
   xlab(NULL) +
-  geom_col(position="identity", alpha=.7) +
   facet_grid(model~.)
+
+# alloc_plot <- multiloc_dat |>
+#   mutate(state = factor(state, levels = ordered_levels$abbreviation),
+#          name = factor(name, levels = c("observed resource need", "allocation"))) |>
+#   arrange(name) |>
+#   ggplot(aes(x=state, y=value, fill=name)) +
+#   scale_fill_manual(name=NULL, values = c("#CD2626", "goldenrod")) +
+#   theme(axis.text.x=element_text(angle=90,vjust=.5), legend.position = c(.9,.95), legend.justification = c(1,1)) +
+#   ylab("Resource need and allocation") +
+#   xlab(NULL) +
+#   geom_col(position="identity", alpha=.7) +
+#   facet_grid(model~.)
 

### 2024-01-29 alloscore_manuscript/alloscore-application-shorter.tex
updating code for Figure 3A

--- 2024-01-29 alloscore_manuscript/alloscore-application-shorter.tex
@@
 All code and data for the analyses presented in this manuscript are available at \url{https://github.com/aaronger/utility-eval-papers}.
-The analyses were generated using a reproducible workflow using R R version 4.3.2 (2023-10-31) and the {\tt targets} package \citep{Rcore-2023, landau_2021_targets}.
+The analyses were generated using a reproducible workflow using R version 4.3.2 (2023-10-31) and the {\tt targets} package \citep{Rcore-2023, landau_2021_targets}.
 
@@
 % latex table generated in R 4.3.2 by xtable 1.8-4 package
-% Mon Jan 29 11:09:55 2024
+% Mon Jan 29 16:46:53 2024
 \begin{table}[ht]

### 2024-01-30 alloscore_manuscript/alloscore-application-shorter.Rnw
incorporate percap into a few more plots

--- 2024-01-29 alloscore_manuscript/alloscore-application-shorter.Rnw
@@
 tar_load(alloscore_2021.12.20)
-slim_alloscore_2021.12.20 <- alloscore::slim(alloscore_2021.12.20, id_cols = c("model", "reference_date"))
+slim_alloscore_2021.12.20 <- alloscore::slim(alloscore_2021.12.20, id_cols = c("model", "reference_date")) %>% 
+bind_rows(percap %>% filter(reference_date == "2021-12-20"))
 
@@
   scale_x_date(limits=range(score_data_summaries$target_end_date)) +
-  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
+  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4, 6)) +
+  scale_linewidth_manual(values = c(.5,1.5)) +
   ylab("Allocation score") +
-  geom_point() + geom_line()+
+  geom_point() + geom_line(aes(linewidth = as.factor(model == "percap")))+
   theme(legend.pos="none") +
@@
 
-p4 <- ggplot(score_data_summaries,
+p4 <- ggplot(bind_rows(score_data_summaries, percap %>% 
+                         select(model, reference_date) %>% 
+                         mutate(reference_date = as.Date(reference_date))),
              aes(x=target_end_date, y=mwis, color=model,shape=model)) +
   geom_point() + geom_line() +
-  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
+  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4, 6)) +
   xlab("Forecast target date") +

### 2024-01-30 alloscore_manuscript/alloscore-application-shorter.tex
incorporate percap into a few more plots

--- 2024-01-29 alloscore_manuscript/alloscore-application-shorter.tex
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Mon Jan 29 15:31:56 2024
+% Mon Jan 29 20:07:04 2024
 \begin{table}[ht]
@@
 CU-select & 669 & 133 & 774 & 326 \\ 
+  percap & 865 &  & 1029 & 366 \\ 
   COVIDhub-ensemble & 873 & 159 & 1067 & 438 \\ 
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\includegraphics[width=\maxwidth]{figure/metrics-over-time-1} \caption[Hospital admissions and evaluation metrics over time]{Hospital admissions and evaluation metrics over time.
+Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022.
+These are the dates for which forecasts were made and evaluated.
+A horizontal dashed line at 15,000 shows the assumed resource constraint $K$.
+Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states.
+The x-axis corresponds to the date for which the prediction was made.
+AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down).
+Panel C shows the MWIS metric across weeks, averaged across all states.
+MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.}\label{fig:metrics-over-time}
+\end{figure}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in `palette()`:\\\#\# ! Insufficient values in manual scale. 18 needed but only 17 provided.}}\end{kframe}
 \end{knitrout}

### 2024-01-30 R/percap.R

@@ -10,10 +10,13 @@ score_per_capita_allocation <- function(dat, pops, Kmax = 60000) {
 	  slice(1, .by = c(reference_date, K ,abbreviation)) %>% 
 	  mutate(model = "percap") %>% 
 	  left_join(pops[c('POPESTIMATE2021', 'abbreviation')], by = "abbreviation") %>% 
-	  mutate(popprop = POPESTIMATE2021 / sum(POPESTIMATE2021), 
-	         .before = x, .by = c(K, reference_date))  %>% 
-	  mutate(x = K * popprop, components_raw = pmax(y-x,0), 
-	         components = components_raw - components_oracle) %>% 
+	  group_by(K, reference_date) %>% 
+	  mutate(popprop = POPESTIMATE2021 / sum(POPESTIMATE2021), .before = x)  %>% 
+	  mutate(x = K * popprop, components_raw = pmax(y-x,0),
+	  	oracle = y * K / sum(y),
+	  	components_oracle = pmax(y - oracle, 0),
+	  	components = components_raw - components_oracle) %>% 
+	  ungroup() %>% 
 	  select(-popprop, -POPESTIMATE2021)
 
 	return(percap)

### 2024-01-30 alloscore_manuscript/alloscore-application-shorter.Rnw
adding updates to the thermometer plot code.

@@
               geo_value = abbreviation),
-          f_colors = c("#4A708B", "#CD2626"),
+          f_colors = c("thistle", "wheat"),
           one_K = 15000) + labs(x = NULL)

### 2024-01-31 alloscore_manuscript/alloscore-application-shorter.Rnw
adding some additional discussion comments.

@@
 For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts.
-The side-by-side shaded regions show the median (solid horizontal line) and 50%, 80% and 95% prediction intervals for the two selected models.
+The side-by-side shaded regions show the median (solid red horizontal line) and 50%, 80% and 95% prediction intervals for the two selected models.
 The forecasts were made for new hospitalizations on January 3, 2022 (vertical dashed line, with number of hospitalizations indicated by red horizontal line segment at the intersection of the dashed line and the grey line of data).
-The vertical bars with red and yellow shading show the allocations: the red bar goes from zero to the level of the observed number of hospitalizations while the yellow bar shows how many resources were suggested by the model to be allocated to that location.
-The allocations (yellow bar) may be more or less than the observed number, therefore exposed red bar indicates unmet need, while the yellow bar extending above the observed number of hospitalizations means that excess resources were suggested for that given location.">>=
+The vertical bars with purple, blue and red shading show the allocations.
+The purple bar goes from zero to the amount of need that was met by the allocation from that model to a specific location.
+A red bar indicates need that exceeded the resource allocation for a location, and a light blue bar shows suggested resource allocation suggested by that model that were above the need for that location.">>=
 source("R/plot_functions.R")
@@
 With careful thought and collaboration between modelers and public health officials, we argue that scores that are more aligned with public health decisions could be developed to inform specific problems.
-We have demonstrated that forecast evaluation methods that are tied to a specific decision making context can yield model rankings that differ substantively from those based on standard measures of forecast skill.
+We have demonstrated that forecast evaluation methods that are tied to a specific decision making context can yield model rankings that differ substantively from those based on standard measures of forecast skill, a result that aligns with findings in other fields \citep{leitch1991economicForecastEval, murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}.
 In particular, the allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.
@@
 While our example used a hypothetical scenario to illustrate a potential application of the allocation score, resource allocation decisions are a realistic example of the kinds of decisions that could motivate more targeted forecasting exercises.
-One real-world example, which was the motivating example presented in the work above, is allocation of ventilators during a respiratory viral pandemic\citep{huang_stockpiling_2017}.
-Other examples include the allocation of a limited stockpile of vaccinations\citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests\citep{du_optimal_2022,pasco_covid-19_2023}.
+%One real-world example, which was the motivating example presented in the work above, is allocation of ventilators during a respiratory viral pandemic\citep{huang_stockpiling_2017}.
+We used ventilators as a notional example for a resource allocation problem because it (a) has been identified as a scarce resource during outbreaks\citep{huang_stockpiling_2017}, (b) could be allocated across different locations, with logistical considerations depending on the geographical scales, and (c) would likely be correlated (although not perfectly so) with hospitalization data.
+However, we acknowledge the limitations of this example in that not everyone who is hospitalized needs a ventilator, and new hospital admissions may not be the best stand in for a need that corresponds to a prevalence of severe cases.
+Other possible real-world examples include the allocation of a limited stockpile of vaccinations\citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests\citep{du_optimal_2022,pasco_covid-19_2023}.
+While there are limitations in the realism of the pairing of a specific resource and forecasted outcome, we emphasize that the important result from this work is how evaluating forecasts using a fundamentally different kind of score can illuminate previously hidden aspects of value in forecasts.
 
 In practice, there are many potential users of forecasts with many different decision making problems.
-Not all can be easily modeled as an effort to minimize the expectation of a transparent loss function.
+While outcomes of some decisions could potentially be specified by loss functions, other uses for forecasts may be less quantifiable, such as when forecasts are used to inform a general sense of situational awareness that may feed in unmeasurable ways into other decisions.
 And even those that are easily framed as expected loss minimization may differ enough that no single score would be
@@
 In conclusion, we argue that the way modelers and policymakers view and evaluate forecasts should change depending on the specific decision-making context.
-Defaulting to standard forecast evaluation metrics can mask the utility (or disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those for which they were intended.
+Defaulting to standard forecast evaluation metrics can mask the utility (or disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those for which they might offer useful guidance.
 New collaborative work between public health officials and modeling teams is needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or simulation exercises that could be used to inform further development of new or alternative scoring metrics.

### 2024-01-31 alloscore_manuscript/alloscore-application-shorter.tex
adding some additional discussion comments.

@@
 
-The hospitalization data used for scoring forecasts were downloaded on December 03, 2023.
+The hospitalization data used for scoring forecasts were downloaded on January 30, 2024.
 
@@
 % latex table generated in R 4.3.2 by xtable 1.8-4 package
-% Mon Jan 29 16:46:53 2024
+% Wed Jan 31 15:25:28 2024
 \begin{table}[ht]
@@
 For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts.
-The side-by-side shaded regions show the median (solid horizontal line) and 50\%, 80\% and 95\% prediction intervals for the two selected models.
+The side-by-side shaded regions show the median (solid red horizontal line) and 50\%, 80\% and 95\% prediction intervals for the two selected models.
 The forecasts were made for new hospitalizations on January 3, 2022 (vertical dashed line, with number of hospitalizations indicated by red horizontal line segment at the intersection of the dashed line and the grey line of data).
-The vertical bars with red and yellow shading show the allocations: the red bar goes from zero to the level of the observed number of hospitalizations while the yellow bar shows how many resources were suggested by the model to be allocated to that location.
-The allocations (yellow bar) may be more or less than the observed number, therefore exposed red bar indicates unmet need, while the yellow bar extending above the observed number of hospitalizations means that excess resources were suggested for that given location.}\label{fig:thermometer-plot}
+The vertical bars with purple, blue and red shading show the allocations.
+The purple bar goes from zero to the amount of need that was met by the allocation from that model to a specific location.
+A red bar indicates need that exceeded the resource allocation for a location, and a light blue bar shows suggested resource allocation suggested by that model that were above the need for that location.}\label{fig:thermometer-plot}
 \end{figure}

### 2024-02-01 alloscore_manuscript/alloscore-application-shorter.Rnw
updating bibliography and citations.
- changing book chapter to display correctly
- removing URLs and ISSNs (kept DOIs)
- fixing capitalization in journal and article titles
- updating citation item classifications to show
- info correctly (e.g. something classified as @misc)...
udpating name of percap to per-capita and changing throughout manuscript

@@
 
+\subsection{Allocation benchmarks}
+
+To adequately compare allocations from individual models with existing standard practices, we used two benchmarks against which the performance of using forecasts to inform allocations could be evaluated.
+First, we evaluated forecasts from the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, and the allocations that resulted from those forecasts \citep{cramer_evaluation_2022}.
+Second, we generated proposed set of allocations on a per capita basis by state, referred to below as \texttt{per-capita}.
+These two approaches generally reflect common choices for ``best practices'' of allocation: either allocating resources based on the most recent observed data, or in proportion to the population of each location.
+As such, they serve as good comparisons for allocations based on forecast models.
+
 \subsection{Data and code availability}
 
-All forecast data used in this evaluation are available through the COVID-19 Forecast Hub \citep{cramer_reichlabcovid19-forecast-hub_2021}.
+All forecast data used in this evaluation are available through the COVID-19 Forecast Hub \citep{cramer_united_2022}.
 An R package implementing the allocation score is available at \url{https://github.com/aaronger/alloscore}.
@@
   arrange(AS) |>
-  xtable::xtable(caption = "Comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS).
+  xtable::xtable(caption = "For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS).
 All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03.
@@
 ref_date_for_plot <- "2021-12-20"
-models_to_plot <- c("JHUAPL-Bucky", "CU-select", "percap")
+models_to_plot <- c("JHUAPL-Bucky", "CU-select", "per-capita")
 
@@
   ylab("Allocation score") +
-  geom_point() + geom_line(aes(linewidth = as.factor(model == "percap")))+
+  geom_point() + geom_line(aes(linewidth = as.factor(model == "per-capita")))+
   theme(legend.pos="none") +
@@
 all_scores |>
-  filter(as.Date(reference_date.x) < as.Date("2022-02-01") ) |>
+  filter(as.Date(reference_date.x) < as.Date("2022-02-01"),
+         model != "per-capita") |>
   dplyr::mutate(target_end_date = as.Date(target_end_date)) |>

### 2024-02-01 alloscore_manuscript/alloscore-application-shorter.tex
updating bibliography and citations.
- changing book chapter to display correctly
- removing URLs and ISSNs (kept DOIs)
- fixing capitalization in journal and article titles
- updating citation item classifications to show
- info correctly (e.g. something classified as @misc)...
udpating name of percap to per-capita and changing throughout manuscript

@@
 % latex table generated in R 4.3.2 by xtable 1.8-4 package
-% Wed Jan 31 15:25:28 2024
+% Thu Feb  1 12:24:44 2024
 \begin{table}[ht]
@@
 CU-select & 669 & 133 & 774 & 326 \\ 
-  percap & 865 &  & 1029 & 366 \\ 
+  per-capita & 865 &  & 1029 & 366 \\ 
   COVIDhub-ensemble & 873 & 159 & 1067 & 438 \\ 
@@
 \end{tabular}
-\caption{Comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS).
+\caption{For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS).
 All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03.

### 2024-02-01 R/percap.R

@@ -2,21 +2,21 @@
 score_per_capita_allocation <- function(dat, pops, Kmax = 60000) {
   require(tidyverse)
 
-	percap <- dat %>% 
-	  filter(model == "COVIDhub-baseline") %>% 
-	  select(-K) %>% 
-	  tidyr::expand_grid(K = make_K_grid(Kmax)) %>% 
+	percap <- dat %>%
+	  filter(model == "COVIDhub-baseline") %>%
+	  select(-K) %>%
+	  tidyr::expand_grid(K = make_K_grid(Kmax)) %>%
 	  relocate(K, .after = reference_date) %>%
-	  slice(1, .by = c(reference_date, K ,abbreviation)) %>% 
-	  mutate(model = "percap") %>% 
-	  left_join(pops[c('POPESTIMATE2021', 'abbreviation')], by = "abbreviation") %>% 
-	  group_by(K, reference_date) %>% 
-	  mutate(popprop = POPESTIMATE2021 / sum(POPESTIMATE2021), .before = x)  %>% 
+	  slice(1, .by = c(reference_date, K ,abbreviation)) %>%
+	  mutate(model = "per-capita") %>%
+	  left_join(pops[c('POPESTIMATE2021', 'abbreviation')], by = "abbreviation") %>%
+	  group_by(K, reference_date) %>%
+	  mutate(popprop = POPESTIMATE2021 / sum(POPESTIMATE2021), .before = x)  %>%
 	  mutate(x = K * popprop, components_raw = pmax(y-x,0),
 	  	oracle = y * K / sum(y),
 	  	components_oracle = pmax(y - oracle, 0),
-	  	components = components_raw - components_oracle) %>% 
-	  ungroup() %>% 
+	  	components = components_raw - components_oracle) %>%
+	  ungroup() %>%
 	  select(-popprop, -POPESTIMATE2021)
 
 	return(percap)

### 2024-02-01 alloscore_manuscript/alloscore-application-shorter.Rnw
updating NA string for tables.

@@
 For all metrics, lower scores indicate better accuracy.",
+                 NA.string = "-",
                  digits = 0,
@@
 
+<<table-score-summary-all-weeks, results='asis'>>=
+all_scores |>
+  filter( !(reference_date.x %in% c("2021-11-22", "2022-02-28"))) |>
+  group_by(model) |>
+  summarize(N = sum(!is.na(alloscore)),
+            MWIS = mean(mwis, na.rm=TRUE),
+            MAS = mean(alloscore, na.rm=TRUE)) |>
+  filter(N > 12) |>
+  arrange(MAS) |>
+  select(model, MWIS, MAS) |>
+  xtable::xtable(caption = "Mean weighted interval scores (MWIS) and mean allocation scores (MAS) by model across 13 weeks.
+For the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, plus a per-capita allocation, these results show the average performance of each model across all weeks.
+Lower scores indicate better performance, and models are sorted with the most accurate allocations at the top.",
+                 digits = 0,
+                 label = "tab:multi-week-performance-summary") |>
+  print(include.rownames = FALSE, NA.string = "-")
+@
+
 \subsubsection{Integrated allocation score across values of K}

### 2024-02-01 alloscore_manuscript/alloscore-application-shorter.tex
updating NA string for tables.

@@
 % latex table generated in R 4.3.2 by xtable 1.8-4 package
-% Thu Feb  1 12:24:44 2024
+% Thu Feb  1 13:48:15 2024
 \begin{table}[ht]
@@
 
+% latex table generated in R 4.3.2 by xtable 1.8-4 package
+% Thu Feb  1 13:48:21 2024
+\begin{table}[ht]
+\centering
+\begin{tabular}{lrr}
+  \hline
+model & MWIS & MAS \\ 
+  \hline
+COVIDhub-ensemble & 70 & 389 \\ 
+  per-capita &  & 464 \\ 
+  COVIDhub-trained\_ensemble & 87 & 483 \\ 
+  CU-select & 81 & 502 \\ 
+  JHUAPL-SLPHospEns & 67 & 526 \\ 
+  COVIDhub-baseline & 93 & 594 \\ 
+  JHUAPL-Bucky & 112 & 643 \\ 
+  MUNI-ARIMA & 116 & 707 \\ 
+  JHUAPL-Gecko & 110 & 929 \\ 
+  USC-SI\_kJalpha & 155 & 1473 \\ 
+   \hline
+\end{tabular}
+\caption{Mean weighted interval scores (MWIS) and mean allocation scores (MAS) by model across 13 weeks.
+For the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, plus a per-capita allocation, these results show the average performance of each model across all weeks.
+Lower scores indicate better performance, and models are sorted with the most accurate allocations at the top.}
+\label{tab:multi-week-performance-summary}
+\end{table}
+
 \subsubsection{Integrated allocation score across values of K}

### 2024-02-01 alloscore_manuscript/alloscore-application-shorter.Rnw
grab pop directly from census.gov add ref for using census data for pwr-capita adjusted legend in fig 3 so it doesn't block smaller states in CU-select in panel A

--- 2024-02-01 alloscore_manuscript/alloscore-application-shorter.Rnw
@@
 First, we evaluated forecasts from the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, and the allocations that resulted from those forecasts \citep{cramer_evaluation_2022}.
-Second, we generated proposed set of allocations on a per capita basis by state, referred to below as \texttt{per-capita}.
+Second, we generated a proposed set of allocations on a per capita basis by state (using US Census data of vintage 2022 \citep{us_census_nst_est2022}), referred to below as \texttt{per-capita}.
 These two approaches generally reflect common choices for ``best practices'' of allocation: either allocating resources based on the most recent observed data, or in proportion to the population of each location.
@@
   scale_fill_manual(name=NULL, values = c("#fb6a4a", "#9ecae1", "#807dba")) +
-  theme(axis.text.x=element_text(angle=90,vjust=.5), legend.position = c(.9,.95), legend.justification = c(1,1)) +
+  theme(axis.text.x=element_text(angle=90,vjust=.5), legend.position = c(.95,.98), legend.justification = c(1,1), legend.key.size = unit(0.45, "cm"), legend.background = element_rect(fill = "transparent")) +
   ylab("Resource need and allocation") +
@@
   facet_grid(model~.) +
-  theme(axis.text.x=element_text(angle=90,vjust=.5), legend.position = c(.9,.95), legend.justification = c(1,1)) +
+  theme(axis.text.x=element_text(angle=90,vjust=.5), legend.position = c(.82,.95), legend.justification = c(1,1), legend.key.size = unit(0.45, "cm"), legend.background = element_rect(fill = "transparent")) +
   xlab(NULL) +

### 2024-02-01 alloscore_manuscript/alloscore-application-shorter.tex
grab pop directly from census.gov add ref for using census data for pwr-capita adjusted legend in fig 3 so it doesn't block smaller states in CU-select in panel A

--- 2024-02-01 alloscore_manuscript/alloscore-application-shorter.tex
@@
 
-The hospitalization data used for scoring forecasts were downloaded on January 30, 2024.
+The hospitalization data used for scoring forecasts were downloaded on February 01, 2024.
 
@@
 All code and data for the analyses presented in this manuscript are available at \url{https://github.com/aaronger/utility-eval-papers}.
-The analyses were generated using a reproducible workflow using R version 4.3.2 (2023-10-31) and the {\tt targets} package \citep{Rcore-2023, landau_2021_targets}.
+The analyses were generated using a reproducible workflow using R version 4.3.1 (2023-06-16) and the {\tt targets} package \citep{Rcore-2023, landau_2021_targets}.
 
@@
 
-% latex table generated in R 4.3.2 by xtable 1.8-4 package
-% Thu Feb  1 12:24:44 2024
+% latex table generated in R 4.3.1 by xtable 1.8-4 package
+% Thu Feb  1 16:28:56 2024
 \begin{table}[ht]

### 2024-02-01 _targets.R

@@ -71,7 +71,7 @@ setup <- list(
   ),
   tar_target(
     name = pops22,
-    command = readr::read_csv("https://raw.githubusercontent.com/reichlab/flusion/main/data-raw/us-census/NST-EST2022-ALLDATA.csv") |> 
+    command = readr::read_csv("https://www2.census.gov/programs-surveys/popest/datasets/2020-2022/state/totals/NST-EST2022-ALLDATA.csv") |> 
       dplyr::select(full_location_name = NAME, POPESTIMATE2021) |>
       dplyr::inner_join(hub_locations, by = join_by(full_location_name))
   )

### 2024-02-01 alloscore_manuscript/alloscore-application-shorter.tex

@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Thu Feb  1 16:28:56 2024
+% Thu Feb  1 16:41:23 2024
 \begin{table}[ht]
@@
 
+% latex table generated in R 4.3.1 by xtable 1.8-4 package
+% Thu Feb  1 16:41:28 2024
+\begin{table}[ht]
+\centering
+\begin{tabular}{lrr}
+  \hline
+model & MWIS & MAS \\ 
+  \hline
+COVIDhub-ensemble & 70 & 389 \\ 
+  per-capita & - & 464 \\ 
+  COVIDhub-trained\_ensemble & 87 & 483 \\ 
+  CU-select & 81 & 502 \\ 
+  JHUAPL-SLPHospEns & 67 & 526 \\ 
+  COVIDhub-baseline & 93 & 594 \\ 
+  JHUAPL-Bucky & 112 & 643 \\ 
+  MUNI-ARIMA & 116 & 707 \\ 
+  JHUAPL-Gecko & 110 & 929 \\ 
+  USC-SI\_kJalpha & 155 & 1473 \\ 
+   \hline
+\end{tabular}
+\caption{Mean weighted interval scores (MWIS) and mean allocation scores (MAS) by model across 13 weeks.
+For the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, plus a per-capita allocation, these results show the average performance of each model across all weeks.
+Lower scores indicate better performance, and models are sorted with the most accurate allocations at the top.}
+\label{tab:multi-week-performance-summary}
+\end{table}
+
 \subsubsection{Integrated allocation score across values of K}

### 2024-02-05 alloscore_manuscript/oup-authoring-template/doc/oup-authoring-template-doc.tex
add series A papers

%%
%% Copyright 2020 OXFORD UNIVERSITY PRESS
%%
%% This file is part of the 'oup-authoring-template Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'oup-authoring-template Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for OXFORD UNIVERSITY PRESS's document class `oup-authoring-template'
%% with bibliographic references
%%
\documentclass{article}

\usepackage{xspace}
\usepackage{fancyvrb}
\usepackage{hyperref}

\begin{document}

\title{\bfseries User Manual\\
OUP Journals---Authoring template}

\author{Copyright Oxford University Press 2020\\
\small Prepared by SPi \TeX\ Support}

\maketitle

\tableofcontents

\section{Introduction }
Oxford University Press has developed this authoring template to
help authors in preparing articles. Authors are encouraged to use
this template to produce \LaTeX\ manuscripts which conform to OUP
Journal styles. This document is a manual for authors to help using
this template during article preparation. It features general
guidelines and contains descriptions regarding various elements that
can be used while preparing manuscripts. Authors are requested to
refer to the file ``readme'' for details of files available for
reference.

Please utilize the OUP authoring template to the maximum, rather
than adding further packages or macros if possible. Please use
semantic mark-up instead of formatting markup where there is a
choice, for instance using \verb+\emph+ rather than \verb+\it+ or
\verb+\textit+ to indicate emphasis.

This documentation is not intended to give an introduction to \LaTeX. For questions concerning \TeX\ systems/installations or the \LaTeX\ mark-up language in general, please visit \url{http://tug.ctan.org/} or any other \TeX\ user group worldwide. The essential reference for \LaTeX\ is Mittelbach F., Goossens M. (2004) \textit{The \LaTeX\ Companion. 2nd edn.}, but there are many other good books about \LaTeX.

\section{Quick reference guide}
\begin{itemize}
\item How to download and install - \url{https://miktex.org/download}
\item Any system requirements - \url{https://miktex.org/kb/prerequisites-2-9}
\item Quick steps to get started - \url{http://users.dickinson.edu/~richesod/latex/latexcheatsheet.pdf}\newline
\url{http://wch.github.io/latexsheet/latexsheet.pdf}\newline
\url{https://www.overleaf.com/learn}
\end{itemize}

\vfill\eject

\section{How to start and prepare your article}
It is assumed that you possess basic knowledge of \LaTeX. Unless
using \LaTeX\ online (for example via Overleaf.com) ensure that you
have \LaTeX2e version installed on your computer. We suggest
employing a recent \TeX installation. You are provided with a class
file ``oup-authoring-template.cls''. This template can be kept with
your manuscript files. Note that the class file depends on the
following packages which are standard and are available along with
\LaTeX installation:

\begin{center}\begin{tabular}{|l|l|l|l|l|}\hline
graphicx & multirow & amsmath & amssymb & amsfonts \\\hline
array & flushend & stfloats & color & xcolor \\\hline
rotating & chngpage & totcount & fix-cm & algorithm \\\hline
algorithmicx & algpseudocode & listings & url & crop \\\hline
\end{tabular}\end{center}

Apart from the above-listed packages, the following additional packages are used in the class file for providing add-on functionalities to the template:

\begin{center}\begin{tabular}{|c|c|c|c|c|}\hline
subfloat & subfig & tikz & hyperref & footnote \\\hline
mathrsfs & natbib & wrapfig & amsthm & apacite \\\hline
\end{tabular}\end{center}

To learn more about the underlying packages, please read the respective documentations (try, e.g., texdoc [package name] at your shell prompt or visit \url{http://tug.ctan.org/}).

\medskip

\noindent You are given with sample tex file: oup-authoring-template.tex

\medskip

Based on your journal style, we would suggest you to use the corresponding sample template file to start your manuscript preparation; other templates available in the package can be removed. Please save a copy of the template file before you start editing the file as per your requirement. These samples contain the lines for calling class files, the preamble area, and the start/end of document where major sample elements required for an article are placed. Comments are included for each element and are self-explanatory. You can add your actual manuscript content in place of these sample elements. The standard structure of each element for an article is explained in detail in the following sections.

To use the OUP authoring template, put all the package files in your
working directory, edit the file ``sample template file'' in your
preferred text editor, and run \LaTeX\ as usual. The resulting
layout is similar but not identical to the layout of the final
article.

Please note that you are not responsible for any final page layout. It is not necessary (and is even sometimes undesirable) to do any fine-tuning with commands like \verb+\break+, \verb+\pagebreak+, \verb+\vspace+, \verb+\clearpage+, etc. Please use semantic mark-up as far as possible and avoid additional formatting commands.

\vfill\eject

\section{Package features and some important settings}

\subsection{Language}
English is the default language used for typesetting rules.

\subsection{Fonts}
Please refrain from using custom fonts.

\medskip

\noindent Text fonts: Unlike the final published version, the authoring template uses non-commercial fonts: Computer Modern. These fonts are free version of the PostScript standard fonts and are supplied as part of all standard \TeX\ distributions.

\medskip

\noindent Math fonts: The standard Computer Modern math fonts are used.

\section{Preamble}
The preamble part comes between the document class line---\newline\verb+\documentclass{...}+---and the beginning of your document---\verb+\begin{document}+. Use this preamble area to include additional packages and customized macros if any.

\section{Details on document class options}

\begin{enumerate}
\item[a.] Options available to select three different types of layout:
\begin{itemize}
\item Modern
\item Traditional
\item Contemporary
\end{itemize}
\item[b.] Options available to select three different paper sizes:
\begin{itemize}
\item Large
\item Medium
\item Small
\end{itemize}
\item[c.] namedate --- for authoryear citation style; default reference citation style is numbered reference style
\item[d.] webpdf --- for cropped paper size in the PDF output
\item[e.] unnumsec --- to get unnumbered section heads
\end{enumerate}

\section{Major structures/elements}

Article contents are divided into three main elements—front matter, main matter, and back matter. The elements preceding the \verb+\maketitle+ tag are considered as front matter elements and the elements placed below the \verb+\maketitle+ tag are main matter elements. Elements found after the section ``Conclusion'' are considered as back matter elements.

\begin{center}\tabcolsep2pt\fontsize{6}{9}\selectfont\begin{tabular}{|l|l|l|}\hline
\textbf{Front matter}&   \textbf{Main matter}&    \textbf{Back matter}\\
\hline
\verb+\title{...}+&     \verb+\section{...}+&   \verb+\begin{appendices}...\end{appendices}+\\\hline
\verb+\author{...}+&    \verb+\subsection{...}+&        \verb+\bibliographystyle{...}+\\\hline
\verb+\authormark{...}+&  \verb+\subsubsection{...}+&     \verb+\bibliography{...}+\\\hline
\verb+\address{...}+&   \verb+\paragraph{...}+& \verb+\begin{biography}{...}{...}\end{biography}+\\\hline
\verb+\corresp[]{...}+&   \verb+\begin{algorithm}...\end{algorithm}+&     \\\hline
\verb+\received{...}{...}{...}+&    \verb+\begin{lstlisting}...\end{lstlisting}+&      \\\hline
\verb+\revised{...}{...}{...}+&     \verb+\begin{table}...\end{table}+&      \\\hline
\verb+\accepted{...}{...}{...}+&    \verb+\begin{figure}...\end{figure}+&    \\\hline
\verb+\abstract{...}+&  \verb+\begin{enumerate}...\end{enumerate}+&      \\\hline
\verb+\keywords{...}+&  \verb+\begin{unlist}...\end{unlist}+&    \\\hline
\verb+\boxedtext{...}+&   \verb+\begin{itemize}...\end{itemize}+&  \\\hline
\verb+\editor{...}+&      \verb+\begin{theorem}...\end{theorem}+&  \\\hline
\verb+ \maketitle+&     \verb+\begin{proposition}...\end{proposition}+&    \\\hline
 &      \verb+\begin{example}...\end{example}+&    \\\hline
 &      \verb+\begin{definition}...\end{definition}+&    \\\hline
 &      \verb+\begin{proof}...\end{proof}+&      \\\hline
&       \verb+\begin{equation}...\end{equation}+&        \\
\hline
\end{tabular}\end{center}

\section{Front matter elements }
Required elements may vary depending on the journal to which you plan to submit. Check the instructions on the journal’s webpage carefully. The tagging details of article opener elements are as follows:
\begin{enumerate}
\item \verb+\title[<short-form-of-article-title>]{<article-title>} +

This tag contains two parameters, first one is optional and the second argument is mandatory. By default, the article title is printed as running heads on both odd/even pages. In case of lengthy article title, provide the short form of article title in the optional argument.

\item \verb+\author[<address-num>]{<author-name>}+---to be used for the authors other than the corresponding author.

\verb+\author[<address-num>,$\ast$]{<author-name>}+---to be used for the corresponding author who is nominated as being responsible for the\break manuscript as it moves through the entire publication process. He is the “time keeper” during each phase of the publication process and the primary contact between the journal and all the other authors of the paper.

\item \verb+\address[<sequence>]{<address-details>}+---affiliation/address details are provided inside this tag. In case of multiple addresses, just provide sequential Arabic numerals in the optional argument of this tag. This numeral is used to denote the affiliation for the respective authors. In case of single author/address, this optional argument can be ignored. For example, refer below:

\verb+\author{...}+

\verb+\address{...}+

\item \verb+\corresp[$\ast$]{...}+— provide corresponding author's email id inside this tag.

\item The other tags listed below are self-explanatory. Note the `received', `revised' and `accepted' dates are placeholders for the final publication: these can be left blank:

{\tabcolsep3pt\fontsize{7}{10}\selectfont\begin{tabular}{|l|l|l|l|l|}
\hline \verb+\orgdiv{. . .}+ & \verb+\orgname{. . .}+ &
\verb+\orgaddress{. . .}+ & \verb+\country{. . .}+ &
\verb+\postcode{. . .}+ \\ \hline \verb+\street{. . .}+ &
\verb+\city{. . .}+ & \verb+\state{. . .}+ & \verb+\abstract{. . .}+
& \verb+\keywords{. . .}+ \\ \hline \verb+\received{. . .}+ &
\verb+\revised{. . .}+ & \verb+\accepted{. . .}+ &  &  \\ \hline
\end{tabular}}

\item \verb+\ORCID{...}+-- provide ORCID link with logo inside this tag.

\item \verb+\maketitle+---this tag is mandatory to print the front matter elements in the output.

\end{enumerate}

\section{Main matter elements }
\subsection{Section headings}
The template allows four levels of headings in different styles:

\medskip

\verb+\section{<First level heading>} +

\verb+\subsection{<Second level heading>} +

\verb+\subsubsection{<Third level heading>}+

\verb+\paragraph{<Fourth level heading>} +

\medskip

\noindent To get unnumbered level heads, provide ``unnumsec'' option to\newline \verb+\documentclass[unnumsec]{oup-authoring-template}+.

\subsection{Mathematical formulae}

As the ``amsmath'' package provides various features for displayed
equations and other mathematical constructs and you are strongly
encouraged to use the mark-ups provided by this package. If
possible, avoid using manual skips to align an equation.

\subsection{Figures and Tables}
The standard interface for graphic inclusion is the \verb+\includegraphics+ command provided by the graphicx package. The ``draft'' option provided by this package locally switches to draft mode, i.e. does not include the graphic, but leaves the correct space, and prints the filename. This option may be used to save the processing time during compilation. Note that the \verb+\graphicspath+ command allows you to declare one or more folders where the graphicx package looks for the image files; hence, it is not necessary to write the path into each \verb+\includegraphics+ command.

The format used for numbered ``figures/tables'' is similar to the basic \LaTeX\ format:

\begin{verbatim}
\begin{figure}[t]
\centering
\includegraphics{<image-file-name>}
\caption{<figure caption text>}\label{...}
\end{figure}
\end{verbatim}

In case of double column layout, the above format puts figure captions/images to a single column width. To get spanned images, we need to provide the environment
\verb+\begin{figure*}...\end{figure*}+.

For the purpose of the sample, we have included the width of images in the optional argument of the \verb+\includegraphics+ tag. Please ignore this.

Images exceeding the text width should be set as rotated images. For this, we need to use \verb+\begin{sidewaysfigure}...\end{sidewaysfigure}+ instead of the \verb+\begin{figure}...\end{figure}+ environment. In case of double column layout, this format puts figure captions/images to single column width. To get spanned rotated images, use\newline \verb+\begin{sidewaysfigure*}...\end{sidewaysfigure*}+.

If the journal requires that you include a graphical abstract, include this as an unnumbered figure directly after the text abstract with no caption, but adding the \verb+\caption+ package to the preamble and including the image as follows (note the file name for the graphical abstract):

\begin{verbatim}
\begin{figure}[t]
\centering
\includegraphics{graphical_abstract.jpg}
\captionsetup{labelformat=empty}
\caption{}
\label{...}
\end{figure}
\end{verbatim}

The format for table is as follows:

\begin{verbatim}
\begin{table}[<float-position>]
\begin{center}
\begin{minipage}{<specify-table-width>}
\caption{...}\label{<table-label>}
\begin{tabular}{<column-alignment-preamble>}
\toprule
...  & ... & ... & ... \\
\midrule
... & ... & ... & ...\footnotemark{1} \\
... & ... & ... & ... \\
... & ... & ... & ...\footnotemark{2} \\
\botrule
\end{tabular}
\footnotetext{...}
\footnotetext[1]{...}
\footnotetext[2]{...}
\end{minipage}
\end{center}
\end{table}
\end{verbatim}

Command to be used for rotated tables:

\begin{verbatim}
\begin{sidewaystable}...\end{sidewaystable}
\end{verbatim}

To span tables across columns in double column layout:

\begin{verbatim}
\begin{table*}...\end{table*}
\end{verbatim}

To span rotated tables across columns in double column layout:

\begin{verbatim}
\begin{sidewaystable*}...\end{sidewaystable*}
\end{verbatim}

\subsection{Lists}
The default list commands available in \LaTeX\ can be used to set different types of lists:

\begin{enumerate}
\item numbered: \verb+\begin{enumerate}...\end{enumerate}+
\item[] first level Arabic numerals;
\item[]second level lowercase alphabet;
\item[]third level lowercase roman numerals;
\item unnumbered: \verb+\begin{unlist}...\end{unlist}+
\item custom list: \verb+\begin{itemize}...\end{itemize}+
\item[]first level bulleted;
\item[]second level bulleted.
\end{enumerate}
Nested lists are allowed for numbered and custom lists.

\subsection{Theorem-like environments}

For theorem-like environments, we require the ``amsthm'' package.
There are three types of predefined theorem styles: thmstyleone,
thmstyletwo, and thmstylethree. The below table shows the output
details for each style:
\begin{center}\begin{tabular}{|l|p{22pc}|}
\hline thmstyleone &Numbered, theorem head in bold font and theorem text in italic style\\
\hline
thmstyletwo & Numbered, theorem head in italic font and theorem text in roman style\\
\hline
thmstylethree & Numbered, theorem head in bold font and theorem text in roman style\\
\hline
\end{tabular}\end{center}
As per the output requirement, the corresponding new theorem style should be defined in the preamble area. For example, if you require a ``Proposition'' environment to be set with ``thmstyletwo,'' then we need to include the below lines in the preamble area:

\begin{verbatim}
\theoremstyle{thmstyletwo}
\newtheorem{proposition}{Proposition}
\end{verbatim}

Refer to the ``amsthm'' package documentation for more details about the additional features available for new theorem styles. As well as the above, a predefined ``proof'' environment is available. This environment prints the ``Proof'' head in italic style and ``body text'' in Roman style with an open square at the end of each proof environment.

\begin{verbatim}
\begin{proof}...\end{proof}
\end{verbatim}

\subsection{Footnotes}

Footnotes are produced with the standard \LaTeX\ command \verb+\footnote{<Some text>}+. This typesets a numerical flag at the location of the footnote command and places the footnote text at the bottom of the page.

\subsection{Algorithms, Program codes, and Listings}

The packages ``algorithm,'' ``algorithmicx,'' and ``algpseudocode'' are used for setting algorithms in \LaTeX. For algorithms, use the below format:

\begin{verbatim}
\begin{algorithm}
\caption{<alg-caption>}\label{<alg-label>}
\begin{algorithmic}[1]
...
\end{algorithmic}
\end{algorithm}
\end{verbatim}
Refer the above-listed package documentation for more details before setting algorithms.

To set program codes, the ``program'' package is required and the command to be used is \verb+\begin{program}...\end{program}+.

The command \verb+\begin{lstlisting}...\end{lstlisting}+ is used to set ``verbatim'' like environments. Refer to the "lstlisting" package documentation for more details.

\subsection{Cross references}

\LaTeX\ is able to automatically insert hypertext links within a document:

\begin{itemize}
\item the \verb+\ref{}+ command adds a clickable link to the referred object;
\item the \verb+\label{}+ command automatically inserts a target;
\end{itemize}

\subsection{Cross citations }
To make a citation in the text, use \verb+\citep{...}+ for a parenthetical citation (Jones et al., 1990), \verb+\citet{...}+ for a textual one, as Jones et al. (1990). Both \verb+\citep+ and \verb+\citet+ are defined by ``natbib'' and are thus not standard. The standard \LaTeX\ command \verb+\cite+ should be avoided, because it behaves like \verb+\citet+ for author–year citations, but like \verb+\citep+ for numerical ones. There also exist the starred versions \verb+\citet*+ and \verb+\citep*+ that print the full author list, and not just the abbreviated one. All of these may take one or two optional arguments to add some text before and after the citation:

\begin{center}\begin{tabular}{|l|l|}\hline
\verb+\citet{jon90}+&   \verb+Jones et al. (1990) +\\\hline
\verb+\citet[chap.~2]{jon90}+&  \verb+Jones et al. (1990, chap. 2) +\\\hline
\verb+\citep{jon90}+&   \verb+(Jones et al., 1990) +\\\hline
\verb+\citep[chap.~2]{jon90}+&  \verb+(Jones et al., 1990, chap. 2) +\\ \hline
\verb+\citep[see][]{jon90}+&    \verb+(see Jones et al., 1990) +\\\hline
\verb+\citep[see][chap.~2]{jon90}+&     \verb+(see Jones et al., 1990, chap. 2) +\\\hline
\verb+\citet*{jon90}+&  \verb+Jones, Baker, and Williams (1990) +\\\hline
\verb+\citep*{jon90}+&  \verb+(Jones, Baker, and Williams, 1990)+\\
\hline
\end{tabular}\end{center}
Please refer to the ``natbib'' package documentation for guidance on other citation commands.

\section{Back matter elements }
\subsection{Appendices}
This section is set with a \verb+\begin{appendices}...\end{appendices}+ environment. All the other commands used to set section heads, tables, and figures inside this section remain the same as the main body text.

\subsection{References}
The basic bibliography environment is accepted for setting a reference section:

\begin{verbatim}
\begin{thebibliography}{9}
\bibitem{bib1} ...
\bibitem{bib2} ...
\end{thebibliography}
\end{verbatim}
However, BiB\TeX\ is the preferred format for references. BiB\TeX\ automates most of the work involved in references for articles. Using BiB\TeX\ options, both citations and references can be automatically updated to the preferred reference style. That is, you need not apply reference style tags for each element manually; it promotes structured writing. Basically, BiB\TeX\ works with two parts of the references: \textit{content} and \textit{style}. The \textit{content} is stored separately in a plain text database file called .bib, in which each entry is structured to distingiush diferent types of entries and fields. The \textit{style} and presentation of the database content are processed with the help of BiB\TeX\ program using a style file called .bst (bibliography style file).

You are requested to use the sample bib file ``reference.bib'' provided as a base for preparing your own .bib file. For the ``author-year'' citation style, use ``abbrvnat'' bst and for ``numbered'' citation style use ``plain'' bst, by adding the following instructions to the preamble.

\begin{enumerate}
\item author-year citation style = \verb+\bibliographystyle{abbrvnat}+
\item numbered citation style = \verb+\bibliographystyle{unsrt}+
\end{enumerate}

Then include your .bib file at the end of your document as shown below:
\begin{verbatim}
\bibliography{<bib-file-without-extension>}
\end{verbatim}
To generate the .bbl file, you should only need to compile \LaTeX/PDF\LaTeX once, then BIB\TeX, then \LaTeX/PDF\LaTeX twice more. The resulting bibliography is ready for typesetting with all formatting tags rendered according to the chosen reference style. For more details, please visit \url{http://www.bibtex.org}.
\section{Author biography}
If this section is required by the journal, please use the below command:
\begin{verbatim}
\begin{biography}
{\includegraphics{<image-file>}}
{\author{<author-Name>} biography-text}
\end{biography}
\end{verbatim}

\section{Author supports}
General support for \LaTeX\ related questions can be obtained from the Internet newsgroup comp.text.tex. Frequently asked questions are available in various web sites dealing with \LaTeX.

\section{Revision History}

\begin{tabular}{|l|l|l|l|}
\hline
Revision State & Revision Date &  Version Number & Revision History\\
\hline
 00 &    16 March, 2020 & Version 1.0   &   \\
 \hline
 01 &    18 March, 2022 & Version 1.1   &   \\
 \hline
\end{tabular}

\end{document}

### 2024-02-05 alloscore_manuscript/oup-authoring-template/oup-authoring-template.tex
add series A papers

%%
%% Copyright 2022 OXFORD UNIVERSITY PRESS
%%
%% This file is part of the 'oup-authoring-template Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'oup-authoring-template Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for OXFORD UNIVERSITY PRESS's document class `oup-authoring-template'
%% with bibliographic references
%%

%%%CONTEMPORARY%%%
%\documentclass[unnumsec,webpdf,contemporary,large]{oup-authoring-template}%
%\documentclass[unnumsec,webpdf,contemporary,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
%\documentclass[unnumsec,webpdf,contemporary,medium]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,contemporary,small]{oup-authoring-template}

%%%MODERN%%%
%\documentclass[unnumsec,webpdf,modern,large]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,modern,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
\documentclass[unnumsec,webpdf,modern,medium]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,modern,small]{oup-authoring-template}

%%%TRADITIONAL%%%
%\documentclass[unnumsec,webpdf,traditional,large]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,traditional,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
%\documentclass[unnumsec,namedate,webpdf,traditional,medium]{oup-authoring-template}
%\documentclass[namedate,webpdf,traditional,small]{oup-authoring-template}

\onecolumn % for one column layouts

%\usepackage{showframe}

\graphicspath{{Fig/}}

% line numbers
%\usepackage[mathlines, switch]{lineno}
%\usepackage[right]{lineno}

\theoremstyle{thmstyleone}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}%
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%
\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}

\begin{document}

\journaltitle{Journal Title Here}
\DOI{DOI HERE}
\copyrightyear{2022}
\pubyear{2019}
\access{Advance Access Publication Date: Day Month Year}
\appnotes{Paper}

\firstpage{1}

%\subtitle{Subject Section}

\title[Short Article Title]{Article Title}

\author[1,$\ast$]{First Author}
\author[2]{Second Author}
\author[3]{Third Author}
\author[3]{Fourth Author}
\author[4]{Fifth Author\ORCID{0000-0000-0000-0000}}

\authormark{Author Name et al.}

\address[1]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}
\address[2]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}
\address[3]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}
\address[4]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}

\corresp[$\ast$]{Corresponding author. \href{email:email-id.com}{email-id.com}}

\received{Date}{0}{Year}
\revised{Date}{0}{Year}
\accepted{Date}{0}{Year}

%\editor{Associate Editor: Name}

%\abstract{
%\textbf{Motivation:} .\\
%\textbf{Results:} .\\
%\textbf{Availability:} .\\
%\textbf{Contact:} \href{name@email.com}{name@email.com}\\
%\textbf{Supplementary information:} Supplementary data are available at \textit{Journal Name}
%online.}

\abstract{Abstracts must be able to stand alone and so cannot contain citations to
the paper's references, equations, etc. An abstract must consist of a single
paragraph and be concise. Because of online formatting, abstracts must appear
as plain as possible.}
\keywords{keyword1, Keyword2, Keyword3, Keyword4}

% \boxedtext{
% \begin{itemize}
% \item Key boxed text here.
% \item Key boxed text here.
% \item Key boxed text here.
% \end{itemize}}

\maketitle

\section{Introduction}
The introduction introduces the context and summarizes the manuscript. It is importantly to clearly state the contributions
of this piece of work. Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

This is an example of a new parapgraph with a numbered footnote\footnote{\url{https://data.gov.uk/}} and a second footnote marker.\footnote{Example of footnote text.}

\section{This is an example for first level head - section head}\label{sec2}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum (refer Section~\ref{sec5}).

\subsection{This is an example for second level head - subsection head}\label{subsec1}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\subsubsection{This is an example for third level head - subsubsection head}\label{subsubsec1}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. %Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\paragraph{This is an example for fourth level head - paragraph head}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\section{This is an example for first level head}\label{sec3}

\subsection{This is an example for second level head - subsection head}\label{subsec2}

\subsubsection{This is an example for third level head - subsubsection head}\label{subsubsec2}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\paragraph{This is an example for fourth level head - paragraph head}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\section{Equations}\label{sec4}

Equations in \LaTeX{} can either be inline or set as display equations. For
inline equations use the \verb+$...$+ commands. Eg: the equation
$H\psi = E \psi$ is written via the command \verb+$H \psi = E \psi$+.

For display equations (with auto generated equation numbers)
one can use the equation or eqnarray environments:
\begin{equation}
\|\tilde{X}(k)\|^2 \leq\frac{\sum\limits_{i=1}^{p}\left\|\tilde{Y}_i(k)\right\|^2+\sum\limits_{j=1}^{q}\left\|\tilde{Z}_j(k)\right\|^2 }{p+q},\label{eq1}
\end{equation}
where,
\begin{align}
D_\mu &=  \partial_\mu - ig \frac{\lambda^a}{2} A^a_\mu \nonumber \\
F^a_{\mu\nu} &= \partial_\mu A^a_\nu - \partial_\nu A^a_\mu + g f^{abc} A^b_\mu A^a_\nu.\label{eq2}
\end{align}
Notice the use of \verb+\nonumber+ in the align environment at the end
of each line, except the last, so as not to produce equation numbers on
lines where no equation numbers are required. The \verb+\label{}+ command
should only be used at the last line of an align environment where
\verb+\nonumber+ is not used.
\begin{equation}
Y_\infty = \left( \frac{m}{\textrm{GeV}} \right)^{-3}
    \left[ 1 + \frac{3 \ln(m/\textrm{GeV})}{15}
    + \frac{\ln(c_2/5)}{15} \right].
\end{equation}
The class file also supports the use of \verb+\mathbb{}+, \verb+\mathscr{}+ and
\verb+\mathcal{}+ commands. As such \verb+\mathbb{R}+, \verb+\mathscr{R}+
and \verb+\mathcal{R}+ produces $\mathbb{R}$, $\mathscr{R}$ and $\mathcal{R}$
respectively (refer Subsubsection~\ref{subsubsec3}).

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum. 
Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.

\section{Tables}\label{sec5}

Tables can be inserted via the normal table and tabular environment. To put
footnotes inside tables one has to Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum. use the additional ``tablenotes" environment
enclosing the tabular environment. The footnote appears just below the table
itself (refer Tables~\ref{tab1} and \ref{tab2}).

\begin{verbatim}
\begin{table}[t]
\begin{center}
\begin{minipage}{<width>}
\caption{<table-caption>\label{<table-label>}}%
\begin{tabular}{@{}llll@{}}
\toprule
column 1 & column 2 & column 3 & column 4\\
\midrule
row 1 & data 1 & data 2          & data 3 \\
row 2 & data 4 & data 5$^{1}$ & data 6 \\
row 3 & data 7 & data 8      & data 9$^{2}$\\
\botrule
\end{tabular}
\begin{tablenotes}%
\item Source: Example for source.
\item[$^{1}$] Example for a 1st table footnote.
\item[$^{2}$] Example for a 2nd table footnote.
\end{tablenotes}
\end{minipage}
\end{center}
\end{table}
\end{verbatim}

Lengthy tables which do not fit within textwidth should be set as rotated tables. For this, we need to use \verb+\begin{sidewaystable}...+ \verb+\end{sidewaystable}+ instead of\break \verb+\begin{table}...+ \verb+\end{table}+ environment.

\begin{table}[!t]
\caption{Caption text\label{tab1}}%
\begin{tabular*}{\columnwidth}{@{\extracolsep\fill}llll@{\extracolsep\fill}}
\toprule
column 1 & column 2  & column 3 & column 4\\
\midrule
row 1    & data 1   & data 2  & data 3  \\
row 2    & data 4   & data 5$^{1}$  & data 6  \\
row 3    & data 7   & data 8  & data 9$^{2}$  \\
\botrule
\end{tabular*}
\begin{tablenotes}%
\item Source: This is an example of table footnote this is an example of table footnote this is an example of table footnote this is an example of~table footnote this is an example of table footnote
\item[$^{1}$] Example for a first table footnote.
\item[$^{2}$] Example for a second table footnote.
\end{tablenotes}
\end{table}

\begin{table*}[t]
\caption{Example of a lengthy table which is set to full textwidth.\label{tab2}}
\tabcolsep=0pt%%
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lcccccc@{\extracolsep{\fill}}}
\toprule%
& \multicolumn{3}{@{}c@{}}{Element 1$^{1}$} & \multicolumn{3}{@{}c@{}}{Element 2$^{2}$} \\
\cline{2-4}\cline{5-7}%
Project & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ \\
\midrule
Element 3  & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$\\
Element 4  & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$\\
\botrule
\end{tabular*}
\begin{tablenotes}%
\item Note: This is an example of table footnote this is an example of table footnote this is an example of table footnote this is an example of~table footnote this is an example of table footnote
\item[$^{1}$] Example for a first table footnote.
\item[$^{2}$] Example for a second table footnote.\vspace*{6pt}
\end{tablenotes}
\end{table*}

\begin{sidewaystable}%[!p]
\caption{Tables which are too long to fit, should be written using the ``sidewaystable" environment as shown here\label{tab3}}
\tabcolsep=0pt%
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lcccccc@{\extracolsep{\fill}}}
\toprule%
& \multicolumn{3}{@{}c@{}}{Element 1$^{1}$}& \multicolumn{3}{@{}c@{}}{Element$^{2}$} \\
\cline{2-4}\cline{5-7}%
Projectile & Energy     & $\sigma_{calc}$ & $\sigma_{expt}$ & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ \\
\midrule
Element 3 & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$ \\
Element 4 & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$ \\
\botrule
\end{tabular*}
\begin{tablenotes}%
\item Note: This is an example of a table footnote this is an example of a table footnote this is an example of a table footnote this is an example of a table footnote this is an example of a table footnote
\item[$^{1}$] This is an example of a table footnote
\end{tablenotes}
\end{sidewaystable}

\section{Figures}\label{sec6}

As per display \LaTeX\ standards one has to use eps images for \verb+latex+ compilation and \verb+pdf/jpg/png+ images for
\verb+pdflatex+ compilation. This is one of the major differences between \verb+latex+
and \verb+pdflatex+. The images should be single-page documents. The command for inserting images
for \verb+latex+ and \verb+pdflatex+ can be generalized. The package used to insert images in \verb+latex/pdflatex+ is the
graphicx package. Figures can be inserted via the normal figure environment as shown in the below example:

\begin{figure}[!t]%
\centering
{\color{black!20}\rule{213pt}{37pt}}
\caption{This is a widefig. This is an example of a long caption this is an example of a long caption  this is an example of a long caption this is an example of a long caption}\label{fig1}
\end{figure}

\begin{figure*}[!t]%
\centering
{\color{black!20}\rule{438pt}{74pt}}
\caption{This is a widefig. This is an example of a long caption this is an example of a long caption  this is an example of a long caption this is an example of a long caption}\label{fig2}
\end{figure*}

\begin{verbatim}
\begin{figure}[t]
        \centering\includegraphics{<eps-file>}
        \caption{<figure-caption>}
        \label{<figure-label>}
\end{figure}
\end{verbatim}

Test text here.

For sample purposes, we have included the width of images in the
optional argument of \verb+\includegraphics+ tag. Please ignore this.
Lengthy figures which do not fit within textwidth should be set in rotated mode. For rotated figures, we need to use \verb+\begin{sidewaysfigure}+ \verb+...+ \verb+\end{sidewaysfigure}+ instead of the \verb+\begin{figure}+ \verb+...+ \verb+\end{figure}+ environment.

\begin{sidewaysfigure}%
\centering
{\color{black!20}\rule{610pt}{102pt}}
\caption{This is an example for a sideways figure. This is an example of a long caption this is an example of a long caption  this is an example of a long caption this is an example of a long caption}\label{fig3}
\end{sidewaysfigure}

\section{Algorithms, Program codes and Listings}\label{sec7}

Packages \verb+algorithm+, \verb+algorithmicx+ and \verb+algpseudocode+ are used for setting algorithms in latex.
For this, one has to use the below format:

\begin{verbatim}
\begin{algorithm}
\caption{<alg-caption>}\label{<alg-label>}
\begin{algorithmic}[1]
. . .
\end{algorithmic}
\end{algorithm}
\end{verbatim}

You may need to refer to the above-listed package documentations for more details before setting an \verb+algorithm+ environment.
To set program codes, one has to use the \verb+program+ package. We need to use the \verb+\begin{program}+ \verb+...+
\verb+\end{program}+ environment to set program codes.

\begin{algorithm}[!t]
\caption{Calculate $y = x^n$}\label{algo1}
\begin{algorithmic}[1]
\Require $n \geq 0 \vee x \neq 0$
\Ensure $y = x^n$
\State $y \Leftarrow 1$
\If{$n < 0$}
        \State $X \Leftarrow 1 / x$
        \State $N \Leftarrow -n$
\Else
        \State $X \Leftarrow x$
        \State $N \Leftarrow n$
\EndIf
\While{$N \neq 0$}
        \If{$N$ is even}
            \State $X \Leftarrow X \times X$
            \State $N \Leftarrow N / 2$
        \Else[$N$ is odd]
            \State $y \Leftarrow y \times X$
            \State $N \Leftarrow N - 1$
        \EndIf
\EndWhile
\end{algorithmic}
\end{algorithm}

Similarly, for \verb+listings+, one has to use the \verb+listings+ package. The \verb+\begin{lstlisting}+ \verb+...+ \verb+\end{lstlisting}+ environment is used to set environments similar to the \verb+verbatim+ environment. Refer to the \verb+lstlisting+ package documentation for more details on this.

\begin{minipage}{\hsize}%
\lstset{language=Pascal}% Set your language (you can change the language for each code-block optionally)
\begin{lstlisting}[frame=single,framexleftmargin=-1pt,framexrightmargin=-17pt,framesep=12pt,linewidth=0.98\textwidth]
for i:=maxint to 0 do
begin
{ do nothing }
end;
Write('Case insensitive ');
Write('Pascal keywords.');
\end{lstlisting}
\end{minipage}

\section{Cross referencing}\label{sec8}

Environments such as figure, table, equation, and align can have a label
declared via the \verb+\label{#label}+ command. For figures and table
environments one should use the \verb+\label{}+ command inside or just
below the \verb+\caption{}+ command.  One can then use the
\verb+\ref{#label}+ command to cross-reference them. As an example, consider
the label declared for Figure \ref{fig1} which is
\verb+\label{fig1}+. To cross-reference it, use the command
\verb+ Figure \ref{fig1}+, for which it comes up as
``Figure~\ref{fig1}".

\subsection{Details on reference citations}\label{subsec3}

With standard numerical .bst files, only numerical citations are possible.
With an author-year .bst file, both numerical and author-year citations are possible.

If author-year citations are selected, \verb+\bibitem+ must have one of the following forms:

{\footnotesize%
\begin{verbatim}
\bibitem[Jones et al.(1990)]{key}...
\bibitem[Jones et al.(1990)Jones,
                Baker, and Williams]{key}...
\bibitem[Jones et al., 1990]{key}...
\bibitem[\protect\citeauthoryear{Jones,
                Baker, and Williams}
                {Jones et al.}{1990}]{key}...
\bibitem[\protect\citeauthoryear{Jones et al.}
                {1990}]{key}...
\bibitem[\protect\astroncite{Jones et al.}
                {1990}]{key}...
\bibitem[\protect\citename{Jones et al., }
                1990]{key}...
\harvarditem[Jones et al.]{Jones, Baker, and
                Williams}{1990}{key}...
\end{verbatim}}

This is either to be made up manually, or to be generated by an
appropriate .bst file with BibTeX. Then,

{%
\begin{verbatim}
                    Author-year mode
                        || Numerical mode
\citet{key} ==>>  Jones et al. (1990)
                        || Jones et al. [21]
\citep{key} ==>> (Jones et al., 1990) || [21]
\end{verbatim}}

\noindent
Multiple citations as normal:

{%
\begin{verbatim}
\citep{key1,key2} ==> (Jones et al., 1990;
                         Smith, 1989)||[21,24]
        or (Jones et al., 1990, 1991)||[21,24]
        or (Jones et al., 1990a,b)   ||[21,24]
\end{verbatim}}

\noindent
\verb+\cite{key}+ is the equivalent of \verb+\citet{key}+ in author-year mode
and  of \verb+\citep{key}+ in numerical mode. Full author lists may be forced with
\verb+\citet*+ or \verb+\citep*+, e.g.

{%
\begin{verbatim}
\citep*{key} ==>> (Jones, Baker, and Mark, 1990)
\end{verbatim}}

\noindent
Optional notes as:

{%
\begin{verbatim}
\citep[chap. 2]{key}     ==>>
        (Jones et al., 1990, chap. 2)
\citep[e.g.,][]{key}     ==>>
        (e.g., Jones et al., 1990)
\citep[see][pg. 34]{key} ==>>
        (see Jones et al., 1990, pg. 34)
\end{verbatim}}

\noindent
(Note: in standard LaTeX, only one note is allowed, after the ref.
Here, one note is like the standard, two make pre- and post-notes.)

{%
\begin{verbatim}
\citealt{key}   ==>> Jones et al. 1990
\citealt*{key}  ==>> Jones, Baker, and
                        Williams 1990
\citealp{key}   ==>> Jones et al., 1990
\citealp*{key}  ==>> Jones, Baker, and
                        Williams, 1990
\end{verbatim}}

\noindent
Additional citation possibilities (both author-year and numerical modes):

{%
\begin{verbatim}
\citeauthor{key}       ==>> Jones et al.
\citeauthor*{key}      ==>> Jones, Baker, and
                                Williams
\citeyear{key}         ==>> 1990
\citeyearpar{key}      ==>> (1990)
\citetext{priv. comm.} ==>> (priv. comm.)
\citenum{key}          ==>> 11 [non-superscripted]
\end{verbatim}}

\noindent
Note: full author lists depend on whether the bib style supports them;
if not, the abbreviated list is printed even when full is requested.

\noindent
For names like della Robbia at the start of a sentence, use

{%
\begin{verbatim}
\Citet{dRob98}      ==>> Della Robbia (1998)
\Citep{dRob98}      ==>> (Della Robbia, 1998)
\Citeauthor{dRob98} ==>> Della Robbia
\end{verbatim}}

\noindent
The following is an example for \verb+\cite{...}+: \cite{rahman2019centroidb}. Another example for \verb+\citep{...}+: \citep{bahdanau2014neural,imboden2018cardiorespiratory,motiian2017unified,murphy2012machine,ji20123d}.
Sample cites here \cite{krizhevsky2012imagenet,horvath2018dna} and \cite{pyrkov2018quantitative}, \cite{wang2018face}, \cite{lecun2015deep,zhang2018fine,ravi2016deep}.

\section{Lists}\label{sec9}

List in \LaTeX{} can be of three types: numbered, bulleted and unnumbered. The ``enumerate'' environment produces a numbered list, the 
``itemize'' environment produces a bulleted list and the ``unlist''
environment produces an unnumbered list.
In each environment, a new entry is added via the \verb+\item+ command.
\begin{enumerate}[1.]
\item This is the 1st item

\item Enumerate creates numbered lists, itemize creates bulleted lists and
unnumerate creates unnumbered lists.
\begin{enumerate}[(a)]
\item Second level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.

\item Second level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.
\begin{enumerate}[(ii)]
\item Third level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.

\item Third level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.
\end{enumerate}

\item Second level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.

\item Second level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.
\end{enumerate}

\item Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.

\item Numbered lists continue.
\end{enumerate}
Lists in \LaTeX{} can be of three types: enumerate, itemize and description.
In each environment, a new entry is added via the \verb+\item+ command.
\begin{itemize}
\item First level bulleted list. This is the 1st item

\item First level bulleted list. Itemize creates bulleted lists and description creates unnumbered lists.
\begin{itemize}
\item Second level dashed list. Itemize creates bulleted lists and description creates unnumbered lists.

\item Second level dashed list. Itemize creates bulleted lists and description creates unnumbered lists.

\item Second level dashed list. Itemize creates bulleted lists and description creates unnumbered lists.
\end{itemize}

\item First level bulleted list. Itemize creates bulleted lists and description creates unnumbered lists.

\item First level bulleted list. Bullet lists continue.
\end{itemize}

\noindent
Example for unnumbered list items:

\begin{unlist}
\item Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text.

\item Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text.

\item sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text.
\end{unlist}

\section{Examples for theorem-like environments}\label{sec10}

For theorem-like environments, we require the \verb+amsthm+ package. There are three types of predefined theorem styles - \verb+thmstyleone+, \verb+thmstyletwo+ and \verb+thmstylethree+   (check your journal's instructions page in case a specific style is required).

\medskip
\noindent\begin{tabular}{|l|p{13pc}|}
\hline
\verb+thmstyleone+ & Numbered, theorem head in bold font and theorem text in italic style \\\hline
\verb+thmstyletwo+ & Numbered, theorem head in roman font and theorem text in italic style \\\hline
\verb+thmstylethree+ & Numbered, theorem head in bold font and theorem text in roman style \\\hline
\end{tabular}

\begin{theorem}[Theorem subhead]\label{thm1}
Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text.
Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text.
Example theorem text.
\end{theorem}

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue.

\begin{proposition}
Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text.
Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text.
\end{proposition}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante.

\begin{example}
Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem.
\end{example}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante.

\begin{remark}
Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem.
\end{remark}

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst.

\begin{definition}[Definition sub head]
Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text.
\end{definition}

Apart from the above styles, we have the \verb+\begin{proof}+ \verb+...+ \verb+\end{proof}+ environment - with the proof head in italic style and the body text in roman font with an open square at the end of each proof environment.

\begin{proof}Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text.
\end{proof}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi.

\begin{proof}[Proof of Theorem~{\upshape\ref{thm1}}]
Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text.
\end{proof}

\noindent
For a quote environment, one has to use\newline \verb+\begin{quote}...\end{quote}+
\begin{quote}
Quoted text example. Aliquam porttitor quam a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
\end{quote}
Donec congue. Maecenas urna mi, suscipit in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl (refer Figure~\ref{fig3}). Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Etiam ligula arcu,
elementum a, venenatis quis, sollicitudin sed, metus. Donec nunc pede, tincidunt in, venenatis vitae, faucibus vel (refer Table~\ref{tab3}).

\section{Conclusion}

Some Conclusions here.

%%%%%%%%%%%%%%

\begin{appendices}

\section{Section title of first appendix}\label{sec11}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

\subsection{Subsection title of first appendix}\label{subsec4}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

\subsubsection{Subsubsection title of first appendix}\label{subsubsec3}

Example for an unnumbered figure:

\begin{figure}[!h]
\centering
{\color{black!20}\rule{85pt}{92pt}}
\end{figure}

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis.

\section{Section title of second appendix}\label{sec12}%

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.

\begin{figure}[b]
\centering
{\color{black!20}\rule{217pt}{120pt}}
\caption{This is an example for appendix figure\label{fig4}}
\end{figure}

\begin{table}[t]%
\begin{center}
\begin{minipage}{.52\columnwidth}
\caption{This is an example of Appendix table showing food requirements of army, navy and airforce\label{tab4}}%
\begin{tabular}{@{}lcc@{}}%
\toprule
col1 head & col2 head & col3 head \\
\midrule
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
\botrule
\end{tabular}
\end{minipage}
\end{center}
\end{table}

\subsection{Subsection title of second appendix}\label{subsec5}

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non odio.
Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce sed justo
eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor. Donec
pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin et quam.
Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien turpis,
fermentum vel, eleifend faucibus, vehicula eu, lacus.

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non odio.
Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce sed justo
eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor. Donec
pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin et quam.
Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien turpis,
fermentum vel, eleifend faucibus, vehicula eu, lacus.

\subsubsection{Subsubsection title of second appendix}\label{subsubsec4}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
vehicula augue eu neque.

Example for an equation inside the appendix:
\begin{align}
 p &= \frac{\gamma^{2} - (n_{C} -1)H}{(n_{C} - 1) + H - 2\gamma}, \label{1eq:hybobo:pfromgH} \\
 \theta &= \frac{(\gamma - H)^{2}(\gamma - n_{C} -1)^{2}}{(n_{C} - 1 + H - 2\gamma)^{2}} \label{2eq:hybobo:tfromgH}\; .
\end{align}

\section{Example of another appendix section}\label{sec13}%

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris
\begin{equation}
\mathcal{L} = i \bar{\psi} \gamma^\mu D_\mu \psi
    - \frac{1}{4} F_{\mu\nu}^a F^{a\mu\nu} - m \bar{\psi} \psi.
\label{eq26}
\end{equation}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

%% Example for unnumbered table inside appendix
\begin{table}
\begin{center}
\begin{minipage}{.52\columnwidth}
\caption{}{%
\begin{tabular}{lcc}%
\toprule
col1 head & col2 head & col3 head \\
\midrule
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
\botrule
\end{tabular}}{}
\end{minipage}
\end{center}
\end{table}

\end{appendices}

\section{Competing interests}
No competing interest is declared.

\section{Author contributions statement}

Must include all authors, identified by initials, for example:
S.R. and D.A. conceived the experiment(s),  S.R. conducted the experiment(s), S.R. and D.A. analysed the results.  S.R. and D.A. wrote and reviewed the manuscript.

\section{Acknowledgments}
The authors thank the anonymous reviewers for their valuable suggestions. This work is supported in part by funds from the National Science Foundation (NSF: \# 1636933 and \# 1920920).

%\bibliographystyle{plain}
%\bibliography{reference}

\begin{thebibliography}{10}

\bibitem{bahdanau2014neural}
Dzmitry Bahdanau, Kyunghyun Cho, and Yoshua Bengio.
\newblock Neural machine translation by jointly learning to align and
  translate.
\newblock {\em arXiv preprint arXiv:1409.0473}, 2014.

\bibitem{horvath2018dna}
Steve Horvath and Kenneth Raj.
\newblock Dna methylation-based biomarkers and the epigenetic clock theory of
  ageing.
\newblock {\em Nature Reviews Genetics}, 19(6):371, 2018.

\bibitem{imboden2018cardiorespiratory}
Mary~T Imboden, Matthew~P Harber, Mitchell~H Whaley, W~Holmes Finch, Derron~L
  Bishop, and Leonard~A Kaminsky.
\newblock Cardiorespiratory fitness and mortality in healthy men and women.
\newblock {\em Journal of the American College of Cardiology},
  72(19):2283--2292, 2018.

\bibitem{ji20123d}
Shuiwang Ji, Wei Xu, Ming Yang, and Kai Yu.
\newblock 3d convolutional neural networks for human action recognition.
\newblock {\em IEEE Transactions on Pattern Analysis and Machine Intelligence},
  35(1):221--231, 2012.

\bibitem{krizhevsky2012imagenet}
Alex Krizhevsky, Ilya Sutskever, and Geoffrey~E Hinton.
\newblock Imagenet classification with deep convolutional neural networks.
\newblock In {\em Advances in Neural Information Processing Systems}, pages
  1097--1105, 2012.

\bibitem{lecun2015deep}
Yann LeCun, Yoshua Bengio, and Geoffrey Hinton.
\newblock Deep learning.
\newblock {\em Nature}, 521(7553):436, 2015.

\bibitem{motiian2017unified}
Saeid Motiian, Marco Piccirilli, Donald~A Adjeroh, and Gianfranco Doretto.
\newblock Unified deep supervised domain adaptation and generalization.
\newblock In {\em Proceedings of the IEEE International Conference on Computer
  Vision}, pages 5715--5725, 2017.

\bibitem{murphy2012machine}
Kevin~P Murphy.
\newblock {\em Machine learning: A probabilistic perspective}.
\newblock MIT press, 2012.

\bibitem{american2013acsm}
American~College of~Sports~Medicine et~al.
\newblock {\em ACSM's guidelines for exercise testing and prescription}.
\newblock Lippincott Williams \& Wilkins, 2013.

\bibitem{pyrkov2018quantitative}
Timothy~V Pyrkov, Evgeny Getmantsev, Boris Zhurov, Konstantin Avchaciov,
  Mikhail Pyatnitskiy, Leonid Menshikov, Kristina Khodova, Andrei~V Gudkov, and
  Peter~O Fedichev.
\newblock Quantitative characterization of biological age and frailty based on
  locomotor activity records.
\newblock {\em Aging (Albany NY)}, 10(10):2973, 2018.

\bibitem{rahman2019centroidb}
Syed~Ashiqur Rahman and Donald Adjeroh.
\newblock Centroid of age neighborhoods: A generalized approach to estimate
  biological age.
\newblock In {\em 2019 IEEE EMBS International Conference on Biomedical \&
  Health Informatics (BHI)}, pages 1--4. IEEE, 2019.

\bibitem{ravi2016deep}
Daniele Rav{\`\i}, Charence Wong, Fani Deligianni, Melissa Berthelot, Javier
  Andreu-Perez, Benny Lo, and Guang-Zhong Yang.
\newblock Deep learning for health informatics.
\newblock {\em IEEE {J}ournal of {B}iomedical and {H}ealth {I}nformatics},
  21(1):4--21, 2016.

\bibitem{wang2018face}
Zongwei Wang, Xu~Tang, Weixin Luo, and Shenghua Gao.
\newblock Face aging with identity-preserved conditional generative adversarial
  networks.
\newblock In {\em Proceedings of the IEEE Conference on Computer Vision and
  Pattern Recognition}, pages 7939--7947, 2018.

\bibitem{zhang2018fine}
Ke~Zhang, Na~Liu, Xingfang Yuan, Xinyao Guo, Ce~Gao, and Zhenbing Zhao.
\newblock Fine-grained age estimation in the wild with attention {LSTM}
  networks.
\newblock {\em arXiv preprint arXiv:1805.10445}, 2018.

\end{thebibliography}

%USE THE BELOW OPTIONS IN CASE YOU NEED AUTHOR YEAR FORMAT.
%\bibliographystyle{abbrvnat}
%\bibliography{reference}

%% sample for biography with author's image
\begin{biography}{{\color{black!20}\rule{77pt}{77pt}}}{\author{Author Name.} This is sample author biography text. The values provided in the optional argument are meant for sample purposes. There is no need to include the width and height of an image in the optional argument for live articles. This is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text.}
\end{biography}

%% sample for biography without author's image
\begin{biography}{}{\author{Author Name.} This is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text.}
\end{biography}

\end{document}

### 2024-02-05 alloscore_manuscript/alloscore-application-shorter.Rnw
adding content in results/discsusion about the per-capita allocation.

--- 2024-02-01 alloscore_manuscript/alloscore-application-shorter.Rnw
@@
 %For many locations, forecasts predicted lower values than were eventually observed, as this was during the period of steep increase of viral transmission across many states.
-Of the 10 models evaluated, the CU-select model had the most accurate forecasts according to the allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table \ref{tab:multi-k-scores}).
+Of the 10 models evaluated for this one week, the CU-select model had the most accurate forecasts according to the allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table \ref{tab:multi-k-scores}).
 The JHUAPL-Bucky model had the second best MWIS but the third worst allocation score.
@@
                  label = "tab:multi-k-scores") |>
-  print(include.rownames = FALSE, )
+  print(include.rownames = FALSE, NA.string = "-")
 @
@@
 Predictions from the JHUAPL-Bucky model were wider, and included the observed level of daily hospital admissions more often.
-Therefore, WIS did not as often assign that model severe penalties for forecasts that underpredicted or overpredicted the actual hospitalization count.
+Therefore, it did not receive as severe penalties for forecasts that underpredicted or overpredicted the actual hospitalization count.
 
@@
 
-\subsubsection{Forecast scores showed differences over time}
+\subsubsection{Forecast scores showed differences in aggregate and over time}
 <<grob-hosps-over-time>>=
@@
 These are the dates for which forecasts were made and evaluated.
-A horizontal dashed line at 15,000 shows the assumed resource constraint $K$.
+A horizontal dashed line at 15,000 shows the hypothetical resource constraint $K$.
 Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states.
-The x-axis corresponds to the date for which the prediction was made.
+The x-axis corresponds to the date of the observation that a model's prediction was targeting (e.g., the date the forecast was made plus the forecast horizon).
 AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down).
+In Panel B, the per-capital allocation is drawn in a heavier solid line.
 Panel C shows the MWIS metric across weeks, averaged across all states.
@@
             range = maxWIS-minWIS)
+@
+
+Averaging across all weeks, improvements in mean allocation score (MAS) were observed when comparing the ensemble forecast to two benchmark comparisons.
+The COVIDhub-ensemble had the best MAS across all weeks, the per-capita allocation had the second best MAS, and the COVIDhub-baseline model had the sixth best allocation (Table \ref{tab:multi-week-performance-summary}).
+No individual model had a better MAS than the per-capita allocation.
+
+Most models were ranked similarly when evaluated using mean allocation scores (MAS) and mean weighted interval scores (MWIS), with some places of disagreement.
+The the four most accurate and the five least accurate models were the same according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
+(This comparison excludes the per-capita method which can be used to suggest allocations but does not make probabilistic forecasts and therefore cannot have a WIS.)
+Notably, the best model according to MWIS was the JHUAPL-SLPHospEns model, but this model only had the fourth best allocation scores.
+
+<<table-score-summary-all-weeks, results='asis'>>=
+all_scores |>
+  filter( !(reference_date.x %in% c("2021-11-22", "2022-02-28"))) |>
+  group_by(model) |>
+  summarize(N = sum(!is.na(alloscore)),
+            MAS = mean(alloscore, na.rm=TRUE),
+            MWIS = mean(mwis, na.rm=TRUE)) |>
+  filter(N > 12) |>
+  arrange(MAS) |>
+  select(model, MAS, MWIS) |>
+  xtable::xtable(caption = "Mean weighted interval scores (MWIS) and mean allocation scores (MAS) by model across 13 weeks.
+For the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, plus a per-capita allocation, these results show the average performance of each model across all weeks.
+Lower scores indicate better performance, and models are sorted with the most accurate allocations at the top.
+WIS is not defined for the per-capita model as this approach only provides a mechanism to allocate resources, it does not provide probabilistic forecasts.",
+                 digits = 0,
+                 label = "tab:multi-week-performance-summary") |>
+  print(include.rownames = FALSE, NA.string = "-")
 @
@@
 
-Models showed differing levels of correlation between their allocation scores and MWIS values (Figure \ref{fig:metrics-correlation}).
+Models showed differing levels of correlation between their allocation scores and MWIS values (Figure \ref{fig:metrics-correlation}), explaining in part why aggregate MWIS and MAS values did not always agree.
 Here are some examples of the different model-specific patterns observed:
@@
 \item Several models had consistently strong MWIS ranks but also had highly variable allocation score ranks with no clear association between the two (e.g., \texttt{JHUAPL-SLPHospEns} and \texttt{CU-select})
-\item One model performed consistently well for both metrics with no clear association (\texttt{COVIDhub-ensemble})
+\item One model performed consistently well for both metrics (\texttt{COVIDhub-ensemble})
 \item One model performed consistently well for allocation score but had only middlings ranks for MWIS (\texttt{CMU-TimeSeries})
@@
 
-AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 (Figure \ref{fig:multi-k}A).
+AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 as well as the per-capital allocation (Figure \ref{fig:multi-k}A).
 These calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of \Sexpr{format(tot_y, big.mark = ",")} that day.
@@
 Both versions of the IAS were correlated with the original AS conducted at $K=15,000$, with the higher correlation coming from the distribution that was centered at $K=15,000$ (Figure \ref{fig:multi-k}C).
-Model rankings based on the AC and the centered IAS were roughly similar, with the top and bottom two models being the same for both scores (Table \ref{tab:multi-k-scores}).
+Model rankings based on the AC and the centered IAS were roughly similar, with the top and bottom three approaches being the same for both scores (Table \ref{tab:multi-k-scores}).
 
@@
 We have demonstrated that forecast evaluation methods that are tied to a specific decision making context can yield model rankings that differ substantively from those based on standard measures of forecast skill, a result that aligns with findings in other fields \citep{leitch1991economicForecastEval, murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}.
-In particular, the allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.
+Additionally, we show that an existing ensemble forecast approach was the only method to outperform two benchmark allocation approaches in a hypothetical application.
+
+%In particular, the allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.
 
@@
 %One real-world example, which was the motivating example presented in the work above, is allocation of ventilators during a respiratory viral pandemic\citep{huang_stockpiling_2017}.
-We used ventilators as a notional example for a resource allocation problem because it (a) has been identified as a scarce resource during outbreaks\citep{huang_stockpiling_2017}, (b) could be allocated across different locations, with logistical considerations depending on the geographical scales, and (c) would likely be correlated (although not perfectly so) with hospitalization data.
-However, we acknowledge the limitations of this example in that not everyone who is hospitalized needs a ventilator, and new hospital admissions may not be the best stand in for a need that corresponds to a prevalence of severe cases.
+We used ventilators as a notional example for a resource allocation problem because it (a) has been identified as a scarce resource during outbreaks \citep{huang_stockpiling_2017}, (b) could be allocated across different locations, with logistical considerations depending on the geographical scales, and (c) would likely be correlated (although not perfectly so) with hospitalization data.
+However, this example has limitations in that not everyone who is hospitalized needs a ventilator, and new hospital admissions may not be the best stand in for a need that corresponds to a prevalence of severe cases.
 Other possible real-world examples include the allocation of a limited stockpile of vaccinations\citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests\citep{du_optimal_2022,pasco_covid-19_2023}.
-While there are limitations in the realism of the pairing of a specific resource and forecasted outcome, we emphasize that the important result from this work is how evaluating forecasts using a fundamentally different kind of score can illuminate previously hidden aspects of value in forecasts.
+While there may be limitations in this specific pairing of a resource and forecasted outcome, we emphasize that the important result from this work is how evaluating forecasts using a fundamentally different kind of score can illuminate previously hidden aspects of value in forecasts.
+In this setting, the feature of a forecast that emerges as being important is that the probabilistic forecasts predict the relative magnitudes between locations accurately, even if the absolute magnitudes are inaccurate.
 
@@
 While outcomes of some decisions could potentially be specified by loss functions, other uses for forecasts may be less quantifiable, such as when forecasts are used to inform a general sense of situational awareness that may feed in unmeasurable ways into other decisions.
-And even those that are easily framed as expected loss minimization may differ enough that no single score would be
+Even those that are easily framed as expected loss minimization may differ enough that no single score would be
 appropriate for all users.
@@
 There are several important limitations to the current work.
-The allocation score we developed here does not directly account for important considerations such as fairness or equity of allocations.
+%NGR - removing the sentence below because it seems like not a central concern/limitation, and I was triyng to cut. Happy to be overruled.
+%The allocation score we developed here does not directly account for important considerations such as fairness or equity of allocations.
 Nor does the proposed framework attempt to capture the broader context of decision making.
@@
 An opportunity for further investigation is to more carefully evaluate whether a forecast adds value to existing decision making processes.
-In the context of decisions about allocations, standard procedures might involve extrapolating need based on a current observed data (similar to the `baseline' model presented above), with or without adjustments based on other political or real-world considerations.
+In the context of decisions about allocations, standard procedures might involve extrapolating need based on a current observed data (similar to the benchmark approaches presented above), with or without adjustments based on other political or real-world considerations.
 For example, in many settings public health stakeholders will make decisions after synthesizing information from a variety of quantitative and qualitative sources coupled with expert judgment.

### 2024-02-05 alloscore_manuscript/alloscore-application-shorter.tex
adding content in results/discsusion about the per-capita allocation.

--- 2024-02-01 alloscore_manuscript/alloscore-application-shorter.tex
@@
 All code and data for the analyses presented in this manuscript are available at \url{https://github.com/aaronger/utility-eval-papers}.
-The analyses were generated using a reproducible workflow using R version 4.3.1 (2023-06-16) and the {\tt targets} package \citep{Rcore-2023, landau_2021_targets}.
+The analyses were generated using a reproducible workflow using R version 4.3.2 (2023-10-31) and the {\tt targets} package \citep{Rcore-2023, landau_2021_targets}.
 
@@
 %For many locations, forecasts predicted lower values than were eventually observed, as this was during the period of steep increase of viral transmission across many states.
-Of the 10 models evaluated, the CU-select model had the most accurate forecasts according to the allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table \ref{tab:multi-k-scores}).
+Of the 10 models evaluated for this one week, the CU-select model had the most accurate forecasts according to the allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table \ref{tab:multi-k-scores}).
 The JHUAPL-Bucky model had the second best MWIS but the third worst allocation score.
 
-% latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Thu Feb  1 16:28:56 2024
+% latex table generated in R 4.3.2 by xtable 1.8-4 package
+% Mon Feb  5 14:04:06 2024
 \begin{table}[ht]
@@
 CU-select & 669 & 133 & 774 & 326 \\ 
-  per-capita & 865 &  & 1029 & 366 \\ 
+  per-capita & 865 & - & 1029 & 366 \\ 
   COVIDhub-ensemble & 873 & 159 & 1067 & 438 \\ 
@@
 
-\subsubsection{Forecast scores showed differences over time}
+\subsubsection{Forecast scores showed differences in aggregate and over time}
 
@@
 
+Averaging across all weeks, improvements in mean allocation score (MAS) were observed when comparing the ensemble forecast to two benchmark comparisons.
+The COVIDhub-ensemble had the best MAS across all weeks, the per-capita allocation had the second best MAS, and the COVIDhub-baseline model had the sixth best allocation (Table \ref{tab:multi-week-performance-summary}).
+No individual model had a better MAS than the per-capita allocation.
+
+Most models were ranked similarly when evaluated using mean allocation scores (MAS) and mean weighted interval scores (MWIS), with some places of disagreement.
+The the four most accurate and the five least accurate models were the same according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
+(This comparison excludes the per-capita method which can be used to suggest allocations but does not make probabilistic forecasts and therefore cannot have a WIS.)
+Notably, the best model according to MWIS was the JHUAPL-SLPHospEns model, but this model only had the fourth best allocation scores.
+
+% latex table generated in R 4.3.2 by xtable 1.8-4 package
+% Mon Feb  5 14:04:11 2024
+\begin{table}[ht]
+\centering
+\begin{tabular}{lrr}
+  \hline
+model & MAS & MWIS \\ 
+  \hline
+COVIDhub-ensemble & 389 & 70 \\ 
+  per-capita & 464 & - \\ 
+  COVIDhub-trained\_ensemble & 483 & 87 \\ 
+  CU-select & 502 & 81 \\ 
+  JHUAPL-SLPHospEns & 526 & 67 \\ 
+  COVIDhub-baseline & 594 & 93 \\ 
+  JHUAPL-Bucky & 643 & 112 \\ 
+  MUNI-ARIMA & 707 & 116 \\ 
+  JHUAPL-Gecko & 929 & 110 \\ 
+  USC-SI\_kJalpha & 1473 & 155 \\ 
+   \hline
+\end{tabular}
+\caption{Mean weighted interval scores (MWIS) and mean allocation scores (MAS) by model across 13 weeks.
+For the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, plus a per-capita allocation, these results show the average performance of each model across all weeks.
+Lower scores indicate better performance, and models are sorted with the most accurate allocations at the top.
+WIS is not defined for the per-capita model as this approach only provides a mechanism to allocate resources, it does not provide probabilistic forecasts.}
+\label{tab:multi-week-performance-summary}
+\end{table}
+
 \subsubsection{Metrics were not consistently correlated over time}
@@
 
-AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 (Figure \ref{fig:multi-k}A).
+AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 as well as the per-capital allocation (Figure \ref{fig:multi-k}A).
 These calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of 19,581 that day.

### 2024-02-06 alloscore_manuscript/alloscore-application-shorter.Rnw
some updates

@@
 Predictions of disease dynamics have been used to inform decision making about a wide variety of measures to reduce disease spread and/or mitigate the severity of disease outcomes.
-For example, estimates of expected onset of flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola and diptheria dynamics have been used to allocate resources \citep{meltzer2014estimating, rainisch2015regional,camacho2015-ebola-bed,finger_real-time_2019}.
+For example, estimates of the onset time of the flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola and diptheria dynamics have been used to allocate resources \citep{meltzer2014estimating, rainisch2015regional,camacho2015-ebola-bed,finger_real-time_2019}.
 \cite{bertsimas2021predictionsCOVID} developed tools to guide decision making from infectious disease forecasts, which have been used to inform allocation of limited resources such as ventilators, ICU capacity, and vaccine supply.
@@
 
-In the wake of the COVID-19 pandemic, researchers have been calling for evaluating forecasts through their impact on policy \citep{marshall2023predictions, bilinski_adaptive_2023}.
-The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid}, log-score \citep{mcgowan_collaborative_2019}, or weighted interval score (WIS) \citep{bracher2021evaluating}.
-However, as we show in this paper, these metrics only evaluate one specific dimension of accuracy, and often do not have a clear correspondence to features that would make a forecast valuable in the context of a specific decision.
+In the wake of the COVID-19 pandemic, researchers have been calling for evaluating infectious disease forecasts through their impact on policy \citep{marshall2023predictions, bilinski_adaptive_2023}.
+Historically, the quality of disease forecasts has typically been measured using scores such as root mean square error (RMSE) \citep{papastefanopoulos2020covid}, log score \citep{mcgowan_collaborative_2019}, or weighted interval score (WIS) \citep{bracher2021evaluating}.
+These scores can be interpreted as measuring how well the forecast corresponds to the observed disease outcome.
+However, as we show in this paper, these scores often do not clearly measure the value of a forecast in the context of a specific decision.
 While these ideas have received little attention in the context of infectious disease forecasting, the results are well known from other fields, such as economics and finance, supply chain management, and meteorology.
@@
 In practice, probabilistic infectious disease forecasts have most often been made for observations that emerge from public health surveillance systems and have typically been evaluated with common scoring rules that have implementations available in existing software packages.
-For example, seasonal influenza forecasts in the US and dengue forecasts for Peru and Puerto Rico targeted public health surveillance measures of incidence over time and space, and used log-score and mean absolute errors to evaluate forecast skill \citep{mcgowan_collaborative_2019,reich_collaborative_2019,johansson_open_2019}.
+For example, seasonal influenza forecasts in the US and dengue forecasts for Peru and Puerto Rico targeted public health surveillance measures of incidence over time and space, and used log score and mean absolute error to evaluate forecast skill \citep{mcgowan_collaborative_2019,reich_collaborative_2019,johansson_open_2019}.
 Pandemic COVID-19 forecasts of observed cases, hospitalizations and/or deaths in the US and Europe, as reported by municipal, state, or federal surveillance systems, were evaluated using the weighted interval score (WIS, which is an approximation of the continuous ranked probability score, or CRPS), and prediction interval coverage \citep{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.
@@
 %Unlike the traditional measures of forecast skill such as WIS, CRPS, and log score, the proposed allocation score measures the suitability of a forecast as an input to this specific decision making task.
-
-%Briefly, the allocation score of a forecast is the avoidable unmet need that results from using that forecast to set resource allocations by minimizing expected unmet need.
-%For example, suppose that a decision maker is provided with forecasts of the level of need for medical resources in each of several states or hospital systems.
-%If there is a limited amount of the medical resource that is available to distribute, a decision maker could choose an allocation of that resource across locations that minimizes the expected unmet need according to that forecast.
-%As measured by the allocation score, one forecast is better than another if it would lead decision makers to an allocation that results in less unmet need.
-%By ``unnecessary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
-%If the amount of resources that is available to distribute is less than the actual need, some amount of unmet need is unavoidable.
-%The allocation score for a forecast does not include the unmet need that was unavoidable given the resource constraint, and so it measures only the amount of unmet need that could have been prevented by using a different allocation of available resources than that suggested by the forecast.
-%We elaborate on these ideas in Section~\ref{sec:methods}.
-
 In section \ref{sec:application}, we present an illustrative analysis using the allocation score to evaluate forecasts of hospital admissions in the US leading up to the Omicron wave in winter 2022.
 This analysis is ``synthetic'' in that it does not correspond to an actual analysis that supported decision making in real time.
-However, the framework described in this paper corresponds to real-world decisions that must be made by public health administrators around the globe, and could be adapted in the future for such real-time situations such as informing bed demand, ventilator usage, or doses needed for specific medications.
+However, the framework described in this paper corresponds to real-world decisions that must be made by public health administrators around the globe, and could be adapted in the future for real-time situations such as informing bed demand, ventilator usage, or doses needed for specific medications.
 %For example, forecasts for districts in Sierra Leone of bed demand to care for patients with Ebola was the subject of a real-time modeling study in late 2014 and early 2015 \citep{camacho2015-ebola-bed}.
@@
 %These examples illustrate the potential for forecasts to inform decisions about how to allocate limited supplies such as temporary hospital beds, ventilators, personal protective equipment, or other supplies that are known to be effective at reducing transmission or severity of disease.
-However, we emphasize again that these studies did not take the step of evaluating forecasts based on the quality of the allocation decisions that they supported or could have been used to support.
+%However, we emphasize again that these studies did not take the step of evaluating forecasts based on the quality of the allocation decisions that they supported or could have been used to support.
 
@@
 These scores illustrate a general result: allocation scores for a forecast will tend to be larger when the resource constraint is close to the observed need, because this is when it matters most which locations are allocated more or less resources.
-If the available resources are small compared to the observed need, any allocation of those limited resources will result in a large amount of unmet need.
-If the available resources are comparatively larger, it becomes less important which locations receive relatively more or fewer resources because all locations will receive enough resources to meet their need.
+If the amount of available resources is small relative to the observed need, any allocation of those limited resources will result in a large amount of unmet need.
+If the amount of available resources is comparatively large, it becomes less important which locations receive relatively more or fewer resources because all locations will receive enough resources to meet their need.
 In either of these extremes of resource availability, the avoidable unmet need that arises from the allocation suggested by a forecast (i.e., the forecast's allocation score) will tend to be small.
@@
 
-To adequately compare allocations from individual models with existing standard practices, we used two benchmarks against which the performance of using forecasts to inform allocations could be evaluated.
-First, we evaluated forecasts from the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, and the allocations that resulted from those forecasts \citep{cramer_evaluation_2022}.
-Second, we generated a proposed set of allocations on a per capita basis by state (using US Census data of vintage 2022 \citep{us_census_nst_est2022}), referred to below as \texttt{per-capita}.
+We used two benchmark methods as references to allow for comparison of model-derived allocations with existing standard practices.
+First, we evaluated forecasts from the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk \citep{cramer_evaluation_2022}.
+Second, we generated a proposed set of allocations where the quantity allocated to each state was proportional to that state's population (using US Census data of vintage 2022 \citep{us_census_nst_est2022}), referred to below as \texttt{per-capita}.
 These two approaches generally reflect common choices for ``best practices'' of allocation: either allocating resources based on the most recent observed data, or in proportion to the population of each location.
-As such, they serve as good comparisons for allocations based on forecast models.
+%As such, they serve as good comparisons for allocations based on forecast models.
 
@@
 The purple bar goes from zero to the amount of need that was met by the allocation from that model to a specific location.
-A red bar indicates need that exceeded the resource allocation for a location, and a light blue bar shows suggested resource allocation suggested by that model that were above the need for that location.">>=
+A red bar indicates need that exceeded the resource allocation for a location, and a light blue bar shows the amount by which the resource allocation suggested by that model exceeded the need for that location.">>=
 source("R/plot_functions.R")
@@
 
-Averaging across all weeks, improvements in mean allocation score (MAS) were observed when comparing the ensemble forecast to two benchmark comparisons.
+Averaging across all weeks, the ensemble forecast achieved a better mean allocation score (MAS) than two benchmark methods.
 The COVIDhub-ensemble had the best MAS across all weeks, the per-capita allocation had the second best MAS, and the COVIDhub-baseline model had the sixth best allocation (Table \ref{tab:multi-week-performance-summary}).
@@
 \item Several models showed a positive association between allocation score and MWIS ranks (e.g., \texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha}).
-\item Several models had consistently strong MWIS ranks but also had highly variable allocation score ranks with no clear association between the two (e.g., \texttt{JHUAPL-SLPHospEns} and \texttt{CU-select})
+\item One model had consistently strong MWIS ranks but also had highly variable allocation score ranks with no clear association between the two (e.g., \texttt{JHUAPL-SLPHospEns})
 \item One model performed consistently well for both metrics (\texttt{COVIDhub-ensemble})
-\item One model performed consistently well for allocation score but had only middlings ranks for MWIS (\texttt{CMU-TimeSeries})
+\item One model performed consistently well for allocation score but had only middling ranks for MWIS (\texttt{CMU-TimeSeries})
 \end{itemize}
@@
 
-AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 as well as the per-capital allocation (Figure \ref{fig:multi-k}A).
+AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 as well as the per-capita allocation (Figure \ref{fig:multi-k}A).
 These calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of \Sexpr{format(tot_y, big.mark = ",")} that day.
@@
 %The allocation score we developed here does not directly account for important considerations such as fairness or equity of allocations.
-Nor does the proposed framework attempt to capture the broader context of decision making.
+The proposed framework does not attempt to capture the broader context of decision making.
 For example, in practice it may be possible to increase the resource constraint $K$ by shifting funding from other disease mitigation measures.

### 2024-02-06 alloscore_manuscript/alloscore-application-shorter.Rnw

@@
 
-In the wake of the COVID-19 pandemic, researchers have been calling for evaluating infectious disease forecasts through their impact on policy \citep{marshall2023predictions, bilinski_adaptive_2023}.
+In the wake of the COVID-19 pandemic, researchers have been calling for evaluating infectious disease forecasts in ways that align more specifically with how forecasts can be used to inform policy decisions \citep{marshall2023predictions, bilinski_adaptive_2023}.
 Historically, the quality of disease forecasts has typically been measured using scores such as root mean square error (RMSE) \citep{papastefanopoulos2020covid}, log score \citep{mcgowan_collaborative_2019}, or weighted interval score (WIS) \citep{bracher2021evaluating}.
-These scores can be interpreted as measuring how well the forecast corresponds to the observed disease outcome.
-However, as we show in this paper, these scores often do not clearly measure the value of a forecast in the context of a specific decision.
-While these ideas have received little attention in the context of infectious disease forecasting, the results are well known from other fields, such as economics and finance, supply chain management, and meteorology.
+These scores can be interpreted as measuring how well the forecast corresponds to the observed disease outcome but often do not clearly measure the value of a forecast in the context of a specific decision, a result that is well known from other fields, such as economics and finance, supply chain management, and meteorology.
+However, as we show in this paper, it is possible to construct formal methods for evaluating forecasts that are directly and uniquely tied to policy decisions (in our hypothetical setting, a question of resource allocation).
+Furthermore, such scores can illuminate important features of forecasts, and their value as decision-making inputs, that can be missed by traditional accuracy measures.
 

### 2024-02-09 alloscore_manuscript/alloscore-application-shorter.Rnw
intro rewrite

--- 2024-02-06 alloscore_manuscript/alloscore-application-shorter.Rnw
@@
 Infectious disease forecasting models have emerged as important tools in public health outbreak response.
-Predictions of disease dynamics have been used to inform decision making about a wide variety of measures to reduce disease spread and/or mitigate the severity of disease outcomes.
-For example, estimates of the onset time of the flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola and diptheria dynamics have been used to allocate resources \citep{meltzer2014estimating, rainisch2015regional,camacho2015-ebola-bed,finger_real-time_2019}.
-\cite{bertsimas2021predictionsCOVID} developed tools to guide decision making from infectious disease forecasts, which have been used to inform allocation of limited resources such as ventilators, ICU capacity, and vaccine supply.
-Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping (University of Texas, 2022)\nocite{utnews2022}.
+The predictions they provide increasingly inform decisions regarding a wide variety of countermeasures intended to reduce transmission and mitigate the severity of disease outcomes.
+For example, estimates of the onset time of the flu season have been used in developing national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola and diptheria dynamics have been made with the clearly stated goal of helping local
+public health workers choose the timing and location of interventions in settings where resources are severely constrained \citep{meltzer2014estimating, rainisch2015regional,camacho2015-ebola-bed,finger_real-time_2019}.
+More recently, in the context of the outbreak of COVID-19 across the US, \cite{bertsimas2021predictionsCOVID} explicitly based triage decision tools for the interstate reallocation
+of ventilators and ICU capacity on forecast inputs and used these forecasts as well to recommend vaccine trial sites to a 
+major trial sponsor.
+\cite{fox_real-time_2022} similarly used predictive models to inform intrastate resource and care site planning, as well as local community guidelines for masking, traveling, dining and shopping (University of Texas, 2022)\nocite{utnews2022}.
 %In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
 
-In the wake of the COVID-19 pandemic, researchers have been calling for evaluating infectious disease forecasts through their impact on policy \citep{marshall2023predictions, bilinski_adaptive_2023}.
-Historically, the quality of disease forecasts has typically been measured using scores such as root mean square error (RMSE) \citep{papastefanopoulos2020covid}, log score \citep{mcgowan_collaborative_2019}, or weighted interval score (WIS) \citep{bracher2021evaluating}.
-These scores can be interpreted as measuring how well the forecast corresponds to the observed disease outcome.
-However, as we show in this paper, these scores often do not clearly measure the value of a forecast in the context of a specific decision.
-While these ideas have received little attention in the context of infectious disease forecasting, the results are well known from other fields, such as economics and finance, supply chain management, and meteorology.
-
-In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to (see details in Section \ref{sec:methods.detailed.decisiontheory}).
-%There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
-We refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \cite{murphy1993whatisagoodforecast} for discussions focused on applications to economics and meteorology, respectively.
-%In finance, the value of forecasts can often be measured by the profits generated by trading decisions informed by the forecasts, perhaps adjusted for risk levels \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
+In the wake of the COVID-19 pandemic, this trend has been followed by calls for infectious disease forecasts to 
+be not only designed, but 
+also evaluated in ways that align specifically with how forecasts can be used to inform such outbreak control
+decisions \citep{marshall2023predictions, bilinski_adaptive_2023}.
+This contrasts, however, with the historically standard practice of measuring the quality of disease forecasts using general purpose
+accuracy and skill scores, especially those that have implementations available in existing software when the relevant outbreak occurs. For point forecasts,
+the root mean square error (RMSE) (e.g., \cite{papastefanopoulos2020covid}) and the mean absolute error (MAE) (e.g., \cite{johansson2016evaluating}) 
+are common choices.
+For probabilistic forecasts, (which are the focus of this paper), researchers have often relied on the logarithmic score (LS), which
+has been used, for example, to evaluate (with respect to both time and space) the skill of US seasonal influenza forecasts \citep{mcgowan_collaborative_2019,reich_collaborative_2019} as well as forecasts targeting surveillance measures of dengue incidence in Peru and Puerto Rico
+ \citep{johansson_open_2019}.
+More recently, the continuous ranked probability score (CRPS) and a discretized version adapted to multi-quantile forecasts, the weighted interval 
+score (WIS) \citep{bracher2021evaluating} (named for its expression in terms of central prediction interval coverage), have gained prominence.
+For example, the CRPS was used to assess probabilistic forecasts (based on random effect models) of dengue incidence at the district level 
+in Vietnam \citep{colon-gonzalez_probabilistic_2021}. And the WIS has been used during the COVID-19 pandemic (through efforts with which we 
+have been intimately involved) to evaluate forecasts of 
+observed cases, hospitalizations and deaths in the US and Europe, as reported by municipal, state, and federal surveillance systems 
+\citep{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.
+
+While it should be noted that there are ways to interpret any of these scores abstractly through the lens of decision theory, and that all of 
+the application-specific papers cited above benefited from direct collaboration with public health agencies, a key impetus for the
+present work has been that we were not able to find in any of them, nor in the literature they represent, explicit connections between how a 
+forecast was evaluated and how that forecast was used in practice.
+
+A general phenomenon at play here --- one that has been observed repeatedly over the past few decades in other fields such as finance, supply chain management, and meteorology --- is that while scores such as RMSE, MAE, LS, CRPS, and WIS can describe the \emph{quality} of a forecast
+at corresponding to the observed disease outcome, they will often fail to register the \emph{value} (or lack thereof) 
+a forecast has in the context of a specific decision. We refer the reader to \cite{yardley2021utility_cost_forecasts} and the references collected 
+therein (especially the foundational \cite{murphy1993whatisagoodforecast}) for a general overview touching on a wide range of forecasting contexts and  
+also to \cite{pesaran2002decision_based_eval} for a clear discussion from an econometric perspective.
+
+%------ literature that seems to be aware of the problem we see and how it falls short ------%
+This is to say that the quality-value distinction is not a new concept in the larger forecasting community. We are aware, however, of only a 
+limited literature attempting to connect the value of infectious disease forecasts to their impact on and through policy.  And
+within this body of work we have found the discussion of such a connection to still be at a formally and quantitively imprecise stage. 
+In \cite{ioannidis2022forecastingCOVIDfailed}, the possible negative consequences of inaccurate forecasts of infectious disease are discussed, but there is no attempt to quantify the utility or loss incurred as a result of those forecasts.
+\cite{bilinski_adaptive_2023} explore ways in which predictive classifiers of local COVID-19 risk levels in the US could be tuned to policymaker preferences for different
+costs associated with over- and under-reaction to disease dynamics, but they do not clearly identify the source of these costs or how they
+depend on quantifiable policy choices. A similar discussion related to dengue countermeasures in Vietnam appears in
+\cite{colon-gonzalez_probabilistic_2021}.  A novel version of the WIS informally motivated by utility considerations is devoloped in 
+\cite{marshall2023predictions}, but the score is not derived in a decision-theoretic manner.
+There is also a thread of literature that frames infectious disease modeling as a component of a larger system for understanding how policy goals, means, and choices interact and constrain
+one another. As an example, \cite{Probert2016decisionMakingFootMouth} explore how policy recommendations ought to flow from a possibly 
+incongruous set of simulation-based projection models of a hypothetical
+foot-and-mouth disease outbreak when there are ranges of plausible responses and stakeholder interests. Decision theory plays a prominent
+role here but not explicitly as a way to evaluate the choices made in developing or running the models.
+
+% However, as we show in this paper, it is possible, using standard tools of decision theory, to construct formal methods (detailed in Section 
+% \ref{sec:methods.detailed.decisiontheory})) for evaluating probabilistic forecasts that are directly and uniquely tied to policy decisions, 
+% in particular those related to resource allocation.  In essence, we will define the score (as a measure of value) of a forecast as  
+% Furthermore, we find evidence that such scores can illuminate important features of forecasts related to their value as decision-making inputs which
+% are missed by traditional quality (that is, accuracy and skill) measures.
+
+% %------ what you can do instead ------%
+% In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to (see details in Section \ref{sec:methods.detailed.decisiontheory}).
+
+% %In finance, the value of forecasts can often be measured by the profits generated by trading decisions informed by the forecasts, perhaps adjusted for risk levels \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
 In applications to supply chain management and meteorology, the value of forecasts has typically been operationalized by considering the costs associated with decisions regarding the amount of inventory to hold or the level of protection against the impacts of extreme weather events to enact \cite[e.g.,][]{catt2007assessingcostofforecasterror, petropoulos2019inventoryperformanceforecasting, palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
-In this framework, a forecast has value if it leads to decisions with low total costs.
-As one example,
-% in supply chain management these decisions may incur costs related to holding inventory, labor, or providing poor service, while
-in meteorology, policy makers need to balance the costs of implementing protective measures with the costs of potentially preventable weather damages \citep{palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
-Across many fields, analyses have consistently found that common measures of forecast accuracy do not necessarily correspond directly to measures of the value of forecasts as an input to decision making \cite[e.g.,][]{leitch1991economicForecastEval, murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}.
-
-We are aware of only a limited body of work that explicitly attempts to measure the value of infectious disease forecasts through their impact on policy.
-For example, \cite{ioannidis2022forecastingCOVIDfailed} discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts.
-\cite{bilinski_adaptive_2023} explore ways in which policymaker preferences could inform risk thresholds for predictive models using a framework for measuring the costs and losses associated with taking an action that is similar to methods that have been used in meteorology and elsewhere.
-\cite{marshall2023predictions} develop a forecast scoring rule that is informally motivated by utility considerations, but the score is not derived from a decision-theoretic set up.
-Separately, there is a thread of literature that quantifies the link between infectious disease modeling and policy making outside of a forecasting context.
-As an example, \cite{Probert2016decisionMakingFootMouth} develop measures of the cost of actions designed to control a hypothetical outbreak of foot-and-mouth disease and use this framework to explore policy recommendations from a variety of simulation-based projection models.
-
-In practice, probabilistic infectious disease forecasts have most often been made for observations that emerge from public health surveillance systems and have typically been evaluated with common scoring rules that have implementations available in existing software packages.
-For example, seasonal influenza forecasts in the US and dengue forecasts for Peru and Puerto Rico targeted public health surveillance measures of incidence over time and space, and used log score and mean absolute error to evaluate forecast skill \citep{mcgowan_collaborative_2019,reich_collaborative_2019,johansson_open_2019}.
-Pandemic COVID-19 forecasts of observed cases, hospitalizations and/or deaths in the US and Europe, as reported by municipal, state, or federal surveillance systems, were evaluated using the weighted interval score (WIS, which is an approximation of the continuous ranked probability score, or CRPS), and prediction interval coverage \citep{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.
-Similarly, CRPS was also used to assess probabilistic forecasts of dengue incidence at the district level in Vietnam \citep{colon-gonzalez_probabilistic_2021}.
-While some of these scores can be interpreted through the lens of decision theory, and all of the application-specific papers cited above have benefited from direct collaboration with public health agencies, none of them make explicit connections between forecast evaluation and how a forecast was used in practice.
-
+
+% In this framework, a forecast has value if it leads to decisions with low total costs.
+% As one example,
+% % in supply chain management these decisions may incur costs related to holding inventory, labor, or providing poor service, while
+% in meteorology, policy makers need to balance the costs of implementing protective measures with the costs of potentially preventable weather damages \citep{palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
+
+%------ our game plan ------%
 In this work, we begin to fill this gap between the ways that infectious disease forecasts have traditionally been evaluated and the ways that they have been used to support public health policy.
-We consider a setting in which forecasts are used to help determine the allocation of a limited quantity of medical supplies across multiple regions.
+To do so, we consider a setting in which forecasts are used to help determine the allocation of a limited quantity of medical supplies across multiple regions.
 In section~\ref{sec:methods} of the paper, we define a new forecast scoring rule --- the {\em allocation score} --- that evaluates forecasts based on how beneficial resource allocations derived from them would turn out to be.
 %Unlike the traditional measures of forecast skill such as WIS, CRPS, and log score, the proposed allocation score measures the suitability of a forecast as an input to this specific decision making task.
-In section \ref{sec:application}, we present an illustrative analysis using the allocation score to evaluate forecasts of hospital admissions in the US leading up to the Omicron wave in winter 2022.
-This analysis is ``synthetic'' in that it does not correspond to an actual analysis that supported decision making in real time.
-However, the framework described in this paper corresponds to real-world decisions that must be made by public health administrators around the globe, and could be adapted in the future for real-time situations such as informing bed demand, ventilator usage, or doses needed for specific medications.
-%For example, forecasts for districts in Sierra Leone of bed demand to care for patients with Ebola was the subject of a real-time modeling study in late 2014 and early 2015 \citep{camacho2015-ebola-bed}.
-%And, in 2020, a model developed by an academic research group turned predictions of COVID-19 hospitalizations into estimates of ventilator usage and shortages.
-This framework was used by the Hartford HealthCare system in Connecticut ``to align ventilator supply with projected demand at a time where the [COVID-19] pandemic was on the rise'' \citep{bertsimas2021predictionsCOVID}.
-%These examples illustrate the potential for forecasts to inform decisions about how to allocate limited supplies such as temporary hospital beds, ventilators, personal protective equipment, or other supplies that are known to be effective at reducing transmission or severity of disease.
-%However, we emphasize again that these studies did not take the step of evaluating forecasts based on the quality of the allocation decisions that they supported or could have been used to support.
+In section \ref{sec:application}, we present an illustrative analysis using the allocation score to evaluate forecasts of hospital admissions in the US that were made leading up to and during the Omicron wave that peaked in January of 2022.
+This analysis is ``synthetic'' insofar as it is not intended to correspond to any specific historical record of allocation decisions 
+that could have been supported by hospitalization forecasts during this period.
+However, we view the general allocation problem on which our framework is based as a versatile template for 
+formalizing real-world decisions that must constantly be made in real-time by public health administrators around the globe, especially those
+related to hospital capacity, ventilator usage, doses needed for specific medications and other situations where an outbreak creates sudden and
+highly variable demand for potentially scarce resources.
 

### 2024-02-09 alloscore_manuscript/alloscore-application-shorter.tex
intro rewrite

@@
 Infectious disease forecasting models have emerged as important tools in public health outbreak response.
-Predictions of disease dynamics have been used to inform decision making about a wide variety of measures to reduce disease spread and/or mitigate the severity of disease outcomes.
-For example, estimates of expected onset of flu season have been used to aid national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola and diptheria dynamics have been used to allocate resources \citep{meltzer2014estimating, rainisch2015regional,camacho2015-ebola-bed,finger_real-time_2019}.
-\cite{bertsimas2021predictionsCOVID} developed tools to guide decision making from infectious disease forecasts, which have been used to inform allocation of limited resources such as ventilators, ICU capacity, and vaccine supply.
-Models developed by \cite{fox_real-time_2022} have been used to inform resource and care site planning, as well as community guidelines for masking, traveling, dining and shopping (University of Texas, 2022)\nocite{utnews2022}.
+The predictions they provide increasingly inform decisions regarding a wide variety of countermeasures intended to reduce transmission and mitigate the severity of disease outcomes.
+For example, estimates of the onset time of the flu season have been used in developing national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola and diptheria dynamics have been made with the clearly stated goal of helping local
+public health workers choose the timing and location of interventions in settings where resources are severely constrained \citep{meltzer2014estimating, rainisch2015regional,camacho2015-ebola-bed,finger_real-time_2019}.
+More recently, in the context of the outbreak of COVID-19 across the US, \cite{bertsimas2021predictionsCOVID} explicitly based triage decision tools for the interstate reallocation
+of ventilators and ICU capacity on forecast inputs and used these forecasts as well to recommend vaccine trial sites to a 
+major trial sponsor.
+\cite{fox_real-time_2022} similarly used predictive models to inform intrastate resource and care site planning, as well as local community guidelines for masking, traveling, dining and shopping (University of Texas, 2022)\nocite{utnews2022}.
 %In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
 
-In the wake of the COVID-19 pandemic, researchers have been calling for evaluating forecasts through their impact on policy \citep{marshall2023predictions, bilinski_adaptive_2023}.
-The value of infectious disease forecasts has typically been measured by how closely they predict disease outcomes such as cases, hospitalizations or deaths using, for example, root mean square error (RMSE) \citep{papastefanopoulos2020covid}, log-score \citep{mcgowan_collaborative_2019}, or weighted interval score (WIS) \citep{bracher2021evaluating}.
-However, as we show in this paper, these metrics only evaluate one specific dimension of accuracy, and often do not have a clear correspondence to features that would make a forecast valuable in the context of a specific decision.
-While these ideas have received little attention in the context of infectious disease forecasting, the results are well known from other fields, such as economics and finance, supply chain management, and meteorology.
+In the wake of the COVID-19 pandemic, this trend has been followed by calls for infectious disease forecasts to 
+be not only designed, but 
+also evaluated in ways that align specifically with how forecasts can be used to inform such outbreak control
+decisions \citep{marshall2023predictions, bilinski_adaptive_2023}.
+This contrasts, however, with the historically standard practice of measuring the quality of disease forecasts using general purpose
+accuracy and skill scores, especially those that have implementations available in existing software when the relevant outbreak occurs. For point forecasts,
+the root mean square error (RMSE) (e.g., \cite{papastefanopoulos2020covid}) and the mean absolute error (MAE) (e.g., \cite{johansson2016evaluating}) 
+are common choices.
+For probabilistic forecasts, (which are the focus of this paper), researchers have often relied on the logarithmic score (LS), which
+has been used, for example, to evaluate (with respect to both time and space) the skill of US seasonal influenza forecasts \citep{mcgowan_collaborative_2019,reich_collaborative_2019} as well as forecasts targeting surveillance measures of dengue incidence in Peru and Puerto Rico
+ \citep{johansson_open_2019}.
+More recently, the continuous ranked probability score (CRPS) and a discretized version adapted to multi-quantile forecasts, the weighted interval 
+score (WIS) \citep{bracher2021evaluating} (named for its expression in terms of central prediction interval coverage), have gained prominence.
+For example, the CRPS was used to assess probabilistic forecasts (based on random effect models) of dengue incidence at the district level 
+in Vietnam \citep{colon-gonzalez_probabilistic_2021}. And the WIS has been used during the COVID-19 pandemic (through efforts with which we 
+have been intimately involved) to evaluate forecasts of 
+observed cases, hospitalizations and deaths in the US and Europe, as reported by municipal, state, and federal surveillance systems 
+\citep{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.
 
-In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to (see details in Section \ref{sec:methods.detailed.decisiontheory}).
-%There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision making context, primarily in fields such as economics and finance, supply chain management, and meteorology.
-We refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \cite{murphy1993whatisagoodforecast} for discussions focused on applications to economics and meteorology, respectively.
-%In finance, the value of forecasts can often be measured by the profits generated by trading decisions informed by the forecasts, perhaps adjusted for risk levels \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
+While it should be noted that there are ways to interpret any of these scores abstractly through the lens of decision theory, and that all of 
+the application-specific papers cited above benefited from direct collaboration with public health agencies, a key impetus for the
+present work has been that we were not able to find in any of them, nor in the literature they represent, explicit connections between how a 
+forecast was evaluated and how that forecast was used in practice.
+
+A general phenomenon at play here --- one that has been observed repeatedly over the past few decades in other fields such as finance, supply chain management, and meteorology --- is that while scores such as RMSE, MAE, LS, CRPS, and WIS can describe the \emph{quality} of a forecast
+at corresponding to the observed disease outcome, they will often fail to register the \emph{value} (or lack thereof) 
+a forecast has in the context of a specific decision. We refer the reader to \cite{yardley2021utility_cost_forecasts} and the references collected 
+therein (especially the foundational \cite{murphy1993whatisagoodforecast}) for a general overview touching on a wide range of forecasting contexts and  
+also to \cite{pesaran2002decision_based_eval} for a clear discussion from an econometric perspective.
+
+%------ literature that seems to be aware of the problem we see and how it falls short ------%
+This is to say that the quality-value distinction is not a new concept in the larger forecasting community. We are aware, however, of only a 
+limited literature attempting to connect the value of infectious disease forecasts to their impact on and through policy.  And
+within this body of work we have found the discussion of such a connection to still be at a formally and quantitively imprecise stage. 
+In \cite{ioannidis2022forecastingCOVIDfailed}, the possible negative consequences of inaccurate forecasts of infectious disease are discussed, but there is no attempt to quantify the utility or loss incurred as a result of those forecasts.
+\cite{bilinski_adaptive_2023} explore ways in which predictive classifiers of local COVID-19 risk levels in the US could be tuned to policymaker preferences for different
+costs associated with over- and under-reaction to disease dynamics, but they do not clearly identify the source of these costs or how they
+depend on quantifiable policy choices. A similar discussion related to dengue countermeasures in Vietnam appears in
+\cite{colon-gonzalez_probabilistic_2021}.  A novel version of the WIS informally motivated by utility considerations is devoloped in 
+\cite{marshall2023predictions}, but the score is not derived in a decision-theoretic manner.
+There is also a thread of literature that frames infectious disease modeling as a component of a larger system for understanding how policy goals, means, and choices interact and constrain
+one another. As an example, \cite{Probert2016decisionMakingFootMouth} explore how policy recommendations ought to flow from a possibly 
+incongruous set of simulation-based projection models of a hypothetical
+foot-and-mouth disease outbreak when there are ranges of plausible responses and stakeholder interests. Decision theory plays a prominent
+role here but not explicitly as a way to evaluate the choices made in developing or running the models.
+
+% However, as we show in this paper, it is possible, using standard tools of decision theory, to construct formal methods (detailed in Section 
+% \ref{sec:methods.detailed.decisiontheory})) for evaluating probabilistic forecasts that are directly and uniquely tied to policy decisions, 
+% in particular those related to resource allocation.  In essence, we will define the score (as a measure of value) of a forecast as  
+% Furthermore, we find evidence that such scores can illuminate important features of forecasts related to their value as decision-making inputs which
+% are missed by traditional quality (that is, accuracy and skill) measures.
+
+% %------ what you can do instead ------%
+% In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to (see details in Section \ref{sec:methods.detailed.decisiontheory}).
+
+% %In finance, the value of forecasts can often be measured by the profits generated by trading decisions informed by the forecasts, perhaps adjusted for risk levels \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
 In applications to supply chain management and meteorology, the value of forecasts has typically been operationalized by considering the costs associated with decisions regarding the amount of inventory to hold or the level of protection against the impacts of extreme weather events to enact \cite[e.g.,][]{catt2007assessingcostofforecasterror, petropoulos2019inventoryperformanceforecasting, palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
-In this framework, a forecast has value if it leads to decisions with low total costs.
-As one example,
-% in supply chain management these decisions may incur costs related to holding inventory, labor, or providing poor service, while
-in meteorology, policy makers need to balance the costs of implementing protective measures with the costs of potentially preventable weather damages \citep{palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
-Across many fields, analyses have consistently found that common measures of forecast accuracy do not necessarily correspond directly to measures of the value of forecasts as an input to decision making \cite[e.g.,][]{leitch1991economicForecastEval, murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}.
 
-We are aware of only a limited body of work that explicitly attempts to measure the value of infectious disease forecasts through their impact on policy.
-For example, \cite{ioannidis2022forecastingCOVIDfailed} discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts.
-\cite{bilinski_adaptive_2023} explore ways in which policymaker preferences could inform risk thresholds for predictive models using a framework for measuring the costs and losses associated with taking an action that is similar to methods that have been used in meteorology and elsewhere.
-\cite{marshall2023predictions} develop a forecast scoring rule that is informally motivated by utility considerations, but the score is not derived from a decision-theoretic set up.
-Separately, there is a thread of literature that quantifies the link between infectious disease modeling and policy making outside of a forecasting context.
-As an example, \cite{Probert2016decisionMakingFootMouth} develop measures of the cost of actions designed to control a hypothetical outbreak of foot-and-mouth disease and use this framework to explore policy recommendations from a variety of simulation-based projection models.
+% In this framework, a forecast has value if it leads to decisions with low total costs.
+% As one example,
+% % in supply chain management these decisions may incur costs related to holding inventory, labor, or providing poor service, while
+% in meteorology, policy makers need to balance the costs of implementing protective measures with the costs of potentially preventable weather damages \citep{palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
 
-In practice, probabilistic infectious disease forecasts have most often been made for observations that emerge from public health surveillance systems and have typically been evaluated with common scoring rules that have implementations available in existing software packages.
-For example, seasonal influenza forecasts in the US and dengue forecasts for Peru and Puerto Rico targeted public health surveillance measures of incidence over time and space, and used log-score and mean absolute errors to evaluate forecast skill \citep{mcgowan_collaborative_2019,reich_collaborative_2019,johansson_open_2019}.
-Pandemic COVID-19 forecasts of observed cases, hospitalizations and/or deaths in the US and Europe, as reported by municipal, state, or federal surveillance systems, were evaluated using the weighted interval score (WIS, which is an approximation of the continuous ranked probability score, or CRPS), and prediction interval coverage \citep{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.
-Similarly, CRPS was also used to assess probabilistic forecasts of dengue incidence at the district level in Vietnam \citep{colon-gonzalez_probabilistic_2021}.
-While some of these scores can be interpreted through the lens of decision theory, and all of the application-specific papers cited above have benefited from direct collaboration with public health agencies, none of them make explicit connections between forecast evaluation and how a forecast was used in practice.
-
+%------ our game plan ------%
 In this work, we begin to fill this gap between the ways that infectious disease forecasts have traditionally been evaluated and the ways that they have been used to support public health policy.
-We consider a setting in which forecasts are used to help determine the allocation of a limited quantity of medical supplies across multiple regions.
+To do so, we consider a setting in which forecasts are used to help determine the allocation of a limited quantity of medical supplies across multiple regions.
 In section~\ref{sec:methods} of the paper, we define a new forecast scoring rule --- the {\em allocation score} --- that evaluates forecasts based on how beneficial resource allocations derived from them would turn out to be.
 %Unlike the traditional measures of forecast skill such as WIS, CRPS, and log score, the proposed allocation score measures the suitability of a forecast as an input to this specific decision making task.
-
-%Briefly, the allocation score of a forecast is the avoidable unmet need that results from using that forecast to set resource allocations by minimizing expected unmet need.
-%For example, suppose that a decision maker is provided with forecasts of the level of need for medical resources in each of several states or hospital systems.
-%If there is a limited amount of the medical resource that is available to distribute, a decision maker could choose an allocation of that resource across locations that minimizes the expected unmet need according to that forecast.
-%As measured by the allocation score, one forecast is better than another if it would lead decision makers to an allocation that results in less unmet need.
-%By ``unnecessary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
-%If the amount of resources that is available to distribute is less than the actual need, some amount of unmet need is unavoidable.
-%The allocation score for a forecast does not include the unmet need that was unavoidable given the resource constraint, and so it measures only the amount of unmet need that could have been prevented by using a different allocation of available resources than that suggested by the forecast.
-%We elaborate on these ideas in Section~\ref{sec:methods}.
-
-In section \ref{sec:application}, we present an illustrative analysis using the allocation score to evaluate forecasts of hospital admissions in the US leading up to the Omicron wave in winter 2022.
-This analysis is ``synthetic'' in that it does not correspond to an actual analysis that supported decision making in real time.
-However, the framework described in this paper corresponds to real-world decisions that must be made by public health administrators around the globe, and could be adapted in the future for such real-time situations such as informing bed demand, ventilator usage, or doses needed for specific medications.
-%For example, forecasts for districts in Sierra Leone of bed demand to care for patients with Ebola was the subject of a real-time modeling study in late 2014 and early 2015 \citep{camacho2015-ebola-bed}.
-%And, in 2020, a model developed by an academic research group turned predictions of COVID-19 hospitalizations into estimates of ventilator usage and shortages.
-This framework was used by the Hartford HealthCare system in Connecticut ``to align ventilator supply with projected demand at a time where the [COVID-19] pandemic was on the rise'' \citep{bertsimas2021predictionsCOVID}.
-%These examples illustrate the potential for forecasts to inform decisions about how to allocate limited supplies such as temporary hospital beds, ventilators, personal protective equipment, or other supplies that are known to be effective at reducing transmission or severity of disease.
-However, we emphasize again that these studies did not take the step of evaluating forecasts based on the quality of the allocation decisions that they supported or could have been used to support.
+In section \ref{sec:application}, we present an illustrative analysis using the allocation score to evaluate forecasts of hospital admissions in the US that were made leading up to and during the Omicron wave that peaked in January of 2022.
+This analysis is ``synthetic'' insofar as it is not intended to correspond to any specific historical record of allocation decisions 
+that could have been supported by hospitalization forecasts during this period.
+However, we view the general allocation problem on which our framework is based as a versatile template for 
+formalizing real-world decisions that must constantly be made in real-time by public health administrators around the globe, especially those
+related to hospital capacity, ventilator usage, doses needed for specific medications and other situations where an outbreak creates sudden and
+highly variable demand for potentially scarce resources.
 
@@
 
-The hospitalization data used for scoring forecasts were downloaded on February 01, 2024.
+The hospitalization data used for scoring forecasts were downloaded on February 07, 2024.
 
@@
 
-% latex table generated in R 4.3.2 by xtable 1.8-4 package
-% Mon Feb  5 14:04:06 2024
+% latex table generated in R 4.3.1 by xtable 1.8-4 package
+% Thu Feb  8 21:06:12 2024
 \begin{table}[ht]
@@
 The purple bar goes from zero to the amount of need that was met by the allocation from that model to a specific location.
-A red bar indicates need that exceeded the resource allocation for a location, and a light blue bar shows suggested resource allocation suggested by that model that were above the need for that location.}\label{fig:thermometer-plot}
+A red bar indicates need that exceeded the resource allocation for a location, and a light blue bar shows the amount by which the resource allocation suggested by that model exceeded the need for that location.}\label{fig:thermometer-plot}
 \end{figure}
@@
 
-% latex table generated in R 4.3.2 by xtable 1.8-4 package
-% Mon Feb  5 14:04:11 2024
+% latex table generated in R 4.3.1 by xtable 1.8-4 package
+% Thu Feb  8 21:06:16 2024
 \begin{table}[ht]
@@
 
-AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 as well as the per-capital allocation (Figure \ref{fig:multi-k}A).
+AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 as well as the per-capita allocation (Figure \ref{fig:multi-k}A).
 These calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of 19,581 that day.

### 2024-02-09 alloscore_manuscript/alloscore-application-shorter.Rnw
some minor updates to intro for wording and sentence structure

@@
 public health workers choose the timing and location of interventions in settings where resources are severely constrained \citep{meltzer2014estimating, rainisch2015regional,camacho2015-ebola-bed,finger_real-time_2019}.
-More recently, in the context of the outbreak of COVID-19 across the US, \cite{bertsimas2021predictionsCOVID} explicitly based triage decision tools for the interstate reallocation
-of ventilators and ICU capacity on forecast inputs and used these forecasts as well to recommend vaccine trial sites to a 
+More recently, in the context of the outbreak of COVID-19 across the US, \cite{bertsimas2021predictionsCOVID} used forecasts as inputs to decision tools for the interstate reallocation
+of ventilators and ICU capacity, and to recommend vaccine trial sites to a 
 major trial sponsor.
@@
 are common choices.
-For probabilistic forecasts, (which are the focus of this paper), researchers have often relied on the logarithmic score (LS), which
-has been used, for example, to evaluate (with respect to both time and space) the skill of US seasonal influenza forecasts \citep{mcgowan_collaborative_2019,reich_collaborative_2019} as well as forecasts targeting surveillance measures of dengue incidence in Peru and Puerto Rico
+For probabilistic forecasts, which are the focus of this paper, researchers have often relied on the logarithmic score (LS).
+For example, the LS has been used to evaluate the skill of US seasonal influenza forecasts \citep{mcgowan_collaborative_2019,reich_collaborative_2019} as well as forecasts targeting surveillance measures of dengue incidence in Peru and Puerto Rico
  \citep{johansson_open_2019}.
 More recently, the continuous ranked probability score (CRPS) and a discretized version adapted to multi-quantile forecasts, the weighted interval 
-score (WIS) \citep{bracher2021evaluating} (named for its expression in terms of central prediction interval coverage), have gained prominence.
+score (WIS) \citep{bracher2021evaluating}, have gained prominence.
 For example, the CRPS was used to assess probabilistic forecasts (based on random effect models) of dengue incidence at the district level 
-in Vietnam \citep{colon-gonzalez_probabilistic_2021}. And the WIS has been used during the COVID-19 pandemic (through efforts with which we 
-have been intimately involved) to evaluate forecasts of 
+in Vietnam \citep{colon-gonzalez_probabilistic_2021}. And the WIS has been used during the COVID-19 pandemic to evaluate forecasts of 
 observed cases, hospitalizations and deaths in the US and Europe, as reported by municipal, state, and federal surveillance systems 
@@
 A general phenomenon at play here --- one that has been observed repeatedly over the past few decades in other fields such as finance, supply chain management, and meteorology --- is that while scores such as RMSE, MAE, LS, CRPS, and WIS can describe the \emph{quality} of a forecast
-at corresponding to the observed disease outcome, they will often fail to register the \emph{value} (or lack thereof) 
-a forecast has in the context of a specific decision. We refer the reader to \cite{yardley2021utility_cost_forecasts} and the references collected 
+in terms of how well it corresponds to the observed disease outcome, they will often fail to register the \emph{value} of a forecast in the context of a specific decision.
+We refer the reader to \cite{yardley2021utility_cost_forecasts} and the references collected
 therein (especially the foundational \cite{murphy1993whatisagoodforecast}) for a general overview touching on a wide range of forecasting contexts and  
@@
 %------ literature that seems to be aware of the problem we see and how it falls short ------%
-This is to say that the quality-value distinction is not a new concept in the larger forecasting community. We are aware, however, of only a 
-limited literature attempting to connect the value of infectious disease forecasts to their impact on and through policy.  And
+Although there has been discussion of the quality-value distinction in the larger forecasting community, we are aware of only a 
+limited literature attempting to connect the value of \emph{infectious disease} forecasts to their impact on and through policy.  And
 within this body of work we have found the discussion of such a connection to still be at a formally and quantitively imprecise stage. 
@@
 depend on quantifiable policy choices. A similar discussion related to dengue countermeasures in Vietnam appears in
-\cite{colon-gonzalez_probabilistic_2021}.  A novel version of the WIS informally motivated by utility considerations is devoloped in 
+\cite{colon-gonzalez_probabilistic_2021}.  A novel version of the WIS informally motivated by utility considerations is developed in 
 \cite{marshall2023predictions}, but the score is not derived in a decision-theoretic manner.
 There is also a thread of literature that frames infectious disease modeling as a component of a larger system for understanding how policy goals, means, and choices interact and constrain
-one another. As an example, \cite{Probert2016decisionMakingFootMouth} explore how policy recommendations ought to flow from a possibly 
-incongruous set of simulation-based projection models of a hypothetical
+one another. As an example, \cite{Probert2016decisionMakingFootMouth} explore how policy recommendations ought to flow from a set of simulation-based projection models of a hypothetical
 foot-and-mouth disease outbreak when there are ranges of plausible responses and stakeholder interests. Decision theory plays a prominent
-role here but not explicitly as a way to evaluate the choices made in developing or running the models.
+role here, but not explicitly as a way to evaluate the choices made in developing the models.
 

### 2024-02-09 alloscore_manuscript/alloscore-application-shorter.Rnw
merge elr edits; cut out some commented text

@@
 
-Recent years have seen increasing efforts to forecast infectious disease burdens, with a primary goal being to help public health workers make informed policy decisions.
-However, there has only been limited discussion of how predominant forecast evaluation metrics might indicate the success of policies based in part on those forecasts.
-We explore one possible tether between forecasts and policy: the allocation of limited medical resources so as to minimize unmet need.
-We use probabilistic forecasts of disease burden in each of several regions to determine optimal resource allocations, and then we score forecasts according to how much unmet need their associated allocations would have allowed.
-We illustrate with forecasts of COVID-19 hospitalizations in the US, and we find that the forecast skill ranking given by this allocation scoring rule can vary substantially from the ranking given by the weighted interval score.
-We see this as evidence that the allocation scoring rule detects forecast value that is missed by traditional accuracy measures and that the general strategy of designing scoring rules that are directly linked to policy performance is a promising direction for epidemic forecast evaluation.
+Recent years have seen increasing efforts to forecast infectious disease burdens, with a primary goal being to help
+public health workers make informed policy decisions. However, there has only been limited discussion of how
+predominant forecast evaluation metrics might indicate the success of policies based in part on those forecasts. We
+explore one possible tether between forecasts and policy: the allocation of limited medical resources so as to minimize
+unmet need. We use probabilistic forecasts of disease burden in each of several regions to determine optimal resource
+allocations, and then we score forecasts according to how much unmet need their associated allocations would have
+allowed. We illustrate with forecasts of COVID-19 hospitalizations in the US, and we find that the forecast skill
+ranking given by this allocation scoring rule can vary substantially from the ranking given by the weighted interval
+score. We see this as evidence that the allocation scoring rule detects forecast value that is missed by traditional
+accuracy measures and that the general strategy of designing scoring rules that are directly linked to policy
+performance is a promising direction for epidemic forecast evaluation.
 
@@
 
-Infectious disease forecasting models have emerged as important tools in public health outbreak response.
-The predictions they provide increasingly inform decisions regarding a wide variety of countermeasures intended to reduce transmission and mitigate the severity of disease outcomes.
-For example, estimates of the onset time of the flu season have been used in developing national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola and diptheria dynamics have been made with the clearly stated goal of helping local
-public health workers choose the timing and location of interventions in settings where resources are severely constrained \citep{meltzer2014estimating, rainisch2015regional,camacho2015-ebola-bed,finger_real-time_2019}.
-More recently, in the context of the outbreak of COVID-19 across the US, \cite{bertsimas2021predictionsCOVID} used forecasts as inputs to decision tools for the interstate reallocation
-of ventilators and ICU capacity, and to recommend vaccine trial sites to a 
-major trial sponsor.
-\cite{fox_real-time_2022} similarly used predictive models to inform intrastate resource and care site planning, as well as local community guidelines for masking, traveling, dining and shopping (University of Texas, 2022)\nocite{utnews2022}.
-%In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
-
-In the wake of the COVID-19 pandemic, this trend has been followed by calls for infectious disease forecasts to 
-be not only designed, but 
-also evaluated in ways that align specifically with how forecasts can be used to inform such outbreak control
-decisions \citep{marshall2023predictions, bilinski_adaptive_2023}.
-This contrasts, however, with the historically standard practice of measuring the quality of disease forecasts using general purpose
-accuracy and skill scores, especially those that have implementations available in existing software when the relevant outbreak occurs. For point forecasts,
-the root mean square error (RMSE) (e.g., \cite{papastefanopoulos2020covid}) and the mean absolute error (MAE) (e.g., \cite{johansson2016evaluating}) 
-are common choices.
-For probabilistic forecasts, which are the focus of this paper, researchers have often relied on the logarithmic score (LS).
-For example, the LS has been used to evaluate the skill of US seasonal influenza forecasts \citep{mcgowan_collaborative_2019,reich_collaborative_2019} as well as forecasts targeting surveillance measures of dengue incidence in Peru and Puerto Rico
- \citep{johansson_open_2019}.
-More recently, the continuous ranked probability score (CRPS) and a discretized version adapted to multi-quantile forecasts, the weighted interval 
-score (WIS) \citep{bracher2021evaluating}, have gained prominence.
-For example, the CRPS was used to assess probabilistic forecasts (based on random effect models) of dengue incidence at the district level 
-in Vietnam \citep{colon-gonzalez_probabilistic_2021}. And the WIS has been used during the COVID-19 pandemic to evaluate forecasts of 
-observed cases, hospitalizations and deaths in the US and Europe, as reported by municipal, state, and federal surveillance systems 
-\citep{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.
-
-While it should be noted that there are ways to interpret any of these scores abstractly through the lens of decision theory, and that all of 
-the application-specific papers cited above benefited from direct collaboration with public health agencies, a key impetus for the
-present work has been that we were not able to find in any of them, nor in the literature they represent, explicit connections between how a 
-forecast was evaluated and how that forecast was used in practice.
-
-A general phenomenon at play here --- one that has been observed repeatedly over the past few decades in other fields such as finance, supply chain management, and meteorology --- is that while scores such as RMSE, MAE, LS, CRPS, and WIS can describe the \emph{quality} of a forecast
-in terms of how well it corresponds to the observed disease outcome, they will often fail to register the \emph{value} of a forecast in the context of a specific decision.
-We refer the reader to \cite{yardley2021utility_cost_forecasts} and the references collected
-therein (especially the foundational \cite{murphy1993whatisagoodforecast}) for a general overview touching on a wide range of forecasting contexts and  
-also to \cite{pesaran2002decision_based_eval} for a clear discussion from an econometric perspective.
-
-%------ literature that seems to be aware of the problem we see and how it falls short ------%
-Although there has been discussion of the quality-value distinction in the larger forecasting community, we are aware of only a 
-limited literature attempting to connect the value of \emph{infectious disease} forecasts to their impact on and through policy.  And
-within this body of work we have found the discussion of such a connection to still be at a formally and quantitively imprecise stage. 
-In \cite{ioannidis2022forecastingCOVIDfailed}, the possible negative consequences of inaccurate forecasts of infectious disease are discussed, but there is no attempt to quantify the utility or loss incurred as a result of those forecasts.
-\cite{bilinski_adaptive_2023} explore ways in which predictive classifiers of local COVID-19 risk levels in the US could be tuned to policymaker preferences for different
-costs associated with over- and under-reaction to disease dynamics, but they do not clearly identify the source of these costs or how they
-depend on quantifiable policy choices. A similar discussion related to dengue countermeasures in Vietnam appears in
-\cite{colon-gonzalez_probabilistic_2021}.  A novel version of the WIS informally motivated by utility considerations is developed in 
-\cite{marshall2023predictions}, but the score is not derived in a decision-theoretic manner.
-There is also a thread of literature that frames infectious disease modeling as a component of a larger system for understanding how policy goals, means, and choices interact and constrain
-one another. As an example, \cite{Probert2016decisionMakingFootMouth} explore how policy recommendations ought to flow from a set of simulation-based projection models of a hypothetical
-foot-and-mouth disease outbreak when there are ranges of plausible responses and stakeholder interests. Decision theory plays a prominent
-role here, but not explicitly as a way to evaluate the choices made in developing the models.
-
-% However, as we show in this paper, it is possible, using standard tools of decision theory, to construct formal methods (detailed in Section 
-% \ref{sec:methods.detailed.decisiontheory})) for evaluating probabilistic forecasts that are directly and uniquely tied to policy decisions, 
-% in particular those related to resource allocation.  In essence, we will define the score (as a measure of value) of a forecast as  
-% Furthermore, we find evidence that such scores can illuminate important features of forecasts related to their value as decision-making inputs which
-% are missed by traditional quality (that is, accuracy and skill) measures.
-
-% %------ what you can do instead ------%
-% In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to (see details in Section \ref{sec:methods.detailed.decisiontheory}).
-
-% %In finance, the value of forecasts can often be measured by the profits generated by trading decisions informed by the forecasts, perhaps adjusted for risk levels \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
-In applications to supply chain management and meteorology, the value of forecasts has typically been operationalized by considering the costs associated with decisions regarding the amount of inventory to hold or the level of protection against the impacts of extreme weather events to enact \cite[e.g.,][]{catt2007assessingcostofforecasterror, petropoulos2019inventoryperformanceforecasting, palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
-
-% In this framework, a forecast has value if it leads to decisions with low total costs.
-% As one example,
-% % in supply chain management these decisions may incur costs related to holding inventory, labor, or providing poor service, while
-% in meteorology, policy makers need to balance the costs of implementing protective measures with the costs of potentially preventable weather damages \citep{palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
-
-%------ our game plan ------%
-In this work, we begin to fill this gap between the ways that infectious disease forecasts have traditionally been evaluated and the ways that they have been used to support public health policy.
-To do so, we consider a setting in which forecasts are used to help determine the allocation of a limited quantity of medical supplies across multiple regions.
-In section~\ref{sec:methods} of the paper, we define a new forecast scoring rule --- the {\em allocation score} --- that evaluates forecasts based on how beneficial resource allocations derived from them would turn out to be.
-%Unlike the traditional measures of forecast skill such as WIS, CRPS, and log score, the proposed allocation score measures the suitability of a forecast as an input to this specific decision making task.
-In section \ref{sec:application}, we present an illustrative analysis using the allocation score to evaluate forecasts of hospital admissions in the US that were made leading up to and during the Omicron wave that peaked in January of 2022.
-This analysis is ``synthetic'' insofar as it is not intended to correspond to any specific historical record of allocation decisions 
-that could have been supported by hospitalization forecasts during this period.
-However, we view the general allocation problem on which our framework is based as a versatile template for 
-formalizing real-world decisions that must constantly be made in real-time by public health administrators around the globe, especially those
-related to hospital capacity, ventilator usage, doses needed for specific medications and other situations where an outbreak creates sudden and
-highly variable demand for potentially scarce resources.
-
-%The remainder of this article is organized as follows.
-%We describe the allocation score in Section \ref{sec:methods}, and in Section \ref{sec:application} we illustrate the use of the score in an application to evaluate short-term forecasts of COVID-19 hospital admissions in the US.
-%Section \ref{sec:discussion} summarizes our contributions and discusses opportunities for further extensions in future work.
+Infectious disease forecasting models have emerged as important tools in public health outbreak response. The
+predictions they provide increasingly inform decisions regarding a wide variety of countermeasures intended to reduce
+transmission and mitigate the severity of disease outcomes. For example, estimates of the onset time of the flu season
+have been used in developing national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola and
+diptheria dynamics have been made with the clearly stated goal of helping local public health workers choose the timing
+and location of interventions in settings where resources are severely constrained 
+\citep{meltzer2014estimating, rainisch2015regional, camacho2015-ebola-bed,finger_real-time_2019}. More recently, in the
+context of the outbreak of COVID-19 across the US, \cite{bertsimas2021predictionsCOVID} used forecasts as inputs to
+decision tools for the interstate reallocation of ventilators and ICU capacity, and to recommend vaccine trial sites to
+a major trial sponsor. \cite{fox_real-time_2022} similarly used predictive models to inform intrastate resource and
+care site planning, as well as local community guidelines for masking, traveling, dining and shopping (University of
+Texas, 2022)\nocite{utnews2022}.
+
+In the wake of the COVID-19 pandemic, this trend has been followed by calls for infectious disease forecasts to be not
+only designed, but also evaluated in ways that align specifically with how forecasts can be used to inform such outbreak
+control decisions \citep{marshall2023predictions, bilinski_adaptive_2023}. This contrasts, however, with the
+historically standard practice of measuring the quality of disease forecasts using general purpose accuracy and skill
+scores, especially those that have implementations available in existing software when the relevant outbreak occurs. For
+point forecasts, the root mean square error (RMSE) (e.g., \cite{papastefanopoulos2020covid}) and the mean absolute error
+(MAE) (e.g., \cite{johansson2016evaluating}) are common choices. For probabilistic forecasts, which are the focus of
+this paper, researchers have often relied on the logarithmic score (LS). For example, the LS has been used to evaluate
+the skill of US seasonal influenza forecasts \citep{mcgowan_collaborative_2019,reich_collaborative_2019} as well as
+forecasts targeting surveillance measures of dengue incidence in Peru and Puerto Rico \citep{johansson_open_2019}. More
+recently, the continuous ranked probability score (CRPS) and a discretized version adapted to multi-quantile forecasts,
+the weighted interval score (WIS) \citep{bracher2021evaluating}, have gained prominence. For example, the CRPS was used
+to assess probabilistic forecasts (based on random effect models) of dengue incidence at the district level in Vietnam
+\citep{colon-gonzalez_probabilistic_2021}. And the WIS has been used during the COVID-19 pandemic to evaluate forecasts
+of observed cases, hospitalizations and deaths in the US and Europe, as reported by municipal, state, and federal
+surveillance systems \citep{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.
+
+While it should be noted that there are ways to interpret any of these scores abstractly through the lens of decision
+theory, and that all of the application-specific papers cited above benefited from direct collaboration with public
+health agencies, a key impetus for the present work has been that we were not able to find in any of them, nor in the
+literature they represent, explicit connections between how a forecast was evaluated and how that forecast was used in
+practice.
+
+A general phenomenon at play here --- one that has been observed repeatedly over the past few decades in other fields
+such as finance, supply chain management, and meteorology --- is that while scores such as RMSE, MAE, LS, CRPS, and WIS
+can describe the \emph{quality} of a forecast in terms of how well it corresponds to the observed disease outcome, they
+will often fail to register the \emph{value} of a forecast in the context of a specific decision. We refer the reader to
+\cite{yardley2021utility_cost_forecasts} and the references collected therein (especially the foundational
+\cite{murphy1993whatisagoodforecast}) for a general overview touching on a wide range of forecasting contexts and  also
+to \cite{pesaran2002decision_based_eval} for a clear discussion from an econometric perspective.
+
+Despite this now well-developed discussion of the quality-value distinction in the larger forecasting community, we are
+aware of only a limited literature attempting to connect the value of \emph{infectious disease} forecasts to their
+impact on and through policy.  And within this body of work we have found the discussion of such a connection to still
+be at a formally and quantitively imprecise stage. In \cite{ioannidis2022forecastingCOVIDfailed}, the possible negative
+consequences of inaccurate forecasts of infectious disease are discussed, but there is no attempt to quantify the
+utility or loss incurred as a result of those forecasts. \cite{bilinski_adaptive_2023} explore ways in which predictive
+classifiers of local COVID-19 risk levels in the US could be tuned to policymaker preferences for different costs
+associated with over- and under-reaction to disease dynamics, but they do not clearly identify the source of these
+costs or how they depend on quantifiable policy choices. A similar discussion related to dengue countermeasures in
+Vietnam appears in \cite{colon-gonzalez_probabilistic_2021}.  A novel version of the WIS informally motivated by
+utility considerations is developed in \cite{marshall2023predictions}, but the score is not derived in a
+decision-theoretic manner.  There is also a thread of literature that frames infectious disease modeling as a component
+of a larger system for understanding how policy goals, means, and choices interact and constrain one another. As an
+example, \cite{Probert2016decisionMakingFootMouth} explore how policy recommendations ought to flow from a possibly 
+incongruous set of simulation-based projection models of a hypothetical foot-and-mouth disease outbreak when there are 
+ranges of plausible responses and stakeholder interests. Decision theory plays a prominent role here, but not explicitly
+as a way to evaluate the choices made in developing the models.
+
+In this work, we begin to fill this gap between the ways that infectious disease forecasts have traditionally been
+evaluated and the ways that they have been used to support public health policy. To do so, we consider a setting in
+which forecasts are used to help determine the allocation of a limited quantity of medical supplies across multiple
+regions. In section~\ref{sec:methods} of the paper, we define a new forecast scoring rule --- the {\em allocation score}
+--- that evaluates forecasts based on how beneficial resource allocations derived from them would turn out to be. In
+section \ref{sec:application}, we present an illustrative analysis using the allocation score to evaluate forecasts of
+hospital admissions in the US that were made leading up to and during the Omicron wave that peaked in January of 2022.
+This analysis is ``synthetic'' insofar as it is not intended to correspond to any specific historical record of
+allocation decisions that could have been supported by hospitalization forecasts during this period. However, we view
+the general allocation problem on which our framework is based as a versatile template for formalizing real-world
+decisions that must constantly be made in real-time by public health administrators around the globe, especially those
+related to hospital capacity, ventilator usage, doses needed for specific medications and other situations where an
+outbreak creates sudden and highly variable demand for potentially scarce resources.
 
@@
 
-We begin with an informal description of the allocation score and some examples illustrating its key characteristics in section \ref{sec:methods.overview}.
-In section \ref{sec:methods.detailed} we develop the allocation score more carefully, using a decision theoretic procedure for deriving proper scoring rules.
-(See section \ref{sec:proper} of the supplement for a definition of a proper scoring rule
-and a more technical discussion of the procedure).
-In section \ref{sec:methods.related}, we note that another group of common scores including the quantile score, WIS, and CRPS, can also be derived from decision theoretic foundations \textemdash starting from a different decision making context.
+We begin with an informal description of the allocation score and some examples illustrating its key characteristics in
+section \ref{sec:methods.overview}. In section \ref{sec:methods.detailed} we develop the allocation score more
+carefully, using a decision theoretic procedure for deriving proper scoring rules. (See section \ref{sec:proper} of the
+supplement for a definition of a proper scoring rule and a more technical discussion of the procedure). In section
+\ref{sec:methods.related}, we note that another group of common scores including the quantile score, WIS, and CRPS, can
+also be derived from decision theoretic foundations \textemdash starting from a different decision making context.
 
@@
 
-Suppose that a decision maker is tasked with determining how to allocate $K$ available units of a resource across $N$ locations.
-If the decision maker is provided with a multivariate forecast $F$ where each marginal forecast distribution $F_i$ predicts resource need in a particular location, one option is to choose the resource allocation that minimizes the expected total unmet need according to the forecast.
-We will give a more precise mathematical statement in section \ref{sec:methods.detailed}, but informally, the total expected unmet need according to the forecast is{}
+Suppose that a decision maker is tasked with determining how to allocate $K$ available units of a resource across $N$
+locations. If the decision maker is provided with a multivariate forecast $F$ where each marginal forecast distribution
+$F_i$ predicts resource need in a particular location, one option is to choose the resource allocation that minimizes
+the expected total unmet need according to the forecast. We will give a more precise mathematical statement in section
+\ref{sec:methods.detailed}, but informally, the total expected unmet need according to the forecast is
 \begin{align}
@@
 \end{align}
-where the unmet need in a particular location is the difference between resource need in that location and the number of resources that were allocated there.
-This allocation problem has an intuitively appealing solution: allocate so that the probabilities of need exceeding allocation in various locations are as close to each other as possible.
-This will lead to the allocations provided by $F$ being quantiles of the marginal distributions $F_i$ for some \emph{single} probability level $\tau$ that is shared in common for all locations.
-
-After time passes and the actual level of resource need has been observed, the quality of a selected allocation can be measured by comparing the actual need in each location to the amount of resources that were sent there.
-Specifically, we compute the total unmet need that resulted from the selected allocation:
+where the unmet need in a particular location is the difference between resource need in that location and the number of
+resources that were allocated there. This allocation problem has an intuitively appealing solution: allocate so that the
+probabilities of need exceeding allocation in various locations are as close to each other as possible. This will lead
+to the allocations provided by $F$ being quantiles of the marginal distributions $F_i$ for some \emph{single}
+probability level $\tau$ that is shared in common for all locations.
+
+After time passes and the actual level of resource need has been observed, the quality of a selected allocation can be
+measured by comparing the actual need in each location to the amount of resources that were sent there. Specifically, we
+compute the total unmet need that resulted from the selected allocation:
 \begin{align}
@@
 %We emphasize that in Equation \eqref{eqn:informal_loss} the calculation of unmet need is based on the actual resource need that was realized in each location, while in Equation \eqref{eqn:informal_objective} the calculation of unmet need was based on the forecast distribution of future levels of resource need.
-One allocation is better than another if it results in lower total unmet need, and one forecast is better than another if the allocation derived from it results in lower total unmet need.
-
-The \textbf{allocation score} of the forecast $F$ is the avoidable unmet need that results from using the allocation that minimizes the expected unmet need according to that forecast.
-By ``avoidable unmet need'', we mean that the allocation score does not include the amount of unmet need that was inevitable simply because the amount of available resources $K$ was less than the need for resources.
-Rather, the allocation score measures the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
-An allocation score of 0 is optimal, and indicates that no other allocation of resources could have met need better than the allocation suggested by $F$.
-A larger allocation score indicates that it would have been possible to improve upon the allocation suggested by $F$.
-
-\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$ and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$.
-The means of these distributions are given by the scale parameters $\sigma_i$.
-When the marginal forecasts are exponential distributions, it can be shown that the optimal allocation divides the available resources among the locations proportionally to the scale parameters $\sigma_i$ (see section
-\ref{sec:bayes-quantiles} of the supplemental materials).
-If $K = 5$ units of the resource are available, the optimal allocation according to $F$ would be 1 unit of resources in location 1 and 4 units of resources in location 2.
-If, on the other hand, $K = 10$ units are available, we will allocate 2 units of resources to location 1 and 8 units to location 2.
-Figure~\ref{fig:exp_alloc_example} illustrates the situation.
+One allocation is better than another if it results in lower total unmet need, and one forecast is better than another
+if the allocation derived from it results in lower total unmet need.
+
+The \textbf{allocation score} of the forecast $F$ is the avoidable unmet need that results from using the allocation
+that minimizes the expected unmet need according to that forecast. By ``avoidable unmet need'', we mean that the
+allocation score does not include the amount of unmet need that was inevitable simply because the amount of available
+resources $K$ was less than the need for resources. Rather, the allocation score measures the unmet need that could have
+been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so
+that nothing is wasted in one location while it could be put to use in another. An allocation score of 0 is optimal, and
+indicates that no other allocation of resources could have met need better than the allocation suggested by $F$. A
+larger allocation score indicates that it would have been possible to improve upon the allocation suggested by $F$.
+
+\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$
+and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$. The means of these distributions are
+given by the scale parameters $\sigma_i$. When the marginal forecasts are exponential distributions, it can be shown
+that the optimal allocation divides the available resources among the locations proportionally to the scale parameters
+$\sigma_i$ (see section \ref{sec:bayes-quantiles} of the supplemental materials). If $K = 5$ units of the resource are
+available, the optimal allocation according to $F$ would be 1 unit of resources in location 1 and 4 units of resources
+in location 2. If, on the other hand, $K = 10$ units are available, we will allocate 2 units of resources to location 1
+and 8 units to location 2. Figure~\ref{fig:exp_alloc_example} illustrates the situation.
 
@@
     \includegraphics[width=\textwidth]{../figures/exponential_pred_expected_loss.pdf}
-    \caption{An illustration of the resource allocation problem in Example 1.
-There are $N = 2$ locations, with predictive distributions $F_1 = \mathrm{Exp}(1)$ and $F_2 = \mathrm{Exp}(1/4)$.
-The cumulative distribution functions of these distributions are illustrated in the panels at bottom and right.
-In the center panel, the background shading corresponds to the expected loss according to these forecasts.
-Diagonal black lines indicate resource constraints at $K=5$ and $K=10$ units; any point along those lines corresponds to an allocation that meets the resource constraint.
-For these forecasts, the optimal allocations are $(1, 4)$ for $K=5$ and $(2, 8)$ for $K=10$.
-These allocations are at the point on the constraint line where the expected loss is smallest, which also corresponds to the point where a level set of the expected loss surface (blue curve) is tangent to the constraint.}
+    \caption{An illustration of the resource allocation problem in Example 1. There are $N = 2$ locations, with
+    predictive distributions $F_1 = \mathrm{Exp}(1)$ and $F_2 = \mathrm{Exp}(1/4)$. The cumulative distribution
+    functions of these distributions are illustrated in the panels at bottom and right. In the center panel, the
+    background shading corresponds to the expected loss according to these forecasts. Diagonal black lines indicate
+    resource constraints at $K=5$ and $K=10$ units; any point along those lines corresponds to an allocation that meets
+    the resource constraint. For these forecasts, the optimal allocations are $(1, 4)$ for $K=5$ and $(2, 8)$ for
+    $K=10$. These allocations are at the point on the constraint line where the expected loss is smallest, which also
+    corresponds to the point where a level set of the expected loss surface (blue curve) is tangent to the constraint.}
     \label{fig:exp_alloc_example}
@@
 
-Next suppose that we observe resource needs of 1 and 10 in locations 1 and 2, respectively.
-Based on these observed needs, we can measure the quality of the allocation suggested by the forecast by calculating the amount of unmet need that resulted from that allocation over and above what was unavoidable given the resource constraint.
-With $K = 5$ units of the resource, the allocation based on the forecast exactly meets the observed need in location 1, but it leaves 6 units of need unmet in location 2.
-However, working within the resource constraint, no other allocation could have done better: for example, allocating 0 units of resources to location 1 and 5 to location 2 still results in a total unmet need of 6 across both locations.
-Therefore, the forecast's allocation score is 0 with $K = 5$.
-On the other hand, when $K = 10$, the forecast's allocation results in $10 - 8 = 2$ units of unmet need in location 2 despite leaving no need unmet in location 1.
-In this case, the oracle would be able to prevent all but 1 of the total 11 units of need from going unmet, for example by allocating 1 unit of resources to location 1 and the remaining 9 units of resources to location 2.
-The allocation score for the forecast when $K = 10$ would therefore be 1 (= 2 realized $-$ 1 unavoidable) in units of avoidable unmet need.
-
-These scores illustrate a general result: allocation scores for a forecast will tend to be larger when the resource constraint is close to the observed need, because this is when it matters most which locations are allocated more or less resources.
-If the amount of available resources is small relative to the observed need, any allocation of those limited resources will result in a large amount of unmet need.
-If the amount of available resources is comparatively large, it becomes less important which locations receive relatively more or fewer resources because all locations will receive enough resources to meet their need.
-In either of these extremes of resource availability, the avoidable unmet need that arises from the allocation suggested by a forecast (i.e., the forecast's allocation score) will tend to be small.
-
-\paragraph{Example 2} Now consider a different forecast that also has exponential distributions for resource need in each location, but that has the scale parameters $\sigma_1 = 2$ and $\sigma_2 = 8$, twice as large as the scale parameters of the forecast in Example 1.
-Because the optimal allocation is proportional to the scale parameters, this forecast would lead to the same allocations as the forecast in Example 1, and would therefore be assigned the same allocation score.
-
-Examination of these results leads to two observations.
-First, the reason that these forecasts had a positive (i.e., non-optimal) allocation score at $K=10$ is that they did not get the relative magnitude of resource need across the two locations right: the realized need was 10 times as large in location 2 as in location 1, but the forecasts only indicated that the resource allocation for location 2 should be 4 times the allocation for location 1.
-At its core, the allocation score measures whether the forecast accurately captures the relative magnitudes of resource need across different locations, which is precisely the information that is needed to allocate resources to those locations subject to a fixed resource constraint.
-
-A second observation is that the forecasts in examples 1 and 2 predicted different mean levels of resource needs, but had the same allocation score.
-The allocation score does not directly measure whether the forecasts correctly capture the absolute magnitude of resource need in each individual location.
-This stands in contrast to other common scoring methods that aggregate scores such as log score, CRPS, or WIS for each location, where a poor forecast for one location is penalized regardless of alignments in other units.
+Next suppose that we observe resource needs of 1 and 10 in locations 1 and 2, respectively. Based on these observed
+needs, we can measure the quality of the allocation suggested by the forecast by calculating the amount of unmet need
+that resulted from that allocation over and above what was unavoidable given the resource constraint. With $K = 5$ units
+of the resource, the allocation based on the forecast exactly meets the observed need in location 1, but it leaves 6
+units of need unmet in location 2. However, working within the resource constraint, no other allocation could have done
+better: for example, allocating 0 units of resources to location 1 and 5 to location 2 still results in a total unmet
+need of 6 across both locations. Therefore, the forecast's allocation score is 0 with $K = 5$. On the other hand, when
+$K = 10$, the forecast's allocation results in $10 - 8 = 2$ units of unmet need in location 2 despite leaving no need
+unmet in location 1. In this case, the oracle would be able to prevent all but 1 of the total 11 units of need from
+going unmet, for example by allocating 1 unit of resources to location 1 and the remaining 9 units of resources to
+location 2. The allocation score for the forecast when $K = 10$ would therefore be 1 (= 2 realized $-$ 1 unavoidable) in
+units of avoidable unmet need.
+
+These scores illustrate a general result: allocation scores for a forecast will tend to be larger when the resource
+constraint is close to the observed need, because this is when it matters most which locations are allocated more or
+less resources. If the amount of available resources is small relative to the observed need, any allocation of those
+limited resources will result in a large amount of unmet need. If the amount of available resources is comparatively
+large, it becomes less important which locations receive relatively more or fewer resources because all locations will
+receive enough resources to meet their need. In either of these extremes of resource availability, the avoidable unmet
+need that arises from the allocation suggested by a forecast (i.e., the forecast's allocation score) will tend to be
+small.
+
+\paragraph{Example 2} Now consider a different forecast that also has exponential distributions for resource need in
+each location, but that has the scale parameters $\sigma_1 = 2$ and $\sigma_2 = 8$, twice as large as the scale
+parameters of the forecast in Example 1. Because the optimal allocation is proportional to the scale parameters, this
+forecast would lead to the same allocations as the forecast in Example 1, and would therefore be assigned the same
+allocation score.
+
+Examination of these results leads to two observations. First, the reason that these forecasts had a positive (i.e.,
+non-optimal) allocation score at $K=10$ is that they did not get the relative magnitude of resource need across the two
+locations right: the realized need was 10 times as large in location 2 as in location 1, but the forecasts only
+indicated that the resource allocation for location 2 should be 4 times the allocation for location 1. At its core, the
+allocation score measures whether the forecast accurately captures the relative magnitudes of resource need across
+different locations, which is precisely the information that is needed to allocate resources to those locations subject
+to a fixed resource constraint.
+
+A second observation is that the forecasts in examples 1 and 2 predicted different mean levels of resource needs, but
+had the same allocation score. The allocation score does not directly measure whether the forecasts correctly capture
+the absolute magnitude of resource need in each individual location. This stands in contrast to other common scoring
+methods that aggregate scores such as log score, CRPS, or WIS for each location, where a poor forecast for one location
+is penalized regardless of alignments in other units.
 %Note that we do not claim that the allocation score is generically preferable to these other scores \textemdash rather, it provides a view of forecast performance that is specifically tuned to the context of decision making about resource allocations.

### 2024-02-09 alloscore_manuscript/alloscore-application-shorter.tex
merge elr edits; cut out some commented text

@@
 
-Infectious disease forecasting models have emerged as important tools in public health outbreak response.
-The predictions they provide increasingly inform decisions regarding a wide variety of countermeasures intended to reduce transmission and mitigate the severity of disease outcomes.
-For example, estimates of the onset time of the flu season have been used in developing national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola and diptheria dynamics have been made with the clearly stated goal of helping local
-public health workers choose the timing and location of interventions in settings where resources are severely constrained \citep{meltzer2014estimating, rainisch2015regional,camacho2015-ebola-bed,finger_real-time_2019}.
-More recently, in the context of the outbreak of COVID-19 across the US, \cite{bertsimas2021predictionsCOVID} explicitly based triage decision tools for the interstate reallocation
-of ventilators and ICU capacity on forecast inputs and used these forecasts as well to recommend vaccine trial sites to a 
-major trial sponsor.
-\cite{fox_real-time_2022} similarly used predictive models to inform intrastate resource and care site planning, as well as local community guidelines for masking, traveling, dining and shopping (University of Texas, 2022)\nocite{utnews2022}.
-%In April of 2022, the Centers for Disease Control and Prevention (CDC) announced the launch of the Center for Forecasting and Outbreak Analytics (CFA) to translate disease forecasts into decision making (CDC, 2022)\nocite{cdc2022cfa}, indicating that this has been identified as an important direction at the highest levels of government public health response.
+Infectious disease forecasting models have emerged as important tools in public health outbreak response. The
+predictions they provide increasingly inform decisions regarding a wide variety of countermeasures intended to reduce
+transmission and mitigate the severity of disease outcomes. For example, estimates of the onset time of the flu season
+have been used in developing national vaccination strategies \citep{igboh2023timing}, and forecasts of Ebola and
+diptheria dynamics have been made with the clearly stated goal of helping local public health workers choose the timing
+and location of interventions in settings where resources are severely constrained 
+\citep{meltzer2014estimating, rainisch2015regional, camacho2015-ebola-bed,finger_real-time_2019}. More recently, in the
+context of the outbreak of COVID-19 across the US, \cite{bertsimas2021predictionsCOVID} used forecasts as inputs to
+decision tools for the interstate reallocation of ventilators and ICU capacity, and to recommend vaccine trial sites to
+a major trial sponsor. \cite{fox_real-time_2022} similarly used predictive models to inform intrastate resource and
+care site planning, as well as local community guidelines for masking, traveling, dining and shopping (University of
+Texas, 2022)\nocite{utnews2022}.
 
-In the wake of the COVID-19 pandemic, this trend has been followed by calls for infectious disease forecasts to 
-be not only designed, but 
-also evaluated in ways that align specifically with how forecasts can be used to inform such outbreak control
-decisions \citep{marshall2023predictions, bilinski_adaptive_2023}.
-This contrasts, however, with the historically standard practice of measuring the quality of disease forecasts using general purpose
-accuracy and skill scores, especially those that have implementations available in existing software when the relevant outbreak occurs. For point forecasts,
-the root mean square error (RMSE) (e.g., \cite{papastefanopoulos2020covid}) and the mean absolute error (MAE) (e.g., \cite{johansson2016evaluating}) 
-are common choices.
-For probabilistic forecasts, (which are the focus of this paper), researchers have often relied on the logarithmic score (LS), which
-has been used, for example, to evaluate (with respect to both time and space) the skill of US seasonal influenza forecasts \citep{mcgowan_collaborative_2019,reich_collaborative_2019} as well as forecasts targeting surveillance measures of dengue incidence in Peru and Puerto Rico
- \citep{johansson_open_2019}.
-More recently, the continuous ranked probability score (CRPS) and a discretized version adapted to multi-quantile forecasts, the weighted interval 
-score (WIS) \citep{bracher2021evaluating} (named for its expression in terms of central prediction interval coverage), have gained prominence.
-For example, the CRPS was used to assess probabilistic forecasts (based on random effect models) of dengue incidence at the district level 
-in Vietnam \citep{colon-gonzalez_probabilistic_2021}. And the WIS has been used during the COVID-19 pandemic (through efforts with which we 
-have been intimately involved) to evaluate forecasts of 
-observed cases, hospitalizations and deaths in the US and Europe, as reported by municipal, state, and federal surveillance systems 
-\citep{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.
+In the wake of the COVID-19 pandemic, this trend has been followed by calls for infectious disease forecasts to be not
+only designed, but also evaluated in ways that align specifically with how forecasts can be used to inform such outbreak
+control decisions \citep{marshall2023predictions, bilinski_adaptive_2023}. This contrasts, however, with the
+historically standard practice of measuring the quality of disease forecasts using general purpose accuracy and skill
+scores, especially those that have implementations available in existing software when the relevant outbreak occurs. For
+point forecasts, the root mean square error (RMSE) (e.g., \cite{papastefanopoulos2020covid}) and the mean absolute error
+(MAE) (e.g., \cite{johansson2016evaluating}) are common choices. For probabilistic forecasts, which are the focus of
+this paper, researchers have often relied on the logarithmic score (LS). For example, the LS has been used to evaluate
+the skill of US seasonal influenza forecasts \citep{mcgowan_collaborative_2019,reich_collaborative_2019} as well as
+forecasts targeting surveillance measures of dengue incidence in Peru and Puerto Rico \citep{johansson_open_2019}. More
+recently, the continuous ranked probability score (CRPS) and a discretized version adapted to multi-quantile forecasts,
+the weighted interval score (WIS) \citep{bracher2021evaluating}, have gained prominence. For example, the CRPS was used
+to assess probabilistic forecasts (based on random effect models) of dengue incidence at the district level in Vietnam
+\citep{colon-gonzalez_probabilistic_2021}. And the WIS has been used during the COVID-19 pandemic to evaluate forecasts
+of observed cases, hospitalizations and deaths in the US and Europe, as reported by municipal, state, and federal
+surveillance systems \citep{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.
 
-While it should be noted that there are ways to interpret any of these scores abstractly through the lens of decision theory, and that all of 
-the application-specific papers cited above benefited from direct collaboration with public health agencies, a key impetus for the
-present work has been that we were not able to find in any of them, nor in the literature they represent, explicit connections between how a 
-forecast was evaluated and how that forecast was used in practice.
+While it should be noted that there are ways to interpret any of these scores abstractly through the lens of decision
+theory, and that all of the application-specific papers cited above benefited from direct collaboration with public
+health agencies, a key impetus for the present work has been that we were not able to find in any of them, nor in the
+literature they represent, explicit connections between how a forecast was evaluated and how that forecast was used in
+practice.
 
-A general phenomenon at play here --- one that has been observed repeatedly over the past few decades in other fields such as finance, supply chain management, and meteorology --- is that while scores such as RMSE, MAE, LS, CRPS, and WIS can describe the \emph{quality} of a forecast
-at corresponding to the observed disease outcome, they will often fail to register the \emph{value} (or lack thereof) 
-a forecast has in the context of a specific decision. We refer the reader to \cite{yardley2021utility_cost_forecasts} and the references collected 
-therein (especially the foundational \cite{murphy1993whatisagoodforecast}) for a general overview touching on a wide range of forecasting contexts and  
-also to \cite{pesaran2002decision_based_eval} for a clear discussion from an econometric perspective.
+A general phenomenon at play here --- one that has been observed repeatedly over the past few decades in other fields
+such as finance, supply chain management, and meteorology --- is that while scores such as RMSE, MAE, LS, CRPS, and WIS
+can describe the \emph{quality} of a forecast in terms of how well it corresponds to the observed disease outcome, they
+will often fail to register the \emph{value} of a forecast in the context of a specific decision. We refer the reader to
+\cite{yardley2021utility_cost_forecasts} and the references collected therein (especially the foundational
+\cite{murphy1993whatisagoodforecast}) for a general overview touching on a wide range of forecasting contexts and  also
+to \cite{pesaran2002decision_based_eval} for a clear discussion from an econometric perspective.
 
-%------ literature that seems to be aware of the problem we see and how it falls short ------%
-This is to say that the quality-value distinction is not a new concept in the larger forecasting community. We are aware, however, of only a 
-limited literature attempting to connect the value of infectious disease forecasts to their impact on and through policy.  And
-within this body of work we have found the discussion of such a connection to still be at a formally and quantitively imprecise stage. 
-In \cite{ioannidis2022forecastingCOVIDfailed}, the possible negative consequences of inaccurate forecasts of infectious disease are discussed, but there is no attempt to quantify the utility or loss incurred as a result of those forecasts.
-\cite{bilinski_adaptive_2023} explore ways in which predictive classifiers of local COVID-19 risk levels in the US could be tuned to policymaker preferences for different
-costs associated with over- and under-reaction to disease dynamics, but they do not clearly identify the source of these costs or how they
-depend on quantifiable policy choices. A similar discussion related to dengue countermeasures in Vietnam appears in
-\cite{colon-gonzalez_probabilistic_2021}.  A novel version of the WIS informally motivated by utility considerations is devoloped in 
-\cite{marshall2023predictions}, but the score is not derived in a decision-theoretic manner.
-There is also a thread of literature that frames infectious disease modeling as a component of a larger system for understanding how policy goals, means, and choices interact and constrain
-one another. As an example, \cite{Probert2016decisionMakingFootMouth} explore how policy recommendations ought to flow from a possibly 
-incongruous set of simulation-based projection models of a hypothetical
-foot-and-mouth disease outbreak when there are ranges of plausible responses and stakeholder interests. Decision theory plays a prominent
-role here but not explicitly as a way to evaluate the choices made in developing or running the models.
+Despite this now well-developed discussion of the quality-value distinction in the larger forecasting community, we are
+aware of only a limited literature attempting to connect the value of \emph{infectious disease} forecasts to their
+impact on and through policy.  And within this body of work we have found the discussion of such a connection to still
+be at a formally and quantitively imprecise stage. In \cite{ioannidis2022forecastingCOVIDfailed}, the possible negative
+consequences of inaccurate forecasts of infectious disease are discussed, but there is no attempt to quantify the
+utility or loss incurred as a result of those forecasts. \cite{bilinski_adaptive_2023} explore ways in which predictive
+classifiers of local COVID-19 risk levels in the US could be tuned to policymaker preferences for different costs
+associated with over- and under-reaction to disease dynamics, but they do not clearly identify the source of these
+costs or how they depend on quantifiable policy choices. A similar discussion related to dengue countermeasures in
+Vietnam appears in \cite{colon-gonzalez_probabilistic_2021}.  A novel version of the WIS informally motivated by
+utility considerations is developed in \cite{marshall2023predictions}, but the score is not derived in a
+decision-theoretic manner.  There is also a thread of literature that frames infectious disease modeling as a component
+of a larger system for understanding how policy goals, means, and choices interact and constrain one another. As an
+example, \cite{Probert2016decisionMakingFootMouth} explore how policy recommendations ought to flow from a possibly 
+incongruous set of simulation-based projection models of a hypothetical foot-and-mouth disease outbreak when there are 
+ranges of plausible responses and stakeholder interests. Decision theory plays a prominent role here, but not explicitly
+as a way to evaluate the choices made in developing the models.
 
-% However, as we show in this paper, it is possible, using standard tools of decision theory, to construct formal methods (detailed in Section 
-% \ref{sec:methods.detailed.decisiontheory})) for evaluating probabilistic forecasts that are directly and uniquely tied to policy decisions, 
-% in particular those related to resource allocation.  In essence, we will define the score (as a measure of value) of a forecast as  
-% Furthermore, we find evidence that such scores can illuminate important features of forecasts related to their value as decision-making inputs which
-% are missed by traditional quality (that is, accuracy and skill) measures.
-
-% %------ what you can do instead ------%
-% In decision making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to (see details in Section \ref{sec:methods.detailed.decisiontheory}).
-
-% %In finance, the value of forecasts can often be measured by the profits generated by trading decisions informed by the forecasts, perhaps adjusted for risk levels \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}.
-In applications to supply chain management and meteorology, the value of forecasts has typically been operationalized by considering the costs associated with decisions regarding the amount of inventory to hold or the level of protection against the impacts of extreme weather events to enact \cite[e.g.,][]{catt2007assessingcostofforecasterror, petropoulos2019inventoryperformanceforecasting, palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
-
-% In this framework, a forecast has value if it leads to decisions with low total costs.
-% As one example,
-% % in supply chain management these decisions may incur costs related to holding inventory, labor, or providing poor service, while
-% in meteorology, policy makers need to balance the costs of implementing protective measures with the costs of potentially preventable weather damages \citep{palmer2002economic, pappenberger2015monetarybenefitfloodwarnings}.
-
-%------ our game plan ------%
-In this work, we begin to fill this gap between the ways that infectious disease forecasts have traditionally been evaluated and the ways that they have been used to support public health policy.
-To do so, we consider a setting in which forecasts are used to help determine the allocation of a limited quantity of medical supplies across multiple regions.
-In section~\ref{sec:methods} of the paper, we define a new forecast scoring rule --- the {\em allocation score} --- that evaluates forecasts based on how beneficial resource allocations derived from them would turn out to be.
-%Unlike the traditional measures of forecast skill such as WIS, CRPS, and log score, the proposed allocation score measures the suitability of a forecast as an input to this specific decision making task.
-In section \ref{sec:application}, we present an illustrative analysis using the allocation score to evaluate forecasts of hospital admissions in the US that were made leading up to and during the Omicron wave that peaked in January of 2022.
-This analysis is ``synthetic'' insofar as it is not intended to correspond to any specific historical record of allocation decisions 
-that could have been supported by hospitalization forecasts during this period.
-However, we view the general allocation problem on which our framework is based as a versatile template for 
-formalizing real-world decisions that must constantly be made in real-time by public health administrators around the globe, especially those
-related to hospital capacity, ventilator usage, doses needed for specific medications and other situations where an outbreak creates sudden and
-highly variable demand for potentially scarce resources.
-
-%The remainder of this article is organized as follows.
-%We describe the allocation score in Section \ref{sec:methods}, and in Section \ref{sec:application} we illustrate the use of the score in an application to evaluate short-term forecasts of COVID-19 hospital admissions in the US.
-%Section \ref{sec:discussion} summarizes our contributions and discusses opportunities for further extensions in future work.
+In this work, we begin to fill this gap between the ways that infectious disease forecasts have traditionally been
+evaluated and the ways that they have been used to support public health policy. To do so, we consider a setting in
+which forecasts are used to help determine the allocation of a limited quantity of medical supplies across multiple
+regions. In section~\ref{sec:methods} of the paper, we define a new forecast scoring rule --- the {\em allocation score}
+--- that evaluates forecasts based on how beneficial resource allocations derived from them would turn out to be. In
+section \ref{sec:application}, we present an illustrative analysis using the allocation score to evaluate forecasts of
+hospital admissions in the US that were made leading up to and during the Omicron wave that peaked in January of 2022.
+This analysis is ``synthetic'' insofar as it is not intended to correspond to any specific historical record of
+allocation decisions that could have been supported by hospitalization forecasts during this period. However, we view
+the general allocation problem on which our framework is based as a versatile template for formalizing real-world
+decisions that must constantly be made in real-time by public health administrators around the globe, especially those
+related to hospital capacity, ventilator usage, doses needed for specific medications and other situations where an
+outbreak creates sudden and highly variable demand for potentially scarce resources.
 
@@
 \end{align}
-where the unmet need in a particular location is the difference between resource need in that location and the number of resources that were allocated there.
-This allocation problem has an intuitively appealing solution: allocate so that the probabilities of need exceeding allocation in various locations are as close to each other as possible.
-This will lead to the allocations provided by $F$ being quantiles of the marginal distributions $F_i$ for some \emph{single} probability level $\tau$ that is shared in common for all locations.
+where the unmet need in a particular location is the difference between resource need in that location and the number of
+resources that were allocated there. This allocation problem has an intuitively appealing solution: allocate so that the
+probabilities of need exceeding allocation in various locations are as close to each other as possible. This will lead
+to the allocations provided by $F$ being quantiles of the marginal distributions $F_i$ for some \emph{single}
+probability level $\tau$ that is shared in common for all locations.
 
-After time passes and the actual level of resource need has been observed, the quality of a selected allocation can be measured by comparing the actual need in each location to the amount of resources that were sent there.
-Specifically, we compute the total unmet need that resulted from the selected allocation:
+After time passes and the actual level of resource need has been observed, the quality of a selected allocation can be
+measured by comparing the actual need in each location to the amount of resources that were sent there. Specifically, we
+compute the total unmet need that resulted from the selected allocation:
 \begin{align}
@@
 %We emphasize that in Equation \eqref{eqn:informal_loss} the calculation of unmet need is based on the actual resource need that was realized in each location, while in Equation \eqref{eqn:informal_objective} the calculation of unmet need was based on the forecast distribution of future levels of resource need.
-One allocation is better than another if it results in lower total unmet need, and one forecast is better than another if the allocation derived from it results in lower total unmet need.
+One allocation is better than another if it results in lower total unmet need, and one forecast is better than another
+if the allocation derived from it results in lower total unmet need.
 
-The \textbf{allocation score} of the forecast $F$ is the avoidable unmet need that results from using the allocation that minimizes the expected unmet need according to that forecast.
-By ``avoidable unmet need'', we mean that the allocation score does not include the amount of unmet need that was inevitable simply because the amount of available resources $K$ was less than the need for resources.
-Rather, the allocation score measures the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
-An allocation score of 0 is optimal, and indicates that no other allocation of resources could have met need better than the allocation suggested by $F$.
-A larger allocation score indicates that it would have been possible to improve upon the allocation suggested by $F$.
+The \textbf{allocation score} of the forecast $F$ is the avoidable unmet need that results from using the allocation
+that minimizes the expected unmet need according to that forecast. By ``avoidable unmet need'', we mean that the
+allocation score does not include the amount of unmet need that was inevitable simply because the amount of available
+resources $K$ was less than the need for resources. Rather, the allocation score measures the unmet need that could have
+been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so
+that nothing is wasted in one location while it could be put to use in another. An allocation score of 0 is optimal, and
+indicates that no other allocation of resources could have met need better than the allocation suggested by $F$. A
+larger allocation score indicates that it would have been possible to improve upon the allocation suggested by $F$.
 
-\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$ and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$.
-The means of these distributions are given by the scale parameters $\sigma_i$.
-When the marginal forecasts are exponential distributions, it can be shown that the optimal allocation divides the available resources among the locations proportionally to the scale parameters $\sigma_i$ (see section
-\ref{sec:bayes-quantiles} of the supplemental materials).
-If $K = 5$ units of the resource are available, the optimal allocation according to $F$ would be 1 unit of resources in location 1 and 4 units of resources in location 2.
-If, on the other hand, $K = 10$ units are available, we will allocate 2 units of resources to location 1 and 8 units to location 2.
-Figure~\ref{fig:exp_alloc_example} illustrates the situation.
+\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$
+and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$. The means of these distributions are
+given by the scale parameters $\sigma_i$. When the marginal forecasts are exponential distributions, it can be shown
+that the optimal allocation divides the available resources among the locations proportionally to the scale parameters
+$\sigma_i$ (see section \ref{sec:bayes-quantiles} of the supplemental materials). If $K = 5$ units of the resource are
+available, the optimal allocation according to $F$ would be 1 unit of resources in location 1 and 4 units of resources
+in location 2. If, on the other hand, $K = 10$ units are available, we will allocate 2 units of resources to location 1
+and 8 units to location 2. Figure~\ref{fig:exp_alloc_example} illustrates the situation.
 
@@
 
-Next suppose that we observe resource needs of 1 and 10 in locations 1 and 2, respectively.
-Based on these observed needs, we can measure the quality of the allocation suggested by the forecast by calculating the amount of unmet need that resulted from that allocation over and above what was unavoidable given the resource constraint.
-With $K = 5$ units of the resource, the allocation based on the forecast exactly meets the observed need in location 1, but it leaves 6 units of need unmet in location 2.
-However, working within the resource constraint, no other allocation could have done better: for example, allocating 0 units of resources to location 1 and 5 to location 2 still results in a total unmet need of 6 across both locations.
-Therefore, the forecast's allocation score is 0 with $K = 5$.
-On the other hand, when $K = 10$, the forecast's allocation results in $10 - 8 = 2$ units of unmet need in location 2 despite leaving no need unmet in location 1.
-In this case, the oracle would be able to prevent all but 1 of the total 11 units of need from going unmet, for example by allocating 1 unit of resources to location 1 and the remaining 9 units of resources to location 2.
-The allocation score for the forecast when $K = 10$ would therefore be 1 (= 2 realized $-$ 1 unavoidable) in units of avoidable unmet need.
+Next suppose that we observe resource needs of 1 and 10 in locations 1 and 2, respectively. Based on these observed
+needs, we can measure the quality of the allocation suggested by the forecast by calculating the amount of unmet need
+that resulted from that allocation over and above what was unavoidable given the resource constraint. With $K = 5$ units
+of the resource, the allocation based on the forecast exactly meets the observed need in location 1, but it leaves 6
+units of need unmet in location 2. However, working within the resource constraint, no other allocation could have done
+better: for example, allocating 0 units of resources to location 1 and 5 to location 2 still results in a total unmet
+need of 6 across both locations. Therefore, the forecast's allocation score is 0 with $K = 5$. On the other hand, when
+$K = 10$, the forecast's allocation results in $10 - 8 = 2$ units of unmet need in location 2 despite leaving no need
+unmet in location 1. In this case, the oracle would be able to prevent all but 1 of the total 11 units of need from
+going unmet, for example by allocating 1 unit of resources to location 1 and the remaining 9 units of resources to
+location 2. The allocation score for the forecast when $K = 10$ would therefore be 1 (= 2 realized $-$ 1 unavoidable) in
+units of avoidable unmet need.
 
-These scores illustrate a general result: allocation scores for a forecast will tend to be larger when the resource constraint is close to the observed need, because this is when it matters most which locations are allocated more or less resources.
-If the amount of available resources is small relative to the observed need, any allocation of those limited resources will result in a large amount of unmet need.
-If the amount of available resources is comparatively large, it becomes less important which locations receive relatively more or fewer resources because all locations will receive enough resources to meet their need.
-In either of these extremes of resource availability, the avoidable unmet need that arises from the allocation suggested by a forecast (i.e., the forecast's allocation score) will tend to be small.
+These scores illustrate a general result: allocation scores for a forecast will tend to be larger when the resource
+constraint is close to the observed need, because this is when it matters most which locations are allocated more or
+less resources. If the amount of available resources is small relative to the observed need, any allocation of those
+limited resources will result in a large amount of unmet need. If the amount of available resources is comparatively
+large, it becomes less important which locations receive relatively more or fewer resources because all locations will
+receive enough resources to meet their need. In either of these extremes of resource availability, the avoidable unmet
+need that arises from the allocation suggested by a forecast (i.e., the forecast's allocation score) will tend to be
+small.
 
-\paragraph{Example 2} Now consider a different forecast that also has exponential distributions for resource need in each location, but that has the scale parameters $\sigma_1 = 2$ and $\sigma_2 = 8$, twice as large as the scale parameters of the forecast in Example 1.
-Because the optimal allocation is proportional to the scale parameters, this forecast would lead to the same allocations as the forecast in Example 1, and would therefore be assigned the same allocation score.
+\paragraph{Example 2} Now consider a different forecast that also has exponential distributions for resource need in
+each location, but that has the scale parameters $\sigma_1 = 2$ and $\sigma_2 = 8$, twice as large as the scale
+parameters of the forecast in Example 1. Because the optimal allocation is proportional to the scale parameters, this
+forecast would lead to the same allocations as the forecast in Example 1, and would therefore be assigned the same
+allocation score.
 
-Examination of these results leads to two observations.
-First, the reason that these forecasts had a positive (i.e., non-optimal) allocation score at $K=10$ is that they did not get the relative magnitude of resource need across the two locations right: the realized need was 10 times as large in location 2 as in location 1, but the forecasts only indicated that the resource allocation for location 2 should be 4 times the allocation for location 1.
-At its core, the allocation score measures whether the forecast accurately captures the relative magnitudes of resource need across different locations, which is precisely the information that is needed to allocate resources to those locations subject to a fixed resource constraint.
+Examination of these results leads to two observations. First, the reason that these forecasts had a positive (i.e.,
+non-optimal) allocation score at $K=10$ is that they did not get the relative magnitude of resource need across the two
+locations right: the realized need was 10 times as large in location 2 as in location 1, but the forecasts only
+indicated that the resource allocation for location 2 should be 4 times the allocation for location 1. At its core, the
+allocation score measures whether the forecast accurately captures the relative magnitudes of resource need across
+different locations, which is precisely the information that is needed to allocate resources to those locations subject
+to a fixed resource constraint.
 
-A second observation is that the forecasts in examples 1 and 2 predicted different mean levels of resource needs, but had the same allocation score.
-The allocation score does not directly measure whether the forecasts correctly capture the absolute magnitude of resource need in each individual location.
-This stands in contrast to other common scoring methods that aggregate scores such as log score, CRPS, or WIS for each location, where a poor forecast for one location is penalized regardless of alignments in other units.
+A second observation is that the forecasts in examples 1 and 2 predicted different mean levels of resource needs, but
+had the same allocation score. The allocation score does not directly measure whether the forecasts correctly capture
+the absolute magnitude of resource need in each individual location. This stands in contrast to other common scoring
+methods that aggregate scores such as log score, CRPS, or WIS for each location, where a poor forecast for one location
+is penalized regardless of alignments in other units.
 %Note that we do not claim that the allocation score is generically preferable to these other scores \textemdash rather, it provides a view of forecast performance that is specifically tuned to the context of decision making about resource allocations.
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Thu Feb  8 21:06:12 2024
+% Fri Feb  9 13:33:13 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Thu Feb  8 21:06:16 2024
+% Fri Feb  9 13:33:17 2024
 \begin{table}[ht]

### 2024-02-14 alloscore_manuscript/alloscore-application-shorter.Rnw
various edits, rephrasing

@@
 where the unmet need in a particular location is the difference between resource need in that location and the number of
-resources that were allocated there. This allocation problem has an intuitively appealing solution: allocate so that the
+resource units that were allocated there. This allocation problem has an intuitively appealing solution: allocate so that the
 probabilities of need exceeding allocation in various locations are as close to each other as possible. This will lead
@@
 
-After time passes and the actual level of resource need has been observed, the quality of a selected allocation can be
+After time passes and the actual level of resource need has been observed, the value of a selected allocation can be
 measured by comparing the actual need in each location to the amount of resources that were sent there. Specifically, we
@@
 %We emphasize that in Equation \eqref{eqn:informal_loss} the calculation of unmet need is based on the actual resource need that was realized in each location, while in Equation \eqref{eqn:informal_objective} the calculation of unmet need was based on the forecast distribution of future levels of resource need.
-One allocation is better than another if it results in lower total unmet need, and one forecast is better than another
-if the allocation derived from it results in lower total unmet need.
+We regard one allocation as better than another (with respect to the realized need) if it results in lower total 
+unmet need, and accordingly regard one forecast as better than another (again with respect to the realized need)
+if the allocation derived from the former results in lower total unmet need than does the allocation derived from the 
+latter.
 
@@
 been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so
-that nothing is wasted in one location while it could be put to use in another. An allocation score of 0 is optimal, and
-indicates that no other allocation of resources could have met need better than the allocation suggested by $F$. A
-larger allocation score indicates that it would have been possible to improve upon the allocation suggested by $F$.
+that nothing is wasted in one location while it could be put to use in another. The best and lowest possible allocation 
+score is 0.  A positive allocation score indicates some way in which an oracle could have used perfect foreknowledge to
+meet more of the need than did the allocation suggested by $F$.
 
@@
 $\sigma_i$ (see section \ref{sec:bayes-quantiles} of the supplemental materials). If $K = 5$ units of the resource are
-available, the optimal allocation according to $F$ would be 1 unit of resources in location 1 and 4 units of resources
-in location 2. If, on the other hand, $K = 10$ units are available, we will allocate 2 units of resources to location 1
-and 8 units to location 2. Figure~\ref{fig:exp_alloc_example} illustrates the situation.
+available, the optimal allocation according to $F$ would be 1 and 4 resource units in locations 1 and 2, respectively. 
+Increasing $K$ to $10$ changes this optimal allocation to 2 and 8. Figure~\ref{fig:exp_alloc_example} illustrates the 
+situation.
 
@@
 Next suppose that we observe resource needs of 1 and 10 in locations 1 and 2, respectively. Based on these observed
-needs, we can measure the quality of the allocation suggested by the forecast by calculating the amount of unmet need
-that resulted from that allocation over and above what was unavoidable given the resource constraint. With $K = 5$ units
-of the resource, the allocation based on the forecast exactly meets the observed need in location 1, but it leaves 6
-units of need unmet in location 2. However, working within the resource constraint, no other allocation could have done
-better: for example, allocating 0 units of resources to location 1 and 5 to location 2 still results in a total unmet
-need of 6 across both locations. Therefore, the forecast's allocation score is 0 with $K = 5$. On the other hand, when
+needs, we evaluate the allocation suggested by the forecast by calculating the amount of unmet need
+that resulted from that allocation over and above what was unavoidable given the resource constraint. With $K = 5$ 
+available resource units, the allocation based on the forecast exactly meets the observed need in location 1, but it leaves 6
+units of need unmet in location 2.
+But these 6 units of unmet need are unaviodable under the constraint $K=5$: allocating 0 units of resources to location 1 and 5 to location 2, for example, still results in a total unmet
+need of 6 across both locations. This gives an allocation score of 0 for the forecast when $K = 5$. On the other hand, when
 $K = 10$, the forecast's allocation results in $10 - 8 = 2$ units of unmet need in location 2 despite leaving no need
@@
 going unmet, for example by allocating 1 unit of resources to location 1 and the remaining 9 units of resources to
-location 2. The allocation score for the forecast when $K = 10$ would therefore be 1 (= 2 realized $-$ 1 unavoidable) in
+location 2. So for $K = 10$, the forecast gets an allocaton score of 1 (= 2 realized $-$ 1 unavoidable) in
 units of avoidable unmet need.
@@
 limited resources will result in a large amount of unmet need. If the amount of available resources is comparatively
-large, it becomes less important which locations receive relatively more or fewer resources because all locations will
+large, it becomes less important which locations receive relatively more or fewer resources because all locations are
+likely to
 receive enough resources to meet their need. In either of these extremes of resource availability, the avoidable unmet
@@
 
-We give a high-level review of a general procedure for developing proper scoring rules that are tailored to specific decision making tasks in section \ref{sec:methods.detailed.decisiontheory}, and then in section \ref{sec:methods.detailed.specific_allocation} we apply that procedure to develop the allocation score based on the task of deciding on how to allocate a fixed supply of resources across multiple locations.
-In \ref{sec:methods.detailed.integrated_allocation} we consider a small extension where the resource constraint is not known, or it is desired to consider the value of forecasts across a range of decision making scenarios.
-This gives rise to the \emph{integrated allocation score}.
+We give a high-level review of a general procedure for developing proper scoring rules that are tailored to specific
+decision making tasks in section \ref{sec:methods.detailed.decisiontheory}, and then in section
+\ref{sec:methods.detailed.specific_allocation} we apply that procedure to develop the allocation score based on the task
+of deciding on how to allocate a fixed supply of resources across multiple locations. In
+\ref{sec:methods.detailed.integrated_allocation} we consider a small extension where the resource constraint is not
+known, or it is desired to consider the value of forecasts across a range of decision making scenarios. This gives rise
+to the \emph{integrated allocation score}.
 
@@
 
-%In this section, we give an overview of the decision theoretic setup for developing proper scoring rules that measure the value of a forecast as an input to decision making.
-We keep the discussion here at a somewhat informal level; we refer the reader to [some subset of Brehmer and Gneiting; Grünwald and Dawid; Dawid; Granger and Pesaran 2000; Granger and Machina 2006; Ehm et al.
-2016] for more technically precise discussion.
-
-In the framework of decision theory, a decision corresponds to the selection of an action $x$ from some set of possible actions $\mathcal{X}$.
-For example, $x$ may correspond to the level of investment in a measure designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with $\mathcal{X}$ being the set of all possible levels of investment that we might select.
-The quality of a decision to take a particular action $x$ is measured in relation to an outcome $y$ that is unknown at the time the decision is made.
-For example, $y$ may correspond to the number of individuals who eventually become sick and would benefit from the mitigation measure, and informally, an action $x$ is successful to the extent that it meets the realized need.
-In the face of uncertainty, a decision maker may use a forecast $F$ of the random variable $Y$ to help inform the selection of the action to take.
-We measure the value of a forecast as an input to this decision making process by the quality of the decisions that it leads to.
-
-We can formalize the preceding discussion with the following three-step procedure for developing scoring rules for probabilistic forecasts:
+%In this section, we give an overview of the decision theoretic setup for developing proper scoring rules that measure
+%the value of a forecast as an input to decision making. We keep the discussion here at a somewhat informal level; we
+%refer the reader to [some subset of Brehmer and Gneiting; Grünwald and Dawid; Dawid; Granger and Pesaran 2000; Granger
+%and Machina 2006; Ehm et al. 2016] for more technically precise discussion.
+
+Decision theory is concerned with understanding how an \emph{action} $x$ both is and ought to be selected from a set of
+possible actions $\mathcal{X}$ --- that is, how a decision is and ought to be made --- based on the anticipated
+consequences of $x$ under future states. We represent such future states as a random variable $Y$ and refer to a
+realization $y$ of $Y$ as an \emph{outcome}. In this paper $x$ will be related to the level of investment in a measure
+designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with
+$\mathcal{X}$ being the set of all feasible levels of investment, while an outcome $y$ will correspond to the number of
+individuals who eventually become sick and would benefit from the mitigation measure. A basic strategy in decision
+theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numerical value, or
+\emph{loss}, measuring the success of $x$, which in our examples will be the extent to which $x$ meets the realized need
+$y$.  One way that a decision maker can try to anticipate the consequence of an action $x$ is by using a forecast $F$ of
+$Y$. The loss of the action $x$ chosen according to $F$ can, in turn, be intepreted as the value of $F$ as an input to
+the decision making process under the realized outcome $y$ of $Y$.
+
+We arrive through this discussion at a three-step procedure for developing scoring rules for probabilistic forecasts:
 \begin{enumerate}
-\item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$ eventually occurs.
-\item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under the distribution $F$.
-\item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: $S(F, y) = s(x^F, y)$.
+  \item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$
+    eventually occurs.
+  \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under
+    the distribution $F$.
+  \item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: 
+    $S(F, y) = s(x^F, y)$.
 \end{enumerate}
-This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss function.
-We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} and in section \ref{sec:proper} of the
-supplement we demonstrate that Bayes scoring rules are proper by construction (a result also shown by \cite{dawid2007geometry,gneiting2007strictly}).
+This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss
+function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} and in section
+\ref{sec:proper} of the supplement we review the basic fact from the scoring rule literature (e.g.,
+\cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring rules are proper by construction.
 
@@
 
-In the decision making setting that we consider, an action $x = (x_1, \ldots, x_N)$ is a vector specifying the amount that is allocated to each of $N$ locations.
-We require that $0 \leq x$, i.e., that each $x_i$ is non-negative, and that the total allocation across all locations equals the amount of available resources, $K$: $\sum_{i=1}^N x_i = K$.
-The set $\mathcal{X}$ consists of all possible allocations that satisfy these constraints.
-The eventually realized resource need in each location is denoted by $y = (y_1, \ldots, y_N)$.
-These levels of need are not known at the time of decision making, so we define the random vector $Y = (Y_1, \ldots, Y_N)$ where $Y_i$ represents the as-yet-unknown level of resource need in location $i$.
-Forecasts of need in each location are collected in $F = (F_1, \ldots, F_N)$.
-We assume that resource need is non-negative and the forecasts reflect that, i.e.
-the support of each $F_i$ is a subset of $\mathbb{R}^+$.
-Finally, we assume that each unit of unmet need incurs a loss denoted by $L$, so that if the selected resource level $x_i$ in location $i$ is less than the realized need $y_i$, a loss of $L \cdot (y_i - x_i)$ results.
-A variety of extensions to this setup are possible; for example, we might account for storage costs for resources that go unused, allow for a different loss per unit of unmet need in each location, or account for resource transportation costs.
-In this work, we choose to keep the loss function relatively simple to focus on the core ideas.
-
-It is helpful to clearly distinguish between the time $t_d$ when a \emph{d}ecision is made about a public health resource allocation and the time $t_r$ when \emph{r}esource needs that might be addressed by that allocation occur.
-Our setup assumes that $t_d < t_r$.
-Additionally, the structure of our loss captures a setting where the resource in question does not impact the amount of demand that will materialize at time $t_r$, but rather it is a resource that satisfies that demand.
-In the context of infectious disease, this means that we do not consider resources that are intended to reduce the number of people who will become sick at some point in the future, such as a preventative influenza or COVID-19 vaccine.
-Instead, our setup addresses resources like hospital beds, oxygen supply, or ventilators which are intended to meet the medical needs of patients who are already sick.
-Additionally, our setup addresses decision-making that is related to resource needs only at the time $t_r$; we do not explicitly consider sequences of multiple decisions that are made over time or account for the impact of decisions on resource needs at any time other than $t_r$.
-We outline some opportunities to extend our work to more complex decision making settings in the discussion.
-
-With this problem formulation in place, we can develop a proper scoring rule following the outline in section \ref{sec:methods.detailed.decisiontheory}.
-
-\paragraph{Step 1: specify a loss function.} The loss associated with a particular allocation is calculated by summing contributions from unmet need in each location:
+In our setting an action is a vector $x = (x_1, \ldots, x_N)$ specifying the amount of resources allocated to each of
+$N$ locations. We require that $0 \leq x$, i.e., that each $x_i$ is non-negative, and that the total allocation across
+all locations equals the amount of available resources, $K$: $\sum_{i=1}^N x_i = K$. The set $\mathcal{X}$ consists of
+all possible allocations that satisfy these constraints. The eventually realized resource need in each location is
+denoted by $y = (y_1, \ldots, y_N)$. The future outcome that will determine the loss caused by choosing the allocation
+$x$ is the vector of resource needs $y = (y_1, \ldots, y_N)$ for each location which is a realization (unknown at the
+time of decision making) of the random vector $Y = (Y_1, \ldots, Y_N)$. A forecast of the random need $Y$ for each
+location is written as $F = (F_1, \ldots, F_N)$. We assume need is non-negative and the forecasts reflect that, i.e. the
+support of each $F_i$ is a subset of $\mathbb{R}^+$. Finally, we assume that any unmet unit of need in any location
+incurs a fixed loss of $L>0$, so that if the selected resource level $x_i$ in location $i$ is less than the realized
+need $y_i$, a loss of $L \cdot (y_i - x_i)$ results. A variety of extensions to this setup are possible; for example, we
+might account for storage costs for resources that go unused, allow for a different loss per unit of unmet need in each
+location, or account for resource transportation costs. In this work, we choose to keep the loss function relatively
+simple to focus on the core ideas.
+
+It is helpful to clearly distinguish between the time $t_d$ when a \emph{d}ecision is made about a public health
+resource allocation and the time $t_r$ when \emph{r}esource needs that might be addressed by that allocation occur. Our
+setup assumes that $t_d < t_r$ and that the resource in question can only meet the need realized at $t_r$, not prevent
+it. That is, we stipulate that the random vector $Y$ of need is independent from the chosen allocation $x$. In the
+context of infectious disease, this means that we do not consider resources, such as vaccines, that are intended to
+reduce the number of people who will become sick at some point in the future. Instead, our setup addresses resources
+like hospital beds, oxygen supply, or ventilators which are intended to meet the medical needs of patients who are
+already sick. Additionally, our setup addresses decision-making that is related to resource needs only at the time
+$t_r$; we do not consider sequences of multiple decisions that are made over time or account for the impact of decisions
+on resource needs at any time other than $t_r$. We outline some opportunities to extend our work to more complex
+decision making settings in the discussion.
+
+With this problem formulation in place, we can develop a proper scoring rule following the procedure in section
+\ref{sec:methods.detailed.decisiontheory}.
+
+\paragraph{Step 1: specify a loss function.} The loss incurred by an allocation $x$ under an outcome $y$ is the sum of
+contributions from unmet need in each location:
 \begin{equation}
@@
 \end{equation}
-Here, $\max(0, y_i - x_i)$ is the unmet need in location $i$, which is given by $y_i - x_i$ if the realized need $y_i$ in location $i$ is greater than the amount $x_i$ allocated to that location, or $0$ if the amount $x_i$ allocated to unit $i$ is greater than or equal to the realized need.
-Also, $L$ is a constant scalar value, the same across all locations, specifying the ``cost'' of one unit of unmet need.
-
-\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the forecast, $x^{F,K}$, is the allocation that minimizes the expected loss, that is, the solution of the \emph{allocation problem} associated with $K$:
+Here, $\max(0, y_i - x_i)$ is the unmet need in location $i$, which is given by $y_i - x_i$ if the realized need $y_i$
+in location $i$ is greater than the amount $x_i$ allocated to that location, or $0$ if the amount $x_i$ allocated to
+unit $i$ is greater than or equal to the realized need. Also, $L$ is a constant scalar value, the same across all
+locations, specifying the ``cost'' of one unit of unmet need.
+
+\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the
+forecast, $x^{F,K}$, is the allocation that minimizes the expected loss, that is, the solution of the \emph{allocation
+problem} associated with $K$:
 \begin{align}
@@
 \end{align}
-where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]$ sums the expected loss due to unmet need across all locations.
-
-In the supplement we show that the components of the Bayes act are quantiles $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource constraint $K$, but is shared across all locations.
-This probability level is the level at which the resource constraint is satisfied: $\sum_{i=1}^N F_i^{-1}(\tau^{F,K}) = K$.
-This tells us that in order to allocate optimally (according to $F$), we must divide resources among the locations so that there is an equal forecasted probability in every location that the allocation is sufficient to meet resource need.
-This solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
-
-\paragraph{Step 3: Define the scoring rule.} We can now use the Bayes act to define a proper scoring rule for the probabilistic forecast $F$.
-Consider first the ``raw'' score defined as
+where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]$ sums the expected loss
+due to unmet need across all locations.
+
+In the supplement we show that the components of the Bayes act are quantiles $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a
+probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource constraint $K$, but is shared across
+all locations. This probability level is the level at which the resource constraint is satisfied: $\sum_{i=1}^N
+F_i^{-1}(\tau^{F,K}) = K$. This tells us that in order to allocate optimally (according to $F$), we must divide
+resources among the locations so that there is an equal forecasted probability in every location that the allocation is
+sufficient to meet resource need. This solution to the allocation problem is well-known in inventory management and is
+often attributed to \cite{hadleywhitin1963}.
+
+\paragraph{Step 3: Define the scoring rule.} We can now use the Bayes act to define a proper scoring rule for the
+probabilistic forecast $F$. Consider first the ``raw'' score defined as
 \begin{align}
@@
 \end{align}
-This measures the total unmet need across all locations that results from using the Bayes allocation associated with the forecast $F$ when the actual level of need in each location is observed to be $y_i$.
-
-To make this a more easily interpreted measure of forecast performance, we will adjust the raw score by subtracting the minimum loss
-achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$.
-When the oracle has sufficient resources to meet the total need, i.e., when $\sum_{i=1}^{N}y_i \leq K$, the oracle's loss is zero
-and allocation score coincides with the raw score.
-On the other hand, when the oracle cannot cover all need and incurs a loss of $L \cdot \left(\sum_{i=1}^{N}y_i - K \right) > 0$, we adjust
-the raw score by this loss.
+This measures the total unmet need across all locations that results from using the Bayes allocation associated with the
+forecast $F$ when the actual level of need in each location is observed to be $y_i$.
+
+To make this a more easily interpreted measure of forecast performance, we will adjust the raw score by subtracting the
+minimum loss achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$. When the
+oracle has sufficient resources to meet the total need, i.e., when $\sum_{i=1}^{N}y_i \leq K$, the oracle's loss is zero
+and allocation score coincides with the raw score. On the other hand, when the oracle cannot cover all need and incurs a
+loss of $L \cdot \left(\sum_{i=1}^{N}y_i - K \right) > 0$, we adjust the raw score by this loss.
 The oracle-adjusted score can therefore be written as
@@
 \end{align}
-The oracle adjustment aligns with a common theme in economic decision theory that \emph{opportunity loss} (often known as \emph{regret} or (negative) \emph{relative utility}) is often a more important quantity than absolute loss (see e.g., \cite{DIECIDUE201788}).
+The oracle adjustment aligns with a common theme in economic decision theory that \emph{opportunity loss} (often known
+as \emph{regret} or (negative) \emph{relative utility}) is often a more important quantity than absolute loss (see e.g.,
+\cite{DIECIDUE201788}).
 
@@
 
-The allocation score $S_A$ that we developed in the previous section measures the skill of the forecast distributions $F$ based on a single probability level $\tau^{F,K}$.
-This is appropriate if the resource constraint $K$ is a known constant.
-However, if $K$ is not precisely known at the time of decision making or there is interest in measuring the value of forecasts across a range of decision making scenarios with different resource constraints, we can use an \emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a distribution $p$:
+The allocation score $S_A$ developed in the previous section evaluates the forecast distributions
+$F$ based on a single probability level $\tau^{F,K}$. This is appropriate if the resource constraint $K$ is a known
+constant. However, if $K$ is not precisely known at the time of decision making or there is interest in measuring the
+value of forecasts across a range of decision making scenarios with different resource constraints, we can use an
+\emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a
+distribution $p$:
 $$S_{IAS}(F, y) = \int S_A(F,y; K) p(K) \, dK$$
-We note that the device of considering a range of hypothetical decision makers or decision making problems with different problem parameters has been employed in the past \cite[e.g.,][]{murphy1993whatisagoodforecast}.
+We note that the device of considering a range of hypothetical decision makers or decision making problems with
+different problem parameters has been employed in the past \cite[e.g.,][]{murphy1993whatisagoodforecast}.
 
@@
 
-The weighted interval score (WIS) was proposed in 2020 as a way to score forecasts that were being made in the early stages of the COVID-19 pandemic \citep{bracher2021evaluating}; equivalent scores had also been used in previous forecast evaluation efforts \cite[e.g.,][]{hong2016probabilisticEnergyForecasting}.
-The WIS is a proper scoring rule for forecasts that use a set of quantiles to represent a probabilistic forecast distribution.
-While pointing a reader interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS is a weighted sum of interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...).
-Larger interval scores indicate less skillful forecasts.
-An interval score consists of (a) the width of the interval, with larger intervals receiving higher scores (higher scores indicate less accuracy), and (b) a penalty if the interval does not cover the eventual observation, which increases the further away the interval is from the observed value.
-Equivalently, the WIS can also be characterized as a weighted sum of quantile scores for each individual predictive quantile.
-The quantile score for a particular quantile level assigns an asymmetric penalty to predictions that are too high or too low, with the relative sizes of the penalties set so that in expectation the score is minimized by the given quantile of the distribution.
-The most commonly used version of WIS is one that uses an equal weighting of all quantile levels, in which case WIS approximates the continuous ranked probability score (CRPS), a commonly used score for probabilistic forecasts.
-\emph{It is important to note that this weighting was proposed because the resulting score approximates the CRPS, and not because it aligned with any particular public health decision-making rationale.}
-
-That said, the quantile score and WIS can be derived using the same decision theoretic procedure that we outlined in section \ref{sec:methods.detailed}.
-In fields such as meteorology and supply chain management, a great deal of attention has been given to the problem where a decision must be made about the quantity of a resource to purchase for a single location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each unit of unmet need.
-This leads to the quantile score for the probability level $\tau = 1 - C/L$.
-From this point, the WIS or CRPS can be obtained by averaging across a range of decision making settings with different cost and loss parameters, using a similar motivation that we used to obtain the IAS from the AS in section \ref{sec:methods.detailed.integrated_allocation} \citep{gneiting2011weightedScoringRules}.
+The weighted interval score (WIS) was proposed in 2020 as a way to score forecasts that were being made in the early
+stages of the COVID-19 pandemic \citep{bracher2021evaluating}; equivalent scores had also been used in previous forecast
+evaluation efforts \cite[e.g.,][]{hong2016probabilisticEnergyForecasting}. The WIS is a proper scoring rule for
+forecasts that use a set of quantiles to represent a probabilistic forecast distribution. While pointing a reader
+interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS is a weighted sum of
+interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...). Larger
+interval scores indicate less skillful forecasts. An interval score consists of (a) the width of the interval, with
+larger intervals receiving higher scores (higher scores indicate less accuracy), and (b) a penalty if the interval does
+not cover the eventual observation, which increases the further away the interval is from the observed value.
+Equivalently, the WIS can also be characterized as a weighted sum of quantile scores for each individual predictive
+quantile. The quantile score for a particular quantile level assigns an asymmetric penalty to predictions that are too
+high or too low, with the relative sizes of the penalties set so that in expectation the score is minimized by the given
+quantile of the distribution. The most commonly used version of WIS is one that uses an equal weighting of all quantile
+levels, in which case WIS approximates the continuous ranked probability score (CRPS), a commonly used score for
+probabilistic forecasts. \emph{It is important to note that this weighting was proposed because the resulting score
+approximates the CRPS, and not because it aligned with any particular public health decision-making rationale.}
+
+That said, the quantile score and WIS can be derived using the same decision theoretic procedure that we outlined in
+section \ref{sec:methods.detailed}. In fields such as meteorology and supply chain management, a great deal of attention
+has been given to the problem where a decision must be made about the quantity of a resource to purchase for a single
+location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each
+unit of unmet need. This leads to the quantile score for the probability level $\tau = 1 - C/L$. From this point, the
+WIS or CRPS can be obtained by averaging across a range of decision making settings with different cost and loss
+parameters, using a similar motivation that we used to obtain the IAS from the AS in section
+\ref{sec:methods.detailed.integrated_allocation} \citep{gneiting2011weightedScoringRules}.
 
@@
 
-We illustrate with an application to hospital admissions in the U.S., considering a hypothetical problem of allocating a limited supply of medical resources to states.
+We illustrate with an application to hospital admissions in the U.S., considering a hypothetical problem of allocating a
+limited supply of medical resources to states.
 
@@
 For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution functions based on the provided quantiles.
-On the interior of the provided quantiles we used a monotonic cubic spline to interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as to match the two lowest and two highest quantiles (see the supplement for further details).
-In the supplement, we show that if this evaluation procedure had been specified prospectively, the resulting score would be proper \textemdash but a post hoc application of this procedure is improper.
-We use the procedure here to illustrate the properties of the score, and note that a forecaster or collaborative forecasting exercise interested in using the allocation score for evaluation could circumvent issues with propriety through thoughtful elicitation of representations of forecast distributions or by collecting the allocations at specified resource levels as part of forecast submissions.
+On the interior of the provided quantiles we used a monotonic cubic spline to interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as to match the two lowest and two highest quantiles (see section \ref{sec:distfromq} the supplement for further details).
+In
+section \ref{sec:propriety_of_parametric_approximation} of the supplement, we show that if this evaluation procedure had been specified prospectively, the resulting score would be proper \textemdash but a post hoc application of this procedure is improper.
+We use the procedure here to illustrate the properties of the AS, and note that a collaborative forecasting hub interested in using the AS for evaluation could circumvent issues with propriety by communicating the definition of the AS to forecasters and then collecting allocations at specified resource levels as part of forecast submissions.
 
@@
 Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states.
-The x-axis corresponds to the date of the observation that a model's prediction was targeting (e.g., the date the forecast was made plus the forecast horizon).
+The $x$-axis corresponds to the date of the observation that a model's prediction was targeting (e.g., the date the forecast was made plus the forecast horizon).
 AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down).
-In Panel B, the per-capital allocation is drawn in a heavier solid line.
+In Panel B, the per-capita allocation is drawn in a heavier solid line.
 Panel C shows the MWIS metric across weeks, averaged across all states.
@@
 Within each facet, each point corresponds to a week.
-The x- and y-values correspond to the MWIS standardized rank and the allocation score standardized rank for that week.
+The $x$- and $y$-values correspond to the MWIS standardized rank and the allocation score standardized rank for that week.
 Points corresponding to earlier dates have darker shading.
@@
 
-<<multi-k, fig.height=10, fig.width = 8, fig.cap = "Allocation scores (AS) across different resource constraints (K) for 10 models that made forecasts on 2021-12-20.
-Panel A shows, for each model, the AS for values of K between 200 and 60,000 at increments of 200.
+<<multi-k, fig.height=10, fig.width = 8, fig.cap = "Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20.
+Panel A shows, for each model, the AS for values of $K$ between 200 and 60,000 at increments of 200.
 The AS show a sharp peak just under 20,000, near the eventually observed number of hospitalizations.
@@
 The first (dark blue circles) computes weights proportional to a normal distribution centered at 15,000 (solid vertical gray line) with a standard deviation of 3,000, and truncated to be between 5,000 and 25,000.
-The second (light orange triangles) uses a uniform weight for all possible values of K.
-Note that the AS used in earlier sections of the application uses the single fixed value of K=15,000.
-Panel C shows how either method of computing the IAS (y-axis) is correlated with AS at K=15,000 (x-axis).
+The second (light orange triangles) uses a uniform weight for all possible values of $K$.
+Note that the AS used in earlier sections of the application uses the single fixed value of $K=15,000$.
+Panel C shows how either method of computing the IAS ($y$-axis) is correlated with AS at $K=15,000$ ($x$-axis).
 Every point represents the AS and IAS for one model.
@@
   geom_abline(slope=1, intercept=0, color="gray") +
-  xlab("Allocation score") +
+  xlab(expression(paste('Allocation score at ', italic('K'), ' = 15,000'))) +
   ylab("Integrated allocation score") +
@@
   geom_vline(xintercept = 15000, color="gray") +
+  xlab(expression(paste(italic('K')))) +
   ylab("Weight")
@@
   geom_line() +
+  xlab(expression(paste(italic('K')))) +
   theme(legend.position = c(0.01,.99), legend.justification = c(0,1),

### 2024-02-14 alloscore_manuscript/alloscore-application-shorter.tex
various edits, rephrasing

@@
 
-%In this section, we give an overview of the decision theoretic setup for developing proper scoring rules that measure the value of a forecast as an input to decision making.
-We keep the discussion here at a somewhat informal level; we refer the reader to [some subset of Brehmer and Gneiting; Grünwald and Dawid; Dawid; Granger and Pesaran 2000; Granger and Machina 2006; Ehm et al.
-2016] for more technically precise discussion.
+%In this section, we give an overview of the decision theoretic setup for developing proper scoring rules that measure
+%the value of a forecast as an input to decision making. We keep the discussion here at a somewhat informal level; we
+%refer the reader to [some subset of Brehmer and Gneiting; Grünwald and Dawid; Dawid; Granger and Pesaran 2000; Granger
+%and Machina 2006; Ehm et al. 2016] for more technically precise discussion.
 
-In the framework of decision theory, a decision corresponds to the selection of an action $x$ from some set of possible actions $\mathcal{X}$.
-For example, $x$ may correspond to the level of investment in a measure designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with $\mathcal{X}$ being the set of all possible levels of investment that we might select.
-The quality of a decision to take a particular action $x$ is measured in relation to an outcome $y$ that is unknown at the time the decision is made.
-For example, $y$ may correspond to the number of individuals who eventually become sick and would benefit from the mitigation measure, and informally, an action $x$ is successful to the extent that it meets the realized need.
-In the face of uncertainty, a decision maker may use a forecast $F$ of the random variable $Y$ to help inform the selection of the action to take.
-We measure the value of a forecast as an input to this decision making process by the quality of the decisions that it leads to.
+Decision theory is concerned with understanding how an \emph{action} $x$ both is and ought to be selected from a set of
+possible actions $\mathcal{X}$ --- that is, how a decision is and ought to be made --- based on the anticipated
+consequences of $x$ under future states. We represent such future states as a random variable $Y$ and refer to a
+realization $y$ of $Y$ as an \emph{outcome}. In this paper $x$ will be related to the level of investment in a measure
+designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with
+$\mathcal{X}$ being the set of all feasible levels of investment, while an outcome $y$ will correspond to the number of
+individuals who eventually become sick and would benefit from the mitigation measure. A basic strategy in decision
+theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numerical value, or
+\emph{loss}, measuring the success of $x$, which in our examples will be the extent to which $x$ meets the realized need
+$y$.  One way that a decision maker can try to anticipate the consequence of an action $x$ is by using a forecast $F$ of
+$Y$. The loss of the action $x$ chosen according to $F$ can, in turn, be intepreted as the value of $F$ as an input to
+the decision making process under the realized outcome $y$ of $Y$.
 
-We can formalize the preceding discussion with the following three-step procedure for developing scoring rules for probabilistic forecasts:
+We arrive through this discussion at a three-step procedure for developing scoring rules for probabilistic forecasts:
 \begin{enumerate}
-\item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$ eventually occurs.
-\item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under the distribution $F$.
-\item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: $S(F, y) = s(x^F, y)$.
+  \item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$
+    eventually occurs.
+  \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under
+    the distribution $F$.
+  \item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: 
+    $S(F, y) = s(x^F, y)$.
 \end{enumerate}
-This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss function.
-We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} and in section \ref{sec:proper} of the
-supplement we demonstrate that Bayes scoring rules are proper by construction (a result also shown by \cite{dawid2007geometry,gneiting2007strictly}).
+This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss
+function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} and in section
+\ref{sec:proper} of the supplement we review the basic fact from the scoring rule literature (e.g.,
+\cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring rules are proper by construction.
 
@@
 
-In the decision making setting that we consider, an action $x = (x_1, \ldots, x_N)$ is a vector specifying the amount that is allocated to each of $N$ locations.
-We require that $0 \leq x$, i.e., that each $x_i$ is non-negative, and that the total allocation across all locations equals the amount of available resources, $K$: $\sum_{i=1}^N x_i = K$.
-The set $\mathcal{X}$ consists of all possible allocations that satisfy these constraints.
-The eventually realized resource need in each location is denoted by $y = (y_1, \ldots, y_N)$.
-These levels of need are not known at the time of decision making, so we define the random vector $Y = (Y_1, \ldots, Y_N)$ where $Y_i$ represents the as-yet-unknown level of resource need in location $i$.
-Forecasts of need in each location are collected in $F = (F_1, \ldots, F_N)$.
-We assume that resource need is non-negative and the forecasts reflect that, i.e.
-the support of each $F_i$ is a subset of $\mathbb{R}^+$.
-Finally, we assume that each unit of unmet need incurs a loss denoted by $L$, so that if the selected resource level $x_i$ in location $i$ is less than the realized need $y_i$, a loss of $L \cdot (y_i - x_i)$ results.
-A variety of extensions to this setup are possible; for example, we might account for storage costs for resources that go unused, allow for a different loss per unit of unmet need in each location, or account for resource transportation costs.
-In this work, we choose to keep the loss function relatively simple to focus on the core ideas.
+In our setting an action is a vector $x = (x_1, \ldots, x_N)$ specifying the amount of resources allocated to each of
+$N$ locations. We require that $0 \leq x$, i.e., that each $x_i$ is non-negative, and that the total allocation across
+all locations equals the amount of available resources, $K$: $\sum_{i=1}^N x_i = K$. The set $\mathcal{X}$ consists of
+all possible allocations that satisfy these constraints. The eventually realized resource need in each location is
+denoted by $y = (y_1, \ldots, y_N)$. The future outcome that will determine the loss caused by choosing the allocation
+$x$ is the vector of resource needs $y = (y_1, \ldots, y_N)$ for each location which is a realization (unknown at the
+time of decision making) of the random vector $Y = (Y_1, \ldots, Y_N)$. A forecast of the random need $Y$ for each
+location is written as $F = (F_1, \ldots, F_N)$. We assume need is non-negative and the forecasts reflect that, i.e. the
+support of each $F_i$ is a subset of $\mathbb{R}^+$. Finally, we assume that any unmet unit of need in any location
+incurs a fixed loss of $L>0$, so that if the selected resource level $x_i$ in location $i$ is less than the realized
+need $y_i$, a loss of $L \cdot (y_i - x_i)$ results. A variety of extensions to this setup are possible; for example, we
+might account for storage costs for resources that go unused, allow for a different loss per unit of unmet need in each
+location, or account for resource transportation costs. In this work, we choose to keep the loss function relatively
+simple to focus on the core ideas.
 
-It is helpful to clearly distinguish between the time $t_d$ when a \emph{d}ecision is made about a public health resource allocation and the time $t_r$ when \emph{r}esource needs that might be addressed by that allocation occur.
-Our setup assumes that $t_d < t_r$.
-Additionally, the structure of our loss captures a setting where the resource in question does not impact the amount of demand that will materialize at time $t_r$, but rather it is a resource that satisfies that demand.
-In the context of infectious disease, this means that we do not consider resources that are intended to reduce the number of people who will become sick at some point in the future, such as a preventative influenza or COVID-19 vaccine.
-Instead, our setup addresses resources like hospital beds, oxygen supply, or ventilators which are intended to meet the medical needs of patients who are already sick.
-Additionally, our setup addresses decision-making that is related to resource needs only at the time $t_r$; we do not explicitly consider sequences of multiple decisions that are made over time or account for the impact of decisions on resource needs at any time other than $t_r$.
-We outline some opportunities to extend our work to more complex decision making settings in the discussion.
+It is helpful to clearly distinguish between the time $t_d$ when a \emph{d}ecision is made about a public health
+resource allocation and the time $t_r$ when \emph{r}esource needs that might be addressed by that allocation occur. Our
+setup assumes that $t_d < t_r$ and that the resource in question can only meet the need realized at $t_r$, not prevent
+it. That is, we stipulate that the random vector $Y$ of need is independent from the chosen allocation $x$. In the
+context of infectious disease, this means that we do not consider resources, such as vaccines, that are intended to
+reduce the number of people who will become sick at some point in the future. Instead, our setup addresses resources
+like hospital beds, oxygen supply, or ventilators which are intended to meet the medical needs of patients who are
+already sick. Additionally, our setup addresses decision-making that is related to resource needs only at the time
+$t_r$; we do not consider sequences of multiple decisions that are made over time or account for the impact of decisions
+on resource needs at any time other than $t_r$. We outline some opportunities to extend our work to more complex
+decision making settings in the discussion.
 
-With this problem formulation in place, we can develop a proper scoring rule following the outline in section \ref{sec:methods.detailed.decisiontheory}.
+With this problem formulation in place, we can develop a proper scoring rule following the procedure in section
+\ref{sec:methods.detailed.decisiontheory}.
 
-\paragraph{Step 1: specify a loss function.} The loss associated with a particular allocation is calculated by summing contributions from unmet need in each location:
+\paragraph{Step 1: specify a loss function.} The loss incurred by an allocation $x$ under an outcome $y$ is the sum of
+contributions from unmet need in each location:
 \begin{equation}
@@
 \end{equation}
-Here, $\max(0, y_i - x_i)$ is the unmet need in location $i$, which is given by $y_i - x_i$ if the realized need $y_i$ in location $i$ is greater than the amount $x_i$ allocated to that location, or $0$ if the amount $x_i$ allocated to unit $i$ is greater than or equal to the realized need.
-Also, $L$ is a constant scalar value, the same across all locations, specifying the ``cost'' of one unit of unmet need.
+Here, $\max(0, y_i - x_i)$ is the unmet need in location $i$, which is given by $y_i - x_i$ if the realized need $y_i$
+in location $i$ is greater than the amount $x_i$ allocated to that location, or $0$ if the amount $x_i$ allocated to
+unit $i$ is greater than or equal to the realized need. Also, $L$ is a constant scalar value, the same across all
+locations, specifying the ``cost'' of one unit of unmet need.
 
-\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the forecast, $x^{F,K}$, is the allocation that minimizes the expected loss, that is, the solution of the \emph{allocation problem} associated with $K$:
+\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the
+forecast, $x^{F,K}$, is the allocation that minimizes the expected loss, that is, the solution of the \emph{allocation
+problem} associated with $K$:
 \begin{align}
@@
 \end{align}
-where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]$ sums the expected loss due to unmet need across all locations.
+where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]$ sums the expected loss
+due to unmet need across all locations.
 
-In the supplement we show that the components of the Bayes act are quantiles $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource constraint $K$, but is shared across all locations.
-This probability level is the level at which the resource constraint is satisfied: $\sum_{i=1}^N F_i^{-1}(\tau^{F,K}) = K$.
-This tells us that in order to allocate optimally (according to $F$), we must divide resources among the locations so that there is an equal forecasted probability in every location that the allocation is sufficient to meet resource need.
-This solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}.
+In the supplement we show that the components of the Bayes act are quantiles $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a
+probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource constraint $K$, but is shared across
+all locations. This probability level is the level at which the resource constraint is satisfied: $\sum_{i=1}^N
+F_i^{-1}(\tau^{F,K}) = K$. This tells us that in order to allocate optimally (according to $F$), we must divide
+resources among the locations so that there is an equal forecasted probability in every location that the allocation is
+sufficient to meet resource need. This solution to the allocation problem is well-known in inventory management and is
+often attributed to \cite{hadleywhitin1963}.
 
-\paragraph{Step 3: Define the scoring rule.} We can now use the Bayes act to define a proper scoring rule for the probabilistic forecast $F$.
-Consider first the ``raw'' score defined as
+\paragraph{Step 3: Define the scoring rule.} We can now use the Bayes act to define a proper scoring rule for the
+probabilistic forecast $F$. Consider first the ``raw'' score defined as
 \begin{align}
@@
 \end{align}
-This measures the total unmet need across all locations that results from using the Bayes allocation associated with the forecast $F$ when the actual level of need in each location is observed to be $y_i$.
+This measures the total unmet need across all locations that results from using the Bayes allocation associated with the
+forecast $F$ when the actual level of need in each location is observed to be $y_i$.
 
-To make this a more easily interpreted measure of forecast performance, we will adjust the raw score by subtracting the minimum loss
-achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$.
-When the oracle has sufficient resources to meet the total need, i.e., when $\sum_{i=1}^{N}y_i \leq K$, the oracle's loss is zero
-and allocation score coincides with the raw score.
-On the other hand, when the oracle cannot cover all need and incurs a loss of $L \cdot \left(\sum_{i=1}^{N}y_i - K \right) > 0$, we adjust
-the raw score by this loss.
+To make this a more easily interpreted measure of forecast performance, we will adjust the raw score by subtracting the
+minimum loss achievable by an \emph{oracle} allocator which has precise foreknowledge of the outcomes $y_i$. When the
+oracle has sufficient resources to meet the total need, i.e., when $\sum_{i=1}^{N}y_i \leq K$, the oracle's loss is zero
+and allocation score coincides with the raw score. On the other hand, when the oracle cannot cover all need and incurs a
+loss of $L \cdot \left(\sum_{i=1}^{N}y_i - K \right) > 0$, we adjust the raw score by this loss.
 The oracle-adjusted score can therefore be written as
@@
 
-The weighted interval score (WIS) was proposed in 2020 as a way to score forecasts that were being made in the early stages of the COVID-19 pandemic \citep{bracher2021evaluating}; equivalent scores had also been used in previous forecast evaluation efforts \cite[e.g.,][]{hong2016probabilisticEnergyForecasting}.
-The WIS is a proper scoring rule for forecasts that use a set of quantiles to represent a probabilistic forecast distribution.
-While pointing a reader interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS is a weighted sum of interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...).
-Larger interval scores indicate less skillful forecasts.
-An interval score consists of (a) the width of the interval, with larger intervals receiving higher scores (higher scores indicate less accuracy), and (b) a penalty if the interval does not cover the eventual observation, which increases the further away the interval is from the observed value.
-Equivalently, the WIS can also be characterized as a weighted sum of quantile scores for each individual predictive quantile.
-The quantile score for a particular quantile level assigns an asymmetric penalty to predictions that are too high or too low, with the relative sizes of the penalties set so that in expectation the score is minimized by the given quantile of the distribution.
-The most commonly used version of WIS is one that uses an equal weighting of all quantile levels, in which case WIS approximates the continuous ranked probability score (CRPS), a commonly used score for probabilistic forecasts.
-\emph{It is important to note that this weighting was proposed because the resulting score approximates the CRPS, and not because it aligned with any particular public health decision-making rationale.}
+The weighted interval score (WIS) was proposed in 2020 as a way to score forecasts that were being made in the early
+stages of the COVID-19 pandemic \citep{bracher2021evaluating}; equivalent scores had also been used in previous forecast
+evaluation efforts \cite[e.g.,][]{hong2016probabilisticEnergyForecasting}. The WIS is a proper scoring rule for
+forecasts that use a set of quantiles to represent a probabilistic forecast distribution. While pointing a reader
+interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS is a weighted sum of
+interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...). Larger
+interval scores indicate less skillful forecasts. An interval score consists of (a) the width of the interval, with
+larger intervals receiving higher scores (higher scores indicate less accuracy), and (b) a penalty if the interval does
+not cover the eventual observation, which increases the further away the interval is from the observed value.
+Equivalently, the WIS can also be characterized as a weighted sum of quantile scores for each individual predictive
+quantile. The quantile score for a particular quantile level assigns an asymmetric penalty to predictions that are too
+high or too low, with the relative sizes of the penalties set so that in expectation the score is minimized by the given
+quantile of the distribution. The most commonly used version of WIS is one that uses an equal weighting of all quantile
+levels, in which case WIS approximates the continuous ranked probability score (CRPS), a commonly used score for
+probabilistic forecasts. \emph{It is important to note that this weighting was proposed because the resulting score
+approximates the CRPS, and not because it aligned with any particular public health decision-making rationale.}
 
-That said, the quantile score and WIS can be derived using the same decision theoretic procedure that we outlined in section \ref{sec:methods.detailed}.
-In fields such as meteorology and supply chain management, a great deal of attention has been given to the problem where a decision must be made about the quantity of a resource to purchase for a single location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each unit of unmet need.
-This leads to the quantile score for the probability level $\tau = 1 - C/L$.
-From this point, the WIS or CRPS can be obtained by averaging across a range of decision making settings with different cost and loss parameters, using a similar motivation that we used to obtain the IAS from the AS in section \ref{sec:methods.detailed.integrated_allocation} \citep{gneiting2011weightedScoringRules}.
+That said, the quantile score and WIS can be derived using the same decision theoretic procedure that we outlined in
+section \ref{sec:methods.detailed}. In fields such as meteorology and supply chain management, a great deal of attention
+has been given to the problem where a decision must be made about the quantity of a resource to purchase for a single
+location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each
+unit of unmet need. This leads to the quantile score for the probability level $\tau = 1 - C/L$. From this point, the
+WIS or CRPS can be obtained by averaging across a range of decision making settings with different cost and loss
+parameters, using a similar motivation that we used to obtain the IAS from the AS in section
+\ref{sec:methods.detailed.integrated_allocation} \citep{gneiting2011weightedScoringRules}.
 
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Feb  9 13:33:13 2024
+% Wed Feb 14 01:08:30 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Feb  9 13:33:17 2024
+% Wed Feb 14 01:08:35 2024
 \begin{table}[ht]
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/multi-k-1} \caption[Allocation scores (AS) across different resource constraints (K) for 10 models that made forecasts on 2021-12-20]{Allocation scores (AS) across different resource constraints (K) for 10 models that made forecasts on 2021-12-20.
-Panel A shows, for each model, the AS for values of K between 200 and 60,000 at increments of 200.
+\includegraphics[width=\maxwidth]{figure/multi-k-1} \caption[Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20]{Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20.
+Panel A shows, for each model, the AS for values of $K$ between 200 and 60,000 at increments of 200.
 The AS show a sharp peak just under 20,000, near the eventually observed number of hospitalizations.

### 2024-02-15 alloscore_manuscript/alloscore-application-shorter.Rnw
plot reworking and some figure caption edits

@@
 
+model_palette <- setNames(
+  hcl(h = seq(15, 375, length = 18), l = 65, c = 100),
+  c(unique(forecast_data$model), "per-capita"))
+model_palette["JHUAPL-Bucky"] <- "thistle"
+model_palette["CU-select"] <- "wheat"
+
 Kat15k_alloscores <- bind_rows(Kat15k_alloscores, percap |> filter(K == 15000))
@@
 
-<<thermometer-plot, fig.height=7, fig.width=7.5, fig.cap="Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the ten highest hospitalization counts.
+<<thermometer-plot, fig.height=7, fig.width=7.5, fig.cap="Probabilistic forecasts for new COVID-19 hospital admissions on January 3, 2022 and their suggested resource allocations for the states with the ten highest (eventually observed) hospitalization counts.
 For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts.
@@
               geo_value = abbreviation),
-          f_colors = c("thistle", "wheat"),
+          f_colors = model_palette[c("JHUAPL-Bucky", "CU-select")],
           one_K = 15000) + labs(x = NULL)
@@
 
-A comparison of the forecasts from the JHUAPL-Bucky and CU-select models shows that while the JHUAPL-Bucky forecast distributions were closer to the eventual observations in many states, the allocations suggested by those forecasts often were more inefficient than those from the CU-select model (Figure \ref{fig:thermometer-plot}).
-In the example of this one week, JHUAPL-Bucky had a worse allocation score than CU-select because its forecasts led to allocations that sent excess resources to several states, such as Ohio, Pennsylvania, and Michigan, that would have been more effectively allocated to states that did receive enough resources to meet their needs, such as Florida and California (Figure \ref{fig:multi-loc-ranks}A).
-These allocation errors resulted because that model’s forecasts did not consistently capture the relative resource needs across different states.
-(Figure \ref{fig:thermometer-plot} \& \ref{fig:multi-loc-ranks}A).
-The CU-select model made some similar errors — most prominently, over-allocating resources to Ohio — but overall, it did a better job of forecasting the relative resource demands across different locations.
-
-On the other hand, CU-select had worse performance as measured by WIS.
+A state-wise comparison using Figures \ref{fig:thermometer-plot} and \ref{fig:multi-loc-ranks}A of the JHUAPL-Bucky and CU-select models shows that while the JHUAPL-Bucky forecast distributions were more centered on the eventual observations in many states, the suggested allocations of the JHUAPL-Bucky model were less efficient than those of the CU-select model.
+Thus JHUAPL-Bucky achieved a worse allocation score than CU-select (on this particular week) by allocating excess resources to several states, such as Pennsylvania and Michigan, rather than to states such as Florida and California where
+these resources would have reduced unmet need.
+These allocation errors of JHUAPL-Bucky resulted from forecasts failing to consistently capture the relative resource needs across different states.
+The CU-select model made some similar errors — most prominently, over-allocating resources to Ohio — but overall, it more successfully forecasted the resource demands across different locations in relative terms.
+
+On the other hand, CU-select had worse performance as measured by MWIS.
 Its forecasts were biased downwards, and it consistently incurred a large penalty for underprediction (Figure \ref{fig:multi-loc-ranks}B).
-Predictions from the JHUAPL-Bucky model were wider, and included the observed level of daily hospital admissions more often.
-Therefore, it did not receive as severe penalties for forecasts that underpredicted or overpredicted the actual hospitalization count.
+The JHUAPL-Bucky model had wider predictive intervals which more often included the observed level of daily hospital admissions.
+Its MWIS was therefore less severely penalized by under- or over-predictions of the actual hospitalization counts.
 
@@
   scale_x_date(limits=range(score_data_summaries$target_end_date)) +
+  scale_color_manual(values = model_palette) +
   scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4, 6)) +
@@
   geom_point() + geom_line() +
+  scale_color_manual(values = model_palette) +
   scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4, 6)) +
@@
 Most models were ranked similarly when evaluated using mean allocation scores (MAS) and mean weighted interval scores (MWIS), with some places of disagreement.
-The the four most accurate and the five least accurate models were the same according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
-(This comparison excludes the per-capita method which can be used to suggest allocations but does not make probabilistic forecasts and therefore cannot have a WIS.)
-Notably, the best model according to MWIS was the JHUAPL-SLPHospEns model, but this model only had the fourth best allocation scores.
+The four most accurate and the five least accurate models were the same according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
+(This comparison excludes the per-capita allocation which does not use forecasts and therefore cannot have a WIS.)
+Notably, the best model according to MWIS was the JHUAPL-SLPHospEns model, but this model had only the fourth best MAS.
 
@@
   xtable::xtable(caption = "Mean weighted interval scores (MWIS) and mean allocation scores (MAS) by model across 13 weeks.
-For the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, plus a per-capita allocation, these results show the average performance of each model across all weeks.
-Lower scores indicate better performance, and models are sorted with the most accurate allocations at the top.
-WIS is not defined for the per-capita model as this approach only provides a mechanism to allocate resources, it does not provide probabilistic forecasts.",
+These results show the average performance across time for the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, as well
+  as for a per-capita allocation. WIS is not defined for the per-capita allocation which does not use forecasts.",
                  digits = 0,
@@
 
-<<multi-k, fig.height=10, fig.width = 8, fig.cap = "Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20.
+<<multi-k, fig.height=5, fig.width = 8, fig.cap = "Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20.
 Panel A shows, for each model, the AS for values of $K$ between 200 and 60,000 at increments of 200.
 The AS show a sharp peak just under 20,000, near the eventually observed number of hospitalizations.
-Panel B shows two possible weighting functions for the Integrated Allocation Score (IAS).
-The first (dark blue circles) computes weights proportional to a normal distribution centered at 15,000 (solid vertical gray line) with a standard deviation of 3,000, and truncated to be between 5,000 and 25,000.
-The second (light orange triangles) uses a uniform weight for all possible values of $K$.
+Two possible weighting functions for the Integrated Allocation Score (IAS) are shown.
+The first (blue shaded area) computes weights proportional to a normal distribution centered at 15,000 with a standard deviation of 3,000, and truncated to be between 5,000 and 25,000.
+The second (orange shaded area) uses a uniform weight for all possible values of $K$.
 Note that the AS used in earlier sections of the application uses the single fixed value of $K=15,000$.
-Panel C shows how either method of computing the IAS ($y$-axis) is correlated with AS at $K=15,000$ ($x$-axis).
-Every point represents the AS and IAS for one model.
-The IAS centered at 15,000 are more closely correlated with the AS values.">>=
+Panel B shows how the two versions of the IAS ($y$-axis) correlate with the AS at $K=15,000$ ($x$-axis).
+Every point represents the AS at 15,000 and a version of the IAS for one model.
+As we would expect, the IAS centered at 15,000 (circles) is more closely correlated with the AS at 15,000 than is the uniform IAS (triangles).">>=
 
@@
 
-p1 <- ggplot(ias_to_plot, aes(x=AS, y=IAS, color=IAS_type, shape=IAS_type)) +
+model_palette_ias <- model_palette[unique(ias_to_plot$model)]
+model_linetypes <- setNames(rep(
+  c("solid",
+    "dashed",
+    "dotted",
+    "dotdash",
+    "longdash", "twodash"),
+  length.out = length(model_palette_ias)
+), names(model_palette_ias))
+model_linetypes["per-capita"] <- "solid"
+model_linewidths <- setNames(rep(.5,
+  length.out = length(model_palette_ias)
+), names(model_palette_ias))
+model_linewidths["per-capita"] <- 1.5
+
+p1 <- ggplot(ias_to_plot, aes(x=AS, y=IAS, group=IAS_type, color=model, shape=IAS_type)) +
   geom_point() +
-  geom_smooth(method="lm", se=FALSE, linetype=2) +
-  scale_color_manual(name = "Integrated allocation score weighting", values = c("#386cb0", "#fdc086")) +
+  geom_smooth(aes(color = IAS_type), method="lm", se=FALSE, linetype=2) +
+  scale_color_manual(guide = "none", values = c(model_palette, "uniform" = "#fdc086", "centered at 15,000" = "#386cb0")) +
+  #scale_color_manual(name = "Integrated allocation score weighting", values = c("#386cb0", "#fdc086")) +
   scale_shape_manual(name = "Integrated allocation score weighting", values = c(1, 2)) +
@@
   ylab("Integrated allocation score") +
-  theme(legend.position = c(0.05,.9), legend.justification = c(0,1),
+  theme(legend.position = c(0.03,.95), legend.justification = c(0,1),
         legend.background=element_rect(fill = alpha("white", 0.2)))
-
-p2 <- slim_alloscore_2021.12.20 |>
-  group_by(model, K) |>
-  summarize(`Allocation score` = sum(components)) |>
-  left_join(kprobs) |>
-  ggplot(aes(x=K)) +
-  geom_point(aes(y=weight), color = "#386cb0") +
-  geom_point(y=1/length(unique(slim_alloscore_2021.12.20$K)), color="#fdc086", shape = 2) +
-  geom_vline(xintercept = 15000, color="gray") +
-  xlab(expression(paste(italic('K')))) +
-  ylab("Weight")
 
@@
   mutate(model = factor(model, levels = rev(ias$model))) |>
-  ggplot(aes(x=K, y=`Allocation score`, color=model, linetype=model)) +
-  geom_line() +
+  ggplot(aes(x=K)) +
+  geom_line(aes(y=`Allocation score`, color=model, linetype = model, 
+                linewidth = model)) +
+  geom_area(data = kprobs, aes(x = K, y=200000*weight), fill = "#386cb0", alpha = .3) +	
+  geom_area(aes(y=200000/length(unique(slim_alloscore_2021.12.20$K))), fill="#fdc086", alpha = .3) +
   xlab(expression(paste(italic('K')))) +
-  theme(legend.position = c(0.01,.99), legend.justification = c(0,1),
-        legend.background=element_rect(fill = alpha("white", 0.2)),
-        legend.title = element_blank(), legend.text = element_text(size=7))
-
-cowplot::plot_grid(p3, p2, p1,
-                   nrow=3,
+  scale_color_manual(values = model_palette) +
+  scale_linetype_manual(values = model_linetypes) +
+  scale_linewidth_manual(values = model_linewidths) +
+  theme(legend.position = "none", legend.title = element_blank())
+
+p3_leg <- p3 + theme(legend.position = "bottom", legend.title = element_blank())
+
+leg3 <- cowplot::get_legend(p3_leg)
+
+cowplot::plot_grid(
+  cowplot::plot_grid(
+    p3,p1,
+    ncol = 2,
     align = "h",
-                   labels = "AUTO")
+    labels = "AUTO",
+    hjust = -2
+  ),
+  leg3,
+  ncol = 1,
+  rel_heights = c(1, 0.3)
+)
 @

### 2024-02-15 alloscore_manuscript/alloscore-application-shorter.tex
plot reworking and some figure caption edits

@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Wed Feb 14 01:08:30 2024
+% Thu Feb 15 02:01:02 2024
 \begin{table}[ht]
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/thermometer-plot-1} \caption[Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the ten highest hospitalization counts]{Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the ten highest hospitalization counts.
+\includegraphics[width=\maxwidth]{figure/thermometer-plot-1} \caption[Probabilistic forecasts for new COVID-19 hospital admissions on January 3, 2022 and their suggested resource allocations for the states with the ten highest (eventually observed) hospitalization counts]{Probabilistic forecasts for new COVID-19 hospital admissions on January 3, 2022 and their suggested resource allocations for the states with the ten highest (eventually observed) hospitalization counts.
 For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts.
@@
 
-A comparison of the forecasts from the JHUAPL-Bucky and CU-select models shows that while the JHUAPL-Bucky forecast distributions were closer to the eventual observations in many states, the allocations suggested by those forecasts often were more inefficient than those from the CU-select model (Figure \ref{fig:thermometer-plot}).
-In the example of this one week, JHUAPL-Bucky had a worse allocation score than CU-select because its forecasts led to allocations that sent excess resources to several states, such as Ohio, Pennsylvania, and Michigan, that would have been more effectively allocated to states that did receive enough resources to meet their needs, such as Florida and California (Figure \ref{fig:multi-loc-ranks}A).
-These allocation errors resulted because that model’s forecasts did not consistently capture the relative resource needs across different states.
-(Figure \ref{fig:thermometer-plot} \& \ref{fig:multi-loc-ranks}A).
-The CU-select model made some similar errors — most prominently, over-allocating resources to Ohio — but overall, it did a better job of forecasting the relative resource demands across different locations.
+A state-wise comparison using Figures \ref{fig:thermometer-plot} and \ref{fig:multi-loc-ranks}A of the JHUAPL-Bucky and CU-select models shows that while the JHUAPL-Bucky forecast distributions were more centered on the eventual observations in many states, the suggested allocations of the JHUAPL-Bucky model were less efficient than those of the CU-select model.
+Thus JHUAPL-Bucky achieved a worse allocation score than CU-select (on this particular week) by allocating excess resources to several states, such as Pennsylvania and Michigan, rather than to states such as Florida and California where
+these resources would have reduced unmet need.
+These allocation errors of JHUAPL-Bucky resulted from forecasts failing to consistently capture the relative resource needs across different states.
+The CU-select model made some similar errors — most prominently, over-allocating resources to Ohio — but overall, it more successfully forecasted the resource demands across different locations in relative terms.
 
-On the other hand, CU-select had worse performance as measured by WIS.
+On the other hand, CU-select had worse performance as measured by MWIS.
 Its forecasts were biased downwards, and it consistently incurred a large penalty for underprediction (Figure \ref{fig:multi-loc-ranks}B).
-Predictions from the JHUAPL-Bucky model were wider, and included the observed level of daily hospital admissions more often.
-Therefore, it did not receive as severe penalties for forecasts that underpredicted or overpredicted the actual hospitalization count.
+The JHUAPL-Bucky model had wider predictive intervals which more often included the observed level of daily hospital admissions.
+Its MWIS was therefore less severely penalized by under- or over-predictions of the actual hospitalization counts.
 
@@
 Most models were ranked similarly when evaluated using mean allocation scores (MAS) and mean weighted interval scores (MWIS), with some places of disagreement.
-The the four most accurate and the five least accurate models were the same according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
-(This comparison excludes the per-capita method which can be used to suggest allocations but does not make probabilistic forecasts and therefore cannot have a WIS.)
-Notably, the best model according to MWIS was the JHUAPL-SLPHospEns model, but this model only had the fourth best allocation scores.
+The four most accurate and the five least accurate models were the same according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
+(This comparison excludes the per-capita allocation which does not use forecasts and therefore cannot have a WIS.)
+Notably, the best model according to MWIS was the JHUAPL-SLPHospEns model, but this model had only the fourth best MAS.
 
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Wed Feb 14 01:08:35 2024
+% Thu Feb 15 02:01:06 2024
 \begin{table}[ht]
@@
 \caption{Mean weighted interval scores (MWIS) and mean allocation scores (MAS) by model across 13 weeks.
-For the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, plus a per-capita allocation, these results show the average performance of each model across all weeks.
-Lower scores indicate better performance, and models are sorted with the most accurate allocations at the top.
-WIS is not defined for the per-capita model as this approach only provides a mechanism to allocate resources, it does not provide probabilistic forecasts.}
+These results show the average performance across time for the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, as well
+  as for a per-capita allocation. WIS is not defined for the per-capita allocation which does not use forecasts.} 
 \label{tab:multi-week-performance-summary}
@@
 The AS show a sharp peak just under 20,000, near the eventually observed number of hospitalizations.
-Panel B shows two possible weighting functions for the Integrated Allocation Score (IAS).
-The first (dark blue circles) computes weights proportional to a normal distribution centered at 15,000 (solid vertical gray line) with a standard deviation of 3,000, and truncated to be between 5,000 and 25,000.
-The second (light orange triangles) uses a uniform weight for all possible values of $K$.
+Two possible weighting functions for the Integrated Allocation Score (IAS) are shown.
+The first (blue shaded area) computes weights proportional to a normal distribution centered at 15,000 with a standard deviation of 3,000, and truncated to be between 5,000 and 25,000.
+The second (orange shaded area) uses a uniform weight for all possible values of $K$.
 Note that the AS used in earlier sections of the application uses the single fixed value of $K=15,000$.
-Panel C shows how either method of computing the IAS ($y$-axis) is correlated with AS at $K=15,000$ ($x$-axis).
-Every point represents the AS and IAS for one model.
-The IAS centered at 15,000 are more closely correlated with the AS values.}\label{fig:multi-k}
+Panel B shows how the two versions of the IAS ($y$-axis) correlate with the AS at $K=15,000$ ($x$-axis).
+Every point represents the AS at 15,000 and a version of the IAS for one model.
+As we would expect, the IAS centered at 15,000 (circles) is more closely correlated with the AS at 15,000 than is the uniform IAS (triangles).}\label{fig:multi-k}
 \end{figure}

### 2024-02-16 alloscore_manuscript/alloscore-application-shorter.Rnw
more edits and reformatting... Discussion paragraphs 2,3,5 still need work.

@@
 impact on and through policy.  And within this body of work we have found the discussion of such a connection to still
-be at a formally and quantitively imprecise stage. In \cite{ioannidis2022forecastingCOVIDfailed}, the possible negative
+be at a formally and quantitatively imprecise stage. In \cite{ioannidis2022forecastingCOVIDfailed}, the possible negative
 consequences of inaccurate forecasts of infectious disease are discussed, but there is no attempt to quantify the
@@
 $y$.  One way that a decision maker can try to anticipate the consequence of an action $x$ is by using a forecast $F$ of
-$Y$. The loss of the action $x$ chosen according to $F$ can, in turn, be intepreted as the value of $F$ as an input to
+$Y$. The loss of the action $x$ chosen according to $F$ can, in turn, be interpreted as the value of $F$ as an input to
 the decision making process under the realized outcome $y$ of $Y$.
@@
 This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss
-function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} and in section
-\ref{sec:proper} of the supplement we review the basic fact from the scoring rule literature (e.g.,
-\cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring rules are proper by construction.
+function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} --- a terminology we find to
+be convenient but which is not, to our knowledge, standard.  In section \ref{sec:proper} of the supplement we review the
+basic fact from the scoring rule literature (e.g., \cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring
+rules are proper by construction.
 
@@
 setup assumes that $t_d < t_r$ and that the resource in question can only meet the need realized at $t_r$, not prevent
-it. That is, we stipulate that the random vector $Y$ of need is independent from the chosen allocation $x$. In the
+it. That is, we stipulate that the random vector $Y$ of need is independent of the chosen allocation $x$. In the
 context of infectious disease, this means that we do not consider resources, such as vaccines, that are intended to
@@
 
-We illustrate with an application to hospital admissions in the U.S., considering a hypothetical problem of allocating a
-limited supply of medical resources to states.
+We illustrate our new forecast evaluation framework with an application to hospital admissions in the U.S., considering
+a hypothetical problem of allocating a limited supply of medical resources to states.
 
@@
 
-%\subsubsection{Hospitalization data}
-%Starting in the summer of 2020, the US Health and Human Services (HHS) began reporting counts of daily new admissions to hospitals for individuals with COVID-19 \citep{healthdatagov_covid-19_nodate}.
-%These daily counts were available for the US as a whole, and all states and several additional jurisdictions such as Puerto Rico and Washington DC.
-%The data were updated daily and were available for download by the public through the HHS HealthData.gov website.
-%For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data \cite{wang-covidhubutils}.
-
-%\subsubsection{Forecast data}
-The US COVID-19 Forecast Hub collected short-term forecasts of daily new hospital admissions for individuals with COVID-19 starting in December 2020 \citep{cramer_united_2022}.
-The target data for these forecasts were hospital admissions as reported by the US Department of Health and Human Services through the HealthData.gov website.
-Forecasts were probabilistic predictions of the number of new hospital admissions on a particular day in the future, in a specific jurisdiction of the US (national level, state, or territory).
-Probability distributions were represented using a set of 23 quantiles for each individual prediction, including a median and the lower and upper limits of 11 central prediction intervals, from a 99\% to a 10\% prediction interval.
-
-The analysis in this work focuses on forecasts made before and during the first wave of the Omicron SARS-CoV-2 variant in the US.
-As such, we analyzed forecasts for the 15 weeks starting with Monday November 22, 2021 through Monday February 28, 2022.
-%Forecast data were downloaded on for analysis on \Sexpr{format(as.Date(metadata$time[which(metadata$name == "forecast_data")]), "%B %d, %Y")}.
-
-Submission to the Forecast Hub followed a weekly cycle, and each Monday the Hub collected the most recent forecasts submitted by all teams that met certain inclusion criteria and created ensemble forecasts using quantile averaging \citep{ray_comparing_2023}.
-Our analysis includes these ensembles (COVIDhub-ensemble and COVIDhub-trained\_ensemble) as well as one other ensemble of hub models created by another team (JHUAPL-SLPHospEns) and several other individual models.
-Models were eligible to be included in the analysis if they were designated as a ``primary'' model from a team.
-For a model to have a complete, eligible submission in a given week, it had to have a 14 day-ahead forecast for all 50 states plus Washington DC.
-Models had to have a complete forecast for at least 4 of the 15 weeks in the analysis to be eligible for inclusion.
+The US COVID-19 Forecast Hub collected short-term forecasts of daily new hospital admissions for individuals with
+COVID-19 starting in December 2020 \citep{cramer_united_2022}. The target data for these forecasts were hospital
+admissions as reported by the US Department of Health and Human Services through the HealthData.gov website. Forecasts
+were probabilistic predictions of the number of new hospital admissions on a particular day in the future, in a specific
+jurisdiction of the US (national level, state, or territory). Probability distributions were represented using a set of
+23 quantiles for each individual prediction, including a median and the lower and upper limits of 11 central prediction
+intervals, from a 99\% to a 10\% prediction interval.
+
+The analysis in this work focuses on forecasts made before and during the first wave of the Omicron SARS-CoV-2 variant
+in the US. As such, we analyzed forecasts for the 15 weeks starting with Monday November 22, 2021 through Monday
+February 28, 2022.
+
+Submission to the Forecast Hub followed a weekly cycle, and each Monday the Hub collected the most recent forecasts
+submitted by all teams that met certain inclusion criteria and created ensemble forecasts using quantile averaging
+\citep{ray_comparing_2023}. Our analysis includes these ensembles (COVIDhub-ensemble and COVIDhub-trained\_ensemble) as
+well as one other ensemble of hub models created by another team (JHUAPL-SLPHospEns) and several other individual
+models. Models were eligible to be included in the analysis if they were designated as a ``primary'' model from a team.
+For a model to have a complete, eligible submission in a given week, it had to have a 14 day-ahead forecast for all 50
+states plus Washington DC. Models had to have a complete forecast for at least 4 of the 15 weeks in the analysis to be
+eligible for inclusion.
 
@@
 
-We measured forecast skill using two forecast scores, the allocation score (AS) and the weighted interval score (WIS), both defined above.
-We computed these scores for the 14 day ahead forecasts made each week.
-For this analysis, we fixed the resource constraint used for the AS to be $K=15,000$, based roughly on a reported number of ventilators available for reallocation in the US \citep{ajao_assessing_2015}.
-We computed the mean WIS (MWIS) across all of the forecasted state-level locations.
-%For each week, we computed the allocation score for the 14 day-ahead forecast.
-
-For both scores, we also computed models standardized rank among all models that submitted forecasts each week.
-The standardized rank is between 0 and 1, where 0 corresponds to the worst rank and 1 to the best.
-In the case of a tie between one or more models, all models received the better rank.
-
-As described above, predictions were submitted to the Forecast Hub in the form of a set of 23 quantiles of the predictive distribution.
-The WIS can be directly calculated from these quantiles.
-However, calculation of the AS requires that the full cumulative distribution functions of the forecast distributions for each location are available.
-For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution functions based on the provided quantiles.
-On the interior of the provided quantiles we used a monotonic cubic spline to interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as to match the two lowest and two highest quantiles (see section \ref{sec:distfromq} the supplement for further details).
-In
-section \ref{sec:propriety_of_parametric_approximation} of the supplement, we show that if this evaluation procedure had been specified prospectively, the resulting score would be proper \textemdash but a post hoc application of this procedure is improper.
-We use the procedure here to illustrate the properties of the AS, and note that a collaborative forecasting hub interested in using the AS for evaluation could circumvent issues with propriety by communicating the definition of the AS to forecasters and then collecting allocations at specified resource levels as part of forecast submissions.
+We evaluated forecasts using the allocation score (AS) and the weighted interval score (WIS), computed at a horizon of
+14 days. For this analysis, we fixed the resource constraint used for the AS to be $K=15,000$, based roughly on a
+reported number of ventilators available for reallocation in the US \citep{ajao_assessing_2015}. We computed the mean
+WIS (MWIS) across all of the forecasted state-level locations.
+
+For both scores, we also computed standardized ranks among all models that submitted forecasts each week. The
+standardized rank lies in the interval $[0, 1]$, where 0 corresponds to the worst rank and 1 to the best. In the case of
+a tie between one or more models, all models received the better rank.
+
+As described above, predictions were submitted to the Forecast Hub in the form of a set of 23 quantiles of the
+predictive distribution. The WIS can be directly calculated from these quantiles. However, calculation of the AS
+requires that the full cumulative distribution functions of the forecast distributions for each location are available.
+For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution
+functions based on the provided quantiles. On the interior of the provided quantiles we used a monotonic cubic spline to
+interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as
+to match the two lowest and two highest quantiles (see section \ref{sec:distfromq} the supplement for further details).
+In section \ref{sec:propriety_of_parametric_approximation} of the supplement, we show that if this evaluation procedure
+had been specified prospectively, the resulting score would be proper, \textemdash but that a post hoc application of this
+procedure is improper. We use the procedure here to illustrate the properties of the AS, and note that a collaborative
+forecasting hub interested in using the AS for evaluation could circumvent propriety issues by communicating the
+definition of the AS to forecasters and then collecting allocations at specified resource levels as part of forecast
+submissions.
 
@@
 
-We used two benchmark methods as references to allow for comparison of model-derived allocations with existing standard practices.
-First, we evaluated forecasts from the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk \citep{cramer_evaluation_2022}.
-Second, we generated a proposed set of allocations where the quantity allocated to each state was proportional to that state's population (using US Census data of vintage 2022 \citep{us_census_nst_est2022}), referred to below as \texttt{per-capita}.
-These two approaches generally reflect common choices for ``best practices'' of allocation: either allocating resources based on the most recent observed data, or in proportion to the population of each location.
-%As such, they serve as good comparisons for allocations based on forecast models.
+In order to explore how model-derived allocations might be compare to ``standard operating procedures'' for allocation
+used by public health officials, we scored two simple benchmark methods. First, we evaluated forecasts from the
+\texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds
+based on a random walk \citep{cramer_evaluation_2022}. Second, we generated a proposed set of allocations where the
+quantity allocated to each state was proportional to that state's population (using US Census data of vintage 2022
+\citep{us_census_nst_est2022}), referred to below as \texttt{per-capita}. These two approaches generally reflect common
+choices for ``best practices'' of allocation: either allocating resources based on the most recent observed data, or in
+proportion to the population of each location.
 
@@
 
-All forecast data used in this evaluation are available through the COVID-19 Forecast Hub \citep{cramer_united_2022}.
-An R package implementing the allocation score is available at \url{https://github.com/aaronger/alloscore}.
-All code and data for the analyses presented in this manuscript are available at \url{https://github.com/aaronger/utility-eval-papers}.
-The analyses were generated using a reproducible workflow using \Sexpr{version$version.string} and the {\tt targets} package \citep{Rcore-2023, landau_2021_targets}.
+All forecast data used in this evaluation are available through the COVID-19 Forecast Hub \citep{cramer_united_2022}. An
+R package implementing the allocation score is available at \url{https://github.com/aaronger/alloscore}. All code and
+data for the analyses presented in this manuscript are available at
+\url{https://github.com/aaronger/utility-eval-papers}. The analyses were generated using a reproducible workflow using
+\Sexpr{version$version.string} and the {\tt targets} package \citep{Rcore-2023, landau_2021_targets}.
 
@@
 tar_load("percap")
-
-model_palette <- setNames(
-  hcl(h = seq(15, 375, length = 18), l = 65, c = 100),
-  c(unique(forecast_data$model), "per-capita"))
-model_palette["JHUAPL-Bucky"] <- "thistle"
-model_palette["CU-select"] <- "wheat"
 
@@
 
+<<normalize-model-aesthetics>>=
+model_palette <- setNames(
+  hcl(h = seq(15, 375, length = 18), l = 65, c = 100),
+  c(unique(forecast_data$model), "per-capita"))
+
+model_palette["JHUAPL-Bucky"] <- "thistle"
+model_palette["CU-select"] <- "wheat"
+
+model_linewidths <- setNames(rep(.5,
+  length.out = length(model_palette)
+), names(model_palette))
+model_linewidths["per-capita"] <- 1.5
+@
+
 \subsubsection{Anatomy of forecast scores for one week}
 
-To illustrate the mechanics of allocation scoring, we start by focusing on how forecasts generated on or before December 20, 2021, with predictions for January 03, 2022, were scored by different metrics.
-This week was around the peak of the Omicon wave nationally, with individual states typically observing a peak at or after January 3, 2022.
-
-%For many locations, forecasts predicted lower values than were eventually observed, as this was during the period of steep increase of viral transmission across many states.
-Of the 10 models evaluated for this one week, the CU-select model had the most accurate forecasts according to the allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table \ref{tab:multi-k-scores}).
-The JHUAPL-Bucky model had the second best MWIS but the third worst allocation score.
+To illustrate the mechanics of allocation scoring, we start by focusing on how forecasts generated on or before December
+20, 2021, with predictions for January 03, 2022, were scored by different metrics. This week was around the peak of the
+Omicron wave nationally, with individual states typically observing a peak at or after January 3, 2022.
+
+Of the 10 models evaluated for this one week, the CU-select model had the most accurate forecasts according to the
+allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table
+\ref{tab:multi-k-scores}). The JHUAPL-Bucky model had the second best MWIS but the third worst allocation score.
 
@@
   arrange(AS) |>
-  xtable::xtable(caption = "For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS).
+  xtable::xtable(caption = paste("For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS), (see Section \\ref{sec:ias_examp}).
 All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03.
 Results are sorted by AS.
-For all metrics, lower scores indicate better accuracy.",
+For all metrics, lower scores indicate better accuracy."),
                  digits = 0,
@@
 @
-
-%Forecasts and allocations for a selection of states with high numbers of hospitalizations on January 3, 2022 reveal mechanics about how allocations are made (Figure \ref{fig:thermometer-plot}).
-%As described in the methods above, allocations for a given location from a specific model are generated by finding the single quantile of the probabilistic forecast across all locations that reaches the resource allocation limit (in this case, assumed to be 15,000).
-%A direct comparison between the forecasts from the JHUAPL-Bucky and CU-select models show that while the JHUAPL-Bucky forecast distributions were closer to the eventual observations in many states, the allocations suggested by those forecasts often were more inefficient than those from the CU-select model (Figure \ref{fig:thermometer-plot}).
 
@@
 
-A state-wise comparison using Figures \ref{fig:thermometer-plot} and \ref{fig:multi-loc-ranks}A of the JHUAPL-Bucky and CU-select models shows that while the JHUAPL-Bucky forecast distributions were more centered on the eventual observations in many states, the suggested allocations of the JHUAPL-Bucky model were less efficient than those of the CU-select model.
-Thus JHUAPL-Bucky achieved a worse allocation score than CU-select (on this particular week) by allocating excess resources to several states, such as Pennsylvania and Michigan, rather than to states such as Florida and California where
-these resources would have reduced unmet need.
-These allocation errors of JHUAPL-Bucky resulted from forecasts failing to consistently capture the relative resource needs across different states.
-The CU-select model made some similar errors — most prominently, over-allocating resources to Ohio — but overall, it more successfully forecasted the resource demands across different locations in relative terms.
-
-On the other hand, CU-select had worse performance as measured by MWIS.
-Its forecasts were biased downwards, and it consistently incurred a large penalty for underprediction (Figure \ref{fig:multi-loc-ranks}B).
-The JHUAPL-Bucky model had wider predictive intervals which more often included the observed level of daily hospital admissions.
-Its MWIS was therefore less severely penalized by under- or over-predictions of the actual hospitalization counts.
-
-%Together, these results corroborate the understanding that the allocation score and weighted interval score measure different aspects of forecast performance.
-The allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.
+A state-wise comparison using Figures \ref{fig:thermometer-plot} and \ref{fig:multi-loc-ranks}A of the JHUAPL-Bucky and
+CU-select models shows that while the JHUAPL-Bucky forecast distributions were more centered on the eventual
+observations in many states, the suggested allocations of the JHUAPL-Bucky model were less efficient than those of the
+CU-select model. Thus JHUAPL-Bucky achieved a worse allocation score than CU-select (on this particular week) by
+allocating excess resources to several states, such as Pennsylvania and Michigan, rather than to states such as Florida
+and California where these resources would have reduced unmet need.  These allocation errors of JHUAPL-Bucky resulted
+from forecasts failing to consistently capture the relative resource needs across different states. The CU-select model
+made some similar errors — most prominently, over-allocating resources to Ohio — but overall, it more successfully
+forecasted the resource demands across different locations in relative terms.
+
+On the other hand, CU-select had worse performance as measured by MWIS. Its forecasts were biased downwards, and it
+consistently incurred a large penalty for underprediction (Figure \ref{fig:multi-loc-ranks}B). The JHUAPL-Bucky model
+had wider predictive intervals which more often included the observed level of daily hospital admissions. Its MWIS was
+therefore less severely penalized by under- or over-predictions of the actual hospitalization counts.
 
@@
 p2 <- ggplot(alloscore_data_summaries,
-             aes(x=target_end_date, y=alloscore, color=model, shape=model)) +
+             aes(x=target_end_date, y=alloscore, color=model, shape=model, linewidth = model)) +
+  geom_point() + geom_line()+
   scale_x_date(limits=range(score_data_summaries$target_end_date)) +
@@
   scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4, 6)) +
-  scale_linewidth_manual(values = c(.5,1.5)) +
+  scale_linewidth_manual(values = model_linewidths) +
   ylab("Allocation score") +
-  geom_point() + geom_line(aes(linewidth = as.factor(model == "per-capita")))+
   theme(legend.pos="none") +
@@
                          mutate(reference_date = as.Date(reference_date))),
-             aes(x=target_end_date, y=mwis, color=model,shape=model)) +
+             aes(x=target_end_date, y=mwis, color=model,shape=model, linewidth = model)) +
   geom_point() + geom_line() +
@@
   scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4, 6)) +
+  scale_linewidth_manual(values = model_linewidths) +
   xlab("Forecast target date") +
   ylab("MWIS")+
-  theme(legend.pos="bottom")+
   guides(colour = guide_legend(title=NULL, ncol = 4),
-         shape = guide_legend(title=NULL, ncol = 4))
+         shape = guide_legend(title=NULL, ncol = 4),
+         linewidth = guide_legend(title=NULL, ncol = 4)) +
+  theme(legend.pos="bottom")
 
@@
 
-Allocation scores varied substantially by date and by model (Figure \ref{fig:metrics-over-time}).
-For predictions made for the first three Mondays in December 2021 and the last three Mondays in February 2022 all models had allocation scores under 500 (and the mean across all models was less than 100), indicating that unnecessary unmet need was fairly low on those days.
-The allocation scores were on the whole highest when the observed number of new hospital admissions was closest to the resource threshold of 15,000, as those are the times when any mistakes in allocation are costly in terms of wasting resources in one location that could have been used in another.
-%Predictions made during the peak week and just after showed the highest variation in allocation scores, with some models having allocation scores under 1000 and others having values over 3500.
+Allocation scores varied substantially by date and by model (Figure \ref{fig:metrics-over-time}). For predictions made
+for the first three Mondays in December 2021 and the last three Mondays in February 2022 all models had allocation
+scores under 500 (and the mean across all models was less than 100), indicating that unnecessary unmet need was fairly
+low on those days. The allocation scores were on the whole highest when the observed number of new hospital admissions
+was closest to the resource threshold of 15,000.  It was on these occasions that models were most likely to waste resources
+in one location that the oracle would have put to use in another.
 
@@
 
-Averaging across all weeks, the ensemble forecast achieved a better mean allocation score (MAS) than two benchmark methods.
-The COVIDhub-ensemble had the best MAS across all weeks, the per-capita allocation had the second best MAS, and the COVIDhub-baseline model had the sixth best allocation (Table \ref{tab:multi-week-performance-summary}).
-No individual model had a better MAS than the per-capita allocation.
-
-Most models were ranked similarly when evaluated using mean allocation scores (MAS) and mean weighted interval scores (MWIS), with some places of disagreement.
-The four most accurate and the five least accurate models were the same according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
-(This comparison excludes the per-capita allocation which does not use forecasts and therefore cannot have a WIS.)
-Notably, the best model according to MWIS was the JHUAPL-SLPHospEns model, but this model had only the fourth best MAS.
+Averaging across all weeks, the ensemble forecast achieved a better mean allocation score (MAS) than two benchmark
+methods. The COVIDhub-ensemble had the best MAS across all weeks, the per-capita allocation had the second best MAS, and
+the COVIDhub-baseline model had the sixth best allocation (Table \ref{tab:multi-week-performance-summary}). No
+individual model had a better MAS than the per-capita allocation.
+
+Model rankings on mean allocation score (MAS) and time-averaged mean weighted interval score
+(taMWIS) were broadly similar, with some places of disagreement. The four most accurate and the five least accurate models were the same
+according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
+(This comparison excludes the per-capita allocation which does not use forecasts and therefore cannot be assigned a MWIS.)
+Notably, the JHUAPL-SLPHospEns model ranked best on taMWIS but had only the fourth best MAS.
 
@@
             MAS = mean(alloscore, na.rm=TRUE),
-            MWIS = mean(mwis, na.rm=TRUE)) |>
+            taMWIS = mean(mwis, na.rm=TRUE)) |>
   filter(N > 12) |>
+  mutate(`taMWIS rank` = rank(taMWIS, na.last = "keep")) |>
   arrange(MAS) |>
-  select(model, MAS, MWIS) |>
-  xtable::xtable(caption = "Mean weighted interval scores (MWIS) and mean allocation scores (MAS) by model across 13 weeks.
+  select(model, MAS, taMWIS, `taMWIS rank`) |>
+  xtable::xtable(caption = "Time averaged mean weighted interval scores (taMWIS) and mean allocation scores (MAS) by model across 13 weeks.
 These results show the average performance across time for the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, as well
-  as for a per-capita allocation. WIS is not defined for the per-capita allocation which does not use forecasts.",
+  as for a per-capita allocation. MWIS is not defined for the per-capita allocation which does not use forecasts.",
                  digits = 0,
@@
 \subsubsection{Metrics were not consistently correlated over time}
+
+We can gain some insight into the discordance between MAS and taMWIS by examining the correlation between time-specific AS and MWIS values of the various models in Figure
+\ref{fig:metrics-correlation}. We find, for example, no clear association between the consistent
+and high MWIS rank of the JHUAPL-SLPHospEns model and its highly variable AS rank. This contrasts 
+with a clearly positive correlation between MWIS and AS ranks in other models such as 
+Karlen-pypm and USC-SI\_kJalpha.
+We also can see in more detail the consistently high performance COVIDhub-ensemble on both scores.
+And while it did not submit forecasts for enough weeks to be included in Table \ref{tab:multi-week-performance-summary}, CMU-TimeSeries is an interesting example of a model that 
+performed consistently well on AS but had only middling MWIS ranks.
 
@@
 
-Models showed differing levels of correlation between their allocation scores and MWIS values (Figure \ref{fig:metrics-correlation}), explaining in part why aggregate MWIS and MAS values did not always agree.
-Here are some examples of the different model-specific patterns observed:
-\begin{itemize}
-\item Several models showed a positive association between allocation score and MWIS ranks (e.g., \texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha}).
-\item One model had consistently strong MWIS ranks but also had highly variable allocation score ranks with no clear association between the two (e.g., \texttt{JHUAPL-SLPHospEns})
-\item One model performed consistently well for both metrics (\texttt{COVIDhub-ensemble})
-\item One model performed consistently well for allocation score but had only middling ranks for MWIS (\texttt{CMU-TimeSeries})
-\end{itemize}
-
 <<eval=FALSE>>=
@@
 \subsubsection{Integrated allocation score across values of K}
+\label{sec:ias_examp}
 
@@
 
-AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 as well as the per-capita allocation (Figure \ref{fig:multi-k}A).
-These calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of \Sexpr{format(tot_y, big.mark = ",")} that day.
-Rankings of AS from all models were fairly stable across a range of values for K, with some crossings, especially at $K$ values further away from the observed value.
-
-The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the constraint ($K$), possibly taking weights into account for different values of $K$ that might be more or less likely (Section \ref{sec:methods.detailed.integrated_allocation}).
-IAS was computed for two distributions on $K$, one uniform across the entire range and the other a symmetric distribution around 15,000 (Figure \ref{fig:multi-k}B).
-Both versions of the IAS were correlated with the original AS conducted at $K=15,000$, with the higher correlation coming from the distribution that was centered at $K=15,000$ (Figure \ref{fig:multi-k}C).
-Model rankings based on the AC and the centered IAS were roughly similar, with the top and bottom three approaches being the same for both scores (Table \ref{tab:multi-k-scores}).
-
-<<multi-k, fig.height=5, fig.width = 8, fig.cap = "Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20.
+AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of
+hospitalizations on January 3, 2021 as well as for the per-capita allocation (Figure \ref{fig:multi-k}A). These
+calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of
+\Sexpr{format(tot_y, big.mark = ",")} that day. Model ranks by AS were fairly stable across all $K$, and there appears
+to be an substantial interval around the observed value over which ranks are constant.
+
+The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the
+constraint ($K$), possibly assigning higher weights to more likely values of $K$ (Section
+\ref{sec:methods.detailed.integrated_allocation}). IAS was computed for two weight distributions on $K$, one uniform
+across the entire range and the other a symmetric distribution around 15,000 (the orange and blue shaded areas in Figure
+\ref{fig:multi-k}A, respectively). Both versions of the IAS were correlated with the original AS for $K=15,000$, with
+the higher correlation coming from the weighting centered at $K=15,000$ (Figure \ref{fig:multi-k}B). Model rankings
+based on the AS and the centered IAS were roughly similar, with the top and bottom three approaches being the same for
+both scores (Table \ref{tab:multi-k-scores}).
+
+<<multi-k, fig.height=4, fig.width = 8, fig.cap = "Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20.
 Panel A shows, for each model, the AS for values of $K$ between 200 and 60,000 at increments of 200.
@@
 model_linetypes["per-capita"] <- "solid"
-model_linewidths <- setNames(rep(.5,
-  length.out = length(model_palette_ias)
-), names(model_palette_ias))
-model_linewidths["per-capita"] <- 1.5
 
@@
 
-In epidemiological forecasting, well-known proper scoring rules such as the log score or variations on the continuous ranked probability score (such as the weighted interval score, or WIS) have been frequently utilized to evaluate probabilistic forecasts.
-Often models are ranked with respect to accuracy according to a particular score, but without reference to any underlying decision-making process for which that score was designed or from which it might be derivable as a Bayes scoring rule.
-With careful thought and collaboration between modelers and public health officials, we argue that scores that are more aligned with public health decisions could be developed to inform specific problems.
-We have demonstrated that forecast evaluation methods that are tied to a specific decision making context can yield model rankings that differ substantively from those based on standard measures of forecast skill, a result that aligns with findings in other fields \citep{leitch1991economicForecastEval, murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}.
-Additionally, we show that an existing ensemble forecast approach was the only method to outperform two benchmark allocation approaches in a hypothetical application.
-
-%In particular, the allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.
-
-%We often conceive of infectious disease forecasts as being useful for decision making purposes, but it is rare for forecast evaluation to be tied directly to the value of the forecasts for informing those decisions.
-This work seeks to address that gap.
-%However, we do note that the decision-making context presented in this work, while motivated by examples from real-world public health resource allocation problems, has not been used to inform an actual real-time decision-making process.
-
-While our example used a hypothetical scenario to illustrate a potential application of the allocation score, resource allocation decisions are a realistic example of the kinds of decisions that could motivate more targeted forecasting exercises.
+Probabilistic forecasts of infectious disease outbreaks have typically been evaluated using well-known proper scoring
+rules such as the LS, CRPS, and WIS. Often models are ranked according to such a score, but without reference to any
+underlying decision problem from which the score might be derivable as a Bayes scoring rule or be otherwise motivated.
+We argue that the value of collaborative forecasting efforts, such as the COVID-19 Forecasting Hub, could be enhanced by
+refocussing the attention of modelers and public health officials involved in these efforts on the elicitation of
+forecasts designed to support specific public health decision problems, and moreover, the evaluation of forecasts with
+scoring rules that directly consider the success of forecasts in supporting these decisions. As evidence for this
+possibility, we have demonstrated how tying forecast evaluation to a specific decision problem (via the allocation
+score) can yield model rankings that differ substantively from those based a current standard measure of forecast
+accuracy, the WIS, which does not refer to any decision problem that has been clearly connected to outbreak response.
+This result aligns with findings in other fields, especially those where the monetary consequences of using a forecast
+can be clearly juxtaposed with standard accuracy measures of the forecast \citep{leitch1991economicForecastEval,
+murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}. Additionally, we show that an existing
+ensemble forecast approach was the only method to outperform (in terms of this new evaluation method) two benchmark
+allocation approaches in a hypothetical application, suggesting that there is room for innovation of current
+epidemiological modeling techniques that might be spurred on by a reorientation of the field toward practical decision
+problems. For example, in order to surpass benchmark allocation scores, forecasters need to devote effort to capturing
+the \emph{relative} magnitude of future resource need across different locations.  Success at this task is not, however,
+directly rewarded by the WIS.
+
+While our example used a hypothetical scenario to illustrate a potential application of the allocation score, resource
+allocation decisions are a realistic example of the kinds of decisions that could motivate more targeted forecasting
+exercises.
 %One real-world example, which was the motivating example presented in the work above, is allocation of ventilators during a respiratory viral pandemic\citep{huang_stockpiling_2017}.
-We used ventilators as a notional example for a resource allocation problem because it (a) has been identified as a scarce resource during outbreaks \citep{huang_stockpiling_2017}, (b) could be allocated across different locations, with logistical considerations depending on the geographical scales, and (c) would likely be correlated (although not perfectly so) with hospitalization data.
-However, this example has limitations in that not everyone who is hospitalized needs a ventilator, and new hospital admissions may not be the best stand in for a need that corresponds to a prevalence of severe cases.
-Other possible real-world examples include the allocation of a limited stockpile of vaccinations\citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests\citep{du_optimal_2022,pasco_covid-19_2023}.
-While there may be limitations in this specific pairing of a resource and forecasted outcome, we emphasize that the important result from this work is how evaluating forecasts using a fundamentally different kind of score can illuminate previously hidden aspects of value in forecasts.
-In this setting, the feature of a forecast that emerges as being important is that the probabilistic forecasts predict the relative magnitudes between locations accurately, even if the absolute magnitudes are inaccurate.
-
-In practice, there are many potential users of forecasts with many different decision making problems.
-While outcomes of some decisions could potentially be specified by loss functions, other uses for forecasts may be less quantifiable, such as when forecasts are used to inform a general sense of situational awareness that may feed in unmeasurable ways into other decisions.
-Even those that are easily framed as expected loss minimization may differ enough that no single score would be
-appropriate for all users.
-Ideally, targeted forecasting tools could be developed through close collaboration between modelers and public health officials.
-However, this may only be possible in settings with sufficient staffing on both an analytics and a public health team.
-Increasingly, collaborative modeling hubs are being used to generate ``one-size-fits-all'' forecasts for many locations at once.
-In these settings, where tailored models are not available, it still could be possible to evaluate contributed models using a set of multiple scores to support public health end users in understanding the value of forecasts as an input to their particular decision making contexts.
-% NGR: leaving this out for now as it feels like too much detail?
-% This may be tricky to operationalize in the setting of a general forecast hub. It matters how you elicit and represent probabilistic forecasts (quantiles? samples? cdfs?).
-
-%In our specific application, we made several key observations that we believe should inform future work in this area.
-%First, it is clear that different metrics (in our case allocation score and mean weighted interval score) captured different aspects of forecast performance.
-%The allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.
-%A forecast with a good allocation score can have a poor weighted interval score and vice versa.
-%The mean WIS (MWIS) was strongly dependent on the scale of the forecasted quantity (e.g., when hospitalizations were high, so were MWIS scores).
-%The allocation score was not as scale-dependent, as the highest/worst allocation scores were observed when the number of hospitalizations was closest to the allocation constraint.
-%We also note that contributed models found it hard to achieve a better allocation score than a na\"ive baseline model that just predicted a flat line from the last previous observation with wide uncertainty.
-%This suggests suggests that contributed models (other than the ensemble, which combined forecasts from all contributed models) were not consistently adding value to allocation decisions over just extrapolating from the current levels.
-
-There are several important limitations to the current work.
-%NGR - removing the sentence below because it seems like not a central concern/limitation, and I was triyng to cut. Happy to be overruled.
-%The allocation score we developed here does not directly account for important considerations such as fairness or equity of allocations.
-The proposed framework does not attempt to capture the broader context of decision making.
-For example, in practice it may be possible to increase the resource constraint $K$ by shifting funding from other disease mitigation measures.
-We also note that in some settings, a ``successful'' epidemiological forecast may lead to policy decisions that change the distribution of the predicted outcome $Y$.
-Our framework would need considerable enhancements before being applicable to forecast evaluation beyond horizons
-for which causal feedback can be neglected.
-
-An opportunity for further investigation is to more carefully evaluate whether a forecast adds value to existing decision making processes.
-In the context of decisions about allocations, standard procedures might involve extrapolating need based on a current observed data (similar to the benchmark approaches presented above), with or without adjustments based on other political or real-world considerations.
-For example, in many settings public health stakeholders will make decisions after synthesizing information from a variety of quantitative and qualitative sources coupled with expert judgment.
-The allocation score presented in this work does not directly measure whether a given forecast adds useful information to such an existing decision-making process.
-While the scoring procedures as presented do not directly address this question, they could be modified (say, by comparison to a `baseline' model or expert-elicited allocations in the absence of forecast data) to quantify the benefit of using a forecast to inform a specific decision.
-
-%We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling applied epidemiological forecasting with public health decision making.
-%We note that a few papers have begun exploring similar linkages as described in the literature review{\textendash}but we see much room for additional work in this area.
-%In some situations, individual or ensemble models could be developed to optimize scores that are attuned to a particular decision making setting.
-%This area is also largely unexplored in the realm of public health to date, with some initial methodological development in econometrics \citep{loaiza-maya_focused_2021}.
-
-In conclusion, we argue that the way modelers and policymakers view and evaluate forecasts should change depending on the specific decision-making context.
-Defaulting to standard forecast evaluation metrics can mask the utility (or disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those for which they might offer useful guidance.
-New collaborative work between public health officials and modeling teams is needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or simulation exercises that could be used to inform further development of new or alternative scoring metrics.
-We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling applied epidemiological forecasting with public health decision making.
+We used ventilators as a notional example for a resource allocation problem because it (a) has been identified as a
+scarce resource during outbreaks \citep{huang_stockpiling_2017}, (b) could be allocated across different locations, with
+logistical considerations depending on the geographical scales, and (c) would likely be correlated (although not
+perfectly so) with hospitalization data. However, this example has limitations in that not everyone who is hospitalized
+needs a ventilator, and new hospital admissions may not be the best stand in for a need that corresponds to a prevalence
+of severe cases. Other possible real-world examples include the allocation of a limited stockpile of vaccinations
+\citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests \citep{du_optimal_2022,pasco_covid-19_2023}. While
+there may be limitations in this specific pairing of a resource and forecasted outcome, we emphasize that the important
+result from this work is how evaluating forecasts using a fundamentally different kind of score can illuminate
+previously hidden aspects of value in forecasts. In this setting, the feature of a forecast that emerges as being
+important is that the probabilistic forecasts predict the relative magnitudes between locations accurately, even if the
+absolute magnitudes are inaccurate.
+
+In practice, there are many potential users of forecasts with many different decision making problems. While outcomes of
+some decisions could potentially be specified by loss functions, other uses for forecasts may be less quantifiable, such
+as when forecasts are used to inform a general sense of situational awareness that may feed in unmeasurable ways into
+other decisions. Even those that are easily framed as expected loss minimization may differ enough that no single score
+would be appropriate for all users. Ideally, targeted forecasting tools could be developed through close collaboration
+between modelers and public health officials. However, this may only be possible in settings with sufficient staffing on
+both an analytics and a public health team. Increasingly, collaborative modeling hubs are being used to generate
+``one-size-fits-all'' forecasts for many locations at once. But it could still be beneficial in these settings to expand
+the set of scoring methods used by the hub to include scores based on specific public health decision problems, such as
+AS, allowing end users to better understand the value of available forecasts as inputs to their particular
+decision making contexts. (We note that issues of propriety are raised when forecasts are evaluated with scoring rules
+for which they were not elicited or via parameters imputed by a hub after submission, both of which are done in our
+application example.  We consider these issues briefly in supplement sections \ref{sec:distfromq} and
+\ref{sec:propriety_of_parametric_approximation}.)
+
+There are several important limitations to the current work. The allocation score we developed here does not directly
+account for important considerations such as fairness or equity of allocations, or more broadly, individual and group
+preference relations that are difficult or even impossible to encode into a loss or utility function.  Ambiguity aversion
+(in the sense of \cite{ellsberg1961risk}), for example, seems especially relevant to the use of forecasts in outbreak
+response, where there can be immense social and political pressure on public health officials to maintain a transparent
+base of evidence for their policy choices and recommendations. The proposed framework also does not attempt to capture
+the broader context of decision making. For example, in practice it may be possible to increase the resource constraint
+$K$ by shifting funding from other disease mitigation measures. Finally, we emphasize that we opted at this stage of
+development to explicitly rule out situations where a ``successful''
+epidemiological forecast could lead to policy decisions that change the distribution of the predicted outcome $Y$. Our
+framework would need considerable enhancements before being applicable to forecast evaluation beyond horizons for which
+causal feedback can be neglected.
+
+An opportunity for further investigation is to more carefully evaluate whether forecasts add value to 
+``standard operating procedures'' already in place for making public health decisions. In the context of allocations,
+such a procedure might extrapolate need from current observed need and population levels (similarly to the two benchmark
+approaches presented above), with adjustments based on other political or real-world considerations. For example, in
+many settings public health stakeholders will make decisions after synthesizing information from a variety of
+quantitative and qualitative sources coupled with expert judgment. The allocation score presented in this work does not
+directly measure whether a given forecast adds useful information to such an existing decision-making process. While the
+scoring procedures as presented do not directly address this question, they could be modified (say, by comparison to a
+`baseline' model or expert-elicited allocations in the absence of forecast data) to quantify the benefit of using a
+forecast to inform a specific decision.
+
+In conclusion, we argue that the way modelers and policymakers view and evaluate forecasts should more explicitly depend on
+the specific decision-making context. Defaulting to standard forecast evaluation metrics can mask the utility (or
+disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those
+for which they are able to offer useful guidance. New collaborative work between public health officials and modeling teams is
+needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or
+simulation exercises that could be used to inform further development of new or alternative scoring metrics. We see this
+work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling
+applied epidemiological forecasting with public health decision making.
 
@@
 
-We wish to thank the following individuals who contributed valuable comments and feedback on early versions of this work: Matthew Biggerstaff, Rebecca Borchering, Sebastian Funk, Melissa Kerr, and Jeffrey Shaman.
+We wish to thank the following individuals who contributed valuable comments and feedback on early versions of this
+work: Matthew Biggerstaff, Rebecca Borchering, Sebastian Funk, Melissa Kerr, and Jeffrey Shaman.
 
 This work has been supported by the National Institutes of General Medical Sciences (R35GM119582) and the U.S.
-CDC(1U01IP001122).
-The content is solely the responsibility of the authors and does not necessarily represent the official views of NIGMS, the National Institutes of Health, or CDC.
+CDC(1U01IP001122). The content is solely the responsibility of the authors and does not necessarily represent the
+official views of NIGMS, the National Institutes of Health, or CDC.
 

### 2024-02-16 alloscore_manuscript/alloscore-application-shorter.tex
more edits and reformatting... Discussion paragraphs 2,3,5 still need work.

@@
 
-%\subsubsection{Hospitalization data}
-%Starting in the summer of 2020, the US Health and Human Services (HHS) began reporting counts of daily new admissions to hospitals for individuals with COVID-19 \citep{healthdatagov_covid-19_nodate}.
-%These daily counts were available for the US as a whole, and all states and several additional jurisdictions such as Puerto Rico and Washington DC.
-%The data were updated daily and were available for download by the public through the HHS HealthData.gov website.
-%For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data \cite{wang-covidhubutils}.
+The US COVID-19 Forecast Hub collected short-term forecasts of daily new hospital admissions for individuals with
+COVID-19 starting in December 2020 \citep{cramer_united_2022}. The target data for these forecasts were hospital
+admissions as reported by the US Department of Health and Human Services through the HealthData.gov website. Forecasts
+were probabilistic predictions of the number of new hospital admissions on a particular day in the future, in a specific
+jurisdiction of the US (national level, state, or territory). Probability distributions were represented using a set of
+23 quantiles for each individual prediction, including a median and the lower and upper limits of 11 central prediction
+intervals, from a 99\% to a 10\% prediction interval.
 
-%\subsubsection{Forecast data}
-The US COVID-19 Forecast Hub collected short-term forecasts of daily new hospital admissions for individuals with COVID-19 starting in December 2020 \citep{cramer_united_2022}.
-The target data for these forecasts were hospital admissions as reported by the US Department of Health and Human Services through the HealthData.gov website.
-Forecasts were probabilistic predictions of the number of new hospital admissions on a particular day in the future, in a specific jurisdiction of the US (national level, state, or territory).
-Probability distributions were represented using a set of 23 quantiles for each individual prediction, including a median and the lower and upper limits of 11 central prediction intervals, from a 99\% to a 10\% prediction interval.
+The analysis in this work focuses on forecasts made before and during the first wave of the Omicron SARS-CoV-2 variant
+in the US. As such, we analyzed forecasts for the 15 weeks starting with Monday November 22, 2021 through Monday
+February 28, 2022.
 
-The analysis in this work focuses on forecasts made before and during the first wave of the Omicron SARS-CoV-2 variant in the US.
-As such, we analyzed forecasts for the 15 weeks starting with Monday November 22, 2021 through Monday February 28, 2022.
-%Forecast data were downloaded on for analysis on format(as.Date(metadata$time[which(metadata$name == "forecast_data")]), "%B %d, %Y").
-
-Submission to the Forecast Hub followed a weekly cycle, and each Monday the Hub collected the most recent forecasts submitted by all teams that met certain inclusion criteria and created ensemble forecasts using quantile averaging \citep{ray_comparing_2023}.
-Our analysis includes these ensembles (COVIDhub-ensemble and COVIDhub-trained\_ensemble) as well as one other ensemble of hub models created by another team (JHUAPL-SLPHospEns) and several other individual models.
-Models were eligible to be included in the analysis if they were designated as a ``primary'' model from a team.
-For a model to have a complete, eligible submission in a given week, it had to have a 14 day-ahead forecast for all 50 states plus Washington DC.
-Models had to have a complete forecast for at least 4 of the 15 weeks in the analysis to be eligible for inclusion.
+Submission to the Forecast Hub followed a weekly cycle, and each Monday the Hub collected the most recent forecasts
+submitted by all teams that met certain inclusion criteria and created ensemble forecasts using quantile averaging
+\citep{ray_comparing_2023}. Our analysis includes these ensembles (COVIDhub-ensemble and COVIDhub-trained\_ensemble) as
+well as one other ensemble of hub models created by another team (JHUAPL-SLPHospEns) and several other individual
+models. Models were eligible to be included in the analysis if they were designated as a ``primary'' model from a team.
+For a model to have a complete, eligible submission in a given week, it had to have a 14 day-ahead forecast for all 50
+states plus Washington DC. Models had to have a complete forecast for at least 4 of the 15 weeks in the analysis to be
+eligible for inclusion.
 
@@
 
-We measured forecast skill using two forecast scores, the allocation score (AS) and the weighted interval score (WIS), both defined above.
-We computed these scores for the 14 day ahead forecasts made each week.
-For this analysis, we fixed the resource constraint used for the AS to be $K=15,000$, based roughly on a reported number of ventilators available for reallocation in the US \citep{ajao_assessing_2015}.
-We computed the mean WIS (MWIS) across all of the forecasted state-level locations.
-%For each week, we computed the allocation score for the 14 day-ahead forecast.
+We evaluated forecasts using the allocation score (AS) and the weighted interval score (WIS), computed at a horizon of
+14 days. For this analysis, we fixed the resource constraint used for the AS to be $K=15,000$, based roughly on a
+reported number of ventilators available for reallocation in the US \citep{ajao_assessing_2015}. We computed the mean
+WIS (MWIS) across all of the forecasted state-level locations.
 
-For both scores, we also computed models standardized rank among all models that submitted forecasts each week.
-The standardized rank is between 0 and 1, where 0 corresponds to the worst rank and 1 to the best.
-In the case of a tie between one or more models, all models received the better rank.
+For both scores, we also computed standardized ranks among all models that submitted forecasts each week. The
+standardized rank lies in the interval $[0, 1]$, where 0 corresponds to the worst rank and 1 to the best. In the case of
+a tie between one or more models, all models received the better rank.
 
-As described above, predictions were submitted to the Forecast Hub in the form of a set of 23 quantiles of the predictive distribution.
-The WIS can be directly calculated from these quantiles.
-However, calculation of the AS requires that the full cumulative distribution functions of the forecast distributions for each location are available.
-For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution functions based on the provided quantiles.
-On the interior of the provided quantiles we used a monotonic cubic spline to interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as to match the two lowest and two highest quantiles (see section \ref{sec:distfromq} the supplement for further details).
-In
-section \ref{sec:propriety_of_parametric_approximation} of the supplement, we show that if this evaluation procedure had been specified prospectively, the resulting score would be proper \textemdash but a post hoc application of this procedure is improper.
-We use the procedure here to illustrate the properties of the AS, and note that a collaborative forecasting hub interested in using the AS for evaluation could circumvent issues with propriety by communicating the definition of the AS to forecasters and then collecting allocations at specified resource levels as part of forecast submissions.
+As described above, predictions were submitted to the Forecast Hub in the form of a set of 23 quantiles of the
+predictive distribution. The WIS can be directly calculated from these quantiles. However, calculation of the AS
+requires that the full cumulative distribution functions of the forecast distributions for each location are available.
+For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution
+functions based on the provided quantiles. On the interior of the provided quantiles we used a monotonic cubic spline to
+interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as
+to match the two lowest and two highest quantiles (see section \ref{sec:distfromq} the supplement for further details).
+In section \ref{sec:propriety_of_parametric_approximation} of the supplement, we show that if this evaluation procedure
+had been specified prospectively, the resulting score would be proper, \textemdash but that a post hoc application of this
+procedure is improper. We use the procedure here to illustrate the properties of the AS, and note that a collaborative
+forecasting hub interested in using the AS for evaluation could circumvent propriety issues by communicating the
+definition of the AS to forecasters and then collecting allocations at specified resource levels as part of forecast
+submissions.
 
@@
 
-All forecast data used in this evaluation are available through the COVID-19 Forecast Hub \citep{cramer_united_2022}.
-An R package implementing the allocation score is available at \url{https://github.com/aaronger/alloscore}.
-All code and data for the analyses presented in this manuscript are available at \url{https://github.com/aaronger/utility-eval-papers}.
-The analyses were generated using a reproducible workflow using R version 4.3.1 (2023-06-16) and the {\tt targets} package \citep{Rcore-2023, landau_2021_targets}.
+All forecast data used in this evaluation are available through the COVID-19 Forecast Hub \citep{cramer_united_2022}. An
+R package implementing the allocation score is available at \url{https://github.com/aaronger/alloscore}. All code and
+data for the analyses presented in this manuscript are available at
+\url{https://github.com/aaronger/utility-eval-papers}. The analyses were generated using a reproducible workflow using
+R version 4.3.1 (2023-06-16) and the {\tt targets} package \citep{Rcore-2023, landau_2021_targets}.
 
@@
 
-To illustrate the mechanics of allocation scoring, we start by focusing on how forecasts generated on or before December 20, 2021, with predictions for January 03, 2022, were scored by different metrics.
-This week was around the peak of the Omicon wave nationally, with individual states typically observing a peak at or after January 3, 2022.
+To illustrate the mechanics of allocation scoring, we start by focusing on how forecasts generated on or before December
+20, 2021, with predictions for January 03, 2022, were scored by different metrics. This week was around the peak of the
+Omicron wave nationally, with individual states typically observing a peak at or after January 3, 2022.
 
-%For many locations, forecasts predicted lower values than were eventually observed, as this was during the period of steep increase of viral transmission across many states.
-Of the 10 models evaluated for this one week, the CU-select model had the most accurate forecasts according to the allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table \ref{tab:multi-k-scores}).
-The JHUAPL-Bucky model had the second best MWIS but the third worst allocation score.
+Of the 10 models evaluated for this one week, the CU-select model had the most accurate forecasts according to the
+allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table
+\ref{tab:multi-k-scores}). The JHUAPL-Bucky model had the second best MWIS but the third worst allocation score.
 
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Thu Feb 15 02:01:02 2024
+% Fri Feb 16 13:15:57 2024
 \begin{table}[ht]
@@
 \end{tabular}
-\caption{For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS).
+\caption{For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS), (see Section \ref{sec:ias_examp}).
 All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03.
@@
 \end{table}
-
-%Forecasts and allocations for a selection of states with high numbers of hospitalizations on January 3, 2022 reveal mechanics about how allocations are made (Figure \ref{fig:thermometer-plot}).
-%As described in the methods above, allocations for a given location from a specific model are generated by finding the single quantile of the probabilistic forecast across all locations that reaches the resource allocation limit (in this case, assumed to be 15,000).
-%A direct comparison between the forecasts from the JHUAPL-Bucky and CU-select models show that while the JHUAPL-Bucky forecast distributions were closer to the eventual observations in many states, the allocations suggested by those forecasts often were more inefficient than those from the CU-select model (Figure \ref{fig:thermometer-plot}).
 
@@
 
-A state-wise comparison using Figures \ref{fig:thermometer-plot} and \ref{fig:multi-loc-ranks}A of the JHUAPL-Bucky and CU-select models shows that while the JHUAPL-Bucky forecast distributions were more centered on the eventual observations in many states, the suggested allocations of the JHUAPL-Bucky model were less efficient than those of the CU-select model.
-Thus JHUAPL-Bucky achieved a worse allocation score than CU-select (on this particular week) by allocating excess resources to several states, such as Pennsylvania and Michigan, rather than to states such as Florida and California where
-these resources would have reduced unmet need.
-These allocation errors of JHUAPL-Bucky resulted from forecasts failing to consistently capture the relative resource needs across different states.
-The CU-select model made some similar errors — most prominently, over-allocating resources to Ohio — but overall, it more successfully forecasted the resource demands across different locations in relative terms.
+A state-wise comparison using Figures \ref{fig:thermometer-plot} and \ref{fig:multi-loc-ranks}A of the JHUAPL-Bucky and
+CU-select models shows that while the JHUAPL-Bucky forecast distributions were more centered on the eventual
+observations in many states, the suggested allocations of the JHUAPL-Bucky model were less efficient than those of the
+CU-select model. Thus JHUAPL-Bucky achieved a worse allocation score than CU-select (on this particular week) by
+allocating excess resources to several states, such as Pennsylvania and Michigan, rather than to states such as Florida
+and California where these resources would have reduced unmet need.  These allocation errors of JHUAPL-Bucky resulted
+from forecasts failing to consistently capture the relative resource needs across different states. The CU-select model
+made some similar errors — most prominently, over-allocating resources to Ohio — but overall, it more successfully
+forecasted the resource demands across different locations in relative terms.
 
-On the other hand, CU-select had worse performance as measured by MWIS.
-Its forecasts were biased downwards, and it consistently incurred a large penalty for underprediction (Figure \ref{fig:multi-loc-ranks}B).
-The JHUAPL-Bucky model had wider predictive intervals which more often included the observed level of daily hospital admissions.
-Its MWIS was therefore less severely penalized by under- or over-predictions of the actual hospitalization counts.
-
-%Together, these results corroborate the understanding that the allocation score and weighted interval score measure different aspects of forecast performance.
-The allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.
+On the other hand, CU-select had worse performance as measured by MWIS. Its forecasts were biased downwards, and it
+consistently incurred a large penalty for underprediction (Figure \ref{fig:multi-loc-ranks}B). The JHUAPL-Bucky model
+had wider predictive intervals which more often included the observed level of daily hospital admissions. Its MWIS was
+therefore less severely penalized by under- or over-predictions of the actual hospitalization counts.
 
@@
 
-Averaging across all weeks, the ensemble forecast achieved a better mean allocation score (MAS) than two benchmark methods.
-The COVIDhub-ensemble had the best MAS across all weeks, the per-capita allocation had the second best MAS, and the COVIDhub-baseline model had the sixth best allocation (Table \ref{tab:multi-week-performance-summary}).
-No individual model had a better MAS than the per-capita allocation.
+Averaging across all weeks, the ensemble forecast achieved a better mean allocation score (MAS) than two benchmark
+methods. The COVIDhub-ensemble had the best MAS across all weeks, the per-capita allocation had the second best MAS, and
+the COVIDhub-baseline model had the sixth best allocation (Table \ref{tab:multi-week-performance-summary}). No
+individual model had a better MAS than the per-capita allocation.
 
-Most models were ranked similarly when evaluated using mean allocation scores (MAS) and mean weighted interval scores (MWIS), with some places of disagreement.
-The four most accurate and the five least accurate models were the same according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
-(This comparison excludes the per-capita allocation which does not use forecasts and therefore cannot have a WIS.)
-Notably, the best model according to MWIS was the JHUAPL-SLPHospEns model, but this model had only the fourth best MAS.
+Model rankings on mean allocation score (MAS) and time-averaged mean weighted interval score
+(taMWIS) were broadly similar, with some places of disagreement. The four most accurate and the five least accurate models were the same
+according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
+(This comparison excludes the per-capita allocation which does not use forecasts and therefore cannot be assigned a MWIS.)
+Notably, the JHUAPL-SLPHospEns model ranked best on taMWIS but had only the fourth best MAS.
 
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Thu Feb 15 02:01:06 2024
+% Fri Feb 16 13:16:01 2024
 \begin{table}[ht]
 \centering
-\begin{tabular}{lrr}
+\begin{tabular}{lrrr}
   \hline
-model & MAS & MWIS \\ 
+model & MAS & taMWIS & taMWIS rank \\ 
   \hline
-COVIDhub-ensemble & 389 & 70 \\ 
-  per-capita & 464 & - \\ 
-  COVIDhub-trained\_ensemble & 483 & 87 \\ 
-  CU-select & 502 & 81 \\ 
-  JHUAPL-SLPHospEns & 526 & 67 \\ 
-  COVIDhub-baseline & 594 & 93 \\ 
-  JHUAPL-Bucky & 643 & 112 \\ 
-  MUNI-ARIMA & 707 & 116 \\ 
-  JHUAPL-Gecko & 929 & 110 \\ 
-  USC-SI\_kJalpha & 1473 & 155 \\ 
+COVIDhub-ensemble & 389 & 70 & 2 \\ 
+  per-capita & 464 & - & - \\ 
+  COVIDhub-trained\_ensemble & 483 & 87 & 4 \\ 
+  CU-select & 502 & 81 & 3 \\ 
+  JHUAPL-SLPHospEns & 526 & 67 & 1 \\ 
+  COVIDhub-baseline & 594 & 93 & 5 \\ 
+  JHUAPL-Bucky & 643 & 112 & 7 \\ 
+  MUNI-ARIMA & 707 & 116 & 8 \\ 
+  JHUAPL-Gecko & 929 & 110 & 6 \\ 
+  USC-SI\_kJalpha & 1473 & 155 & 9 \\ 
    \hline
 \end{tabular}
-\caption{Mean weighted interval scores (MWIS) and mean allocation scores (MAS) by model across 13 weeks.
+\caption{Time averaged mean weighted interval scores (taMWIS) and mean allocation scores (MAS) by model across 13 weeks.
 These results show the average performance across time for the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, as well
-  as for a per-capita allocation. WIS is not defined for the per-capita allocation which does not use forecasts.} 
+  as for a per-capita allocation. MWIS is not defined for the per-capita allocation which does not use forecasts.} 
 \label{tab:multi-week-performance-summary}
@@
 
-Models showed differing levels of correlation between their allocation scores and MWIS values (Figure \ref{fig:metrics-correlation}), explaining in part why aggregate MWIS and MAS values did not always agree.
-Here are some examples of the different model-specific patterns observed:
-\begin{itemize}
-\item Several models showed a positive association between allocation score and MWIS ranks (e.g., \texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha}).
-\item One model had consistently strong MWIS ranks but also had highly variable allocation score ranks with no clear association between the two (e.g., \texttt{JHUAPL-SLPHospEns})
-\item One model performed consistently well for both metrics (\texttt{COVIDhub-ensemble})
-\item One model performed consistently well for allocation score but had only middling ranks for MWIS (\texttt{CMU-TimeSeries})
-\end{itemize}
-
 \subsubsection{Integrated allocation score across values of K}
+\label{sec:ias_examp}
 
@@
 
-AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 as well as the per-capita allocation (Figure \ref{fig:multi-k}A).
-These calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of 19,581 that day.
-Rankings of AS from all models were fairly stable across a range of values for K, with some crossings, especially at $K$ values further away from the observed value.
+AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of
+hospitalizations on January 3, 2021 as well as for the per-capita allocation (Figure \ref{fig:multi-k}A). These
+calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of
+19,581 that day. Model ranks by AS were fairly stable across all $K$, and there appears
+to be an substantial interval around the observed value over which ranks are constant.
 
-The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the constraint ($K$), possibly taking weights into account for different values of $K$ that might be more or less likely (Section \ref{sec:methods.detailed.integrated_allocation}).
-IAS was computed for two distributions on $K$, one uniform across the entire range and the other a symmetric distribution around 15,000 (Figure \ref{fig:multi-k}B).
-Both versions of the IAS were correlated with the original AS conducted at $K=15,000$, with the higher correlation coming from the distribution that was centered at $K=15,000$ (Figure \ref{fig:multi-k}C).
-Model rankings based on the AC and the centered IAS were roughly similar, with the top and bottom three approaches being the same for both scores (Table \ref{tab:multi-k-scores}).
+The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the
+constraint ($K$), possibly assigning higher weights to more likely values of $K$ (Section
+\ref{sec:methods.detailed.integrated_allocation}). IAS was computed for two weight distributions on $K$, one uniform
+across the entire range and the other a symmetric distribution around 15,000 (the orange and blue shaded areas in Figure
+\ref{fig:multi-k}A, respectively). Both versions of the IAS were correlated with the original AS for $K=15,000$, with
+the higher correlation coming from the weighting centered at $K=15,000$ (Figure \ref{fig:multi-k}B). Model rankings
+based on the AS and the centered IAS were roughly similar, with the top and bottom three approaches being the same for
+both scores (Table \ref{tab:multi-k-scores}).
 
@@
 
-In epidemiological forecasting, well-known proper scoring rules such as the log score or variations on the continuous ranked probability score (such as the weighted interval score, or WIS) have been frequently utilized to evaluate probabilistic forecasts.
-Often models are ranked with respect to accuracy according to a particular score, but without reference to any underlying decision-making process for which that score was designed or from which it might be derivable as a Bayes scoring rule.
-With careful thought and collaboration between modelers and public health officials, we argue that scores that are more aligned with public health decisions could be developed to inform specific problems.
-We have demonstrated that forecast evaluation methods that are tied to a specific decision making context can yield model rankings that differ substantively from those based on standard measures of forecast skill, a result that aligns with findings in other fields \citep{leitch1991economicForecastEval, murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}.
-Additionally, we show that an existing ensemble forecast approach was the only method to outperform two benchmark allocation approaches in a hypothetical application.
+Probabilistic forecasts of infectious disease outbreaks have typically been evaluated using well-known proper scoring
+rules such as the LS, CRPS, and WIS. Often models are ranked according to such a score, but without reference to any
+underlying decision problem from which the score might be derivable as a Bayes scoring rule or be otherwise motivated.
+We argue that the value of collaborative forecasting efforts, such as the COVID-19 Forecasting Hub, could be enhanced by
+refocussing the attention of modelers and public health officials involved in these efforts on the elicitation of
+forecasts designed to support specific public health decision problems, and moreover, the evaluation of forecasts with
+scoring rules that directly consider the success of forecasts in supporting these decisions. As evidence for this
+possibility, we have demonstrated how tying forecast evaluation to a specific decision problem (via the allocation
+score) can yield model rankings that differ substantively from those based a current standard measure of forecast
+accuracy, the WIS, which does not refer to any decision problem that has been clearly connected to outbreak response.
+This result aligns with findings in other fields, especially those where the monetary consequences of using a forecast
+can be clearly juxtaposed with standard accuracy measures of the forecast \citep{leitch1991economicForecastEval,
+murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}. Additionally, we show that an existing
+ensemble forecast approach was the only method to outperform (in terms of this new evaluation method) two benchmark
+allocation approaches in a hypothetical application, suggesting that there is room for innovation of current
+epidemiological modeling techniques that might be spurred on by a reorientation of the field toward practical decision
+problems. For example, in order to surpass benchmark allocation scores, forecasters need to devote effort to capturing
+the \emph{relative} magnitude of future resource need across different locations.  Success at this task is not, however,
+directly rewarded by the WIS.
 
-%In particular, the allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.
+While our example used a hypothetical scenario to illustrate a potential application of the allocation score, resource
+allocation decisions are a realistic example of the kinds of decisions that could motivate more targeted forecasting
+exercises.
+%One real-world example, which was the motivating example presented in the work above, is allocation of ventilators during a respiratory viral pandemic\citep{huang_stockpiling_2017}.
+We used ventilators as a notional example for a resource allocation problem because it (a) has been identified as a
+scarce resource during outbreaks \citep{huang_stockpiling_2017}, (b) could be allocated across different locations, with
+logistical considerations depending on the geographical scales, and (c) would likely be correlated (although not
+perfectly so) with hospitalization data. However, this example has limitations in that not everyone who is hospitalized
+needs a ventilator, and new hospital admissions may not be the best stand in for a need that corresponds to a prevalence
+of severe cases. Other possible real-world examples include the allocation of a limited stockpile of vaccinations
+\citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests \citep{du_optimal_2022,pasco_covid-19_2023}. While
+there may be limitations in this specific pairing of a resource and forecasted outcome, we emphasize that the important
+result from this work is how evaluating forecasts using a fundamentally different kind of score can illuminate
+previously hidden aspects of value in forecasts. In this setting, the feature of a forecast that emerges as being
+important is that the probabilistic forecasts predict the relative magnitudes between locations accurately, even if the
+absolute magnitudes are inaccurate.
 
-%We often conceive of infectious disease forecasts as being useful for decision making purposes, but it is rare for forecast evaluation to be tied directly to the value of the forecasts for informing those decisions.
-This work seeks to address that gap.
-%However, we do note that the decision-making context presented in this work, while motivated by examples from real-world public health resource allocation problems, has not been used to inform an actual real-time decision-making process.
+In practice, there are many potential users of forecasts with many different decision making problems. While outcomes of
+some decisions could potentially be specified by loss functions, other uses for forecasts may be less quantifiable, such
+as when forecasts are used to inform a general sense of situational awareness that may feed in unmeasurable ways into
+other decisions. Even those that are easily framed as expected loss minimization may differ enough that no single score
+would be appropriate for all users. Ideally, targeted forecasting tools could be developed through close collaboration
+between modelers and public health officials. However, this may only be possible in settings with sufficient staffing on
+both an analytics and a public health team. Increasingly, collaborative modeling hubs are being used to generate
+``one-size-fits-all'' forecasts for many locations at once. But it could still be beneficial in these settings to expand
+the set of scoring methods used by the hub to include scores based on specific public health decision problems, such as
+AS, allowing end users to better understand the value of available forecasts as inputs to their particular
+decision making contexts. (We note that issues of propriety are raised when forecasts are evaluated with scoring rules
+for which they were not elicited or via parameters imputed by a hub after submission, both of which are done in our
+application example.  We consider these issues briefly in supplement sections \ref{sec:distfromq} and
+\ref{sec:propriety_of_parametric_approximation}.)
 
-While our example used a hypothetical scenario to illustrate a potential application of the allocation score, resource allocation decisions are a realistic example of the kinds of decisions that could motivate more targeted forecasting exercises.
-%One real-world example, which was the motivating example presented in the work above, is allocation of ventilators during a respiratory viral pandemic\citep{huang_stockpiling_2017}.
-We used ventilators as a notional example for a resource allocation problem because it (a) has been identified as a scarce resource during outbreaks \citep{huang_stockpiling_2017}, (b) could be allocated across different locations, with logistical considerations depending on the geographical scales, and (c) would likely be correlated (although not perfectly so) with hospitalization data.
-However, this example has limitations in that not everyone who is hospitalized needs a ventilator, and new hospital admissions may not be the best stand in for a need that corresponds to a prevalence of severe cases.
-Other possible real-world examples include the allocation of a limited stockpile of vaccinations\citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests\citep{du_optimal_2022,pasco_covid-19_2023}.
-While there may be limitations in this specific pairing of a resource and forecasted outcome, we emphasize that the important result from this work is how evaluating forecasts using a fundamentally different kind of score can illuminate previously hidden aspects of value in forecasts.
-In this setting, the feature of a forecast that emerges as being important is that the probabilistic forecasts predict the relative magnitudes between locations accurately, even if the absolute magnitudes are inaccurate.
+There are several important limitations to the current work. The allocation score we developed here does not directly
+account for important considerations such as fairness or equity of allocations, or more broadly, individual and group
+preference relations that are difficult or even impossible to encode into a loss or utility function.  Ambiguity aversion
+(in the sense of \cite{ellsberg1961risk}), for example, seems especially relevant to the use of forecasts in outbreak
+response, where there can be immense social and political pressure on public health officials to maintain a transparent
+base of evidence for their policy choices and recommendations. The proposed framework also does not attempt to capture
+the broader context of decision making. For example, in practice it may be possible to increase the resource constraint
+$K$ by shifting funding from other disease mitigation measures. Finally, we emphasize that we opted at this stage of
+development to explicitly rule out situations where a ``successful''
+epidemiological forecast could lead to policy decisions that change the distribution of the predicted outcome $Y$. Our
+framework would need considerable enhancements before being applicable to forecast evaluation beyond horizons for which
+causal feedback can be neglected.
 
-In practice, there are many potential users of forecasts with many different decision making problems.
-While outcomes of some decisions could potentially be specified by loss functions, other uses for forecasts may be less quantifiable, such as when forecasts are used to inform a general sense of situational awareness that may feed in unmeasurable ways into other decisions.
-Even those that are easily framed as expected loss minimization may differ enough that no single score would be
-appropriate for all users.
-Ideally, targeted forecasting tools could be developed through close collaboration between modelers and public health officials.
-However, this may only be possible in settings with sufficient staffing on both an analytics and a public health team.
-Increasingly, collaborative modeling hubs are being used to generate ``one-size-fits-all'' forecasts for many locations at once.
-In these settings, where tailored models are not available, it still could be possible to evaluate contributed models using a set of multiple scores to support public health end users in understanding the value of forecasts as an input to their particular decision making contexts.
-% NGR: leaving this out for now as it feels like too much detail?
-% This may be tricky to operationalize in the setting of a general forecast hub. It matters how you elicit and represent probabilistic forecasts (quantiles? samples? cdfs?).
+An opportunity for further investigation is to more carefully evaluate whether forecasts add value to 
+``standard operating procedures'' already in place for making public health decisions. In the context of allocations,
+such a procedure might extrapolate need from current observed need and population levels (similarly to the two benchmark
+approaches presented above), with adjustments based on other political or real-world considerations. For example, in
+many settings public health stakeholders will make decisions after synthesizing information from a variety of
+quantitative and qualitative sources coupled with expert judgment. The allocation score presented in this work does not
+directly measure whether a given forecast adds useful information to such an existing decision-making process. While the
+scoring procedures as presented do not directly address this question, they could be modified (say, by comparison to a
+`baseline' model or expert-elicited allocations in the absence of forecast data) to quantify the benefit of using a
+forecast to inform a specific decision.
 
-%In our specific application, we made several key observations that we believe should inform future work in this area.
-%First, it is clear that different metrics (in our case allocation score and mean weighted interval score) captured different aspects of forecast performance.
-%The allocation score penalizes models that mischaracterize the relative magnitude of resource need across different locations, while WIS penalizes forecasts that do not capture the absolute magnitude of the target quantity in each location.
-%A forecast with a good allocation score can have a poor weighted interval score and vice versa.
-%The mean WIS (MWIS) was strongly dependent on the scale of the forecasted quantity (e.g., when hospitalizations were high, so were MWIS scores).
-%The allocation score was not as scale-dependent, as the highest/worst allocation scores were observed when the number of hospitalizations was closest to the allocation constraint.
-%We also note that contributed models found it hard to achieve a better allocation score than a na\"ive baseline model that just predicted a flat line from the last previous observation with wide uncertainty.
-%This suggests suggests that contributed models (other than the ensemble, which combined forecasts from all contributed models) were not consistently adding value to allocation decisions over just extrapolating from the current levels.
-
-There are several important limitations to the current work.
-%NGR - removing the sentence below because it seems like not a central concern/limitation, and I was triyng to cut. Happy to be overruled.
-%The allocation score we developed here does not directly account for important considerations such as fairness or equity of allocations.
-The proposed framework does not attempt to capture the broader context of decision making.
-For example, in practice it may be possible to increase the resource constraint $K$ by shifting funding from other disease mitigation measures.
-We also note that in some settings, a ``successful'' epidemiological forecast may lead to policy decisions that change the distribution of the predicted outcome $Y$.
-Our framework would need considerable enhancements before being applicable to forecast evaluation beyond horizons
-for which causal feedback can be neglected.
-
-An opportunity for further investigation is to more carefully evaluate whether a forecast adds value to existing decision making processes.
-In the context of decisions about allocations, standard procedures might involve extrapolating need based on a current observed data (similar to the benchmark approaches presented above), with or without adjustments based on other political or real-world considerations.
-For example, in many settings public health stakeholders will make decisions after synthesizing information from a variety of quantitative and qualitative sources coupled with expert judgment.
-The allocation score presented in this work does not directly measure whether a given forecast adds useful information to such an existing decision-making process.
-While the scoring procedures as presented do not directly address this question, they could be modified (say, by comparison to a `baseline' model or expert-elicited allocations in the absence of forecast data) to quantify the benefit of using a forecast to inform a specific decision.
-
-%We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling applied epidemiological forecasting with public health decision making.
-%We note that a few papers have begun exploring similar linkages as described in the literature review{\textendash}but we see much room for additional work in this area.
-%In some situations, individual or ensemble models could be developed to optimize scores that are attuned to a particular decision making setting.
-%This area is also largely unexplored in the realm of public health to date, with some initial methodological development in econometrics \citep{loaiza-maya_focused_2021}.
-
-In conclusion, we argue that the way modelers and policymakers view and evaluate forecasts should change depending on the specific decision-making context.
-Defaulting to standard forecast evaluation metrics can mask the utility (or disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those for which they might offer useful guidance.
-New collaborative work between public health officials and modeling teams is needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or simulation exercises that could be used to inform further development of new or alternative scoring metrics.
-We see this work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling applied epidemiological forecasting with public health decision making.
+In conclusion, we argue that the way modelers and policymakers view and evaluate forecasts should more explicitly depend on
+the specific decision-making context. Defaulting to standard forecast evaluation metrics can mask the utility (or
+disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those
+for which they are able to offer useful guidance. New collaborative work between public health officials and modeling teams is
+needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or
+simulation exercises that could be used to inform further development of new or alternative scoring metrics. We see this
+work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling
+applied epidemiological forecasting with public health decision making.
 

### 2024-02-17 alloscore_manuscript/alloscore-application-shorter.Rnw
attempting to rework/excise ventilator narrative

@@
 We evaluated forecasts using the allocation score (AS) and the weighted interval score (WIS), computed at a horizon of
-14 days. For this analysis, we fixed the resource constraint used for the AS to be $K=15,000$, based roughly on a
-reported number of ventilators available for reallocation in the US \citep{ajao_assessing_2015}. We computed the mean
+14 days. We chose to focus much of our analysis on the AS computed for a resource constraint of $K=15,000$. This level provides
+an anchor for interesting comparisons between phases of the Omicron wave by representing a national shortage at the peak and
+a national excess before and after the wave (see Figure \ref{fig:metrics-over-time}A), assuming that the resource need corresponds 
+directly to new hospital admissions.
+We additionally computed the mean
 WIS (MWIS) across all of the forecasted state-level locations.
@@
                    rel_heights = c(1, 1, 1.85),
-                   labels = "AUTO")
+                   labels = "AUTO", hjust = -1)
 @
@@
 
-While our example used a hypothetical scenario to illustrate a potential application of the allocation score, resource
-allocation decisions are a realistic example of the kinds of decisions that could motivate more targeted forecasting
-exercises.
-%One real-world example, which was the motivating example presented in the work above, is allocation of ventilators during a respiratory viral pandemic\citep{huang_stockpiling_2017}.
+Our example application of the AS refers to a hypothetical and unspecified resource need that is supposed to 
+pair with our forecasted outcome (new hospital admissions) in a manner that might appear overly simplistic and 
+artificially direct.
+A concrete decision problem, which motivated our work on allocation scoring, is
+
+allocation of ventilators during a respiratory viral pandemic
+\citep{huang_stockpiling_2017}.
 We used ventilators as a notional example for a resource allocation problem because it (a) has been identified as a
@@
 of severe cases. Other possible real-world examples include the allocation of a limited stockpile of vaccinations
-\citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests \citep{du_optimal_2022,pasco_covid-19_2023}. While
+\citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests \citep{du_optimal_2022,pasco_covid-19_2023}.
+
+While
 there may be limitations in this specific pairing of a resource and forecasted outcome, we emphasize that the important
@@
 
-In practice, there are many potential users of forecasts with many different decision making problems. While outcomes of
-some decisions could potentially be specified by loss functions, other uses for forecasts may be less quantifiable, such
-as when forecasts are used to inform a general sense of situational awareness that may feed in unmeasurable ways into
-other decisions. Even those that are easily framed as expected loss minimization may differ enough that no single score
-would be appropriate for all users. Ideally, targeted forecasting tools could be developed through close collaboration
-between modelers and public health officials. However, this may only be possible in settings with sufficient staffing on
-both an analytics and a public health team. Increasingly, collaborative modeling hubs are being used to generate
-``one-size-fits-all'' forecasts for many locations at once. But it could still be beneficial in these settings to expand
-the set of scoring methods used by the hub to include scores based on specific public health decision problems, such as
-AS, allowing end users to better understand the value of available forecasts as inputs to their particular
-decision making contexts. (We note that issues of propriety are raised when forecasts are evaluated with scoring rules
-for which they were not elicited or via parameters imputed by a hub after submission, both of which are done in our
-application example.  We consider these issues briefly in supplement sections \ref{sec:distfromq} and
-\ref{sec:propriety_of_parametric_approximation}.)
+In practice, epidemiological forecasts are often directed at a diverse set of end-users. It may be easy for some of
+these users to summarize the consequences of how they use a forecast with a numerical loss function, but for others,
+such as officials deciding how best to update public ``situational awareness'' of an outbreak, this may be essentially
+impossible. Even those forecast uses that are easily framed as expected loss minimization may differ enough that no
+single score would be appropriate for all users. Ideally, targeted forecasting tools could be developed through close
+collaboration between modelers and public health officials. However, this may only be possible in settings with
+sufficient staffing on both an analytics and a public health team. Increasingly, collaborative modeling hubs are being
+used to generate ``one-size-fits-all'' forecasts for many locations at once. But it could still be beneficial in these
+settings to expand the set of scoring methods used by the hub to include scores, such as the AS, which are based on
+specific public health decision problems.  This could help end users better understand the value of available forecasts
+as inputs to their particular decision making contexts. (We note that issues of propriety are raised when forecasts are
+evaluated with scoring rules for which they were not elicited or via parameters imputed by a hub after submission, both
+of which are done in our application example.  We consider these issues briefly in supplement sections
+\ref{sec:distfromq} and \ref{sec:propriety_of_parametric_approximation}.)
 

### 2024-02-17 alloscore_manuscript/alloscore-application-shorter.tex
attempting to rework/excise ventilator narrative

@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Feb 16 13:15:57 2024
+% Fri Feb 16 20:39:10 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Feb 16 13:16:01 2024
+% Fri Feb 16 20:39:15 2024
 \begin{table}[ht]
@@
 
-In practice, there are many potential users of forecasts with many different decision making problems. While outcomes of
-some decisions could potentially be specified by loss functions, other uses for forecasts may be less quantifiable, such
-as when forecasts are used to inform a general sense of situational awareness that may feed in unmeasurable ways into
-other decisions. Even those that are easily framed as expected loss minimization may differ enough that no single score
+In practice, epidemiological forecasts are often directed at a diverse set of end-users. It may be easy for some of these
+users to summarize the consequences of how they use a forecast with a numerical loss function, but for others, such
+as officials deciding how best to update public ``situational awareness'' of an outbreak, this may be essentially impossible.
+Even those forecast uses that are easily framed as expected loss minimization 
+may differ enough 
+that no single score
 would be appropriate for all users. Ideally, targeted forecasting tools could be developed through close collaboration
@@
 ``one-size-fits-all'' forecasts for many locations at once. But it could still be beneficial in these settings to expand
-the set of scoring methods used by the hub to include scores based on specific public health decision problems, such as
-AS, allowing end users to better understand the value of available forecasts as inputs to their particular
+the set of scoring methods used by the hub to include scores, such as the AS, which are based on specific public health decision
+problems.  This could help end users better understand the value of available forecasts as inputs to their particular
 decision making contexts. (We note that issues of propriety are raised when forecasts are evaluated with scoring rules

### 2024-02-17 alloscore_manuscript/alloscore-application-shorter.Rnw
convert supplement to appendices

@@
 \usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
-
+\usepackage{appendix}
 \usepackage{amsmath, amsfonts, amssymb}
@@
 \usepackage{xurl}
-\externaldocument{supplement_short}
 
@@
 section \ref{sec:methods.overview}. In section \ref{sec:methods.detailed} we develop the allocation score more
-carefully, using a decision theoretic procedure for deriving proper scoring rules. (See section \ref{sec:proper} of the
-supplement for a definition of a proper scoring rule and a more technical discussion of the procedure). In section
+carefully, using a decision theoretic procedure for deriving proper scoring rules. (See appendix \ref{sec:proper}
+for a definition of a proper scoring rule and a more technical discussion of the procedure). In section
 \ref{sec:methods.related}, we note that another group of common scores including the quantile score, WIS, and CRPS, can
@@
 that the optimal allocation divides the available resources among the locations proportionally to the scale parameters
-$\sigma_i$ (see section \ref{sec:bayes-quantiles} of the supplemental materials). If $K = 5$ units of the resource are
+$\sigma_i$ (see appendices \ref{sec:bayes-quantiles}). If $K = 5$ units of the resource are
 available, the optimal allocation according to $F$ would be 1 and 4 resource units in locations 1 and 2, respectively. 
@@
 function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} --- a terminology we find to
-be convenient but which is not, to our knowledge, standard.  In section \ref{sec:proper} of the supplement we review the
+be convenient but which is not, to our knowledge, standard.  In appendix \ref{sec:proper} we review the
 basic fact from the scoring rule literature (e.g., \cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring
@@
 
-In the supplement we show that the components of the Bayes act are quantiles $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a
+In appendix \ref{sec:bayes-quantiles} we show that the components of the Bayes act are quantiles $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a
 probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource constraint $K$, but is shared across
@@
 submitted by all teams that met certain inclusion criteria and created ensemble forecasts using quantile averaging
-\citep{ray_comparing_2023}. Our analysis includes these ensembles (COVIDhub-ensemble and COVIDhub-trained\_ensemble) as
-well as one other ensemble of hub models created by another team (JHUAPL-SLPHospEns) and several other individual
+\citep{ray_comparing_2023}. Our analysis includes these ensembles (\texttt{COVIDhub-ensemble} and \texttt{COVIDhub-trained\_ensemble}) as
+well as one other ensemble of hub models created by another team (\texttt{JHUAPL-SLPHospEns}) and several other individual
 models. Models were eligible to be included in the analysis if they were designated as a ``primary'' model from a team.
@@
 interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as
-to match the two lowest and two highest quantiles (see section \ref{sec:distfromq} the supplement for further details).
-In section \ref{sec:propriety_of_parametric_approximation} of the supplement, we show that if this evaluation procedure
+to match the two lowest and two highest quantiles (see appendix \ref{sec:distfromq} for further details).
+In appendix \ref{sec:propriety_of_parametric_approximation}, we show that if this evaluation procedure
 had been specified prospectively, the resulting score would be proper, \textemdash but that a post hoc application of this
@@
 allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table
-\ref{tab:multi-k-scores}). The JHUAPL-Bucky model had the second best MWIS but the third worst allocation score.
+\ref{tab:multi-k-scores}). The \texttt{JHUAPL-Bucky} model had the second best MWIS but the third worst allocation score.
 
@@
 
-A state-wise comparison using Figures \ref{fig:thermometer-plot} and \ref{fig:multi-loc-ranks}A of the JHUAPL-Bucky and
-CU-select models shows that while the JHUAPL-Bucky forecast distributions were more centered on the eventual
+A state-wise comparison using Figures \ref{fig:thermometer-plot} and \ref{fig:multi-loc-ranks}A of the \texttt{JHUAPL-Bucky} and
+\texttt{CU-select} models shows that while the \texttt{JHUAPL-Bucky} forecast distributions were more centered on the eventual
 observations in many states, the suggested allocations of the JHUAPL-Bucky model were less efficient than those of the
-CU-select model. Thus JHUAPL-Bucky achieved a worse allocation score than CU-select (on this particular week) by
+\texttt{CU-select} model. Thus \texttt{JHUAPL-Bucky} achieved a worse allocation score than CU-select (on this particular week) by
 allocating excess resources to several states, such as Pennsylvania and Michigan, rather than to states such as Florida
-and California where these resources would have reduced unmet need.  These allocation errors of JHUAPL-Bucky resulted
-from forecasts failing to consistently capture the relative resource needs across different states. The CU-select model
+and California where these resources would have reduced unmet need.  These allocation errors of \texttt{JHUAPL-Bucky} resulted
+from forecasts failing to consistently capture the relative resource needs across different states. The \texttt{CU-select} model
 made some similar errors — most prominently, over-allocating resources to Ohio — but overall, it more successfully
@@
 
-On the other hand, CU-select had worse performance as measured by MWIS. Its forecasts were biased downwards, and it
-consistently incurred a large penalty for underprediction (Figure \ref{fig:multi-loc-ranks}B). The JHUAPL-Bucky model
+On the other hand, \texttt{CU-select} had worse performance as measured by MWIS. Its forecasts were biased downwards, and it
+consistently incurred a large penalty for underprediction (Figure \ref{fig:multi-loc-ranks}B). The \texttt{JHUAPL-Bucky} model
 had wider predictive intervals which more often included the observed level of daily hospital admissions. Its MWIS was
@@
 
-<<multi-loc-ranks, fig.width=8, fig.cap="Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select).
+<<multi-loc-ranks, fig.width=8, fig.cap="Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (\\texttt{JHUAPL-Bucky} and \\texttt{CU-select}).
 Panel A shows the observed resource need, in this case the observed number of hospitalizations, for each state, along with the hypothetical number of resources allocated to the given location based on the forecasts from each model.
@@
 AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down).
-In Panel B, the per-capita allocation is drawn in a heavier solid line.
+In Panel B, the \\texttt{per-capita} allocation is drawn in a heavier solid line.
 Panel C shows the MWIS metric across weeks, averaged across all states.
@@
 Averaging across all weeks, the ensemble forecast achieved a better mean allocation score (MAS) than two benchmark
-methods. The COVIDhub-ensemble had the best MAS across all weeks, the per-capita allocation had the second best MAS, and
-the COVIDhub-baseline model had the sixth best allocation (Table \ref{tab:multi-week-performance-summary}). No
+methods. The \texttt{COVIDhub-ensemble} had the best MAS across all weeks, the \texttt{per-capita} allocation had the second best MAS, and
+the \texttt{COVIDhub-baseline} model had the sixth best allocation (Table \ref{tab:multi-week-performance-summary}). No
 individual model had a better MAS than the per-capita allocation.
@@
 according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
-(This comparison excludes the per-capita allocation which does not use forecasts and therefore cannot be assigned a MWIS.)
-Notably, the JHUAPL-SLPHospEns model ranked best on taMWIS but had only the fourth best MAS.
+(This comparison excludes the \texttt{per-capita} allocation which does not use forecasts and therefore cannot be assigned a MWIS.)
+Notably, the \texttt{JHUAPL-SLPHospEns} model ranked best on taMWIS but had only the fourth best MAS.
 
@@
 These results show the average performance across time for the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, as well
-  as for a per-capita allocation. MWIS is not defined for the per-capita allocation which does not use forecasts.",
+  as for a per-capita allocation. MWIS is not defined for the \\texttt{per-capita} allocation which does not use forecasts.",
                  digits = 0,
@@
 \ref{fig:metrics-correlation}. We find, for example, no clear association between the consistent
-and high MWIS rank of the JHUAPL-SLPHospEns model and its highly variable AS rank. This contrasts 
+and high MWIS rank of the \texttt{JHUAPL-SLPHospEns} model and its highly variable AS rank. This contrasts 
 with a clearly positive correlation between MWIS and AS ranks in other models such as 
-Karlen-pypm and USC-SI\_kJalpha.
-We also can see in more detail the consistently high performance COVIDhub-ensemble on both scores.
-And while it did not submit forecasts for enough weeks to be included in Table \ref{tab:multi-week-performance-summary}, CMU-TimeSeries is an interesting example of a model that 
+\texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha}.
+We also can see in more detail the consistently high performance \texttt{COVIDhub-ensemble} on both scores.
+And while it did not submit forecasts for enough weeks to be included in Table \ref{tab:multi-week-performance-summary}, \texttt{CMU-TimeSeries} is an interesting example of a model that 
 performed consistently well on AS but had only middling MWIS ranks.
@@
 AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of
-hospitalizations on January 3, 2021 as well as for the per-capita allocation (Figure \ref{fig:multi-k}A). These
+hospitalizations on January 3, 2021 as well as for the \texttt{per-capita} allocation (Figure \ref{fig:multi-k}A). These
 calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of
@@
 evaluated with scoring rules for which they were not elicited or via parameters imputed by a hub after submission, both
-of which are done in our application example.  We consider these issues briefly in supplement sections
+of which are done in our application example.  We consider these issues briefly in appendices
 \ref{sec:distfromq} and \ref{sec:propriety_of_parametric_approximation}.)
@@
 
+\appendix
+\appendixpage
+\addappheadtotoc
+
+\section{Introduction}
+\label{sec:intro}
+
+We address some technical and methodological points from the main text.
+We begin in section \ref{sec:proper} by defining proper scoring rules and showing that the allocation score is proper.
+In section \ref{sec:ex-shortage}, we formalize the concept of a \emph{shortage} of resources, give some key results about expected resource shortages under a distribution characterizing uncertainty about (future) levels of resource need, and illustrate how quantiles arise as the Bayes act in a decision making problem about the quantity of a resource to purchase.
+Section \ref{sec:bayes-quantiles} gives a justification for the result that the Bayes act for the allocation problem is given by a vector of quantiles of the forecast distributions in each location at a shared probability level, which was stated in section \ref{sec:methods.detailed.specific_allocation} of the main text.
+We describe the algorithm that we use to compute allocations given a forecast distribution in each location in section \ref{sec:numeric}.
+In section \ref{sec:distfromq}, we describe the methods for approximating a full distribution from a set of quantiles, implemented in the R package \verb`distfromq`, that we used to support computation of allocation scores from quantile forecasts that were submitted to the US COVID-19 Forecast Hub for the application in the main text.
+Finally, in section \ref{sec:propriety_of_parametric_approximation} we examine implications for propriety of an analysis that uses summaries of forecasts (such as predictive quantiles) rather than the full forecast distributions for computation of allocation scores.
+
+\section{Proper scoring rules}
+\label{sec:proper}
+
+In decision theory, a loss function $l$ is used to formalize a decision problem by assigning numerical value $l(x,y)$ to the
+\emph{result} of taking an \emph{action} $x$ in preparation for an \emph{outcome} $y$. A \emph{scoring rule} $S$ is a
+loss function for a decision problem where the action is a probabilistic forecast $F$ of the outcome $y$ (or the statement of $F$ by a forecaster).
+% As does any loss fun $S$ transforms a random outcome variable $Y$ into a random loss $S(F,Y)$.
+We refer to the realized loss $S(F,y)$ as the \emph{score} of $F$ at $y$.
+
+Probabilistic forecasts can be seen as a unique kind of action in that they can be used to generate their
+own (simulated) outcome data, against which they can be scored using $S$. A probabilistic forecast
+$F$ is thus committed to the ``self-assessment'' $\Ex_F [S(F, Y)] := \Ex [S(F, Y^F)]$, where $Y^F \sim F$ is the random variable defined
+by sampling from $F$, as well to an assessment $\Ex_F [S(G, Y)]$ of any alternative forecast $G$.
+
+A natural consistency criterion for $S$ is that, for observations assumed to be drawn from $F$, it will not assess any other forecast $G$
+as being better than $F$ itself, that is, that
+\begin{align}
+\Ex_F [S(F, Y)] \leq \Ex_F [S(G, Y)] \label{eqn:prop_ineq}
+\end{align}
+for any $F,G$. A scoring rule meeting this criterion is called \emph{proper}. If $S$ were improper, then from the perspective of
+a forecaster focussed (solely) on expected loss minimization, the decision to state
+a forecast $G$ other than the forecast $F$ which they believe describes $Y$ could be superior to the decision to state $F$.
+$S$ is \emph{strictly proper} when
+\eqref{eqn:prop_ineq} is a strict inequality, in which case the
+\emph{only} optimal decision for a forecaster seeking to minimize their expected loss is to state the forecast they believe to be true.
+
+\subsection{The allocation score is proper}
+\label{sec:alloscore_proper}
+
+Our primary decision theoretical procedure, outlined in section \ref{sec:methods.detailed.decisiontheory} of the main text,
+uses a decision problem with loss function $s(x,y)$ to define a scoring rule
+\begin{align}
+S(F,y) := s(x^F,y) \label{eqn:bayes_sr}
+\end{align}
+where $x^F := \argmin_{x} \Ex_F[s(x,Y)]$ is the Bayes act for $F$ with respect to $s$.
+Such scoring rules, which we call \emph{Bayes scoring rules},
+are proper by construction since
+\begin{align}
+\Ex_F [S(F, Y)] &= \Ex_F [ s(x^F, Y) ] \nonumber \\
+ &= \mathrm{min}_{x} \Ex_F [ s(x, Y) ] \quad \text{ (by definition of $x^F$)} \\
+ &\leq \Ex_F [ s(x^G, Y) ] \label{eqn:dt_proper_key} \\
+ &= \Ex_F [ S(G, Y)]. \nonumber
+\end{align}
+
+The allocation scoring rule is Bayes and therefore proper.
+
+We note that in the probabilistic forecasting literature (see e.g., \cite{gneiting2011making}, Theorem 3) what we have
+termed Bayes scoring rules typically appear via \eqref{eqn:bayes_sr} where $x^F$ is some given functional of $F$ which
+can be shown to be \emph{elicitable}, that is, to be the Bayes act for some loss function $s$.
+Such a loss function is said to be a \emph{consistent loss (or scoring) function} for the functional $F \mapsto x^F$, and many important
+recent results in the literature (e.g., \cite{fisslerziegel2016consistency}) address whether there \emph{exists} any loss
+function that is consistent for $x^F$. Our orientation
+is different from this insofar as we \emph{begin} by specifying a decision problem and a loss function of subject matter relevance
+and use the Bayes act only as a bridge to a proper scoring rule.  Consistency is never in doubt.
+
+\section{Expected shortages}
+\label{sec:ex-shortage}
+
+A key feature of loss functions for decision problems used to define quantiles and related scoring rules such as the
+CRPS and the WIS (see e.g., \cite{gneiting2011quantiles}, \cite{jose2009evaluating},
+and \cite{royset2022optimization}, sections 1.C and 3.C), as well as the allocation loss function presented in this work,
+is the presence of a \emph{shortage}:
+the amount $\max\{0,y-x\}$ by which a resource demand
+$y$ exceeds a supply decision variable $x$, which, for convenience, we write as $(y-x)_{+}$. In particular, a quantile at
+probability level $\alpha$ of a distribution $F$ on $\mathbb{R}^1$ (which we assume to have a well-defined density $f(x)$)
+is a Bayes act for the loss function
+\[
+l(x,y) = Cx + L(y-x)_{+}
+\]
+where $\alpha = 1-C/L$ and $C$ and $L$ can be interpreted as the cost per unit of a resource (such as medicine) and the loss
+incurred when a unit of demand (such as illness) cannot be met due to the shortage $(y-x)_{+}$.  This follows because a
+Bayes act, as a minimizer of $\Ex_F[l(x,Y)]$, must also be a vanishing point of the derivative
+\begin{align}
+\dby{}{x} \Ex_F\left[l(x,Y)\right] &= \Ex_F\left[\dby{}{x}l(x,Y)\right] \nonumber\\
+&= C + L\Ex_F\left[\dby{}{x}(Y-x)_+\right] \nonumber\\
+&= C - L\Ex_F\left[\mathbf{1}\{Y > x\}\right] \nonumber\\
+&= C + L(F(x) - 1), \label{eqn:q_deriv}
+\end{align}
+so that $1-C/L = F(x)$.
+The formula $\dby{}{x}\Ex_F\left[(Y-x)_+\right] = F(x) - 1$ for the derivative of the shortage,
+used above in \eqref{eqn:q_deriv}, can be obtained from an application of the ``Leibniz Rule'':
+\begin{align}
+	\frac{d}{dx} \Ex_F [(Y-x)_{+}] &= \frac{d}{dx} \int_{x}^{\infty} (y-x) f_Y(y)dy \nonumber\\
+	&= \int_{x}^{\infty} \frac{d}{dx}(y-x) f_Y(y)dy - (x-x) f_Y(x) = -\int_{x}^{\infty} f_Y(y)dy = F(x)-1. \label{eqn:shortage_deriv}
+\end{align}
+Note that more care is required when $F$ does not have a density.
+We will also use this result below in deriving the Bayes act for the allocation loss.
+
+\section{Allocation Bayes acts as vectors of marginal quantiles.}
+\label{sec:bayes-quantiles}
+
+Here we study the form of the Bayes act for the allocation problem (AP) (equation \eqref{eqn:loss_fn} in section \ref{sec:methods.detailed.specific_allocation}) of the text:
+\begin{align}
+    \underset{0 \leq x}{\mathrm{minimize}}\,\, \mathbb{E}_{F} [s_A(x, Y)]= \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[(Y_i - x_i)_{+}]
+     \text{ subject to }
+     \, \sum_{i=1}^N x_i = K, \label{AP-long}
+\end{align}
+where the marginal forecasts $F_i$ for $i=1,\dots,N$ represent forecasts for $N$ distinct locations.
+We show that the Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$ and resource constraint level $K$
+is a vector of quantiles of the marginal forecast distributions $F_i$ at a single probability level
+$\tau^{F,K}$, that is, $x_i^{F,K} = q_{F_i,\tau^{F,K}}$. An immediate consequence used in the examples in Section \ref
+{sec:methods.overview} in the main text is that if $F_i = \mathrm{Exp}(1/\sigma_i)$ for all $i$, then the Bayes act is
+proportional to $(\sigma_1,\ldots,\sigma_N)$, since $q_{\mathrm{Exp}(1/\sigma),\tau} = -\sigma \log(1-\tau)$.
+
+In order for $x^{\star} \in \mathbb{R}^N_{+}$ to solve the AP it must be true that reallocating $\delta > 0$ units of the resource from location $i$ to location $j$ will lead to a net increase in expected shortage --- in other words, the reallocation increases the expected shortage in location $i$ is at least as much as it decreases the expected shortage in location $j$:
+\begin{align*}
+&\mathbb{E}_{F_i}[(Y_i - (x^{\star}_i - \delta))_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}]
+\text{ (increase in $i$) } \nonumber \\
+&\qquad \geq
+\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - (x^{\star}_j + \delta))_{+}]
+\text{ (decrease in $j$) }.
+\end{align*}
+
+Dividing by $\delta$ and letting $\delta \searrow 0$, this implies from
+\eqref{eqn:shortage_deriv} that
+\begin{align}
+1-F_i(x^{\star}_i) &= -\frac{d}{dx_i}\mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}] \nonumber \\
+&= \lim_{\delta \searrow 0} \frac{1}{\delta}
+\left\{\mathbb{E}_{F_i}[(Y_i - (x^{\star}_i - \delta))_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}]\right\}
+\text{ (increase in $i$) } \nonumber \\
+&\geq
+\lim_{\delta \searrow 0} \frac{1}{\delta}
+\left\{\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - (x^{\star}_j + \delta))_{+}]\right\}
+\text{ (decrease in $j$) } \nonumber \\
+ &= -\dby{}{x_j}\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] = 1-F_j(x^{\star}_j) \label{eqn:ASoptimal1}
+\end{align}
+
+Note that negative derivatives appear because our optimality condition addresses how a \emph{decrease} in resources will
+\emph{increase} the expected shortage in $i$ and vice versa in $j$. Since \eqref{eqn:ASoptimal1} also holds with $i$ and $j$
+reversed, a number $\lambda$ (a \emph{Lagrange multiplier}) exists such that
+$L(1-F_k(x^{\star}_k)) = \lambda$ for all $k \in 1,\ldots,N$.
+(We scale by $L$ to facilitate possible future interpretations of $\lambda$ in terms of the partial derivatives
+of $\mathbb{E}_{F} [s_A(x, Y)]$.)
+That is, $x^{\star}_k$ is a quantile $q_{\tau,F_k}$ for
+$\tau = 1 - \lambda/L$. The value of $\tau$ is then determined by the constraint equation
+\begin{align}
+\sum_{i=1}^N q_{\tau,F_i} = K. \label{eqn:quantiles-sum-to-K}
+\end{align}
+It is important to note that $\tau$ depends on $F$ and $K$ and is \emph{not} a fixed parameter
+of the allocation scoring rule.
+
+\section{Numerical computation of allocation Bayes acts}
+\label{sec:numeric}
+
+To compute an allocation score $S_A(F,y;K) := s_A(x^{F,K},y)$, we require numerical values for a
+Bayes act solving the AP \eqref{AP-long} --- that is, we must find the specific resource allocations for each location that are determined by the forecast $F$ under the resource constraint $K$.
+Assuming we have reliable means of calculating quantiles $q_{\alpha,F_i}$
+of the marginal forecasts $F_i$,
+these allocations are given by $q_{\tau^{\star},F_i}$ where $\tau^{\star}$ solves the equation \eqref{eqn:quantiles-sum-to-K}.
+However, this equation is not analytically tractable and we must resort to a numerical method
+for finding an approximation $\tilde{\tau}$ of $\tau^{\star}$.
+
+We have implemented an iterative bisection method that makes use of the fact that $\sum_{i=1}^N q_{\tau,F_i}$
+is an increasing function of $\tau$.
+The algorithm begins with an initial search interval $[\tau_{L,1}, \tau_{U,1}]$ (such as $[0,\max_{i}F_i(K)]$) that clearly contains
+the solution $\tau^{\star}$.
+% This can be confirmed by verifying that the resource allocations corresponding to the lower endpoint of the search interval do not exceed the resource constraint ($\sum_{i=1}^N q_{\tau_L,F_i} \leq K$) and the resource allocations corresponding to the upper endpoint match or exceed the available resources ($\sum_{i=1}^N q_{\tau_U,F_i} \geq K$).
+At each step $j$ of the algorithm, we evaluate the total allocation $\sum_{i=1}^N q_{\tau_{M,j},F_i}$ at the midpoint of the search interval,
+$\tau_{M,j} = \frac{1}{2}(\tau_{L,j} + \tau_{U,j})$ and continue the search on the narrowed sub-interval
+\begin{align}
+[\tau_{L,j+1},\tau_{U,j+1}] =
+\begin{cases}
+[\tau_{L,j}, \tau_{M,j}] & \text{if } \sum_{i=1}^N q_{\tau_{M,j},F_i} \geq K \\
+[\tau_{M,j}, \tau_{U,j}] & \text{if } \sum_{i=1}^N q_{\tau_{M,j},F_i} < K.
+\end{cases} \nonumber
+\end{align}
+% We then narrow the search to the sub-interval that contains the solution:
+% \begin{itemize}
+%   \item If $\sum_{i=1}^N q_{\tau_{M},F_i} \geq K$, we continue the search on the interval $[\tau_L, \tau_M]$.
+%   \item If $\sum_{i=1}^N q_{\tau_M, F_i} \leq K$, we continue the search on the interval $[\tau_M, \tau_U]$.
+% \end{itemize}
+This search continues until $\tau_{U,j+1} < (1+\varepsilon)\tau_{L,j+1}$ for a suitably small $\varepsilon>0$.
+We have implemented this procedure along with the resulting score computations in the
+R package \verb`alloscore` \citep{gerding-alloscore} which provided all allocation score values used in the analysis of section \ref{sec:application}
+in the main text.
+
+Subtleties can arise when the forecast
+densities $f_i$ vanish or are very small, in which case quantiles are non-unique or highly variable near a probability level,
+leading to ambiguity or numerical instabilities{} in the evaluation of $\sum_{i=1}^N q_{\tau,F_i}$. Additionally, if point masses are present in any of the $F_i$,
+\eqref{eqn:quantiles-sum-to-K} will not have a unique solution for some discrete set of constraint levels $K$.
+We have adopted conventions for detecting such levels and enforcing consistency in score calculations
+near them.
+Through extensive experimentation, we have determined that these conditions seem to address these challenges with the forecasts we are working with, but we leave a more rigorous approximation error analysis for later work.
+
+\section{Computing allocations from finite quantile forecast representations}
+\label{sec:distfromq}
+
+In section \ref{sec:application} of the manuscript, we used the allocation score to evaluate forecasts of COVID-19 hospitalizations that have been submitted to the US COVID-19 Forecast Hub.
+These forecasts are submitted to the Hub using a set of 23 quantiles of the forecast distribution at the 23 probability levels in the set $\mathcal{T} = \{0.01, 0.025, 0.05, 0.1, 0.15, \ldots, \allowbreak 0.9, 0.95, 0.975, 0.99\}$, which specify a predictive median and the endpoints of central $(1 - \alpha) \times 100\%$ prediction intervals at levels $\alpha = 0.02, 0.05, 0.1, 0.2, \allowbreak 0.3, 0.4, 0.5, 0.6, \allowbreak 0.7, 0.8, 0.9$.
+For a given week and target date, we use $q_{i,k}$ to denote the submitted quantiles for location $i$ and probability level $\tau_k \in \mathcal{T}$, $k = 1, \ldots, 23$.
+
+In the event that there is some $k \in \{1, \ldots, 23\}$ for which $\sum_i q_{i,k} = K$, i.e., the provided predictive quantiles at level $\tau_k$ sum across locations to the resource constraint $K$, the solution to the allocation problem is given by those quantiles.
+However, generally this will not be the case; the optimal allocation will typically be at some probability level $\tau^\star \notin \mathcal{T}$.
+
+To address this situation and support the numerical allocation algorithm outlined in section \ref{sec:numeric}, we need a mechanism to approximate the full cumulative distribution functions $F_i$, $i = 1, \ldots, N$ based on the provided quantiles.
+We have developed functionality for this purpose in the \verb`distfromq` package for R.\citep{ray-distfromq} This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:
+\begin{enumerate}
+  \item Identify a discrete component of the distribution consisting of zero or more point masses, and create an adjusted set of predictive quantiles for the continuous part of the distribution by subtracting the point mass probabilities and rescaling.
+  \item For the continuous part of the distribution, different approaches are used on the interior and exterior of the provided quantiles:
+  \begin{enumerate}
+    \item On the interior, a monotonic cubic spline interpolates the adjusted quantiles representing the continuous part of the distribution.
+    \item A location-scale parametric family is used to extrapolate beyond the provided quantiles.
+The location and scale parameters are estimated separately for the lower and upper tails so as to obtain a tail distribution that matches the two most extreme quantiles in each tail.
+In this work, we use normal distributions for the tails.
+  \end{enumerate}
+\end{enumerate}
+The resulting distributional estimate exactly matches all of the predictive quantiles provided by the forecaster.
+We use the cumulative distribution function resulting from this procedure as an input to the allocation score algorithm.
+
+We refer the reader to the \verb`distfromq` documentation for further detail.\citep{ray-distfromq}
+
+% \begin{enumerate}
+%   \item First, we handle settings where the provided quantiles imply the existence of one or more point masses.
+Specifically, if there are distinct probability levels $\tau_k < \tau_{k^\prime}$ that have the same quantiles, $q_{l,k} = q_{l, k^\prime}$, the distribution contains a point mass at $q_{l,k}$.
+We split into three cases depending on the number of distinct provided quantiles:
+%   \begin{enumerate}
+%     \item If all provided quantiles are equal to each other, we infer that the distribution consists of a single point mass at that quantile value.
+%     \item If there are two distinct quantiles across all of the probability levels $\tau_k \in \mathcal{T}$, our estimated distribution consists of a mixture of two point masses.
+The probability assigned to the point mass at $q$ is proportional to $\tau_q^\max - \tau_q^\min$, where
+%     $$\tau_q^\max = \begin{cases}
+%       \max_{\{\}} \tau_k
+%     \end{cases}$$ is the largest probability level $\tau_k$ for which $q_{l,k} = q$, and $\tau_q^\min$ is defined similarly.
+%     \item Otherwise, our final estimate will consist of a mixture of a discrete distribution with point masses at the duplicated quantiles and a continuous distribution elsewhere.
+In this step, we  In this case, the discrete distribution has point mass probabilities calculated as $d_q = \tau_q^\max - \tau_q^\min$, where $\tau_q^\max$ and $\tau_q^\min$ are defined as in the previous point.
+%   \end{enumerate}
+% \end{enumerate}
+
+\section{Propriety of parametric approximation} % (fold)
+\label{sec:propriety_of_parametric_approximation}
+
+In practice, open forecasting exercises are generally not able to collect a perfect description of the forecast distribution $F$ other than in simple settings such as for a categorical variable with a relatively small number of categories.
+In settings where the outcome being forecasted is a continuous quantity (such as the proportion of outpatient doctor visits where the patient has influenza-like illness) or a count (such as influenza hospitalizations), forecasting exercises have therefore resorted to collecting summaries of a forecast distribution such as bin probabilities or predictive quantiles.
+In this section, we address two practical concerns raised by this.
+First, we discuss conditions under which it is possible to calculate the allocation score when only summaries of a forecast distribution are recorded in a submission to a forecast hub.
+Second, we show that a post hoc attempt to compute the allocation score based on submitted predictive quantiles may in fact compute an alternative score that is not proper.
+
+\subsection{Propriety when scoring methods are announced prospectively}
+
+We consider a setting where a forecasting exercise (such as a forecast hub) pre-specifies that forecasts will be represented using a parametric family of forecast distributions $G_\theta(y)$, and the task of the forecaster is to select a particular parameter value $\theta$.
+We use $\mathcal{P}$ to denote the collection of all distributions $G_\theta$ in the given parametric family.
+For instance, it has recently been proposed that mixture distributions could be used to represent forecast distributions \citep{wadsworth2023mixture}.
+Additionally, we note that the functionality in \verb`distfromq` can be viewed as specifying a parametric family $\mathcal{P}_{\mathrm{dfq}}$ where the parameters $\theta$ of $G_\theta$
+are its quantiles at pre-specified probability levels, and where the shape of any $G_\theta \in \mathcal{P}_{\mathrm{dfq}}$ over the full range of its support is entirely controlled by these quantiles.
+
+We find it helpful now to formally distinguish between two decision making problems.
+The first is the public health decision maker's allocation problem where the task is to select an allocation $x$, with the allocation loss $s_A(x, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i)$ as described in section \ref{sec:methods.detailed.specific_allocation} of the main text.
+The second is the forecaster's reporting problem where the task is to select parameter values $\theta$ to report. The forecaster's loss is given by
+\begin{align}
+s_R(\theta, y) = s_A(x^{G_\theta}, y), \label{eqn:forecaster_theta_loss}
+\end{align}
+where $x^{G_\theta}$ is the Bayes act for the allocation problem under the distribution $G_\theta$.
+In words, the loss associated with reporting $\theta$ is equal to the loss associated with taking the Bayes allocation corresponding to the distribution $G_\theta$.
+
+Following our usual construction, the Bayes act for the forecast reporting problem is the parameter set that minimizes the forecaster's expected loss.
+Breaking with our earlier notation for improved legibility, we use $\theta^\star(F)$ to denote this Bayes act:
+\begin{align*}
+\theta^\star(F) &= \text{argmin}_\theta \Ex_F [s_R(\theta, Y)] \\
+&= \text{argmin}_\theta \Ex_F [s_A(x^{G_\theta}, Y)]
+\end{align*}
+
+We then arrive at the scoring rule
+$$S_R(F, y) = s_R(\theta^\star(F), y) = s_A(x^{G_{\theta^\star(F)}}, y).$$
+It follows from the discussion in section \ref{sec:alloscore_proper} that this is a proper scoring rule for $F$.
+Although the full forecast distribution $F$ is not available in the forecast submission, the score $S_R(F, y)$ can be calculated from the reported parameter values as long as the forecaster submits the optimal parameters $\theta^\star(F)$.
+
+We emphasize that the forecaster's true predictive distribution $F$ does not need to be a member of the specified parametric family $\mathcal{P}$ for this construction to yield a proper score.
+It is, however, necessary to specify the parametric family to use and the foundational scoring rule $s_A$ (including any relevant problem parameters such as the resource constraint $K$) in advance, so that forecasters can identify the Bayes act parameter set $\theta^\star(F)$ to report.
+
+If the parametric family used to represent forecast distributions is flexible enough, the reporting scoring rule $S_R$ and the allocation score are equivalent in the sense that they will yield the same score for any distribution $F$.
+Suppose that for a given resource constraint $K$, for any forecast distribution $F$ it is possible to find a member $G_{\theta^\star}$ of the specified parametric family $\mathcal{P}$ with the same allocation as $F$ (i.e., $x^F = x^{G_{\theta^\star}}$).
+Then $\theta^\star$ is a Bayes act for the reporting problem since for any other parameter value $\theta$,
+\begin{align*}
+\Ex_F[s_R(\theta^\star, Y)] &= \Ex_F[ s_A(x^{G_{\theta^\star}}, Y) ] \\
+&= \Ex_F[s_A(x^F, Y)]  \quad \text{ (since $x^F = x^{G_{\theta^\star}}$)} \\
+&\leq \Ex_F[ s_A(x^{G_\theta}, Y) ]  \quad \text{ (by definition of $x^F$)} \\
+&= \Ex_F[ s_R(\theta, Y)].
+\end{align*}
+It therefore follows that
+\begin{align*}
+S_R(F, y) &= s_R(\theta^\star, y) \\
+&= s_A(x^{G_{\theta^\star}}, y) \\
+&= s_A(x^F, y) \\
+&= S_A(F, y).
+\end{align*}
+
+For the particular choice of the parametric family $\mathcal{P}_{\text{dfq}}$ (i.e., using the \verb`distfromq` package), this flexibility requirement is satisfied.
+For instance, the forecaster could pick one required quantile level (such as 0.5, for which the corresponding predictions are predictive medians), and set the submitted quantiles of their forecast distribution in each location at that level to be the desired allocations, which sum to $K$ across all locations.
+However, this representation of the forecast may be quite different from the actual forecast distribution $F$.
+For example, for the actual forecast distribution $F$ the allocations may occur at some quantile level other than 0.5.
+
+As another alternative for practical forecasting exercises, a forecast hub could ask forecasters to directly provide the Bayes allocations associated with their forecasts for one or more specified resource constraints $K$.
+At the cost of increasing the number of quantities solicited by the forecast hub, this would have several advantages: it would prevent any artificial distortion of the forecast distributions, allow for direct calculation of scores, and narrow the gap between model outputs and public health end users.
+For this to be feasible, implementations of the allocation algorithm would have to be provided to participating forecasters in the computational languages being used for modeling.
+
+\subsection{Impropriety of post hoc allocation scoring with quantile forecasts}
+
+A \emph{post hoc} evaluation of quantile forecasts that combines the parametric family specified by \verb`distfromq` with the allocation score does not yield the allocation score of the forecast distribution $F$.
+Instead, it computes an alternative score that is improper.
+This is because the forecast distribution $F$ and the distribution $G^q \in \mathcal{P}_{\text{dfq}}$ with the same quantiles as $F$ may determine different resource allocations.
+In our investigations, these discrepancies appear to be relatively minor on the interior of the provided quantiles, but could be severe if the tail extrapolations performed by \verb`distfromq` do not match the tail behavior of $F$ and the allocations are in the tails of the predictive distribution.
+
+We define
+$$G^{\star}(F) := \argmin_{G \in \mathcal{P}_{\mathrm{dfq}}} E_{F}[S_A(G, Y)].$$
+Since $S_R$ is defined as the Bayes scoring rule for the forecaster's loss \eqref{eqn:forecaster_theta_loss}, $G^{\star}(F)$ coincides
+with $G_{\theta^{\star}(F)}$, the distribution in $\mathcal{P}_{\mathrm{dfq}}$ given by the
+optimal submission parameters $\theta^{\star}(F)$ for the forecaster with predictive distribution $F$.
+In general, $G^q(F)$ and $G^\star(F)$ will be different distributions:  matching $F$ at specific quantiles
+does not require $G^q(F)$ to match $F$ at the quantiles for $\tau^{F,K}$ (c.f. \eqref{eqn:quantiles-sum-to-K}),
+which would be necessary for it to share $x^F$ as an optimal allocation.
+
+When an analyst attempts a post hoc computation of the allocation score using $G^q(F)$ (implicitly assuming that $G^q(F) = G^\star(F)$), they in fact compute the alternative score
+$$\tilde{S}(F, y) = S_A(G^q(F), y) = s_A(x^{G^q(F)}, y).$$
+This score is improper because $E_{F}[S_A(G, Y)]$ is minimized by $G^\star(F)$, not $G^q(F)$.
+In general, we have
+\begin{align}
+E_{F}[\tilde{S}(G^\star(F), Y)] &\leq E_{F}[ S_A(G^q(F), Y) ] \label{eqn:tilde_s_improper} \\
+  &= E_{F}[\tilde{S}(F, Y)] \nonumber
+\end{align}
+However, the inequality in \eqref{eqn:tilde_s_improper} will typically be strict.
+For example, if $F$ has heavy upper tails (such as for a lognormal distribution), but normal distributions are used for tail extrapolations in \verb`distfromq`, then the resource allocations based on the distribution $G^q(F)$ may be quite different from the optimal allocations under the distribution $F$, leading to a strict inequality.
+This demonstrates that $\tilde{S}$ is improper.
+
 \bibliography{allocation}

### 2024-02-17 alloscore_manuscript/alloscore-application-shorter.tex
convert supplement to appendices

@@
 allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table
-\ref{tab:multi-k-scores}). The JHUAPL-Bucky model had the second best MWIS but the third worst allocation score.
+\ref{tab:multi-k-scores}). The \texttt{JHUAPL-Bucky} model had the second best MWIS but the third worst allocation score.
 
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Feb 16 20:39:10 2024
+% Sat Feb 17 02:22:59 2024
 \begin{table}[ht]
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/multi-loc-ranks-1} \caption[Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select)]{Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select).
+\includegraphics[width=\maxwidth]{figure/multi-loc-ranks-1} \caption[Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (\texttt{JHUAPL-Bucky} and \texttt{CU-select})]{Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (\texttt{JHUAPL-Bucky} and \texttt{CU-select}).
 Panel A shows the observed resource need, in this case the observed number of hospitalizations, for each state, along with the hypothetical number of resources allocated to the given location based on the forecasts from each model.
@@
 AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down).
-In Panel B, the per-capita allocation is drawn in a heavier solid line.
+In Panel B, the \texttt{per-capita} allocation is drawn in a heavier solid line.
 Panel C shows the MWIS metric across weeks, averaged across all states.
@@
 according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
-(This comparison excludes the per-capita allocation which does not use forecasts and therefore cannot be assigned a MWIS.)
-Notably, the JHUAPL-SLPHospEns model ranked best on taMWIS but had only the fourth best MAS.
+(This comparison excludes the \texttt{per-capita} allocation which does not use forecasts and therefore cannot be assigned a MWIS.)
+Notably, the \texttt{JHUAPL-SLPHospEns} model ranked best on taMWIS but had only the fourth best MAS.
 
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Feb 16 20:39:15 2024
+% Sat Feb 17 02:23:03 2024
 \begin{table}[ht]
@@
 These results show the average performance across time for the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, as well
-  as for a per-capita allocation. MWIS is not defined for the per-capita allocation which does not use forecasts.} 
+  as for a per-capita allocation. MWIS is not defined for the \texttt{per-capita} allocation which does not use forecasts.} 
 \label{tab:multi-week-performance-summary}
@@
 
-In practice, epidemiological forecasts are often directed at a diverse set of end-users. It may be easy for some of these
-users to summarize the consequences of how they use a forecast with a numerical loss function, but for others, such
-as officials deciding how best to update public ``situational awareness'' of an outbreak, this may be essentially impossible.
-Even those forecast uses that are easily framed as expected loss minimization 
-may differ enough 
-that no single score
-would be appropriate for all users. Ideally, targeted forecasting tools could be developed through close collaboration
-between modelers and public health officials. However, this may only be possible in settings with sufficient staffing on
-both an analytics and a public health team. Increasingly, collaborative modeling hubs are being used to generate
-``one-size-fits-all'' forecasts for many locations at once. But it could still be beneficial in these settings to expand
-the set of scoring methods used by the hub to include scores, such as the AS, which are based on specific public health decision
-problems.  This could help end users better understand the value of available forecasts as inputs to their particular
-decision making contexts. (We note that issues of propriety are raised when forecasts are evaluated with scoring rules
-for which they were not elicited or via parameters imputed by a hub after submission, both of which are done in our
-application example.  We consider these issues briefly in supplement sections \ref{sec:distfromq} and
-\ref{sec:propriety_of_parametric_approximation}.)
+In practice, epidemiological forecasts are often directed at a diverse set of end-users. It may be easy for some of
+these users to summarize the consequences of how they use a forecast with a numerical loss function, but for others,
+such as officials deciding how best to update public ``situational awareness'' of an outbreak, this may be essentially
+impossible. Even those forecast uses that are easily framed as expected loss minimization may differ enough that no
+single score would be appropriate for all users. Ideally, targeted forecasting tools could be developed through close
+collaboration between modelers and public health officials. However, this may only be possible in settings with
+sufficient staffing on both an analytics and a public health team. Increasingly, collaborative modeling hubs are being
+used to generate ``one-size-fits-all'' forecasts for many locations at once. But it could still be beneficial in these
+settings to expand the set of scoring methods used by the hub to include scores, such as the AS, which are based on
+specific public health decision problems.  This could help end users better understand the value of available forecasts
+as inputs to their particular decision making contexts. (We note that issues of propriety are raised when forecasts are
+evaluated with scoring rules for which they were not elicited or via parameters imputed by a hub after submission, both
+of which are done in our application example.  We consider these issues briefly in appendices
+\ref{sec:distfromq} and \ref{sec:propriety_of_parametric_approximation}.)
 

### 2024-02-17 alloscore_manuscript/alloscore-application-shorter.Rnw
get more model names into monspace

@@
   select(Model, AS, MWIS, `IAS centered at 15k`, `IAS uniform`) |>
+  mutate(Model = gsub("_", "\\\\_", Model)) |>
+  mutate(Model = sprintf("\\texttt{%s}", Model)) |>
   arrange(AS) |>
@@
                  label = "tab:multi-k-scores") |>
-  print(include.rownames = FALSE, NA.string = "-")
+  print(include.rownames = FALSE, NA.string = "-", sanitize.text.function = function(x) {x})
 @
@@
 \texttt{CU-select} models shows that while the \texttt{JHUAPL-Bucky} forecast distributions were more centered on the eventual
-observations in many states, the suggested allocations of the JHUAPL-Bucky model were less efficient than those of the
-\texttt{CU-select} model. Thus \texttt{JHUAPL-Bucky} achieved a worse allocation score than CU-select (on this particular week) by
+observations in many states, the suggested allocations of the \texttt{JHUAPL-Bucky} model were less efficient than those of the
+\texttt{CU-select} model. Thus \texttt{JHUAPL-Bucky} achieved a worse allocation score than \texttt{CU-select} (on this particular week) by
 allocating excess resources to several states, such as Pennsylvania and Michigan, rather than to states such as Florida
@@
   filter(N > 12) |>
-  mutate(`taMWIS rank` = rank(taMWIS, na.last = "keep")) |>
+  mutate(`taMWIS rank` = rank(taMWIS, na.last = "keep"),
+         model = gsub("_", "\\\\_", model) %>% 
+           sprintf("\\texttt{%s}", .)) |>
   arrange(MAS) |>
@@
                  label = "tab:multi-week-performance-summary") |>
-  print(include.rownames = FALSE, NA.string = "-")
+  print(include.rownames = FALSE, NA.string = "-", sanitize.text.function = function(x) {x})
 @

### 2024-02-17 alloscore_manuscript/alloscore-application-shorter.tex
get more model names into monspace

@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Sat Feb 17 02:22:59 2024
+% Sat Feb 17 02:56:05 2024
 \begin{table}[ht]
@@
   \hline
-CU-select & 669 & 133 & 774 & 326 \\ 
-  per-capita & 865 & - & 1029 & 366 \\ 
-  COVIDhub-ensemble & 873 & 159 & 1067 & 438 \\ 
-  USC-SI\_kJalpha & 995 & 91 & 1216 & 1097 \\ 
-  JHUAPL-Gecko & 1034 & 164 & 1141 & 418 \\ 
-  MUNI-ARIMA & 1084 & 169 & 1248 & 440 \\ 
-  COVIDhub-trained\_ensemble & 1089 & 169 & 1271 & 823 \\ 
-  COVIDhub-baseline & 1175 & 170 & 1317 & 535 \\ 
-  JHUAPL-Bucky & 1358 & 102 & 1566 & 1214 \\ 
-  JHUAPL-SLPHospEns & 1540 & 129 & 1604 & 1102 \\ 
-  UVA-Ensemble & 2469 & 213 & 2635 & 2494 \\ 
+\texttt{CU-select} & 669 & 133 & 774 & 326 \\ 
+  \texttt{per-capita} & 865 & - & 1029 & 366 \\ 
+  \texttt{COVIDhub-ensemble} & 873 & 159 & 1067 & 438 \\ 
+  \texttt{USC-SI\_kJalpha} & 995 & 91 & 1216 & 1097 \\ 
+  \texttt{JHUAPL-Gecko} & 1034 & 164 & 1141 & 418 \\ 
+  \texttt{MUNI-ARIMA} & 1084 & 169 & 1248 & 440 \\ 
+  \texttt{COVIDhub-trained\_ensemble} & 1089 & 169 & 1271 & 823 \\ 
+  \texttt{COVIDhub-baseline} & 1175 & 170 & 1317 & 535 \\ 
+  \texttt{JHUAPL-Bucky} & 1358 & 102 & 1566 & 1214 \\ 
+  \texttt{JHUAPL-SLPHospEns} & 1540 & 129 & 1604 & 1102 \\ 
+  \texttt{UVA-Ensemble} & 2469 & 213 & 2635 & 2494 \\ 
    \hline
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Sat Feb 17 02:23:03 2024
+% Sat Feb 17 02:56:09 2024
 \begin{table}[ht]
@@
   \hline
-COVIDhub-ensemble & 389 & 70 & 2 \\ 
-  per-capita & 464 & - & - \\ 
-  COVIDhub-trained\_ensemble & 483 & 87 & 4 \\ 
-  CU-select & 502 & 81 & 3 \\ 
-  JHUAPL-SLPHospEns & 526 & 67 & 1 \\ 
-  COVIDhub-baseline & 594 & 93 & 5 \\ 
-  JHUAPL-Bucky & 643 & 112 & 7 \\ 
-  MUNI-ARIMA & 707 & 116 & 8 \\ 
-  JHUAPL-Gecko & 929 & 110 & 6 \\ 
-  USC-SI\_kJalpha & 1473 & 155 & 9 \\ 
+\texttt{COVIDhub-ensemble} & 389 & 70 & 2 \\ 
+  \texttt{per-capita} & 464 & - & - \\ 
+  \texttt{COVIDhub-trained\_ensemble} & 483 & 87 & 4 \\ 
+  \texttt{CU-select} & 502 & 81 & 3 \\ 
+  \texttt{JHUAPL-SLPHospEns} & 526 & 67 & 1 \\ 
+  \texttt{COVIDhub-baseline} & 594 & 93 & 5 \\ 
+  \texttt{JHUAPL-Bucky} & 643 & 112 & 7 \\ 
+  \texttt{MUNI-ARIMA} & 707 & 116 & 8 \\ 
+  \texttt{JHUAPL-Gecko} & 929 & 110 & 6 \\ 
+  \texttt{USC-SI\_kJalpha} & 1473 & 155 & 9 \\ 
    \hline

### 2024-02-21 alloscore_manuscript/alloscore-application-shorter.Rnw
finish discussion

@@
 
-Our example application of the AS refers to a hypothetical and unspecified resource need that is supposed to 
+Our example application of the AS refers to a hypothetical and unspecified resource need that is assumed to 
 pair with our forecasted outcome (new hospital admissions) in a manner that might appear overly simplistic and 
-artificially direct.
-A concrete decision problem, which motivated our work on allocation scoring, is
-
-allocation of ventilators during a respiratory viral pandemic
-\citep{huang_stockpiling_2017}.
-We used ventilators as a notional example for a resource allocation problem because it (a) has been identified as a
-scarce resource during outbreaks \citep{huang_stockpiling_2017}, (b) could be allocated across different locations, with
-logistical considerations depending on the geographical scales, and (c) would likely be correlated (although not
-perfectly so) with hospitalization data. However, this example has limitations in that not everyone who is hospitalized
-needs a ventilator, and new hospital admissions may not be the best stand in for a need that corresponds to a prevalence
+artificially direct.  We are optimistic, however, that by augmenting this scheme with a probabilistic measurement
+model for resource need and more detailed domain knowledge, it could be meaningfully aligned with a 
+more concrete decision problem such as the allocation of ventilators during a respiratory viral pandemic.  This problem 
+was explored from a stochastic optimization perspective (not from a forecast evaluation perspective) in 
+\cite{huang_stockpiling_2017} and provided an initial motivation for this project.
+We see it as promising focus for further development of the AS framework since it (a) has been identified as a prime
+manifestation of scarcity during respiratory outbreaks (b) tansparently involves
+logistical considerations depending on geographical and population scales, and (c) would likely imply a central role for 
+hospitalization data. But again, additional modeling and data synthesis would be required since not everyone who is hospitalized
+needs a ventilator, and new hospital admissions will not fully describe for a need that corresponds to a prevalence
 of severe cases. Other possible real-world examples include the allocation of a limited stockpile of vaccinations
 \citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests \citep{du_optimal_2022,pasco_covid-19_2023}.
-
-While
-there may be limitations in this specific pairing of a resource and forecasted outcome, we emphasize that the important
-result from this work is how evaluating forecasts using a fundamentally different kind of score can illuminate
-previously hidden aspects of value in forecasts. In this setting, the feature of a forecast that emerges as being
-important is that the probabilistic forecasts predict the relative magnitudes between locations accurately, even if the
-absolute magnitudes are inaccurate.
 
@@
 In section \ref{sec:ex-shortage}, we formalize the concept of a \emph{shortage} of resources, give some key results about expected resource shortages under a distribution characterizing uncertainty about (future) levels of resource need, and illustrate how quantiles arise as the Bayes act in a decision making problem about the quantity of a resource to purchase.
-Section \ref{sec:bayes-quantiles} gives a justification for the result that the Bayes act for the allocation problem is given by a vector of quantiles of the forecast distributions in each location at a shared probability level, which was stated in section \ref{sec:methods.detailed.specific_allocation} of the main text.
+Section \ref{sec:bayes-quantiles} gives a justification for the result that the Bayes act for the allocation problem is given by a vector of quantiles of the forecast distributions in each location at a shared probability level, which was stated in section
+\ref{sec:methods.detailed.specific_allocation}.
 We describe the algorithm that we use to compute allocations given a forecast distribution in each location in section \ref{sec:numeric}.
-In section \ref{sec:distfromq}, we describe the methods for approximating a full distribution from a set of quantiles, implemented in the R package \verb`distfromq`, that we used to support computation of allocation scores from quantile forecasts that were submitted to the US COVID-19 Forecast Hub for the application in the main text.
-Finally, in section \ref{sec:propriety_of_parametric_approximation} we examine implications for propriety of an analysis that uses summaries of forecasts (such as predictive quantiles) rather than the full forecast distributions for computation of allocation scores.
+In section \ref{sec:distfromq}, we describe the methods for approximating a full distribution from a set of quantiles, implemented in the R package \verb`distfromq`, that we used to support computation of allocation scores from quantile forecasts that were submitted to the US COVID-19 Forecast Hub for the application in section \ref{sec:application}.
+Finally, in section
+\ref{sec:propriety_of_parametric_approximation} we examine implications for propriety of an analysis that uses summaries of forecasts (such as predictive quantiles) rather than the full forecast distributions for computation of allocation scores.
 
@@
 
-Our primary decision theoretical procedure, outlined in section \ref{sec:methods.detailed.decisiontheory} of the main text,
+Our primary decision theoretical procedure, outlined in section \ref{sec:methods.detailed.decisiontheory},
 uses a decision problem with loss function $s(x,y)$ to define a scoring rule
@@
 $\tau^{F,K}$, that is, $x_i^{F,K} = q_{F_i,\tau^{F,K}}$. An immediate consequence used in the examples in Section \ref
-{sec:methods.overview} in the main text is that if $F_i = \mathrm{Exp}(1/\sigma_i)$ for all $i$, then the Bayes act is
+{sec:methods.overview} is that if $F_i = \mathrm{Exp}(1/\sigma_i)$ for all $i$, then the Bayes act is
 proportional to $(\sigma_1,\ldots,\sigma_N)$, since $q_{\mathrm{Exp}(1/\sigma),\tau} = -\sigma \log(1-\tau)$.
@@
 We have implemented this procedure along with the resulting score computations in the
-R package \verb`alloscore` \citep{gerding-alloscore} which provided all allocation score values used in the analysis of section \ref{sec:application}
-in the main text.
+R package \verb`alloscore` \citep{gerding-alloscore} which provided all allocation score values used in the analysis of section \ref{sec:application}.
 
@@
 We find it helpful now to formally distinguish between two decision making problems.
-The first is the public health decision maker's allocation problem where the task is to select an allocation $x$, with the allocation loss $s_A(x, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i)$ as described in section \ref{sec:methods.detailed.specific_allocation} of the main text.
+The first is the public health decision maker's allocation problem where the task is to select an allocation $x$, with the allocation loss $s_A(x, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i)$ as described in section \ref{sec:methods.detailed.specific_allocation}.
 The second is the forecaster's reporting problem where the task is to select parameter values $\theta$ to report. The forecaster's loss is given by

### 2024-02-21 alloscore_manuscript/alloscore-application-shorter.tex
finish discussion

@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Sat Feb 17 02:56:05 2024
+% Tue Feb 20 19:44:31 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Sat Feb 17 02:56:09 2024
+% Tue Feb 20 19:44:36 2024
 \begin{table}[ht]

### 2024-02-21 alloscore_manuscript/alloscore-application-shorter.Rnw
clean up Rnw

@@
 \end{align}
-%We emphasize that in Equation \eqref{eqn:informal_loss} the calculation of unmet need is based on the actual resource need that was realized in each location, while in Equation \eqref{eqn:informal_objective} the calculation of unmet need was based on the forecast distribution of future levels of resource need.
 We regard one allocation as better than another (with respect to the realized need) if it results in lower total 
@@
 Next suppose that we observe resource needs of 1 and 10 in locations 1 and 2, respectively. Based on these observed
-needs, we evaluate the allocation suggested by the forecast by calculating the amount of unmet need
-that resulted from that allocation over and above what was unavoidable given the resource constraint. With $K = 5$ 
-available resource units, the allocation based on the forecast exactly meets the observed need in location 1, but it leaves 6
-units of need unmet in location 2.
-But these 6 units of unmet need are unaviodable under the constraint $K=5$: allocating 0 units of resources to location 1 and 5 to location 2, for example, still results in a total unmet
-need of 6 across both locations. This gives an allocation score of 0 for the forecast when $K = 5$. On the other hand, when
-$K = 10$, the forecast's allocation results in $10 - 8 = 2$ units of unmet need in location 2 despite leaving no need
-unmet in location 1. In this case, the oracle would be able to prevent all but 1 of the total 11 units of need from
-going unmet, for example by allocating 1 unit of resources to location 1 and the remaining 9 units of resources to
-location 2. So for $K = 10$, the forecast gets an allocaton score of 1 (= 2 realized $-$ 1 unavoidable) in
-units of avoidable unmet need.
+needs, we evaluate the allocation suggested by the forecast by calculating the amount of unmet need that resulted from
+that allocation over and above what was unavoidable given the resource constraint. With $K = 5$ available resource
+units, the allocation based on the forecast exactly meets the observed need in location 1, but it leaves 6 units of need
+unmet in location 2. But these 6 units of unmet need are unaviodable under the constraint $K=5$: allocating 0 units of
+resources to location 1 and 5 to location 2, for example, still results in a total unmet need of 6 across both
+locations. This gives an allocation score of 0 for the forecast when $K = 5$. On the other hand, when $K = 10$, the
+forecast's allocation results in $10 - 8 = 2$ units of unmet need in location 2 despite leaving no need unmet in
+location 1. In this case, the oracle would be able to prevent all but 1 of the total 11 units of need from going unmet,
+for example by allocating 1 unit of resources to location 1 and the remaining 9 units of resources to location 2. So for
+$K = 10$, the forecast gets an allocaton score of 1 (= 2 realized $-$ 1 unavoidable) in units of avoidable unmet need.
 
@@
 is penalized regardless of alignments in other units.
-%Note that we do not claim that the allocation score is generically preferable to these other scores \textemdash rather, it provides a view of forecast performance that is specifically tuned to the context of decision making about resource allocations.
 
@@
 \label{sec:methods.detailed.decisiontheory}
-
-%In this section, we give an overview of the decision theoretic setup for developing proper scoring rules that measure
-%the value of a forecast as an input to decision making. We keep the discussion here at a somewhat informal level; we
-%refer the reader to [some subset of Brehmer and Gneiting; Grünwald and Dawid; Dawid; Granger and Pesaran 2000; Granger
-%and Machina 2006; Ehm et al. 2016] for more technically precise discussion.
 
@@
 
-In appendix \ref{sec:bayes-quantiles} we show that the components of the Bayes act are quantiles $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a
-probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource constraint $K$, but is shared across
-all locations. This probability level is the level at which the resource constraint is satisfied: $\sum_{i=1}^N
-F_i^{-1}(\tau^{F,K}) = K$. This tells us that in order to allocate optimally (according to $F$), we must divide
-resources among the locations so that there is an equal forecasted probability in every location that the allocation is
-sufficient to meet resource need. This solution to the allocation problem is well-known in inventory management and is
-often attributed to \cite{hadleywhitin1963}.
+In appendix \ref{sec:bayes-quantiles} we show that the components of the Bayes act are quantiles 
+$x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource
+constraint $K$, but is shared across all locations. This probability level is the level at which the resource constraint
+is satisfied: $\sum_{i=1}^N F_i^{-1}(\tau^{F,K}) = K$. This tells us that in order to allocate optimally (according to
+$F$), we must divide resources among the locations so that there is an equal forecasted probability in every location
+that the allocation is sufficient to meet resource need. This solution to the allocation problem is well-known in
+inventory management and is often attributed to \cite{hadleywhitin1963}.
 
@@
 submitted by all teams that met certain inclusion criteria and created ensemble forecasts using quantile averaging
-\citep{ray_comparing_2023}. Our analysis includes these ensembles (\texttt{COVIDhub-ensemble} and \texttt{COVIDhub-trained\_ensemble}) as
-well as one other ensemble of hub models created by another team (\texttt{JHUAPL-SLPHospEns}) and several other individual
-models. Models were eligible to be included in the analysis if they were designated as a ``primary'' model from a team.
-For a model to have a complete, eligible submission in a given week, it had to have a 14 day-ahead forecast for all 50
-states plus Washington DC. Models had to have a complete forecast for at least 4 of the 15 weeks in the analysis to be
-eligible for inclusion.
+\citep{ray_comparing_2023}. Our analysis includes these ensembles (\texttt{COVIDhub-ensemble} and
+\texttt{COVIDhub-trained\_ensemble}) as well as one other ensemble of hub models created by another team
+(\texttt{JHUAPL-SLPHospEns}) and several other individual models. Models were eligible to be included in the analysis if
+they were designated as a ``primary'' model from a team. For a model to have a complete, eligible submission in a given
+week, it had to have a 14 day-ahead forecast for all 50 states plus Washington DC. Models had to have a complete
+forecast for at least 4 of the 15 weeks in the analysis to be eligible for inclusion.
 
@@
 We evaluated forecasts using the allocation score (AS) and the weighted interval score (WIS), computed at a horizon of
-14 days. We chose to focus much of our analysis on the AS computed for a resource constraint of $K=15,000$. This level provides
-an anchor for interesting comparisons between phases of the Omicron wave by representing a national shortage at the peak and
-a national excess before and after the wave (see Figure \ref{fig:metrics-over-time}A), assuming that the resource need corresponds 
-directly to new hospital admissions.
-We additionally computed the mean
-WIS (MWIS) across all of the forecasted state-level locations.
+14 days. We chose to focus much of our analysis on the AS computed for a resource constraint of $K=15,000$. This level
+provides an anchor for interesting comparisons between phases of the Omicron wave by representing a national shortage at
+the peak and a national excess before and after the wave (see Figure \ref{fig:metrics-over-time}A), assuming that the
+resource need corresponds directly to new hospital admissions. We additionally computed the mean WIS (MWIS) across all
+of the forecasted state-level locations.
 
@@
 submissions.
-
-%We also computed a standardized rank for the allocation score for each model $m$ and week $w$.
-%First, we computed the number of models that forecasted that week ($n_w$) and the rank of model $m$ among the $n_w$ models ($r_{m,w}^{AS}$).
-%The model with the best allocation score received a rank of 1 and the worst received a rank of $n_w$.
-%In the case of a tie between one or more models, all models received the better rank.
-%We then rescaled these rankings to compute the allocation score standardized rank ($sr_{m,w}^AS$) between 0 and 1, where 0 corresponds to the worst rank and 1 to the best.
-
-%\begin{equation}
-%sr_{m,w}^{AS} = 1 - \frac{r^{AS}_{m,w}-1}{n_w-1}
-%\end{equation}
-
-%\subsubsection{Weighted Interval Score (WIS)}
-
-%The Weighted Interval Score (WIS), as described in earlier sections, measures the alignment of a single probabilistic forecast ($F$) with an observation ($y$).
-%We computed the mean WIS across all $L$ locations for each model and each forecasted week as
-%\begin{equation}
-%MWIS_{m,w} = \frac{1}{L}\sum_{l=1}^L WIS(F_{l,m,w},y)
-%\end{equation}
-%where $F_{l,m,w}$ is the probabilistic forecast from model $m$ for location $l$ and week $w$.
-%Using the same procedure as for allocation scores described above, we computed standardized ranks for MWIS ($sr_{m,w}^{MWIS}$).
 
@@
 
-%For JHUAPL-Bucky, the allocations were assigned at the \Sexpr{Buck_quantile}th percentile of the predictive distribution and for CU-select they were at the \Sexpr{CU_quantile}th percentile.
-%This reflects that in general the JHUAPL-Bucky forecasts assigned more probability to higher values and thus a low quantile value across all locations reached the allocation limit of 15,000.
-%In California, New York and Texas, the eventually observed number of hospital admissions was closer to the predicted median from JHUAPL-Bucky than the median from CU-select, and was contained in JHUAPL-Bucky's 80\% PI but not in the 80\% PI from CU-select.
-However for each of these three states, the resource allocation was lower for JHUAPL-Bucky than for CU-select and therefore resulted in more hypothetical unmet need.
-%In Florida, which was the state that saw the largest number of reported hospitalizations on this target date, the forecast from JHUAPL-Bucky missed by a wide margin (as did the forecast from CU-select) and the allocation was \Sexpr{FL_diff} units lower than the allocation derived from the CU-select model.
-
 <<thermometer-plot, fig.height=7, fig.width=7.5, fig.cap="Probabilistic forecasts for new COVID-19 hospital admissions on January 3, 2022 and their suggested resource allocations for the states with the ten highest (eventually observed) hospitalization counts.
 For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts.
-The side-by-side shaded regions show the median (solid red horizontal line) and 50%, 80% and 95% prediction intervals for the two selected models.
+The side-by-side shaded regions show the median (solid red horizontal line) and 50\\%, 80\\% and 95\\% prediction intervals for the two selected models.
 The forecasts were made for new hospitalizations on January 3, 2022 (vertical dashed line, with number of hospitalizations indicated by red horizontal line segment at the intersection of the dashed line and the grey line of data).
@@
 
-A state-wise comparison using Figures \ref{fig:thermometer-plot} and \ref{fig:multi-loc-ranks}A of the \texttt{JHUAPL-Bucky} and
-\texttt{CU-select} models shows that while the \texttt{JHUAPL-Bucky} forecast distributions were more centered on the eventual
-observations in many states, the suggested allocations of the \texttt{JHUAPL-Bucky} model were less efficient than those of the
-\texttt{CU-select} model. Thus \texttt{JHUAPL-Bucky} achieved a worse allocation score than \texttt{CU-select} (on this particular week) by
-allocating excess resources to several states, such as Pennsylvania and Michigan, rather than to states such as Florida
-and California where these resources would have reduced unmet need.  These allocation errors of \texttt{JHUAPL-Bucky} resulted
-from forecasts failing to consistently capture the relative resource needs across different states. The \texttt{CU-select} model
-made some similar errors — most prominently, over-allocating resources to Ohio — but overall, it more successfully
-forecasted the resource demands across different locations in relative terms.
-
-On the other hand, \texttt{CU-select} had worse performance as measured by MWIS. Its forecasts were biased downwards, and it
-consistently incurred a large penalty for underprediction (Figure \ref{fig:multi-loc-ranks}B). The \texttt{JHUAPL-Bucky} model
-had wider predictive intervals which more often included the observed level of daily hospital admissions. Its MWIS was
-therefore less severely penalized by under- or over-predictions of the actual hospitalization counts.
+A state-wise comparison using Figures \ref{fig:thermometer-plot} and \ref{fig:multi-loc-ranks}A of the
+\texttt{JHUAPL-Bucky} and \texttt{CU-select} models shows that while the \texttt{JHUAPL-Bucky} forecast distributions
+were more centered on the eventual observations in many states, the suggested allocations of the \texttt{JHUAPL-Bucky}
+model were less efficient than those of the \texttt{CU-select} model. Thus \texttt{JHUAPL-Bucky} achieved a worse
+allocation score than \texttt{CU-select} (on this particular week) by allocating excess resources to several states,
+such as Pennsylvania and Michigan, rather than to states such as Florida and California where these resources would have
+reduced unmet need.  These allocation errors of \texttt{JHUAPL-Bucky} resulted from forecasts failing to consistently
+capture the relative resource needs across different states. The \texttt{CU-select} model made some similar errors —
+most prominently, over-allocating resources to Ohio — but overall, it more successfully forecasted the resource demands
+across different locations in relative terms.
+
+On the other hand, \texttt{CU-select} had worse performance as measured by MWIS. Its forecasts were biased downwards,
+and it consistently incurred a large penalty for underprediction (Figure \ref{fig:multi-loc-ranks}B). The
+\texttt{JHUAPL-Bucky} model had wider predictive intervals which more often included the observed level of daily
+hospital admissions. Its MWIS was therefore less severely penalized by under- or over-predictions of the actual
+hospitalization counts.
 
@@
 
-%Overall, across the first 11 weeks evaluated (we excluded the last four since nearly all the models achieved an allocation score of zero), the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, had the highest rank for allocation score in four weeks, more than any other model except the \texttt{COVIDhub-ensemble} which also had four.
-
 <<eval=FALSE>>=
@@
 @
-
-%Mean weighted interval scores (MWIS) also varied by date and model, and more clearly were dependent on the scale of the observed data.
-%MWIS values were low (all models under 100) for all Mondays in December 2021 and the final four Mondays evaluated.
-%Across all models both the average and median MWIS value for every Monday in January was above 100, with the largest errors occurring one and two weeks after the peak was observed.
 
@@
 Averaging across all weeks, the ensemble forecast achieved a better mean allocation score (MAS) than two benchmark
-methods. The \texttt{COVIDhub-ensemble} had the best MAS across all weeks, the \texttt{per-capita} allocation had the second best MAS, and
-the \texttt{COVIDhub-baseline} model had the sixth best allocation (Table \ref{tab:multi-week-performance-summary}). No
-individual model had a better MAS than the per-capita allocation.
-
-Model rankings on mean allocation score (MAS) and time-averaged mean weighted interval score
-(taMWIS) were broadly similar, with some places of disagreement. The four most accurate and the five least accurate models were the same
+methods. The \texttt{COVIDhub-ensemble} had the best MAS across all weeks, the \texttt{per-capita} allocation had the
+second best MAS, and the \texttt{COVIDhub-baseline} model had the sixth best allocation (Table
+\ref{tab:multi-week-performance-summary}). No individual model had a better MAS than the per-capita allocation.
+
+Model rankings on mean allocation score (MAS) and time-averaged mean weighted interval score (taMWIS) were broadly
+similar, with some places of disagreement. The four most accurate and the five least accurate models were the same
 according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
-(This comparison excludes the \texttt{per-capita} allocation which does not use forecasts and therefore cannot be assigned a MWIS.)
-Notably, the \texttt{JHUAPL-SLPHospEns} model ranked best on taMWIS but had only the fourth best MAS.
+(This comparison excludes the \texttt{per-capita} allocation which does not use forecasts and therefore cannot be
+assigned a MWIS.) Notably, the \texttt{JHUAPL-SLPHospEns} model ranked best on taMWIS but had only the fourth best MAS.
 
@@
 
-We can gain some insight into the discordance between MAS and taMWIS by examining the correlation between time-specific AS and MWIS values of the various models in Figure
-\ref{fig:metrics-correlation}. We find, for example, no clear association between the consistent
-and high MWIS rank of the \texttt{JHUAPL-SLPHospEns} model and its highly variable AS rank. This contrasts 
-with a clearly positive correlation between MWIS and AS ranks in other models such as 
-\texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha}.
-We also can see in more detail the consistently high performance \texttt{COVIDhub-ensemble} on both scores.
-And while it did not submit forecasts for enough weeks to be included in Table \ref{tab:multi-week-performance-summary}, \texttt{CMU-TimeSeries} is an interesting example of a model that 
+We can gain some insight into the discordance between MAS and taMWIS by examining the correlation between time-specific
+AS and MWIS values of the various models in Figure \ref{fig:metrics-correlation}. We find, for example, no clear
+association between the consistent and high MWIS rank of the \texttt{JHUAPL-SLPHospEns} model and its highly variable AS
+rank. This contrasts with a clearly positive correlation between MWIS and AS ranks in other models such as
+\texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha}. We also can see in more detail the consistently high performance
+\texttt{COVIDhub-ensemble} on both scores. And while it did not submit forecasts for enough weeks to be included in
+Table \ref{tab:multi-week-performance-summary}, \texttt{CMU-TimeSeries} is an interesting example of a model that
 performed consistently well on AS but had only middling MWIS ranks.
@@
 
-%The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the constraint ($K$), possibly taking weights into account for different values of $K$ that might be more or less likely (Section \ref{sec:methods.detailed.integrated_allocation}).
-%This could be useful in situations where the actual constraint may not be known precisely, or where we wish to consider results across a range of decision making settings with different constraints.
-%Unlike in the above analyses where we conducted analyses assuming that $K$ was known to be 15,000, the analyses in this subsection are conducted assuming different distributions on $K$.
-
 <<calc-tot-y>>=
@@
 
-Our example application of the AS refers to a hypothetical and unspecified resource need that is assumed to 
-pair with our forecasted outcome (new hospital admissions) in a manner that might appear overly simplistic and 
-artificially direct.  We are optimistic, however, that by augmenting this scheme with a probabilistic measurement
-model for resource need and more detailed domain knowledge, it could be meaningfully aligned with a 
-more concrete decision problem such as the allocation of ventilators during a respiratory viral pandemic.  This problem 
-was explored from a stochastic optimization perspective (not from a forecast evaluation perspective) in 
-\cite{huang_stockpiling_2017} and provided an initial motivation for this project.
-We see it as promising focus for further development of the AS framework since it (a) has been identified as a prime
-manifestation of scarcity during respiratory outbreaks (b) tansparently involves
+Our example application of the AS refers to a hypothetical and unspecified resource need that is assumed to pair with
+our forecasted outcome (new hospital admissions) in a manner that might appear overly simplistic and artificially
+direct.  We are optimistic, however, that by augmenting this scheme with a probabilistic measurement model for resource
+need and more detailed domain knowledge, it could be meaningfully aligned with a more concrete decision problem such as
+the allocation of ventilators during a respiratory viral pandemic.  This problem was explored from a stochastic
+optimization perspective (not from a forecast evaluation perspective) in \cite{huang_stockpiling_2017} and provided an
+initial motivation for this project. We see it as promising focus for further development of the AS framework since it
+(a) has been identified as a prime manifestation of scarcity during respiratory outbreaks (b) tansparently involves
 logistical considerations depending on geographical and population scales, and (c) would likely imply a central role for
-hospitalization data. But again, additional modeling and data synthesis would be required since not everyone who is hospitalized
-needs a ventilator, and new hospital admissions will not fully describe for a need that corresponds to a prevalence
-of severe cases. Other possible real-world examples include the allocation of a limited stockpile of vaccinations
-\citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests \citep{du_optimal_2022,pasco_covid-19_2023}.
+hospitalization data. But again, additional modeling and data synthesis would be required since not everyone who is
+hospitalized needs a ventilator, and new hospital admissions will not fully describe for a need that corresponds to a
+prevalence of severe cases. Other possible real-world examples include the allocation of a limited stockpile of
+vaccinations \citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests
+\citep{du_optimal_2022,pasco_covid-19_2023}.
 
@@
 
-We address some technical and methodological points from the main text.
-We begin in section \ref{sec:proper} by defining proper scoring rules and showing that the allocation score is proper.
-In section \ref{sec:ex-shortage}, we formalize the concept of a \emph{shortage} of resources, give some key results about expected resource shortages under a distribution characterizing uncertainty about (future) levels of resource need, and illustrate how quantiles arise as the Bayes act in a decision making problem about the quantity of a resource to purchase.
-Section \ref{sec:bayes-quantiles} gives a justification for the result that the Bayes act for the allocation problem is given by a vector of quantiles of the forecast distributions in each location at a shared probability level, which was stated in section
-\ref{sec:methods.detailed.specific_allocation}.
-We describe the algorithm that we use to compute allocations given a forecast distribution in each location in section \ref{sec:numeric}.
-In section \ref{sec:distfromq}, we describe the methods for approximating a full distribution from a set of quantiles, implemented in the R package \verb`distfromq`, that we used to support computation of allocation scores from quantile forecasts that were submitted to the US COVID-19 Forecast Hub for the application in section \ref{sec:application}.
-Finally, in section
-\ref{sec:propriety_of_parametric_approximation} we examine implications for propriety of an analysis that uses summaries of forecasts (such as predictive quantiles) rather than the full forecast distributions for computation of allocation scores.
+We address some technical and methodological points from the main text. We begin in section \ref{sec:proper} by defining
+proper scoring rules and showing that the allocation score is proper. In section \ref{sec:ex-shortage}, we formalize the
+concept of a \emph{shortage} of resources, give some key results about expected resource shortages under a distribution
+characterizing uncertainty about (future) levels of resource need, and illustrate how quantiles arise as the Bayes act
+in a decision making problem about the quantity of a resource to purchase. Section \ref{sec:bayes-quantiles} gives a
+justification for the result that the Bayes act for the allocation problem is given by a vector of quantiles of the
+forecast distributions in each location at a shared probability level, which was stated in section
+\ref{sec:methods.detailed.specific_allocation}. We describe the algorithm that we use to compute allocations given a
+forecast distribution in each location in section \ref{sec:numeric}. In section \ref{sec:distfromq}, we describe the
+methods for approximating a full distribution from a set of quantiles, implemented in the R package \verb`distfromq`,
+that we used to support computation of allocation scores from quantile forecasts that were submitted to the US COVID-19
+Forecast Hub for the application in section \ref{sec:application}. Finally, in section
+\ref{sec:propriety_of_parametric_approximation} we examine implications for propriety of an analysis that uses summaries
+of forecasts (such as predictive quantiles) rather than the full forecast distributions for computation of allocation
+scores.
 
@@
 
-In decision theory, a loss function $l$ is used to formalize a decision problem by assigning numerical value $l(x,y)$ to the
-\emph{result} of taking an \emph{action} $x$ in preparation for an \emph{outcome} $y$. A \emph{scoring rule} $S$ is a
-loss function for a decision problem where the action is a probabilistic forecast $F$ of the outcome $y$ (or the statement of $F$ by a forecaster).
-% As does any loss fun $S$ transforms a random outcome variable $Y$ into a random loss $S(F,Y)$.
-We refer to the realized loss $S(F,y)$ as the \emph{score} of $F$ at $y$.
-
-Probabilistic forecasts can be seen as a unique kind of action in that they can be used to generate their
-own (simulated) outcome data, against which they can be scored using $S$. A probabilistic forecast
-$F$ is thus committed to the ``self-assessment'' $\Ex_F [S(F, Y)] := \Ex [S(F, Y^F)]$, where $Y^F \sim F$ is the random variable defined
-by sampling from $F$, as well to an assessment $\Ex_F [S(G, Y)]$ of any alternative forecast $G$.
-
-A natural consistency criterion for $S$ is that, for observations assumed to be drawn from $F$, it will not assess any other forecast $G$
-as being better than $F$ itself, that is, that
+In decision theory, a loss function $l$ is used to formalize a decision problem by assigning numerical value $l(x,y)$ to
+the \emph{result} of taking an \emph{action} $x$ in preparation for an \emph{outcome} $y$. A \emph{scoring rule} $S$ is
+a loss function for a decision problem where the action is a probabilistic forecast $F$ of the outcome $y$ (or the
+statement of $F$ by a forecaster). We refer to the realized loss $S(F,y)$ as the \emph{score} of $F$ at $y$.
+
+Probabilistic forecasts can be seen as a unique kind of action in that they can be used to generate their own
+(simulated) outcome data, against which they can be scored using $S$. A probabilistic forecast $F$ is thus committed to
+the ``self-assessment'' $\Ex_F [S(F, Y)] := \Ex [S(F, Y^F)]$, where $Y^F \sim F$ is the random variable defined by
+sampling from $F$, as well to an assessment $\Ex_F [S(G, Y)]$ of any alternative forecast $G$.
+
+A natural consistency criterion for $S$ is that, for observations assumed to be drawn from $F$, it will not assess any
+other forecast $G$ as being better than $F$ itself, that is, that
 \begin{align}
@@
 \end{align}
-for any $F,G$. A scoring rule meeting this criterion is called \emph{proper}. If $S$ were improper, then from the perspective of
-a forecaster focussed (solely) on expected loss minimization, the decision to state
-a forecast $G$ other than the forecast $F$ which they believe describes $Y$ could be superior to the decision to state $F$.
-$S$ is \emph{strictly proper} when
-\eqref{eqn:prop_ineq} is a strict inequality, in which case the
-\emph{only} optimal decision for a forecaster seeking to minimize their expected loss is to state the forecast they believe to be true.
+for any $F,G$. A scoring rule meeting this criterion is called \emph{proper}. If $S$ were improper, then from the
+perspective of a forecaster focussed (solely) on expected loss minimization, the decision to state a forecast $G$ other
+than the forecast $F$ which they believe describes $Y$ could be superior to the decision to state $F$. $S$ is
+\emph{strictly proper} when \eqref{eqn:prop_ineq} is a strict inequality, in which case the \emph{only} optimal decision
+for a forecaster seeking to minimize their expected loss is to state the forecast they believe to be true.
 
@@
 termed Bayes scoring rules typically appear via \eqref{eqn:bayes_sr} where $x^F$ is some given functional of $F$ which
-can be shown to be \emph{elicitable}, that is, to be the Bayes act for some loss function $s$.
-Such a loss function is said to be a \emph{consistent loss (or scoring) function} for the functional $F \mapsto x^F$, and many important
-recent results in the literature (e.g., \cite{fisslerziegel2016consistency}) address whether there \emph{exists} any loss
-function that is consistent for $x^F$. Our orientation
-is different from this insofar as we \emph{begin} by specifying a decision problem and a loss function of subject matter relevance
-and use the Bayes act only as a bridge to a proper scoring rule.  Consistency is never in doubt.
+can be shown to be \emph{elicitable}, that is, to be the Bayes act for some loss function $s$. Such a loss function is
+said to be a \emph{consistent loss (or scoring) function} for the functional $F \mapsto x^F$, and many important recent
+results in the literature (e.g., \cite{fisslerziegel2016consistency}) address whether there \emph{exists} any loss
+function that is consistent for $x^F$. Our orientation is different from this insofar as we \emph{begin} by specifying a
+decision problem and a loss function of subject matter relevance and use the Bayes act only as a bridge to a proper
+scoring rule.  Consistency is never in doubt.
 
@@
 A key feature of loss functions for decision problems used to define quantiles and related scoring rules such as the
-CRPS and the WIS (see e.g., \cite{gneiting2011quantiles}, \cite{jose2009evaluating},
-and \cite{royset2022optimization}, sections 1.C and 3.C), as well as the allocation loss function presented in this work,
-is the presence of a \emph{shortage}:
-the amount $\max\{0,y-x\}$ by which a resource demand
-$y$ exceeds a supply decision variable $x$, which, for convenience, we write as $(y-x)_{+}$. In particular, a quantile at
-probability level $\alpha$ of a distribution $F$ on $\mathbb{R}^1$ (which we assume to have a well-defined density $f(x)$)
-is a Bayes act for the loss function
+CRPS and the WIS (see e.g., \cite{gneiting2011quantiles}, \cite{jose2009evaluating}, and \cite{royset2022optimization},
+sections 1.C and 3.C), as well as the allocation loss function presented in this work, is the presence of a
+\emph{shortage}: the amount $\max\{0,y-x\}$ by which a resource demand $y$ exceeds a supply decision variable $x$,
+which, for convenience, we write as $(y-x)_{+}$. In particular, a quantile at probability level $\alpha$ of a
+distribution $F$ on $\mathbb{R}^1$ (which we assume to have a well-defined density $f(x)$) is a Bayes act for the loss
+function
 \[
@@
 \]
-where $\alpha = 1-C/L$ and $C$ and $L$ can be interpreted as the cost per unit of a resource (such as medicine) and the loss
-incurred when a unit of demand (such as illness) cannot be met due to the shortage $(y-x)_{+}$.  This follows because a
-Bayes act, as a minimizer of $\Ex_F[l(x,Y)]$, must also be a vanishing point of the derivative
+where $\alpha = 1-C/L$ and $C$ and $L$ can be interpreted as the cost per unit of a resource (such as medicine) and the
+loss incurred when a unit of demand (such as illness) cannot be met due to the shortage $(y-x)_{+}$.  This follows
+because a Bayes act, as a minimizer of $\Ex_F[l(x,Y)]$, must also be a vanishing point of the derivative
 \begin{align}
@@
 so that $1-C/L = F(x)$.
-The formula $\dby{}{x}\Ex_F\left[(Y-x)_+\right] = F(x) - 1$ for the derivative of the shortage,
-used above in \eqref{eqn:q_deriv}, can be obtained from an application of the ``Leibniz Rule'':
+The formula $\dby{}{x}\Ex_F\left[(Y-x)_+\right] = F(x) - 1$ for the derivative of the shortage, used above in
+\eqref{eqn:q_deriv}, can be obtained from an application of the ``Leibniz Rule'':
 \begin{align}
@@
 
-Here we study the form of the Bayes act for the allocation problem (AP) (equation \eqref{eqn:loss_fn} in section \ref{sec:methods.detailed.specific_allocation}) of the text:
+Here we study the form of the Bayes act for the allocation problem (AP) (equation \eqref{eqn:loss_fn} in section
+\ref{sec:methods.detailed.specific_allocation}) of the text:
 \begin{align}
@@
 \end{align}
-where the marginal forecasts $F_i$ for $i=1,\dots,N$ represent forecasts for $N$ distinct locations.
-We show that the Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$ and resource constraint level $K$
-is a vector of quantiles of the marginal forecast distributions $F_i$ at a single probability level
-$\tau^{F,K}$, that is, $x_i^{F,K} = q_{F_i,\tau^{F,K}}$. An immediate consequence used in the examples in Section \ref
-{sec:methods.overview} is that if $F_i = \mathrm{Exp}(1/\sigma_i)$ for all $i$, then the Bayes act is
-proportional to $(\sigma_1,\ldots,\sigma_N)$, since $q_{\mathrm{Exp}(1/\sigma),\tau} = -\sigma \log(1-\tau)$.
-
-In order for $x^{\star} \in \mathbb{R}^N_{+}$ to solve the AP it must be true that reallocating $\delta > 0$ units of the resource from location $i$ to location $j$ will lead to a net increase in expected shortage --- in other words, the reallocation increases the expected shortage in location $i$ is at least as much as it decreases the expected shortage in location $j$:
+where the marginal forecasts $F_i$ for $i=1,\dots,N$ represent forecasts for $N$ distinct locations. We show that the
+Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$ and resource constraint level $K$ is a vector of
+quantiles of the marginal forecast distributions $F_i$ at a single probability level $\tau^{F,K}$, that is, $x_i^{F,K} =
+q_{F_i,\tau^{F,K}}$. An immediate consequence used in the examples in Section \ref {sec:methods.overview} is that if
+$F_i = \mathrm{Exp}(1/\sigma_i)$ for all $i$, then the Bayes act is proportional to $(\sigma_1,\ldots,\sigma_N)$, since
+$q_{\mathrm{Exp}(1/\sigma),\tau} = -\sigma \log(1-\tau)$.
+
+In order for $x^{\star} \in \mathbb{R}^N_{+}$ to solve the AP it must be true that reallocating $\delta > 0$ units of
+the resource from location $i$ to location $j$ will lead to a net increase in expected shortage --- in other words, the
+reallocation increases the expected shortage in location $i$ is at least as much as it decreases the expected shortage
+in location $j$:
 \begin{align*}
@@
 
-To compute an allocation score $S_A(F,y;K) := s_A(x^{F,K},y)$, we require numerical values for a
-Bayes act solving the AP \eqref{AP-long} --- that is, we must find the specific resource allocations for each location that are determined by the forecast $F$ under the resource constraint $K$.
-Assuming we have reliable means of calculating quantiles $q_{\alpha,F_i}$
-of the marginal forecasts $F_i$,
-these allocations are given by $q_{\tau^{\star},F_i}$ where $\tau^{\star}$ solves the equation \eqref{eqn:quantiles-sum-to-K}.
-However, this equation is not analytically tractable and we must resort to a numerical method
-for finding an approximation $\tilde{\tau}$ of $\tau^{\star}$.
-
-We have implemented an iterative bisection method that makes use of the fact that $\sum_{i=1}^N q_{\tau,F_i}$
-is an increasing function of $\tau$.
-The algorithm begins with an initial search interval $[\tau_{L,1}, \tau_{U,1}]$ (such as $[0,\max_{i}F_i(K)]$) that clearly contains
-the solution $\tau^{\star}$.
-% This can be confirmed by verifying that the resource allocations corresponding to the lower endpoint of the search interval do not exceed the resource constraint ($\sum_{i=1}^N q_{\tau_L,F_i} \leq K$) and the resource allocations corresponding to the upper endpoint match or exceed the available resources ($\sum_{i=1}^N q_{\tau_U,F_i} \geq K$).
-At each step $j$ of the algorithm, we evaluate the total allocation $\sum_{i=1}^N q_{\tau_{M,j},F_i}$ at the midpoint of the search interval,
+To compute an allocation score $S_A(F,y;K) := s_A(x^{F,K},y)$, we require numerical values for a Bayes act solving the
+AP \eqref{AP-long} --- that is, we must find the specific resource allocations for each location that are determined by
+the forecast $F$ under the resource constraint $K$. Assuming we have reliable means of calculating quantiles
+$q_{\alpha,F_i}$ of the marginal forecasts $F_i$, these allocations are given by $q_{\tau^{\star},F_i}$ where
+$\tau^{\star}$ solves the equation \eqref{eqn:quantiles-sum-to-K}. However, this equation is not analytically tractable
+and we must resort to a numerical method for finding an approximation $\tilde{\tau}$ of $\tau^{\star}$.
+
+We have implemented an iterative bisection method that makes use of the fact that $\sum_{i=1}^N q_{\tau,F_i}$ is an
+increasing function of $\tau$. The algorithm begins with an initial search interval $[\tau_{L,1}, \tau_{U,1}]$ (such as
+$[0,\max_{i}F_i(K)]$) that clearly contains the solution $\tau^{\star}$. At each step $j$ of the algorithm, we evaluate
+the total allocation $\sum_{i=1}^N q_{\tau_{M,j},F_i}$ at the midpoint of the search interval,
 $\tau_{M,j} = \frac{1}{2}(\tau_{L,j} + \tau_{U,j})$ and continue the search on the narrowed sub-interval
@@
 \end{align}
-% We then narrow the search to the sub-interval that contains the solution:
-% \begin{itemize}
-%   \item If $\sum_{i=1}^N q_{\tau_{M},F_i} \geq K$, we continue the search on the interval $[\tau_L, \tau_M]$.
-%   \item If $\sum_{i=1}^N q_{\tau_M, F_i} \leq K$, we continue the search on the interval $[\tau_M, \tau_U]$.
-% \end{itemize}
-This search continues until $\tau_{U,j+1} < (1+\varepsilon)\tau_{L,j+1}$ for a suitably small $\varepsilon>0$.
-We have implemented this procedure along with the resulting score computations in the
-R package \verb`alloscore` \citep{gerding-alloscore} which provided all allocation score values used in the analysis of section \ref{sec:application}.
-
-Subtleties can arise when the forecast
-densities $f_i$ vanish or are very small, in which case quantiles are non-unique or highly variable near a probability level,
-leading to ambiguity or numerical instabilities{} in the evaluation of $\sum_{i=1}^N q_{\tau,F_i}$. Additionally, if point masses are present in any of the $F_i$,
-\eqref{eqn:quantiles-sum-to-K} will not have a unique solution for some discrete set of constraint levels $K$.
-We have adopted conventions for detecting such levels and enforcing consistency in score calculations
-near them.
-Through extensive experimentation, we have determined that these conditions seem to address these challenges with the forecasts we are working with, but we leave a more rigorous approximation error analysis for later work.
+This search continues until $\tau_{U,j+1} < (1+\varepsilon)\tau_{L,j+1}$ for a suitably small $\varepsilon>0$. We have
+implemented this procedure along with the resulting score computations in the R package \verb`alloscore`
+\citep{gerding-alloscore} which provided all allocation score values used in the analysis of section
+\ref{sec:application}.
+
+Subtleties can arise when the forecast densities $f_i$ vanish or are very small, in which case quantiles are non-unique
+or highly variable near a probability level, leading to ambiguity or numerical instabilities{} in the evaluation of
+$\sum_{i=1}^N q_{\tau,F_i}$. Additionally, if point masses are present in any of the $F_i$,
+\eqref{eqn:quantiles-sum-to-K} will not have a unique solution for some discrete set of constraint levels $K$. We have
+adopted conventions for detecting such levels and enforcing consistency in score calculations near them. Through
+extensive experimentation, we have determined that these conditions seem to address these challenges with the forecasts
+we are working with, but we leave a more rigorous approximation error analysis for later work.
 
@@
 
-In section \ref{sec:application} of the manuscript, we used the allocation score to evaluate forecasts of COVID-19 hospitalizations that have been submitted to the US COVID-19 Forecast Hub.
-These forecasts are submitted to the Hub using a set of 23 quantiles of the forecast distribution at the 23 probability levels in the set $\mathcal{T} = \{0.01, 0.025, 0.05, 0.1, 0.15, \ldots, \allowbreak 0.9, 0.95, 0.975, 0.99\}$, which specify a predictive median and the endpoints of central $(1 - \alpha) \times 100\%$ prediction intervals at levels $\alpha = 0.02, 0.05, 0.1, 0.2, \allowbreak 0.3, 0.4, 0.5, 0.6, \allowbreak 0.7, 0.8, 0.9$.
-For a given week and target date, we use $q_{i,k}$ to denote the submitted quantiles for location $i$ and probability level $\tau_k \in \mathcal{T}$, $k = 1, \ldots, 23$.
-
-In the event that there is some $k \in \{1, \ldots, 23\}$ for which $\sum_i q_{i,k} = K$, i.e., the provided predictive quantiles at level $\tau_k$ sum across locations to the resource constraint $K$, the solution to the allocation problem is given by those quantiles.
-However, generally this will not be the case; the optimal allocation will typically be at some probability level $\tau^\star \notin \mathcal{T}$.
-
-To address this situation and support the numerical allocation algorithm outlined in section \ref{sec:numeric}, we need a mechanism to approximate the full cumulative distribution functions $F_i$, $i = 1, \ldots, N$ based on the provided quantiles.
-We have developed functionality for this purpose in the \verb`distfromq` package for R.\citep{ray-distfromq} This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:
+In section \ref{sec:application} of the manuscript, we used the allocation score to evaluate forecasts of COVID-19
+hospitalizations that have been submitted to the US COVID-19 Forecast Hub. These forecasts are submitted to the Hub
+using a set of 23 quantiles of the forecast distribution at the 23 probability levels in the set $\mathcal{T} = \{0.01,
+0.025, 0.05, 0.1, 0.15, \ldots, \allowbreak 0.9, 0.95, 0.975, 0.99\}$, which specify a predictive median and the
+endpoints of central $(1 - \alpha) \times 100\%$ prediction intervals at levels $\alpha = 0.02, 0.05, 0.1, 0.2,
+\allowbreak 0.3, 0.4, 0.5, 0.6, \allowbreak 0.7, 0.8, 0.9$. For a given week and target date, we use $q_{i,k}$ to denote
+the submitted quantiles for location $i$ and probability level $\tau_k \in \mathcal{T}$, $k = 1, \ldots, 23$.
+
+In the event that there is some $k \in \{1, \ldots, 23\}$ for which $\sum_i q_{i,k} = K$, i.e., the provided predictive
+quantiles at level $\tau_k$ sum across locations to the resource constraint $K$, the solution to the allocation problem
+is given by those quantiles. However, generally this will not be the case; the optimal allocation will typically be at
+some probability level $\tau^\star \notin \mathcal{T}$.
+
+To address this situation and support the numerical allocation algorithm outlined in section \ref{sec:numeric}, we need
+a mechanism to approximate the full cumulative distribution functions $F_i$, $i = 1, \ldots, N$ based on the provided
+quantiles. We have developed functionality for this purpose in the \verb`distfromq` package for R.\citep{ray-distfromq}
+This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:
 \begin{enumerate}
-  \item Identify a discrete component of the distribution consisting of zero or more point masses, and create an adjusted set of predictive quantiles for the continuous part of the distribution by subtracting the point mass probabilities and rescaling.
-  \item For the continuous part of the distribution, different approaches are used on the interior and exterior of the provided quantiles:
+  \item Identify a discrete component of the distribution consisting of zero or more point masses, and create an
+    adjusted set of predictive quantiles for the continuous part of the distribution by subtracting the point mass
+    probabilities and rescaling.
+  \item For the continuous part of the distribution, different approaches are used on the interior and exterior of the
+    provided quantiles:
   \begin{enumerate}
-    \item On the interior, a monotonic cubic spline interpolates the adjusted quantiles representing the continuous part of the distribution.
-    \item A location-scale parametric family is used to extrapolate beyond the provided quantiles.
-The location and scale parameters are estimated separately for the lower and upper tails so as to obtain a tail distribution that matches the two most extreme quantiles in each tail.
-In this work, we use normal distributions for the tails.
+    \item On the interior, a monotonic cubic spline interpolates the adjusted quantiles representing the continuous part
+      of the distribution.
+    \item A location-scale parametric family is used to extrapolate beyond the provided quantiles. The location and
+      scale parameters are estimated separately for the lower and upper tails so as to obtain a tail distribution that
+      matches the two most extreme quantiles in each tail. In this work, we use normal distributions for the tails.
   \end{enumerate}
 \end{enumerate}
-The resulting distributional estimate exactly matches all of the predictive quantiles provided by the forecaster.
-We use the cumulative distribution function resulting from this procedure as an input to the allocation score algorithm.
-
-We refer the reader to the \verb`distfromq` documentation for further detail.\citep{ray-distfromq}
-
-% \begin{enumerate}
-%   \item First, we handle settings where the provided quantiles imply the existence of one or more point masses.
-Specifically, if there are distinct probability levels $\tau_k < \tau_{k^\prime}$ that have the same quantiles, $q_{l,k} = q_{l, k^\prime}$, the distribution contains a point mass at $q_{l,k}$.
-We split into three cases depending on the number of distinct provided quantiles:
-%   \begin{enumerate}
-%     \item If all provided quantiles are equal to each other, we infer that the distribution consists of a single point mass at that quantile value.
-%     \item If there are two distinct quantiles across all of the probability levels $\tau_k \in \mathcal{T}$, our estimated distribution consists of a mixture of two point masses.
-The probability assigned to the point mass at $q$ is proportional to $\tau_q^\max - \tau_q^\min$, where
-%     $$\tau_q^\max = \begin{cases}
-%       \max_{\{\}} \tau_k
-%     \end{cases}$$ is the largest probability level $\tau_k$ for which $q_{l,k} = q$, and $\tau_q^\min$ is defined similarly.
-%     \item Otherwise, our final estimate will consist of a mixture of a discrete distribution with point masses at the duplicated quantiles and a continuous distribution elsewhere.
-In this step, we  In this case, the discrete distribution has point mass probabilities calculated as $d_q = \tau_q^\max - \tau_q^\min$, where $\tau_q^\max$ and $\tau_q^\min$ are defined as in the previous point.
-%   \end{enumerate}
-% \end{enumerate}
+The resulting distributional estimate exactly matches all of the predictive quantiles provided by the forecaster. We use
+the cumulative distribution function resulting from this procedure as an input to the allocation score algorithm.
+
+We refer the reader to the \verb`distfromq` documentation for further detail \citep{ray-distfromq}.
 
@@
 
-In practice, open forecasting exercises are generally not able to collect a perfect description of the forecast distribution $F$ other than in simple settings such as for a categorical variable with a relatively small number of categories.
-In settings where the outcome being forecasted is a continuous quantity (such as the proportion of outpatient doctor visits where the patient has influenza-like illness) or a count (such as influenza hospitalizations), forecasting exercises have therefore resorted to collecting summaries of a forecast distribution such as bin probabilities or predictive quantiles.
-In this section, we address two practical concerns raised by this.
-First, we discuss conditions under which it is possible to calculate the allocation score when only summaries of a forecast distribution are recorded in a submission to a forecast hub.
-Second, we show that a post hoc attempt to compute the allocation score based on submitted predictive quantiles may in fact compute an alternative score that is not proper.
+In practice, open forecasting exercises are generally not able to collect a perfect description of the forecast
+distribution $F$ other than in simple settings such as for a categorical variable with a relatively small number of
+categories. In settings where the outcome being forecasted is a continuous quantity (such as the proportion of
+outpatient doctor visits where the patient has influenza-like illness) or a count (such as influenza hospitalizations),
+forecasting exercises have therefore resorted to collecting summaries of a forecast distribution such as bin
+probabilities or predictive quantiles. In this section, we address two practical concerns raised by this. First, we
+discuss conditions under which it is possible to calculate the allocation score when only summaries of a forecast
+distribution are recorded in a submission to a forecast hub. Second, we show that a post hoc attempt to compute the
+allocation score based on submitted predictive quantiles may in fact compute an alternative score that is not proper.
 
@@
 
-We consider a setting where a forecasting exercise (such as a forecast hub) pre-specifies that forecasts will be represented using a parametric family of forecast distributions $G_\theta(y)$, and the task of the forecaster is to select a particular parameter value $\theta$.
-We use $\mathcal{P}$ to denote the collection of all distributions $G_\theta$ in the given parametric family.
-For instance, it has recently been proposed that mixture distributions could be used to represent forecast distributions \citep{wadsworth2023mixture}.
-Additionally, we note that the functionality in \verb`distfromq` can be viewed as specifying a parametric family $\mathcal{P}_{\mathrm{dfq}}$ where the parameters $\theta$ of $G_\theta$
-are its quantiles at pre-specified probability levels, and where the shape of any $G_\theta \in \mathcal{P}_{\mathrm{dfq}}$ over the full range of its support is entirely controlled by these quantiles.
-
-We find it helpful now to formally distinguish between two decision making problems.
-The first is the public health decision maker's allocation problem where the task is to select an allocation $x$, with the allocation loss $s_A(x, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i)$ as described in section \ref{sec:methods.detailed.specific_allocation}.
-The second is the forecaster's reporting problem where the task is to select parameter values $\theta$ to report. The forecaster's loss is given by
+We consider a setting where a forecasting exercise (such as a forecast hub) pre-specifies that forecasts will be
+represented using a parametric family of forecast distributions $G_\theta(y)$, and the task of the forecaster is to
+select a particular parameter value $\theta$. We use $\mathcal{P}$ to denote the collection of all distributions
+$G_\theta$ in the given parametric family. For instance, it has recently been proposed that mixture distributions could
+be used to represent forecast distributions \citep{wadsworth2023mixture}. Additionally, we note that the functionality
+in \verb`distfromq` can be viewed as specifying a parametric family $\mathcal{P}_{\mathrm{dfq}}$ where the parameters
+$\theta$ of $G_\theta$ are its quantiles at pre-specified probability levels, and where the shape of any $G_\theta \in
+\mathcal{P}_{\mathrm{dfq}}$ over the full range of its support is entirely controlled by these quantiles.
+
+We find it helpful now to formally distinguish between two decision making problems. The first is the public health
+decision maker's allocation problem where the task is to select an allocation $x$, with the allocation loss $s_A(x, y) =
+\sum_{i=1}^N L \cdot \max(0, y_i - x_i)$ as described in section \ref{sec:methods.detailed.specific_allocation}. The
+second is the forecaster's reporting problem where the task is to select parameter values $\theta$ to report. The
+forecaster's loss is given by
 \begin{align}
@@
 \end{align}
-where $x^{G_\theta}$ is the Bayes act for the allocation problem under the distribution $G_\theta$.
-In words, the loss associated with reporting $\theta$ is equal to the loss associated with taking the Bayes allocation corresponding to the distribution $G_\theta$.
-
-Following our usual construction, the Bayes act for the forecast reporting problem is the parameter set that minimizes the forecaster's expected loss.
-Breaking with our earlier notation for improved legibility, we use $\theta^\star(F)$ to denote this Bayes act:
+where $x^{G_\theta}$ is the Bayes act for the allocation problem under the distribution $G_\theta$. In words, the loss
+associated with reporting $\theta$ is equal to the loss associated with taking the Bayes allocation corresponding to the
+distribution $G_\theta$.
+
+Following our usual construction, the Bayes act for the forecast reporting problem is the parameter set that minimizes
+the forecaster's expected loss. Breaking with our earlier notation for improved legibility, we use $\theta^\star(F)$ to
+denote this Bayes act:
 \begin{align*}
@@
 It follows from the discussion in section \ref{sec:alloscore_proper} that this is a proper scoring rule for $F$.
-Although the full forecast distribution $F$ is not available in the forecast submission, the score $S_R(F, y)$ can be calculated from the reported parameter values as long as the forecaster submits the optimal parameters $\theta^\star(F)$.
-
-We emphasize that the forecaster's true predictive distribution $F$ does not need to be a member of the specified parametric family $\mathcal{P}$ for this construction to yield a proper score.
-It is, however, necessary to specify the parametric family to use and the foundational scoring rule $s_A$ (including any relevant problem parameters such as the resource constraint $K$) in advance, so that forecasters can identify the Bayes act parameter set $\theta^\star(F)$ to report.
-
-If the parametric family used to represent forecast distributions is flexible enough, the reporting scoring rule $S_R$ and the allocation score are equivalent in the sense that they will yield the same score for any distribution $F$.
-Suppose that for a given resource constraint $K$, for any forecast distribution $F$ it is possible to find a member $G_{\theta^\star}$ of the specified parametric family $\mathcal{P}$ with the same allocation as $F$ (i.e., $x^F = x^{G_{\theta^\star}}$).
-Then $\theta^\star$ is a Bayes act for the reporting problem since for any other parameter value $\theta$,
+Although the full forecast distribution $F$ is not available in the forecast submission, the score $S_R(F, y)$ can be
+calculated from the reported parameter values as long as the forecaster submits the optimal parameters
+$\theta^\star(F)$.
+
+We emphasize that the forecaster's true predictive distribution $F$ does not need to be a member of the specified
+parametric family $\mathcal{P}$ for this construction to yield a proper score. It is, however, necessary to specify the
+parametric family to use and the foundational scoring rule $s_A$ (including any relevant problem parameters such as the
+resource constraint $K$) in advance, so that forecasters can identify the Bayes act parameter set $\theta^\star(F)$ to
+report.
+
+If the parametric family used to represent forecast distributions is flexible enough, the reporting scoring rule $S_R$
+and the allocation score are equivalent in the sense that they will yield the same score for any distribution $F$.
+Suppose that for a given resource constraint $K$, for any forecast distribution $F$ it is possible to find a member
+$G_{\theta^\star}$ of the specified parametric family $\mathcal{P}$ with the same allocation as $F$ (i.e., 
+$x^F = x^{G_{\theta^\star}}$). Then $\theta^\star$ is a Bayes act for the reporting problem since for any other
+parameter value
+$\theta$,
 \begin{align*}
@@
 
-For the particular choice of the parametric family $\mathcal{P}_{\text{dfq}}$ (i.e., using the \verb`distfromq` package), this flexibility requirement is satisfied.
-For instance, the forecaster could pick one required quantile level (such as 0.5, for which the corresponding predictions are predictive medians), and set the submitted quantiles of their forecast distribution in each location at that level to be the desired allocations, which sum to $K$ across all locations.
-However, this representation of the forecast may be quite different from the actual forecast distribution $F$.
-For example, for the actual forecast distribution $F$ the allocations may occur at some quantile level other than 0.5.
-
-As another alternative for practical forecasting exercises, a forecast hub could ask forecasters to directly provide the Bayes allocations associated with their forecasts for one or more specified resource constraints $K$.
-At the cost of increasing the number of quantities solicited by the forecast hub, this would have several advantages: it would prevent any artificial distortion of the forecast distributions, allow for direct calculation of scores, and narrow the gap between model outputs and public health end users.
-For this to be feasible, implementations of the allocation algorithm would have to be provided to participating forecasters in the computational languages being used for modeling.
+For the particular choice of the parametric family $\mathcal{P}_{\text{dfq}}$ (i.e., using the \verb`distfromq`
+package), this flexibility requirement is satisfied. For instance, the forecaster could pick one required quantile level
+(such as 0.5, for which the corresponding predictions are predictive medians), and set the submitted quantiles of their
+forecast distribution in each location at that level to be the desired allocations, which sum to $K$ across all
+locations. However, this representation of the forecast may be quite different from the actual forecast distribution
+$F$. For example, for the actual forecast distribution $F$ the allocations may occur at some quantile level other than
+0.5.
+
+As another alternative for practical forecasting exercises, a forecast hub could ask forecasters to directly provide the
+Bayes allocations associated with their forecasts for one or more specified resource constraints $K$. At the cost of
+increasing the number of quantities solicited by the forecast hub, this would have several advantages: it would prevent
+any artificial distortion of the forecast distributions, allow for direct calculation of scores, and narrow the gap
+between model outputs and public health end users. For this to be feasible, implementations of the allocation algorithm
+would have to be provided to participating forecasters in the computational languages being used for modeling.
 
@@
 
-A \emph{post hoc} evaluation of quantile forecasts that combines the parametric family specified by \verb`distfromq` with the allocation score does not yield the allocation score of the forecast distribution $F$.
-Instead, it computes an alternative score that is improper.
-This is because the forecast distribution $F$ and the distribution $G^q \in \mathcal{P}_{\text{dfq}}$ with the same quantiles as $F$ may determine different resource allocations.
-In our investigations, these discrepancies appear to be relatively minor on the interior of the provided quantiles, but could be severe if the tail extrapolations performed by \verb`distfromq` do not match the tail behavior of $F$ and the allocations are in the tails of the predictive distribution.
+A \emph{post hoc} evaluation of quantile forecasts that combines the parametric family specified by \verb`distfromq`
+with the allocation score does not yield the allocation score of the forecast distribution $F$. Instead, it computes an
+alternative score that is improper. This is because the forecast distribution $F$ and the distribution $G^q \in
+\mathcal{P}_{\text{dfq}}$ with the same quantiles as $F$ may determine different resource allocations. In our
+investigations, these discrepancies appear to be relatively minor on the interior of the provided quantiles, but could
+be severe if the tail extrapolations performed by \verb`distfromq` do not match the tail behavior of $F$ and the
+allocations are in the tails of the predictive distribution.
 
@@
 
-When an analyst attempts a post hoc computation of the allocation score using $G^q(F)$ (implicitly assuming that $G^q(F) = G^\star(F)$), they in fact compute the alternative score
+When an analyst attempts a post hoc computation of the allocation score using $G^q(F)$ (implicitly assuming that $G^q(F)
+= G^\star(F)$), they in fact compute the alternative score
 $$\tilde{S}(F, y) = S_A(G^q(F), y) = s_A(x^{G^q(F)}, y).$$
@@
 \end{align}
-However, the inequality in \eqref{eqn:tilde_s_improper} will typically be strict.
-For example, if $F$ has heavy upper tails (such as for a lognormal distribution), but normal distributions are used for tail extrapolations in \verb`distfromq`, then the resource allocations based on the distribution $G^q(F)$ may be quite different from the optimal allocations under the distribution $F$, leading to a strict inequality.
-This demonstrates that $\tilde{S}$ is improper.
+However, the inequality in \eqref{eqn:tilde_s_improper} will typically be strict. For example, if $F$ has heavy upper
+tails (such as for a lognormal distribution), but normal distributions are used for tail extrapolations in
+\verb`distfromq`, then the resource allocations based on the distribution $G^q(F)$ may be quite different from the
+optimal allocations under the distribution $F$, leading to a strict inequality. This demonstrates that $\tilde{S}$ is
+improper.
 

### 2024-02-21 alloscore_manuscript/alloscore-application-shorter.tex
clean up Rnw

@@
 
-%We also computed a standardized rank for the allocation score for each model $m$ and week $w$.
-%First, we computed the number of models that forecasted that week ($n_w$) and the rank of model $m$ among the $n_w$ models ($r_{m,w}^{AS}$).
-%The model with the best allocation score received a rank of 1 and the worst received a rank of $n_w$.
-%In the case of a tie between one or more models, all models received the better rank.
-%We then rescaled these rankings to compute the allocation score standardized rank ($sr_{m,w}^AS$) between 0 and 1, where 0 corresponds to the worst rank and 1 to the best.
-
-%\begin{equation}
-%sr_{m,w}^{AS} = 1 - \frac{r^{AS}_{m,w}-1}{n_w-1}
-%\end{equation}
-
-%\subsubsection{Weighted Interval Score (WIS)}
-
-%The Weighted Interval Score (WIS), as described in earlier sections, measures the alignment of a single probabilistic forecast ($F$) with an observation ($y$).
-%We computed the mean WIS across all $L$ locations for each model and each forecasted week as
-%\begin{equation}
-%MWIS_{m,w} = \frac{1}{L}\sum_{l=1}^L WIS(F_{l,m,w},y)
-%\end{equation}
-%where $F_{l,m,w}$ is the probabilistic forecast from model $m$ for location $l$ and week $w$.
-%Using the same procedure as for allocation scores described above, we computed standardized ranks for MWIS ($sr_{m,w}^{MWIS}$).
-
 \subsection{Allocation benchmarks}
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Tue Feb 20 19:44:31 2024
+% Tue Feb 20 23:42:19 2024
 \begin{table}[ht]
@@
 
-%For JHUAPL-Bucky, the allocations were assigned at the Buck_quantileth percentile of the predictive distribution and for CU-select they were at the CU_quantileth percentile.
-%This reflects that in general the JHUAPL-Bucky forecasts assigned more probability to higher values and thus a low quantile value across all locations reached the allocation limit of 15,000.
-%In California, New York and Texas, the eventually observed number of hospital admissions was closer to the predicted median from JHUAPL-Bucky than the median from CU-select, and was contained in JHUAPL-Bucky's 80\% PI but not in the 80\% PI from CU-select.
-However for each of these three states, the resource allocation was lower for JHUAPL-Bucky than for CU-select and therefore resulted in more hypothetical unmet need.
-%In Florida, which was the state that saw the largest number of reported hospitalizations on this target date, the forecast from JHUAPL-Bucky missed by a wide margin (as did the forecast from CU-select) and the allocation was FL_diff units lower than the allocation derived from the CU-select model.
-
 \begin{knitrout}
@@
 
-%Overall, across the first 11 weeks evaluated (we excluded the last four since nearly all the models achieved an allocation score of zero), the \texttt{COVIDhub-baseline} model, which predicts a flat line from the most recent observation with uncertainty bounds based on a random walk, had the highest rank for allocation score in four weeks, more than any other model except the \texttt{COVIDhub-ensemble} which also had four.
-
-%Mean weighted interval scores (MWIS) also varied by date and model, and more clearly were dependent on the scale of the observed data.
-%MWIS values were low (all models under 100) for all Mondays in December 2021 and the final four Mondays evaluated.
-%Across all models both the average and median MWIS value for every Monday in January was above 100, with the largest errors occurring one and two weeks after the peak was observed.
-
 Averaging across all weeks, the ensemble forecast achieved a better mean allocation score (MAS) than two benchmark
-methods. The \texttt{COVIDhub-ensemble} had the best MAS across all weeks, the \texttt{per-capita} allocation had the second best MAS, and
-the \texttt{COVIDhub-baseline} model had the sixth best allocation (Table \ref{tab:multi-week-performance-summary}). No
-individual model had a better MAS than the per-capita allocation.
-
-Model rankings on mean allocation score (MAS) and time-averaged mean weighted interval score
-(taMWIS) were broadly similar, with some places of disagreement. The four most accurate and the five least accurate models were the same
+methods. The \texttt{COVIDhub-ensemble} had the best MAS across all weeks, the \texttt{per-capita} allocation had the
+second best MAS, and the \texttt{COVIDhub-baseline} model had the sixth best allocation (Table
+\ref{tab:multi-week-performance-summary}). No individual model had a better MAS than the per-capita allocation.
+
+Model rankings on mean allocation score (MAS) and time-averaged mean weighted interval score (taMWIS) were broadly
+similar, with some places of disagreement. The four most accurate and the five least accurate models were the same
 according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
-(This comparison excludes the \texttt{per-capita} allocation which does not use forecasts and therefore cannot be assigned a MWIS.)
-Notably, the \texttt{JHUAPL-SLPHospEns} model ranked best on taMWIS but had only the fourth best MAS.
+(This comparison excludes the \texttt{per-capita} allocation which does not use forecasts and therefore cannot be
+assigned a MWIS.) Notably, the \texttt{JHUAPL-SLPHospEns} model ranked best on taMWIS but had only the fourth best MAS.
 
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Tue Feb 20 19:44:36 2024
+% Tue Feb 20 23:42:23 2024
 \begin{table}[ht]
@@
 \label{sec:ias_examp}
-
-%The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the constraint ($K$), possibly taking weights into account for different values of $K$ that might be more or less likely (Section \ref{sec:methods.detailed.integrated_allocation}).
-%This could be useful in situations where the actual constraint may not be known precisely, or where we wish to consider results across a range of decision making settings with different constraints.
-%Unlike in the above analyses where we conducted analyses assuming that $K$ was known to be 15,000, the analyses in this subsection are conducted assuming different distributions on $K$.
 

### 2024-02-22 alloscore_manuscript/alloscore-application-shorter.Rnw
some edits to manuscript

@@
 that nothing is wasted in one location while it could be put to use in another. The best and lowest possible allocation 
-score is 0.  A positive allocation score indicates some way in which an oracle could have used perfect foreknowledge to
-meet more of the need than did the allocation suggested by $F$.
+score is 0.  A positive allocation score indicates that an oracle with perfect foreknowledge could have selected an allocation 
+that met more need than the allocation suggested by $F$.
 
@@
 units, the allocation based on the forecast exactly meets the observed need in location 1, but it leaves 6 units of need
-unmet in location 2. But these 6 units of unmet need are unaviodable under the constraint $K=5$: allocating 0 units of
+unmet in location 2. But these 6 units of unmet need are unavoidable under the constraint $K=5$: allocating 0 units of
 resources to location 1 and 5 to location 2, for example, still results in a total unmet need of 6 across both
@@
 
-Decision theory is concerned with understanding how an \emph{action} $x$ both is and ought to be selected from a set of
-possible actions $\mathcal{X}$ --- that is, how a decision is and ought to be made --- based on the anticipated
-consequences of $x$ under future states. We represent such future states as a random variable $Y$ and refer to a
-realization $y$ of $Y$ as an \emph{outcome}. In this paper $x$ will be related to the level of investment in a measure
+Decision theory is concerned with how to decide on an \emph{action} $x$ to select from a set of
+possible actions $\mathcal{X}$ in the face of uncertainty about the future.
+We use the random variable $Y$ to represent future states and refer to a
+realization $y$ of $Y$ as an \emph{outcome}. For example, $x$ might correspond to the level of investment in a measure that is
 designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with
-$\mathcal{X}$ being the set of all feasible levels of investment, while an outcome $y$ will correspond to the number of
-individuals who eventually become sick and would benefit from the mitigation measure. A basic strategy in decision
-theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numerical value, or
-\emph{loss}, measuring the success of $x$, which in our examples will be the extent to which $x$ meets the realized need
-$y$.  One way that a decision maker can try to anticipate the consequence of an action $x$ is by using a forecast $F$ of
-$Y$. The loss of the action $x$ chosen according to $F$ can, in turn, be interpreted as the value of $F$ as an input to
+$\mathcal{X}$ being the set of all feasible levels of investment, while an outcome $y$ might correspond to the number of
+individuals who eventually become sick and would benefit from the mitigation measure. A foundational strategy in decision
+theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numeric value, or
+\emph{loss}, measuring the (lack of) success of $x$. In our example, the loss measures the extent to which $x$ meets the realized need
+$y$, with greater unmet need incurring a larger loss.
+One way that a decision maker can try to anticipate the consequence of an action $x$ is by using a forecast $F$ of
+$Y$.
+This can help inform the selection of the action to take; for example, the decision maker might choose an action with the lowest expected loss according to the forecast.
+The loss of the action $x$ chosen according to $F$ can, in turn, be interpreted as the value of $F$ as an input to
 the decision making process under the realized outcome $y$ of $Y$.
 
-We arrive through this discussion at a three-step procedure for developing scoring rules for probabilistic forecasts:
+Through this discussion, we arrive at a three-step procedure for developing scoring rules for probabilistic forecasts:
 \begin{enumerate}
@@
 This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss
-function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} --- a terminology we find to
-be convenient but which is not, to our knowledge, standard.  In appendix \ref{sec:proper} we review the
-basic fact from the scoring rule literature (e.g., \cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring
+function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} (noting that to our knowledge, this terminology is not standard).  In appendix \ref{sec:proper} we review the
+result from the scoring rule literature (e.g., \cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring
 rules are proper by construction.
@@
 all possible allocations that satisfy these constraints. The eventually realized resource need in each location is
-denoted by $y = (y_1, \ldots, y_N)$. The future outcome that will determine the loss caused by choosing the allocation
-$x$ is the vector of resource needs $y = (y_1, \ldots, y_N)$ for each location which is a realization (unknown at the
+denoted by $y = (y_1, \ldots, y_N)$; this is a realization (unknown at the
 time of decision making) of the random vector $Y = (Y_1, \ldots, Y_N)$. A forecast of the random need $Y$ for each
 location is written as $F = (F_1, \ldots, F_N)$. We assume need is non-negative and the forecasts reflect that, i.e. the
-support of each $F_i$ is a subset of $\mathbb{R}^+$. Finally, we assume that any unmet unit of need in any location
+support of each $F_i$ is a subset of $\mathbb{R}^+$. Finally, we assume that each unmet unit of need in any location
 incurs a fixed loss of $L>0$, so that if the selected resource level $x_i$ in location $i$ is less than the realized
@@
 setup assumes that $t_d < t_r$ and that the resource in question can only meet the need realized at $t_r$, not prevent
-it. That is, we stipulate that the random vector $Y$ of need is independent of the chosen allocation $x$. In the
+it. That is, we stipulate that the choice of allocation $x$ has no effect on the distribution of the random vector $Y$. In the
 context of infectious disease, this means that we do not consider resources, such as vaccines, that are intended to
@@
 For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts.
-The side-by-side shaded regions show the median (solid red horizontal line) and 50\\%, 80\\% and 95\\% prediction intervals for the two selected models.
+The side-by-side shaded regions show the median (grey horizontal line) and 50\\%, 80\\% and 95\\% prediction intervals for the two selected models.
 The forecasts were made for new hospitalizations on January 3, 2022 (vertical dashed line, with number of hospitalizations indicated by red horizontal line segment at the intersection of the dashed line and the grey line of data).
@@
 
-A state-wise comparison using Figures \ref{fig:thermometer-plot} and \ref{fig:multi-loc-ranks}A of the
+A state-wise comparison of the
 \texttt{JHUAPL-Bucky} and \texttt{CU-select} models shows that while the \texttt{JHUAPL-Bucky} forecast distributions
 were more centered on the eventual observations in many states, the suggested allocations of the \texttt{JHUAPL-Bucky}
-model were less efficient than those of the \texttt{CU-select} model. Thus \texttt{JHUAPL-Bucky} achieved a worse
+model were less efficient than those of the \texttt{CU-select} model (Figures \ref{fig:thermometer-plot} and \ref{fig:multi-loc-ranks}A). Thus \texttt{JHUAPL-Bucky} achieved a worse
 allocation score than \texttt{CU-select} (on this particular week) by allocating excess resources to several states,
 such as Pennsylvania and Michigan, rather than to states such as Florida and California where these resources would have
-reduced unmet need.  These allocation errors of \texttt{JHUAPL-Bucky} resulted from forecasts failing to consistently
+reduced unmet need.  These allocation errors resulted from forecasts failing to consistently
 capture the relative resource needs across different states. The \texttt{CU-select} model made some similar errors —
@@
 The number of available resources was fixed at 15,000 and forecasts from each model were used to determine an optimal allocation strategy before the resource need was known.
-For most locations the resource need exceeded the resources allocated, indicated by the `observed resource need' bar being larger than the `allocation' bar.
+For most locations the resource need exceeded the resources allocated, indicated by some amount of `unmet resource need' above the `need met by allocated resources'.
 Panel B shows the breakdown of the weighted interval score (WIS) into components of underprediction, overprediction and dispersion.
@@
 Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states.
-The $x$-axis corresponds to the date of the observation that a model's prediction was targeting (e.g., the date the forecast was made plus the forecast horizon).
+The $x$-axis corresponds to the date of the observation that a model's prediction was targeting (i.e., the date the forecast was made plus the forecast horizon).
 AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down).
@@
 initial motivation for this project. We see it as promising focus for further development of the AS framework since it
-(a) has been identified as a prime manifestation of scarcity during respiratory outbreaks (b) tansparently involves
+(a) has been identified as a setting in which allocation decisions are made during respiratory disease outbreaks, (b) involves
 logistical considerations depending on geographical and population scales, and (c) would likely imply a central role for
 hospitalization data. But again, additional modeling and data synthesis would be required since not everyone who is
-hospitalized needs a ventilator, and new hospital admissions will not fully describe for a need that corresponds to a
-prevalence of severe cases. Other possible real-world examples include the allocation of a limited stockpile of
+hospitalized needs a ventilator. Other possible real-world examples include the allocation of a limited stockpile of
 vaccinations \citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests
@@
 
-In conclusion, we argue that the way modelers and policymakers view and evaluate forecasts should more explicitly depend on
+In conclusion, we argue that when possible, the way modelers and policymakers view and evaluate forecasts should more explicitly depend on
 the specific decision-making context. Defaulting to standard forecast evaluation metrics can mask the utility (or
@@
 the resource from location $i$ to location $j$ will lead to a net increase in expected shortage --- in other words, the
-reallocation increases the expected shortage in location $i$ is at least as much as it decreases the expected shortage
+reallocation increases the expected shortage in location $i$ at least as much as it decreases the expected shortage
 in location $j$:
@@
 
-In section \ref{sec:application} of the manuscript, we used the allocation score to evaluate forecasts of COVID-19
+In section \ref{sec:application}, we used the allocation score to evaluate forecasts of COVID-19
 hospitalizations that have been submitted to the US COVID-19 Forecast Hub. These forecasts are submitted to the Hub
@@
 a mechanism to approximate the full cumulative distribution functions $F_i$, $i = 1, \ldots, N$ based on the provided
-quantiles. We have developed functionality for this purpose in the \verb`distfromq` package for R.\citep{ray-distfromq}
+quantiles. We have developed functionality for this purpose in the \verb`distfromq` package for R \citep{ray-distfromq}.
 This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:

### 2024-02-22 alloscore_manuscript/alloscore-application-shorter.tex
remove ref to manuscript in appendix

@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Tue Feb 20 23:42:19 2024
+% Wed Feb 21 13:49:14 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Tue Feb 20 23:42:23 2024
+% Wed Feb 21 13:49:19 2024
 \begin{table}[ht]

### 2024-02-22 alloscore_manuscript/alloscore-application-shorter.Rnw
merging more elr edits

--- 2024-02-22 alloscore_manuscript/alloscore-application-shorter.Rnw
@@
 
-Decision theory is concerned with how to decide on an \emph{action} $x$ to select from a set of
-possible actions $\mathcal{X}$ in the face of uncertainty about the future.
-We use the random variable $Y$ to represent future states and refer to a
-realization $y$ of $Y$ as an \emph{outcome}. For example, $x$ might correspond to the level of investment in a measure that is
+Decision theory is concerned with understanding how an \emph{action} $x$ both is and ought to be selected from a set of
+possible actions $\mathcal{X}$ --- that is, how a decision is and ought to be made --- based on the anticipated
+consequences of $x$ under future states. We represent such future states as a random variable $Y$ and refer to a
+realization $y$ of $Y$ as an \emph{outcome}. In this paper $x$ will be related to the level of investment in a measure
 designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with
-$\mathcal{X}$ being the set of all feasible levels of investment, while an outcome $y$ might correspond to the number of
-individuals who eventually become sick and would benefit from the mitigation measure. A foundational strategy in decision
-theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numeric value, or
-\emph{loss}, measuring the (lack of) success of $x$. In our example, the loss measures the extent to which $x$ meets the realized need
-$y$, with greater unmet need incurring a larger loss.
-One way that a decision maker can try to anticipate the consequence of an action $x$ is by using a forecast $F$ of
-$Y$.
-This can help inform the selection of the action to take; for example, the decision maker might choose an action with the lowest expected loss according to the forecast.
-The loss of the action $x$ chosen according to $F$ can, in turn, be interpreted as the value of $F$ as an input to
+$\mathcal{X}$ being the set of all feasible levels of investment, while an outcome $y$ will correspond to the number of
+individuals who eventually become sick and would benefit from the mitigation measure. A basic strategy in decision
+theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numerical value, or
+\emph{loss}, measuring the success of $x$, which in our examples will be the extent to which $x$ meets the realized need
+$y$.  One way that a decision maker can try to anticipate the consequence of an action $x$ is by using a forecast $F$ of
+$Y$. The loss of the action $x$ chosen according to $F$ can, in turn, be interpreted as the value of $F$ as an input to
 the decision making process under the realized outcome $y$ of $Y$.
 
-Through this discussion, we arrive at a three-step procedure for developing scoring rules for probabilistic forecasts:
+We arrive through this discussion at a three-step procedure for developing scoring rules for probabilistic forecasts:
 \begin{enumerate}

### 2024-02-22 alloscore_manuscript/alloscore-application-shorter.Rnw
withing hunk acception (partly as git experiment)
a few more fixes

@@
 
-We arrive through this discussion at a three-step procedure for developing scoring rules for probabilistic forecasts:
+Through this discussion, we arrive at a three-step procedure for developing scoring rules for probabilistic forecasts:
 \begin{enumerate}
@@
 <<thermometer-plot, fig.height=7, fig.width=7.5, fig.cap="Probabilistic forecasts for new COVID-19 hospital admissions on January 3, 2022 and their suggested resource allocations for the states with the ten highest (eventually observed) hospitalization counts.
-For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts.
+For each state, the dark black line shows the data observed when the forecast was made, the grey line shows counts observed after the forecast was made, and the horizontal dark red segment shows the count observed on the target date.
 The side-by-side shaded regions show the median (grey horizontal line) and 50\\%, 80\\% and 95\\% prediction intervals for the two selected models.
@@
 In Panel B, the \\texttt{per-capita} allocation is drawn in a heavier solid line.
-Panel C shows the MWIS metric across weeks, averaged across all states.
-MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.">>=
+Panel C shows the MWIS  across weeks.
+MWIS tends to scale with the observed and predicted values, peaking just after the peak of the Omicron wave.">>=
 p2 <- ggplot(alloscore_data_summaries,

### 2024-02-22 alloscore_manuscript/alloscore-application-shorter.tex
a few more fixes

@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Wed Feb 21 13:49:14 2024
+% Thu Feb 22 16:13:50 2024
 \begin{table}[ht]
@@
 \includegraphics[width=\maxwidth]{figure/thermometer-plot-1} \caption[Probabilistic forecasts for new COVID-19 hospital admissions on January 3, 2022 and their suggested resource allocations for the states with the ten highest (eventually observed) hospitalization counts]{Probabilistic forecasts for new COVID-19 hospital admissions on January 3, 2022 and their suggested resource allocations for the states with the ten highest (eventually observed) hospitalization counts.
-For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts.
-The side-by-side shaded regions show the median (solid red horizontal line) and 50\%, 80\% and 95\% prediction intervals for the two selected models.
+For each state, the dark black line shows the data observed when the forecast was made, the grey line shows counts observed after the forecast was made, and the horizontal dark red segment shows the count observed on the target date.
+The side-by-side shaded regions show the median (grey horizontal line) and 50\%, 80\% and 95\% prediction intervals for the two selected models.
 The forecasts were made for new hospitalizations on January 3, 2022 (vertical dashed line, with number of hospitalizations indicated by red horizontal line segment at the intersection of the dashed line and the grey line of data).
@@
 Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states.
-The $x$-axis corresponds to the date of the observation that a model's prediction was targeting (e.g., the date the forecast was made plus the forecast horizon).
+The $x$-axis corresponds to the date of the observation that a model's prediction was targeting (i.e., the date the forecast was made plus the forecast horizon).
 AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down).
 In Panel B, the \texttt{per-capita} allocation is drawn in a heavier solid line.
-Panel C shows the MWIS metric across weeks, averaged across all states.
-MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.}\label{fig:metrics-over-time}
+Panel C shows the MWIS  across weeks.
+MWIS tends to scale with the observed and predicted values, peaking just after the peak of the Omicron wave.}\label{fig:metrics-over-time}
 \end{figure}
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Wed Feb 21 13:49:19 2024
+% Thu Feb 22 16:13:55 2024
 \begin{table}[ht]

### 2024-02-23 alloscore_manuscript/alloscore-application-shorter.Rnw
next decision theory intro iteration

@@
 
-Decision theory is concerned with understanding how an \emph{action} $x$ both is and ought to be selected from a set of
-possible actions $\mathcal{X}$ --- that is, how a decision is and ought to be made --- based on the anticipated
-consequences of $x$ under future states. We represent such future states as a random variable $Y$ and refer to a
-realization $y$ of $Y$ as an \emph{outcome}. In this paper $x$ will be related to the level of investment in a measure
-designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with
-$\mathcal{X}$ being the set of all feasible levels of investment, while an outcome $y$ will correspond to the number of
-individuals who eventually become sick and would benefit from the mitigation measure. A basic strategy in decision
+Decision theory investigates how decisions are made by formalizing a decision as the process of selecting an
+\emph{action} $x$ from a specified set $\mathcal{X}$ of potential actions while taking into consideration the possible
+consequences of $x$ under future states of the world. Let $Y$ be a quantifiable aspect of the future which will help to
+determine the consequences of $x$. $Y$ is a random variable inasmuch as it is indeterminate when the decision is made.
+We refer to a realization $y$ of $Y$ as an \emph{outcome}. When a decision maker assumes $Y$ to have a specified
+distribution $F_Y$ and $Y$ is the only aspect of the future relevant to the selection of $x$, the process is referred to
+as a decision under risk.  The selection of $x$ without a full specification of the distribution of $Y$ is referred to
+as a decision under uncertainty. Clearly almost any real-world decision will be a decision under uncertainty.  Yet many
+foundational results in decision theory were guided (as we will be) by the intuition that much can be learned about a
+decision making process by building a ``small world'' approximation of the real-world (after \cite{savage1954foundations})
+in which the beliefs of a decision
+maker are equivalent to a fully specified distribution of a $Y$ that exhaustiveley quantifies all relevant aspects of
+the future, converting a decision under uncertainty to a decison under risk.
+
+Examples of actions with which this paper is concerned are levels of investment in measures designed to mitigate severe
+disease outcomes such as hospital beds, ventilators, medication, or medical staff, with $\mathcal{X}$ being the set of
+all feasible levels of investment. An outcome $y$ determining the consequences of such an action might be the number of
+individuals who eventually become sick and would benefit from the mitigation measure. A fundamental strategy in decision
 theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numerical value, or
-\emph{loss}, measuring the success of $x$, which in our examples will be the extent to which $x$ meets the realized need
-$y$.  One way that a decision maker can try to anticipate the consequence of an action $x$ is by using a forecast $F$ of
-$Y$. The loss of the action $x$ chosen according to $F$ can, in turn, be interpreted as the value of $F$ as an input to
-the decision making process under the realized outcome $y$ of $Y$.
+\emph{loss}, measuring the (lack of) success of $x$. In our example, $x$ will succeed to the extent that it meets the
+realized need $y$, so that greater unmet need will incur greater loss.
+
+One tool a decision maker can use to anticipate the consequences of an action $x$ is a forecast
+$F$ of $Y$. And within a ``small world'' where $Y$ exhaustively quantifies the future, the decision maker might choose that
+$x$ which yields the lowest expected loss according to $F$. The loss of the action $x$ chosen according to $F$ can then, 
+in turn, be interpreted as the value of $F$ as an
+input to the decision making process under the realized outcome $y$ of $Y$.
 

### 2024-02-23 alloscore_manuscript/alloscore-application-shorter.tex
next decision theory intro iteration

@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Thu Feb 22 16:13:50 2024
+% Fri Feb 23 13:37:38 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Thu Feb 22 16:13:55 2024
+% Fri Feb 23 13:37:42 2024
 \begin{table}[ht]

### 2024-02-23 alloscore_manuscript/alloscore-application-shorter.Rnw
edits to text

@@
 determine the consequences of $x$. $Y$ is a random variable inasmuch as it is indeterminate when the decision is made.
-We refer to a realization $y$ of $Y$ as an \emph{outcome}. When a decision maker assumes $Y$ to have a specified
-distribution $F_Y$ and $Y$ is the only aspect of the future relevant to the selection of $x$, the process is referred to
-as a decision under risk.  The selection of $x$ without a full specification of the distribution of $Y$ is referred to
-as a decision under uncertainty. Clearly almost any real-world decision will be a decision under uncertainty.  Yet many
-foundational results in decision theory were guided (as we will be) by the intuition that much can be learned about a
-decision making process by building a ``small world'' approximation of the real-world (after \cite{savage1954foundations})
-in which the beliefs of a decision
-maker are equivalent to a fully specified distribution of a $Y$ that exhaustiveley quantifies all relevant aspects of
-the future, converting a decision under uncertainty to a decison under risk.
-
-Examples of actions with which this paper is concerned are levels of investment in measures designed to mitigate severe
+We refer to a realization $y$ of $Y$ as an \emph{outcome}. A fundamental strategy in decision
+theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numeric value, or
+\emph{loss}, measuring the (lack of) success of $x$.
+
+Some examples of actions are levels of investment in measures designed to mitigate severe
 disease outcomes such as hospital beds, ventilators, medication, or medical staff, with $\mathcal{X}$ being the set of
 all feasible levels of investment. An outcome $y$ determining the consequences of such an action might be the number of
-individuals who eventually become sick and would benefit from the mitigation measure. A fundamental strategy in decision
-theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numerical value, or
-\emph{loss}, measuring the (lack of) success of $x$. In our example, $x$ will succeed to the extent that it meets the
-realized need $y$, so that greater unmet need will incur greater loss.
-
-One tool a decision maker can use to anticipate the consequences of an action $x$ is a forecast
-$F$ of $Y$. And within a ``small world'' where $Y$ exhaustively quantifies the future, the decision maker might choose that
-$x$ which yields the lowest expected loss according to $F$. The loss of the action $x$ chosen according to $F$ can then, 
-in turn, be interpreted as the value of $F$ as an
+individuals who eventually become sick and would benefit from the mitigation measure. In our example, $x$ will succeed to
+the extent that it meets the realized need $y$, so that greater unmet need will incur greater loss.
+
+When a forecast $F$ of $Y$ is available, a decision maker might choose that action
+$x$ which yields the lowest expected loss according to $F$.
+The loss of the action $x$ chosen according to $F$ can then, in turn, be interpreted as the value of $F$ as an
 input to the decision making process under the realized outcome $y$ of $Y$.
+
+The strategy of minimizing the expected loss according to $F$ might be used if the
+decision maker trusted that $F$ was accurate and did not have any additional information that was not captured
+in $F$. However, here we use it only as a device for evaluating forecasts by examining the consequences of
+hypothetical decisions that would be made when using only $F$ as an input.
+We return to this point in the discussion.
 
@@
 We argue that the value of collaborative forecasting efforts, such as the COVID-19 Forecasting Hub, could be enhanced by
-refocussing the attention of modelers and public health officials involved in these efforts on the elicitation of
-forecasts designed to support specific public health decision problems, and moreover, the evaluation of forecasts with
-scoring rules that directly consider the success of forecasts in supporting these decisions. As evidence for this
+including evaluations of forecasts with
+scoring rules that directly consider the success of forecasts in supporting specific public health decisions. As evidence for this
 possibility, we have demonstrated how tying forecast evaluation to a specific decision problem (via the allocation

### 2024-02-23 alloscore_manuscript/alloscore-application-shorter.Rnw
adjust correlation plot

@@
 determine the consequences of $x$. $Y$ is a random variable inasmuch as it is indeterminate when the decision is made.
-We refer to a realization $y$ of $Y$ as an \emph{outcome}. A fundamental strategy in decision
-theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numeric value, or
-\emph{loss}, measuring the (lack of) success of $x$.
-
-Some examples of actions are levels of investment in measures designed to mitigate severe
-disease outcomes such as hospital beds, ventilators, medication, or medical staff, with $\mathcal{X}$ being the set of
-all feasible levels of investment. An outcome $y$ determining the consequences of such an action might be the number of
-individuals who eventually become sick and would benefit from the mitigation measure. In our example, $x$ will succeed to
-the extent that it meets the realized need $y$, so that greater unmet need will incur greater loss.
-
-When a forecast $F$ of $Y$ is available, a decision maker might choose that action
-$x$ which yields the lowest expected loss according to $F$.
-The loss of the action $x$ chosen according to $F$ can then, in turn, be interpreted as the value of $F$ as an
-input to the decision making process under the realized outcome $y$ of $Y$.
-
-The strategy of minimizing the expected loss according to $F$ might be used if the
-decision maker trusted that $F$ was accurate and did not have any additional information that was not captured
-in $F$. However, here we use it only as a device for evaluating forecasts by examining the consequences of
-hypothetical decisions that would be made when using only $F$ as an input.
-We return to this point in the discussion.
-
-Through this discussion, we arrive at a three-step procedure for developing scoring rules for probabilistic forecasts:
+We refer to a realization $y$ of $Y$ as an \emph{outcome}. A fundamental strategy in decision theory is to assume that
+the consequences of an action $x$ under outcome $y$ can be assigned a numeric value, or \emph{loss}, measuring the (lack
+of) success of $x$.
+
+Some examples of actions are levels of investment in measures designed to mitigate severe disease outcomes such as
+hospital beds, ventilators, medication, or medical staff, with $\mathcal{X}$ being the set of all feasible levels of
+investment. An outcome $y$ determining the consequences of such an action might be the number of individuals who
+eventually become sick and would benefit from the mitigation measure. In our example, $x$ will succeed to the extent
+that it meets the realized need $y$, so that greater unmet need will incur greater loss.
+
+When a forecast $F$ of $Y$ is available, a decision maker might choose that action $x$ which yields the lowest expected
+loss according to $F$. The loss of the action $x$ chosen according to $F$ can then, in turn, be interpreted as the value
+of $F$ as an input to the decision making process under the realized outcome $y$ of $Y$.
+
+The strategy of minimizing the expected loss according to $F$ might be used if the decision maker trusted that $F$ was
+accurate and did not have any additional information that was not captured in $F$. However, here we use it only as a
+device for evaluating forecasts by examining the consequences of hypothetical decisions that would be made when using
+only $F$ as an input. We return to this point in the discussion.
+
+We arrive at a three-step procedure for developing scoring rules for probabilistic forecasts:
 \begin{enumerate}
@@
 This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss
-function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} (noting that to our knowledge, this terminology is not standard).  In appendix \ref{sec:proper} we review the
+function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} (noting that to our knowledge, 
+this terminology is not standard).  In appendix \ref{sec:proper} we review the
 result from the scoring rule literature (e.g., \cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring
@@
 
-<<thermometer-plot, fig.height=7, fig.width=7.5, fig.cap="Probabilistic forecasts for new COVID-19 hospital admissions on January 3, 2022 and their suggested resource allocations for the states with the ten highest (eventually observed) hospitalization counts.
+<<thermometer-plot, fig.height=7, fig.pos = "H", fig.width=7.5, fig.cap="Probabilistic forecasts for new COVID-19 hospital admissions on January 3, 2022 and their suggested resource allocations for the states with the ten highest (eventually observed) hospitalization counts.
 For each state, the dark black line shows the data observed when the forecast was made, the grey line shows counts observed after the forecast was made, and the horizontal dark red segment shows the count observed on the target date.
@@
 
-<<metrics-correlation, fig.width=8, fig.height=9, fig.cap="Association of standardized ranks for MWIS and allocation score by model and week.
+<<metrics-correlation, fig.width=8, fig.height=6, fig.pos="H", fig.cap="Association of standardized ranks for MWIS and allocation score by model and week.
 Each facet of the plot corresponds to one model.
@@
   filter(as.Date(reference_date.x) < as.Date("2022-02-01"),
-         model != "per-capita") |>
+         model != "per-capita", model != "JHU_IDD-CovidSP") |>
   dplyr::mutate(target_end_date = as.Date(target_end_date)) |>
@@
     geom_point(alpha=.7) +
-    scale_x_continuous("MWIS standardized rank", limits=c(0,1)) +
-    scale_y_continuous("Allocation score standardized rank", limits=c(0,1)) +
+    scale_x_continuous("MWIS standardized rank", limits=c(0,1), 
+                       breaks = c(0,.5,1), labels = c("0",".5","1")) +
+    scale_y_continuous("Allocation score standardized rank", limits=c(0,1), 
+                       breaks = c(0,.5,1), labels = c("0",".5","1")) +
 #    scale_color_gradient("Target date") +
@@
     scale_size_continuous("Observed\nhospitalizations") +
-    facet_wrap(.~model, ncol=3) +
-    geom_abline(slope=1, intercept=0, linetype=2, color="gray") #+
-#    theme(legend.pos="bottom")
+    facet_wrap(.~model, ncol=4) +
+    geom_abline(slope=1, intercept=0, linetype=2, color="gray") +
+    theme(strip.text = element_text(size = 7))
 
@@
 
-<<multi-k, fig.height=4, fig.width = 8, fig.cap = "Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20.
+<<multi-k, fig.height=4, fig.width = 8, fig.pos="H", fig.cap = "Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20.
 Panel A shows, for each model, the AS for values of $K$ between 200 and 60,000 at increments of 200.

### 2024-02-23 alloscore_manuscript/alloscore-application-shorter.tex
adjust correlation plot

@@
 determine the consequences of $x$. $Y$ is a random variable inasmuch as it is indeterminate when the decision is made.
-We refer to a realization $y$ of $Y$ as an \emph{outcome}. When a decision maker assumes $Y$ to have a specified
-distribution $F_Y$ and $Y$ is the only aspect of the future relevant to the selection of $x$, the process is referred to
-as a decision under risk.  The selection of $x$ without a full specification of the distribution of $Y$ is referred to
-as a decision under uncertainty. Clearly almost any real-world decision will be a decision under uncertainty.  Yet many
-foundational results in decision theory were guided (as we will be) by the intuition that much can be learned about a
-decision making process by building a ``small world'' approximation of the real-world (after \cite{savage1954foundations})
-in which the beliefs of a decision
-maker are equivalent to a fully specified distribution of a $Y$ that exhaustiveley quantifies all relevant aspects of
-the future, converting a decision under uncertainty to a decison under risk.
-
-Examples of actions with which this paper is concerned are levels of investment in measures designed to mitigate severe
-disease outcomes such as hospital beds, ventilators, medication, or medical staff, with $\mathcal{X}$ being the set of
-all feasible levels of investment. An outcome $y$ determining the consequences of such an action might be the number of
-individuals who eventually become sick and would benefit from the mitigation measure. A fundamental strategy in decision
-theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numerical value, or
-\emph{loss}, measuring the (lack of) success of $x$. In our example, $x$ will succeed to the extent that it meets the
-realized need $y$, so that greater unmet need will incur greater loss.
-
-One tool a decision maker can use to anticipate the consequences of an action $x$ is a forecast
-$F$ of $Y$. And within a ``small world'' where $Y$ exhaustively quantifies the future, the decision maker might choose that
-$x$ which yields the lowest expected loss according to $F$. The loss of the action $x$ chosen according to $F$ can then, 
-in turn, be interpreted as the value of $F$ as an
-input to the decision making process under the realized outcome $y$ of $Y$.
-
-Through this discussion, we arrive at a three-step procedure for developing scoring rules for probabilistic forecasts:
+We refer to a realization $y$ of $Y$ as an \emph{outcome}. A fundamental strategy in decision theory is to assume that
+the consequences of an action $x$ under outcome $y$ can be assigned a numeric value, or \emph{loss}, measuring the (lack
+of) success of $x$.
+
+Some examples of actions are levels of investment in measures designed to mitigate severe disease outcomes such as
+hospital beds, ventilators, medication, or medical staff, with $\mathcal{X}$ being the set of all feasible levels of
+investment. An outcome $y$ determining the consequences of such an action might be the number of individuals who
+eventually become sick and would benefit from the mitigation measure. In our example, $x$ will succeed to the extent
+that it meets the realized need $y$, so that greater unmet need will incur greater loss.
+
+When a forecast $F$ of $Y$ is available, a decision maker might choose that action $x$ which yields the lowest expected
+loss according to $F$. The loss of the action $x$ chosen according to $F$ can then, in turn, be interpreted as the value
+of $F$ as an input to the decision making process under the realized outcome $y$ of $Y$.
+
+The strategy of minimizing the expected loss according to $F$ might be used if the decision maker trusted that $F$ was
+accurate and did not have any additional information that was not captured in $F$. However, here we use it only as a
+device for evaluating forecasts by examining the consequences of hypothetical decisions that would be made when using
+only $F$ as an input. We return to this point in the discussion.
+
+We arrive at a three-step procedure for developing scoring rules for probabilistic forecasts:
 \begin{enumerate}
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Feb 23 13:37:38 2024
+% Fri Feb 23 18:40:36 2024
 \begin{table}[ht]
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
 \includegraphics[width=\maxwidth]{figure/thermometer-plot-1} \caption[Probabilistic forecasts for new COVID-19 hospital admissions on January 3, 2022 and their suggested resource allocations for the states with the ten highest (eventually observed) hospitalization counts]{Probabilistic forecasts for new COVID-19 hospital admissions on January 3, 2022 and their suggested resource allocations for the states with the ten highest (eventually observed) hospitalization counts.
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Feb 23 13:37:42 2024
+% Fri Feb 23 18:40:41 2024
 \begin{table}[ht]
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
 \includegraphics[width=\maxwidth]{figure/metrics-correlation-1} \caption[Association of standardized ranks for MWIS and allocation score by model and week]{Association of standardized ranks for MWIS and allocation score by model and week.
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
 \includegraphics[width=\maxwidth]{figure/multi-k-1} \caption[Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20]{Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20.

### 2024-02-24 alloscore_manuscript/alloscore-application-shorter.Rnw
author fix, save some space in appendix

@@
 \usepackage{appendix}
+\usepackage{authblk}
+\usepackage{authblk}
 \usepackage{amsmath, amsfonts, amssymb}
@@
 
+\author{Aaron Gerding\thanks{Corresponding Author: \texttt{agerding@umass.edu}}}
+\author{Nicholas G. Reich}
+\author{Benjamin Rogers}
+\author{Evan L. Ray}
+\affil{Department of Biostatistics and Epidemiology, School of Public Health and Health
+Sciences, University of Massachusetts at Amherst}
+
 \title{Evaluating infectious disease forecasts \\ with allocation scoring rules}
-\author{Aaron Gerding, Nicholas G. Reich, Benjamin Rogers, Evan L. Ray}
 
@@
 If the parametric family used to represent forecast distributions is flexible enough, the reporting scoring rule $S_R$
-and the allocation score are equivalent in the sense that they will yield the same score for any distribution $F$.
+and the allocation score will coincide.
 Suppose that for a given resource constraint $K$, for any forecast distribution $F$ it is possible to find a member
@@
 \end{align*}
-It therefore follows that
-\begin{align*}
-S_R(F, y) &= s_R(\theta^\star, y) \\
-&= s_A(x^{G_{\theta^\star}}, y) \\
-&= s_A(x^F, y) \\
-&= S_A(F, y).
-\end{align*}
+Thus $S_R(F, y) = s_R(\theta^\star, y) 
+= s_A(x^{G_{\theta^\star}}, y) 
+= s_A(x^F, y) 
+= S_A(F, y)$.
 

### 2024-02-24 alloscore_manuscript/alloscore-application-shorter.tex
author fix, save some space in appendix

@@
 
+\author{Aaron Gerding\thanks{Corresponding Author: \texttt{agerding@umass.edu}}}
+\author{Nicholas G. Reich}
+\author{Benjamin Rogers}
+\author{Evan L. Ray}
+\affil{Department of Biostatistics and Epidemiology, School of Public Health and Health
+Sciences, University of Massachusetts at Amherst}
+
 \title{Evaluating infectious disease forecasts \\ with allocation scoring rules}
-\author{Aaron Gerding, Nicholas G. Reich, Benjamin Rogers, Evan L. Ray}
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Feb 23 18:40:36 2024
+% Fri Feb 23 19:04:47 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Feb 23 18:40:41 2024
+% Fri Feb 23 19:04:51 2024
 \begin{table}[ht]

### 2024-02-26 alloscore_manuscript/alloscore-application-shorter.Rnw

@@
 rank. This contrasts with a clearly positive correlation between MWIS and AS ranks in other models such as
-\texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha}. We also can see in more detail the consistently high performance
+\texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha}. We also can see in more detail the consistently high performance of the
 \texttt{COVIDhub-ensemble} on both scores. And while it did not submit forecasts for enough weeks to be included in
@@
 scoring rules that directly consider the success of forecasts in supporting specific public health decisions. As evidence for this
-possibility, we have demonstrated how tying forecast evaluation to a specific decision problem (via the allocation
+possibility, we have demonstrated how tying forecast evaluation to a specific decision problem (e.g., via the allocation
 score) can yield model rankings that differ substantively from those based a current standard measure of forecast

### 2024-02-27 alloscore_manuscript/alloscore-application-shorter.Rnw
reorg to get all appendices referenced incrementally in main text.

--- 2024-02-24 alloscore_manuscript/alloscore-application-shorter.Rnw
@@
 section \ref{sec:methods.overview}. In section \ref{sec:methods.detailed} we develop the allocation score more
-carefully, using a decision theoretic procedure for deriving proper scoring rules. (See appendix \ref{sec:proper}
+carefully, using a decision theoretic procedure for deriving proper scoring rules. (See appendix \ref{sec:a:proper}
 for a definition of a proper scoring rule and a more technical discussion of the procedure). In section
@@
 that the optimal allocation divides the available resources among the locations proportionally to the scale parameters
-$\sigma_i$ (see appendices \ref{sec:bayes-quantiles}). If $K = 5$ units of the resource are
+$\sigma_i$ (see appendix \ref{sec:a:bayes-quantiles}). If $K = 5$ units of the resource are
 available, the optimal allocation according to $F$ would be 1 and 4 resource units in locations 1 and 2, respectively. 
@@
 function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} (noting that to our knowledge, 
-this terminology is not standard).  In appendix \ref{sec:proper} we review the
+this terminology is not standard).  In appendix \ref{sec:a:proper} we review the
 result from the scoring rule literature (e.g., \cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring
@@
   \underset{0 \leq x}{\mathrm{minimize}}\,\, \mathbb{E}_{F} [s_A(x, Y)] \text{ subject to }
-  \, \sum_{i=1}^N x_i = K, \label{AP}
+  \, \sum_{i=1}^N x_i = K, \label{eqn:AP}
 \end{align}
@@
 
-In appendix \ref{sec:bayes-quantiles} we show that the components of the Bayes act are quantiles 
+In appendix \ref{sec:a:bayes-quantiles} we show that the components of the Bayes act are quantiles 
 $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource
@@
 As described above, predictions were submitted to the Forecast Hub in the form of a set of 23 quantiles of the
-predictive distribution. The WIS can be directly calculated from these quantiles. However, calculation of the AS
-requires that the full cumulative distribution functions of the forecast distributions for each location are available.
-For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution
-functions based on the provided quantiles. On the interior of the provided quantiles we used a monotonic cubic spline to
-interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as
-to match the two lowest and two highest quantiles (see appendix \ref{sec:distfromq} for further details).
-In appendix \ref{sec:propriety_of_parametric_approximation}, we show that if this evaluation procedure
+predictive distribution. The WIS can be directly calculated from these quantiles. However, our numerical method for calculating 
+the AS (outlined in appendix \ref{sec:a:numeric})
+requires a full cumulative distribution function (CDF) for the forecast in each location.
+For the purpose of this analysis, we imputed CDFs based on the provided quantiles following a procedure detailed in 
+appendix \ref{sec:a:distfromq}. Briefly, this involves interpolating the provided quantiles between the lowest (.01) and highest 
+(.99) probability levels using monotonic cubic splinea, and then extending outside of these level with normal distribution tails
+parametrized to match the two lowest and two highest quantiles. In appendix \ref{sec:a:propriety_of_parametric_approximation}, 
+we show that if this evaluation procedure
 had been specified prospectively, the resulting score would be proper, \textemdash but that a post hoc application of this
@@
 of which are done in our application example.  We consider these issues briefly in appendices
-\ref{sec:distfromq} and \ref{sec:propriety_of_parametric_approximation}.)
+\ref{sec:a:distfromq} and \ref{sec:a:propriety_of_parametric_approximation}.)
 
@@
 
-\section{Introduction}
-\label{sec:intro}
-
-We address some technical and methodological points from the main text. We begin in section \ref{sec:proper} by defining
-proper scoring rules and showing that the allocation score is proper. In section \ref{sec:ex-shortage}, we formalize the
-concept of a \emph{shortage} of resources, give some key results about expected resource shortages under a distribution
-characterizing uncertainty about (future) levels of resource need, and illustrate how quantiles arise as the Bayes act
-in a decision making problem about the quantity of a resource to purchase. Section \ref{sec:bayes-quantiles} gives a
+We address some technical and methodological points from the main text. We begin in section \ref{sec:a:proper} by defining
+proper scoring rules and showing that the allocation score is proper. Section \ref{sec:a:bayes-quantiles} gives a
 justification for the result that the Bayes act for the allocation problem is given by a vector of quantiles of the
@@
 \ref{sec:methods.detailed.specific_allocation}. We describe the algorithm that we use to compute allocations given a
-forecast distribution in each location in section \ref{sec:numeric}. In section \ref{sec:distfromq}, we describe the
+forecast distribution in each location in section \ref{sec:a:numeric}. In section \ref{sec:a:distfromq}, we describe the
 methods for approximating a full distribution from a set of quantiles, implemented in the R package \verb`distfromq`,
@@
 Forecast Hub for the application in section \ref{sec:application}. Finally, in section
-\ref{sec:propriety_of_parametric_approximation} we examine implications for propriety of an analysis that uses summaries
+\ref{sec:a:propriety_of_parametric_approximation} we examine implications for propriety of an analysis that uses summaries
 of forecasts (such as predictive quantiles) rather than the full forecast distributions for computation of allocation
@@
 \section{Proper scoring rules}
-\label{sec:proper}
+\label{sec:a:proper}
 
@@
 \subsection{The allocation score is proper}
-\label{sec:alloscore_proper}
+\label{sec:a:alloscore_proper}
 
@@
 
-\section{Expected shortages}
-\label{sec:ex-shortage}
-
-A key feature of loss functions for decision problems used to define quantiles and related scoring rules such as the
+\section{Allocation Bayes acts as vectors of marginal quantiles.}
+\label{sec:a:bayes-quantiles}
+
+Here we study the form of the Bayes act for the allocation problem (AP) (equation \eqref{eqn:AP} in section
+\ref{sec:methods.detailed.specific_allocation}) of the text:
+\begin{align}
+    \underset{0 \leq x}{\mathrm{minimize}}\,\, 
+    \mathbb{E}_{F} [s_A(x, Y)]= \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]
+     \text{ subject to }
+     \, \sum_{i=1}^N x_i = K, \label{eqn:AP-long}
+\end{align}
+where the marginal forecasts $F_i$ for $i=1,\dots,N$ represent forecasts for $N$ distinct locations. We show that the
+Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$ and resource constraint level $K$ is a vector of
+quantiles of the marginal forecast distributions $F_i$ at a single probability level $\tau^{F,K}$, that is, $x_i^{F,K} =
+q_{F_i,\tau^{F,K}}$. An immediate consequence used in the examples in Section \ref {sec:methods.overview} is that if
+$F_i = \mathrm{Exp}(1/\sigma_i)$ for all $i$, then the Bayes act is proportional to $(\sigma_1,\ldots,\sigma_N)$, since
+$q_{\mathrm{Exp}(1/\sigma),\tau} = -\sigma \log(1-\tau)$.
+
+We begin by noting that a key feature of each term of the loss function $s_A(x, Y)$ defining the AP is the presence of 
+a \emph{shortage}: an amount $\max\{0,y-x\}$ by which a resource demand $y$ exceeds a supply decision variable $x$,
+which, for convenience, we write as $(y-x)_{+}$.  This is a feature shared with decision problems used to define quantiles 
+and related scoring rules such as the
 CRPS and the WIS (see e.g., \cite{gneiting2011quantiles}, \cite{jose2009evaluating}, and \cite{royset2022optimization},
-sections 1.C and 3.C), as well as the allocation loss function presented in this work, is the presence of a
-\emph{shortage}: the amount $\max\{0,y-x\}$ by which a resource demand $y$ exceeds a supply decision variable $x$,
-which, for convenience, we write as $(y-x)_{+}$. In particular, a quantile at probability level $\alpha$ of a
-distribution $F$ on $\mathbb{R}^1$ (which we assume to have a well-defined density $f(x)$) is a Bayes act for the loss
+sections 1.C and 3.C). In particular, 
+a quantile at probability level $\alpha$ of a
+distribution $F$ on $\mathbb{R}$ (which we assume to have a well-defined density $f(x)$) is a Bayes act for the loss
 function
@@
 Note that more care is required when $F$ does not have a density.
-We will also use this result below in deriving the Bayes act for the allocation loss.
-
-\section{Allocation Bayes acts as vectors of marginal quantiles.}
-\label{sec:bayes-quantiles}
-
-Here we study the form of the Bayes act for the allocation problem (AP) (equation \eqref{eqn:loss_fn} in section
-\ref{sec:methods.detailed.specific_allocation}) of the text:
-\begin{align}
-    \underset{0 \leq x}{\mathrm{minimize}}\,\, \mathbb{E}_{F} [s_A(x, Y)]= \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[(Y_i - x_i)_{+}]
-     \text{ subject to }
-     \, \sum_{i=1}^N x_i = K, \label{AP-long}
-\end{align}
-where the marginal forecasts $F_i$ for $i=1,\dots,N$ represent forecasts for $N$ distinct locations. We show that the
-Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$ and resource constraint level $K$ is a vector of
-quantiles of the marginal forecast distributions $F_i$ at a single probability level $\tau^{F,K}$, that is, $x_i^{F,K} =
-q_{F_i,\tau^{F,K}}$. An immediate consequence used in the examples in Section \ref {sec:methods.overview} is that if
-$F_i = \mathrm{Exp}(1/\sigma_i)$ for all $i$, then the Bayes act is proportional to $(\sigma_1,\ldots,\sigma_N)$, since
-$q_{\mathrm{Exp}(1/\sigma),\tau} = -\sigma \log(1-\tau)$.
-
-In order for $x^{\star} \in \mathbb{R}^N_{+}$ to solve the AP it must be true that reallocating $\delta > 0$ units of
+
+Returning to the AP \eqref{eqn:AP-long}, notice that in order for $x^{\star} \in \mathbb{R}^N_{+}$ to be a Bayes act it must be true that reallocating $\delta > 0$ units of
 the resource from location $i$ to location $j$ will lead to a net increase in expected shortage --- in other words, the
@@
 &\mathbb{E}_{F_i}[(Y_i - (x^{\star}_i - \delta))_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}]
-\text{ (increase in $i$) } \nonumber \\
+\text{ (detriment in $i$) } \nonumber \\
 &\qquad \geq
 \mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - (x^{\star}_j + \delta))_{+}]
-\text{ (decrease in $j$) }.
+\text{ (benefit in $j$) }.
 \end{align*}
@@
 \begin{align}
-1-F_i(x^{\star}_i) &= -\frac{d}{dx_i}\mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}] \nonumber \\
+1-F_i(x^{\star}_i) &= -\frac{d}{dx_i}\mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}] 
+\text{ (rate of detriment in $i$) }\nonumber \\
 &= \lim_{\delta \searrow 0} \frac{1}{\delta}
 \left\{\mathbb{E}_{F_i}[(Y_i - (x^{\star}_i - \delta))_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}]\right\}
-\text{ (increase in $i$) } \nonumber \\
+\nonumber \\
 &\geq
@@
 \left\{\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - (x^{\star}_j + \delta))_{+}]\right\}
-\text{ (decrease in $j$) } \nonumber \\
- &= -\dby{}{x_j}\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] = 1-F_j(x^{\star}_j) \label{eqn:ASoptimal1}
+ \nonumber \\
+ &= -\dby{}{x_j}\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] = 1-F_j(x^{\star}_j) 
+ \text{ (rate of benefit in $j$).}\label{eqn:ASoptimal1}
 \end{align}
-
-Note that negative derivatives appear because our optimality condition addresses how a \emph{decrease} in resources will
-\emph{increase} the expected shortage in $i$ and vice versa in $j$. Since \eqref{eqn:ASoptimal1} also holds with $i$ and $j$
+(Negative derivatives appear here because our optimality condition addresses how a \emph{decrease} in resources will
+ \emph{increase} the expected shortage in $i$ and vice versa in $j$.)
+
+Since \eqref{eqn:ASoptimal1} also holds with $i$ and $j$
 reversed, a number $\lambda$ (a \emph{Lagrange multiplier}) exists such that
@@
 \section{Numerical computation of allocation Bayes acts}
-\label{sec:numeric}
+\label{sec:a:numeric}
 
 To compute an allocation score $S_A(F,y;K) := s_A(x^{F,K},y)$, we require numerical values for a Bayes act solving the
-AP \eqref{AP-long} --- that is, we must find the specific resource allocations for each location that are determined by
+AP \eqref{eqn:AP-long} --- that is, we must find the specific resource allocations for each location that are determined by
 the forecast $F$ under the resource constraint $K$. Assuming we have reliable means of calculating quantiles
@@
 \section{Computing allocations from finite quantile forecast representations}
-\label{sec:distfromq}
+\label{sec:a:distfromq}
 
@@
 
-To address this situation and support the numerical allocation algorithm outlined in section \ref{sec:numeric}, we need
+To address this situation and support the numerical allocation algorithm outlined in section \ref{sec:a:numeric}, we need
 a mechanism to approximate the full cumulative distribution functions $F_i$, $i = 1, \ldots, N$ based on the provided
@@
 \section{Propriety of parametric approximation} % (fold)
-\label{sec:propriety_of_parametric_approximation}
+\label{sec:a:propriety_of_parametric_approximation}
 
@@
 $$S_R(F, y) = s_R(\theta^\star(F), y) = s_A(x^{G_{\theta^\star(F)}}, y).$$
-It follows from the discussion in section \ref{sec:alloscore_proper} that this is a proper scoring rule for $F$.
+It follows from the discussion in section \ref{sec:a:alloscore_proper} that this is a proper scoring rule for $F$.
 Although the full forecast distribution $F$ is not available in the forecast submission, the score $S_R(F, y)$ can be
@@
 We define
-$$G^{\star}(F) := \argmin_{G \in \mathcal{P}_{\mathrm{dfq}}} E_{F}[S_A(G, Y)].$$
+$$G^{\star}(F) := \argmin_{G \in \mathcal{P}_{sec:propriety_of_parametric_approximation\mathrm{dfq}}} E_{F}[S_A(G, Y)].$$
 Since $S_R$ is defined as the Bayes scoring rule for the forecaster's loss \eqref{eqn:forecaster_theta_loss}, $G^{\star}(F)$ coincides

### 2024-02-27 alloscore_manuscript/alloscore-application-shorter.tex
reorg to get all appendices referenced incrementally in main text.

@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Feb 23 19:04:47 2024
+% Tue Feb 27 15:06:55 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Feb 23 19:04:51 2024
+% Tue Feb 27 15:06:59 2024
 \begin{table}[ht]
@@
 \left\{\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - (x^{\star}_j + \delta))_{+}]\right\}
-\text{ (decrease in $j$) } \nonumber \\
- &= -\dby{}{x_j}\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] = 1-F_j(x^{\star}_j) \label{eqn:ASoptimal1}
-\end{align}
-
-Note that negative derivatives appear because our optimality condition addresses how a \emph{decrease} in resources will
-\emph{increase} the expected shortage in $i$ and vice versa in $j$. Since \eqref{eqn:ASoptimal1} also holds with $i$ and $j$
+ \nonumber \\
+ &= -\dby{}{x_j}\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] = 1-F_j(x^{\star}_j) 
+ \text{ (rate of benefit in $j$).}\label{eqn:ASoptimal1}
+\end{align}
+(Negative derivatives appear here because our optimality condition addresses how a \emph{decrease} in resources will
+ \emph{increase} the expected shortage in $i$ and vice versa in $j$.)
+
+Since \eqref{eqn:ASoptimal1} also holds with $i$ and $j$
 reversed, a number $\lambda$ (a \emph{Lagrange multiplier}) exists such that

### 2024-02-27 arxivsub2/submission.tex
add arxiv2 sub

--- 2024-02-24 alloscore_manuscript/alloscore-application-shorter.tex
@@
 \begin{figure}
-    \includegraphics[width=\textwidth]{../figures/exponential_pred_expected_loss.pdf}
+    \includegraphics[width=\textwidth]{figure/exponential_pred_expected_loss.pdf}
     \caption{An illustration of the resource allocation problem in Example 1. There are $N = 2$ locations, with
@@
 
-\bibliography{allocation}
+\begin{thebibliography}{51}
+\providecommand{\natexlab}[1]{#1}
+\providecommand{\url}[1]{\texttt{#1}}
+\expandafter\ifx\csname urlstyle\endcsname\relax
+  \providecommand{\doi}[1]{doi: #1}\else
+  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
+
+\bibitem[Igboh et~al.(2023)Igboh, Roguski, Marcenac, Emukule, Charles, Tempia,
+  Herring, Vandemaele, Moen, Olsen, et~al.]{igboh2023timing}
+Ledor~S Igboh, Katherine Roguski, Perrine Marcenac, et~al.
+\newblock Timing of seasonal influenza epidemics for 25 countries in {A}frica
+  during 2010--19: a retrospective analysis.
+\newblock \emph{The Lancet Global Health}, 11\penalty0 (5):\penalty0
+  e729--e739, 2023.
+
+\bibitem[Meltzer et~al.(2014)Meltzer, Atkins, Santibanez, Knust, Petersen,
+  Ervin, Nichol, Damon, and Washington]{meltzer2014estimating}
+Martin~I Meltzer, Charisma~Y Atkins, Scott Santibanez, et~al.
+\newblock Estimating the future number of cases in the {E}bola
+  epidemic--{L}iberia and {S}ierra {L}eone, 2014--2015.
+\newblock \emph{MMWR}, 63:\penalty0 1--14, Sep 2014.
+
+\bibitem[Rainisch et~al.(2015)Rainisch, Shankar, Wellman, Merlin, and
+  Meltzer]{rainisch2015regional}
+Gabriel Rainisch, Manjunath Shankar, Michael Wellman, et~al.
+\newblock Regional spread of {E}bola virus, {W}est {A}frica, 2014.
+\newblock \emph{Emerging Infectious Diseases}, 21\penalty0 (3):\penalty0 444,
+  2015.
+
+\bibitem[Camacho et~al.(2015)Camacho, Kucharski, Aki-Sawyerr, White, Flasche,
+  Baguelin, Pollington, Carney, Glover, Smout, et~al.]{camacho2015-ebola-bed}
+Anton Camacho, Adam Kucharski, Yvonne Aki-Sawyerr, et~al.
+\newblock Temporal changes in {E}bola transmission in {S}ierra {L}eone and
+  implications for control requirements: a real-time modelling study.
+\newblock \emph{PLOS Currents}, 7, 2015.
+
+\bibitem[Finger et~al.(2019)Finger, Funk, White, Siddiqui, Edmunds, and
+  Kucharski]{finger_real-time_2019}
+Flavio Finger, Sebastian Funk, Kate White, et~al.
+\newblock Real-time analysis of the diphtheria outbreak in forcibly displaced
+  {Myanmar} nationals in {Bangladesh}.
+\newblock \emph{BMC Medicine}, 17\penalty0 (1):\penalty0 58, March 2019.
+\newblock \doi{10.1186/s12916-019-1288-7}.
+
+\bibitem[Bertsimas et~al.(2021)Bertsimas, Boussioux, Cory-Wright, Delarue,
+  Digalakis, Jacquillat, Kitane, Lukin, Li, Mingardi, Nohadani, Orfanoudaki,
+  Papalexopoulos, Paskov, Pauphilet, Lami, Stellato, Bouardi, Carballo, Wiberg,
+  and Zeng]{bertsimas2021predictionsCOVID}
+Dimitris Bertsimas, Leonard Boussioux, Ryan Cory-Wright, et~al.
+\newblock From predictions to prescriptions: A data-driven response to
+  {COVID}-19.
+\newblock \emph{Health Care Management Science}, 24:\penalty0 253--272, 2021.
+
+\bibitem[Fox et~al.(2022)Fox, Lachmann, Tec, Pasco, Woody, Du, Wang, Ingle,
+  Javan, Dahan, Gaither, Escott, Adler, Johnston, Scott, and
+  Meyers]{fox_real-time_2022}
+Spencer~J. Fox, Michael Lachmann, Mauricio Tec, et~al.
+\newblock Real-time pandemic surveillance using hospital admissions and
+  mobility data.
+\newblock \emph{Proceedings of the National Academy of Sciences}, 119\penalty0
+  (7):\penalty0 e2111870119, February 2022.
+\newblock \doi{10.1073/pnas.2111870119}.
+
+\bibitem[{University of Texas at Austin}(2022)]{utnews2022}
+{University of Texas at Austin}.
+\newblock {COVID} forecasting method using hospital and cellphone data proves
+  it can reliably guide us cities through pandemic threats.
+\newblock Available at
+  \url{https://news.utexas.edu/2022/02/02/covid-forecasting-method-using-hospital-and-cellphone-data-proves-it-can-reliably-guide-us-cities-through-pandemic-threats/}
+  (2023/05/26), February 2022.
+
+\bibitem[Marshall et~al.(2023)Marshall, Parker, and
+  Gardner]{marshall2023predictions}
+Maximilian Marshall, Felix Parker, and Lauren~Marie Gardner.
+\newblock When are predictions useful? {A} new method for evaluating epidemic
+  forecasts.
+\newblock \emph{medRxiv}, 2023.
+\newblock \doi{10.1101/2023.06.29.23292042}.
+
+\bibitem[Bilinski et~al.(2023)Bilinski, Salomon, and
+  Hatfield]{bilinski_adaptive_2023}
+Alyssa~M. Bilinski, Joshua~A. Salomon, and Laura~A. Hatfield.
+\newblock Adaptive metrics for an evolving pandemic: {A} dynamic approach to
+  area-level {COVID}-19 risk designations.
+\newblock \emph{Proceedings of the National Academy of Sciences}, 120\penalty0
+  (32):\penalty0 e2302528120, August 2023.
+\newblock \doi{10.1073/pnas.2302528120}.
+
+\bibitem[Papastefanopoulos et~al.(2020)Papastefanopoulos, Linardatos, and
+  Kotsiantis]{papastefanopoulos2020covid}
+Vasilis Papastefanopoulos, Pantelis Linardatos, and Sotiris Kotsiantis.
+\newblock {COVID-19}: a comparison of time series methods to forecast
+  percentage of active cases per population.
+\newblock \emph{Applied Sciences}, 10\penalty0 (11):\penalty0 3880, 2020.
+
+\bibitem[Johansson et~al.(2016)Johansson, Reich, Hota, Brownstein, and
+  Santillana]{johansson2016evaluating}
+Michael~A Johansson, Nicholas~G Reich, Aditi Hota, et~al.
+\newblock Evaluating the performance of infectious disease forecasts: A
+  comparison of climate-driven and seasonal dengue forecasts for mexico.
+\newblock \emph{Scientific reports}, 6\penalty0 (1):\penalty0 33707, 2016.
+
+\bibitem[McGowan et~al.(2019)McGowan, Biggerstaff, Johansson, Apfeldorf,
+  Ben-Nun, Brooks, Convertino, Erraguntla, Farrow, Freeze, Ghosh, Hyun,
+  Kandula, Lega, Liu, Michaud, Morita, Niemi, Ramakrishnan, Ray, Reich, Riley,
+  Shaman, Tibshirani, Vespignani, Zhang, and Reed]{mcgowan_collaborative_2019}
+Craig~J. McGowan, Matthew Biggerstaff, Michael Johansson, et~al.
+\newblock Collaborative efforts to forecast seasonal influenza in the {United}
+  {States}, 2015–2016.
+\newblock \emph{Scientific Reports}, 9\penalty0 (1):\penalty0 683, January
+  2019.
+\newblock \doi{10.1038/s41598-018-36361-9}.
+
+\bibitem[Reich et~al.(2019)Reich, Brooks, Fox, Kandula, McGowan, Moore, Osthus,
+  Ray, Tushar, Yamana, Biggerstaff, Johansson, Rosenfeld, and
+  Shaman]{reich_collaborative_2019}
+Nicholas~G. Reich, Logan~C. Brooks, Spencer~J. Fox, et~al.
+\newblock A collaborative multiyear, multimodel assessment of seasonal
+  influenza forecasting in the {United} {States}.
+\newblock \emph{Proceedings of the National Academy of Sciences}, 116\penalty0
+  (8):\penalty0 3146--3154, February 2019.
+\newblock \doi{10.1073/pnas.1812594116}.
+
+\bibitem[Johansson et~al.(2019)Johansson, Apfeldorf, Dobson, Devita, Buczak,
+  Baugher, Moniz, Bagley, Babin, Guven, Yamana, Shaman, Moschou, Lothian, Lane,
+  Osborne, Jiang, Brooks, Farrow, Hyun, Tibshirani, Rosenfeld, Lessler, Reich,
+  Cummings, Lauer, Moore, Clapham, Lowe, Bailey, García-Díez, Carvalho,
+  Rodó, Sardar, Paul, Ray, Sakrejda, Brown, Meng, Osoba, Vardavas, Manheim,
+  Moore, Rao, Porco, Ackley, Liu, Worden, Convertino, Liu, Reddy, Ortiz,
+  Rivero, Brito, Juarrero, Johnson, Gramacy, Cohen, Mordecai, Murdock, Rohr,
+  Ryan, Stewart-Ibarra, Weikel, Jutla, Khan, Poultney, Colwell, Rivera-García,
+  Barker, Bell, Biggerstaff, Swerdlow, Mier-Y-Teran-Romero, Forshey, Trtanj,
+  Asher, Clay, Margolis, Hebbeler, George, and Chretien]{johansson_open_2019}
+Michael~A. Johansson, Karyn~M. Apfeldorf, Scott Dobson, et~al.
+\newblock An open challenge to advance probabilistic forecasting for dengue
+  epidemics.
+\newblock \emph{Proceedings of the National Academy of Sciences}, 116\penalty0
+  (48):\penalty0 24268--24274, November 2019.
+\newblock \doi{10.1073/pnas.1909865116}.
+
+\bibitem[Bracher et~al.(2021)Bracher, Ray, Gneiting, and
+  Reich]{bracher2021evaluating}
+Johannes Bracher, Evan~L Ray, Tilmann Gneiting, and Nicholas~G Reich.
+\newblock Evaluating epidemic forecasts in an interval format.
+\newblock \emph{PLOS Computational Biology}, 17\penalty0 (2):\penalty0
+  e1008618, 2021.
+
+\bibitem[Colón-González et~al.(2021)Colón-González, Bastos, Hofmann,
+  Hopkin, Harpham, Crocker, Amato, Ferrario, Moschini, James, Malde, Ainscoe,
+  Nam, Tan, Khoa, Harrison, Tsarouchi, Lumbroso, Brady, and
+  Lowe]{colon-gonzalez_probabilistic_2021}
+Felipe~J. Colón-González, Leonardo~Soares Bastos, Barbara Hofmann, et~al.
+\newblock Probabilistic seasonal dengue forecasting in {Vietnam}: {A} modelling
+  study using superensembles.
+\newblock \emph{PLOS Medicine}, 18\penalty0 (3):\penalty0 e1003542, March 2021.
+\newblock \doi{10.1371/journal.pmed.1003542}.
+
+\bibitem[Cramer et~al.(2022{\natexlab{a}})Cramer, Ray, Lopez, Bracher, Brennen,
+  Castro~Rivadeneira, Gerding, Gneiting, House, Huang, Jayawardena, Kanji,
+  Khandelwal, Le, Mühlemann, Niemi, Shah, Stark, Wang, Wattanachit, Zorn, Gu,
+  Jain, Bannur, Deva, Kulkarni, Merugu, Raval, Shingi, Tiwari, White,
+  Abernethy, Woody, Dahan, Fox, Gaither, Lachmann, Meyers, Scott, Tec,
+  Srivastava, George, Cegan, Dettwiller, England, Farthing, Hunter, Lafferty,
+  Linkov, Mayo, Parno, Rowland, Trump, Zhang-James, Chen, Faraone, Hess,
+  Morley, Salekin, Wang, Corsetti, Baer, Eisenberg, Falb, Huang, Martin,
+  McCauley, Myers, Schwarz, Sheldon, Gibson, Yu, Gao, Ma, Wu, Yan, Jin, Wang,
+  Chen, Guo, Zhao, Gu, Chen, Wang, Xu, Zhang, Zou, Biegel, Lega, McConnell,
+  Nagraj, Guertin, Hulme-Lowe, Turner, Shi, Ban, Walraven, Hong, Kong, van~de
+  Walle, Turtle, Ben-Nun, Riley, Riley, Koyluoglu, DesRoches, Forli, Hamory,
+  Kyriakides, Leis, Milliken, Moloney, Morgan, Nirgudkar, Ozcan, Piwonka, Ravi,
+  Schrader, Shakhnovich, Siegel, Spatz, Stiefeling, Wilkinson, Wong, Cavany,
+  España, Moore, Oidtman, Perkins, Kraus, Kraus, Gao, Bian, Cao,
+  Lavista~Ferres, Li, Liu, Xie, Zhang, Zheng, Vespignani, Chinazzi, Davis, Mu,
+  Pastore~y Piontti, Xiong, Zheng, Baek, Farias, Georgescu, Levi, Sinha, Wilde,
+  Perakis, Bennouna, Nze-Ndong, Singhvi, Spantidakis, Thayaparan, Tsiourvas,
+  Sarker, Jadbabaie, Shah, Della~Penna, Celi, Sundar, Wolfinger, Osthus,
+  Castro, Fairchild, Michaud, Karlen, Kinsey, Mullany, Rainwater-Lovett, Shin,
+  Tallaksen, Wilson, Lee, Dent, Grantz, Hill, Kaminsky, Kaminsky, Keegan,
+  Lauer, Lemaitre, Lessler, Meredith, Perez-Saez, Shah, Smith, Truelove, Wills,
+  Marshall, Gardner, Nixon, Burant, Wang, Gao, Gu, Kim, Li, Wang, Wang, Yu,
+  Reiner, Barber, Gakidou, Hay, Lim, Murray, Pigott, Gurung, Baccam, Stage,
+  Suchoski, Prakash, Adhikari, Cui, Rodríguez, Tabassum, Xie, Keskinocak,
+  Asplund, Baxter, Oruc, Serban, Arik, Dusenberry, Epshteyn, Kanal, Le, Li,
+  Pfister, Sava, Sinha, Tsai, Yoder, Yoon, Zhang, Abbott, Bosse, Funk,
+  Hellewell, Meakin, Sherratt, Zhou, Kalantari, Yamana, Pei, Shaman, Li,
+  Bertsimas, Skali~Lami, Soni, Tazi~Bouardi, Ayer, Adee, Chhatwal, Dalgic,
+  Ladd, Linas, Mueller, Xiao, Wang, Wang, Xie, Zeng, Green, Bien, Brooks, Hu,
+  Jahja, McDonald, Narasimhan, Politsch, Rajanala, Rumack, Simon, Tibshirani,
+  Tibshirani, Ventura, Wasserman, O’Dea, Drake, Pagano, Tran, Ho, Huynh,
+  Walker, Slayton, Johansson, Biggerstaff, and Reich]{cramer_evaluation_2022}
+Estee~Y. Cramer, Evan~L. Ray, Velma~K. Lopez, et~al.
+\newblock Evaluation of individual and ensemble probabilistic forecasts of
+  {COVID}-19 mortality in the {United} {States}.
+\newblock \emph{Proceedings of the National Academy of Sciences}, 119\penalty0
+  (15):\penalty0 e2113561119, April 2022{\natexlab{a}}.
+\newblock \doi{10.1073/pnas.2113561119}.
+
+\bibitem[Sherratt et~al.(2023)Sherratt, Gruson, Grah, Johnson, Niehus, Prasse,
+  Sandmann, Deuschel, Wolffram, Abbott, et~al.]{sherratt2023predictive}
+Katharine Sherratt, Hugo Gruson, Rok Grah, et~al.
+\newblock Predictive performance of multi-model ensemble forecasts of
+  {COVID}-19 across {E}uropean nations.
+\newblock \emph{e{L}ife}, 12:\penalty0 e81916, 2023.
+
+\bibitem[Yardley and Petropoulos(2021)]{yardley2021utility_cost_forecasts}
+Elizabeth Yardley and Fotios Petropoulos.
+\newblock Beyond error measures to the utility and cost of the forecasts.
+\newblock \emph{Foresight: The International Journal of Applied Forecasting},
+  \penalty0 (63):\penalty0 36--45, 2021.
+
+\bibitem[Murphy(1993)]{murphy1993whatisagoodforecast}
+Allan~H Murphy.
+\newblock What is a good forecast? an essay on the nature of goodness in
+  weather forecasting.
+\newblock \emph{Weather and Forecasting}, 8\penalty0 (2):\penalty0 281--293,
+  1993.
+
+\bibitem[Pesaran and Skouras(2002)]{pesaran2002decision_based_eval}
+M~Hashem Pesaran and Spyros Skouras.
+\newblock Decision-based methods for forecast evaluation.
+\newblock In Michael~P. Clements and David~F. Hendry, editors, \emph{A
+  {C}ompanion to {E}conomic {F}orecasting}, chapter~11, pages 241--267.
+  Blackwell, Oxford, 2002.
+
+\bibitem[Ioannidis et~al.(2022)Ioannidis, Cripps, and
+  Tanner]{ioannidis2022forecastingCOVIDfailed}
+John~PA Ioannidis, Sally Cripps, and Martin~A Tanner.
+\newblock Forecasting for {COVID}-19 has failed.
+\newblock \emph{International Journal of Forecasting}, 38\penalty0
+  (2):\penalty0 423--438, 2022.
+
+\bibitem[Probert et~al.(2016)Probert, Shea, Fonnesbeck, Runge, Carpenter,
+  Dürr, Garner, Harvey, Stevenson, Webb, Werkman, Tildesley, and
+  Ferrari]{Probert2016decisionMakingFootMouth}
+William~J.M. Probert, Katriona Shea, Christopher~J. Fonnesbeck, et~al.
+\newblock Decision-making for foot-and-mouth disease control: {O}bjectives
+  matter.
+\newblock \emph{Epidemics}, 15:\penalty0 10--19, 2016.
+\newblock \doi{10.1016/j.epidem.2015.11.002}.
+
+\bibitem[Dawid(2007)]{dawid2007geometry}
+A~Philip Dawid.
+\newblock The geometry of proper scoring rules.
+\newblock \emph{Annals of the Institute of Statistical Mathematics},
+  59:\penalty0 77--93, 2007.
+
+\bibitem[Gneiting and Raftery(2007)]{gneiting2007strictly}
+Tilmann Gneiting and Adrian~E Raftery.
+\newblock Strictly proper scoring rules, prediction, and estimation.
+\newblock \emph{Journal of the American Statistical Association}, 102\penalty0
+  (477):\penalty0 359--378, 2007.
+
+\bibitem[Hadley and Whitin(1963)]{hadleywhitin1963}
+G.~Hadley and Thomson~M. Whitin.
+\newblock \emph{Analysis of {I}nventory {S}ystems.}
+\newblock Prentice-Hall international series in management. Prentice-Hall,
+  1963.
+
+\bibitem[Diecidue and Somasundaram(2017)]{DIECIDUE201788}
+Enrico Diecidue and Jeeva Somasundaram.
+\newblock Regret theory: A new foundation.
+\newblock \emph{Journal of Economic Theory}, 172:\penalty0 88--119, 2017.
+\newblock \doi{10.1016/j.jet.2017.08.006}.
+
+\bibitem[Hong et~al.(2016)Hong, Pinson, Fan, Zareipour, Troccoli, and
+  Hyndman]{hong2016probabilisticEnergyForecasting}
+Tao Hong, Pierre Pinson, Shu Fan, et~al.
+\newblock Probabilistic energy forecasting: {G}lobal energy forecasting
+  competition 2014 and beyond.
+\newblock \emph{International Journal of Forecasting}, 32\penalty0
+  (3):\penalty0 896--913, 2016.
+
+\bibitem[Gneiting and Ranjan(2011)]{gneiting2011weightedScoringRules}
+Tilmann Gneiting and Roopesh Ranjan.
+\newblock Comparing density forecasts using threshold- and quantile-weighted
+  scoring rules.
+\newblock \emph{Journal of Business \& Economic Statistics}, 29\penalty0
+  (3):\penalty0 411--422, 2011.
+\newblock \doi{10.1198/jbes.2010.08110}.
+
+\bibitem[Cramer et~al.(2022{\natexlab{b}})Cramer, Huang, Wang, Ray, Cornell,
+  Bracher, Brennen, Rivadeneira, Gerding, House, Jayawardena, Kanji,
+  Khandelwal, Le, Mody, Mody, Niemi, Stark, Shah, Wattanchit, Zorn, and
+  Reich]{cramer_united_2022}
+Estee~Y. Cramer, Yuxin Huang, Yijin Wang, et~al.
+\newblock The {United} {States} {COVID}-19 {Forecast} {Hub} dataset.
+\newblock \emph{Scientific Data}, 9\penalty0 (1):\penalty0 462, August
+  2022{\natexlab{b}}.
+\newblock \doi{10.1038/s41597-022-01517-w}.
+
+\bibitem[Ray et~al.(2023)Ray, Brooks, Bien, Biggerstaff, Bosse, Bracher,
+  Cramer, Funk, Gerding, Johansson, Rumack, Wang, Zorn, Tibshirani, and
+  Reich]{ray_comparing_2023}
+Evan~L. Ray, Logan~C. Brooks, Jacob Bien, et~al.
+\newblock Comparing trained and untrained probabilistic ensemble forecasts of
+  {COVID}-19 cases and deaths in the {United} {States}.
+\newblock \emph{International Journal of Forecasting}, 39\penalty0
+  (3):\penalty0 1366--1383, July 2023.
+\newblock \doi{10.1016/j.ijforecast.2022.06.005}.
+
+\bibitem[{United States Census Bureau}(2022)]{us_census_nst_est2022}
+{United States Census Bureau}.
+\newblock {Annual Population Estimates, Estimated Components of Resident
+  Population Change, and Rates of the Components of Resident Population Change
+  for the United States, States, District of Columbia, and Puerto Rico: April
+  1, 2020 to July 1, 2022}.
+\newblock US Census Bureau, 2022.
+\newblock URL
+  \url{https://www2.census.gov/programs-surveys/popest/datasets/2020-2022/state/totals/NST-EST2022-ALLDATA.csv}.
+\newblock Accessed: 2024-02-01.
+
+\bibitem[{R Core Team}(2023)]{Rcore-2023}
+{R Core Team}.
+\newblock \emph{R: A Language and Environment for Statistical Computing}.
+\newblock R Foundation for Statistical Computing, Vienna, Austria, 2023.
+\newblock URL \url{https://www.R-project.org/}.
+
+\bibitem[Landau(2021)]{landau_2021_targets}
+William~Michael Landau.
+\newblock The targets r package: a dynamic make-like function-oriented pipeline
+  toolkit for reproducibility and high-performance computing.
+\newblock \emph{Journal of Open Source Software}, 6\penalty0 (57):\penalty0
+  2959, 2021.
+\newblock \doi{10.21105/joss.02959}.
+
+\bibitem[Leitch and Tanner(1991)]{leitch1991economicForecastEval}
+Gordon Leitch and J~Ernest Tanner.
+\newblock Economic forecast evaluation: profits versus the conventional error
+  measures.
+\newblock \emph{The American Economic Review}, 81\penalty0 (3):\penalty0
+  580--590, 1991.
+
+\bibitem[Cenesizoglu and
+  Timmermann(2012)]{cenesizoglu2012returnPredictionEconValue}
+Tolga Cenesizoglu and Allan Timmermann.
+\newblock Do return prediction models add economic value?
+\newblock \emph{Journal of Banking \& Finance}, 36\penalty0 (11):\penalty0
+  2974--2987, 2012.
+
+\bibitem[Huang et~al.(2017)Huang, Araz, Morton, Johnson, Damien, Clements, and
+  Meyers]{huang_stockpiling_2017}
+Hsin-Chan Huang, Ozgur~M. Araz, David~P. Morton, et~al.
+\newblock Stockpiling {Ventilators} for {Influenza} {Pandemics}.
+\newblock \emph{Emerging Infectious Diseases}, 23\penalty0 (6), 2017.
+\newblock \doi{10.3201/eid2306.161417}.
+
+\bibitem[Araz et~al.(2012)Araz, Galvani, and Meyers]{araz_geographic_2012}
+Ozgur~M. Araz, Alison Galvani, and Lauren~A. Meyers.
+\newblock Geographic prioritization of distributing pandemic influenza
+  vaccines.
+\newblock \emph{Health Care Management Science}, 15\penalty0 (3):\penalty0
+  175--187, September 2012.
+\newblock \doi{10.1007/s10729-012-9199-6}.
+
+\bibitem[Persad et~al.(2023)Persad, Leland, Ottersen, Richardson, Saenz,
+  Schaefer, and Emanuel]{persad_fair_2023}
+Govind Persad, R.~J. Leland, Trygve Ottersen, et~al.
+\newblock Fair domestic allocation of monkeypox virus countermeasures.
+\newblock \emph{The Lancet Public Health}, 8\penalty0 (5):\penalty0 e378--e382,
+  May 2023.
+\newblock \doi{10.1016/S2468-2667(23)00061-0}.
+
+\bibitem[Du et~al.(2022)Du, J~Beesley, Lee, Zhou, Dempsey, and
+  Mukherjee]{du_optimal_2022}
+Jiacong Du, Lauren J~Beesley, Seunggeun Lee, et~al.
+\newblock Optimal diagnostic test allocation strategy during the {COVID}-19
+  pandemic and beyond.
+\newblock \emph{Statistics in Medicine}, 41\penalty0 (2):\penalty0 310--327,
+  2022.
+\newblock \doi{10.1002/sim.9238}.
+
+\bibitem[Pasco et~al.(2023)Pasco, Johnson, Fox, Pierce, Johnson-León,
+  Lachmann, Morton, and Meyers]{pasco_covid-19_2023}
+Remy Pasco, Kaitlyn Johnson, Spencer~J. Fox, et~al.
+\newblock {COVID}-19 {Test} {Allocation} {Strategy} to {Mitigate}
+  {SARS}-{CoV}-2 {Infections} across {School} {Districts}.
+\newblock \emph{Emerging Infectious Diseases}, 29\penalty0 (3), 2023.
+\newblock \doi{10.3201/eid2903.220761}.
+
+\bibitem[Ellsberg(1961)]{ellsberg1961risk}
+Daniel Ellsberg.
+\newblock Risk, ambiguity, and the savage axioms.
+\newblock \emph{The Quarterly Journal of Economics}, 75\penalty0 (4):\penalty0
+  643--669, 1961.
+
+\bibitem[Gneiting(2011{\natexlab{a}})]{gneiting2011making}
+Tilmann Gneiting.
+\newblock Making and evaluating point forecasts.
+\newblock \emph{Journal of the American Statistical Association}, 106\penalty0
+  (494):\penalty0 746--762, 2011{\natexlab{a}}.
+
+\bibitem[Fissler and Ziegel(2016)]{fisslerziegel2016consistency}
+Tobias Fissler and Johanna~F. Ziegel.
+\newblock {Higher order elicitability and Osband's principle}.
+\newblock \emph{The Annals of Statistics}, 44\penalty0 (4):\penalty0 1680 --
+  1707, 2016.
+
+\bibitem[Gneiting(2011{\natexlab{b}})]{gneiting2011quantiles}
+Tilmann Gneiting.
+\newblock Quantiles as optimal point forecasts.
+\newblock \emph{International Journal of forecasting}, 27\penalty0
+  (2):\penalty0 197--207, 2011{\natexlab{b}}.
+
+\bibitem[Jose and Winkler(2009)]{jose2009evaluating}
+Victor Richmond~R Jose and Robert~L Winkler.
+\newblock Evaluating quantile assessments.
+\newblock \emph{Operations research}, 57\penalty0 (5):\penalty0 1287--1297,
+  2009.
+
+\bibitem[Royset and Wets(2022)]{royset2022optimization}
+Johannes~O Royset and Roger J-B Wets.
+\newblock \emph{An optimization primer}.
+\newblock Springer, 2022.
+
+\bibitem[Gerding and Ray(2023)]{gerding-alloscore}
+Aaron Gerding and Evan Ray.
+\newblock \emph{alloscore: Tools for Implementing Allocation Scoring Rules},
+  2023.
+\newblock URL \url{https://github.com/aaronger/alloscore}.
+\newblock R package version 0.0.9001.
+
+\bibitem[Ray and Gerding(2023)]{ray-distfromq}
+Evan~L Ray and Aaron Gerding.
+\newblock \emph{distfromq: Reconstruct a Distribution from a Collection of
+  Quantiles}, 2023.
+\newblock URL \url{https://github.com/reichlab/distfromq}.
+\newblock R package version 1.0.2.
+
+\bibitem[Wadsworth et~al.(2023)Wadsworth, Niemi, and
+  Reich]{wadsworth2023mixture}
+Spencer Wadsworth, Jarad Niemi, and Nicholas~G Reich.
+\newblock Mixture distributions for probabilistic forecasts of disease
+  outbreaks.
+\newblock \emph{arXiv preprint arXiv:2310.11939}, 2023.
+\newblock \doi{10.48550/arXiv.2310.11939}.
+
+\end{thebibliography}
 

### 2024-02-28 alloscore_manuscript/alloscore-application-shorter.tex
Some edits to text in alloscore_application_shorter file

@@
 
-\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$
-and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$. The means of these distributions are
+\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations such that $F_1$ and $F_2$ have marginal exponential distributions with scale parameters $\sigma_1 = 1$ and $\sigma_2 = 4$, respectively.
+The means of these distributions are
 given by the scale parameters $\sigma_i$. When the marginal forecasts are exponential distributions, it can be shown
@@
 In our setting an action is a vector $x = (x_1, \ldots, x_N)$ specifying the amount of resources allocated to each of
-$N$ locations. We require that $0 \leq x$, i.e., that each $x_i$ is non-negative, and that the total allocation across
+$N$ locations. We require that each $x_i$ is non-negative, and that the total allocation across
 all locations equals the amount of available resources, $K$: $\sum_{i=1}^N x_i = K$. The set $\mathcal{X}$ consists of
@@
 time of decision making) of the random vector $Y = (Y_1, \ldots, Y_N)$. A forecast of the random need $Y$ for each
-location is written as $F = (F_1, \ldots, F_N)$. We assume need is non-negative and the forecasts reflect that, i.e. the
+location is written as $F = (F_1, \ldots, F_N)$. We assume need is non-negative and that the forecasts reflect that, i.e. the
 support of each $F_i$ is a subset of $\mathbb{R}^+$. Finally, we assume that each unmet unit of need in any location
@@
 
-\subsection{Connections to Other Scores}
+\subsection{Connections to Weighted Interval Score}
 \label{sec:methods.related}
 
-The weighted interval score (WIS) was proposed in 2020 as a way to score forecasts that were being made in the early
+The weighted interval score (WIS) was proposed in 2021 as a way to score forecasts that were being made in the early
 stages of the COVID-19 pandemic \citep{bracher2021evaluating}; equivalent scores had also been used in previous forecast
 evaluation efforts \cite[e.g.,][]{hong2016probabilisticEnergyForecasting}. The WIS is a proper scoring rule for
-forecasts that use a set of quantiles to represent a probabilistic forecast distribution. While pointing a reader
-interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS is a weighted sum of
-interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...). Larger
-interval scores indicate less skillful forecasts. An interval score consists of (a) the width of the interval, with
+forecasts that use a set of quantiles to represent a probabilistic forecast distribution and scores forecasts as a weighted sum of interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...).
+Larger interval scores indicate less skillful forecasts.
+An interval score consists of (a) the width of the interval, with
 larger intervals receiving higher scores (higher scores indicate less accuracy), and (b) a penalty if the interval does
@@
 probabilistic forecasts. \emph{It is important to note that this weighting was proposed because the resulting score
-approximates the CRPS, and not because it aligned with any particular public health decision-making rationale.}
+approximates the CRPS, and not because it aligned with any particular public health decision-making rationale.} For more mathematical detail, we point readers to \cite{bracher2021evaluating}.
 
@@
 The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the
-constraint ($K$), possibly assigning higher weights to more likely values of $K$ (Section
+constraint ($K$), allowing one to place higher weights to more likely values of $K$ (Section
 \ref{sec:methods.detailed.integrated_allocation}). IAS was computed for two weight distributions on $K$, one uniform
-across the entire range and the other a symmetric distribution around 15,000 (the orange and blue shaded areas in Figure
+across the entire range and the other a truncated normal distribution centered at 15,000 (the orange and blue shaded areas in Figure
 \ref{fig:multi-k}A, respectively). Both versions of the IAS were correlated with the original AS for $K=15,000$, with

### 2024-02-28 alloscore_manuscript/alloscore-application-shorter.tex
knit Ben's changes

--- 2024-02-27 alloscore_manuscript/alloscore-application-shorter.tex
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Tue Feb 27 15:06:55 2024
+% Wed Feb 28 17:15:20 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Tue Feb 27 15:06:59 2024
+% Wed Feb 28 17:15:24 2024
 \begin{table}[ht]

### 2024-02-29 alloscore_manuscript/alloscore-application-shorter.Rnw
BWR edits made to .Rnw file

@@
 
-\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$
-and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$. The means of these distributions are
+\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations such that $F_1$ and $F_2$ have exponential marginal distribution with scale parameters $\sigma_1 = 1$ and $\sigma_2 = 4$.
+The means of these distributions are
 given by the scale parameters $\sigma_i$. When the marginal forecasts are exponential distributions, it can be shown
@@
 
-It is helpful to clearly distinguish between the time $t_d$ when a \emph{d}ecision is made about a public health
-resource allocation and the time $t_r$ when \emph{r}esource needs that might be addressed by that allocation occur. Our
+It is helpful to clearly distinguish between the time $t_d$ when a decision is made about a public health
+resource allocation and the time $t_r$ when resource needs that might be addressed by that allocation occur. Our
 setup assumes that $t_d < t_r$ and that the resource in question can only meet the need realized at $t_r$, not prevent
@@
 \paragraph{Step 3: Define the scoring rule.} We can now use the Bayes act to define a proper scoring rule for the
-probabilistic forecast $F$. Consider first the ``raw'' score defined as
+probabilistic forecast $F$. Consider first the raw score defined as
 \begin{align}
@@
 \end{align}
-The oracle adjustment aligns with a common theme in economic decision theory that \emph{opportunity loss} (often known
-as \emph{regret} or (negative) \emph{relative utility}) is often a more important quantity than absolute loss (see e.g.,
+The oracle adjustment aligns with a common theme in economic decision theory that opportunity loss (often known
+as regret or (negative) relative utility) is often a more important quantity than absolute loss (see e.g.,
 \cite{DIECIDUE201788}).
@@
 evaluation efforts \cite[e.g.,][]{hong2016probabilisticEnergyForecasting}. The WIS is a proper scoring rule for
-forecasts that use a set of quantiles to represent a probabilistic forecast distribution. While pointing a reader
-interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS is a weighted sum of
+forecasts that use a set of quantiles to represent a probabilistic forecast distribution. Scores are calculated as a weighted sum of
 interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...). Larger
@@
 levels, in which case WIS approximates the continuous ranked probability score (CRPS), a commonly used score for
-probabilistic forecasts. \emph{It is important to note that this weighting was proposed because the resulting score
-approximates the CRPS, and not because it aligned with any particular public health decision-making rationale.}
+probabilistic forecasts. It is important to note that this weighting was proposed because the resulting score
+approximates the CRPS, and not because it aligned with any particular public health decision-making rationale.
+For more mathematical detail on the WIS, we point readers to \cite{bracher2021evaluating}.
 
@@
 The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the
-constraint ($K$), possibly assigning higher weights to more likely values of $K$ (Section
+constraint ($K$), allowing one to assign higher weights to more likely values of $K$ (Section
 \ref{sec:methods.detailed.integrated_allocation}). IAS was computed for two weight distributions on $K$, one uniform
-across the entire range and the other a symmetric distribution around 15,000 (the orange and blue shaded areas in Figure
+across the entire range and the other a truncated normal distribution centered at 15,000 (the orange and blue shaded areas in Figure
 \ref{fig:multi-k}A, respectively). Both versions of the IAS were correlated with the original AS for $K=15,000$, with
@@
 $K$ by shifting funding from other disease mitigation measures. Finally, we emphasize that we opted at this stage of
-development to explicitly rule out situations where a ``successful''
+development to explicitly rule out situations where a successful
 epidemiological forecast could lead to policy decisions that change the distribution of the predicted outcome $Y$. Our
@@
 An opportunity for further investigation is to more carefully evaluate whether forecasts add value to 
-``standard operating procedures'' already in place for making public health decisions. In the context of allocations,
+standard  procedures already in place for making public health decisions. In the context of allocations,
 such a procedure might extrapolate need from current observed need and population levels (similarly to the two benchmark
@@
 scoring procedures as presented do not directly address this question, they could be modified (say, by comparison to a
-`baseline' model or expert-elicited allocations in the absence of forecast data) to quantify the benefit of using a
+baseline model or expert-elicited allocations in the absence of forecast data) to quantify the benefit of using a
 forecast to inform a specific decision.

### 2024-02-29 alloscore_manuscript/alloscore-application-shorter.tex
merge rest of Ben's changes

@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Wed Feb 28 17:15:20 2024
+% Thu Feb 29 18:53:54 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Wed Feb 28 17:15:24 2024
+% Thu Feb 29 18:53:58 2024
 \begin{table}[ht]

### 2024-03-01 alloscore_manuscript/alloscore-application-shorter.Rnw
small change

@@
 appendix \ref{sec:a:distfromq}. Briefly, this involves interpolating the provided quantiles between the lowest (.01) and highest 
-(.99) probability levels using monotonic cubic splinea, and then extending outside of these level with normal distribution tails
-parametrized to match the two lowest and two highest quantiles. In appendix \ref{sec:a:propriety_of_parametric_approximation}, 
-we show that if this evaluation procedure
-had been specified prospectively, the resulting score would be proper, \textemdash but that a post hoc application of this
+(.99) probability levels using monotonic cubic splines, and then extending outside these levels with normal distribution tails
+parametrized to match the two lowest and two highest quantiles. We show in appendix \ref{sec:a:propriety_of_parametric_approximation}
+that if this evaluation procedure
+had been specified prospectively, the resulting score would be proper, but that a post hoc application of this
 procedure is improper. We use the procedure here to illustrate the properties of the AS, and note that a collaborative

### 2024-07-25 arxivsub3/submission.tex
cleaning up

--- 2024-02-27 arxivsub2/submission.tex
@@
 \usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
+\usepackage[T1]{fontenc}
 \usepackage{appendix}
@@
 \usepackage{algorithm}
-\usepackage[noend]{algpseudocode}
 \usepackage{amsthm}
@@
 \usepackage{setspace}
-\onehalfspacing
+\singlespacing
 
@@
 \begin{abstract}
-
 Recent years have seen increasing efforts to forecast infectious disease burdens, with a primary goal being to help
@@
 performance is a promising direction for epidemic forecast evaluation.
-
 \end{abstract}
+
+\noindent \textbf{Keywords:} public health, forecast evaluation, proper scoring rules, resource allocation, epidemiology
 
@@
 
-\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$
-and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$. The means of these distributions are
+\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations such that $F_1$ and $F_2$ have exponential marginal distribution with scale parameters $\sigma_1 = 1$ and $\sigma_2 = 4$.
+The means of these distributions are
 given by the scale parameters $\sigma_i$. When the marginal forecasts are exponential distributions, it can be shown
 that the optimal allocation divides the available resources among the locations proportionally to the scale parameters
-$\sigma_i$ (see appendices \ref{sec:bayes-quantiles}). If $K = 5$ units of the resource are
+$\sigma_i$ (see appendix \ref{sec:a:bayes-quantiles}). If $K = 5$ units of the resource are
 available, the optimal allocation according to $F$ would be 1 and 4 resource units in locations 1 and 2, respectively. 
@@
 As described above, predictions were submitted to the Forecast Hub in the form of a set of 23 quantiles of the
-predictive distribution. The WIS can be directly calculated from these quantiles. However, calculation of the AS
-requires that the full cumulative distribution functions of the forecast distributions for each location are available.
-For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution
-functions based on the provided quantiles. On the interior of the provided quantiles we used a monotonic cubic spline to
-interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as
-to match the two lowest and two highest quantiles (see appendix \ref{sec:distfromq} for further details).
-In appendix \ref{sec:propriety_of_parametric_approximation}, we show that if this evaluation procedure
-had been specified prospectively, the resulting score would be proper, \textemdash but that a post hoc application of this
+predictive distribution. The WIS can be directly calculated from these quantiles. However, our numerical method for calculating 
+the AS (outlined in appendix \ref{sec:a:numeric})
+requires a full cumulative distribution function (CDF) for the forecast in each location.
+For the purpose of this analysis, we imputed CDFs based on the provided quantiles following a procedure detailed in 
+appendix \ref{sec:a:distfromq}. Briefly, this involves interpolating the provided quantiles between the lowest (.01) and highest 
+(.99) probability levels using monotonic cubic splines, and then extending outside these levels with normal distribution tails
+parametrized to match the two lowest and two highest quantiles. We show in appendix \ref{sec:a:propriety_of_parametric_approximation}
+that if this evaluation procedure
+had been specified prospectively, the resulting score would be proper, but that a post hoc application of this
 procedure is improper. We use the procedure here to illustrate the properties of the AS, and note that a collaborative
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Feb 23 19:04:47 2024
+% Fri Mar  1 16:18:50 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Feb 23 19:04:51 2024
+% Fri Mar  1 16:18:55 2024
 \begin{table}[ht]
@@
 \end{thebibliography}
-
 \end{document}

### 2024-07-25 alloscore_manuscript/alloscore-application-shorter.tex

@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Thu Feb 29 18:53:54 2024
+% Fri Mar  1 16:18:50 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Thu Feb 29 18:53:58 2024
+% Fri Mar  1 16:18:55 2024
 \begin{table}[ht]
@@
 We define
-$$G^{\star}(F) := \argmin_{G \in \mathcal{P}_{sec:propriety_of_parametric_approximation\mathrm{dfq}}} E_{F}[S_A(G, Y)].$$
+$$G^{\star}(F) := \argmin_{G \in \mathcal{P}_{\mathrm{dfq}}} E_{F}[S_A(G, Y)].$$
 Since $S_R$ is defined as the Bayes scoring rule for the forecaster's loss \eqref{eqn:forecaster_theta_loss}, $G^{\star}(F)$ coincides

### 2024-07-26 notes/alloscore-application.tex
more cleanup

--- 2024-07-26 notes/alloscore-application.Rnw
@@
 For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data \cite{wang-covidhubutils}.
-The hospitalization data were downloaded for analysis on \Sexpr{format(as.Date(metadata$time[which(metadata$name == "truth_data")]), "%B %d, %Y")}.
+The hospitalization data were downloaded for analysis on December 03, 2023.
 
@@
 
-<<ias-comparison>>=
-tar_load(alloscore_2021.12.20)
-slim_alloscore_2021.12.20 <- alloscore::slim(alloscore_2021.12.20, id_cols = c("model", "reference_date"))
-
-## for visual context
-## curve(dnorm(x, mean=15000, sd=3000), from=0, to=30000)
-ks <- seq(5000, 25000, by=200)
-kprobs <- data.frame(K = ks,
-                     weight = dnorm(ks, mean=15000, sd=3000)/sum(dnorm(ks, mean=15000, sd=3000)))
-
-ias <- slim_alloscore_2021.12.20 |>
-  group_by(model, K) |>
-  summarize(`Allocation score` = sum(components)) |>
-  left_join(kprobs) |>
-  group_by(model) |>
-  summarize(IAS_wt = sum(`Allocation score`*weight, na.rm=TRUE),
-            IAS = mean(`Allocation score`)) |>
-  arrange(IAS)
-
-as <- slim_alloscore_2021.12.20 |>
-  filter(K == 15000) |>
-  group_by(model) |>
-  summarize(AS = sum(components)) |>
-  arrange(AS)
-@
-
-<<multi-k-scores, results='asis'>>=
-wis_vals <- all_scores |>
-  filter(target_end_date == "2022-01-03", !is.na(mwis)) |>
-  select(model,
-# can uncomment the two lines below to check they match
-#         AS = alloscore,
-#         `AS rank` = allo_rank,
-         MWIS = mwis)
-
-left_join(ias, as) |>
-  left_join(wis_vals, by="model") |>
-  rename(`IAS centered at 15k` = IAS_wt,
-         `IAS uniform` = IAS) |>
-  select(model, AS, MWIS, `IAS centered at 15k`, `IAS uniform`) |>
-  arrange(AS) |>
-  xtable::xtable(caption = "Comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS).
+% latex table generated in R 4.3.1 by xtable 1.8-4 package
+% Thu Dec 21 23:53:38 2023
+\begin{table}[ht]
+\centering
+\begin{tabular}{lrrrr}
+  \hline
+model & AS & MWIS & IAS centered at 15k & IAS uniform \\ 
+  \hline
+CU-select & 669 & 133 & 774 & 326 \\ 
+  COVIDhub-ensemble & 873 & 159 & 1067 & 438 \\ 
+  USC-SI\_kJalpha & 995 & 91 & 1216 & 1097 \\ 
+  JHUAPL-Gecko & 1034 & 164 & 1141 & 418 \\ 
+  MUNI-ARIMA & 1084 & 169 & 1248 & 440 \\ 
+  COVIDhub-trained\_ensemble & 1089 & 169 & 1271 & 823 \\ 
+  COVIDhub-baseline & 1175 & 170 & 1317 & 535 \\ 
+  JHUAPL-Bucky & 1358 & 102 & 1566 & 1214 \\ 
+  JHUAPL-SLPHospEns & 1540 & 129 & 1604 & 1102 \\ 
+  UVA-Ensemble & 2469 & 213 & 2635 & 2494 \\ 
+   \hline
+\end{tabular}
+\caption{Comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS).
 All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03.
 Results are sorted by AS.
-For all metrics, lower scores indicate better accuracy.",
-                 digits = 0,
-                 label = "tab:multi-k-scores") |>
-  print(include.rownames = FALSE, )
-@
-
-<<metrics-rankings, results = "asis", eval=FALSE>>=
-## NOTE: this content now shown in multi-k-scores chunk
-## for highlighting cells, if we want that https://stackoverflow.com/questions/37448348/color-code-cells-of-xtable
-all_scores |>
-  filter(target_end_date == "2022-01-03", !is.na(mwis)) |>
-  select(model,
-         AS = alloscore,
-         `AS rank` = allo_rank,
-         MWIS = mwis,
-         `MWIS rank` = mwis_rank) |>
-  arrange(`AS rank`) |>
-  xtable::xtable(label = "tab:one-week-results",
-                 caption = "Allocation scores (AS) and mean weighted interval scores (MWIS) for each model' as forecasted's forecasts for ") |>
-  print(include.rownames = FALSE)
-@
+For all metrics, lower scores indicate better accuracy.}
+\label{tab:multi-k-scores}
+\end{table}
 
@@
 
-<<detailedASinfo>>=
-source("R/run-alloscore.R")
-as_Buck <- run_alloscore(
-  forecast_data = forecast_data,
-  truth_data = truth_data,
-  one_reference_date = "2021-12-20",
-  one_model = "JHUAPL-Bucky",
-  one_K = 15000,
-  slim = FALSE)
-
-as_CU <- run_alloscore(
-  forecast_data = forecast_data,
-  truth_data = truth_data,
-  one_reference_date = "2021-12-20",
-  one_model = "CU-select",
-  one_K = 15000,
-  slim = FALSE)
-
-Buck_quantile <- round(1 - as_Buck$lam,2)*100
-CU_quantile <- round(1 - as_CU$lam,2)*100
-
-FL_allocs <- Kat15k_alloscores |>
-   dplyr::filter(
-     model %in% c("JHUAPL-Bucky", "CU-select"),
-     reference_date == "2021-12-20",
-     abbreviation == "FL") |>
-  dplyr::select(model, x) |>
-  tibble::deframe()
-FL_diff <- round(FL_allocs["CU-select"] - FL_allocs["JHUAPL-Bucky"])
-@
-
-For JHUAPL-Bucky, the allocations were assigned at the \Sexpr{Buck_quantile}th percentile of the predictive distribution and for CU-select they were at the \Sexpr{CU_quantile}th percentile.
+For JHUAPL-Bucky, the allocations were assigned at the 28th percentile of the predictive distribution and for CU-select they were at the 89th percentile.
 This reflects that in general the JHUAPL-Bucky forecasts assigned more probability to higher values and thus a low quantile value across all locations reached the allocation limit of 15,000.
@@
 However for each of these three states, the resource allocation was lower for JHUAPL-Bucky than for CU-select and therefore resulted in more hypothetical unmet need.
-In Florida, which was the state that saw the largest number of reported hospitalizations on this target date, the forecast from JHUAPL-Bucky missed by a wide margin (as did the forecast from CU-select) and the allocation was \Sexpr{FL_diff} units lower than the allocation derived from the CU-select model.
+In Florida, which was the state that saw the largest number of reported hospitalizations on this target date, the forecast from JHUAPL-Bucky missed by a wide margin (as did the forecast from CU-select) and the allocation was 716 units lower than the allocation derived from the CU-select model.
 
-<<thermometer-plot, fig.height=3.5, fig.width=7.5, fig.cap="Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the nine highest hospitalization counts.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\includegraphics[width=\maxwidth]{figure/thermometer-plot-1} \caption[Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the nine highest hospitalization counts]{Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the nine highest hospitalization counts.
 For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts.
-The side-by-side shaded regions show the median (solid horizontal line) and 50%, 80% and 95% prediction intervals for the two selected models.
+The side-by-side shaded regions show the median (solid horizontal line) and 50\%, 80\% and 95\% prediction intervals for the two selected models.
 The forecasts were made for new hospitalizations on January 3, 2022 (vertical dashed line, with number of hospitalizations indicated by red horizontal line segment at the intersection of the dashed line and the grey line of data).
 The vertical bars with red and yellow shading show the allocations: the red bar goes from zero to the level of the observed number of hospitalizations while the yellow bar shows how many resources were suggested by the model to be allocated to that location.
-The allocations (yellow bar) may be more or less than the observed number, therefore exposed red bar indicates unmet need, while the yellow bar extending above the observed number of hospitalizations means that excess resources were suggested for that given location.">>=
-source("R/plot_functions.R")
-plot_hosp(models = c("JHUAPL-Bucky", "CU-select"),
-          start_date = "2021-12-01",
-          stop_date = "2022-01-25",
-          f_date = "2021-12-20",
-          te_date = "2022-01-03",
-          ## highest locations in terms of hospitalizations
-          locations = c("FL", "NY", "TX", "CA"),
-          f_width1 = 4, f_alpha = .4,
-          allocations = TRUE,
-          forecasts_hosp = forecast_data %>%  filter(reference_date == "2021-12-20"),
-          alloscores = slim_alloscore_2021.12.20,
-          truth = truth_data,
-          f_colors = c("#4A708B", "#CD2626"),
-          one_K = 15000) + labs(x = NULL)
-@
+The allocations (yellow bar) may be more or less than the observed number, therefore exposed red bar indicates unmet need, while the yellow bar extending above the observed number of hospitalizations means that excess resources were suggested for that given location.}\label{fig:thermometer-plot}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<calc-tot-y>>=
-tot_y <- slim_alloscore_2021.12.20 |>
-  filter(model == "COVIDhub-ensemble", K==200) |>
-  pull(y) |>
-  sum()
-@
-
 AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 (Figure \ref{fig:multi-k}A).
-These calculations highlight that AS was highest at the $K$ nearest to the observed value of \Sexpr{format(tot_y, big.mark = ",")}.
+These calculations highlight that AS was highest at the $K$ nearest to the observed value of 19,581.
 Rankings of AS from all models were fairly stable across a range of values for K, with some crossings, especially at $K$ values further away from the observed value.

### 2024-07-26 notes/methods_new.rmd
more cleanup

--- 2023-04-28 alloscore_manuscript/methods_new.rmd
@@
 
-Unit-level decision problem: allocate to potential case in location $a$ or $b$. With a decision variable $d$ we write this as $d=a$ or $d=b$.
+### The unit-level decision problem:
 
-- elicit forecast $P(Y_a=1) = p$ where $Y_a=1$ if a case in addition to what is currently provided for at horizon $h$ occurs during the day at horizon $h$ in $a$, and 0 otherwise.
-$p$ can in particular come from a distributional incidence forecast $F_a$ via $p = 1-F_a(x)$ where $x$ is the number of cases currently provided for in $h$ days.
+Assume that there are currently resources allocated to care for $x_a$ and $x_b$ many cases in locations $a$ and $b$.
+Another resource unit becomes availible and we must decide whether to allocate it to a new potential case in location $a$ or $b$.
+In terms of a decision variable $d$ the decision is whether to set $d=a$ or $d=b$.
+
+To support our decision making process we elicit forecasts $p_a$ of $P(Y_a=1)$ and $p_b$ of $P(Y_b=1)$ where $Y_a, Y_b = 1$ if a case in addition to what is currently provided for at horizon $h$ occurs during the day at horizon $h$ in $a,b$, and 0 otherwise.
+
+$p_{a,b}$ can in particular come from a distributional incidence forecast $F_a$ via $p = 1-F_a(x)$ where $x$ is the number of cases currently provided for in $h$ days.
 
@@
 This is the function of the decision $d$ and outcome $Y_a$
+
 $$

### 2024-07-26 notes/supplement.tex
more cleanup

--- 2023-12-14 alloscore_manuscript/supplement.rnw
@@
 &= L(1-\alpha)(x - Y)_{+} + L\alpha(Y-x)_{+} \\
-&= L(\alpha - \mathbf{1}\{Y < x\})(Y-x),
+&= L(\alpha - \mathbf{1}\{Y < x\})(Y-x) \\
+&= L \max \left\{(\alpha-1)(Y-x), \alpha(Y-x)\right\}
 \end{align}

### 2024-07-26 alloscore_manuscript/manuscript.tex
change manuscript name and add some morae instructions to the readme

--- 2024-07-25 alloscore_manuscript/alloscore-application-shorter.tex
@@
 \newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlsng}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
 \newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
 \newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hldef}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
 \newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
@@
 
-The hospitalization data used for scoring forecasts were downloaded on February 07, 2024.
+The hospitalization data used for scoring forecasts were downloaded on July 26, 2024.
 
@@
 \url{https://github.com/aaronger/utility-eval-papers}. The analyses were generated using a reproducible workflow using
-R version 4.3.1 (2023-06-16) and the {\tt targets} package \citep{Rcore-2023, landau_2021_targets}.
+R version 4.4.1 (2024-06-14) and the {\tt targets} package \citep{Rcore-2023, landau_2021_targets}.
 
@@
 
-% latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Mar  1 16:18:50 2024
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Fri Jul 26 10:21:35 2024
 \begin{table}[ht]
@@
 
-% latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Fri Mar  1 16:18:55 2024
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Fri Jul 26 10:21:39 2024
 \begin{table}[ht]

### 2024-08-01 alloscore_manuscript/manuscript.tex
fix legend.pos->legend.position in metrics-over-time (fig4) components
sensitivity analysis  and discussion

@@
 
+\subsection{Sensitivity of allocation score to values of K}
+\label{sec:sensitivity}
+
+To assess the sensitivity of AS to choice of $K$, AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 as well as for the \texttt{per-capita} allocation (Figure \ref{fig:multi-k}A).
+These calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of 19,581 that day.
+Model ranks by AS were fairly stable across all $K$, and there appears to be an substantial interval around the observed value over which ranks are constant.
+
 \subsubsection{Integrated allocation score across values of K}
 \label{sec:ias_examp}
-
-AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of
-hospitalizations on January 3, 2021 as well as for the \texttt{per-capita} allocation (Figure \ref{fig:multi-k}A). These
-calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of
-19,581 that day. Model ranks by AS were fairly stable across all $K$, and there appears
-to be an substantial interval around the observed value over which ranks are constant.
 
@@
 \ref{sec:a:distfromq} and \ref{sec:a:propriety_of_parametric_approximation}.)
+
+Additionally, during an epidemic, public health decision makers may be faced with an ongoing series of decision over time in response to changes in conditions such as resource availability or the current allocations of resources.
+The AS presented here looks at a single snapshot in time  and ignores many of the complexities related to adjusting existing allocations.
+The example presented in section \ref{sec:application} is based on a prospective estimate of bed availability, but one could use observed total resource availability to derive an
 

### 2024-08-01 alloscore_manuscript/manuscript.Rnw

@@
   ylab("Allocation score") +
-  theme(legend.pos="none") +
+  theme(legend.position="none") +
   xlab(NULL)
@@
          linewidth = guide_legend(title=NULL, ncol = 4)) +
-  theme(legend.pos="bottom")
+  theme(legend.position="bottom")
 

### 2024-08-01 alloscore_manuscript/manuscript.tex
addressing Referee 1 concerns 1, 3, 4, 5, 6

@@
 distribution $p$:
-$$S_{IAS}(F, y) = \int S_A(F,y; K) p(K) \, dK$$
+\begin{align} 
+S_{IAS}(F, y) &= \int S_A(F,y; K) p(K) \, dK \label{eq:IASintegral}\\
+	& = \sum_{K=1}^{\infty} S_A(F,y; K) p(K) \label{eq:IASsum}.
+\end{align}
 We note that the device of considering a range of hypothetical decision makers or decision making problems with
@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Fri Jul 26 10:21:35 2024
+% Thu Aug  1 13:13:10 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Fri Jul 26 10:21:39 2024
+% Thu Aug  1 13:13:14 2024
 \begin{table}[ht]
@@
 
-\subsection{Sensitivity of allocation score to values of K}
+\subsubsection{Sensitivity of allocation score to values of K}
 \label{sec:sensitivity}
@@
 constraint ($K$), allowing one to assign higher weights to more likely values of $K$ (Section
-\ref{sec:methods.detailed.integrated_allocation}). IAS was computed for two weight distributions on $K$, one uniform
+\ref{sec:methods.detailed.integrated_allocation}).
+Calculating exact IAS as defined in equation \eqref{eq:IASsum} is computationally intensive so an approximate IAS, denoted $\hat{S}(F,y)$, was used such that
+\begin{align*}
+	\hat{S}(F,y) = \sum_{k=1}^{600} S_A(F,y; 100k) p(100k),
+\end{align*}
+where summation terms were calculated at every 100 beds with an upper bound of 60,000, or four times the estimated available beds.
+
+Approximate IAS was computed for two weight distributions on $K$, one uniform
 across the entire range and the other a truncated normal distribution centered at 15,000 (the orange and blue shaded areas in Figure
@@
 including evaluations of forecasts with
-scoring rules that directly consider the success of forecasts in supporting specific public health decisions. As evidence for this
-possibility, we have demonstrated how tying forecast evaluation to a specific decision problem (e.g., via the allocation
+scoring rules that directly consider the success of forecasts in supporting specific public health decisions.
+The work presented here has applications in many other fields such as economics, where the monetary consequences of using a forecast
+can be clearly juxtaposed with standard accuracy measures of the forecast \cite{bannigidadmath2016stock, zhang2019forecasting}, resource allocation in urban planning \citep{burnett2014urban, liang2013resource}, sustainability initiatives such as water and energy allocation \cite{syme1999defining, gebre2021multi, colett2016using}.
+We have demonstrated how tying forecast evaluation to a specific decision problem (e.g., via the allocation
 score) can yield model rankings that differ substantively from those based a current standard measure of forecast
 accuracy, the WIS, which does not refer to any decision problem that has been clearly connected to outbreak response.
-This result aligns with findings in other fields, especially those where the monetary consequences of using a forecast
-can be clearly juxtaposed with standard accuracy measures of the forecast \citep{leitch1991economicForecastEval,
+This result aligns with findings in other fields \citep{leitch1991economicForecastEval,
 murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}. Additionally, we show that an existing
@@
 The AS presented here looks at a single snapshot in time  and ignores many of the complexities related to adjusting existing allocations.
-The example presented in section \ref{sec:application} is based on a prospective estimate of bed availability, but one could use observed total resource availability to derive an
+The example presented in section \ref{sec:application} is based on a prospective estimate of bed availability, but one could also use observed total resource availability to derive an optimal set of prospective AS parameters.
+In the case of allocation of preventative resources, such as vaccines, one may which to account for the negative feedback loop between resource allocation and disease cases.
+Alternatively, the impact of unmet need often differs across different locations and demographics, which could be accounted for by incorporating a function for the cost of unmet need, which differs by location or other covariates.
+Future work is needed to explore the use of AS while accounting for the dynamic nature of disease spread and resource availability.
 
@@
 disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those
-for which they are able to offer useful guidance. New collaborative work between public health officials and modeling teams is
+for which they are able to offer useful guidance. The work presented here New collaborative work between decision-makers and modeling teams is
 needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or

### 2024-08-05 alloscore_manuscript/manuscript.Rnw
add Ben's edits to Rnw

@@
 distribution $p$:
-$$S_{IAS}(F, y) = \int S_A(F,y; K) p(K) \, dK$$
+\begin{align} 
+  S_{IAS}(F, y) &= \int S_A(F,y; K) p(K) \, dK \label{eq:IASintegral}\\
+    & = \sum_{K=1}^{\infty} S_A(F,y; K) p(K) \quad (\text{assumming non-divisible resource units})\label{eq:IASsum}.
+\end{align}
 We note that the device of considering a range of hypothetical decision makers or decision making problems with
@@
 
-\subsubsection{Integrated allocation score across values of K}
-\label{sec:ias_examp}
+\subsubsection{Sensitivity of allocation score to values of K}
+\label{sec:sensitivity}
 
@@
 
+\subsubsection{Integrated allocation score across values of K}
+\label{sec:ias_examp}
+
 The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the
 constraint ($K$), allowing one to assign higher weights to more likely values of $K$ (Section
-\ref{sec:methods.detailed.integrated_allocation}). IAS was computed for two weight distributions on $K$, one uniform
+\ref{sec:methods.detailed.integrated_allocation}). 
+Calculating exact IAS as defined in equation \eqref{eq:IASsum} is computationally intensive so an approximate IAS, denoted $\hat{S}(F,y)$, was used such that
+\begin{align*}
+	\hat{S}(F,y) = \sum_{k=1}^{600} S_A(F,y; 200k) p(200k),
+\end{align*}
+where summation terms were calculated at every 200 beds with an upper bound of 60,000, or four times the estimated available beds.
+An approximate IAS was computed for two weight distributions on $K$, one uniform
 across the entire range and the other a truncated normal distribution centered at 15,000 (the orange and blue shaded areas in Figure
@@
 
+Additionally, during an epidemic, public health decision makers may be faced with an ongoing series of decisions over time in response to changes in conditions such as resource availability or the current effective allocations of resources.
+The AS presented here looks at a single snapshot in time  and ignores many of the possible complexities related to adjusting existing allocations.
+The example presented in section
+\ref{sec:application} is based on a prospective estimate of bed availability, but one could also use observed total resource availability to target an optimal set of prospective AS parameters. 
+Alternatively, the impact of unmet need often differs across different locations and demographics, which could be accounted for by incorporating a function of location and demographics which defines the cost of unmet need.
+Future work is needed to explore the use of AS while accounting for the dynamic nature of disease spread and resource availability.
+
 There are several important limitations to the current work. The allocation score we developed here does not directly
@@
 disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those
-for which they are able to offer useful guidance. New collaborative work between public health officials and modeling teams is
+for which they are able to offer useful guidance. New collaborative work between decision-makers and modeling teams is
 needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or

### 2024-08-05 alloscore_manuscript/manuscript.Rnw
just added a missing 'or'

@@
 The work presented here has applications in many other fields such as economics, where the monetary consequences of using a forecast
-can be clearly juxtaposed with standard accuracy measures of the forecast \cite{bannigidadmath2016stock, zhang2019forecasting}, resource allocation in urban planning \citep{burnett2014urban, liang2013resource}, sustainability initiatives such as water and energy allocation \cite{syme1999defining, gebre2021multi, colett2016using}.
+can be clearly juxtaposed with standard accuracy measures of the forecast \cite{bannigidadmath2016stock, zhang2019forecasting}, resource allocation in urban planning \citep{burnett2014urban, liang2013resource}, or sustainability initiatives such as water and energy allocation \cite{syme1999defining, gebre2021multi, colett2016using}.
 We have demonstrated how tying forecast evaluation to a specific decision problem (e.g., via the allocation

### 2024-08-07 alloscore_manuscript/manuscript.Rnw
reviewer 2 edits

--- 2024-08-05 alloscore_manuscript/manuscript.Rnw
@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
 
@@
 
-While it should be noted that there are ways to interpret any of these scores abstractly through the lens of decision
-theory, and that all of the application-specific papers cited above benefited from direct collaboration with public
-health agencies, a key impetus for the present work has been that we were not able to find in any of them, nor in the
-literature they represent, explicit connections between how a forecast was evaluated and how that forecast was used in
-practice.
+While any of these scores can be interpreted through the lens of decision theory, they lack explicit connections between how a forecast is evaluated and how that forecast was used in practice.
+A key impetus for the present work was to develop a score that is interpretable to decision makers by measuring the extent to which forecasts lead to improved decisions.
+In addition, the allocation score we propose is expressed in units that are intelligible to decision makers.
 
@@
 rules such as the LS, CRPS, and WIS. Often models are ranked according to such a score, but without reference to any
-underlying decision problem from which the score might be derivable as a Bayes scoring rule or be otherwise motivated.
+underlying decision problem from which the score might be derivable as a Bayes scoring rule or be otherwise motivated.The 
 We argue that the value of collaborative forecasting efforts, such as the COVID-19 Forecasting Hub, could be enhanced by
 including evaluations of forecasts with
-scoring rules that directly consider the success of forecasts in supporting specific public health decisions.
-The work presented here has applications in many other fields such as economics, where the monetary consequences of using a forecast
-can be clearly juxtaposed with standard accuracy measures of the forecast \cite{bannigidadmath2016stock, zhang2019forecasting}, resource allocation in urban planning \citep{burnett2014urban, liang2013resource}, sustainability initiatives such as water and energy allocation \cite{syme1999defining, gebre2021multi, colett2016using}.
+scoring rules that directly consider the success of forecasts in supporting specific decisions.
+The allocation score was developed for this explicit purpose, evaluating forecasts in the context of the decisions they lead to.
+The AS is expressed in units that are directly relevant to decisions, which helps not only with ranking forecasts in the context of a specific action, but also allows decision-makers to contextualize the magnitude of differences between forecasts.
+The benefits of the work presented here generalize beyond public health, to any application of probability forecasts for decision-making, such as economics, where the monetary consequences of using a forecast
+can be clearly juxtaposed with standard accuracy measures of the forecast \cite{bannigidadmath2016stock, zhang2019forecasting}, resource allocation in urban planning \citep{burnett2014urban, liang2013resource}, or sustainability initiatives such as water and energy allocation \cite{syme1999defining, gebre2021multi, colett2016using}.
+
 We have demonstrated how tying forecast evaluation to a specific decision problem (e.g., via the allocation

### 2024-08-07 alloscore_manuscript/manuscript.Rnw
bisection explanation

--- 2024-08-05 alloscore_manuscript/manuscript.Rnw
@@
 
-\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations such that $F_1$ and $F_2$ have exponential marginal distribution with scale parameters $\sigma_1 = 1$ and $\sigma_2 = 4$.
+\paragraph{Example 1} \label{examp1} Suppose we have a forecast $F$ for need in two locations such that $F_1$ and $F_2$ have exponential marginal distribution with scale parameters $\sigma_1 = 1$ and $\sigma_2 = 4$.
 The means of these distributions are
@@
 
-\paragraph{Example 2} Now consider a different forecast that also has exponential distributions for resource need in
+\paragraph{Example 2} \label{examp2} Now consider a different forecast that also has exponential distributions for resource need in
 each location, but that has the scale parameters $\sigma_1 = 2$ and $\sigma_2 = 8$, twice as large as the scale
@@
 $F_i = \mathrm{Exp}(1/\sigma_i)$ for all $i$, then the Bayes act is proportional to $(\sigma_1,\ldots,\sigma_N)$, since
-$q_{\mathrm{Exp}(1/\sigma),\tau} = -\sigma \log(1-\tau)$.
+\begin{align}
+q_{\mathrm{Exp}(1/\sigma),\tau} = -\sigma \log(1-\tau). \label{eqn:exp-quantile}
+\end{align}
 
@@
 $q_{\alpha,F_i}$ of the marginal forecasts $F_i$, these allocations are given by $q_{\tau^{\star},F_i}$ where
-$\tau^{\star}$ solves the equation \eqref{eqn:quantiles-sum-to-K}. However, this equation is not analytically tractable
-and we must resort to a numerical method for finding an approximation $\tilde{\tau}$ of $\tau^{\star}$.
-
-We have implemented an iterative bisection method that makes use of the fact that $\sum_{i=1}^N q_{\tau,F_i}$ is an
-increasing function of $\tau$. The algorithm begins with an initial search interval $[\tau_{L,1}, \tau_{U,1}]$ (such as
+$\tau^{\star}$ solves the equation \eqref{eqn:quantiles-sum-to-K}.
+This equation is not analytically tractable in general, but because the left-hand side of \eqref{eqn:quantiles-sum-to-K} is non-decreasing in $\tau$, it is straightforward to find arbitrarily accurate approximations to a solution using an iterative bisection method.
+
+We have
+implemented such a method along with the resulting score computations in the R package \verb`alloscore`
+\citep{gerding-alloscore} which provided all allocation score values used in the analysis of section
+\ref{sec:application}.
+The procedure begins with an initial search interval $[\tau_{L,1}, \tau_{U,1}]$ (such as
 $[0,\max_{i}F_i(K)]$) that clearly contains the solution $\tau^{\star}$. At each step $j$ of the algorithm, we evaluate
@@
 \end{align}
-This search continues until $\tau_{U,j+1} < (1+\varepsilon)\tau_{L,j+1}$ for a suitably small $\varepsilon>0$. We have
-implemented this procedure along with the resulting score computations in the R package \verb`alloscore`
-\citep{gerding-alloscore} which provided all allocation score values used in the analysis of section
-\ref{sec:application}.
+This search continues until $\tau_{U,j+1} < (1+\varepsilon)\tau_{L,j+1}$ 
+for a suitably small $\varepsilon>0$.
+
+Figure \ref{fig:bisection} illustrates this bisection method in the context of the examples of Section \ref {sec:methods.overview}, where it quickly finds close approximations to the analytic solutions $\tau^{\star} = 1-e^{-1}, 1-e^{-2}$ (for $K= 5,10$ respectively) which, in this case, are available (c.f.
+\eqref{eqn:exp-quantile}).
+
+<<bisection, fig.height=4, fig.width = 8, fig.pos="H", fig.cap = "Illustrations of bisection method for the examples of Section \\ref{sec:methods.overview}.
+In each plot the thinner black lines are the quantile functions of the exponential forecasts for Location 1, with $\\sigma_1=1$ (lower), and Location 2, with $\\sigma_2=4$ (upper), while the thicker black line gives the sum of these quantiles, which must meet the resource constraints of $K = 5$ on the left and $K=10$ on the right.
+The red segments show the iterates $\\tau_{M,j}$ and their associated total resource demands as given by the function \\texttt{allocate} from the package \\texttt{alloscore} applied to these examples and using an initial search interval $[0,1]$.
+(The $\\tau_{M,j}$ are in this case just sums of terms $\\pm 2^{-j}$.) The horizontal dashed lines show the resulting allocations (c.f.
+Figure \\ref{fig:exp_alloc_example} ).">>=
+expexamp <- tibble(
+  target_names = LETTERS[1:2],
+  dist = "exp",
+  rate = c(1,.25),
+) %>% add_pdqr_funs(types = c("p", "d", "q"))
+
+alloc <- allocate(df = expexamp, K = c(5,10))
+
+q1 <- expexamp$Q[[1]]
+q2 <- expexamp$Q[[2]]
+q_sum <- function(tau) q1(tau) + q2(tau)
+iters5 <- tibble(taus = 1-alloc$lam_seq[[1]], qtot = q_sum(taus))
+iters10 <- tibble(taus = 1-alloc$lam_seq[[2]], qtot = q_sum(taus))
+
+p <- ggplot() +
+  geom_function(fun = q1, n=1000) +
+  geom_function(fun = q2, n=1000) +
+  geom_function(fun = q_sum, n=1000, linewidth = 1) +
+  labs(
+    x = expression(tau),
+    y = expression(paste(italic(q[tau * "," * F[1]]),
+                         ", ", 
+                         italic(q[tau * "," * F[2]]),
+                         "  and  ", 
+                         italic(q[tau * "," * F[1]] + q[tau * "," * F[2]])))) +
+  theme(panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        axis.ticks.length.x = unit(.25, "cm"))
+
+p5 <- p + geom_hline(aes(yintercept = 5)) +
+  geom_hline(aes(yintercept = c(1,4)), linetype = 2) +
+  geom_segment(data = iters5, aes(x=taus, xend = taus, y=0, yend = qtot), color = "red", alpha = .5) +
+  scale_x_continuous(limits = c(.45,1), expand = c(0,0), 
+                     breaks = c(.5,.6,0.632,.7,.8,.9,1),
+                     labels = c("0.5","0.6",expression(italic(atop(" ", 1 - e^{-1} %~~% 0.632))),
+                                "0.7","0.8","0.9","1")) +
+  scale_y_continuous(limits = c(0,12), breaks = c(0,1,4,5,10,15), expand = c(0,0))
+
+p10 <- p + geom_hline(aes(yintercept = 10)) +
+  geom_hline(aes(yintercept = c(2,8)), linetype = 2) +
+  geom_segment(data = iters10, aes(x=taus, xend = taus, y=0, yend = qtot), color = "red", alpha = .5) +
+  scale_x_continuous(limits = c(.45,1), expand = c(0,0), 
+                     breaks = c(.5,.6,.7,.8,0.865,.9,1),
+                     labels = c("0.5","0.6",
+                                "0.7","0.8", 
+                                expression(italic(atop(" ", 1 - e^{-1} %~~% 0.865)), "0.9", "1"))) +
+  scale_y_continuous(limits = c(0,12), breaks = c(0,2,5,8,10,15), expand = c(0,0), 
+                     name = element_blank())
+
+cowplot::plot_grid(p5, p10, align = "h")
+
+@
 

### 2024-08-07 alloscore_manuscript/manuscript.tex
bisection explanation

@@
   S_{IAS}(F, y) &= \int S_A(F,y; K) p(K) \, dK \label{eq:IASintegral}\\
-	& = \sum_{K=1}^{\infty} S_A(F,y; K) p(K) \label{eq:IASsum}.
+    & = \sum_{K=1}^{\infty} S_A(F,y; K) p(K) \quad (\text{assumming non-divisible resource units})\label{eq:IASsum}.
 \end{align}
@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu Aug  1 13:13:10 2024
+% Wed Aug  7 14:47:41 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu Aug  1 13:13:14 2024
+% Wed Aug  7 14:47:46 2024
 \begin{table}[ht]
@@
 
-To assess the sensitivity of AS to choice of $K$, AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 as well as for the \texttt{per-capita} allocation (Figure \ref{fig:multi-k}A).
-These calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of 19,581 that day.
-Model ranks by AS were fairly stable across all $K$, and there appears to be an substantial interval around the observed value over which ranks are constant.
+AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of
+hospitalizations on January 3, 2021 as well as for the \texttt{per-capita} allocation (Figure \ref{fig:multi-k}A). These
+calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of
+19,581 that day. Model ranks by AS were fairly stable across all $K$, and there appears
+to be an substantial interval around the observed value over which ranks are constant.
 
@@
 \begin{align*}
-	\hat{S}(F,y) = \sum_{k=1}^{600} S_A(F,y; 100k) p(100k),
+	\hat{S}(F,y) = \sum_{k=1}^{600} S_A(F,y; 200k) p(200k),
 \end{align*}
-where summation terms were calculated at every 100 beds with an upper bound of 60,000, or four times the estimated available beds.
-
-Approximate IAS was computed for two weight distributions on $K$, one uniform
+where summation terms were calculated at every 200 beds with an upper bound of 60,000, or four times the estimated available beds.
+An approximate IAS was computed for two weight distributions on $K$, one uniform
 across the entire range and the other a truncated normal distribution centered at 15,000 (the orange and blue shaded areas in Figure
@@
 
-Additionally, during an epidemic, public health decision makers may be faced with an ongoing series of decision over time in response to changes in conditions such as resource availability or the current allocations of resources.
-The AS presented here looks at a single snapshot in time  and ignores many of the complexities related to adjusting existing allocations.
-The example presented in section \ref{sec:application} is based on a prospective estimate of bed availability, but one could also use observed total resource availability to derive an optimal set of prospective AS parameters.
-In the case of allocation of preventative resources, such as vaccines, one may which to account for the negative feedback loop between resource allocation and disease cases.
-Alternatively, the impact of unmet need often differs across different locations and demographics, which could be accounted for by incorporating a function for the cost of unmet need, which differs by location or other covariates.
+Additionally, during an epidemic, public health decision makers may be faced with an ongoing series of decisions over time in response to changes in conditions such as resource availability or the current effective allocations of resources.
+The AS presented here looks at a single snapshot in time  and ignores many of the possible complexities related to adjusting existing allocations.
+The example presented in section
+\ref{sec:application} is based on a prospective estimate of bed availability, but one could also use observed total resource availability to target an optimal set of prospective AS parameters. 
+Alternatively, the impact of unmet need often differs across different locations and demographics, which could be accounted for by incorporating a function of location and demographics which defines the cost of unmet need.
 Future work is needed to explore the use of AS while accounting for the dynamic nature of disease spread and resource availability.
@@
 disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those
-for which they are able to offer useful guidance. The work presented here New collaborative work between decision-makers and modeling teams is
+for which they are able to offer useful guidance. New collaborative work between decision-makers and modeling teams is
 needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or
@@
 \end{align}
-This search continues until $\tau_{U,j+1} < (1+\varepsilon)\tau_{L,j+1}$ for a suitably small $\varepsilon>0$. We have
-implemented this procedure along with the resulting score computations in the R package \verb`alloscore`
-\citep{gerding-alloscore} which provided all allocation score values used in the analysis of section
-\ref{sec:application}.
+This search continues until $\tau_{U,j+1} < (1+\varepsilon)\tau_{L,j+1}$ 
+for a suitably small $\varepsilon>0$.
+
+Figure \ref{fig:bisection} illustrates this bisection method in the context of the examples of Section \ref {sec:methods.overview}, where it quickly finds close approximations to the analytic solutions $\tau^{\star} = 1-e^{-1}, 1-e^{-2}$ (for $K= 5,10$ respectively) which, in this case, are available (c.f.
+\eqref{eqn:exp-quantile}).
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
+\includegraphics[width=\maxwidth]{figure/bisection-1} \caption{Illustrations of bisection method for the examples of Section \ref{sec:methods.overview}.
+In each plot the thinner black lines are the quantile functions of the exponential forecasts for Location 1, with $\sigma_1=1$ (lower), and Location 2, with $\sigma_2=4$ (upper), while the thicker black line gives the sum of these quantiles, which must meet the resource constraints of $K = 5$ on the left and $K=10$ on the right.
+The red segments show the iterates $\tau_{M,j}$ and their associated total resource demands as given by the function \texttt{allocate} from the package \texttt{alloscore} applied to these examples and using an initial search interval $[0,1]$.
+(The $\tau_{M,j}$ are in this case just sums of terms $\pm 2^{-j}$.) The horizontal dashed lines show the resulting allocations (c.f.
+Figure \ref{fig:exp_alloc_example} ).}\label{fig:bisection}
+\end{figure}
+
+\end{knitrout}
 

### 2024-08-07 random_undocumented_experiments/iter_pic.R

@@ -22,7 +22,13 @@ make_iter_pic(K = 20,
                 list(dist = "unif", min = 7, max = 15)
               ))
 
-make_iter_pic(K = 2,
+pk10 <- make_iter_pic(K = 10,
+              dists_and_params = list(
+                list(dist = "exp", rate = 1),
+                list(dist = "exp", rate = 1/4)
+              ))
+
+pk5 <- make_iter_pic(K = 5,
               dists_and_params = list(
                 list(dist = "exp", rate = 1),
                 list(dist = "exp", rate = 1/4)

### 2024-08-07 random_undocumented_experiments/iter_pic2.R

@@ -18,9 +18,9 @@ library(alloscore)
     add_pdqr_funs() %>% mutate(
       q_scale = map2_dbl(Q, alpha, exec),
       # add marginal expected benefit functions scaled to fit on single [0,1] interval,
-      Lambda_scaled = pmap(., margexb_fun),
+      Lambda_scaled = pmap(select(., F, alpha), alloscore::dexp_gpl_loss),
       # and unscaled as well, which are not yet being used
-      Lambda = pmap(.[names(.) != "q_scale"], margexb_fun),
+      Lambda = pmap(select(., F, alpha),  alloscore::dexp_gpl_loss),
       #Eloss = pmap(.[c("alpha", "f")], gpl_loss_exp_fun)
     )
 )

### 2024-08-07 random_undocumented_experiments/iter_pic_funs.R

@@ -41,12 +41,13 @@ make_iter_pic <- function(K, alpha = 1, w = 1, dists_and_params) {
          p1 <- ggplot() + xlim(c(0,xmax)) + ylim(0, ymax) +
            geom_hline(yintercept = 0) +
            map(lam_seq[[1]], ~ geom_hline(yintercept = ., alpha = .3)) +
-           geom_label(data = enframe(lam_seq[[1]]), aes(x=0,y=value,label=name)) +
+           geom_label(data = enframe(lam_seq[[1]]), aes(x=0,y=value,label=name), label.size = .05) +
            map(1:n, ~ geom_function(aes(color = dat$name[[.]]), fun = dat$Lambda[[.]])) +
            map(1:n, ~ geom_vline(aes(xintercept = x[[1]][.], color = dat$name[[.]]), alpha=.5)) +
            labs(color = "Distribution",
-                y = expression(lambda(x)),
+                y = expression(1-tau),
                 x = expression(x[i])) +
+           scale_x_continuous(breaks = seq(0, xmax, by = 1)) +
            theme_classic() +
            theme(legend.position = "left")
 
@@ -61,7 +62,8 @@ make_iter_pic <- function(K, alpha = 1, w = 1, dists_and_params) {
            # #geom_line(aes(x =  K_funfac(forecasts = c(1,2,4))(y))) +
            #map(1:n, ~ geom_line(aes(x = K_funfac(forecasts = c(.))(y), color = name[[.]]))) +
            geom_vline(xintercept = K) +
-           labs(x = expression(w ^ T * Q(alpha - w * lambda))) +
+           labs(x =expression(sum( q[tau * ", " * F[i]],i == 1, N))
+                ) +
            theme(
              axis.text.y = element_blank(),
              axis.ticks.y = element_blank(),

### 2024-08-07 alloscore_manuscript/manuscript.Rnw
fixed typo in intro

--- 2024-08-07 alloscore_manuscript/manuscript.Rnw
@@
 rules such as the LS, CRPS, and WIS. Often models are ranked according to such a score, but without reference to any
-underlying decision problem from which the score might be derivable as a Bayes scoring rule or be otherwise motivated.The 
+underlying decision problem from which the score might be derivable as a Bayes scoring rule or be otherwise motivated. 
 We argue that the value of collaborative forecasting efforts, such as the COVID-19 Forecasting Hub, could be enhanced by

### 2024-08-07 alloscore_manuscript/manuscript.Rnw

--- 2024-08-07 alloscore_manuscript/manuscript.Rnw
@@
 including evaluations of forecasts with
-scoring rules that directly consider the success of forecasts in supporting specific public health decisions.
-The work presented here has applications in many other fields such as economics, where the monetary consequences of using a forecast
-can be clearly juxtaposed with standard accuracy measures of the forecast \cite{bannigidadmath2016stock, zhang2019forecasting}, resource allocation in urban planning \citep{burnett2014urban, liang2013resource}, sustainability initiatives such as water and energy allocation \cite{syme1999defining, gebre2021multi, colett2016using}.
+scoring rules that directly consider the success of forecasts in supporting specific decisions.
+The allocation score was developed for this explicit purpose, evaluating forecasts in the context of the decisions they lead to.
+The AS is expressed in units that are directly relevant to decisions, which helps not only with ranking forecasts in the context of a specific action, but also allows decision-makers to contextualize the magnitude of differences between forecasts.
+The benefits of the work presented here generalize beyond public health, to any application of probability forecasts for decision-making, such as economics, where the monetary consequences of using a forecast
+can be clearly juxtaposed with standard accuracy measures of the forecast \cite{bannigidadmath2016stock, zhang2019forecasting}, resource allocation in urban planning \citep{burnett2014urban, liang2013resource}, or sustainability initiatives such as water and energy allocation \cite{syme1999defining, gebre2021multi, colett2016using}.
+
 We have demonstrated how tying forecast evaluation to a specific decision problem (e.g., via the allocation

### 2024-08-07 alloscore_manuscript/manuscript.Rnw
Made most of the 'minor edits' from Associate Editor

--- 2024-08-07 alloscore_manuscript/manuscript.Rnw
@@
 to assess probabilistic forecasts (based on random effect models) of dengue incidence at the district level in Vietnam
-\citep{colon-gonzalez_probabilistic_2021}. And the WIS has been used during the COVID-19 pandemic to evaluate forecasts
+\citep{colon-gonzalez_probabilistic_2021}. Likewise, the WIS has been used during the COVID-19 pandemic to evaluate forecasts
 of observed cases, hospitalizations and deaths in the US and Europe, as reported by municipal, state, and federal
@@
 aware of only a limited literature attempting to connect the value of \emph{infectious disease} forecasts to their
-impact on and through policy.  And within this body of work we have found the discussion of such a connection to still
+impact on and through policy.  Even within this body of work we have found the discussion of such a connection to
 be at a formally and quantitatively imprecise stage. In \cite{ioannidis2022forecastingCOVIDfailed}, the possible negative

### 2024-08-07 alloscore_manuscript/manuscript.Rnw

@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
 
@@
 \ref{sec:application} is based on a prospective estimate of bed availability, but one could also use observed total resource availability to target an optimal set of prospective AS parameters. 
-Alternatively, the impact of unmet need often differs across different locations and demographics, which could be accounted for by incorporating a function of location and demographics which defines the cost of unmet need.
+Alternatively, the impact of unmet need often differs across different locations and demographics, thus one may wish to define a varying cost which depends on some covariates.
 Future work is needed to explore the use of AS while accounting for the dynamic nature of disease spread and resource availability.

### 2024-08-07 alloscore_manuscript/manuscript.Rnw
typo in bisection fig

--- 2024-08-07 alloscore_manuscript/manuscript.Rnw
@@
                                 "0.7","0.8", 
-                                expression(italic(atop(" ", 1 - e^{-1} %~~% 0.865)), "0.9", "1"))) +
+                                expression(italic(atop(" ", 1 - e^{-2} %~~% 0.865)), "0.9", "1"))) +
   scale_y_continuous(limits = c(0,12), breaks = c(0,2,5,8,10,15), expand = c(0,0), 

### 2024-08-07 alloscore_manuscript/manuscript.tex
typo in bisection fig

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Wed Aug  7 14:47:41 2024
+% Wed Aug  7 16:57:25 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Wed Aug  7 14:47:46 2024
+% Wed Aug  7 16:57:30 2024
 \begin{table}[ht]

### 2024-08-08 alloscore_manuscript/manuscript.Rnw
update diff.tex

@@
 
-Figure \ref{fig:bisection} illustrates this bisection method in the context of the examples of Section \ref {sec:methods.overview}, where it quickly finds close approximations to the analytic solutions $\tau^{\star} = 1-e^{-1}, 1-e^{-2}$ (for $K= 5,10$ respectively) which, in this case, are available (c.f.
+Figure \ref{fig:bisection} illustrates this bisection method in the context of the examples of Section \ref{sec:methods.overview}, where it quickly finds close approximations to the analytic solutions $\tau^{\star} = 1-e^{-1}, 1-e^{-2}$ (for $K= 5,10$ respectively) which, in this case, are available (c.f.
 \eqref{eqn:exp-quantile}).
@@
 <<bisection, fig.height=4, fig.width = 8, fig.pos="H", fig.cap = "Illustrations of bisection method for the examples of Section \\ref{sec:methods.overview}.
-In each plot the thinner black lines are the quantile functions of the exponential forecasts for Location 1, with $\\sigma_1=1$ (lower), and Location 2, with $\\sigma_2=4$ (upper), while the thicker black line gives the sum of these quantiles, which must meet the resource constraints of $K = 5$ on the left and $K=10$ on the right.
+In each plot the thinner black curves are the quantile functions of the exponential forecasts for Location 1, with $\\sigma_1=1$ (lower), and Location 2, with $\\sigma_2=4$ (upper), while the thicker black curve gives the sum of these quantiles, which must meet the resource constraints of $K = 5$ on the left and $K=10$ on the right.
 The red segments show the iterates $\\tau_{M,j}$ and their associated total resource demands as given by the function \\texttt{allocate} from the package \\texttt{alloscore} applied to these examples and using an initial search interval $[0,1]$.
@@
   scale_y_continuous(limits = c(0,12), breaks = c(0,2,5,8,10,15), expand = c(0,0), 
-                     name = element_blank())
+                     name = "  ")
 

### 2024-08-08 alloscore_manuscript/manuscript.tex
update diff.tex

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Wed Aug  7 16:57:25 2024
+% Wed Aug  7 23:13:42 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Wed Aug  7 16:57:30 2024
+% Wed Aug  7 23:13:46 2024
 \begin{table}[ht]
@@
 \includegraphics[width=\maxwidth]{figure/bisection-1} \caption{Illustrations of bisection method for the examples of Section \ref{sec:methods.overview}.
-In each plot the thinner black lines are the quantile functions of the exponential forecasts for Location 1, with $\sigma_1=1$ (lower), and Location 2, with $\sigma_2=4$ (upper), while the thicker black line gives the sum of these quantiles, which must meet the resource constraints of $K = 5$ on the left and $K=10$ on the right.
+In each plot the thinner black curves are the quantile functions of the exponential forecasts for Location 1, with $\sigma_1=1$ (lower), and Location 2, with $\sigma_2=4$ (upper), while the thicker black curve gives the sum of these quantiles, which must meet the resource constraints of $K = 5$ on the left and $K=10$ on the right.
 The red segments show the iterates $\tau_{M,j}$ and their associated total resource demands as given by the function \texttt{allocate} from the package \texttt{alloscore} applied to these examples and using an initial search interval $[0,1]$.

### 2024-08-13 alloscore_manuscript/manuscript.Rnw
enumerate limitations

@@
 context of infectious disease, this means that we do not consider resources, such as vaccines, that are intended to
-reduce the number of people who will become sick at some point in the future. Instead, our setup addresses resources
+reduce the number of people who will become sick within the forecast horizon. Instead, our setup addresses resources
 like hospital beds, oxygen supply, or ventilators which are intended to meet the medical needs of patients who are
@@
 
-\paragraph{Step 1: specify a loss function.} The loss incurred by an allocation $x$ under an outcome $y$ is the sum of
+\paragraph{Step 1: Specify a loss function.} The loss incurred by an allocation $x$ under an outcome $y$ is the sum of
 contributions from unmet need in each location:
@@
 \begin{align}
-  S_A^{\text{raw}}(F, y; K) = s_A(x^{F,K}, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i^{F,K}).
+  S_A^{\text{raw}}(F, y; K) = s_A(x^{F,K}, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i^{F,K}). \label{eqn:AS-raw}
 \end{align}
@@
   S_A(F, y; K)  &= S_A^{\text{raw}}(F, y; K) - L \cdot \max\left(0,\sum_{i=1}^{N}y_i - K\right) \\
-  &= L\left\{\sum_{i=1}^N \max(0, y_i - x_i^{F,K}) -  \max\left(0,\sum_{i=1}^{N}y_i - K\right)\right\}.
+  &= L\left\{\sum_{i=1}^N \max(0, y_i - x_i^{F,K}) -  \max\left(0,\sum_{i=1}^{N}y_i - K\right)\right\}. \label{eqn:AS-orc}
 \end{align}
@@
 
-\subsection{Connections to Other Scores}
+\subsection{Comparisons with Other Scores}
 \label{sec:methods.related}
@@
 \ref{sec:methods.detailed.integrated_allocation} \citep{gneiting2011weightedScoringRules}.
+
+We also note that the AS, just as the (aggregate or mean) quantile score and WIS, depends only on the marginal distributions for each location of a multi-location forecast (cf.
+equations \eqref{eqn:AS-raw} and \eqref{eqn:quantiles-sum-to-K} in Section \ref{sec:a:bayes-quantiles} of the appendix).
+It is however the case that when a forecaster modifies a single marginal forecast (holding those for other locations fixed), the aggregate WIS of thier forecast will change only according to the WIS of that marginal forecast, whereas every term in the sum defining the AS (in \eqref{eqn:AS-raw}) will change as the Bayes probability level defined by \eqref{eqn:quantiles-sum-to-K} changes.
+In this sense, the AS depends “jointly” on the marginal forecasts while the WIS does not.
 
@@
 the \emph{relative} magnitude of future resource need across different locations.  Success at this task is not, however,
-directly rewarded by the WIS.
+directly rewarded by the WIS (cf. the discussion at the end of Section \ref{sec:methods.related}).
 
@@
 optimization perspective (not from a forecast evaluation perspective) in \cite{huang_stockpiling_2017} and provided an
-initial motivation for this project. We see it as promising focus for further development of the AS framework since it
+initial motivation for this project. We see this as a promising focus for further development of the AS framework since it
 (a) has been identified as a setting in which allocation decisions are made during respiratory disease outbreaks, (b) involves
@@
 
-Additionally, during an epidemic, public health decision makers may be faced with an ongoing series of decisions over time in response to changes in conditions such as resource availability or the current effective allocations of resources.
-The AS presented here looks at a single snapshot in time  and ignores many of the possible complexities related to adjusting existing allocations.
-The example presented in section
-\ref{sec:application} is based on a prospective estimate of bed availability, but one could also use observed total resource availability to target an optimal set of prospective AS parameters. 
-Alternatively, the impact of unmet need often differs across different locations and demographics, thus one may wish to define a varying cost which depends on some covariates.
-Future work is needed to explore the use of AS while accounting for the dynamic nature of disease spread and resource availability.
-
-There are several important limitations to the current work. The allocation score we developed here does not directly
-account for important considerations such as fairness or equity of allocations, or more broadly, individual and group
+We reiterate three important limitations to the current work which we noted when introducing the AS in 
+Section \ref{sec:methods.detailed.specific_allocation}.  We view these issues as promising avenues for 
+future investigation, but beyond the scope of this paper.
+
+First, the terms in our loss function  
+\eqref{eqn:loss_fn} do not depend on geographic or demographic covariates.  Policy makers may, however, wish to incorporate 
+differences in the impact of unmet need across populations into a forecast evaluation methodology.  For example, in the state-level forecasting context of Section \ref{sec:application}, it might be
+appropriate to make the loss for failing to
+procuring a bed for a hospitalizable illness an increasing function of the state's median age. 
+Along these lines, it should also be mentioned that the AS does not directly
+take into consideration the fairness or equity of allocations, or more broadly, individual and group
 preference relations that are difficult or even impossible to encode into a loss or utility function.  Ambiguity aversion
@@
 response, where there can be immense social and political pressure on public health officials to maintain a transparent
-base of evidence for their policy choices and recommendations. The proposed framework also does not attempt to capture
-the broader context of decision making. For example, in practice it may be possible to increase the resource constraint
-$K$ by shifting funding from other disease mitigation measures. Finally, we emphasize that we opted at this stage of
+base of evidence for their policy choices and recommendations.
+
+Secondly, our formulation of the AS does not take into
+consideration the temporal inter-dependencies between sequential decisions that public health decision makers are likely to face as an epidemic unfolds.
+Such dependencies could arise from time-varying inputs such as resource availability or the current effective allocations of resources.
+The sequential decision problem could also be combined with others allowing constraints for different resources to be modified by time-varying funding shifts between various disease mitigation measures.
+Incorporating these possibilities into an AS framework would likely require technical developments involving either dynamic programming
+(see e.g., \cite{bertsekas2012dynamic}) or a scenario-based approach (see e.g., \cite{pflug2014multistage}).
+
+Thirdly, we emphasize that we opted at this stage of
 development to explicitly rule out situations where a successful
-epidemiological forecast could lead to policy decisions that change the distribution of the predicted outcome $Y$. Our
-framework would need considerable enhancements before being applicable to forecast evaluation beyond horizons for which
-causal feedback can be neglected.
-
-An opportunity for further investigation is to more carefully evaluate whether forecasts add value to 
+epidemiological forecast could lead to policy decisions that change the distribution of the predicted outcome $Y$. 
+This excludes the fundamentally important context of vaccine allocation for certain pathogens and
+spatio-temporal scales, but we note that supply-chain decisions made at a national level in the US determining the state-level
+availability of current COVID-19 vaccines are unlikely to affect incidence or exposure levels at a 14-day horizon. In any event,
+we are not yet prepared to suggest how valid forecast evaluation using scores given by \eqref{eqn:AS-orc} would proceed when 
+both disease burden and unmet need are considered as endogenous variables.
+
+Another opportunity for further investigation is to more carefully evaluate whether forecasts add value to 
 standard  procedures already in place for making public health decisions. In the context of allocations,
@@
 
-Figure \ref{fig:bisection} illustrates this bisection method in the context of the examples of Section \ref{sec:methods.overview}, where it quickly finds close approximations to the analytic solutions $\tau^{\star} = 1-e^{-1}, 1-e^{-2}$ (for $K= 5,10$ respectively) which, in this case, are available (c.f.
+Figure \ref{fig:bisection} illustrates this bisection method in the context of the examples of Section \ref{sec:methods.overview}, where it quickly finds close approximations to the analytic solutions $\tau^{\star} = 1-e^{-1}, 1-e^{-2}$ (for $K= 5,10$ respectively) which, in this case, are available (cf.
 \eqref{eqn:exp-quantile}).
@@
 The red segments show the iterates $\\tau_{M,j}$ and their associated total resource demands as given by the function \\texttt{allocate} from the package \\texttt{alloscore} applied to these examples and using an initial search interval $[0,1]$.
-(The $\\tau_{M,j}$ are in this case just sums of terms $\\pm 2^{-j}$.) The horizontal dashed lines show the resulting allocations (c.f.
+(The $\\tau_{M,j}$ are in this case just sums of terms $\\pm 2^{-j}$.) The horizontal dashed lines show the resulting allocations (cf.
 Figure \\ref{fig:exp_alloc_example} ).">>=
@@
 In general, $G^q(F)$ and $G^\star(F)$ will be different distributions:  matching $F$ at specific quantiles
-does not require $G^q(F)$ to match $F$ at the quantiles for $\tau^{F,K}$ (c.f. \eqref{eqn:quantiles-sum-to-K}),
+does not require $G^q(F)$ to match $F$ at the quantiles for $\tau^{F,K}$ (cf. \eqref{eqn:quantiles-sum-to-K}),
 which would be necessary for it to share $x^F$ as an optimal allocation.

### 2024-08-13 alloscore_manuscript/manuscript.tex
enumerate limitations

@@
 
+We also note that the AS, just as the (aggregate or mean) quantile score and WIS, depends only on the marginal distributions for each location of a multi-location forecast (cf.
+equations \eqref{eqn:AS-raw} and \eqref{eqn:quantiles-sum-to-K} in Section \ref{sec:a:bayes-quantiles} of the appendix).
+It is however the case that when a forecaster modifies a single marginal forecast (holding those for other locations fixed), the aggregate WIS of thier forecast will change only according to the WIS of that marginal forecast, whereas every term in the sum defining the AS (in \eqref{eqn:AS-raw}) will change as the Bayes probability level defined by \eqref{eqn:quantiles-sum-to-K} changes.
+In this sense, the AS depends “jointly” on the marginal forecasts while the WIS does not.
+
 \section{Evaluating forecasts of COVID hospitalizations using the allocation score}
@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Wed Aug  7 23:13:42 2024
+% Tue Aug 13 15:15:48 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Wed Aug  7 23:13:46 2024
+% Tue Aug 13 15:15:52 2024
 \begin{table}[ht]
@@
 The red segments show the iterates $\tau_{M,j}$ and their associated total resource demands as given by the function \texttt{allocate} from the package \texttt{alloscore} applied to these examples and using an initial search interval $[0,1]$.
-(The $\tau_{M,j}$ are in this case just sums of terms $\pm 2^{-j}$.) The horizontal dashed lines show the resulting allocations (c.f.
+(The $\tau_{M,j}$ are in this case just sums of terms $\pm 2^{-j}$.) The horizontal dashed lines show the resulting allocations (cf.
 Figure \ref{fig:exp_alloc_example} ).}\label{fig:bisection}

### 2024-08-15 alloscore_manuscript/manuscript.Rnw
some updates in response to reviewer comments

@@
 context of infectious disease, this means that we do not consider resources, such as vaccines, that are intended to
-reduce the number of people who will become sick within the forecast horizon. Instead, our setup addresses resources
+reduce the number of people who will become sick at some point in the future. Instead, our setup addresses resources
 like hospital beds, oxygen supply, or ventilators which are intended to meet the medical needs of patients who are
@@
 distribution $p$:
-\begin{align} 
-  S_{IAS}(F, y) &= \int S_A(F,y; K) p(K) \, dK \label{eq:IASintegral}\\
-    & = \sum_{K=1}^{\infty} S_A(F,y; K) p(K) \quad (\text{assumming non-divisible resource units})\label{eq:IASsum}.
-\end{align}
+\begin{equation}
+S_{IAS}(F, y) = \int S_A(F,y; K) p(K) \, dK \label{eqn:IAS_definition}
+\end{equation}
 We note that the device of considering a range of hypothetical decision makers or decision making problems with
 different problem parameters has been employed in the past \cite[e.g.,][]{murphy1993whatisagoodforecast}.
+
+In practice, it may be challenging to compute the integral in \eqref{eqn:IAS_definition}.
+In the applied investigations below we resort to a discretization, evaluating at a grid of values of the resource constraint $K$.
+A discrete treatment along these lines will often be natural, in settings where the resource in question is non-divisible (e.g., ventilators, masks).
 
@@
 
-AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of
+AS was computed for a range of $K$ from 200 to 60,000 at increments of 200 for forecasts made on December 20, 2020 predicting levels of
 hospitalizations on January 3, 2021 as well as for the \texttt{per-capita} allocation (Figure \ref{fig:multi-k}A). These
@@
 \ref{sec:methods.detailed.integrated_allocation}). 
-Calculating exact IAS as defined in equation \eqref{eq:IASsum} is computationally intensive so an approximate IAS, denoted $\hat{S}(F,y)$, was used such that
-\begin{align*}
-	\hat{S}(F,y) = \sum_{k=1}^{600} S_A(F,y; 200k) p(200k),
-\end{align*}
-where summation terms were calculated at every 200 beds with an upper bound of 60,000, or four times the estimated available beds.
-An approximate IAS was computed for two weight distributions on $K$, one uniform
-across the entire range and the other a truncated normal distribution centered at 15,000 (the orange and blue shaded areas in Figure
-\ref{fig:multi-k}A, respectively). Both versions of the IAS were correlated with the original AS for $K=15,000$, with
+IAS was computed for two weight distributions on the same grid of values of $K$ used in the sensitivity analysis above,
+one uniform across the entire range and the other with weights proportional a truncated normal distribution centered at 15,000
+(the orange and blue shaded areas in Figure
+\ref{fig:multi-k}A, respectively).
+With these weight distributions, the IAS defined in Equation~\eqref{eqn:IAS_definition} can be computed as the sum
+$$S_{IAS}(F, y) = \sum_K S_A(F,y; K) p(K)$$
+
+Both versions of the IAS were correlated with the original AS for $K=15,000$, with
 the higher correlation coming from the weighting centered at $K=15,000$ (Figure \ref{fig:multi-k}B). Model rankings
@@
 scoring rules that directly consider the success of forecasts in supporting specific decisions.
-The allocation score was developed for this explicit purpose, evaluating forecasts in the context of the decisions they lead to.
-The AS is expressed in units that are directly relevant to decisions, which helps not only with ranking forecasts in the context of a specific action, but also allows decision-makers to contextualize the magnitude of differences between forecasts.
-The benefits of the work presented here generalize beyond public health, to any application of probability forecasts for decision-making, such as economics, where the monetary consequences of using a forecast
-can be clearly juxtaposed with standard accuracy measures of the forecast \cite{bannigidadmath2016stock, zhang2019forecasting}, resource allocation in urban planning \citep{burnett2014urban, liang2013resource}, or sustainability initiatives such as water and energy allocation \cite{syme1999defining, gebre2021multi, colett2016using}.
+The allocation score was developed for this purpose.
+The AS is expressed in units that are directly relevant to decisions, which helps not only with ranking forecasts in the context of a specific action, but also allows decision-makers to contextualize the magnitude of differences between forecast scores.
+The benefits of the approach to forecast evaluation here generalize beyond public health, to any application of probability forecasts for decision-making, such as economics, where the monetary consequences of using a forecast
+can be clearly juxtaposed with standard accuracy measures of the forecast \citep{bannigidadmath2016stock, zhang2019forecasting}, resource allocation in urban planning \citep{burnett2014urban, liang2013resource}, or sustainability initiatives such as water and energy allocation \citep{syme1999defining, gebre2021multi, colett2016using}.
 
@@
 the \emph{relative} magnitude of future resource need across different locations.  Success at this task is not, however,
-directly rewarded by the WIS (cf. the discussion at the end of Section \ref{sec:methods.related}).
+directly rewarded by the WIS (see the discussion at the end of Section \ref{sec:methods.related}).
 
@@
 
-We reiterate three important limitations to the current work which we noted when introducing the AS in 
+There are three important limitations to the current work which we noted when introducing the AS in 
 Section \ref{sec:methods.detailed.specific_allocation}.  We view these issues as promising avenues for 
-future investigation, but beyond the scope of this paper.
+future investigation.
 
@@
 differences in the impact of unmet need across populations into a forecast evaluation methodology.  For example, in the state-level forecasting context of Section \ref{sec:application}, it might be
-appropriate to make the loss for failing to
-procuring a bed for a hospitalizable illness an increasing function of the state's median age. 
+appropriate to adjust the loss so that the costs of unmet need are larger in locations with greater population vulnerability.
 Along these lines, it should also be mentioned that the AS does not directly
@@
 epidemiological forecast could lead to policy decisions that change the distribution of the predicted outcome $Y$. 
-This excludes the fundamentally important context of vaccine allocation for certain pathogens and
-spatio-temporal scales, but we note that supply-chain decisions made at a national level in the US determining the state-level
-availability of current COVID-19 vaccines are unlikely to affect incidence or exposure levels at a 14-day horizon. In any event,
-we are not yet prepared to suggest how valid forecast evaluation using scores given by \eqref{eqn:AS-orc} would proceed when 
-both disease burden and unmet need are considered as endogenous variables.
+For example, this excludes the important problem of allocation of doses of a preventative vaccine.
+We have not developed a formal approach to forecast evaluation using scores given by \eqref{eqn:AS-orc} when both
+disease burden and unmet need are considered as endogenous variables. Our sense is that such an approach would require some form of
+accounting for the causal effect of the allocation decision on observed measures of disease burden, e.g. estimating the (counterfactual)
+cases or deaths averted from one or another allocation strategy through the available observed data.
 

### 2024-08-15 alloscore_manuscript/manuscript.Rnw
wording cleanup
Update alloscore_manuscript/manuscript.Rnw

@@
 We have not developed a formal approach to forecast evaluation using scores given by \eqref{eqn:AS-orc} when both
-disease burden and unmet need are considered as endogenous variables. Our sense is that such an approach would require some form of
-accounting for the causal effect of the allocation decision on observed measures of disease burden, e.g. estimating the (counterfactual)
-cases or deaths averted from one or another allocation strategy through the available observed data.
+disease burden and unmet need are considered as endogenous variables. Such an approach would require some form of
+accounting for the causal effect of the allocation decision on observed measures of disease burden, e.g. using the available observed data to estimate the (counterfactual)
+number of cases or deaths averted from one or another allocation strategy.
 

### 2024-08-15 alloscore_manuscript/manuscript.Rnw
Update alloscore_manuscript/manuscript.Rnw

@@
 IAS was computed for two weight distributions on the same grid of values of $K$ used in the sensitivity analysis above,
-one uniform across the entire range and the other with weights proportional a truncated normal distribution centered at 15,000
+one uniform across the entire range and the other with weights proportional to a truncated normal distribution centered at 15,000
 (the orange and blue shaded areas in Figure

### 2024-08-15 alloscore_manuscript/manuscript.Rnw
Update alloscore_manuscript/manuscript.Rnw

@@
 The AS is expressed in units that are directly relevant to decisions, which helps not only with ranking forecasts in the context of a specific action, but also allows decision-makers to contextualize the magnitude of differences between forecast scores.
-The benefits of the approach to forecast evaluation here generalize beyond public health, to any application of probability forecasts for decision-making, such as economics, where the monetary consequences of using a forecast
+The benefits of the approach to forecast evaluation discussed here generalize beyond public health, to any application of probability forecasts for decision-making, such as economics, where the monetary consequences of using a forecast
 can be clearly juxtaposed with standard accuracy measures of the forecast \citep{bannigidadmath2016stock, zhang2019forecasting}, resource allocation in urban planning \citep{burnett2014urban, liang2013resource}, or sustainability initiatives such as water and energy allocation \citep{syme1999defining, gebre2021multi, colett2016using}.

### 2024-08-16 alloscore_manuscript/manuscript.tex
knit/latexdiff Evan's changes

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Tue Aug 13 15:15:48 2024
+% Thu Aug 15 23:41:50 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Tue Aug 13 15:15:52 2024
+% Thu Aug 15 23:41:54 2024
 \begin{table}[ht]
@@
 \ref{sec:methods.detailed.integrated_allocation}). 
-Calculating exact IAS as defined in equation \eqref{eq:IASsum} is computationally intensive so an approximate IAS, denoted $\hat{S}(F,y)$, was used such that
-\begin{align*}
-	\hat{S}(F,y) = \sum_{k=1}^{600} S_A(F,y; 200k) p(200k),
-\end{align*}
-where summation terms were calculated at every 200 beds with an upper bound of 60,000, or four times the estimated available beds.
-An approximate IAS was computed for two weight distributions on $K$, one uniform
-across the entire range and the other a truncated normal distribution centered at 15,000 (the orange and blue shaded areas in Figure
-\ref{fig:multi-k}A, respectively). Both versions of the IAS were correlated with the original AS for $K=15,000$, with
+IAS was computed for two weight distributions on the same grid of values of $K$ used in the sensitivity analysis above,
+one uniform across the entire range and the other with weights proportional to a truncated normal distribution centered at 15,000
+(the orange and blue shaded areas in Figure
+\ref{fig:multi-k}A, respectively).
+With these weight distributions, the IAS defined in Equation~\eqref{eqn:IAS_definition} can be computed as the sum
+$$S_{IAS}(F, y) = \sum_K S_A(F,y; K) p(K)$$
+
+Both versions of the IAS were correlated with the original AS for $K=15,000$, with
 the higher correlation coming from the weighting centered at $K=15,000$ (Figure \ref{fig:multi-k}B). Model rankings
@@
 scoring rules that directly consider the success of forecasts in supporting specific decisions.
-The allocation score was developed for this explicit purpose, evaluating forecasts in the context of the decisions they lead to.
-The AS is expressed in units that are directly relevant to decisions, which helps not only with ranking forecasts in the context of a specific action, but also allows decision-makers to contextualize the magnitude of differences between forecasts.
-The benefits of the work presented here generalize beyond public health, to any application of probability forecasts for decision-making, such as economics, where the monetary consequences of using a forecast
-can be clearly juxtaposed with standard accuracy measures of the forecast \cite{bannigidadmath2016stock, zhang2019forecasting}, resource allocation in urban planning \citep{burnett2014urban, liang2013resource}, or sustainability initiatives such as water and energy allocation \cite{syme1999defining, gebre2021multi, colett2016using}.
+The allocation score was developed for this purpose.
+The AS is expressed in units that are directly relevant to decisions, which helps not only with ranking forecasts in the context of a specific action, but also allows decision-makers to contextualize the magnitude of differences between forecast scores.
+The benefits of the approach to forecast evaluation discussed here generalize beyond public health, to any application of probability forecasts for decision-making, such as economics, where the monetary consequences of using a forecast
+can be clearly juxtaposed with standard accuracy measures of the forecast \citep{bannigidadmath2016stock, zhang2019forecasting}, resource allocation in urban planning \citep{burnett2014urban, liang2013resource}, or sustainability initiatives such as water and energy allocation \citep{syme1999defining, gebre2021multi, colett2016using}.
 
@@
 epidemiological forecast could lead to policy decisions that change the distribution of the predicted outcome $Y$. 
-This excludes the fundamentally important context of vaccine allocation for certain pathogens and
-spatio-temporal scales, but we note that supply-chain decisions made at a national level in the US determining the state-level
-availability of current COVID-19 vaccines are unlikely to affect incidence or exposure levels at a 14-day horizon. In any event,
-we are not yet prepared to suggest how valid forecast evaluation using scores given by \eqref{eqn:AS-orc} would proceed when 
-both disease burden and unmet need are considered as endogenous variables.
+For example, this excludes the important problem of allocation of doses of a preventative vaccine.
+We have not developed a formal approach to forecast evaluation using scores given by \eqref{eqn:AS-orc} when both
+disease burden and unmet need are considered as endogenous variables. Such an approach would require some form of
+accounting for the causal effect of the allocation decision on observed measures of disease burden, e.g. using the available observed data to estimate the (counterfactual)
+number of cases or deaths averted from one or another allocation strategy.
 

### 2024-08-21 alloscore_manuscript/manuscript.Rnw
splice Nick's unsrtnat.bst hack for author trunc into rss.bst
fix a few last glitches

@@
 \usepackage{natbib}
-\bibliographystyle{unsrtnat-abbr}
+\bibliographystyle{rss-abbr}
 
@@
 equations \eqref{eqn:AS-raw} and \eqref{eqn:quantiles-sum-to-K} in Section \ref{sec:a:bayes-quantiles} of the appendix).
-It is however the case that when a forecaster modifies a single marginal forecast (holding those for other locations fixed), the aggregate WIS of thier forecast will change only according to the WIS of that marginal forecast, whereas every term in the sum defining the AS (in \eqref{eqn:AS-raw}) will change as the Bayes probability level defined by \eqref{eqn:quantiles-sum-to-K} changes.
+It is however the case that when a forecaster modifies a single marginal forecast (holding those for other locations fixed), the aggregate WIS of their forecast will change only according to the WIS of that marginal forecast, whereas every term in the sum defining the AS (in \eqref{eqn:AS-raw}) will change as the Bayes probability level defined by \eqref{eqn:quantiles-sum-to-K} changes.
 In this sense, the AS depends “jointly” on the marginal forecasts while the WIS does not.

### 2024-08-21 alloscore_manuscript/manuscript.tex
splice Nick's unsrtnat.bst hack for author trunc into rss.bst
fix a few last glitches

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu Aug 15 23:41:50 2024
+% Tue Aug 20 15:20:10 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu Aug 15 23:41:54 2024
+% Tue Aug 20 15:20:15 2024
 \begin{table}[ht]

### 2024-08-30 alloscore_manuscript/manuscript.Rnw
typo fix in joint dependence comment

@@
 
-We also note that the AS, just as the (aggregate or mean) quantile score and WIS, depends only on the marginal distributions for each location of a multi-location forecast (cf.
-equations \eqref{eqn:AS-raw} and \eqref{eqn:quantiles-sum-to-K} in Section \ref{sec:a:bayes-quantiles} of the appendix).
+We also note that the AS, just as the (aggregate or mean) quantile score and WIS, depends only on the marginal distributions for each location of a multi-location forecast (see equations \eqref{eqn:AS-raw} and \eqref{eqn:quantiles-sum-to-K} in Section \ref{sec:a:bayes-quantiles} of the appendix).
 It is however the case that when a forecaster modifies a single marginal forecast (holding those for other locations fixed), the aggregate WIS of their forecast will change only according to the WIS of that marginal forecast, whereas every term in the sum defining the AS (in \eqref{eqn:AS-raw}) will change as the Bayes probability level defined by \eqref{eqn:quantiles-sum-to-K} changes.
@@
 q_sum <- function(tau) q1(tau) + q2(tau)
-iters5 <- tibble(taus = 1-alloc$lam_seq[[1]], qtot = q_sum(taus))
-iters10 <- tibble(taus = 1-alloc$lam_seq[[2]], qtot = q_sum(taus))
+iters5 <- tibble(taus = 1-alloc$lam_seq[[1]], qtot = q_sum(taus), ind = seq_along(alloc$lam_seq[[1]]))
+iters10 <- tibble(taus = 1-alloc$lam_seq[[2]], qtot = q_sum(taus), ind = seq_along(alloc$lam_seq[[2]]))
 
@@
   geom_hline(aes(yintercept = c(1,4)), linetype = 2) +
-  geom_segment(data = iters5, aes(x=taus, xend = taus, y=0, yend = qtot), color = "red", alpha = .5) +
+  geom_segment(data = iters5, aes(x=taus, xend = taus, y=0, yend = qtot, color = ind), alpha = .5) +
+  scale_color_viridis_c("Iteration\n index", end = 0.75, direction=-1, limits = c(1, 18), option="B") +
   scale_x_continuous(limits = c(.45,1), expand = c(0,0), 
@@
                                 "0.7","0.8","0.9","1")) +
-  scale_y_continuous(limits = c(0,12), breaks = c(0,1,4,5,10,15), expand = c(0,0))
+  scale_y_continuous(limits = c(0,12), breaks = c(0,1,4,5,10,15), expand = c(0,0)) +
+  theme(legend.position = "none")
 
@@
   geom_hline(aes(yintercept = c(2,8)), linetype = 2) +
-  geom_segment(data = iters10, aes(x=taus, xend = taus, y=0, yend = qtot), color = "red", alpha = .5) +
+  geom_segment(data = iters10, aes(x=taus, xend = taus, y=0, yend = qtot, color = ind), alpha = .5) +
+  scale_color_viridis_c("Iteration\nindex", end = 0.75, direction=-1, limits = c(1, 18), option="B") +
   scale_x_continuous(limits = c(.45,1), expand = c(0,0), 
@@
 
-cowplot::plot_grid(p5, p10, align = "h")
-
+legend <- get_legend(p10 + theme(legend.box.margin = margin(0, 0, 0, 12)))
+p10 <- p10 + theme(legend.position = "none")
+
+plot_grid(
+  cowplot::plot_grid(p5, p10, align = "h"),
+  legend,
+  rel_widths = c(3, 0.4)
+)
 @

### 2024-08-31 alloscore_manuscript/manuscript.Rnw

@@
 
-We also note that the AS, just as the (aggregate or mean) quantile score and WIS, depends only on the marginal distributions for each location of a multi-location forecast (see equations \eqref{eqn:AS-raw} and \eqref{eqn:quantiles-sum-to-K} in Section \ref{sec:a:bayes-quantiles} of the appendix).
+We also note that the AS, just as the (aggregate or mean) quantile score and WIS, depends only on the marginal distributions for each location of a multi-location forecast (cf.
+equations \eqref{eqn:AS-raw} and \eqref{eqn:quantiles-sum-to-K} in Section \ref{sec:a:bayes-quantiles} of the appendix).
 It is however the case that when a forecaster modifies a single marginal forecast (holding those for other locations fixed), the aggregate WIS of their forecast will change only according to the WIS of that marginal forecast, whereas every term in the sum defining the AS (in \eqref{eqn:AS-raw}) will change as the Bayes probability level defined by \eqref{eqn:quantiles-sum-to-K} changes.

### 2024-09-09 alloscore_manuscript/manuscript.tex
fix 'highest at at'
response2, using hyperlinks to diff

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Tue Aug 20 15:20:10 2024
+% Mon Sep  9 16:35:24 2024
 \begin{table}[ht]
@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Tue Aug 20 15:20:15 2024
+% Mon Sep  9 16:35:29 2024
 \begin{table}[ht]
@@
 19,581 that day. Model ranks by AS were fairly stable across all $K$, and there appears
-to be an substantial interval around the observed value over which ranks are constant.
+to be a substantial interval around the observed value over which ranks are constant.
 
@@
 The AS is expressed in units that are directly relevant to decisions, which helps not only with ranking forecasts in the context of a specific action, but also allows decision-makers to contextualize the magnitude of differences between forecast scores.
+For example, it may be the case that differences in AS are small relative to the amount of available resources or the accuracy with which planned resource allocations
+can be executed.
 The benefits of the approach to forecast evaluation discussed here generalize beyond public health, to any application of probability forecasts for decision-making, such as economics, where the monetary consequences of using a forecast
